### 2015-11-02 analyze_cfec_data.Rmd
init commit

---
title: "Analyzing portfolio effects in commercial catch data"
author: "NCEAS commercial catch portfolio working group"
date: "October 30, 2015"
output:
  pdf_document:
    fig_caption: yes
---

## Overview
We are interested in multiple metrics of diversity. For this first paper, we plan to divide the analysis into several components:  
1. Characterize trends in diversity over time (using GAMs). The metrics of diversity we might want to include are  
- variation in species diversity  
- variation in gear diversity  
- variation in spatial diversity (number of commercial areas fished?)  
- variation in permit holdings over time  
2. Using diversity as the predictor, look at absolute or relative change revenue / vessel (or revenue / vessel / days fished) as the response.

For both of the above, there's a large number of other perturbations including EVOS, management changes, climate shifts, changes in gear or processsing capacity, hatchery production of salmon, market conditions (such as Japanese crash in mid-1990s), constraints on limited entry or quotas, and biological conditions or shifts (abundance, distribution, size structure, etc).

We'll initially restrict our analysis to data after 1984, since the data from 1975-1984 were dominated by foreign vessels (which we don't have). For more information, see AFSC technical reports (such as http://www.afsc.noaa.gov/Publications/AFSC-TM/NOAA-TM-AFSC-54.pdf). The cutoff of 1984/1985 is also consistent with the ADFG use. Further, the species reporting changes drastically before / after 1985 -- examples include no rockfish (demersal, pelagic, slope categories below), flatfish (shallow, deep categories below).

Additional information on CFEC data here,  
Gear descriptions: https://www.cfec.state.ak.us/misc/GEARDESC.HTM  
Fishery codes: https://www.cfec.state.ak.us/misc/FshyDesC.htm  
Species and gear codes:   https://www.adfg.alaska.gov/static/license/fishing/pdfs/elanding_codes.pdf  
Commercial areas: http://www.adfg.alaska.gov/index.cfm?adfg=fishingCommercialByFishery.statmaps  
Halibut areas: http://www.iphc.int/publications/techrep/tech0049.pdf

```{r chunkSet, cache=FALSE, echo=FALSE,warning=FALSE,message=FALSE}
library(knitr)
library(dplyr)
library(ggplot2)
library(vegan)
library(mgcv)
opts_chunk$set(tidy=TRUE, message=FALSE)
```

## Data processing
Load the data from the saved workspace. This is the raw cfec data that has been merged to have the same columns across years, 1975 - present. The key columns are the species labels ("spec") and species number codes ("specn"), the gear number codes ("gearn"), the vessel id ("adfg"), the area where the catch occurred ("stat6"), the pounds harvested ("g_pounds"), gross earnings ("g_earn").

### Note: for any confidentiality concerns, the workspace loaded here has been slightly jittered. This will need to be replaced with the real data for the final model. 
```{r loadData}
load("cfec_jitter.Rdata")
# initially only use data 1985+
cfec = cfec[cfec$year>1984,]
```

The first data processing step is to aggregate some of the species. This is largely needed because of changes in reporting over time. For example, some of the rockfishes weren't reported to the same species level. For other species, there are unidentified groups, so there's additional aggregation required.

```{r specLump}
# Lump all skates together
cfec$spec = as.character(cfec$spec)
cfec$specn[cfec$specn==700] = 700
cfec$specn[cfec$specn==701] = 700
cfec$specn[cfec$specn==702] = 700
cfec$spec[cfec$specn==700] = "SKATE"

# Lump all sharks together
cfec$specn[cfec$specn==691] = 689
cfec$specn[cfec$specn==689] = 689
cfec$spec[cfec$specn==689] = "SHARK"

# Lump flatfish (Shallow)
cfec$specn[cfec$specn==123] = 119 # rock
cfec$specn[cfec$specn==129] = 119 # starry flounder
cfec$specn[cfec$specn==127] = 119 # yellowfin
cfec$specn[cfec$specn==126] = 119 # butter
cfec$spec[cfec$specn==119] = "FLAT.shallow"

# Lump flatfish (deep)
cfec$specn[cfec$specn==124] = 118 # dover
cfec$specn[cfec$specn==134] = 118 # greenland turbot
cfec$spec[cfec$specn==118] = "FLAT.deep"

# Lump rockfish, codes from here:
cfec$specn[cfec$specn==141] = 144 # POP
cfec$specn[cfec$specn==136] = 144 # northern
cfec$specn[cfec$specn==151] = 144 # rougheye
cfec$specn[cfec$specn==166] = 144 # sharpchin
cfec$specn[cfec$specn==152] = 144 # shortraker
cfec$specn[cfec$specn==153] = 144 # redbanded
cfec$specn[cfec$specn==157] = 144 # silvergray
cfec$specn[cfec$specn==158] = 144 # redstripe
cfec$specn[cfec$specn==137] = 144 # bocaccio
cfec$specn[cfec$specn==144] = 144 # unid
cfec$spec[cfec$specn==144] = "ROCK.slope"

cfec$specn[cfec$specn==138] = 168 # copper
cfec$specn[cfec$specn==145] = 168 # yelloweye
cfec$specn[cfec$specn==146] = 168 # Canary
cfec$specn[cfec$specn==147] = 168 # quillback
cfec$specn[cfec$specn==148] = 168 # tiger
cfec$specn[cfec$specn==149] = 168 # China
cfec$specn[cfec$specn==150] = 168 # rosethorn
cfec$spec[cfec$specn==168] = "ROCK.demersal"

cfec$specn[cfec$specn==154] = 169 # dusky
cfec$specn[cfec$specn==155] = 169 # yellowtail
cfec$specn[cfec$specn==156] = 169 # widow
cfec$specn[cfec$specn==172] = 169 # dusky
cfec$specn[cfec$specn==173] = 169 # dark
cfec$specn[cfec$specn==142] = 169 # black
cfec$specn[cfec$specn==167] = 169 # blue
cfec$spec[cfec$specn==169] = "ROCK.pelagic"

# Lump king crabs
cfec$spec[cfec$spec%in%c("GKCR","BKCR","UKCR","RKCR")] = "KCRB" # king
# Lump tanner crabs
cfec$spec[cfec$spec%in%c("UTCR","BTCR")] = "TCRB" # tanner
# Lump shrimp - spot prawn, pink shrimp, 
cfec$spec[cfec$spec%in%c("SPRW","PSHP","USRM")] = "USRM" # shrimp

# drop records with no spp
cfec = cfec[-which(cfec$spec==""),]
```

Next, we'll deal with processing the areas. We have multiple types of area codes. The 6-digit codes represent groundfish management areas, the 5-digit codes are state salmon/shellfish areas, and the 3-digit codes are the halibut areas. For the groundfish and salmon areas, there are sub-areas that are also reported (as the last 2-digits), so we'll lump sub-areas together by stripping off the last 2-digits.

```{r}
# Filter out sub-districts and group things at larger areas
# Lump statistical areas by district -- groundfish
cfec$stat6 = as.character(as.numeric(cfec$stat6)) # as.numeric needed because of dumb integer
cfec$areaType = NA
cfec$stat6_char = NA
indx = which(cfec$stat6 > 99999 & cfec$stat6 < 999999)
cfec$stat6_char[indx] = nchar(cfec$stat6[indx])
cfec$areaType[indx] = "groundfish"
# This is just turning the last digit - sub region - into 0
cfec$stat6[indx] = paste(substr(cfec$stat6[indx],1,5), "0",sep="")
cfec$stat6[indx] = as.numeric(cfec$stat6[indx])
# Now do salmon / shellfish - This is just turning the last 2 digits - sub region - into 00
indx = which(cfec$stat6 > 9999 & cfec$stat6 < 100000)
cfec$stat6[indx] = paste(substr(cfec$stat6[indx],1,3), "00",sep="")
cfec$areaType[indx] = "salmonShellfish"

cfec$stat6 = as.numeric(cfec$stat6)
# Deal with the IPHC areas too. We can verify that nearly all spp caught in 3-digit areas are halibut,
#sort(table(cfec$spec[which(cfec$stat6 < 1000)]))
```

Next, we'll aggregate the stat6 codes into larger geographic regions. The focus here is on Prince William Sound, but for contrast we'll include Southeast Alaska, Kodiak W / E, and Cook Inlet.

```{r}
# Add the areas -- 5 digit are salmon/shellfish, 6-digit are groundfish
cfec$area = NA
cfec$area[cfec$stat6==485900] = "PWS"
cfec$area[cfec$stat6==475900] = "PWS"
cfec$area[cfec$stat6==465900] = "PWS"
cfec$area[cfec$stat6==455900] = "PWS"
cfec$area[cfec$stat6==445900] = "PWS"
cfec$area[cfec$stat6==495930] = "PWS"
cfec$area[cfec$stat6==485930] = "PWS"
cfec$area[cfec$stat6==475930] = "PWS"
cfec$area[cfec$stat6==465930] = "PWS"
cfec$area[cfec$stat6==455930] = "PWS"
cfec$area[cfec$stat6==445930] = "PWS"
cfec$area[cfec$stat6==486000] = "PWS"
cfec$area[cfec$stat6==476000] = "PWS"
cfec$area[cfec$stat6==466000] = "PWS"
cfec$area[cfec$stat6==456000] = "PWS"
cfec$area[cfec$stat6==446000] = "PWS"
cfec$area[cfec$stat6==486030] = "PWS"
cfec$area[cfec$stat6==476030] = "PWS"
cfec$area[cfec$stat6==466030] = "PWS"
cfec$area[cfec$stat6==456030] = "PWS"
cfec$area[cfec$stat6==486100] = "PWS"
cfec$area[cfec$stat6==476100] = "PWS"
cfec$area[cfec$stat6==466100] = "PWS"
cfec$area[which(cfec$stat6 < 20099 & cfec$stat6 >= 20000)] = "PWS"
cfec$area[which(cfec$stat6 < 21299 & cfec$stat6 >= 21200)] = "PWS"
cfec$area[which(cfec$stat6 < 22799 & cfec$stat6 >= 22700)] = "PWS"  
cfec$area[which(cfec$stat6 < 22699 & cfec$stat6 >= 22600)] = "PWS"
cfec$area[which(cfec$stat6 < 22899 & cfec$stat6 >= 22800)] = "PWS"
cfec$area[which(cfec$stat6 < 22199 & cfec$stat6 >= 22100)] = "PWS"
cfec$area[which(cfec$stat6 < 22299 & cfec$stat6 >= 22200)] = "PWS"  
cfec$area[which(cfec$stat6 < 22599 & cfec$stat6 >= 22500)] = "PWS"
cfec$area[which(cfec$stat6 < 22499 & cfec$stat6 >= 22400)] = "PWS"
cfec$area[which(cfec$stat6 < 22399 & cfec$stat6 >= 22300)] = "PWS" 
# Add halibut areas
cfec$area[cfec$stat6 %in% c(242,232,220,230,240)] = "PWS"

cfec$area[cfec$stat6==525900] = "CookInlet"
cfec$area[cfec$stat6==535900] = "CookInlet"
cfec$area[cfec$stat6==515930] = "CookInlet"
cfec$area[cfec$stat6==525930] = "CookInlet"
cfec$area[cfec$stat6==535930] = "CookInlet"
cfec$area[cfec$stat6==516000] = "CookInlet"
cfec$area[cfec$stat6==526000] = "CookInlet"
cfec$area[cfec$stat6==496030] = "CookInlet"
cfec$area[cfec$stat6==506030] = "CookInlet"
cfec$area[cfec$stat6==516030] = "CookInlet"
cfec$area[cfec$stat6==526030] = "CookInlet"
cfec$area[cfec$stat6==506100] = "CookInlet"
cfec$area[cfec$stat6==516100] = "CookInlet"
cfec$area[which(cfec$stat6 < 24799 & cfec$stat6 >= 24400)] = "CookInlet" 
cfec$area[which(cfec$stat6%in%c(261, 272))] = "CookInlet"

cfec$area[which(cfec$stat6 < 11599 & cfec$stat6 >= 11100)] = "Southeast"
cfec$area[which(cfec$stat6 < 10499 & cfec$stat6 >= 10100)] = "Southeast"
cfec$area[which(cfec$stat6 < 11099 & cfec$stat6 >= 10500)] = "Southeast" 
cfec$area[which(cfec$stat6 >= 305430 & cfec$stat6 <= 305530)] = "Southeast"
cfec$area[which(cfec$stat6 >= 315430 & cfec$stat6 <= 315600)] = "Southeast"
cfec$area[which(cfec$stat6 >= 325430 & cfec$stat6 <= 325700)] = "Southeast"
cfec$area[which(cfec$stat6 >= 335430 & cfec$stat6 <= 335800)] = "Southeast"
cfec$area[which(cfec$stat6 >= 345400 & cfec$stat6 <= 345830)] = "Southeast"
cfec$area[which(cfec$stat6 >= 355400 & cfec$stat6 <= 355900)] = "Southeast"
cfec$area[which(cfec$stat6 > 365400 & cfec$stat6 < 365830)] = "Southeast"
# For IPHC see Fig. 20 here: http://www.iphc.int/publications/techrep/tech0049.pdf
cfec$area[which(cfec$stat6%in%c(140, 141, 142, 143, 144, 150, 151, 152, 153, 160, 161, 162, 163, 170, 171, 172, 174, 174, 181, 182, 183, 184, 185))] = "Southeast"

# Kodiak east area is defined by halibut areas 270/280, includes ADFG salmon / 
# shellfish on E side of kodiak, and groundfish areas extending to NMFS Area 630
# boundary
cfec$area[cfec$stat6%in%c(270,280)] = "Kodiak.east"
cfec$area[which(cfec$stat6 < 25899 & cfec$stat6 >= 25800)] = "Kodiak.east"
cfec$area[which(cfec$stat6 < 25999 & cfec$stat6 >= 25900)] = "Kodiak.east"
cfec$area[cfec$stat6==535630] = "Kodiak.east"
cfec$area[cfec$stat6==525630] = "Kodiak.east"
cfec$area[cfec$stat6==515630] = "Kodiak.east"
cfec$area[cfec$stat6==505630] = "Kodiak.east"
cfec$area[cfec$stat6==495630] = "Kodiak.east"
cfec$area[cfec$stat6==535700] = "Kodiak.east"
cfec$area[cfec$stat6==525700] = "Kodiak.east"
cfec$area[cfec$stat6==515700] = "Kodiak.east"
cfec$area[cfec$stat6==505700] = "Kodiak.east"
cfec$area[cfec$stat6==495700] = "Kodiak.east"
cfec$area[cfec$stat6==525730] = "Kodiak.east"
cfec$area[cfec$stat6==515730] = "Kodiak.east"
cfec$area[cfec$stat6==505730] = "Kodiak.east"
cfec$area[cfec$stat6==495730] = "Kodiak.east"
cfec$area[cfec$stat6==525800] = "Kodiak.east"
cfec$area[cfec$stat6==515800] = "Kodiak.east"
cfec$area[cfec$stat6==505800] = "Kodiak.east"
cfec$area[cfec$stat6==495800] = "Kodiak.east"

cfec$area[cfec$stat6%in%c(271,281)] = "Kodiak.west"
cfec$area[cfec$stat6==535730] = "Kodiak.west"
cfec$area[cfec$stat6==545730] = "Kodiak.west"
cfec$area[cfec$stat6==525800] = "Kodiak.west"
cfec$area[cfec$stat6==535800] = "Kodiak.west"
cfec$area[cfec$stat6==545800] = "Kodiak.west"
cfec$area[cfec$stat6==535830] = "Kodiak.west"
cfec$area[cfec$stat6==525830] = "Kodiak.west"
cfec$area[which(cfec$stat6 < 25699 & cfec$stat6 >= 25500)] = "Kodiak.west"
cfec$area[which(cfec$stat6 < 25499 & cfec$stat6 >= 25300)] = "Kodiak.west"  
cfec$area[which(cfec$stat6 < 26299 & cfec$stat6 >= 26000)] = "Kodiak.west"

# Bristol Bay, defined as going west to the boundary between NMFS areas 512 / 516
cfec$area[cfec$stat6%in%c(31600, 31700, 31800, 32000, 32100, 32200, 32400, 32500, 32600)] = "BristolBay"
cfec$area[cfec$stat6%in%c(575730, 575800, 575830)] = "BristolBay"
cfec$area[cfec$stat6%in%c(585700, 585730, 585800, 585830)] = "BristolBay"
cfec$area[cfec$stat6%in%c(595630, 595700, 595730, 595800, 595830)] = "BristolBay"
cfec$area[cfec$stat6%in%c(605600, 605630, 605700, 605730, 605800, 605830)] = "BristolBay"
cfec$area[cfec$stat6%in%c(615600, 615630, 615700, 615730, 615800, 615830)] = "BristolBay"

# Area to add - Alaska peninsula - Halibut areas 300-340?
```

## Exploratory data analysis

Drop all species records that have less than 1000 total observations. This is discarding about 5000 of 20 million records. 
```{r discardSpec}
cfecClean = group_by(cfec, spec) %>%
  mutate(nspecies = n()) %>%
  filter(nspecies > 1000) %>%
  data.frame

# Also drop out unid groundfish, unid flatfish, smelts, grenadier, english sole -- all very rare
cfecClean = cfecClean[-which(cfecClean$spec%in%c("UFLT","UGRN","SMLT","GRDR","EGLS")),]

```

Drop all species records with gears occuring < 1000 times. This is discarding about 1250 records.
```{r discardGear}
cfecClean = group_by(cfecClean, gearn) %>%
  mutate(ngears = n()) %>%
  filter(ngears > 1000) %>%
  data.frame
```

Delete catch from fisheries that have less than 100 total records. This simplifies the dimension of the fishery codes a lot (from 378 to 234), removes about 3050 records, and doesn't affect the total earnings (these fisheries represent 0.0009 of total g_earn).  
```{r discardFishery}
cfecClean = group_by(cfecClean, p_fshy) %>%
  mutate(nfshy = n()) %>%
  filter(nfshy > 100) %>%
  data.frame
```

Subset the data to only include vessels that fished at any point in Prince William Sound.  
```{r useOnlyPWS}
cfecPWS = group_by(cfecClean, adfg) %>%
  mutate(in_PWS = any(area%in%"PWS")) %>%
  as.data.frame() %>%
  filter(in_PWS) %>%
  data.frame
```

Alan suggested looking at some coarse plots. First, we can show participation of vessels over time. Halibut fishing pre-1995 needed to get quota by catching halibut in 1988-1990.  
```{r,fig.cap="Time series of participation by year",fig.pos="placeHere"}
plot(min(cfecPWS$year):max(cfecPWS$year), specnumber(table(cfecPWS$year,cfecPWS$adfg)), xlab="",ylab="# Vessels", type="b")
```

We can also look at species by year

```{r,fig.cap="Time series of species caught by year",fig.pos="placeHere"}
# total species
par(mfrow = c(2,1), mgp = c(2,1,0), mai=c(0.5,0.5,0.3,0.1))
plot(min(cfecPWS$year):max(cfecPWS$year), specnumber(table(cfecPWS$year, cfecPWS$spec)), xlab="",ylab="# Species", type="b",main="Catch from vessels that ever fished in PWS")

plot(min(cfecPWS$year):max(cfecPWS$year), specnumber(table(cfecPWS$year[cfecPWS$area=="PWS"], cfecPWS$spec[cfecPWS$area=="PWS"])), xlab="",ylab="# Species", type="b",main="Catch only in PWS")
```

Look at species diversity using Vegan. How has diversity in # species fished changed over time?   
```{r veganSpecies,echo=FALSE, fig.cap="Time series of species diversity",fig.pos="placeHere"}
nSpecies.mean = NA # mean richness
nSpecies.cv = NA # cv richness
simp.mean = NA # mean simpson
simp.cv = NA # cv simpson
shannon.mean = NA # mean shannon
shannon.cv = NA # cv shannon
for(yr in min(cfecPWS$year):max(cfecPWS$year)) {
  nSpecies.mean[yr+1-min(cfecPWS$year)] = mean(specnumber(table(cfecPWS$adfg[cfecPWS$year==yr], cfecPWS$spec[cfecPWS$year==yr])))
   nSpecies.cv[yr+1-min(cfecPWS$year)] = sd(specnumber(table(cfecPWS$adfg[cfecPWS$year==yr], cfecPWS$spec[cfecPWS$year==yr])))/nSpecies.mean[yr+1-min(cfecPWS$year)]
  simp.mean[yr+1-min(cfecPWS$year)] = mean(diversity(table(cfecPWS$adfg[cfecPWS$year==yr], cfecPWS$spec[cfecPWS$year==yr]),index="invsimpson"))
  simp.cv[yr+1-min(cfecPWS$year)] = sd(diversity(table(cfecPWS$adfg[cfecPWS$year==yr], cfecPWS$spec[cfecPWS$year==yr]),index="invsimpson"))/simp.mean[yr+1-min(cfecPWS$year)]
  shannon.mean[yr+1-min(cfecPWS$year)] = mean(diversity(table(cfecPWS$adfg[cfecPWS$year==yr], cfecPWS$spec[cfecPWS$year==yr])))
  shannon.cv[yr+1-min(cfecPWS$year)] = sd(diversity(table(cfecPWS$adfg[cfecPWS$year==yr], cfecPWS$spec[cfecPWS$year==yr])))/simp.mean[yr+1-min(cfecPWS$year)]  
}

par(mfrow = c(3,2), mgp=c(2,1,0),mai=c(0.3,0.5,0.1,0.1))
plot(min(cfecPWS$year):max(cfecPWS$year),nSpecies.mean,ylab="Mean species richness",type="b")
plot(min(cfecPWS$year):max(cfecPWS$year),nSpecies.cv,ylab="CV species richness",type="b")
plot(min(cfecPWS$year):max(cfecPWS$year),simp.mean,ylab="Mean simpson div",type="b")
plot(min(cfecPWS$year):max(cfecPWS$year),simp.cv,ylab="CV simpson div",type="b")
plot(min(cfecPWS$year):max(cfecPWS$year),shannon.mean,ylab="Mean Shannon div",type="b")
plot(min(cfecPWS$year):max(cfecPWS$year),shannon.cv,ylab="CV Shannon div",type="b")
```

We can do similar plots for the number of gears used / vessel  
```{r veganGears,echo=FALSE, fig.cap="Time series of diversity of number of fishing gears",fig.pos="placeHere"}
nGears.mean = NA # mean richness
nGears.cv = NA # cv richness
simp.mean = NA # mean simpson
simp.cv = NA # cv simpson
shannon.mean = NA # mean shannon
shannon.cv = NA # cv shannon
for(yr in min(cfecPWS$year):max(cfecPWS$year)) {
  nGears.mean[yr+1-min(cfecPWS$year)] = mean(specnumber(table(cfecPWS$adfg[cfecPWS$year==yr], cfecPWS$gearn[cfecPWS$year==yr])))
   nGears.cv[yr+1-min(cfecPWS$year)] = sd(specnumber(table(cfecPWS$adfg[cfecPWS$year==yr], cfecPWS$gearn[cfecPWS$year==yr])))/nGears.mean[yr+1-min(cfecPWS$year)]
  simp.mean[yr+1-min(cfecPWS$year)] = mean(diversity(table(cfecPWS$adfg[cfecPWS$year==yr], cfecPWS$gearn[cfecPWS$year==yr]),index="invsimpson"))
  simp.cv[yr+1-min(cfecPWS$year)] = sd(diversity(table(cfecPWS$adfg[cfecPWS$year==yr], cfecPWS$gearn[cfecPWS$year==yr]),index="invsimpson"))/simp.mean[yr+1-min(cfecPWS$year)]
  shannon.mean[yr+1-min(cfecPWS$year)] = mean(diversity(table(cfecPWS$adfg[cfecPWS$year==yr], cfecPWS$gearn[cfecPWS$year==yr])))
  shannon.cv[yr+1-min(cfecPWS$year)] = sd(diversity(table(cfecPWS$adfg[cfecPWS$year==yr], cfecPWS$gearn[cfecPWS$year==yr])))/simp.mean[yr+1-min(cfecPWS$year)]  
}

par(mfrow = c(3,2), mai=c(0.3,0.5,0.1,0.1))
plot(min(cfecPWS$year):max(cfecPWS$year),nGears.mean,ylab="Mean gear richness",type="b")
plot(min(cfecPWS$year):max(cfecPWS$year),nGears.cv,ylab="CV gear richness",type="b")
plot(min(cfecPWS$year):max(cfecPWS$year),simp.mean,ylab="Mean simpson div",type="b")
plot(min(cfecPWS$year):max(cfecPWS$year),simp.cv,ylab="CV simpson div",type="b")
plot(min(cfecPWS$year):max(cfecPWS$year),shannon.mean,ylab="Mean Shannon div",type="b")
plot(min(cfecPWS$year):max(cfecPWS$year),shannon.cv,ylab="CV Shannon div",type="b")
```

And we can do similar plots for permit types per vessel  
```{r veganPermits,echo=FALSE, fig.cap="Time series of diversity of number of permits used",fig.pos="placeHere"}
nPermits.mean = NA # mean richness
nPermits.cv = NA # cv richness
simp.mean = NA # mean simpson
simp.cv = NA # cv simpson
shannon.mean = NA # mean shannon
shannon.cv = NA # cv shannon
for(yr in min(cfecPWS$year):max(cfecPWS$year)) {
  nPermits.mean[yr+1-min(cfecPWS$year)] = mean(specnumber(table(cfecPWS$adfg[cfecPWS$year==yr], cfecPWS$p_fshy[cfecPWS$year==yr])))
   nPermits.cv[yr+1-min(cfecPWS$year)] = sd(specnumber(table(cfecPWS$adfg[cfecPWS$year==yr], cfecPWS$p_fshy[cfecPWS$year==yr])))/nPermits.mean[yr+1-min(cfecPWS$year)]
  simp.mean[yr+1-min(cfecPWS$year)] = mean(diversity(table(cfecPWS$adfg[cfecPWS$year==yr], cfecPWS$p_fshy[cfecPWS$year==yr]),index="invsimpson"))
  simp.cv[yr+1-min(cfecPWS$year)] = sd(diversity(table(cfecPWS$adfg[cfecPWS$year==yr], cfecPWS$p_fshy[cfecPWS$year==yr]),index="invsimpson"))/simp.mean[yr+1-min(cfecPWS$year)]
  shannon.mean[yr+1-min(cfecPWS$year)] = mean(diversity(table(cfecPWS$adfg[cfecPWS$year==yr], cfecPWS$p_fshy[cfecPWS$year==yr])))
  shannon.cv[yr+1-min(cfecPWS$year)] = sd(diversity(table(cfecPWS$adfg[cfecPWS$year==yr], cfecPWS$p_fshy[cfecPWS$year==yr])))/simp.mean[yr+1-min(cfecPWS$year)]  
}

par(mfrow = c(3,2), mai=c(0.3,0.5,0.1,0.1))
plot(min(cfecPWS$year):max(cfecPWS$year),nPermits.mean,ylab="Mean permit richness",type="b")
plot(min(cfecPWS$year):max(cfecPWS$year),nPermits.cv,ylab="CV permit richness",type="b")
plot(min(cfecPWS$year):max(cfecPWS$year),simp.mean,ylab="Mean simpson div",type="b")
plot(min(cfecPWS$year):max(cfecPWS$year),simp.cv,ylab="CV simpson div",type="b")
plot(min(cfecPWS$year):max(cfecPWS$year),shannon.mean,ylab="Mean Shannon div",type="b")
plot(min(cfecPWS$year):max(cfecPWS$year),shannon.cv,ylab="CV Shannon div",type="b")
```

```{r, echo=FALSE,include=FALSE}
group_by(cfecPWS, adfg) %>%
  summarize(n = length(unique(year))) %>%
  ggplot(aes(n)) + geom_histogram()
# To do: 7 - year peaks are weird. Survey? 
```

Make exploratory plot of vessels that fished only before and after 1989. These aren't shown -- but we looked at them in Santa Barbara, and the major take home point is that more vessels left the fishery after 1989 than joined.

```{r,fig.cap="Histogram of vessels that only fished in PWS before 1990",fig.pos="placeHere", include=FALSE,echo=FALSE}
group_by(cfecPWS, adfg) %>%
mutate(before = ifelse(min(year) < 1990, 1, 0)) %>% 
  filter(before == 1) %>%
  summarize(n = length(unique(year))) %>%
  ggplot(aes(n)) + geom_histogram()
```

```{r,fig.cap="Histogram of vessels that only fished in PWS after 1990",fig.pos="placeHere", include=FALSE,echo=FALSE}
group_by(cfecPWS, adfg) %>%
mutate(after = ifelse(min(year) >= 1990, 1, 0)) %>% 
  filter(after == 1) %>%
  summarize(n = length(unique(year))) %>%
  ggplot(aes(n)) + geom_histogram()
```

## Metric of change
We can use absolute or relative earnings over time, by vessel. We adopted the relative measure, because we're comparing vessels of very different size classes. Sizes of individual vessels may change slowly over time. The earning metric is calculated as the first-differenced log earnings from time (t+1) to time (t), $log(earn_{t+1}) - log(earn_{t})$.

The key challenge here is that we need a metric of earnings that will allow us to compare different types of vessels (e.g. large, small) that don't overlap in time. For example, we need a metric that standardizes catches or earnings from a small vessel fishing 1985-1990 to that of a large vessel fishing 2000-2005. The relative change is universal across vessels, and can be interpreted as percent change (e.g. a value of -1 = exp(-1) = 0.367, or a ~ 63% decline in earnings between years).

We'll need to be cautious in putting together the covariates if the response is first differenced. If we code up indicators for press / pulse perturbations, those need to be first differenced too.

## Aggregating data
The dataset currently is individual records. This needs to be aggregated by vessel / year, so there are totals for each individual vessel. The 3 key dimensions we want to explore with variability are diversity in (1) species, (2) gears, and (3) areas. Diversity in fishery permits is also included for exploration, but permits confounds areas, species, and gears. We need to do some lumping, so we'll initially aggregate  
- 10 or more species  
- 2 or more areas  
- 3 or more gears  
- 4 or more fisheries

```{r aggregatePWSVessels}
totalPWS = group_by(cfecPWS, adfg, year) %>%
  summarize(totalearnings = sum(g_earn), 
            nspecies = length(unique(spec)),  
            ngear = length(unique(gearn)),  
            narea = length(unique(area)), 
            nfishery = length(unique(p_fshy))) 
# Lump larger groups
totalPWS$nspecies[totalPWS$nspecies>10] = 10
totalPWS$narea[totalPWS$narea>2] = 2
totalPWS$ngear[totalPWS$ngear>3] = 3
totalPWS$nfishery[totalPWS$nfishery>4] = 4
```

Add in NAs for missing years. We need these values to be included so that when we first difference, we're not taking differences across > 1 year. When we take the differences, we're also having variables for number of species, gears, etc (along with the differences of each predictor). For the undifferenced data, the number of species, gears, etc. corresponds to the value in the second year. We may want to replace this with a 2-year mean or something.

```{r mergeYears}
dummyDF = expand.grid(adfg = unique(totalPWS$adfg), year = min(totalPWS$year):max(totalPWS$year), stringsAsFactors = F)
# expand the data frame w/missing years
totalPWSExp = left_join(dummyDF, totalPWS) %>%
  arrange(adfg, year)
totalPWSExp = group_by(totalPWSExp, adfg) %>%
  mutate(difflog = c(NA, diff(log(totalearnings))), 
         diffspec = c(NA, diff(nspecies)),
         diffgear = c(NA, diff(ngear)),
         diffarea = c(NA, diff(narea)),
         difffishery = c(NA, diff(nfishery))) %>%
  filter(!is.na(difflog))
```

Now with our aggregated dataset, we can make histograms of each predictor.

```{r,fig.cap="Histogram of predictors (across all years and data)",fig.pos="placeHere"}
par(mfrow = c(2,2), mai=c(0.6,0.6,0.1,0.1))
#hist(log(totalPWSExp$totalearnings), main="Log of earnings",xlab="",col="grey")
hist(totalPWSExp$nspecies, main="Number of species", xlab="", col="grey")
hist(totalPWSExp$ngear, main="Number of gears", xlab="", col="grey")
hist(totalPWSExp$narea, main="Number of areas", xlab="", col="grey")
hist(totalPWSExp$nfishery, main="Number of fisheries", xlab="", col="grey")
```

We can also look at change in these predictors over time -- we can also look at the response, which is change in earnings from time t to time t+1.
```{r,fig.cap="Histogram of change in predictors (across all years and data)",fig.pos="placeHere"}
par(mfrow = c(3,2), mai=c(0.6,0.6,0.2,0.1))
hist(totalPWSExp$difflog, main="Log of change in earnings", xlab="", col="grey")
hist(totalPWSExp$diffspec, main="Change in species", xlab="", col="grey")
hist(totalPWSExp$diffgear, main="Change in gears", xlab="", col="grey")
hist(totalPWSExp$diffarea, main="Change in areas", xlab="", col="grey")
hist(totalPWSExp$difffishery, main="Change in fisheries", xlab="", col="grey")
```

## Plotting change in earnings through time

We can start by making plots for individual data points where the dimension doesn't change (e.g. individuals fish 1 species in both time t and time t+1).

```{r,fig.cap="Median and quartiles of change in earnings, by species, for individuals that don't change number of species fished through time",fig.pos="placeHere"}
group_by(totalPWSExp[totalPWSExp$diffspec==0,], year, nspecies) %>% 
  summarize(medDiffLog = median(difflog),  
  lower = quantile(difflog,probs = 0.25), 
  upper = quantile(difflog,probs = 0.75)) %>% 
ggplot(aes(year, medDiffLog)) + geom_line() + facet_wrap(~nspecies) +  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.2)
```

```{r,fig.cap="Median and quartiles of change in earnings, by gears, for individuals that don't change number of gears fished through time",fig.pos="placeHere"}
group_by(totalPWSExp[totalPWSExp$diffgear==0,], year, ngear) %>% 
  summarize(medDiffLog = median(difflog), 
  lower = quantile(difflog,probs = 0.25), 
  upper = quantile(difflog,probs = 0.75)) %>% 
ggplot(aes(year, medDiffLog)) + geom_line() + facet_wrap(~ngear) +  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.2)
```

```{r,fig.cap="Median and quartiles of change in earnings, by areas, for individuals that don't change number of areas fished through time",fig.pos="placeHere"}
group_by(totalPWSExp[totalPWSExp$diffarea==0,], year, narea) %>% 
  summarize(medDiffLog = median(difflog), 
  lower = quantile(difflog,probs = 0.25), 
  upper = quantile(difflog,probs = 0.75)) %>% 
ggplot(aes(year, medDiffLog)) + geom_line() + facet_wrap(~narea) +  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.2)
```

```{r,fig.cap="Median and quartiles of change in earnings, by areas, for individuals that don't change number of fisheries fished through time",fig.pos="placeHere"}
group_by(totalPWSExp[totalPWSExp$difffishery==0,], year, nfishery) %>% 
  summarize(medDiffLog = median(difflog), 
  lower = quantile(difflog,probs = 0.25), 
  upper = quantile(difflog,probs = 0.75)) %>% 
ggplot(aes(year, medDiffLog)) + geom_line() + facet_wrap(~nfishery) +  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.2)
```

We can break the differences out further, by species (across columns) and by number of areas fished. The areas are grouped by vessels fishing in one area (just PWS) versus vessels that fish in PWS and 1 or more other regions. This plot includes some confounding, so don't read too much into it. 
```{r,fig.cap="Median and quartiles of change in earnings, by species x areas",fig.pos="placeHere"}
totalPWSExp2 = totalPWSExp
totalPWSExp2$nspecies[totalPWSExp2$nspecies > 5] = 5
group_by(totalPWSExp2, year, nspecies, narea) %>% 
  summarize(medDiffLog = median(difflog), 
            lower = quantile(difflog,probs = 0.25), 
            upper = quantile(difflog,probs = 0.75)) %>% 
  ggplot(aes(year, medDiffLog)) + geom_line() + facet_grid(narea~nspecies) +  geom_ribbon(aes(ymin=lower, ymax=upper), alpha=0.2) 
```

Make some summary tables of vessels fishing in 1988 / 1989 that only landed one species. Make tables of the most common species landed, both by records and earnings. 
```{r}
vOne = filter(totalPWSExp,nspecies==1 & diffspec==0) %>%
  select(adfg) %>% #select(year,adfg)
  unique

LJ = left_join(vOne, cfecPWS)

# Look at species that fishers are specializing on
head(table(LJ$spec[LJ$year%in%seq(1988,1989)]) %>%
  sort %>%
  rev)

# Look at earnings of fishers that people are specializing on
LJ89 = LJ[LJ$year==1989,]
LJ89table = aggregate(g_earn ~ spec, LJ89, sum)
head(LJ89table[with(LJ89table, order(-g_earn)), ])

```

We can also get really into the weeds and make plots of individual annual landings, by vessel and species. For confidentiality, these aren't shown but the code can be run.  
```{r,include=FALSE,echo=FALSE}
#LJsummarized = group_by(LJ, year, spec,adfg) %>%
#  summarize(total=sum(g_earn)) %>%
#  as.data.frame() %>%
#  arrange(adfg, year, spec)
#ggplot(LJsummarized[1:10000,], aes(year,total,color=spec)) + geom_line()+facet_wrap(~adfg, scales="free_y")
```

## Fitting models to data

Before we fit anything, we need to calculate the mean of each predictor (# species, gears, areas) over the 2 year period. This will basically scale the effect, and will be included alongside predictors representing the change in each of these variables.

Are a priori expectation is that if diversity (species, gear, area) is beneficial, all of the differenced predictors (diffspec, diffgear, diffarea) will have positive effects on the response (change in earnings). But if the differenced values are small, it means individual strategy is relatively constant, and we'd also expect the effects of diverification on the mean values (nspecies, narea, ngear).

```{r runningMean}
totalPWSExp$nspecies = (totalPWSExp$nspecies + (totalPWSExp$nspecies + totalPWSExp$diffspec))/2
totalPWSExp$ngear = (totalPWSExp$ngear + (totalPWSExp$ngear + totalPWSExp$diffgear))/2
totalPWSExp$narea = (totalPWSExp$narea + (totalPWSExp$narea + totalPWSExp$diffarea))/2
totalPWSExp$narea = (totalPWSExp$narea + (totalPWSExp$narea + totalPWSExp$diffarea))/2
```

Start with simple linear model, no random effects  
```{r linearModel}
lm.simple = lm(difflog~as.factor(year) + nspecies + ngear + narea + diffspec + diffgear + diffarea,data=totalPWSExp)
summary(lm.simple)
```

We can also fit a gam, with spline over year, and basically get the same result. These models show that changing strategy (increasing gears used or areas fished correlates with more earnings, but the overall strategy of more areas or gears doesn't have a positive effect). In contrast, the effect of species is positive in both, suggesting that fishing more species translates to positive revenue. Other splines can be included (nspecies, diffspec) but the effects are basically linear (not shown).

```{r simpleGam}
gam.simple = gam(difflog~s(year) + nspecies + ngear + narea + diffspec + diffgear + diffarea,data=totalPWSExp)
summary(gam.simple)
```

## Extending model across areas

First, it's useful to look at the number of vessels that can be uniquely assigned to each of our areas. This shows that > 80% of vessels never fish in more than 1 of our coarse areas (Prince William Sound, Cook Inlet, Bristol Bay, E/W Kodiak, etc), so we can start our analysis focusing on those.  
```{r tableVesselArea}
tableVesselArea = table(cfecClean$adfg, cfecClean$area)
tableVesselArea = ceiling(tableVesselArea/1.0e10)
T = table(apply(tableVesselArea,1,sum))
m = matrix(NA, 3, 2)
colnames(m) = c("# Vessels", "% Total")
rownames(m) = c("1","2","3+")
m[,1] = c(T[2],T[3],sum(T[-c(1:3)]))
m[,2] = round(m[,1]/sum(m[,1]),3)

kable(m) # Make simple table w/ knitr
```

Next, we'll extract the vessels that are only assigned to one area, subsetting the entire dataset to just these vessels.
```{r }
singleAreaVessels = rownames(tableVesselArea)[which( apply(tableVesselArea, 1, sum) == 1)]
cfecClean$adfg = as.character(cfecClean$adfg)
cfecSingles = group_by(cfecClean, adfg) %>% 
  mutate(is_Single = any(adfg %in% singleAreaVessels)) %>%
  as.data.frame() %>%
  filter(is_Single) %>%
  data.frame
```

We can re-run the code above to total the response / predictors by calendar year. We'll add a column for area. 
```{r aggregateSingleVessel}
totalSingles = group_by(cfecSingles, adfg, year) %>%
  summarize(totalearnings = sum(g_earn), 
            nspecies = length(unique(spec)),  
            ngear = length(unique(gearn)),  
            narea = length(unique(area)), 
            nfishery = length(unique(p_fshy))) 
# Lump larger groups
totalSingles$nspecies[totalSingles$nspecies > 10] = 10
totalSingles$narea[totalSingles$narea > 2] = 2
totalSingles$ngear[totalSingles$ngear > 3] = 3
totalSingles$nfishery[totalSingles$nfishery > 4] = 4

# Add areas
totalSingles$area = NA
for(i in 1:length(colnames(tableVesselArea))) {
  totalSingles$area[totalSingles$adfg %in% rownames(tableVesselArea)[which(tableVesselArea[, i] == 1)]] = colnames(tableVesselArea)[i]
}

# Fill in missing years and difference
dummyDF = expand.grid(adfg = unique(totalSingles$adfg), year =  min(totalSingles$year):max(totalSingles$year), stringsAsFactors = F)
# expand the data frame w/missing years
totalSingles = left_join(dummyDF, totalSingles) %>%
  arrange(adfg, year)
totalSingles = group_by(totalSingles, adfg) %>%
  mutate(difflog = c(NA, diff(log(totalearnings))), 
         diffspec = c(NA, diff(nspecies)), 
         diffgear = c(NA, diff(ngear)), 
         diffarea = c(NA, diff(narea)), 
         difffishery = c(NA, diff(nfishery))) %>%
  filter(!is.na(difflog))
```

Fit another gam, with shared coefficients across areas, but different splines over time by area (Note 'narea' and 'diffarea' aren't included because this model only includes vessels fishing in one area). These don't include quartiles like previous plots, but show that annual patterns are different (and more variable, as function of sample size) across regions. We'll use something like hierarchical linear models to do a better job of modeling these parameters.

```{r simpleAreaGam, fig.pos="placeHere",fig.cap="Simple GAM model, grouped by area, with area specific effects of species and change in species fished. Gears, change in gears, areas, and change in areas fished also included. Plots show mean smooth year term, with SE."}
gam.area = gam(difflog~area + s(year,by=as.factor(area)) + area*nspecies + area*diffspec + ngear + diffgear,data=totalSingles)
summary(gam.area)
par(mfrow = c(3,2),mgp=c(2,1,0),mai=c(0.5,0.5,0.3,0.1))
plot(gam.area)
```

## To do list
1. Fit GAMs to diversity metrics to characterize change over time. This could be done with the full dataset, 1975-present. 
2. We're going to repeat this kind of analysis on different areas. Some things to consider are:  
- for estimation of random effects, we may need to only include vessels with 3+ records  
- we may want to think about avoiding double counting. The way we have the sorting set up now, a vessel fishing in SEAK and PWS would be counted in both areas / analyses
3. For comparison between areas, we're interested in characterizing differences in year to year patterns, whether or not diversity relationships/coefficients are shared or random across areas (shared vs random effects), and possibly whether impacts of diversity are changing over time (think DLMs).

## Data issues:
1. Why are there no data for longline (gearn == 6) in 2009-2010?

### 2015-11-02 process_cfec_data.R

@@ -1,0 +1,87 @@
+########################################################
+# Process raw data files, first merging them together
+# with common field names. Directory structure is 3 folders,
+# cfecft75_89
+# cfecft90_99
+# cfecft00_09
+########################################################
+
+# Data are separately loaded by year. Create an .Rdata file that merges them all together
+cfec = read.csv("cfecft75_89/cfecft1975.csv")
+
+indx2drop = which(names(cfec)%in%c("corradfg","origadfg","f_fshery"))# drop names not present after 1990
+cfec = cfec[,-c(indx2drop)]
+
+# Merge the other years
+for(yr in 76:89) {
+  newdata = read.csv(paste("cfecft75_89/cfecft19",yr,".csv",sep=""))
+  indx2drop = which(names(newdata)%in%c("corradfg","origadfg","f_fshery"))# drop names not present after 1990
+  newdata = newdata[,-c(indx2drop)]
+  
+  cfec = rbind(cfec,newdata)
+}
+for(yr in 90:90) {
+  newdata = read.csv(paste("cfecft90_99/cfecft19",yr,".csv",sep=""))
+  indx2drop = which(names(newdata)%in%c("corradfg","origadfg","f_fshery"))# drop names not present after 1990
+  newdata = newdata[,-c(indx2drop)]
+  
+  cfec = rbind(cfec,newdata)
+}
+
+cfec = cfec[,-which(names(cfec)%in%c("f_pounds","f_netlbs","f_value","c_landst"))]
+
+# These years have more field codes, and the names change relative to above
+for(yr in 91:99) {
+  newdata = read.csv(paste("cfecft90_99/cfecft19",yr,".csv",sep=""))
+  names(newdata) = tolower(names(newdata))
+  # only include GOA data - get rid of BSAI and INTL
+  #newdata = newdata[-which(newdata$fmp_area%in%c("BSAI","INTL")),]
+  newsub = newdata[,c("year","startdt","lnddate","landdate","procid","port","stat6","adfg","f_cdq","gearn",
+                      "specn","prod","size","g_pounds","g_earn","g_price","p_fshy","p_serial",
+                      "p_check","spec","g_spcs","harvest")]
+  cfec = rbind(cfec,newsub)
+}
+
+# These years have more field codes, and the names change relative to above
+for(yr in 0:9) {
[…]

### 2016-03-18 analyze_cfec_data.Rmd
update

@@
 author: "NCEAS commercial catch portfolio working group"
-date: "October 30, 2015"
+date: "February 22, 2016"
 output:
@@
     fig_caption: yes
+  html_document: default
 ---
@@
 ```{r chunkSet, cache=FALSE, echo=FALSE,warning=FALSE,message=FALSE}
+library(MASS)
 library(knitr)
@@
 library(mgcv)
+library(lme4)
+library(scales)
+library(fields)
+library(reshape)
+library(reshape2)
+library(cluster)
 opts_chunk$set(tidy=TRUE, message=FALSE)
@@
 ```{r loadData}
-load("cfec_jitter.Rdata")
+load("cfec.Rdata")
 # initially only use data 1985+
@@
 The first data processing step is to aggregate some of the species.
-This is largely needed because of changes in reporting over time.
-For example, some of the rockfishes weren't reported to the same species level.
+This is largely needed because of changes in reporting over time (before 1985 for example, lots of the catch was from foreign fleets & not reported to species level -- for example, some of the rockfishes).
 For other species, there are unidentified groups, so there's additional aggregation required.
@@
 cfec$spec[cfec$specn==119] = "FLAT.shallow"
-
+# 5, 6, 7, 15, 26, 47, 61, 
 # Lump flatfish (deep)
@@
 cfec$spec[cfec$spec%in%c("SPRW","PSHP","USRM")] = "USRM" # shrimp
-
+# Correct REX and "REX "
+cfec$spec[cfec$spec%in%c("REX ")] = "REX"
 # drop records with no spp
 cfec = cfec[-which(cfec$spec==""),]
+```
+
+Next, we have to clean up a few of the gears. Drop a few gears that are sporadically used, and combine longline gears.
+
+```{r}
+# 8 = fishwheel, used at start / end
+# 13 = dipnet, only last 2 years
+# 99 = other, stopped reporting
+cfec = cfec[-which(cfec$gearn%in%c(8, 13, 99)),]
+# Merge 6 and 61 into longline
+cfec$gearn[which(cfec$gearn==61)] = 6
 ```
@@
 # Lump statistical areas by district -- groundfish
-cfec$stat6 = as.character(as.numeric(cfec$stat6)) # as.numeric needed because of dumb integer
+cfec$stat6 = as.numeric(cfec$stat6) # as.numeric needed because of integer
 cfec$areaType = NA
@@
 indx = which(cfec$stat6 > 99999 & cfec$stat6 < 999999)
-cfec$stat6_char[indx] = nchar(cfec$stat6[indx])
+cfec$stat6_char[indx] = nchar(as.character(cfec$stat6[indx]))
 cfec$areaType[indx] = "groundfish"
 # This is just turning the last digit - sub region - into 0
-cfec$stat6[indx] = paste(substr(cfec$stat6[indx],1,5), "0",sep="")
-cfec$stat6[indx] = as.numeric(cfec$stat6[indx])
+cfec$stat6[indx] = as.numeric(paste(substr(as.character(cfec$stat6[indx]),1,5), "0",sep=""))
 # Now do salmon / shellfish - This is just turning the last 2 digits - sub region - into 00
 indx = which(cfec$stat6 > 9999 & cfec$stat6 < 100000)
-cfec$stat6[indx] = paste(substr(cfec$stat6[indx],1,3), "00",sep="")
+cfec$stat6[indx] = as.numeric(paste(substr(as.character(cfec$stat6[indx]), 1, 3), "00", sep=""))
 cfec$areaType[indx] = "salmonShellfish"
 
-cfec$stat6 = as.numeric(cfec$stat6)
 # Deal with the IPHC areas too. We can verify that nearly all spp caught in 3-digit areas are halibut,
@@
 
-Next, we'll aggregate the stat6 codes into larger geographic regions. The focus here is on Prince William Sound, but for contrast we'll include Southeast Alaska, Kodiak W / E, and Cook Inlet.
+Next, we'll aggregate the stat6 codes into larger geographic regions.
+The focus here is on Prince William Sound, but for contrast we'll include Southeast Alaska, Kodiak W / E, Cook Inlet, and Bristol Bay.
 
@@
 
+Another alternative approach to grouping would be to perform a cluster analysis on vessels / areas.
+Here we'll make a table of vessels (rows) and areas (column) then use 'clara' to construct 8 groups.
+This is somewhat arbitrary, but reflects roughly the number constructed above.
+
+```{r}
+T = table(cfec$adfg, cfec$stat6)
+CC = clara(T, k = 8)
+cluster_vesselArea = t(CC$medoids)
+```
+
+Next we can sort out the mutually exclusive areas in the medoids to identify how the areas are being clustered.  
+```{r}
+cluster_vesselArea = ceiling(cluster_vesselArea / 1.0e10)
+cluster_vesselArea = cluster_vesselArea[-which(apply(cluster_vesselArea,1,sum)!=1),]
+```
+
+For example, the 6th medoid is clearly pulling out vessels in Southeast, 
+```{r}
+names(which(cluster_vesselArea[,6]==1))
+```
+
+Similarly, medoid 1 is PWS, medoid 2 is Cook Inlet, medoid 4 is Bristol Bay, medoid 5 is ~ 700 records of halibut (from a southern area not in GoA), medoid 6 is Southeast, medoid 7 is Kodiak, medoids 3 and 8 are Bering.
+So this seems to agree with our clustering, maybe with the exception of not splitting Kodiak into W/E.
+
+We can look at how the species landed varies by area (table in 1000s), 
+```{r}
+kable(round(table(cfec$spec, cfec$area)/1000, 2))
+```
+
+We can filter out rare species, and turn this table into catch composition (by occurence, not weight or earnings) to highlight how the areas are different,
+
+```{r}
+specArea = table(cfec$spec, cfec$area)
+# normalize by area
+for(i in 1:dim(specArea)[2]) {
+specArea[,i] = specArea[,i]/sum(specArea[,i])
+}
+kable(round(specArea[-which(apply(specArea,1,max) < 0.01),], 2))
+```
+
 ## Exploratory data analysis
@@
   data.frame
-
 # Also drop out unid groundfish, unid flatfish, smelts, grenadier, english sole -- all very rare
 cfecClean = cfecClean[-which(cfecClean$spec%in%c("UFLT","UGRN","SMLT","GRDR","EGLS")),]
-
 ```
@@
 Delete catch from fisheries that have less than 100 total records.
-This simplifies the dimension of the fishery codes a lot (from 378 to 234), removes about 3050 records, and doesn't affect the total earnings (these fisheries represent 0.0009 of total g_earn).
+This simplifies the dimension of the fishery codes a lot (from 378 to 234), removes about 3050 records, and doesn't affect the total revenue (these fisheries represent 0.0009 of total g_earn).
 ```{r discardFishery}
@@
   data.frame
+
+# Write the SSL related sites to a workspace
+sslStat6 = c(355630,
+365730, 375830, 345530, 335430, 335430, 335430,
+335430, 335430, 365800, 355700, 345530, 355630,
+355700, 335700, 335700, 355700, 445930, 495930, 
+476000, 485930, 495930, 325600, 495930, 476000,
+475930, 555530, 565600, 525700, 525700, 525830,
+525700, 515800, 515830, 535630, 365800, 575630,
+535630, 565630, 525830, 595500, 595430, 595500,
+595430, 635430, 635430, 605430, 615430, 635504,
+585500)
+
+# calculate areas in circle around square of interest
+sslStat6.square = c(sslStat6, sslStat6 +- 9970, sslStat6 +- 10000, 
+sslStat6 +- 10070, sslStat6 +- 70, sslStat6 +- 9930, sslStat6 +- 10000, 
+sslStat6 +- 10030, sslStat6 +- 30)
+sslStat6.label = rep(sslStat6,9)
+
+cfecSSL = cfecClean[cfecClean$stat6%in%sslStat6.square,]
+cfecSSL$label = sslStat6.label[match(cfecSSL$stat6, sslStat6.square)]
+save(cfecSSL, file="cfecSSL.Rdata")
+
 ```
 
-Subset the data to only include vessels that fished at any point in Prince William Sound.  
+Tabulate the gears and species remaining.  
+```{r tableOfGears}
+m = as.matrix(sort(table(cfecClean$gearn)), ncol=1)
+# Normalize percent of all records
+m = cbind(m, round(100*m[,1]/sum(m[,1]),2))
+colnames(m) = c("Records","Percent")
+kable(m)
+```
+
+This shows that 79% of the records in the database are salmon!   
+```{r tableOfSpec}
+m = as.matrix(sort(table(cfecClean$spec)), ncol=1)
+# Normalize percent of all records
+m = cbind(m, round(100*m[,1]/sum(m[,1]),2))
+colnames(m) = c("Records","Percent")
+kable(m)
+```
+
+Subset the data to only include vessels that ever fished in Prince William Sound. This data frame becomes 'cfecPWS'.
+
 ```{r useOnlyPWS}
@@
 ```{r,fig.cap="Time series of participation by year",fig.pos="placeHere"}
-plot(min(cfecPWS$year):max(cfecPWS$year), specnumber(table(cfecPWS$year,cfecPWS$adfg)), xlab="",ylab="# Vessels", type="b")
+plot(min(cfecPWS$year):max(cfecPWS$year), specnumber(table(cfecPWS$year,cfecPWS$adfg)), xlab="",ylab="# Vessels", type="b", main = "Participation in PWS region")
 ```
 
-We can also look at species by year
+We can also look at species landed by year.
+This is not terribly informative -- for example the total number of species landed in the dataset has ranged from 38-41.
+An alternative would be to take the average (below).
 
@@
 # total species
-par(mfrow = c(2,1), mgp = c(2,1,0), mai=c(0.5,0.5,0.3,0.1))
-plot(min(cfecPWS$year):max(cfecPWS$year), specnumber(table(cfecPWS$year, cfecPWS$spec)), xlab="",ylab="# Species", type="b",main="Catch from vessels that ever fished in PWS")
+par(mfrow = c(3,1), mgp = c(2,1,0), mai=c(0.5,0.5,0.3,0.1))
+plot(min(cfecPWS$year) : max(cfecPWS$year) , 
+     specnumber( table (cfecPWS$year, cfecPWS$spec)), xlab="", 
+     ylab="# Species", type="b", 
+     main="Catch from vessels that ever fished in PWS")
 
-plot(min(cfecPWS$year):max(cfecPWS$year), specnumber(table(cfecPWS$year[cfecPWS$area=="PWS"], cfecPWS$spec[cfecPWS$area=="PWS"])), xlab="",ylab="# Species", type="b",main="Catch only in PWS")
+plot(min(cfecPWS$year) : max(cfecPWS$year), 
+     specnumber(table(cfecPWS$year[cfecPWS$area=="PWS"], 
+                      cfecPWS$spec[cfecPWS$area=="PWS"])), 
+     xlab="",ylab="# Species", type="b",main="Catch only in PWS")
+
+plot(min(cfecClean$year) : max(cfecClean$year), 
+     specnumber(table(cfecClean$year,  cfecClean$spec)), 
+     xlab="",ylab="# Species", type="b", 
+     main="All vessels in dataset")
+```
+
+Now we can calculate the average species landed / year, where the average is calculated as the total number of species landed divided by the total number of vessels fishing.
+
+```{r,fig.cap="Time series of avg species caught by year per vessel",fig.pos="placeHere"}
+# total species
+par(mfrow = c(3,1), mgp = c(2,1,0), mai=c(0.5,0.5,0.3,0.1))
+plot(min(cfecPWS$year):max(cfecPWS$year), 
+     specnumber(table(cfecPWS$year, cfecPWS$spec)) / 
+       specnumber(table(cfecPWS$year,  cfecPWS$adfg)), 
+     xlab="", ylab="Avg # Species / Ves", 
+     type="b", main="Catch from vessels that ever fished in PWS")
+
+plot(min(cfecPWS$year):max(cfecPWS$year), 
+     specnumber(table(cfecPWS$year[cfecPWS$area=="PWS"], 
+     cfecPWS$spec[cfecPWS$area=="PWS"])) / 
+       specnumber(table(cfecPWS$year[cfecPWS$area=="PWS"], 
+     cfecPWS$adfg[cfecPWS$area=="PWS"])), 
+     xlab="",ylab="Avg # Species / Ves", 
+     type="b",main="Catch only in PWS")
+
+plot(min(cfecClean$year):max(cfecClean$year), 
+     specnumber(table(cfecClean$year,  cfecClean$spec)) / 
+       specnumber(table(cfecClean$year,  cfecClean$adfg)), 
+     xlab="",ylab="Avg # Species / Ves",
+     type="b", main="All vessels in dataset")
+```
+
+Another way to calculate the average is total # species landed divided by the total number of fishing records per year.
+
+```{r,fig.cap="Time series of avg species caught by year per vessel",fig.pos="placeHere"}
+# total species
+par(mfrow = c(3,1), mgp = c(2,1,0), mai=c(0.5,0.5,0.3,0.1))
+plot(min(cfecPWS$year):max(cfecPWS$year), 
+     specnumber(table(cfecPWS$year, cfecPWS$spec)) / 
+       as.numeric(table(cfecPWS$year)), xlab="", 
+     ylab = "Avg # Species / Haul", type = "b",  
+     main = "Catch from vessels that ever fished in PWS")
+
+plot(min(cfecPWS$year):max(cfecPWS$year), 
+     specnumber(table(cfecPWS$year[cfecPWS$area=="PWS"], 
+    cfecPWS$spec[cfecPWS$area=="PWS"])) / 
+      as.numeric(table(cfecPWS$year[cfecPWS$area=="PWS"])), 
+    xlab = "", ylab = "Avg # Species / Haul", 
+    type = "b",main = "Catch only in PWS")
+
+plot(min(cfecClean$year):max(cfecClean$year), 
+     specnumber(table(cfecClean$year,  cfecClean$spec)) / 
+       as.numeric(table(cfecClean$year)), xlab = "", 
+     ylab = "Avg # Species / Haul", 
+     type = "b", main = "All vessels in dataset")
 ```
@@
 Look at species diversity using Vegan. How has diversity in # species fished changed over time?   
-```{r veganSpecies,echo=FALSE, fig.cap="Time series of species diversity",fig.pos="placeHere"}
+```{r veganSpecies,echo=FALSE, fig.cap="Time series of species diversity (for vessels that have ever fished in Prince William Sound)",fig.pos="placeHere"}
 nSpecies.mean = NA # mean richness
@@
 We can do similar plots for the number of gears used / vessel  
-```{r veganGears,echo=FALSE, fig.cap="Time series of diversity of number of fishing gears",fig.pos="placeHere"}
+```{r veganGears,echo=FALSE, fig.cap="Time series of diversity of number of fishing gears (for vessels that have ever fished in Prince William Sound)",fig.pos="placeHere"}
 nGears.mean = NA # mean richness
@@
 ## Metric of change
-We can use absolute or relative earnings over time, by vessel.
+We can use absolute or relative revenue over time, by vessel.
 We adopted the relative measure, because we're comparing vessels of very different size classes.
 Sizes of individual vessels may change slowly over time.
-The earning metric is calculated as the first-differenced log earnings from time (t+1) to time (t), $log(earn_{t+1}) - log(earn_{t})$.
+The revenue metric is calculated as the first-differenced log revenue from time (t+1) to time (t), $log(earn_{t+1}) - log(earn_{t})$.
 
-The key challenge here is that we need a metric of earnings that will allow us to compare different types of vessels (e.g.
+The key challenge here is that we need a metric of revenue that will allow us to compare different types of vessels (e.g.
 large, small) that don't overlap in time.
-For example, we need a metric that standardizes catches or earnings from a small vessel fishing 1985-1990 to that of a large vessel fishing 2000-2005.
+For example, we need a metric that standardizes catches or revenue from a small vessel fishing 1985-1990 to that of a large vessel fishing 2000-2005.
 The relative change is universal across vessels, and can be interpreted as percent change (e.g.
-a value of -1 = exp(-1) = 0.367, or a ~ 63% decline in earnings between years).
+a value of -1 = exp(-1) = 0.367, or a ~ 63% decline in revenue between years).
 
@@
 ```{r mergeYears}
-dummyDF = expand.grid(adfg = unique(totalPWS$adfg), year = min(totalPWS$year):max(totalPWS$year), stringsAsFactors = F)
+dummyDF = expand.grid(adfg = unique(totalPWS$adfg), 
+  year = min(totalPWS$year) : max(totalPWS$year), 
+  stringsAsFactors = F)
 # expand the data frame w/missing years
@@
 par(mfrow = c(2,2), mai=c(0.6,0.6,0.1,0.1))
-#hist(log(totalPWSExp$totalearnings), main="Log of earnings",xlab="",col="grey")
+#hist(log(totalPWSExp$totalearnings), main="Log of revenue",xlab="",col="grey")
 hist(totalPWSExp$nspecies, main="Number of species", xlab="", col="grey")
@@
 
-We can also look at change in these predictors over time -- we can also look at the response, which is change in earnings from time t to time t+1.
+We can also look at change in these predictors over time -- we can also look at the response, which is change in revenue from time t to time t+1.
 ```{r,fig.cap="Histogram of change in predictors (across all years and data)",fig.pos="placeHere"}
 par(mfrow = c(3,2), mai=c(0.6,0.6,0.2,0.1))
-hist(totalPWSExp$difflog, main="Log of change in earnings", xlab="", col="grey")
+hist(totalPWSExp$difflog, main="Log of change in revenue", xlab="", col="grey")
 hist(totalPWSExp$diffspec, main="Change in species", xlab="", col="grey")
@@
 
-## Plotting change in earnings through time
+## Plotting change in revenue through time
 
@@
 
-```{r,fig.cap="Median and quartiles of change in earnings, by species, for individuals that don't change number of species fished through time",fig.pos="placeHere"}
-group_by(totalPWSExp[totalPWSExp$diffspec==0,], year, nspecies) %>% 
+```{r,fig.cap="Median and quartiles of change in revenue, by species, for individuals that don't change number of species fished through time",fig.pos="placeHere"}
+group_by(totalPWSExp[totalPWSExp$diffspec==0, ], year, nspecies) %>%  
   summarize(medDiffLog = median(difflog),  
@@
 
-```{r,fig.cap="Median and quartiles of change in earnings, by gears, for individuals that don't change number of gears fished through time",fig.pos="placeHere"}
-group_by(totalPWSExp[totalPWSExp$diffgear==0,], year, ngear) %>% 
+```{r,fig.cap="Median and quartiles of change in revenue, by gears, for individuals that don't change number of gears fished through time",fig.pos="placeHere"}
+group_by(totalPWSExp[totalPWSExp$diffgear==0, ], year, ngear) %>%  
   summarize(medDiffLog = median(difflog),  
@@
 
-```{r,fig.cap="Median and quartiles of change in earnings, by areas, for individuals that don't change number of areas fished through time",fig.pos="placeHere"}
-group_by(totalPWSExp[totalPWSExp$diffarea==0,], year, narea) %>% 
+```{r,fig.cap="Median and quartiles of change in revenue, by areas, for individuals that don't change number of areas fished through time",fig.pos="placeHere"}
+group_by(totalPWSExp[totalPWSExp$diffarea==0, ], year, narea) %>%  
   summarize(medDiffLog = median(difflog),  
@@
 
-```{r,fig.cap="Median and quartiles of change in earnings, by areas, for individuals that don't change number of fisheries fished through time",fig.pos="placeHere"}
-group_by(totalPWSExp[totalPWSExp$difffishery==0,], year, nfishery) %>% 
+```{r,fig.cap="Median and quartiles of change in revenue, by areas, for individuals that don't change number of fisheries fished through time",fig.pos="placeHere"}
+group_by(totalPWSExp[totalPWSExp$difffishery==0, ], year, nfishery) %>%  
   summarize(medDiffLog = median(difflog),  
@@
 This plot includes some confounding, so don't read too much into it.
-```{r,fig.cap="Median and quartiles of change in earnings, by species x areas",fig.pos="placeHere"}
+```{r,fig.cap="Median and quartiles of change in revenue, by species x areas",fig.pos="placeHere"}
 totalPWSExp2 = totalPWSExp 
@@
 
-Make some summary tables of vessels fishing in 1988 / 1989 that only landed one species. Make tables of the most common species landed, both by records and earnings. 
+Make some summary tables of vessels fishing in 1988 / 1989 that only landed one species. Make tables of the most common species landed, both by records and revenue. 
 ```{r}
@@
 
-# Look at earnings of fishers that people are specializing on
+# Look at revenue of fishers that people are specializing on
 LJ89 = LJ[LJ$year==1989,]
@@
 
-Are a priori expectation is that if diversity (species, gear, area) is beneficial, all of the differenced predictors (diffspec, diffgear, diffarea) will have positive effects on the response (change in earnings).
+Are a priori expectation is that if diversity (species, gear, area) is beneficial, all of the differenced predictors (diffspec, diffgear, diffarea) will have positive effects on the response (change in revenue).
 But if the differenced values are small, it means individual strategy is relatively constant, and we'd also expect the effects of diverification on the mean values (nspecies, narea, ngear).
@@
 We can also fit a gam, with spline over year, and basically get the same result.
-These models show that changing strategy (increasing gears used or areas fished correlates with more earnings, but the overall strategy of more areas or gears doesn't have a positive effect).
+These models show that changing strategy (increasing gears used or areas fished correlates with more revenue, but the overall strategy of more areas or gears doesn't have a positive effect).
 In contrast, the effect of species is positive in both, suggesting that fishing more species translates to positive revenue.
@@
 m = matrix(NA, 3, 2)
-colnames(m) = c("# Vessels", "% Total")
+colnames(m) = c("# Vessels", "Percent")
 rownames(m) = c("1","2","3+")
 m[,1] = c(T[2],T[3],sum(T[-c(1:3)]))
-m[,2] = round(m[,1]/sum(m[,1]),3)
+m[,2] = round(100*m[,1]/sum(m[,1]),3)
 
@@
 # Fill in missing years and difference
-dummyDF = expand.grid(adfg = unique(totalSingles$adfg), year =  min(totalSingles$year):max(totalSingles$year), stringsAsFactors = F)
+dummyDF = expand.grid(adfg = unique(totalSingles$adfg), 
+ year =  min(totalSingles$year) : max(totalSingles$year), 
+ stringsAsFactors = F)
 # expand the data frame w/missing years
@@
 ## To do list
-1. Fit GAMs to diversity metrics to characterize change over time. This could be done with the full dataset, 1975-present. 
+1. Fit GAMs to diversity metrics to characterize change over time. This could be done with the full dataset, 1985-present. 
 2. We're going to repeat this kind of analysis on different areas. Some things to consider are:  
@@
 
-## Data issues:
-1. Why are there no data for longline (gearn == 6) in 2009-2010?
+******
+# More rigorous analysis
+
+Above, we focued on breaking the paper into 2 parts: (1) analyzing patterns over time in diversity by fitting GAMs, and then (2) doing statistical modeling of change in revenue as a function of diversification metrics.
+
+## Change in diversity metrics through time
+
+We're going to do some harder filtering, first only retaining vessels / permits that ever fished in one of the designated areas. In other words, they can't have an NA in the record - ever.
+
+```{r}
+cfecCleaned = cfecClean
+cfecCleaned$isNA = 0 # is an area ever NA?
+cfecCleaned$isNA[is.na(cfecCleaned$area)] = 1
+vessel.isna = table(cfecCleaned$adfg,cfecCleaned$isNA)
+cfecCleaned = cfecCleaned[cfecCleaned$adfg %in% names(which(vessel.isna[, 2] ==
+0)), ]
+```
+
+```{r}
+# Of this restricted dataset, 93% of vessels don't move between areas -
+# ever. So the primary focus of our analysis can be on gears and species
+# landed. Just so the area switching won't matter, let's delete those
+# records that have multiple areas
+t = ceiling(table(cfecCleaned$adfg, cfecCleaned$area)/1e+10)
+# delete vessels that fish in 2+ areas
+cfecCleaned = cfecCleaned[cfecCleaned$adfg %in% names(which(apply(t, 1, sum) ==
+1)), ]
+
+# Now tabulate summaries of species, gears, fisheries, etc
+totalVessels = group_by(cfecCleaned, adfg, year) %>%  
+  summarize(totalearnings = sum(g_earn),  
+  nspecies = length(unique(spec)),  
+  ngear = length(unique(gearn)),  
+  nfishery = length(unique(p_fshy)),
+  area = area[1])
+
+# calculate landings by species, year, vessel
+landBySpec = group_by(cfecCleaned, adfg, year, spec) %>%  
+summarize(totalpounds = sum(g_pounds))  
+# calculate effective number of species by exp() for shannon
+# http://jonlefcheck.net/2012/10/23/diversity-as-effective-numbers/
+landBySpec = group_by(landBySpec, adfg, year) %>%  
+   summarize(effectiveSpec = exp(diversity(totalpounds)))
+
+totalVessels$effectiveSpec = landBySpec$effectiveSpec
+
+# calculate landings by gear, year, vessel
+landByGear = group_by(cfecCleaned, adfg, year, gearn) %>%  
+  summarize(totalpounds = sum(g_pounds))  
+landByGear = group_by(landByGear, adfg, year) %>%  
+  summarize(effectiveGear = exp(diversity(totalpounds)))
+
+totalVessels$effectiveGear = landByGear$effectiveGear
+
+# calculate landings by fishery, year, vessel
+landByFshy = group_by(cfecCleaned, adfg, year, p_fshy) %>%  
+summarize(totalpounds = sum(g_pounds))  
+landByFshy = group_by(landByFshy, adfg, year) %>%  
+summarize(effectiveFshy = exp(diversity(totalpounds)))
+
+totalVessels$effectiveFshy = landByFshy$effectiveFshy
+
+# Lump larger groups
+totalVessels$nspecies[totalVessels$nspecies > 10] = 10
+totalVessels$ngear[totalVessels$ngear > 3] = 3
+totalVessels$nfishery[totalVessels$nfishery > 4] = 4
+```
+
+### GAM fitting
+
+Without area included, our 3 dimensions are (1) species, (2) gears, and (3) fisheries.
+For both metrics, we can fit GAMS examining either the raw counts, or use the effective number, which we calculated above.
+For each diversity metric, we can also do model selection to ask whether the spline over time is constant or shared across areas.
+
+#### Change in species over time
+
+We can fit 4 models to the species data.
+We're not comparing Raw / Effective counts, but constant versus area-specific splines.
+We can see that the area splines are strongly favored, suggesting regional patterns in diversity with respect to species.
+```{r}
+rawspec.ConstantSpline = gam(log(nspecies) ~ s(year), data=totalVessels)
+rawspec.AreaSpline = gam(log(nspecies) ~ s(year, by = as.factor(area)), data=totalVessels)
+effspec.ConstantSpline = gam(log(effectiveSpec) ~ s(year), data=totalVessels)
+effspec.AreaSpline = gam(log(effectiveSpec) ~ s(year, by = as.factor(area)), data=totalVessels)
+spec.mat = matrix(c("Raw","Raw","Effective","Effective", "Constant", "Area", "Constant", "Area", 
+             round(AIC(rawspec.ConstantSpline),2), round(AIC(rawspec.AreaSpline),2), round(AIC(effspec.ConstantSpline),2), round(AIC(effspec.AreaSpline),2)), 4, 3)
+colnames(spec.mat) = c("Counts", "Spline", "AIC")
+kable(spec.mat)
+```
+
+#### Change in gears over time
+
+We can fit the same 4 models to the gear data.
+We're not comparing Raw / Effective counts, but constant versus area-specific splines.
+Like with species, the area splines are strongly favored, suggesting regional patterns in diversity with respect to gears fished.
+```{r}
+rawgear.ConstantSpline = gam(log(ngear) ~ s(year), data=totalVessels)
+rawgear.AreaSpline = gam(log(ngear) ~ s(year, by = as.factor(area)), data=totalVessels)
+effgear.ConstantSpline = gam(log(effectiveGear) ~ s(year), data=totalVessels)
+effgear.AreaSpline = gam(log(effectiveGear) ~ s(year, by = as.factor(area)), data=totalVessels)
+gear.mat = matrix(c("Raw","Raw","Effective","Effective", "Constant", "Area", "Constant", "Area", 
+             round(AIC(rawgear.ConstantSpline),2), round(AIC(rawgear.AreaSpline),2), round(AIC(effgear.ConstantSpline),2), round(AIC(effgear.AreaSpline),2)), 4, 3)
+colnames(gear.mat) = c("Counts", "Spline", "AIC")
+kable(gear.mat)
+```
+
+#### Change in fisheries over time
+
+We can fit the same 4 models to the fishery data.
+We're not comparing Raw / Effective counts, but constant versus area-specific splines.
+Like with species and gears, the area splines are strongly favored, suggesting regional patterns in diversity with respect to fisheries fished.
+```{r}
+rawfshy.ConstantSpline = gam(log(nfishery) ~ s(year), data=totalVessels)
+rawfshy.AreaSpline = gam(log(nfishery) ~ s(year, by = as.factor(area)), data=totalVessels)
+efffshy.ConstantSpline = gam(log(effectiveFshy) ~ s(year), data=totalVessels)
+efffshy.AreaSpline = gam(log(effectiveFshy) ~ s(year, by = as.factor(area)), data=totalVessels)
+fshy.mat = matrix(c("Raw","Raw","Effective","Effective", "Constant", "Area", "Constant", "Area", round(AIC(rawfshy.ConstantSpline),2), round(AIC(rawfshy.AreaSpline),2), round(AIC(efffshy.ConstantSpline),2), round(AIC(efffshy.AreaSpline),2)), 4, 3)
+colnames(fshy.mat) = c("Counts", "Spline", "AIC")
+kable(fshy.mat)
+```
+
+So in all cases, the diversity changes over time varies spatially. Finally, we'll plot the model output for the best models.
+
+```{r, fig.pos="placeHere",fig.cap="GAM predictions of responses (species, gears, fisheries) by metric (raw, effective). Colors represent regions."}
+# make a 3 x 2 plot with diversity measure on rows, and raw/effect on cols
+areas = levels(as.factor(totalVessels$area))
+ys = 1985:2014
+par(mfrow = c(3,2), mgp=c(2,1,0), mai=c(0.5,0.5,0.05,0.05))
+# BristolBay CookInlet Kodiak.east Kodiak.west PWS Southeast
+
+plot(ys, predict(rawspec.AreaSpline, 
+newdata=expand.grid(area="BristolBay",year=ys)), type="l", 
+col = tim.colors(6)[1], ylim=c(0, max(rawspec.AreaSpline$fitted.values)), 
+xlab="", ylab = "Predicted log(raw spec)")
+for(i in 2:length(areas)) {
+lines(ys, predict(rawspec.AreaSpline, newdata=expand.grid(area=areas[i],year=ys)), col = tim.colors(6)[i])
+}
+legend("topleft", areas, col = tim.colors(6), lty=1, bty="n")
+
+plot(ys, predict(effspec.AreaSpline, 
+newdata=expand.grid(area="BristolBay",year=ys)), type="l", 
+col = tim.colors(6)[1], ylim=c(0, max(effspec.AreaSpline$fitted.values)), 
+xlab="", ylab = "Predicted log(eff. spec)")
+for(i in 2:length(areas)) {
+lines(ys, predict(effspec.AreaSpline, newdata=expand.grid(area=areas[i],year=ys)), col = tim.colors(6)[i])
+}
+
+plot(ys, predict(rawgear.AreaSpline, 
+newdata=expand.grid(area="BristolBay",year=ys)), type="l", 
+col = tim.colors(6)[1], ylim=c(0, max(rawgear.AreaSpline$fitted.values)), 
+xlab="", ylab = "Predicted log(raw gears)")
+for(i in 2:length(areas)) {
+lines(ys, predict(rawgear.AreaSpline, newdata=expand.grid(area=areas[i],year=ys)), col = tim.colors(6)[i])
+}
+
+plot(ys, predict(effgear.AreaSpline, 
+newdata=expand.grid(area="BristolBay",year=ys)), type="l", 
+col = tim.colors(6)[1], ylim=c(0, max(effgear.AreaSpline$fitted.values)), 
+xlab="", ylab = "Predicted log(eff gears)")
+for(i in 2:length(areas)) {
+lines(ys, predict(effgear.AreaSpline, newdata=expand.grid(area=areas[i],year=ys)), col = tim.colors(6)[i])
+}
+
+plot(ys, predict(rawfshy.AreaSpline, 
+newdata=expand.grid(area="BristolBay",year=ys)), type="l", 
+col = tim.colors(6)[1], ylim=c(0, max(rawfshy.AreaSpline$fitted.values)), 
+xlab="", ylab = "Predicted log(raw fshy)")
+for(i in 2:length(areas)) {
+lines(ys, predict(rawfshy.AreaSpline, newdata=expand.grid(area=areas[i],year=ys)), col = tim.colors(6)[i])
+}
+
+plot(ys, predict(efffshy.AreaSpline, 
+newdata=expand.grid(area="BristolBay",year=ys)), type="l", 
+col = tim.colors(6)[1], ylim=c(0, max(efffshy.AreaSpline$fitted.values)), 
+xlab="", ylab = "Predicted log(eff fshy)")
+for(i in 2:length(areas)) {
+lines(ys, predict(efffshy.AreaSpline, newdata=expand.grid(area=areas[i],year=ys)), col = tim.colors(6)[i])
+}
+```
+
+## Clustering vessel-years by diversity in species
+
+Our response (revenue) and predictors are on an annual scale, so one analysis that might be interesting to look at would be to try to reduce the dimension of the dataset.
+In other words, is it possible to classify vessel-year landings into salmon v non-salmon fishers?
+Or are other patterns of specialists evident?
+
+One way to do this is to create a matrix with rows = vessel-year combinations, and columns = species landed.
+
+```{r}
+# summarize landings by year/spec/
+totalLandings = group_by(cfecCleaned, adfg, year, spec) %>%  
+  summarize(totallandings = sum(g_pounds))
+# cast to wide format with species across cols
+totalLandings = as.data.frame(totalLandings)
+#totalLandings = totalLandings[-which(is.na(totalLandings$adfg)),]
+lmlt <- melt(totalLandings, id.vars = c("adfg", "year","spec"))
+newdf <- dcast(lmlt, year + adfg ~ spec)
+```
+
+Now we can apply MDS to look at clusters. Turn all NAs to 0s.
+
+```{r}
+for(i in 3:dim(newdf)[2]) {
+  newdf[which(is.na(newdf[,i])),i] = 0
+  #newdf[,i] = as.numeric(newdf[,i])
+}
+# Let's sample ~ 10000 data points for starters
+s = sample(seq(1, dim(newdf)[1]), size=5000, replace=F)
+#distMat = dist(log(newdf[s,-c(1:2)]+1))
+
+#avgWidth = 100
+#for(i in 2:10) {
+#  clust = pam(log(newdf[s,-c(1:2)]+1), k = i)
+#  avgWidth[i] = clust$silinfo$avg.width
+#}
+#bestClust = which.min(avgWidth)
+
+# Using mean(silhouette(clust)) we find that 3 clusters
+# has the lowest mean silhouette width
+bestClust = 3
+clust = pam(log(newdf[s,-c(1:2)]+1), k = bestClust)
+
+# We can look at what the medoids represent, 
+m = t(clust$medoids)
+colnames(m) = paste("Group",seq(1,dim(m)[2]))
+kable(m)
+```
+
+What follows out here is that the data are dividing annual landings into 3 larger groups: Sockeye fishing, Non-sockeye salmon fishing, and halibut fishing (PHLB).
+
+Just for general interest, we could basically repeat the same analysis after dropping out the salmon records.
+This shows that herring and halibut landings are the largest groups (and everything else as 3rd group).
+
+```{r}
+#avgWidth = 100
+dropnames = which(colnames(newdf)%in%c("year", "adfg", "CHNK", "CHUM", "COHO", "PINK", "SOCK"))
+#for(i in 2:10) {
+#  clust = pam(log(newdf[s,-dropnames]+1), k = i)
+#  avgWidth[i] = clust$silinfo$avg.width
+#}
+#bestClust = which.min(avgWidth)
+# Using mean(silhouette(clust)) we find that 3 clusters
+# has the lowest mean silhouette width
+bestClust=3
+clust = pam(log(newdf[s,-dropnames]+1), k = bestClust)
+
+# We can look at what the medoids represent, 
+m = t(clust$medoids)
+colnames(m) = paste("Group",seq(1,dim(m)[2]))
+kable(m)
+```
+
+One potential criticism of this approach is that the separation / group medoids are largely affected by landed weight.
+An alternative approach would be to do the same clustering using binary classification with 'mona'.
+
+```{r}
+newdf01 = ceiling(newdf[,-c(1:2)]/1.0e10)
+avgWidth = 100
+dropnames = which(colnames(newdf)%in%c("year", "adfg", "CHNK", "CHUM", "COHO", "PINK", "SOCK"))
+for(i in 2:10) {
+  clust = pam(log(newdf01[s,-dropnames]+1), k = i)
+  avgWidth[i] = clust$silinfo$avg.width
+}
+
+bestClust = which.min(avgWidth)
+clust = pam(log(newdf01[s,-dropnames]+1), k = bestClust)
+
+# We can look at what the medoids represent, 
+m = t(clust$medoids)
+colnames(m) = paste("Group",seq(1,dim(m)[2]))
+kable(m)
+```
+
+So basically the groups are the same before, with 2 kinds of salmon fishermen and halibut defining the 3rd cluster.
+
+## Linear modeling of changing diversity on revenue
+
+Add in NAs for missing years. For the undifferenced data, the number of species, gears, etc. corresponds to the value in the second year. We may want to replace this with a 2-year mean or something.
+
+```{r}
+dummyDF = expand.grid(adfg = unique(totalVessels$adfg), 
+  year = min(totalVessels$year) : max(totalVessels$year),
+  stringsAsFactors = F)
+
+# expand the data frame w/missing years
+totalVesselsExp = left_join(dummyDF, totalVessels) %>%
+  arrange(adfg, year)
+totalVesselsExp = group_by(totalVesselsExp, adfg) %>%
+  mutate(difflog = c(NA, diff(log(totalearnings))),  
+         diffspec = c(NA, diff(nspecies)),  
+         diffgear = c(NA, diff(ngear)),  
+         difffishery = c(NA, diff(nfishery)),  
+         diffeffspec = c(NA, diff(effectiveSpec)),  
+         diffeffgear = c(NA, diff(effectiveGear)),  
+         diffefffishery = c(NA, diff(effectiveFshy))) %>%  
+  filter(!is.na(difflog))
+```
+
+We may also want to calculate the weighted average of each of the predictors, with weights being pounds landed.
+For example, if someone catches 2000 kg of sockeye and 5 kg of chum, the weights may better reflect divserification.
+
+## Inclusion of vessel random effects
+
+We’re modeling revenue (or change in revenue) across a large number of vessels (5000+).
+One way to make this model hierarchical would be to include random effects in vessel.
+The intercept here represents Bristol Bay in 1986 – all factors (year / area) are offsets.
+
+```{r}
+lme.vess.1trend = lmer(difflog ~ area + as.factor(year) + 
+  nspecies + diffspec + ngear + diffgear + (1 | adfg), 
+  data = totalVesselsExp)
+summary(lme.vess.1trend)
+```
+
+We can plot the overall annual trend for Bristol Bay, the patterns for other regions are all just offset from the Bristol Bay trend.
+
+```{r, fig.pos="placeHere",fig.cap="Mixed effects model output of fixed year effects for the Bristol Bay region."}
+ycoef = c(0, fixef(lme.vess.1trend)[7:34])
+ycoef = ycoef + fixef(lme.vess.1trend)[1]
+plot(1986:2014, ycoef, type = "b", xlab = "", 
+     ylab = "Year effect across vessels in Bristol Bay")
+```
+
+The key point here is that inclusion of random effects, or more complicated interactions (below) doesn’t change the magnitude or sign of the fixed coefficients we care about – namely, the number of species fished (and change) or number of gears fished (and change).
+These more complicated models are largely a way to think about how the larger trends / year-to-year variability are changing across space.
+
+Next, we can make the model maybe a little more realistic by allowing the year effects to vary over space, by including a year x area interaction.
+
+```{r}
+lme.vess.int = lmer(difflog ~ area * as.factor(year) + nspecies + diffspec + ngear + diffgear + (1 | adfg), data = totalVesselsExp)
+summary(lme.vess.int)
+```
+
+Does this improve the model?
+Based on the AIC values, the answer is yes.
+One of the interesting things with both models is that both 'diffspec' and 'diffgear' are statistically significant, and have positive effects [interpretation: increasing either species landed or gears fished from one year to the next increases revenue].
+
+```{r}
+m = matrix(c("1 trend", "Regional trends", round(AIC(lme.vess.1trend), 1), round(AIC(lme.vess.int), 1)), 2, 2)
+colnames(m) = c("Model", "AIC")
+kable(m)
+```
+
+We can look at how the year effects vary by region:
+
+```{r, fig.pos="placeHere",fig.cap="Fixed effects by region from the mixed effects model (represents area x time interaction)."}
+coef.bb = c(0, fixef(lme.vess.int)[7:34])# Bristol Bay
+coef.ci = coef.bb + fixef(lme.vess.int)[2] + 
+  c(0, fixef(lme.vess.int)[grep("areaCookInlet",  
+  names(fixef(lme.vess.int)))[-1]])
+coef.ke = coef.bb + fixef(lme.vess.int)[3] + 
+  c(0, fixef(lme.vess.int)[grep("areaKodiak.east", 
+  names(fixef(lme.vess.int)))[-1]])
+coef.kw = coef.bb + fixef(lme.vess.int)[4] + 
+  c(0, fixef(lme.vess.int)[grep("areaKodiak.west", 
+  names(fixef(lme.vess.int)))[-1]])
+coef.pw = coef.bb + fixef(lme.vess.int)[5] + 
+  c(0, fixef(lme.vess.int)[grep("areaPWS", 
+  names(fixef(lme.vess.int)))[-1]])
+coef.se = coef.bb + fixef(lme.vess.int)[6] + 
+  c(0, fixef(lme.vess.int)[grep("areaSoutheast", 
+  names(fixef(lme.vess.int)))[-1]])
+
+plot(1986:2014, coef.bb, type = "l", col = tim.colors(6)[1], ylim = c(-2, 2), ylab = "Year effects by region")
+lines(1986:2014, coef.ci, type = "l", col = tim.colors(6)[2])
+lines(1986:2014, coef.kw, type = "l", col = tim.colors(6)[3])
+lines(1986:2014, coef.ke, type = "l", col = tim.colors(6)[4])
+lines(1986:2014, coef.se, type = "l", col = tim.colors(6)[5])
+lines(1986:2014, coef.pw, type = "l", col = tim.colors(6)[6])
+legend("topleft", c("Bristol Bay", "Cook Inlet", "Kodiak - east", "Kodiak - west",
+"Southeast", "PWS"), col = tim.colors(6), lty = 1)
+```
+
+But if we look at the model output more closely, we see that the variance of the vessel specific random effects is driven to 0.
+This may be because of the large number of vessels with single values (remember, each value indicates a vessel that fishes in consecutive years, because our response is based on the difference).
+
+```{r}
+ # Calculate proportion of vessels with 1 observation
+round(length(table(totalVesselsExp$adfg))/length(which(table(totalVesselsExp$adfg) == 1)), 2)
+```
+
+We could also refine the model to concentrate on vessels that we more active – e.g. vessels with 10 or more years of observations.
+
+```{r}
+activeVess = names(which(table(totalVesselsExp$adfg) >= 10))
+lme.vess.int10 = lmer(difflog ~ area + as.factor(year) + nspecies + diffspec +
+    ngear + diffgear + (1 | adfg), data = totalVesselsExp[totalVesselsExp$adfg %in%
+activeVess, ])
+summary(lme.vess.int10)
+```
+
+It might be worth looking at these active vessels, and going back to the original data to get a better sense for the relationship between increasing species fished and increasing profit.
+
+```{r, fig.pos="placeHere",fig.cap="Individual performance (log total revenue in red) and species landed (raw = black, effective = blue)."}
+activeVess = names(which(table(totalVesselsExp$adfg) >= 10))
+activeVess.id = names(rev(sort(table(totalVesselsExp[totalVesselsExp$adfg %in% activeVess, ]$adfg))))
+
+par(mfrow = c(5, 5), mai = c(0.01, 0.01, 0.01, 0.01))
+for (i in 1:25) {
+  df = as.data.frame(totalVesselsExp[totalVesselsExp$adfg == activeVess.id[i], ])
+  plot(df$nspecies, type = "l")
+  lines(rescale(log(df$totalearnings), range(df$nspecies)), col = "red")
+  lines(rescale(log(df$effectiveSpec), range(df$nspecies)), col = "blue")
+}
+```
+
+Another way to make the model more complex would be to allow for time varying effects.
+For all models above, the variable ‘diffspec’ is the strongest, so we could allow this variable to vary over time and / or space.
+
+```{r}
+lme.vess.int.area = lmer(difflog ~ area * as.factor(year) + nspecies + area * diffspec + ngear + diffgear + (1 | adfg), data = totalVesselsExp)
+lme.vess.int.time = lmer(difflog ~ area * as.factor(year) + nspecies + as.factor(year) * diffspec + ngear + diffgear + (1 | adfg), data = totalVesselsExp)
+lme.vess.int.timearea = lmer(difflog ~ area * as.factor(year) + nspecies + area * as.factor(year) * diffspec + ngear + diffgear + (1 | adfg), data = totalVesselsExp)
+```
+
+In this case, the comparison among the three models suggests that there’s more support for the coefficient varying spatially and temporally.
+[interpretation: the benefit of increasing species landed varies over time, and has a different pattern for each spatial region included here].
+
+```{r}
+m = matrix(c("Area", "Time", "Area-Time", round(AIC(lme.vess.int.area), 1), round(AIC(lme.vess.int.time), 1), round(AIC(lme.vess.int.timearea), 1)), 3, 2)
+colnames(m) = c("Model of diffspec", "AIC")
+kable(m)
+```
+
+We can also explore whether the effective number of species fished is a better proxy for species diversity, in comparing our two variables.
+
+```{r}
+lme.vess.int.area.effective = lmer(difflog ~ area * as.factor(year) + effectiveSpec + area * diffspec + effectiveGear + diffeffgear + (1 | adfg), data = totalVesselsExp)
+```

### 2016-03-18 process_raw_data/process_raw_cfec_data.Rmd
update

---
title: "Description of raw CFEC data processing for NCEAS portfolio effects working group"
author: "Jordan Watson & Eric Ward"
date: "March 16, 2016"
output: pdf_document
---

```{r setup, include=FALSE}
library(knitr)
library(dplyr)
library(ggplot2)
library(mgcv)
knitr::opts_chunk$set(echo = TRUE)
```

## Objective

The objective of this document is to summarize the data cleaning that was done to clean the raw CFEC data (1985 - 2014) for use by the NCEAS catch portfolio working group.

## Raw data summary

We'll initially restrict our analysis to data after 1984, since the data from 1975-1984 were dominated by foreign vessels (data we don't have). For more information, see AFSC technical reports (such as http://www.afsc.noaa.gov/Publications/AFSC-TM/NOAA-TM-AFSC-54.pdf). The cutoff of 1984/1985 is also consistent with the ADFG use. Further, the species reporting changes drastically before / after 1985 -- examples include no rockfish (demersal, pelagic, slope categories below), flatfish (shallow, deep categories below).

Additional information on CFEC data here,  
Gear descriptions: https://www.cfec.state.ak.us/misc/GEARDESC.HTM  
Fishery codes: https://www.cfec.state.ak.us/misc/FshyDesC.htm  
Species and gear codes:   https://www.adfg.alaska.gov/static/license/fishing/pdfs/elanding_codes.pdf  
Commercial areas: http://www.adfg.alaska.gov/index.cfm?adfg=fishingCommercialByFishery.statmaps  
Halibut areas: http://www.iphc.int/publications/techrep/tech0049.pdf

The file "cfec_w_IDs" is a zip file of an R workspace created by Jordan Watson on 03/15/2016. This file is a large workspace (~ 15 GB) that contains a single names object ('cfecnew'). The size of the data frame is approximately 19.8 million rows by 77 variables.

```{r dataload, warnings=FALSE, message = FALSE}
load("cfec_w_IDs")
cfec = cfecnew
rm(cfecnew) # remove the old named object
```

For clarity, this workspace contains some variables that have been renamed. The original variable names were changed with the following commented out code,  
```{r}
#cfecnew <- rename(cfecnew, year=F_YEAR,         # ADFG_B_BATCH_YEAR
#               port=F_PORT,         # ADFG_H_PORT_CODE
#               p_holder=P_FILENO,   # CFEC_FILE_NUMBER - unique permit holder ID
#               p_serial=P_SERIAL,   # CFEC_PMT_SERIAL  - unique permit ID
#               g_earn=G_EARN,       # CFEC_VALUE
#               g_pounds=F_NETLBS,   # ADFG_I_POUNDS
#               f_ticket=F_TICKET,   # ADFG_H_SEQ_TICKET_NUMBER - unique fish tiket ID
#               adfg=F_ADFG,         # ADFG_H_ADFG_NUMBER - don't use this one
#               cadfg=C_ADFG,        # CFEC_CORRECTED_ADFG
#               harvest=F_HARVST,    # ADFG_I_HARVEST_CODE
#               procid=F_PROCSS,     # ADFG_H_PROCESSOR_CODE
#               g_spcs=G_SPCS,       # CFEC_SPECIES_CODE
#               stat6=F_STAT,        # CFEC_STAT_AREA
#               specn=F_SPCS,        # ADFG_I_SPECIES_CODE
#               p_fshy=P_FSHY,       # CFEC_PMT_FSHY
#               p_check=P_CHECK)     # CFEC_PMT_CHECK

# cfecnew$g_price=ifelse(cfecnew$g_pounds>0,cfecnew$g_earn/cfecnew$g_pounds,0)
```

## Deleting variables

There's a number of variables that can be deleted either because they have no information (null values), have been used inconsistently (the double permit variables 'D_OPER', etc. were used 2008-2013), or are either redundant or not informative with respect to the questions we're interested in, so we'll drop these out to save memory. More details about these variables can be found in the associated metadata (.xlsx workbook).

```{r delColumns}
isNull = c("F_FG_MGT","C_CFACT", "F_DISTRC", "F_ERRFLG", "F_AYK", "G_FED_AR","F_BTAREA",
"A_CITY", "A_STATE", "A_ZIP", "P_YEAR", "I_NAME", "I_NAMTYP", "F_POT_DR", "F_PRCENT", 
"C_ADFGST", "F_CTDATE", "F_LNDATE", "F_NUMBER", "F_VALUE", "F_POUNDS", "F_TNUMB",  
"F_BATCH", "F_PERIOD",  "F_EMBSEQ", "F_PROCCK", "F_WEEK",  "F_TTYPE",    
"F_ITEM", "F_FSHERY",  "G_GEAR", "G_AREA", "G_PRC_AR",  "P_YEAR", "P_PMTSEQ", 
"p_check", "P_ID", "P_FEETYP", "P_ADFG", "P_FEEFLG", "P_FSHBLE", "F_DISPOS", 
"F_MPGMNM", "D_OPER", "D_O_FSHY", "D_O_PNUM", "D_O_PCHK", "D_O_YSEQ", "F_SIZE", 
"D_O_YCHK", "adfg")
cfec = cfec[,-which(names(cfec)%in%isNull)]
```

This significantly reduces the size of the dataset, to the remaining 29 variables,  
```{r}
names(cfec)
```

The fish ticket field (f_ticket) provides unique fish ticket numbers by year. To make unique fish tickets across all years, affix the last two digits of each year to the fish ticket ID number. We subesquently overwrite the original column with the new one.
```{r renameFtix}
cfec <- group_by(cfec,year) %>% mutate(ft=paste0(gsub("19","",year),f_ticket,p_holder)) %>% dplyr::select(-f_ticket) %>% rename(f_ticket=ft)
```

## Deleting records

The CFEC data contains a number of records that we don't want to be included in our analyses. The first obvious group includes landings that are non-commercial, from test-fisheries, discards, etc. There are several fields we can use to do this.

First, we'll use 'C_LANDST' to remove all non-commercial records. We won't want to use commercial landings that are bycatch (Pacific Salmon Commission) or discards, because they'd show up as having no value. As a second filter, we can use the corrected ADFG number 'cadfg' to remove those problematic records (missing values, values < 10000, values > 98000). Third, we'll delete when the 'P_STATUS' is blank or 'O' (15 records). We may want to filter on the remaining codes,  
*A*: revoked permit holder  
*C*: current permit holder  
*E*: former emergency transferor  
*F*: former permit holder (permanent transferor)  
*I*: inactive emergency transferee

Finally, we can use the 'P_TYPE' field to include the following 4 types of permits:  
*E*: interim-entry in a limited fishery  
*I*: interim-use in an unlimited fishery  
*M*: moratorium  
*P*: permanent in a limited fishery

```{r}
cfec = cfec[which(cfec$C_LANDST == "C"), ]

cfec = cfec[-which(cfec$cadfg == ""), ]
cfec = cfec[-which(cfec$cadfg < 10000 | cfec$cadfg >= 98000), ]

cfec = cfec[-which(cfec$P_STATUS%in%c("", "O")), ]

table(cfec$P_STATUS)
```

## Merging species and gears

### Gears
We have to clean up a few of the gears. Drop a few gears that are sporadically used, and combine longline and pot gears because of changes in how those gears were reported. This doesn't destroy information, because we still have data on vessel length, etc.

```{r gearLump}
cfec$gearn = as.numeric(cfec$F_GEAR)
cfec = cfec[, -which(names(cfec)=="F_GEAR")]
# 8 = fishwheel, used at start / end, 676 recs
# 13 = dipnet, only last 2 years, 329 recs
# 23 = Hydro/mechanical dredge, only 74 recs
# 27 = Double otter trawl, 37 recs 
# 37 = Pair trawl, 16 recs
# 41 = Sunken gillnet, 59 recs
# 85 = ??, 5 recs
# 90 = trap, 1 rec
# 99 = other, stopped reporting, 175 recs
cfec = cfec[-which(cfec$gearn%in%c(8, 13, 23, 27, 37, 41, 85, 90, 99)),]
# Merge 6 and 61 into longline
cfec$gearn[which(cfec$gearn==61)] = 6
# Merge all pot gears together -- note only 9 and 91 are reported
cfec$gearn[which(cfec$gearn%in%c(9, 19, 29, 39, 49, 59, 69, 91))] = 9
```

### Species  
Because of changes in species reporting (examples = rockfish, the use of unidentified species), we need to do some lumping some species. Groupings were consistent to be consistent with the ADFG codes, so for example rockfish are grouped into a 'pelagic', 'demersal', and 'slope' group.  
```{r specLump}

# Lump all skates together
cfec$specn[cfec$specn==700] = 700
cfec$specn[cfec$specn==701] = 700
cfec$specn[cfec$specn==702] = 700
cfec$specn[cfec$specn==703] = 700
cfec$specn[cfec$specn==704] = 700
cfec$specn[cfec$specn==705] = 700

# Lump all sharks together
cfec$specn[cfec$specn==691] = 689
cfec$specn[cfec$specn==689] = 689
cfec$specn[cfec$specn==692] = 689

# Lump flatfish (Shallow)
cfec$specn[cfec$specn==123] = 119 # rock
cfec$specn[cfec$specn==129] = 119 # starry flounder
cfec$specn[cfec$specn==127] = 119 # yellowfin
cfec$specn[cfec$specn==126] = 119 # butter

# 5, 6, 7, 15, 26, 47, 61, 
# Lump flatfish (deep)
cfec$specn[cfec$specn==124] = 118 # dover
cfec$specn[cfec$specn==134] = 118 # greenland turbot

# Lump rockfish, codes from here:
# Slope rockfish
cfec$specn[cfec$specn==141] = 144 # POP
cfec$specn[cfec$specn==136] = 144 # northern
cfec$specn[cfec$specn==151] = 144 # rougheye
cfec$specn[cfec$specn==166] = 144 # sharpchin
cfec$specn[cfec$specn==152] = 144 # shortraker
cfec$specn[cfec$specn==153] = 144 # redbanded
cfec$specn[cfec$specn==157] = 144 # silvergray
cfec$specn[cfec$specn==158] = 144 # redstripe
cfec$specn[cfec$specn==137] = 144 # bocaccio
cfec$specn[cfec$specn==144] = 144 # unid
# Demersal rockfish
cfec$specn[cfec$specn==138] = 168 # copper
cfec$specn[cfec$specn==145] = 168 # yelloweye
cfec$specn[cfec$specn==146] = 168 # Canary
cfec$specn[cfec$specn==147] = 168 # quillback
cfec$specn[cfec$specn==148] = 168 # tiger
cfec$specn[cfec$specn==149] = 168 # China
cfec$specn[cfec$specn==150] = 168 # rosethorn
# Pelagic rockfish
cfec$specn[cfec$specn==154] = 169 # dusky
cfec$specn[cfec$specn==155] = 169 # yellowtail
cfec$specn[cfec$specn==156] = 169 # widow
cfec$specn[cfec$specn==172] = 169 # dusky
cfec$specn[cfec$specn==173] = 169 # dark
cfec$specn[cfec$specn==142] = 169 # black
cfec$specn[cfec$specn==167] = 169 # blue

#  Whitefish and trout did not have text labels in the species database.
#  We created our own for easier reference but given this lack of detail
#  it seems reasonable to lump these species groups for the sake of clarity
# Whitefish
cfec$specn[cfec$specn==580] <- 580  # Whitefish
cfec$specn[cfec$specn==583] <- 580  # Least cisco
cfec$specn[cfec$specn==584] <- 580  # Arctic cisco
cfec$specn[cfec$specn==585] <- 580  # Bering cisco
cfec$specn[cfec$specn==589] <- 580  # Humpback whitefish
# Trout - convert all to the same number - we'll label as "trout" later
cfec$specn[cfec$specn==520] <- 520  # Arctic char
cfec$specn[cfec$specn==530] <- 520  # Dolly varden
cfec$specn[cfec$specn==531] <- 520  # Dolly varden anadromous
cfec$specn[cfec$specn==540] <- 520  # Steelhead
cfec$specn[cfec$specn==560] <- 520  # Cutthroat

# Halibut - appear under several different historic species codes so we consolidate these
cfec$specn[cfec$specn %in% c(200:204)] <- 200

# Salmon roe (by species)
# We consolidate roe of different species into a single roe category
cfec$specn[cfec$specn %in% c(400:405)] <- 400
# In early years, there were separate specn for salmon roe and salmon. In later years,
# salmon roe got absorbed into the specn for salmon but given a F_DELIVR==14.
# We may want to change this at some point, but for now, if F_DELIVR==14, then change
# the specn to that for roe instead of the specn "salmon."
cfec$specn[(cfec$specn %in% c(410,420,430,440,450)) & cfec$F_DELIVR=="14"] <- 400

# We no longer need the F_DELIVR column and can remove it
cfec <- dplyr::select(cfec,-F_DELIVR)

#------------------------------------------
# To add text species codes, load and join with Species lookup table
load("SpeciesTable.RData")

#  000,033 and 095 don't show up in the species codes (they account for only 7 records)
#  Remove them
cfec <- filter(cfec,!specn %in% c("000","033","095"))
#  To avoid issues with different factor levels, convert specn to numeric for now.
cfec$specn <- as.numeric(as.character(cfec$specn))

#  Join the cfec dataset with the "spec" 
cfec <- left_join(cfec,spec.tbl)

#  Rename some of the groups
cfec$spec <- as.character(cfec$spec)
cfec$spec[cfec$specn==700] = "SKATE"
cfec$spec[cfec$specn==689] = "SHARK"
cfec$spec[cfec$specn==119] = "FLAT.shallow"
cfec$spec[cfec$specn==118] = "FLAT.deep"
cfec$spec[cfec$specn==144] = "ROCK.slope"
cfec$spec[cfec$specn==168] = "ROCK.demersal"
cfec$spec[cfec$specn==169] = "ROCK.pelagic"
cfec$spec[cfec$specn==520] = "TROUT"

# Lump king crabs
cfec$spec[cfec$spec%in%c("GKCR","BKCR","UKCR","RKCR")] = "KCRB" # king
# Lump tanner crabs
cfec$spec[cfec$spec%in%c("UTCR","BTCR","aTCGR","aTCTR")] = "TCRB" # tanner
# Lump shrimp - spot prawn, pink shrimp, coonstriped 
cfec$spec[cfec$spec%in%c("SPRW","PSHP","USRM","aSSHP","aCOOS")] = "USRM" # shrimp

cfec$spec[cfec$spec==""] <- NA

# drop records with no spp
cfec = cfec[-which(is.na(cfec$spec)),]
```

## Assigning coarse spatial regions  
Examining the spatial diversity of fishing behavior is complicated because we have 3 types of area codes. The 6-digit codes represent groundfish management areas (federal), the 5-digit codes are state salmon/shellfish areas, and the 3-digit codes are halibut areas. For the groundfish and salmon areas, there are sub-areas that are also reported (as the last 2-digits), so we'll lump sub-areas together by stripping off the last 2-digits. The challenge too is that all three of these areas overlap in space -- so in a single location (latitude, longitude), a vessel may land a fish and record it in any of the 3 areas (for examples of this occuring, see species like sablefish).

For larger spatial comparisons, we'll create a field called 'area'. We'll group groundfish / salmon-shellfish / halibut areas into the following regions: Prince William Sound, Kodiak (east and west), Cook Inlet, and Southeast Alaska. More regions could be added below, and records that don't fit into one of these regions are left as NAs. The justification of these boundaries are largely based on halibut areas (which are the largest spatially) and by a series of k-means cluster analyses that identified similar groupings.

```{r}
cfec$area = NA
cfec$area[cfec$stat6==485900] = "PWS"
cfec$area[cfec$stat6==475900] = "PWS"
cfec$area[cfec$stat6==465900] = "PWS"
cfec$area[cfec$stat6==455900] = "PWS"
cfec$area[cfec$stat6==445900] = "PWS"
cfec$area[cfec$stat6==495930] = "PWS"
cfec$area[cfec$stat6==485930] = "PWS"
cfec$area[cfec$stat6==475930] = "PWS"
cfec$area[cfec$stat6==465930] = "PWS"
cfec$area[cfec$stat6==455930] = "PWS"
cfec$area[cfec$stat6==445930] = "PWS"
cfec$area[cfec$stat6==486000] = "PWS"
cfec$area[cfec$stat6==476000] = "PWS"
cfec$area[cfec$stat6==466000] = "PWS"
cfec$area[cfec$stat6==456000] = "PWS"
cfec$area[cfec$stat6==446000] = "PWS"
cfec$area[cfec$stat6==486030] = "PWS"
cfec$area[cfec$stat6==476030] = "PWS"
cfec$area[cfec$stat6==466030] = "PWS"
cfec$area[cfec$stat6==456030] = "PWS"
cfec$area[cfec$stat6==486100] = "PWS"
cfec$area[cfec$stat6==476100] = "PWS"
cfec$area[cfec$stat6==466100] = "PWS"
cfec$area[which(cfec$stat6 < 20099 & cfec$stat6 >= 20000)] = "PWS"
cfec$area[which(cfec$stat6 < 21299 & cfec$stat6 >= 21200)] = "PWS"
cfec$area[which(cfec$stat6 < 22799 & cfec$stat6 >= 22700)] = "PWS"  
cfec$area[which(cfec$stat6 < 22699 & cfec$stat6 >= 22600)] = "PWS"
cfec$area[which(cfec$stat6 < 22899 & cfec$stat6 >= 22800)] = "PWS"
cfec$area[which(cfec$stat6 < 22199 & cfec$stat6 >= 22100)] = "PWS"
cfec$area[which(cfec$stat6 < 22299 & cfec$stat6 >= 22200)] = "PWS"  
cfec$area[which(cfec$stat6 < 22599 & cfec$stat6 >= 22500)] = "PWS"
cfec$area[which(cfec$stat6 < 22499 & cfec$stat6 >= 22400)] = "PWS"
cfec$area[which(cfec$stat6 < 22399 & cfec$stat6 >= 22300)] = "PWS" 
# Add halibut areas
cfec$area[cfec$stat6 %in% c(242,232,220,230,240)] = "PWS"

cfec$area[cfec$stat6==525900] = "CookInlet"
cfec$area[cfec$stat6==535900] = "CookInlet"
cfec$area[cfec$stat6==515930] = "CookInlet"
cfec$area[cfec$stat6==525930] = "CookInlet"
cfec$area[cfec$stat6==535930] = "CookInlet"
cfec$area[cfec$stat6==516000] = "CookInlet"
cfec$area[cfec$stat6==526000] = "CookInlet"
cfec$area[cfec$stat6==496030] = "CookInlet"
cfec$area[cfec$stat6==506030] = "CookInlet"
cfec$area[cfec$stat6==516030] = "CookInlet"
cfec$area[cfec$stat6==526030] = "CookInlet"
cfec$area[cfec$stat6==506100] = "CookInlet"
cfec$area[cfec$stat6==516100] = "CookInlet"
cfec$area[which(cfec$stat6 < 24799 & cfec$stat6 >= 24400)] = "CookInlet" 
cfec$area[which(cfec$stat6%in%c(261, 272))] = "CookInlet"

cfec$area[which(cfec$stat6 < 11599 & cfec$stat6 >= 11100)] = "Southeast"
cfec$area[which(cfec$stat6 < 10499 & cfec$stat6 >= 10100)] = "Southeast"
cfec$area[which(cfec$stat6 < 11099 & cfec$stat6 >= 10500)] = "Southeast" 
cfec$area[which(cfec$stat6 >= 305430 & cfec$stat6 <= 305530)] = "Southeast"
cfec$area[which(cfec$stat6 >= 315430 & cfec$stat6 <= 315600)] = "Southeast"
cfec$area[which(cfec$stat6 >= 325430 & cfec$stat6 <= 325700)] = "Southeast"
cfec$area[which(cfec$stat6 >= 335430 & cfec$stat6 <= 335800)] = "Southeast"
cfec$area[which(cfec$stat6 >= 345400 & cfec$stat6 <= 345830)] = "Southeast"
cfec$area[which(cfec$stat6 >= 355400 & cfec$stat6 <= 355900)] = "Southeast"
cfec$area[which(cfec$stat6 > 365400 & cfec$stat6 < 365830)] = "Southeast"
# For IPHC see Fig. 20 here: http://www.iphc.int/publications/techrep/tech0049.pdf
cfec$area[which(cfec$stat6%in%c(140, 141, 142, 143, 144, 150, 151, 152, 153, 160, 161, 
162, 163, 170, 171, 172, 174, 174, 181, 182, 183, 184, 185))] = "Southeast"

# Kodiak east area is defined by halibut areas 270/280, includes ADFG salmon / 
# shellfish on E side of kodiak, and groundfish areas extending to NMFS Area 630
# boundary
cfec$area[cfec$stat6%in%c(270,280)] = "Kodiak.east"
cfec$area[which(cfec$stat6 < 25899 & cfec$stat6 >= 25800)] = "Kodiak.east"
cfec$area[which(cfec$stat6 < 25999 & cfec$stat6 >= 25900)] = "Kodiak.east"
cfec$area[cfec$stat6==535630] = "Kodiak.east"
cfec$area[cfec$stat6==525630] = "Kodiak.east"
cfec$area[cfec$stat6==515630] = "Kodiak.east"
cfec$area[cfec$stat6==505630] = "Kodiak.east"
cfec$area[cfec$stat6==495630] = "Kodiak.east"
cfec$area[cfec$stat6==535700] = "Kodiak.east"
cfec$area[cfec$stat6==525700] = "Kodiak.east"
cfec$area[cfec$stat6==515700] = "Kodiak.east"
cfec$area[cfec$stat6==505700] = "Kodiak.east"
cfec$area[cfec$stat6==495700] = "Kodiak.east"
cfec$area[cfec$stat6==525730] = "Kodiak.east"
cfec$area[cfec$stat6==515730] = "Kodiak.east"
cfec$area[cfec$stat6==505730] = "Kodiak.east"
cfec$area[cfec$stat6==495730] = "Kodiak.east"
cfec$area[cfec$stat6==525800] = "Kodiak.east"
cfec$area[cfec$stat6==515800] = "Kodiak.east"
cfec$area[cfec$stat6==505800] = "Kodiak.east"
cfec$area[cfec$stat6==495800] = "Kodiak.east"

cfec$area[cfec$stat6%in%c(271,281)] = "Kodiak.west"
cfec$area[cfec$stat6==535730] = "Kodiak.west"
cfec$area[cfec$stat6==545730] = "Kodiak.west"
cfec$area[cfec$stat6==525800] = "Kodiak.west"
cfec$area[cfec$stat6==535800] = "Kodiak.west"
cfec$area[cfec$stat6==545800] = "Kodiak.west"
cfec$area[cfec$stat6==535830] = "Kodiak.west"
cfec$area[cfec$stat6==525830] = "Kodiak.west"
cfec$area[which(cfec$stat6 < 25699 & cfec$stat6 >= 25500)] = "Kodiak.west"
cfec$area[which(cfec$stat6 < 25499 & cfec$stat6 >= 25300)] = "Kodiak.west"  
cfec$area[which(cfec$stat6 < 26299 & cfec$stat6 >= 26000)] = "Kodiak.west"

# Bristol Bay, defined as going west to the boundary between NMFS areas 512 / 516
cfec$area[cfec$stat6%in%c(31600, 31700, 31800, 32000, 32100, 
  32200, 32400, 32500, 32600)] = "BristolBay"
cfec$area[cfec$stat6%in%c(575730, 575800, 575830)] = "BristolBay"
cfec$area[cfec$stat6%in%c(585700, 585730, 585800, 585830)] = "BristolBay"
cfec$area[cfec$stat6%in%c(595630, 595700, 595730, 595800, 595830)] = "BristolBay"
cfec$area[cfec$stat6%in%c(605600, 605630, 605700, 605730, 605800, 605830)] = "BristolBay"
cfec$area[cfec$stat6%in%c(615600, 615630, 615700, 615730, 615800, 615830)] = "BristolBay"

```

## Strip the permit holder's identity

Just for safety (because 'p_holder' is technically traceable to a person), we're going to convert 'p_holder', 'cadfg', 'p_serial' to a non-traceable identifier. We'll also re-number fish tickets 1:n, so there's no information here that could be used to identify individuals and/or vessels.

```{r}
cfec$p_holder = as.numeric(as.factor(cfec$p_holder))
cfec$cadfg = as.numeric(as.factor(cfec$cadfg))
cfec$p_serial = as.numeric(as.factor(cfec$p_serial))

# number fish tickets sequentially
cfec$f_ticket = seq(1, dim(cfec)[1])
cfec$F_PERMIT = as.numeric(as.factor(cfec$F_PERMIT))
```

## Output data file for use on other projects
```{r}
save(cfec, file = "cfecCleaned.Rdata")
```

# Meta-data for saved variables

*F_REGION*

*year* Year of landed catch

*g_pounds* This was *F_NETLBS* in original CFEC data. Net pounds what gets delivered to dock.

*F_PERMIT* permit fishery, serial, then check digit.

*F_INPFC* An identification number assigned to a CDQ group by NMFS. Only applicable if we're interested in CDQ quota.

*port* This was *F_PORT* in original CFEC data. Port code.

*procid* This was *F_PROCSS* in original CFEC data. Processor code.

*f_ticket* This fish ticket identification number is stamped on the fish tickets at the ADFG local area offices and is annually unique. First 2 digits = office #, last 6 = serial. Our data only have 6 digits total and I'm not sure why. However, in order to make values that are unique across all years and not just within a year, we have appended the last two digits of the year to the front of the ticket.

*harvest* This was *F_HARVST* in original CFEC data. This code is used to separate the common property commercial catch from a variety of specialty fisheries. Examples of specialty fisheries are test fisheries, educational fisheries, hatchery controlled fisheries, and derbies.

*specn* This was *F_SPCS* in original CFEC data. Species number. As a note of caution, care should be taken in merging our cleaned databases with other datasets because of how we changed the species codes in some cases (e.g. all halibut 200-204 became 200, or see rockfish groups above).

*stat6* This was *F_STAT* in original CFEC data. The 3, 5, or 6 digit ADF&G statistical code denoting a specific area of catch.

*g_spcs* G_SPCS in original CFEC data. A species code assignment based on grouped ADF&G species codes (*ADFG_I_SPECIES_CODE*). 'A' is abalone, 'B' is halibut, 'C' is sablefish, 'D' is Dungeness crab, 'E' is hair crab, 'F' is fresh water fish, 'G' is sac roe herring, 'H' is herring, bait herring, and food herring, 'I' is lingcod, 'J' is geoduck, 'K' is king crab, 'L' is herring roe or herring roe on kelp or other substrate, 'M' is groundfish, 'N' is snail, 'O' is octopus or squid, 'P' is shrimp, 'Q' is sea cucumber, 'R' is clam (except geoduck), 'S' is salmon, 'T' is Tanner cab, 'U' is sea urchin, 'W' is scallops, 'Y' is Southeast Alaska demersal shelf rockfish, and 'Z' are miscellaneous marine invertebrates.  
*g_earn* G_EARN in original CFEC data. Estimated ex-vessel value (gross earnings) of a fish ticket item based on the net pounds (ADFG_I_POUNDS) multiplied by the price per pound.

*p_serial* P_SERIAL in original CFEC data. Permit serial number. Different serial number ranges are used depending upon the permit type (*CFEC_PMT_TYPE*). Mariculture (aquatic farm) is 10000-10499, experimental is 10500-10599, test fishing is 10600-10699, educational is 10700-10799, reservation is 10800-10899, hatchery cost recovery is 10900-10999, interim-use in an unlimited fishery is 11000-49999, interim-entry in a limited fishery is 50000-54999, moratorium is 50000-54999, vessel moratorium is 50000-54999, permanent in a limited fishery is 55000-99999, vessel permanent in a vessel limited fishery is 55000-99999.

*p_holder* This was *P_FILENO* in original CFEC data. A unique number assigned to each permit holder. These file numbers are generated from the ID number and range from 000000 to 999999.

*p_fshy* P_FSHY in original CFEC data. Permit fishery code. The 6 byte fishery code consists of a 2 byte resource code; a 2 byte gear code or a 1 byte gear code and a 1 byte vessel restriction or 1 byte gear code and 1 byte gear restriction; and a 1 byte CFEC administrative area. The 6th byte indicates a landing permit ('L'), vessel permit ('V'), or CDQ group ('A', 'B', 'C', 'D', 'E', 'F', or 'G').

*P_TYPE* The type of permit. 'F' is a mariculture (aquatic farm) permit, 'X' is experimental, 'T' is test fishing, 'C' is educational, 'R' is reservation, 'H' is hatchery cost recovery, 'I' is interim-use in an unlimited fishery, 'E' is interim-entry in a limited fishery, 'M' is moratorium, 'V' is vessel moratorium, 'P' is permanent in a limited fishery, and 'L' is permanent vessel permit in a vessel limited fishery.

*P_STATUS* Permit holder status. 'A' is revoked permit holder, 'C' is current permit holder, 'D' is current permit holder that emergency transferred permit (transferor, valid in current year only), 'E' is former emergency transferor, 'F' is former permit holder (permanent transferor), 'I' is inactive emergency transferee, 'O' is old permit owner (permit never active, this indicates who paid renewal fees when a transfer is done after fees have been paid but before the calendar year of the permit), 'T' is current permit holder of emergency transfer (transferee, valid in current year only).

*A_RES* The declared residency of the permit holder. 'R' is Alaska resident, 'N' is nonresident, and 'U' indicates the person has never signed a statement regarding residency.

*cadfg* This was *C_ADFG* in original data. Corrected ADFG code.

*C_LANDST* A flag indicating whether or not the landing is defined as commercial or non-commercial catch by CFEC. 'C' indicates commercial catch and 'N' indicates non-commercial catch.

*F_MPGMID* Indicates the program under which fishing occurred. It is the management program abbreviation such as LE (limited entry), CDQ, AFA, or ADAK. For IFQ crab, halibut, or sablefish, the management program is IFQ. For groundfish with no other management program the value should be OA, for open access. Populated 2005 on in shellfish and groundfish. Populated 2009 on for  Salmon database data with landing report number.

*startdt* The original data had a field *F_CTDATE* - Date gear was deployed in the water or harvest of fish or shellfish was initiated. This field was only in the MMDD format. startdt is a concatenation of the *F_CTDATE* and the year of fishing *F_YEAR*.

*landdate* The original data had a field *F_LNDATE* - The date fish or shellfish are off-loaded or trans-shipped from the catcher vessel to the first purchaser. The land date is opposed to Catch Date however; these two dates can be the same. If off-loading takes more than one day to complete, the land date is the day off-load is completed. For catcher-seller vessels the land date is the date the product is initially. This field was only in the MMDD format, so landdate is a concatenation of the *F_LNDATE* and the year of fishing *F_YEAR*.

*g_price* Calculated as g_earn/g_pounds

*gearn* Numeric code used to represent gears.

*spec* Derived species code. This should be used for coarse filtering, but metrics of diversity should be calculated from the numeric version, *specn*.

*area* Coarse areas for comparison across the Gulf of Alaska.

### 2016-03-18 process_raw_data/SpeciesTable.R

@@ -1,0 +1,59 @@
+load("cfec.Rdata")
+
+cfec$spec[cfec$spec=="REX "] <- "REX"
+cfec$spec[cfec$spec=="POP "] <- "POP"
+cfec$spec[cfec$spec==" POP"] <- "POP"
+
+spec.tbl <- distinct(cfec,specn,spec) %>% dplyr::select(specn,spec) %>% arrange(specn)
+spec.tbl$spec <- as.character(spec.tbl$spec)
+
+#  For those species with total landed values greater than $1000
+#  and for which a spec code is missing, create one.
+#  Species letter codes that I've created start with "a"
+spec.tbl$spec[spec.tbl$specn==161] <- "aSCLP1"  # Coastrange sculpin
+spec.tbl$spec[spec.tbl$specn==163] <- "aSCLP2"  # Fourhorn sculpin
+spec.tbl$spec[spec.tbl$specn==207] <- "aGUNN"   # Gunnel
+spec.tbl$spec[spec.tbl$specn==208] <- "aWARB"   # Warbonnet, prickleback, etc.
+spec.tbl$spec[spec.tbl$specn==214] <- "aGREN"   # Giant grenadier
+spec.tbl$spec[spec.tbl$specn==216] <- "aLUMP"   # Smooth lumpsucker
+spec.tbl$spec[spec.tbl$specn==250] <- "aTOMC"   # Tomcod
+spec.tbl$spec[spec.tbl$specn==260] <- "aFLAT"   # Pacific flatnose
+spec.tbl$spec[spec.tbl$specn==400] <- "aUROE"   # Unknown salmon roe
+spec.tbl$spec[spec.tbl$specn==401] <- "aKROE"   # Chinook roe
+spec.tbl$spec[spec.tbl$specn==402] <- "aSROE"   # Sockeye roe
+spec.tbl$spec[spec.tbl$specn==403] <- "aCROE"   # Coho roe
+spec.tbl$spec[spec.tbl$specn==404] <- "aPROE"   # Pink roe
+spec.tbl$spec[spec.tbl$specn==405] <- "aHROE"   # Chum roe
+spec.tbl$spec[spec.tbl$specn==500] <- "aPIKE"   # Northern pike
+spec.tbl$spec[spec.tbl$specn==516] <- "CPLN"    # Capelin 
[…]

### 2016-03-18 process_raw_data/old/process_cfec_data.R

@@ -1,0 +1,111 @@
+########################################################
+# Process raw data files, first merging them together
+# with common field names. Directory structure is 3 folders,
+# cfecft75_89
+# cfecft90_99
+# cfecft00_09
+########################################################
+
+# Data are separately loaded by year. Create an .Rdata file that merges them all together
+cfec = read.csv("cfecft75_89/cfecft1975.csv")
+
+indx2drop = which(names(cfec)%in%c("corradfg","origadfg","f_fshery"))# drop names not present after 1990
+cfec = cfec[,-c(indx2drop)]
+
+# Merge the other years
+for(yr in 76:89) {
+  newdata = read.csv(paste("cfecft75_89/cfecft19",yr,".csv",sep=""), stringsAsFactors=FALSE)
+  indx2drop = which(names(newdata)%in%c("corradfg","origadfg","f_fshery"))# drop names not present after 1990
+  newdata = newdata[,-c(indx2drop)]
+  
+  cfec = rbind(cfec,newdata)
+}
+for(yr in 90:90) {
+  newdata = read.csv(paste("cfecft90_99/cfecft19",yr,".csv",sep=""), stringsAsFactors=FALSE)
+  indx2drop = which(names(newdata)%in%c("corradfg","origadfg","f_fshery"))# drop names not present after 1990
+  newdata = newdata[,-c(indx2drop)]
+  
+  cfec = rbind(cfec,newdata)
+}
+
+cfec = cfec[,-which(names(cfec)%in%c("f_pounds","f_netlbs","f_value","c_landst"))]
+
+# These years have more field codes, and the names change relative to above
+for(yr in 91:99) {
+  newdata = read.csv(paste("cfecft90_99/cfecft19",yr,".csv",sep=""), stringsAsFactors=FALSE)
+  names(newdata) = tolower(names(newdata))
+  # only include GOA data - get rid of BSAI and INTL
+  #newdata = newdata[-which(newdata$fmp_area%in%c("BSAI","INTL")),]
+  newsub = newdata[,c("year","startdt","lnddate","landdate","procid","port","stat6","adfg","f_cdq","gearn",
+                      "specn","prod","size","g_pounds","g_earn","g_price","p_fshy","p_serial",
+                      "p_check","spec","g_spcs","harvest")]
+  cfec = rbind(cfec,newsub)
+}
+
[…]

### 2016-03-21 process_raw_data/process_raw_cfec_data.Rmd
bug fix in areas

@@
 cfec$area = NA
+cfec$stat6 = as.numeric(cfec$stat6)
+
 cfec$area[cfec$stat6==485900] = "PWS"

### 2016-03-22 process_raw_data/process_raw_cfec_data.Rmd
bug fix

@@
 rm(cfecnew) # remove the old named object
+
+# Remove A80 and CP vessels
+A80<-c(68870,29923,22011,39798,59870,8500,41219,62545,56965,54693,1119,57228,69038,61081,61083,55921,57211,56964,36202,43260,54392,51873,48183,55767)
+CP <-c(56987,59378,57621,59170,55767,57450,54886,56991,60407,60795,60202,60660,56789,56618,59503)
+cfec <- filter(cfec,!adfg %in% c(A80,CP))
 ```
@@
 ```{r specLump}
+cfec$specn.orig = cfec$specn
 
@@
 # number fish tickets sequentially
-cfec$f_ticket = seq(1, dim(cfec)[1])
+cfec$f_ticket = as.numeric(as.factor(cfec$f_ticket))
 cfec$F_PERMIT = as.numeric(as.factor(cfec$F_PERMIT))

### 2016-03-25 process_raw_data/process_raw_cfec_data.Rmd
commit

@@
 ```{r dataload, warnings=FALSE, message = FALSE}
-load("cfec_w_IDs")
+load("Data/cfec_w_IDs.RData")
 cfec = cfecnew
@@
 
+#idx = grep("S 04T", cfec$p_fshy)
+#group_by(cfec[idx,], year) %>% 
+#  summarize(nSetnet = length(unique(p_holder)))
 # Remove A80 and CP vessels
@@
 cfec <- filter(cfec,!adfg %in% c(A80,CP))
+rm(A80,CP)
 ```
@@
 cfec = cfec[,-which(names(cfec)%in%isNull)]
+rm(isNull)
 ```
@@
 
-cfec = cfec[-which(cfec$cadfg == ""), ]
-cfec = cfec[-which(cfec$cadfg < 10000 | cfec$cadfg >= 98000), ]
+#cfec = cfec[-which(cfec$cadfg == ""), ]
+#cfec = cfec[-which(cfec$cadfg < 10000 | cfec$cadfg >= 98000), ]
 
@@
 # To add text species codes, load and join with Species lookup table
-load("SpeciesTable.RData")
+load("Data/SpeciesTable.RData")
 

### 2016-03-25 catch_portfolios_mtg3.Rmd
init commit of meeting 3 code

---
title: "Results for working group papers"
author: "Catch portfolios subgroup: NCEAS meeting 3"
date: "March 18, 2016"
output:
  pdf_document:
    fig_caption: yes
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=FALSE, tidy.opts=list(width.cutoff=60), warning = FALSE, message = FALSE)
```

## Load cleaned data  
```{r dataLoad, warning = FALSE, message=FALSE}
library(dplyr)
library(knitr)
library(mgcv)
library(ggplot2)
library(randomForest)
library(MASS)
library(MatchIt)
load("cfecCleaned.Rdata")
```

```{r, warning = FALSE, message=FALSE}
# This is to investigate Bristol Bay set nets for Milo
# Check BB setnet records / year
#idx = grep("S 04T", cfec$p_fshy)
#group_by(cfec[idx,], year) %>% 
#  summarize(nSetnet = length(unique(p_holder)))
#cfecnew = cfec[idx,]
```

## Function definiton  
We need to define a few functions for calculating diversity. We'll use the Simpson's diversity index (HHH) to calculate the entropy-based inverse, which is a proxy for effective species diversity. Note the C++ version is commented out below,   
```{r, warning = FALSE, message=FALSE}

simp.div = function(x) {
  1/sum((x/sum(x))^2)
}

#library(Rcpp)
#cppFunction('double simpCpp(NumericVector x) {
#  double y = 1/sum((x/sum(x))^2);
#  return y;
#}')
```

# Paper 2: Investigating relative effects of management changes (IFQ) relative to EVOS
### Alan's tenatively co-leading

We'll restrict our analysis initially to a few years before and after the IFQ change, in 1995. People had to land halibut or sablefish in 1988, 1989, or 1990 to qualify for quota, and 5 of their best 6 years fished were used to calculate quota (so continuous fishing not a requirement). Alan's previously used 1992:1994 versus 1995:1997. We'll slightly expand this range initially, but may want to change it. [Side note: Eric and Sean already explored this, and there's no qualitative difference between the shorter/longer time periods].

Alan defined 4 primary groups of individuals (or vessels):  
1. Individuals who fished halibut / sablefish before and after 1995 (IFQ group)  
2. Individuals who only fished halibut / sablefish before 1995 ["Pre"]
3. Individuals who only fished halibut / sablefish after 1995 ["Post"]  
4. Individuals who never fished halibut / sablefish ["Control group"]

```{r, warning = FALSE, message=FALSE}
# We'll use 3 years before/after IFQ. This is consistent with the work that
# Alan has done. Years up to 1990 were included in the calculation of future
# quota, and 1998 is problematic because the ratio of our halibut earnings to 
# AKFIN's is ~ 60% -- suggesting there may be something systemic that we're
# missing. 
cfec$year = as.numeric(cfec$year)

dat = cfec[cfec$year %in% seq(1992,1997),]
#dat$p_holder = NULL
#dat = rename(dat, p_holder = cadfg)

# Create an indicator for species codes halibut (200) and sablefish (710)
dat$HS = 0
dat$HS[dat$g_earn > 0 & dat$specn%in%c(200, 710)] = 1

# Create a data frame that we'll use to assign years to pre- and post- groups  
df.years = data_frame(year=1992:1997, 
  ind = c(rep("Pre",3), rep("Post",3)))
dat =left_join(dat, df.years)
dat$ind = factor(dat$ind, levels = c("Pre","Post"))

dat.hs = group_by(dat, p_holder, ind) %>% 
  summarize(HS = sum(g_earn[HS==1]) > 0 ) 
hs.tbl = reshape2::dcast(dat.hs , p_holder ~ ind)
hs.tbl$Post[is.na(hs.tbl$Post)] = FALSE
hs.tbl$Pre[is.na(hs.tbl$Pre)] = FALSE

# Assign groups
hs.tbl = mutate(hs.tbl, group = ifelse(Post & Pre, "IFQ", 
  ifelse(Post & !Pre, "Post.only", 
    ifelse(!Post & Pre, "Pre.only", "Cntl"))))

dat = left_join(dat, hs.tbl)

# Even though we initially defined a control group, get rid of it because
# after a lot of group discussion, these people are not true controls:
dat = dat[dat$group != "Cntl", ]
```

## IFQ Results:  
We want to look at a number of different responses, based on revenue (mean revenue, and variability in revenue) as well as metrics of diversity (species) and metrics of effort or participation (days fished, etc).

```{r, cache = TRUE, warning = FALSE, message=FALSE}

# Convert date string to a character -- we only care about unique values
dat$day = substr(dat$landdate, 6, 10)

# We're first going to group over the indicator (pre-post), group, individual (p_holder), 
# and year. We'll calculate total revenue, total revenue made up of halibut or sablefish, 
# total pounds, total pounds made up of halibut or sablefish, effective species diversity, 
# effective gear diversity (which is likely not that interesting since most of these 
# individuals are landing halibut/sablefish on longlines), # days fished per year. 
# Second, we'll then aggregate these annual metrics across years to calculate means and 
# CVs, when possible (the CV == 0 when 2 years of participation present, NA with 1 year). 
# We'll make no restriction about # of years that someone needs to participate either 
# before or after 1995 [side note: Eric and Sean looked at this, and qualitatively, 
# results don't change]. Taking the mean of the 'annual means' will also allow us to 
# compare the revenue of someone who fishes say ~ 4 years pre-IFQ and 1 year post-IFQ.

hs.summary = group_by(dat, ind, group, p_holder, year) %>% 
  summarize(tot.revenue.ind = sum(g_earn), 
    tot.revenue.ind.h = sum(g_earn[specn==200]), 
    tot.revenue.ind.s = sum(g_earn[specn==710]),
    tot.pounds.ind = sum(g_pounds), 
    tot.pounds.ind.h = sum(g_pounds[specn==200]), 
    tot.pounds.ind.s = sum(g_pounds[specn==710]),
    eff.spec.div = simp.div(table(specn)), 
    eff.gear.div = simp.div(table(gearn)),
    numDays = length(unique(day)),
    numHalibutDays = length(unique(day[specn==200]))) %>% 
  as.data.frame %>% 
  group_by(ind, group, p_holder) %>% 
  summarize(cv.revenue.ind = sd(tot.revenue.ind,na.rm=T) / mean(tot.revenue.ind,na.rm=T),
    mean.revenue.ind = mean(tot.revenue.ind,na.rm=T), 
    mean.revenue.ind.h = mean(tot.revenue.ind.h,na.rm=T),
    mean.revenue.ind.s = mean(tot.revenue.ind.s,na.rm=T),
    mean.pounds.ind = mean(tot.pounds.ind,na.rm=T),
    mean.pounds.ind.h = mean(tot.pounds.ind.h,na.rm=T),
    mean.pounds.ind.s = mean(tot.pounds.ind.s,na.rm=T),
    eff.spec.div.cv = sd(eff.spec.div,na.rm=T)/mean(eff.spec.div,na.rm=T),
    eff.spec.div.mean = mean(eff.spec.div,na.rm=T),
    eff.gear.div.cv = sd(eff.gear.div,na.rm=T)/mean(eff.gear.div,na.rm=T),
    eff.gear.div.mean = mean(eff.gear.div,na.rm=T),
    records.per.ind = n(), nYears = length(unique(year)), 
    nDays = mean(numDays, na.rm=T),
    nHalibutDays = mean(numHalibutDays, na.rm=T))

# Here we calculate the weighted diversity metrics. 
# We'll generate 3 versions of the estimates to make sure the results are robust:  
# 1. Diversity based on frequency in fish tickets  
# 2. Diversity weighted by sum of earnings  
# 3. Diversity weighted by sum of biomass (catch, in pounds)  
hs.div = group_by(dat, ind, group, p_holder, year, specn) %>% 
  summarize(tot.weight.sp = sum(g_pounds), 
    tot.revenue.sp = sum(g_earn)) %>% 
  group_by(ind, group, p_holder, year) %>% 
  summarize(eff.spec.div.mean.earn = simp.div(tot.revenue.sp),
    eff.spec.div.mean.weight = simp.div(tot.weight.sp)) %>%
  group_by(ind, group, p_holder) %>% 
  summarize(eff.spec.div.cv.earn = 
      sd(eff.spec.div.mean.earn) / mean(eff.spec.div.mean.earn), 
    eff.spec.div.mean.earn = mean(eff.spec.div.mean.earn), 
    eff.spec.div.cv.weight = 
      sd(eff.spec.div.mean.weight) / mean(eff.spec.div.mean.weight),
    eff.spec.div.mean.weight = mean(eff.spec.div.mean.weight))
hs.summary = left_join(hs.summary, hs.div)

# Calculate vessels used per individual and join
vessels.per.ind = group_by(dat, ind, group, p_holder) %>% 
  summarize(boat.per.ind = length(unique(cadfg)))
hs.summary = left_join(hs.summary, vessels.per.ind)

# We proposed calculating diversity across fish tix -- but that's impossible because
# each fish ticket is delivery of a single species. But we can calculate the diversity 
# by day (lumping all catch / fish tickets for that day).

# This is a bit complicated, so we'll do it using a series of group_by statements that are 
# broken out here for clarity. First, calculate total weight, earnings, and 
# frequency (counts) over species by fish ticket, year, indicator, and group
hs.day = group_by(dat, ind, group, p_holder, year, day, specn) %>% 
  summarize(tot.weight.sp = sum(g_pounds), 
    tot.earn.sp = sum(g_earn), 
    tot.freq.sp = n()) %>%
  as.data.frame

# Second, calculate the diversity metrics integrating across species. We can also calculate 
# the derived proportion of halibut and sablefish that are contributing to individual 
# fish tickets (so that we can examine the distribution of these across days)
hs.day = group_by(hs.day, ind, group, p_holder, year, day) %>% 
  summarize(pHal.weight = sum(tot.weight.sp[specn==200])/sum(tot.weight.sp), 
    pHal.earn = sum(tot.earn.sp[specn==200])/sum(tot.earn.sp), 
    pHal.freq = sum(tot.freq.sp[specn==200])/sum(tot.freq.sp), 
    pSabl.weight = sum(tot.weight.sp[specn==710])/sum(tot.weight.sp), 
    pSabl.earn = sum(tot.earn.sp[specn==710])/sum(tot.earn.sp), 
    pSabl.freq = sum(tot.freq.sp[specn==710])/sum(tot.freq.sp), 
    tot.weight.sp = simp.div(tot.weight.sp), 
    tot.earn.sp = simp.div(tot.earn.sp), 
    tot.freq.sp = simp.div(tot.freq.sp)) %>% 
  as.data.frame()

# Third, we'll calculate the quantiles of these metrics across fish tickets. This will 
# help us plot the CDF style plots  
hs.day.summary = group_by(hs.day, ind, group, year) %>% 
  summarize(l.weight = quantile(tot.weight.sp, 0.25, na.rm=TRUE), 
    m.weight = quantile(tot.weight.sp, 0.5, na.rm=TRUE), 
    u.weight = quantile(tot.weight.sp, 0.75, na.rm=TRUE),
    l.freq = quantile(tot.freq.sp, 0.25, na.rm=TRUE), 
    m.freq = quantile(tot.freq.sp, 0.5, na.rm=TRUE), 
    u.freq = quantile(tot.freq.sp, 0.75, na.rm=TRUE), 
    l.earn = quantile(tot.earn.sp, 0.25, na.rm=TRUE), 
    m.earn = quantile(tot.earn.sp, 0.5, na.rm=TRUE), 
    u.earn = quantile(tot.earn.sp, 0.75, na.rm=TRUE), 
    l.hal.earn = quantile(pHal.earn, 0.25, na.rm=TRUE),
    m.hal.earn = quantile(pHal.earn, 0.5, na.rm=TRUE), 
    u.hal.earn = quantile(pHal.earn, 0.75, na.rm=TRUE), 
    l.hal.weight = quantile(pHal.weight, 0.25, na.rm=TRUE), 
    m.hal.weight = quantile(pHal.weight, 0.5, na.rm=TRUE), 
    u.hal.weight = quantile(pHal.weight, 0.75, na.rm=TRUE),
    l.hal.freq = quantile(pHal.freq, 0.25, na.rm=TRUE), 
    m.hal.freq = quantile(pHal.freq, 0.5, na.rm=TRUE), 
    u.hal.freq = quantile(pHal.freq, 0.75, na.rm=TRUE),
    l.sab.earn = quantile(pSabl.earn, 0.25, na.rm=TRUE), 
    m.sab.earn = quantile(pSabl.earn, 0.5, na.rm=TRUE), 
    u.sab.earn = quantile(pSabl.earn, 0.75, na.rm=TRUE),
    l.sab.weight = quantile(pSabl.weight, 0.25, na.rm=TRUE), 
    m.sab.weight = quantile(pSabl.weight, 0.5, na.rm=TRUE), 
    u.sab.weight = quantile(pSabl.weight, 0.75, na.rm=TRUE),
    l.sab.freq = quantile(pSabl.freq, 0.25, na.rm=TRUE), 
    m.sab.freq = quantile(pSabl.freq, 0.5, na.rm=TRUE), 
    u.sab.freq = quantile(pSabl.freq, 0.75, na.rm=TRUE))

```

```{r, fig.pos="placeHere", fig.cap="Plot of species diversity metrics (calculated by fish ticket freq.) through time for each of our 4 groups during the pre-post IFQ period. "}
ggplot(hs.day.summary, aes(year, m.freq)) + geom_line() + facet_wrap(~group) + 
  geom_ribbon(aes(ymin=l.freq, ymax=u.freq), alpha=0.2)
```

```{r, fig.pos="placeHere", fig.cap="Plot of species diversity metrics (calculated by earnings) through time for each of our 4 groups during the pre-post IFQ period. "}
ggplot(hs.day.summary, aes(year, m.earn)) + geom_line() + facet_wrap(~group) + 
  geom_ribbon(aes(ymin=l.earn, ymax=u.earn), alpha=0.2)
```

```{r, fig.pos="placeHere", fig.cap="Plot of species diversity metrics (calculated by pounds of catch) through time for each of our 4 groups during the pre-post IFQ period. "}
ggplot(hs.day.summary, aes(year, m.weight)) + geom_line() + facet_wrap(~group) + 
  geom_ribbon(aes(ymin=l.weight, ymax=u.weight), alpha=0.2)
```

Getting back to the comparisons between groups (aggregated over years in our pre-post IFQ groupings), we can convert the landed halibut and sablefish revenue and weight into a proportion of the total catch.

First, make a dotplot (with quartile range) of the pre-post comparison for each metric of revenue.  
```{r, fig.pos="placeHere", fig.cap = "Metric of annual revenue, with interquartile range, for each of the 4 IFQ groups."}

hs.summary$mean.revenue.ind.h = 
  hs.summary$mean.revenue.ind.h / hs.summary$mean.revenue.ind
hs.summary$mean.revenue.ind.s = 
  hs.summary$mean.revenue.ind.s / hs.summary$mean.revenue.ind
hs.summary$mean.pounds.ind.h = 
  hs.summary$mean.pounds.ind.h / hs.summary$mean.pounds.ind
hs.summary$mean.pounds.ind.s = 
  hs.summary$mean.pounds.ind.s / hs.summary$mean.pounds.ind

dplyr::select(hs.summary, group, ind, p_holder, 
  contains("mean"), -contains("div")) %>%
  reshape2::melt(id.vars = c("group", "ind","p_holder")) %>%
  group_by(group, ind, variable) %>% 
  filter(is.na(value)==F) %>% 
  summarize(l = quantile(value, 0.25, na.rm=T), 
    u = quantile(value, 0.75, na.rm=T), 
    m = quantile(value, 0.5, na.rm=T)) %>%
  ggplot(aes(x = ind, y = m, ymax = u, ymin = l)) + 
  geom_pointrange() + facet_grid(variable ~ group, scales="free")

```

Look at the mean number of days where catch was recorded as landed, as well as how many of those days are when halibut are landed.  
```{r}
dplyr::select(hs.summary, group, ind, p_holder, 
  contains("nDays"), contains("nHalibutDays")) %>%
  reshape2::melt(id.vars = c("group", "ind","p_holder")) %>%
  group_by(group, ind, variable) %>% 
  filter(is.na(value)==F) %>% 
  summarize(l = quantile(value, 0.25, na.rm=T), 
    u = quantile(value, 0.75, na.rm=T), 
    m = quantile(value, 0.5, na.rm=T)) %>%
  ggplot(aes(x = ind, y = m, ymax = u, ymin = l)) + 
  geom_pointrange() + facet_grid(variable ~ group, scales="free")
```

```{r, fig.pos="placeHere", fig.cap = "Boxplots of revenue metrics (with interquartile range) by group and pre-post-IFQ. ", include=FALSE, echo=FALSE, message = FALSE, warning = FALSE}
#dplyr::select(hs.summary, group, ind, p_holder, contains("mean"), -contains("div")) %>% 
#  reshape2::melt(id.vars = c("group", "ind","p_holder")) %>%
#  ggplot(aes(ind, value)) + geom_boxplot(aes(fill=ind), alpha=0.5, outlier.size=NA) + 
#  facet_grid(variable ~ group, scales="free")
```

```{r, fig.pos="placeHere", fig.cap = "Boxplots of diversity metrics (with interquartile range) by group and pre-post-IFQ. "}
dplyr::select(hs.summary, group, ind, p_holder, contains("eff.spec.div.mean"), contains("eff.gear.div.mean"), contains("eff.spec.div.mean")) %>% 
  reshape2::melt(id.vars = c("group", "ind","p_holder")) %>%
  ggplot(aes(ind, value)) + 
  geom_boxplot(aes(fill=ind), alpha=0.5, outlier.size=NA) + 
  facet_grid(variable ~ group, scales="free")
# +  geom_point(alpha = 0.02, position = position_jitter(width = 0.4), cex = 0.2)

```

```{r, fig.pos="placeHere", fig.cap="Violin plot of individual differences (post-IFQ - pre-IFQ for the revenue and diversity metrics included in our analysis."}
hs.long = reshape2::melt(hs.summary, id.vars = c("group", "ind","p_holder"))
hs.long = reshape2::dcast(hs.long, variable + p_holder + group ~ ind) %>%
  mutate(diff_ = (Post-Pre)/Pre)

ggplot(hs.long, aes(group, diff_)) + 
  geom_violin(aes(fill=group), alpha=0.5) + 
  facet_wrap(~ variable, scales="free") +
  geom_point(alpha = 0.02, position = 
      position_jitter(width = 0.4), cex = 0.2)
```

## Changes in gear diversity through time  
Let's look at the change in gear diversity through time for each of the 4 groups in our analysis. 
We'll start with effective gear diversity, by

```{r, fig.pos = "placeHere", fig.cap = "Gear diversity through time for each of the 3 groups, calculated based on the catch (black) and earnings (red)."}
hs.summary.gear = group_by(dat, ind, group, year, gearn) %>% 
  summarize(n = n(), nEarn = sum(g_earn), nWeight = sum(g_pounds)) %>% 
  as.data.frame %>% 
  group_by(ind, group, year) %>% 
  summarize(gearDiv.freq = simp.div(n), 
    gearDiv.earn = simp.div(nEarn), 
    gearDiv.weight = simp.div(nWeight))

ggplot(hs.summary.gear, aes(year, gearDiv.weight)) + 
  geom_line() + facet_wrap(~ group, scales="free") + 
  geom_line(aes(year, gearDiv.earn), col="red")
```

We'll break down the proportion used for the top 8 gears,
```{r, fig.pos="placeHere", fig.cap = "Time series of proportion of gears used"}
hs.gear.tot = group_by(dat, ind, group, year) %>% 
  summarize(nTot = n())
hs.summary.gear = group_by(dat, ind, group, year, gearn) %>% 
  summarize(n = n()) %>%
  left_join(hs.gear.tot)
hs.summary.gear$p = hs.summary.gear$n/hs.summary.gear$nTot
hs.summary.gear = filter(hs.summary.gear, gearn %in% c(1,6,15,7,9,47,4,5)) %>% 
ggplot(aes(year, p)) + geom_line() + facet_grid(gearn ~ group, scales="free")

```

Can we predict effective species diversity as function of gear diversity?
```{r}
#hs.geardiv = group_by(dat, ind, group, p_holder, year, gearn) %>% 
#  summarize(tot.weight.sp = sum(g_pounds), 
#    tot.revenue.sp = sum(g_earn)) %>% 
#  group_by(ind, group, p_holder, year) %>% 
#  summarize(eff.gear.div.mean.earn = simp.div(tot.revenue.sp),
#    eff.gear.div.mean.weight = simp.div(tot.weight.sp)) %>%
#  group_by(ind, group, p_holder) %>% 
#  summarize(eff.gear.div.cv.earn = 
#      sd(eff.gear.div.mean.earn) / mean(eff.gear.div.mean.earn), 
#    eff.gear.div.mean.earn = mean(eff.gear.div.mean.earn), 
#    eff.gear.div.cv.weight = 
#      sd(eff.gear.div.mean.weight) / mean(eff.gear.div.mean.weight),
#    eff.gear.div.mean.weight = mean(eff.gear.div.mean.weight))
#hs.summary = left_join(hs.summary, hs.geardiv) %>% 
#  as.data.frame()
```

## EVOS comparison

We'll restrict our analysis initially to a few years before and after EVOS (1989). To make the comparison similar to the IFQ comparison, we will define the following groups:

1. Individuals who fished in closed areas before and after spill  
2. Individuals who fished in closed areas before spill ["Pre"]
3. Individuals who only fished in closed areas after 1995 ["Post"]  
4. Individuals who never fished in closed areas ["Control group"]

## Propensity matching for EVOS
We're interested in forming a control group for the comparison of the EVOS effects...
```{r, warning = FALSE, message=FALSE}

# Don't include 1989, but include before / after
dat = cfec[cfec$year %in% c(1985:1988, 1990:1991),]
dat$day = substr(dat$landdate, 6, 10)
# Create an indicator based on the salmon / shellfish araes that were closed in PWS
dat$exposed = 0
dat$exposed[dat$stat6%in%c(22710, 22720, 22730, 22610, 22620, 22630, 22640, 22650, 22510, 22520, 22530)] = 1

# Exposed variable is the response, and we want to predict based on traits
dat.ev = group_by(dat, p_holder, year) %>% 
  summarize(tot.revenue.ind = sum(g_earn), 
    tot.revenue.ind.her = sum(g_earn[specn==230]), 
    tot.revenue.ind.sal = sum(g_earn[specn%in%c(400,410,420,430,440,450)]),
    tot.pounds.ind = sum(g_pounds), 
    tot.pounds.ind.her = sum(g_pounds[specn==230]), 
    tot.pounds.ind.sal = sum(g_pounds[specn%in%c(400,410,420,430,440,450)]),
    eff.spec.div = simp.div(table(specn)), 
    eff.gear.div = simp.div(table(gearn)),
    numDays = length(unique(day)),
    exposed = max(exposed), 
    ind = ifelse(year[1] < 1989, "Pre", "Post")) %>% 
    as.data.frame() %>%
  group_by(p_holder, ind) %>% 
  summarize(tot.revenue.ind = mean(tot.revenue.ind), 
    tot.revenue.ind.her = mean(tot.revenue.ind.her), 
    tot.revenue.ind.sal = mean(tot.revenue.ind.sal),
    tot.pounds.ind = mean(tot.pounds.ind), 
    tot.pounds.ind.her = mean(tot.pounds.ind.her), 
    tot.pounds.ind.sal = mean(tot.pounds.ind.sal),
    eff.spec.div = mean(eff.spec.div), 
    eff.gear.div = mean(eff.gear.div),
    exposed = max(exposed),
    meanDays = mean(numDays)) %>% 
    as.data.frame()

# fix exposure column 
dat.ev = group_by(dat.ev, p_holder) %>% 
  mutate(exposed = max(exposed)) %>% 
  as.data.frame()

# 1. Diversity based on frequency in fish tickets  
# 2. Diversity weighted by sum of earnings  
# 3. Diversity weighted by sum of biomass (catch, in pounds)  
ev.div = group_by(dat, p_holder, year, specn) %>% 
  summarize(ind = ifelse(year[1] < 1989, "Pre", "Post"), 
    tot.weight.sp = sum(g_pounds), 
    tot.revenue.sp = sum(g_earn)) %>% 
  group_by(ind, p_holder, year) %>% 
  summarize(eff.spec.div.mean.earn = simp.div(tot.revenue.sp),
    eff.spec.div.mean.weight = simp.div(tot.weight.sp)) %>%
  group_by(ind, p_holder) %>% 
  summarize(eff.spec.div.cv.earn = 
      sd(eff.spec.div.mean.earn) / mean(eff.spec.div.mean.earn), 
    eff.spec.div.mean.earn = mean(eff.spec.div.mean.earn), 
    eff.spec.div.cv.weight = 
      sd(eff.spec.div.mean.weight) / mean(eff.spec.div.mean.weight),
    eff.spec.div.mean.weight = mean(eff.spec.div.mean.weight))
dat.ev = left_join(dat.ev, ev.div)

# Filter out only individulas with pre / post (2 rows)
dat.ev.count = group_by(dat.ev, p_holder) %>% 
  summarize(n = n()) %>%
  filter(n == 2) %>% 
  as.data.frame

dat.ev = inner_join(dat.ev, dplyr::select(dat.ev.count, p_holder))

# Only find matches based on pre-EVOS data. We can then compare pre/post effects

dat.ev$eff.spec.div.cv.earn = NULL
dat.ev$eff.spec.div.cv.weight = NULL

dat.ev.rf = filter(dat.ev, ind == "Pre")

#rf = randomForest(x = dat.ev.rf, y = as.factor(dat.ev.rf$exposed))

dat.ev.rf[,3:8] = log(dat.ev.rf[,3:8] + 1)

m.glm = glm(exposed ~ ., data = dat.ev.rf[dat.ev.rf$ind=="Pre", -c(1,2)], family = "binomial")
stepAIC(m.glm, direction="both")

bestModel = glm(exposed ~ tot.revenue.ind.her + tot.revenue.ind.sal + 
    tot.pounds.ind + tot.pounds.ind.her + tot.pounds.ind.sal + 
    eff.spec.div + meanDays + eff.spec.div.mean.weight, data = 
    dat.ev.rf[dat.ev.rf$ind == "Pre", ])

m.out <- matchit(exposed ~ tot.revenue.ind + tot.revenue.ind.her + 
    tot.revenue.ind.sal + tot.pounds.ind + tot.pounds.ind.her + 
    tot.pounds.ind.sal + eff.spec.div + eff.gear.div + meanDays + 
    eff.spec.div.mean.earn, method = "nearest", data = dat.ev.rf[dat.ev.rf$ind == 
    "Pre", ])

matches = match.data(m.out)

#m.out <- matchit(exposed, method = "nearest", data = dat.ev[dat.ev$ind==0,], distance = rf$votes[,2])

dat.matched = dat.ev[dat.ev$p_holder %in% unique(matches$p_holder), ]

fn_bal <- function(dta, variable) {
    dta$variable <- dta[, variable]
    ggplot(dta, aes(y = distance, x = variable, color = as.factor(exposed))) +
      geom_point(alpha = 0.2, size = 1.5) +
      geom_smooth(method = "loess", se = F) +
      ylab("Propensity score") +
      xlab(variable) + 
      theme_bw()  
}
fn_bal(matches, "tot.revenue.ind")

dat.matched$plotind = ifelse(dat.matched$ind == "Pre", 0, 1)
dat.matched$ind2 = dat.matched$plotind
dat.matched$ind2[dat.matched$exposed==1] = dat.matched$ind2[dat.matched$exposed==1] + 0.2

dat.matched.summary = reshape2::melt(dat.matched, id.vars = c("exposed","ind", "p_holder")) %>%
  group_by(ind, exposed, variable) %>% 
  filter(is.na(value)==F) %>% 
  summarize(l = quantile(value, 0.25, na.rm=T), 
    u = quantile(value, 0.75, na.rm=T), 
    m = quantile(value, 0.5, na.rm=T))
dat.matched.summary$ind2 = as.numeric(as.factor(dat.matched.summary$ind)) - 1
dat.matched.summary$ind2[dat.matched.summary$exposed==1] = dat.matched.summary$ind2[dat.matched.summary$exposed==1] + 0.2

ggplot(dat.matched.summary, aes(x = ind2, y = m, ymax = u, ymin = l, col = as.factor(exposed))) + geom_pointrange() + facet_wrap(~ variable, scales="free")

```

```{r}

ev.long = reshape2::melt(dat.matched, id.vars = c("exposed","ind","p_holder"))
ev.long = reshape2::dcast(ev.long, variable + p_holder + exposed ~ ind) %>%
  mutate(diff_ = (Post-Pre)/Pre)

ev.long = filter(ev.long, !is.na(diff_), is.finite(diff_))
ggplot(ev.long, aes(exposed, diff_)) + 
  geom_violin(aes(fill=exposed), alpha=0.5) + 
  facet_wrap(~ variable, scales="free") +
  geom_point(alpha = 0.02, position = 
  position_jitter(width = 0.4), cex = 0.2)

```

```{r}
#cfec$areaPermit = 0
#cfec$areaPermit[grep("B", cfec$p_fshy)] = 1
#cfec$areaPermit[grep("G", cfec$p_fshy)] = 1

#permitTypes = group_by(cfec, p_holder) %>% 
#summarize(nState = length(which(areaPermit==1)), n = length(which(areaPermit==0))) %>% as.data.frame()

# 41113 individuals
# 18383 just have regional permits, 12251 just have statewide/GOA, 10479 have both
#length(which(permitTypes$nState==0 & permitTypes$n > 0))
```

### 2016-03-28 catch_portfolios_mtg3.Rmd
updated Rmd file to include filtering example

@@
 library(MASS)
+library(zoo)
 library(MatchIt)
@@
 #}')
+```
+# Paper 1:
+
+To start, we can coarsely summarize the variability in total revenue by species through time, 
+```{r, fig.pos="placeHere", fig.cap = "Log(total) revenue of each spp"}
+spTot = group_by(cfec, specn, year) %>%
+  summarize(totRev.bySpec = log(sum(g_earn, na.rm=T))) %>% 
+  as.data.frame() %>%
+  filter(is.finite(totRev.bySpec)) %>%   
+  as.data.frame
+ggplot(spTot, aes(x=year,y=totRev.bySpec,group=as.factor(specn))) +geom_line(aes(x=year,y=totRev.bySpec, color=as.factor(specn)))
+```
+
+```{r, fig.pos="placeHere", fig.cap = "CV of log(total revenue) across species (sd in log space)"}
+spTot = group_by(spTot, year) %>%
+  summarize(sd_yr = sd(totRev.bySpec, na.rm=T)) %>% 
+  as.data.frame()
+ggplot(spTot, aes(x=year,y=sd_yr)) + geom_point()
+
+```
+
+## Filtering out unique vessel - person combinations  
+Filter out people who only use 1 vessel / year , and similarly filter out boats with more than 1 person / year.
+Some people might fish on multiple vessels within a year, and vessels might be used by multiple people within a year.
+
+For each individual in each year, we can calculate the effective number of vessels used, and trends in that (mean across people).   
+```{r}
+# calculate diversity of individuals across vessels / time
+cfec.boatDiv = group_by(cfec, p_holder, year) %>% 
+  summarize(effBoats = 1 / sum((table(cadfg) / sum(table(cadfg)))^2) ) %>% 
+  data.frame
+
+cfec.boatDiv = as.data.frame(cfec.boatDiv)
+g = gam(as.numeric(effBoats) ~ s(as.numeric(year)), data=cfec.boatDiv)
+gpred = predict(g, newdata=list(year=1985:2014), se.fit=TRUE)
+plot(1985:2014, gpred$fit, type="l", lwd=3, ylim=c(1.03,1.1))
+lines(1985:2014, gpred$fit+2*gpred$se.fit)
+lines(1985:2014, gpred$fit-2*gpred$se.fit)
+```
+
+Now we'll filter out individuals who only used 1 boat in a year, and boats with only one permit holder fishing in a year.  
+```{r}
+cfecnew = cfec
+cfec.1boat = group_by(cfecnew, p_holder, year) %>% 
+  summarize(nBoat = length(unique(cadfg))) %>%
+  filter(nBoat==1) %>% 
+  left_join(cfecnew)
+
+# Deal with setnet fisheries, where lots of p_holder
+# share same codes
+cfec.1boat = group_by(cfec.1boat, cadfg, year) %>% 
+  summarize(nPerson = length(unique(p_holder))) %>%
+  filter((nPerson==1 & cadfg %in% c(27960,27961) == FALSE) | cadfg %in% c(27960,27961)) %>% 
+  left_join(cfec.1boat)
+
+cfecnew = cfec.1boat
+dim(cfecnew)[1]
+```
+
+So there's about 50% of the records remaining -- these are unique individial / vessel combinations.
+Side note: all of the set net / razor clam permits and individuals have been included.
+Not sure we want to drop them out [Milo - thoughts?
+]
+
+Now we can create a unique *id* column for individual - vessel combinations.
+
+```{r}
+cfecnew$id = paste(cfecnew$p_holder,cfecnew$cadfg)
+# calculate number of unique person - vessel pairs ~ 40700
+length(unique(cfecnew$id))
+```
+
+Just for demo purposes, we can start by plotting mean - CV relationships for each of these individual- vessel combinations (across years).
+
+```{r, fig.pos="placeHere", fig.cap="Mean and variability (CV) of annual earnings (across time) for unique people-vessel combinations."}
+meanVar = group_by(cfecnew, id, year) %>% 
+  summarize(totRev = sum(g_earn)) %>% 
+  as.data.frame() %>%
+  group_by(id) %>%
+  summarize(m = mean(log(totRev+1),na.rm=T), cv = sd(totRev,na.rm=T)/mean(totRev,na.rm=T))
+plot(meanVar$m, meanVar$cv, xlab="Mean log(annual revenue)", ylab = "CV annual revenue")
+```
+
+One question we also were chatting about was what the distribution of the percent change metric looked like. 
+```{r mergeYears}
+
+totalcfec = group_by(cfecnew, id, year) %>% 
+  summarize(totrev = sum(g_earn))
+
+totalcfec$year = as.numeric(totalcfec$year)
+dummyDF = expand.grid(id = unique(totalcfec$id), 
+  year = min(totalcfec$year) : max(totalcfec$year), 
+  stringsAsFactors = F)
+# expand the data frame w/missing years
+cfecExp = left_join(dummyDF, totalcfec) %>%
+  arrange(id, year)
+cfecExp = group_by(cfecExp, id) %>%
+  mutate(difflog = c(NA, diff(log(totrev))), meanRev = c(NA, zoo::rollmean(totrev, 2))) %>%
+  filter(!is.na(difflog))
+```
+
+This shows the pattern we'd hope to expect -- no trend as a function of mean revenue, .
+Namely that people making less than about 35K / year tend to have negative returns from one year to the next, and people with revenues > 35K -- especially the high earners -- tend to have positive returns.
+We also see decreased variability as a function of returns -- maybe because investments in technology, permits, etc.
+```{r, fig.pos="placeHere", fig.cap = "Distribution of percent change in total revenue + relationship (or lack of) with total annual revenue"}
+par(mfrow = c(2,1), mai=c(0.5,0.5,0.1,0.1))
+plot(log(cfecExp$meanRev), cfecExp$difflog, xlab = "log(revenue)", ylab="percent change in revenue", cex=0.7)
+abline(0, 0, col="red")
+
+hist(cfecExp$difflog, 80, xlab="Percent return (empirical)", col="grey", xlim= quantile(cfecExp$difflog, c(.005,0.995)), main="")
+
 ```

### 2016-04-07 process_raw_data/process_raw_cfec_data.Rmd
Revised species and vessel info

@@
 author: "Jordan Watson & Eric Ward"
-date: "March 16, 2016"
+date: "March 16, 2016 (Update 4/7/2016)""
 output: pdf_document
@@
 library(ggplot2)
-library(mgcv)
+#library(mgcv)
 knitr::opts_chunk$set(echo = TRUE)
@@
 ```{r dataload, warnings=FALSE, message = FALSE}
+setwd("//nmfs.local/AKC-ABL/Users2/jordan.watson/Desktop/AFSC/GOA/AKFIN/Meeting")
 load("Data/cfec_w_IDs.RData")
@@
 "F_ITEM", "F_FSHERY",  "G_GEAR", "G_AREA", "G_PRC_AR",  "P_YEAR", "P_PMTSEQ", 
-"p_check", "P_ID", "P_FEETYP", "P_ADFG", "P_FEEFLG", "P_FSHBLE", "F_DISPOS", 
+"p_check", "P_ID", "P_FEETYP", "P_ADFG", "P_FEEFLG", "P_FSHBLE", 
 "F_MPGMNM", "D_OPER", "D_O_FSHY", "D_O_PNUM", "D_O_PCHK", "D_O_YSEQ", "F_SIZE", 
@@
 
-table(cfec$P_STATUS)
+#table(cfec$P_STATUS)
 ```
@@
 # Lump all sharks together
+cfec$specn[cfec$specn==689] = 689
+cfec$specn[cfec$specn==690] = 689
 cfec$specn[cfec$specn==691] = 689
-cfec$specn[cfec$specn==689] = 689
 cfec$specn[cfec$specn==692] = 689
@@
 cfec$specn[cfec$specn==126] = 119 # butter
+cfec$specn[cfec$specn==122] = 119 # flathead
+cfec$specn[cfec$specn==125] = 119 # rex
+cfec$specn[cfec$specn==128] = 119 # English
+cfec$specn[cfec$specn==131] = 119 # petrale
+cfec$specn[cfec$specn==132] = 119 # sand
+cfec$specn[cfec$specn==133] = 119 # Alaska plaice
+cfec$specn[cfec$specn==116] = 119 # Bering flounder
+cfec$specn[cfec$specn==117] = 119 # Kamchatka flounder
 
@@
 # Slope rockfish
+cfec$specn[cfec$specn==136] = 144 # northern
+cfec$specn[cfec$specn==137] = 144 # bocaccio
 cfec$specn[cfec$specn==141] = 144 # POP
-cfec$specn[cfec$specn==136] = 144 # northern
+cfec$specn[cfec$specn==143] = 144 # thornyhead
+cfec$specn[cfec$specn==144] = 144 # unid
 cfec$specn[cfec$specn==151] = 144 # rougheye
-cfec$specn[cfec$specn==166] = 144 # sharpchin
 cfec$specn[cfec$specn==152] = 144 # shortraker
@@
 cfec$specn[cfec$specn==158] = 144 # redstripe
-cfec$specn[cfec$specn==137] = 144 # bocaccio
-cfec$specn[cfec$specn==144] = 144 # unid
+cfec$specn[cfec$specn==159] = 144 # darkblotched
+cfec$specn[cfec$specn==166] = 144 # sharpchin
+cfec$specn[cfec$specn==176] = 144 # harlequin
+cfec$specn[cfec$specn==177] = 144 # blackgill
+cfec$specn[cfec$specn==179] = 144 # pygmy
+cfec$specn[cfec$specn==181] = 144 # shortbelly
+cfec$specn[cfec$specn==182] = 144 # splitnose
+cfec$specn[cfec$specn==183] = 144 # stripe tail
+cfec$specn[cfec$specn==184] = 144 # vermillion
+cfec$specn[cfec$specn==185] = 144 # aurora
+
 # Demersal rockfish
+cfec$specn[cfec$specn==135] = 168 # greenstripe
 cfec$specn[cfec$specn==138] = 168 # copper
+cfec$specn[cfec$specn==139] = 168 # unid
+cfec$specn[cfec$specn==140] = 168 # red
 cfec$specn[cfec$specn==145] = 168 # yelloweye
@@
 cfec$specn[cfec$specn==150] = 168 # rosethorn
+cfec$specn[cfec$specn==171] = 168 # shortraker/rougheye
+cfec$specn[cfec$specn==175] = 168 # yellowmouth
+cfec$specn[cfec$specn==178] = 168 # chilipepper
+
 # Pelagic rockfish
+cfec$specn[cfec$specn==142] = 169 # black
 cfec$specn[cfec$specn==154] = 169 # dusky
@@
 cfec$specn[cfec$specn==156] = 169 # widow
+cfec$specn[cfec$specn==167] = 169 # blue
 cfec$specn[cfec$specn==172] = 169 # dusky
 cfec$specn[cfec$specn==173] = 169 # dark
-cfec$specn[cfec$specn==142] = 169 # black
-cfec$specn[cfec$specn==167] = 169 # blue
+
+# Group sculpins
+cfec$specn[cfec$specn %in% c(160:165)] = 160
+
+# Group greenlings (not AMCK or lingcod)
+cfec$specn[cfec$specn %in% c(190,191,192,194)] = 190
+
+# Group all osmerids together
+cfec$specn[cfec$specn %in% c(510:516)] <- 510
+
+# Group all char
+cfec$specn[cfec$specn %in% c(520:532)] <- 520  # char
+
+# Group all trout
+cfec$specn[cfec$specn%in%c(540:562)] <- 540  # trout
 
@@
 cfec$specn[cfec$specn==589] <- 580  # Humpback whitefish
-# Trout - convert all to the same number - we'll label as "trout" later
-cfec$specn[cfec$specn==520] <- 520  # Arctic char
-cfec$specn[cfec$specn==530] <- 520  # Dolly varden
-cfec$specn[cfec$specn==531] <- 520  # Dolly varden anadromous
-cfec$specn[cfec$specn==540] <- 520  # Steelhead
-cfec$specn[cfec$specn==560] <- 520  # Cutthroat
 
@@
 
-# Salmon roe (by species)
-# We consolidate roe of different species into a single roe category
-cfec$specn[cfec$specn %in% c(400:405)] <- 400
-# In early years, there were separate specn for salmon roe and salmon. In later years,
-# salmon roe got absorbed into the specn for salmon but given a F_DELIVR==14.
-# We may want to change this at some point, but for now, if F_DELIVR==14, then change
-# the specn to that for roe instead of the specn "salmon."
-cfec$specn[(cfec$specn %in% c(410,420,430,440,450)) & cfec$F_DELIVR=="14"] <- 400
+# Chinook have two species codes. Consolidate.
+cfec$specn[cfec$specn==411] <- 410
+
+# Salmon roe (by species) get lumped into a single salmon roe category.
+This was a topic of much discussion but ultimately, since much of the roe is unidentified, any other assignments would end up omitting a large percentage of the roe landings.
+cfec$specn[cfec$specn %in% c(401:405)] <- 400
+cfec$specn[cfec$specn%in%c(410:450) & cfec$F_DELIVR==14] <- 400
+
+# Herring are listed under 5 different species codes but 231:235 only account for a few thousand bucks. They are mostly bycatch or non-commercial.
+# We'll lump them into the sac-roe species code because they will have no effect on the numbers. But this way we are consolidating codes instead
+# of making a bunch of extra ones.
+cfec$specn[cfec$specn %in% c(231:235)] <- 23043
+
+# Most herring landings occur under species 230, with different F_DELIVR codes
+# Create a species code for the sac-roe fishery
+cfec$specn[cfec$specn==230 & cfec$F_DELIVR==43] <- 23043
+# Create a species code for the bait/food fishery
+cfec$specn[cfec$specn==230 & cfec$F_DELIVR==44] <- 23044
+# Create a species code for the pound fishery
+cfec$specn[cfec$specn==230 & cfec$F_DELIVR==45] <- 23045
+# There may be 1 or 2 records with other codes, and which probably have no dollar value.
+# Lump them into the sac roe fishery to avoid having extra codes
+cfec$specn[cfec$specn==230] <- 23043
+
+# Lump all king crabs
+cfec$specn[cfec$specn %in% c(920:925)] <- 920
+
+# Lump all king crabs
+cfec$specn[cfec$specn %in% c(930:934)] <- 930
+
+# Lump all remaining crabs (not Dungeness)
+cfec$specn[cfec$specn %in% c(900,940:953)] <- 940
+
+# Lump all shrimp
+cfec$specn[cfec$specn %in% c(960:965)] <- 960
+
+# Lump non-geoduck and non-razor clams together
+cfec$specn[cfec$specn %in% c(810,812,820,825,840,842)] <- 810
+
+# Lump weathervane and pink scallops
+cfec$specn[cfec$specn %in% c(850:851)] <- 850
+
+# Lump urchins
+cfec$specn[cfec$specn %in% c(892:893,896)] <- 892
+
+# Lump sea cucumbers
+cfec$specn[cfec$specn %in% c(894:895)] <- 895
+
+# Create an "OTHER" category that includes misc freshwater and a bunch of random trawl caught spec
+cfec$specn[cfec$specn %in% c(100,101,112,170,180,190,206:220,250,260,310:380,490,500,520,540,570,580,590,600:683,714:800,880,890,855,899,998,999)] <- 206
 
@@
 cfec$spec <- as.character(cfec$spec)
-cfec$spec[cfec$specn==700] = "SKATE"
-cfec$spec[cfec$specn==689] = "SHARK"
+
+#  Rename some of the groups
+cfec$spec <- as.character(cfec$spec)
+cfec$spec[cfec$specn==118] = "FLAT.deep"
 cfec$spec[cfec$specn==119] = "FLAT.shallow"
-cfec$spec[cfec$specn==118] = "FLAT.deep"
 cfec$spec[cfec$specn==144] = "ROCK.slope"
@@
 cfec$spec[cfec$specn==169] = "ROCK.pelagic"
-cfec$spec[cfec$specn==520] = "TROUT"
-
-# Lump king crabs
-cfec$spec[cfec$spec%in%c("GKCR","BKCR","UKCR","RKCR")] = "KCRB" # king
-# Lump tanner crabs
-cfec$spec[cfec$spec%in%c("UTCR","BTCR","aTCGR","aTCTR")] = "TCRB" # tanner
-# Lump shrimp - spot prawn, pink shrimp, coonstriped 
-cfec$spec[cfec$spec%in%c("SPRW","PSHP","USRM","aSSHP","aCOOS")] = "USRM" # shrimp
+cfec$spec[cfec$specn==206] = "OTHER"
+cfec$spec[cfec$specn==400] = "aUROE"
+cfec$spec[cfec$specn==689] = "SHARK"
+cfec$spec[cfec$specn==700] = "SKATE"
+cfec$spec[cfec$specn==810] = "OCLM"
+cfec$spec[cfec$specn==850] = "SCAL"
+cfec$spec[cfec$specn==892] = "URCH"
+cfec$spec[cfec$specn==920] = "KCRB"
+cfec$spec[cfec$specn==930] = "TCRB"
+cfec$spec[cfec$specn==940] = "OCRB"
+cfec$spec[cfec$specn==960] = "USRM"
+cfec$spec[cfec$specn==23043] = "SAC"
+cfec$spec[cfec$specn==23044] = "BAIT"
+cfec$spec[cfec$specn==23045] = "POUND"
 
@@
 cfec = cfec[-which(is.na(cfec$spec)),]
+rm(spec.tbl)
 ```
@@
 
+cfec <- as.data.frame(cfec)   
+```
+
+Add vessel lengths to data
+```{r}
+#  Load the vessel characteristics data that were created using "Vessel_Characteristics.Rmd."
+load("Data/vessel.RData")
+
+#  For now, we only really need the vessel length data but different/additional fields can be added later by changing the select statement below.
+Note, I use "dplyr::select" instead of just "select" because if you have the plyr package loaded also, the select statement may get confused.
+
+#  Convert fields to make them compatible with the vessel data.
+cfec$cadfg2 <- as.numeric(cfec$cadfg)
+cfec$year <- as.numeric(cfec$year)
+
+#  I use a right join here instead of left join for continuity since "vessel" was the already loaded data frame which automatically puts it as the first argument of the piped command.
+cfec <- dplyr::select(vessel,year,adfg,VLength) %>% right_join(cfec,by=c("adfg"="cadfg2","year"="year")) %>% rename(cadfg2=adfg)
+
+#  Create an index value so that we can link back to the original (unscrambled data) if necessary.
+cfec$INDEX <- 1:nrow(cfec)
+
 ```
@@
 ```{r}
-save(cfec, file = "cfecCleaned.Rdata")
+save(cfec, file = "cfecCleaned2.Rdata")
 ```
@@
 all halibut 200-204 became 200, or see rockfish groups above).
+
+*specn.orig* The original version of the specn field before species were lumped into categories.
+
+*spec* A text code created for each species. Some of these are the original AKFIN names and others were created by us. See the Species_Categories.xls file for descriptions.
 

### 2016-04-13 Create_Summary_Tables_AllFishingActivities.Rmd
Added files via upload

---
title: "Permits, permit holders, vessels 2"
author: "Jordan Watson"
date: "March 9, 2016"
output:
  pdf_document:
    fig_caption: yes
    latex_engine: xelatex
  html_document: default
---

```{r chunkSet, cache=FALSE, echo=FALSE,warning=FALSE,message=FALSE}
library(dplyr)
library(ggplot2)
library(gridExtra)
library(knitr)
library(reshape2)
library(vegan)
#setwd("~/Desktop/NCEAS")
setwd("//nmfs.local/AKC-ABL/Users2/jordan.watson/Desktop/AFSC/GOA/AKFIN/Meeting")

opts_chunk$set(tidy=TRUE, message=FALSE)
```

```{r loadData,warning=FALSE,message=FALSE}

load("Data/cfecCleaned_new.RData")

# initially only use data 1985+
cfec = cfec[cfec$year > 1984, ]
cfec <- rename(cfec,adfg=cadfg) %>% data.frame
cfec$year <- as.numeric(cfec$year)

#  Read-in the CPI adjusted for 2009 base dollars. 
cpi <- read.csv("Data/CPI_2009.csv") %>% filter(year>1984)
cfec <- left_join(cfec,cpi) 
#  Adjust g_earn and g_price to 2009 dollars
cfec[,c("g_earn","g_price")] <- cfec[,c("g_earn","g_price")] / cfec$GDPDEF09_base1
cfec$GDPDEF09_base1 <- NULL
rm(cpi)
```

```{r}
#  Get rid of the spaces in p_fshy
cfec$p_fshy <- gsub(" ","",cfec$p_fshy)

#  Read in the fishing locations by fishery permit (p_fshy)
p_fshy <- read.csv("Data/new.locs.csv")
p_fshy$p_fshy <- as.character(p_fshy$p_fshy)

#  Combine these areas (original and our coarse new area) to the data
cfec <- left_join(cfec,p_fshy)

#  Now for gears...
#  Read in gear description file and link with cfec.
gears <- read.csv("Data/new_gear_cat.csv")
cfec <- left_join(cfec,gears[,c("gearn","Gear")])

#  Troll fisheries only happen in Southeast
cfec$New.Loc[cfec$New.Loc=="Statewide" & cfec$Gear=="Troll"] <- "Southeast"
#  Abalone were only ever caught in southeast
cfec$New.Loc[cfec$New.Loc=="Statewide" & cfec$spec=="ABLN"] <- "Southeast"
```

```{r}
#  Create a temporary (ie., smaller) data frame for calculating with only the necessary fields.
temp2 <- select(cfec,-F_REGION,-F_INPFC,-g_price,-P_TYPE,-P_STATUS,-A_RES,-g_price,-startdt,-landdate,harvest,F_PERMIT)  %>% filter(g_earn>0 & !is.na(Gear) & !is.na(New.Loc))

#  Identify the earliest fishing year for each p_holder
startyear <- group_by(temp2,p_holder) %>% summarise(start=min(year))
#  Join that year to the data
temp2 <- left_join(temp2,startyear)
rm(startyear)

#  Define custom function for calculating the mode
Mode <- function(x) {
     ux <- unique(x)
     ux[which.max(tabulate(match(x, ux)))]
}

#  Calculate initial metrics for each p_holder and year
xx <- group_by(temp2,p_holder,year) %>% 
              summarise(StartYear=start[1],
              permits=length(unique(p_serial)),# Number of permits
              permitTypes=length(unique(p_fshy)),# Number of permit types
              landings=length(unique(f_ticket)), # Number of fish tickets (proxy for # of trips)
              boats=length(unique(adfg)),# Number of boats fished by p_holder
              areaSm=length(unique(Location)), # Number of permit areas held (N=51)
              areaLg=length(unique(New.Loc)), # Number of our aggregated permit areas held (N=13)
              statareas=length(unique(stat6)), # Number of different areas from which fish were reported to have been caught
              vlength=Mode(VLength)
              ) %>% data.frame

rm(gears,p_fshy,Mode)
```

### We now begin to examine Revenues as a function of fishing area and gear for each p_holder and year

#### Aggregate revenue by area and gear for each p_holder and year
```{r}
#temp.dollars <- group_by(temp2,p_holder,year,Gear,New.Loc) %>% summarise(Dollars=sum(g_earn))
#
#for(i in 1:length(unique(temp.dollars$Gear))){     
#  print(ggplot(temp.dollars[temp.dollars$Gear==paste(unique(temp.dollars$Gear)[i]),],aes(x=factor(year),y=log(Dollars))) + geom_boxplot() + facet_wrap(~New.Loc) + ggtitle(paste(unique(temp.dollars$Gear)[i])))
#}
```

```{r}
#  Tally the dollars of fish by each p_holder during each year for each gear type
#  First in long format
tempgear.d <- group_by(temp2,p_holder,year,Gear) %>% summarise(Dollars=sum(g_earn))
#  Now in wide format, with zeros filled in.
tempgearwide.d <- dcast(tempgear.d,p_holder+year~Gear,value.var="Dollars")
tempgearwide.d[is.na(tempgearwide.d)] <- 0

#  Tally the dollars of fish by each p_holder during each year for each permit location
#  First in long format
temparea.d <- group_by(temp2,p_holder,year,New.Loc) %>% summarise(Dollars=sum(g_earn))
#  Now in wide format, with zeros filled in.
tempareawide.d <- dcast(temparea.d,p_holder+year~New.Loc,value.var="Dollars")
tempareawide.d[is.na(tempareawide.d)] <- 0

#  Tally the dollars of fish by each p_holder during each year for each species
#  First in long format
tempspec.d <- group_by(temp2,p_holder,year,spec) %>% summarise(Dollars=sum(g_earn))
tempspecwide.d <- dcast(tempspec.d,p_holder+year~spec,value.var="Dollars")
#  Now in wide format
tempspecwide.d[is.na(tempspecwide.d)] <- 0

# Join each of the above wide format tables. In joining, values of g_earn from different tables will become redundant so do not use these dollar values.   
tempjoin.dollars <- inner_join(xx,tempareawide.d) %>% inner_join(tempgearwide.d) %>% inner_join(tempspecwide.d)
# Convert the dollar values to dummy values
tempjoin.dollars[,-c(which(names(tempjoin.dollars) %in% colnames(xx)))] <- lapply(tempjoin.dollars[,-c(which(names(tempjoin.dollars) %in% colnames(xx)))], function(x) replace(x,x!=0, 1))
# Join the total annual revenue for each p_holder with the above table
revenueMetrics <- group_by(temp2,p_holder,year) %>% summarise(totalRev=round(sum(g_earn)),
                                                              medianRev=round(median(g_earn)),
                                                              cvRev=round(sd(g_earn)/mean(g_earn)))  %>% 
                                                              inner_join(tempjoin.dollars) %>% 
                                                              mutate(yearsFishing=(year-StartYear)+1) %>%
                                                              select(p_holder,year,yearsFishing,everything()) %>% data.frame

# For plotting purposes the long format will be more useful
longjoin.dollars <- inner_join(xx,temparea.d) %>% inner_join(tempgear.d) %>% inner_join(tempspec.d) %>% mutate(yearsFishing=(year-StartYear)+1)
```

Now do the same thing but aggregate the pounds landed (instead of the revenue) landed by area, gear,species for each p_holder and year
```{r}

#  Tally the pounds of fish by each p_holder during each year for each gear type
#  First in long format
tempgear.p <- group_by(temp2,p_holder,year,Gear) %>% summarise(Mass=sum(g_pounds))
#  Now in wide format, with zeros filled in.
tempgearwide.p <- dcast(tempgear.p,p_holder+year~Gear,value.var="Mass")
tempgearwide.p[is.na(tempgearwide.p)] <- 0

#  Tally the pounds of fish by each p_holder during each year for each permit location
#  First in long format
temparea.p <- group_by(temp2,p_holder,year,New.Loc) %>% summarise(Mass=sum(g_pounds))
#  Now in wide format, with zeros filled in.
tempareawide.p <- dcast(temparea.p,p_holder+year~New.Loc,value.var="Mass")
tempareawide.p[is.na(tempareawide.p)] <- 0

#  Tally the pounds of fish by each p_holder during each year for each species
#  First in long format
tempspec.p <- group_by(temp2,p_holder,year,spec) %>% summarise(Mass=sum(g_pounds))
tempspecwide.p <- dcast(tempspec.p,p_holder+year~spec,value.var="Mass")
#  Now in wide format
tempspecwide.p[is.na(tempspecwide.p)] <- 0

# For plotting purposes the long format will be more useful
longjoin.pounds <- inner_join(xx,temparea.p) %>% inner_join(tempgear.p) %>% inner_join(tempspec.p) %>% mutate(yearsFishing=(year-StartYear)+1)

tempjoin.pounds <- inner_join(xx,tempareawide.p) %>% inner_join(tempgearwide.p) %>% inner_join(tempspecwide.p)
tempjoin.pounds[,-c(which(names(tempjoin.pounds) %in% colnames(xx)))] <- lapply(tempjoin.pounds[,-c(which(names(tempjoin.pounds) %in% colnames(xx)))], function(x) replace(x,x!=0, 1))

# Join the total annual revenue for each p_holder with the above table
landedMetrics <- group_by(temp2,p_holder,year) %>% summarise(totalPounds=round(sum(g_pounds)),
                                                              medianPound=round(median(g_pounds)),
                                                              cvPounds=round(sd(g_pounds)/mean(g_pounds)))  %>% 
                                                              inner_join(tempjoin.pounds) %>% 
                                                              mutate(yearsFishing=(year-StartYear)+1) %>%
                                                              select(p_holder,year,yearsFishing,everything()) %>% data.frame

```

Combine metrics into a single table

```{r}

AllMetrics.wide <- inner_join(revenueMetrics,landedMetrics) %>% select(p_holder,StartYear,yearsFishing,vlength,permits:statareas,totalRev:cvRev,totalPounds:cvPounds,everything())
AllMetrics.long <- inner_join(longjoin.pounds,longjoin.dollars) %>% select(p_holder,StartYear,yearsFishing,vlength,permits:New.Loc,Gear,spec,Mass,Dollars)

temparea <- inner_join(temparea.p,temparea.d)
tempgear <- inner_join(tempgear.p,tempgear.d)
tempspec <- inner_join(tempspec.p,tempspec.d)

rm(xx,cfec,temp2,revenueMetrics,landedMetrics,longjoin.pounds,longjoin.dollars,tempjoin.dollars,tempjoin.pounds,tempspec.p,tempspec.d,temparea.p,temparea.d,tempgear.p,tempgear.d)
save.image("cfecModelsetup.RData")
```

## Metadata - what do all of the different objects mean?
For modeling, the object most usefull will likely be "AllMetrics.wide"

*AllMetrics.long* contains combined characteristics for each p_holder and a single column for all species, a single column for all gears, a single column for all areas (New.Loc), a single column for pounds landed, and a single column for dollars earned.

*AllMetrics.wide* contains combined characteristics for each p_holder and a different column for each species, a different column for each gear, a different column for each areas (New.Loc), a single column total,median and cv of revenue, a single column for total, median, and cv of pounds landed.

*temparea* contain dollars & pounds for each p_holder by year and by area (New.Loc) - all areas are in a single column

*tempareawide.d* / *tempareawide.p* contain dollars / pounds for each p_holder by year and by area (New.Loc) - each areas is in a different column

*tempgear* contain dollars & pounds for each p_holder by year and by gear type - all gears are in a single column

*tempgearwide.d* / *tempgearwide.p* contain dollars / pounds for each p_holder by year and by gear - each gear is in a different column

*tempspec* contain dollars & pounds for each p_holder by year and by species - all species are in a single column

*tempspecwide.d* / *tempspecwide.p* contain dollars / pounds for each p_holder by year and by species - each species is in a different column

### 2016-04-13 Create_Summary_Tables_AllFishingActivities.Rmd
Update Create_Summary_Tables_AllFishingActivities.Rmd

@@
 ---
+
+## This file summarizes the cfec catch data into a few different objects.
+These objects contain p_holder annual summaries of landed pounds and earnings by gear type, area fished and by species.
+Both wide and long formats are provided for easier modeling / plotting.
+The output of this file is an RData image called "cfecModelsetup.RData."
 

### 2016-06-09 analysis_diversity/paper1a_diversity_metrics.Rmd
committing updates to all papers

---
title: "Changes in diversity and variability in diversity through time: Catch portfolios paper 1A"
date: "April 7, 2016"
output:
  pdf_document:
    fig_caption: yes
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=FALSE, tidy.opts=list(width.cutoff=60), warning = FALSE, message = FALSE)
```

## Load cleaned data  
```{r dataLoad, warning = FALSE, message=FALSE, results="hide"}
library(dplyr)
library(knitr)
library(mgcv)
library(ggplot2)
library(randomForest)
library(MASS)
library(zoo)
library(MatchIt)
library(metafolio)
load("cfecCleaned2.Rdata")
```

We need to adjust the landed values and prices for inflation. To do this, we'll use the data Alan pulled together (adjusting to 2009 dollars).

https://docs.google.com/spreadsheets/d/1hRLYTX1nZBLKnYXwTCOtU7-XnDoMZf_1ufdTiNQFd4U/edit#gid=0

```{r}
deflationTable = data.frame(year = 1985:2014, defl = c(0.57236, 0.58393, 0.59879, 0.61974,  
0.64388, 0.66774, 0.68993, 0.70564, 0.72244, 0.73781,  
0.75321, 0.76695, 0.78009, 0.78855, 0.80061,  
0.81883, 0.83753, 0.85038, 0.86729, 0.89114,  
0.91981, 0.94812, 0.97334, 0.9925, 1,  
1.01217, 1.03307, 1.05213, 1.06926, 1.08682))

# Adjust price and g_earn for inflation
cfec$year = as.numeric(cfec$year)
cfec$day = substr(cfec$landdate, 6, 10)
cfec$g_price = cfec$g_price/deflationTable$defl[cfec$year-1984]
cfec$g_earn = cfec$g_earn/deflationTable$defl[cfec$year-1984]
```

### Function definiton  
We need to define a few functions for calculating diversity. We'll use the Simpson's diversity index (HHH) to calculate the entropy-based inverse, which is a proxy for effective species diversity. Note the C++ version is commented out below,   
```{r, warning = FALSE, message=FALSE}

simp.div = function(x) {
  1/sum((x/sum(x))^2)
}

#library(Rcpp)
#cppFunction('double simpCpp(NumericVector x) {
#  double y = 1/sum((x/sum(x))^2);
#  return y;
#}')

# Note that the first line of this function uses group_by_ because 
# we need to pass in a string for 'variable'
effectiveDiversity_by_personYear = function(dataFrame, variable) {
  effDiv = group_by_(dataFrame, variable, "year", "p_holder") %>%
  summarize(totRev = sum(g_earn, na.rm=T),
  totN = n(),
  totWeight = sum(g_pounds, na.rm=T),
    days = length(unique(day))) %>% 
  as.data.frame() %>%
  filter(is.finite(totRev) & is.finite(totN) & is.finite(totWeight)) %>%   
  as.data.frame() %>%
  group_by(p_holder, year) %>% 
  summarize(eff.freq = simp.div(totN),
  eff.earn = simp.div(totRev),
  eff.lbs = simp.div(totWeight),
    totIndRev = sum(totRev),
    nCalDays = sum(days)) %>% 
  filter(!is.na(eff.freq) & !is.na(eff.earn) & !is.na(eff.lbs)) %>%
  as.data.frame()
  return(effDiv)
}

# This function takes the output of the one above and 
# summmarizes diversity metrics
summarize_effectiveDiversity = function(dataFrame) {
  effDiv = group_by(dataFrame, year) %>% 
  summarize(mean.freq = mean(eff.freq),
  sd.freq = sd(eff.freq),
  mean.earn = mean(eff.earn),
  sd.earn = sd(eff.earn),
  mean.lbs = mean(eff.lbs),
  sd.lbs = sd(eff.lbs)) %>% 
  as.data.frame()  
  return(effDiv)
}
```

## Diversity by species, gear, and fisheries.

We have 3-4 dimensions we're interested in calculating diversity (species, gears, fisheries, space). And for each dimension, there's 3 ways we can calculate diversity: diversity based on fish ticket frequency, diversity based on revenue of each (weighted by g_earn) and diversity based on the weight of catch (g_pounds).

To calculate the uncertainty (CV, quartiles, whatever) we need to calculate the diversity for each individual in each year (p_holder), and then summarize variability across individuals. We'll do this two ways: first using the whole dataset, and then using Eric's more restricted version (limiting the analysis to single people : vessel combinations)

### Calculations

We can calculate diversity across all metrics in the same block.

```{r}
spDiv_personYear = effectiveDiversity_by_personYear(cfec, variable="specn")
spTot = summarize_effectiveDiversity( spDiv_personYear )

gearDiv_personYear = effectiveDiversity_by_personYear(cfec, variable="gearn")
gearTot = summarize_effectiveDiversity( gearDiv_personYear )

fishDiv_personYear = effectiveDiversity_by_personYear(cfec, variable="p_fshy")
fishTot = summarize_effectiveDiversity( fishDiv_personYear )
```

### Boxplots

We can coarsely group individuals based on their species, gear, and fisheries diversity through time (e.g. 1-1.5, 1.5-2, 2-2.5, etc.) and make boxplots of revenue by group over time (similar to Jordan's boxplots). 
```{r}
brks = c(1, 1.01, 1.5, 2, 2.5, 3, 3.5, 4, 10)
spDiv_personYear$group = cut(spDiv_personYear$eff.earn, breaks = brks, right=FALSE)
gearDiv_personYear$group = cut(gearDiv_personYear$eff.earn, breaks = brks, right=FALSE)
fishDiv_personYear$group = cut(fishDiv_personYear$eff.earn, breaks = brks, right=FALSE)
```

```{r, fig.pos="placeHere", fig.cap="Revenue by effective species diversity and year"}
ggplot(spDiv_personYear, aes(x = factor(year), y = log10(totIndRev), fill = factor(group))) + 
  geom_boxplot(outlier.shape=NA) + 
  ggtitle("Revenue by species diversity ('specn') and year") + 
xlab("Year") + ylab("log10(revenue)")
```

```{r, fig.pos="placeHere", fig.cap="Revenue by effective species diversity across years"}
ggplot(spDiv_personYear, aes(x = factor(group), y = log10(totIndRev))) + 
  geom_boxplot(outlier.shape=NA) + 
  ggtitle("Revenue by species diversity ('specn')") + 
xlab("Year") + ylab("log10(revenue)")
```

```{r, fig.pos="placeHere", fig.cap="Number of unique landed dates by effective species diversity across years"}
ggplot(spDiv_personYear, aes(x = factor(group), y = log(nCalDays))) + 
  geom_boxplot(outlier.shape=NA) + 
  ggtitle("Unique landed dates by species diversity ('specn')") + 
xlab("Year") + ylab("log(unique days)")
```

```{r, fig.pos="placeHere", fig.cap="Revenue by effective gear diversity and year"}
ggplot(gearDiv_personYear, aes(x = factor(year), y = log10(totIndRev), fill = factor(group))) + 
  geom_boxplot(outlier.shape=NA) + 
  ggtitle("Revenue by gear diversity ('gearn') and year") + 
xlab("Year") + ylab("log10(revenue)")
```

```{r, fig.pos="placeHere", fig.cap="Revenue by effective gear diversity across years"}
ggplot(gearDiv_personYear, aes(x = factor(group), y = log10(totIndRev))) + 
  geom_boxplot(outlier.shape=NA) + 
  ggtitle("Revenue by gear diversity ('gearn')") + 
xlab("Year") + ylab("log10(revenue)")
```

```{r, fig.pos="placeHere", fig.cap="Number of unique landed dates by effective gear diversity across years"}
ggplot(gearDiv_personYear, aes(x = factor(group), y = log(nCalDays))) + 
  geom_boxplot(outlier.shape=NA) + 
  ggtitle("Unique landed dates by gear diversity ('gearn')") + 
xlab("Year") + ylab("log(unique days)")
```

```{r, fig.pos="placeHere", fig.cap="Revenue by effective permit diversity and year"}
ggplot(fishDiv_personYear, aes(x = factor(year), y = log10(totIndRev), fill = factor(group))) + 
  geom_boxplot(outlier.shape=NA) + 
  ggtitle("Revenue by permit diversity ('p_fshy') and year") + 
xlab("Year") + ylab("log10(revenue)")
```

```{r, fig.pos="placeHere", fig.cap="Revenue by effective permit diversity across years"}
ggplot(fishDiv_personYear, aes(x = factor(group), y = log10(totIndRev))) + 
  geom_boxplot(outlier.shape=NA) + 
  ggtitle("Revenue by permit diversity ('p_fshy')") + 
xlab("Year") + ylab("log10(revenue)")
```

```{r, fig.pos="placeHere", fig.cap="Number of unique landed dates by effective permit diversity across years"}
ggplot(fishDiv_personYear, aes(x = factor(group), y = log(nCalDays))) + 
  geom_boxplot(outlier.shape=NA) + 
  ggtitle("Unique landed dates by permit diversity ('p_fshy')") + 
xlab("Year") + ylab("log(unique days)")
```

### Species

```{r, fig.pos="placeHere", fig.cap="Species diversity (all individuals/vessels) by frequency in fish ticket data"}
ggplot(spTot, aes(x=year, y=mean.freq)) + geom_line(col="blue") +  
geom_ribbon(aes(ymin=mean.freq - sd.freq/mean.freq, 
  ymax=mean.freq + sd.freq/mean.freq), 
  alpha=0.2, fill="blue") + 
ggtitle("Effective species diversity ('specn') by fishTix") + 
xlab("Year") + ylab("Mean (+/- CV)")
```

```{r, fig.pos="placeHere", fig.cap="Species diversity (all individuals/vessels) weighted by landed value and catch weights"}
# Plot by earning / weight on diff scale
ggplot(spTot, aes(x=year, y=mean.earn)) + geom_line(col="red") + 
geom_ribbon(aes(x = year, ymin=mean.earn - sd.earn/mean.earn, 
  ymax = mean.earn + sd.earn/mean.earn), 
  alpha=0.2, fill="red") + 
geom_line(aes(x=year,y=mean.lbs), col="green") + 
geom_ribbon(aes(x = year, ymin=mean.lbs - sd.lbs/mean.lbs, 
  ymax=mean.lbs + sd.lbs/mean.lbs),  alpha=0.2, fill="green") + 
ggtitle("Effective species diversity ('specn') [red = $, green=weight]") + 
xlab("Year") + ylab("Mean (+/- CV)")

```

### Gears

```{r, fig.pos="placeHere", fig.cap="Gear diversity (all individuals/vessels) by frequency in fish ticket data"}
ggplot(gearTot, aes(x=year, y=mean.freq)) + geom_line(col="blue") + 
geom_ribbon(aes(ymin=mean.freq - sd.freq/mean.freq, 
  ymax=mean.freq + sd.freq/mean.freq), 
  alpha=0.2, fill="blue") + 
ggtitle("Effective gear diversity ('gearn') by fishTix") + 
xlab("Year") + ylab("Mean (+/- CV)")
```

```{r, fig.pos="placeHere", fig.cap="Gear diversity (all individuals/vessels) weighted by landed value and weight of catch"}
# Plot by earning / weight on diff scale
ggplot(gearTot, aes(x=year, y=mean.earn)) + geom_line(col="red") + 
geom_ribbon(aes(x = year, ymin=mean.earn - sd.earn/mean.earn, 
  ymax = mean.earn + sd.earn/mean.earn), alpha=0.2, fill="red") + 
geom_line(aes(x=year,y=mean.lbs), col="green") + 
geom_ribbon(aes(x = year, ymin=mean.lbs - sd.lbs/mean.lbs, 
  ymax=mean.lbs + sd.lbs/mean.lbs), alpha=0.2, fill="green") + 
ggtitle("Effective gear diversity ('gearn') [red = $, green=weight]") + 
xlab("Year") + ylab("Mean (+/- CV)")
```

### Fisheries

```{r, fig.pos="placeHere", fig.cap="Permit diversity (all individuals/vessels) by frequency in fish ticket data"}
ggplot(fishTot, aes(x=year, y=mean.freq)) + geom_line(col="blue") + 
geom_ribbon(aes(ymin=mean.freq - sd.freq/mean.freq, 
  ymax=mean.freq + sd.freq/mean.freq), alpha=0.2, fill="blue") + 
ggtitle("Effective permit diversity ('p_fshy') by fishTix") + 
xlab("Year") + ylab("Mean (+/- CV)")
```

```{r, fig.pos="placeHere", fig.cap="Permit diversity (all individuals/vessels) weighted by landed value and weight of catch"} 
# Plot by earning / weight on diff scale
ggplot(fishTot, aes(x=year, y=mean.earn)) + geom_line(col="red") + 
geom_ribbon(aes(x = year, ymin=mean.earn - sd.earn/mean.earn, 
  ymax = mean.earn + sd.earn/mean.earn), alpha=0.2, fill="red") + 
geom_line(aes(x=year,y=mean.lbs), col="green") + 
geom_ribbon(aes(x = year, ymin=mean.lbs - sd.lbs/mean.lbs, 
  ymax=mean.lbs + sd.lbs/mean.lbs), alpha=0.2, fill="green") +
ggtitle("Effective permit diversity ('p_fshy') [red = $, green=weight]") + 
xlab("Year") + ylab("Mean (+/- CV)")
```

## Filtering of unique people:vessel combinations

Filter out people who only use 1 vessel / year , and similarly filter out boats with more than 1 person / year. Some people might fish on multiple vessels within a year, and vessels might be used by multiple people within a year.

Now we'll filter out individuals who only used 1 boat in a year, and boats with only one permit holder fishing in a year.  
```{r}
cfecnew = cfec
cfec.1boat = group_by(cfecnew, p_holder, year) %>% 
  summarize(nBoat = length(unique(cadfg))) %>%
  filter(nBoat==1) %>% 
  left_join(cfecnew)

# Deal with setnet fisheries, where lots of p_holder
# share same codes
cfec.1boat = group_by(cfec.1boat, cadfg, year) %>% 
  summarize(nPerson = length(unique(p_holder))) %>%
  filter((nPerson==1 & cadfg %in% c(27960,27961) == FALSE) | cadfg %in% c(27960,27961)) %>% 
  left_join(cfec.1boat)

cfecnew = cfec.1boat
dim(cfecnew)[1]

#Now we can create a unique *id* column for individual - vessel combinations.  
cfecnew$id = paste(cfecnew$p_holder,cfecnew$cadfg)
# calculate number of unique person - vessel pairs ~ 40700
length(unique(cfecnew$id))
```
### Re-calculating diversity

```{r}
spDiv_personYear = effectiveDiversity_by_personYear(cfecnew, variable="specn")
spTot = summarize_effectiveDiversity( spDiv_personYear )

gearDiv_personYear = effectiveDiversity_by_personYear(cfecnew, variable="gearn")
gearTot = summarize_effectiveDiversity( gearDiv_personYear )

fishDiv_personYear = effectiveDiversity_by_personYear(cfecnew, variable="p_fshy")
fishTot = summarize_effectiveDiversity( fishDiv_personYear )
```

### Species

```{r, fig.pos = "placeHere", fig.cap = "Species diversity by frequency of fish tickets of unique id:vessel combinations (excluding people who fish multiple vessels / year, or vessels that are fished by multiple individuals in a year."}

ggplot(spTot, aes(x=year, y=mean.freq)) + geom_line(col="blue") +
  geom_ribbon(aes(ymin=mean.freq - sd.freq/mean.freq, 
  ymax=mean.freq + sd.freq/mean.freq), alpha=0.2, fill="blue") + 
ggtitle("Effective species diversity ('specn') by fishTix") + 
xlab("Year") + ylab("Mean (+/- CV)")
```

```{r, fig.pos = "placeHere", fig.cap = "Species diversity by revenue and landed weight of unique id:vessel combinations (excluding people who fish multiple vessels / year, or vessels that are fished by multiple individuals in a year."}

# Plot by earning / weight on diff scale
ggplot(spTot, aes(x=year, y=mean.earn)) + geom_line(col="red") + 
geom_ribbon(aes(x = year, ymin=mean.earn - sd.earn/mean.earn, 
  ymax = mean.earn + sd.earn/mean.earn), alpha=0.2, fill="red") + 
geom_line(aes(x=year,y=mean.lbs), col="green") + 
geom_ribbon(aes(x = year, ymin=mean.lbs - sd.lbs/mean.lbs, 
  ymax=mean.lbs + sd.lbs/mean.lbs), alpha=0.2, fill="green") +
ggtitle("Effective species diversity ('specn') [red = $, green=weight]") + 
xlab("Year") + ylab("Mean (+/- CV)")
```

### Gears

```{r, fig.pos = "placeHere", fig.cap = "Gear diversity by fish tickets of unique id:vessel combinations (excluding people who fish multiple vessels / year, or vessels that are fished by multiple individuals in a year."}
ggplot(gearTot, aes(x=year, y=mean.freq)) + geom_line(col="blue") + 
geom_ribbon(aes(ymin=mean.freq - sd.freq/mean.freq, 
  ymax=mean.freq + sd.freq/mean.freq), alpha=0.2, fill="blue") + 
ggtitle("Effective gear diversity ('gearn') by fishTix") + 
xlab("Year") + ylab("Mean (+/- CV)")
```

```{r, fig.pos = "placeHere", fig.cap = "Gear diversity by revenue and landed weight of unique id:vessel combinations (excluding people who fish multiple vessels / year, or vessels that are fished by multiple individuals in a year."}
# Plot by earning / weight on diff scale
ggplot(gearTot, aes(x=year, y=mean.earn)) + geom_line(col="red") + 
geom_ribbon(aes(x = year, ymin=mean.earn - sd.earn/mean.earn, 
  ymax = mean.earn + sd.earn/mean.earn), alpha=0.2, fill="red") + 
geom_line(aes(x=year,y=mean.lbs), col="green") + 
geom_ribbon(aes(x = year, ymin=mean.lbs - sd.lbs/mean.lbs, 
  ymax=mean.lbs + sd.lbs/mean.lbs), alpha=0.2, fill="green") + 
ggtitle("Effective gear diversity ('gearn') [red = $, green=weight]") + 
xlab("Year") + ylab("Mean (+/- CV)")
```

### Fisheries 
```{r, fig.pos = "placeHere", fig.cap = "Permit diversity by fish ticket frequency of unique id:vessel combinations (excluding people who fish multiple vessels / year, or vessels that are fished by multiple individuals in a year."}

ggplot(fishTot, aes(x=year, y=mean.freq)) + geom_line(col="blue") +
geom_ribbon(aes(ymin=mean.freq - sd.freq/mean.freq, 
  ymax=mean.freq + sd.freq/mean.freq), alpha=0.2, fill="blue") + 
ggtitle("Effective permit diversity ('p_fshy') by fishTix") + 
xlab("Year") + ylab("Mean (+/- CV)")
```

```{r, fig.pos = "placeHere", fig.cap = "Permit diversity by revenue and landed weight of unique id:vessel combinations (excluding people who fish multiple vessels / year, or vessels that are fished by multiple individuals in a year."}

# Plot by earning / weight on diff scale
ggplot(fishTot, aes(x=year, y=mean.earn)) + geom_line(col="red") + 
geom_ribbon(aes(x = year, ymin=mean.earn - sd.earn/mean.earn, 
  ymax = mean.earn + sd.earn/mean.earn), alpha=0.2, fill="red") + 
geom_line(aes(x=year,y=mean.lbs), col="green") + 
geom_ribbon(aes(x = year, ymin=mean.lbs - sd.lbs/mean.lbs, 
  ymax=mean.lbs + sd.lbs/mean.lbs), alpha=0.2, fill="green") + 
ggtitle("Effective permit diversity ('p_fshy') [red = $, green=weight]") + 
xlab("Year") + ylab("Mean (+/- CV)")
```

## Comparing diversity between vessels and individuals

Just out of curiousity, we can summarize diversity by year taking the approach used by Kasperski and Holland (based on CADFG number) and compare that to the same criteria using people ("p_holder"). We'll only do this comparison on the species diversity, but use the same restriction: use only vessels or people who had participated continuously for the entire time period (this throws out 90%+ of the data).

First, based on vessels:
```{r}
contVessels = group_by(cfec, cadfg) %>% 
  summarize(nYear = length(unique(year))) %>% 
  filter(nYear == 30)

cfecnew = left_join(contVessels, cfec)

# calculate diversity by species by CADFG (vessel #)
effDiv.vessel = group_by(cfecnew, specn, year, cadfg) %>%
  summarize(totRev = sum(g_earn, na.rm=T),
  totN = n(),
  totWeight = sum(g_pounds, na.rm=T),
    days = length(unique(day))) %>% 
  as.data.frame() %>%
  filter(is.finite(totRev) & is.finite(totN) & is.finite(totWeight)) %>%   
  as.data.frame() %>%
  group_by(cadfg, year) %>% 
  summarize(eff.freq = simp.div(totN),
  eff.earn = simp.div(totRev),
  eff.lbs = simp.div(totWeight),
    totIndRev = sum(totRev),
    nCalDays = sum(days)) %>% 
  filter(!is.na(eff.freq) & !is.na(eff.earn) & !is.na(eff.lbs)) %>%
  as.data.frame()

effDiv.vessel = summarize_effectiveDiversity(effDiv.vessel)

effDiv.person = effectiveDiversity_by_personYear(cfecnew, variable="specn")
effDiv.person = summarize_effectiveDiversity( effDiv.person )
```

```{r, fig.pos="placeHere", fig.cap="Comparison of species diversity using vessel versus individuals"}
ggplot(effDiv.vessel, aes(x=year, y=mean.earn)) + geom_line(col="red") + 
geom_ribbon(aes(x = year, ymin=mean.earn - sd.earn/mean.earn, 
  ymax = mean.earn + sd.earn/mean.earn), alpha=0.2, fill="red") + 
geom_line(aes(x=year,y=effDiv.person$mean.earn), col="green") + 
geom_ribbon(aes(x = year, ymin=effDiv.person$mean.earn - 
    effDiv.person$sd.earn/effDiv.person$mean.earn, 
  ymax=effDiv.person$mean.earn + effDiv.person$sd.earn/effDiv.person$mean.earn), 
  alpha=0.2, fill="green") + 
ggtitle("Effective species diversity ('specn') [red = vessel, green=person]") + 
xlab("Year") + ylab("Mean (+/- CV)")
```

Now make the portfolio plots from the metafolio package:

```{r}
species_diversity_by_year <-
  effectiveDiversity_by_personYear(cfec, variable = "specn")
breaks = c(1, 1.01, 1.5, 2, 2.5, 3, 3.5, 4, 10)
species_diversity <-
  group_by(species_diversity_by_year, p_holder) %>%
  summarize(
    # diversity_by_frequency = mean(eff.freq),
    diversity_by_earnings = mean(eff.earn),
    # diversity_by_weight = mean(eff.lbs),
    m = log10(mean(totIndRev)),
    v = sd(totIndRev) / mean(totIndRev)
  ) %>%
  filter(!is.na(v)) %>%
  mutate(diversity_group = cut(diversity_by_earnings, breaks = breaks,
    right = FALSE))

species_diversity_list <- split(species_diversity,
  species_diversity$diversity_group)

#pdf("earnings_portfolio_by_species_diversity.pdf", height = 5, width = 5)
par(cex = 0.8)
metafolio::plot_cons_plans(
  species_diversity_list,
  plans_name = names(species_diversity_list),
  cols = RColorBrewer::brewer.pal(length(names(species_diversity_list)),
    "YlOrRd"),
  xlab = "CV of gross earnings",
  ylab = "log10 of gross earnings",
  add_all_efs = FALSE)
#dev.off()
```

### 2016-06-09 analysis_diversity/paper1a_diversity_metrics_4-20.Rmd
committing updates to all papers

---
title: 'Changes in diversity and variability in diversity through time: Catch portfolios
  paper 1A'
output:
  pdf_document:
    fig_caption: yes
  html_document: default
date: "April 7, 2016"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=FALSE, tidy.opts=list(width.cutoff=60), warning = FALSE, message = FALSE)
```

## Load cleaned data  
```{r dataLoad, warning = FALSE, message=FALSE, results="hide", echo=FALSE}
library(dplyr)
library(knitr)
library(mgcv)
library(ggplot2)
library(randomForest)
library(MASS)
library(zoo)
library(MatchIt)
library(metafolio)
library(randomForest)
load("cfecCleaned2.Rdata")

# convert 6-dig p_fshy codes to 5 digits
# removing blank space - now they're all length(5)
idx = which(nchar(cfec$p_fshy)==6)
cfec$p_fshy[idx] = paste(substr(cfec$p_fshy[idx], 1, 1), 
  substr(cfec$p_fshy[idx], 3, 6), sep="")

# remove white space in 2nd position
idx = which(substr(cfec$p_fshy, 2, 2) == " ")
cfec$p_fshy[idx] = paste(substr(cfec$p_fshy[idx], 1, 1), 
  substr(cfec$p_fshy[idx], 3, length(cfec$p_fshy[idx])), sep="")

# Pull in "new.locs" from Anne/Jordan/Rich/Ben/Ole
newlocs = read.csv("new.locs.csv")
newlocs$p_fshy = as.character(newlocs$p_fshy)

# Match cfec codes in newlocs to identify new 
# locations (Anne/Jordan/Rich/Ben/Ole)
cfec$landedLoc = newlocs$New.Loc[match(cfec$p_fshy, newlocs$p_fshy)]
cfec$landedLoc[which(is.na(cfec$landedLoc))] = "Statewide" # C04B

```

We need to adjust the landed values and prices for inflation. To do this, we'll use the data Alan pulled together (adjusting to 2009 dollars).

https://docs.google.com/spreadsheets/d/1hRLYTX1nZBLKnYXwTCOtU7-XnDoMZf_1ufdTiNQFd4U/edit#gid=0

```{r, echo=FALSE}
deflationTable = data.frame(year = 1985:2014, defl = c(0.57236, 0.58393, 0.59879, 0.61974,  
0.64388, 0.66774, 0.68993, 0.70564, 0.72244, 0.73781,  
0.75321, 0.76695, 0.78009, 0.78855, 0.80061,  
0.81883, 0.83753, 0.85038, 0.86729, 0.89114,  
0.91981, 0.94812, 0.97334, 0.9925, 1,  
1.01217, 1.03307, 1.05213, 1.06926, 1.08682))

# Adjust price and g_earn for inflation
cfec$year = as.numeric(cfec$year)
cfec$day = substr(cfec$landdate, 6, 10)
cfec$g_price = cfec$g_price/deflationTable$defl[cfec$year-1984]
cfec$g_earn = cfec$g_earn/deflationTable$defl[cfec$year-1984]
```

### Function definiton  
We need to define a few functions for calculating diversity. We'll use the Simpson's diversity index (HHI) to calculate the entropy-based inverse, which is a proxy for effective species diversity. Note the C++ version is commented out below,   
```{r, warning = FALSE, message=FALSE}

simp.div = function(x) {
  1/sum((x/sum(x))^2)
}

#library(Rcpp)
#cppFunction('double simpCpp(NumericVector x) {
#  double y = 1/sum((x/sum(x))^2);
#  return y;
#}')

# Note that the first line of this function uses group_by_ because 
# we need to pass in a string for 'variable'
effectiveDiversity_by_personYear = function(dataFrame, variable) {
  effDiv = group_by_(dataFrame, variable, "year", "p_holder") %>%
  summarize(totRev = sum(g_earn, na.rm=T),
  totN = n(),
  totWeight = sum(g_pounds, na.rm=T),
    days = length(unique(day))) %>% 
  as.data.frame() %>%
  filter(is.finite(totRev) & is.finite(totN) & is.finite(totWeight)) %>%   
  as.data.frame() %>%
  group_by(p_holder, year) %>% 
  summarize(eff.freq = simp.div(totN),
  eff.earn = simp.div(totRev),
  eff.lbs = simp.div(totWeight),
    totIndRev = sum(totRev),
    nCalDays = sum(days)) %>% 
  filter(!is.na(eff.freq) & !is.na(eff.earn) & !is.na(eff.lbs)) %>%
  as.data.frame()
  return(effDiv)
}

# Note that the first line of this function uses group_by_ because 
# we need to pass in a string for 'variable'
effectiveDiversity_by_personYear2 = function(dataFrame, variable1, variable2) {
  effDiv = group_by_(dataFrame, variable1, variable2, "year", "p_holder") %>%
  summarize(totRev = sum(g_earn, na.rm=T),
  totN = n(),
  totWeight = sum(g_pounds, na.rm=T),
    days = length(unique(day))) %>% 
  as.data.frame() %>%
  filter(is.finite(totRev) & is.finite(totN) & is.finite(totWeight)) %>%   
  as.data.frame() %>% 
  group_by_("p_holder", "year", variable2) %>% 
  summarize(eff.freq = simp.div(totN),
  eff.earn = simp.div(totRev),
  eff.lbs = simp.div(totWeight),
    totIndRev = sum(totRev),
    nCalDays = sum(days)) %>% 
  filter(!is.na(eff.freq) & !is.na(eff.earn) & !is.na(eff.lbs)) %>%
  as.data.frame()
  return(effDiv)
}

effectiveDiversity_by_Year = function(dataFrame, variable) {
  effDiv = group_by_(dataFrame, variable, "year") %>%
  summarize(totRev = sum(g_earn, na.rm=T),
  totN = n(),
  totWeight = sum(g_pounds, na.rm=T),
    days = length(unique(day))) %>% 
  as.data.frame() %>%
  filter(is.finite(totRev) & is.finite(totN) & is.finite(totWeight)) %>%   
  as.data.frame() %>%
  group_by(year) %>% 
  summarize(eff.freq = simp.div(totN),
  eff.earn = simp.div(totRev),
  eff.lbs = simp.div(totWeight),
    totIndRev = sum(totRev),
    nCalDays = sum(days)) %>% 
  filter(!is.na(eff.freq) & !is.na(eff.earn) & !is.na(eff.lbs)) %>%
  as.data.frame()
  return(effDiv)
}

# This function takes the output of the one above and 
# summmarizes diversity metrics
summarize_effectiveDiversity = function(dataFrame) {
  effDiv = group_by(dataFrame, year) %>% 
  summarize(mean.freq = mean(eff.freq),
  sd.freq = sd(eff.freq),
  mean.earn = mean(eff.earn),
  sd.earn = sd(eff.earn),
  mean.lbs = mean(eff.lbs),
  sd.lbs = sd(eff.lbs)) %>% 
  as.data.frame()  
  return(effDiv)
}

# This function takes the output of the one above and 
# summmarizes diversity metrics
summarize_effectiveDiversity2 = function(dataFrame, variable2) {
  effDiv = group_by_(dataFrame, "year", variable2) %>% 
  summarize(mean.freq = mean(eff.freq),
  sd.freq = sd(eff.freq),
  mean.earn = mean(eff.earn),
  sd.earn = sd(eff.earn),
  mean.lbs = mean(eff.lbs),
  sd.lbs = sd(eff.lbs)) %>% 
  as.data.frame()  
  return(effDiv)
}
```

## Diversity by species, gear, and fisheries.

We have 4 dimensions we're interested in calculating diversity (species, gears, fisheries, space). And for each dimension, there's 3 ways we can calculate diversity: diversity based on fish ticket frequency, diversity based on revenue of each (weighted by g_earn) and diversity based on the weight of catch (g_pounds).

To calculate the uncertainty (CV, quartiles, whatever) we need to calculate the diversity for each individual in each year (p_holder), and then summarize variability across individuals. We'll do this two ways: first using the whole dataset, and then using Eric's more restricted version (limiting the analysis to single people : vessel combinations)

### Calculations

We can calculate diversity across all metrics in the same block.

```{r}
cfec$permit_coarse = substr(cfec$p_fshy,1,1)
spDiv_personYear = effectiveDiversity_by_personYear(cfec, variable="specn")
spTot = summarize_effectiveDiversity( spDiv_personYear )

gearDiv_personYear = effectiveDiversity_by_personYear(cfec, variable="gearn")
gearTot = summarize_effectiveDiversity( gearDiv_personYear )

fishDiv_personYear = effectiveDiversity_by_personYear(cfec, variable="permit_coarse")
fishTot = summarize_effectiveDiversity( fishDiv_personYear )

areaDiv_personYear = effectiveDiversity_by_personYear(cfec, variable="area")
areaTot = summarize_effectiveDiversity( areaDiv_personYear )
```

### Summarizing landings and value

First, it's helpful to identify the broad patterns in the CFEC data, showing time series of people, landings, and revenue by year. The general patterns are consistent with Kasperski & Holland's analysis that showed diversity declining. More specificially,

- Total participation decreased after ~ 1990, both in terms of people + boats  
- Participation has been flat since early 2000s, but since 2000
- both catch (lbs landed) / person, and revenue / person have been increasing

```{r}
cfecSummarize = group_by(cfec, year) %>%
  summarize(people = length(unique(p_holder)), 
    revenue = sum(g_earn), catch = sum(g_pounds),
    vessels = length(unique(cadfg))) %>%
  as.data.frame
par(mfrow = c(3,2), mgp=c(2,1,0), mai=c(0.5,0.5,0.05,0.05))
plot(cfecSummarize$year, cfecSummarize$people/1000, xlab="", ylab="People (1000s)", type="b", lwd=3)
plot(cfecSummarize$year, cfecSummarize$revenue/1e+09, xlab="", ylab="Billion $s", type="b", lwd=3)
plot(cfecSummarize$year, cfecSummarize$catch/2e+06, xlab="", ylab="Weight (tons)", type="b", lwd=3)
plot(cfecSummarize$year, cfecSummarize$revenue/cfecSummarize$people, xlab="", ylab="$s / person", type="b", lwd=3)
plot(cfecSummarize$year, cfecSummarize$catch/cfecSummarize$people, xlab="", ylab="Lbs / person", type="b", lwd=3)
plot(cfecSummarize$year, cfecSummarize$vessels/cfecSummarize$people, xlab="", ylab="Boats / person", type="b", lwd=3)

```

### Patterns in diversity metrics

We can coarsely group individuals based on their species, gear, and fisheries diversity through time (e.g. 1-1.5, 1.5-2, 2-2.5, etc.) and make boxplots of revenue by group over time (similar to Jordan's boxplots). 
```{r}
brks = c(1, 1.0001, 1.5, 2, 2.5, 3, 3.5, 4, 10)

spDiv_personYear$group = cut(spDiv_personYear$eff.earn, breaks = brks, right=FALSE)
gearDiv_personYear$group = cut(gearDiv_personYear$eff.earn, breaks = brks, right=FALSE)
fishDiv_personYear$group = cut(fishDiv_personYear$eff.earn, breaks = brks, right=FALSE)
areaDiv_personYear$group = cut(areaDiv_personYear$eff.earn, breaks = brks, right=FALSE)
```

We can start by just plotting changes in group membership over time. The interesting patterns here are:
1. Species: If we lump the lowest diversity groups (1 - 1.5 spp), there's been a slight uptick from ~ 55% - 65% (consistent with specialization).  
2. Gears / Permits: people fishing 1 gear or 1 permit have also increased since ~ 1991, from ~ 70% to > 80% of all records.  
3. There's no interesting trend with respect to regions; 75-80% of the people fish 1 region, and there's no trend. [Revisit after getting Anne's areas]

```{r, fig.pos="placeHere", fig.cap="Proportion of people by species diversity groups"}
group.sum = group_by(spDiv_personYear, year, group) %>% 
  summarize(n = length(unique(p_holder))) %>% 
  as.data.frame()
tot = group_by(group.sum, year) %>% 
  summarize(tot = sum(n))
group.sum.spec = left_join(group.sum, tot)
ggplot(group.sum.spec, aes(x=year, y=n/tot, group=group, colour=group ) ) + geom_line() +
  ggtitle("P_holder by species diversity groups") + 
xlab("Year") + ylab("Percent of total") 
```

```{r, fig.pos="placeHere", fig.cap="Proportion of people by gear diversity groups"}
group.sum = group_by(gearDiv_personYear, year, group) %>% 
  summarize(n = length(unique(p_holder))) %>% 
  as.data.frame()
tot = group_by(group.sum, year) %>% 
  summarize(tot = sum(n))
group.sum.gear = left_join(group.sum, tot)
ggplot(group.sum.gear, aes(x=year, y=n/tot, group=group, colour=group ) ) + geom_line() +
  ggtitle("P_holder by gear diversity groups") + 
xlab("Year") + ylab("Percent of total") 
```

```{r, fig.pos="placeHere", fig.cap="Proportion of people by area diversity groups"}
group.sum = group_by(areaDiv_personYear, year, group) %>% 
  summarize(n = length(unique(p_holder))) %>% 
  as.data.frame()
tot = group_by(group.sum, year) %>% 
  summarize(tot = sum(n))
group.sum.area = left_join(group.sum, tot)
ggplot(group.sum.area, aes(x=year, y=n/tot, group=group, colour=group ) ) + geom_line() +
  ggtitle("P_holder by area diversity groups") + 
xlab("Year") + ylab("Percent of total") 
```

```{r, fig.pos="placeHere", fig.cap="Proportion of people by permit diversity groups"}
group.sum = group_by(fishDiv_personYear, year, group) %>% 
  summarize(n = length(unique(p_holder))) %>% 
  as.data.frame()
tot = group_by(group.sum, year) %>% 
  summarize(tot = sum(n))
group.sum.fish = left_join(group.sum, tot)
ggplot(group.sum.fish, aes(x=year, y=n/tot, group=group, colour=group ) ) + geom_line() +
  ggtitle("P_holder by permit diversity groups") + 
xlab("Year") + ylab("Percent of total") 
```

It's pretty easy as a next step to focus on specialists (gears, permits). Which gears and permits are being used, and how does this break down by area? This shows some pretty interesting patterns that we can characterize about 'specialists':

- longline specialists greatly decreased after 2004 (going to 0 in alaska peninsula, cook inlet, huge drops in kodiak, smaller drops in PWS / Southeast).  
- we've seen slight increases in drift gillnet specialists (gearn = 3) in Bristol Bay, and larger increases in PWS / Cook Inlet (presumably copper river sockeye)

```{r}
# Find all people in group 1 (specialists)
sub = gearDiv_personYear[which(as.numeric(gearDiv_personYear$group)%in%c(1,2)),]
LJ = left_join(sub, cfec)
# Calculate number of people using each gear type
LJ = group_by(LJ, gearn, year, landedLoc) %>% 
  summarize(n = length(unique(p_holder)))
# Calculate total to normalize to %
tot = group_by(LJ, year, landedLoc) %>% 
  summarize(tot = sum(n))
LJ = left_join(LJ, tot)
# Plot proportions
ggplot(LJ, aes(x=year, y=n/tot, group=as.factor(gearn), colour=as.factor(gearn) ) ) + geom_line() + facet_wrap(~ landedLoc, scales="free_y") + ylab("Proportion of people")
```

The next is to pull out the specialists, by gear or species. For example we can pull out people specializing on salmon. For example, we could look at the trends in some of these metrics for people who are only using a gear specific to salmon (1, 3, 4). There's a few interesting things here:

-- Nobody uses > 2 gears

```{r}
salmonSpecialists = group_by(cfec, year, p_holder, landedLoc) %>% 
  summarize(gears = length(unique(gearn)), 
    salmonGears = length(unique(gearn%in%c(1,3,4))),
    salmonSpec = length(unique(specn[specn%in%c(410,420,430,440,450)]))) %>% 
  filter(salmonGears == gears & salmonSpec >= 1)

salmonSummary = group_by(salmonSpecialists, year, landedLoc) %>% 
  summarize(ngear = mean(gears), ngear2 = length(which(gears == 2)), npeople = length(unique(p_holder)), nspec = mean(salmonSpec)) %>% 
  filter(landedLoc %in% c("Bering Sea", "Aleutian Islands", "Dutch Harbor", "Statewide", "GOA", "EEZ")==FALSE)

ggplot(salmonSummary, aes(x=year, y=ngear2/npeople) ) + geom_line() + facet_wrap(~ landedLoc, scales="free_y") + ylab("% People using 2 salmon gears")

ggplot(salmonSummary, aes(x=year, y=npeople) ) + geom_line() + facet_wrap(~ landedLoc, scales="free_y") + ylab("People exclusively using gears 1,3,4")

ggplot(salmonSummary, aes(x=year, y=nspec) ) + geom_line() + facet_wrap(~ landedLoc, scales="free_y") + ylab("People exclusively using gears 1,3,4")

salmonSpecialists = group_by(cfec, year, p_holder) %>% 
  summarize(gears = length(unique(gearn)), 
    salmonGears = length(unique(gearn%in%c(1,3,4))),
    salmonSpec = length(unique(specn[specn%in%c(410,420,430,440,450)]))) %>% 
  filter(salmonGears == gears & salmonSpec >= 1) 
salmonSpecialists = left_join(salmonSpecialists, cfec)
```

```{r, fig.pos="placeHere", fig.cap="Revenue by effective species diversity and year", echo=FALSE}
ggplot(spDiv_personYear, aes(x = factor(year), y = log10(totIndRev), fill = factor(group))) + 
  geom_boxplot(outlier.shape=NA) + 
  ggtitle("Revenue by species diversity ('specn') and year") + 
xlab("Year") + ylab("log10(revenue)")
```

We can also try to calculate downside risk, by group. We'll start with the lower 95% for each group. 
```{r, fig.pos="placeHere", fig.cap="Downside risk metrics: lower 5% and 20% of total revenue, by species diversity and year", echo=FALSE}
spDiv_lower = group_by(spDiv_personYear, year, group) %>% 
  summarize(lower95 = quantile(log10(totIndRev),0.05),
    lower80 = quantile(log10(totIndRev),0.2))
ggplot(spDiv_lower, aes(x=year, y=lower95, group=group, colour=group ) ) + 
            geom_line(size=1) + geom_line(aes(x=year, y=lower80, group=group, colour=group ), size=1, linetype = 2) +
  ggtitle("Lower 5% (solid) of and 20% (dashed) log10(revenue)") + 
xlab("Year") + ylab("Lower 5% and 20%")  
```

```{r, fig.pos="placeHere", fig.cap="Revenue by effective species diversity across years", echo=FALSE}
ggplot(spDiv_personYear, aes(x = factor(group), y = log10(totIndRev))) + 
  geom_boxplot(outlier.shape=NA) + 
  ggtitle("Revenue by species diversity ('specn')") + 
xlab("Effective diversity") + ylab("log10(revenue)")
```

```{r, fig.pos="placeHere", fig.cap="Number of unique landed dates by effective species diversity across years", echo=FALSE, include=FALSE}
ggplot(spDiv_personYear, aes(x = factor(group), y = log(nCalDays))) + 
  geom_boxplot(outlier.shape=NA) + 
  ggtitle("Unique landed dates by species diversity ('specn')") + 
xlab("Effective diversity") + ylab("log(unique days)")
```

```{r, fig.pos="placeHere", fig.cap="Revenue by effective gear diversity and year", echo=FALSE}
ggplot(gearDiv_personYear, aes(x = factor(year), y = log10(totIndRev), fill = factor(group))) + 
  geom_boxplot(outlier.shape=NA) + 
  ggtitle("Revenue by gear diversity ('gearn') and year") + 
xlab("Year") + ylab("log10(revenue)")
```

```{r, fig.pos="placeHere", fig.cap="Downside risk metrics: lower 5% and 20% of total revenue, by gear diversity and year", echo=FALSE, include=FALSE}
gearDiv_lower = group_by(gearDiv_personYear, year, group) %>% 
  summarize(lower95 = quantile(log10(totIndRev),0.05),
    lower80 = quantile(log10(totIndRev),0.2))
ggplot(gearDiv_lower, aes(x=year, y=lower95, group=group, colour=group ) ) + 
            geom_line(size=1) + geom_line(aes(x=year, y=lower80, group=group, colour=group ), size=1, linetype = 2) +
  ggtitle("Lower 5% (solid) of and 20% (dashed) log10(revenue)") + 
xlab("Year") + ylab("Lower 5% and 20%")  
```

```{r, fig.pos="placeHere", fig.cap="Revenue by effective gear diversity across years", echo=FALSE, include=FALSE}
ggplot(gearDiv_personYear, aes(x = factor(group), y = log10(totIndRev))) + 
  geom_boxplot(outlier.shape=NA) + 
  ggtitle("Revenue by gear diversity ('gearn')") + 
xlab("Effective diversity") + ylab("log10(revenue)")
```

```{r, fig.pos="placeHere", fig.cap="Number of unique landed dates by effective gear diversity across years", echo=FALSE, include=FALSE}
ggplot(gearDiv_personYear, aes(x = factor(group), y = log(nCalDays))) + 
  geom_boxplot(outlier.shape=NA) + 
  ggtitle("Unique landed dates by gear diversity ('gearn')") + 
xlab("Effective diversity") + ylab("log(unique days)")
```

```{r, fig.pos="placeHere", fig.cap="Revenue by effective permit diversity and year", echo=FALSE}
ggplot(fishDiv_personYear, aes(x = factor(year), y = log10(totIndRev), fill = factor(group))) + 
  geom_boxplot(outlier.shape=NA) + 
  ggtitle("Revenue by permit diversity ('p_fshy') and year") + 
xlab("Year") + ylab("log10(revenue)")
```

```{r, fig.pos="placeHere", fig.cap="Downside risk metrics: lower 5% and 20% of total revenue, by permit diversity and year", echo=FALSE, include=FALSE}
fishDiv_lower = group_by(fishDiv_personYear, year, group) %>% 
  summarize(lower95 = quantile(log10(totIndRev),0.05),
    lower80 = quantile(log10(totIndRev),0.2))
ggplot(fishDiv_lower, aes(x=year, y=lower95, group=group, colour=group ) ) + 
            geom_line(size=1) + geom_line(aes(x=year, y=lower80, group=group, colour=group ), size=1, linetype = 2) +
  ggtitle("Lower 5% (solid) of and 20% (dashed) log10(revenue)") + 
xlab("Year") + ylab("Lower 5% and 20%")  
```

```{r, fig.pos="placeHere", fig.cap="Revenue by effective permit diversity across years", echo=FALSE, include=FALSE}
ggplot(fishDiv_personYear, aes(x = factor(group), y = log10(totIndRev))) + 
  geom_boxplot(outlier.shape=NA) + 
  ggtitle("Revenue by permit diversity ('p_fshy')") + 
xlab("Effective diversity") + ylab("log10(revenue)")
```

```{r, fig.pos="placeHere", fig.cap="Number of unique landed dates by effective permit diversity across years", echo=FALSE, include=FALSE}
ggplot(fishDiv_personYear, aes(x = factor(group), y = log(nCalDays))) + 
  geom_boxplot(outlier.shape=NA) + 
  ggtitle("Unique landed dates by permit diversity ('p_fshy')") + 
xlab("Effective diversity") + ylab("log(unique days)")
```

```{r, fig.pos="placeHere", fig.cap="Revenue by effective area diversity and year", echo=FALSE}
ggplot(areaDiv_personYear, aes(x = factor(year), y = log10(totIndRev), fill = factor(group))) + 
  geom_boxplot(outlier.shape=NA) + 
  ggtitle("Revenue by area diversity ('landedLoc') and year") + 
xlab("Year") + ylab("log10(revenue)")
```

```{r, fig.pos="placeHere", fig.cap="Revenue by area permit diversity across years", include=FALSE, echo=FALSE}
ggplot(areaDiv_personYear, aes(x = factor(group), y = log10(totIndRev))) + 
  geom_boxplot(outlier.shape=NA) + 
  ggtitle("Revenue by area diversity ('landedLoc')") + 
xlab("Effective diversity") + ylab("log10(revenue)")
```

```{r, fig.pos="placeHere", fig.cap="Number of unique landed dates by area permit diversity across years", echo=FALSE, include=FALSE}
ggplot(fishDiv_personYear, aes(x = factor(group), y = log(nCalDays))) + 
  geom_boxplot(outlier.shape=NA) + 
  ggtitle("Unique landed dates by area diversity ('landedLoc')") + 
xlab("Effective diversity") + ylab("log(unique days)")
```

### Species

```{r, fig.pos="placeHere", fig.cap="Species diversity (all individuals/vessels). For scaling purposes, 1.2 is subtracted from the diversity based on fish ticket frequency.", echo=FALSE}
# Plot by earning / weight on diff scale
ggplot(spTot, aes(x=year, y=mean.earn)) + geom_line(col="green") + 
geom_line(aes(x=year,y=mean.lbs), col="red") + 
geom_line(aes(x=year,y=mean.freq-1.2), col="blue") + 
ggtitle("Effective species diversity ('specn') [green='$', red=weight, blue = %]") + 
xlab("Year") + ylab("Effective diversity")
```

```{r, fig.pos="placeHere", fig.cap="CV of Species diversity (all individuals/vessels)", echo=FALSE}
# Plot by earning / weight on diff scale
ggplot(spTot, aes(x=year, y=sd.earn/mean.earn)) + geom_line(col="green") + 
geom_line(aes(x=year,y=sd.lbs/mean.lbs), col="red") + 
geom_line(aes(x=year,y=sd.freq/mean.freq), col="blue") + 
ggtitle("CV of species diversity ('specn') [green='$', red=weight, blue = %]") + 
xlab("Year") + ylab("CV")
```

### Gears

```{r, fig.pos="placeHere", fig.cap="Gear diversity (all individuals/vessels)", echo=FALSE}
# Plot by earning / weight on diff scale
ggplot(gearTot, aes(x=year, y=mean.earn)) + geom_line(col="green") + 
geom_line(aes(x=year,y=mean.lbs), col="red") + 
geom_line(aes(x=year,y=mean.freq), col="blue") + 
ggtitle("Effective gear diversity ('gearn') [green='$', red=weight, blue = %]") + 
xlab("Year") + ylab("Effective diversity")
```

```{r, fig.pos="placeHere", fig.cap="CV of gear diversity (all individuals/vessels)", echo=FALSE}
# Plot by earning / weight on diff scale
ggplot(gearTot, aes(x=year, y=sd.earn/mean.earn)) + geom_line(col="green") + 
geom_line(aes(x=year,y=sd.lbs/mean.lbs), col="red") + 
geom_line(aes(x=year,y=sd.freq/mean.freq), col="blue") + 
ggtitle("CV of gear diversity ('gearn') [green='$', red=weight, blue = %]") + 
xlab("Year") + ylab("CV")
```

### Fisheries

```{r, fig.pos="placeHere", fig.cap="Permit diversity (all individuals/vessels)", echo=FALSE}
# Plot by earning / weight on diff scale
ggplot(fishTot, aes(x=year, y=mean.earn)) + geom_line(col="green") + 
geom_line(aes(x=year,y=mean.lbs), col="red") + 
geom_line(aes(x=year,y=mean.freq), col="blue") + 
ggtitle("Effective permit diversity ('p_fshy') [green='$', red=weight, blue = %]") + 
xlab("Year") + ylab("Effective diversity")
```

```{r, fig.pos="placeHere", fig.cap="CV of permit diversity (all individuals/vessels)", echo=FALSE}
# Plot by earning / weight on diff scale
ggplot(fishTot, aes(x=year, y=sd.earn/mean.earn)) + geom_line(col="green") + 
geom_line(aes(x=year,y=sd.lbs/mean.lbs), col="red") + 
geom_line(aes(x=year,y=sd.freq/mean.freq), col="blue") + 
ggtitle("CV of permit diversity ('p_fshy') [green='$', red=weight, blue = %]") + 
xlab("Year") + ylab("CV")
```

### Areas (including EEZ / Statewide / GOA categories)

```{r, fig.pos="placeHere", fig.cap="Area diversity (all individuals/vessels)", echo=FALSE}
# Plot by earning / weight on diff scale
ggplot(areaTot, aes(x=year, y=mean.earn)) + geom_line(col="green") + 
geom_line(aes(x=year,y=mean.lbs), col="red") + 
geom_line(aes(x=year,y=mean.freq), col="blue") + 
ggtitle("Effective area diversity ('p_fshy') [green='$', red=weight, blue = %]") + 
xlab("Year") + ylab("Effective diversity")
```

```{r, fig.pos="placeHere", fig.cap="CV of area diversity (all individuals/vessels)", echo=FALSE}
# Plot by earning / weight on diff scale
ggplot(areaTot, aes(x=year, y=sd.earn/mean.earn)) + geom_line(col="green") + 
geom_line(aes(x=year,y=sd.lbs/mean.lbs), col="red") + 
geom_line(aes(x=year,y=sd.freq/mean.freq), col="blue") + 
ggtitle("CV of area diversity ('landedLoc') [green='$', red=weight, blue = %]") + 
xlab("Year") + ylab("CV")
```

## Finer scale exploration of patterns

We can also look further into whether the observed trends in mean and CV of revenue vary by gear type, area, permit type, etc.

```{r, fig.pos="placeHere", fig.cap="Mean species diversity, split by gear type (weighted by catch and revenue).", echo=FALSE}
spDiv_personYear2 = effectiveDiversity_by_personYear2(cfec, variable1="specn", variable2 = "gearn")
spTot2 = summarize_effectiveDiversity2( spDiv_personYear2, variable2 = "gearn" )

# Use gear as the facet here
ggplot(spTot2, aes(year, mean.earn)) + geom_line(col="green") +
geom_line(aes(x=year,y=mean.lbs), col="red") + 
facet_wrap(~ gearn, scales="free_y") + 
ggtitle("Species diversity ('specn') [green='$', red=weight]") + 
xlab("Year") + ylab("Effective diversity")
```

```{r, fig.pos="placeHere", fig.cap="CV species diversity, split by gear type (weighted by catch and revenue).", echo=FALSE}

# Use gear as the facet here
ggplot(spTot2, aes(year, sd.earn/mean.earn)) + geom_line(col="green") +
geom_line(aes(x=year,y=sd.lbs/mean.lbs), col="red") + 
facet_wrap(~ gearn, scales="free_y") + 
ggtitle("CV of species diversity ('specn') [green='$', red=weight]") + 
xlab("Year") + ylab("CV of effective diversity")
```

```{r, fig.pos="placeHere", fig.cap="Mean species diversity, split by location (weighted by catch and revenue).", echo=FALSE}
spDiv_personYear2 = effectiveDiversity_by_personYear2(cfec, variable1="specn", variable2 = "landedLoc")
spTot2 = summarize_effectiveDiversity2( spDiv_personYear2, variable2 = "landedLoc" )

# Use landedLoc as the facet here
ggplot(spTot2, aes(year, mean.earn)) + geom_line(col="green") +
geom_line(aes(x=year,y=mean.lbs), col="red") + 
facet_wrap(~ landedLoc, scales="free_y") + 
ggtitle("Species diversity ('specn') [green='$', red=weight]") + 
xlab("Year") + ylab("Effective diversity")
```

```{r, fig.pos="placeHere", fig.cap="CV species diversity, split by location (weighted by catch and revenue).", echo=FALSE}

# Use gear as the facet here
ggplot(spTot2, aes(year, sd.earn/mean.earn)) + geom_line(col="green") +
geom_line(aes(x=year,y=sd.lbs/mean.lbs), col="red") + 
facet_wrap(~ landedLoc, scales="free_y") + 
ggtitle("CV of species diversity ('specn') [green='$', red=weight]") + 
xlab("Year") + ylab("CV of effective diversity")
```

```{r, fig.pos="placeHere", fig.cap="Mean species diversity, split by fishery (weighted by catch and revenue).", echo=FALSE}
cfec$permit_coarse = substr(cfec$p_fshy,1,1)
spDiv_personYear2 = effectiveDiversity_by_personYear2(cfec, variable1="specn", variable2 = "permit_coarse")
spTot2 = summarize_effectiveDiversity2( spDiv_personYear2, variable2 = "permit_coarse" )

# Use landedLoc as the facet here
ggplot(spTot2, aes(year, mean.earn)) + geom_line(col="green") +
geom_line(aes(x=year,y=mean.lbs), col="red") + 
facet_wrap(~ permit_coarse, scales="free_y") + 
ggtitle("Species diversity ('specn') [green='$', red=weight]") + 
xlab("Year") + ylab("Effective diversity")
```

```{r, fig.pos="placeHere", fig.cap="CV species diversity, split by fishery (weighted by catch and revenue).", echo=FALSE}
# Use gear as the facet here
ggplot(spTot2, aes(year, sd.earn/mean.earn)) + geom_line(col="green") +
geom_line(aes(x=year,y=sd.lbs/mean.lbs), col="red") + 
facet_wrap(~ permit_coarse, scales="free_y") + 
ggtitle("CV of species diversity ('specn') [green='$', red=weight]") + 
xlab("Year") + ylab("CV of effective diversity")
```

```{r, fig.pos="placeHere", fig.cap="Mean gear diversity, split by area (weighted by catch and revenue).", echo=FALSE}
cfec$permit_coarse = substr(cfec$p_fshy,1,1)
spDiv_personYear2 = effectiveDiversity_by_personYear2(cfec, variable1="gearn", variable2 = "landedLoc")
spTot2 = summarize_effectiveDiversity2( spDiv_personYear2, variable2 = "landedLoc" )

# Use landedLoc as the facet here
ggplot(spTot2, aes(year, mean.earn)) + geom_line(col="green") +
geom_line(aes(x=year,y=mean.lbs), col="red") + 
facet_wrap(~ landedLoc, scales="free_y") + 
ggtitle("Gear diversity ('gearn') [green='$', red=weight]") + 
xlab("Year") + ylab("Effective diversity")
```

```{r, fig.pos="placeHere", fig.cap="CV species diversity, split by fishery (weighted by catch and revenue).", echo=FALSE}
# Use gear as the facet here
ggplot(spTot2, aes(year, sd.earn/mean.earn)) + geom_line(col="green") +
geom_line(aes(x=year,y=sd.lbs/mean.lbs), col="red") + 
facet_wrap(~ landedLoc, scales="free_y") + 
ggtitle("CV of gear diversity ('gearn') [green='$', red=weight]") + 
xlab("Year") + ylab("CV of effective diversity")
```

```{r, fig.pos="placeHere", fig.cap="Mean permit diversity, split by area (weighted by catch and revenue).", echo=FALSE}
spDiv_personYear2 = effectiveDiversity_by_personYear2(cfec, variable1="p_fshy", variable2 = "landedLoc")
spTot2 = summarize_effectiveDiversity2( spDiv_personYear2, variable2 = "landedLoc" )

# Use landedLoc as the facet here
ggplot(spTot2, aes(year, mean.earn)) + geom_line(col="green") +
geom_line(aes(x=year,y=mean.lbs), col="red") + 
facet_wrap(~ landedLoc, scales="free_y") + 
ggtitle("Permit diversity ('p_fshy') [green='$', red=weight]") + 
xlab("Year") + ylab("Effective diversity")
```

```{r, fig.pos="placeHere", fig.cap="CV permit diversity, split by fishery (weighted by catch and revenue).", echo=FALSE}
# Use gear as the facet here
ggplot(spTot2, aes(year, sd.earn/mean.earn)) + geom_line(col="green") +
geom_line(aes(x=year,y=sd.lbs/mean.lbs), col="red") + 
facet_wrap(~ landedLoc, scales="free_y") + 
ggtitle("CV of permit diversity ('p_fshy') [green='$', red=weight]") + 
xlab("Year") + ylab("CV of effective diversity")
```

```{r, fig.pos="placeHere", fig.cap="Mean gear diversity, split by permits (weighted by catch and revenue).", echo=FALSE}
spDiv_personYear2 = effectiveDiversity_by_personYear2(cfec, variable1="gearn", variable2 = "permit_coarse")
spTot2 = summarize_effectiveDiversity2( spDiv_personYear2, variable2 = "permit_coarse" )

# Use landedLoc as the facet here
ggplot(spTot2, aes(year, mean.earn)) + geom_line(col="green") +
geom_line(aes(x=year,y=mean.lbs), col="red") + 
facet_wrap(~ permit_coarse, scales="free_y") + 
ggtitle("Permit gear diversity ('gearn') [green='$', red=weight]") + 
xlab("Year") + ylab("Effective diversity")
```

```{r, fig.pos="placeHere", fig.cap="CV gear diversity, split by permit (weighted by catch and revenue).", echo=FALSE}
# Use gear as the facet here
ggplot(spTot2, aes(year, sd.earn/mean.earn)) + geom_line(col="green") +
geom_line(aes(x=year,y=sd.lbs/mean.lbs), col="red") + 
facet_wrap(~ permit_coarse, scales="free_y") + 
ggtitle("CV of gear diversity ('p_fshy') [green='$', red=weight]") + 
xlab("Year") + ylab("CV of effective diversity")
```

## Filtering of unique people:vessel combinations

Filter out people who only use 1 vessel / year , and similarly filter out boats with more than 1 person / year. Some people might fish on multiple vessels within a year, and vessels might be used by multiple people within a year.

Now we'll filter out individuals who only used 1 boat in a year, and boats with only one permit holder fishing in a year.  
```{r}
cfecnew = cfec
cfec.1boat = group_by(cfecnew, p_holder, year) %>% 
  summarize(nBoat = length(unique(cadfg))) %>%
  filter(nBoat==1) %>% 
  left_join(cfecnew)

# Deal with setnet fisheries, where lots of p_holder
# share same codes
cfec.1boat = group_by(cfec.1boat, cadfg, year) %>% 
  summarize(nPerson = length(unique(p_holder))) %>%
  filter((nPerson==1 & cadfg %in% c(27960,27961) == FALSE) | cadfg %in% c(27960,27961)) %>% 
  left_join(cfec.1boat)

cfecnew = cfec.1boat
dim(cfecnew)[1]

#Now we can create a unique *id* column for individual - vessel combinations.  
cfecnew$id = paste(cfecnew$p_holder,cfecnew$cadfg)
# calculate number of unique person - vessel pairs ~ 40700
length(unique(cfecnew$id))
```
### Re-calculating diversity

```{r}
spDiv_personYear = effectiveDiversity_by_personYear(cfecnew, variable="specn")
spTot = summarize_effectiveDiversity( spDiv_personYear )

gearDiv_personYear = effectiveDiversity_by_personYear(cfecnew, variable="gearn")
gearTot = summarize_effectiveDiversity( gearDiv_personYear )

fishDiv_personYear = effectiveDiversity_by_personYear(cfecnew, variable="p_fshy")
fishTot = summarize_effectiveDiversity( fishDiv_personYear )
```

### Species

```{r, fig.pos = "placeHere", fig.cap = "Species diversity by frequency of fish tickets of unique id:vessel combinations (excluding people who fish multiple vessels / year, or vessels that are fished by multiple individuals in a year.", echo=FALSE}

ggplot(spTot, aes(x=year, y=mean.freq)) + geom_line(col="blue") +
  geom_ribbon(aes(ymin=mean.freq - sd.freq/mean.freq, 
  ymax=mean.freq + sd.freq/mean.freq), alpha=0.2, fill="blue") + 
ggtitle("Effective species diversity ('specn') by fishTix") + 
xlab("Year") + ylab("Mean (+/- CV)")
```

```{r, fig.pos = "placeHere", fig.cap = "Species diversity by revenue and landed weight of unique id:vessel combinations (excluding people who fish multiple vessels / year, or vessels that are fished by multiple individuals in a year.", echo=FALSE}

# Plot by earning / weight on diff scale
ggplot(spTot, aes(x=year, y=mean.earn)) + geom_line(col="red") + 
geom_ribbon(aes(x = year, ymin=mean.earn - sd.earn/mean.earn, 
  ymax = mean.earn + sd.earn/mean.earn), alpha=0.2, fill="red") + 
geom_line(aes(x=year,y=mean.lbs), col="green") + 
geom_ribbon(aes(x = year, ymin=mean.lbs - sd.lbs/mean.lbs, 
  ymax=mean.lbs + sd.lbs/mean.lbs), alpha=0.2, fill="green") +
ggtitle("Effective species diversity ('specn') [red = $, green=weight]") + 
xlab("Year") + ylab("Mean (+/- CV)")
```

### Gears

```{r, fig.pos = "placeHere", fig.cap = "Gear diversity by fish tickets of unique id:vessel combinations (excluding people who fish multiple vessels / year, or vessels that are fished by multiple individuals in a year.", echo=FALSE}
ggplot(gearTot, aes(x=year, y=mean.freq)) + geom_line(col="blue") + 
geom_ribbon(aes(ymin=mean.freq - sd.freq/mean.freq, 
  ymax=mean.freq + sd.freq/mean.freq), alpha=0.2, fill="blue") + 
ggtitle("Effective gear diversity ('gearn') by fishTix") + 
xlab("Year") + ylab("Mean (+/- CV)")
```

```{r, fig.pos = "placeHere", fig.cap = "Gear diversity by revenue and landed weight of unique id:vessel combinations (excluding people who fish multiple vessels / year, or vessels that are fished by multiple individuals in a year.", echo=FALSE}
# Plot by earning / weight on diff scale
ggplot(gearTot, aes(x=year, y=mean.earn)) + geom_line(col="red") + 
geom_ribbon(aes(x = year, ymin=mean.earn - sd.earn/mean.earn, 
  ymax = mean.earn + sd.earn/mean.earn), alpha=0.2, fill="red") + 
geom_line(aes(x=year,y=mean.lbs), col="green") + 
geom_ribbon(aes(x = year, ymin=mean.lbs - sd.lbs/mean.lbs, 
  ymax=mean.lbs + sd.lbs/mean.lbs), alpha=0.2, fill="green") + 
ggtitle("Effective gear diversity ('gearn') [red = $, green=weight]") + 
xlab("Year") + ylab("Mean (+/- CV)")
```

### Fisheries 
```{r, fig.pos = "placeHere", fig.cap = "Permit diversity by fish ticket frequency of unique id:vessel combinations (excluding people who fish multiple vessels / year, or vessels that are fished by multiple individuals in a year.", echo=FALSE}

ggplot(fishTot, aes(x=year, y=mean.freq)) + geom_line(col="blue") +
geom_ribbon(aes(ymin=mean.freq - sd.freq/mean.freq, 
  ymax=mean.freq + sd.freq/mean.freq), alpha=0.2, fill="blue") + 
ggtitle("Effective permit diversity ('p_fshy') by fishTix") + 
xlab("Year") + ylab("Mean (+/- CV)")
```

```{r, fig.pos = "placeHere", fig.cap = "Permit diversity by revenue and landed weight of unique id:vessel combinations (excluding people who fish multiple vessels / year, or vessels that are fished by multiple individuals in a year.", echo=FALSE}

# Plot by earning / weight on diff scale
ggplot(fishTot, aes(x=year, y=mean.earn)) + geom_line(col="red") + 
geom_ribbon(aes(x = year, ymin=mean.earn - sd.earn/mean.earn, 
  ymax = mean.earn + sd.earn/mean.earn), alpha=0.2, fill="red") + 
geom_line(aes(x=year,y=mean.lbs), col="green") + 
geom_ribbon(aes(x = year, ymin=mean.lbs - sd.lbs/mean.lbs, 
  ymax=mean.lbs + sd.lbs/mean.lbs), alpha=0.2, fill="green") + 
ggtitle("Effective permit diversity ('p_fshy') [red = $, green=weight]") + 
xlab("Year") + ylab("Mean (+/- CV)")
```

## Comparing diversity between vessels and individuals

Just out of curiousity, we can summarize diversity by year taking the approach used by Kasperski and Holland (based on CADFG number) and compare that to the same criteria using people ("p_holder"). We'll only do this comparison on the species diversity, but use the same restriction: use only vessels or people who had participated continuously for the entire time period (this throws out 90%+ of the data).

First, based on vessels:
```{r}
contVessels = group_by(cfec, cadfg) %>% 
  summarize(nYear = length(unique(year))) %>% 
  filter(nYear == 30)

cfecnew = left_join(contVessels, cfec)

# calculate diversity by species by CADFG (vessel #)
effDiv.vessel = group_by(cfecnew, specn, year, cadfg) %>%
  summarize(totRev = sum(g_earn, na.rm=T),
  totN = n(),
  totWeight = sum(g_pounds, na.rm=T),
    days = length(unique(day))) %>% 
  as.data.frame() %>%
  filter(is.finite(totRev) & is.finite(totN) & is.finite(totWeight)) %>%   
  as.data.frame() %>%
  group_by(cadfg, year) %>% 
  summarize(eff.freq = simp.div(totN),
  eff.earn = simp.div(totRev),
  eff.lbs = simp.div(totWeight),
    totIndRev = sum(totRev),
    nCalDays = sum(days)) %>% 
  filter(!is.na(eff.freq) & !is.na(eff.earn) & !is.na(eff.lbs)) %>%
  as.data.frame()

effDiv.vessel = summarize_effectiveDiversity(effDiv.vessel)

effDiv.person = effectiveDiversity_by_personYear(cfecnew, variable="specn")
effDiv.person = summarize_effectiveDiversity( effDiv.person )
```

```{r, fig.pos="placeHere", fig.cap="Comparison of species diversity using vessel versus individuals", echo=FALSE}
ggplot(effDiv.vessel, aes(x=year, y=mean.earn)) + geom_line(col="red") + 
geom_ribbon(aes(x = year, ymin=mean.earn - sd.earn/mean.earn, 
  ymax = mean.earn + sd.earn/mean.earn), alpha=0.2, fill="red") + 
geom_line(aes(x=year,y=effDiv.person$mean.earn), col="green") + 
geom_ribbon(aes(x = year, ymin=effDiv.person$mean.earn - 
    effDiv.person$sd.earn/effDiv.person$mean.earn, 
  ymax=effDiv.person$mean.earn + effDiv.person$sd.earn/effDiv.person$mean.earn), 
  alpha=0.2, fill="green") + 
ggtitle("Effective species diversity ('specn') [red = vessel, green=person]") + 
xlab("Year") + ylab("Mean (+/- CV)")
```

Now make the portfolio plots from the metafolio package:

```{r, echo=FALSE}
species_diversity_by_year <-
  effectiveDiversity_by_personYear(cfec, variable = "specn")
breaks = c(1, 1.01, 1.5, 2, 2.5, 3, 3.5, 4, 10)
species_diversity <-
  group_by(species_diversity_by_year, p_holder) %>%
  summarize(
    # diversity_by_frequency = mean(eff.freq),
    diversity_by_earnings = mean(eff.earn),
    # diversity_by_weight = mean(eff.lbs),
    m = log10(mean(totIndRev)),
    v = sd(totIndRev) / mean(totIndRev)
  ) %>%
  filter(!is.na(v)) %>%
  mutate(diversity_group = cut(diversity_by_earnings, breaks = breaks,
    right = FALSE))

species_diversity_list <- split(species_diversity,
  species_diversity$diversity_group)

#pdf("earnings_portfolio_by_species_diversity.pdf", height = 5, width = 5)
par(cex = 0.8)
metafolio::plot_cons_plans(
  species_diversity_list,
  plans_name = names(species_diversity_list),
  cols = RColorBrewer::brewer.pal(length(names(species_diversity_list)),
    "YlOrRd"),
  xlab = "CV of gross earnings",
  ylab = "log10 of gross earnings",
  add_all_efs = FALSE)
#dev.off()
```

### 2016-06-09 analysis_portfolio/1.5-createData_linear_modeling.Rmd
committing updates to all papers

--- 2016-06-09 analysis_diversity/paper1a_diversity_metrics_4-20.Rmd
@@
 ---
-title: 'Changes in diversity and variability in diversity through time: Catch portfolios
-  paper 1A'
+title: "createData_linear_modeling"
+author: ""
+date: "June 05, 2016"
 output:
@@
   html_document: default
-date: "April 7, 2016"
 ---
 
-```{r setup, include=FALSE}
+```{r setup, include=FALSE, echo=FALSE}
 knitr::opts_chunk$set(echo = TRUE, tidy=FALSE, tidy.opts=list(width.cutoff=60), warning = FALSE, message = FALSE)
@@
 library(randomForest)
-load("cfecCleaned2.Rdata")
+library(date)
+
+#load("cfecCleaned2.Rdata")
+cfec <- feather::read_feather("data/cfec.feather")
 
@@
 
-# Pull in "new.locs" from Anne/Jordan/Rich/Ben/Ole
-newlocs = read.csv("new.locs.csv")
-newlocs$p_fshy = as.character(newlocs$p_fshy)
-
-# Match cfec codes in newlocs to identify new 
-# locations (Anne/Jordan/Rich/Ben/Ole)
-cfec$landedLoc = newlocs$New.Loc[match(cfec$p_fshy, newlocs$p_fshy)]
-cfec$landedLoc[which(is.na(cfec$landedLoc))] = "Statewide" # C04B
-
+region<-read.csv("areas/regions.csv")
+region<-region[,c(1:2)] #don’t include “assigned by” column
+cfec<-left_join(cfec,region)
+#cfec$region <- replace(cfec$region, is.na(cfec$region), "Unassigned")
 ```
@@
 Note the C++ version is commented out below,
-```{r, warning = FALSE, message=FALSE}
+
+```{r, warning = FALSE, message=FALSE, echo=FALSE}
 
@@
 
-## Diversity by species, gear, and fisheries.
-
-We have 4 dimensions we're interested in calculating diversity (species, gears, fisheries, space).
-And for each dimension, there's 3 ways we can calculate diversity: diversity based on fish ticket frequency, diversity based on revenue of each (weighted by g_earn) and diversity based on the weight of catch (g_pounds).
-
-To calculate the uncertainty (CV, quartiles, whatever) we need to calculate the diversity for each individual in each year (p_holder), and then summarize variability across individuals.
-We'll do this two ways: first using the whole dataset, and then using Eric's more restricted version (limiting the analysis to single people : vessel combinations)
-
-### Calculations
-
-We can calculate diversity across all metrics in the same block.
-
-```{r}
-cfec$permit_coarse = substr(cfec$p_fshy,1,1)
-spDiv_personYear = effectiveDiversity_by_personYear(cfec, variable="specn")
-spTot = summarize_effectiveDiversity( spDiv_personYear )
-
-gearDiv_personYear = effectiveDiversity_by_personYear(cfec, variable="gearn")
-gearTot = summarize_effectiveDiversity( gearDiv_personYear )
-
-fishDiv_personYear = effectiveDiversity_by_personYear(cfec, variable="permit_coarse")
-fishTot = summarize_effectiveDiversity( fishDiv_personYear )
-
-areaDiv_personYear = effectiveDiversity_by_personYear(cfec, variable="area")
-areaTot = summarize_effectiveDiversity( areaDiv_personYear )
-```
-
-### Summarizing landings and value
-
-First, it's helpful to identify the broad patterns in the CFEC data, showing time series of people, landings, and revenue by year.
-The general patterns are consistent with Kasperski & Holland's analysis that showed diversity declining.
-More specificially,
-
-- Total participation decreased after ~ 1990, both in terms of people + boats  
-- Participation has been flat since early 2000s, but since 2000
-- both catch (lbs landed) / person, and revenue / person have been increasing
-
-```{r}
-cfecSummarize = group_by(cfec, year) %>%
-  summarize(people = length(unique(p_holder)), 
-    revenue = sum(g_earn), catch = sum(g_pounds),
-    vessels = length(unique(cadfg))) %>%
-  as.data.frame
-par(mfrow = c(3,2), mgp=c(2,1,0), mai=c(0.5,0.5,0.05,0.05))
-plot(cfecSummarize$year, cfecSummarize$people/1000, xlab="", ylab="People (1000s)", type="b", lwd=3)
-plot(cfecSummarize$year, cfecSummarize$revenue/1e+09, xlab="", ylab="Billion $s", type="b", lwd=3)
-plot(cfecSummarize$year, cfecSummarize$catch/2e+06, xlab="", ylab="Weight (tons)", type="b", lwd=3)
-plot(cfecSummarize$year, cfecSummarize$revenue/cfecSummarize$people, xlab="", ylab="$s / person", type="b", lwd=3)
-plot(cfecSummarize$year, cfecSummarize$catch/cfecSummarize$people, xlab="", ylab="Lbs / person", type="b", lwd=3)
-plot(cfecSummarize$year, cfecSummarize$vessels/cfecSummarize$people, xlab="", ylab="Boats / person", type="b", lwd=3)
-
-```
-
-### Patterns in diversity metrics
-
-We can coarsely group individuals based on their species, gear, and fisheries diversity through time (e.g.
-1-1.5, 1.5-2, 2-2.5, etc.) and make boxplots of revenue by group over time (similar to Jordan's boxplots).
-```{r}
-brks = c(1, 1.0001, 1.5, 2, 2.5, 3, 3.5, 4, 10)
-
-spDiv_personYear$group = cut(spDiv_personYear$eff.earn, breaks = brks, right=FALSE)
-gearDiv_personYear$group = cut(gearDiv_personYear$eff.earn, breaks = brks, right=FALSE)
-fishDiv_personYear$group = cut(fishDiv_personYear$eff.earn, breaks = brks, right=FALSE)
-areaDiv_personYear$group = cut(areaDiv_personYear$eff.earn, breaks = brks, right=FALSE)
-```
-
-We can start by just plotting changes in group membership over time. The interesting patterns here are:
-1. Species: If we lump the lowest diversity groups (1 - 1.5 spp), there's been a slight uptick from ~ 55% - 65% (consistent with specialization).  
-2. Gears / Permits: people fishing 1 gear or 1 permit have also increased since ~ 1991, from ~ 70% to > 80% of all records.  
-3. There's no interesting trend with respect to regions; 75-80% of the people fish 1 region, and there's no trend. [Revisit after getting Anne's areas]
-
-```{r, fig.pos="placeHere", fig.cap="Proportion of people by species diversity groups"}
-group.sum = group_by(spDiv_personYear, year, group) %>% 
-  summarize(n = length(unique(p_holder))) %>% 
-  as.data.frame()
-tot = group_by(group.sum, year) %>% 
-  summarize(tot = sum(n))
-group.sum.spec = left_join(group.sum, tot)
-ggplot(group.sum.spec, aes(x=year, y=n/tot, group=group, colour=group ) ) + geom_line() +
-  ggtitle("P_holder by species diversity groups") + 
-xlab("Year") + ylab("Percent of total") 
-```
-
-```{r, fig.pos="placeHere", fig.cap="Proportion of people by gear diversity groups"}
-group.sum = group_by(gearDiv_personYear, year, group) %>% 
-  summarize(n = length(unique(p_holder))) %>% 
-  as.data.frame()
-tot = group_by(group.sum, year) %>% 
-  summarize(tot = sum(n))
-group.sum.gear = left_join(group.sum, tot)
-ggplot(group.sum.gear, aes(x=year, y=n/tot, group=group, colour=group ) ) + geom_line() +
-  ggtitle("P_holder by gear diversity groups") + 
-xlab("Year") + ylab("Percent of total") 
-```
-
-```{r, fig.pos="placeHere", fig.cap="Proportion of people by area diversity groups"}
-group.sum = group_by(areaDiv_personYear, year, group) %>% 
-  summarize(n = length(unique(p_holder))) %>% 
-  as.data.frame()
-tot = group_by(group.sum, year) %>% 
-  summarize(tot = sum(n))
-group.sum.area = left_join(group.sum, tot)
-ggplot(group.sum.area, aes(x=year, y=n/tot, group=group, colour=group ) ) + geom_line() +
-  ggtitle("P_holder by area diversity groups") + 
-xlab("Year") + ylab("Percent of total") 
-```
-
-```{r, fig.pos="placeHere", fig.cap="Proportion of people by permit diversity groups"}
-group.sum = group_by(fishDiv_personYear, year, group) %>% 
-  summarize(n = length(unique(p_holder))) %>% 
-  as.data.frame()
-tot = group_by(group.sum, year) %>% 
-  summarize(tot = sum(n))
-group.sum.fish = left_join(group.sum, tot)
-ggplot(group.sum.fish, aes(x=year, y=n/tot, group=group, colour=group ) ) + geom_line() +
-  ggtitle("P_holder by permit diversity groups") + 
-xlab("Year") + ylab("Percent of total") 
-```
-
-It's pretty easy as a next step to focus on specialists (gears, permits).
-Which gears and permits are being used, and how does this break down by area?
-This shows some pretty interesting patterns that we can characterize about 'specialists':
-
-- longline specialists greatly decreased after 2004 (going to 0 in alaska peninsula, cook inlet, huge drops in kodiak, smaller drops in PWS / Southeast).  
-- we've seen slight increases in drift gillnet specialists (gearn = 3) in Bristol Bay, and larger increases in PWS / Cook Inlet (presumably copper river sockeye)
-
-```{r}
-# Find all people in group 1 (specialists)
-sub = gearDiv_personYear[which(as.numeric(gearDiv_personYear$group)%in%c(1,2)),]
-LJ = left_join(sub, cfec)
-# Calculate number of people using each gear type
-LJ = group_by(LJ, gearn, year, landedLoc) %>% 
-  summarize(n = length(unique(p_holder)))
-# Calculate total to normalize to %
-tot = group_by(LJ, year, landedLoc) %>% 
-  summarize(tot = sum(n))
-LJ = left_join(LJ, tot)
-# Plot proportions
-ggplot(LJ, aes(x=year, y=n/tot, group=as.factor(gearn), colour=as.factor(gearn) ) ) + geom_line() + facet_wrap(~ landedLoc, scales="free_y") + ylab("Proportion of people")
-```
-
-The next is to pull out the specialists, by gear or species.
-For example we can pull out people specializing on salmon.
-For example, we could look at the trends in some of these metrics for people who are only using a gear specific to salmon (1, 3, 4).
-There's a few interesting things here:
-
--- Nobody uses > 2 gears
-
-```{r}
-salmonSpecialists = group_by(cfec, year, p_holder, landedLoc) %>% 
-  summarize(gears = length(unique(gearn)), 
-    salmonGears = length(unique(gearn%in%c(1,3,4))),
-    salmonSpec = length(unique(specn[specn%in%c(410,420,430,440,450)]))) %>% 
-  filter(salmonGears == gears & salmonSpec >= 1)
-
-salmonSummary = group_by(salmonSpecialists, year, landedLoc) %>% 
-  summarize(ngear = mean(gears), ngear2 = length(which(gears == 2)), npeople = length(unique(p_holder)), nspec = mean(salmonSpec)) %>% 
-  filter(landedLoc %in% c("Bering Sea", "Aleutian Islands", "Dutch Harbor", "Statewide", "GOA", "EEZ")==FALSE)
-
-ggplot(salmonSummary, aes(x=year, y=ngear2/npeople) ) + geom_line() + facet_wrap(~ landedLoc, scales="free_y") + ylab("% People using 2 salmon gears")
-
-ggplot(salmonSummary, aes(x=year, y=npeople) ) + geom_line() + facet_wrap(~ landedLoc, scales="free_y") + ylab("People exclusively using gears 1,3,4")
-
-ggplot(salmonSummary, aes(x=year, y=nspec) ) + geom_line() + facet_wrap(~ landedLoc, scales="free_y") + ylab("People exclusively using gears 1,3,4")
-
-salmonSpecialists = group_by(cfec, year, p_holder) %>% 
-  summarize(gears = length(unique(gearn)), 
-    salmonGears = length(unique(gearn%in%c(1,3,4))),
-    salmonSpec = length(unique(specn[specn%in%c(410,420,430,440,450)]))) %>% 
-  filter(salmonGears == gears & salmonSpec >= 1) 
-salmonSpecialists = left_join(salmonSpecialists, cfec)
-```
-
-```{r, fig.pos="placeHere", fig.cap="Revenue by effective species diversity and year", echo=FALSE}
-ggplot(spDiv_personYear, aes(x = factor(year), y = log10(totIndRev), fill = factor(group))) + 
-  geom_boxplot(outlier.shape=NA) + 
-  ggtitle("Revenue by species diversity ('specn') and year") + 
-xlab("Year") + ylab("log10(revenue)")
-```
-
-We can also try to calculate downside risk, by group. We'll start with the lower 95% for each group. 
-```{r, fig.pos="placeHere", fig.cap="Downside risk metrics: lower 5% and 20% of total revenue, by species diversity and year", echo=FALSE}
-spDiv_lower = group_by(spDiv_personYear, year, group) %>% 
-  summarize(lower95 = quantile(log10(totIndRev),0.05),
-    lower80 = quantile(log10(totIndRev),0.2))
-ggplot(spDiv_lower, aes(x=year, y=lower95, group=group, colour=group ) ) + 
-            geom_line(size=1) + geom_line(aes(x=year, y=lower80, group=group, colour=group ), size=1, linetype = 2) +
-  ggtitle("Lower 5% (solid) of and 20% (dashed) log10(revenue)") + 
-xlab("Year") + ylab("Lower 5% and 20%")  
-```
-
-```{r, fig.pos="placeHere", fig.cap="Revenue by effective species diversity across years", echo=FALSE}
-ggplot(spDiv_personYear, aes(x = factor(group), y = log10(totIndRev))) + 
-  geom_boxplot(outlier.shape=NA) + 
-  ggtitle("Revenue by species diversity ('specn')") + 
-xlab("Effective diversity") + ylab("log10(revenue)")
-```
-
-```{r, fig.pos="placeHere", fig.cap="Number of unique landed dates by effective species diversity across years", echo=FALSE, include=FALSE}
-ggplot(spDiv_personYear, aes(x = factor(group), y = log(nCalDays))) + 
-  geom_boxplot(outlier.shape=NA) + 
-  ggtitle("Unique landed dates by species diversity ('specn')") + 
-xlab("Effective diversity") + ylab("log(unique days)")
-```
-
-```{r, fig.pos="placeHere", fig.cap="Revenue by effective gear diversity and year", echo=FALSE}
-ggplot(gearDiv_personYear, aes(x = factor(year), y = log10(totIndRev), fill = factor(group))) + 
-  geom_boxplot(outlier.shape=NA) + 
-  ggtitle("Revenue by gear diversity ('gearn') and year") + 
-xlab("Year") + ylab("log10(revenue)")
-```
-
-```{r, fig.pos="placeHere", fig.cap="Downside risk metrics: lower 5% and 20% of total revenue, by gear diversity and year", echo=FALSE, include=FALSE}
-gearDiv_lower = group_by(gearDiv_personYear, year, group) %>% 
-  summarize(lower95 = quantile(log10(totIndRev),0.05),
-    lower80 = quantile(log10(totIndRev),0.2))
-ggplot(gearDiv_lower, aes(x=year, y=lower95, group=group, colour=group ) ) + 
-            geom_line(size=1) + geom_line(aes(x=year, y=lower80, group=group, colour=group ), size=1, linetype = 2) +
-  ggtitle("Lower 5% (solid) of and 20% (dashed) log10(revenue)") + 
-xlab("Year") + ylab("Lower 5% and 20%")  
-```
-
-```{r, fig.pos="placeHere", fig.cap="Revenue by effective gear diversity across years", echo=FALSE, include=FALSE}
-ggplot(gearDiv_personYear, aes(x = factor(group), y = log10(totIndRev))) + 
-  geom_boxplot(outlier.shape=NA) + 
-  ggtitle("Revenue by gear diversity ('gearn')") + 
-xlab("Effective diversity") + ylab("log10(revenue)")
-```
-
-```{r, fig.pos="placeHere", fig.cap="Number of unique landed dates by effective gear diversity across years", echo=FALSE, include=FALSE}
-ggplot(gearDiv_personYear, aes(x = factor(group), y = log(nCalDays))) + 
-  geom_boxplot(outlier.shape=NA) + 
-  ggtitle("Unique landed dates by gear diversity ('gearn')") + 
-xlab("Effective diversity") + ylab("log(unique days)")
-```
-
-```{r, fig.pos="placeHere", fig.cap="Revenue by effective permit diversity and year", echo=FALSE}
-ggplot(fishDiv_personYear, aes(x = factor(year), y = log10(totIndRev), fill = factor(group))) + 
-  geom_boxplot(outlier.shape=NA) + 
-  ggtitle("Revenue by permit diversity ('p_fshy') and year") + 
-xlab("Year") + ylab("log10(revenue)")
-```
-
-```{r, fig.pos="placeHere", fig.cap="Downside risk metrics: lower 5% and 20% of total revenue, by permit diversity and year", echo=FALSE, include=FALSE}
-fishDiv_lower = group_by(fishDiv_personYear, year, group) %>% 
-  summarize(lower95 = quantile(log10(totIndRev),0.05),
-    lower80 = quantile(log10(totIndRev),0.2))
-ggplot(fishDiv_lower, aes(x=year, y=lower95, group=group, colour=group ) ) + 
-            geom_line(size=1) + geom_line(aes(x=year, y=lower80, group=group, colour=group ), size=1, linetype = 2) +
-  ggtitle("Lower 5% (solid) of and 20% (dashed) log10(revenue)") + 
-xlab("Year") + ylab("Lower 5% and 20%")  
-```
-
-```{r, fig.pos="placeHere", fig.cap="Revenue by effective permit diversity across years", echo=FALSE, include=FALSE}
-ggplot(fishDiv_personYear, aes(x = factor(group), y = log10(totIndRev))) + 
-  geom_boxplot(outlier.shape=NA) + 
-  ggtitle("Revenue by permit diversity ('p_fshy')") + 
-xlab("Effective diversity") + ylab("log10(revenue)")
-```
-
-```{r, fig.pos="placeHere", fig.cap="Number of unique landed dates by effective permit diversity across years", echo=FALSE, include=FALSE}
-ggplot(fishDiv_personYear, aes(x = factor(group), y = log(nCalDays))) + 
-  geom_boxplot(outlier.shape=NA) + 
-  ggtitle("Unique landed dates by permit diversity ('p_fshy')") + 
-xlab("Effective diversity") + ylab("log(unique days)")
-```
-
-```{r, fig.pos="placeHere", fig.cap="Revenue by effective area diversity and year", echo=FALSE}
-ggplot(areaDiv_personYear, aes(x = factor(year), y = log10(totIndRev), fill = factor(group))) + 
-  geom_boxplot(outlier.shape=NA) + 
-  ggtitle("Revenue by area diversity ('landedLoc') and year") + 
-xlab("Year") + ylab("log10(revenue)")
-```
-
-```{r, fig.pos="placeHere", fig.cap="Revenue by area permit diversity across years", include=FALSE, echo=FALSE}
-ggplot(areaDiv_personYear, aes(x = factor(group), y = log10(totIndRev))) + 
-  geom_boxplot(outlier.shape=NA) + 
-  ggtitle("Revenue by area diversity ('landedLoc')") + 
-xlab("Effective diversity") + ylab("log10(revenue)")
-```
-
-```{r, fig.pos="placeHere", fig.cap="Number of unique landed dates by area permit diversity across years", echo=FALSE, include=FALSE}
-ggplot(fishDiv_personYear, aes(x = factor(group), y = log(nCalDays))) + 
-  geom_boxplot(outlier.shape=NA) + 
-  ggtitle("Unique landed dates by area diversity ('landedLoc')") + 
-xlab("Effective diversity") + ylab("log(unique days)")
-```
-
-### Species
-
-```{r, fig.pos="placeHere", fig.cap="Species diversity (all individuals/vessels). For scaling purposes, 1.2 is subtracted from the diversity based on fish ticket frequency.", echo=FALSE}
-# Plot by earning / weight on diff scale
-ggplot(spTot, aes(x=year, y=mean.earn)) + geom_line(col="green") + 
-geom_line(aes(x=year,y=mean.lbs), col="red") + 
-geom_line(aes(x=year,y=mean.freq-1.2), col="blue") + 
-ggtitle("Effective species diversity ('specn') [green='$', red=weight, blue = %]") + 
-xlab("Year") + ylab("Effective diversity")
-```
-
-```{r, fig.pos="placeHere", fig.cap="CV of Species diversity (all individuals/vessels)", echo=FALSE}
-# Plot by earning / weight on diff scale
-ggplot(spTot, aes(x=year, y=sd.earn/mean.earn)) + geom_line(col="green") + 
-geom_line(aes(x=year,y=sd.lbs/mean.lbs), col="red") + 
-geom_line(aes(x=year,y=sd.freq/mean.freq), col="blue") + 
-ggtitle("CV of species diversity ('specn') [green='$', red=weight, blue = %]") + 
-xlab("Year") + ylab("CV")
-```
-
-### Gears
-
-```{r, fig.pos="placeHere", fig.cap="Gear diversity (all individuals/vessels)", echo=FALSE}
-# Plot by earning / weight on diff scale
-ggplot(gearTot, aes(x=year, y=mean.earn)) + geom_line(col="green") + 
-geom_line(aes(x=year,y=mean.lbs), col="red") + 
-geom_line(aes(x=year,y=mean.freq), col="blue") + 
-ggtitle("Effective gear diversity ('gearn') [green='$', red=weight, blue = %]") + 
-xlab("Year") + ylab("Effective diversity")
-```
-
-```{r, fig.pos="placeHere", fig.cap="CV of gear diversity (all individuals/vessels)", echo=FALSE}
-# Plot by earning / weight on diff scale
-ggplot(gearTot, aes(x=year, y=sd.earn/mean.earn)) + geom_line(col="green") + 
-geom_line(aes(x=year,y=sd.lbs/mean.lbs), col="red") + 
-geom_line(aes(x=year,y=sd.freq/mean.freq), col="blue") + 
-ggtitle("CV of gear diversity ('gearn') [green='$', red=weight, blue = %]") + 
-xlab("Year") + ylab("CV")
-```
-
-### Fisheries
-
-```{r, fig.pos="placeHere", fig.cap="Permit diversity (all individuals/vessels)", echo=FALSE}
-# Plot by earning / weight on diff scale
-ggplot(fishTot, aes(x=year, y=mean.earn)) + geom_line(col="green") + 
-geom_line(aes(x=year,y=mean.lbs), col="red") + 
-geom_line(aes(x=year,y=mean.freq), col="blue") + 
-ggtitle("Effective permit diversity ('p_fshy') [green='$', red=weight, blue = %]") + 
-xlab("Year") + ylab("Effective diversity")
-```
-
-```{r, fig.pos="placeHere", fig.cap="CV of permit diversity (all individuals/vessels)", echo=FALSE}
-# Plot by earning / weight on diff scale
-ggplot(fishTot, aes(x=year, y=sd.earn/mean.earn)) + geom_line(col="green") + 
-geom_line(aes(x=year,y=sd.lbs/mean.lbs), col="red") + 
-geom_line(aes(x=year,y=sd.freq/mean.freq), col="blue") + 
-ggtitle("CV of permit diversity ('p_fshy') [green='$', red=weight, blue = %]") + 
-xlab("Year") + ylab("CV")
-```
-
-### Areas (including EEZ / Statewide / GOA categories)
-
-```{r, fig.pos="placeHere", fig.cap="Area diversity (all individuals/vessels)", echo=FALSE}
-# Plot by earning / weight on diff scale
-ggplot(areaTot, aes(x=year, y=mean.earn)) + geom_line(col="green") + 
-geom_line(aes(x=year,y=mean.lbs), col="red") + 
-geom_line(aes(x=year,y=mean.freq), col="blue") + 
-ggtitle("Effective area diversity ('p_fshy') [green='$', red=weight, blue = %]") + 
-xlab("Year") + ylab("Effective diversity")
-```
-
-```{r, fig.pos="placeHere", fig.cap="CV of area diversity (all individuals/vessels)", echo=FALSE}
-# Plot by earning / weight on diff scale
-ggplot(areaTot, aes(x=year, y=sd.earn/mean.earn)) + geom_line(col="green") + 
-geom_line(aes(x=year,y=sd.lbs/mean.lbs), col="red") + 
-geom_line(aes(x=year,y=sd.freq/mean.freq), col="blue") + 
-ggtitle("CV of area diversity ('landedLoc') [green='$', red=weight, blue = %]") + 
-xlab("Year") + ylab("CV")
-```
-
-## Finer scale exploration of patterns
-
-We can also look further into whether the observed trends in mean and CV of revenue vary by gear type, area, permit type, etc.
-
-```{r, fig.pos="placeHere", fig.cap="Mean species diversity, split by gear type (weighted by catch and revenue).", echo=FALSE}
-spDiv_personYear2 = effectiveDiversity_by_personYear2(cfec, variable1="specn", variable2 = "gearn")
-spTot2 = summarize_effectiveDiversity2( spDiv_personYear2, variable2 = "gearn" )
-
-# Use gear as the facet here
-ggplot(spTot2, aes(year, mean.earn)) + geom_line(col="green") +
-geom_line(aes(x=year,y=mean.lbs), col="red") + 
-facet_wrap(~ gearn, scales="free_y") + 
-ggtitle("Species diversity ('specn') [green='$', red=weight]") + 
-xlab("Year") + ylab("Effective diversity")
-```
-
-```{r, fig.pos="placeHere", fig.cap="CV species diversity, split by gear type (weighted by catch and revenue).", echo=FALSE}
-
-# Use gear as the facet here
-ggplot(spTot2, aes(year, sd.earn/mean.earn)) + geom_line(col="green") +
-geom_line(aes(x=year,y=sd.lbs/mean.lbs), col="red") + 
-facet_wrap(~ gearn, scales="free_y") + 
-ggtitle("CV of species diversity ('specn') [green='$', red=weight]") + 
-xlab("Year") + ylab("CV of effective diversity")
-```
-
-```{r, fig.pos="placeHere", fig.cap="Mean species diversity, split by location (weighted by catch and revenue).", echo=FALSE}
-spDiv_personYear2 = effectiveDiversity_by_personYear2(cfec, variable1="specn", variable2 = "landedLoc")
-spTot2 = summarize_effectiveDiversity2( spDiv_personYear2, variable2 = "landedLoc" )
-
-# Use landedLoc as the facet here
-ggplot(spTot2, aes(year, mean.earn)) + geom_line(col="green") +
-geom_line(aes(x=year,y=mean.lbs), col="red") + 
-facet_wrap(~ landedLoc, scales="free_y") + 
-ggtitle("Species diversity ('specn') [green='$', red=weight]") + 
-xlab("Year") + ylab("Effective diversity")
-```
-
-```{r, fig.pos="placeHere", fig.cap="CV species diversity, split by location (weighted by catch and revenue).", echo=FALSE}
-
-# Use gear as the facet here
-ggplot(spTot2, aes(year, sd.earn/mean.earn)) + geom_line(col="green") +
-geom_line(aes(x=year,y=sd.lbs/mean.lbs), col="red") + 
-facet_wrap(~ landedLoc, scales="free_y") + 
-ggtitle("CV of species diversity ('specn') [green='$', red=weight]") + 
-xlab("Year") + ylab("CV of effective diversity")
-```
-
-```{r, fig.pos="placeHere", fig.cap="Mean species diversity, split by fishery (weighted by catch and revenue).", echo=FALSE}
-cfec$permit_coarse = substr(cfec$p_fshy,1,1)
-spDiv_personYear2 = effectiveDiversity_by_personYear2(cfec, variable1="specn", variable2 = "permit_coarse")
-spTot2 = summarize_effectiveDiversity2( spDiv_personYear2, variable2 = "permit_coarse" )
-
-# Use landedLoc as the facet here
-ggplot(spTot2, aes(year, mean.earn)) + geom_line(col="green") +
-geom_line(aes(x=year,y=mean.lbs), col="red") + 
-facet_wrap(~ permit_coarse, scales="free_y") + 
-ggtitle("Species diversity ('specn') [green='$', red=weight]") + 
-xlab("Year") + ylab("Effective diversity")
-```
-
-```{r, fig.pos="placeHere", fig.cap="CV species diversity, split by fishery (weighted by catch and revenue).", echo=FALSE}
-# Use gear as the facet here
-ggplot(spTot2, aes(year, sd.earn/mean.earn)) + geom_line(col="green") +
-geom_line(aes(x=year,y=sd.lbs/mean.lbs), col="red") + 
-facet_wrap(~ permit_coarse, scales="free_y") + 
-ggtitle("CV of species diversity ('specn') [green='$', red=weight]") + 
-xlab("Year") + ylab("CV of effective diversity")
-```
-
-```{r, fig.pos="placeHere", fig.cap="Mean gear diversity, split by area (weighted by catch and revenue).", echo=FALSE}
-cfec$permit_coarse = substr(cfec$p_fshy,1,1)
-spDiv_personYear2 = effectiveDiversity_by_personYear2(cfec, variable1="gearn", variable2 = "landedLoc")
-spTot2 = summarize_effectiveDiversity2( spDiv_personYear2, variable2 = "landedLoc" )
-
-# Use landedLoc as the facet here
-ggplot(spTot2, aes(year, mean.earn)) + geom_line(col="green") +
-geom_line(aes(x=year,y=mean.lbs), col="red") + 
-facet_wrap(~ landedLoc, scales="free_y") + 
-ggtitle("Gear diversity ('gearn') [green='$', red=weight]") + 
-xlab("Year") + ylab("Effective diversity")
-```
-
-```{r, fig.pos="placeHere", fig.cap="CV species diversity, split by fishery (weighted by catch and revenue).", echo=FALSE}
-# Use gear as the facet here
-ggplot(spTot2, aes(year, sd.earn/mean.earn)) + geom_line(col="green") +
-geom_line(aes(x=year,y=sd.lbs/mean.lbs), col="red") + 
-facet_wrap(~ landedLoc, scales="free_y") + 
-ggtitle("CV of gear diversity ('gearn') [green='$', red=weight]") + 
-xlab("Year") + ylab("CV of effective diversity")
-```
-
-```{r, fig.pos="placeHere", fig.cap="Mean permit diversity, split by area (weighted by catch and revenue).", echo=FALSE}
-spDiv_personYear2 = effectiveDiversity_by_personYear2(cfec, variable1="p_fshy", variable2 = "landedLoc")
-spTot2 = summarize_effectiveDiversity2( spDiv_personYear2, variable2 = "landedLoc" )
-
-# Use landedLoc as the facet here
-ggplot(spTot2, aes(year, mean.earn)) + geom_line(col="green") +
-geom_line(aes(x=year,y=mean.lbs), col="red") + 
-facet_wrap(~ landedLoc, scales="free_y") + 
-ggtitle("Permit diversity ('p_fshy') [green='$', red=weight]") + 
-xlab("Year") + ylab("Effective diversity")
-```
-
-```{r, fig.pos="placeHere", fig.cap="CV permit diversity, split by fishery (weighted by catch and revenue).", echo=FALSE}
-# Use gear as the facet here
-ggplot(spTot2, aes(year, sd.earn/mean.earn)) + geom_line(col="green") +
-geom_line(aes(x=year,y=sd.lbs/mean.lbs), col="red") + 
-facet_wrap(~ landedLoc, scales="free_y") + 
-ggtitle("CV of permit diversity ('p_fshy') [green='$', red=weight]") + 
-xlab("Year") + ylab("CV of effective diversity")
-```
-
-```{r, fig.pos="placeHere", fig.cap="Mean gear diversity, split by permits (weighted by catch and revenue).", echo=FALSE}
-spDiv_personYear2 = effectiveDiversity_by_personYear2(cfec, variable1="gearn", variable2 = "permit_coarse")
-spTot2 = summarize_effectiveDiversity2( spDiv_personYear2, variable2 = "permit_coarse" )
-
-# Use landedLoc as the facet here
-ggplot(spTot2, aes(year, mean.earn)) + geom_line(col="green") +
-geom_line(aes(x=year,y=mean.lbs), col="red") + 
-facet_wrap(~ permit_coarse, scales="free_y") + 
-ggtitle("Permit gear diversity ('gearn') [green='$', red=weight]") + 
-xlab("Year") + ylab("Effective diversity")
-```
-
-```{r, fig.pos="placeHere", fig.cap="CV gear diversity, split by permit (weighted by catch and revenue).", echo=FALSE}
-# Use gear as the facet here
-ggplot(spTot2, aes(year, sd.earn/mean.earn)) + geom_line(col="green") +
-geom_line(aes(x=year,y=sd.lbs/mean.lbs), col="red") + 
-facet_wrap(~ permit_coarse, scales="free_y") + 
-ggtitle("CV of gear diversity ('p_fshy') [green='$', red=weight]") + 
-xlab("Year") + ylab("CV of effective diversity")
-```
-
-## Filtering of unique people:vessel combinations
-
+## Filtering out unique vessel - person combinations  
 Filter out people who only use 1 vessel / year , and similarly filter out boats with more than 1 person / year.
@@
 Now we'll filter out individuals who only used 1 boat in a year, and boats with only one permit holder fishing in a year.
+
 ```{r}
@@
 dim(cfecnew)[1]
+```
 
-#Now we can create a unique *id* column for individual - vessel combinations.  
-cfecnew$id = paste(cfecnew$p_holder,cfecnew$cadfg)
-# calculate number of unique person - vessel pairs ~ 40700
-length(unique(cfecnew$id))
-```
-### Re-calculating diversity
+Aggregate annual metrics by person,
 
 ```{r}
-spDiv_personYear = effectiveDiversity_by_personYear(cfecnew, variable="specn")
-spTot = summarize_effectiveDiversity( spDiv_personYear )
 
-gearDiv_personYear = effectiveDiversity_by_personYear(cfecnew, variable="gearn")
-gearTot = summarize_effectiveDiversity( gearDiv_personYear )
+cfecnew$day = mdy.date(as.numeric(lapply(strsplit(cfecnew$day, "-"), getElement, 1)), as.numeric(lapply(strsplit(cfecnew$day, "-"), getElement, 2)), cfecnew$year) - mdy.date(1, 1, cfecnew$year)
 
-fishDiv_personYear = effectiveDiversity_by_personYear(cfecnew, variable="p_fshy")
-fishTot = summarize_effectiveDiversity( fishDiv_personYear )
+cfecAnnual = group_by(cfecnew, p_holder, year) %>%
+  summarize(permit = p_fshy[1], boat = cadfg2[1], 
+    days = diff(range(day) + 1), 
+    ngear = length(unique(gearn)),
+    nspec = length(unique(specn)),
+    npermit = length(unique(p_fshy)),
+    revenue = sum(g_earn),
+    length = mean(VLength, na.rm=T)) %>% 
+  filter(npermit == 1)
+
+# also calculate weighted simpson diversity of gears and species
+cfecAnnual = left_join(cfecAnnual, group_by(cfecnew, p_holder, year, specn) %>% 
+  summarize(totRev.spec = sum(g_earn)) %>% 
+  group_by(p_holder, year) %>% 
+  summarize(specDiv = simp.div(totRev.spec)))
+
+# also calculate weighted simpson diversity of gears and species
+cfecAnnual = left_join(cfecAnnual, group_by(cfecnew, p_holder, year, gearn) %>% 
+  summarize(totRev.gear = sum(g_earn)) %>% 
+  group_by(p_holder, year) %>% 
+  summarize(gearDiv = simp.div(totRev.gear)))
+
+temp = group_by(cfecnew, p_holder, year, specn) %>% 
+  summarize(totRev.spec = sum(g_earn), 
+    permit = p_fshy[1])
+temp = left_join(temp, group_by(cfecnew, p_holder, year) %>% 
+    summarize(totRev = sum(g_earn)))
+temp$totRev.spec = temp$totRev.spec / temp$totRev
+
+mkt = group_by(cfecnew, p_fshy, year, specn) %>% 
+  summarize(totRev.mkt.spec = sum(g_earn)) 
+mkt = left_join(mkt, group_by(mkt, p_fshy, year) %>% 
+  summarize(totRev.mkt = sum(totRev.mkt.spec))) 
+mkt$totRev.mkt.spec = mkt$totRev.mkt.spec / mkt$totRev.mkt
+names(mkt)[1] = "permit"
+
+# combine individual and market datasets
+temp = left_join(temp, mkt)
+
+# Calculate squared proportions
+temp$p2 = (temp$totRev.spec - temp$totRev.mkt.spec)^2
+
+cfecAnnual = left_join(cfecAnnual, dplyr::select(temp, p_holder, year, p2))
+
+# Also merge in previous revenue. Idx stores index of people last year
+idx = match(paste(cfecAnnual$p_holder,cfecAnnual$year-1), paste(cfecAnnual$p_holder,cfecAnnual$year))
+
+cfecAnnual$permit.prev = cfecAnnual$permit[idx]
+cfecAnnual$ngear.prev = cfecAnnual$ngear[idx]
+cfecAnnual$nspec.prev = cfecAnnual$nspec[idx]
+cfecAnnual$npermit.prev = cfecAnnual$npermit[idx]
+cfecAnnual$revenue.prev = cfecAnnual$revenue[idx]
+cfecAnnual$specdiv.prev = cfecAnnual$specDiv[idx]
+cfecAnnual$geardiv.prev = cfecAnnual$gearDiv[idx]
+cfecAnnual$p2.prev = cfecAnnual$p2[idx]
+
+cfecAnnual$specDiv.change = cfecAnnual$specDiv - cfecAnnual$specdiv.prev
+
+# Filter out salmon people
+cfecAnnual$salmon = ifelse(substr(cfecAnnual$permit,1,1)=="S", "Y", "N")
+
+cfecAnnual.salmon = cfecAnnual[cfecAnnual$salmon=="Y",]
+
+cfecAnnual.salmon = cfecAnnual.salmon[is.na(cfecAnnual.salmon$specDiv.change)==F, ]
+saveRDS(cfecAnnual.salmon, file="salmon_linearModeling_complete.rds")
 ```
-
-### Species
-
-```{r, fig.pos = "placeHere", fig.cap = "Species diversity by frequency of fish tickets of unique id:vessel combinations (excluding people who fish multiple vessels / year, or vessels that are fished by multiple individuals in a year.", echo=FALSE}
-
-ggplot(spTot, aes(x=year, y=mean.freq)) + geom_line(col="blue") +
-  geom_ribbon(aes(ymin=mean.freq - sd.freq/mean.freq, 
-  ymax=mean.freq + sd.freq/mean.freq), alpha=0.2, fill="blue") + 
-ggtitle("Effective species diversity ('specn') by fishTix") + 
-xlab("Year") + ylab("Mean (+/- CV)")
-```
-
-```{r, fig.pos = "placeHere", fig.cap = "Species diversity by revenue and landed weight of unique id:vessel combinations (excluding people who fish multiple vessels / year, or vessels that are fished by multiple individuals in a year.", echo=FALSE}
-
-# Plot by earning / weight on diff scale
-ggplot(spTot, aes(x=year, y=mean.earn)) + geom_line(col="red") + 
-geom_ribbon(aes(x = year, ymin=mean.earn - sd.earn/mean.earn, 
-  ymax = mean.earn + sd.earn/mean.earn), alpha=0.2, fill="red") + 
-geom_line(aes(x=year,y=mean.lbs), col="green") + 
-geom_ribbon(aes(x = year, ymin=mean.lbs - sd.lbs/mean.lbs, 
-  ymax=mean.lbs + sd.lbs/mean.lbs), alpha=0.2, fill="green") +
-ggtitle("Effective species diversity ('specn') [red = $, green=weight]") + 
-xlab("Year") + ylab("Mean (+/- CV)")
-```
-
-### Gears
-
-```{r, fig.pos = "placeHere", fig.cap = "Gear diversity by fish tickets of unique id:vessel combinations (excluding people who fish multiple vessels / year, or vessels that are fished by multiple individuals in a year.", echo=FALSE}
-ggplot(gearTot, aes(x=year, y=mean.freq)) + geom_line(col="blue") + 
-geom_ribbon(aes(ymin=mean.freq - sd.freq/mean.freq, 
-  ymax=mean.freq + sd.freq/mean.freq), alpha=0.2, fill="blue") + 
-ggtitle("Effective gear diversity ('gearn') by fishTix") + 
-xlab("Year") + ylab("Mean (+/- CV)")
-```
-
-```{r, fig.pos = "placeHere", fig.cap = "Gear diversity by revenue and landed weight of unique id:vessel combinations (excluding people who fish multiple vessels / year, or vessels that are fished by multiple individuals in a year.", echo=FALSE}
-# Plot by earning / weight on diff scale
-ggplot(gearTot, aes(x=year, y=mean.earn)) + geom_line(col="red") + 
-geom_ribbon(aes(x = year, ymin=mean.earn - sd.earn/mean.earn, 
-  ymax = mean.earn + sd.earn/mean.earn), alpha=0.2, fill="red") + 
-geom_line(aes(x=year,y=mean.lbs), col="green") + 
-geom_ribbon(aes(x = year, ymin=mean.lbs - sd.lbs/mean.lbs, 
-  ymax=mean.lbs + sd.lbs/mean.lbs), alpha=0.2, fill="green") + 
-ggtitle("Effective gear diversity ('gearn') [red = $, green=weight]") + 
-xlab("Year") + ylab("Mean (+/- CV)")
-```
-
-### Fisheries 
-```{r, fig.pos = "placeHere", fig.cap = "Permit diversity by fish ticket frequency of unique id:vessel combinations (excluding people who fish multiple vessels / year, or vessels that are fished by multiple individuals in a year.", echo=FALSE}
-
-ggplot(fishTot, aes(x=year, y=mean.freq)) + geom_line(col="blue") +
-geom_ribbon(aes(ymin=mean.freq - sd.freq/mean.freq, 
-  ymax=mean.freq + sd.freq/mean.freq), alpha=0.2, fill="blue") + 
-ggtitle("Effective permit diversity ('p_fshy') by fishTix") + 
-xlab("Year") + ylab("Mean (+/- CV)")
-```
-
-```{r, fig.pos = "placeHere", fig.cap = "Permit diversity by revenue and landed weight of unique id:vessel combinations (excluding people who fish multiple vessels / year, or vessels that are fished by multiple individuals in a year.", echo=FALSE}
-
-# Plot by earning / weight on diff scale
-ggplot(fishTot, aes(x=year, y=mean.earn)) + geom_line(col="red") + 
-geom_ribbon(aes(x = year, ymin=mean.earn - sd.earn/mean.earn, 
-  ymax = mean.earn + sd.earn/mean.earn), alpha=0.2, fill="red") + 
-geom_line(aes(x=year,y=mean.lbs), col="green") + 
-geom_ribbon(aes(x = year, ymin=mean.lbs - sd.lbs/mean.lbs, 
-  ymax=mean.lbs + sd.lbs/mean.lbs), alpha=0.2, fill="green") + 
-ggtitle("Effective permit diversity ('p_fshy') [red = $, green=weight]") + 
-xlab("Year") + ylab("Mean (+/- CV)")
-```
-
-## Comparing diversity between vessels and individuals
-
-Just out of curiousity, we can summarize diversity by year taking the approach used by Kasperski and Holland (based on CADFG number) and compare that to the same criteria using people ("p_holder").
-We'll only do this comparison on the species diversity, but use the same restriction: use only vessels or people who had participated continuously for the entire time period (this throws out 90%+ of the data).
-
-First, based on vessels:
-```{r}
-contVessels = group_by(cfec, cadfg) %>% 
-  summarize(nYear = length(unique(year))) %>% 
-  filter(nYear == 30)
-
-cfecnew = left_join(contVessels, cfec)
-
-# calculate diversity by species by CADFG (vessel #)
-effDiv.vessel = group_by(cfecnew, specn, year, cadfg) %>%
-  summarize(totRev = sum(g_earn, na.rm=T),
-  totN = n(),
-  totWeight = sum(g_pounds, na.rm=T),
-    days = length(unique(day))) %>% 
-  as.data.frame() %>%
-  filter(is.finite(totRev) & is.finite(totN) & is.finite(totWeight)) %>%   
-  as.data.frame() %>%
-  group_by(cadfg, year) %>% 
-  summarize(eff.freq = simp.div(totN),
-  eff.earn = simp.div(totRev),
-  eff.lbs = simp.div(totWeight),
-    totIndRev = sum(totRev),
-    nCalDays = sum(days)) %>% 
-  filter(!is.na(eff.freq) & !is.na(eff.earn) & !is.na(eff.lbs)) %>%
-  as.data.frame()
-
-effDiv.vessel = summarize_effectiveDiversity(effDiv.vessel)
-
-effDiv.person = effectiveDiversity_by_personYear(cfecnew, variable="specn")
-effDiv.person = summarize_effectiveDiversity( effDiv.person )
-```
-
-```{r, fig.pos="placeHere", fig.cap="Comparison of species diversity using vessel versus individuals", echo=FALSE}
-ggplot(effDiv.vessel, aes(x=year, y=mean.earn)) + geom_line(col="red") + 
-geom_ribbon(aes(x = year, ymin=mean.earn - sd.earn/mean.earn, 
-  ymax = mean.earn + sd.earn/mean.earn), alpha=0.2, fill="red") + 
-geom_line(aes(x=year,y=effDiv.person$mean.earn), col="green") + 
-geom_ribbon(aes(x = year, ymin=effDiv.person$mean.earn - 
-    effDiv.person$sd.earn/effDiv.person$mean.earn, 
-  ymax=effDiv.person$mean.earn + effDiv.person$sd.earn/effDiv.person$mean.earn), 
-  alpha=0.2, fill="green") + 
-ggtitle("Effective species diversity ('specn') [red = vessel, green=person]") + 
-xlab("Year") + ylab("Mean (+/- CV)")
-```
-
-Now make the portfolio plots from the metafolio package:
-
-```{r, echo=FALSE}
-species_diversity_by_year <-
-  effectiveDiversity_by_personYear(cfec, variable = "specn")
-breaks = c(1, 1.01, 1.5, 2, 2.5, 3, 3.5, 4, 10)
-species_diversity <-
-  group_by(species_diversity_by_year, p_holder) %>%
-  summarize(
-    # diversity_by_frequency = mean(eff.freq),
-    diversity_by_earnings = mean(eff.earn),
-    # diversity_by_weight = mean(eff.lbs),
-    m = log10(mean(totIndRev)),
-    v = sd(totIndRev) / mean(totIndRev)
-  ) %>%
-  filter(!is.na(v)) %>%
-  mutate(diversity_group = cut(diversity_by_earnings, breaks = breaks,
-    right = FALSE))
-
-species_diversity_list <- split(species_diversity,
-  species_diversity$diversity_group)
-
-#pdf("earnings_portfolio_by_species_diversity.pdf", height = 5, width = 5)
-par(cex = 0.8)
-metafolio::plot_cons_plans(
-  species_diversity_list,
-  plans_name = names(species_diversity_list),
-  cols = RColorBrewer::brewer.pal(length(names(species_diversity_list)),
-    "YlOrRd"),
-  xlab = "CV of gross earnings",
-  ylab = "log10 of gross earnings",
-  add_all_efs = FALSE)
-#dev.off()
-```

### 2016-06-09 analysis/0-read-and-save.R

@@ -1,46 +1,0 @@
-# Read in, cleanup, and write out a version of the data
-
-library(dplyr)
-load("data/cfecCleaned_new")
-names(cfec) <- tolower(names(cfec))
-cfec <- as_data_frame(cfec)
-names(cfec) <- gsub("\\.", "_", names(cfec))
-cfec$year <- as.numeric(cfec$year)
-cfec$specn_orig <- as.numeric(cfec$specn_orig)
-
-deflationTable = data.frame(year = 1985:2014,
-  defl = c(0.57236, 0.58393, 0.59879, 0.61974,
-    0.64388, 0.66774, 0.68993, 0.70564, 0.72244, 0.73781,
-    0.75321, 0.76695, 0.78009, 0.78855, 0.80061,
-    0.81883, 0.83753, 0.85038, 0.86729, 0.89114,
-    0.91981, 0.94812, 0.97334, 0.9925, 1,
-    1.01217, 1.03307, 1.05213, 1.06926, 1.08682))
-
-# Adjust price and g_earn for inflation
-cfec$year <- as.numeric(cfec$year)
-cfec$day <- substr(cfec$landdate, 6, 10)
-cfec <- inner_join(cfec, deflationTable)
-cfec <- mutate(cfec, g_price = g_price / defl, g_earn = g_earn / defl)
-cfec$defl <- NULL
-
-groupings <- read.csv("data/species-groupings.csv", stringsAsFactors = FALSE)
-names(groupings) <- tolower(names(groupings))
-names(groupings) <- gsub(" \\(spec\\)", "", names(groupings))
-groupings <- select(groupings, specn, group1)
-groupings <- rename(groupings, taxa_broad = group1)
-cfec <- left_join(cfec, groupings)
-cfec <- mutate(cfec, salmon = ifelse(spec %in%
-  c("CHNK", "SOCK", "COHO", "PINK", "CHUM"), TRUE, FALSE))
-cfec <- mutate(cfec, pollock = ifelse(spec == "PLCK", TRUE, FALSE))
-
-region <- read.csv("data/regions.csv", stringsAsFactors = FALSE)
-region <- region[,c(1:2)] #don’t include “assigned by” column
-cfec$region <- NULL
-cfec <- left_join(cfec, region)
-
-# devtools::install_github("wesm/feather/R")
-feather::write_feather(cfec, "data/cfec.feather")
-
-# saveRDS(cfec, file = "../data/cfec.rds")
[…]

### 2016-06-09 analysis/1-explore-portfolio-plots.R

@@ -1,203 +1,0 @@
-library(dplyr)
-library(ggplot2)
-library(viridis)
-source("analysis/downside-risk.R")
-source("analysis/contours.R")
-
-if (!exists("cfec")) {
-  cfec <- feather::read_feather("data/cfec.feather")
-}
-
-simp.div <- function(x) {
-  1 / sum((x / sum(x)) ^ 2)
-}
-
-effectiveDiversity_by_personYear <- function(dataFrame, variable) {
-  group_by_(dataFrame, variable, "year", "p_holder") %>%
-    summarize(
-      totRev = sum(g_earn, na.rm = TRUE),
-      totN = n(),
-      totWeight = sum(g_pounds, na.rm = TRUE),
-      days = length(unique(day)),
-      mean_vessel_length = mean(vlength),
-      area = ifelse(length(unique(area)) == 1, area[1], NA_character_),
-      pollock = ifelse("PLCK" %in% spec, TRUE, FALSE),
-      region = ifelse(length(unique(region)) == 1, region[1], NA_character_),
-      taxa = ifelse(length(unique(taxa_broad)) == 1, taxa_broad[1], NA_character_),
-      salmon = ifelse(length(unique(salmon)) == 1, salmon[1], as.logical(NA))
-    ) %>%
-    as.data.frame() %>%
-    filter(is.finite(totRev) &
-        is.finite(totN) & is.finite(totWeight)) %>%
-    as.data.frame() %>%
-    group_by(p_holder, year) %>%
-    summarize(
-      eff.freq = simp.div(totN),
-      eff.earn = simp.div(totRev),
-      eff.lbs = simp.div(totWeight),
-      totIndRev = sum(totRev),
-      nCalDays = sum(days),
-      mean_vessel_length = mean(mean_vessel_length),
-      area = ifelse(length(unique(area)) == 1, area[1], NA_character_),
-      pollock = ifelse(TRUE %in% pollock, TRUE, FALSE),
-      region = ifelse(length(unique(region)) == 1, region[1], NA_character_),
-      taxa = ifelse(length(unique(taxa)) == 1, taxa[1], NA_character_),
-      salmon = ifelse(length(unique(salmon)) == 1, salmon[1], as.logical(NA))
-    ) %>%
-    filter(!is.na(eff.freq) &
-        !is.na(eff.earn) & !is.na(eff.lbs)) %>%
-    as.data.frame()
-}
-
-if (!file.exists("generated-data/species_diversity_by_year.rds")) {
[…]

### 2016-06-09 analysis/2-try-model.R

@@ -1,109 +1,0 @@
-library(dplyr)
-library(rstan) # >= 2.7.0-1
-rstan::rstan_options(auto_write = TRUE)
-options(mc.cores = parallel::detectCores())
-library(ggplot2)
-
-set.seed(123)
-b0 <- 0.5
-b1 <- 0.3
-sigma0 <- 0.2
-sigma1 <- -0.2
-N <- 500
-s <- base::sample(seq(1, 10), N, replace = TRUE)
-u <- (rnorm(N, b0 + b1*s, sqrt(exp(2*(sigma0 + sigma1*s)))))
-
-plot(s, u)
-
-d <- readRDS("generated-data/species_diversity_metrics.rds")
-ggplot(d, aes(log(diversity_by_earnings), log(10^m))) + geom_point(alpha = 0.1)
-ggplot(d, aes(log(diversity_by_earnings), log(cv))) + geom_point(alpha = 0.1)
-
-library(nlme)
-m_gls1 <- gls(u ~ s, weights = varExp(form = ~s))
-m_gls2 <- gls(log(10^m) ~ log(diversity_by_earnings), weights = varExp(form = ~log(diversity_by_earnings)), data = d)
-summary(m_gls1)
-summary(m_gls2)
-
-de <- seq(1, 5, length.out = 100)
-predicted <- data.frame(diversity_by_earnings = de)
-predicted <- mutate(predicted, pm = predict(m_gls2, newdata = predicted))
-se <- summary(m_gls2)$model[[1]][[1]]
-predicted <- mutate(predicted, pl = pm -1.96*sqrt(exp(2*(se*log(diversity_by_earnings)))))
-predicted <- mutate(predicted, pu = pm +1.96*sqrt(exp(2*(se*log(diversity_by_earnings)))))
-
-ggplot(d, aes(diversity_by_earnings, log(10^m))) + geom_point(alpha = 0.1) +
-  geom_line(data = predicted, aes(x = (diversity_by_earnings), pm), color = "red") +
-  geom_line(data = predicted, aes(x = (diversity_by_earnings), pl), color = "red") +
-  geom_line(data = predicted, aes(x = (diversity_by_earnings), pu), color = "red")
-
-m_stan <- stan("portfolio.stan", iter = 400,
-  data = list(
-    N = N,
-    y = u,
-    x = s
[…]

### 2016-06-09 analysis/contours.R

@@ -1,43 +1,0 @@
-get_contour <- function(df, x_variable, y_variable, prob = 0.8, n = 200, ...) {
-  x <- dplyr::select_(df, x_variable, y_variable) %>%
-    data.matrix()  %>%
-      coda::mcmc() %>%
-        emdbook::HPDregionplot(prob = prob, n = n, ...)
-
-  for (j in length(x)) x[[j]]$contour_group <- j
-
-  x <- x %>% lapply(as.data.frame) %>%
-    bind_rows()
-
-  x <- mutate(x, contour_group =
-    ifelse(!is.na(contour_group), contour_group, 1))
-
-  x[,c("x", "y", "contour_group")]
-}
-
-plot_polygons <- function(polygon_data, x_column, y_column,
-  xlab = "Variance", ylab = "Mean", prob = 0.75,
-  grouping = c("diversity_group"), ...) {
-  contours <- plyr::ddply(polygon_data, grouping,
-    function(x) get_contour(x, x_column, y_column, prob = prob))
-# We need unique IDs in case there are multiple polygons for one group
-  contours$id <- paste(contours$contour_group, contours[,grouping[1]])
-  polygon_data <- mutate(polygon_data, id = 1)
-
-  p <- contours %>%
-    ggplot(aes_string("x", "y", group = "id", color = grouping[1],
-      fill = grouping[1]))
-
-  p <- p + geom_point(data = polygon_data, aes_string(x_column, y_column),
-    alpha = 0.1)
-  p <- p + geom_polygon(alpha = 0.3) +
-    scale_fill_viridis(discrete = TRUE) + scale_color_viridis(discrete = TRUE) +
-    theme_bw() +
-    xlab(xlab) + ylab(ylab)
-
-  if (length(grouping) > 1)
-    p <- p + facet_wrap(grouping[2])
-
-  print(p)
-  invisible(contours)
-}

### 2016-06-09 analysis/downside-risk.R

@@ -1,30 +1,0 @@
-semi_variance <- function(x, mar = mean(x, na.omit = TRUE), full = TRUE,
-  downside = TRUE) {
-
-  x <- na.omit(x)
-
-  if (downside)
-    x_mar <- x[x < mar]
-  else
-    x_mar <- x[x >= mar]
-
-  if (full)
-    len <- length(x)
-  else
-    len <- length(x_mar)
-
-  sum((x_mar - mar)^2)/len
-}
-
-semi_deviation <- function(x, ...) {
-  sqrt(semi_variance(x, ...))
-}
-
-# Variance at Risk:
-var_ <- function(x, prob = 0.05, ...) {
-  quantile(x, probs = prob, ...)
-}
-# Conditional Value at Risk:
-cvar <- function(x, prob = 0.05, ...) {
-  mean(x[x < var_(x, prob = prob, ...)])
-}

### 2016-06-09 analysis_portfolio/0-read-and-save.R

@@ -1,0 +1,46 @@
+# Read in, cleanup, and write out a version of the data
+
+library(dplyr)
+load("data/cfecCleaned_new")
+names(cfec) <- tolower(names(cfec))
+cfec <- as_data_frame(cfec)
+names(cfec) <- gsub("\\.", "_", names(cfec))
+cfec$year <- as.numeric(cfec$year)
+cfec$specn_orig <- as.numeric(cfec$specn_orig)
+
+deflationTable = data.frame(year = 1985:2014,
+  defl = c(0.57236, 0.58393, 0.59879, 0.61974,
+    0.64388, 0.66774, 0.68993, 0.70564, 0.72244, 0.73781,
+    0.75321, 0.76695, 0.78009, 0.78855, 0.80061,
+    0.81883, 0.83753, 0.85038, 0.86729, 0.89114,
+    0.91981, 0.94812, 0.97334, 0.9925, 1,
+    1.01217, 1.03307, 1.05213, 1.06926, 1.08682))
+
+# Adjust price and g_earn for inflation
+cfec$year <- as.numeric(cfec$year)
+cfec$day <- substr(cfec$landdate, 6, 10)
+cfec <- inner_join(cfec, deflationTable)
+cfec <- mutate(cfec, g_price = g_price / defl, g_earn = g_earn / defl)
+cfec$defl <- NULL
+
+groupings <- read.csv("data/species-groupings.csv", stringsAsFactors = FALSE)
+names(groupings) <- tolower(names(groupings))
+names(groupings) <- gsub(" \\(spec\\)", "", names(groupings))
+groupings <- select(groupings, specn, group1)
+groupings <- rename(groupings, taxa_broad = group1)
+cfec <- left_join(cfec, groupings)
+cfec <- mutate(cfec, salmon = ifelse(spec %in%
+  c("CHNK", "SOCK", "COHO", "PINK", "CHUM"), TRUE, FALSE))
+cfec <- mutate(cfec, pollock = ifelse(spec == "PLCK", TRUE, FALSE))
+
+region <- read.csv("data/regions.csv", stringsAsFactors = FALSE)
+region <- region[,c(1:2)] #don’t include “assigned by” column
+cfec$region <- NULL
+cfec <- left_join(cfec, region)
+
+# devtools::install_github("wesm/feather/R")
+feather::write_feather(cfec, "data/cfec.feather")
+
+# saveRDS(cfec, file = "../data/cfec.rds")
[…]

### 2016-06-09 analysis_portfolio/1-explore-portfolio-plots.R

@@ -1,0 +1,203 @@
+library(dplyr)
+library(ggplot2)
+library(viridis)
+source("analysis/downside-risk.R")
+source("analysis/contours.R")
+
+if (!exists("cfec")) {
+  cfec <- feather::read_feather("data/cfec.feather")
+}
+
+simp.div <- function(x) {
+  1 / sum((x / sum(x)) ^ 2)
+}
+
+effectiveDiversity_by_personYear <- function(dataFrame, variable) {
+  group_by_(dataFrame, variable, "year", "p_holder") %>%
+    summarize(
+      totRev = sum(g_earn, na.rm = TRUE),
+      totN = n(),
+      totWeight = sum(g_pounds, na.rm = TRUE),
+      days = length(unique(day)),
+      mean_vessel_length = mean(vlength),
+      area = ifelse(length(unique(area)) == 1, area[1], NA_character_),
+      pollock = ifelse("PLCK" %in% spec, TRUE, FALSE),
+      region = ifelse(length(unique(region)) == 1, region[1], NA_character_),
+      taxa = ifelse(length(unique(taxa_broad)) == 1, taxa_broad[1], NA_character_),
+      salmon = ifelse(length(unique(salmon)) == 1, salmon[1], as.logical(NA))
+    ) %>%
+    as.data.frame() %>%
+    filter(is.finite(totRev) &
+        is.finite(totN) & is.finite(totWeight)) %>%
+    as.data.frame() %>%
+    group_by(p_holder, year) %>%
+    summarize(
+      eff.freq = simp.div(totN),
+      eff.earn = simp.div(totRev),
+      eff.lbs = simp.div(totWeight),
+      totIndRev = sum(totRev),
+      nCalDays = sum(days),
+      mean_vessel_length = mean(mean_vessel_length),
+      area = ifelse(length(unique(area)) == 1, area[1], NA_character_),
+      pollock = ifelse(TRUE %in% pollock, TRUE, FALSE),
+      region = ifelse(length(unique(region)) == 1, region[1], NA_character_),
+      taxa = ifelse(length(unique(taxa)) == 1, taxa[1], NA_character_),
+      salmon = ifelse(length(unique(salmon)) == 1, salmon[1], as.logical(NA))
+    ) %>%
+    filter(!is.na(eff.freq) &
+        !is.na(eff.earn) & !is.na(eff.lbs)) %>%
+    as.data.frame()
+}
+
+if (!file.exists("generated-data/species_diversity_by_year.rds")) {
[…]

### 2016-06-09 analysis_portfolio/2-try-model.R

@@ -1,0 +1,109 @@
+library(dplyr)
+library(rstan) # >= 2.7.0-1
+rstan::rstan_options(auto_write = TRUE)
+options(mc.cores = parallel::detectCores())
+library(ggplot2)
+
+set.seed(123)
+b0 <- 0.5
+b1 <- 0.3
+sigma0 <- 0.2
+sigma1 <- -0.2
+N <- 500
+s <- base::sample(seq(1, 10), N, replace = TRUE)
+u <- (rnorm(N, b0 + b1*s, sqrt(exp(2*(sigma0 + sigma1*s)))))
+
+plot(s, u)
+
+d <- readRDS("generated-data/species_diversity_metrics.rds")
+ggplot(d, aes(log(diversity_by_earnings), log(10^m))) + geom_point(alpha = 0.1)
+ggplot(d, aes(log(diversity_by_earnings), log(cv))) + geom_point(alpha = 0.1)
+
+library(nlme)
+m_gls1 <- gls(u ~ s, weights = varExp(form = ~s))
+m_gls2 <- gls(log(10^m) ~ log(diversity_by_earnings), weights = varExp(form = ~log(diversity_by_earnings)), data = d)
+summary(m_gls1)
+summary(m_gls2)
+
+de <- seq(1, 5, length.out = 100)
+predicted <- data.frame(diversity_by_earnings = de)
+predicted <- mutate(predicted, pm = predict(m_gls2, newdata = predicted))
+se <- summary(m_gls2)$model[[1]][[1]]
+predicted <- mutate(predicted, pl = pm -1.96*sqrt(exp(2*(se*log(diversity_by_earnings)))))
+predicted <- mutate(predicted, pu = pm +1.96*sqrt(exp(2*(se*log(diversity_by_earnings)))))
+
+ggplot(d, aes(diversity_by_earnings, log(10^m))) + geom_point(alpha = 0.1) +
+  geom_line(data = predicted, aes(x = (diversity_by_earnings), pm), color = "red") +
+  geom_line(data = predicted, aes(x = (diversity_by_earnings), pl), color = "red") +
+  geom_line(data = predicted, aes(x = (diversity_by_earnings), pu), color = "red")
+
+m_stan <- stan("portfolio.stan", iter = 400,
+  data = list(
+    N = N,
+    y = u,
+    x = s
[…]

### 2016-06-09 analysis_portfolio/contours.R

@@ -1,0 +1,43 @@
+get_contour <- function(df, x_variable, y_variable, prob = 0.8, n = 200, ...) {
+  x <- dplyr::select_(df, x_variable, y_variable) %>%
+    data.matrix()  %>%
+      coda::mcmc() %>%
+        emdbook::HPDregionplot(prob = prob, n = n, ...)
+
+  for (j in length(x)) x[[j]]$contour_group <- j
+
+  x <- x %>% lapply(as.data.frame) %>%
+    bind_rows()
+
+  x <- mutate(x, contour_group =
+    ifelse(!is.na(contour_group), contour_group, 1))
+
+  x[,c("x", "y", "contour_group")]
+}
+
+plot_polygons <- function(polygon_data, x_column, y_column,
+  xlab = "Variance", ylab = "Mean", prob = 0.75,
+  grouping = c("diversity_group"), ...) {
+  contours <- plyr::ddply(polygon_data, grouping,
+    function(x) get_contour(x, x_column, y_column, prob = prob))
+# We need unique IDs in case there are multiple polygons for one group
+  contours$id <- paste(contours$contour_group, contours[,grouping[1]])
+  polygon_data <- mutate(polygon_data, id = 1)
+
+  p <- contours %>%
+    ggplot(aes_string("x", "y", group = "id", color = grouping[1],
+      fill = grouping[1]))
+
+  p <- p + geom_point(data = polygon_data, aes_string(x_column, y_column),
+    alpha = 0.1)
+  p <- p + geom_polygon(alpha = 0.3) +
+    scale_fill_viridis(discrete = TRUE) + scale_color_viridis(discrete = TRUE) +
+    theme_bw() +
+    xlab(xlab) + ylab(ylab)
+
+  if (length(grouping) > 1)
+    p <- p + facet_wrap(grouping[2])
+
+  print(p)
+  invisible(contours)
+}

### 2016-06-09 analysis_portfolio/downside-risk.R

@@ -1,0 +1,30 @@
+semi_variance <- function(x, mar = mean(x, na.omit = TRUE), full = TRUE,
+  downside = TRUE) {
+
+  x <- na.omit(x)
+
+  if (downside)
+    x_mar <- x[x < mar]
+  else
+    x_mar <- x[x >= mar]
+
+  if (full)
+    len <- length(x)
+  else
+    len <- length(x_mar)
+
+  sum((x_mar - mar)^2)/len
+}
+
+semi_deviation <- function(x, ...) {
+  sqrt(semi_variance(x, ...))
+}
+
+# Variance at Risk:
+var_ <- function(x, prob = 0.05, ...) {
+  quantile(x, probs = prob, ...)
+}
+# Conditional Value at Risk:
+cvar <- function(x, prob = 0.05, ...) {
+  mean(x[x < var_(x, prob = prob, ...)])
+}

### 2016-06-09 analysis_portfolio/quantile-2d.R

@@ -1,0 +1,53 @@
+# Extracted and adapted from the metafolio package
+#' Get quantile contour
+#'
+#' @param x Output from \code{\link[MASS]{kde2d}}.
+#' @param alpha The quantile level.
+get_quantile_contour <- function(x, alpha = 0.8) {
+  zdens <- rev(sort(x$z))
+  Czdens <- cumsum(zdens)
+  Czdens <- (Czdens/Czdens[length(zdens)])
+  crit.val <- zdens[max(which(Czdens<=alpha))]
+  b.full=contourLines(x,levels=crit.val)
+  list(x = b.full[[1]]$x, y = b.full[[1]]$y)
+}
+
+#' Custom bandwidth
+#'
+#' Based on \code{bandwidth.nrd} from \pkg{MASS}. This version takes the
+#' absolute value of \code{var} to avoid errors.
+#'
+#' @param x A numeric vector
+custom_bw <- function(x) {
+    r <- quantile(x, c(0.25, 0.75))
+    h <- (r[2] - r[1])/1.34
+    4 * 1.06 * min(sqrt(abs(var(x))), h) * length(x)^(-1/5)
+}
+
+#' Add a kernel density polygon
+#'
+#' @param x x values
+#' @param y y values
+#' @param col Colour to add polygon with. Will be made into two levels of
+#'   opacity.
+#' @param lwd lwd Line width
+#' @param alpha A numeric vector of length 2 that gives the confidence levels
+#'   for the two kernel density polygons.
+#' @param add_poly Add polygons?
+#' @param add_pts Logical: should points be added?
+add_dens_polygon <- function(x, y, col = "#000000", lwd = 1.7,
+  alpha = c(0.25, 0.75), add_pts = FALSE, add_poly = TRUE,
+  plot = TRUE) {
+  if(add_poly) {
+    x_bw <- custom_bw(na.omit(x))
+    y_bw <- custom_bw(na.omit(y))
+    k <- get_quantile_contour(MASS::kde2d(x,y, h = c(x_bw, y_bw)), alpha = alpha[1])
+    if (plot)
+      polygon(k$x, k$y, border = col, col = paste(col, "40", sep = ""), lwd = lwd)
+    k <- get_quantile_contour(MASS::kde2d(x,y, h = c(x_bw, y_bw)), alpha = alpha[2])
+    if (plot)
[…]

### 2016-06-14 analysis_portfolio/4b-Bayesian_portfolio_modeling.Rmd
commit bayesian code

---
title: "Bayesian_portfolio_modeling"
author: ""
date: "June 05, 2016"
output:
  pdf_document:
    fig_caption: yes
  html_document: default
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=FALSE, tidy.opts=list(width.cutoff=60), warning = FALSE, message = FALSE)
```

## Load packages and such 
```{r dataLoad, warning = FALSE, message=FALSE, results="hide", echo=FALSE}
library(dplyr)
library(knitr)
library(mgcv)
library(ggplot2)
library(randomForest)
library(MASS)
library(zoo)
library(MatchIt)
library(metafolio)
library(randomForest)
library(date)
library(rstan)

```

## LMER Model fitting

Start by fitting some lmer models to revenue for the non-salmon people as an example (smaller dataset).

```{r}
cfecAnnual.salmon = readRDS(file="salmon_linearModeling_complete.rds")

cfecAnnual.salmon$spec = substr(cfecAnnual.salmon$permit,1,1)
cfecAnnual.salmon$area = substr(cfecAnnual.salmon$permit,4,4)

full.mod = lmer(log(revenue) ~ I(log(revenue.prev)) + specDiv +specDiv.change+ (1+specDiv|permit) + length + days, data = cfecAnnual.salmon[complete.cases(cfecAnnual.salmon),])

full.mod.ind = lmer(log(revenue) ~ permit*as.factor(year) + specDiv + (1+specDiv|permit) + length + days + (1|p_holder), data = cfecAnnual.salmon[complete.cases(cfecAnnual.salmon),])

group_by(cfecAnnual.salmon[complete.cases(cfecAnnual.salmon),], p_holder) %>% 
  summarize(earn = mean(log(revenue)), cv = sd(log(revenue)) / earn, div = mean(specDiv), permit = permit[1]) %>%
  as.data.frame %>%
  group_by(permit) %>% 
  mutate(n = n()) %>% 
  filter(n > 50) %>% 
  ggplot(aes(x = div, y = earn, color = cv)) + geom_point() + facet_wrap(~ permit, scales = "free")
```

## Code for testing fitting the model using STAN.

```{r}
cfecAnnual.nonsalmon = readRDS(file="nonsalmon_linearModeling_complete.rds")

# Subset the data to include only p_holder - year combinations with > 100 records, basically the top 20 permits
testData = cfecAnnual.nonsalmon[cfecAnnual.nonsalmon$permit%in%names(rev(sort(table(cfecAnnual.nonsalmon$permit)))[1:20]),]

dim(testData)

# Create numeric representation of characters/factors for linear modeling
P_HOLDER = as.numeric(as.factor(testData$p_holder))
PERMIT = as.numeric(as.factor(testData$permit))
PERMIT_AREA = as.numeric(as.factor(substr(testData$permit, 4, 4)))
PERMIT_SPEC = as.numeric(as.factor(substr(testData$permit, 1, 1)))
SPEC_DIV = testData$specDiv
DAYS = log(testData$days+1)

# Calculate variables / matrices to plug into STAN. These variables / matrices can 
# be changed to add or delete covariates. Covariates can be included in the following:
# fixed effects in slope of mean (b1_cov)
# fixed effects in slope of var (sigma1_cov)
# random effects of intercept of mean (b0_re)
# random effects of intercept of var (sigma0_re)
# random effects of slope of mean (b1_re)
# random effects of slope of var (sigma1_re)

N = dim(testData)[1]
b1_cov = cbind(SPEC_DIV, DAYS) # fixed effect covariate on slope
sigma1_cov = cbind(SPEC_DIV, DAYS) # fixed effect covariate on slope
b1_cov_re = cbind(SPEC_DIV)
sigma1_cov_re = cbind(SPEC_DIV)

b0_re = cbind(P_HOLDER, PERMIT, PERMIT_AREA, PERMIT_SPEC)
b0_re_levels = apply(b0_re,2,max)
b0_re_n = dim(b0_re)[2]

b1_re = cbind(PERMIT, PERMIT_AREA, PERMIT_SPEC)
b1_re_levels = apply(b1_re,2,max)
b1_re_n = dim(b1_re)[2]

sigma0_re = cbind(P_HOLDER, PERMIT, PERMIT_AREA, PERMIT_SPEC)
sigma0_re_levels = apply(b0_re,2,max)
sigma0_re_n = dim(sigma0_re)[2]

sigma1_re = cbind(PERMIT, PERMIT_AREA, PERMIT_SPEC)
sigma1_re_levels = apply(sigma1_re,2,max)
sigma1_re_n = dim(sigma1_re)[2]

K_mean = dim(b1_cov)[2]
K_sig = dim(sigma1_cov)[2]
K_mean_re = dim(b1_cov_re)[2]
K_sig_re = dim(sigma1_cov_re)[2]
y = log(testData$revenue)

### Center all of the covariate matrices
b1_cov = scale(b1_cov, scale=FALSE)
sigma1_cov = scale(sigma1_cov, scale=FALSE)
b1_cov_re = scale(b1_cov_re, scale=FALSE)
sigma1_cov_re = scale(sigma1_cov_re, scale=FALSE)
```

```{r}
# Named list of data
list_data = list("N"=N, "K_mean"=K_mean, "b1_cov"=b1_cov, "sigma1_cov"=sigma1_cov,
  "K_mean_re"=K_mean_re, "K_sig_re"=K_sig_re,"b1_cov_re"=b1_cov_re, "sigma1_cov_re"=sigma1_cov_re,
  "b0_re"=b0_re, "b0_re_levels"=b0_re_levels, "b0_re_n"=b0_re_n,
  "b1_re"=b1_re, "b1_re_levels"=b1_re_levels, "b1_re_n"=b1_re_n,
  "sigma0_re"=sigma0_re, "sigma0_re_levels"=sigma0_re_levels, "sigma0_re_n"=sigma0_re_n,
  "sigma1_re"=sigma1_re, "sigma1_re_levels"=sigma1_re_levels, "sigma1_re_n"=sigma1_re_n, "K_sig"=K_sig, "y"=y)
# parameters to monitor
list_pars = c("b0_mu","b0_dev","b0_tau", "b1_mu","b1_dev","b1_tau", "sigma0_mu","sigma0_dev","sigma0_tau", "sigma1_mu","sigma1_dev","sigma1_tau")

# fit the model
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

stanMod = stan(file = 'portfolio_mixedModel.stan',data = list_data, 
  verbose = TRUE, chains = 4, thin = 1, warmup = 500, iter = 1000, 
  pars = list_pars)
p = extract(stanMod, list_pars, permuted=TRUE)
```

## To-do list:
### 1. model covariance of slope / intercept in random effects. maybe start with bivariate normal
### 2. Make plots of random effects v permits

### 2016-06-14 diversity/salmon_deepDive.Rmd
Start rearranging the files

---
title: "salmon_deepDive"
author: "Eric Ward"
date: "May 18, 2016"
output:
  pdf_document:
    fig_caption: yes
  html_document: default
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=FALSE, tidy.opts=list(width.cutoff=60), warning = FALSE, message = FALSE)
```

## Load cleaned data  
```{r dataLoad, warning = FALSE, message=FALSE, results="hide", echo=FALSE}
library(dplyr)
library(knitr)
library(mgcv)
library(ggplot2)
library(randomForest)
library(MASS)
library(zoo)
library(MatchIt)
library(metafolio)
library(randomForest)
library(date)
load("cfecCleaned2.Rdata")

# convert 6-dig p_fshy codes to 5 digits
# removing blank space - now they're all length(5)
idx = which(nchar(cfec$p_fshy)==6)
cfec$p_fshy[idx] = paste(substr(cfec$p_fshy[idx], 1, 1), 
  substr(cfec$p_fshy[idx], 3, 6), sep="")

# remove white space in 2nd position
idx = which(substr(cfec$p_fshy, 2, 2) == " ")
cfec$p_fshy[idx] = paste(substr(cfec$p_fshy[idx], 1, 1), 
  substr(cfec$p_fshy[idx], 3, length(cfec$p_fshy[idx])), sep="")

# Pull in "new.locs" from Anne/Jordan/Rich/Ben/Ole
newlocs = read.csv("new.locs.csv")
newlocs$p_fshy = as.character(newlocs$p_fshy)

# Match cfec codes in newlocs to identify new 
# locations (Anne/Jordan/Rich/Ben/Ole)
cfec$landedLoc = newlocs$New.Loc[match(cfec$p_fshy, newlocs$p_fshy)]
cfec$landedLoc[which(is.na(cfec$landedLoc))] = "Statewide" # C04B

# Split out permit types
cfec$pType = ""
cfec$pType[substr(cfec$p_fshy,1,2)=="S0"] = "Salm.other"
cfec$pType[cfec$p_fshy == "S01A"] = "SE.ps"
cfec$pType[cfec$p_fshy == "S03A"] = "SE.dn"
cfec$pType[cfec$p_fshy == "S01E"] = "PWS.ps"
cfec$pType[cfec$p_fshy == "S03E"] = "PWS.dn"
cfec$pType[cfec$p_fshy == "S04E"] = "PWS.sn"
cfec$pType[cfec$p_fshy == "S01H"] = "CI.ps"
cfec$pType[cfec$p_fshy == "S03H"] = "CI.dn"
cfec$pType[cfec$p_fshy == "S04H"] = "CI.sn"
cfec$pType[cfec$p_fshy == "S01K"] = "K.ps"
cfec$pType[cfec$p_fshy == "S04K"] = "K.sn"
cfec$pType[cfec$p_fshy == "S01L"] = "CH.ps"
cfec$pType[cfec$p_fshy == "S01M"] = "AP.ps"
cfec$pType[cfec$p_fshy == "S03M"] = "AP.dn"
cfec$pType[cfec$p_fshy == "S04M"] = "AP.sn"
cfec$pType[cfec$p_fshy == "S03T"] = "BB.dn"
cfec$pType[cfec$p_fshy == "S04T"] = "BB.sn"
```

We need to adjust the landed values and prices for inflation. To do this, we'll use the data Alan pulled together (adjusting to 2009 dollars).

https://docs.google.com/spreadsheets/d/1hRLYTX1nZBLKnYXwTCOtU7-XnDoMZf_1ufdTiNQFd4U/edit#gid=0

```{r, echo=FALSE}
deflationTable = data.frame(year = 1985:2014, defl = c(0.57236, 0.58393, 0.59879, 0.61974,  
0.64388, 0.66774, 0.68993, 0.70564, 0.72244, 0.73781,  
0.75321, 0.76695, 0.78009, 0.78855, 0.80061,  
0.81883, 0.83753, 0.85038, 0.86729, 0.89114,  
0.91981, 0.94812, 0.97334, 0.9925, 1,  
1.01217, 1.03307, 1.05213, 1.06926, 1.08682))

# Adjust price and g_earn for inflation
cfec$year = as.numeric(cfec$year)
cfec$day = substr(cfec$landdate, 6, 10)
cfec$g_price = cfec$g_price/deflationTable$defl[cfec$year-1984]
cfec$g_earn = cfec$g_earn/deflationTable$defl[cfec$year-1984]
```

### Function definiton  
We need to define a few functions for calculating diversity. We'll use the Simpson's diversity index (HHI) to calculate the entropy-based inverse, which is a proxy for effective species diversity. Note the C++ version is commented out below,

```{r, warning = FALSE, message=FALSE, echo=FALSE}

simp.div = function(x) {
  1/sum((x/sum(x))^2)
}

#library(Rcpp)
#cppFunction('double simpCpp(NumericVector x) {
#  double y = 1/sum((x/sum(x))^2);
#  return y;
#}')

# Note that the first line of this function uses group_by_ because 
# we need to pass in a string for 'variable'
effectiveDiversity_by_personYear = function(dataFrame, variable) {
  effDiv = group_by_(dataFrame, variable, "year", "p_holder") %>%
  summarize(totRev = sum(g_earn, na.rm=T),
  totN = n(),
  totWeight = sum(g_pounds, na.rm=T),
    days = length(unique(day))) %>% 
  as.data.frame() %>%
  filter(is.finite(totRev) & is.finite(totN) & is.finite(totWeight)) %>%   
  as.data.frame() %>%
  group_by(p_holder, year) %>% 
  summarize(eff.freq = simp.div(totN),
  eff.earn = simp.div(totRev),
  eff.lbs = simp.div(totWeight),
    totIndRev = sum(totRev),
    nCalDays = sum(days)) %>% 
  filter(!is.na(eff.freq) & !is.na(eff.earn) & !is.na(eff.lbs)) %>%
  as.data.frame()
  return(effDiv)
}

# Note that the first line of this function uses group_by_ because 
# we need to pass in a string for 'variable'
effectiveDiversity_by_personYear2 = function(dataFrame, variable1, variable2) {
  effDiv = group_by_(dataFrame, variable1, variable2, "year", "p_holder") %>%
  summarize(totRev = sum(g_earn, na.rm=T),
  totN = n(),
  totWeight = sum(g_pounds, na.rm=T),
    days = length(unique(day))) %>% 
  as.data.frame() %>%
  filter(is.finite(totRev) & is.finite(totN) & is.finite(totWeight)) %>%   
  as.data.frame() %>% 
  group_by_("p_holder", "year", variable2) %>% 
  summarize(eff.freq = simp.div(totN),
  eff.earn = simp.div(totRev),
  eff.lbs = simp.div(totWeight),
    totIndRev = sum(totRev),
    nCalDays = sum(days)) %>% 
  filter(!is.na(eff.freq) & !is.na(eff.earn) & !is.na(eff.lbs)) %>%
  as.data.frame()
  return(effDiv)
}

effectiveDiversity_by_Year = function(dataFrame, variable) {
  effDiv = group_by_(dataFrame, variable, "year") %>%
  summarize(totRev = sum(g_earn, na.rm=T),
  totN = n(),
  totWeight = sum(g_pounds, na.rm=T),
    days = length(unique(day))) %>% 
  as.data.frame() %>%
  filter(is.finite(totRev) & is.finite(totN) & is.finite(totWeight)) %>%   
  as.data.frame() %>%
  group_by(year) %>% 
  summarize(eff.freq = simp.div(totN),
  eff.earn = simp.div(totRev),
  eff.lbs = simp.div(totWeight),
    totIndRev = sum(totRev),
    nCalDays = sum(days)) %>% 
  filter(!is.na(eff.freq) & !is.na(eff.earn) & !is.na(eff.lbs)) %>%
  as.data.frame()
  return(effDiv)
}

# This function takes the output of the one above and 
# summmarizes diversity metrics
summarize_effectiveDiversity = function(dataFrame) {
  effDiv = group_by(dataFrame, year) %>% 
  summarize(mean.freq = mean(eff.freq),
  sd.freq = sd(eff.freq),
  mean.earn = mean(eff.earn),
  sd.earn = sd(eff.earn),
  mean.lbs = mean(eff.lbs),
  sd.lbs = sd(eff.lbs)) %>% 
  as.data.frame()  
  return(effDiv)
}

# This function takes the output of the one above and 
# summmarizes diversity metrics
summarize_effectiveDiversity2 = function(dataFrame, variable2) {
  effDiv = group_by_(dataFrame, "year", variable2) %>% 
  summarize(mean.freq = mean(eff.freq),
  sd.freq = sd(eff.freq),
  mean.earn = mean(eff.earn),
  sd.earn = sd(eff.earn),
  mean.lbs = mean(eff.lbs),
  sd.lbs = sd(eff.lbs)) %>% 
  as.data.frame()  
  return(effDiv)
}
```

# Overview

Based on the larger coarse plots, I thought it'd worth diving into the bigget group of specialists (= salmon fishermen). I'm defining this as people who are fishing with gillnets / seinenets (gearn = 1, 3, 4) and landing salmon.

```{r, echo=FALSE}
# Add effective species diversity, weighted by earnings
effSpec = group_by(cfec, year, p_holder, specn) %>% 
  summarize(earn = sum(g_earn)) %>% 
  group_by(year, p_holder) %>% 
  summarize(effSpec = simp.div(earn), p.pink=sum(earn[specn==440])/sum(earn), p.chum=sum(earn[specn==450])/sum(earn))

cfec = left_join(cfec, effSpec)

salmonSpecialists = group_by(cfec[cfec$pType!="",], year, p_holder, pType) %>% 
  summarize(gears = length(unique(gearn)), 
    salmonGears = length(unique(gearn[gearn%in%c(1,3,4)])),
    salmonSpec = length(unique(specn[specn%in%c(410,420,430,440,450)])),
    totRev = sum(g_earn), 
    effSpec = mean(effSpec),
    p.pink = mean(p.pink,na.rm=T),
    p.chum = mean(p.chum,na.rm=T)) %>% 
  filter(salmonGears == gears & salmonSpec >= 1)

# Calculate total participation
n.salmonSpecialists  = group_by(salmonSpecialists, year) %>% 
  summarize(n = length(unique(p_holder)))
n.tot= group_by(cfec, year) %>% 
  summarize(n = length(unique(p_holder)))

# In each year, we can also categorize our population of salmon specialists as:
# 1. new fisherman
# 2. pre-existing salmon specialist
# 3. someone who was fishing salmon and something else
# 4. someone who wasn't fishing salmon
# 5. someone who didn't fish last year, but did in the past
salmonParticipation.nofilter = group_by(cfec, p_holder, year) %>% 
  summarize(gears = length(unique(gearn)), 
    salmonGears = length(unique(gearn[gearn%in%c(1,3,4)])),
    salmonSpec = length(unique(specn[specn%in%c(410,420,430,440,450)])))

salmonParticipation.filter = filter(salmonParticipation.nofilter, salmonGears == gears & salmonSpec > 0)

# Find matches in previous years 
idx = match(paste(salmonParticipation.filter$p_holder, salmonParticipation.filter$year-1), paste(salmonParticipation.nofilter$p_holder, salmonParticipation.nofilter$year))
salmonParticipation.filter$prev.gear = salmonParticipation.nofilter$gears[idx]
salmonParticipation.filter$prev.salmgear = salmonParticipation.nofilter$salmonGears[idx]
salmonParticipation.filter$prev.spec = salmonParticipation.nofilter$salmonSpec[idx]

# Find matches in later years 
idx = match(paste(salmonParticipation.filter$p_holder, salmonParticipation.filter$year+1), paste(salmonParticipation.nofilter$p_holder, salmonParticipation.nofilter$year))
salmonParticipation.filter$next.gear = salmonParticipation.nofilter$gears[idx]
salmonParticipation.filter$next.salmgear = salmonParticipation.nofilter$salmonGears[idx]
salmonParticipation.filter$next.spec = salmonParticipation.nofilter$salmonSpec[idx]

startYear = group_by(cfec, p_holder) %>% 
  summarize(start = min(year))
endYear = group_by(cfec, p_holder) %>% 
  summarize(end = max(year))
salmonParticipation.filter = left_join(salmonParticipation.filter, startYear)
salmonParticipation.filter = left_join(salmonParticipation.filter, endYear)

salmonParticipation.filter$group = 5
salmonParticipation.filter$group[which(salmonParticipation.filter$year == salmonParticipation.filter$start)] = 1
salmonParticipation.filter$group[which(salmonParticipation.filter$prev.salmgear == salmonParticipation.filter$prev.gear & salmonParticipation.filter$prev.spec > 0)] = 2
salmonParticipation.filter$group[which(salmonParticipation.filter$prev.salmgear < salmonParticipation.filter$prev.gear & salmonParticipation.filter$prev.spec > 0)] = 3
salmonParticipation.filter$group[which(salmonParticipation.filter$prev.spec==0)] = 4

salmonParticipation.filter$nextgroup = 5
salmonParticipation.filter$nextgroup[which(salmonParticipation.filter$year == salmonParticipation.filter$end)] = 1
salmonParticipation.filter$nextgroup[which(salmonParticipation.filter$next.salmgear == salmonParticipation.filter$next.gear & salmonParticipation.filter$next.spec > 0)] = 2
salmonParticipation.filter$nextgroup[which(salmonParticipation.filter$next.salmgear != salmonParticipation.filter$next.gear & salmonParticipation.filter$next.spec > 0)] = 3
salmonParticipation.filter$nextgroup[which(salmonParticipation.filter$next.salmgear == salmonParticipation.filter$next.gear & salmonParticipation.filter$next.spec == 0)] = 4

salmonSummary = group_by(salmonSpecialists, year, pType) %>% 
  summarize(ngear = mean(gears), ngear2 = length(which(gears == 2)), npeople = length(unique(p_holder)), nspec = mean(salmonSpec), neffspec = mean(effSpec),
    p.pink = mean(p.pink, na.rm=T), p.chum = mean(p.chum, na.rm=T)) 
```

```{r, fig.cap="Mean effective species diversity, by permit type", fig.pos="placeHere", echo=FALSE, warning=FALSE, message = FALSE}
group_by(cfec[cfec$pType!="", ], year, pType) %>% 
  summarize(meanEff = mean(effSpec)) %>% 
  ggplot(aes(x = year, y = meanEff, color = factor(pType))) + geom_line() + facet_wrap(~ pType) + ylab("Effective species diversity")
```

## Interesting observations:  
### 1. Like trends in total fishermen, salmon specialists have declined since the 1990s, but at a much slower rate. Thus, more people are becoming salmon specialists -- using purse-seine / gillnet gears exclusively. Participation overall has been flat since rationalization (early 2000s) but a greater proportion of fishermen are specializing on salmon.

```{r, fig.cap = "Trends in salmon specialists (defined above), and total people fishing, 1985-2014", fig.pos = "placeHere", echo=FALSE}
par(mfrow = c(2,1), mai=c(0.5,0.6,0.1,0.1), mgp=c(2,1,0))
plot(n.salmonSpecialists, col="tomato1", lwd=3, type="b", ylim=c(5000,15000), xlab = "", ylab="People")
lines(n.tot$year, n.tot$n, col="grey", lwd=3)
points(n.tot$year, n.tot$n, col="grey", lwd=3)
legend('topright', c("Salmon","Total"), col=c("tomato1", "grey"), bty='n', lwd=3)
plot(n.tot$year, n.salmonSpecialists$n/n.tot$n, lwd=3, type="b", xlab="", ylab="% Salmon specialists")
```

### Perhaps the drop in Bristol Bay drift / set nets in the early 2000s is a response to low prices & low abundance (and a consequence of portfolio not being diversified).

```{r, fig.pos="placeHere", fig.cap = "Participation by permit fishery (number of unique permit holders actively fishing). For the purposes of display, people fishing in other areas (Yakutat, etc) or with troll gear not included.", echo=FALSE}
group_by(cfec[cfec$pType!="" & cfec$pType!="Salm.other",], year, pType) %>% 
summarize(n = length(unique(p_holder))) %>%
ggplot(aes(x = year, y = n, color = factor(pType))) + geom_line()
```

Coincident with rationalization, the participation in the early 2000s may have also been influenced by price drops of all salmon species,

```{r, fig.cap = "Trends in salmon prices, 1985-2014", fig.pos = "placeHere", echo=FALSE}
# https://www.cfec.state.ak.us/pmtvalue/mnusalm.htm
group_by(cfec[cfec$specn%in%c(410:450) & cfec$pType != "",], spec, year, pType) %>% 
summarize(price = mean(g_price)) %>% 
ggplot(aes(x=year, y=(price), color = factor(spec)) ) + geom_line() + ylab("Price (dollars)") + facet_wrap(~ pType, scales="free_y")
```

```{r, fig.cap = "Trends in natural-log salmon prices, 1985-2014", fig.pos = "placeHere", echo=FALSE}
group_by(cfec[cfec$specn%in%c(410:450) & cfec$pType != "",], spec, year, pType) %>% 
summarize(price = mean(g_price)) %>% 
ggplot(aes(x=year, y=log(price), color = factor(spec)) ) + geom_line() + ylab("Ln Price (dollars)") + facet_wrap(~ pType, scales="free_y")
```

```{r, fig.cap = "Tradeoff between average prices, and year-to-year variation in prices, 1985-2014", fig.pos = "placeHere", echo=FALSE}
grp = group_by(cfec[cfec$specn%in%c(410:450) & cfec$pType != "",], spec, year, pType) %>% 
summarize(logprice = log(mean(g_price)), meanprice = mean(g_price)) %>%
group_by(spec, pType) %>% 
  summarize(sd = sd(diff(logprice)), mu = mean(meanprice), cv = sd(meanprice)/mean(meanprice))

ggplot(grp, aes(x=log(mu), y=log(cv), color = factor(spec)) ) + geom_point() + xlab("Ln avg price") + ylab("Ln CV")
#ggplot(grp, aes(x=log(mu), y=log(cv), color = factor(spec)) ) + geom_point() + xlab("Ln avg price") + ylab("Ln CV")
```

We can also do some work looking at price elasticity (basically how much landings of one species goes up in relation to prices of others).

```{r, echo=FALSE, include=FALSE}
grp = group_by(cfec[cfec$specn%in%c(410:450) & cfec$pType != "",], spec, year, pType) %>% 
summarize(abund = sum(g_pounds)) %>% 
  data.frame

price.spec = group_by(cfec[cfec$specn%in%c(410:450) & cfec$pType != "",], year, pType) %>% 
  summarize(price.chum=mean(g_price[spec=="CHUM"]),
  price.coho=mean(g_price[spec=="COHO"]),
  price.chnk=mean(g_price[spec=="CHNK"]),
  price.pink=mean(g_price[spec=="PINK"]),
  price.sock=mean(g_price[spec=="SOCK"]))

grp = left_join(grp, price.spec)

group_by(grp, pType, spec) %>% 
  summarize(d.chum = diff(price.chum),
    d.coho = diff(price.coho),
    d.sock = diff(price.sock),
    d.pink = diff(price.pink),
    d.chnk = diff(price.chnk))

grp$abund = log(grp$abund)
grp$price.chum=log(grp$price.chum)
grp$price.coho=log(grp$price.coho)
grp$price.pink=log(grp$price.pink)
grp$price.sock=log(grp$price.sock)
grp$price.chnk=log(grp$price.chnk)
library(lme4)
chum.model = lm(abund ~ price.sock + price.coho + price.chnk + price.pink, grp[grp$spec=="CHUM",])
coho.model = lm(abund ~ price.chum + price.sock + price.chnk + price.pink, grp[grp$spec=="COHO",])
chnk.model = lm(abund ~ price.chum + price.coho + price.sock + price.pink, grp[grp$spec=="CHNK",])
pink.model = lm(abund ~ price.chum + price.coho + price.chnk + price.sock, grp[grp$spec=="PINK",])

coho.model = lmer(abund ~ price.sock + (1+price.sock|pType) + price.chnk + (1+price.chnk|pType) + price.chum + (1+price.chum|pType)+price.pink + (1+price.pink|pType), grp[grp$spec=="COHO",])

pink.model = lmer(abund ~ price.sock + (1+price.sock|pType) + price.chnk + (1+price.chnk|pType), grp[grp$spec=="PINK",])

chum.model = lmer(abund ~ price.sock + (1+price.sock|pType) + price.chnk + (1+price.chnk|pType), grp[grp$spec=="CHUM",])

chnk.model = lmer(abund ~ price.chnk + (1+price.chnk|pType) + price.coho + (1+price.coho|pType), grp[grp$spec=="SOCK",])

# Chum abund positively affected by price of pink, coho, negatively affected by sockeye price
# - chum abund goes down when sockeye price goes up
# Coho abund goes up with price of pink / chnk, down with price of chum
# Chnk abund goes up as price of chum/sockeye drop, goes up with price of pink
# Pink abund goes up with price of chum, down as price of coho/chnk increase
# Sock abund goes up as price of chum/chnk drops, or as price of coho/pink increase

```

### Is the increase in people targeting salmon a result of new people entering the fishery, or people switching from other things? New fisherman entering as salmon specialists has declined, but remained constant since ~ 2005 (at 400 people / year). People switching from non-salmon (groundfish / shellfish) to salmon is relatively low compared to the number of people who were fishing salmon + non-salmon species. This latter group represents people specializing toward salmon from a more diverse portfolio + peaked in the mid-1990s.

```{r, fig.pos="placeHere", echo=FALSE, fig.cap = "Separating salmon specialists into what they were doing the previous year"}
salmonParticipation.filter$group[which(salmonParticipation.filter$group==1)] = "New entrants"
salmonParticipation.filter$group[which(salmonParticipation.filter$group==2)] = "Was salmon specialist"
salmonParticipation.filter$group[which(salmonParticipation.filter$group==3)] = "Fished salmon & other"
salmonParticipation.filter$group[which(salmonParticipation.filter$group==4)] = "Fished, but not salmon"
salmonParticipation.filter$group[which(salmonParticipation.filter$group==5)] = "Skipped at least a year"

group_by(salmonParticipation.filter[salmonParticipation.filter$year > 1985 & is.na(salmonParticipation.filter$group)==FALSE,], year, group) %>% 
  summarize(n = n()) %>% 
  filter(is.na(n)==FALSE) %>% 
ggplot(aes(x = year, y = n, color = factor(group))) + geom_line() + facet_wrap(~ group, scales="free_y")
```

```{r, fig.pos="placeHere", echo=FALSE, fig.cap = "Separating salmon specialists into what they were doing the next year"}
salmonParticipation.filter$nextgroup[which(salmonParticipation.filter$nextgroup==1)] = "Never fished again"
salmonParticipation.filter$nextgroup[which(salmonParticipation.filter$nextgroup==2)] = "Stayed salmon specialist"
salmonParticipation.filter$nextgroup[which(salmonParticipation.filter$nextgroup==3)] = "Fished salmon & other"
salmonParticipation.filter$nextgroup[which(salmonParticipation.filter$nextgroup==4)] = "Fished, but no salmon"
salmonParticipation.filter$nextgroup[which(salmonParticipation.filter$nextgroup==5)] = "Skipped at least a year"

group_by(salmonParticipation.filter[salmonParticipation.filter$year < 2013 & is.na(salmonParticipation.filter$nextgroup)==FALSE,], year, nextgroup) %>% 
  summarize(n = n()) %>% 
  filter(is.na(n)==FALSE) %>% 
ggplot(aes(x = year, y = n, color = factor(nextgroup))) + geom_line() + facet_wrap(~ nextgroup, scales="free_y")
```

The latent permits can be broken out as well (not exactly latent permits, but people who are salmon specialists who aren't fishing at all).

```{r, fig.cap = "Proportion of salmon specialists not fishing", fig.pos = "placeHere", echo=FALSE}

dummy = expand.grid("p_holder"=unique(salmonParticipation.filter$p_holder), "year"=unique(salmonParticipation.filter$year))
latentSalmon = left_join(dummy, salmonParticipation.filter)
latentSalmon$active = 0
latentSalmon$active[which(latentSalmon$gear > 0)] =1

startEnd=group_by(latentSalmon, p_holder) %>%
  summarize(endYr = max(end, na.rm=T), startYr = min(start, na.rm=T)) 
latentSalmon = left_join(latentSalmon, startEnd)

latentSalmon=latentSalmon[which(latentSalmon$year <= latentSalmon$endYr & latentSalmon$year >= latentSalmon$startYr),]

group_by(latentSalmon[latentSalmon$year < 2014 & latentSalmon$year > 1985,], year) %>% 
  summarize(n = n(), nActive = length(which(active==1))) %>% 
  ggplot(aes(x = year, y = (n-nActive)/n)) + geom_line() + ylab("Proportion of salmon specialists not fishing")

#group_by(latentSalmon, startYr) %>% 
#  summarize(n = length(which(endYr==startYr))) %>% 
#  ggplot(aes(x = startYr, y = n)) + geom_line()
```

```{r, fig.cap = "Proportion of salmon specialists not fishing, by area", fig.pos = "placeHere", echo=FALSE}

# add the fishing location to the latentSalmon object
loc = group_by(cfec[cfec$pType!="",], p_holder, year) %>% 
  summarize(loc = pType[1])
latentSalmon = left_join(latentSalmon, loc)

group_by(latentSalmon[is.na(latentSalmon$loc) == FALSE & latentSalmon$loc != "" & latentSalmon$year < 2014 & latentSalmon$year > 1985,], year, loc) %>% 
  summarize(n = n(), nActive = length(which(active==1))) %>% 
  ggplot(aes(x = year, y = (n-nActive)/n, , color=factor(loc))) + geom_line() + ylab("Proportion of salmon specialists not fishing") + facet_wrap(~loc)
```

\newpage

### 2. PWS is an outlier in that there are now more salmon specialists than there ever have been

```{r, fig.cap = "Trends in salmon specialists (defined above), by region, 1985-2014", fig.pos = "placeHere", echo=FALSE}

salmonSpecialists = group_by(cfec, year, p_holder, landedLoc) %>% 
  summarize(gears = length(unique(gearn)), 
    salmonGears = length(unique(gearn[gearn%in%c(1,3,4)])),
    salmonSpec = length(unique(specn[specn%in%c(410,420,430,440,450)])),
    totRev = sum(g_earn), 
    effSpec = mean(effSpec),
    p.pink = mean(p.pink,na.rm=T),
    p.chum = mean(p.chum,na.rm=T),
    pType = pType[1]) %>% 
  filter(salmonGears == gears & salmonSpec >= 1)

salmonSummary = group_by(salmonSpecialists, year, landedLoc) %>% 
  summarize(ngear = mean(gears), ngear2 = length(which(gears == 2)), npeople = length(unique(p_holder)), nspec = mean(salmonSpec), neffspec = mean(effSpec),
    p.pink = mean(p.pink, na.rm=T), p.chum = mean(p.chum, na.rm=T), pType=pType[1]) %>% 
  filter(landedLoc %in% c("Bering Sea", "Aleutian Islands", "Dutch Harbor", "Statewide", "GOA", "EEZ")==FALSE)

ggplot(salmonSummary, aes(x=year, y=npeople) ) + geom_line() + facet_wrap(~ landedLoc, scales="free_y") + ylab("People exclusively using gears 1,3,4")
```

```{r, fig.cap = "Trends in salmon specialists (defined above), by region, 1985-2014", fig.pos = "placeHere", echo=FALSE}
group_by(cfec[cfec$pType!="",], year, pType) %>% 
  summarize(npeople = length(unique(p_holder))) %>%
ggplot(aes(x=year, y=npeople, color=factor(pType)) ) + geom_line() + facet_wrap(~ pType, scales="free_y") + ylab("People")
```

\newpage

### 3. As we saw in previous plots, trends in gear diversity and permit diversity are also declining. The number of people fishing 2 of the 3 gear types (set gillnets, drift gillnets, seine) has declined over time -- so people are reducing their portfolio of gears (+ potential species caught).

```{r, fig.cap = "Trends in salmon specialists using 2 gears, 1985-2014", fig.pos = "placeHere", echo=FALSE}
ggplot(salmonSummary, aes(x=year, y=ngear2/npeople) ) + geom_line() + facet_wrap(~ landedLoc, scales="free_y") + ylab("% People using 2 salmon gears")
```

\newpage

### 4. Effective species diversity is variable from year to year, and considerably lower in Bristol Bay than other places. However, decreased diversity of landings in Bristol Bay doesn't translate to more variable year-to-year revenue. Bristol Bay is among the lower CVs (below)

```{r, fig.cap = "Trends in effective species diversity (incl. non-salmon landed by gears 1,3,4, weighted by revenue), 1985-2014", fig.pos = "placeHere", echo=FALSE}
ggplot(salmonSummary, aes(x=year, y=neffspec) ) + geom_line() + facet_wrap(~ landedLoc) + ylab("Mean effective species diversity")
```

We can calculate the CV of each person (across years) and then plot these by region,

```{r, fig.cap = "CV of revenue, 1985-2014", fig.pos = "placeHere", echo=FALSE}
group_by(salmonSpecialists[salmonSpecialists$landedLoc%in%c("Aleutian Islands", "Bering Sea", "GOA")==FALSE,], p_holder) %>% 
  summarize(cv = sd(totRev, na.rm=T)/mean(totRev, na.rm=T), landedLoc = landedLoc[1]) %>% 
ggplot(aes(x = factor(landedLoc), y = cv)) + 
  geom_boxplot()
```

It may be also helpful to show the quartiles of the distribution,

```{r, fig.cap = "Quartile of revenue, 1985-2014", fig.pos = "placeHere", echo=FALSE}
salmonSpecialists$landedLoc = as.character(salmonSpecialists$landedLoc)
salmonQuantile = group_by(salmonSpecialists[salmonSpecialists$landedLoc%in%c("Aleutian Islands", "Bering Sea", "GOA")==FALSE,], p_holder) %>% 
  summarize(cv = sd(totRev, na.rm=T)/mean(totRev, na.rm=T), landedLoc = landedLoc[1]) %>% 
  group_by(landedLoc) %>% 
  summarize(lower25 = quantile(cv, 0.25, na.rm=T), 
    median50 = quantile(cv, 0.5, na.rm=T), 
    upper75 = quantile(cv,0.75, na.rm=T))

plot(as.factor(salmonQuantile$landedLoc), salmonQuantile$median50, ylab="CV", lwd=3, ylim=c(0,1.2))
points(as.factor(salmonQuantile$landedLoc), salmonQuantile$lower25, lwd=3)
points(as.factor(salmonQuantile$landedLoc), salmonQuantile$upper75, lwd=3)

```

We can also split this same trends in diversity out by gear / area (see Figure 2).

\newpage

### 5. The salmon specialists highlight the need to standardize by effort (days fished). We can calculate days fished as max (calendar day) - min (calendar day). The strongest trend is with respect to sockeye -- more days fished equals higher revenue (particularly in Alaska Peninsula, Bristol Bay, Kodiak, Cook Inlet).

```{r, echo=FALSE, fig.cap = "Time series of average days fished (max - min calendar days)", fig.pos = "placeHere"}
# Calculate season length
cfec$month = as.numeric(substr(cfec$day, 1, 2))
cfec$day = as.numeric(substr(cfec$day, 4,5))
cfec$calday = mdy.date(cfec$month, cfec$day, cfec$year) - mdy.date(1, 1, cfec$year)

salmonSpecialists = group_by(cfec, year, p_holder) %>% 
  summarize(gears = length(unique(gearn)), 
    salmonGears = length(unique(gearn[gearn%in%c(1,3,4)])),
    salmonSpec = length(unique(specn[specn%in%c(410,420,430,440,450)]))) %>% 
  filter(salmonGears == gears & salmonSpec >= 1) 
salmonSpecialists = left_join(salmonSpecialists, cfec)

# for each person find max day
maxDay = group_by(salmonSpecialists, p_holder, year) %>%
  summarize(maxDate = max(calday), minDate = min(calday))
salmonSpecialists = left_join(salmonSpecialists, maxDay)

group_by(salmonSpecialists[salmonSpecialists$landedLoc%in%c("Aleutian Islands", "GOA", "Statewide", "Bering Sea")==FALSE,], year, landedLoc) %>% 
  summarize(meanDays = mean(maxDate - minDate)) %>% 
  ggplot(aes(x=year, y=meanDays) ) + geom_line() + facet_wrap(~ landedLoc, scales="free") + ylab("Mean days fished")
```

```{r, echo=FALSE, fig.cap = "Time series of average days fished (max - min calendar days) by permit", fig.pos = "placeHere"}
group_by(salmonSpecialists[salmonSpecialists$pType!="",], year, pType) %>% 
  summarize(meanDays = mean(maxDate - minDate)) %>% 
  ggplot(aes(x=year, y=meanDays, color = factor(pType)) ) + geom_line() + facet_wrap(~ pType, scales="free") + ylab("Mean days fished")
```

```{r, echo=FALSE, fig.cap = "Time series of salmon revenue versus days fished", fig.pos="placeHere"}
# calculate total earnings of pink salmon versus mean cal days
salmSummary = group_by(salmonSpecialists[salmonSpecialists$specn%in%c(410:450),], year, p_holder, spec) %>% 
  summarize(totalDollars = sum(g_earn), 
    days = maxDate[1] - minDate[1],
    landedLoc = landedLoc[1]) %>% 
  group_by(year, landedLoc, spec) %>% 
  summarize(dollars = mean(totalDollars),
    days = mean(days))

ggplot(salmSummary, aes(x=log(dollars), y = days, color=factor(spec))) + geom_line() + facet_wrap(~ landedLoc, scales="free") + xlab("Ln Dollars") + ylab("Days fished")
```

```{r, echo=FALSE, fig.cap = "Time series of sockeye salmon revenue versus days fished", fig.pos="placeHere"}
ggplot(salmSummary[salmSummary$spec=="SOCK" & salmSummary$landedLoc%in%c("Bristol Bay", "Alaska Peninsula", "Prince William Sound", "Kodiak", "Southeast", "Cook Inlet"),], aes(x=log(dollars), y = days, color=factor(landedLoc))) + geom_line() + xlab("Ln Dollars") + ylab("Days fished")
```

```{r, echo=FALSE, fig.cap = "Time series of salmon revenue versus days fished", fig.pos="placeHere"}
# calculate total earnings of pink salmon versus mean cal days
salmSummary = group_by(salmonSpecialists[salmonSpecialists$specn%in%c(410:450),], year, p_holder, spec) %>% 
  summarize(totalDollars = sum(g_earn), 
    days = maxDate[1] - minDate[1],
    pType = pType[1]) %>% 
  group_by(year, pType) %>% 
  summarize(dollars = mean(totalDollars),
    days = mean(days))

ggplot(salmSummary, aes(x=log(dollars), y = days, color=factor(pType))) + geom_line() + facet_wrap(~ pType, scales="free") + xlab("Ln Dollars") + ylab("Days fished")
```

\newpage

## Focus on pink / chum salmon

To look at potential hatchery influences, we can examine the fraction of pink / chum landings relative to the total.

### This shows that the percentage landings of pink / sockeye salmon have been in the opposite direction for Kodiak / PWS / SEAK, increasing a bit for Cook Inlet / Alaska Peninsula.

```{r, fig.cap="Total percentage of earnings by species in each region", fig.pos="placeHere", echo=FALSE}
summary = group_by(salmonSpecialists[salmonSpecialists$specn%in%c(410:450) & salmonSpecialists$landedLoc%in%c("Aleutian Islands","Bering Sea","Statewide","GOA")==FALSE,], landedLoc, year, spec) %>% 
  summarize(spec.earn = sum(g_earn))
summary.tot = group_by(summary, year, landedLoc) %>% 
  summarize(n = sum(spec.earn))
summary = left_join(summary, summary.tot)

ggplot(summary, aes(x=year, y=spec.earn/n, color=factor(spec))) + facet_wrap(~ landedLoc) + geom_line() + ylab("Proportion of earnings")
```

```{r, fig.cap="Total percentage of earnings by species and permit", fig.pos="placeHere", echo=FALSE}
summary = group_by(salmonSpecialists[salmonSpecialists$specn%in%c(410:450),],  pType, year, spec) %>% 
  summarize(spec.earn = sum(g_earn))
summary.tot = group_by(summary, year, pType) %>% 
  summarize(n = sum(spec.earn))
summary = left_join(summary, summary.tot)

ggplot(summary[summary$pType!="",], aes(x=year, y=spec.earn/n, color=factor(spec))) + facet_wrap(~ pType) + geom_line() + ylab("Proportion of earnings")
```

```{r, fig.cap="Total earnings by species in each region", echo=FALSE, fig.pos="placeHere"}
group_by(salmonSpecialists[salmonSpecialists$specn%in%c(410:450) & salmonSpecialists$landedLoc%in%c("Aleutian Islands","Bering Sea","Statewide","GOA")==FALSE,], landedLoc, year, spec) %>% 
  summarize(dollars = sum(g_earn)) %>% 
ggplot(aes(x=year, y=log(dollars), color=factor(spec))) + geom_line() + facet_wrap(~ landedLoc)

#group_by(salmonSpecialists[salmonSpecialists$specn%in%c(410:450) & #salmonSpecialists$landedLoc%in%c("Aleutian Islands","Bering Sea","Statewide","GOA")==FALSE,], #landedLoc, year) %>% 
#  summarize(dollars = sum(g_earn)) %>% 
#ggplot(aes(x=year, y=log(dollars))) + geom_line() + facet_wrap(~ landedLoc)
```

```{r, fig.cap="Total earnings by species by permit", echo=FALSE, fig.pos="placeHere"}
group_by(salmonSpecialists[salmonSpecialists$specn%in%c(410:450) & salmonSpecialists$pType!="",], pType, year, spec) %>% 
  summarize(dollars = sum(g_earn)) %>% 
ggplot(aes(x=year, y=log(dollars), color=factor(spec))) + geom_line() + facet_wrap(~ pType)
```

```{r, fig.cap="Total landings by species in each region", echo=FALSE, fig.pos="placeHere"}
group_by(salmonSpecialists[salmonSpecialists$specn%in%c(410:450) & salmonSpecialists$landedLoc%in%c("Aleutian Islands","Bering Sea","Statewide","GOA")==FALSE,], landedLoc, year, spec) %>% 
  summarize(dollars = sum(g_pounds)) %>% 
ggplot(aes(x=year, y=log(dollars), color=factor(spec))) + geom_line() + facet_wrap(~ landedLoc) + ylab("Ln Pounds")
```

```{r, fig.cap="Total landings by species by permit", echo=FALSE, fig.pos="placeHere"}
group_by(salmonSpecialists[salmonSpecialists$specn%in%c(410:450) & salmonSpecialists$pType!="",], pType, year, spec) %>% 
  summarize(pounds = sum(g_pounds)) %>% 
ggplot(aes(x=year, y=log(pounds), color=factor(spec))) + geom_line() + facet_wrap(~ pType) + ylab("Ln Pounds")
```

One way to look at the effect of this diversification would be to compare the CV of revenue of people only fishing in the early 1990s to people only fishing in recent years (say 2005+).

```{r, echo=FALSE, fig.pos="placeHere"}
pinkChumFishers = group_by(salmonSpecialists[salmonSpecialists$pType!="",], p_holder, year) %>% 
  summarize(meanEarn = sum(g_earn),
    pType = pType[1]) %>% 
  group_by(p_holder, pType) %>% 
  summarize(cv.early = sd(meanEarn[year%in%c(1990:1999)]) / mean(meanEarn[year%in%c(1990:1999)]),
    cv.late = sd(meanEarn[year%in%c(2005:2014)]) / mean(meanEarn[year%in%c(2005:2014)])) %>%
  group_by(pType) %>% 
  summarize(cv.early = mean(cv.early,na.rm=T), cv.late = mean(cv.late,na.rm=T))
```

```{r, echo=FALSE}
kable(pinkChumFishers)
```

```{r, echo=FALSE, fig.pos="placeHere"}
ggplot(salmonSummary, aes(x = p.pink+p.chum, y = neffspec, color = factor(landedLoc))) + geom_line() + xlab("Proportion of pink and chum salmon") + ylab("Effective species diversity")
```

### 2016-06-14 evos-ifq/catch_portfolios_EVOSIFQ.Rmd
Start rearranging the files

--- 2016-03-25 catch_portfolios_mtg3.Rmd
@@
 title: "Results for working group papers"
-author: "Catch portfolios subgroup: NCEAS meeting 3"
+author: 'Catch portfolios subgroup: NCEAS meeting 3'
 date: "March 18, 2016"
@@
   html_document: default
+  word_document: default
 ---
 
-```{r setup, include=FALSE}
+```{r setup, include=FALSE, echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
 knitr::opts_chunk$set(echo = TRUE, tidy=FALSE, tidy.opts=list(width.cutoff=60), warning = FALSE, message = FALSE)
@@
 ## Load cleaned data  
-```{r dataLoad, warning = FALSE, message=FALSE}
+```{r dataLoad, warning = FALSE, message=FALSE, results="hide"}
 library(dplyr)
@@
 library(MASS)
+library(zoo)
 library(MatchIt)
-load("cfecCleaned.Rdata")
+load("cfecCleaned2.Rdata")
 ```
 
-```{r, warning = FALSE, message=FALSE}
-# This is to investigate Bristol Bay set nets for Milo
-# Check BB setnet records / year
-#idx = grep("S 04T", cfec$p_fshy)
-#group_by(cfec[idx,], year) %>% 
-#  summarize(nSetnet = length(unique(p_holder)))
-#cfecnew = cfec[idx,]
-```
+We need to adjust the landed values and prices for inflation. To do this, we'll use the data Alan pulled together (adjusting to 2009 dollars).
 
-## Function definiton  
-We need to define a few functions for calculating diversity.
-We'll use the Simpson's diversity index (HHH) to calculate the entropy-based inverse, which is a proxy for effective species diversity.
-Note the C++ version is commented out below,
-```{r, warning = FALSE, message=FALSE}
+https://docs.google.com/spreadsheets/d/1hRLYTX1nZBLKnYXwTCOtU7-XnDoMZf_1ufdTiNQFd4U/edit#gid=0
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
+deflationTable = data.frame(year = 1985:2014, defl = c(0.57236, 0.58393, 0.59879, 0.61974,  
+0.64388, 0.66774, 0.68993, 0.70564, 0.72244, 0.73781,  
+0.75321, 0.76695, 0.78009, 0.78855, 0.80061,  
+0.81883, 0.83753, 0.85038, 0.86729, 0.89114,  
+0.91981, 0.94812, 0.97334, 0.9925, 1,  
+1.01217, 1.03307, 1.05213, 1.06926, 1.08682))
+
+# Adjust price and g_earn for inflation
+cfec$year = as.numeric(cfec$year)
+cfec$g_price = cfec$g_price/deflationTable$defl[cfec$year-1984]
+cfec$g_earn = cfec$g_earn/deflationTable$defl[cfec$year-1984]
 
@@
 }
+```
 
-#library(Rcpp)
-#cppFunction('double simpCpp(NumericVector x) {
-#  double y = 1/sum((x/sum(x))^2);
-#  return y;
-#}')
+## Filtering out unique vessel - person combinations  
+Filter out people who only use 1 vessel / year , and similarly filter out boats with more than 1 person / year.
+Some people might fish on multiple vessels within a year, and vessels might be used by multiple people within a year.
+
+```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
+cfecnew = cfec
+cfec.1boat = group_by(cfecnew, p_holder, year) %>% 
+  summarize(nBoat = length(unique(cadfg))) %>%
+  filter(nBoat==1) %>% 
+  left_join(cfecnew)
+
+# Deal with setnet fisheries, where lots of p_holder
+# share same codes
+cfec.1boat = group_by(cfec.1boat, cadfg, year) %>% 
+  summarize(nPerson = length(unique(p_holder))) %>%
+  filter((nPerson==1 & cadfg %in% c(27960,27961) == FALSE) | cadfg %in% c(27960,27961)) %>% 
+  left_join(cfec.1boat)
+
+cfecnew = cfec.1boat
+dim(cfecnew)[1]
 ```
@@
 
-```{r, warning = FALSE, message=FALSE}
+```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
 # We'll use 3 years before/after IFQ. This is consistent with the work that
@@
 
-```{r, cache = TRUE, warning = FALSE, message=FALSE}
+```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
 
@@
 
-```{r, fig.pos="placeHere", fig.cap="Plot of species diversity metrics (calculated by fish ticket freq.) through time for each of our 4 groups during the pre-post IFQ period. "}
+```{r, fig.pos="placeHere", fig.cap="Plot of species diversity metrics (calculated by fish ticket freq.) through time for each of our 4 groups during the pre-post IFQ period.
+", echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
 ggplot(hs.day.summary, aes(year, m.freq)) + geom_line() + facet_wrap(~group) + 
@@
 
-```{r, fig.pos="placeHere", fig.cap="Plot of species diversity metrics (calculated by earnings) through time for each of our 4 groups during the pre-post IFQ period. "}
+```{r, fig.pos="placeHere", fig.cap="Plot of species diversity metrics (calculated by earnings) through time for each of our 4 groups during the pre-post IFQ period.
+", echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
 ggplot(hs.day.summary, aes(year, m.earn)) + geom_line() + facet_wrap(~group) + 
@@
 
-```{r, fig.pos="placeHere", fig.cap="Plot of species diversity metrics (calculated by pounds of catch) through time for each of our 4 groups during the pre-post IFQ period. "}
+```{r, fig.pos="placeHere", fig.cap="Plot of species diversity metrics (calculated by pounds of catch) through time for each of our 4 groups during the pre-post IFQ period.
+", echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
 ggplot(hs.day.summary, aes(year, m.weight)) + geom_line() + facet_wrap(~group) + 
@@
 First, make a dotplot (with quartile range) of the pre-post comparison for each metric of revenue.  
-```{r, fig.pos="placeHere", fig.cap = "Metric of annual revenue, with interquartile range, for each of the 4 IFQ groups."}
+```{r, fig.pos="placeHere", fig.cap = "Metric of annual revenue, with interquartile range, for each of the 4 IFQ groups.", echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
 
@@
 Look at the mean number of days where catch was recorded as landed, as well as how many of those days are when halibut are landed.  
-```{r}
+```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
 dplyr::select(hs.summary, group, ind, p_holder, 
@@
 
-```{r, fig.pos="placeHere", fig.cap = "Boxplots of revenue metrics (with interquartile range) by group and pre-post-IFQ. ", include=FALSE, echo=FALSE, message = FALSE, warning = FALSE}
+```{r, fig.pos="placeHere", fig.cap = "Boxplots of revenue metrics (with interquartile range) by group and pre-post-IFQ. ", echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
 #dplyr::select(hs.summary, group, ind, p_holder, contains("mean"), -contains("div")) %>% 
@@
 
-```{r, fig.pos="placeHere", fig.cap = "Boxplots of diversity metrics (with interquartile range) by group and pre-post-IFQ. "}
+```{r, fig.pos="placeHere", fig.cap = "Boxplots of diversity metrics (with interquartile range) by group and pre-post-IFQ. ", echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
 dplyr::select(hs.summary, group, ind, p_holder, contains("eff.spec.div.mean"), contains("eff.gear.div.mean"), contains("eff.spec.div.mean")) %>% 
@@
 
-```{r, fig.pos="placeHere", fig.cap="Violin plot of individual differences (post-IFQ - pre-IFQ for the revenue and diversity metrics included in our analysis."}
+```{r, fig.pos="placeHere", fig.cap="Violin plot of individual differences (post-IFQ - pre-IFQ for the revenue and diversity metrics included in our analysis.", echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
 hs.long = reshape2::melt(hs.summary, id.vars = c("group", "ind","p_holder"))
@@
 
-```{r, fig.pos = "placeHere", fig.cap = "Gear diversity through time for each of the 3 groups, calculated based on the catch (black) and earnings (red)."}
+```{r, fig.pos = "placeHere", fig.cap = "Gear diversity through time for each of the 3 groups, calculated based on the catch (black) and earnings (red).", echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
 hs.summary.gear = group_by(dat, ind, group, year, gearn) %>% 
@@
 We'll break down the proportion used for the top 8 gears,
-```{r, fig.pos="placeHere", fig.cap = "Time series of proportion of gears used"}
+```{r, fig.pos="placeHere", fig.cap = "Time series of proportion of gears used", echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
 hs.gear.tot = group_by(dat, ind, group, year) %>% 
@@
 Can we predict effective species diversity as function of gear diversity?
-```{r}
+```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
 #hs.geardiv = group_by(dat, ind, group, p_holder, year, gearn) %>% 
@@
 
+\newpage
+
 ## EVOS comparison
 
-We'll restrict our analysis initially to a few years before and after EVOS (1989). To make the comparison similar to the IFQ comparison, we will define the following groups:
+We'll restrict our analysis initially to a few years before and after EVOS (1989).
+To make the comparison similar to the IFQ comparison, we will have an impacted group.
+We'll define 'impacted' as people who were landing catch in statistical areas over 1985-1988 that were later closed because of the spill (1989).
+These areas were then opened up for fishing in 1990.
+We'll restrict the analysis to only looking at immediate impacts, so we're interested in how affected / unaffected individuals fared in 1990-1991.
 
-1. Individuals who fished in closed areas before and after spill  
-2. Individuals who fished in closed areas before spill ["Pre"]
-3. Individuals who only fished in closed areas after 1995 ["Post"]  
-4. Individuals who never fished in closed areas ["Control group"]
+To create a control group, we'll use propensity matching (other clustering methods also an option). This procedure involves:  
+(1) defining attributes for all individuals during 1985-1988 (total catch, total revenue, % revenue from salmon, etc).  
+(2) Doing some simple model selection with a binomial GLM to reduce the number of variables included in our model (some will be redundant with others, and others will be uninformative)  
+(3) Run propensity matching to identify among the unaffected individuals a best match for each impacted individual. Note that this will only identify 1 match for each case. 
+(4) Using the affected individuals and new control group of 'matches', examine how the two groups fared post spill, in 1990-1991.
+(As an aside, Rich thought many individuals fishing salmon, etc.
+in these areas were going right back to the previously closed areas).
 
 ## Propensity matching for EVOS
-We're interested in forming a control group for the comparison of the EVOS effects...
-```{r, warning = FALSE, message=FALSE}
+
+We're interested in forming a control group for the comparison of the EVOS effects.  To do this, we adopted the following approach:  
+1.  Filter the data to be only 1985:1988, and 1990:1991 as our pre/post group. Only use data for people who fished before and after.    
+2.  Create an 'exposed' treatment effect for anyone who fished one of the following permits before 1989: Kodiak setnet, Kodiak purse seine, Cook Inlet drift net  
+3.
+Calculate the pounds landed per species and earnings per species for the dominant ~ 15 species from these groups (mostly salmon, shellfish, etc).
+Also calculate total annual revenue, mean days fished (landed days) and diversity metrics (species, gear, etc).
+4.  Fit a GLM (logistic regression) to the pre-spill data to identify the best predictors of classifying exposure.    
+5.  With the same model in (4) use propensity score matching to find best matches for individual fisherman in the "pre" & exposed group, to create a control group  
+6. Calculate BACI style metrics, similar to the IFQ comparison to look at changes in groups (control, treatment) pre/post 1989
+
+### Take home: there doesn't seem to be a negative effect of revenue on the treatement / exposed group.
+But pounds (and revenue) of pink (#440) and chum (#450) salmon decreased more so for this group, and species diversity also appears to be affected.
+People in the not exposed group fished more coho post-EVOS.
+
+```{r, warning = FALSE, message=FALSE, results = "hide", echo=FALSE}
 
@@
 dat$day = substr(dat$landdate, 6, 10)
+
 # Create an indicator based on the salmon / shellfish araes that were closed in PWS
+# If someone was fishing in the area before EVOS, they are affected
+#dat$exposed = 0
+#dat$exposed[dat$stat6%in%c(22710, 22720, 22730, 22610, 22620, 22630, 22640, 
+#  22650, 22510, 22520, 22530) & dat$year < 1989] = 1
+
+# For the new analysis (June 2016), based on Eric's plots of the salmon fisherment, the 
+# most obvious 'exposed' group are people who had the following permits:
+# Kodiak setnet, Kodiak purse seine, Cook Inlet drift net. PWS set net could potentially be incl,
+# but participation in 1989 dropped from 560 to 500 people -- didn't go to 0
 dat$exposed = 0
-dat$exposed[dat$stat6%in%c(22710, 22720, 22730, 22610, 22620, 22630, 22640, 22650, 22510, 22520, 22530)] = 1
+dat$exposed[which(dat$p_fshy%in%c("S 03H","S 04K","S 01K") & dat$year <= 1989)] = 1
 
@@
   summarize(tot.revenue.ind = sum(g_earn), 
-    tot.revenue.ind.her = sum(g_earn[specn==230]), 
-    tot.revenue.ind.sal = sum(g_earn[specn%in%c(400,410,420,430,440,450)]),
+    tot.rev.230 = sum(g_earn[specn==230]),
+    tot.rev.420 = sum(g_earn[specn==420]), 
+    tot.rev.450 = sum(g_earn[specn==450]), 
+    tot.rev.430 = sum(g_earn[specn==430]), 
+    tot.rev.440 = sum(g_earn[specn==440]), 
+    tot.rev.410 = sum(g_earn[specn==410]),
+    tot.rev.710 = sum(g_earn[specn==710]),
+    tot.rev.910 = sum(g_earn[specn==910]), 
+    tot.rev.400 = sum(g_earn[specn==400]), 
+    tot.rev.168 = sum(g_earn[specn==168]), 
+    tot.rev.270 = sum(g_earn[specn==270]),    
+    tot.rev.130 = sum(g_earn[specn==130]),
+    tot.rev.931 = sum(g_earn[specn==931]),
+    tot.rev.830 = sum(g_earn[specn==830]),
+    tot.rev.520 = sum(g_earn[specn==520]),    
     tot.pounds.ind = sum(g_pounds), 
-    tot.pounds.ind.her = sum(g_pounds[specn==230]), 
-    tot.pounds.ind.sal = sum(g_pounds[specn%in%c(400,410,420,430,440,450)]),
+    tot.pnds.230 = sum(g_pounds[specn==230]),
+    tot.pnds.420 = sum(g_pounds[specn==420]), 
+    tot.pnds.450 = sum(g_pounds[specn==450]), 
+    tot.pnds.430 = sum(g_pounds[specn==430]), 
+    tot.pnds.440 = sum(g_pounds[specn==440]), 
+    tot.pnds.410 = sum(g_pounds[specn==410]),
+    tot.pnds.710 = sum(g_pounds[specn==710]),
+    tot.pnds.910 = sum(g_pounds[specn==910]), 
+    tot.pnds.400 = sum(g_pounds[specn==400]), 
+    tot.pnds.168 = sum(g_pounds[specn==168]), 
+    tot.pnds.270 = sum(g_pounds[specn==270]),    
+    tot.pnds.130 = sum(g_pounds[specn==130]),
+    tot.pnds.931 = sum(g_pounds[specn==931]),
+    tot.pnds.830 = sum(g_pounds[specn==830]),
+    tot.pnds.520 = sum(g_pounds[specn==520]),    
     eff.spec.div = simp.div(table(specn)), 
@@
     ind = ifelse(year[1] < 1989, "Pre", "Post")) %>% 
-    as.data.frame() %>%
-  group_by(p_holder, ind) %>% 
+    as.data.frame()
+
+# Now summarize over years 
+  dat.ev = group_by(dat.ev, p_holder, ind) %>% 
   summarize(tot.revenue.ind = mean(tot.revenue.ind), 
-    tot.revenue.ind.her = mean(tot.revenue.ind.her), 
-    tot.revenue.ind.sal = mean(tot.revenue.ind.sal),
     tot.pounds.ind = mean(tot.pounds.ind), 
-    tot.pounds.ind.her = mean(tot.pounds.ind.her), 
-    tot.pounds.ind.sal = mean(tot.pounds.ind.sal),
     eff.spec.div = mean(eff.spec.div), 
@@
     exposed = max(exposed),
-    meanDays = mean(numDays)) %>% 
+    meanDays = mean(numDays),
+    tot.rev.230 = mean(tot.rev.230)/tot.revenue.ind,
+    tot.rev.420 = mean(tot.rev.420)/tot.revenue.ind, 
+    tot.rev.450 = mean(tot.rev.450)/tot.revenue.ind, 
+    tot.rev.430 = mean(tot.rev.430)/tot.revenue.ind, 
+    tot.rev.440 = mean(tot.rev.440)/tot.revenue.ind, 
+    tot.rev.410 = mean(tot.rev.410)/tot.revenue.ind,
+    tot.rev.710 = mean(tot.rev.710)/tot.revenue.ind,
+    tot.rev.910 = mean(tot.rev.910)/tot.revenue.ind, 
+    tot.rev.400 = mean(tot.rev.400)/tot.revenue.ind, 
+    tot.rev.168 = mean(tot.rev.168)/tot.revenue.ind, 
+    tot.rev.270 = mean(tot.rev.270)/tot.revenue.ind,    
+    tot.rev.130 = mean(tot.rev.130)/tot.revenue.ind,
+    tot.rev.931 = mean(tot.rev.931)/tot.revenue.ind,
+    tot.rev.830 = mean(tot.rev.830)/tot.revenue.ind,
+    tot.rev.520 = mean(tot.rev.520)/tot.revenue.ind,    
+    tot.pnds.230 = mean(tot.pnds.230)/tot.pounds.ind,
+    tot.pnds.420 = mean(tot.pnds.420)/tot.pounds.ind, 
+    tot.pnds.450 = mean(tot.pnds.450)/tot.pounds.ind, 
+    tot.pnds.430 = mean(tot.pnds.430)/tot.pounds.ind, 
+    tot.pnds.440 = mean(tot.pnds.440)/tot.pounds.ind, 
+    tot.pnds.410 = mean(tot.pnds.410)/tot.pounds.ind,
+    tot.pnds.710 = mean(tot.pnds.710)/tot.pounds.ind,
+    tot.pnds.910 = mean(tot.pnds.910)/tot.pounds.ind, 
+    tot.pnds.400 = mean(tot.pnds.400)/tot.pounds.ind, 
+    tot.pnds.168 = mean(tot.pnds.168)/tot.pounds.ind, 
+    tot.pnds.270 = mean(tot.pnds.270)/tot.pounds.ind,    
+    tot.pnds.130 = mean(tot.pnds.130)/tot.pounds.ind,
+    tot.pnds.931 = mean(tot.pnds.931)/tot.pounds.ind,
+    tot.pnds.830 = mean(tot.pnds.830)/tot.pounds.ind,
+    tot.pnds.520 = mean(tot.pnds.520)/tot.pounds.ind
+    ) %>% 
     as.data.frame()
@@
 
+# We've calculated pounds and revenue of each of the top 15 spp -- including
+# both is potentially problematic though because they're correlated ,as expected.
+# So for now we'll just focus on revenue. 
+dat.ev.rf = dat.ev.rf[,-c(4, 24:38, 40)]
+
 #rf = randomForest(x = dat.ev.rf, y = as.factor(dat.ev.rf$exposed))
 
-dat.ev.rf[,3:8] = log(dat.ev.rf[,3:8] + 1)
-
+dat.ev.rf[,c(3:5, 23)] = log(dat.ev.rf[,c(3:5, 23)] + 1)
+# Select variables that can help explain variation in exposure 
 m.glm = glm(exposed ~ ., data = dat.ev.rf[dat.ev.rf$ind=="Pre", -c(1,2)], family = "binomial")
@@
 
-bestModel = glm(exposed ~ tot.revenue.ind.her + tot.revenue.ind.sal + 
-    tot.pounds.ind + tot.pounds.ind.her + tot.pounds.ind.sal + 
-    eff.spec.div + meanDays + eff.spec.div.mean.weight, data = 
+bestModel = glm(exposed ~ tot.revenue.ind + eff.spec.div + eff.gear.div + meanDays + 
+    tot.rev.420 + tot.rev.450 + tot.rev.430 + tot.rev.410 + 
+    tot.rev.710 + tot.rev.400 + tot.rev.168 + tot.rev.270, data = 
     dat.ev.rf[dat.ev.rf$ind == "Pre", ])
 
-m.out <- matchit(exposed ~ tot.revenue.ind + tot.revenue.ind.her + 
-    tot.revenue.ind.sal + tot.pounds.ind + tot.pounds.ind.her + 
-    tot.pounds.ind.sal + eff.spec.div + eff.gear.div + meanDays + 
-    eff.spec.div.mean.earn, method = "nearest", data = dat.ev.rf[dat.ev.rf$ind == 
-    "Pre", ])
+m.out <- matchit(exposed ~ tot.revenue.ind + eff.spec.div + eff.gear.div + meanDays + 
+    tot.rev.420 + tot.rev.450 + tot.rev.430 + tot.rev.410 + 
+    tot.rev.710 + tot.rev.400 + tot.rev.168 + tot.rev.270, method = "nearest", data = dat.ev.rf[dat.ev.rf$ind == "Pre", ])
 
@@
 dat.matched$ind2 = dat.matched$plotind
-dat.matched$ind2[dat.matched$exposed==1] = dat.matched$ind2[dat.matched$exposed==1] + 0.2
+dat.matched$ind2[dat.matched$exposed==1] = 
+  dat.matched$ind2[dat.matched$exposed==1] + 0.2
 
@@
 
-ggplot(dat.matched.summary, aes(x = ind2, y = m, ymax = u, ymin = l, col = as.factor(exposed))) + geom_pointrange() + facet_wrap(~ variable, scales="free")
+# break this into 2 plots because of the panel size
+ggplot(dat.matched.summary[dat.matched.summary$variable %in% unique(dat.matched.summary$variable)[1:10], ], aes(x = ind2, y = m, ymax = u, ymin = l, col = as.factor(exposed))) + 
+  geom_pointrange() + facet_wrap(~ variable, scales="free")
+
+ggplot(dat.matched.summary[dat.matched.summary$variable %in% unique(dat.matched.summary$variable)[11:19], ], aes(x = ind2, y = m, ymax = u, ymin = l, col = as.factor(exposed))) + 
+  geom_pointrange() + facet_wrap(~ variable, scales="free")
+
+ggplot(dat.matched.summary[dat.matched.summary$variable %in% unique(dat.matched.summary$variable)[20:29], ], aes(x = ind2, y = m, ymax = u, ymin = l, col = as.factor(exposed))) + 
+  geom_pointrange() + facet_wrap(~ variable, scales="free")
+
+ggplot(dat.matched.summary[dat.matched.summary$variable %in% unique(dat.matched.summary$variable)[30:39], ], aes(x = ind2, y = m, ymax = u, ymin = l, col = as.factor(exposed))) + 
+  geom_pointrange() + facet_wrap(~ variable, scales="free")
 
@@
 
-```{r}
+```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
 
-ev.long = reshape2::melt(dat.matched, id.vars = c("exposed","ind","p_holder"))
-ev.long = reshape2::dcast(ev.long, variable + p_holder + exposed ~ ind) %>%
-  mutate(diff_ = (Post-Pre)/Pre)
+#ev.long = reshape2::melt(dat.matched, id.vars = c("exposed","ind","p_holder"))
+#ev.long = reshape2::dcast(ev.long, variable + p_holder + exposed ~ ind) %>%
+#  mutate(diff_ = (Post-Pre)/Pre)
 
-ev.long = filter(ev.long, !is.na(diff_), is.finite(diff_))
-ggplot(ev.long, aes(exposed, diff_)) + 
-  geom_violin(aes(fill=exposed), alpha=0.5) + 
-  facet_wrap(~ variable, scales="free") +
-  geom_point(alpha = 0.02, position = 
-  position_jitter(width = 0.4), cex = 0.2)
+#ev.long = filter(ev.long, !is.na(diff_), is.finite(diff_))
+#ggplot(ev.long, aes(exposed, diff_)) + 
+#  geom_violin(aes(fill=exposed), alpha=0.5) + 
+#  facet_wrap(~ variable, scales="free") +
+#  geom_point(alpha = 0.02, position = 
+#  position_jitter(width = 0.4), cex = 0.2)
 
@@
 
-```{r}
+```{r, echo=FALSE, warning=FALSE, message=FALSE, results="hide"}
 #cfec$areaPermit = 0

### 2016-06-14 portfolio/analysis/1.5-createData_linear_modeling.Rmd
Start rearranging the files

--- 2016-06-09 analysis_portfolio/1.5-createData_linear_modeling.Rmd
@@
     revenue = sum(g_earn),
-    length = mean(VLength, na.rm=T)) %>% 
+    length = mean(VLength, na.rm=T))
+
+cfecAnnual = group_by(cfecAnnual, p_holder, year) %>%
   filter(npermit == 1)
@@
 
-temp = group_by(cfecnew, p_holder, year, specn) %>% 
-  summarize(totRev.spec = sum(g_earn), 
-    permit = p_fshy[1])
-temp = left_join(temp, group_by(cfecnew, p_holder, year) %>% 
-    summarize(totRev = sum(g_earn)))
-temp$totRev.spec = temp$totRev.spec / temp$totRev
+#temp = group_by(cfecnew, p_holder, year, specn) %>% 
+#  summarize(totRev.spec = sum(g_earn), 
+#    permit = p_fshy[1])
+#temp = left_join(temp, group_by(cfecnew, p_holder, year) %>% 
+#    summarize(totRev = sum(g_earn)))
+#temp$totRev.spec = temp$totRev.spec / temp$totRev
 
-mkt = group_by(cfecnew, p_fshy, year, specn) %>% 
-  summarize(totRev.mkt.spec = sum(g_earn)) 
-mkt = left_join(mkt, group_by(mkt, p_fshy, year) %>% 
-  summarize(totRev.mkt = sum(totRev.mkt.spec))) 
-mkt$totRev.mkt.spec = mkt$totRev.mkt.spec / mkt$totRev.mkt
-names(mkt)[1] = "permit"
+#mkt = group_by(cfecnew, p_fshy, year, specn) %>% 
+#  summarize(totRev.mkt.spec = sum(g_earn)) 
+#mkt = left_join(mkt, group_by(mkt, p_fshy, year) %>% 
+#  summarize(totRev.mkt = sum(totRev.mkt.spec))) 
+#mkt$totRev.mkt.spec = mkt$totRev.mkt.spec / mkt$totRev.mkt
+#names(mkt)[1] = "permit"
 
 # combine individual and market datasets
-temp = left_join(temp, mkt)
+#temp = left_join(temp, mkt)
 
 # Calculate squared proportions
-temp$p2 = (temp$totRev.spec - temp$totRev.mkt.spec)^2
+#temp$p2 = (temp$totRev.spec - temp$totRev.mkt.spec)^2
 
-cfecAnnual = left_join(cfecAnnual, dplyr::select(temp, p_holder, year, p2))
+#cfecAnnual = left_join(cfecAnnual, dplyr::select(temp, p_holder, year, p2))
 
@@
 cfecAnnual$geardiv.prev = cfecAnnual$gearDiv[idx]
-cfecAnnual$p2.prev = cfecAnnual$p2[idx]
+#cfecAnnual$p2.prev = cfecAnnual$p2[idx]
 
@@
 cfecAnnual.salmon = cfecAnnual[cfecAnnual$salmon=="Y",]
+cfecAnnual.nonsalmon = cfecAnnual[cfecAnnual$salmon=="N",]
 
 cfecAnnual.salmon = cfecAnnual.salmon[is.na(cfecAnnual.salmon$specDiv.change)==F, ]
+cfecAnnual.nonsalmon = cfecAnnual.nonsalmon[is.na(cfecAnnual.nonsalmon$specDiv.change)==F, ]
+
 saveRDS(cfecAnnual.salmon, file="salmon_linearModeling_complete.rds")
+saveRDS(cfecAnnual.nonsalmon, file="salmon_linearModeling_complete.rds")
 ```
+
+We can look at whether there's clear combinations of permit types that are combinations -- most of these people either have a single local permit (PWS, Bristol Bay, etc) or have multiple statewide permits.
+
+```{r}
+
+cfecnew$area = substr(cfecnew$p_fshy,4,4)
+cfecnew$area[cfecnew$area %in% c("A","C")] = "Southeast"
+cfecnew$area[cfecnew$area == "B"] = "Statewide"
+cfecnew$area[cfecnew$area == "D"] = "Yakutat"
+cfecnew$area[cfecnew$area == "E"] = "PWS"
+cfecnew$area[cfecnew$area == "H"] = "CookInlet"
+cfecnew$area[cfecnew$area == "J"] = "Westward"
+cfecnew$area[cfecnew$area == "K"] = "Kodiak"
+cfecnew$area[cfecnew$area == "L"] = "Chignik"
+cfecnew$area[cfecnew$area == "M"] = "AlaskaPeninsula"
+cfecnew$area[cfecnew$area == "N"] = "Nelson"
+cfecnew$area[cfecnew$area == "Q"] = "BeringSea"
+cfecnew$area[cfecnew$area == "T"] = "BristolBay"
+cfecnew$area[cfecnew$area == "X"] = "Kotzebue"
+cfecnew$area[cfecnew$area == "Z"] = "NortonSound"
+
+cfecnew$spec = substr(cfecnew$p_fshy,1,1)
+cfecnew$spec[cfecnew$spec == "B"] = "Halibut"
+cfecnew$spec[cfecnew$spec == "C"] = "Sablefish"
+cfecnew$spec[cfecnew$spec %in% c("D","E","K","T")] = "Crab"
+cfecnew$spec[cfecnew$spec %in% c("F")] = "Freshwater"
+cfecnew$spec[cfecnew$spec %in% c("G","H","L")] = "Herring"
+cfecnew$spec[cfecnew$spec %in% c("I")] = "Lingcod"
+cfecnew$spec[cfecnew$spec %in% c("A","J","N","R","W")] = "Shellfish"
+cfecnew$spec[cfecnew$spec %in% c("O","P","Q","U","Z")] = "Invert"
+cfecnew$spec[cfecnew$spec %in% c("M")] = "Misc"
+cfecnew$spec[cfecnew$spec %in% c("S")] = "Salmon"
+cfecnew$area.spec = paste(cfecnew$area, cfecnew$spec, sep="_")
+
+cfecAnnual.multiplePermits = group_by(cfecnew, p_holder, year) %>%
+  summarize(npermit = length(unique(p_fshy)), 
+    allpermits = paste(sort(unique(p_fshy)), collapse = " "),
+    areaspec = paste(sort(unique(area.spec)), collapse = " ")) %>%
+  filter(npermit > 1)
+
+```
+
+## Model fitting
+
+Start by fitting some lmer models to revenue for the non-salmon people as an example (smaller dataset).
+
+```{r}
+library(lme4)
+
+cfecAnnual.salmon$spec = substr(cfecAnnual.salmon$permit,1,1)
+cfecAnnual.salmon$area = substr(cfecAnnual.salmon$permit,4,4)
+
+full.mod = lmer(log(revenue) ~ I(log(revenue.prev)) + specDiv +specDiv.change+ (1+specDiv|permit) + length + days, data = cfecAnnual.salmon[complete.cases(cfecAnnual.salmon),])
+
+full.mod.ind = lmer(log(revenue) ~ permit*as.factor(year) + specDiv + (1+specDiv|permit) + length + days + (1|p_holder), data = cfecAnnual.salmon[complete.cases(cfecAnnual.salmon),])
+
+group_by(cfecAnnual.salmon[complete.cases(cfecAnnual.salmon),], p_holder) %>% 
+  summarize(earn = mean(log(revenue)), cv = sd(log(revenue)) / earn, div = mean(specDiv), permit = permit[1]) %>%
+  as.data.frame %>%
+  group_by(permit) %>% 
+  mutate(n = n()) %>% 
+  filter(n > 50) %>% 
+  ggplot(aes(x = div, y = earn, color = cv)) + geom_point() + facet_wrap(~ permit, scales = "free")
+```

### 2016-06-14 portfolio/analysis/4a-createData_linear_modeling.Rmd
Start rearranging the files

--- 2016-06-14 portfolio/analysis/1.5-createData_linear_modeling.Rmd
@@
 saveRDS(cfecAnnual.salmon, file="salmon_linearModeling_complete.rds")
-saveRDS(cfecAnnual.nonsalmon, file="salmon_linearModeling_complete.rds")
+saveRDS(cfecAnnual.nonsalmon, file="nonsalmon_linearModeling_complete.rds")
+
 ```
 
+## Code for looking at permit 
 We can look at whether there's clear combinations of permit types that are combinations -- most of these people either have a single local permit (PWS, Bristol Bay, etc) or have multiple statewide permits.
@@
 ```
-
-## Model fitting
-
-Start by fitting some lmer models to revenue for the non-salmon people as an example (smaller dataset).
-
-```{r}
-library(lme4)
-
-cfecAnnual.salmon$spec = substr(cfecAnnual.salmon$permit,1,1)
-cfecAnnual.salmon$area = substr(cfecAnnual.salmon$permit,4,4)
-
-full.mod = lmer(log(revenue) ~ I(log(revenue.prev)) + specDiv +specDiv.change+ (1+specDiv|permit) + length + days, data = cfecAnnual.salmon[complete.cases(cfecAnnual.salmon),])
-
-full.mod.ind = lmer(log(revenue) ~ permit*as.factor(year) + specDiv + (1+specDiv|permit) + length + days + (1|p_holder), data = cfecAnnual.salmon[complete.cases(cfecAnnual.salmon),])
-
-group_by(cfecAnnual.salmon[complete.cases(cfecAnnual.salmon),], p_holder) %>% 
-  summarize(earn = mean(log(revenue)), cv = sd(log(revenue)) / earn, div = mean(specDiv), permit = permit[1]) %>%
-  as.data.frame %>%
-  group_by(permit) %>% 
-  mutate(n = n()) %>% 
-  filter(n > 50) %>% 
-  ggplot(aes(x = div, y = earn, color = cv)) + geom_point() + facet_wrap(~ permit, scales = "free")
-```

### 2016-06-14 archive/catch_portfolios_mtg2.Rmd
More file rearranging

--- 2016-03-18 analyze_cfec_data.Rmd
@@
 
-## Overview
+# Overview
 We are interested in multiple metrics of diversity. For this first paper, we plan to divide the analysis into several components:  
@@
 opts_chunk$set(tidy=TRUE, message=FALSE)
-```
-
-## Data processing
-Load the data from the saved workspace.
-This is the raw cfec data that has been merged to have the same columns across years, 1975 - present.
-The key columns are the species labels ("spec") and species number codes ("specn"), the gear number codes ("gearn"), the vessel id ("adfg"), the area where the catch occurred ("stat6"), the pounds harvested ("g_pounds"), gross earnings ("g_earn").
-
-### Note: for any confidentiality concerns, the workspace loaded here has been slightly jittered. This will need to be replaced with the real data for the final model. 
-```{r loadData}
-load("cfec.Rdata")
-# initially only use data 1985+
-cfec = cfec[cfec$year>1984,]
-```
-
-The first data processing step is to aggregate some of the species.
-This is largely needed because of changes in reporting over time (before 1985 for example, lots of the catch was from foreign fleets & not reported to species level -- for example, some of the rockfishes).
-For other species, there are unidentified groups, so there's additional aggregation required.
-
-```{r specLump}
-# Lump all skates together
-cfec$spec = as.character(cfec$spec)
-cfec$specn[cfec$specn==700] = 700
-cfec$specn[cfec$specn==701] = 700
-cfec$specn[cfec$specn==702] = 700
-cfec$spec[cfec$specn==700] = "SKATE"
-
-# Lump all sharks together
-cfec$specn[cfec$specn==691] = 689
-cfec$specn[cfec$specn==689] = 689
-cfec$spec[cfec$specn==689] = "SHARK"
-
-# Lump flatfish (Shallow)
-cfec$specn[cfec$specn==123] = 119 # rock
-cfec$specn[cfec$specn==129] = 119 # starry flounder
-cfec$specn[cfec$specn==127] = 119 # yellowfin
-cfec$specn[cfec$specn==126] = 119 # butter
-cfec$spec[cfec$specn==119] = "FLAT.shallow"
-# 5, 6, 7, 15, 26, 47, 61, 
-# Lump flatfish (deep)
-cfec$specn[cfec$specn==124] = 118 # dover
-cfec$specn[cfec$specn==134] = 118 # greenland turbot
-cfec$spec[cfec$specn==118] = "FLAT.deep"
-
-# Lump rockfish, codes from here:
-cfec$specn[cfec$specn==141] = 144 # POP
-cfec$specn[cfec$specn==136] = 144 # northern
-cfec$specn[cfec$specn==151] = 144 # rougheye
-cfec$specn[cfec$specn==166] = 144 # sharpchin
-cfec$specn[cfec$specn==152] = 144 # shortraker
-cfec$specn[cfec$specn==153] = 144 # redbanded
-cfec$specn[cfec$specn==157] = 144 # silvergray
-cfec$specn[cfec$specn==158] = 144 # redstripe
-cfec$specn[cfec$specn==137] = 144 # bocaccio
-cfec$specn[cfec$specn==144] = 144 # unid
-cfec$spec[cfec$specn==144] = "ROCK.slope"
-
-cfec$specn[cfec$specn==138] = 168 # copper
-cfec$specn[cfec$specn==145] = 168 # yelloweye
-cfec$specn[cfec$specn==146] = 168 # Canary
-cfec$specn[cfec$specn==147] = 168 # quillback
-cfec$specn[cfec$specn==148] = 168 # tiger
-cfec$specn[cfec$specn==149] = 168 # China
-cfec$specn[cfec$specn==150] = 168 # rosethorn
-cfec$spec[cfec$specn==168] = "ROCK.demersal"
-
-cfec$specn[cfec$specn==154] = 169 # dusky
-cfec$specn[cfec$specn==155] = 169 # yellowtail
-cfec$specn[cfec$specn==156] = 169 # widow
-cfec$specn[cfec$specn==172] = 169 # dusky
-cfec$specn[cfec$specn==173] = 169 # dark
-cfec$specn[cfec$specn==142] = 169 # black
-cfec$specn[cfec$specn==167] = 169 # blue
-cfec$spec[cfec$specn==169] = "ROCK.pelagic"
-
-# Lump king crabs
-cfec$spec[cfec$spec%in%c("GKCR","BKCR","UKCR","RKCR")] = "KCRB" # king
-# Lump tanner crabs
-cfec$spec[cfec$spec%in%c("UTCR","BTCR")] = "TCRB" # tanner
-# Lump shrimp - spot prawn, pink shrimp, 
-cfec$spec[cfec$spec%in%c("SPRW","PSHP","USRM")] = "USRM" # shrimp
-# Correct REX and "REX "
-cfec$spec[cfec$spec%in%c("REX ")] = "REX"
-# drop records with no spp
-cfec = cfec[-which(cfec$spec==""),]
-```
-
-Next, we have to clean up a few of the gears. Drop a few gears that are sporadically used, and combine longline gears.
-
-```{r}
-# 8 = fishwheel, used at start / end
-# 13 = dipnet, only last 2 years
-# 99 = other, stopped reporting
-cfec = cfec[-which(cfec$gearn%in%c(8, 13, 99)),]
-# Merge 6 and 61 into longline
-cfec$gearn[which(cfec$gearn==61)] = 6
-```
-
-Next, we'll deal with processing the areas.
-We have multiple types of area codes.
-The 6-digit codes represent groundfish management areas, the 5-digit codes are state salmon/shellfish areas, and the 3-digit codes are the halibut areas.
-For the groundfish and salmon areas, there are sub-areas that are also reported (as the last 2-digits), so we'll lump sub-areas together by stripping off the last 2-digits.
-
-```{r}
-# Filter out sub-districts and group things at larger areas
-# Lump statistical areas by district -- groundfish
-cfec$stat6 = as.numeric(cfec$stat6) # as.numeric needed because of integer
-cfec$areaType = NA
-cfec$stat6_char = NA
-indx = which(cfec$stat6 > 99999 & cfec$stat6 < 999999)
-cfec$stat6_char[indx] = nchar(as.character(cfec$stat6[indx]))
-cfec$areaType[indx] = "groundfish"
-# This is just turning the last digit - sub region - into 0
-cfec$stat6[indx] = as.numeric(paste(substr(as.character(cfec$stat6[indx]),1,5), "0",sep=""))
-# Now do salmon / shellfish - This is just turning the last 2 digits - sub region - into 00
-indx = which(cfec$stat6 > 9999 & cfec$stat6 < 100000)
-cfec$stat6[indx] = as.numeric(paste(substr(as.character(cfec$stat6[indx]), 1, 3), "00", sep=""))
-cfec$areaType[indx] = "salmonShellfish"
-
-# Deal with the IPHC areas too. We can verify that nearly all spp caught in 3-digit areas are halibut,
-#sort(table(cfec$spec[which(cfec$stat6 < 1000)]))
-```
-
-Next, we'll aggregate the stat6 codes into larger geographic regions.
-The focus here is on Prince William Sound, but for contrast we'll include Southeast Alaska, Kodiak W / E, Cook Inlet, and Bristol Bay.
-
-```{r}
-# Add the areas -- 5 digit are salmon/shellfish, 6-digit are groundfish
-cfec$area = NA
-cfec$area[cfec$stat6==485900] = "PWS"
-cfec$area[cfec$stat6==475900] = "PWS"
-cfec$area[cfec$stat6==465900] = "PWS"
-cfec$area[cfec$stat6==455900] = "PWS"
-cfec$area[cfec$stat6==445900] = "PWS"
-cfec$area[cfec$stat6==495930] = "PWS"
-cfec$area[cfec$stat6==485930] = "PWS"
-cfec$area[cfec$stat6==475930] = "PWS"
-cfec$area[cfec$stat6==465930] = "PWS"
-cfec$area[cfec$stat6==455930] = "PWS"
-cfec$area[cfec$stat6==445930] = "PWS"
-cfec$area[cfec$stat6==486000] = "PWS"
-cfec$area[cfec$stat6==476000] = "PWS"
-cfec$area[cfec$stat6==466000] = "PWS"
-cfec$area[cfec$stat6==456000] = "PWS"
-cfec$area[cfec$stat6==446000] = "PWS"
-cfec$area[cfec$stat6==486030] = "PWS"
-cfec$area[cfec$stat6==476030] = "PWS"
-cfec$area[cfec$stat6==466030] = "PWS"
-cfec$area[cfec$stat6==456030] = "PWS"
-cfec$area[cfec$stat6==486100] = "PWS"
-cfec$area[cfec$stat6==476100] = "PWS"
-cfec$area[cfec$stat6==466100] = "PWS"
-cfec$area[which(cfec$stat6 < 20099 & cfec$stat6 >= 20000)] = "PWS"
-cfec$area[which(cfec$stat6 < 21299 & cfec$stat6 >= 21200)] = "PWS"
-cfec$area[which(cfec$stat6 < 22799 & cfec$stat6 >= 22700)] = "PWS"  
-cfec$area[which(cfec$stat6 < 22699 & cfec$stat6 >= 22600)] = "PWS"
-cfec$area[which(cfec$stat6 < 22899 & cfec$stat6 >= 22800)] = "PWS"
-cfec$area[which(cfec$stat6 < 22199 & cfec$stat6 >= 22100)] = "PWS"
-cfec$area[which(cfec$stat6 < 22299 & cfec$stat6 >= 22200)] = "PWS"  
-cfec$area[which(cfec$stat6 < 22599 & cfec$stat6 >= 22500)] = "PWS"
-cfec$area[which(cfec$stat6 < 22499 & cfec$stat6 >= 22400)] = "PWS"
-cfec$area[which(cfec$stat6 < 22399 & cfec$stat6 >= 22300)] = "PWS" 
-# Add halibut areas
-cfec$area[cfec$stat6 %in% c(242,232,220,230,240)] = "PWS"
-
-cfec$area[cfec$stat6==525900] = "CookInlet"
-cfec$area[cfec$stat6==535900] = "CookInlet"
-cfec$area[cfec$stat6==515930] = "CookInlet"
-cfec$area[cfec$stat6==525930] = "CookInlet"
-cfec$area[cfec$stat6==535930] = "CookInlet"
-cfec$area[cfec$stat6==516000] = "CookInlet"
-cfec$area[cfec$stat6==526000] = "CookInlet"
-cfec$area[cfec$stat6==496030] = "CookInlet"
-cfec$area[cfec$stat6==506030] = "CookInlet"
-cfec$area[cfec$stat6==516030] = "CookInlet"
-cfec$area[cfec$stat6==526030] = "CookInlet"
-cfec$area[cfec$stat6==506100] = "CookInlet"
-cfec$area[cfec$stat6==516100] = "CookInlet"
-cfec$area[which(cfec$stat6 < 24799 & cfec$stat6 >= 24400)] = "CookInlet" 
-cfec$area[which(cfec$stat6%in%c(261, 272))] = "CookInlet"
-
-cfec$area[which(cfec$stat6 < 11599 & cfec$stat6 >= 11100)] = "Southeast"
-cfec$area[which(cfec$stat6 < 10499 & cfec$stat6 >= 10100)] = "Southeast"
-cfec$area[which(cfec$stat6 < 11099 & cfec$stat6 >= 10500)] = "Southeast" 
-cfec$area[which(cfec$stat6 >= 305430 & cfec$stat6 <= 305530)] = "Southeast"
-cfec$area[which(cfec$stat6 >= 315430 & cfec$stat6 <= 315600)] = "Southeast"
-cfec$area[which(cfec$stat6 >= 325430 & cfec$stat6 <= 325700)] = "Southeast"
-cfec$area[which(cfec$stat6 >= 335430 & cfec$stat6 <= 335800)] = "Southeast"
-cfec$area[which(cfec$stat6 >= 345400 & cfec$stat6 <= 345830)] = "Southeast"
-cfec$area[which(cfec$stat6 >= 355400 & cfec$stat6 <= 355900)] = "Southeast"
-cfec$area[which(cfec$stat6 > 365400 & cfec$stat6 < 365830)] = "Southeast"
-# For IPHC see Fig. 20 here: http://www.iphc.int/publications/techrep/tech0049.pdf
-cfec$area[which(cfec$stat6%in%c(140, 141, 142, 143, 144, 150, 151, 152, 153, 160, 161, 162, 163, 170, 171, 172, 174, 174, 181, 182, 183, 184, 185))] = "Southeast"
-
-# Kodiak east area is defined by halibut areas 270/280, includes ADFG salmon / 
-# shellfish on E side of kodiak, and groundfish areas extending to NMFS Area 630
-# boundary
-cfec$area[cfec$stat6%in%c(270,280)] = "Kodiak.east"
-cfec$area[which(cfec$stat6 < 25899 & cfec$stat6 >= 25800)] = "Kodiak.east"
-cfec$area[which(cfec$stat6 < 25999 & cfec$stat6 >= 25900)] = "Kodiak.east"
-cfec$area[cfec$stat6==535630] = "Kodiak.east"
-cfec$area[cfec$stat6==525630] = "Kodiak.east"
-cfec$area[cfec$stat6==515630] = "Kodiak.east"
-cfec$area[cfec$stat6==505630] = "Kodiak.east"
-cfec$area[cfec$stat6==495630] = "Kodiak.east"
-cfec$area[cfec$stat6==535700] = "Kodiak.east"
-cfec$area[cfec$stat6==525700] = "Kodiak.east"
-cfec$area[cfec$stat6==515700] = "Kodiak.east"
-cfec$area[cfec$stat6==505700] = "Kodiak.east"
-cfec$area[cfec$stat6==495700] = "Kodiak.east"
-cfec$area[cfec$stat6==525730] = "Kodiak.east"
-cfec$area[cfec$stat6==515730] = "Kodiak.east"
-cfec$area[cfec$stat6==505730] = "Kodiak.east"
-cfec$area[cfec$stat6==495730] = "Kodiak.east"
-cfec$area[cfec$stat6==525800] = "Kodiak.east"
-cfec$area[cfec$stat6==515800] = "Kodiak.east"
-cfec$area[cfec$stat6==505800] = "Kodiak.east"
-cfec$area[cfec$stat6==495800] = "Kodiak.east"
-
-cfec$area[cfec$stat6%in%c(271,281)] = "Kodiak.west"
-cfec$area[cfec$stat6==535730] = "Kodiak.west"
-cfec$area[cfec$stat6==545730] = "Kodiak.west"
-cfec$area[cfec$stat6==525800] = "Kodiak.west"
-cfec$area[cfec$stat6==535800] = "Kodiak.west"
-cfec$area[cfec$stat6==545800] = "Kodiak.west"
-cfec$area[cfec$stat6==535830] = "Kodiak.west"
-cfec$area[cfec$stat6==525830] = "Kodiak.west"
-cfec$area[which(cfec$stat6 < 25699 & cfec$stat6 >= 25500)] = "Kodiak.west"
-cfec$area[which(cfec$stat6 < 25499 & cfec$stat6 >= 25300)] = "Kodiak.west"  
-cfec$area[which(cfec$stat6 < 26299 & cfec$stat6 >= 26000)] = "Kodiak.west"
-
-# Bristol Bay, defined as going west to the boundary between NMFS areas 512 / 516
-cfec$area[cfec$stat6%in%c(31600, 31700, 31800, 32000, 32100, 32200, 32400, 32500, 32600)] = "BristolBay"
-cfec$area[cfec$stat6%in%c(575730, 575800, 575830)] = "BristolBay"
-cfec$area[cfec$stat6%in%c(585700, 585730, 585800, 585830)] = "BristolBay"
-cfec$area[cfec$stat6%in%c(595630, 595700, 595730, 595800, 595830)] = "BristolBay"
-cfec$area[cfec$stat6%in%c(605600, 605630, 605700, 605730, 605800, 605830)] = "BristolBay"
-cfec$area[cfec$stat6%in%c(615600, 615630, 615700, 615730, 615800, 615830)] = "BristolBay"
-
-# Area to add - Alaska peninsula - Halibut areas 300-340?
+load("cfecCleaned.Rdata")
 ```
@@
 ```{r}
-T = table(cfec$adfg, cfec$stat6)
+if(useADFG == TRUE) T = table(cfec$adfg, cfec$stat6)
+if(useADFG == FALSE) T = table(cfec$cfecid, cfec$stat6)
 CC = clara(T, k = 8)
@@
 }
+# look at only species that make up > 1% in any single area
 kable(round(specArea[-which(apply(specArea,1,max) < 0.01),], 2))
@@
 
-## Exploratory data analysis
+## Identifying data problems with permits  
+Doing some additional sorting, we can see that there are a few unique ADFG numbers that have a huge number of permits associated with them. These are likely data issues, and should be removed.  
+```{r}
+head(group_by(cfec,adfg) %>% 
+       summarise(permits=length(unique(p_serial))) %>% 
+       arrange(-permits) %>% 
+       data.frame,20)
+```
+
+So we can delete these extreme permit holders. Per R. Brenner / J. Shriver, many of these are research vessels, out of state groundfish, PSC-related bycatch, etc.
+
+```{r}
+cfec = cfec[-which(cfec$adfg >= 98000),]
+cfec = cfec[-which(cfec$adfg == 0),]
+cfec = cfec[-which(is.na(cfec$adfg)),] # drop NAs
+```
+
+# Exploratory data analysis (Fall 2015)
 
@@
 cfecClean = cfecClean[-which(cfecClean$spec%in%c("UFLT","UGRN","SMLT","GRDR","EGLS")),]
+cfecClean = cfecClean[-which(cfecClean$gearn%in%c(0, 18, 22, 27, 37, 41, 62, 85)), ]
 ```
@@
 # Write the SSL related sites to a workspace
-sslStat6 = c(355630,
-365730, 375830, 345530, 335430, 335430, 335430,
-335430, 335430, 365800, 355700, 345530, 355630,
-355700, 335700, 335700, 355700, 445930, 495930, 
-476000, 485930, 495930, 325600, 495930, 476000,
-475930, 555530, 565600, 525700, 525700, 525830,
-525700, 515800, 515830, 535630, 365800, 575630,
-535630, 565630, 525830, 595500, 595430, 595500,
-595430, 635430, 635430, 605430, 615430, 635504,
-585500)
+#sslStat6 = c(355630,
+#365730, 375830, 345530, 335430, 335430, 335430,
+#335430, 335430, 365800, 355700, 345530, 355630,
+#355700, 335700, 335700, 355700, 445930, 495930, 
+#476000, 485930, 495930, 325600, 495930, 476000,
+#475930, 555530, 565600, 525700, 525700, 525830,
+#525700, 515800, 515830, 535630, 365800, 575630,
+#535630, 565630, 525830, 595500, 595430, 595500,
+#595430, 635430, 635430, 605430, 615430, 635504,
+#585500)
 
 # calculate areas in circle around square of interest
-sslStat6.square = c(sslStat6, sslStat6 +- 9970, sslStat6 +- 10000, 
-sslStat6 +- 10070, sslStat6 +- 70, sslStat6 +- 9930, sslStat6 +- 10000, 
-sslStat6 +- 10030, sslStat6 +- 30)
-sslStat6.label = rep(sslStat6,9)
-
-cfecSSL = cfecClean[cfecClean$stat6%in%sslStat6.square,]
-cfecSSL$label = sslStat6.label[match(cfecSSL$stat6, sslStat6.square)]
-save(cfecSSL, file="cfecSSL.Rdata")
-
+#sslStat6.square = c(sslStat6, sslStat6 +- 9970, sslStat6 +- 10000, 
+#sslStat6 +- 10070, sslStat6 +- 70, sslStat6 +- 9930, sslStat6 +- 10000, 
+#sslStat6 +- 10030, sslStat6 +- 30)
+#sslStat6.label = rep(sslStat6,9)
+
+#cfecSSL = cfecClean[cfecClean$stat6%in%sslStat6.square,]
+#cfecSSL$label = sslStat6.label[match(cfecSSL$stat6, sslStat6.square)]
+#save(cfecSSL, file="cfecSSL.Rdata")
+
+# Just for general interest, aggregate earnings by year and vessel
+# Similar to Kasperski and Holland
+#earnings_byVesselYear = aggregate(cfecClean$g_earn, by = list("Vessel"=cfecClean$adfg, "Year"= cfecClean$year), FUN = sum, na.rm=T)
+#median(earnings_byVesselYear$x) # 40884
+#hist(log(earnings_byVesselYear$x), 40, col="grey")
 ```
@@
 
-```{r runningMean}
+```{r runningMean,include=FALSE,echo=FALSE}
 totalPWSExp$nspecies = (totalPWSExp$nspecies + (totalPWSExp$nspecies + totalPWSExp$diffspec))/2
@@
 Start with simple linear model, no random effects  
-```{r linearModel}
+```{r linearModel,include=FALSE,echo=FALSE}
 lm.simple = lm(difflog~as.factor(year) + nspecies + ngear + narea + diffspec + diffgear + diffarea,data=totalPWSExp)
@@
 
-```{r simpleGam}
+```{r simpleGam,include=FALSE,echo=FALSE}
 gam.simple = gam(difflog~s(year) + nspecies + ngear + narea + diffspec + diffgear + diffarea,data=totalPWSExp)
@@
 This shows that > 80% of vessels never fish in more than 1 of our coarse areas (Prince William Sound, Cook Inlet, Bristol Bay, E/W Kodiak, etc), so we can start our analysis focusing on those.
-```{r tableVesselArea}
+```{r tableVesselArea,include=FALSE,echo=FALSE}
 tableVesselArea = table(cfecClean$adfg, cfecClean$area)
@@
 Next, we'll extract the vessels that are only assigned to one area, subsetting the entire dataset to just these vessels.
-```{r }
+```{r ,include=FALSE,echo=FALSE}
 singleAreaVessels = rownames(tableVesselArea)[which( apply(tableVesselArea, 1, sum) == 1)]
@@
 We can re-run the code above to total the response / predictors by calendar year. We'll add a column for area. 
-```{r aggregateSingleVessel}
+```{r aggregateSingleVessel,include=FALSE,echo=FALSE}
 totalSingles = group_by(cfecSingles, adfg, year) %>%
@@
 Gears, change in gears, areas, and change in areas fished also included.
-Plots show mean smooth year term, with SE."}
-gam.area = gam(difflog~area + s(year,by=as.factor(area)) + area*nspecies + area*diffspec + ngear + diffgear,data=totalSingles)
-summary(gam.area)
-par(mfrow = c(3,2),mgp=c(2,1,0),mai=c(0.5,0.5,0.3,0.1))
-plot(gam.area)
+Plots show mean smooth year term, with SE.",include=FALSE,echo=FALSE}
+#gam.area = gam(difflog~area + s(year,by=as.factor(area)) + area*nspecies + area*diffspec + ngear + diffgear,data=totalSingles)
+#summary(gam.area)
+#par(mfrow = c(3,2),mgp=c(2,1,0),mai=c(0.5,0.5,0.3,0.1))
+#plot(gam.area)
 ```
@@
 ******
-# More rigorous analysis
+\newpage
+
+# More rigorous analysis (Spring 2016)
 
@@
 cfecCleaned$isNA[is.na(cfecCleaned$area)] = 1
+if(useADFG==TRUE) {
   vessel.isna = table(cfecCleaned$adfg,cfecCleaned$isNA)
@@
 0)), ]  
-```
-
-```{r}
+}
+
+if(useADFG==FALSE) {
+  vessel.isna = table(cfecCleaned$cfecid,cfecCleaned$isNA)
+  cfecCleaned = cfecCleaned[cfecCleaned$cfecid %in% names(which(vessel.isna[, 2] ==
+0)), ] 
+}
+
+```
+
+```{r}
+if(useADFG == TRUE) {
 # Of this restricted dataset, 93% of vessels don't move between areas -
@@
 
+# also make a table of continuous vessels
+T = ceiling(table(cfecCleaned$adfg, cfecCleaned$year)/1.0e10)
+contVessels = names(which(apply(T,1,sum)==dim(T)[2]))
+
 # Now tabulate summaries of species, gears, fisheries, etc
@@
 
+contVess.df = aggregate(cbind(nspecies,ngear,nfishery,totalearnings) ~ adfg, data = totalVessels[totalVessels$adfg%in%contVessels, ], FUN = mean)
+contVess.df$cv = aggregate(totalearnings ~ adfg, data = totalVessels[totalVessels$adfg%in%contVessels, ], FUN = sd)$totalearnings / contVess.df$totalearnings
+
 # Lump larger groups
@@
 totalVessels$nfishery[totalVessels$nfishery > 4] = 4
+} # end useADFG
+
+```
+
+```{r}
+if(useADFG == FALSE) {
+# Of this restricted dataset, 93% of vessels don't move between areas -
+# ever. So the primary focus of our analysis can be on gears and species
+# landed. Just so the area switching won't matter, let's delete those
+# records that have multiple areas
+t = ceiling(table(cfecCleaned$cfecid, cfecCleaned$area)/1e+10)
+# delete vessels that fish in 2+ areas
+cfecCleaned = cfecCleaned[cfecCleaned$cfecid %in% names(which(apply(t, 1, sum) ==
+1)), ]
+
+# also make a table of continuous vessels
+T = ceiling(table(cfecCleaned$cfecid, cfecCleaned$year)/1.0e10)
+contVessels = names(which(apply(T,1,sum)==dim(T)[2]))
+
+# Now tabulate summaries of species, gears, fisheries, etc
+totalVessels = group_by(cfecCleaned, cfecid, year) %>%  
+  summarize(totalearnings = sum(g_earn),  
+  nspecies = length(unique(spec)),  
+  ngear = length(unique(gearn)),  
+  nfishery = length(unique(p_fshy)),
+  area = area[1])
+
+# calculate landings by species, year, vessel
+landBySpec = group_by(cfecCleaned, cfecid, year, spec) %>%  
+summarize(totalpounds = sum(g_pounds))  
+# calculate effective number of species by exp() for shannon
+# http://jonlefcheck.net/2012/10/23/diversity-as-effective-numbers/
+landBySpec = group_by(landBySpec, cfecid, year) %>%  
+   summarize(effectiveSpec = exp(diversity(totalpounds)))
+
+totalVessels$effectiveSpec = landBySpec$effectiveSpec
+
+# calculate landings by gear, year, vessel
+landByGear = group_by(cfecCleaned, cfecid, year, gearn) %>%  
+  summarize(totalpounds = sum(g_pounds))  
+landByGear = group_by(landByGear, cfecid, year) %>%  
+  summarize(effectiveGear = exp(diversity(totalpounds)))
+
+totalVessels$effectiveGear = landByGear$effectiveGear
+
+# calculate landings by fishery, year, vessel
+landByFshy = group_by(cfecCleaned, cfecid, year, p_fshy) %>%  
+summarize(totalpounds = sum(g_pounds))  
+landByFshy = group_by(landByFshy, cfecid, year) %>%  
+summarize(effectiveFshy = exp(diversity(totalpounds)))
+
+totalVessels$effectiveFshy = landByFshy$effectiveFshy
+
+contVess.df = aggregate(cbind(nspecies,ngear,nfishery,totalearnings) ~ cfecid, data = totalVessels[totalVessels$cfecid%in%contVessels, ], FUN = mean)
+contVess.df$cv = aggregate(totalearnings ~ cfecid, data = totalVessels[totalVessels$cfecid%in%contVessels, ], FUN = sd)$totalearnings / contVess.df$totalearnings
+
+# Lump larger groups
+totalVessels$nspecies[totalVessels$nspecies > 10] = 10
+totalVessels$ngear[totalVessels$ngear > 3] = 3
+totalVessels$nfishery[totalVessels$nfishery > 4] = 4
+} # end useADFG
+
 ```
@@
 ```{r}
+if(useADFG == TRUE) {
 # summarize landings by year/spec/
@@
 newdf <- dcast(lmlt, year + adfg ~ spec)
+}
+if(useADFG == FALSE) {
+# summarize landings by year/spec/
+totalLandings = group_by(cfecCleaned, cfecid, year, spec) %>%  
+  summarize(totallandings = sum(g_pounds))
+# cast to wide format with species across cols
+totalLandings = as.data.frame(totalLandings)
+#totalLandings = totalLandings[-which(is.na(totalLandings$adfg)),]
+lmlt <- melt(totalLandings, id.vars = c("cfecid", "year","spec"))
+newdf <- dcast(lmlt, year + cfecid ~ spec)
+}
 ```
@@
 ```{r}
+if(useADFG==TRUE) {
 dummyDF = expand.grid(adfg = unique(totalVessels$adfg), 
@@
          diffefffishery = c(NA, diff(effectiveFshy))) %>%  
-  filter(!is.na(difflog))
+  filter(!is.na(difflog)) %>%
+  filter(is.finite(difflog))
+}
+
+if(useADFG==FALSE) {
+dummyDF = expand.grid(cfecid = unique(totalVessels$cfecid), 
+  year = min(totalVessels$year) : max(totalVessels$year),
+  stringsAsFactors = F)
+
+# expand the data frame w/missing years
+totalVesselsExp = left_join(dummyDF, totalVessels) %>%
+  arrange(cfecid, year)
+totalVesselsExp = group_by(totalVesselsExp, cfecid) %>%
+  mutate(difflog = c(NA, diff(log(totalearnings))),  
+         diffspec = c(NA, diff(nspecies)),  
+         diffgear = c(NA, diff(ngear)),  
+         difffishery = c(NA, diff(nfishery)),  
+         diffeffspec = c(NA, diff(effectiveSpec)),  
+         diffeffgear = c(NA, diff(effectiveGear)),  
+         diffefffishery = c(NA, diff(effectiveFshy))) %>%  
+  filter(!is.na(difflog)) %>%
+  filter(is.finite(difflog))
+}
 ```
@@
 ```{r}
+if(useADFG == TRUE) {
 lme.vess.1trend = lmer(difflog ~ area + as.factor(year) + 
@@
 summary(lme.vess.1trend)
+}
+if(useADFG == FALSE) {
+lme.vess.1trend = lmer(difflog ~ area + as.factor(year) + 
+  nspecies + diffspec + ngear + diffgear + (1 | cfecid), 
+  data = totalVesselsExp)
+summary(lme.vess.1trend)
+}
 ```
@@
 ```{r}
+if(useADFG==TRUE) {
 lme.vess.int = lmer(difflog ~ area * as.factor(year) + nspecies + diffspec + ngear + diffgear + (1 | adfg), data = totalVesselsExp)
 summary(lme.vess.int)
+}
+if(useADFG==FALSE) {
+lme.vess.int = lmer(difflog ~ area * as.factor(year) + nspecies + diffspec + ngear + diffgear + (1 | cfecid), data = totalVesselsExp)
+summary(lme.vess.int)
+}
 ```
@@
  # Calculate proportion of vessels with 1 observation
+if(useADFG==TRUE) {
 round(length(table(totalVesselsExp$adfg))/length(which(table(totalVesselsExp$adfg) == 1)), 2)
+}
+if(useADFG==FALSE) {
+round(length(table(totalVesselsExp$cfecid))/length(which(table(totalVesselsExp$cfecid) == 1)), 2)
+}
 ```
@@
 ```{r}
+if(useADFG==TRUE) {
 activeVess = names(which(table(totalVesselsExp$adfg) >= 10))
@@
 summary(lme.vess.int10)
+}
+
+if(useADFG==FALSE) {
+activeVess = names(which(table(totalVesselsExp$cfecid) >= 10))
+lme.vess.int10 = lmer(difflog ~ area + as.factor(year) + nspecies + diffspec +
+    ngear + diffgear + (1 | cfecid), data = totalVesselsExp[totalVesselsExp$cfecid %in%
+activeVess, ])
+summary(lme.vess.int10)
+}
 ```
@@
 ```{r, fig.pos="placeHere",fig.cap="Individual performance (log total revenue in red) and species landed (raw = black, effective = blue)."}
+if(useADFG==TRUE) {
 activeVess = names(which(table(totalVesselsExp$adfg) >= 10))
@@
 }
+}
+
+if(useADFG==FALSE) {
+activeVess = names(which(table(totalVesselsExp$cfecid) >= 10))
+activeVess.id = names(rev(sort(table(totalVesselsExp[totalVesselsExp$cfecid %in% activeVess, ]$cfecid))))
+
+par(mfrow = c(5, 5), mai = c(0.01, 0.01, 0.01, 0.01))
+for (i in 1:25) {
+  df = as.data.frame(totalVesselsExp[totalVesselsExp$cfecid == activeVess.id[i], ])
+  plot(df$nspecies, type = "l")
+  lines(rescale(log(df$totalearnings), range(df$nspecies)), col = "red")
+  lines(rescale(log(df$effectiveSpec), range(df$nspecies)), col = "blue")
+}
+}
 ```
@@
 ```{r}
+if(useADFG==TRUE) {
 lme.vess.int.area = lmer(difflog ~ area * as.factor(year) + nspecies + area * diffspec + ngear + diffgear + (1 | adfg), data = totalVesselsExp)
@@
 lme.vess.int.timearea = lmer(difflog ~ area * as.factor(year) + nspecies + area * as.factor(year) * diffspec + ngear + diffgear + (1 | adfg), data = totalVesselsExp)
+}
+
+if(useADFG==FALSE) {
+lme.vess.int.area = lmer(difflog ~ area * as.factor(year) + nspecies + area * diffspec + ngear + diffgear + (1 | cfecid), data = totalVesselsExp)
+lme.vess.int.time = lmer(difflog ~ area * as.factor(year) + nspecies + as.factor(year) * diffspec + ngear + diffgear + (1 | cfecid), data = totalVesselsExp)
+lme.vess.int.timearea = lmer(difflog ~ area * as.factor(year) + nspecies + area * as.factor(year) * diffspec + ngear + diffgear + (1 | cfecid), data = totalVesselsExp)
+}
 ```
@@
 ```{r}
-lme.vess.int.area.effective = lmer(difflog ~ area * as.factor(year) + effectiveSpec + area * diffspec + effectiveGear + diffeffgear + (1 | adfg), data = totalVesselsExp)
-```
+if(useADFG==TRUE) {
+lme.vess.int.area.effective = lmer(difflog ~ area * as.factor(year) + effectiveSpec + area * diffeffspec + effectiveGear + diffeffgear + (1 | adfg), data = totalVesselsExp)
+}
+if(useADFG==FALSE) {
+lme.vess.int.area.effective = lmer(difflog ~ area * as.factor(year) + effectiveSpec + area * diffeffspec + effectiveGear + diffeffgear + (1 | cfecid), data = totalVesselsExp)
+}
+```
+
+## Relationship between CV and metrics of diversity.
+
+Similar to Kasperski and Holland (2013), we can reduce all of the individual vessels to single metrics of revenue and variability.
+Following their analysis, we can take the records of all vessels that fished continuously, 1985-2014.
+For each, we can calculate the mean and CV of revenue, and as covariates the average number of gears, species landed, and gears fished.
+(This data frame actually created above as contVess.df).
+[Note: these vessels represent all vessels in the dataset, not those that are constrained to fish in our designated spatial zones.]
+
+```{r}
+par(mfrow = c(3,2), mai=c(0.4,0.4,0.05,0.05))
+hist(log(contVess.df$totalearnings), 40, main="Ln(avg revenue)", col="grey")
+hist(contVess.df$cv, 40, main="CV(revenue)", col="grey")
+hist(contVess.df$ngear, 20, main="Mean gears", col="grey")
+hist(contVess.df$nspecies, 20, main="Mean species", col="grey")
+hist(contVess.df$nfishery, 20, main="Mean fisheries", col="grey")
+```
+
+### Do increases in diversity translate to lower variability in revenue?
+
+We can treat CV as the response, and species / gears / fisheries as the predictors. Stepwise model selection suggests that increasing gears 
+```{r}
+gmod = glm(log(contVess.df$cv) ~ log(nspecies) + log(ngear) + log(nfishery), data = contVess.df)
+stepAIC(gmod, direction = "both")
+```

### 2016-06-14 data/process_raw_data/Vessel_Characteristics.Rmd
Add raw data processing folder to main data folder

---
title: "Vessel Characteristics"
author: "Anne Beaudreau & Jordan Watson"
date: "March 25, 2016"
output: pdf_document
---

We want to add vessel information to the catch data.These characteristics lie in annual files pulled from the cfec dataset via individual queries. These data are in the "Vessel" folder within the Commercial Catch Portfolios folder in the Google Doc.

```{r setup, include=FALSE}

setwd("~/Desktop/NCEAS")
years<-c(1985:2014)
#  Create an empty data.frame in which to store the data
vessel=data.frame()

for (i in (1:length(years))){
  #  All the data files have the same naming structure so we can paste the names via a loop
  xx=read.csv(paste("Data/Vessels/Vessels",years[i],".csv",sep=""))
  #  The zip codes are funky and cause an error if we don't specify them. 
  xx$Zip.Code<-as.factor(xx$Zip.Code)
  #  Bind each successive year into the same data frame. 
  vessel=bind_rows(vessel,xx) %>% data.frame
}
rm(i,years,xx)
#  The vessel data.frame will be joined with the cfec data using the year and adfg columns. The vessel data have different names though so rename the columns to make them consistent with the cfec dataset. I also rename the "Length" field to "VLength" so as to avoid any issues associated with using a column name that is so similar to the R function length().

vessel <- rename(vessel,year=Year,adfg=ADFG.Number,VLength=Length)
```

A single vessel may have multiple owners in one year. When this happens, a single adfg number may have multiple different entries where the vessel data are duplicated but the owner data are unique. The problem is that this creates duplicate records for the vessel components that we are matching to the cfec data. Filter out the duplicated vessel data by unique combinations of "year" and "adfg" and then select only the vessel characterstic columns.

```{r}
vessel<-distinct(vessel,year,adfg) %>% dplyr::select(year,adfg,Year.Built,VLength,Gross.Tons,Net.Tons,Horse.Power,Hold.Tank.Capacity,Live.Tank.Capacity,Fuel.Capacity)

save.image("Data/vessel.RData")
```

### 2016-06-15 portfolio/analysis/4b-Bayesian_portfolio_modeling.Rmd
fixes
I modified the code to allow the 4 arrays of ints to still be an array
with 1 random effect. Previously, a dimension error was being thrown by
STAN.

--- 2016-06-14 analysis_portfolio/4b-Bayesian_portfolio_modeling.Rmd
@@
 N = dim(testData)[1]
-b1_cov = cbind(SPEC_DIV, DAYS) # fixed effect covariate on slope
-sigma1_cov = cbind(SPEC_DIV, DAYS) # fixed effect covariate on slope
+b1_cov = cbind(SPEC_DIV) # fixed effect covariate on slope
+sigma1_cov = cbind(SPEC_DIV) # fixed effect covariate on slope
 b1_cov_re = cbind(SPEC_DIV)
@@
 
-b0_re = cbind(P_HOLDER, PERMIT, PERMIT_AREA, PERMIT_SPEC)
+b0_re = cbind(PERMIT) #cbind(P_HOLDER, PERMIT, PERMIT_AREA, PERMIT_SPEC)
 b0_re_levels = apply(b0_re,2,max)
@@
 
-b1_re = cbind(PERMIT, PERMIT_AREA, PERMIT_SPEC)
+b1_re = cbind(PERMIT)#cbind(P_HOLDER, PERMIT, PERMIT_AREA, PERMIT_SPEC)
 b1_re_levels = apply(b1_re,2,max)
@@
 
-sigma0_re = cbind(P_HOLDER, PERMIT, PERMIT_AREA, PERMIT_SPEC)
-sigma0_re_levels = apply(b0_re,2,max)
+sigma0_re = cbind(PERMIT)
+sigma0_re_levels = apply(sigma0_re,2,max)
 sigma0_re_n = dim(sigma0_re)[2]
 
-sigma1_re = cbind(PERMIT, PERMIT_AREA, PERMIT_SPEC)
+sigma1_re = cbind(PERMIT)
 sigma1_re_levels = apply(sigma1_re,2,max)
@@
 ### Center all of the covariate matrices
-b1_cov = scale(b1_cov, scale=FALSE)
-sigma1_cov = scale(sigma1_cov, scale=FALSE)
-b1_cov_re = scale(b1_cov_re, scale=FALSE)
-sigma1_cov_re = scale(sigma1_cov_re, scale=FALSE)
+#b1_cov = scale(b1_cov, scale=FALSE)
+#sigma1_cov = scale(sigma1_cov, scale=FALSE)
+#b1_cov_re = scale(b1_cov_re, scale=FALSE)
+#sigma1_cov_re = scale(sigma1_cov_re, scale=FALSE)
 ```
@@
   "K_mean_re"=K_mean_re, "K_sig_re"=K_sig_re,"b1_cov_re"=b1_cov_re, "sigma1_cov_re"=sigma1_cov_re,
-  "b0_re"=b0_re, "b0_re_levels"=b0_re_levels, "b0_re_n"=b0_re_n,
-  "b1_re"=b1_re, "b1_re_levels"=b1_re_levels, "b1_re_n"=b1_re_n,
-  "sigma0_re"=sigma0_re, "sigma0_re_levels"=sigma0_re_levels, "sigma0_re_n"=sigma0_re_n,
-  "sigma1_re"=sigma1_re, "sigma1_re_levels"=sigma1_re_levels, "sigma1_re_n"=sigma1_re_n, "K_sig"=K_sig, "y"=y)
+  "b0_re"=b0_re, "b0_re_levels"=c(b0_re_levels,0), "b0_re_n"=b0_re_n,
+  "b1_re"=b1_re, "b1_re_levels"=c(b1_re_levels,0), "b1_re_n"=b1_re_n,
+  "sigma0_re"=sigma0_re, "sigma0_re_levels"=c(sigma0_re_levels,0), "sigma0_re_n"=sigma0_re_n,
+  "sigma1_re"=sigma1_re, "sigma1_re_levels"=c(sigma1_re_levels,0), "sigma1_re_n"=sigma1_re_n, "K_sig"=K_sig, "y"=y)
 # parameters to monitor
@@
 stanMod = stan(file = 'portfolio_mixedModel.stan',data = list_data, 
-  verbose = TRUE, chains = 4, thin = 1, warmup = 500, iter = 1000, 
+  verbose = TRUE, chains = 1, thin = 1, warmup = 20, iter = 50, 
   pars = list_pars)

### 2016-06-16 portfolio/analysis/4a-createData_linear_modeling.Rmd
Clean up data creation script
Note that since we have different raw data
a couple of these things might not work

@@
 library(knitr)
-library(mgcv)
+# library(mgcv)
 library(ggplot2)
-library(randomForest)
-library(MASS)
-library(zoo)
-library(MatchIt)
-library(metafolio)
-library(randomForest)
+# library(randomForest)
+# library(MASS)
+# library(zoo)
+# library(MatchIt)
+# library(metafolio)
+# library(randomForest)
 library(date)
 
+devtools::load_all("../../pfxr/")
+
 #load("cfecCleaned2.Rdata")
-cfec <- feather::read_feather("data/cfec.feather")
+cfec <- feather::read_feather("../data-generated/cfec.feather")
 
@@
   substr(cfec$p_fshy[idx], 3, length(cfec$p_fshy[idx])), sep="")
-
-region<-read.csv("areas/regions.csv")
-region<-region[,c(1:2)] #don’t include “assigned by” column
-cfec<-left_join(cfec,region)
-#cfec$region <- replace(cfec$region, is.na(cfec$region), "Unassigned")
 ```
@@
 
-```{r, echo=FALSE}
-deflationTable = data.frame(year = 1985:2014, defl = c(0.57236, 0.58393, 0.59879, 0.61974,  
-0.64388, 0.66774, 0.68993, 0.70564, 0.72244, 0.73781,  
-0.75321, 0.76695, 0.78009, 0.78855, 0.80061,  
-0.81883, 0.83753, 0.85038, 0.86729, 0.89114,  
-0.91981, 0.94812, 0.97334, 0.9925, 1,  
-1.01217, 1.03307, 1.05213, 1.06926, 1.08682))
-
-# Adjust price and g_earn for inflation
-cfec$year = as.numeric(cfec$year)
-cfec$day = substr(cfec$landdate, 6, 10)
-cfec$g_price = cfec$g_price/deflationTable$defl[cfec$year-1984]
-cfec$g_earn = cfec$g_earn/deflationTable$defl[cfec$year-1984]
-```
-
-### Function definiton  
+### Function definition  
 We need to define a few functions for calculating diversity.
@@
 ```{r, warning = FALSE, message=FALSE, echo=FALSE}
-
-simp.div = function(x) {
-  1/sum((x/sum(x))^2)
-}
-
-#library(Rcpp)
-#cppFunction('double simpCpp(NumericVector x) {
-#  double y = 1/sum((x/sum(x))^2);
-#  return y;
-#}')
 
@@
 cfec.1boat = group_by(cfecnew, p_holder, year) %>% 
-  summarize(nBoat = length(unique(cadfg))) %>%
-  filter(nBoat==1) %>% 
-  left_join(cfecnew)
+  mutate(nBoat = length(unique(cadfg))) %>%
+  filter(nBoat==1)
 
@@
 cfec.1boat = group_by(cfec.1boat, cadfg, year) %>% 
-  summarize(nPerson = length(unique(p_holder))) %>%
-  filter((nPerson==1 & cadfg %in% c(27960,27961) == FALSE) | cadfg %in% c(27960,27961)) %>% 
-  left_join(cfec.1boat)
+  mutate(nPerson = length(unique(p_holder))) %>%
+  filter((nPerson==1 & cadfg %in% c(27960,27961) == FALSE) | cadfg %in% c(27960,27961))
 
@@
 cfecAnnual = group_by(cfecnew, p_holder, year) %>%
-  summarize(permit = p_fshy[1], boat = cadfg2[1], 
+  summarize(permit = p_fshy[1], boat = cadfg[1], # TODO cadfg2?
     days = diff(range(day) + 1), 
@@
     revenue = sum(g_earn),
-    length = mean(VLength, na.rm=T))
+    length = mean(vlength, na.rm=T))
 
-cfecAnnual = group_by(cfecAnnual, p_holder, year) %>%
-  filter(npermit == 1)
+cfecAnnual = filter(cfecAnnual, npermit == 1)
 
@@
 
-saveRDS(cfecAnnual.salmon, file="salmon_linearModeling_complete.rds")
-saveRDS(cfecAnnual.nonsalmon, file="nonsalmon_linearModeling_complete.rds")
+saveRDS(cfecAnnual.salmon, file="../data-generated/salmon_linearModeling_complete.rds")
+saveRDS(cfecAnnual.nonsalmon, file="../data-generated/nonsalmon_linearModeling_complete.rds")
 

### 2016-06-16 portfolio/analysis/2-try-model.R
Add in progress and messy code

@@ -23,7 +23,6 @@ library(nlme)
 m_gls1 <- gls(u ~ s, weights = varExp(form = ~s))
 
 library(TMB)
-setwd("analysis")
 compile("revenue1.cpp")
 dyn.load(dynlib("revenue1"))
 
@@ -188,7 +187,165 @@ abline(a = 0, b = 1)
 plot(b0_eta, filter(r, parameter == "b0_k")$Estimate)
 abline(a = 0, b = 1)
 ##################################################
+compile("revenue6.cpp")
+dyn.load(dynlib("revenue6"))
 
+mm <- model.matrix(~ s)
+obj <- MakeADFun(
+  data = list(x_ij = mm, y_i = u, k_i = k_i, n_k = n_k, 
+    b1_cov_re_i = s, sigma1_cov_re_i = s),
+  parameters = list(b_j = c(0, 0), sigma_j = c(0, 0), log_b0_sigma = -1, b0_k = rep(0, n_k),
+    b1_k = rep(0, n_k), log_b1_sigma = -1, sigma0_k = rep(0, n_k), sigma1_k = rep(0, n_k),
+    log_sigma0_sigma = -1, log_sigma1_sigma = -1),
+  random = c("b0_k", "b_j", "b1_k", "sigma0_k", "sigma1_k"),
+  DLL = "revenue6")
+
+opt <- nlminb( start=obj$par, objective=obj$fn, gradient=obj$gr, control=list("trace"=1) )
+obj$gr(opt$par)
+sd_report <- sdreport(obj)
+# rep <- obj$report()
+
+(r <- summary(sd_report, "random"))
+(f <- summary(sd_report, "fixed"))
+parameters <- row.names(r)
+r <- as.data.frame(r)
+r$parameter <- parameters
+
+ggplot(r, aes(Estimate, parameter)) + geom_point()
+
+par(mfrow = c(2, 2))
+plot(sigma1_eta, filter(r, parameter == "sigma1_k")$Estimate)
+abline(a = 0, b = 1)
+plot(sigma0_eta, filter(r, parameter == "sigma0_k")$Estimate)
+abline(a = 0, b = 1)
+
+plot(b1_eta, filter(r, parameter == "b1_k")$Estimate)
+abline(a = 0, b = 1)
+plot(b0_eta, filter(r, parameter == "b0_k")$Estimate)
+abline(a = 0, b = 1)
+##################################################
+##################################################
+
[…]

### 2016-06-17 portfolio/analysis/4a-createData_linear_modeling.Rmd

@@
 
+## Do some exploration by permit types + diversity
+
+For example, when someone is diversifying with a herring, halibut, or crab permit, what does that diversification look like?
+
+```{r, echo=FALSE}
+cfecnew$taxa = substr(cfecnew$p_fshy, 1, 1)
+specPermit = group_by(cfecnew, taxa, specn) %>% 
+  summarize(percent.revenue = sum(g_earn))
+specPermit.tot = group_by(specPermit, taxa) %>% 
+  summarize(tot = sum(percent.revenue))
+specPermit = left_join(specPermit, specPermit.tot)
+specPermit$percent.revenue = specPermit$percent.revenue/specPermit$tot
+specPermit.df = group_by(specPermit, taxa, specn) %>% 
+  filter(percent.revenue > 0.02) %>% 
+  arrange(taxa) %>% 
+  as.data.frame
+specPermit.df$spec = cfec$spec[match(specPermit.df$specn, cfec$specn)]
+kable(specPermit.df)
+```
+
 ## Code for looking at permit 

### 2016-06-17 portfolio/analysis/4a-createData_linear_modeling.Rmd
cleaned up code, added more filtering

--- 2016-06-14 portfolio/analysis/4a-createData_linear_modeling.Rmd
@@
 library(knitr)
-library(mgcv)
 library(ggplot2)
-library(randomForest)
-library(MASS)
-library(zoo)
-library(MatchIt)
-library(metafolio)
-library(randomForest)
 library(date)
@@
 ```{r, warning = FALSE, message=FALSE, echo=FALSE}
-
 simp.div = function(x) {
@@
 }
-
-#library(Rcpp)
-#cppFunction('double simpCpp(NumericVector x) {
-#  double y = 1/sum((x/sum(x))^2);
-#  return y;
-#}')
-
-# Note that the first line of this function uses group_by_ because 
-# we need to pass in a string for 'variable'
-effectiveDiversity_by_personYear = function(dataFrame, variable) {
-  effDiv = group_by_(dataFrame, variable, "year", "p_holder") %>%
-  summarize(totRev = sum(g_earn, na.rm=T),
-  totN = n(),
-  totWeight = sum(g_pounds, na.rm=T),
-    days = length(unique(day))) %>% 
-  as.data.frame() %>%
-  filter(is.finite(totRev) & is.finite(totN) & is.finite(totWeight)) %>%   
-  as.data.frame() %>%
-  group_by(p_holder, year) %>% 
-  summarize(eff.freq = simp.div(totN),
-  eff.earn = simp.div(totRev),
-  eff.lbs = simp.div(totWeight),
-    totIndRev = sum(totRev),
-    nCalDays = sum(days)) %>% 
-  filter(!is.na(eff.freq) & !is.na(eff.earn) & !is.na(eff.lbs)) %>%
-  as.data.frame()
-  return(effDiv)
-}
-
-# Note that the first line of this function uses group_by_ because 
-# we need to pass in a string for 'variable'
-effectiveDiversity_by_personYear2 = function(dataFrame, variable1, variable2) {
-  effDiv = group_by_(dataFrame, variable1, variable2, "year", "p_holder") %>%
-  summarize(totRev = sum(g_earn, na.rm=T),
-  totN = n(),
-  totWeight = sum(g_pounds, na.rm=T),
-    days = length(unique(day))) %>% 
-  as.data.frame() %>%
-  filter(is.finite(totRev) & is.finite(totN) & is.finite(totWeight)) %>%   
-  as.data.frame() %>% 
-  group_by_("p_holder", "year", variable2) %>% 
-  summarize(eff.freq = simp.div(totN),
-  eff.earn = simp.div(totRev),
-  eff.lbs = simp.div(totWeight),
-    totIndRev = sum(totRev),
-    nCalDays = sum(days)) %>% 
-  filter(!is.na(eff.freq) & !is.na(eff.earn) & !is.na(eff.lbs)) %>%
-  as.data.frame()
-  return(effDiv)
-}
-
-effectiveDiversity_by_Year = function(dataFrame, variable) {
-  effDiv = group_by_(dataFrame, variable, "year") %>%
-  summarize(totRev = sum(g_earn, na.rm=T),
-  totN = n(),
-  totWeight = sum(g_pounds, na.rm=T),
-    days = length(unique(day))) %>% 
-  as.data.frame() %>%
-  filter(is.finite(totRev) & is.finite(totN) & is.finite(totWeight)) %>%   
-  as.data.frame() %>%
-  group_by(year) %>% 
-  summarize(eff.freq = simp.div(totN),
-  eff.earn = simp.div(totRev),
-  eff.lbs = simp.div(totWeight),
-    totIndRev = sum(totRev),
-    nCalDays = sum(days)) %>% 
-  filter(!is.na(eff.freq) & !is.na(eff.earn) & !is.na(eff.lbs)) %>%
-  as.data.frame()
-  return(effDiv)
-}
-
-# This function takes the output of the one above and 
-# summmarizes diversity metrics
-summarize_effectiveDiversity = function(dataFrame) {
-  effDiv = group_by(dataFrame, year) %>% 
-  summarize(mean.freq = mean(eff.freq),
-  sd.freq = sd(eff.freq),
-  mean.earn = mean(eff.earn),
-  sd.earn = sd(eff.earn),
-  mean.lbs = mean(eff.lbs),
-  sd.lbs = sd(eff.lbs)) %>% 
-  as.data.frame()  
-  return(effDiv)
-}
-
-# This function takes the output of the one above and 
-# summmarizes diversity metrics
-summarize_effectiveDiversity2 = function(dataFrame, variable2) {
-  effDiv = group_by_(dataFrame, "year", variable2) %>% 
-  summarize(mean.freq = mean(eff.freq),
-  sd.freq = sd(eff.freq),
-  mean.earn = mean(eff.earn),
-  sd.earn = sd(eff.earn),
-  mean.lbs = mean(eff.lbs),
-  sd.lbs = sd(eff.lbs)) %>% 
-  as.data.frame()  
-  return(effDiv)
-}
 ```
 
-## Filtering out unique vessel - person combinations  
-Filter out people who only use 1 vessel / year , and similarly filter out boats with more than 1 person / year.
-Some people might fish on multiple vessels within a year, and vessels might be used by multiple people within a year.
+## Filtering out unique vessel - person combinations. Drops rows from 17 to 13 million.
 
@@
 
-# Deal with setnet fisheries, where lots of p_holder
-# share same codes
-cfec.1boat = group_by(cfec.1boat, cadfg, year) %>% 
-  summarize(nPerson = length(unique(p_holder))) %>%
-  filter((nPerson==1 & cadfg %in% c(27960,27961) == FALSE) | cadfg %in% c(27960,27961)) %>% 
-  left_join(cfec.1boat)
-
 cfecnew = cfec.1boat
 dim(cfecnew)[1]
+```
+
+## Filtering out people with 10+ years of fishing. Drops rows from 13.7 to 10 million. 
+```{r}
+cfecnew = group_by(cfecnew, p_holder) %>% 
+  mutate(nyr = length(unique(year))) %>% 
+  filter(nyr >= 10)
+```
+
+## Identify most popular strategies by person-year combination
+
+The top 120 permits cuts the rows down to 8.6 million. The top 50 gets it down to 7.8 million.
+
+```{r}
+topstrategies = group_by(cfecnew, p_holder, year) %>% 
+  summarize(strategy = paste(sort(unique(p_fshy)), collapse = " ")) %>%
+  as.data.frame()
+cfecnew = left_join(cfecnew, topstrategies) # adds strategy column
+
+CUTOFF = 50
+cfecnew = cfecnew[cfecnew$strategy%in%names(rev(sort(table(topstrategies$strategy)))[1:CUTOFF]), ]
 ```
@@
   summarize(specDiv = simp.div(totRev.spec)))
+
+# random slope by p_holder
+# Random intercepts in people, random intercepts in strategies, random slopes in strategies
+# Group level predictors on intercepts 
+# Indicator variable for EVOS years. Or whether
 

### 2016-06-17 portfolio/analysis/4a-createData_linear_modeling.Rmd
updated code to require 10% filter
a permit needs to account for 10% of a person’s earnings need to be
included

@@
 
+###For variability, any person in a year will need to have 10% or more of their revenue to come from each permit (otherwise doesn't count as a strategy)
+
 The top 120 permits cuts the rows down to 8.6 million. The top 50 gets it down to 7.8 million.
@@
 ```{r}
+permit_rev = group_by(cfecnew, p_holder, year, p_fshy) %>% 
+  summarize(permit.rev = sum(g_earn)) %>% 
+  as.data.frame() %>% 
+  group_by(p_holder, year) %>% 
+  mutate(tot.rev = sum(permit.rev))
+permit_rev$permit.rev = permit_rev$permit.rev/permit_rev$tot.rev # fraction of total
+permit_rev = permit_rev[permit_rev$permit.rev < 0.1,]
+
+# Delete these minimal earnings
+cfecnew = cfecnew[-which(paste(cfecnew$p_holder, cfecnew$year, cfecnew$p_fshy) %in% paste(permit_rev$p_holder, permit_rev$year, permit_rev$p_fshy))]
+
 topstrategies = group_by(cfecnew, p_holder, year) %>% 
@@
     revenue = sum(g_earn),
-    length = mean(VLength, na.rm=T))
+    length = mean(VLength, na.rm=T),
+    strategy = strategy[1])
 
@@
 cfecAnnual$geardiv.prev = cfecAnnual$gearDiv[idx]
+cfecAnnual$strategy.prev = cfecAnnual$strategy[idx]
 #cfecAnnual$p2.prev = cfecAnnual$p2[idx]
@@
 cfecAnnual$specDiv.change = cfecAnnual$specDiv - cfecAnnual$specdiv.prev
+# idx = which(cfecAnnual$strategy != cfecAnnual$strategy.prev)
 
@@
 
-cfecAnnual.salmon = cfecAnnual.salmon[is.na(cfecAnnual.salmon$specDiv.change)==F, ]
-cfecAnnual.nonsalmon = cfecAnnual.nonsalmon[is.na(cfecAnnual.nonsalmon$specDiv.change)==F, ]
+#cfecAnnual.salmon = cfecAnnual.salmon[is.na(cfecAnnual.salmon$specDiv.change)==F, ]
+#cfecAnnual.nonsalmon = cfecAnnual.nonsalmon[is.na(cfecAnnual.nonsalmon$specDiv.change)==F, ]
 

### 2016-06-19 portfolio/analysis/6a-createData_linear_modeling_diff.Rmd
upload differenced model

--- 2016-06-17 portfolio/analysis/4a-createData_linear_modeling.Rmd
@@
 ---
+
+## Modifications from un-differenced filtering:  
+### 1. No threshold for people to be included (e.g. 10+ years)
+### 2. Inclusion of everyone, regardless of permits fished
+### 3. Excluding observations that we don't have consecutive years for
+### 4. Some permits lumped -- especially those based on vessel sizes
 
@@
 
-## Filtering out people with 10+ years of fishing. Drops rows from 13.7 to 10 million. 
+## Aggregate some permits, especially some of the longline/trawl ones based on vessel size, for which reporting is also variable
+
 ```{r}
-cfecnew = group_by(cfecnew, p_holder) %>% 
-  mutate(nyr = length(unique(year))) %>% 
-  filter(nyr >= 10)
+cfecnew$p_fshy[cfecnew$p_fshy=="B61B"] = "B06B"
+cfecnew$p_fshy[cfecnew$p_fshy=="C61B"] = "C06B"
+cfecnew$p_fshy[cfecnew$p_fshy=="C91B"] = "C09B"
+cfecnew$p_fshy[cfecnew$p_fshy=="M61B"] = "M06B"
+cfecnew$p_fshy[cfecnew$p_fshy=="E09B"] = "E91B"
+cfecnew$p_fshy[cfecnew$p_fshy=="I06B"] = "I61B"
+cfecnew$p_fshy[cfecnew$p_fshy=="Y06A"] = "Y61A"
+cfecnew$p_fshy[cfecnew$p_fshy=="O06B"] = "O61B"
+cfecnew$p_fshy[cfecnew$p_fshy=="O09B"] = "O91B"
+
+cfecnew$p_fshy[cfecnew$p_fshy%in%c("M7FB","M7GB", "M7HB","M7IB")] = "M7IB"
+cfecnew$p_fshy[cfecnew$p_fshy%in%c("M7FG","M7GG", "M7HG","M7IG")] = "M7IG"
+
+# crab and shrimp vessels 
+cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="D91")] = paste("D09", substr(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="D91")],4,length(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="D91")])), sep="")
+cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="T91")] = paste("T09", substr(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="T91")],4,length(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="T91")])), sep="")
+cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="P91")] = paste("P09", substr(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="P91")],4,length(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="P91")])), sep="")
+
+cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="M91")] = paste("M09", substr(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="M91")],4,length(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="M91")])), sep="")
+cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="M61")] = paste("M06", substr(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="M61")],4,length(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="M61")])), sep="")
 ```
@@
 
-CUTOFF = 50
-cfecnew = cfecnew[cfecnew$strategy%in%names(rev(sort(table(topstrategies$strategy)))[1:CUTOFF]), ]
+#CUTOFF = 100
+#cfecnew = cfecnew[cfecnew$strategy%in%names(rev(sort(table(topstrategies$strategy)))[1:CUTOFF]), ]
 ```
@@
     length = mean(VLength, na.rm=T),
-    strategy = strategy[1])
-
-cfecAnnual = group_by(cfecAnnual, p_holder, year) %>%
-  filter(npermit == 1)
+    strategy = strategy[1],
+    weight = sum(g_pounds))
 
@@
 
-# random slope by p_holder
-# Random intercepts in people, random intercepts in strategies, random slopes in strategies
-# Group level predictors on intercepts 
-# Indicator variable for EVOS years. Or whether
-
 # also calculate weighted simpson diversity of gears and species
@@
   summarize(gearDiv = simp.div(totRev.gear)))
-
-#temp = group_by(cfecnew, p_holder, year, specn) %>% 
-#  summarize(totRev.spec = sum(g_earn), 
-#    permit = p_fshy[1])
-#temp = left_join(temp, group_by(cfecnew, p_holder, year) %>% 
-#    summarize(totRev = sum(g_earn)))
-#temp$totRev.spec = temp$totRev.spec / temp$totRev
-
-#mkt = group_by(cfecnew, p_fshy, year, specn) %>% 
-#  summarize(totRev.mkt.spec = sum(g_earn)) 
-#mkt = left_join(mkt, group_by(mkt, p_fshy, year) %>% 
-#  summarize(totRev.mkt = sum(totRev.mkt.spec))) 
-#mkt$totRev.mkt.spec = mkt$totRev.mkt.spec / mkt$totRev.mkt
-#names(mkt)[1] = "permit"
-
-# combine individual and market datasets
-#temp = left_join(temp, mkt)
-
-# Calculate squared proportions
-#temp$p2 = (temp$totRev.spec - temp$totRev.mkt.spec)^2
-
-#cfecAnnual = left_join(cfecAnnual, dplyr::select(temp, p_holder, year, p2))
 
@@
 cfecAnnual$strategy.prev = cfecAnnual$strategy[idx]
+cfecAnnual$weight.prev = cfecAnnual$weight[idx]
+cfecAnnual$days.prev = cfecAnnual$days[idx]
 #cfecAnnual$p2.prev = cfecAnnual$p2[idx]
@@
 cfecAnnual$specDiv.change = cfecAnnual$specDiv - cfecAnnual$specdiv.prev
-# idx = which(cfecAnnual$strategy != cfecAnnual$strategy.prev)
 
-# Filter out salmon people
-cfecAnnual$salmon = ifelse(substr(cfecAnnual$permit,1,1)=="S", "Y", "N")
+cfecAnnual.diff = cfecAnnual[is.na(cfecAnnual$specDiv.change)==F, ]
 
-cfecAnnual.salmon = cfecAnnual[cfecAnnual$salmon=="Y",]
-cfecAnnual.nonsalmon = cfecAnnual[cfecAnnual$salmon=="N",]
+saveRDS(cfecAnnual.diff, file="diff_linearModeling_complete.rds")
 
-#cfecAnnual.salmon = cfecAnnual.salmon[is.na(cfecAnnual.salmon$specDiv.change)==F, ]
-#cfecAnnual.nonsalmon = cfecAnnual.nonsalmon[is.na(cfecAnnual.nonsalmon$specDiv.change)==F, ]
+# 4200 different strategies, need to model only most common, 
+top.strategies = names(rev(sort(table(cfecAnnual.diff$strategy)))[1:100])
 
-saveRDS(cfecAnnual.salmon, file="salmon_linearModeling_complete.rds")
-saveRDS(cfecAnnual.nonsalmon, file="nonsalmon_linearModeling_complete.rds")
+cfecAnnual.diff = cfecAnnual.diff[cfecAnnual.diff$strategy%in%top.strategies, ]
+# idx = which(cfecAnnual.diff$strategy != cfecAnnual.diff$strategy.prev)
+# paste(cfecAnnual.diff$strategy.prev[idx], cfecAnnual.diff$strategy[idx], sep=":")
+cfecAnnual.diff = cfecAnnual.diff[which(cfecAnnual.diff$strategy == cfecAnnual.diff$strategy.prev), ]
+
+saveRDS(cfecAnnual.diff, file="diff_linearModeling_subset.rds")
 
 ```
-
-## Do some exploration by permit types + diversity
-
-For example, when someone is diversifying with a herring, halibut, or crab permit, what does that diversification look like?
-
-```{r, echo=FALSE}
-cfecnew$taxa = substr(cfecnew$p_fshy, 1, 1)
-specPermit = group_by(cfecnew, taxa, specn) %>% 
-  summarize(percent.revenue = sum(g_earn))
-specPermit.tot = group_by(specPermit, taxa) %>% 
-  summarize(tot = sum(percent.revenue))
-specPermit = left_join(specPermit, specPermit.tot)
-specPermit$percent.revenue = specPermit$percent.revenue/specPermit$tot
-specPermit.df = group_by(specPermit, taxa, specn) %>% 
-  filter(percent.revenue > 0.02) %>% 
-  arrange(taxa) %>% 
-  as.data.frame
-specPermit.df$spec = cfec$spec[match(specPermit.df$specn, cfec$specn)]
-kable(specPermit.df)
-```
-
-## Code for looking at permit 
-We can look at whether there's clear combinations of permit types that are combinations -- most of these people either have a single local permit (PWS, Bristol Bay, etc) or have multiple statewide permits.
-
-```{r}
-
-cfecnew$area = substr(cfecnew$p_fshy,4,4)
-cfecnew$area[cfecnew$area %in% c("A","C")] = "Southeast"
-cfecnew$area[cfecnew$area == "B"] = "Statewide"
-cfecnew$area[cfecnew$area == "D"] = "Yakutat"
-cfecnew$area[cfecnew$area == "E"] = "PWS"
-cfecnew$area[cfecnew$area == "H"] = "CookInlet"
-cfecnew$area[cfecnew$area == "J"] = "Westward"
-cfecnew$area[cfecnew$area == "K"] = "Kodiak"
-cfecnew$area[cfecnew$area == "L"] = "Chignik"
-cfecnew$area[cfecnew$area == "M"] = "AlaskaPeninsula"
-cfecnew$area[cfecnew$area == "N"] = "Nelson"
-cfecnew$area[cfecnew$area == "Q"] = "BeringSea"
-cfecnew$area[cfecnew$area == "T"] = "BristolBay"
-cfecnew$area[cfecnew$area == "X"] = "Kotzebue"
-cfecnew$area[cfecnew$area == "Z"] = "NortonSound"
-
-cfecnew$spec = substr(cfecnew$p_fshy,1,1)
-cfecnew$spec[cfecnew$spec == "B"] = "Halibut"
-cfecnew$spec[cfecnew$spec == "C"] = "Sablefish"
-cfecnew$spec[cfecnew$spec %in% c("D","E","K","T")] = "Crab"
-cfecnew$spec[cfecnew$spec %in% c("F")] = "Freshwater"
-cfecnew$spec[cfecnew$spec %in% c("G","H","L")] = "Herring"
-cfecnew$spec[cfecnew$spec %in% c("I")] = "Lingcod"
-cfecnew$spec[cfecnew$spec %in% c("A","J","N","R","W")] = "Shellfish"
-cfecnew$spec[cfecnew$spec %in% c("O","P","Q","U","Z")] = "Invert"
-cfecnew$spec[cfecnew$spec %in% c("M")] = "Misc"
-cfecnew$spec[cfecnew$spec %in% c("S")] = "Salmon"
-cfecnew$area.spec = paste(cfecnew$area, cfecnew$spec, sep="_")
-
-cfecAnnual.multiplePermits = group_by(cfecnew, p_holder, year) %>%
-  summarize(npermit = length(unique(p_fshy)), 
-    allpermits = paste(sort(unique(p_fshy)), collapse = " "),
-    areaspec = paste(sort(unique(area.spec)), collapse = " ")) %>%
-  filter(npermit > 1)
-
-```

### 2016-06-19 portfolio/analysis/6b-tmb-modeling_diff.R

@@ -1,0 +1,159 @@
+# Experiment with models 
+# by fitting them quickly with TMB
+rm(list = ls())
+library(TMB)
+library(dplyr)
+library(ggplot2)
+compile("revenue_diff.cpp")
+dyn.load(dynlib("revenue_diff"))
+
+# LOAD DATA
+cfecAnnual.diff = readRDS(file="diff_linearModeling_complete.rds")
+
+# 4200 different strategies, need to model only most common, 
+top.strategies = names(rev(sort(table(cfecAnnual.diff$strategy)))[1:100])
+cfecAnnual.diff = cfecAnnual.diff[cfecAnnual.diff$strategy%in%top.strategies, ]
+# restrict analysis to people who don't change strategies (keeps ~ 90%)
+testData = cfecAnnual.diff[which(cfecAnnual.diff$strategy == cfecAnnual.diff$strategy.prev), ]
+
+#P_HOLDER = as.numeric(as.factor(testData$p_holder))
+#STRATEGY = as.numeric(as.factor(testData$strategy))
+#RESPONSE = log(testData$revenue / testData$revenue.prev)
+#PREV_REV = log(testData$revenue.prev)
+#SPECDIV = log(testData$specdiv.prev)
+#SPEC_DIFF = log(testData$specDiv / testData$specdiv.prev)
+#WEIGHT_DIFF = log(testData$weight / testData$weight.prev)
+#DAYS_DIFF = log((testData$days + 1) / (testData$days.prev+1))
+#OFFSET = WEIGHT_DIFF
+
+format_data <- function(x, n_obs = 100) {
+  keep <- group_by(x, strategy)  %>% summarize(n = n())  %>% filter(n>=n_obs)
+  x <- x[x$strategy %in% keep$strategy, ]
+  #x <- x[!is.na(x$length), ]
+  x$permit_id <- as.numeric(as.factor(x$strategy))
+  x$log_spec_diff <- log(x$specDiv / x$specdiv.prev)
+  x$log_spec = log(x$specdiv.prev)
+  #x$log_length <- scale(log(x$length + 1))
+  x$log_days <- log((x$days + 1) / (x$days.prev+1))
+  x$offset = log(x$weight / x$weight.prev)
+  
+  mm <- model.matrix(~ log_spec_diff, data = x)
+  n_k <- max(x$permit_id)
+  n_fe <- ncol(mm)
+  
+  list(
+    raw_data = x,
[…]

### 2016-06-19 portfolio/analysis/revenue_diff.cpp

@@ -1,0 +1,105 @@
+#include <TMB.hpp>
+
+template<class Type>
+Type objective_function<Type>::operator() () {
+// data:
+DATA_MATRIX(x_ij);
+DATA_VECTOR(y_i);
+DATA_IVECTOR(k_i); // vector of IDs
+DATA_INTEGER(n_k); // number of IDs
+DATA_INTEGER(n_j); // number of IDs
+DATA_VECTOR(b1_cov_re_i); // predictor data for random slope
+DATA_VECTOR(sigma1_cov_re_i); // predictor data for random slope
+DATA_VECTOR(offset); // offset
+//DATA_VECTOR(sigma2_cov_re_i); // predictor data for random slope
+
+// parameters:
+PARAMETER_VECTOR(b_j)
+PARAMETER_VECTOR(sigma_j);
+PARAMETER(log_b1_sigma);
+PARAMETER_VECTOR(b1_k);
+PARAMETER(log_sigma0_sigma);
+PARAMETER(log_sigma1_sigma);
+PARAMETER_VECTOR(sigma0_k);
+PARAMETER_VECTOR(sigma1_k);
+//PARAMETER(log_b0_sigma);
+//PARAMETER_VECTOR(b0_k);
+//PARAMETER(log_sigma2_sigma);
+//PARAMETER_VECTOR(sigma2_k);
+
+int n_data = y_i.size(); // get number of data points to loop over
+
+// Linear predictor
+vector<Type> linear_predictor_i(n_data);
+vector<Type> linear_predictor_sigma_i(n_data);
+linear_predictor_i = x_ij*b_j;
+linear_predictor_sigma_i = x_ij*sigma_j;
+
+Type nll = 0.0; // initialize negative log likelihood
+
+for(int i = 0; i < n_data; i++){
+  nll -= dnorm(
+      y_i(i),
+
+      b1_k(k_i(i)) * b1_cov_re_i(i) +
+      linear_predictor_i(i) + offset(i),
+
+      sqrt(exp(
+          sigma0_k(k_i(i)) +
+          sigma1_k(k_i(i)) * sigma1_cov_re_i(i) +
+          linear_predictor_sigma_i(i))),
+
+      true);
+}
+for(int k = 0; k < n_k; k++){
[…]

### 2016-06-20 portfolio/analysis/4a-createData_linear_modeling.Rmd
Add Eric's pfishy amalgamation

@@
   filter(nyr >= 10)
+```
+
+```{r}
+cfecnew$p_fshy[cfecnew$p_fshy=="B61B"] = "B06B"
+cfecnew$p_fshy[cfecnew$p_fshy=="C61B"] = "C06B"
+cfecnew$p_fshy[cfecnew$p_fshy=="C91B"] = "C09B"
+cfecnew$p_fshy[cfecnew$p_fshy=="M61B"] = "M06B"
+cfecnew$p_fshy[cfecnew$p_fshy=="E09B"] = "E91B"
+cfecnew$p_fshy[cfecnew$p_fshy=="I06B"] = "I61B"
+cfecnew$p_fshy[cfecnew$p_fshy=="Y06A"] = "Y61A"
+cfecnew$p_fshy[cfecnew$p_fshy=="O06B"] = "O61B"
+cfecnew$p_fshy[cfecnew$p_fshy=="O09B"] = "O91B"
+
+cfecnew$p_fshy[cfecnew$p_fshy%in%c("M7FB","M7GB", "M7HB","M7IB")] = "M7IB"
+cfecnew$p_fshy[cfecnew$p_fshy%in%c("M7FG","M7GG", "M7HG","M7IG")] = "M7IG"
+
+# crab and shrimp vessels
+cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="D91")] = paste("D09", substr(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="D91")],4,length(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="D91")])), sep="")
+cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="T91")] = paste("T09", substr(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="T91")],4,length(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="T91")])), sep="")
+cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="P91")] = paste("P09", substr(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="P91")],4,length(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="P91")])), sep="")
+
+cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="M91")] = paste("M09", substr(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="M91")],4,length(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="M91")])), sep="")
+cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="M61")] = paste("M06", substr(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="M61")],4,length(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="M61")])), sep="")
 ```

### 2016-06-20 portfolio/analysis/4a-createData_linear_modeling.Rmd
Make script run universally; don't cut off strategies

@@
 #load("cfecCleaned2.Rdata")
+
+if (Sys.info()[["user"]] == "seananderson") {
+  cfec <- feather::read_feather("../data-generated/cfec.feather")
+} else {
   cfec <- feather::read_feather("data/cfec.feather")
+}
 
@@
 
+if (Sys.info()[["user"]] != "seananderson") {
   region<-read.csv("areas/regions.csv")
@@
 #cfec$region <- replace(cfec$region, is.na(cfec$region), "Unassigned")
+}
 ```
@@
 ```{r, echo=FALSE}
+if (Sys.info()[["user"]] != "seananderson") {
 deflationTable = data.frame(year = 1985:2014, defl = c(0.57236, 0.58393, 0.59879, 0.61974,
@@
 cfec$g_earn = cfec$g_earn/deflationTable$defl[cfec$year-1984]
+}
 ```
@@
 ### Function definiton
-We need to define a few functions for calculating diversity.
-We'll use the Simpson's diversity index (HHI) to calculate the entropy-based inverse, which is a proxy for effective species diversity.
-Note the C++ version is commented out below,
+We need to define a few functions for calculating diversity. We'll use the Simpson's diversity index (HHI) to calculate the entropy-based inverse, which is a proxy for effective species diversity.
 
@@
 # Delete these minimal earnings
-cfecnew = cfecnew[-which(paste(cfecnew$p_holder, cfecnew$year, cfecnew$p_fshy) %in% paste(permit_rev$p_holder, permit_rev$year, permit_rev$p_fshy))]
+cfecnew = cfecnew[-which(paste(cfecnew$p_holder, cfecnew$year, cfecnew$p_fshy) %in% paste(permit_rev$p_holder, permit_rev$year, permit_rev$p_fshy)), ]
 
@@
 CUTOFF = 50
-cfecnew = cfecnew[cfecnew$strategy%in%names(rev(sort(table(topstrategies$strategy)))[1:CUTOFF]), ]
+write.csv(rev(sort(table(topstrategies$strategy))), file = "../data-generated/strategies.csv")
+# cfecnew = cfecnew[cfecnew$strategy%in%names(rev(sort(table(topstrategies$strategy)))[1:CUTOFF]), ]
 ```
@@
 
+# fixme: lubridate?
 cfecnew$day = mdy.date(as.numeric(lapply(strsplit(cfecnew$day, "-"), getElement, 1)), as.numeric(lapply(strsplit(cfecnew$day, "-"), getElement, 2)), cfecnew$year) - mdy.date(1, 1, cfecnew$year)
 
+# fixme What is cadfg2?
+if ("VLength" %in% names(cfecnew)) {
+  cfecnew <- rename(cfecnew, vlength = VLength)
+}
 cfecAnnual = group_by(cfecnew, p_holder, year) %>%
-  summarize(permit = p_fshy[1], boat = cadfg2[1], 
+  summarize(permit = p_fshy[1], boat = cadfg[1], # fixme this was cadfg2
     days = diff(range(day) + 1),
@@
     revenue = sum(g_earn),
-    length = mean(VLength, na.rm=T),
+    weight = sum(g_pounds),
+    length = mean(vlength, na.rm=TRUE),
     strategy = strategy[1])
 
-cfecAnnual = group_by(cfecAnnual, p_holder, year) %>%
-  filter(npermit == 1)
+# fixme I assume we no longer want to do this? -SA
+#cfecAnnual = group_by(cfecAnnual, p_holder, year) %>%
+#  filter(npermit == 1)
 
@@
 # Filter out salmon people
-cfecAnnual$salmon = ifelse(substr(cfecAnnual$permit,1,1)=="S", "Y", "N")
+cfecAnnual$salmon = ifelse(substr(cfecAnnual$permit,1,1)=="S", TRUE, FALSE)
 
-cfecAnnual.salmon = cfecAnnual[cfecAnnual$salmon=="Y",]
-cfecAnnual.nonsalmon = cfecAnnual[cfecAnnual$salmon=="N",]
+# cfecAnnual.salmon = cfecAnnual[cfecAnnual$salmon=="Y",]
+# cfecAnnual.nonsalmon = cfecAnnual[cfecAnnual$salmon=="N",]
 
@@
 
-saveRDS(cfecAnnual.salmon, file="salmon_linearModeling_complete.rds")
-saveRDS(cfecAnnual.nonsalmon, file="nonsalmon_linearModeling_complete.rds")
+# saveRDS(cfecAnnual.salmon, file="salmon_linearModeling_complete.rds")
+# saveRDS(cfecAnnual.nonsalmon, file="nonsalmon_linearModeling_complete.rds")
+cfecAnnual <- as_data_frame(cfecAnnual)
+nrow(cfecAnnual)
+saveRDS(cfecAnnual, file="../data-generated/cfec-for-modeling.rds")
 
@@
 
-cfecnew$area = substr(cfecnew$p_fshy,4,4)
-cfecnew$area[cfecnew$area %in% c("A","C")] = "Southeast"
-cfecnew$area[cfecnew$area == "B"] = "Statewide"
-cfecnew$area[cfecnew$area == "D"] = "Yakutat"
-cfecnew$area[cfecnew$area == "E"] = "PWS"
-cfecnew$area[cfecnew$area == "H"] = "CookInlet"
-cfecnew$area[cfecnew$area == "J"] = "Westward"
-cfecnew$area[cfecnew$area == "K"] = "Kodiak"
-cfecnew$area[cfecnew$area == "L"] = "Chignik"
-cfecnew$area[cfecnew$area == "M"] = "AlaskaPeninsula"
-cfecnew$area[cfecnew$area == "N"] = "Nelson"
-cfecnew$area[cfecnew$area == "Q"] = "BeringSea"
-cfecnew$area[cfecnew$area == "T"] = "BristolBay"
-cfecnew$area[cfecnew$area == "X"] = "Kotzebue"
-cfecnew$area[cfecnew$area == "Z"] = "NortonSound"
-
-cfecnew$spec = substr(cfecnew$p_fshy,1,1)
-cfecnew$spec[cfecnew$spec == "B"] = "Halibut"
-cfecnew$spec[cfecnew$spec == "C"] = "Sablefish"
-cfecnew$spec[cfecnew$spec %in% c("D","E","K","T")] = "Crab"
-cfecnew$spec[cfecnew$spec %in% c("F")] = "Freshwater"
-cfecnew$spec[cfecnew$spec %in% c("G","H","L")] = "Herring"
-cfecnew$spec[cfecnew$spec %in% c("I")] = "Lingcod"
-cfecnew$spec[cfecnew$spec %in% c("A","J","N","R","W")] = "Shellfish"
-cfecnew$spec[cfecnew$spec %in% c("O","P","Q","U","Z")] = "Invert"
-cfecnew$spec[cfecnew$spec %in% c("M")] = "Misc"
-cfecnew$spec[cfecnew$spec %in% c("S")] = "Salmon"
-cfecnew$area.spec = paste(cfecnew$area, cfecnew$spec, sep="_")
-
-cfecAnnual.multiplePermits = group_by(cfecnew, p_holder, year) %>%
-  summarize(npermit = length(unique(p_fshy)), 
-    allpermits = paste(sort(unique(p_fshy)), collapse = " "),
-    areaspec = paste(sort(unique(area.spec)), collapse = " ")) %>%
-  filter(npermit > 1)
+# cfecnew$area = substr(cfecnew$p_fshy,4,4)
+# cfecnew$area[cfecnew$area %in% c("A","C")] = "Southeast"
+# cfecnew$area[cfecnew$area == "B"] = "Statewide"
+# cfecnew$area[cfecnew$area == "D"] = "Yakutat"
+# cfecnew$area[cfecnew$area == "E"] = "PWS"
+# cfecnew$area[cfecnew$area == "H"] = "CookInlet"
+# cfecnew$area[cfecnew$area == "J"] = "Westward"
+# cfecnew$area[cfecnew$area == "K"] = "Kodiak"
+# cfecnew$area[cfecnew$area == "L"] = "Chignik"
+# cfecnew$area[cfecnew$area == "M"] = "AlaskaPeninsula"
+# cfecnew$area[cfecnew$area == "N"] = "Nelson"
+# cfecnew$area[cfecnew$area == "Q"] = "BeringSea"
+# cfecnew$area[cfecnew$area == "T"] = "BristolBay"
+# cfecnew$area[cfecnew$area == "X"] = "Kotzebue"
+# cfecnew$area[cfecnew$area == "Z"] = "NortonSound"
+#
+# cfecnew$spec = substr(cfecnew$p_fshy,1,1)
+# cfecnew$spec[cfecnew$spec == "B"] = "Halibut"
+# cfecnew$spec[cfecnew$spec == "C"] = "Sablefish"
+# cfecnew$spec[cfecnew$spec %in% c("D","E","K","T")] = "Crab"
+# cfecnew$spec[cfecnew$spec %in% c("F")] = "Freshwater"
+# cfecnew$spec[cfecnew$spec %in% c("G","H","L")] = "Herring"
+# cfecnew$spec[cfecnew$spec %in% c("I")] = "Lingcod"
+# cfecnew$spec[cfecnew$spec %in% c("A","J","N","R","W")] = "Shellfish"
+# cfecnew$spec[cfecnew$spec %in% c("O","P","Q","U","Z")] = "Invert"
+# cfecnew$spec[cfecnew$spec %in% c("M")] = "Misc"
+# cfecnew$spec[cfecnew$spec %in% c("S")] = "Salmon"
+# cfecnew$area.spec = paste(cfecnew$area, cfecnew$spec, sep="_")
+#
+# cfecAnnual.multiplePermits = group_by(cfecnew, p_holder, year) %>%
+#   summarize(npermit = length(unique(p_fshy)),
+#     allpermits = paste(sort(unique(p_fshy)), collapse = " "),
+#     areaspec = paste(sort(unique(area.spec)), collapse = " ")) %>%
+#   filter(npermit > 1)
 

### 2016-06-29 portfolio/analysis/4a-createData_linear_modeling.Rmd
Merge annual and differenced data creation scripts

@@
 ---
-title: "createData_linear_modeling"
-author: ""
-date: "June 05, 2016"
+title: "Create data sets for modeling"
 output:
@@
 
+## Differenced vs. undifferenced datasets:
+
+1. Excluding observations that we don't have consecutive years for
+
 ```{r setup, include=FALSE, echo=FALSE}
@@
 ## Load cleaned data
+
 ```{r dataLoad, warning = FALSE, message=FALSE, results="hide", echo=FALSE}
@@
 library(date)
-
-#load("cfecCleaned2.Rdata")
 
@@
   cfec<-left_join(cfec,region)
-#cfec$region <- replace(cfec$region, is.na(cfec$region), "Unassigned")
 }
@@
 
-We need to adjust the landed values and prices for inflation. To do this, we'll use the data Alan pulled together (adjusting to 2009 dollars).
+We need to adjust the landed values and prices for inflation. To do this,
+we'll use the data Alan pulled together (adjusting to 2009 dollars).
 
-https://docs.google.com/spreadsheets/d/1hRLYTX1nZBLKnYXwTCOtU7-XnDoMZf_1ufdTiNQFd4U/edit#gid=0
+<https://docs.google.com/spreadsheets/d/1hRLYTX1nZBLKnYXwTCOtU7-XnDoMZf_1ufdTiNQFd4U/edit#gid=0>
 
@@
 if (Sys.info()[["user"]] != "seananderson") {
-deflationTable = data.frame(year = 1985:2014, defl = c(0.57236, 0.58393, 0.59879, 0.61974,
+deflationTable = data.frame(year = 1985:2014, 
+  defl = c(0.57236, 0.58393, 0.59879, 0.61974,
 0.64388, 0.66774, 0.68993, 0.70564, 0.72244, 0.73781,
@@
 
-### Function definiton
-We need to define a few functions for calculating diversity. We'll use the Simpson's diversity index (HHI) to calculate the entropy-based inverse, which is a proxy for effective species diversity.
+## Function definition
+
+We need to define a few functions for calculating diversity. We'll use the
+Simpson's diversity index (HHI) to calculate the entropy-based inverse, which
+is a proxy for effective species diversity.
 
@@
 
-Now we'll filter out individuals who only used 1 boat in a year, and boats with only one permit holder fishing in a year.
+Now we'll filter out (i.e. keep) individuals who only used 1 boat in a year, and boats with
+only one permit holder fishing in a year.
 
@@
 
-## Filtering out people with 10+ years of fishing. Drops rows from 13.7 to 10 million.
-```{r}
-cfecnew = group_by(cfecnew, p_holder) %>%
-  mutate(nyr = length(unique(year))) %>%
-  filter(nyr >= 10)
-```
+## Combine some strategies
 
@@
 
-###For variability, any person in a year will need to have 10% or more of their revenue to come from each permit (otherwise doesn't count as a strategy)
+Any person in a year will need to have 10% or more of their revenue to come
+from each permit (otherwise doesn't count as a strategy).
 
-The top 120 permits cuts the rows down to 8.6 million. The top 50 gets it down to 7.8 million.
+The top 120 permits cuts the rows down to 8.6 million. The top 50 gets it down
+to 7.8 million.
 
@@
 
-CUTOFF = 50
 write.csv(rev(sort(table(topstrategies$strategy))), file = "../data-generated/strategies.csv")
-# cfecnew = cfecnew[cfecnew$strategy%in%names(rev(sort(table(topstrategies$strategy)))[1:CUTOFF]), ]
 ```
@@
 
-# fixme I assume we no longer want to do this? -SA
-#cfecAnnual = group_by(cfecAnnual, p_holder, year) %>%
-#  filter(npermit == 1)
-
 # also calculate weighted simpson diversity of gears and species
@@
 
-# random slope by p_holder
-# Random intercepts in people, random intercepts in strategies, random slopes in strategies
-# Group level predictors on intercepts
-# Indicator variable for EVOS years. Or whether
-
 # also calculate weighted simpson diversity of gears and species
@@
   summarize(gearDiv = simp.div(totRev.gear)))
-
-#temp = group_by(cfecnew, p_holder, year, specn) %>%
-#  summarize(totRev.spec = sum(g_earn),
-#    permit = p_fshy[1])
-#temp = left_join(temp, group_by(cfecnew, p_holder, year) %>%
-#    summarize(totRev = sum(g_earn)))
-#temp$totRev.spec = temp$totRev.spec / temp$totRev
-
-#mkt = group_by(cfecnew, p_fshy, year, specn) %>%
-#  summarize(totRev.mkt.spec = sum(g_earn))
-#mkt = left_join(mkt, group_by(mkt, p_fshy, year) %>%
-#  summarize(totRev.mkt = sum(totRev.mkt.spec)))
-#mkt$totRev.mkt.spec = mkt$totRev.mkt.spec / mkt$totRev.mkt
-#names(mkt)[1] = "permit"
-
-# combine individual and market datasets
-#temp = left_join(temp, mkt)
-
-# Calculate squared proportions
-#temp$p2 = (temp$totRev.spec - temp$totRev.mkt.spec)^2
-
-#cfecAnnual = left_join(cfecAnnual, dplyr::select(temp, p_holder, year, p2))
 
@@
 cfecAnnual$strategy.prev = cfecAnnual$strategy[idx]
-#cfecAnnual$p2.prev = cfecAnnual$p2[idx]
+cfecAnnual$weight.prev = cfecAnnual$weight[idx]
+cfecAnnual$days.prev = cfecAnnual$days[idx]
+
+# idx = which(cfecAnnual$strategy != cfecAnnual$strategy.prev)
+```
+
+Add some identifying columns:
+
+```{r}
+cfecAnnual$area = substr(cfecAnnual$p_fshy,4,4)
+cfecAnnual$area[cfecAnnual$area %in% c("A","C")] = "Southeast"
+cfecAnnual$area[cfecAnnual$area == "B"] = "Statewide"
+cfecAnnual$area[cfecAnnual$area == "D"] = "Yakutat"
+cfecAnnual$area[cfecAnnual$area == "E"] = "PWS"
+cfecAnnual$area[cfecAnnual$area == "H"] = "CookInlet"
+cfecAnnual$area[cfecAnnual$area == "J"] = "Westward"
+cfecAnnual$area[cfecAnnual$area == "K"] = "Kodiak"
+cfecAnnual$area[cfecAnnual$area == "L"] = "Chignik"
+cfecAnnual$area[cfecAnnual$area == "M"] = "AlaskaPeninsula"
+cfecAnnual$area[cfecAnnual$area == "N"] = "Nelson"
+cfecAnnual$area[cfecAnnual$area == "Q"] = "BeringSea"
+cfecAnnual$area[cfecAnnual$area == "T"] = "BristolBay"
+cfecAnnual$area[cfecAnnual$area == "X"] = "Kotzebue"
+cfecAnnual$area[cfecAnnual$area == "Z"] = "NortonSound"
+
+cfecAnnual$spec = substr(cfecAnnual$p_fshy,1,1)
+cfecAnnual$spec[cfecAnnual$spec == "B"] = "Halibut"
+cfecAnnual$spec[cfecAnnual$spec == "C"] = "Sablefish"
+cfecAnnual$spec[cfecAnnual$spec %in% c("D","E","K","T")] = "Crab"
+cfecAnnual$spec[cfecAnnual$spec %in% c("F")] = "Freshwater"
+cfecAnnual$spec[cfecAnnual$spec %in% c("G","H","L")] = "Herring"
+cfecAnnual$spec[cfecAnnual$spec %in% c("I")] = "Lingcod"
+cfecAnnual$spec[cfecAnnual$spec %in% c("A","J","N","R","W")] = "Shellfish"
+cfecAnnual$spec[cfecAnnual$spec %in% c("O","P","Q","U","Z")] = "Invert"
+cfecAnnual$spec[cfecAnnual$spec %in% c("M")] = "Misc"
+cfecAnnual$spec[cfecAnnual$spec %in% c("S")] = "Salmon"
+cfecAnnual$area_spec = paste(cfecAnnual$area, cfecAnnual$spec, sep="_")
+```
+
+Save the data for modeling:
+
+```{r}
+# make sure the data are tidy:
+cfecAnnual <- as_data_frame(cfecAnnual)
+
+## Filtering out people with 10+ years of fishing. Drops rows from 13.7 to 10 million.
+## (now doing this at the beginning of the modeling scripts)
+# cfecAnnual = group_by(cfecAnnual, p_holder) %>%
+#   mutate(nyr = length(unique(year))) %>%
+#   filter(nyr >= 10)
+
+saveRDS(cfecAnnual, file="../data-generated/cfec-annual-for-modeling.rds")
 
 cfecAnnual$specDiv.change = cfecAnnual$specDiv - cfecAnnual$specdiv.prev
-# idx = which(cfecAnnual$strategy != cfecAnnual$strategy.prev)
-
-# Filter out salmon people
-cfecAnnual$salmon = ifelse(substr(cfecAnnual$permit,1,1)=="S", TRUE, FALSE)
-
-# cfecAnnual.salmon = cfecAnnual[cfecAnnual$salmon=="Y",]
-# cfecAnnual.nonsalmon = cfecAnnual[cfecAnnual$salmon=="N",]
-
-#cfecAnnual.salmon = cfecAnnual.salmon[is.na(cfecAnnual.salmon$specDiv.change)==F, ]
-#cfecAnnual.nonsalmon = cfecAnnual.nonsalmon[is.na(cfecAnnual.nonsalmon$specDiv.change)==F, ]
-
-# saveRDS(cfecAnnual.salmon, file="salmon_linearModeling_complete.rds")
-# saveRDS(cfecAnnual.nonsalmon, file="nonsalmon_linearModeling_complete.rds")
-cfecAnnual <- as_data_frame(cfecAnnual)
-nrow(cfecAnnual)
-saveRDS(cfecAnnual, file="../data-generated/cfec-for-modeling.rds")
-
+cfecAnnual.diff = cfecAnnual[is.na(cfecAnnual$specDiv.change)==FALSE, ]
+saveRDS(cfecAnnual.diff, file="../data-generated/cfec-diff-for-modeling.rds")
 ```

### 2016-06-29 portfolio/analysis/4a-createData_linear_modeling.Rmd
Switch to lubridate for date parsing

@@
 
-# fixme: lubridate?
-cfecnew$day = mdy.date(as.numeric(lapply(strsplit(cfecnew$day, "-"), getElement, 1)), as.numeric(lapply(strsplit(cfecnew$day, "-"), getElement, 2)), cfecnew$year) - mdy.date(1, 1, cfecnew$year)
+cfecnew$yday <- paste(cfecnew$year, cfecnew$day, sep = "-") %>%
+  lubridate::ymd() %>%
+  lubridate::yday() - 1
 
@@
   summarize(permit = p_fshy[1], boat = cadfg[1], # fixme this was cadfg2
-    days = diff(range(day) + 1),
+    days = diff(range(yday) + 1),
     ngear = length(unique(gearn)),

### 2016-06-29 portfolio/text/supp.tex
Add Eric's notes on the differenced model

\documentclass[12pt]{article}
\usepackage{geometry}
\geometry{verbose, letterpaper, tmargin = 2.54cm, bmargin = 2.54cm,
  lmargin = 2.54cm, rmargin = 2.54cm}
\geometry{letterpaper}
\usepackage{amsmath}
\usepackage{fixltx2e}

\begin{document}

\section{Log-difference model}

The log-difference model has several pros and cons over the undifferenced model. Negatives are that it uses a smaller subset of the data; positives are that it doesn't require estimating people-specific random effects (n = 3000+) and removes possible time effects. The basic regression model is

\begin{equation}
  { Y }_{ i,t } \sim \mathrm{N}(\hat { { Y }}_{ i,t } ,\sigma^2 )
\end{equation}

\subsection{Model for the mean}

The model's expectation is written as a hierarchical random effects linear model with random effects in the intercept and slope.

\begin{equation}
{ \hat { Y }  }_{ t,i }={ B }_{ 0,j }+{ B }_{ 1,j }\cdot { s }_{ t,i }+{o}_{ t,i}
\end{equation}

where ${ B }_{ 0,j }$ is a normally distributed random effect intercept in the mean (j subscripting different strategies). The interpretation of ${B}_{ 0,j}$ is that each after log-differencing, each strategy has different effects on changes in revenue (some tend to be winners, others tend to be losers). Initial versions of this model included estimating ${ B }_{ 0,j }$, which were largely centered on 0 (as expected), so these random effects were set to 0.  
\newline

The covariate ${ s }_{ t,i }$ is input as data, and represents the percent change in species diversity over the same time period that the percent change in revenue is being modeled, $log\left( \frac { { div }_{ t+1,i } }{ { div }_{ t,i } }  \right)$ where ${ div }_{ t,i }$ is the effective diversity of individual *i* at time *t*. The coefficient ${ B }_{ 1,j }$ is a random effect on the slope, by strategy, representing how much an X% change in species diversity within a strategy translates to a Y% change in revenue.

\subsection{Offset}

To account for variable effort over time (changes in revenue could be attributed to diversification, but also to more fishing) we included an offset term ${o}_{t,i}$, representing the percent change in landing dates between years. We experminted with different versions of the offset and when included in the model as a predictor, the coefficient is estimated to be ~ 1, so this is a reasonable assumption. The offset is calculated as $log\left( \frac { { day }_{ t+1,i } }{ { day }_{ t,i } }  \right)$. 
\newline

\subsection{Equation for the variance}

Finally, the residual error term is also modeled as a linear function of covariates,

${ \sigma  }_{ t,i }=\sqrt { \left( \gamma _{ 0,j }+{ \gamma  }_{ 1,j }\cdot { s }_{ t,i } \right)  }$

where the $gamma _{ 0,j }$ and $gamma _{ 1,j }$ represent random intercepts and slopes in the strategies. I had been interpreting the intercepts as the between-group variability in strategies (some strategies will have more variable % change in revenue than others, similar to some stocks having higher variability in returns). I was interpreting the random slope $gamma _{ 1,j }$ as the effect of within - strategy diversification; in other words, the effect of increases in species diversification varied among different strategies.

\end{document}

### 2016-06-29 portfolio/text/supp.tex
Edit log difference model notes

@@
 
-The log-difference model has several pros and cons over the undifferenced model.
-Negatives are that it uses a smaller subset of the data; positives are that it doesn't require estimating people-specific random effects (n = 3000+) and removes possible time effects.
-The basic regression model is
+The log-difference model has several pros and cons over the undifferenced
+model. Negatives are that it uses a smaller subset of the data; positives are
+that it doesn't require estimating people-specific random effects (n = 3000+)
+and removes possible time effects. The basic regression model is
 
@@
 
-The model's expectation is written as a hierarchical random effects linear model with random effects in the intercept and slope.
+The model's expectation is written as a hierarchical random effects linear
+model with random effects in the intercept and slope.
 
@@
 
-where ${ B }_{ 0,j }$ is a normally distributed random effect intercept in the mean (j subscripting different strategies).
-The interpretation of ${B}_{ 0,j}$ is that each after log-differencing, each strategy has different effects on changes in revenue (some tend to be winners, others tend to be losers).
-Initial versions of this model included estimating ${ B }_{ 0,j }$, which were largely centered on 0 (as expected), so these random effects were set to 0.
-\newline
+\noindent
+where ${ B }_{ 0,j }$ is a normally distributed random effect intercept in the
+mean (j subscripting different strategies). The interpretation of ${B}_{ 0,j}$
+is that each after log-differencing, each strategy has different effects on
+changes in revenue (some tend to be winners, others tend to be losers). Initial
+versions of this model included estimating ${ B }_{ 0,j }$, which were largely
+centered on 0 (as expected), so these random effects were set to 0.  \newline
 
-The covariate ${ s }_{ t,i }$ is input as data, and represents the percent change in species diversity over the same time period that the percent change in revenue is being modeled, $log\left( \frac { { div }_{ t+1,i } }{ { div }_{ t,i } }  \right)$ where ${ div }_{ t,i }$ is the effective diversity of individual *i* at time *t*.
-The coefficient ${ B }_{ 1,j }$ is a random effect on the slope, by strategy, representing how much an X% change in species diversity within a strategy translates to a Y% change in revenue.
+The covariate ${ s }_{ t,i }$ is input as data, and represents the percent
+change in species diversity over the same time period that the percent change
+in revenue is being modeled, $\log\left( \frac { { div }_{ t+1,i } }{ { div }_{
+t,i } }  \right)$ where ${ div }_{ t,i }$ is the effective diversity of
+individual $i$ at time $t$. The coefficient ${ B }_{ 1,j }$ is a random effect
+on the slope, by strategy, representing how much an X\% change in species
+diversity within a strategy translates to a Y\% change in revenue.
 
@@
 
-To account for variable effort over time (changes in revenue could be attributed to diversification, but also to more fishing) we included an offset term ${o}_{t,i}$, representing the percent change in landing dates between years.
-We experminted with different versions of the offset and when included in the model as a predictor, the coefficient is estimated to be ~ 1, so this is a reasonable assumption.
-The offset is calculated as $log\left( \frac { { day }_{ t+1,i } }{ { day }_{ t,i } }  \right)$.
-\newline
+To account for variable effort over time (changes in revenue could be
+attributed to diversification, but also to more fishing) we included an offset
+term ${o}_{t,i}$, representing the percent change in landing dates between
+years. We experimented with different versions of the offset and when included
+in the model as a predictor, the coefficient is estimated to be ~ 1, so this is
+a reasonable assumption. The offset is calculated as $\log\left( \frac { { day
+}_{ t+1,i } }{ { day }_{ t,i } }  \right)$. \newline
 
@@
 
-Finally, the residual error term is also modeled as a linear function of covariates,
+Finally, the scale of the residual error term is also modeled as a linear
+function of covariates,
 
-${ \sigma  }_{ t,i }=\sqrt { \left( \gamma _{ 0,j }+{ \gamma  }_{ 1,j }\cdot { s }_{ t,i } \right)  }$
+\begin{equation}
+{ \sigma  }_{ t,i }=
+\sqrt { \exp \left( \gamma _{ 0,j }+
+{ \gamma  }_{ 1,j }\cdot { s }_{ t,i } \right)  }
+\end{equation}
 
-where the $gamma _{ 0,j }$ and $gamma _{ 1,j }$ represent random intercepts and slopes in the strategies.
-I had been interpreting the intercepts as the between-group variability in strategies (some strategies will have more variable % change in revenue than others, similar to some stocks having higher variability in returns).
-I was interpreting the random slope $gamma _{ 1,j }$ as the effect of within - strategy diversification; in other words, the effect of increases in species diversification varied among different strategies.
+\noindent
+where the $\gamma_{ 0,j }$ and $\gamma_{ 1,j }$ represent random intercepts and
+slopes in the strategies. I had been interpreting the intercepts as the
+between-group variability in strategies (some strategies will have more
+variable \% change in revenue than others, similar to some stocks having higher
+variability in returns). I was interpreting the random slope $\gamma_{ 1,j }$
+as the effect of within-strategy diversification; in other words, the effect
+of increases in species diversification varied among different strategies.
 

### 2016-06-29 portfolio/analysis/4a-createData_linear_modeling.Rmd
Remove incorrect code

@@
 
-Add some identifying columns:
-
-```{r}
-cfecAnnual$area = substr(cfecAnnual$p_fshy,4,4)
-cfecAnnual$area[cfecAnnual$area %in% c("A","C")] = "Southeast"
-cfecAnnual$area[cfecAnnual$area == "B"] = "Statewide"
-cfecAnnual$area[cfecAnnual$area == "D"] = "Yakutat"
-cfecAnnual$area[cfecAnnual$area == "E"] = "PWS"
-cfecAnnual$area[cfecAnnual$area == "H"] = "CookInlet"
-cfecAnnual$area[cfecAnnual$area == "J"] = "Westward"
-cfecAnnual$area[cfecAnnual$area == "K"] = "Kodiak"
-cfecAnnual$area[cfecAnnual$area == "L"] = "Chignik"
-cfecAnnual$area[cfecAnnual$area == "M"] = "AlaskaPeninsula"
-cfecAnnual$area[cfecAnnual$area == "N"] = "Nelson"
-cfecAnnual$area[cfecAnnual$area == "Q"] = "BeringSea"
-cfecAnnual$area[cfecAnnual$area == "T"] = "BristolBay"
-cfecAnnual$area[cfecAnnual$area == "X"] = "Kotzebue"
-cfecAnnual$area[cfecAnnual$area == "Z"] = "NortonSound"
-
-cfecAnnual$spec = substr(cfecAnnual$p_fshy,1,1)
-cfecAnnual$spec[cfecAnnual$spec == "B"] = "Halibut"
-cfecAnnual$spec[cfecAnnual$spec == "C"] = "Sablefish"
-cfecAnnual$spec[cfecAnnual$spec %in% c("D","E","K","T")] = "Crab"
-cfecAnnual$spec[cfecAnnual$spec %in% c("F")] = "Freshwater"
-cfecAnnual$spec[cfecAnnual$spec %in% c("G","H","L")] = "Herring"
-cfecAnnual$spec[cfecAnnual$spec %in% c("I")] = "Lingcod"
-cfecAnnual$spec[cfecAnnual$spec %in% c("A","J","N","R","W")] = "Shellfish"
-cfecAnnual$spec[cfecAnnual$spec %in% c("O","P","Q","U","Z")] = "Invert"
-cfecAnnual$spec[cfecAnnual$spec %in% c("M")] = "Misc"
-cfecAnnual$spec[cfecAnnual$spec %in% c("S")] = "Salmon"
-cfecAnnual$area_spec = paste(cfecAnnual$area, cfecAnnual$spec, sep="_")
-```
-
 Save the data for modeling:

### 2016-07-01 portfolio/analysis/7-separate-models-mean.Rmd
Add first pass at 2-model testing

---
title: "Dive into separate models"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

Let's take a look at the simpler model that uses the CV of revenue for each permit holder as a response. If our more complicated models with structure in the error term are doing what we think then we should be able to replicate similar conclusions.

```{r}
library(dplyr)
library(ggplot2)
library(lme4)
rm(list = ls())
dm <- readRDS("../data-generated/cfec-annual-for-modeling.rds")
nrow(dm)
```

We'll filter the data to keep permit holders that fished for at least ten years. We'll also have to filter out any permit holders that have no variability in the diversity fished.

```{r}
dm <- group_by(dm, p_holder) %>%
  mutate(nyr = length(unique(year))) %>%
  mutate(range_div = diff(range(specDiv))) %>%
  as_data_frame() %>%
  filter(nyr >= 10) %>%
  filter(range_div > 0.5)
nrow(dm)

top.strategies = names(rev(sort(table(dm$strategy)))[1:100])
dm = dm[dm$strategy%in%top.strategies, ]
nrow(dm)

dm_holder <- group_by(dm, p_holder) %>%
  summarize(mean_diversity = mean(log(specDiv)))%>%
  mutate(p_holder = as.character(p_holder))

dm_strategy <- group_by(dm, strategy) %>%
  summarize(mean_diversity = mean(log(specDiv)))

dm$log_spec_div <- scale(log(dm$specDiv))
dm$log_length <- scale(log(dm$length + 1))
dm$log_weight <- scale(log(dm$weight + 1))
dm$log_days <- scale(log(dm$days + 1))
```

We'll fit a model with `lmer()`.

```{r}
m5 <- lmer(log(revenue) ~ log_spec_div * log_days * log_length +
    (1 + log_spec_div|strategy) + (1 + log_spec_div|p_holder), data = dm)
co_strategy <- coef(m5)$strategy
co_strategy$strategy <- row.names(co_strategy)
names(co_strategy)[1:2] <- c("intercept", "log_diversity_re")

co_holder <- coef(m5)$p_holder
co_holder$p_holder <- row.names(co_holder)
names(co_holder)[1:2] <- c("intercept", "log_diversity_re")

arm::display(m5)
```

```{r}
co_holder <- co_holder %>%
  inner_join(dm_holder %>% unique())
ggplot(co_holder, aes(mean_diversity, intercept)) + geom_point() + geom_smooth()
ggplot(co_holder, aes(mean_diversity, log_diversity_re)) + geom_point() + geom_smooth()
```

```{r}
co_strategy <- co_strategy %>%
  inner_join(dm_strategy %>% unique())
ggplot(co_strategy, aes(mean_diversity, intercept, label = strategy)) + geom_text() + geom_smooth()
ggplot(co_strategy, aes(mean_diversity, log_diversity_re, label = strategy)) + geom_text() + geom_smooth()
```

```{r, fig.width=10, fig.height=6}
strategy_grouped = rbind(cbind(co_strategy[grep("M",co_strategy$strategy),], group="Groundfish"),
  cbind(co_strategy[grep("S",co_strategy$strategy),], group="Salmon"),
  cbind(co_strategy[grep("B|C",co_strategy$strategy),], group="Halibut-Sablefish"),
  cbind(co_strategy[grep("K|T|D",co_strategy$strategy),], group="Crab"),
  cbind(co_strategy[grep("G",co_strategy$strategy),], group="Herring"))

ggplot(strategy_grouped, aes(mean_diversity, intercept, label = strategy)) +
  geom_text() + geom_smooth(se = FALSE, method = "lm") +
  facet_wrap(~group, scales = "free_y")
ggplot(strategy_grouped, aes(mean_diversity, log_diversity_re, label = strategy)) +
  geom_text() + geom_smooth(se = FALSE, method = "lm") +
  facet_wrap(~group, scales = "free_y")
```

### 2016-07-01 portfolio/analysis/7-separate-models.Rmd
Add first pass at 2-model testing

---
title: "Dive into separate models"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Let's take a look at the simpler model that uses the CV of revenue for each permit holder as a response. If our more complicated models with structure in the error term are doing what we think then we should be able to replicate similar conclusions.

```{r}
library(dplyr)
library(ggplot2)
library(lme4)
rm(list = ls())
dm <- readRDS("../data-generated/cfec-annual-for-modeling.rds")
nrow(dm)
```

We'll filter the data to keep permit holders that fished for at least ten years. We'll also have to filter out any permit holders that have no variability in the diversity fished.

```{r}
dm <- group_by(dm, p_holder) %>%
  mutate(nyr = length(unique(year))) %>%
  mutate(range_div = diff(range(specDiv))) %>%
  as_data_frame() %>%
  filter(nyr >= 10) %>%
  filter(range_div > 0)
nrow(dm)

# # we don't need to crop the strategies at this point
# top.strategies = names(rev(sort(table(dm$strategy)))[1:100])
# dm = dm[dm$strategy%in%top.strategies, ]
# nrow(dm)

dm$log_spec_div <- scale(log(dm$specDiv))
dm$log_length <- scale(log(dm$length + 1))
dm$log_weight <- scale(log(dm$weight + 1))
dm$log_days <- scale(log(dm$days + 1))
```

We'll consider strategies taxonomic groups here and keep all strategies that have at least 5 permit holders.

```{r}
dmv <- dm %>% group_by(p_holder) %>%
  summarise(cv = sd(revenue)/mean(revenue), log_specDiv = mean(log(specDiv)),
    # summarise(cv = mean(revenue), log_specDiv = mean(log(specDiv)),
    length = mean(length, na.rm = TRUE), days = mean(days, na.rm = TRUE),
    strategy = paste(sort(unique(substr(strategy, 1, 1))), collapse = " ")) %>%
  as_data_frame() %>%
  group_by(strategy) %>%
  mutate(n_strategy = n()) %>%
  filter(n_strategy > 5) %>%
  mutate(mean_log_spec = mean(log_specDiv)) %>%
  as_data_frame() %>%
  mutate(scaled_log_days = scale(log(days + 1)), scaled_log_spec = scale(log_specDiv))
dmv$strategy <- as.factor(dmv$strategy)
```

This is how much the last step filtered out.

```{r}
length(unique(dm$p_holder))
nrow(dmv)
```

We'll fit a model with `lmer()`. Random intercepts and random slopes for diversity group by strategy. I've also now added a quadratic term for species diversity.

```{r}
m4 <- lmer(log(cv) ~ scaled_log_spec + scaled_log_days + I(scaled_log_spec^2) +
    (1 + scaled_log_spec+I(scaled_log_spec^2)|strategy), data = dmv)
co <- coef(m4)$strategy
co$strategy <- row.names(co)
names(co) <- c("intercept", "log_diversity", "log_days", "log_diversity_squared", "strategy")
arm::display(m4)
```

Because number of days fished is related to species diversity CV decreases both with increases in species diversity and with increases in number of days fished. Therefore for the purposes of visualization I will predict at a few levels of mean species diversity.

```{r}
newdata <- data.frame(select(dmv, -scaled_log_days), scaled_log_days = 0)
newdata <- dmv %>%
  mutate(diversity_cut = cut(log_specDiv, breaks = c(-1, 0.25, 0.5, 0.75, 1e6))) %>%
  group_by(strategy, diversity_cut) %>%
  mutate(scaled_log_days = mean(scaled_log_days))
```

```{r, fig.width=10, fig.height=9}
dmv$prediction <- predict(m4, newdata = newdata)
dmv %>%
  arrange(strategy, log_specDiv) %>%
  ggplot(aes(scaled_log_spec, log(cv), color = scaled_log_days)) + facet_wrap(~strategy) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = FALSE, color = "grey50") +
  geom_line(aes(y = prediction), color = "orange", lwd = 1)
```

```{r}
dmv %>%
  arrange(strategy, log_specDiv) %>%
  ggplot(aes(scaled_log_spec, log(cv), color = scaled_log_days, group = strategy)) +
  geom_point(alpha = 0.7) +
  geom_line(aes(y = prediction), color = "orange", lwd = 1)
```

Let's get an idea of what a group level predictors would look like. In other words what the effect is like across strategies.

```{r}
co <- co %>%
  mutate(n_strategy = nchar(gsub(" ", "", strategy))) %>%
  inner_join(select(dmv, strategy, mean_log_spec) %>% unique())
ggplot(co, aes(mean_log_spec, log_diversity, label = strategy)) + geom_text() +
  ylab("diversity slope random effect") +
  geom_smooth()
ggplot(co, aes(mean_log_spec, intercept, label = strategy)) + geom_text() +
  ylab("intercept random effect") +
  geom_smooth()
ggplot(co, aes(mean_log_spec, log_diversity_squared, label = strategy)) + geom_text() +
  ylab("log_diversity_squared random effect") +
  geom_smooth()
```

```{r, fig.width=10, fig.height=6}
strategy_grouped = rbind(cbind(co[grep("M",co$strategy),], group="Groundfish"),
  cbind(co[grep("S",co$strategy),], group="Salmon"),
  cbind(co[grep("B|C",co$strategy),], group="Halibut-Sablefish"),
  cbind(co[grep("K|T|D",co$strategy),], group="Crab"),
  cbind(co[grep("G",co$strategy),], group="Herring"))
ggplot(strategy_grouped, aes(mean_log_spec, intercept, label = strategy)) +
  geom_text() + geom_smooth(se = FALSE) +
  facet_wrap(~group, scales = "free_y")
ggplot(strategy_grouped, aes(mean_log_spec, log_diversity, label = strategy)) +
  geom_text() + geom_smooth(se = FALSE) +
   ylab("diversity slope random effect") +
  facet_wrap(~group, scales = "free_y")
ggplot(strategy_grouped, aes(mean_log_spec, log_diversity_squared, label = strategy)) +
  geom_text() + geom_smooth(se = FALSE) +
   ylab("log_diversity_squared random effect") +
  facet_wrap(~group, scales = "free_y")
```

### 2016-07-01 portfolio/analysis/6b-tmb-modeling_diff.R
Add back diff modeling code (with additional plots)
Was accidentally deleted in an earlier commit

@@ -1,0 +1,360 @@
+# Experiment with models
+# by fitting them quickly with TMB
+rm(list = ls())
+library(TMB)
+library(dplyr)
+library(ggplot2)
+library(lme4)
+compile("revenue_diff.cpp")
+dyn.load(dynlib("revenue_diff"))
+
+# LOAD DATA
+cfecAnnual.diff = readRDS(file="../data-generated/cfec-diff-for-modeling.rds")
+
+# 4200 different strategies, need to model only most common,
+top.strategies = names(rev(sort(table(cfecAnnual.diff$strategy)))[1:100])
+cfecAnnual.diff = cfecAnnual.diff[cfecAnnual.diff$strategy%in%top.strategies, ]
+# restrict analysis to people who don't change strategies (keeps ~ 90%)
+testData = cfecAnnual.diff[which(cfecAnnual.diff$strategy == cfecAnnual.diff$strategy.prev), ]
+
+p <- cfecAnnual.diff %>%
+  ggplot(aes(log(specDiv/specdiv.prev),
+    log(revenue/revenue.prev),
+    color = specDiv)) +
+  geom_point(alpha = 0.2) +
+  geom_smooth(method = "lm", se = FALSE) +
+  facet_wrap(~strategy,scales = "free_y") +
+  ggsave("../figs/rev-diff-vs-spec-diff-strategy.pdf",width = 33, height = 24)
+
+p <- cfecAnnual.diff %>%
+  group_by(p_holder) %>%
+  summarize(mean_spec_div_diff = mean(log(specDiv/specdiv.prev)),
+    sd_revenue_diff = sd(log(revenue/revenue.prev)),
+    strategy = strategy[1]) %>%
+  ggplot(aes(mean_spec_div_diff,
+    sd_revenue_diff,
+    color = mean_spec_div_diff)) +
+  geom_point(alpha = 0.2) +
+  geom_smooth(method = "lm", se = FALSE)# +
+  # facet_wrap(~strategy,scales = "free_y")
+print(p)
+ggsave("../figs/rev-diff-sd-vs-spec-diff-strategy.pdf",width = 33, height = 24)
+
+p <- cfecAnnual.diff %>%
+  group_by(strategy) %>%
+  summarize(mean_spec_div_diff = mean(log(specDiv/specdiv.prev)),
+    sd_revenue_diff = sd(log(revenue/revenue.prev))) %>%
+  ggplot(aes(mean_spec_div_diff,
+    sd_revenue_diff,
[…]

### 2016-07-01 portfolio/text/supp.tex
Fix latex mistake

@@
 years. We experimented with different versions of the offset and when included
-in the model as a predictor, the coefficient is estimated to be ~ 1, so this is
+in the model as a predictor, the coefficient is estimated to be $\sim$ 1, so this is
 a reasonable assumption. The offset is calculated as $\log\left( \frac { { day

### 2016-07-01 portfolio/analysis/7-separate-models.Rmd
Rework separate model analysis and notes

@@
 ---
-title: "Dive into separate models"
+title: "Analysis of variability"
 output: html_document
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, cache = TRUE, autodep = TRUE, warning = FALSE)
 ```
@@
 
-```{r}
+```{r load-data}
 library(dplyr)
@@
 library(lme4)
-rm(list = ls())
 dm <- readRDS("../data-generated/cfec-annual-for-modeling.rds")
@@
 
-```{r}
+```{r filter-data}
 dm <- group_by(dm, p_holder) %>%
@@
   as_data_frame() %>%
-  filter(nyr >= 10) %>%
+  filter(nyr >= 5) %>%
   filter(range_div > 0)
@@
 
-```{r}
+```{r condense-data-by-permit-holder}
 dmv <- dm %>% group_by(p_holder) %>%
@@
     # summarise(cv = mean(revenue), log_specDiv = mean(log(specDiv)),
-    length = mean(length, na.rm = TRUE), days = mean(days, na.rm = TRUE),
+    length = mean(log(length+1), na.rm = TRUE), days = mean(days, na.rm = TRUE),
     strategy = paste(sort(unique(substr(strategy, 1, 1))), collapse = " ")) %>%
@@
   mutate(n_strategy = n()) %>%
-  filter(n_strategy > 5) %>%
+  filter(n_strategy > 4) %>%
   mutate(mean_log_spec = mean(log_specDiv)) %>%
   as_data_frame() %>%
-  mutate(scaled_log_days = scale(log(days + 1)), scaled_log_spec = scale(log_specDiv))
+  mutate(scaled_log_days = scale(log(days + 1)), scaled_log_spec = scale(log_specDiv),
+    scaled_log_length = scale(length))
 dmv$strategy <- as.factor(dmv$strategy)
@@
 
-```{r}
+```{r check-n}
 length(unique(dm$p_holder))
 nrow(dmv)
+# dmv <- na.omit(dmv) # some permit holders are lacking vessel length 
+# nrow(dmv)
 ```
 
-We'll fit a model with `lmer()`. Random intercepts and random slopes for diversity group by strategy. I've also now added a quadratic term for species diversity.
+Note that we are losing the most permit holders by needing to have data for vessel length.
 
-```{r}
-m4 <- lmer(log(cv) ~ scaled_log_spec + scaled_log_days + I(scaled_log_spec^2) +
-    (1 + scaled_log_spec+I(scaled_log_spec^2)|strategy), data = dmv)
+We'll fit a model with `lmer()`.
+
+```{r fit-model}
+# m3 <- lmer(log(cv) ~ scaled_log_spec + scaled_log_days + scale(I(log_specDiv^2)) +
+#     (1 + scaled_log_spec+scale(I(log_specDiv^2))|strategy), data = dmv)
+
+# m3 <- lmer(log(cv) ~ scaled_log_spec + scaled_log_days + scaled_log_length +
+#     (1 + (scaled_log_spec + scaled_log_days + scaled_log_length)|strategy), data = dmv)
+
+    # scale(I(log_specDiv^2)) +
+
+# m4 <- lmer(log(cv) ~ (scaled_log_spec+scaled_log_days+scaled_log_length)^2 + 
+#     (1 + (scaled_log_spec+scaled_log_days+scaled_log_length)^2|strategy), data = dmv)
+
+m4 <- lmer(log(cv) ~ (scaled_log_spec+scaled_log_days)^2 + 
+    (1 + (scaled_log_spec+scaled_log_days)^2|strategy), data = dmv)
+```
+
+```{r extract-coefficients}
 co <- coef(m4)$strategy
 co$strategy <- row.names(co)
-names(co) <- c("intercept", "log_diversity", "log_days", "log_diversity_squared", "strategy")
-arm::display(m4)
+# names(co)[1:4] <- c("intercept", "log_diversity", "log_days", "strategy")
+# arm::display(m4)
+# summary(m4)
 ```
 
-Because number of days fished is related to species diversity CV decreases both with increases in species diversity and with increases in number of days fished.
-Therefore for the purposes of visualization I will predict at a few levels of mean species diversity.
+Let's look at the residuals in a few ways.
 
-```{r}
-newdata <- data.frame(select(dmv, -scaled_log_days), scaled_log_days = 0)
-newdata <- dmv %>%
-  mutate(diversity_cut = cut(log_specDiv, breaks = c(-1, 0.25, 0.5, 0.75, 1e6))) %>%
-  group_by(strategy, diversity_cut) %>%
-  mutate(scaled_log_days = mean(scaled_log_days))
+```{r check-residuals-groups, fig.width=12, fig.height=9}
+dmv$residuals <- residuals(m4)
+dmv %>%
+  arrange(strategy, log_specDiv) %>%
+  ggplot(aes(scaled_log_spec, residuals, color = scaled_log_days)) + facet_wrap(~strategy) +
+  geom_point(alpha = 0.7) +
+  geom_smooth(se = FALSE, color = "grey50")
 ```
 
-```{r, fig.width=10, fig.height=9}
-dmv$prediction <- predict(m4, newdata = newdata)
+These look pretty good with the exception that we might be missing a quadratic effect of species diversity for the strategies involving salmon (`S`).
+See the curve downward towards the right in the panels that include salmon, especially the salmon-only panel.
+
+```{r check-residuals}
 dmv %>%
   arrange(strategy, log_specDiv) %>%
-  ggplot(aes(scaled_log_spec, log(cv), color = scaled_log_days)) + facet_wrap(~strategy) +
-  geom_point(alpha = 0.7) +
-  geom_smooth(method = "lm", se = FALSE, color = "grey50") +
-  geom_line(aes(y = prediction), color = "orange", lwd = 1)
+  ggplot(aes(scaled_log_spec, residuals, color = scaled_log_days)) + 
+  geom_point(alpha = 0.1) +
+  geom_smooth(se = FALSE, color = "grey50")
 ```
 
-```{r}
+```{r check-residuals-fitted, fig.width=12, fig.height=9}
+dmv$fitted <- fitted(m4)
 dmv %>%
-  arrange(strategy, log_specDiv) %>%
-  ggplot(aes(scaled_log_spec, log(cv), color = scaled_log_days, group = strategy)) +
+  ggplot(aes(fitted, residuals, color = scaled_log_spec)) + facet_wrap(~strategy) +
   geom_point(alpha = 0.7) +
-  geom_line(aes(y = prediction), color = "orange", lwd = 1)
+  geom_smooth(se = FALSE, color = "grey50")
 ```
 
-Let's get an idea of what a group level predictors would look like. In other words what the effect is like across strategies.
+These look pretty good with the exception of the `G S` (herring-salmon) and `S` (salmon-only) where we are under predicting CV in some cases where there is actually high CV.
 
-```{r}
-co <- co %>%
-  mutate(n_strategy = nchar(gsub(" ", "", strategy))) %>%
-  inner_join(select(dmv, strategy, mean_log_spec) %>% unique())
-ggplot(co, aes(mean_log_spec, log_diversity, label = strategy)) + geom_text() +
-  ylab("diversity slope random effect") +
-  geom_smooth()
-ggplot(co, aes(mean_log_spec, intercept, label = strategy)) + geom_text() +
-  ylab("intercept random effect") +
-  geom_smooth()
-ggplot(co, aes(mean_log_spec, log_diversity_squared, label = strategy)) + geom_text() +
-  ylab("log_diversity_squared random effect") +
-  geom_smooth()
+Across permit holder effects:
+
+```{r coefficient-plot}
+broom::tidy(m4, conf.int = TRUE) %>%
+  filter(term != "(Intercept)") %>%
+  filter(!grepl("cor_", term)) %>%
+  mutate(conf.low = ifelse(is.na(std.error), estimate, conf.low)) %>%
+  mutate(conf.high = ifelse(is.na(std.error), estimate, conf.high)) %>%
+  ggplot(aes(y = estimate, ymax = conf.low, ymin = conf.high, x = term)) +
+  geom_pointrange() +
+  coord_flip() +
+  geom_hline(yintercept = 0, lty = 2)
 ```
 
-```{r, fig.width=10, fig.height=6}
-strategy_grouped = rbind(cbind(co[grep("M",co$strategy),], group="Groundfish"),
-  cbind(co[grep("S",co$strategy),], group="Salmon"),
-  cbind(co[grep("B|C",co$strategy),], group="Halibut-Sablefish"),
-  cbind(co[grep("K|T|D",co$strategy),], group="Crab"),
-  cbind(co[grep("G",co$strategy),], group="Herring"))
-ggplot(strategy_grouped, aes(mean_log_spec, intercept, label = strategy)) +
-  geom_text() + geom_smooth(se = FALSE) +
-  facet_wrap(~group, scales = "free_y")
-ggplot(strategy_grouped, aes(mean_log_spec, log_diversity, label = strategy)) +
-  geom_text() + geom_smooth(se = FALSE) +
-   ylab("diversity slope random effect") +
-  facet_wrap(~group, scales = "free_y")
-ggplot(strategy_grouped, aes(mean_log_spec, log_diversity_squared, label = strategy)) +
-  geom_text() + geom_smooth(se = FALSE) +
-   ylab("log_diversity_squared random effect") +
-  facet_wrap(~group, scales = "free_y")
+What I see:
+
+- increased diversity = lower variability (somewhat quadratically for those including salmon in their strategy)
+- longer fishing seasons = lower variability
+- benefit of diversifying is much greater for those with longer fishing seasons
+
+Across strategy effects:
+
+```{r group-level-predictors, fig.width=10, fig.height=8}
+co <- mutate(co, salmon = grepl("S", strategy))
+
+co_strategy_long <- tidyr::gather(co, parameter, estimate, -salmon, -strategy) %>%
+  inner_join(select(dmv, strategy, mean_log_spec) %>% unique())
+
+ggplot(co_strategy_long, aes(mean_log_spec, estimate, label = strategy, color = salmon)) +
+  geom_text(size = 3) +
+  facet_wrap(~parameter, scales = "free_x") +
+  geom_hline(yintercept = 0, lty = 2) +
+  geom_smooth(method = "lm", se = FALSE)
 ```
+
+What I see:
+
+- strategies with higher mean diversity may have very slightly higher CV (but see below where that isn't as clear across most strategy groups)
+- strategies with higher mean diversity have more benefit to diversifying if salmon is included in the strategy (this remains consistent across the groups in the below figure)
+
+```{r group-level-predictors-grouped, fig.width=14, fig.height=7}
+strategy_grouped <- rbind(cbind(co_strategy_long[grep("M",co_strategy_long$strategy),], group="Groundfish"),
+  cbind(co_strategy_long[grep("S",co_strategy_long$strategy),], group="Salmon"),
+  cbind(co_strategy_long[grep("B|C",co_strategy_long$strategy),], group="Halibut-Sablefish"),
+  cbind(co_strategy_long[grep("K|T|D",co_strategy_long$strategy),], group="Crab"),
+  cbind(co_strategy_long[grep("G",co_strategy_long$strategy),], group="Herring"))
+
+strategy_grouped <- filter(strategy_grouped, 
+  parameter %in% c("(Intercept)", "scaled_log_days", "scaled_log_spec"))
+ggplot(strategy_grouped, aes(mean_log_spec, estimate, label = strategy)) +
+  geom_text() + geom_smooth(se = FALSE, method = "lm") +
+  geom_hline(yintercept = 0, lty = 2) +
+  facet_grid(parameter~group, scales = "free_y")
+
+```

### 2016-07-11 portfolio/analysis/3-createData_linear_modeling.Rmd
cleaning
Changed all the files to meet the new data requirements, and added the
lmer() models

--- 2016-06-29 portfolio/analysis/4a-createData_linear_modeling.Rmd
@@
 if (Sys.info()[["user"]] != "seananderson") {
-  region<-read.csv("areas/regions.csv")
+  region<-read.csv("data/regions.csv")
   region<-region[,c(1:2)] #don’t include “assigned by” column
@@
 
-## Filtering out unique vessel - person combinations. Drops rows from 17 to 13 million.
-
-Now we'll filter out (i.e. keep) individuals who only used 1 boat in a year, and boats with
-only one permit holder fishing in a year.
+## Previously, we excluded people using 2 boats in a year, but now we'll include them.
 
 ```{r}
+# Now we'll filter out (i.e. keep) individuals who only used 1 boat in a year, and boats with
+# only one permit holder fishing in a year.
 cfecnew = cfec
-cfec.1boat = group_by(cfecnew, p_holder, year) %>%
-  summarize(nBoat = length(unique(cadfg))) %>%
-  filter(nBoat==1) %>%
-  left_join(cfecnew)
+#cfec.1boat = group_by(cfecnew, p_holder, year) %>%
+#  summarize(nBoat = length(unique(cadfg))) %>%
+#  filter(nBoat==1) %>%
+#  left_join(cfecnew)
+#cfecnew = cfec.1boat
 
-cfecnew = cfec.1boat
 dim(cfecnew)[1]
@@
 
-# fixme What is cadfg2?
 if ("VLength" %in% names(cfecnew)) {
@@
 cfecAnnual = group_by(cfecnew, p_holder, year) %>%
-  summarize(permit = p_fshy[1], boat = cadfg[1], # fixme this was cadfg2
+  summarize(permit = p_fshy[1], boat = cadfg[1],
     days = diff(range(yday) + 1),
@@
 cfecAnnual$days.prev = cfecAnnual$days[idx]
-
-# idx = which(cfecAnnual$strategy != cfecAnnual$strategy.prev)
 ```

### 2016-07-11 portfolio/analysis/0-read-and-save.R

@@ -1,7 +1,7 @@
 # Read in, cleanup, and write out a version of the data
 
 library(dplyr)
-load("../../data/cfecCleaned_new")
+load("../../data/cfec_070616")
 names(cfec) <- tolower(names(cfec))
 cfec <- as_data_frame(cfec)
 names(cfec) <- gsub("\\.", "_", names(cfec))

### 2016-07-11 portfolio/analysis/6-lmer_models.R

@@ -1,0 +1,292 @@
+# Experiment with models 
+# by fitting them quickly with TMB
+
+# 2 separate but more complicated models. More interactions:
+# - lmer() model to full dataset: interactions allowed to vary?
+# - (1) model for mean, (2) model for variance (log sd)
+# random effects in strategy? 
+# days as a covariate?
+# 1. What's benefit for individual person
+# 2. Benefit across people within a strategy [does this change with div]
+# 3. benefit across strategies? [does this change with div]
+
+rm(list = ls())
+library(dplyr)
+library(ggplot2)
+library(lme4)
+library(MuMIn)
+
+#####################################################################################
+# UNTRANSFORMED DATA. With threshold of 5 yrs / person, about the same size as
+# differenced dataset. Most pasted from Sean's "7-Separate-models.Rmd" files
+#####################################################################################
+dat = readRDS(file="data/cfec-annual-for-modeling.rds")
+dat <- group_by(dat, p_holder) %>%
+  mutate(nyr = length(unique(year))) %>%
+  mutate(range_div = diff(range(specDiv))) %>%
+  as_data_frame() %>%
+  filter(nyr >= 5) %>%
+  filter(range_div > 0)
+nrow(dat)
+
+dat$log_spec_div <- scale(log(dat$specDiv))
+dat$log_length <- scale(log(dat$length + 1))
+dat$log_weight <- scale(log(dat$weight + 1))
+dat$log_days <- scale(log(dat$days + 1))
+
+# Filters: remove people making < 5000 / year, and 
+dat = dat[which(dat$revenue >= 5000), ]
+
+# 4200 different strategies, need to model only most common, 
+top.strategies = names(rev(sort(table(dat$strategy)))[1:100])
+dat = dat[dat$strategy%in%top.strategies, ]
+nrow(dat)
+
+# LMER models here -- conditional R^2 ~ 0.8
+# I also played with including length, but some of that is already in strategy intercepts
+mod <- lmer(log(revenue) ~ log_spec_div + log_days + as.factor(npermit) + 
+    (1 + log_spec_div + log_days|strategy), data = dat)
+r.squaredGLMM(mod)
+
[…]

### 2016-07-11 portfolio/analysis/6c-residuals_using_lmer.R

@@ -1,279 +1,0 @@
-# Experiment with models
-# by fitting them quickly with TMB
-
-# 2 separate but more complicated models. More interactions:
-# - lmer() model to full dataset: interactions allowed to vary?
-# - (1) model for mean, (2) model for variance (log sd)
-# random effects in strategy?
-# days as a covariate?
-# 1. What's benefit for individual person
-# 2. Benefit across people within a strategy [does this change with div]
-# 3. benefit across strategies? [does this change with div]
-
-rm(list = ls())
-library(dplyr)
-library(ggplot2)
-library(lme4)
-library(MuMIn)
-
-# LOAD DATA
-#cfecAnnual.diff = readRDS(file="diff_linearModeling_complete.rds")
-if (file.exists("diff_linearModeling_complete_all.rds")) {
-  cfecAnnual.diff = readRDS(file="diff_linearModeling_complete_all.rds")
-} else {
-  cfecAnnual.diff = readRDS(file="../data-generated/cfec-diff-for-modeling.rds")
-}
-
-# Calculate some differenced/derived variables
-cfecAnnual.diff$specDiv.pctChange = log(cfecAnnual.diff$specDiv / cfecAnnual.diff$specdiv.prev)
-cfecAnnual.diff$days.pctChange = log((cfecAnnual.diff$days+1) / (cfecAnnual.diff$days.prev+1))
-cfecAnnual.diff$rev.pctChange = log(cfecAnnual.diff$revenue / cfecAnnual.diff$revenue.prev)
-
-# restrict analysis to people who don't change strategies (keeps ~ 90%)
-cfecAnnual.diff = cfecAnnual.diff[which(cfecAnnual.diff$strategy == cfecAnnual.diff$strategy.prev), ]
-
-# also filter out people making < 1000 / year
-cfecAnnual.diff = cfecAnnual.diff[which(cfecAnnual.diff$revenue >= 5000), ]
-
-# 4200 different strategies, need to model only most common,
-top.strategies = names(rev(sort(table(cfecAnnual.diff$strategy)))[1:100])
-cfecAnnual.diff = cfecAnnual.diff[cfecAnnual.diff$strategy%in%top.strategies, ]
-testData = cfecAnnual.diff
-
-# Convert skewed data to something more like normal
-fit.t = MASS::fitdistr(testData$rev.pctChange, "t")
-testData$rev.pctChange.raw = testData$rev.pctChange
[…]

### 2016-07-12 portfolio/analysis/3-createData_linear_modeling.Rmd
Switch strategies to species groupings

@@
 
+```{r}
+codes <- readr::read_csv("../../data/cfec-codes.csv")
+# convert 6-dig p_fshy codes to 5 digits
+# removing blank space - now they're all length(5)
+idx = which(nchar(codes$p_fshy)==6)
+codes$p_fshy[idx] = paste(substr(codes$p_fshy[idx], 1, 1),
+  substr(codes$p_fshy[idx], 3, 6), sep="")
+
+# remove white space in 2nd position
+idx = which(substr(codes$p_fshy, 2, 2) == " ")
+codes$p_fshy[idx] = paste(substr(codes$p_fshy[idx], 1, 1),
+  substr(codes$p_fshy[idx], 3, length(codes$p_fshy[idx])), sep="")
+
+codes$description <- gsub(", VESSEL OVER 80", " VESSEL OVER 80", codes$description)
+codes$description <- gsub(";97m", "", codes$description)
+codes$description <- gsub(", VESSEL PERMIT OVER 80", " VESSEL PERMIT OVER 80", codes$description)
+codes$description <- gsub(", VESSEL PERMIT TO 80", " VESSEL PERMIT TO 80", codes$description)
+reg <- "([A-Za-z0-9/&(). -]+),([A-Za-z0-9()%/&.' -]+),([A-Za-z0-9/&.' ,()-]+)"
+codes$pfshy_species <- tolower(gsub(reg, "\\1", codes$description))
+codes$pfshy_gear <- tolower(gsub(reg, "\\2", codes$description))
+codes$pfshy_area <- tolower(gsub(reg, "\\3", codes$description))
+codes$pfshy_species <- sub(" +$", "", codes$pfshy_species)
+readr::write_csv(codes, path = "../data-generated/cfec-codes.csv")
+# as.data.frame(codes)
+```
 ## Identify most popular strategies by person-year combination
@@
 Any person in a year will need to have 10% or more of their revenue to come
-from each permit (otherwise doesn't count as a strategy).
-
-The top 120 permits cuts the rows down to 8.6 million. The top 50 gets it down
-to 7.8 million.
+from each strategy (otherwise doesn't count as a strategy).
 
 ```{r}
-permit_rev = group_by(cfecnew, p_holder, year, p_fshy) %>%
+cfecnew$p_spec = substr(cfecnew$p_fshy,1,1)
+cfecnew$p_spec[cfecnew$p_spec == "B"] = "Halibut"
+cfecnew$p_spec[cfecnew$p_spec == "C"] = "Sablefish"
+cfecnew$p_spec[cfecnew$p_spec %in% c("D","E","K","T")] = "Crab"
+cfecnew$p_spec[cfecnew$p_spec %in% c("F")] = "Freshwater"
+cfecnew$p_spec[cfecnew$p_spec %in% c("G","H","L")] = "Herring"
+cfecnew$p_spec[cfecnew$p_spec %in% c("I")] = "Lingcod"
+cfecnew$p_spec[cfecnew$p_spec %in% c("A","J","N","R","W")] = "Shellfish"
+cfecnew$p_spec[cfecnew$p_spec %in% c("O","P","Q","U","Z")] = "Invert"
+cfecnew$p_spec[cfecnew$p_spec %in% c("M")] = "Misc"
+cfecnew$p_spec[cfecnew$p_spec %in% c("S")] = "Salmon"
+cfecnew$p_spec[cfecnew$p_spec %in% c("Y")] = "Rockfish"
+cfecnew <- filter(cfecnew, p_spec != "Freshwater") # only 1400 rows
+rev(sort(table(cfecnew$p_spec)))
+
+permit_rev = group_by(cfecnew, p_holder, year, p_spec) %>%
   summarize(permit.rev = sum(g_earn)) %>%
-  as.data.frame() %>%
+  as_data_frame() %>%
   group_by(p_holder, year) %>%
@@
 topstrategies = group_by(cfecnew, p_holder, year) %>%
-  summarize(strategy = paste(sort(unique(p_fshy)), collapse = " ")) %>%
-  as.data.frame()
+  summarize(strategy = paste(sort(unique(p_spec)), collapse = ", ")) %>%
+  as_data_frame()
+rev(sort(table(topstrategies$strategy)))[1:50]
 cfecnew = left_join(cfecnew, topstrategies) # adds strategy column

### 2016-07-13 portfolio/analysis/3-createData_linear_modeling.Rmd
Filter based on strategy column

@@
 # Delete these minimal earnings
-cfecnew = cfecnew[-which(paste(cfecnew$p_holder, cfecnew$year, cfecnew$p_fshy) %in% paste(permit_rev$p_holder, permit_rev$year, permit_rev$p_fshy)), ]
+nrow(cfecnew)
+cfecnew = cfecnew[-which(paste(cfecnew$p_holder, cfecnew$year, cfecnew$p_spec) %in% 
+  paste(permit_rev$p_holder, permit_rev$year, permit_rev$p_spec)), ]
+nrow(cfecnew)
 

### 2016-07-13 portfolio/analysis/6-lmer_models.R
Include random effect standard deviations in coefficient plots

@@ -133,6 +133,16 @@ plot_coefficients_tmb <- function(model) {
   ci <- as_data_frame(ci)  %>% mutate(term = term)
   ci$estimate <- fixef(model)[[1]]
   ci <- filter(ci, !grepl("Intercept", term))
+
+  re_s <- attr(summary(model)$varcor$cond$strategy, "stddev")
+  re_p <- attr(summary(model)$varcor$cond$p_holder, "stddev")
+  re <- data.frame(term = c(paste0("re.sd.strategy.", names(re_s)), 
+      paste0("re.sd.p_holder.", names(re_p))))
+  re$estimate <- c(re_s, re_p)
+  re$`2.5 %` <- re$estimate
+  re$`97.5 %` <- re$estimate
+
+  ci <- bind_rows(ci, re)
   ggplot(ci, aes(y = estimate, ymax = `2.5 %`, ymin = `97.5 %`, x = term)) +
     geom_pointrange() + coord_flip() + geom_hline(yintercept = 0, lty = 2)
 }

### 2016-07-14 portfolio/analysis/3-createData_linear_modeling.Rmd
Add gear strategies to data

@@
 
+Calculate gear based strategies c:
+
+```{r}
+codes <- mutate(codes, p_gear = substr(codes$p_fshy, 2, 3))
+gears <- select(codes, p_gear, pfshy_gear) %>% distinct()
+cfecnew$p_gear = substr(cfecnew$p_fshy,2,3)
+gear_freq <- sort(table(cfecnew$p_gear)) %>% as_data_frame()  %>% 
+  rename(p_gear = Var1)
+gears <- left_join(gear_freq, gears)
+trim <- function (x) gsub("^\\s+|\\s+$", "", x)
+gears <- mutate(gears, pfshy_gear = trim(pfshy_gear))
+gears <- group_by(gears, p_gear) %>% 
+  summarize(n = n[1], pfshy_gear = paste(pfshy_gear, collapse = "-")) %>% 
+    as_data_frame() %>% 
+      arrange(-n)
+gears %>% readr::write_csv(path = "../data-generated/gear-strategies.csv")
+
+# hand edited with groups:
+gears <- readr::read_csv("../../data/gear-strategies.csv")
+# adds gear strategy column:
+cfecnew <- left_join(cfecnew, select(gears, p_gear, gear_group))
+```
+
 Aggregate annual metrics by person,

### 2016-07-14 portfolio/analysis/3-createData_linear_modeling.Rmd
Cull minimal earnings based on p_fshy (as was originally done)

@@
 
-permit_rev = group_by(cfecnew, p_holder, year, p_spec) %>%
+permit_rev = group_by(cfecnew, p_holder, year, p_fshy) %>%
   summarize(permit.rev = sum(g_earn)) %>%
@@
 nrow(cfecnew)
-cfecnew = cfecnew[-which(paste(cfecnew$p_holder, cfecnew$year, cfecnew$p_spec) %in% 
-  paste(permit_rev$p_holder, permit_rev$year, permit_rev$p_spec)), ]
+cfecnew = cfecnew[-which(paste(cfecnew$p_holder, cfecnew$year, cfecnew$p_fshy) %in% 
+  paste(permit_rev$p_holder, permit_rev$year, permit_rev$p_fshy)), ]
 nrow(cfecnew)

### 2016-07-14 portfolio/analysis/3-createData_linear_modeling.Rmd
Fixed typo
Also calculate unique days

@@
 
-Calculate gear based strategies c:
+Calculate gear based strategies:
 
@@
     days = diff(range(yday) + 1),
+    days_n = sum(unique(yday)),
     ngear = length(unique(gearn)),

### 2016-07-14 portfolio/analysis/3-createData_linear_modeling.Rmd
Group the gear strategies by permit holder - year

@@
 gears <- readr::read_csv("../../data/gear-strategies.csv")
-# adds gear strategy column:
+gears <- mutate(gears, p_gear = 
+  ifelse(nchar(p_gear) == 1, paste0("0", p_gear), p_gear))
+# adds gear group column:
+cfecnew$gear_group <- NULL
 cfecnew <- left_join(cfecnew, select(gears, p_gear, gear_group))
+
+topstrategies_gear <- group_by(cfecnew, p_holder, year) %>%
+  summarize(strategy_gear = paste(sort(unique(gear_group)), collapse = ", ")) %>%
+  as_data_frame()
+rev(sort(table(topstrategies_gear$strategy_gear)))[1:50]
+
+cfecnew <- left_join(cfecnew, topstrategies_gear) # adds gear strategy column
+
+# xx <- filter(topstrategies_gear, strategy_gear == "")
+# filter(cfecnew, p_holder == 138, year == 1990) %>% as.data.frame()
 ```

### 2016-07-14 portfolio/analysis/3-createData_linear_modeling.Rmd
Fix 2 mistakes with gear strategies

@@
     days = diff(range(yday) + 1),
-    days_n = sum(unique(yday)),
+    days_n = length(unique(yday)),
     ngear = length(unique(gearn)),
@@
     length = mean(vlength, na.rm=TRUE),
-    strategy = strategy[1])
+    strategy = strategy[1],
+    strategy = strategy_gear[1])
 

### 2016-07-14 portfolio/analysis/3-createData_linear_modeling.Rmd
Stop combining some vessel lengths

@@
 ```{r}
-cfecnew$p_fshy[cfecnew$p_fshy=="B61B"] = "B06B"
-cfecnew$p_fshy[cfecnew$p_fshy=="C61B"] = "C06B"
-cfecnew$p_fshy[cfecnew$p_fshy=="C91B"] = "C09B"
-cfecnew$p_fshy[cfecnew$p_fshy=="M61B"] = "M06B"
-cfecnew$p_fshy[cfecnew$p_fshy=="E09B"] = "E91B"
-cfecnew$p_fshy[cfecnew$p_fshy=="I06B"] = "I61B"
-cfecnew$p_fshy[cfecnew$p_fshy=="Y06A"] = "Y61A"
-cfecnew$p_fshy[cfecnew$p_fshy=="O06B"] = "O61B"
-cfecnew$p_fshy[cfecnew$p_fshy=="O09B"] = "O91B"
-
-cfecnew$p_fshy[cfecnew$p_fshy%in%c("M7FB","M7GB", "M7HB","M7IB")] = "M7IB"
-cfecnew$p_fshy[cfecnew$p_fshy%in%c("M7FG","M7GG", "M7HG","M7IG")] = "M7IG"
-
-# crab and shrimp vessels
-cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="D91")] = paste("D09", substr(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="D91")],4,length(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="D91")])), sep="")
-cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="T91")] = paste("T09", substr(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="T91")],4,length(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="T91")])), sep="")
-cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="P91")] = paste("P09", substr(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="P91")],4,length(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="P91")])), sep="")
-
-cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="M91")] = paste("M09", substr(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="M91")],4,length(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="M91")])), sep="")
-cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="M61")] = paste("M06", substr(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="M61")],4,length(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="M61")])), sep="")
+# cfecnew$p_fshy[cfecnew$p_fshy=="B61B"] = "B06B"
+# cfecnew$p_fshy[cfecnew$p_fshy=="C61B"] = "C06B"
+# cfecnew$p_fshy[cfecnew$p_fshy=="C91B"] = "C09B"
+# cfecnew$p_fshy[cfecnew$p_fshy=="M61B"] = "M06B"
+# cfecnew$p_fshy[cfecnew$p_fshy=="E09B"] = "E91B"
+# cfecnew$p_fshy[cfecnew$p_fshy=="I06B"] = "I61B"
+# cfecnew$p_fshy[cfecnew$p_fshy=="Y06A"] = "Y61A"
+# cfecnew$p_fshy[cfecnew$p_fshy=="O06B"] = "O61B"
+# cfecnew$p_fshy[cfecnew$p_fshy=="O09B"] = "O91B"
+# 
+# cfecnew$p_fshy[cfecnew$p_fshy%in%c("M7FB","M7GB", "M7HB","M7IB")] = "M7IB"
+# cfecnew$p_fshy[cfecnew$p_fshy%in%c("M7FG","M7GG", "M7HG","M7IG")] = "M7IG"
+# 
+# # crab and shrimp vessels
+# cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="D91")] = paste("D09", substr(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="D91")],4,length(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="D91")])), sep="")
+# cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="T91")] = paste("T09", substr(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="T91")],4,length(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="T91")])), sep="")
+# cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="P91")] = paste("P09", substr(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="P91")],4,length(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="P91")])), sep="")
+# 
+# cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="M91")] = paste("M09", substr(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="M91")],4,length(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="M91")])), sep="")
+# cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="M61")] = paste("M06", substr(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="M61")],4,length(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="M61")])), sep="")
 ```

### 2016-07-14 portfolio/analysis/3-createData_linear_modeling.Rmd
Add missing line

@@
 ```
+
 ## Identify most popular strategies by person-year combination

### 2016-07-14 portfolio/analysis/3-createData_linear_modeling.Rmd
Divide some gears by taxonomic groups

@@
 
+cfecnew$gear_group[cfecnew$p_spec == "Salmon" & 
+  cfecnew$gear_group == "drift gillnet"] <- "drift gillnet salmon"
+cfecnew$gear_group[cfecnew$p_spec == "Salmon" & 
+  cfecnew$gear_group == "purse seine"] <- "purse seine salmon"
+cfecnew$gear_group[(cfecnew$p_spec == "Sablefish" | cfecnew$p_spec == "Halibut") & 
+  cfecnew$gear_group == "pot gear"] <- "pot gear sablefish/halibut"
+cfecnew$gear_group[(cfecnew$p_spec == "Sablefish" | cfecnew$p_spec == "Halibut") & 
+  cfecnew$gear_group == "longline under 60"] <- "longline under 60 sablefish/halibut"
+cfecnew$gear_group[(cfecnew$p_spec == "Sablefish" | cfecnew$p_spec == "Halibut") & 
+  cfecnew$gear_group == "longline over 60"] <- "longline over 60 sablefish/halibut"
+cfecnew$gear_group[(cfecnew$p_spec == "Sablefish" | cfecnew$p_spec == "Halibut") & 
+  cfecnew$gear_group == "longline over 90"] <- "longline over 90 sablefish/halibut"
+
 topstrategies_gear <- group_by(cfecnew, p_holder, year) %>%

### 2016-07-14 portfolio/analysis/3-createData_linear_modeling.Rmd
Miscellaneous fixes

@@
 # hand edited with groups:
-gears <- readr::read_csv("../../data/gear-strategies.csv")
+gears <- readr::read_csv("../../data/gear-strategies-edited.csv")
 gears <- mutate(gears, p_gear = 
   ifelse(nchar(p_gear) == 1, paste0("0", p_gear), p_gear))
+
 # adds gear group column:
@@
     strategy = strategy[1],
-    strategy = strategy_gear[1])
+    strategy_gear = strategy_gear[1])
 

### 2016-07-27 portfolio/analysis/3-createData_linear_modeling.Rmd
added new calc for days

@@
 }
+
+# calculate 
+permit_days = group_by(cfecnew, p_holder, year, p_fshy) %>%
+  summarize(permitDays = diff(range(yday)+1)) %>%
+  group_by(p_holder, year) %>% 
+  summarize(tot.permitDays = sum(permitDays))
+
 cfecAnnual = group_by(cfecnew, p_holder, year) %>%
@@
 
-# also calculate weighted simpson diversity of gears and species
+cfecAnnual = left_join(cfecAnnual, permit_days)
+# also calculate weighted simpson diversity of species
 cfecAnnual = left_join(cfecAnnual, group_by(cfecnew, p_holder, year, specn) %>%
@@
 
-# also calculate weighted simpson diversity of gears and species
+# also calculate weighted simpson diversity of gears
 cfecAnnual = left_join(cfecAnnual, group_by(cfecnew, p_holder, year, gearn) %>%

### 2016-07-27 portfolio/analysis/3-createData_linear_modeling.Rmd
updated files

@@
 
+topstrategies.bypermit = group_by(cfecnew, p_holder, year) %>%
+  summarize(strategy.permit = paste(sort(unique(p_fshy)), collapse = " ")) %>%
+  as_data_frame()
+cfecnew = left_join(cfecnew, topstrategies.bypermit) # adds strategy.permit column
+
 write.csv(rev(sort(table(topstrategies$strategy))), file = "../data-generated/strategies.csv")
@@
 ```{r}
-
 cfecnew$yday <- paste(cfecnew$year, cfecnew$day, sep = "-") %>%
@@
   summarize(tot.permitDays = sum(permitDays))
+cfecnew = left_join(cfecnew, permit_days)
 
@@
     days_n = length(unique(yday)),
+    days_permit = tot.permitDays[1], 
     ngear = length(unique(gearn)),
@@
     strategy = strategy[1],
-    strategy_gear = strategy_gear[1])
+    strategy_gear = strategy_gear[1],
+    strategy_permit = strategy.permit[1])
 
@@
 cfecAnnual$strategy.prev = cfecAnnual$strategy[idx]
+cfecAnnual$strategy_gear.prev = cfecAnnual$strategy_gear[idx]
+cfecAnnual$strategy_permit.prev = cfecAnnual$strategy_permit[idx]
 cfecAnnual$weight.prev = cfecAnnual$weight[idx]

### 2016-07-27 portfolio/analysis/6-lmer_models.R

@@ -29,6 +29,8 @@ dat <- group_by(dat, p_holder) %>%
 # Filters: remove people-year combinations making < $5000
 dat = dat[which(dat$revenue >= 5000), ]
 
+# note: grouping here is based on strategies defined by permits
+dat$strategy = dat$strategy_permit
 dat <- group_by(dat, strategy) %>%
   mutate(npeople = length(unique(p_holder)))
 
@@ -63,7 +65,12 @@ top_permits$new[which(substr(top_permits$new,1,3)=="T09")] = "T09" # tanner crab
 # remaining, and only 2 can be combined: M26B/G (mechanical jig, GOA/statewide)
 top_permits$new[which(substr(top_permits$new,1,3)=="M26")] = "M26" # mechanical jig
 
-# 6. Combining salmon permits is a little more difficult, because the species composition
+# 6. Combine some of the longline categories based on vessel size.
+top_permits$new[top_permits$new=="C06B"] = "C61B"
+top_permits$new[top_permits$new=="B06B"] = "B61B"
+top_permits$new[top_permits$new=="S05B"] = "S15B"
+
+# 7. Combining salmon permits is a little more difficult, because the species composition
 # varies widely based on geography. For the drift gillnet (S03) permits, we
 # combined S03 permits from Cook Inlet, Bristol Bay, and Alaska Peninsula because sockeye
 # represents > 92% of species revenue
@@ -79,7 +86,7 @@ top_permits$new[which(top_permits$new%in%c("S04E","S04H","S04K","S04M","S04T"))]
 top_permits$new[which(top_permits$new%in%c("S04W","S04Z"))] = "S04b"
 # There's a few remaining permits (S04D, S04X, S04P, S04Y) - but they're so different they can't
 # be grouped
-#7. combine the new permit groupings into new strategies
+#8. combine the new permit groupings into new strategies
 top_permits$orig = as.character(top_permits$orig)
 top_strategies$new.strategy = NA
 for(i in 1:nrow(top_strategies)) {
@@ -91,8 +98,16 @@ dat = left_join(dat, top_strategies) %>% select(-c(n, earn))
 dat$strategy=dat$new.strategy
 dat = dat[is.na(dat$strategy)==FALSE,]
 
[…]

### 2016-07-27 portfolio/analysis/3-createData_linear_modeling.Rmd
added lookup table for species percentages by strategy
largely for interpretation

@@
 cfecnew = left_join(cfecnew, topstrategies) # adds strategy column
+write.csv(rev(sort(table(topstrategies$strategy))), file = "../data-generated/strategies.csv")
 
@@
 
-write.csv(rev(sort(table(topstrategies$strategy))), file = "../data-generated/strategies.csv")
+# Summarize key species landings by strategy.permit -- this is for interpretation later
+key.spec = c("SOCK","COHO","CHUM","PINK","CHNK","PCOD","PHLB","SABL","TCRB","KCRB","DCRB","PLCK","SAC","POUND")
+species_by_strategy = group_by(cfecnew, strategy.permit, spec) %>%
+  summarize(earn.bySpec = sum(g_earn)) %>% 
+  data.frame %>% 
+  group_by(strategy.permit) %>% 
+  mutate(totEarn = sum(earn.bySpec)) %>% 
+  filter(spec %in% key.spec)
+species_by_strategy$earn.bySpec = species_by_strategy$earn.bySpec/species_by_strategy$totEarn
+species_by_strategy = filter(species_by_strategy, earn.bySpec > 0.1) %>% select(-totEarn)
+# convert to wide format
+library(reshape2)
+melted_sbs <- melt(species_by_strategy, id.vars = c("strategy.permit", "spec"))
+species_by_strategy <- dcast(melted_sbs, strategy.permit ~ spec + variable)
+write.csv(species_by_strategy, file = "../data-generated/species_by_strategy.csv")
+
 ```

### 2016-07-27 portfolio/analysis/6-lmer_models.R

@@ -147,9 +147,8 @@ nrow(dat)
 library(glmmTMB)
 
 mod <- glmmTMB(log(revenue) ~ scaled_spec_div*log_days +
-  I(log_days^2) + I(scaled_spec_div^2) + (1 + scaled_spec_div + log_days + I(scaled_spec_div^2)|strategy),
+  I(log_days^2) * I(scaled_spec_div^2) + (1 + scaled_spec_div + log_days|strategy),
     data = dat)
-
 summary(mod)
 AIC(mod)
 

### 2016-07-27 portfolio/analysis/3-createData_linear_modeling.Rmd
re-ran data creation file
included previous days

@@
 cfecAnnual$days.prev = cfecAnnual$days[idx]
+cfecAnnual$days_permit.prev = cfecAnnual$days_permit[idx]
 ```

### 2016-07-27 portfolio/analysis/6-lmer_models_diff.R

@@ -1,0 +1,400 @@
+# Experiment with models
+# by fitting them quickly with lme4
+
+# 2 separate but more complicated models. More interactions:
+# - lmer() model to full dataset: interactions allowed to vary?
+# - (1) model for mean, (2) model for variance (log sd)
+# random effects in strategy?
+# days as a covariate?
+# 1. What's benefit for individual person
+# 2. Benefit across people within a strategy [does this change with div]
+# 3. benefit across strategies? [does this change with div]
+
+rm(list = ls())
+library(dplyr)
+library(ggplot2)
+library(lme4)
+library(MuMIn)
+
+dat = readRDS(file="portfolio/data-generated/cfec-diff-for-modeling.rds")
+
+### Eric's culling:
+#1. We restricted our analysis to p_holders who fished for 5 or more years
+dat <- group_by(dat, p_holder) %>%
+  mutate(nyr = length(unique(year))) %>%
+  mutate(range_div = diff(range(specDiv))) %>%
+  as_data_frame() %>%
+  filter(nyr >= 5)
+
+# Filters: remove people-year combinations making < $5000
+dat = dat[which(dat$revenue >= 5000), ]
+
+# note: grouping here is based on strategies defined by permits
+dat$strategy = dat$strategy_permit
+dat <- group_by(dat, strategy) %>%
+  mutate(npeople = length(unique(p_holder)))
+
+#2. For each person-year combination, we created 'strategies' by concatenating
+# all permits fished, and only retaining 'strategies' with >= 200 data points. Previously,
+# this was done by people-year, but Eric changed this to be including > 200 people / strategy
+top_strategies = group_by(dat, strategy) %>%
+  summarize(n = length(unique(p_holder)),
+    earn=sum(revenue)) %>%
+  filter(n>=100) # note -- the code should work fine with this >= 100 too
+
+#3. We then tabulated the permits that make up these strategies, and there are only
+# 56.
+top_permits = data.frame("orig"=names(table(unlist(lapply(top_strategies$strategy, strsplit, " ")))))
+
+# 4. of these top permits, we proceeded to group permits that were targeting a single species
[…]

### 2016-07-27 portfolio/figs/portfolio-risk-return.tex
Add old TeX file

\documentclass[12pt]{article}
\usepackage{geometry}
\geometry{letterpaper}
\usepackage{graphicx}

\author{Sean Anderson}
\title{Risk returns plots broken out by dimension}
% \date{}

\begin{document}
\maketitle

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=6in]{portfolio-gross-earnings-cv.pdf}
\caption{The original plot from  before.}
\label{portfolio-gross-earnings-cv.pdf}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=6in]{portfolio-gross-earnings-semideviation.pdf}
\caption{This version shows downside semideviation.}
\label{portfolio-semideviation-gross-earnings.pdf}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=6in]{portfolio-gross-earnings-semideviation-upside.pdf}
\caption{This version shows only upside semideviation.}
\label{portfolio-gross-earnings-semideviation-upside.pdf}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=6.5in]{portfolio-gross-earnings-cv-vessel-length.pdf}
\caption{These are broken out by mean vessel length as shown in the header of
each box. The panel NA represents permit holders that are associated with
multiple vessel lengths.}
\label{portfolio-gross-earnings-cv-vessel-length.pdf}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=6in]{portfolio-gross-earnings-cv-pollock.pdf}
\caption{The panel
marked TRUE should be permit holders that ever finished pollock.}
\label{port}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=6.5in]{portfolio-gross-earnings-cv-taxa.pdf}
\caption{Split up in broad taxonomic categories. NA represents permit holders
that don't fall into just one of these categories.}
\label{portfolio-gross-earnings-cv-taxa.pdf}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=6.5in]{portfolio-gross-earnings-cv-time-window.pdf}
\caption{The original plot broken out into a ten year
moving window. The year indicated in each panel represents the starting year.}
\label{portfolio-gross-earnings-cv-time-window.pdf}
\end{center}
\end{figure}

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=6in]{portfolio-gross-earnings-cv-salmon.pdf}
\caption{The right panel shows salmon permit holders and the left panel shows
everything else.}
\label{portfolio-gross-earnings-cv-salmon.pdf}
\end{center}
\end{figure}

\end{document}

### 2016-07-30 portfolio/analysis/3-createData_linear_modeling.Rmd
messing around with additional output / vars

--- 2016-07-27 portfolio/analysis/3-createData_linear_modeling.Rmd
@@
     strategy_gear = strategy_gear[1],
-    strategy_permit = strategy.permit[1])
+    strategy_permit = strategy.permit[1],
+    chnk = sum(g_earn[spec=="CHNK"]),
+    coho = sum(g_earn[spec=="COHO"]),
+    pink = sum(g_earn[spec=="PINK"]),
+    sock = sum(g_earn[spec=="SOCK"]),
+    chum = sum(g_earn[spec=="CHUM"]))
 
@@
 cfecAnnual$days.prev = cfecAnnual$days[idx]
+cfecAnnual$days_permit.prev = cfecAnnual$days_permit[idx]
+cfecAnnual$chnk.prev = cfecAnnual$chnk[idx]
+cfecAnnual$coho.prev = cfecAnnual$coho[idx]
+cfecAnnual$chum.prev = cfecAnnual$chum[idx]
+cfecAnnual$pink.prev = cfecAnnual$pink[idx]
+cfecAnnual$sock.prev = cfecAnnual$sock[idx]
 ```

### 2016-08-03 portfolio/analysis/3-createData_linear_modeling.Rmd
directories

@@
 
-if (Sys.info()[["user"]] == "seananderson") {
-  cfec <- feather::read_feather("../data-generated/cfec.feather")
-} else {
-  cfec <- feather::read_feather("data/cfec.feather")
-}
+cfec <- feather::read_feather("portfolio/data-generated/cfec.feather")
 
@@
 
-if (Sys.info()[["user"]] != "seananderson") {
-  region<-read.csv("data/regions.csv")
-  region<-region[,c(1:2)] #don’t include “assigned by” column
-  cfec<-left_join(cfec,region)
-}
-```
-
-We need to adjust the landed values and prices for inflation. To do this,
-we'll use the data Alan pulled together (adjusting to 2009 dollars).
-
-<https://docs.google.com/spreadsheets/d/1hRLYTX1nZBLKnYXwTCOtU7-XnDoMZf_1ufdTiNQFd4U/edit#gid=0>
-
-```{r, echo=FALSE}
-if (Sys.info()[["user"]] != "seananderson") {
-deflationTable = data.frame(year = 1985:2014, 
-  defl = c(0.57236, 0.58393, 0.59879, 0.61974,
-0.64388, 0.66774, 0.68993, 0.70564, 0.72244, 0.73781,
-0.75321, 0.76695, 0.78009, 0.78855, 0.80061,
-0.81883, 0.83753, 0.85038, 0.86729, 0.89114,
-0.91981, 0.94812, 0.97334, 0.9925, 1,
-1.01217, 1.03307, 1.05213, 1.06926, 1.08682))
-
-# Adjust price and g_earn for inflation
-cfec$year = as.numeric(cfec$year)
-cfec$day = substr(cfec$landdate, 6, 10)
-cfec$g_price = cfec$g_price/deflationTable$defl[cfec$year-1984]
-cfec$g_earn = cfec$g_earn/deflationTable$defl[cfec$year-1984]
-}
 ```
@@
 ```{r}
-# cfecnew$p_fshy[cfecnew$p_fshy=="B61B"] = "B06B"
-# cfecnew$p_fshy[cfecnew$p_fshy=="C61B"] = "C06B"
-# cfecnew$p_fshy[cfecnew$p_fshy=="C91B"] = "C09B"
-# cfecnew$p_fshy[cfecnew$p_fshy=="M61B"] = "M06B"
-# cfecnew$p_fshy[cfecnew$p_fshy=="E09B"] = "E91B"
-# cfecnew$p_fshy[cfecnew$p_fshy=="I06B"] = "I61B"
-# cfecnew$p_fshy[cfecnew$p_fshy=="Y06A"] = "Y61A"
-# cfecnew$p_fshy[cfecnew$p_fshy=="O06B"] = "O61B"
-# cfecnew$p_fshy[cfecnew$p_fshy=="O09B"] = "O91B"
-# 
-# cfecnew$p_fshy[cfecnew$p_fshy%in%c("M7FB","M7GB", "M7HB","M7IB")] = "M7IB"
-# cfecnew$p_fshy[cfecnew$p_fshy%in%c("M7FG","M7GG", "M7HG","M7IG")] = "M7IG"
-# 
-# # crab and shrimp vessels
-# cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="D91")] = paste("D09", substr(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="D91")],4,length(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="D91")])), sep="")
-# cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="T91")] = paste("T09", substr(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="T91")],4,length(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="T91")])), sep="")
-# cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="P91")] = paste("P09", substr(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="P91")],4,length(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="P91")])), sep="")
-# 
-# cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="M91")] = paste("M09", substr(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="M91")],4,length(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="M91")])), sep="")
-# cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="M61")] = paste("M06", substr(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="M61")],4,length(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="M61")])), sep="")
-```
-
-```{r}
-codes <- readr::read_csv("../../data/cfec-codes.csv")
+codes <- readr::read_csv("data/cfec-codes.csv")
 # convert 6-dig p_fshy codes to 5 digits
@@
 codes$pfshy_species <- sub(" +$", "", codes$pfshy_species)
-readr::write_csv(codes, path = "../data-generated/cfec-codes.csv")
+readr::write_csv(codes, path = "portfolio/data-generated/cfec-codes.csv")
 # as.data.frame(codes)
@@
 cfecnew = left_join(cfecnew, topstrategies) # adds strategy column
-write.csv(rev(sort(table(topstrategies$strategy))), file = "../data-generated/strategies.csv")
+write.csv(rev(sort(table(topstrategies$strategy))), file = "portfolio/data-generated/strategies.csv")
 
@@
 species_by_strategy <- dcast(melted_sbs, strategy.permit ~ spec + variable)
-write.csv(species_by_strategy, file = "../data-generated/species_by_strategy.csv")
+write.csv(species_by_strategy, file = "portfolio/data-generated/species_by_strategy.csv")
 
@@
       arrange(-n)
-gears %>% readr::write_csv(path = "../data-generated/gear-strategies.csv")
+gears %>% readr::write_csv(path = "portfolio/data-generated/gear-strategies.csv")
 
 # hand edited with groups:
-gears <- readr::read_csv("../../data/gear-strategies-edited.csv")
+gears <- readr::read_csv("data/gear-strategies-edited.csv")
 gears <- mutate(gears, p_gear = 
@@
 
-saveRDS(cfecAnnual, file="../data-generated/cfec-annual-for-modeling.rds")
+saveRDS(cfecAnnual, file="portfolio/data-generated/cfec-annual-for-modeling.rds")
 
@@
 cfecAnnual.diff = cfecAnnual[is.na(cfecAnnual$specDiv.change)==FALSE, ]
-saveRDS(cfecAnnual.diff, file="../data-generated/cfec-diff-for-modeling.rds")
+saveRDS(cfecAnnual.diff, file="portfolio/data-generated/cfec-diff-for-modeling.rds")
 ```
-
-## Do some exploration by permit types + diversity
-
-For example, when someone is diversifying with a herring, halibut, or crab permit, what does that diversification look like?
-
-```{r, echo=FALSE}
-cfecnew$taxa = substr(cfecnew$p_fshy, 1, 1)
-specPermit = group_by(cfecnew, taxa, specn) %>%
-  summarize(percent.revenue = sum(g_earn))
-specPermit.tot = group_by(specPermit, taxa) %>%
-  summarize(tot = sum(percent.revenue))
-specPermit = left_join(specPermit, specPermit.tot)
-specPermit$percent.revenue = specPermit$percent.revenue/specPermit$tot
-specPermit.df = group_by(specPermit, taxa, specn) %>%
-  filter(percent.revenue > 0.02) %>%
-  arrange(taxa) %>%
-  as.data.frame
-specPermit.df$spec = cfec$spec[match(specPermit.df$specn, cfec$specn)]
-kable(specPermit.df)
-```
-
-## Code for looking at permit
-We can look at whether there's clear combinations of permit types that are combinations -- most of these people either have a single local permit (PWS, Bristol Bay, etc) or have multiple statewide permits.
-
-```{r}
-
-# cfecnew$area = substr(cfecnew$p_fshy,4,4)
-# cfecnew$area[cfecnew$area %in% c("A","C")] = "Southeast"
-# cfecnew$area[cfecnew$area == "B"] = "Statewide"
-# cfecnew$area[cfecnew$area == "D"] = "Yakutat"
-# cfecnew$area[cfecnew$area == "E"] = "PWS"
-# cfecnew$area[cfecnew$area == "H"] = "CookInlet"
-# cfecnew$area[cfecnew$area == "J"] = "Westward"
-# cfecnew$area[cfecnew$area == "K"] = "Kodiak"
-# cfecnew$area[cfecnew$area == "L"] = "Chignik"
-# cfecnew$area[cfecnew$area == "M"] = "AlaskaPeninsula"
-# cfecnew$area[cfecnew$area == "N"] = "Nelson"
-# cfecnew$area[cfecnew$area == "Q"] = "BeringSea"
-# cfecnew$area[cfecnew$area == "T"] = "BristolBay"
-# cfecnew$area[cfecnew$area == "X"] = "Kotzebue"
-# cfecnew$area[cfecnew$area == "Z"] = "NortonSound"
-#
-# cfecnew$spec = substr(cfecnew$p_fshy,1,1)
-# cfecnew$spec[cfecnew$spec == "B"] = "Halibut"
-# cfecnew$spec[cfecnew$spec == "C"] = "Sablefish"
-# cfecnew$spec[cfecnew$spec %in% c("D","E","K","T")] = "Crab"
-# cfecnew$spec[cfecnew$spec %in% c("F")] = "Freshwater"
-# cfecnew$spec[cfecnew$spec %in% c("G","H","L")] = "Herring"
-# cfecnew$spec[cfecnew$spec %in% c("I")] = "Lingcod"
-# cfecnew$spec[cfecnew$spec %in% c("A","J","N","R","W")] = "Shellfish"
-# cfecnew$spec[cfecnew$spec %in% c("O","P","Q","U","Z")] = "Invert"
-# cfecnew$spec[cfecnew$spec %in% c("M")] = "Misc"
-# cfecnew$spec[cfecnew$spec %in% c("S")] = "Salmon"
-# cfecnew$area.spec = paste(cfecnew$area, cfecnew$spec, sep="_")
-#
-# cfecAnnual.multiplePermits = group_by(cfecnew, p_holder, year) %>%
-#   summarize(npermit = length(unique(p_fshy)),
-#     allpermits = paste(sort(unique(p_fshy)), collapse = " "),
-#     areaspec = paste(sort(unique(area.spec)), collapse = " ")) %>%
-#   filter(npermit > 1)
-
-```

### 2016-08-03 portfolio/analysis/0-read-and-save.R

@@ -28,9 +28,11 @@ cfec <- mutate(cfec, salmon = ifelse(spec %in%
 cfec <- mutate(cfec, pollock = ifelse(spec == "PLCK", TRUE, FALSE))
 
 region <- read.csv("data/regions.csv", stringsAsFactors = FALSE)
-region <- region[,c(1:2)] #don’t include “assigned by” column
+
+region <- region[,c("stat6","final")] #don’t include “assigned by” column
+names(region)[2] = "region"
 cfec$region <- NULL
 cfec <- left_join(cfec, region)
 
 # devtools::install_github("wesm/feather/R")
-feather::write_feather(cfec, "data-generated/cfec.feather")
+feather::write_feather(cfec, "portfolio/data-generated/cfec.feather")

### 2016-08-09 portfolio/text/ms.Rmd
Start paper skeleton

---
title: Title here
author: |
  | Authors here
date: ""
geometry: margin=1.1in
fontfamily: libertine
fontsize: 12pt
output: pdf_document
---

# Introduction

# Methods

# Results

# Discussion

# References

# Figures

\clearpage

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.6\textwidth]{../figs/stan-across-strategy-variability.pdf}
\caption{Caption here}
\label{fig:across}
\end{center}
\end{figure}

\clearpage

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=0.7\textwidth]{../figs/stan-gg-spoke2.pdf}
\caption{Caption here}
\label{fig:spokes}
\end{center}
\end{figure}

\clearpage

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=\textwidth]{../figs/offset-break-anti-spaghetti.pdf}
\caption{Caption here}
\label{fig:within}
\end{center}
\end{figure}

### 2016-08-09 portfolio/analysis/9.3-within-strategy-plot.R
Add Stan version of within strategy plot

@@ -1,0 +1,61 @@
+library(dplyr)
+library(ggplot2)
+library(viridis)
+source("portfolio/analysis/prep-stan-model-matrix.R")
+source("portfolio/analysis/grid_arrange_shared_legend.R")
+load("portfolio/data-generated/m.rda")
+devtools::load_all("pfxr")
+
+b <- broom::tidyMCMC(m, conf.int = T, estimate.method = "median",
+  conf.level = 0.8, conf.method = "quantile")
+
+g1 <- b[grepl("coef_g1", b$term), ]
+b1 <- b[grepl("coef_b1", b$term), ]
+g2 <- b[grepl("coef_g2", b$term), ]
+b2 <- b[grepl("coef_b1", b$term), ]
+
+res <- data.frame(strategy_id = 1:nrow(g1)) %>%
+  inner_join(md) %>%
+  mutate(strategy = as.character(strategy))
+res$inc <- g2$estimate
+res$dec <- g1$estimate
+res$inc_rev <- b2$estimate
+res$dec_rev <- b1$estimate
+
+ns <- dat %>% group_by(strategy) %>%
+  summarise(nn = length(unique(p_holder)))
+res <- inner_join(res, ns)
+
+p1 <- res %>%
+  mutate(strategy_label = ifelse(nn > 250, str_label, NA)) %>%
+  mutate(inc = -inc) %>%
+  ggplot(aes(x = inc_rev, y = inc)) +
+  geom_hline(yintercept = 0, lty = 2, col = "grey60") +
+  geom_vline(xintercept = 0, lty = 2, col = "grey60") +
+  geom_text_repel(aes(label = strategy_label, x = inc_rev, y = inc), nudge_y = 0.05,
+    size = 3, colour = "grey50") +
+  scale_color_viridis() +
+  geom_point(aes(color = strategy_mean_div, size = nn)) +
+  theme_gg() +
+  xlab("Effect of generalizing on revenue") +
+  ylab("Effect of generalizing on variability") +
+  labs(colour = "Mean sp.\ndiversity", size = "Number of\npermits")
+
+p2 <- res %>%
+  mutate(dec = dec, dec_rev = -dec_rev) %>%
+  mutate(strategy_label = ifelse(nn > 250, str_label, NA)) %>%
+  ggplot(aes(x = dec_rev, y = dec)) +
+  geom_hline(yintercept = 0, lty = 2, col = "grey60") +
+  geom_vline(xintercept = 0, lty = 2, col = "grey60") +
[…]

### 2016-08-11 9.3-plot-overall-ts2.R
Figures!

@@ -1,0 +1,137 @@
+# This figure is meant to illustrate increased specialization
+
+# number of permits / person
+library(dplyr)
+library(ggplot2)
+devtools::load_all("pfxr")
+
+# cfec.feather already inflation adjusted
+if (!exists("cfec")) {
+  cfec <- feather::read_feather("portfolio/data-generated/cfec.feather")
+  cfec$year <- as.numeric(cfec$year)
+  # add indicator for single permit landings
+  cfec = group_by(cfec, year, p_holder) %>%
+    mutate(n.permit = length(unique(p_fshy))) %>%
+    mutate(permits = ifelse(n.permit==1, "Single","Multiple")) %>%
+    select(-n.permit)
+}
+
+# ------------------------------------------------
+cfec <- mutate(cfec, sp_id = substr(p_fshy, 1, 1))
+es = group_by(cfec, year, p_holder, specn) %>%
+  summarize(earn = sum(g_earn), spp = paste(sp_id, collapse = ""))
+es2 = group_by(es, year, p_holder) %>%
+  summarize(effSpec = simp.div(earn), spp = paste(spp, collapse = ""))
+
+usp <- sort(table(cfec$sp_id))
+usp <- names(usp[usp > 75000])
+usp <- c("x", usp)
+
+es2$unique_spp <- unlist(lapply(lapply(
+      strsplit(es2$spp, ""), unique),
+    paste, collapse = ""))
+grepl("^S$", es2$unique_spp[1:10])
+
+out_div <- plyr::ldply(usp, function(x) {
+  message(x)
+  filter(es2, !grepl(x, unique_spp)) %>%
+  # filter(es2, !grepl(paste0("^", x, "$"), unique_spp)) %>%
+  group_by(year) %>%
+  summarize(mean_spdiv = mean(effSpec, na.rm = T), jk = x)
+})
+
+# ------------------------------------------------
+permits_person_year <- group_by(cfec, p_holder, year, sp_id) %>%
+  filter(g_earn > 0) %>%
+  summarize(npermit = length(unique(p_fshy)))
+
+permits_person_year2 <- group_by(permits_person_year, p_holder, year) %>%
+  summarize(npermit = sum(npermit), spp = paste(sp_id, collapse = ""))
+
+out_permits <- plyr::ldply(usp, function(x) {
+  message(x)
[…]

### 2016-08-11 pfxr/R/gg.R

@@ -9,9 +9,10 @@ theme_gg <- function(base_size = 11, base_family = "") {
     axis.title = element_text(colour = "grey30"),
     legend.title = element_text(colour = "grey30"),
     panel.border = element_rect(fill = NA, colour = "grey70", size = 1),
-    legend.key.size = unit(1, "lines"),
-    legend.text = element_text(size = rel(0.7)),
-    legend.title = element_text(size = rel(0.7))
+    legend.key.size = unit(0.8, "lines"),
+    legend.text = element_text(size = rel(0.7), colour = "grey30"),
+    legend.title = element_text(size = rel(0.6)),
+    legend.key = element_rect(colour = NA)
   )
 }
 
@@ -38,4 +39,20 @@ grid_arrange_shared_legend <- function(..., ncol = length(list(...)),
       ncol = 2,
       widths =  grid::unit.c(grid::unit(1, "npc") - lwidth, lwidth)))
   grid::grid.draw(combined)
+}
+
+#' ggplot2-like colour scale in HCL space
+#'
+#' @param n Number of colours to return.
+#' @param hue_min Minimum hue value in the range [0,360]
+#' @param hue_max Maximum hue value in the range [0,360]
+#' @param l Luminance in the range [0,100]
+#' @param c Chroma of the colour.
+#' @details See the \code{hcl} function for details.
+#' @examples
+#' gg_color_hue(10)
+#'
+gg_color_hue <- function(n, hue_min = 10, hue_max = 280, l = 62, c = 100) {
+  hues = seq(hue_min, hue_max, length=n+1)
+  hcl(h=hues, l=l, c=c)[1:n]
 }

### 2016-08-11 portfolio/analysis/9.1-offset-stan.R

@@ -22,7 +22,7 @@ standat <- list(
   str_yr_i = dat$str_yr_id,
 
   b1_cov_i = b1(dat$spec_change),
-  b2_cov_i = b2(dat$days_change),
+  b2_cov_i = b2(dat$spec_change),
   g1_cov_i = b1(dat$spec_change),
   g2_cov_i = b2(dat$spec_change),
 
@@ -61,8 +61,9 @@ init_fun <- function() {
 }
 
 m <- stan("portfolio/analysis/portfolio-offset.stan",
-  data = standat, iter = 90, chains = 4,
-  pars = c("mu", "sigma", "b0_str_yr"), include = FALSE, init = init_fun)
+  data = standat, iter = 400, chains = 4,
+  pars = c("mu", "sigma", "b0_str_yr"), include = FALSE) #, init = init_fun)
+save(m, file = "portfolio/data-generated/m.rda")
 b <- broom::tidy(m, conf.int = T, estimate.method = "median", rhat = T, ess = T)
 filter(b, rhat > 1.15)
 filter(b, ess < 40)

### 2016-08-11 portfolio/analysis/9.2-check-offset-stan.R

@@ -1,5 +1,6 @@
 library(dplyr)
 library(ggplot2)
+library(rstan)
 
 source("portfolio/analysis/prep-stan-model-matrix.R")
 
@@ -9,13 +10,13 @@ load("portfolio/data-generated/m.rda")
 b <- broom::tidy(m, conf.int = T, estimate.method = "median", rhat = T, ess = T)
 filter(b, rhat > 1.1)
 filter(b, ess < 100)
-filter(b, grepl("^h1", term))
-filter(b, grepl("^b0", term))
-filter(b, grepl("^g0", term))
-filter(b, grepl("^g_k", term))
+# filter(b, grepl("^h1", term))
+# filter(b, grepl("^b0", term))
+# filter(b, grepl("^g0", term))
+# filter(b, grepl("^g_k", term))
 filter(b, grepl("^b_j", term))
-filter(b, grepl("*_tau$", term))
-filter(b, grepl("^coef_*", term)) %>% select(rhat, ess)
+# filter(b, grepl("*_tau$", term))
+# filter(b, grepl("^coef_*", term)) %>% select(rhat, ess)
 
 # check mixing:
 pdf("portfolio/figs/stan-traces.pdf", width = 10, height = 8)
@@ -25,7 +26,7 @@ traceplot(m, pars = b$term[grepl("^b2_strategy\\[", b$term)], inc_warmup = F)
 traceplot(m, pars = b$term[grepl("^g1_strategy\\[", b$term)], inc_warmup = F)
 traceplot(m, pars = b$term[grepl("^g2_strategy\\[", b$term)], inc_warmup = F)
 traceplot(m, pars = b$term[grepl("^g0_strategy\\[", b$term)], inc_warmup = F)
-traceplot(m, pars = b$term[grepl("^g_k|^b_j|^b0$|^g0$", b$term)], inc_warmup = F)
+traceplot(m, pars = b$term[grepl("^g_k|^b_j|^g0$|^h1$", b$term)], inc_warmup = F)
 dev.off()
 
 # ----------------------------------------
@@ -59,7 +60,7 @@ tau_samples <- function(p) {
   hist(extract(m)[[p]], xlim = c(0, 5), main = p)
   t_prior()
 }
-normal_prior <- function(mu = 0, sigma=1) {
+normal_prior <- function(mu = 0, sigma=2) {
[…]

### 2016-08-11 portfolio/analysis/9.3-map.R

@@ -1,0 +1,12 @@
+library(maps)
+library(mapdata)
+map("worldHires",".", xlim=c(-178,-130), ylim=c(51,73), col="gray80",
+  fill=TRUE, mar = c(1,1,1,1), boundary = F, border = "grey80")
+box(col = "grey40")
+
+p <- readr::read_csv("data/map-points.csv")
+
+for (i in 1:nrow(p)) {
+  points(-1*p$long[i], p$lat[i], col = "red", pch = 19, cex = 1)
+  graphics::text(-1*p$long[i], p$lat[i], labels = p$name[i], pos = 4)
+}

### 2016-08-11 portfolio/analysis/9.3-plot-overall-ts.R

@@ -73,7 +73,14 @@ p2_dat = group_by(effSpec, year) %>%
     sd.other = sd(effSpec.nosal[is.finite(effSpec.nosal)],na.rm=T),
     lower = mean - sd/mean, upper = mean + sd/mean,
     lower.other = mean.other - sd.other/mean.other,
-    upper.other = mean.other + sd.other/mean.other)
+    upper.other = mean.other + sd.other/mean.other,
+    med = median(effSpec, na.rm = T),
+    lq = quantile(effSpec, probs = 0.25, na.rm = T),
+    uq = quantile(effSpec, probs = 0.75, na.rm = T),
+    med.other = median(effSpec.nosal[is.finite(effSpec.nosal)],na.rm=T),
+    lq.other = quantile(effSpec.nosal[is.finite(effSpec.nosal)],probs=0.25,na.rm=T),
+    uq.other = quantile(effSpec.nosal[is.finite(effSpec.nosal)],probs=0.75,na.rm=T)
+    )
 
 # --------------------------------------------
 this_fig_theme <- theme_gg() +
@@ -114,6 +121,190 @@ p2 <- p2_dat %>%
   # ggtitle("Species diversity") +
   this_fig_theme
 
-pdf("portfolio/figs/fig1.pdf", width = 6.5, height = 4.75)
-gridExtra::grid.arrange(p0a, p0b, p1, p2, ncol=2)
+# pdf("portfolio/figs/fig1.pdf", width = 6.5, height = 4.75)
+# gridExtra::grid.arrange(p0a, p0b, p1, p2, ncol=2)
+# grid_arrange_shared_legend(p0a, p0b, p1, p2, ncol = 2, nrow = 2, position = "right")
+# grid::grid.draw(rbind(cbind(ggplotGrob(p1), ggplotGrob(p2), size="last"))
+# dev.off()
+
+# par(mfrow = c(2, 2))
+pdf("portfolio/figs/fig1-map.pdf", width = 5, height = 5)
+# layout(rbind(c(1, 2), c(1, 3), c(1, 4)))
+# plot(1)
+# layout.show(4)
+par(mfrow = c(2, 2))
+par(cex = 0.6)
+par(mar = c(4, 4, 1, 1))
+par(oma = c(1, 1, 1, 1))
+library(maps)
+library(mapdata)
+x <- map("worldHires",".", xlim=c(-178,-130), ylim=c(51,73), col="gray80",
[…]

### 2016-08-11 portfolio/analysis/9.3-plot-overall-ts2.R

@@ -1,0 +1,135 @@
+# This figure is meant to illustrate increased specialization
+
+# number of permits / person
+library(dplyr)
+library(ggplot2)
+devtools::load_all("pfxr")
+
+# cfec.feather already inflation adjusted
+if (!exists("cfec")) {
+  cfec <- feather::read_feather("portfolio/data-generated/cfec.feather")
+  cfec$year <- as.numeric(cfec$year)
+  # add indicator for single permit landings
+  cfec = group_by(cfec, year, p_holder) %>%
+    mutate(n.permit = length(unique(p_fshy))) %>%
+    mutate(permits = ifelse(n.permit==1, "Single","Multiple")) %>%
+    select(-n.permit)
+}
+
+# ------------------------------------------------
+cfec <- mutate(cfec, sp_id = substr(p_fshy, 1, 1))
+es = group_by(cfec, year, p_holder, specn) %>%
+  summarize(earn = sum(g_earn), spp = paste(sp_id, collapse = ""))
+es2 = group_by(es, year, p_holder) %>%
+  summarize(effSpec = simp.div(earn), spp = paste(spp, collapse = ""))
+
+usp <- sort(table(cfec$sp_id))
+usp <- names(usp[usp > 75000])
+usp <- c("x", usp)
+
+es2$unique_spp <- unlist(lapply(lapply(
+      strsplit(es2$spp, ""), unique),
+    paste, collapse = ""))
+grepl("^S$", es2$unique_spp[1:10])
+
+out_div <- plyr::ldply(usp, function(x) {
+  message(x)
+  filter(es2, !grepl(x, unique_spp)) %>%
+  # filter(es2, !grepl(paste0("^", x, "$"), unique_spp)) %>%
+  group_by(year) %>%
+  summarize(mean_spdiv = mean(effSpec, na.rm = T), jk = x)
+})
+
+# ------------------------------------------------
+permits_person_year <- group_by(cfec, p_holder, year, sp_id) %>%
+  filter(g_earn > 0) %>%
+  summarize(npermit = length(unique(p_fshy)))
+
+permits_person_year2 <- group_by(permits_person_year, p_holder, year) %>%
+  summarize(npermit = sum(npermit), spp = paste(sp_id, collapse = ""))
+
+out_permits <- plyr::ldply(usp, function(x) {
+  message(x)
[…]

### 2016-08-11 portfolio/analysis/9.3-plot-stan-beans.R

@@ -3,6 +3,7 @@ library(dplyr)
 source("portfolio/analysis/prep-stan-model-matrix.R")
 
 load("portfolio/data-generated/m.rda")
+devtools::load_all("pfxr")
 
 # -----------------------------------------------
 # posterior beanplots
@@ -31,10 +32,40 @@ b2 <- tidyr::gather(p$coef_b2_strategy, term, posterior) %>%
   mutate(parameter = "b2")
 po <- bind_rows(g0, g1, g2, b1, b2)
 
-pl <- ggplot(po, aes(term, posterior)) + geom_violin() +
-  facet_wrap(~parameter, nrow=1) + ylab("") +
-  coord_flip() + theme_gg() +
-  theme(panel.grid.major = element_line(colour = "grey90",
-    size = 0.2), panel.grid.minor = element_line(colour = "grey98"))
-ggsave("portfolio/figs/stan-str-posteriors.pdf", width = 15, height = 9)
+md2 <- md
+md2$str_label <- as.factor(md2$str_label)
+md2$str_label <- factor(md2$str_label, levels = md2$str_label[order(md2$strategy_mean_div)])
 
+po <- rename(po, strategy = term) %>%
+  inner_join(select(md2, strategy, str_label, strategy_mean_div))
+
+# pl <- ggplot(po, aes(str_label, posterior)) + geom_violin() +
+#   facet_wrap(~parameter, nrow=1) + ylab("") +
+#   coord_flip() + theme_gg() +
+#   theme(panel.grid.major = element_line(colour = "grey90",
+#     size = 0.2), panel.grid.minor = element_line(colour = "grey98"))
+# ggsave("portfolio/figs/stan-str-posteriors.pdf", width = 15, height = 9)
+
+pos <- group_by(po, str_label, parameter) %>% summarise(
+  l = quantile(posterior, probs = 0.025),
+  l.5 = quantile(posterior, probs = 0.25),
+  m = quantile(posterior, probs = 0.5),
+  u.5 = quantile(posterior, probs = 0.75),
+  u = quantile(posterior, probs = 0.975),
+  strategy_mean_div = strategy_mean_div[1])
+
+pos$line <- 0
+pos[pos$parameter == "g0", "line"] <- NA
+pos <- as_data_frame(pos)
+
+pos$parameter <- factor(pos$parameter, levels = c("g0", "g1", "g2", "b1", "b2"))
[…]

### 2016-08-11 portfolio/analysis/9.3-within-strategy-plot.R

@@ -1,8 +1,8 @@
 library(dplyr)
 library(ggplot2)
 library(viridis)
+library(ggrepel)
 source("portfolio/analysis/prep-stan-model-matrix.R")
-source("portfolio/analysis/grid_arrange_shared_legend.R")
 load("portfolio/data-generated/m.rda")
 devtools::load_all("pfxr")
 
@@ -27,34 +27,44 @@ ns <- dat %>% group_by(strategy) %>%
 res <- inner_join(res, ns)
 
 p1 <- res %>%
-  mutate(strategy_label = ifelse(nn > 250, str_label, NA)) %>%
-  mutate(inc = -inc) %>%
+  mutate(strategy_label = ifelse(nn > 100, str_label, NA)) %>%
+  mutate(inc = inc) %>%
   ggplot(aes(x = inc_rev, y = inc)) +
-  geom_hline(yintercept = 0, lty = 2, col = "grey60") +
-  geom_vline(xintercept = 0, lty = 2, col = "grey60") +
-  geom_text_repel(aes(label = strategy_label, x = inc_rev, y = inc), nudge_y = 0.05,
-    size = 3, colour = "grey50") +
-  scale_color_viridis() +
+  geom_hline(yintercept = 0, lty = 2, col = "grey65") +
+  geom_vline(xintercept = 0, lty = 2, col = "grey65") +
+  geom_text_repel(aes(label = strategy_label, x = inc_rev, y = inc),
+    size = 2.8, colour = "grey60", segment.color = "grey80",
+    point.padding = unit(0.3, "lines"), max.iter = 6e3, segment.size = 0.3) +
   geom_point(aes(color = strategy_mean_div, size = nn)) +
-  theme_gg() +
   xlab("Effect of generalizing on revenue") +
   ylab("Effect of generalizing on variability") +
-  labs(colour = "Mean sp.\ndiversity", size = "Number of\npermits")
+  scale_color_viridis() +
+  theme_gg() +
+  guides(
+    colour = guide_legend(override.aes = list(size=3.5), order = 1),
+    size = guide_legend(order = 2, override.aes = list(pch = 21))) +
+  annotate("text", x = min(res$inc_rev), y = max(res$inc), label = "A",
+    fontface = "bold", size = 5) +
+  labs(colour = "Mean sp.\ndiversity", size = "Number of\npermits") +
+  theme(legend.title = element_text(size = rel(0.85)))
 
 p2 <- res %>%
[…]

### 2016-08-11 portfolio/analysis/9.4-plot-across-strategies.R

@@ -1,5 +1,6 @@
 library(dplyr)
 library(ggplot2)
+library(ggrepel)
 devtools::load_all("pfxr")
 source("portfolio/analysis/prep-stan-model-matrix.R")
 load("portfolio/data-generated/m.rda")
@@ -29,33 +30,34 @@ dd <- plyr::ldply(x, function(xx)
     u.5 = quantile(p$h1[[1]] * xx + p$g0[[1]], probs = 0.75)
   ))
 
-mr <- group_by(dat, strategy) %>% summarise(strategy_med_rev = median(revenue)/1e3)
+mr <- group_by(dat, strategy) %>%
+  summarise(strategy_med_rev = median(revenue)/1e3, nn = length(unique(p_holder)))
 g0$strategy_med_rev <- NULL
+g0$nn <- NULL
 g0 <- inner_join(g0, mr)
 
-pl <- ggplot(g0, aes(x = strategy_mean_div, y = exp(g0.m))) +
-
+pl <-  g0 %>% mutate(str_label = ifelse(nn > 50, str_label, NA)) %>%
+  ggplot(aes(x = strategy_mean_div, y = exp(g0.m))) +
   geom_ribbon(data = dd, aes(strategy_mean_div, y = exp(m), ymax=exp(u), ymin=exp(l)),
-    fill = "grey93") +
+    # fill = "#65B2E7", alpha = 0.3) +
+    fill = "grey85", alpha = 0.3) +
   geom_ribbon(data = dd, aes(strategy_mean_div, y = exp(m), ymax=exp(u.5), ymin=exp(l.5)),
-    fill = "grey80") +
-  geom_line(data = dd, aes(strategy_mean_div, exp(m)), lwd = 0.7) +
+    # fill = "#65B2E7", alpha = 0.8) +
+    fill = "grey80", alpha = 0.8) +
+  geom_line(data = dd, aes(strategy_mean_div, exp(m)), lwd = 0.8) +
   geom_segment(aes(y = exp(g0.l), yend = exp(g0.u), xend = strategy_mean_div),
     lwd = 0.2, color = "grey60") +
   geom_segment(aes(y = exp(g0.l.5), yend = exp(g0.u.5), xend = strategy_mean_div),
-    lwd = 0.7, color = "grey30") +
-  # geom_point(aes(size = strategy_med_rev)) +
-  # scale_size(range = c(0.5, 3)) +
-
[…]

### 2016-08-11 portfolio/analysis/9.4-spokes.R

@@ -80,30 +80,62 @@ library(ggrepel)
 
 labs <- data.frame(single_permit = c("S03", "G34", "K91", "B61B"),
   single_permit_clean =
-    c("Salmon, drift gillnet", "Herring roe, gillnet", "King crab", "Halibut"))
+    c("Salmon, drift gillnet", "bold(B)~Herring~roe~gillnet", "King crab", "bold(A)~Halibut"))
 sp <- inner_join(sp, labs)
 sp$single_permit_clean <- factor(sp$single_permit_clean,
   levels =  c(
+    "bold(A)~Halibut",
     "Salmon, drift gillnet",
-    "Herring roe, gillnet",
-    "King crab",
-    "Halibut"))
+    "bold(B)~Herring~roe~gillnet",
+    "King crab"
+    ))
+ns <- dat %>% group_by(strategy) %>%
+  summarise(nn = length(unique(p_holder)))
+sp <- inner_join(sp, ns)
 
-pl <- ggplot(sp,
+let <- data_frame(x = rep(min(sp$strategy_med_rev[sp$single_permit == "B61B"]), 2),
+  y = rep(.7, 2),
+  single_permit_clean = unique(sp$single_permit_clean)[c(3, 1)],
+  labs=LETTERS[1:2])
+
+pl <- filter(sp, !single_permit %in% c("S03", "K91")) %>%
+  mutate(str_label = ifelse(nn > 5, str_label, NA)) %>%
+  ggplot(
   aes(strategy_med_rev, exp(estimate), yend = exp(g0_less), xend = b0_less,
-    label = str_label, colour = strategy_mean_div)) +
+    label = str_label, colour = as.factor(n_permits))) +
   geom_segment(aes(x = strategy_med_rev, xend = strategy_med_rev,
-    y = exp(conf.low), yend = exp(conf.high)), alpha = 0.6, colour = "grey50") +
-  # geom_point() +
-  geom_segment() +
-  geom_text(size = 3) +
-  facet_wrap(~single_permit_clean, scales = "free") +
+    y = exp(conf.low), yend = exp(conf.high)), alpha = 0.5, size = 0.3,
+    colour = "grey60") +
+  geom_segment(colour = "grey70", alpha = 0.6, size = 0.5) +
+  geom_point(aes(size = nn)) +
+  geom_text_repel(size = 2.7,
+    point.padding = unit(0.15, "lines"), max.iter = 9e3, segment.size = 0,
+    box.padding = unit(0.15, "lines"), nudge_y = -0.03,
[…]

### 2016-08-11 portfolio/analysis/prep-stan-model-matrix.R

@@ -6,8 +6,8 @@ if (!exists("mm") | !exists("dat") | !exists("md")) {
 
   # downsample for fast testing
   unique_holders <- unique(dat$p_holder)
-  n_sample <- round(length(unique_holders)*0.50)
-  set.seed(1)
+  n_sample <- round(length(unique_holders)*0.5)
+  set.seed(121)
   dat <- dplyr::filter(dat, p_holder %in% base::sample(unique_holders, n_sample))
   nrow(dat)
 
@@ -44,7 +44,7 @@ if (!exists("mm") | !exists("dat") | !exists("md")) {
   look_str <- function(str) {
     x <- strsplit(str, " ")[[1]]
     x <- sapply(x, function(xx) labels$str_label[labels$strategy == xx])
-    paste(x, collapse = ", ")
+    paste(x, collapse = " + ")
   }
 
   md$str_label <- sapply(md$strategy, function(x) look_str(x))

### 2016-08-11 salmon/analysis/salmon_paper.Rmd
Added rmd file for salmon analysis (effect of within-permit diversification)

---
title: "Diversification of salmon fishermen within strategies (permits)"
author: ""
date: ""
output:
  pdf_document:
    fig_caption: yes
  html_document: default
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=FALSE, tidy.opts=list(width.cutoff=60), warning = FALSE, message = FALSE)
library(dplyr)
library(knitr)
library(ggplot2)
library(date)
library(ggtern)
library(reshape2)
library(mgcv)
library(fpc)
library(gridExtra)
library(glmmTMB)
```

## Intro / thoughts
- Ideally we'd compare lots of strategies across space, but most are concentrated in particular regions (herring, crab, etc).  
- Focusing people who exclusively adopt salmon as a strategy is important because of overall trends toward more people doing that (Fig. 1 in Sean's paper). For these fisheries, general trends we're seeing are (1) more people participating, (2) median revenue increasing, and (3) within-permit diversity decreasing.  
- Across strategies, we see little effect of species diversity on mean revenue or CV -- the strongest relationship might be a slight negative effect of increased species diversity on the CV. Figure \ref{fig:pairsplot} . There's certainly some strategies (permits) that are not very diverse, but have high revenue + low CVs.  
- There's a number of salmon permits that allow people to diversify.  Some don't (sockeye in Bristol Bay, Cook Inlet, etc), so we'll focus on the permits where people may be able to choose different within-permit strategies.  
- It's ok to be specialized if both abundance and prices are high. If one or the other is high, that may be ok too -- but if abundance and prices drop, things are in bad quadrant.  
- Substitution of high value species (Chinook, sockeye) for lower value species (pink,chum) has analogs in terrestrial farming (e.g. tobacco or cocoa substitution). Extremes are high-value low volume crops (e.g. cereals) vs low-volume high value (e.g. spices).  
- Diversifying has analog in terrestrial world -- specialization increases the CV and may or may not have benefit to the mean. Purdy, Barry M; Langemeier, Michael R; Featherstone, Allen M. Journal of Agricultural and Applied Economics29.1 (Jul 1, 1997): 149.  
- Hook: Specialization on certain spcecies a result of management (hatchery releases) + economics (+ marketing); reduced diversification on abundant low-value spp

## Methods

### Does diversification exist within permits and if so, do we see similar advantages of diversification on revenue + CV?

### Is there a changing benefit to diversification over time?

### How resilient are strategies to fluctuating environments (environments = fluctuations in abundance, prices, oil spill)
Key things to look at: 
- EVOS (closure for Kodiak)  
- PWS pink salmon crash in 1992/1993: Pacific Salmon & their Ecosystems: Status and Future Options

## Discussion

- Prices synchronous for coho/chum/pink. Have increased more quickly for these than sockeye and Chinook in most areas since 2002. Projecting future prices are complex -- Gunnar Knapp has a good presentation on this   http://www.adfg.alaska.gov/static/fishing/PDFs/commercial/gk_trends_4-12-12.pdf  
- We're seeing increased specialization within permits. People in PWS are trending toward increased specialization on pink salmon. People in SEAK trending toward increased specialization on chum. Both of these are switches from higher value species (sockeye, chum) to lower value species. Both pink and chum have higher variability in year to year returns than sockeye do.

\break

```{r, echo = FALSE, cache=TRUE, eval = FALSE}
# Effect of Diversification Across Space

#Using the same strategies as paper # 1, the comparisons across areas are mostly related to #salmon fisheries (these are areas with 2 or more strategies present and strategies with 2 or #more areas represented). And as in paper # 1, this is for people making >= 5K / year.

dat = readRDS(file="../../portfolio/data-generated/cfec-diff-for-modeling.rds")
### Eric's culling:
# Filters: remove people-year combinations making < $10000
dat = dat[-which(dat$nregion> 1 | dat$nregion.prev > 1 | (dat$region!=dat$region.prev)),]
dat$region[which(dat$region%in%c("SEAK Outside","SEAK Inside"))] = "SEAK"
dat = dat[dat$region%in%names(rev(sort(table(dat$region)))[1:10]), ]

dat = dat[which(dat$revenue >= 5000 & dat$revenue.prev >= 5000), ]

# note: grouping here is based on strategies defined by permits
dat$strategy = dat$strategy_permit
dat <- group_by(dat, strategy) %>%
  mutate(npeople = length(unique(p_holder)))

dat[dat$length<15 & !is.na(dat$length), ]$length <- NA

dat <- dat %>% group_by(boat) %>%
  mutate(median_length_boat = median(length, na.rm = T))

dat <- dat %>% group_by(strategy) %>%
  mutate(median_length_permit = median(length, na.rm = T))

dat <- dat %>% mutate(length = ifelse(is.na(length), median_length_boat, length))
dat <- dat %>% mutate(length = ifelse(is.na(length), median_length_permit, length))

#2. For each person-year combination, we created 'strategies' by concatenating
# all permits fished, and only retaining 'strategies' with >= 200 data points. Previously,
# this was done by people-year, but Eric changed this to be including > 200 people / strategy
top_strategies = group_by(dat, strategy) %>%
  summarize(n = length(unique(p_holder)),
    earn=sum(revenue)) %>%
  filter(n>=50) # note -- the code should work fine with this >= 100 too

# A few of these can be deleted (small in river fisheries, experimental, etc)
top_strategies = top_strategies[-which(top_strategies$strategy %in% c("S04X","S04Y","Z12B","R18B")),]

#3. We then tabulated the permits that make up these strategies, and there are only
# 56.
top_permits = data.frame("orig"=names(table(unlist(lapply(top_strategies$strategy, strsplit, " ")))))

# 4. of these top permits, we proceeded to group permits that were targeting a single species
# and other than area, were otherwise the same. In other words, someone fishing herring roe
# in kodiak (G34K) and alaska peninsula (G34M) have a herring roe strategy. This combining
# reduces permits to 59
top_permits$new = as.character(top_permits$orig)
top_permits$new[which(substr(top_permits$new,1,3)%in%c("D09","D9C","D9D"))] = "D09" # dungeness
top_permits$new[which(substr(top_permits$new,1,3)=="G01")] = "G01" # herring roe, purse seine
top_permits$new[which(substr(top_permits$new,1,3)=="G34")] = "G34" # herring roe, gillnet
top_permits$new[which(substr(top_permits$new,1,3)=="K91")] = "K91" # king crab
top_permits$new[which(substr(top_permits$new,1,3)=="L12")] = "L12" # herring, hand pick
top_permits$new[which(substr(top_permits$new,1,3)=="L21")] = "L21" # herring, pound
top_permits$new[which(substr(top_permits$new,1,3)=="P09")] = "P09" # shrimp pot gear
top_permits$new[which(substr(top_permits$new,1,3)=="Q11")] = "Q11" # sea cucumber
top_permits$new[which(substr(top_permits$new,1,3)=="T09")] = "T09" # tanner crab

# 5. Combining multispecies finfish permits (M). There are 8 miscellaneous finfish permits (M)
# remaining, and only 2 can be combined: M26B/G (mechanical jig, GOA/statewide)
top_permits$new[which(substr(top_permits$new,1,3)=="M26")] = "M26" # mechanical jig

# 6. Combine some of the longline categories based on vessel size.
top_permits$new[top_permits$new=="C06B"] = "C61B"
top_permits$new[top_permits$new=="B06B"] = "B61B"
top_permits$new[top_permits$new=="S05B"] = "S15B"

# 7. Combining salmon permits is a little more difficult, because the species composition
# varies widely based on geography. For the drift gillnet (S03) permits, we
# combined S03 permits from Cook Inlet, Bristol Bay, and Alaska Peninsula because sockeye
# represents > 92% of species revenue
top_permits$new[which(top_permits$new%in%c("S03H","S03M","S03T"))] = "S03" # drift gillnet
# For the purse seine permits (S01), again there's wide variety in species landings by
# area (85% pink in PWS to 85% sockeye on alaska peninsula). Of the 5 permits, only 2 had
# similar species compositions, so we grouped Kodiak and Cook Inlet
top_permits$new[which(top_permits$new%in%c("S01H","S01K"))] = "S01" # purse seine
# For the set gillnet, S04 permits, there's some variety but also clear sockeye specialists. We
# grouped permits that had > 80% sockeye
top_permits$new[which(top_permits$new%in%c("S04E","S04H","S04K","S04M","S04T"))] = "S04a" # purse seine
# We grouped permits from norton sound and kuskokwim because of similar species %
top_permits$new[which(top_permits$new%in%c("S04W","S04Z"))] = "S04b"
# There's a few remaining permits (S04D, S04X, S04P, S04Y) - but they're so different they can't be grouped

#8. combine the new permit groupings into new strategies
top_permits$orig = as.character(top_permits$orig)
top_strategies$new.strategy = NA
for(i in 1:nrow(top_strategies)) {
  top_strategies$new.strategy[i] = paste(top_permits$new[match(lapply(lapply(top_strategies$strategy, strsplit, " "), unlist)[[i]],
  top_permits$orig)], collapse=" ")
}
# join this into the data frame -- this is leaving ~ 70 strategies
dat = left_join(dat, top_strategies) %>% select(-c(n, earn))
dat$strategy=dat$new.strategy
dat = dat[is.na(dat$strategy)==FALSE,]

# 9. Remove data points where people only did a strategy in one year. Below
# we try to include people-strategy random effects
dat = group_by(dat, strategy, p_holder) %>%
  mutate(nsp = n()) %>% filter(nsp > 1) %>% select(-nsp)

# Derived variables
scale2 <- function(x) {
  x <- x - mean(x, na.rm = TRUE)
  x / (2 * sd(x, na.rm = TRUE))
}
dat$log_spec_div <- scale2(log(dat$specDiv))
dat$scaled_spec_div <- scale2(dat$specDiv)
dat$log_length <- scale2(log(dat$length + 1))
dat$log_weight <- scale2(log(dat$weight + 1))
dat$log_days <- scale2(log(dat$days + 1))
dat$log_npermit <- scale2(log(dat$npermit))
dat$log_days_permit <- scale2(log(dat$days_permit+1))

# delete strategy: region combinations with less than 50 observations
dat$strategy_region = paste(dat$strategy, dat$region)
dat = group_by(dat, strategy_region) %>% 
  mutate(n = n()) %>% 
  filter(n > 50) %>% select(-n)

# delete strategies in one region
dat = group_by(dat, strategy) %>% 
  mutate(n = length(unique(region))) %>% 
  filter(n > 1) %>% select(-n)

# drop regions with < 3 strategies
dat = group_by(dat, region) %>% 
  mutate(n = length(unique(strategy))) %>% 
  filter(n > 2) %>% select(-n)

dat$logdiff = log(dat$revenue/dat$revenue.prev)
dat$specdiff = log(dat$specDiv/dat$specdiv.prev)
dat$daysdiff = log((dat$days_permit+1)/(dat$days_permit.prev+1))
#mu.mod = glmmTMB(log(revenue) ~ specdiff * daysdiff + (1 + specdiff * daysdiff| strategy) + offset(log(revenue.prev)), data = dat)
m = ceiling(table(dat$strategy, dat$region)/1.0e10)
colnames(m) = c("AP", "Bering","BB","CookInlt","Kodiak","PWS","SEAK","Yakutat")
kable(m)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
dat.annual = readRDS(file="../../portfolio/data-generated/cfec-annual-for-modeling.rds")

# single permit holders, salmon permits
dat.annual  = filter(dat.annual ,npermit==1 & substr(strategy_permit,1,1)=="S") %>%
  group_by(strategy_permit) %>% 
  mutate(nPeople = length(unique(p_holder))) %>% 
  filter(nPeople > 200) %>% 
  dplyr::select(-nPeople)
# drop troll
dat.annual = dat.annual[dat.annual$strategy_permit%in%c("S05B","S15B")==FALSE,]
# drop S04X, Y, Z because low participation (< 20) in many years
# S04W deleted because low median earnings
dat.annual = dat.annual[dat.annual$strategy_permit%in%c("S04W", "S04X","S04Y","S04Z")==FALSE,]

agg.annual = group_by(dat.annual, strategy_permit, year) %>% 
  summarize(people = length(unique(p_holder)),
    specDiv = mean(specDiv),
    meanrev = median(revenue),
    low = quantile(revenue,0.25),
    upp = quantile(revenue,0.75))

dat.diff = readRDS(file="../../portfolio/data-generated/cfec-diff-for-modeling.rds")
dat.diff = dat.diff[which(dat.diff$revenue >= 5000 & dat.diff$revenue.prev >= 5000), ]
dat.diff = dat.diff[dat.diff$strategy_permit%in%c("S01A","S01E","S01K","S01L","S01M",
  "S03A","S03E","S03H","S03M","S03T","S04D","S04H","S04K","S04M","S04T"),]

agg.diff = group_by(dat.diff, strategy_permit) %>% 
  summarize(specDiv = mean(specDiv), meanrev = mean(revenue),
    cv = sd(log(revenue/revenue.prev)))

# Also load in raw CFEC data for plots
cfec <- feather::read_feather("../../portfolio/data-generated/cfec.feather")
cfec = cfec[cfec$p_fshy%in%c("S 01A", "S 01E", "S 01K", "S 03A", "S 03E", "S 01M", "S 04K"),]
cfec$year = as.numeric(substr(cfec$landdate, 1, 4))

# only use people who are using just that permit - diversification within permit
cfec.raw = cfec
dat.salmon = dat.diff[dat.diff$strategy_permit%in%c("S01A", "S01E", "S01K", "S03A", "S03E","S01M","S04K"),]

dat.salmon$days.change = log((dat.salmon$days_permit+1) / (dat.salmon$days_permit.prev+1))
dat.salmon$year_permit = as.factor(paste(dat.salmon$year, dat.salmon$strategy_permit))

cfec = cfec[which(paste(cfec$p_holder,cfec$year) %in% paste(dat.salmon$p_holder,dat.salmon$year)),]

```

\break

# Overall trends across permits

```{r, fig.cap="Participation (unique people)\\label{fig:participationplot}", fig.pos="placeHere", echo=FALSE}
ggplot(agg.annual, aes(year, people, colour = specDiv)) + geom_line() + facet_wrap(~ strategy_permit, scale="free") + ylab("Participation")
```

```{r, fig.cap="Species diversity by year\\label{fig:diversitybyyear}", fig.pos="placeHere", echo=FALSE, eval=FALSE}
#p1 = ggplot(agg, aes(year, specDiv, colour = strategy_permit)) + geom_line() + ylab("Species diversity")
ggplot(agg.annual, aes(year, specDiv, colour = specDiv)) + geom_line() + facet_wrap(~ strategy_permit, scale="free") + ylab("Species diversity")
```

```{r, fig.cap="Mean revenue by year\\label{fig:medianrevenuebyyear}", fig.pos="placeHere", echo=FALSE, eval=FALSE}
#p1 = ggplot(agg, aes(year, log(meanrev), color = strategy_permit)) + ylab("Ln median revenue") + geom_line()
ggplot(agg.annual, aes(year, log(meanrev))) + facet_wrap(~ strategy_permit, scale="free") + ylab("Ln median revenue") + geom_ribbon(aes(ymin= log(low), ymax= log(upp)), alpha=0.4) + geom_line()
```

```{r, fig.cap="Diversity, mean revenue and CV (aggregated across years)\\label{fig:pairsplot}", fig.pos="placeHere", echo=FALSE}

p1 = ggplot(agg.diff, aes(specDiv, log(meanrev), colour = cv, label=strategy_permit)) + xlab("Species diversity") +ylab("Ln (mean revenue)") + geom_text(size=2)
p2 = ggplot(agg.diff, aes(specDiv, cv, colour = log(meanrev), label=strategy_permit)) + xlab("Species diversity") +ylab("CV") + geom_text(size=2)
p3 = ggplot(agg.diff, aes(log(meanrev), cv, colour = specDiv, label=strategy_permit)) + ylab("CV") + xlab("Ln (mean revenue)") + geom_text(size=2)
gridExtra::grid.arrange(p1,p2,p3, ncol=2)
```

```{r, fig.cap="Diversity, mean revenue and CV (aggregated across years)", fig.pos="placeHere", echo=FALSE, eval=FALSE}
# We can also do the same calculations by calculating means for each year and then averaging across years, 
agg = group_by(dat, strategy_permit, year) %>% 
  summarize(specDiv = mean(specDiv), meanrev = mean(revenue),
    cv = sd(log(revenue/revenue.prev))) %>% 
  group_by(strategy_permit) %>%
  summarize(specDiv = mean(specDiv,na.rm=T), meanrev=mean(meanrev,na.rm=T),
    cv = mean(cv,na.rm=T))

p1 = ggplot(agg, aes(specDiv, log(meanrev), colour = cv, label=strategy_permit)) + xlab("Species diversity") +ylab("Ln (mean revenue)") + geom_text()
p2 = ggplot(agg, aes(log(meanrev), cv, colour = specDiv, label=strategy_permit)) + ylab("CV") + xlab("Ln (mean revenue)") + geom_text()
p3 = ggplot(agg, aes(specDiv, cv, colour = log(meanrev), label=strategy_permit)) + xlab("Species diversity") +ylab("CV") + geom_text()
gridExtra::grid.arrange(p1,p2,p3, ncol=2)
```

```{r, fig.cap="CV of revenue by year", fig.pos="placeHere", echo=FALSE, eval=FALSE}
#p1 = group_by(dat.diff, strategy_permit, year) %>% 
#  summarize(sd_logdiff = sd(log(revenue/revenue.prev))) %>% 
#  ggplot(aes(year, sd_logdiff, colour = strategy_permit)) + geom_line() + ylab("CV revenue")
p2 = group_by(dat.diff, strategy_permit, year) %>% 
  summarize(CV = sd(log(revenue/revenue.prev))) %>% 
  ggplot(aes(year, CV, colour = CV)) + geom_line() + 
  facet_wrap(~ strategy_permit, scale="free") + ylab("CV revenue")
p2
```

```{r, fig.cap="Mean revenue versus CV", fig.pos="placeHere", echo=FALSE, eval=FALSE}
group_by(dat.diff, strategy_permit, year) %>% 
  summarize(meanrev = mean(revenue), sd_logdiff = sd(log(revenue/revenue.prev))) %>% 
  ggplot(aes(log(meanrev), sd_logdiff, colour = sd_logdiff)) + 
  geom_point() + facet_wrap(~ strategy_permit, scale="free") + 
  ylab("CV revenue") + xlab("Ln mean revenue") + geom_smooth()
```

```{r, fig.cap="Species diversity versus ln (mean revenue)", fig.pos="placeHere", echo=FALSE, eval=FALSE}
group_by(dat.diff, strategy_permit, year) %>% 
  summarize(specDiv = mean(specDiv), meanrev = mean(revenue)) %>% 
  ggplot(aes(specDiv, log(meanrev), colour = meanrev)) + 
  geom_point() + facet_wrap(~ strategy_permit, scale="free") + 
  ylab("Ln (mean revenue)") + xlab("Mean species diversity") + geom_smooth()
```

```{r, fig.cap="Species diversity versus CV", fig.pos="placeHere", echo=FALSE, eval=FALSE}
group_by(dat.diff, strategy_permit, year) %>% 
  summarize(specDiv = mean(specDiv), sd_logdiff = sd(log(revenue/revenue.prev))) %>% 
  ggplot(aes(specDiv, sd_logdiff, colour = sd_logdiff)) + 
  geom_point() + facet_wrap(~ strategy_permit, scale="free") + 
  ylab("CV revenue") + xlab("Mean species diversity") + geom_smooth()
```

\break

# Can individuals diversify within strategies?

To examine the ability of individuals to diversify within strategies (permits) we can examine patterns in the following fisheries.

* S01A: Purse seine in SEAK  
* S01E: Purse seine in PWS  
* S01K: Purse seine in Kodiak  
* S01M: Purse seine on the Alaska Peninsula  
* S03A: Gillnet in SEAK  
* S03E: Gillnet in PWS  
* S03K: Gillnet in Kodiak

### For each of these permits, we can plot the proportion of species landed. There's 3 species that dominate the earnings (chum / sockeye / pink).

```{r, fig.cap="Proportion of species landed, weighted by earnings", fig.pos="placeHere", echo=FALSE}
# plot proportion of species landed by salmon permit
group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p)) %>%
  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() + facet_wrap(~p_fshy) + 
  ylab("Species % (total) ") + xlab("Year")
```

### This increased specialization on chum/pink/sockeye corresponds with a decline in diversity.

```{r, fig.cap="Effective species diversity across people", fig.pos="placeHere", echo=FALSE, fig.height = 3.5}
group_by(dat.salmon, year, strategy_permit) %>%
  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
  ggplot(aes(year, meanDiv)) + geom_line() + facet_wrap(~strategy_permit, scale="free") + 
  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Effective diversity (across people)")
```

### One way to visualize the change is with ternary plots over the 3 axes (chum / sockeye / pink)

```{r, fig.cap="Species composition for 1988 and 2014 (each dot is a person) in the S03E fishery", fig.pos="placeHere", echo=FALSE, fig.height = 2}
daty = group_by(cfec[cfec$p_fshy=="S 03E" & cfec$spec %in% c("CHUM","PINK","SOCK","COHO","CHNK"),], spec, p_holder, year) %>%
  summarize(p = sum(g_earn)) %>%
  group_by(p_holder, year) %>% mutate(p = p/sum(p))
daty <- melt(daty, id.vars = c("spec", "p_holder", "year"))
daty <- dcast(daty, p_holder + year ~ spec + variable)[,-1]
daty[is.na(daty)]=0
ggtern(data=daty[which(daty$year%in%c(1988,2014)),c("year", "CHUM_p","PINK_p","SOCK_p")], aes(x=CHUM_p,y=PINK_p,SOCK_p,color=as.factor(year))) +
  geom_point(alpha=0.3,size=1) +
  labs(x="CHUM",y="PINK",z="SOCK",title="Proportion of annual revenue (S03E)")
```

\break

# Does effect of diversification change over time?  
- The idea here could be to do something like the modeling for Sean's paper, but include time varying interactions on species diversity.

Approach here: extend portfolio model to include random effects by strategy : year. Also include random quadratic terms on 'specdiv.prev' allowing the effect of specDiv.change to change over the range of effective species fished (similar to the breakpoint model in Sean's paper).

```{r, eval=FALSE}
glmmTMB(log(revenue) ~ specDiv.change + specDiv.change:specdiv.prev + 
    specDiv.change:I(specdiv.prev^2) + days.change + 
    (1+days.change|strategy_permit) + 
    (-1+specDiv.change + specDiv.change:specdiv.prev + 
        specDiv.change:I(specdiv.prev^2)|year_permit), 
  offset = log(revenue.prev), dat.salmon)
```

```{r, echo=FALSE}

mu.mod = glmmTMB(log(revenue) ~ specDiv.change + specDiv.change:specdiv.prev + specDiv.change:I(specdiv.prev^2) + days.change + (1+days.change|strategy_permit) + (-1+specDiv.change + specDiv.change:specdiv.prev + specDiv.change:I(specdiv.prev^2)|year_permit), offset = log(revenue.prev), dat.salmon)

cv.mod = glmmTMB(log(abs(residuals(mu.mod))) ~ specDiv.change + specDiv.change:specdiv.prev + specDiv.change:I(specdiv.prev^2) + days.change + (1+days.change|strategy_permit) + (-1+specDiv.change + specDiv.change:specdiv.prev + specDiv.change:I(specdiv.prev^2)|year_permit), offset = log(revenue.prev), dat.salmon)
# to ask whether benefit of diversificaiton is changing via strategy, we want to summarize (1) direction of 
# benefit (e.g. is the effect negative or positive), and (2) the magnitude (how steep / shallow it is). With the
# quadratic model, 
# 'squared term' indicates concave up/down
# derivative is magnitude, 2a * x + b

df = ranef(mu.mod)$cond$year_permit
colnames(df) = c("c","b","a")
df$year = as.numeric(substr(rownames(df), 1, 4))
df$strategy = substr(rownames(df), 6, 9)
df$deriv = 2*df$a + df$b
# also calculate change of going from 1 to 2 spp
df$pred1 = df$a*(1^2) + df$b*1 + df$c + fixef(mu.mod)$cond["specDiv.change"] + fixef(mu.mod)$cond["specDiv.change:specdiv.prev"]*1
# add fixed effects into predictions

# We can demonstrate the quadratic curves by year for a single strategy / set of years
df.s01e = df[df$strategy=="S01E",]
demo = expand.grid("speciesDiversity" = seq(1,4,length.out=100), "year" = 1986:2014)
demo = left_join(demo, df.s01e)
demo$effect = demo$a * (demo$speciesDiversity^2) + demo$b * (demo$speciesDiversity) + demo$c
```

###We can make quadratic plots of these effects by year and strategy. These plots represent the effect of increasing diversity (as relative change) on mean revenue. For example, in 1986 if you were fishing 3-4 species there would be little effect of adding more spp; but if you were fishing 1 species, there would be a slight negative effect of adding more. In 1991, there would be a huge benefit of increasing diversity, and that effect would taper off as more were added.

```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Variation in effects of diversification by year for the S01E (PWS purse seine) permit holders. "}
ggplot(demo, aes(speciesDiversity, effect, group = year)) + geom_line() + facet_wrap(~ year) + xlab("Species Diversity") + ylab("Effect of increasing diversity (relative change)")
```

###This plot shows the time series of becoming increasingly diversified from 1 species on mean revenue. For S01E for example, there's huge benefits of diversifying in 1991, and huge costs in 2010. 2010 was a huge pink salmon year for PWS (with high prices specialization was good) and 1991 was the start of pink salmon stock + price collapse. Perhaps slight declines in some of these in recent years?

```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Time series of benefits of diversification by permit / strategy."}
ggplot(df, aes(year, pred1, group = strategy, colour = as.factor(strategy))) + geom_line() + xlab("Year") + ylab("Benefit of increasing species diversity from 1 species") + facet_wrap(~ strategy, scale="free") + geom_smooth() + geom_hline(aes(yintercept=0)) + ggtitle("Effect on Mean")
```

```{r, echo=FALSE}
df.cv = ranef(cv.mod)$cond$year_permit
colnames(df.cv) = c("c.cv","b.cv","a.cv")
df.cv$year = as.numeric(substr(rownames(df.cv), 1, 4))
df.cv$strategy = substr(rownames(df.cv), 6, 9)
# also calculate change of going from 1 to 2 spp
df.cv$pred1.cv = df.cv$a.cv*(1^2) + df.cv$b.cv*1 + df.cv$c.cv + fixef(cv.mod)$cond["specDiv.change"] + fixef(cv.mod)$cond["specDiv.change:specdiv.prev"]*1

#ggplot(df.cv, aes(year, pred1.cv, group = strategy, colour = as.factor(strategy))) + geom_line() + xlab("Year") + ylab("Benefit of increasing species diversity from 1 species") + facet_wrap(~ strategy, scale="free") + geom_smooth() + geom_hline(aes(yintercept=0)) + ggtitle("Effect on CV")

ggplot(df.cv, aes(year, pred1.cv, group = strategy, colour = as.factor(strategy))) + geom_line(linetype="dashed") + geom_line(data=df, aes(year, pred1, group = strategy, colour = as.factor(strategy))) + xlab("Year") + ylab("Effect of increasing species diversity from 1 species") + facet_wrap(~ strategy, scale="free") + ggtitle("Effects on mean (solid) and CV (dashed)") + geom_hline(aes(yintercept=0), color="grey30")

#df = left_join(df,df.cv)
#ggplot(df, aes(pred1, pred1.cv, label=year)) + facet_wrap(~strategy, scale="free")+ #geom_vline(aes(xintercept=0)) + geom_hline(aes(yintercept=0)) + xlab("Benefit on the mean") + #ylab("Benefit on the CV") + geom_text(size=2)
```

\break

# Robustness of strategies to changing environments

It may be interesting to separate out the effect of changing strategies from changing conditions. In other words, if you hold a strategy contstant, how resilient is your income to changing environment?? One approach to do this:

1. Group people into generalist and specialist categories (by permit, aggregated over all years). This will yield 2-3 groups per permit, allowing us to look at changes in these groups through time.

2. Avoid modeling people who switch strategies from one year to the next (confounded with flucuating portfolios of fish) and isolate individuals within a group that keep approximately the same species composition from one year to the next.

3.  For each group, we can then calculate average revenue and the CV by year.

4. The groupings for the strategies are roughly as follows:

S01E: Pink specialists v pink / chum  
S01A: More gradual, diversification on pink-chum gradient  
S01K: Sockeye specialists v sockeye/pink/chum  
S03A: Chum specialists v sockeye/other  
S03E: Sockeye specialists v sockeye/chinook/chum

```{r, echo = FALSE}
dat = dat.salmon
# create other category, lumping coho/chinook
dat$salm = dat$coho+dat$chnk
dat$salm.prev = dat$coho.prev+dat$chnk.prev

dat$totsalm = dat$salm + dat$chum + dat$sock + dat$pink
dat$totsalm.prev = dat$salm.prev + dat$chum.prev + dat$sock.prev + dat$pink.prev

dat$salm = dat$salm/dat$totsalm
dat$sock = dat$sock/dat$totsalm
dat$pink = dat$pink/dat$totsalm
dat$chum = dat$chum/dat$totsalm
dat$chnk = dat$chnk/dat$totsalm
dat$coho = dat$coho/dat$totsalm
dat$salm.prev = dat$salm.prev/dat$totsalm.prev
dat$sock.prev = dat$sock.prev/dat$totsalm.prev
dat$pink.prev = dat$pink.prev/dat$totsalm.prev
dat$chum.prev = dat$chum.prev/dat$totsalm.prev
dat$chnk.prev = dat$chnk.prev/dat$totsalm.prev
dat$coho.prev = dat$coho.prev/dat$totsalm.prev

dat = dat[-which(is.na(dat$salm+dat$sock+dat$pink+dat$chum)),]
dat$logdiff = log(dat$revenue/dat$revenue.prev)

dat_old = dat
dat$sumP2 = abs(dat$salm - dat$salm.prev) + 
  abs(dat$pink-dat$pink.prev) + 
  abs(dat$chum-dat$chum.prev) + 
  abs(dat$sock-dat$sock.prev)
dat = dat[which(dat$sumP2 < 0.3),]
```

```{r, echo = FALSE, fig.cap="Results for PWS Purse seine fishery", echo=FALSE, fig.pos="placeHere", cache=TRUE}

df = dat[which(dat$strategy_permit=="S01E"),]
pamk_s01e <- pamk(df[,c("chum","pink","sock","salm")])
df$group = c("Pink-chum","Pink")[pamk_s01e$pamobject$clustering]

p3 = ggtern(data=df[,c("chum","pink","sock","group")], aes(x=chum,y=pink,sock,color=as.factor(group))) +
  geom_point(alpha=0.3) +
  labs(x="CHUM",y="PINK",z="SOCK") + theme(
        axis.title=element_text(size=8)) + theme(legend.position = "none")

p0 = group_by(df, year, group) %>% 
  summarize(n = length(unique(p_holder))) %>% 
ggplot(aes(x = year, y = n, group=group, col = group)) + geom_line() + ylab("People") + ggtitle("S01E")

p1 = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T)) %>% 
ggplot(aes(x = year, y = meanrev, group=group, col = group)) + geom_line() + ylab("Mean revenue") + ggtitle("S01E")

p2 = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T)) %>% 
ggplot(aes(x = year, y = cv, group=group, col = group)) + geom_line() + ylab("CV revenue") + ggtitle("S01E")

g = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T))

gridExtra::grid.arrange(p3,p0,p1,p2,ncol=2)
```

```{r, echo = FALSE, fig.cap="SEAK purse seine fishery", cache=TRUE, echo=FALSE, fig.pos="placeHere"}

df = dat[which(dat$strategy_permit=="S01A"),]
pamk_s01a <- pamk(df[,c("chum","pink","sock","salm")])
df$group = pamk_s01a$pamobject$clustering

p3 = ggtern(data=df[,c("chum","pink","sock","group")], aes(x=chum,y=pink,sock,color=as.factor(group))) +
  geom_point(alpha=0.3) +
  labs(x="CHUM",y="PINK",z="SOCK") + theme(
        axis.title=element_text(size=8)) + theme(legend.position = "none")

p0 = group_by(df, year, group) %>% 
  summarize(n = length(unique(p_holder))) %>% 
ggplot(aes(x = year, y = n, group=group, col = as.factor(group))) + geom_line() + ylab("People") + ggtitle("S01A")

p1 = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T)) %>% 
ggplot(aes(x = year, y = meanrev, group=group, col = as.factor(group))) + geom_line() + ylab("Mean revenue") + ggtitle("S01A")

p2 = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T)) %>% 
ggplot(aes(x = year, y = cv, group=group, col = as.factor(group))) + geom_line() + ylab("CV revenue") + ggtitle("S01A")

gridExtra::grid.arrange(p3, p0, p1, p2, ncol=2)
```

```{r, echo = FALSE, fig.cap="SEAK gillnet (clustering primarily in chum-sockeye-chinokook/coho dimensions)", echo=FALSE, fig.pos="placeHere", cache=TRUE, eval=FALSE}
fishery = "S03A"
df = dat[which(dat$strategy_permit==fishery),]
pamk_s03a <- pamk(df[,c("chum","pink","sock","salm")], 3)
df$group = pamk_s03a$pamobject$clustering

p3 = ggtern(data=df[,c("chum","salm","sock","group")], aes(x=chum,y=salm,sock,color=as.factor(group))) +
  geom_point(alpha=0.3) +
  labs(x="CHUM",y="OTHER",z="SOCK") + theme(
        axis.title=element_text(size=8)) + theme(legend.position = "none")

p0 = group_by(df, year, group) %>% 
  summarize(n = length(unique(p_holder))) %>% 
ggplot(aes(x = year, y = n, group=group, col = as.factor(group))) + geom_line() + ylab("People") + ggtitle(fishery)

p1 = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T)) %>% 
ggplot(aes(x = year, y = meanrev, group=group, col = as.factor(group))) + geom_line() + ylab("Mean revenue") + ggtitle(fishery)

p2 = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T)) %>% 
ggplot(aes(x = year, y = cv, group=group, col = as.factor(group))) + geom_line() + ylab("CV revenue") + ggtitle(fishery)

gridExtra::grid.arrange(p3, p0, p1, p2, ncol=2)
```

```{r, echo = FALSE, fig.cap="SEAK gillnet fishery (clustering primarily in sockeye-chinook/coho dimension)", echo=FALSE, fig.pos="placeHere", cache=TRUE, eval=FALSE}
fishery = "S03E"
df = dat[which(dat$strategy_permit==fishery),]
pamk_s03e <- pamk(df[,c("chum","pink","sock","salm")])
df$group = pamk_s03e$pamobject$clustering

p3 = ggtern(data=df[,c("chum","salm","sock","group")], aes(x=chum,y=salm,sock,color=as.factor(group))) +
  geom_point(alpha=0.3) +
  labs(x="CHUM",y="OTHER",z="SOCK") + theme(
        axis.title=element_text(size=8)) + theme(legend.position = "none")

p0 = group_by(df, year, group) %>% 
  summarize(n = length(unique(p_holder))) %>% 
ggplot(aes(x = year, y = n, group=group, col = as.factor(group))) + geom_line() + ylab("People") + ggtitle(fishery)

p1 = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T)) %>% 
ggplot(aes(x = year, y = meanrev, group=group, col = as.factor(group))) + geom_line() + ylab("Mean revenue") + ggtitle(fishery)

p2 = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T)) %>% 
ggplot(aes(x = year, y = cv, group=group, col = as.factor(group))) + geom_line() + ylab("CV revenue") + ggtitle(fishery)

gridExtra::grid.arrange(p3, p0, p1, p2, ncol=2)
```

```{r, echo = FALSE, fig.cap="Kodiak purse seine (clustering primarily in sockeye-pink dimension)", echo=FALSE, fig.pos="placeHere", eval=FALSE, cache=TRUE}
fishery = "S01K"
df = dat[which(dat$strategy_permit==fishery),]
pamk_s01k <- pamk(df[,c("chum","pink","sock","salm")])
df$group = pamk_s01k$pamobject$clustering

p3 = ggtern(data=df[,c("chum","salm","sock","group")], aes(x=chum,y=salm,sock,color=as.factor(group))) +
  geom_point(alpha=0.3) +
  labs(x="CHUM",y="OTHER",z="SOCK") + theme(
        axis.title=element_text(size=8)) + theme(legend.position = "none")

p0 = group_by(df, year, group) %>% 
  summarize(n = length(unique(p_holder))) %>% 
ggplot(aes(x = year, y = n, group=group, col = as.factor(group))) + geom_line() + ylab("People") + ggtitle(fishery)

p1 = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T)) %>% 
ggplot(aes(x = year, y = meanrev, group=group, col = as.factor(group))) + geom_line() + ylab("Mean revenue") + ggtitle(fishery)

p2 = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T)) %>% 
ggplot(aes(x = year, y = cv, group=group, col = as.factor(group))) + geom_line() + ylab("CV revenue") + ggtitle(fishery)

gridExtra::grid.arrange(p3, p0, p1, p2, ncol=2)
```

```{r, echo = FALSE, fig.cap="Kodiak gillnet (clustering primarily in sockeye-pink dimension)", echo=FALSE, fig.pos="placeHere", cache=TRUE, eval=FALSE}
fishery = "S04K"
df = dat[which(dat$strategy_permit==fishery),]
pamk_s04k <- pamk(df[,c("chum","pink","sock","salm")])
df$group = pamk_s04k$pamobject$clustering

p3 = ggtern(data=df[,c("chum","salm","sock","group")], aes(x=chum,y=salm,sock,color=as.factor(group))) +
  geom_point(alpha=0.3) +
  labs(x="CHUM",y="OTHER",z="SOCK") + theme(
        axis.title=element_text(size=8)) + theme(legend.position = "none")

p0 = group_by(df, year, group) %>% 
  summarize(n = length(unique(p_holder))) %>% 
ggplot(aes(x = year, y = n, group=group, col = as.factor(group))) + geom_line() + ylab("People") + ggtitle(fishery)

p1 = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T)) %>% 
ggplot(aes(x = year, y = meanrev, group=group, col = as.factor(group))) + geom_line() + ylab("Mean revenue") + ggtitle(fishery)

p2 = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T)) %>% 
ggplot(aes(x = year, y = cv, group=group, col = as.factor(group))) + geom_line() + ylab("CV revenue") + ggtitle(fishery)

gridExtra::grid.arrange(p3, p0, p1, p2, ncol=2)
```

```{r, echo = FALSE, fig.cap="SEAK Purse seine (clustering primarily in sockeye-pink dimension)", echo=FALSE, fig.pos="placeHere", eval=FALSE, cache=TRUE}
fishery = "S01M"
df = dat[which(dat$strategy_permit==fishery),]
pamk_s01m <- pamk(df[,c("chum","pink","sock","salm")])
df$group = pamk_s01m$pamobject$clustering

p3 = ggtern(data=df[,c("chum","pink","sock","group")], aes(x=chum,y=pink,sock,color=as.factor(group))) +
  geom_point(alpha=0.3) +
  labs(x="CHUM",y="PINK",z="SOCK") + theme(
        axis.title=element_text(size=8)) + theme(legend.position = "none")

p0 = group_by(df, year, group) %>% 
  summarize(n = length(unique(p_holder))) %>% 
ggplot(aes(x = year, y = n, group=group, col = as.factor(group))) + geom_line() + ylab("People") + ggtitle(fishery)

p1 = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T)) %>% 
ggplot(aes(x = year, y = meanrev, group=group, col = as.factor(group))) + geom_line() + ylab("Mean revenue") + ggtitle(fishery)

p2 = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T)) %>% 
ggplot(aes(x = year, y = cv, group=group, col = as.factor(group))) + geom_line() + ylab("CV revenue") + ggtitle(fishery)

gridExtra::grid.arrange(p3, p0, p1, p2, ncol=2)
```

# Trends in revenue and prices

Across regions, we see lots of synchrony in prices between coho, chum, and pink salmon, with them increasing more quickly than sockeye/Chinook (meaning it's easier to substitute). There's a bit of diversity in patterns across areas for sockeye and Chinook. Sockeye/Chinook have been the highest priced historically, with coho > chum > pink.  For most permit types, median revenue / person has generally increased, especially since early 2000s. For some permits, variability appears to have changed through time (quartile range getting smaller).

```{r, fig.cap="Total revenue by area (across people and permits within each)", fig.pos="placeHere", echo=FALSE}
# plot prices of species landed by salmon permit
cfec.raw$area = NA
cfec.raw$area[cfec.raw$p_fshy%in%c("S 01A", "S 03A")] = "Southeast"
cfec.raw$area[cfec.raw$p_fshy%in%c("S 01E", "S 03E")] = "PWS"
cfec.raw$area[cfec.raw$p_fshy%in%c("S 01K", "S 04K")] = "Kodiak"
cfec.raw$area[cfec.raw$p_fshy%in%c("S 01M")] = "Alaska Peninsula"

group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, area) %>%
  summarize(p = sum(g_earn)) %>% group_by(year, area) %>%
  ggplot(aes(year, (p)/1e6, group=spec, color = spec)) + geom_line() + facet_wrap(~area, scale="free") +
  ylab("Revenue (millions) ")
```

```{r, fig.cap="Inflation adjusted prices / species by area, relative to 2002 (the year corresponding to the lowest average prices).", fig.pos="placeHere", echo=FALSE}
group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, area) %>%
  summarize(p = mean(g_price)) %>% 
  group_by(spec, area) %>% 
  mutate(p = p/p[which(year==2002)]) %>% 
  group_by(year, area) %>%
  ggplot(aes(year, log(p), group=spec, color = spec)) + geom_line() + facet_wrap(~area) +
  ylab("Ln avg price relative to 2002")
```

```{r, fig.cap="Number of people landing fish", fig.pos="placeHere", echo=FALSE, eval=FALSE}
#Participation (people landing fish on one of these permits) generally followed trends in prices, with an uptick since early 2000s,  
group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, p_fshy) %>%
  summarize(people = length(unique(p_holder))) %>%
  ggplot(aes(year, people)) + geom_line() + facet_wrap(~p_fshy, scale="free") + 
  ylab("People")
```

```{r, fig.cap="Trends in revenue and variability across people", fig.pos="placeHere", echo=FALSE}
#p1 = group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], p_holder, year, p_fshy) %>%
#  summarize(gearn = sum(g_earn)) %>%
#  group_by(year, p_fshy) %>%
#  summarize(cv = sd(gearn)/mean(gearn)) %>%
#  ggplot(aes(year, cv, color = p_fshy)) + geom_line() + 
#  ylab("CV revenue (across people")

p2 = group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], p_holder, year, p_fshy) %>%
  summarize(gearn = sum(g_earn)) %>%
  group_by(year, p_fshy) %>%
  summarize(med = median(gearn), low = quantile(gearn, 0.25),
    upp = quantile(gearn,0.75)) %>%
  ggplot(aes(year, log(med))) + geom_line() + facet_wrap(~p_fshy, scale="free") + 
  geom_ribbon(aes(ymin= log(low), ymax= log(upp)), alpha=0.4) + ylab("Ln quartiles of revenue (across people)")
print(p2)
```

Revenue is going up since 2000 -- but each permit holder is fishing more (more days, more total pounds) and the trend in weight is faster than the trend in price.  
```{r}
group_by(cfec, p_fshy, year, p_holder) %>% 
  summarize(meanW = mean(g_pounds/2000, na.rm=T)) %>% 
  group_by(p_fshy, year) %>% 
  summarize(meanW = mean(meanW, na.rm=T)) %>% 
  ggplot(aes(year, meanW)) + geom_line() + facet_wrap(~ p_fshy, scale="free") + ylab("Mean total tons per person") + xlab("Year")

```

```{r, fig.cap="CV of average species composition. This quantifies the synchrony across people -- so low CVs mean everyone's doing the same thing.", fig.pos="placeHere", echo=FALSE, eval=FALSE}
# Another way to think about diversification is to calculate the deviations in species composition from the mean. Bascially what's the variability of euclidian distances? Given trends in total landings and diversity, the 'portfolio' of species available to any of these permit holders has changed through time. Another way to think about the diversification is whether the variability of people around the mean has changed (basically whether there's been trends in the CV of species diveristy across people).

dat.salm = group_by(dat, year, p_holder) %>% 
  mutate(earn = (chnk+coho+chum+pink+sock),
    chnk=chnk/earn, coho=coho/earn,
    chum=chum/earn, sock=sock/earn,
    pink=pink/earn) %>% 
  as.data.frame 
dat.salm$chnk[is.na(dat.salm$chnk)] = 0
dat.salm$coho[is.na(dat.salm$coho)] = 0
dat.salm$chum[is.na(dat.salm$chum)] = 0
dat.salm$pink[is.na(dat.salm$pink)] = 0
dat.salm$sock[is.na(dat.salm$sock)] = 0

dat.salm = group_by(dat.salm, year, strategy_permit) %>% 
  mutate(mean.chnk = mean(chnk,na.rm=T), mean.chum = mean(chum,na.rm=T),
    mean.pink = mean(pink,na.rm=T), mean.coho = mean(coho,na.rm=T), 
    mean.sock = mean(sock,na.rm=T))

dat.salm$eucld = (dat.salm$mean.chnk-dat.salm$chnk)^2 + (dat.salm$mean.coho-dat.salm$coho)^2 + (dat.salm$mean.pink-dat.salm$pink)^2 + (dat.salm$mean.sock-dat.salm$sock)^2 + (dat.salm$mean.chum-dat.salm$chum)^2

p1 = group_by(dat.salm, year, strategy_permit) %>%
  summarize(sd = sd(eucld)) %>%
  ggplot(aes(year, sd, color = as.factor(strategy_permit))) + geom_line() + ylab("SD of euclidian dist.") + facet_wrap(~strategy_permit)

group_by(dat, year, strategy_permit) %>% 
  summarize(cv = sd(specDiv)/mean(specDiv)) %>% 
  ggplot(aes(year, cv, color = as.factor(strategy_permit))) + geom_line() + ylab("CV of species diversity (across people)") + facet_wrap(~ strategy_permit, scale="free") + geom_smooth()
```

### 2016-08-11 portfolio/analysis/9-offset.R
Play with adding specdiv.prev to offset model

@@ -5,7 +5,7 @@ dat <- cullDat(diff = TRUE)
 
 # downsample for speed of testing
 unique_holders <- unique(dat$p_holder)
-n_sample <- round(length(unique_holders)*0.35)
+n_sample <- round(length(unique_holders)*0.30)
 set.seed(12)
 dat <- dplyr::filter(dat, p_holder %in% base::sample(unique_holders, n_sample))
 nrow(dat)
@@ -182,6 +182,24 @@ m.b <- lmer(log(revenue) ~
 MuMIn::r.squaredGLMM(m)
 MuMIn::r.squaredGLMM(m.b)
 AIC(m) - AIC(m.b)
+
+m.c <- lmer(log(revenue) ~ #days_change +
+  b1(spec_change, bp)*log_days +
+  b2(spec_change, bp)*log_days +
+  b1(spec_change, bp)*specdiv.prev +
+  b2(spec_change, bp)*specdiv.prev +
+  (-1 + b1(spec_change, bp) + b2(spec_change, bp) | strategy) +
+  (1|strategy_year),
+  data = dat, offset = log(revenue.prev), REML = F)
+summary(m.c)
+
+m.b <- lmer(log(revenue) ~
+  spec_change*days_change +
+  spec_change*specdiv.prev +
+  (-1 + spec_change + spec_change:specdiv.prev | strategy) +
+  (1|strategy_year),
+  data = dat, offset = log(revenue.prev), REML = F)
+summary(m.b)
 
 m <- lmer(log(revenue) ~
   b1(spec_change, bp)*days_change + 

### 2016-08-11 portfolio/analysis/9.3-within-strategy-plot.R
Tweak label

@@ -45,7 +45,7 @@ p1 <- res %>%
     size = guide_legend(order = 2, override.aes = list(pch = 21))) +
   annotate("text", x = min(res$inc_rev), y = max(res$inc), label = "A",
     fontface = "bold", size = 5) +
-  labs(colour = "Mean sp.\ndiversity", size = "Number of\npermits") +
+  labs(colour = "Mean sp.\ndiversity", size = "Number of\npermit holders") +
   theme(legend.title = element_text(size = rel(0.85)))
 
 p2 <- res %>%

### 2016-08-11 portfolio/text/ms.Rmd
Update figures
Update text
Tweak fig sizes

@@
 \begin{center}
-\includegraphics[width=0.6\textwidth]{../figs/stan-across-strategy-variability.pdf}
+\includegraphics[width=0.7\textwidth]{../figs/fig1-jk.pdf}
+\caption{Caption here}
+\label{fig:ts}
+\end{center}
+\end{figure}
+
+\clearpage
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=0.9\textwidth]{../figs/stan-gg-spoke2.pdf}
+\caption{Caption here}
+\label{fig:spokes}
+\end{center}
+\end{figure}
+
+\clearpage
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=0.7\textwidth]{../figs/stan-across-strategy-variability.pdf}
 \caption{Caption here}
@@
 \begin{center}
-\includegraphics[width=0.7\textwidth]{../figs/stan-gg-spoke2.pdf}
-\caption{Caption here}
-\label{fig:spokes}
-\end{center}
-\end{figure}
-
-\clearpage
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[width=\textwidth]{../figs/offset-break-anti-spaghetti.pdf}
+\includegraphics[width=1.1\textwidth]{../figs/stan-offset-break-anti-spaghetti.pdf}
 \caption{Caption here}

### 2016-08-11 salmon/analysis/salmon_paper_orig.Rmd
Added old version for archive, cleaned junk out of new er update

--- 2016-08-11 salmon/analysis/salmon_paper.Rmd
@@
 2010 was a huge pink salmon year for PWS (with high prices specialization was good) and 1991 was the start of pink salmon stock + price collapse.
-Perhaps slight declines in some of these in recent years?
+Across permits, perhaps slight declines in some of these in recent years?
 
@@
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Time series of effects of diversification on mean and CV by permit."}
 df.cv = ranef(cv.mod)$cond$year_permit
@@
 1. Group people into generalist and specialist categories (by permit, aggregated over all years). This will yield 2-3 groups per permit, allowing us to look at changes in these groups through time.  
-
 2.
@@
 Revenue is going up since 2000 -- but each permit holder is fishing more (more days, more total pounds) and the trend in weight is faster than the trend in price.  
-```{r}
+```{r, echo=FALSE}
 group_by(cfec, p_fshy, year, p_holder) %>% 
@@
   ggplot(aes(year, meanW)) + geom_line() + facet_wrap(~ p_fshy, scale="free") + ylab("Mean total tons per person") + xlab("Year")
-
 ```

### 2016-08-11 salmon/analysis/salmon_paper.Rmd
Cleaned up figures

@@
 
-```{r, echo = FALSE, cache=TRUE, eval = FALSE}
-# Effect of Diversification Across Space
-
-#Using the same strategies as paper # 1, the comparisons across areas are mostly related to #salmon fisheries (these are areas with 2 or more strategies present and strategies with 2 or #more areas represented).
-And as in paper # 1, this is for people making >= 5K / year.
-
-dat = readRDS(file="../../portfolio/data-generated/cfec-diff-for-modeling.rds")
-### Eric's culling:
-# Filters: remove people-year combinations making < $10000
-dat = dat[-which(dat$nregion> 1 | dat$nregion.prev > 1 | (dat$region!=dat$region.prev)),]
-dat$region[which(dat$region%in%c("SEAK Outside","SEAK Inside"))] = "SEAK"
-dat = dat[dat$region%in%names(rev(sort(table(dat$region)))[1:10]), ]
-
-dat = dat[which(dat$revenue >= 5000 & dat$revenue.prev >= 5000), ]
-
-# note: grouping here is based on strategies defined by permits
-dat$strategy = dat$strategy_permit
-dat <- group_by(dat, strategy) %>%
-  mutate(npeople = length(unique(p_holder)))
-
-dat[dat$length<15 & !is.na(dat$length), ]$length <- NA
-
-dat <- dat %>% group_by(boat) %>%
-  mutate(median_length_boat = median(length, na.rm = T))
-
-dat <- dat %>% group_by(strategy) %>%
-  mutate(median_length_permit = median(length, na.rm = T))
-
-dat <- dat %>% mutate(length = ifelse(is.na(length), median_length_boat, length))
-dat <- dat %>% mutate(length = ifelse(is.na(length), median_length_permit, length))
-
-#2. For each person-year combination, we created 'strategies' by concatenating
-# all permits fished, and only retaining 'strategies' with >= 200 data points. Previously,
-# this was done by people-year, but Eric changed this to be including > 200 people / strategy
-top_strategies = group_by(dat, strategy) %>%
-  summarize(n = length(unique(p_holder)),
-    earn=sum(revenue)) %>%
-  filter(n>=50) # note -- the code should work fine with this >= 100 too
-
-# A few of these can be deleted (small in river fisheries, experimental, etc)
-top_strategies = top_strategies[-which(top_strategies$strategy %in% c("S04X","S04Y","Z12B","R18B")),]
-
-#3. We then tabulated the permits that make up these strategies, and there are only
-# 56.
-top_permits = data.frame("orig"=names(table(unlist(lapply(top_strategies$strategy, strsplit, " ")))))
-
-# 4. of these top permits, we proceeded to group permits that were targeting a single species
-# and other than area, were otherwise the same. In other words, someone fishing herring roe
-# in kodiak (G34K) and alaska peninsula (G34M) have a herring roe strategy. This combining
-# reduces permits to 59
-top_permits$new = as.character(top_permits$orig)
-top_permits$new[which(substr(top_permits$new,1,3)%in%c("D09","D9C","D9D"))] = "D09" # dungeness
-top_permits$new[which(substr(top_permits$new,1,3)=="G01")] = "G01" # herring roe, purse seine
-top_permits$new[which(substr(top_permits$new,1,3)=="G34")] = "G34" # herring roe, gillnet
-top_permits$new[which(substr(top_permits$new,1,3)=="K91")] = "K91" # king crab
-top_permits$new[which(substr(top_permits$new,1,3)=="L12")] = "L12" # herring, hand pick
-top_permits$new[which(substr(top_permits$new,1,3)=="L21")] = "L21" # herring, pound
-top_permits$new[which(substr(top_permits$new,1,3)=="P09")] = "P09" # shrimp pot gear
-top_permits$new[which(substr(top_permits$new,1,3)=="Q11")] = "Q11" # sea cucumber
-top_permits$new[which(substr(top_permits$new,1,3)=="T09")] = "T09" # tanner crab
-
-# 5. Combining multispecies finfish permits (M). There are 8 miscellaneous finfish permits (M)
-# remaining, and only 2 can be combined: M26B/G (mechanical jig, GOA/statewide)
-top_permits$new[which(substr(top_permits$new,1,3)=="M26")] = "M26" # mechanical jig
-
-# 6. Combine some of the longline categories based on vessel size.
-top_permits$new[top_permits$new=="C06B"] = "C61B"
-top_permits$new[top_permits$new=="B06B"] = "B61B"
-top_permits$new[top_permits$new=="S05B"] = "S15B"
-
-# 7. Combining salmon permits is a little more difficult, because the species composition
-# varies widely based on geography. For the drift gillnet (S03) permits, we
-# combined S03 permits from Cook Inlet, Bristol Bay, and Alaska Peninsula because sockeye
-# represents > 92% of species revenue
-top_permits$new[which(top_permits$new%in%c("S03H","S03M","S03T"))] = "S03" # drift gillnet
-# For the purse seine permits (S01), again there's wide variety in species landings by
-# area (85% pink in PWS to 85% sockeye on alaska peninsula). Of the 5 permits, only 2 had
-# similar species compositions, so we grouped Kodiak and Cook Inlet
-top_permits$new[which(top_permits$new%in%c("S01H","S01K"))] = "S01" # purse seine
-# For the set gillnet, S04 permits, there's some variety but also clear sockeye specialists. We
-# grouped permits that had > 80% sockeye
-top_permits$new[which(top_permits$new%in%c("S04E","S04H","S04K","S04M","S04T"))] = "S04a" # purse seine
-# We grouped permits from norton sound and kuskokwim because of similar species %
-top_permits$new[which(top_permits$new%in%c("S04W","S04Z"))] = "S04b"
-# There's a few remaining permits (S04D, S04X, S04P, S04Y) - but they're so different they can't be grouped
-
-#8. combine the new permit groupings into new strategies
-top_permits$orig = as.character(top_permits$orig)
-top_strategies$new.strategy = NA
-for(i in 1:nrow(top_strategies)) {
-  top_strategies$new.strategy[i] = paste(top_permits$new[match(lapply(lapply(top_strategies$strategy, strsplit, " "), unlist)[[i]],
-  top_permits$orig)], collapse=" ")
-}
-# join this into the data frame -- this is leaving ~ 70 strategies
-dat = left_join(dat, top_strategies) %>% select(-c(n, earn))
-dat$strategy=dat$new.strategy
-dat = dat[is.na(dat$strategy)==FALSE,]
-
-# 9. Remove data points where people only did a strategy in one year. Below
-# we try to include people-strategy random effects
-dat = group_by(dat, strategy, p_holder) %>%
-  mutate(nsp = n()) %>% filter(nsp > 1) %>% select(-nsp)
-
-# Derived variables
-scale2 <- function(x) {
-  x <- x - mean(x, na.rm = TRUE)
-  x / (2 * sd(x, na.rm = TRUE))
-}
-dat$log_spec_div <- scale2(log(dat$specDiv))
-dat$scaled_spec_div <- scale2(dat$specDiv)
-dat$log_length <- scale2(log(dat$length + 1))
-dat$log_weight <- scale2(log(dat$weight + 1))
-dat$log_days <- scale2(log(dat$days + 1))
-dat$log_npermit <- scale2(log(dat$npermit))
-dat$log_days_permit <- scale2(log(dat$days_permit+1))
-
-# delete strategy: region combinations with less than 50 observations
-dat$strategy_region = paste(dat$strategy, dat$region)
-dat = group_by(dat, strategy_region) %>% 
-  mutate(n = n()) %>% 
-  filter(n > 50) %>% select(-n)
-
-# delete strategies in one region
-dat = group_by(dat, strategy) %>% 
-  mutate(n = length(unique(region))) %>% 
-  filter(n > 1) %>% select(-n)
-
-# drop regions with < 3 strategies
-dat = group_by(dat, region) %>% 
-  mutate(n = length(unique(strategy))) %>% 
-  filter(n > 2) %>% select(-n)
-
-dat$logdiff = log(dat$revenue/dat$revenue.prev)
-dat$specdiff = log(dat$specDiv/dat$specdiv.prev)
-dat$daysdiff = log((dat$days_permit+1)/(dat$days_permit.prev+1))
-#mu.mod = glmmTMB(log(revenue) ~ specdiff * daysdiff + (1 + specdiff * daysdiff| strategy) + offset(log(revenue.prev)), data = dat)
-m = ceiling(table(dat$strategy, dat$region)/1.0e10)
-colnames(m) = c("AP", "Bering","BB","CookInlt","Kodiak","PWS","SEAK","Yakutat")
-kable(m)
-```
-
 ```{r, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
@@
 
-```{r, fig.cap="Species diversity by year\\label{fig:diversitybyyear}", fig.pos="placeHere", echo=FALSE, eval=FALSE}
-#p1 = ggplot(agg, aes(year, specDiv, colour = strategy_permit)) + geom_line() + ylab("Species diversity")
-ggplot(agg.annual, aes(year, specDiv, colour = specDiv)) + geom_line() + facet_wrap(~ strategy_permit, scale="free") + ylab("Species diversity")
-```
-
-```{r, fig.cap="Mean revenue by year\\label{fig:medianrevenuebyyear}", fig.pos="placeHere", echo=FALSE, eval=FALSE}
-#p1 = ggplot(agg, aes(year, log(meanrev), color = strategy_permit)) + ylab("Ln median revenue") + geom_line()
-ggplot(agg.annual, aes(year, log(meanrev))) + facet_wrap(~ strategy_permit, scale="free") + ylab("Ln median revenue") + geom_ribbon(aes(ymin= log(low), ymax= log(upp)), alpha=0.4) + geom_line()
-```
-
 ```{r, fig.cap="Diversity, mean revenue and CV (aggregated across years)\\label{fig:pairsplot}", fig.pos="placeHere", echo=FALSE}
@@
 gridExtra::grid.arrange(p1,p2,p3, ncol=2)
-```
-
-```{r, fig.cap="Diversity, mean revenue and CV (aggregated across years)", fig.pos="placeHere", echo=FALSE, eval=FALSE}
-# We can also do the same calculations by calculating means for each year and then averaging across years, 
-agg = group_by(dat, strategy_permit, year) %>% 
-  summarize(specDiv = mean(specDiv), meanrev = mean(revenue),
-    cv = sd(log(revenue/revenue.prev))) %>% 
-  group_by(strategy_permit) %>%
-  summarize(specDiv = mean(specDiv,na.rm=T), meanrev=mean(meanrev,na.rm=T),
-    cv = mean(cv,na.rm=T))
-
-p1 = ggplot(agg, aes(specDiv, log(meanrev), colour = cv, label=strategy_permit)) + xlab("Species diversity") +ylab("Ln (mean revenue)") + geom_text()
-p2 = ggplot(agg, aes(log(meanrev), cv, colour = specDiv, label=strategy_permit)) + ylab("CV") + xlab("Ln (mean revenue)") + geom_text()
-p3 = ggplot(agg, aes(specDiv, cv, colour = log(meanrev), label=strategy_permit)) + xlab("Species diversity") +ylab("CV") + geom_text()
-gridExtra::grid.arrange(p1,p2,p3, ncol=2)
-```
-
-```{r, fig.cap="CV of revenue by year", fig.pos="placeHere", echo=FALSE, eval=FALSE}
-#p1 = group_by(dat.diff, strategy_permit, year) %>% 
-#  summarize(sd_logdiff = sd(log(revenue/revenue.prev))) %>% 
-#  ggplot(aes(year, sd_logdiff, colour = strategy_permit)) + geom_line() + ylab("CV revenue")
-p2 = group_by(dat.diff, strategy_permit, year) %>% 
-  summarize(CV = sd(log(revenue/revenue.prev))) %>% 
-  ggplot(aes(year, CV, colour = CV)) + geom_line() + 
-  facet_wrap(~ strategy_permit, scale="free") + ylab("CV revenue")
-p2
-```
-
-```{r, fig.cap="Mean revenue versus CV", fig.pos="placeHere", echo=FALSE, eval=FALSE}
-group_by(dat.diff, strategy_permit, year) %>% 
-  summarize(meanrev = mean(revenue), sd_logdiff = sd(log(revenue/revenue.prev))) %>% 
-  ggplot(aes(log(meanrev), sd_logdiff, colour = sd_logdiff)) + 
-  geom_point() + facet_wrap(~ strategy_permit, scale="free") + 
-  ylab("CV revenue") + xlab("Ln mean revenue") + geom_smooth()
-```
-
-```{r, fig.cap="Species diversity versus ln (mean revenue)", fig.pos="placeHere", echo=FALSE, eval=FALSE}
-group_by(dat.diff, strategy_permit, year) %>% 
-  summarize(specDiv = mean(specDiv), meanrev = mean(revenue)) %>% 
-  ggplot(aes(specDiv, log(meanrev), colour = meanrev)) + 
-  geom_point() + facet_wrap(~ strategy_permit, scale="free") + 
-  ylab("Ln (mean revenue)") + xlab("Mean species diversity") + geom_smooth()
-```
-
-```{r, fig.cap="Species diversity versus CV", fig.pos="placeHere", echo=FALSE, eval=FALSE}
-group_by(dat.diff, strategy_permit, year) %>% 
-  summarize(specDiv = mean(specDiv), sd_logdiff = sd(log(revenue/revenue.prev))) %>% 
-  ggplot(aes(specDiv, sd_logdiff, colour = sd_logdiff)) + 
-  geom_point() + facet_wrap(~ strategy_permit, scale="free") + 
-  ylab("CV revenue") + xlab("Mean species diversity") + geom_smooth()
 ```
@@
 2010 was a huge pink salmon year for PWS (with high prices specialization was good) and 1991 was the start of pink salmon stock + price collapse.
-Perhaps slight declines in some of these in recent years?
+Across permits, perhaps slight declines in some of these in recent years?
 
-```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Time series of benefits of diversification by permit / strategy."}
-ggplot(df, aes(year, pred1, group = strategy, colour = as.factor(strategy))) + geom_line() + xlab("Year") + ylab("Benefit of increasing species diversity from 1 species") + facet_wrap(~ strategy, scale="free") + geom_smooth() + geom_hline(aes(yintercept=0)) + ggtitle("Effect on Mean")
+```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Time series of benefits of diversification by permit / strategy.", fig.height = 3}
+ggplot(df, aes(year, pred1, group = strategy, colour = as.factor(strategy))) + geom_line() + xlab("Year") + ylab("Benefit of increasing species diversity from 1 species") + ggtitle("Effect on Mean")
 ```
 
-```{r, echo=FALSE}
+```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Time series of benefits of diversification split out by permit / strategy.", fig.height = 3}
+ggplot(df, aes(year, pred1, group = strategy, colour = as.factor(strategy))) + geom_line() + xlab("Year") + ylab("Benefit of increasing species diversity from 1 species") + facet_wrap(~ strategy, scale="free") + geom_smooth() + geom_hline(aes(yintercept=0))
+```
+
+```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Time series of effects of diversification on mean and CV by permit."}
 df.cv = ranef(cv.mod)$cond$year_permit
@@
 
-#ggplot(df.cv, aes(year, pred1.cv, group = strategy, colour = as.factor(strategy))) + geom_line() + xlab("Year") + ylab("Benefit of increasing species diversity from 1 species") + facet_wrap(~ strategy, scale="free") + geom_smooth() + geom_hline(aes(yintercept=0)) + ggtitle("Effect on CV")
-
 ggplot(df.cv, aes(year, pred1.cv, group = strategy, colour = as.factor(strategy))) + geom_line(linetype="dashed") + geom_line(data=df, aes(year, pred1, group = strategy, colour = as.factor(strategy))) + xlab("Year") + ylab("Effect of increasing species diversity from 1 species") + facet_wrap(~ strategy, scale="free") + ggtitle("Effects on mean (solid) and CV (dashed)") + geom_hline(aes(yintercept=0), color="grey30")
 
-#df = left_join(df,df.cv)
-#ggplot(df, aes(pred1, pred1.cv, label=year)) + facet_wrap(~strategy, scale="free")+ #geom_vline(aes(xintercept=0)) + geom_hline(aes(yintercept=0)) + xlab("Benefit on the mean") + #ylab("Benefit on the CV") + geom_text(size=2)
 ```
@@
 
-```{r, echo = FALSE, fig.cap="Results for PWS Purse seine fishery", echo=FALSE, fig.pos="placeHere", cache=TRUE}
+```{r, echo = FALSE, fig.cap="Results for PWS Purse seine fishery", fig.pos="placeHere", cache=TRUE}
 
@@
 
-```{r, echo = FALSE, fig.cap="SEAK purse seine fishery", cache=TRUE, echo=FALSE, fig.pos="placeHere"}
+```{r, echo = FALSE, fig.cap="SEAK purse seine fishery", cache=TRUE, fig.pos="placeHere"}
 
@@
 
-```{r, echo = FALSE, fig.cap="SEAK gillnet (clustering primarily in chum-sockeye-chinokook/coho dimensions)", echo=FALSE, fig.pos="placeHere", cache=TRUE, eval=FALSE}
-fishery = "S03A"
-df = dat[which(dat$strategy_permit==fishery),]
-pamk_s03a <- pamk(df[,c("chum","pink","sock","salm")], 3)
-df$group = pamk_s03a$pamobject$clustering
-
-p3 = ggtern(data=df[,c("chum","salm","sock","group")], aes(x=chum,y=salm,sock,color=as.factor(group))) +
-  geom_point(alpha=0.3) +
-  labs(x="CHUM",y="OTHER",z="SOCK") + theme(
-        axis.title=element_text(size=8)) + theme(legend.position = "none")
-
-p0 = group_by(df, year, group) %>% 
-  summarize(n = length(unique(p_holder))) %>% 
-ggplot(aes(x = year, y = n, group=group, col = as.factor(group))) + geom_line() + ylab("People") + ggtitle(fishery)
-
-p1 = group_by(df, year, group) %>% 
-  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
-    cv = sd(logdiff,na.rm=T)) %>% 
-ggplot(aes(x = year, y = meanrev, group=group, col = as.factor(group))) + geom_line() + ylab("Mean revenue") + ggtitle(fishery)
-
-p2 = group_by(df, year, group) %>% 
-  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
-    cv = sd(logdiff,na.rm=T)) %>% 
-ggplot(aes(x = year, y = cv, group=group, col = as.factor(group))) + geom_line() + ylab("CV revenue") + ggtitle(fishery)
-
-gridExtra::grid.arrange(p3, p0, p1, p2, ncol=2)
-```
-
-```{r, echo = FALSE, fig.cap="SEAK gillnet fishery (clustering primarily in sockeye-chinook/coho dimension)", echo=FALSE, fig.pos="placeHere", cache=TRUE, eval=FALSE}
-fishery = "S03E"
-df = dat[which(dat$strategy_permit==fishery),]
-pamk_s03e <- pamk(df[,c("chum","pink","sock","salm")])
-df$group = pamk_s03e$pamobject$clustering
-
-p3 = ggtern(data=df[,c("chum","salm","sock","group")], aes(x=chum,y=salm,sock,color=as.factor(group))) +
-  geom_point(alpha=0.3) +
-  labs(x="CHUM",y="OTHER",z="SOCK") + theme(
-        axis.title=element_text(size=8)) + theme(legend.position = "none")
-
-p0 = group_by(df, year, group) %>% 
-  summarize(n = length(unique(p_holder))) %>% 
-ggplot(aes(x = year, y = n, group=group, col = as.factor(group))) + geom_line() + ylab("People") + ggtitle(fishery)
-
-p1 = group_by(df, year, group) %>% 
-  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
-    cv = sd(logdiff,na.rm=T)) %>% 
-ggplot(aes(x = year, y = meanrev, group=group, col = as.factor(group))) + geom_line() + ylab("Mean revenue") + ggtitle(fishery)
-
-p2 = group_by(df, year, group) %>% 
-  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
-    cv = sd(logdiff,na.rm=T)) %>% 
-ggplot(aes(x = year, y = cv, group=group, col = as.factor(group))) + geom_line() + ylab("CV revenue") + ggtitle(fishery)
-
-gridExtra::grid.arrange(p3, p0, p1, p2, ncol=2)
-```
-
-```{r, echo = FALSE, fig.cap="Kodiak purse seine (clustering primarily in sockeye-pink dimension)", echo=FALSE, fig.pos="placeHere", eval=FALSE, cache=TRUE}
-fishery = "S01K"
-df = dat[which(dat$strategy_permit==fishery),]
-pamk_s01k <- pamk(df[,c("chum","pink","sock","salm")])
-df$group = pamk_s01k$pamobject$clustering
-
-p3 = ggtern(data=df[,c("chum","salm","sock","group")], aes(x=chum,y=salm,sock,color=as.factor(group))) +
-  geom_point(alpha=0.3) +
-  labs(x="CHUM",y="OTHER",z="SOCK") + theme(
-        axis.title=element_text(size=8)) + theme(legend.position = "none")
-
-p0 = group_by(df, year, group) %>% 
-  summarize(n = length(unique(p_holder))) %>% 
-ggplot(aes(x = year, y = n, group=group, col = as.factor(group))) + geom_line() + ylab("People") + ggtitle(fishery)
-
-p1 = group_by(df, year, group) %>% 
-  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
-    cv = sd(logdiff,na.rm=T)) %>% 
-ggplot(aes(x = year, y = meanrev, group=group, col = as.factor(group))) + geom_line() + ylab("Mean revenue") + ggtitle(fishery)
-
-p2 = group_by(df, year, group) %>% 
-  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
-    cv = sd(logdiff,na.rm=T)) %>% 
-ggplot(aes(x = year, y = cv, group=group, col = as.factor(group))) + geom_line() + ylab("CV revenue") + ggtitle(fishery)
-
-gridExtra::grid.arrange(p3, p0, p1, p2, ncol=2)
-```
-
-```{r, echo = FALSE, fig.cap="Kodiak gillnet (clustering primarily in sockeye-pink dimension)", echo=FALSE, fig.pos="placeHere", cache=TRUE, eval=FALSE}
-fishery = "S04K"
-df = dat[which(dat$strategy_permit==fishery),]
-pamk_s04k <- pamk(df[,c("chum","pink","sock","salm")])
-df$group = pamk_s04k$pamobject$clustering
-
-p3 = ggtern(data=df[,c("chum","salm","sock","group")], aes(x=chum,y=salm,sock,color=as.factor(group))) +
-  geom_point(alpha=0.3) +
-  labs(x="CHUM",y="OTHER",z="SOCK") + theme(
-        axis.title=element_text(size=8)) + theme(legend.position = "none")
-
-p0 = group_by(df, year, group) %>% 
-  summarize(n = length(unique(p_holder))) %>% 
-ggplot(aes(x = year, y = n, group=group, col = as.factor(group))) + geom_line() + ylab("People") + ggtitle(fishery)
-
-p1 = group_by(df, year, group) %>% 
-  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
-    cv = sd(logdiff,na.rm=T)) %>% 
-ggplot(aes(x = year, y = meanrev, group=group, col = as.factor(group))) + geom_line() + ylab("Mean revenue") + ggtitle(fishery)
-
-p2 = group_by(df, year, group) %>% 
-  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
-    cv = sd(logdiff,na.rm=T)) %>% 
-ggplot(aes(x = year, y = cv, group=group, col = as.factor(group))) + geom_line() + ylab("CV revenue") + ggtitle(fishery)
-
-gridExtra::grid.arrange(p3, p0, p1, p2, ncol=2)
-```
-
-```{r, echo = FALSE, fig.cap="SEAK Purse seine (clustering primarily in sockeye-pink dimension)", echo=FALSE, fig.pos="placeHere", eval=FALSE, cache=TRUE}
-fishery = "S01M"
-df = dat[which(dat$strategy_permit==fishery),]
-pamk_s01m <- pamk(df[,c("chum","pink","sock","salm")])
-df$group = pamk_s01m$pamobject$clustering
-
-p3 = ggtern(data=df[,c("chum","pink","sock","group")], aes(x=chum,y=pink,sock,color=as.factor(group))) +
-  geom_point(alpha=0.3) +
-  labs(x="CHUM",y="PINK",z="SOCK") + theme(
-        axis.title=element_text(size=8)) + theme(legend.position = "none")
-
-p0 = group_by(df, year, group) %>% 
-  summarize(n = length(unique(p_holder))) %>% 
-ggplot(aes(x = year, y = n, group=group, col = as.factor(group))) + geom_line() + ylab("People") + ggtitle(fishery)
-
-p1 = group_by(df, year, group) %>% 
-  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
-    cv = sd(logdiff,na.rm=T)) %>% 
-ggplot(aes(x = year, y = meanrev, group=group, col = as.factor(group))) + geom_line() + ylab("Mean revenue") + ggtitle(fishery)
-
-p2 = group_by(df, year, group) %>% 
-  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
-    cv = sd(logdiff,na.rm=T)) %>% 
-ggplot(aes(x = year, y = cv, group=group, col = as.factor(group))) + geom_line() + ylab("CV revenue") + ggtitle(fishery)
-
-gridExtra::grid.arrange(p3, p0, p1, p2, ncol=2)
-```
-
 # Trends in revenue and prices
@@
 
-```{r, fig.cap="Inflation adjusted prices / species by area, relative to 2002 (the year corresponding to the lowest average prices).", fig.pos="placeHere", echo=FALSE}
-group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, area) %>%
-  summarize(p = mean(g_price)) %>% 
-  group_by(spec, area) %>% 
+```{r, fig.cap="Inflation adjusted prices / species by area, relative to 2002 (the year corresponding to the lowest average prices).", fig.pos="placeHere", echo=FALSE, fig.height = 3.5}
+g = group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, area) %>%
+  summarize(p = mean(g_price))
+p1 = group_by(g, spec, area) %>% 
   mutate(p = p/p[which(year==2002)]) %>% 
@@
   ggplot(aes(year, log(p), group=spec, color = spec)) + geom_line() + facet_wrap(~area) +
-  ylab("Ln avg price relative to 2002")
+  ylab("Ln avg price relative to 2002") + ggtitle("Synchrony in prices: relative change")
+p1
 ```
 
-```{r, fig.cap="Number of people landing fish", fig.pos="placeHere", echo=FALSE, eval=FALSE}
-#Participation (people landing fish on one of these permits) generally followed trends in prices, with an uptick since early 2000s,  
-group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, p_fshy) %>%
-  summarize(people = length(unique(p_holder))) %>%
-  ggplot(aes(year, people)) + geom_line() + facet_wrap(~p_fshy, scale="free") + 
-  ylab("People")
+```{r, fig.cap="Percent change (log-difference) in inflation adjusted prices / species by area", fig.pos="placeHere", echo=FALSE, fig.height = 3.5}
+g = group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, area) %>%
+  summarize(p = mean(g_price))
+# do log-diff of prices -- get price year before
+g$p_t1 = g$p[match(paste(g$year, g$spec, g$area), 
+  paste(g$year+1, g$spec, g$area))]
+p2 = group_by(g, spec, area) %>% 
+  ggplot(aes(year, log(p/p_t1), group=spec, color = spec)) + geom_line() + facet_wrap(~area) +
+  ylab("Ln price (t/t-1)") + ggtitle("Synchrony in prices: percent change") 
+p2
 ```
 
-```{r, fig.cap="Trends in revenue and variability across people", fig.pos="placeHere", echo=FALSE}
-#p1 = group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], p_holder, year, p_fshy) %>%
-#  summarize(gearn = sum(g_earn)) %>%
-#  group_by(year, p_fshy) %>%
-#  summarize(cv = sd(gearn)/mean(gearn)) %>%
-#  ggplot(aes(year, cv, color = p_fshy)) + geom_line() + 
-#  ylab("CV revenue (across people")
-
+```{r, fig.cap="Trends in revenue and variability across people", fig.pos="placeHere", echo=FALSE, fig.height = 3.5}
 p2 = group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], p_holder, year, p_fshy) %>%
@@
   ggplot(aes(year, log(med))) + geom_line() + facet_wrap(~p_fshy, scale="free") + 
-  geom_ribbon(aes(ymin= log(low), ymax= log(upp)), alpha=0.4) + ylab("Ln quartiles of revenue (across people)")
+  geom_ribbon(aes(ymin= log(low), ymax= log(upp)), alpha=0.4) + ylab("Ln quartiles of revenue (across people)") + ggtitle("Trends in revenue")
 print(p2)
@@
 Revenue is going up since 2000 -- but each permit holder is fishing more (more days, more total pounds) and the trend in weight is faster than the trend in price.  
-```{r}
+```{r, echo=FALSE, fig.height = 3.5, fig.cap="Trends in average landed weight per person", fig.pos="placeHere"}
 group_by(cfec, p_fshy, year, p_holder) %>% 
@@
   summarize(meanW = mean(meanW, na.rm=T)) %>% 
-  ggplot(aes(year, meanW)) + geom_line() + facet_wrap(~ p_fshy, scale="free") + ylab("Mean total tons per person") + xlab("Year")
-
+  ggplot(aes(year, meanW)) + geom_line() + facet_wrap(~ p_fshy, scale="free") + ylab("Mean total tons per person") + xlab("Year") + ggtitle("Trends in landed weight")
 ```
-
-```{r, fig.cap="CV of average species composition. This quantifies the synchrony across people -- so low CVs mean everyone's doing the same thing.", fig.pos="placeHere", echo=FALSE, eval=FALSE}
-# Another way to think about diversification is to calculate the deviations in species composition from the mean.
-Bascially what's the variability of euclidian distances?
-Given trends in total landings and diversity, the 'portfolio' of species available to any of these permit holders has changed through time.
-Another way to think about the diversification is whether the variability of people around the mean has changed (basically whether there's been trends in the CV of species diveristy across people).
-
-dat.salm = group_by(dat, year, p_holder) %>% 
-  mutate(earn = (chnk+coho+chum+pink+sock),
-    chnk=chnk/earn, coho=coho/earn,
-    chum=chum/earn, sock=sock/earn,
-    pink=pink/earn) %>% 
-  as.data.frame 
-dat.salm$chnk[is.na(dat.salm$chnk)] = 0
-dat.salm$coho[is.na(dat.salm$coho)] = 0
-dat.salm$chum[is.na(dat.salm$chum)] = 0
-dat.salm$pink[is.na(dat.salm$pink)] = 0
-dat.salm$sock[is.na(dat.salm$sock)] = 0
-
-dat.salm = group_by(dat.salm, year, strategy_permit) %>% 
-  mutate(mean.chnk = mean(chnk,na.rm=T), mean.chum = mean(chum,na.rm=T),
-    mean.pink = mean(pink,na.rm=T), mean.coho = mean(coho,na.rm=T), 
-    mean.sock = mean(sock,na.rm=T))
-
-dat.salm$eucld = (dat.salm$mean.chnk-dat.salm$chnk)^2 + (dat.salm$mean.coho-dat.salm$coho)^2 + (dat.salm$mean.pink-dat.salm$pink)^2 + (dat.salm$mean.sock-dat.salm$sock)^2 + (dat.salm$mean.chum-dat.salm$chum)^2
-
-p1 = group_by(dat.salm, year, strategy_permit) %>%
-  summarize(sd = sd(eucld)) %>%
-  ggplot(aes(year, sd, color = as.factor(strategy_permit))) + geom_line() + ylab("SD of euclidian dist.") + facet_wrap(~strategy_permit)
-
-group_by(dat, year, strategy_permit) %>% 
-  summarize(cv = sd(specDiv)/mean(specDiv)) %>% 
-  ggplot(aes(year, cv, color = as.factor(strategy_permit))) + geom_line() + ylab("CV of species diversity (across people)") + facet_wrap(~ strategy_permit, scale="free") + geom_smooth()
-```

### 2016-08-11 portfolio/text/anderson-etal-rev-pfx.tex
Add text skeleton

\documentclass[12pt,english]{article}
\usepackage{geometry}
\geometry{verbose, letterpaper, tmargin = 2.54cm, bmargin = 2.54cm,
  lmargin = 2.54cm, rmargin = 2.54cm}
\geometry{letterpaper}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{setspace}
\usepackage{booktabs}
\usepackage{ragged2e}
\usepackage{longtable}
\usepackage{multirow}
\usepackage{array}
\usepackage{tabu} % for spacing between rows in longtable
\usepackage{titlesec}
\usepackage{url}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{commath} % for \abs{}

\usepackage{lineno}
\usepackage{xcolor}
\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
\modulolinenumbers[2]

% Linux Libertine:
\usepackage{textcomp}
\usepackage[sb]{libertine}
\usepackage[varqu,varl]{inconsolata}% sans serif typewriter
\usepackage[libertine,bigdelims,vvarbb]{newtxmath} % bb from STIX
\usepackage[cal=boondoxo]{mathalfa} % mathcal
\useosf % osf for text, not math
\usepackage[supstfm=libertinesups,%
  supscaled=1.2,%
  raised=-.13em]{superiors}

\textheight 22.0cm

\usepackage[round,sectionbib]{natbib}
\bibpunct{(}{)}{;}{a}{}{;}

% Fix line numbering and align environment
% http://phaseportrait.blogspot.ca/2007/08/lineno-and-amsmath-compatibility.html
\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
  \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
  \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
  \renewenvironment{#1}%
     {\linenomath\csname old#1\endcsname}%
     {\csname oldend#1\endcsname\endlinenomath}}%
\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
  \patchAmsMathEnvironmentForLineno{#1}%
  \patchAmsMathEnvironmentForLineno{#1*}}%
\AtBeginDocument{%
\patchBothAmsMathEnvironmentsForLineno{equation}%
\patchBothAmsMathEnvironmentsForLineno{align}%
\patchBothAmsMathEnvironmentsForLineno{flalign}%
\patchBothAmsMathEnvironmentsForLineno{alignat}%
\patchBothAmsMathEnvironmentsForLineno{gather}%
\patchBothAmsMathEnvironmentsForLineno{multline}%
}

\title{Benefits and costs of income diversification for 
  Alaskan fishermen
\ldots}
\author{
Sean C. Anderson, 
\ldots}
\date{}

% remove numbers in front of sections:
\makeatletter
\renewcommand\@seccntformat[1]{}
\makeatother

\begin{document}
% \begin{flushleft}

\maketitle

\setlength{\parindent}{0.9cm}
\setlength\parskip{0in}

\vspace{-1cm}

%\input{../analysis/values} % R output

\thispagestyle{empty}

\linenumbers

\begin{spacing}{1.4}
\input{ms}

\end{spacing}
% \end{flushleft}
\end{document}

### 2016-08-11 portfolio/text/supp.tex
Add text skeleton

@@
 \geometry{letterpaper}
+\usepackage{graphicx}
+\usepackage{amssymb}
 \usepackage{amsmath}
-\usepackage{fixltx2e}
+\usepackage{amsfonts}
+\usepackage{setspace}
+\usepackage{booktabs}
+\usepackage{verbatim}
+\usepackage{amsmath}
+\usepackage{url}
+\usepackage{mathtools}
+\usepackage{bm}
+\usepackage{lineno}
+\usepackage{xcolor}
+\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
+\modulolinenumbers[2]
+
+% Linux Libertine:
+\usepackage{textcomp}
+\usepackage[sb]{libertine}
+\usepackage[varqu,varl]{inconsolata}% sans serif typewriter
+\usepackage[libertine,bigdelims,vvarbb]{newtxmath} % bb from STIX
+\usepackage[cal=boondoxo]{mathalfa} % mathcal
+%\useosf % osf for text, not math
+\usepackage[supstfm=libertinesups,%
+  supscaled=1.2,%
+  raised=-.13em]{superiors}
+
+\mathchardef\mhyphen="2D % math hyphen
+
+\textheight 22.0cm
+
+\usepackage[round,sectionbib]{natbib}
+\bibpunct{(}{)}{;}{a}{}{;}
+
+\setlength\parskip{0.10in}
+\setlength\parindent{0in}
+
+% Fix line numbering and align environment
+% http://phaseportrait.blogspot.ca/2007/08/lineno-and-amsmath-compatibility.html
+\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
+  \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
+  \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
+  \renewenvironment{#1}%
+     {\linenomath\csname old#1\endcsname}%
+     {\csname oldend#1\endcsname\endlinenomath}}%
+\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
+  \patchAmsMathEnvironmentForLineno{#1}%
+  \patchAmsMathEnvironmentForLineno{#1*}}%
+\AtBeginDocument{%
+\patchBothAmsMathEnvironmentsForLineno{equation}%
+\patchBothAmsMathEnvironmentsForLineno{align}%
+\patchBothAmsMathEnvironmentsForLineno{flalign}%
+\patchBothAmsMathEnvironmentsForLineno{alignat}%
+\patchBothAmsMathEnvironmentsForLineno{gather}%
+\patchBothAmsMathEnvironmentsForLineno{multline}%
+}
+
+% remove numbers in front of sections:
+\makeatletter
+\renewcommand\@seccntformat[1]{}
+\makeatother
+
+\title{Supporting Information}
+
+\author{
+}
+\date{}
 
@@
 
-\section{Log-difference model}
+\maketitle
 
-The log-difference model has several pros and cons over the undifferenced
-model. Negatives are that it uses a smaller subset of the data; positives are
+\renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\thetable}{S\arabic{table}}
+
+Hierarchical model
+
+\citep{kasperski2013}
+
+\begin{align*}
+\mathrm{Pr}(\nu_i < 10) &\sim \mathrm{Beta}(A_i, B_i)\\
+\mu_i &= \mathrm{logit}^{-1}(\alpha
+  + \alpha^\mathrm{class}_{j[i]}
+  + \alpha^\mathrm{order}_{k[i]}
+  + \alpha^\mathrm{species}_{l[i]}
+  + \bm{X_i} \bm{\beta}),
+  \: \\
+  &\quad \, \text{for } i = 1, \dots, 609\\
+A_i &= \phi_\mathrm{disp} \mu_i\\
+B_i &= \phi_\mathrm{disp} (1 - \mu_i)\\
+\alpha^\mathrm{class}_j &\sim
+  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{class}}),
+  \: \text{for } j = 1, \dots, 6\\
+\alpha^\mathrm{order}_k &\sim
+  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{order}}),
+  \: \text{for } k = 1, \dots, 38\\
+\alpha^\mathrm{species}_l &\sim
+  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{species}}),
+  \: \text{for } l = 1, \dots, 301,
+\end{align*}
+
+Negatives are that it uses a smaller subset of the data;
+
+positives are
 that it doesn't require estimating people-specific random effects (n = 3000+)
@@
 
+\bibliographystyle{apalike}
+\bibliography{refs}
+
+\clearpage
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=\textwidth]{../figs/stan-str-posteriors-dot.pdf}
+\caption{Caption here}
+\label{fig:ts}
+\end{center}
+\end{figure}
+
 \end{document}

### 2016-08-11 salmon/analysis/salmon_paper.Rmd
final version to send out
added code for residual plots

@@
 title: "Diversification of salmon fishermen within strategies (permits)"
-author: ""
-date: ""
+author: ''
+date: ''
 output:
@@
   html_document: default
+  word_document: default
 ---
@@
 ```{r, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
+revenueThreshold = 10000
+
 dat.annual = readRDS(file="../../portfolio/data-generated/cfec-annual-for-modeling.rds")
@@
 dat.diff = readRDS(file="../../portfolio/data-generated/cfec-diff-for-modeling.rds")
-dat.diff = dat.diff[which(dat.diff$revenue >= 5000 & dat.diff$revenue.prev >= 5000), ]
+dat.diff = dat.diff[which(dat.diff$revenue >= revenueThreshold & dat.diff$revenue.prev >= revenueThreshold), ]
 dat.diff = dat.diff[dat.diff$strategy_permit%in%c("S01A","S01E","S01K","S01L","S01M",
@@
 demo$effect = demo$a * (demo$speciesDiversity^2) + demo$b * (demo$speciesDiversity) + demo$c
+```
+
+```{r, echo = FALSE}
+pdf("Salmon_residuals.pdf")
+# Residuals for mean model
+dat.salmon$mu.resid = residuals(mu.mod)
+dat.salmon$mu.fitted = fitted(mu.mod)
+
+ggplot(dat.salmon, aes(mu.fitted, mu.resid, group = strategy_permit)) + geom_point() + facet_wrap(~strategy_permit) + xlab("Fitted") + ylab("Residuals") + ggtitle("Mean model")
+ggplot(dat.salmon, aes(specDiv.change, mu.resid, group = strategy_permit)) + geom_point() + facet_wrap(~strategy_permit) + xlab("Change in species diversification") + ylab("Residuals") + ggtitle("Mean model") + geom_hline(aes(yintercept=0),color="salmon1")
+ggplot(dat.salmon, aes(specdiv.prev, mu.resid, group = strategy_permit)) + geom_point() + facet_wrap(~strategy_permit) + xlab("Previous species diversity") + ylab("Residuals") + ggtitle("Mean model") + geom_hline(aes(yintercept=0),color="salmon1")
+ggplot(dat.salmon, aes(log((revenue+revenue.prev)/2), mu.resid, group = strategy_permit)) + geom_point() + facet_wrap(~strategy_permit) + xlab("Average revenue") + ylab("Residuals") + ggtitle("Mean model") + geom_hline(aes(yintercept=0),color="salmon1")
+
+# Residuals for CV model
+dat.salmon$cv.resid = residuals(cv.mod)
+dat.salmon$cv.fitted = fitted(cv.mod)
+
+ggplot(dat.salmon, aes(cv.fitted, cv.resid, group = strategy_permit)) + geom_point() + facet_wrap(~strategy_permit) + xlab("Fitted") + ylab("Residuals") + ggtitle("CV model")
+ggplot(dat.salmon, aes(specDiv.change, cv.resid, group = strategy_permit)) + geom_point() + facet_wrap(~strategy_permit) + xlab("Change in species diversification") + ylab("Residuals") + ggtitle("CV model") + geom_hline(aes(yintercept=0),color="salmon1")
+ggplot(dat.salmon, aes(specdiv.prev, cv.resid, group = strategy_permit)) + geom_point() + facet_wrap(~strategy_permit) + xlab("Previous species diversity") + ylab("Residuals") + ggtitle("CV model") + geom_hline(aes(yintercept=0),color="salmon1")
+ggplot(dat.salmon, aes(log((revenue+revenue.prev)/2), cv.resid, group = strategy_permit)) + geom_point() + facet_wrap(~strategy_permit) + xlab("Average revenue") + ylab("Residuals") + ggtitle("CV model") + geom_hline(aes(yintercept=0),color="salmon1")
+dev.off()
 ```
@@
 Revenue is going up since 2000 -- but each permit holder is fishing more (more days, more total pounds) and the trend in weight is faster than the trend in price.
+
 ```{r, echo=FALSE, fig.height = 3.5, fig.cap="Trends in average landed weight per person", fig.pos="placeHere"}

### 2016-08-11 salmon/analysis/salmon_paper.Rmd
removed offset in CV model
not much of an effect, but shouldn’t be there

--- 2016-08-11 salmon/analysis/salmon_paper.Rmd
@@
 
-cv.mod = glmmTMB(log(abs(residuals(mu.mod))) ~ specDiv.change + specDiv.change:specdiv.prev + specDiv.change:I(specdiv.prev^2) + days.change + (1+days.change|strategy_permit) + (-1+specDiv.change + specDiv.change:specdiv.prev + specDiv.change:I(specdiv.prev^2)|year_permit), offset = log(revenue.prev), dat.salmon)
+cv.mod = glmmTMB(log(abs(residuals(mu.mod))) ~ specDiv.change + specDiv.change:specdiv.prev + specDiv.change:I(specdiv.prev^2) + days.change + (1+days.change|strategy_permit) + (-1+specDiv.change + specDiv.change:specdiv.prev + specDiv.change:I(specdiv.prev^2)|year_permit), dat.salmon)
 # to ask whether benefit of diversificaiton is changing via strategy, we want to summarize (1) direction of 
@@
 demo$effect = demo$a * (demo$speciesDiversity^2) + demo$b * (demo$speciesDiversity) + demo$c
+```
+
+```{r, echo = FALSE}
+pdf("Salmon_residuals.pdf")
+# Residuals for mean model
+dat.salmon$mu.resid = residuals(mu.mod)
+dat.salmon$mu.fitted = fitted(mu.mod)
+
+ggplot(dat.salmon, aes(mu.fitted, mu.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Fitted") + ylab("Residuals") + ggtitle("Mean model")
+ggplot(dat.salmon, aes(specDiv.change, mu.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Change in species diversification") + ylab("Residuals") + ggtitle("Mean model") + geom_hline(aes(yintercept=0),color="salmon1")
+ggplot(dat.salmon, aes(specdiv.prev, mu.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Previous species diversity") + ylab("Residuals") + ggtitle("Mean model") + geom_hline(aes(yintercept=0),color="salmon1")
+ggplot(dat.salmon, aes(log((revenue+revenue.prev)/2), mu.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Average revenue") + ylab("Residuals") + ggtitle("Mean model") + geom_hline(aes(yintercept=0),color="salmon1")
+ggplot(dat.salmon, aes(specDiv.change, log(abs(mu.resid)), group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Change in species diversification") + ylab("Ln abs(residuals)") + ggtitle("Mean model")
+ggplot(dat.salmon, aes(specdiv.prev, log(abs(mu.resid)), group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Previous species diversity") + ylab("Ln abs(residuals)") + ggtitle("Mean model")
+ggplot(dat.salmon, aes(specdiv.prev, log(abs(mu.resid)), group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Previous species diversity") + ylab("Ln abs(residuals)") + ggtitle("Mean model")
+
+# Residuals for CV model
+dat.salmon$cv.resid = residuals(cv.mod)
+dat.salmon$cv.fitted = fitted(cv.mod)
+
+ggplot(dat.salmon, aes(cv.fitted, cv.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Fitted") + ylab("Residuals") + ggtitle("CV model")
+ggplot(dat.salmon, aes(specDiv.change, cv.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Change in species diversification") + ylab("Residuals") + ggtitle("CV model") + geom_hline(aes(yintercept=0),color="salmon1")
+ggplot(dat.salmon, aes(specdiv.prev, cv.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Previous species diversity") + ylab("Residuals") + ggtitle("CV model") + geom_hline(aes(yintercept=0),color="salmon1")
+ggplot(dat.salmon, aes(log((revenue+revenue.prev)/2), cv.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Average revenue") + ylab("Residuals") + ggtitle("CV model") + geom_hline(aes(yintercept=0),color="salmon1")
+dev.off()
 ```

### 2016-08-11 salmon/analysis/salmon_paper.Rmd
suppress warnings / add residuals

@@
 
-```{r, echo = FALSE}
-pdf("Salmon_residuals.pdf")
+```{r, echo = FALSE, warning=FALSE, message=FALSE, results='hide'}
+pdf("../../salmon/Salmon_residuals.pdf")
 # Residuals for mean model
@@
 ```
+
+\break
 

### 2016-08-13 portfolio/text/supp.tex
Draft model description

@@
 
-Hierarchical model
+\section{Hierarchical model}
 
-\citep{kasperski2013}
+ % mm <- model.matrix(log(revenue) ~ -1 + b1(spec_change) + b2(spec_change) +
+      % days_change + b1(spec_change):days_change + b2(spec_change):days_change,
 
-\begin{align*}
-\mathrm{Pr}(\nu_i < 10) &\sim \mathrm{Beta}(A_i, B_i)\\
-\mu_i &= \mathrm{logit}^{-1}(\alpha
-  + \alpha^\mathrm{class}_{j[i]}
-  + \alpha^\mathrm{order}_{k[i]}
-  + \alpha^\mathrm{species}_{l[i]}
-  + \bm{X_i} \bm{\beta}),
-  \: \\
-  &\quad \, \text{for } i = 1, \dots, 609\\
-A_i &= \phi_\mathrm{disp} \mu_i\\
-B_i &= \phi_\mathrm{disp} (1 - \mu_i)\\
-\alpha^\mathrm{class}_j &\sim
-  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{class}}),
-  \: \text{for } j = 1, \dots, 6\\
-\alpha^\mathrm{order}_k &\sim
-  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{order}}),
-  \: \text{for } k = 1, \dots, 38\\
-\alpha^\mathrm{species}_l &\sim
-  \mathrm{Normal}(0, \sigma^2_{\alpha \; \mathrm{species}}),
-  \: \text{for } l = 1, \dots, 301,
-\end{align*}
-
-Negatives are that it uses a smaller subset of the data;
-
-positives are
-that it doesn't require estimating people-specific random effects (n = 3000+)
-and removes possible time effects. The basic regression model is
+The basic model is a heirarchical linear regression with Gaussian errors
+predicting $\log$ revenue, $\log(R_{i,t})$, for
+each fisher $i$ and time (year) $t$:
 
 \begin{equation}
-  { Y }_{ i,t } \sim \mathrm{N}(\hat { { Y }}_{ i,t } ,\sigma^2 )
+\log(R_{i,t}) \sim \mathrm{N} \left(\widehat{\log(R_{i,t})}, \widehat{\sigma_{i,t}^2} \right).
 \end{equation}
 
-\subsection{Model for the mean}
-
-The model's expectation is written as a hierarchical random effects linear
-model with random effects in the intercept and slope.
+Each year $t$ of revenue for fisher $i$
+is assigned a strategy indexed by $j$ and
+expected $\log(R_{i,t})$ is modelled as:
 
 \begin{equation}
-{ \hat { Y }  }_{ t,i }={ B }_{ 0,j }+{ B }_{ 1,j }\cdot { s }_{ t,i }+{o}_{ t,i}
+\widehat{\log(R_{i,t})} = \beta_{0,j,t} +
+  \beta_{1,j} S_{i,t} I_{i,t} +
+  \beta_{2,j} S_{i,t} I_{i,t} +
+  \beta_3 D_{i,t} +
+  \beta_4 S_{i,t} D_{i,t} I_{i,t} +
+  \beta_5 S_{i,t} D_{i,t} I_{i,t} +
+  \log(R_{i,t-1})
 \end{equation}
@@
 \noindent
-where ${ B }_{ 0,j }$ is a normally distributed random effect intercept in the
-mean (j subscripting different strategies). The interpretation of ${B}_{ 0,j}$
-is that each after log-differencing, each strategy has different effects on
-changes in revenue (some tend to be winners, others tend to be losers). Initial
-versions of this model included estimating ${ B }_{ 0,j }$, which were largely
-centered on 0 (as expected), so these random effects were set to 0.  \newline
+where $S_{i,t}$ is the percent change in species diversity,
+$\log \left(\mathrm{div}_{i,t}/\mathrm{div}_{i,t-1}\right)$,
+$D_{i,t}$ is the percent change in days fished,
+$\log \left(\mathrm{days}_{i,t}/\mathrm{days}_{i,t-1}\right)$, and
+$I_{i,t}$ is an indicator variable that takes a value $0$
+if $\mathrm{div}_{i,t} \ge \mathrm{div}_{i,t-1}$
+(species diversity increasing; generalizing)
+and takes a value $1$ if
+$\mathrm{div}_{i,t} < \mathrm{div}_{i,t-1}$
+(species diversity decreasing; specializing).
+$\log(R_{i,t-1}$) is treated as an offset term:
+it is added on the right side of the equation
+to essentially model percent change in revenue,
+i.e., $\log(R_{i,t}) - \log(R_{i,t-1}) = \log(R_{i,t}/R_{i,t-1})$.
 
-The covariate ${ s }_{ t,i }$ is input as data, and represents the percent
-change in species diversity over the same time period that the percent change
-in revenue is being modeled, $\log\left( \frac { { div }_{ t+1,i } }{ { div }_{
-t,i } }  \right)$ where ${ div }_{ t,i }$ is the effective diversity of
-individual $i$ at time $t$. The coefficient ${ B }_{ 1,j }$ is a random effect
-on the slope, by strategy, representing how much an X\% change in species
-diversity within a strategy translates to a Y\% change in revenue.
-
-\subsection{Offset}
-
-To account for variable effort over time (changes in revenue could be
-attributed to diversification, but also to more fishing) we included an offset
-term ${o}_{t,i}$, representing the percent change in landing dates between
-years. We experimented with different versions of the offset and when included
-in the model as a predictor, the coefficient is estimated to be $\sim$ 1, so this is
-a reasonable assumption. The offset is calculated as $\log\left( \frac { { day
-}_{ t+1,i } }{ { day }_{ t,i } }  \right)$. \newline
-
-\subsection{Equation for the variance}
-
-Finally, the scale of the residual error term is also modeled as a linear
-function of covariates,
+The $\beta$'s represent estimated coefficients, some of which are
+allowed to vary.
+The intercepts $\beta_{0,j,t}$ are allowed to vary by
+strategy-year combinations,
+are centered on zero, and are constrained by a normal distribution:
 
 \begin{equation}
-{ \sigma  }_{ t,i }=
-\sqrt { \exp \left( \gamma _{ 0,j }+
-{ \gamma  }_{ 1,j }\cdot { s }_{ t,i } \right)  }
+\beta_{0,j,t} \sim \mathrm{N} \left(0, \sigma^2_{\beta_{j,t}} \right).
 \end{equation}
 
-\noindent
-where the $\gamma_{ 0,j }$ and $\gamma_{ 1,j }$ represent random intercepts and
-slopes in the strategies. I had been interpreting the intercepts as the
-between-group variability in strategies (some strategies will have more
-variable \% change in revenue than others, similar to some stocks having higher
-variability in returns). I was interpreting the random slope $\gamma_{ 1,j }$
-as the effect of within-strategy diversification; in other words, the effect
-of increases in species diversification varied among different strategies.
+This, combined with the lack of a global intercept, constrains
+the mean change in log revenue across all fishers and years to
+be zero. I.e., we assume that the data are stationary
+after accounting for all covariates and including
+an intercept that varies with strategy-year combinations.
+Initial versions of the model
+confirmed that the global intercept along with stragey-level
+intercepts would be estimated at almost exactly zero.
+
+TODO EXPLAIN WHY DO THIS
+
+The slopes $\beta_{1,j}$ and $\beta_{1,j}$
+are then allowed to vary by strategy with estimated means:
+
+\begin{align}
+\beta_{1,j} &\sim \mathrm{N} \left(\beta_1, \sigma^2_{\beta_1}\right),\\
+\beta_{2,j} &\sim \mathrm{N} \left(\beta_1, \sigma^2_{\beta_2}\right).
+\end{align}
+
+The $\beta_{1,j}$'s represent how much an X\% decrease in species
+diversity within a strategy translates to a Y\% change in revenue.
+The $\beta_{2,j}$'s represent how much an X\% increase in species
+diversity within a strategy translates to a Y\% change in revenue.
+These are after controlling for percent change in days fished, $\beta_3$,
+and the interaction between days fished and percent change in species
+diversity, $\beta_4$ and $\beta_5$.
+
+Whereas a typical heirarchical linear regression model
+would estimate a single residual error scale term, $\sigma$,
+we model the scale of the residual error  with
+another heirarchical model with a similar form:
+
+\begin{equation}
+\widehat{\sigma_{i,t}} =
+\exp \left( \gamma_{0,j} + \eta_1 M_j +
+\gamma_{1,j} S_{i,t} I_{i,t} +
+\gamma_{2,j} S_{i,t} I_{i,t} +
+\gamma_3 D_{i,t} +
+\gamma_4 S_{i,t} D_{i,t} I_{i,t} +
+\gamma_5 S_{i,t} D_{i,t} I_{i,t}\right),
+\end{equation}
+
+where $M_j$ represents the raw mean species diversity for strategy $j$
+and $\eta_1$ represents an estimated strategy-level predictor.
+In other words, $\gamma_{0,j} + \eta_1 M_j$ represents the intercept
+with a built-in strategy-level regression (Fig.~3)
+that incorporates parameter uncertainty.
+Again, $S_{i,t}$ and $I_{i,t}$ represent, respectively,
+percent change in species
+diversity and an indicator variable for increasing or
+decreasing species diversity.
+We exponentiate the equation to ensure all scale parameters, $\sigma_{i,t}$,
+are positive.
+
+The $\gamma$'s are estimated coefficients, three of which
+are allowed to vary by strategy, constrained by normal distributions
+and with estimated means:
+
+\begin{align}
+\gamma_{0,j}  &\sim \mathrm{N} \left(\gamma_0, \sigma^2_{\gamma_0}\right),\\
+\gamma_{1,j}  &\sim \mathrm{N} \left(\gamma_1, \sigma^2_{\gamma_1}\right),\\
+\gamma_{2,j}  &\sim \mathrm{N} \left(\gamma_2, \sigma^2_{\gamma_2}\right).
+\end{align}
+
+We fit our models in Stan version 2.11.1. \ldots
+
+We checked our models with a two-part modelling approach
+using the R package lme4.
+To do this, we fit the initial model of the mean component $\log(R_{i,t})$
+and then fit the second model to the log absolute value of the
+residuals, $\log | \widehat{\log(R_{i,t})} - \log(R_{i,t}) |$.
+Most importantly, (1) this model does not carry uncertainty
+from the first mean-component model to the variance-component model, but
+(2) it also leaves the residuals in the mean-component model strongly
+heteroskedastic and therefore improperly weights individual data points, and
+(3) does not allow for proper modelling of the strategy-level predictor,
+$\eta_1$. Nonetheless, it provides a check that our results
+from the Stan model are approximately correct and allowed us
+to quickly iterate and build intuition about the models.
+
+\subsection{points to include}
+
+Negatives are that it uses a smaller subset of the data;
+positives are
+that it doesn't require estimating people-specific random effects (n = 3000+)
+and removes time effects.
+
+% \verbatiminput{../analysis/portfolio-offset.stan}
 

### 2016-08-13 salmon/analysis/salmon_paper.Rmd

@@
 output:
+  word_document: default
   pdf_document:
@@
   html_document: default
-  word_document: default
 ---
@@
 library(glmmTMB)
+library(zoo)
 ```
@@
 ## Methods
+
+### Hatchery programs were designed to reduce variation in harvest. Has that happened?
 
@@
 gridExtra::grid.arrange(p1,p2,p3, ncol=2)
+```
+
+\break
+
+# Has variability in harvest been reduced? (in part due to increased hatchery production?)
+
+### There's a number of ways we can think about quantifying variation  
+- We can calculate the aggregate (total harvest weight or revenue summed over people)  
+- We can calculate the variability in revenue across individual fishermen  
+- We can compare variability of revenue of individual people
+
+What we see here is that (1) aggregate variability remains high at the fishery level.
+In part because of large swings in prices in late 1990s/early 2000s.
+We also see (2) variability across people flat or increasing for most fisheries, with the exception of S03A.
+Finally, we see (3) Increased CV of revenue for people fishing in the recent decade who were also fishing 1985-1995 (exception: S01A)
+
+```{r, fig.cap="Trends in annual variability by fishery", fig.pos="placeHere", echo=FALSE}
+summary = group_by(cfec.raw[cfec.raw$spec%in%c("CHUM","COHO","CHNK","SOCK","PINK"),], year, p_fshy) %>% 
+  summarize(sumg = log(sum(g_earn)), sumw = log(sum(g_pounds))) %>% 
+  arrange(p_fshy, year) %>% group_by(p_fshy) %>% 
+  mutate(logdiffg = rollapply(c(NA, diff(sumg)), 3, sd, fill=c(NA, NULL, NA)), 
+    logdiffw = rollapply(c(NA, diff(sumw)), 3, sd, fill=c(NA, NULL, NA)))
+
+ggplot(summary, aes(year, logdiffg )) + geom_line(color ="green" ) + facet_wrap(~ p_fshy, scale="free") + geom_line(aes(year, logdiffw), color = "blue") + ylab("3 year rolling CV") + xlab("Year") + ggtitle("Total revenue (green) and weight (blue)")
+```
+
+```{r, fig.cap="Trends in variability across people", fig.pos="placeHere", echo=FALSE, fig.height=3.5}
+summary = group_by(cfec.raw[cfec.raw$spec%in%c("CHUM","COHO","CHNK","SOCK","PINK"),], year, p_fshy, p_holder) %>% 
+  summarize(sumg = sum(g_earn), sumw = sum(g_pounds)) %>% 
+  group_by(p_fshy, year) %>% 
+  summarize(cvg = sd(sumg,na.rm=T)/mean(sumg,na.rm=T),
+    cvw = sd(sumw,na.rm=T)/mean(sumw,na.rm=T))
+
+ggplot(summary, aes(year, cvg )) + geom_line(color ="green" ) + facet_wrap(~ p_fshy, scale="free") + geom_line(aes(year, cvw), color = "blue") + ylab("3 year rolling CV") + xlab("Year") + ggtitle("Variability across people: total revenue (green) and weight (blue)")
+
+```
+
+```{r, fig.cap="Change in CV for individual people", fig.pos="placeHere", echo=FALSE, fig.height=3.5}
+summary = group_by(cfec.raw[cfec.raw$spec%in%c("CHUM","COHO","CHNK","SOCK","PINK"),], year, p_fshy, p_holder) %>% 
+  summarize(sumg = sum(g_earn)) %>% 
+  group_by(p_fshy, p_holder) %>% 
+  mutate(nPre = length(which(year%in%seq(1985,1995))),
+    nPost = length(which(year%in%seq(2005,2014)))) %>% 
+  filter(nPre > 3 & nPost > 3) %>% select(-nPre, -nPost) %>% 
+  summarize(cv.pre = mean(sumg[which(year <= 1995)],na.rm=T)/sd(sumg[which(year <= 1995)],na.rm=T),
+    cv.post = mean(sumg[which(year >= 2005)],na.rm=T)/sd(sumg[which(year >= 2005)],na.rm=T))
+
+ggplot(summary, aes(log(cv.post/cv.pre))) + geom_histogram(bins=100) + facet_wrap(~ p_fshy) + geom_vline(aes(xintercept = 0)) + ylab("Count") + xlab("% Change in CV for individual people (2005-2014 / 1985-1995)") + ggtitle("Change in variability for individual people")
+#summary = left_join(expand.grid(year = unique(summary$year), p_holder = unique(summary$p_holder)), summary)
+
+# add rolling CV for each person
+#summary = group_by(summary, p_holder) %>% 
+#  mutate(logdiffg = rollapply(c(NA, diff(sumg)), 3, sd, fill=c(NA, NULL, NA))) %>%
+#  do(fit = lm(logdiffg ~ year, data = .))
+
 ```

### 2016-08-13 portfolio/analysis/9.2-check-offset-stan.R
Add main effects plot to SI

@@ -146,3 +146,57 @@ filter(b, grepl("b_j\\[", term) | grepl("g_k\\[", term) | grepl("h1", term)) %>%
   geom_hline(yintercept = 0, lty = 2) +
   coord_flip()
 ggsave("portfolio/figs/stan-main-effects.pdf", width = 6.5, height = 6)
+
+p <- stanhelpers::extract_df(m, output = "long_df") %>%
+  filter(grepl("^b_j", variable) | grepl("^g0$", variable) |
+      grepl("^h1", variable) | grepl("^g_k", variable)) %>%
+  mutate(variable = as.character(variable)) %>%
+  group_by(variable) %>%
+  summarize(l = quantile(value, probs = 0.025),
+    l.5 = quantile(value, probs = 0.25),
+    m = quantile(value, probs = 0.5),
+    u.5 = quantile(value, probs = 0.75),
+    u = quantile(value, probs = 0.975))
+
+# changed?
+stopifnot(p$variable == c("b_j_1", "b_j_2", "b_j_3", "b_j_4", "b_j_5",
+  "g_k_1", "g_k_2",
+  "g_k_3", "g_k_4", "g_k_5", "g0", "h1"))
+
+term_lu <- data.frame(
+  variable = c("b_j_1", "b_j_2", "b_j_3", "b_j_4", "b_j_5", "g_k_1", "g_k_2",
+    "g_k_3", "g_k_4", "g_k_5", "g0", "h1"),
+  term_clean = c(
+    "b1 (specializing)",
+    "b2 (generalizing)",
+    "b3 (days fished % change)",
+    "b4 (days fished %:specializing)",
+    "b5 (days fished %:generalizing)",
+
+    "g1 (specializing)",
+    "g2 (generalizing)",
+    "g3 (days fished % change)",
+    "g4 (days fished %:specializing)",
+    "g5 (days fished %:generalizing)",
+
+    "g0 (global scale intercept)",
+    "u1 (strategy-level diversity effect)"
+  )
+  )
+p$term_clean <- NULL
+p <- inner_join(p, term_lu)
+p$term_clean <- factor(p$term_clean,
+  levels = rev(sort(as.character(p$term_clean))))
+
+devtools::load_all("pfxr")
+
+p1 <- ggplot(p, aes(x = term_clean, y = m)) +
+  geom_point() +
+  geom_pointrange(aes(ymin = l, ymax = u), size = 0.2) +
[…]

### 2016-08-13 portfolio/figs/strategy-table.tex
Add strategy table to SI

% latex table generated in R 3.3.1 by xtable 1.8-2 package
Label & Description \\ 
  \midrule
Dun C & Dungeness crab \\ 
  Fin - jig & Finfish, mechanical jig \\ 
  Fin - ll & Finfish, longline, statewide \\ 
  Fin - otter & Finfish, otter trawl, statewide \\ 
  Fin - pot & Finfish, pot gear, statewide \\ 
  Hal & Halibut \\ 
  Herr - pound & Herring, pound \\ 
  Herr roe - gill & Herring roe, gillnet \\ 
  Herr roe - seine & Herring roe, purse seine \\ 
  King C & King crab \\ 
  S - drift - CI, AP, BB (sock) & Salmon (mostly sockeye), drift gill net, CI, AP, BB \\ 
  S - drift - PWS & Salmon, drift gill net, PWS \\ 
  S - drift - SE & Salmon, drift gill net, Southeast Alaska \\ 
  S - seine - AP & Salmon, purse seine, Alaska Penninsula \\ 
  S - seine - CH & Salmon, purse seine, Chignik \\ 
  S - seine - K & Salmon, purse seine, Cook Inlet and Kodiak \\ 
  S - seine - PWS & Salmon, purse seine, PWS \\ 
  S - seine - SE & Salmon, purse seine, Southeast Alaska \\ 
  S - set - NS/K & Salmon, set gillnet, Norton Sound and Kuskokwim \\ 
  S - set - PWS, CI, K, AP, BB (sock) & Salmon (mostly sockeye), set gillnet, PWS, CI, K,  AP, BB \\ 
  S - set - YAK & Salmon, set gillnet, Yakutat \\ 
  S - set - YUK & Salmon, gillnet, lower Yukon \\ 
  S - troll & Salmon , power troll, statewide \\ 
  Sab & Sablefish, longline, statewide \\ 
  Sab - SE & Sablefish , longline, Southeast Alaska \\ 
  Sea cuc & Sea cucumber \\ 
  Shrimp & Shrimp \\ 
  Tan C & Tanner crab, pot gear vessel under 60' \\ 
  Tan C - large & Tanner crab, pot gear vessel over 60' \\ 
   \bottomrule

### 2016-08-13 portfolio/analysis/9.3-str-tabel.R

@@ -1,0 +1,19 @@
+library(dplyr)
+d <- readr::read_csv("data/strategies-labels.csv")
+d <- select(d, str_label, description) %>%
+  rename(Label = str_label, Description = description) %>%
+  arrange(Label)
+
+xtable::print.xtable(
+  xtable::xtable(d,
+    caption = ""),
+  include.rownames = FALSE,
+  file = "portfolio/figs/strategy-table.tex",
+  booktabs = TRUE,
+  caption.placement = "top",
+  size = "small",
+  sanitize.text.function = identity,
+  only.contents = TRUE,
+  timestamp = NULL
+)
+

### 2016-08-13 portfolio/analysis/1-create-modelling-data.Rmd
Tidy file names / archive many files

--- 2016-08-03 portfolio/analysis/3-createData_linear_modeling.Rmd
@@
     sock = sum(g_earn[spec=="SOCK"]),
-    chum = sum(g_earn[spec=="CHUM"]))
+    chum = sum(g_earn[spec=="CHUM"]),
+    region = region[1],
+    nregion = length(unique(region)))
 
@@
 cfecAnnual$sock.prev = cfecAnnual$sock[idx]
+
+cfecAnnual$region.prev = cfecAnnual$region[idx]
+cfecAnnual$nregion.prev = cfecAnnual$nregion[idx]
 ```

### 2016-08-13 portfolio/analysis/archive/7-separate-models-mean.Rmd
Tidy file names / archive many files

---
title: "Analysis of revenue"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE, autodep=TRUE, warning = TRUE)
```

```{r load-data}
library(dplyr)
library(ggplot2)
library(lme4)
dm <- readRDS("../data-generated/cfec-annual-for-modeling.rds")
nrow(dm)
```

We'll filter the data to keep permit holders that fished for at least ten years. We'll also have to filter out any permit holders that have no variability in the diversity fished.

```{r filter-data}
dm <- group_by(dm, p_holder) %>%
  mutate(nyr = length(unique(year))) %>%
  mutate(range_div = diff(range(specDiv))) %>%
  as_data_frame() %>%
  filter(nyr >= 10) %>%
  filter(range_div > 0)
nrow(dm)

top.strategies = names(rev(sort(table(dm$strategy)))[1:100])
dm = dm[dm$strategy%in%top.strategies, ]
nrow(dm)

dm_holder <- group_by(dm, p_holder) %>%
  summarize(mean_diversity = mean(log(specDiv)))%>%
  mutate(p_holder = as.character(p_holder)) %>%
  unique

dm_strategy <- group_by(dm, strategy) %>%
  summarize(mean_diversity = mean(log(specDiv))) %>%
  unique

holder_strategy <- select(dm, strategy, p_holder) %>% distinct() %>%
  mutate(p_holder = as.character(p_holder))

dm$log_spec_div <- scale(log(dm$specDiv))
dm$log_length <- scale(log(dm$length + 1))
dm$log_weight <- scale(log(dm$weight + 1))
dm$log_days <- scale(log(dm$days + 1))
```

We'll fit a model with `lmer()`.

```{r fit-models}
# m5 <- lmer(log(revenue) ~ log_spec_div * log_days * log_length +
#     (1 + log_spec_div|strategy) + (1 + log_spec_div|p_holder), data = dm)
m5 <- lmer(log(revenue) ~ (log_spec_div + log_days + log_length)^2 +
    (1 + (log_spec_div + log_days + log_length)|strategy) + 
    (1 + log_spec_div|p_holder), data = dm)
```

```{r extract-coefficients}
co_strategy <- coef(m5)$strategy
co_strategy$strategy <- row.names(co_strategy)

co_holder <- coef(m5)$p_holder
co_holder$p_holder <- row.names(co_holder)

# arm::display(m5)
# print(m5, digits = 2)
# summary(m5)
```

For a given permit holder effects:

```{r plot-coefficients}
broom::tidy(m5, conf.int = TRUE) %>%
  filter(term != "(Intercept)") %>%
  filter(!grepl("cor_", term)) %>%
  mutate(conf.low = ifelse(is.na(std.error), estimate, conf.low)) %>%
  mutate(conf.high = ifelse(is.na(std.error), estimate, conf.high)) %>%
  ggplot(aes(y = estimate, ymax = conf.low, ymin = conf.high, x = term)) +
  geom_pointrange() +
  coord_flip() +
  geom_hline(yintercept = 0, lty = 2)
```

What I see:

- increased diversity = lower revenue
- longer boats make more money (obviously)s
- longer seasons = more money (obviously)
- no major interaction effects for these 3 predictors

Now let's look at across permit holder effects:

```{r plot-holder-level-coefficients, fig.width=13, fig.height=6}
co_holder_long <- tidyr::gather(co_holder, parameter, estimate, -p_holder) %>%
  filter(parameter %in% c("(Intercept)", "log_spec_div")) %>%
  inner_join(dm_holder) %>%
  inner_join(holder_strategy)

strategy_grouped <- rbind(cbind(co_holder_long[grep("M",co_holder_long$strategy),], group="Groundfish"),
  cbind(co_holder_long[grep("S",co_holder_long$strategy),], group="Salmon"),
  cbind(co_holder_long[grep("B|C",co_holder_long$strategy),], group="Halibut-Sablefish"),
  cbind(co_holder_long[grep("K|T|D",co_holder_long$strategy),], group="Crab"),
  cbind(co_holder_long[grep("G",co_holder_long$strategy),], group="Herring"))

strategy_grouped <- mutate(strategy_grouped, 
  salmon = ifelse(grepl("S", strategy), TRUE, FALSE))
ggplot(strategy_grouped, aes(mean_diversity, estimate, color = salmon)) +
  geom_point(alpha = 0.1) +
  facet_grid(parameter~group, scales = "free_y") +
  geom_smooth(method = "lm", se = FALSE)
```

What I see:

- permit holders with higher mean diversity on average make slightly more money
- the variability around this relationship becomes tighter at higher levels of mean diversity
- not much change in the cost of diversifying across permit holders with different mean diversity levels

Now let's look at across strategy effects:

```{r plot-strategy-level-coefficients, fig.width=13, fig.height=6}
co_strategy_long <- tidyr::gather(co_strategy, parameter, estimate, -strategy) %>%
  filter(parameter %in% c("(Intercept)", "log_spec_div")) %>%
  inner_join(dm_strategy)

strategy_grouped <- rbind(cbind(co_strategy_long[grep("M",co_strategy_long$strategy),], group="Groundfish"),
  cbind(co_strategy_long[grep("S",co_strategy_long$strategy),], group="Salmon"),
  cbind(co_strategy_long[grep("B|C",co_strategy_long$strategy),], group="Halibut-Sablefish"),
  cbind(co_strategy_long[grep("K|T|D",co_strategy_long$strategy),], group="Crab"),
  cbind(co_strategy_long[grep("G",co_strategy_long$strategy),], group="Herring"))

strategy_grouped <- mutate(strategy_grouped, 
  salmon = ifelse(grepl("S", strategy), TRUE, FALSE))

ggplot(strategy_grouped, aes(mean_diversity, estimate, label = strategy, color = salmon)) +
  geom_text() + geom_smooth(se = FALSE, method = "lm") +
  facet_grid(parameter~group, scales = "free_y")
```

What I see:

- strategies with higher mean diversity on average make more money, especially if salmon are amongst the strategy 
- for some groupings, strategies with higher mean diversity have less of a diversifying cost (especially for salmon only, salmon plus crab, and salmon plus herring)

### 2016-08-13 portfolio/analysis/archive/7-separate-models.Rmd
Tidy file names / archive many files

--- 2016-07-01 portfolio/analysis/7-separate-models.Rmd
@@
   as_data_frame() %>%
-  filter(nyr >= 5) %>%
+  filter(nyr >= 10) %>%
   filter(range_div > 0)
@@
     summarise(cv = sd(revenue)/mean(revenue), log_specDiv = mean(log(specDiv)),
+    # summarise(cv = sd(log(revenue)-log(revenue.prev), na.rm = T),
+      log_specDiv = mean(log(specDiv)),
     # summarise(cv = mean(revenue), log_specDiv = mean(log(specDiv)),
@@
   as_data_frame() %>%
+  filter(!is.na(cv)) %>%
   group_by(strategy) %>%
@@
     (1 + (scaled_log_spec+scaled_log_days)^2|strategy), data = dmv)
+```
+
+```{r}
+library(mgcv)
+dmv <- dmv %>% group_by(strategy) %>% mutate(cv_centered = log(cv) - mean(log(cv)))
+m1 <- gam(cv_centered ~ s(scaled_log_spec) + s(scaled_log_days), data = dmv)
+par(mfrow = c(1,2))
+plot(m1)
 ```
@@
 ```
+
+With differenced data
+
+```{r}
+df = readRDS(file="../data-generated/cfec-diff-for-modeling.rds")
+dmv <- df %>% group_by(p_holder) %>%
+    summarise(sd_ = sd(log(revenue/revenue.prev)), log_specDiv = mean(log(specDiv)),
+    length = mean(log(length+1), na.rm = TRUE), days = mean(log(days+1), na.rm = TRUE),
+    strategy = paste(sort(unique(substr(strategy, 1, 1))), collapse = " ")) %>%
+  as_data_frame() %>%
+  filter(!is.na(sd_)) %>%
+  group_by(strategy) %>%
+  mutate(n_strategy = n()) %>%
+  filter(n_strategy > 6) %>%
+  mutate(mean_log_spec = mean(log_specDiv)) %>%
+  as_data_frame() %>%
+  mutate(scaled_log_days = scale(days), scaled_log_spec = scale(log_specDiv),
+    scaled_log_length = scale(length))
+dmv$strategy <- as.factor(dmv$strategy)
+
+library(mgcv)
+dmv <- dmv %>% group_by(strategy) %>% mutate(log_sd = log(sd_) - mean(log(sd_)))
+m1 <- gam(log_sd ~ s(scaled_log_spec) + s(scaled_log_days), data = dmv)
+par(mfrow = c(1,2))
+plot(m1)
+
+```

### 2016-08-13 portfolio/analysis/0.5-standardize.R

@@ -1,54 +1,0 @@
-library(dplyr)
-library(ggplot2)
-library(mgcv)
-if (!exists("cfec")) {
-  cfec <- feather::read_feather("data/cfec.feather")
-}
-
-standardize_smooth <- function(year, response) {
-  response <- response + 1
-  tryCatch({
-  # m <- gam(response ~ s(year), family = Gamma(link = "log"))
-  m <- glm(response ~ year, family = Gamma(link = "log"))
-  p <- predict(m, type = "response")
-  p
-  }, error = function(e) rep(NA, length(response)))
-}
-
-cfec_holder <- cfec %>%
-  group_by(p_holder, year) %>%
-    summarize(total_revenue = sum(g_earn, na.rm = TRUE),
-     species_permit = substr(p_fshy[1],1,1),
-     species_permit_n = length(unique(substr(p_fshy,1,1))),
-     area_permit = substr(p_fshy[1],5,5),
-     area_n = length(unique(substr(p_fshy,5,5))))
-
-ignore <- filter(cfec_holder, species_permit_n > 1 | area_n > 1)  %>%
-  as.data.frame
-
-ggplot(ignore, aes(species_permit_n)) + geom_histogram()
-ggplot(ignore, aes(area_n)) + geom_histogram()
-ggplot(cfec_holder, aes(species_permit_n)) + geom_histogram()
-ggplot(cfec_holder, aes(area_n)) + geom_histogram()
-
-cfec_holder <- filter(cfec_holder, species_permit_n == 1, area_n == 1) %>%
-  as.data.frame
-
-nrow(ignore)/nrow(cfec_holder)
-
-cfec_holder <- mutate(cfec_holder, group = paste(area_permit, species_permit)) %>%
-  mutate(year_factor = as.factor(year))
-
-cfec_holder <-
-  cfec_holder %>% group_by(species_permit, area_permit) %>%
-  mutate(smooth = standardize_smooth(year_factor, total_revenue )) %>%
-  as.data.frame
-
-cfec_holder <- cfec_holder %>% arrange(p_holder, year)
-
-p <- ggplot(filter(cfec_holder, group %in% unique(cfec_holder$group)[1:20]),
-  aes(x = year, y = log(total_revenue +1), group = p_holder)) +
-  geom_line(alpha = 0.05) +
-    geom_point(aes(y = log(smooth)), colour = "red", lwd = 1) +
-    facet_wrap(area_permit~species_permit)
-ggsave("figs/check-smooths.png", width = 8, height = 8)

### 2016-08-13 portfolio/analysis/2-try-model.R

@@ -1,438 +1,0 @@
-library(dplyr)
-library(rstan) # >= 2.7.0-1
-rstan::rstan_options(auto_write = TRUE)
-options(mc.cores = parallel::detectCores())
-library(ggplot2)
-
-set.seed(1234)
-b0 <- 0.5
-b1 <- 0.3
-sigma0 <- 0.2
-sigma1 <- -0.2
-N <- 20000
-s <- base::sample(seq(1, 10), N, replace = TRUE)
-u <- (rnorm(N, b0 + b1*s, sqrt(exp(sigma0 + sigma1*s))))
-
-plot(s, u)
-
-# d <- readRDS("generated-data/species_diversity_metrics.rds")
-# ggplot(d, aes(log(diversity_by_earnings), log(10^m))) + geom_point(alpha = 0.1)
-# ggplot(d, aes(log(diversity_by_earnings), log(cv))) + geom_point(alpha = 0.1)
-
-library(nlme)
-m_gls1 <- gls(u ~ s, weights = varExp(form = ~s))
-
-library(TMB)
-compile("revenue1.cpp")
-dyn.load(dynlib("revenue1"))
-
-obj <- MakeADFun(
-  data = list(x = s, y = u),
-  parameters = list(b0 = 0, b1 = 0, sigma0 = 0, sigma1 = 0),
-  DLL = "revenue1")
-opt <- nlminb(start=obj$env$last.par.best, objective=obj$fn, gradient=obj$gr)
-opt$par
-rep <- sdreport(obj)
-rep
-
-# with matrix notation
-compile("revenue2.cpp")
-dyn.load(dynlib("revenue2"))
-
-mm <- model.matrix(~ s)
-
-obj <- MakeADFun(
-  data = list(x_ij = mm, y_i = u),
-  parameters = list(b_j = c(0, 0), sigma_j = c(0, 0)),
-  DLL = "revenue2")
-
-opt <- nlminb(start=obj$env$last.par.best, objective=obj$fn, gradient=obj$gr)
-opt$par
-rep <- sdreport(obj)
-rep
-
-# Add a random intercept
-set.seed(1234)
-b0 <- 0.5
-b1 <- 0.3
-sigma0 <- 0.2
-sigma1 <- -0.2
-log_b0_sigma <- -1.8
-n_k <- 30
-b0_eta <- rnorm(n_k, 0, exp(log_b0_sigma))
[…]

### 2016-08-13 portfolio/analysis/9.3-map.R

@@ -1,12 +1,0 @@
-library(maps)
-library(mapdata)
-map("worldHires",".", xlim=c(-178,-130), ylim=c(51,73), col="gray80",
-  fill=TRUE, mar = c(1,1,1,1), boundary = F, border = "grey80")
-box(col = "grey40")
-
-p <- readr::read_csv("data/map-points.csv")
-
-for (i in 1:nrow(p)) {
-  points(-1*p$long[i], p$lat[i], col = "red", pch = 19, cex = 1)
-  graphics::text(-1*p$long[i], p$lat[i], labels = p$name[i], pos = 4)
-}

### 2016-08-13 portfolio/text/supp.Rnw
Switch SI to .Rnw

--- 2016-08-13 portfolio/text/supp.tex
@@
 \usepackage{amsmath}
+\usepackage{longtable}
+\usepackage{multirow}
+\usepackage{array}
+\usepackage{tabu} % for spacing between rows in longtable
 \usepackage{url}
@@
 
+<<set-knitr-options, cache=FALSE, echo=FALSE>>=
+library("knitr")
+opts_chunk$set(fig.align='center', fig.pos="htpb", cache=TRUE, echo=FALSE,
+  message=FALSE, autodep=TRUE, results='hide', fig.path='../figs/',
+  fig.width=6, par=TRUE)
+opts_chunk$set(warning=FALSE, message=FALSE, tidy=FALSE, refresh=TRUE,
+  cache.path="../cache/", cache.lazy=TRUE, size="small", cache.comments=FALSE)
+opts_chunk$set(dev = 'pdf')
+opts_knit$set(out.format = "latex", out.width="\\textwidth")
+opts_knit$set(eval.after = 'fig.cap') # evaluate fig.cap after the chunk
+opts_chunk$set(collapse = TRUE, comment = "")
+@
+
+<<get-values, cache=TRUE>>=
+wd <- getwd()
+setwd("../../")
+source("portfolio/analysis/prep-stan-model-matrix.R")
+rows_dat <- nrow(mm)
+setwd(wd)
+@
+
 \maketitle
@@
 \section{Hierarchical model}
+
+Rows of data: \Sexpr{rows_dat}.
 
@@
 
+\begin{table}[h!]
+  \centering
+  \caption{Caption.}
+  \label{tab:strategies}
+  \begin{tabular}{ll}
+  \toprule
+\input{../figs/strategy-table.tex}
+  \end{tabular}
+\end{table}
+
 \begin{figure}[htbp]
 \begin{center}
-\includegraphics[width=\textwidth]{../figs/stan-str-posteriors-dot.pdf}
+\includegraphics[width=5in]{../figs/stan-main-effects.pdf}
+\caption{Caption here}
+\label{fig:ts}
+\end{center}
+\end{figure}
+
+\clearpage
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=1.05\textwidth]{../figs/stan-str-posteriors-dot.pdf}
 \caption{Caption here}

### 2016-08-13 portfolio/text/ms.Rmd
Add text skeleton
Integrate output values

@@
----
-title: Title here
-author: |
-  | Authors here
-date: ""
-geometry: margin=1.1in
-fontfamily: libertine
-fontsize: 12pt
-output: pdf_document
----
+```{r set-knitr-options, cache=FALSE, echo=FALSE}
+library("knitr")
+opts_chunk$set(cache=TRUE, echo=FALSE, message=FALSE, autodep=TRUE,
+  results='hide')
+opts_chunk$set(warning=FALSE, message=FALSE, cache.path="../cache/",
+  cache.lazy=TRUE)
+```
+
+```{r get-dat, cache=FALSE}
+load("../data-generated/output-values.rda")
+```
 
@@
 
+[@kasperski2013]
+[@simpson1949]
+[@sethi2012b]
+[@sethi2014]
+
 # Methods
+
+[@standevelopmentteam2015]
 
@@
 
-# References
+\bibliographystyle{apalike}
+\bibliography{refs}
+
+\clearpage
 
 # Figures
-
-\clearpage
 

### 2016-08-13 portfolio/text/supp.Rnw
Integrate output values

--- 2016-08-13 portfolio/text/supp.tex
@@
 
+<<set-knitr-options, cache=FALSE, echo=FALSE>>=
+library("knitr")
+opts_chunk$set(cache=TRUE, echo=FALSE, message=FALSE, autodep=TRUE,
+  results='hide')
+opts_chunk$set(warning=FALSE, message=FALSE, cache.path="../cache/",
+  cache.lazy=TRUE)
+@
+
+<<get-values, cache=FALSE>>=
+load("../data-generated/output-values.rda")
+@
+
 \maketitle
@@
 
- % mm <- model.matrix(log(revenue) ~ -1 + b1(spec_change) + b2(spec_change) +
-      % days_change + b1(spec_change):days_change + b2(spec_change):days_change,
+Rows of data: \Sexpr{nrows}.
+Number of permit holders: \Sexpr{npholders}.
+Range of years: \Sexpr{range_yrs}.
+Number of strategies: \Sexpr{nstr}
+
+\citep{gelman2014}
 
@@
 We fit our models in Stan version 2.11.1. \ldots
+\citep{hoffman2014, standevelopmentteam2015a, carpenter2015a, standevelopmentteam2015}
 
@@
 
+\begin{table}[h!]
+  \centering
+  \caption{Caption.}
+  \label{tab:strategies}
+  \begin{tabular}{ll}
+  \toprule
+\input{../figs/strategy-table.tex}
+  \end{tabular}
+\end{table}
+
 \begin{figure}[htbp]
 \begin{center}
-\includegraphics[width=\textwidth]{../figs/stan-str-posteriors-dot.pdf}
+\includegraphics[width=5in]{../figs/stan-main-effects.pdf}
 \caption{Caption here}
-\label{fig:ts}
+\label{fig:fe}
+\end{center}
+\end{figure}
+
+\clearpage
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=1.05\textwidth]{../figs/stan-str-posteriors-dot.pdf}
+\caption{Caption here}
+\label{fig:re}
 \end{center}

### 2016-08-13 portfolio/analysis/3.5-output-values.R

@@ -1,0 +1,39 @@
+rm(list = ls())
+source("portfolio/analysis/prep-stan-model-matrix.R")
+load("portfolio/data-generated/fig1-dat.rda")
+
+library(dplyr)
+
+nrows <- nrow(mm)
+npholders <- length(unique(dat$p_holder))
+range_yrs <- range(dat$year)
+nstr <- length(unique(dat$strategy))
+rvers <- paste(R.version$major, R.version$minor, sep = ".")
+
+prop_single_all <- filter(out_permits, jk == "x")$prop_single
+min_prop_single <- min(prop_single_all)
+yr_min_prop_single <- filter(out_permits, jk == "x")$year[prop_single_x == min_prop_single]
+min_prop_single <- min_prop_single %>% round(2) * 100
+max_prop_single <- filter(out_permits, jk == "x")$prop_single %>%
+  max() %>% round(2) * 100
+curr_prop_single_nohal <- filter(out_permits,
+  jk == "B")$prop_single[length(unique(out_permits$year))] %>% round(2) * 100
+curr_prop_single_nosalm <- filter(out_permits,
+  jk == "S")$prop_single[length(unique(out_permits$year))] %>% round(2) * 100
+
+spdiv_all <- filter(out_div, jk == "x")$mean_spdiv
+max_spdiv_all <- max(spdiv_all)
+yr_max_spdiv_all <- filter(out_div, jk == "x")$year[spdiv_all == max_spdiv_all]
+max_spdiv_all <- max_spdiv_all %>% round(2) * 100
+
+min_spdiv_all <- min(spdiv_all)
+yr_min_spdiv_all <- filter(out_div, jk == "x")$year[spdiv_all == min_spdiv_all]
+min_spdiv_all <- min_spdiv_all %>% round(2) * 100
+
+spdiv_salm <- filter(out_div, jk == "S")$mean_spdiv
+max_spdiv_salm <- max(spdiv_salm)
+yr_max_spdiv_salm <- filter(out_div, jk == "S")$year[spdiv_salm == max_spdiv_salm]
+max_spdiv_salm <- max_spdiv_salm %>% round(2) * 100
+
+rm(cullDat, mm, mm2, dat, unique_holders)
+save.image(file = "portfolio/data-generated/output-values.rda")

### 2016-08-14 portfolio/text/ms.Rmd
Develop text

@@
 
+# Abstract
+
 # Introduction
@@
 [@standevelopmentteam2015]
+[@rcoreteam2016]
 
@@
 \includegraphics[width=0.7\textwidth]{../figs/fig1-jk.pdf}
-\caption{Caption here}
+\caption{
+Trends in revenue and diversification over time.
+(A) Map of fishing regions in Alaska, USA.
+(B) Total revenue in billions of 20XX U.S. dollars from `r range_yrs[1]` to
+`r range_yrs[2]` for multiple (grey) and single (blue) permit holders.
+(C) Proportion of single permit holders for all fisheries (thick black line)
+and with permit holders omitted each year who fished any of the
+`r n_jacknifed` top species groups (thin lines).
+(D) Mean effective diversity of species fished (Simpson's diversity index) 
+for all fisheries (thick black line)
+and with permit holders omitted each year who fished any of the
+`r n_jacknifed` top species groups (thin lines).
+}
 \label{fig:ts}
@@
 \includegraphics[width=0.9\textwidth]{../figs/stan-gg-spoke2.pdf}
-\caption{Caption here}
+\caption{Adding permits reduces risk.
+Two examples of expected revenue variability 
+and median revenue for
+owners of (A) individual halibut and (B) gillnet herring roe permits
+(light blue)
+compared with owners of the same permit plus one more (medium blue)
+or two more permits (dark blue).
+On the y-axis, dots and lines represent posterior median and 50\%
+credible intervals.}
 \label{fig:spokes}
@@
 \includegraphics[width=0.7\textwidth]{../figs/stan-across-strategy-variability.pdf}
-\caption{Caption here}
+\caption{Fishing strategies with higher species diversity
+have lower expected revenue variability.
+On the y-axis, dots, thick lines, and thin lines represent posterior medians,
+50\%, and 95\% credible intervals for
+the `r nstr` most common strategies.
+Thick black line, dark, and light grey shaded regions
+indicate median, 50\%, and 95\% credible intervals
+of strategy-level regression built into the heirarchical model.
+}
 \label{fig:across}
@@
 \includegraphics[width=1.1\textwidth]{../figs/stan-offset-break-anti-spaghetti.pdf}
-\caption{Caption here}
+\caption{Effects of generalizing and specializing from year to year for
+individual permit holders across the `r nstr` most common strategies.}
 \label{fig:within}

### 2016-08-14 portfolio/text/ms.Rmd
Edit captions

@@
 \includegraphics[width=0.9\textwidth]{../figs/stan-gg-spoke2.pdf}
-\caption{Adding permits reduces risk.
+\caption{Diversifying permits reduces risk and usually increases revenue.
 Two examples of expected revenue variability 
@@
 \includegraphics[width=0.7\textwidth]{../figs/stan-across-strategy-variability.pdf}
-\caption{Fishing strategies with higher species diversity
+\caption{Individuals who choose
+permits with higher species diversity
 have lower expected revenue variability.
@@
 50\%, and 95\% credible intervals for
-the `r nstr` most common strategies.
-Thick black line, dark, and light grey shaded regions
+the `r nstr` most common permit strategies.
+Thick black line, and dark and light grey shaded regions
 indicate median, 50\%, and 95\% credible intervals
-of strategy-level regression built into the heirarchical model.
+of strastegy-level regression built into the heirarchical model.
 }
@@
 \includegraphics[width=1.1\textwidth]{../figs/stan-offset-break-anti-spaghetti.pdf}
-\caption{Effects of generalizing and specializing from year to year for
-individual permit holders across the `r nstr` most common strategies.}
+\caption{Effects of (A) generalizing and (B) specializing from year to year for
+individual permit holders across the `r nstr` most common strategies.
+Changing species diversity from year to year 
+is associated with higher variability
+for individuals within the strategies above the zero line.
+Generalizing (A) or specializing (B) is associated with
+greater expected revenue to the right of the zero line
+and lower expected revenue to the left of the zero line.
+These effects are estimated holding
+fishing season length constant from year to year.
+}
 \label{fig:within}

### 2016-08-16 portfolio/analysis/1-create-modelling-data.Rmd
Keep track of sample sizes during filtering

@@
     fig_caption: yes
-  html_document: default
 ---
@@
   substr(cfec$p_fshy[idx], 3, length(cfec$p_fshy[idx])), sep="")
-
 ```
@@
 dim(cfecnew)[1]
+ndat <- list()
+ndat$initial <- nrow(cfecnew)
 ```
@@
 permit_rev$permit.rev = permit_rev$permit.rev/permit_rev$tot.rev # fraction of total
+ndat$rev_b4_10perc <- sum(permit_rev$permit.rev)
 permit_rev = permit_rev[permit_rev$permit.rev < 0.1,]
+ndat$rev_after_10perc <- sum(permit_rev$permit.rev)
 
@@
 nrow(cfecnew)
+ndat$before_10_perc <- nrow(cfecnew)
 cfecnew = cfecnew[-which(paste(cfecnew$p_holder, cfecnew$year, cfecnew$p_fshy) %in% 
@@
 nrow(cfecnew)
+ndat$nrow_over_10_perc <- nrow(cfecnew)
 
@@
 cfecAnnual <- as_data_frame(cfecAnnual)
-
-## Filtering out people with 10+ years of fishing. Drops rows from 13.7 to 10 million.
-## (now doing this at the beginning of the modeling scripts)
-# cfecAnnual = group_by(cfecAnnual, p_holder) %>%
-#   mutate(nyr = length(unique(year))) %>%
-#   filter(nyr >= 10)
+ndat$annual <- nrow(cfecAnnual)
 
@@
 saveRDS(cfecAnnual.diff, file="portfolio/data-generated/cfec-diff-for-modeling.rds")
+ndat$annual_diff <- nrow(cfecAnnual.diff)
+saveRDS(ndat, file = "portfolio/data-generated/ndat.rds")
 ```

### 2016-08-16 portfolio/text/supp.Rnw
Develop text
Keep track of sample sizes during filtering

@@
 \usepackage{amsmath}
-\usepackage{longtable}
-\usepackage{multirow}
-\usepackage{array}
-\usepackage{tabu} % for spacing between rows in longtable
 \usepackage{url}
 \usepackage{mathtools}
-\usepackage{bm}
 \usepackage{lineno}
@@
   raised=-.13em]{superiors}
-
-\mathchardef\mhyphen="2D % math hyphen
 
@@
 
+\section{Data filtering}
+
+<<nothing, eval=FALSE>>=
+permit_rev = group_by(cfecnew, p_holder, year, p_fshy) %>%
+  summarize(permit.rev = sum(g_earn)) %>%
+  as_data_frame() %>%
+  group_by(p_holder, year) %>%
+  mutate(tot.rev = sum(permit.rev))
+permit_rev$permit.rev = permit_rev$permit.rev/permit_rev$tot.rev # fraction of total
+permit_rev = permit_rev[permit_rev$permit.rev < 0.1,]
+
+# Delete these minimal earnings
+cfecnew = cfecnew[-which(paste(cfecnew$p_holder, cfecnew$year, cfecnew$p_fshy) %in%
+  paste(permit_rev$p_holder, permit_rev$year, permit_rev$p_fshy)), ]
+
+cfecnew <- filter(cfecnew, p_spec != "Freshwater") # only 1400 rows
+@
+
+Rows of data initial: \Sexpr{round(ndat$initial/1e6,2)} million
+
+Rows of data annual: \Sexpr{ndat$annual}
+
+Rows of data diff: \Sexpr{ndat$annual_diff}
+
+% \Sexpr{ndat$rev_b4_10perc/1e9}
+% \Sexpr{round(ndat$rev_after_10perc/1e9,2)}
+
+\Sexpr{ndat$before_10_perc}
+
+Rows of data after 10,000 cull: \Sexpr{ndat$large_only}
+
+10,000 cull: rev:
+\Sexpr{round(1-ndat$rev_large_only/ndat$rev_b4_large_only,2)*100}
+
+\Sexpr{round(ndat$rev_large_only/1e9,2)}
+
+\section{Assigning strategies}
+
+\Sexpr{nstrat$enough_pholders}
+
+\Sexpr{round(ndat$rev_enough_pholders/1e9,2)}
+
+\Sexpr{nstrat$no_river_exp}
+
+\Sexpr{round(ndat$rev_no_river_exp/1e9,2)}
+
+Number of strategies: \Sexpr{nstr}
+
+Rows of data before no singles: \Sexpr{ndat$before_nosingles}
+
+% \Sexpr{ndat$b4_no_na_strat}
+%
+% \Sexpr{ndat$rev_b4_no_na_strat}
+%
+% \Sexpr{ndat$after_no_na_strat}
+
+lost \% rev no strat:
+\Sexpr{round(ndat$rev_after_no_na_strat/ndat$rev_b4_no_na_strat,2)*100}
+
+lost rows no strat:
+\Sexpr{round(ndat$after_no_na_strat/ndat$b4_no_na_strat,2)*100}
+
+% \Sexpr{ndat$rev_after_no_na_strat}
+%
+% \Sexpr{ndat$rev_final}
+
+Rows of data: \Sexpr{ndat$final}
+
+Number of permit holders: \Sexpr{npholders}
+
+Range of years: \Sexpr{range_yrs}
+
+Number of strategy-years: \Sexpr{nstr_yrs}
+
 \section{Hierarchical model}
 
-Rows of data: \Sexpr{nrows}.
-Number of permit holders: \Sexpr{npholders}.
-Range of years: \Sexpr{range_yrs}.
-Number of strategies: \Sexpr{nstr}
+Overview... why first diff, why 2 part
 
-\citep{gelman2014}
+Negatives are that it uses a smaller subset of the data;
+positives are
+that it doesn't require estimating people-specific random effects (n = 3000+)
+and removes time effects.
 
@@
 
+\citep{gelman2014}
+\citep{rcoreteam2016}
+
 We checked our models with a two-part modelling approach
@@
 to quickly iterate and build intuition about the models.
-
-\subsection{points to include}
-
-Negatives are that it uses a smaller subset of the data;
-positives are
-that it doesn't require estimating people-specific random effects (n = 3000+)
-and removes time effects.
 

### 2016-08-16 portfolio/analysis/3.5-output-values.R

@@ -5,13 +5,15 @@ load("portfolio/data-generated/fig1-dat.rda")
 library(dplyr)
 
 nums <- c("one" = 1, "two" = 2, "three" = 3, "four" = 4, "five" = 5, "six" = 6,
-  "seven" = 7, "eight" = 8, "nine" = 9, "ten" = 10, "11" = 11, "12" = 12, "13" = 13)
+  "seven" = 7, "eight" = 8, "nine" = 9, "ten" = 10, "11" = 11, "12" = 12, "13" = 13,
+  "14" = 14, "15" = 15, "16" = 16)
 
 nrows <- nrow(mm)
 npholders <- length(unique(dat$p_holder))
 range_yrs <- range(dat$year)
 nstr <- length(unique(dat$strategy))
 rvers <- paste(R.version$major, R.version$minor, sep = ".")
+nstr_yrs <- length(unique(dat$strategy_year))
 
 prop_single_all <- filter(out_permits, jk == "x")$prop_single
 min_prop_single <- min(prop_single_all)
@@ -41,5 +43,9 @@ max_spdiv_salm <- max_spdiv_salm %>% round(2) * 100
 n_jacknifed <- length(unique(out_div$jk)) - 1
 n_jacknifed <- names(nums[n_jacknifed])
 
-rm(cullDat, mm, mm2, dat, unique_holders)
+ndat <- readRDS("portfolio/data-generated/ndat.rds")
+ndat_cull <- readRDS("portfolio/data-generated/ndat-cull.rds")
+ndat <- c(ndat, ndat_cull)
+nstrat <- readRDS("portfolio/data-generated/nstrat-cull.rds")
+rm(cullDat, mm, mm2, dat)
 save.image(file = "portfolio/data-generated/output-values.rda")

### 2016-08-16 portfolio/analysis/cull-dat.R

@@ -12,7 +12,11 @@ if(diff==TRUE) dat = readRDS(file="portfolio/data-generated/cfec-diff-for-modeli
 #  filter(nyr >= 5)
 
 # Filters: remove people-year combinations making < $5000
+ndat <- list()
+ndat$rev_b4_large_only <- sum(dat$revenue)
 dat = dat[which(dat$revenue >= 10000), ]
+ndat$large_only <- nrow(dat)
+ndat$rev_large_only <- sum(dat$revenue)
 
 # note: grouping here is based on strategies defined by permits
 dat$strategy = dat$strategy_permit
@@ -34,19 +38,25 @@ dat <- dat %>% mutate(length = ifelse(is.na(length), median_length_permit, lengt
 
 
 #2. For each person-year combination, we created 'strategies' by concatenating
-# all permits fished, and only retaining 'strategies' with >= 200 data points. Previously,
-# this was done by people-year, but Eric changed this to be including > 200 people / strategy
+# all permits fished, and only retaining 'strategies' with >= 100 data points. Previously,
+# this was done by people-year, but Eric changed this to be including > 100 people / strategy
 top_strategies = group_by(dat, strategy) %>%
   summarize(n = length(unique(p_holder)),
     earn=sum(revenue)) %>%
-  filter(n>=100) # note -- the code should work fine with this >= 100 too
+  filter(n>=100)
+nstrat <- list()
+nstrat$enough_pholders <- nrow(top_strategies)
+ndat$rev_enough_pholders <- sum(top_strategies$earn)
 
 # A few of these can be deleted (small in river fisheries, experimental, etc)
 top_strategies = top_strategies[-which(top_strategies$strategy %in% c("S04X","S04Y","Z12B","R18B")),]
+nstrat$no_river_exp <- nrow(top_strategies)
+ndat$rev_no_river_exp <- sum(top_strategies$earn)
 
 #3. We then tabulated the permits that make up these strategies, and there are only
 # 56.
 top_permits = data.frame("orig"=names(table(unlist(lapply(top_strategies$strategy, strsplit, " ")))))
+nstrat$top_permits <- nrow(top_permits)
 
 # 4. of these top permits, we proceeded to group permits that were targeting a single species
[…]

### 2016-08-16 portfolio/text/ms.Rmd
Update references and start listing references in ms

@@
 
+What's been done
+
 [@kasperski2013]
-[@simpson1949]
 [@sethi2012b]
+[@sethi2010b]
 [@sethi2014]
+[@perruso2005]
+[@hilborn2001]
+[@edwards2004]
+[@trenkel2013]
+[@mumford2009]
+[@oostenbrugge2002]
+[@baldursson1997]
+[@jardine2015]
+[@sanchirico2008]
+
+Multiple ways one can diversify: 
+buy another permit, pick diverse permit, diversify within existing constraints.
 
@@
 [@rcoreteam2016]
+[@simpson1949]
 

### 2016-08-16 portfolio/text/anderson-etal-rev-pfx.tex
Develop text
Start adding affiliations

@@
 
-\title{Benefits and costs of income diversification for 
-  Alaskan fishermen
-\ldots}
+\title{Portfolios of fishing permits and species
+reduce risk but adjusting fishing strategy is risky}
 \author{
-Sean C. Anderson, 
-\ldots}
+Sean C. Anderson$^1$,
+Eric J. Ward$^2$,
+Milo D. Adkison$^3$,
+Anne H. Beaudreau$^3$,\\
+Richard E. Brenner$^4$,
+Alan C. Haynie$^5$,
+Andrew O. Shelton$^2$,\\
+Jennifer C. Shriver,
+Jordan T. Watson
+}
+
 \date{}
@@
 \maketitle
+
+\noindent
+$^1$School of Aquatic and Fishery Sciences, Box 355020,
+University of Washington, Seattle, WA 98195, USA
+
+\noindent
+$^2$Conservation Biology Division, Northwest Fisheries Science Center,
+National Marine Fisheries Service, National Oceanographic and
+Atmospheric Administration, 2725 Montlake Blvd E, Seattle, WA 98112, USA
+
+\noindent
+$^5$Alaska Fisheries Science Center,
+National Marine Fisheries Service, National Oceanographic and
+Atmospheric Administration, Bldg 4,
+7600 Sand Point Way NE, Seattle, WA, 98115, USA
+
+\noindent
+Affiliations \ldots
 

### 2016-08-16 salmon/analysis/salmon_paper.Rmd

@@
 output:
-  word_document: default
   pdf_document:
@@
 - It's ok to be specialized if both abundance and prices are high. If one or the other is high, that may be ok too -- but if abundance and prices drop, things are in bad quadrant.  
+- Late 1980s/early 1990s saw pretty strong declines in pink prices + high abundance (tie in with crash of foreign currencies)  
 - Substitution of high value species (Chinook, sockeye) for lower value species (pink,chum) has analogs in terrestrial farming (e.g.
@@
 ### Hatchery programs were designed to reduce variation in harvest. Has that happened?  
+Below we quantified 3 trends in variation: year-to-year variation at the entire fishery level, year-to-year variation at the individual level, and variability between people
 
@@
 Projecting future prices are complex -- Gunnar Knapp has a good presentation on this http://www.adfg.alaska.gov/static/fishing/PDFs/commercial/gk_trends_4-12-12.pdf
+
+- Ability to specialize primarily driven by forces beyond control of fishermen.
+But in some situations, diversification is possible by extending fishing season to exploit additional spp (fishing earlier for chum in PWS S01E).
+
 - We're seeing increased specialization within permits.
@@
 dat.salmon = dat.diff[dat.diff$strategy_permit%in%c("S01A", "S01E", "S01K", "S03A", "S03E","S01M","S04K"),]
+
+# Restrict analysis to people who keep same strategy - complicates days, etc
+dat.salmon = dat.salmon[which(dat.salmon$strategy_permit==dat.salmon$strategy_permit.prev),]
 
@@
 
+```{r, echo=FALSE, eval = FALSE}
+# These are the run  timing plots per Milo's questions
+
+# for 2008, create cumulative density plots of salmon landings by calday
+pinkchum = cfec.raw[cfec.raw$p_fshy=="S 01E" & cfec.raw$spec%in%c("PINK","CHUM","SOCK"),]
+#pinkchum = cfec.raw[cfec.raw$p_fshy=="S 03T" & cfec.raw$spec%in%c("PINK","SOCK","CHNK"),]
+
+# aggregate by calendar day
+pinkchum$month = as.numeric(substr(pinkchum$day, 1, 2))
+pinkchum$day = as.numeric(substr(pinkchum$day, 4, 5))
+pinkchum$jday = date::mdy.date(pinkchum$month, pinkchum$day, pinkchum$year) - date::mdy.date(1,1,pinkchum$year)
+
+group_by(pinkchum, jday, year) %>% 
+  summarize(pink = sum(g_earn[spec=="PINK"]), 
+    chum = sum(g_earn[spec=="CHUM"]),
+    sock = sum(g_earn[spec=="SOCK"])) %>% 
+  group_by(year) %>% 
+  mutate(tot = sum(pink) + sum(chum) + sum(sock), pink = pink/tot, chum = chum/tot, sock = sock/tot) %>% 
+ggplot(aes(jday, pink)) + geom_line(color="pink") + geom_line(aes(jday, chum), color="darkgreen") + geom_line(aes(jday, sock), color="lightblue") + facet_wrap(~ year, scale="free")
+
+# make plot of all earnings or weight by jday
+salm = cfec.raw[cfec.raw$p_fshy=="S 01E",]
+salm$month = as.numeric(substr(salm$day, 1, 2))
+salm$day = as.numeric(substr(salm$day, 4, 5))
+salm$jday = date::mdy.date(salm$month, salm$day, salm$year) - date::mdy.date(1,1,salm$year)
+
+g = group_by(salm, jday, year) %>% 
+  summarize(earn = sum(g_earn), 
+    lbs = sum(g_pounds), 
+    pink = sum(g_earn[spec=="PINK"])/earn,
+    sock = sum(g_earn[spec=="SOCK"])/earn,
+    chum = sum(g_earn[spec=="CHUM"])/earn,
+    coho = sum(g_earn[spec=="COHO"])/earn,
+    chnk = sum(g_earn[spec=="CHNK"])/earn) %>% 
+ggplot(g, aes(jday, (earn))) + geom_line(colour = rgb(g$sock+g$coho+g$chnk, g$chum, g$pink)) + facet_wrap(~ year, scale="free_y")
+
+```
+
 ```{r, echo = FALSE, fig.cap="SEAK purse seine fishery", cache=TRUE, fig.pos="placeHere"}

### 2016-08-17 portfolio/text/supp.Rnw
Supp figs

@@
 
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=1\textwidth]{../figs/stan-gg-spoke-all.pdf}
+\caption{Caption here}
+\label{fig:spoke-all}
+\end{center}
+\end{figure}
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=1\textwidth]{../figs/stan-offset-panel-mean-resid.png}
+\caption{Caption here}
+\label{fig:resid}
+\end{center}
+\end{figure}
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=1\textwidth]{../figs/stan-offset-panel-mean-resid-down.png}
+\caption{Downside only}
+\label{fig:resid-downside}
+\end{center}
+\end{figure}
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=1\textwidth]{../figs/strategy-year-effects.pdf}
+\caption{Caption here}
+\label{fig:str-yrs}
+\end{center}
+\end{figure}
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=0.7\textwidth]{../figs/lmer-vs-stan.pdf}
+\caption{Caption here}
+\label{fig:lmer-stan}
+\end{center}
+\end{figure}
+
 \end{document}

### 2016-08-17 portfolio/analysis/3.3-lmer-stan-comparison.R

@@ -1,0 +1,71 @@
+library(dplyr)
+library(ggplot2)
+source("portfolio/analysis/prep-stan-model-matrix.R")
+load("portfolio/data-generated/m.rda")
+devtools::load_all("pfxr")
+library(lme4)
+
+b1 <- function(x, bp = 0) ifelse(x < bp, x, 0)
+b2 <- function(x, bp = 0) ifelse(x < bp, 0, x)
+
+m_lmer <- lmer(
+  log(revenue) ~ -1 + b1(spec_change) + b2(spec_change) +
+    days_change + b1(spec_change):days_change + b2(spec_change):days_change +
+    (-1 + b1(spec_change) + b2(spec_change) |strategy) + (1|strategy_year),
+  data = dat, offset = log(revenue.prev))
+
+re_lmer <- coef(m_lmer)$strategy
+re_lmer$strategy <- row.names(re_lmer)
+row.names(re_lmer) <- NULL
+library(tidyr)
+
+p <- stanhelpers::extract_df(m)
+colnames(p$coef_g0_strategy) <- md$strategy
+colnames(p$coef_g1_strategy) <- md$strategy
+colnames(p$coef_g2_strategy) <- md$strategy
+colnames(p$coef_b1_strategy) <- md$strategy
+colnames(p$coef_b2_strategy) <- md$strategy
+
+# g0 <- tidyr::gather(p$coef_g0_strategy, term, posterior) %>%
+#   mutate(parameter = "g0")
+# g1 <- tidyr::gather(p$coef_g1_strategy, term, posterior) %>%
+#   mutate(parameter = "g1")
+# g2 <- tidyr::gather(p$coef_g2_strategy, term, posterior) %>%
+#   mutate(parameter = "g2")
+b1 <- tidyr::gather(p$coef_b1_strategy, term, posterior) %>%
+  mutate(parameter = "b1")
+b2 <- tidyr::gather(p$coef_b2_strategy, term, posterior) %>%
+  mutate(parameter = "b2")
+po <- bind_rows(g0, g1, g2, b1, b2)
+
+pos <- group_by(po, term, parameter) %>% summarise(
+  l = quantile(posterior, probs = 0.025),
+  l.5 = quantile(posterior, probs = 0.25),
+  m = quantile(posterior, probs = 0.5),
+  u.5 = quantile(posterior, probs = 0.75),
[…]

### 2016-08-17 portfolio/analysis/3.3-plot-spokes.R

@@ -6,13 +6,16 @@ devtools::load_all("pfxr")
 
 # -----------------------------------------------
 # spoke plots:
-mr <- group_by(dat, strategy) %>% summarise(strategy_med_rev = median(revenue)/1e3)
+mr <- group_by(dat, strategy) %>%
+  summarise(strategy_med_rev = median(revenue)/1e3,
+    n_permits = npermit[1])
 b <- broom::tidy(m, conf.int = T, estimate.method = "median", conf.level = 0.5)
 md2 <- filter(b, grepl("coef_g0_strategy", term)) %>%
   mutate(strategy_id = 1:n()) %>% inner_join(md) %>% inner_join(mr)
 
 
 permit_plot <- function(permit) {
+  message(permit)
   gre1 <- paste0(permit, " ")
   gre2 <- paste0(permit, "$")
 
@@ -68,7 +71,7 @@ sp <- plyr::ldply(c(
   "G01",
   "G34",
   "K91",
-  "T91Q",
+  "T09",
   "B61B",
   "C61B",
   "S01",
@@ -139,3 +142,57 @@ pl <- filter(sp, !single_permit %in% c("S03", "K91")) %>%
 # print(pl)
 # print(pl)
 ggsave("portfolio/figs/stan-gg-spoke2.pdf", width = 7, height = 3)
+
+
+# ------------------------------------------------------
+# more for SI
+sp <- plyr::ldply(c(
+  "G01",
+  "G34",
+  "K91",
+  "T09",
+  "B61B",
+  "C61B",
+  "S01",
+  "S03"
+),  permit_plot)
+
+labs <- readr::read_csv("data/strategies-labels.csv") %>%
+  rename(single_permit_clean = description,
+    single_permit = strategy) %>%
+  select(-str_label)
+sp <- inner_join(sp, labs, by = "single_permit")
+
+ns <- dat %>% group_by(strategy) %>%
+  summarise(nn = length(unique(p_holder)))
+sp <- inner_join(sp, ns)
+
+pl <- sp %>%
+  mutate(str_label = ifelse(nn > 20, str_label, NA)) %>%
+  ggplot(
+    aes(strategy_med_rev, exp(estimate), yend = exp(g0_less), xend = b0_less,
+      label = str_label, colour = as.factor(n_permits))) +
+  geom_segment(aes(x = strategy_med_rev, xend = strategy_med_rev,
+    y = exp(conf.low), yend = exp(conf.high)), alpha = 0.5, size = 0.3,
[…]

### 2016-08-17 portfolio/analysis/3.3-residuals.R

@@ -1,0 +1,139 @@
+library(dplyr)
+library(ggplot2)
+source("portfolio/analysis/prep-stan-model-matrix.R")
+load("portfolio/data-generated/m.rda")
+devtools::load_all("pfxr")
+
+# devtools::install_github("seananderson/stanhelpers")
+p <- stanhelpers::extract_df(m)
+
+dat$str_label <- NULL
+dat <- inner_join(dat, select(md, strategy_id, str_label))
+
+# names(p$b_j) <- colnames(mm)
+# names(p$g_k) <- colnames(mm2)
+
+colnames(p$coef_g0_strategy) <- md$strategy
+colnames(p$coef_g1_strategy) <- md$strategy
+colnames(p$coef_g2_strategy) <- md$strategy
+colnames(p$coef_b1_strategy) <- md$strategy
+colnames(p$coef_b2_strategy) <- md$strategy
+
+library(tidyr)
+
+# g0_strategy <- gather(p$coef_g0_strategy[1,], strategy, g0_strategy)
+# g1_strategy <- gather(p$coef_g1_strategy[1,], strategy, g1_strategy)
+# g2_strategy <- gather(p$coef_g2_strategy[1,], strategy, g2_strategy)
+
+iters <- seq(1, nrow(p$b_j), 10)
+length(iters)
+out <- matrix(ncol = length(iters), nrow = nrow(dat))
+dat_res_select <- select(dat, strategy, str_yr_id, revenue.prev)
+for(i in seq_along(iters)) {
+  message(round(i/length(iters), 2)*100)
+  b1_strategy <- gather(p$coef_b1_strategy[iters[i],], strategy, b1_strategy)
+  b2_strategy <- gather(p$coef_b2_strategy[iters[i],], strategy, b2_strategy)
+  b0_str_yr <- gather(p$b0_str_yr[iters[i],], str_yr_id, b0_str_yr) %>%
+    mutate(str_yr_id = as.numeric(substring(str_yr_id, 11)))
+
+  dat_res <- dat_res_select
+  dat_res <- inner_join(dat_res, b0_str_yr, by = "str_yr_id")
+  dat_res <- inner_join(dat_res, b1_strategy, by = "strategy")
+  dat_res <- inner_join(dat_res, b2_strategy, by = "strategy")
+
+  eta <- as.matrix(mm) %*% t(as.matrix(p$b_j[iters[i],]))
+  eta <- eta + log(dat$revenue.prev)
+  eta <- eta + dat_res$b0_str_yr +
[…]

### 2016-08-18 portfolio/analysis/3.3-make-strategy-table.R
Filter out strategies we aren't using in the supplement table

@@ -1,8 +1,15 @@
 library(dplyr)
 d <- readr::read_csv("data/strategies-labels.csv")
-d <- select(d, str_label, description) %>%
-  rename(Label = str_label, Description = description) %>%
+
+load("portfolio/data-generated/diff-dat-stan.rda")
+d <- d %>% mutate(modeled = ifelse(strategy %in% md$strategy, "Yes", "No"))
+
+d <- select(d, str_label, description, modeled) %>%
+  rename(Label = str_label, Description = description,
+    `Over 100 permit holders` = modeled) %>%
   arrange(Label)
+d <- d %>% filter(`Over 100 permit holders` == "Yes") %>%
+  select(-`Over 100 permit holders`)
 
 xtable::print.xtable(
   xtable::xtable(d,

### 2016-08-18 portfolio/figs/strategy-table.tex
Update strategy table

@@
 Dun C & Dungeness crab \\ 
-  Fin - jig & Finfish, mechanical jig \\ 
+  Fin - jig & Finfish, mechanical jig , statewide \\ 
   Fin - ll & Finfish, longline, statewide \\ 
@@
   Fin - pot & Finfish, pot gear, statewide \\ 
-  Hal & Halibut \\ 
+  Hal - ll & Halibut, longline \\ 
+  Hal - troll & Halibut, hand troll \\ 
+  Herr - hand & Herring, hand pick \\ 
   Herr - pound & Herring, pound \\ 
   Herr roe - gill & Herring roe, gillnet \\ 
-  Herr roe - seine & Herring roe, purse seine \\ 
   King C & King crab \\ 
-  S - drift - CI, AP, BB (sock) & Salmon (mostly sockeye), drift gill net, CI, AP, BB \\ 
-  S - drift - PWS & Salmon, drift gill net, PWS \\ 
-  S - drift - SE & Salmon, drift gill net, Southeast Alaska \\ 
+  S - drift - CI, AP, BB (sock) & Salmon (mostly sockeye), drift gillnet, CI, AP, BB \\ 
+  S - drift - PWS & Salmon, drift gillnet, PWS \\ 
+  S - drift - SE & Salmon, drift gillnet, Southeast Alaska \\ 
   S - seine - AP & Salmon, purse seine, Alaska Penninsula \\ 
@@
   S - set - YAK & Salmon, set gillnet, Yakutat \\ 
-  S - set - YUK & Salmon, gillnet, lower Yukon \\ 
-  S - troll & Salmon , power troll, statewide \\ 
+  S - troll & Salmon, power troll, statewide \\ 
   Sab & Sablefish, longline, statewide \\ 
-  Sab - SE & Sablefish , longline, Southeast Alaska \\ 
   Sea cuc & Sea cucumber \\ 
   Shrimp & Shrimp \\ 
-  Tan C & Tanner crab, pot gear vessel under 60' \\ 
-  Tan C - large & Tanner crab, pot gear vessel over 60' \\ 
+  Tan C & Tanner crab, pot gear \\ 
    \bottomrule

### 2016-08-18 portfolio/text/supp.Rnw
Add draft captions to supplemental figures

@@
 \begin{equation}
+\label{eq:mean}
 \widehat{\log(R_{i,t})} = \beta_{0,j,t} +
@@
 \begin{equation}
+\label{eq:cv}
 \widehat{\sigma_{i,t}} =
@@
   \centering
-  \caption{Caption.}
+  \caption{Labels and descriptions of fishing strategies. In many cases, 
+a particular strategy represents a group of similar permits across areas 
+or vessel sizes. These strategies all had at least 100 
+permit holders.}
   \label{tab:strategies}
@@
 \includegraphics[width=5in]{../figs/stan-main-effects.pdf}
-\caption{Caption here}
+\caption{Main effects from the hierarchical model 
+predicting revenue (b or $\beta$) and variability (g or $\gamma$) in revenue.
+The coefficient $u_1$ represents a strategy level predictor
+relating mean species diversity of a strategy
+to the expected magnitude of residual error.}
 \label{fig:fe}
@@
 \begin{center}
-\includegraphics[width=1.05\textwidth]{../figs/stan-str-posteriors-dot.pdf}
-\caption{Caption here}
+\includegraphics[width=1.06\textwidth]{../figs/stan-str-posteriors-dot.pdf}
+\caption{Estimates of within strategy effects from the hierarchical model. 
+Panels refer in order to $\gamma_0$ (g0), $\gamma_1$ (g1), $\gamma_2$ (g2),
+$\beta_1$ (b1), and $\beta_2$ (b2) from Eqns \ref{eq:mean} and \ref{eq:cv}.
+Shown are medians (dots) and 50\% and 90\% credible intervals
+(thick and thin line segments).
+Strategies are ordered from high to low mean
+species diversity from top to bottom.}
 \label{fig:re}
@@
 \includegraphics[width=1\textwidth]{../figs/stan-gg-spoke-all.pdf}
-\caption{Caption here}
+\caption{The effect of adding a permit to ones strategy
+on estimated revenue variability and median revenue.
+This is an extended version of Fig.~2 that includes all major 
+strategy combinations.}
 \label{fig:spoke-all}
@@
 \includegraphics[width=1\textwidth]{../figs/stan-offset-panel-mean-resid.png}
-\caption{Caption here}
+\caption{The log-absolute value residuals from the mean-component model (Eq.~\ref{eq:mean}).
+It is these residuals that are modeled in the variance-component model
+(Eq.~\ref{eq:cv}).
+Breakpoint linear regression fits are overlaid in red for visual interpretation.}
 \label{fig:resid}
@@
 \includegraphics[width=1\textwidth]{../figs/stan-offset-panel-mean-resid-down.png}
-\caption{Downside only}
+\caption{The log-absolute value of downside (negative) residuals from the main
+component model.
+Figure looks qualitatively the same as Fig.~\ref{fig:resid} and so justifies
+using variability (positive and negative residuals) as a measure of risk.}
 \label{fig:resid-downside}
@@
 \includegraphics[width=1\textwidth]{../figs/strategy-year-effects.pdf}
-\caption{Caption here}
+\caption{Strategy-year intercept estimates ($\beta_{0,j,t}$ in
+Eq.~\ref{eq:mean}) from the hierarchical model. 
+Shown are 40 samples from the posterior distribution.}
 \label{fig:str-yrs}
@@
 
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[width=0.7\textwidth]{../figs/lmer-vs-stan.pdf}
-\caption{Caption here}
-\label{fig:lmer-stan}
-\end{center}
-\end{figure}
+% \begin{figure}[htbp]
+% \begin{center}
+% \includegraphics[width=0.9\textwidth]{../figs/lmer-vs-stan2.pdf}
+% \caption{Comparison of }
+% \label{fig:lmer-stan}
+% \end{center}
+% \end{figure}
 

### 2016-08-19 portfolio/analysis/3.3-plot-strategy-effects.R
Fix bj's and gj's plot

@@ -13,14 +13,17 @@ p <- stanhelpers::extract_df(m)
 names(p$b_j) <- colnames(mm)
 names(p$g_k) <- colnames(mm2)
 
+colnames(p$g0_strategy) <- md$strategy
 colnames(p$coef_g0_strategy) <- md$strategy
 colnames(p$coef_g1_strategy) <- md$strategy
 colnames(p$coef_g2_strategy) <- md$strategy
 colnames(p$coef_b1_strategy) <- md$strategy
 colnames(p$coef_b2_strategy) <- md$strategy
 
-g0 <- tidyr::gather(p$coef_g0_strategy, term, posterior) %>%
-  mutate(parameter = "g0")
+g0 <- tidyr::gather(p$g0_strategy, term, posterior) %>%
+  mutate(parameter = "g0", g0main = rep(p$g0[[1]], length(unique(md$strategy))))
+g0 <- mutate(g0, posterior = posterior + g0main) %>%
+  select(-g0main)
 g1 <- tidyr::gather(p$coef_g1_strategy, term, posterior) %>%
   mutate(parameter = "g1")
 g2 <- tidyr::gather(p$coef_g2_strategy, term, posterior) %>%
@@ -58,13 +61,16 @@ pos[pos$parameter == "g0", "line"] <- NA
 pos <- as_data_frame(pos)
 
 pos$parameter <- factor(pos$parameter, levels = c("g0", "g1", "g2", "b1", "b2"))
+pos$parameter <- factor(pos$parameter,
+  labels = c("gamma[0][j]", "gamma[1][j]", "gamma[2][j]", "beta[1][j]", "beta[2][j]"))
 pl <- ggplot(pos, aes(x = str_label, y = m)) +
   geom_point() +
   geom_pointrange(aes(ymin = l, ymax = u), size = 0.2) +
   geom_linerange(aes(ymin = l.5, ymax = u.5), size = 0.9) +
-  facet_wrap(~parameter, nrow=1) + ylab("Parameter estimate") +
+  facet_wrap(~parameter, nrow=1, labeller = label_parsed) +
+      ylab("Parameter estimate") +
   geom_hline(aes(yintercept = line), lty = 2, col = "grey40") +
   coord_flip() + theme_gg() +
   theme(panel.grid.major = element_line(colour = "grey93",
     size = 0.2), panel.grid.minor = element_blank()) + xlab("")
[…]

### 2016-08-19 portfolio/text/ms.Rmd
Edit captions and equations

@@
 (A) Map of fishing regions in Alaska, USA.
-(B) Total revenue in billions of 20XX U.S. dollars from `r range_yrs[1]` to
+(B) Total revenue in billions of 2009 U.S. dollars from `r range_yrs[1]` to
 `r range_yrs[2]` for multiple (grey) and single (blue) permit holders.
@@
 \includegraphics[width=0.7\textwidth]{../figs/stan-across-strategy-variability.pdf}
-\caption{Individuals who choose
-permits with higher species diversity
+\caption{Individuals who fish
+permits and permit combinations (permit strategies) with higher species diversity
 have lower expected revenue variability.
@@
 indicate median, 50\%, and 95\% credible intervals
-of strastegy-level regression built into the heirarchical model.
+of strategy-level regression built into the hierarchical model.
+Estimated variability represents expected standard deviation
+at an average of about
+`r mean_permit_days` combined fishing days across permits,
+and for a permit holder who does not change
+species diversity or days fished from year to year.
 }

### 2016-08-19 portfolio/text/supp.Rnw
Edit captions and equations

@@
 \begin{align}
-\gamma_{0,j}  &\sim \mathrm{N} \left(\gamma_0, \sigma^2_{\gamma_0}\right),\\
+\gamma_{0,j}  &\sim \mathrm{N} \left(\gamma_0 + \eta_1 M_j + \eta_2 A_j, \sigma^2_{\gamma_0}\right),\\
 \gamma_{1,j}  &\sim \mathrm{N} \left(\gamma_1, \sigma^2_{\gamma_1}\right),\\
@@
 \caption{Main effects from the hierarchical model
-predicting revenue (b or $\beta$) and variability (g or $\gamma$) in revenue.
-The coefficient $u_1$ represents a strategy level predictor
-relating mean species diversity of a strategy
-to the expected magnitude of residual error.}
+predicting revenue ($\beta$) and variability ($\gamma$) in revenue.
+The $\eta$ coefficients represents strategy-level predictors
+relating mean species diversity and mean days fished
+within a strategy
+(both centered by their means and divided by
+two times their standard deivations)
+to the magnitude of residual error.}
 \label{fig:fe}
@@
 \includegraphics[width=1.06\textwidth]{../figs/stan-str-posteriors-dot.pdf}
-\caption{Estimates of within strategy effects from the hierarchical model. 
-Panels refer in order to $\gamma_0$ (g0), $\gamma_1$ (g1), $\gamma_2$ (g2),
-$\beta_1$ (b1), and $\beta_2$ (b2) from Eqns \ref{eq:mean} and \ref{eq:cv}.
-Shown are medians (dots) and 50\% and 90\% credible intervals
+\caption{Estimates of within-strategy effects from the hierarchical model.
+Panels refer to parameters in Eqns \ref{eq:mean} and \ref{eq:cv}.
+Shown are medians (dots) and 50\% and 95\% credible intervals
 (thick and thin line segments).

### 2016-08-22 portfolio/text/ms.Rmd
Early draft of abstract, other text

@@
 ```
-
 ```{r get-dat, cache=FALSE}
@@
 
+\noindent
+Main message:
+Diversified portfolios of fishing permits and species reduce risk for fishermen,
+but changing species diversity can be risky, costly, and limited by regulations.
+
 # Abstract
+
+<!-- Fishing is inherently risky and we have long-sought regulations and fishing behaviors that reduce this risk. --> 
+<!-- The increasing the diversity of species fished has been proposed to reduce risk for fishermen -->
+
+Individuals relying on natural resources as their primary source of income
+often experience high levels of income-variability 
+due to the combined effects of environmental and price variation. 
+Fishing is a prime example and
+recent work has sought to identify regulations and fishing behaviours 
+that can reduce this variability. 
+A top candidate has been that fishing higher levels of species diversity
+decreases income variability, which is concerning 
+given that some fishing regulations and market forces
+may be increasingly promoting specialization. 
+However, it is unknown if this species-diversity effect 
+on income variability holds 
+across multiple approaches to diversifying catch,
+at the level of individual permit holders,
+and across all fisheries. 
+Here, we simultaneously estimate the effects of three 
+methods of diversification for Alaskan fishery permit holders
+on revenue and revenue variability:
+(1) owning multiple permits, 
+(2) owning permits that inherently fish higher species diversity, 
+and (3) adjusting species diversity within the constraints of a permit.
+We control for changes in effort and other 
+residual effects through time.
+We find that 
+fishing an additional permit is associated with lower variability (~XX--XX%) 
+and higher revenue,
+fishing permits with higher species diversity reduces risk 
+(XX--XX% per unit increase in Simpson's diversity index),
+but that changing species diversity within a
+permit's constraints has a fishery-dependent effect on revenue and 
+is usually (XX%) associated with increased revenue variability.
+Our results suggest that the benefit of species diversity
+on reducing revenue variability 
+is mostly an across-permit effect, not
+a benefit that can be easily harnessed by an individual permit holder.
+Effective methods of diversification --- 
+purchasing additional permits or switching permits ---
+is costly and often limited by regulations such as limited entry
+programs and individual transferable quotas (ITQs).
+Our results suggest ...
+
+\bigskip
+\noindent
+Extra:
+
+- there is often limited room for intentional diversification within a permit
+- in fact, specializing can be profitable within some fisheries
+
+<!-- Main message in 25 words or less: -->
+
+<!-- Take 1: -->
+<!-- - There is a trend towards specializing on single salmon permits in Alaska -->
+<!-- - Diversified portfolios of fishing permits and species reduce risk, -->
+<!--   but changing species diversity for an individual permit holder can be risky -->
+<!-- - The types of diversity needed to reduce risk are often restricted by -->
+<!--   cost and regulations -->
+
+<!-- Take 2: -->
+<!-- - The effect of species diversification on revenue and risk for fishermen -->
+<!-- depends on the type of diversification and the fishery. -->
+
+<!-- Introduction, Methods, Results, Discussion, Conclusion:
+<!-- 600, 900, 500, 800, 200, -->
+
+<!-- write out each paragraph’s main message in 15 words or less -->
+
+<!-- circulate skeleton ... this is the time to fix problems with flow, -->
+<!-- ideas/content, main message -->
 

### 2016-08-23 portfolio/text/ms.Rmd
Edit abstract

@@
 due to the combined effects of environmental and price variation. 
-Fishing is a prime example and
+Commercial fishermen are a prime example and
 recent work has sought to identify regulations and fishing behaviours 
@@
 at the level of individual permit holders,
-and across all fisheries. 
+and across different fishing strategies. 
 Here, we simultaneously estimate the effects of three 
@@
 We find that 
-fishing an additional permit is associated with lower variability (~XX--XX%) 
-and higher revenue,
+fishing an additional permit is associated with 
+lower variability (~XX--XX%) and higher revenue,
 fishing permits with higher species diversity reduces risk 
@@
 purchasing additional permits or switching permits ---
-is costly and often limited by regulations such as limited entry
+are costly and often limited by regulations such as limited entry
 programs and individual transferable quotas (ITQs).

### 2016-08-23 portfolio/text/supp.Rnw
Edit text

@@
 \label{eq:mean}
-\widehat{\log(R_{i,t})} = \beta_{0,j,t} +
+\widehat{\log(R_{i,t})} = \beta_{0,j,t} -
   \beta_{1,j} S_{i,t} I_{i,t} +
   \beta_{2,j} S_{i,t} I_{i,t} +
-  \beta_3 D_{i,t} +
+  \beta_3 D_{i,t} -
   \beta_4 S_{i,t} D_{i,t} I_{i,t} +
@@
 \widehat{\sigma_{i,t}} =
-\exp \left( \gamma_{0,j} + \eta_1 M_j +
+\exp \left( \gamma_{0,j} -
 \gamma_{1,j} S_{i,t} I_{i,t} +
 \gamma_{2,j} S_{i,t} I_{i,t} +
-\gamma_3 D_{i,t} +
+\gamma_3 D_{i,t} -
 \gamma_4 S_{i,t} D_{i,t} I_{i,t} +
-\gamma_5 S_{i,t} D_{i,t} I_{i,t}\right),
+\gamma_5 S_{i,t} D_{i,t} I_{i,t}\right).
 \end{equation}
 
-where $M_j$ represents the raw mean species diversity for strategy $j$
-and $\eta_1$ represents an estimated strategy-level predictor.
-In other words, $\gamma_{0,j} + \eta_1 M_j$ represents the intercept
-with a built-in strategy-level regression (Fig.~3)
-that incorporates parameter uncertainty.
 Again, $S_{i,t}$ and $I_{i,t}$ represent, respectively,
@@
 \begin{align}
-\gamma_{0,j}  &\sim \mathrm{N} \left(\gamma_0 + \eta_1 M_j + \eta_2 A_j, \sigma^2_{\gamma_0}\right),\\
+\gamma_{0,j}  &\sim \mathrm{N} \left(\gamma_0 + \eta_1 M_j + \eta_2 E_j, \sigma^2_{\gamma_0}\right),\\
 \gamma_{1,j}  &\sim \mathrm{N} \left(\gamma_1, \sigma^2_{\gamma_1}\right),\\
@@
 \end{align}
+
+The mean of the $\gamma_{0,j}$ coefficients is
+modelled with
+two strategy-level predictors.
+The symbols $M_j$ and $E_j$ represent the raw mean species diversity and
+mean combined fishing season length (effort) for strategy $j$.
+The coefficients $\eta_1$ and $\eta_2$ represent predictors
+in this strategy-level regression.
+The effect of $\eta_1$ is shown in Fig.~3.
 

### 2016-08-23 portfolio/text/supp.Rnw
Typos

@@
 
-The basic model is a heirarchical linear regression with Gaussian errors
+The basic model is a hierarchical linear regression with Gaussian errors
 predicting $\log$ revenue, $\log(R_{i,t})$, for
@@
 
-Whereas a typical heirarchical linear regression model
+Whereas a typical hierarchical linear regression model
 would estimate a single residual error scale term, $\sigma$,
 we model the scale of the residual error  with
-another heirarchical model with a similar form:
+another hierarchical model with a similar form:
 
@@
 The $\gamma$'s are estimated coefficients, three of which
-are allowed to vary by strategy, constrained by normal distributions
-and with estimated means:
+are allowed to vary by strategy, constrained by normal distributions:
 

### 2016-08-23 portfolio/text/ms.Rmd
Edit text
Edit text and abstract

@@
 Individuals relying on natural resources as their primary source of income
-often experience high levels of income-variability 
+often experience highly variable income levels 
 due to the combined effects of environmental and price variation. 
-Commercial fishermen are a prime example and
+Commercial fishers are a prime example and
 recent work has sought to identify regulations and fishing behaviours 
@@
 A top candidate has been that fishing higher levels of species diversity
-decreases income variability, which is concerning 
-given that some fishing regulations and market forces
-may be increasingly promoting specialization. 
+decreases income variability.
 However, it is unknown if this species-diversity effect 
@@
 is usually (XX%) associated with increased revenue variability.
+At the same time, Alaskan commercial fisheries
+have increasingly focussed on 
+single-permit (often salmon) fishing strategies
+with single-permit fishers earning 
+more than multiple-permit fishers
+for the first time in 2014.
 Our results suggest that the benefit of species diversity
@@
 programs and individual transferable quotas (ITQs).
-Our results suggest ...
+Take home ...
 
@@
 
+<!-- , which is concerning  -->
+<!-- given that some fishing regulations and market forces -->
+<!-- may be increasingly promoting specialization. -->
 <!-- Main message in 25 words or less: -->
@@
 fishing season length constant from year to year.
+Dots and line segments represent medians and 
+50\% credible intervals of posteriors.
 }

### 2016-08-23 portfolio/text/supp.Rnw
Edit text and abstract

@@
 heteroskedastic and therefore improperly weights individual data points, and
-(3) does not allow for proper modelling of the strategy-level predictor,
-$\eta_1$. Nonetheless, it provides a check that our results
+(3) does not allow for proper modelling of the strategy-level predictors,
+$\eta_1$ and $\eta_2$. Nonetheless, it provides a check that our results
 from the Stan model are approximately correct and allowed us
 to quickly iterate and build intuition about the models.
+A version of the Stan model without the variance-component
+matched the estimates of the same model in lme4 almost exactly.
 

### 2016-08-24 portfolio/text/ms.Rmd
Copy editing text

@@
 Individuals relying on natural resources as their primary source of income
-often experience highly variable income levels 
+often experience high levels of income variability 
 due to the combined effects of environmental and price variation. 
@@
 more than multiple-permit fishers
-for the first time in 2014.
+since 2013.
 Our results suggest that the benefit of species diversity
@@
 or two more permits (dark blue).
-On the y-axis, dots and lines represent posterior median and 50\%
+On the y-axis, dots and lines represent posterior median and 90\%
 credible intervals.}

### 2016-08-25 diversity/analysis/herring_case_study.Rmd
updated figs and text

---
title: "Herring: case study 1"
author: ''
date: ''
output:
  pdf_document:
    fig_caption: yes
  html_document: default
  word_document: default
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=FALSE, tidy.opts=list(width.cutoff=60), warning = FALSE, message = FALSE)
library(dplyr)
library(knitr)
library(ggplot2)
library(date)
```

```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide"}
cfec <- feather::read_feather("../../portfolio/data-generated/cfec.feather")
cfec$year <- as.numeric(cfec$year)

herring = cfec[substr(cfec$p_fshy,1,1) %in% c("L","G"),]

herring_summary = group_by(herring, p_fshy, year) %>%
  summarize(g = sum(g_earn), p = mean(g_price,na.rm=T), lbs = sum(g_pounds), people = length(unique(p_holder))) %>%
  group_by(p_fshy) %>%
  mutate(n = length(unique(year)), maxpeople = max(people,na.rm=T)) %>%
  filter(n > 9) %>%
  filter(maxpeople > 25)
```

## Overview

First, we can pull out the herring permits by all areas. We'll restrict the analysis to (1) fisheries with 10 or more years of data, and (2) fisheries that at their max had at least 25 people landing catch in a year. We see that people generally decline to 0, except fisheries in SE (G01A, G34A), and spawn on kelp fisheries (L21). Despite catch being flat or maybe increasing, we see prices go down or flat for nearly all permits (and spawn on kelp L21 permits have significantly higher price / lb compared to roe fisheries).

In terms of the diversification of these permit holders, some questions to ask might be (1) if people stopped fishing herring, did they stop fishing entirely, or fish something else? (2) for the people that stayed, how many of them are diversified (other permits) and how much of their revenue comes from other permits?

For a lot of permits, we see decreases in the proportion of people fishing only herring. In terms of spatial diversity, for several of the fisheries that haven't crashed, we see increases in diversity: G01M/T, G34N, L21. In terms of diversifying to other types of herring permits,

```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Herring permit holders by year."}
  ggplot(herring_summary, aes(year, people)) + geom_line() + facet_wrap(~ p_fshy, scale="free_y") + xlab("Year") + ylab("People")
```

\break

```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Trends in landed catch by permit.", fig.height=3.5}
  ggplot(herring_summary, aes(year, log(lbs))) + geom_line() + 
  facet_wrap(~ p_fshy) + ylab("Ln (tons)") + xlab("Year")
```

```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Trends in prices by permit.", fig.height=3.5}
  ggplot(herring_summary, aes(year, p)) + geom_line() + 
  facet_wrap(~ p_fshy, scale="free_y") + ylab("Price (Dollars / lb.)") + xlab("Year")
```

```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Scatterplots of prices versus revenue by permit."}
  ggplot(herring_summary, aes(p, log(lbs/2000), colour = year)) + geom_point() + facet_wrap(~ p_fshy, scale="free") + ylab("Ln (tons)") + xlab("Price (dollars / lb.)")

```

\break

## How much of your revenue is driven by non-herring?

```{r, echo = FALSE}

herring_permits = unique(herring_summary$p_fshy)

# the complicated thing here is that people may hold multiple herring permits, 
# multiple permits, no permits. So for each permit type we need to calculate 
# each of these groups independently. Some people may be counted multiple times.

# next calculate strategies for diversification
calc_permit_summary = function(this_permit) {
people_year = group_by(cfec[cfec$p_fshy == this_permit,], p_holder, year) %>% 
  summarize(g = sum(g_earn)) %>% 
  filter(g > 0) %>% 
  select(-g)
# grab all records pertaining to these people-year combinations
herring = left_join(people_year, cfec)

# aggregate fish tickets to permits
herring = group_by(herring, p_holder, p_fshy, year) %>% 
  summarize(g_earn = sum(g_earn))

permit_summary = group_by(herring, p_holder, year) %>% 
  summarize(npermit = length(unique(p_fshy)),
    npermit_herring = length(which(p_fshy %in% herring_permits)),
    npermit_g34 = length(which(p_fshy %in% herring_permits & substr(p_fshy,1,4)=="G 34")),
    npermit_g01 = length(which(p_fshy %in% herring_permits & substr(p_fshy,1,4)=="G 01")),
    npermit_l12 = length(which(p_fshy %in% herring_permits & substr(p_fshy,1,4)=="L 12")),
    npermit_l21 = length(which(p_fshy %in% herring_permits & substr(p_fshy,1,4)=="L 21")),
    nmult = ceiling(npermit_g34/100) + ceiling(npermit_g01/100) + ceiling(npermit_l21/100) + ceiling(npermit_l12/100)) %>% 
    as.data.frame() %>%
    group_by(year) %>% 
  summarize(only_herring_single = length(which(npermit==1 & npermit_herring == 1)),
    only_herring_multiple = length(which(npermit>1 & npermit_herring >1 & npermit == npermit_herring)), herring_combo = length(which(npermit > 1 & npermit_herring >= 1 & npermit > npermit_herring)), g34 = length(which(npermit_g34 > 1)), g01 = length(which(npermit_g01 > 1)),
    l12 = length(which(npermit_l12 > 1)), l21 = length(which(npermit_l21 > 1)), 
    multiple_fisheries = length(which(nmult > 1)))

permit_summary$p_fshy = this_permit
return(permit_summary)
}

herring_summary = calc_permit_summary(herring_permits[1])
for(i in 2:length(herring_permits)) {
  herring_summary = rbind(herring_summary, calc_permit_summary(herring_permits[i]))
}

```

```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Proportion of permit holders specializing on herring.", fig.height=3.5}
ggplot(herring_summary, aes(year, (only_herring_single+only_herring_multiple)/(only_herring_single+only_herring_multiple+herring_combo))) + facet_wrap(~p_fshy) + geom_line() + xlab("Year") + ylab("Proportion of people only fishing herring")
```

```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Proportion of permit holders diversifying spatially.", fig.height=3.5}
p1 = ggplot(herring_summary, aes(year, (g34)/(only_herring_single+only_herring_multiple+herring_combo), group = p_fshy, colour = p_fshy)) + geom_line() + xlab("Year") + ylab("Proportion of people diversifying spatially")

p2 = ggplot(herring_summary, aes(year, (g01)/(only_herring_single+only_herring_multiple+herring_combo), group = p_fshy, colour = p_fshy)) + geom_line() + xlab("Year") + ylab("Proportion of people diversifying spatially")

p3 = ggplot(herring_summary, aes(year, (l21)/(only_herring_single+only_herring_multiple+herring_combo), group = p_fshy, colour = p_fshy)) + geom_line() + xlab("Year") + ylab("Proportion of people diversifying spatially")

p4 = ggplot(herring_summary, aes(year, (l12)/(only_herring_single+only_herring_multiple+herring_combo), group = p_fshy, colour = p_fshy)) + geom_line() + xlab("Year") + ylab("Proportion of people diversifying spatially")

gridExtra::grid.arrange(p1, p2, p3, p4)
```

\break

```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Proportion of people diversifying w/diff herring permits.", fig.height=3.5}
ggplot(herring_summary, aes(year, (multiple_fisheries)/(only_herring_single+only_herring_multiple+herring_combo))) + facet_wrap(~p_fshy) + geom_line() + xlab("Year") + ylab("Proportion of people diversifying w/diff herring permits")
```

```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Ratio of effective diversity to total number of permit holders (as measure of evenness). Values near 1 indicate everyone's revenue is similar.", fig.height=3.5}
# We can also calculate how herring catch is distributed across fisherment
simp.div = function(x) {1/sum((x/sum(x))^2)}

group_by(cfec[cfec$p_fshy%in%herring_permits,], year, p_holder, p_fshy) %>% 
  summarize(g = sum(g_earn)) %>% 
  group_by(year, p_fshy) %>% 
  summarize(div = simp.div(g), n = length(unique(p_holder))) %>% 
  ggplot(aes(year, div/n)) + facet_wrap(~p_fshy, scale="free_y") + geom_line() + xlab("Year") + ylab("Diversity across people")
```

\break

```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Fraction of revenue made up of herring (median, of per-person fractions).",eval=FALSE}

herring_permits = unique(herring_summary$p_fshy)

herring_people = group_by(cfec, p_fshy, year, p_holder) %>%
  summarize(g = sum(g_earn),
  herring_landing = ifelse(length(which(p_fshy %in% herring_permits)) == 0, 0, 1)) %>% 
  data.frame() %>% 
  group_by(p_holder, year) %>% 
  mutate(herring_personyear = ifelse(length(which(p_fshy %in% herring_permits)) == 0, 0, 1),
  g_herring = sum(g[herring_landing==1]), g_other = sum(g[herring_landing==0]),
  p_fshy = p_fshy[which(p_fshy %in% herring_permits)][1]) %>% 
  group_by(p_holder) %>% 
  mutate(ever_herring = sum(herring_personyear)) %>%   
  filter(ever_herring > 0) %>% 
  select(-g, -herring_landing, -herring_personyear, -ever_herring)

herring_people$p_herring = herring_people$g_herring / (herring_people$g_herring + herring_people$g_other)
group_by(herring_people[herring_people$p_fshy %in% herring_permits, ], p_fshy, year) %>% 
  summarize(g = median(p_herring)) %>% 
  ggplot(aes(year, g)) + geom_line() + facet_wrap(~p_fshy) + ylab("Median % herring (as fraction of total revenue)") + xlab("Year")
```

### 2016-09-07 portfolio/text/ms.Rmd
Continue developing outline/draft of commercial paper

@@
 often experience high levels of income variability
-due to the combined effects of environmental and price variation. 
+due to the combined effects of environmental, biological, and price variation.
 Commercial fishers are a prime example and
@@
 (2) owning permits that inherently fish higher species diversity,
-and (3) adjusting species diversity within the constraints of a permit.
+and (3) increasing species diversity within the constraints of a permit set.
 We control for changes in effort and other
-residual effects through time.
+latent effects through time.
 We find that
@@
 fishing permits with higher species diversity reduces risk
-(XX--XX% per unit increase in Simpson's diversity index),
-but that changing species diversity within a
-permit's constraints has a fishery-dependent effect on revenue and 
+(XX% [XX--XX 95% CI] per unit increase in mean Simpson's diversity index),
+but that increasing species diversity within
+the constraints of a permit set has a fishery-dependent effect on revenue and
 is usually (XX%) associated with increased revenue variability.
 At the same time, Alaskan commercial fisheries
-have increasingly focussed on 
+have been increasingly focussed on
 single-permit (often salmon) fishing strategies
@@
 more than multiple-permit fishers
-since 2013.
+for the first time in 2013.
 Our results suggest that the benefit of species diversity
@@
 
-What's been done
+Natural resources such as forests, land animals, and fishes
+provide a primary source of income for billions(?) of people worldwide [REFS].
+However, environmental,
+biological, and price variation
+combined with regulatory changes
+and human-caused environmental disasters
+generate high levels of income variability and risk
+for these individuals.
+For example, ...
+And EVOS example
 
-[@kasperski2013]
-[@sethi2012b]
-[@sethi2010b]
-[@sethi2014]
-[@perruso2005]
-[@hilborn2001]
-[@edwards2004]
-[@trenkel2013]
-[@mumford2009]
-[@oostenbrugge2002]
-[@baldursson1997]
-[@jardine2015]
-[@sanchirico2008]
+\bigskip
 
-Multiple ways one can diversify: 
-buy another permit, pick diverse permit, diversify within existing constraints.
+Targeting a higher diversity of species
+has been proposed as a risk-buffering mechanism...
+Give examples from other sectors
+
+\bigskip
+
+This effect has been identified for fisheries.
+At community level, ... [@sethi2014]
+At vessel level, ... [@kasperski2013] ...
+
+\bigskip
+
+Despite this existing work,
+key questions remain on the relationship between species diversity and
+fishing revenue variability.
+First, we do not know the effect of changing species diversity
+for individual permit holders on a year-by-year basis ---
+existing approaches have worked at the vessel or community level
+and condensed time series into single metrics of variability or risk.
+Second, we do not know the effect of changing species diversity
+on revenue itself ---
+existing approaches have focused solely on variability or risk.
+Third, we do not know if the approach to
+diversifying affects the relationship between
+species diversity and revenue variability.
+A fisheries permit holder can purchase
+an additional permit to target other species,
+pick a permit that inherently targets many species,
+or increase species diversity
+within the constraints of his or her existing permits.
+Many permit holders may be limited to the last approach
+given that the first two may be costly and limited by regulations,
+but it is unclear if the benefits and costs are the same.
+Here, we address these questions by
+simultaneously estimating the effects of
+these three approaches to species diversification
+for Alaskan fishery permit holders on
+annual revenue and revenue variability.
+Our model makes use of the entire time series,
+controls for effort and latent time effects,
+and allows the effect of species diversity
+to vary by fishing strategy (sets of similar permits)
+in a hierarchical model.
+
+<!-- Multiple ways one can diversify: -->
+<!-- buy another permit, pick diverse permit, diversify within existing constraints. -->
 
@@
 \end{figure}
+
+\clearpage
+
+# Notes
+
+Fishing can be big $ but is risky, for example, Alaska
+
+What's been done
+
+## @kasperski2013
+- vessel-level CFEC data
+- dome-shaped result
+- Herfindahl-Hirschman Index/HHI (Simpson)
+- diversity decreasing for those still fishing in 2010, but not otherwise
+
+## @sethi2012b
+- risk measures to 90 Alaskan fishing fishery/fleets
+- revenue more variable than catch (+ price = extra variability)
+- large catches, big fleets = lower risk
+
+## @sethi2010b
+- risk management for fisheries review
+
+## @sethi2012
+- equal-share catch-pooling cooperatives in king and snow crab
+- "variability reduction depends on the degree of correlation amongst
+  participants’ catches"
+
+## @sethi2014
+- community-level CFEC data
+- number of fisheries strongest predictor of (lower) CV/CVaR
+- also simpson index
+- suggest barriers to commercial entry increase community risk by restricting
+  access to 'variety of fisheries'
+- some communities have limited ability to diversify; may need to diversify
+  outside of fishing
+
+[@perruso2005]
+[@hilborn2001]
+[@edwards2004]
+[@trenkel2013]
+[@mumford2009]
+[@oostenbrugge2002]
+[@baldursson1997]
+[@jardine2015]
+[@sanchirico2008]

### 2016-09-09 portfolio/text/ms.Rmd
Calculate effect sizes for abstracts and results...
Also switch to 95% CIs where possible
Edits to text

@@
 load("../data-generated/output-values.rda")
+h1 <- readRDS("../data-generated/h1_effect.rds")
+spoke <- readRDS("../data-generated/spoke_fraction.rds")
+g2_dens <- readRDS("../data-generated/g2_prob_dens_above_zero.rds")
 ```
@@
 We find that
-fishing an additional permit is associated with
-lower variability (~XX--XX%) and higher revenue,
-fishing permits with higher species diversity reduces risk
-(XX% [XX--XX 95% CI] per unit increase in mean Simpson's diversity index),
+fishing an additional permit is associated with higher revenue
+and lower variability (e.g., 
+`r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR] lower
+expected standard deviation for halibut permit holders), 
+fishing permits with higher species diversity reduces variability
+(`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction per unit 
+increase in mean Simpson's diversity index),
 but that increasing species diversity within
 the constraints of a permit set has a fishery-dependent effect on revenue and
-is usually (XX%) associated with increased revenue variability.
+is usually (`r g2_dens`% of probability density) 
+associated with increased revenue variability.
 At the same time, Alaskan commercial fisheries
@@
 <!-- write out each paragraph’s main message in 15 words or less -->
-
 <!-- circulate skeleton ... this is the time to fix problems with flow, -->
@@
 owners of (A) individual halibut and (B) gillnet herring roe permits
-(light blue)
+(dark blue)
 compared with owners of the same permit plus one more (medium blue)
-or two more permits (dark blue).
-On the y-axis, dots and lines represent posterior median and 90\%
-credible intervals.}
+or two more permits (light blue).
+On the y-axis, dots and lines represent posterior median and 95\%
+credible intervals of expected standard deviation.}
 \label{fig:spokes}
@@
 \begin{center}
-\includegraphics[width=0.7\textwidth]{../figs/stan-across-strategy-variability.pdf}
+\includegraphics[width=0.7\textwidth]{../figs/stan-across-strategy-variability2.pdf}
 \caption{Individuals who fish
@@
 Estimated variability represents expected standard deviation
-at an average of about
-`r mean_permit_days` combined fishing days across permits,
-and for a permit holder who does not change
-species diversity or days fished from year to year.
+for a permit holder who does not change
+species diversity or days fished from year to year
+with the strategy-level effect of mean days fished removed.
 }

### 2016-09-09 portfolio/analysis/3.3-plot-across-strategy-effects.R

@@ -44,6 +44,15 @@ g0_temp <- mutate(g0_temp, resid =
   component_resid_div = resid + g0main + h1 * scaled_strategy_mean_div,
   component_resid_days = resid + g0main + h2 * scaled_strategy_mean_days)
 
+scale_factor <- sd(md$strategy_mean_div) * 2
+# effect for paper:
+h1_effect <- quantile(g0_temp$h1, probs = c(0.025, 0.5, 0.975)) %>% exp() %>%
+  `*`(100) %>% `/`(scale_factor) %>% round(0)
+# # test with figure:
+# max(dd$m) %>% exp * (1-h1_effect[[2]]*0.01)
+# max(dd$m) %>% exp * (1-h1_effect[[2]]*0.01) * (1-h1_effect[[2]]*0.01)
+saveRDS(h1_effect, file = "portfolio/data-generated/h1_effect.rds")
+
 p1 <- ggplot(g0_temp, aes(scaled_strategy_mean_div, component_resid_div,
   group = strategy_id)) +
   geom_point(alpha = 0.01, position = position_jitter(width = 0.05))
@@ -62,8 +71,8 @@ gridExtra::grid.arrange(p1, p3, ncol = 2)
 g0 <- g0_temp %>% group_by(strategy) %>%
   summarise(
     g0.m = median(component_resid_div),
-    g0.l = quantile(component_resid_div, probs = 0.05),
-    g0.u = quantile(component_resid_div, probs = 0.95),
+    g0.l = quantile(component_resid_div, probs = 0.025),
+    g0.u = quantile(component_resid_div, probs = 0.975),
     g0.l.5 = quantile(component_resid_div, probs = 0.25),
     g0.u.5 = quantile(component_resid_div, probs = 0.75)) %>%
   ungroup() %>%
@@ -77,8 +86,8 @@ dd <- plyr::ldply(seq_along(x_scaled), function(i) {
   xx <- x_scaled[i]
   data.frame(strategy_mean_div = x_true[i],
     m = median(p$h1[[1]] * xx + p$g0[[1]]),
-    l = quantile(p$h1[[1]] * xx + p$g0[[1]], probs = 0.05),
-    u = quantile(p$h1[[1]] * xx + p$g0[[1]], probs = 0.95),
+    l = quantile(p$h1[[1]] * xx + p$g0[[1]], probs = 0.025),
[…]

### 2016-09-09 portfolio/analysis/3.3-plot-spokes.R

@@ -31,8 +31,8 @@ g0_temp <- mutate(g0_temp, posterior =
 md2 <- g0_temp %>% group_by(strategy) %>%
   summarise(
     estimate = median(posterior),
-    conf.low = quantile(posterior, probs = 0.05),
-    conf.high = quantile(posterior, probs = 0.95)) %>%
+    conf.low = quantile(posterior, probs = 0.025),
+    conf.high = quantile(posterior, probs = 0.975)) %>%
   ungroup() %>%
   inner_join(md)
 
@@ -174,6 +174,20 @@ pl <- filter(sp, !single_permit %in% c("S03", "K91")) %>%
 # print(pl)
 ggsave("portfolio/figs/stan-gg-spoke2.pdf", width = 7, height = 3)
 
+sp2 <- mutate(sp, fract = exp(estimate)/exp(g0_less),
+  fract_rev = strategy_med_rev/b0_less)
+sp2 <- filter(sp2, single_permit!=strategy)
+# hist(sp2$fract)
+# hist(sp2$fract_rev)
+spoke_fraction <- quantile(sp2$fract, probs = c(0.025, 0.5, 0.975)) %>%
+  round(2)%>% `*`(100)
+spoke_fraction <- quantile(sp2$fract, probs = c(0.25, 0.5, 75)) %>%
+  round(2)%>% `*`(100)
+saveRDS(spoke_fraction, file = "portfolio/data-generated/spoke_fraction.rds")
+
+spoke_fraction_rev <- quantile(sp2$fract_rev, probs = c(0.25, 0.5, 0.75)) %>%
+  round(2)%>% `*`(100)
+saveRDS(spoke_fraction, file = "portfolio/data-generated/spoke_fraction_rev.rds")
 
 # ------------------------------------------------------
 # more for SI

### 2016-09-09 portfolio/analysis/3.3-plot-strategy-effects.R

@@ -36,6 +36,10 @@ b2 <- tidyr::gather(p$coef_b2_strategy, term, posterior) %>%
   mutate(parameter = "b2")
 po <- bind_rows(g0, g1, g2, b1, b2)
 
+g2_prob_dens_above_zero <- round(sum(g2$posterior > 0)/nrow(g2)*100, 0)
+saveRDS(g2_prob_dens_above_zero,
+  file = "portfolio/data-generated/g2_prob_dens_above_zero.rds")
+
 md2 <- md
 md2$str_label <- as.factor(md2$str_label)
 md2$str_label <- factor(md2$str_label, levels = md2$str_label[order(md2$strategy_mean_div)])

### 2016-09-09 portfolio/text/anderson-etal-rev-pfx.tex
Early draft of abstract, other text
Title brainstorm

@@
 \title{Portfolios of fishing permits and species
-reduce risk but adjusting fishing strategy is risky}
+reduce risk but adjusting fishing strategy is risky\\
+Portfolio benefits to fishers
+depend on approach to diversification\\
+Portfolio benefits of income diversification
+are unavailable to many fishermen\\
+Effect of species diversity on income variability
+depends on fishing strategy\\
+Effect of species diversity on revenue risk
+depends on diversification approach\\
+}
 \author{
@@
 Jennifer C. Shriver,
-Jordan T. Watson
+Jordan T. Watson,
+\ldots
 }
@@
 
-\vspace{-1cm}
+% \vspace{-1cm}
 
@@
 \thispagestyle{empty}
+
+\clearpage
 

### 2016-09-09 portfolio/text/ms.Rmd
Tweak abstract

@@
 `r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR] lower
-expected standard deviation for halibut permit holders), 
+expected standard deviation compared to single halibut permits), 
 fishing permits with higher species diversity reduces variability
@@
 Our results suggest that the benefit of species diversity
-on reducing revenue variability
+for revenue variability
 is mostly an across-permit effect, not

### 2016-09-09 portfolio/analysis/3.3-make-strategy-table.R
Save number of unique group permits

@@ -24,3 +24,5 @@ xtable::print.xtable(
   timestamp = NULL
 )
 
+n_grouped_permits <- nrow(d)
+saveRDS(n_grouped_permits, "portfolio/data-generated/n_grouped_permits.rds")

### 2016-09-09 portfolio/text/supp.Rnw
Develop supplement text

@@
 
-<<get-values, cache=FALSE>>=
+<<get-values, cache=FALSE, echo=FALSE>>=
 load("../data-generated/output-values.rda")
+n_grouped_permits <- readRDS("../data-generated/n_grouped_permits.rds")
 @
@@
 
-\section{Data filtering}
+\section{Filtering data and grouping permits into strategies}
 
-<<nothing, eval=FALSE>>=
-permit_rev = group_by(cfecnew, p_holder, year, p_fshy) %>%
-  summarize(permit.rev = sum(g_earn)) %>%
-  as_data_frame() %>%
-  group_by(p_holder, year) %>%
-  mutate(tot.rev = sum(permit.rev))
-permit_rev$permit.rev = permit_rev$permit.rev/permit_rev$tot.rev # fraction of total
-permit_rev = permit_rev[permit_rev$permit.rev < 0.1,]
+We filtered the data in a number of ways before fitting our model. 
+We began with \Sexpr{round(ndat$initial/1e6,2)} million rows of data.
+We removed revenue from permit-year combinations 
+that made up less than 10\% of annual revenue
+for a given permit holder. 
+We removed permit holders whose median revenue
+was less than \$10,000 USD (adjusted to 2009 values)
+to focus on individuals actively engaged 
+in commercial fishing.
+This reduced the total revenue from \Sexpr{round(ndat$rev_b4_large_only/1e9,0)}
+billion 2009 USD
+to \Sexpr{round(ndat$rev_large_only/1e9,0)} billion 2009 USD.
 
-# Delete these minimal earnings
-cfecnew = cfecnew[-which(paste(cfecnew$p_holder, cfecnew$year, cfecnew$p_fshy) %in%
-  paste(permit_rev$p_holder, permit_rev$year, permit_rev$p_fshy)), ]
+To limit the number of permits 
+and permit strategies 
+and focus on the main fisheries, 
+we removed all permits with less than 50 unique permit holders 
+across all years. 
+We also removed permits for freshwater species and
+for experimental fisheries.
+A list of all CFEC permit codes and descriptions is available at:
+\url{https://www.cfec.state.ak.us/misc/FshyDesC.htm}
 
-cfecnew <- filter(cfecnew, p_spec != "Freshwater") # only 1400 rows
-@
+We then proceeded to combine similar permits in the following steps. 
+We combined a number of permits that were the same except for vessel size.
+We combined permits for 
+otter troll, 
+longline finfish,
+pot gear.
+king crab, 
+and
+shrimp.
+We grouped a number of permits that were targeting a single species 
+and differed only in region fished.
+For example, someone fishing herring roe in the Kodiak (G34K)
+and someone fishing herring roe in Alaska Peninsula (G34M) 
+were both considered to have a herring roe permit strategy.
+We combined permits across each of 
+dungeness crab, 
+herring roe purse seining,
+herring roe gillnet,
+herring hand picked,
+herring pound fishing,
+sea cucumber,
+sea urchin,
+tanner crabs
 
-Rows of data initial: \Sexpr{round(ndat$initial/1e6,2)} million
+Combining salmon permits was more difficult because 
+the species composition varies widely based on geography. 
+We grouped salmon permits within gear type-region 
+combinations that shared similar species composition.
+For the drift gillnet (S03) permits, 
+we combined permits from 
+Cook Inlet, Bristol Bay, and Alaska Peninsula because sockeye
+represents greater than 90\% of species revenue.
+For the purse seine permits (S01),
+there was a wide variety in species revenue by
+area ($\sim$ 85\% pink in Prince William Sound to
+85\% sockeye on the Alaskan Peninsula). 
+Of the five permits, only two had similar species compositions, 
+so we grouped Kodiak and Cook Inlet.
+For the set gillnet salmon permits (S04), 
+there was some variety but also clear sockeye specialists. 
+We grouped permits that had greater than 80\% sockeye revenue.
+This combined permits across 
+Prince William Sound, 
+Cook Inlet,
+Kodiak,
+Alaska Peninsula,
+and Bristol Bay.
+We grouped permits from Norton Sound
+and Kuskokwim because of similar species compositions.
+The remaining salmon permits were relatively unique 
+and were left ungrouped.
 
-Rows of data annual: \Sexpr{ndat$annual}
+After this process of combining permits, 
+we were left with \Sexpr{n_grouped_permits}
+unique fishing strategies (Table~\ref{tab:strategies}),
+which we aggregated
+within each person-year combination.
+For example, if someone fished halibut 
+and sablefish in the same year 
+their strategy for that year would be ``halibut-sablefish''.
+We removed people who switched strategies from one year to the next
+since we were working with changes in revenue from year to year
+and so needed consistent strategies across any pair of years
+for a permit holder.
+We removed strategies with less than 100 unique permit holders
+across all years to focus on the major strategies
+leaving us with a total of \Sexpr{nstr} strategies.
+Finally, we removed all person-year combinations without 
+revenue recorded in the following year 
+in order to calculate percent change in revenue
+from year to year.
+This reduced the rows (person-year combinations)
+in the final dataset from \Sexpr{ndat$annual} to
+\Sexpr{ndat$annual_diff}.
 
-Rows of data diff: \Sexpr{ndat$annual_diff}
-
-% \Sexpr{ndat$rev_b4_10perc/1e9}
-% \Sexpr{round(ndat$rev_after_10perc/1e9,2)}
-
-\Sexpr{ndat$before_10_perc}
-
-Rows of data after 10,000 cull: \Sexpr{ndat$large_only}
-
-10,000 cull: rev:
-\Sexpr{round(1-ndat$rev_large_only/ndat$rev_b4_large_only,2)*100}
-
-\Sexpr{round(ndat$rev_large_only/1e9,2)}
-
-\section{Assigning strategies}
-
-\Sexpr{nstrat$enough_pholders}
-
-\Sexpr{round(ndat$rev_enough_pholders/1e9,2)}
-
-\Sexpr{nstrat$no_river_exp}
-
-\Sexpr{round(ndat$rev_no_river_exp/1e9,2)}
-
-Number of strategies: \Sexpr{nstr}
-
-Rows of data before no singles: \Sexpr{ndat$before_nosingles}
-
-% \Sexpr{ndat$b4_no_na_strat}
-%
-% \Sexpr{ndat$rev_b4_no_na_strat}
-%
-% \Sexpr{ndat$after_no_na_strat}
-
-lost \% rev no strat:
-\Sexpr{round(ndat$rev_after_no_na_strat/ndat$rev_b4_no_na_strat,2)*100}
-
-lost rows no strat:
-\Sexpr{round(ndat$after_no_na_strat/ndat$b4_no_na_strat,2)*100}
-
-% \Sexpr{ndat$rev_after_no_na_strat}
-%
-% \Sexpr{ndat$rev_final}
-
-Rows of data: \Sexpr{ndat$final}
-
-Number of permit holders: \Sexpr{npholders}
-
-Range of years: \Sexpr{range_yrs}
-
-Number of strategy-years: \Sexpr{nstr_yrs}
+% Rows of data: \Sexpr{ndat$final}
+% Number of permit holders: \Sexpr{npholders}
+% Range of years: \Sexpr{range_yrs}
+% Number of strategy-years: \Sexpr{nstr_yrs}
+% \Sexpr{nstrat$enough_pholders}
+% \Sexpr{round(ndat$rev_enough_pholders/1e9,2)}
+% lost \% rev no strat:
+% \Sexpr{round(ndat$rev_after_no_na_strat/ndat$rev_b4_no_na_strat,2)*100}
+% lost rows no strat:
+% \Sexpr{round(ndat$after_no_na_strat/ndat$b4_no_na_strat,2)*100}
 
 \section{Hierarchical model}
+
+Describe predictors
+
+We adjusted for inflation by setting all revenues 2009 USD.
+
+We calculated Simpson's diversity index weighted by revenue
+as: 1/sum((x/sum(x)) squared)
 
@@
 a particular strategy represents a group of similar permits across areas
-or vessel sizes. These strategies all had at least 100
-permit holders.}
+or vessel sizes.}
   \label{tab:strategies}

### 2016-09-14 portfolio/figs/strategy-table.tex
Update strategy table after correct 10,000 threshold

@@
 Dun C & Dungeness crab \\ 
-  Fin - jig & Finfish, mechanical jig , statewide \\ 
   Fin - ll & Finfish, longline, statewide \\ 
@@
   Hal - ll & Halibut, longline \\ 
-  Hal - troll & Halibut, hand troll \\ 
-  Herr - hand & Herring, hand pick \\ 
   Herr - pound & Herring, pound \\ 
@@
   Sea cuc & Sea cucumber \\ 
-  Shrimp & Shrimp \\ 
   Tan C & Tanner crab, pot gear \\ 

### 2016-09-14 portfolio/text/ms.Rmd
Tweak abstract

@@
 that can reduce this variability.
-A top candidate has been that fishing higher levels of species diversity
+One candidate has been that fishing higher levels of species diversity
 decreases income variability.
@@
 `r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR] lower
-expected standard deviation compared to single halibut permits), 
+expected standard deviation
+for halibut plus another permit vs. single halibut permits); 
 fishing permits with higher species diversity reduces variability
 (`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction per unit 
-increase in mean Simpson's diversity index),
+increase in mean Simpson's diversity index);
 but that increasing species diversity within
 the constraints of a permit set has a fishery-dependent effect on revenue and
-is usually (`r g2_dens`% of probability density) 
+is usually (`r g2_dens`% probability) 
 associated with increased revenue variability.
@@
 Natural resources such as forests, land animals, and fishes
-provide a primary source of income for billions(?) of people worldwide [REFS].
+provide a primary source of income for millions(?) of people worldwide [REFS].
 However, environmental,

### 2016-09-14 portfolio/text/ms.Rmd
Develop methods
Also moved some of the methods to the main paper

@@
 (`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction per unit 
-increase in mean Simpson's diversity index);
+increase in mean effective species diversity);
 but that increasing species diversity within
@@
 fishing revenue variability.
-First, we do not know the effect of changing species diversity
-for individual permit holders on a year-by-year basis ---
+First, the effect of changing species diversity
+for individual permit holders on a year-by-year basis is unknown ---
 existing approaches have worked at the vessel or community level
 and condensed time series into single metrics of variability or risk.
-Second, we do not know the effect of changing species diversity
-on revenue itself ---
+Second, the effect of changing species diversity
+on revenue itself is unknown ---
 existing approaches have focused solely on variability or risk.
-Third, we do not know if the approach to
-diversifying affects the relationship between
-species diversity and revenue variability.
-A fisheries permit holder can purchase
+Third, it is unknown if the approach to
+diversifying species fished affects the relationship between
+species diversity and revenue variability ---
+a fisheries permit holder can purchase
 an additional permit to target other species,
@@
 
-[@standevelopmentteam2015]
-[@rcoreteam2016]
-[@simpson1949]
+- overview
+
+- describe predictors
+
+- We adjusted for inflation by setting all revenues in 2009 USD 
+based on data from TODO.
+
+We calculated effective species diversity for
+permit holder $i$ and year $t$ as the inverse of 
+Simpson's diversity index $\lambda$ [@simpson1949]
+weighted by revenue $R$ (herein 'species diversity'):
+
+\begin{equation}
+1/\lambda_{i,t} = 1 / \left(\sum\limits_{s=1}^{n_{i,t}}
+  \left(R_{i,s,t}/ \sum\limits_{s=1}^{n_{i,t}} R_{i,s,t}\right)\right)^2,
+\end{equation}
+
+\noindent
+where $s$ indexes species from $1$ through $n$.
+
+The basic model is a hierarchical linear regression with Gaussian errors
+predicting $\log$ revenue, $\log(R_{i,t})$, for
+each fisher $i$ and time (year) $t$:
+
+\begin{equation}
+\log(R_{i,t}) \sim \mathrm{N} \left(\widehat{\log(R_{i,t})}, \widehat{\sigma_{i,t}^2} \right).
+\end{equation}
+
+\noindent
+Each year $t$ of revenue for fisher $i$
+is assigned a strategy indexed by $j$ and
+expected $\log(R_{i,t})$ is modelled as:
+
+\begin{equation}
+\label{eq:mean}
+\widehat{\log(R_{i,t})} = \beta_{0,j,t} -
+  \beta_{1,j} S_{i,t} I_{i,t} +
+  \beta_{2,j} S_{i,t} I_{i,t} +
+  \beta_3 D_{i,t} -
+  \beta_4 S_{i,t} D_{i,t} I_{i,t} +
+  \beta_5 S_{i,t} D_{i,t} I_{i,t} +
+  \log(R_{i,t-1})
+\end{equation}
+
+\noindent
+where $S_{i,t}$ is the percent change in species diversity,
+$\log \left(\mathrm{div}_{i,t}/\mathrm{div}_{i,t-1}\right)$,
+$D_{i,t}$ is the percent change in days fished,
+$\log \left(\mathrm{days}_{i,t}/\mathrm{days}_{i,t-1}\right)$, and
+$I_{i,t}$ is an indicator variable that takes a value $0$
+if $\mathrm{div}_{i,t} \ge \mathrm{div}_{i,t-1}$
+(species diversity increasing; generalizing)
+and takes a value $1$ if
+$\mathrm{div}_{i,t} < \mathrm{div}_{i,t-1}$
+(species diversity decreasing; specializing).
+
+The term $\log(R_{i,t-1}$) is treated as an offset term:
+it is added on the right side of the equation
+to essentially model percent change in revenue,
+i.e., $\log(R_{i,t}) - \log(R_{i,t-1}) = \log(R_{i,t}/R_{i,t-1})$.
+While this approach required filtering the data 
+to remove person-year combinations without a subsequent year of revenue,
+it has a number of advantages: 
+it eliminates having to deal with the nonstationary 
+timeseries and eliminates having to 
+estimate thousands of people-specific intercepts.
+
+The $\beta$'s represent estimated coefficients, some of which are
+allowed to vary.
+The intercepts $\beta_{0,j,t}$ are allowed to vary by
+strategy-year combinations,
+are centered on zero, and are constrained by a normal distribution:
+
+\begin{equation}
+\beta_{0,j,t} \sim \mathrm{N} \left(0, \sigma^2_{\beta_{j,t}} \right).
+\end{equation}
+
+\noindent
+This, combined with the lack of a global intercept, constrains
+the mean change in log revenue across all fishers and years to
+be zero. I.e., we assume that the data are stationary
+after accounting for all covariates and including
+an intercept that varies with strategy-year combinations.
+Initial versions of the model
+confirmed that the global intercept along with stragey-level
+intercepts would be estimated at almost exactly zero.
+
+The slopes $\beta_{1,j}$ and $\beta_{1,j}$
+are then allowed to vary by strategy with estimated means:
+
+\begin{align}
+\beta_{1,j} &\sim \mathrm{N} \left(\beta_1, \sigma^2_{\beta_1}\right),\\
+\beta_{2,j} &\sim \mathrm{N} \left(\beta_1, \sigma^2_{\beta_2}\right).
+\end{align}
+
+\noindent
+The $\beta_{1,j}$'s represent how much an X\% decrease in species
+diversity within a strategy translates to a Y\% change in revenue.
+The $\beta_{2,j}$'s represent how much an X\% increase in species
+diversity within a strategy translates to a Y\% change in revenue.
+These are after controlling for percent change in days fished, $\beta_3$,
+and the interaction between days fished and percent change in species
+diversity, $\beta_4$ and $\beta_5$.
+
+Whereas a typical hierarchical linear regression model
+would estimate a single residual error scale term, $\sigma$,
+we model the scale of the residual error  with
+another hierarchical model with a similar form:
+
+\begin{equation}
+\label{eq:cv}
+\widehat{\sigma_{i,t}} =
+\exp \left( \gamma_{0,j} -
+\gamma_{1,j} S_{i,t} I_{i,t} +
+\gamma_{2,j} S_{i,t} I_{i,t} +
+\gamma_3 D_{i,t} -
+\gamma_4 S_{i,t} D_{i,t} I_{i,t} +
+\gamma_5 S_{i,t} D_{i,t} I_{i,t}\right).
+\end{equation}
+
+\noindent
+Again, $S_{i,t}$ and $I_{i,t}$ represent, respectively,
+percent change in species
+diversity and an indicator variable for increasing or
+decreasing species diversity.
+We exponentiate the equation to ensure all scale parameters, $\sigma_{i,t}$,
+are positive. The $\gamma$'s are estimated coefficients, three of which
+are allowed to vary by strategy, constrained by normal distributions:
+
+\begin{align}
+\gamma_{0,j}  &\sim \mathrm{N} \left(\gamma_0 + \eta_1 M_j + \eta_2 E_j, \sigma^2_{\gamma_0}\right),\\
+\gamma_{1,j}  &\sim \mathrm{N} \left(\gamma_1, \sigma^2_{\gamma_1}\right),\\
+\gamma_{2,j}  &\sim \mathrm{N} \left(\gamma_2, \sigma^2_{\gamma_2}\right).
+\end{align}
+
+\noindent
+The mean of the $\gamma_{0,j}$ coefficients is
+modelled with
+two strategy-level predictors.
+The symbols $M_j$ and $E_j$ represent the raw mean species diversity and
+mean combined fishing season length (effort) for strategy $j$.
+The coefficients $\eta_1$ and $\eta_2$ represent predictors
+in this strategy-level regression
+estimating effects across strategies.
+
+We fit our models with Stan 2.11.1
+[@hoffman2014; @standevelopmentteam2015; @standevelopmentteam2015a; @carpenter2015]
+and R 3.3.1 [@rcoreteam2016].
+Stan implements the 
+No-U-Turn Hamiltonian Markov chain Monte Carlo algorithm [@hoffman2014]
+to perform Bayesian statistical inference.
+We ran four chains and 2,000 iterations and discarded
+the first 1,000 iterations of each chain
+as warm up (i.e. 4,000 final samples). 
+We checked for chain convergence visually with trace plots.
+We also ensured that $\hat{R}<1.05$ (the potential scale reduction factor)
+and that the effective sample size was 
+greater than 200 for all parameters [@gelman2014].
+Code to reproduce our results (with a simulated data set)
+is available at <https://github.com/NCEAS/pfx-commercial>.
 

### 2016-09-14 portfolio/text/supp.Rnw
Strip white space and space out supplement
Develop methods
Also moved some of the methods to the main paper

@@
 \begin{document}
+
+\linenumbers
+
+\setlength{\parindent}{0.9cm}
+\setlength\parskip{0in}
+
+\begin{spacing}{1.3}
 
@@
 
+TODO length(unique(pholders))
 % Rows of data: \Sexpr{ndat$final}
@@
 \section{Hierarchical model}
-
-Describe predictors
-
-We adjusted for inflation by setting all revenues 2009 USD.
-
-We calculated Simpson's diversity index weighted by revenue
-as: 1/sum((x/sum(x)) squared)
-
-Overview... why first diff, why 2 part
-
-Negatives are that it uses a smaller subset of the data;
-positives are
-that it doesn't require estimating people-specific random effects (n = 3000+)
-and removes time effects.
-
-The basic model is a hierarchical linear regression with Gaussian errors
-predicting $\log$ revenue, $\log(R_{i,t})$, for
-each fisher $i$ and time (year) $t$:
-
-\begin{equation}
-\log(R_{i,t}) \sim \mathrm{N} \left(\widehat{\log(R_{i,t})}, \widehat{\sigma_{i,t}^2} \right).
-\end{equation}
-
-Each year $t$ of revenue for fisher $i$
-is assigned a strategy indexed by $j$ and
-expected $\log(R_{i,t})$ is modelled as:
-
-\begin{equation}
-\label{eq:mean}
-\widehat{\log(R_{i,t})} = \beta_{0,j,t} -
-  \beta_{1,j} S_{i,t} I_{i,t} +
-  \beta_{2,j} S_{i,t} I_{i,t} +
-  \beta_3 D_{i,t} -
-  \beta_4 S_{i,t} D_{i,t} I_{i,t} +
-  \beta_5 S_{i,t} D_{i,t} I_{i,t} +
-  \log(R_{i,t-1})
-\end{equation}
-
-\noindent
-where $S_{i,t}$ is the percent change in species diversity,
-$\log \left(\mathrm{div}_{i,t}/\mathrm{div}_{i,t-1}\right)$,
-$D_{i,t}$ is the percent change in days fished,
-$\log \left(\mathrm{days}_{i,t}/\mathrm{days}_{i,t-1}\right)$, and
-$I_{i,t}$ is an indicator variable that takes a value $0$
-if $\mathrm{div}_{i,t} \ge \mathrm{div}_{i,t-1}$
-(species diversity increasing; generalizing)
-and takes a value $1$ if
-$\mathrm{div}_{i,t} < \mathrm{div}_{i,t-1}$
-(species diversity decreasing; specializing).
-$\log(R_{i,t-1}$) is treated as an offset term:
-it is added on the right side of the equation
-to essentially model percent change in revenue,
-i.e., $\log(R_{i,t}) - \log(R_{i,t-1}) = \log(R_{i,t}/R_{i,t-1})$.
-
-The $\beta$'s represent estimated coefficients, some of which are
-allowed to vary.
-The intercepts $\beta_{0,j,t}$ are allowed to vary by
-strategy-year combinations,
-are centered on zero, and are constrained by a normal distribution:
-
-\begin{equation}
-\beta_{0,j,t} \sim \mathrm{N} \left(0, \sigma^2_{\beta_{j,t}} \right).
-\end{equation}
-
-This, combined with the lack of a global intercept, constrains
-the mean change in log revenue across all fishers and years to
-be zero. I.e., we assume that the data are stationary
-after accounting for all covariates and including
-an intercept that varies with strategy-year combinations.
-Initial versions of the model
-confirmed that the global intercept along with stragey-level
-intercepts would be estimated at almost exactly zero.
-
-TODO EXPLAIN WHY DO THIS
-
-The slopes $\beta_{1,j}$ and $\beta_{1,j}$
-are then allowed to vary by strategy with estimated means:
-
-\begin{align}
-\beta_{1,j} &\sim \mathrm{N} \left(\beta_1, \sigma^2_{\beta_1}\right),\\
-\beta_{2,j} &\sim \mathrm{N} \left(\beta_1, \sigma^2_{\beta_2}\right).
-\end{align}
-
-The $\beta_{1,j}$'s represent how much an X\% decrease in species
-diversity within a strategy translates to a Y\% change in revenue.
-The $\beta_{2,j}$'s represent how much an X\% increase in species
-diversity within a strategy translates to a Y\% change in revenue.
-These are after controlling for percent change in days fished, $\beta_3$,
-and the interaction between days fished and percent change in species
-diversity, $\beta_4$ and $\beta_5$.
-
-Whereas a typical hierarchical linear regression model
-would estimate a single residual error scale term, $\sigma$,
-we model the scale of the residual error  with
-another hierarchical model with a similar form:
-
-\begin{equation}
-\label{eq:cv}
-\widehat{\sigma_{i,t}} =
-\exp \left( \gamma_{0,j} -
-\gamma_{1,j} S_{i,t} I_{i,t} +
-\gamma_{2,j} S_{i,t} I_{i,t} +
-\gamma_3 D_{i,t} -
-\gamma_4 S_{i,t} D_{i,t} I_{i,t} +
-\gamma_5 S_{i,t} D_{i,t} I_{i,t}\right).
-\end{equation}
-
-Again, $S_{i,t}$ and $I_{i,t}$ represent, respectively,
-percent change in species
-diversity and an indicator variable for increasing or
-decreasing species diversity.
-We exponentiate the equation to ensure all scale parameters, $\sigma_{i,t}$,
-are positive.
-
-The $\gamma$'s are estimated coefficients, three of which
-are allowed to vary by strategy, constrained by normal distributions:
-
-\begin{align}
-\gamma_{0,j}  &\sim \mathrm{N} \left(\gamma_0 + \eta_1 M_j + \eta_2 E_j, \sigma^2_{\gamma_0}\right),\\
-\gamma_{1,j}  &\sim \mathrm{N} \left(\gamma_1, \sigma^2_{\gamma_1}\right),\\
-\gamma_{2,j}  &\sim \mathrm{N} \left(\gamma_2, \sigma^2_{\gamma_2}\right).
-\end{align}
-
-The mean of the $\gamma_{0,j}$ coefficients is
-modelled with
-two strategy-level predictors.
-The symbols $M_j$ and $E_j$ represent the raw mean species diversity and
-mean combined fishing season length (effort) for strategy $j$.
-The coefficients $\eta_1$ and $\eta_2$ represent predictors
-in this strategy-level regression.
-The effect of $\eta_1$ is shown in Fig.~3.
-
-We fit our models in Stan version 2.11.1. \ldots
-\citep{hoffman2014, standevelopmentteam2015a, carpenter2015a, standevelopmentteam2015}
-
-\citep{gelman2014}
-\citep{rcoreteam2016}
 
@@
 \bibliography{refs}
+
+\end{spacing}
 

### 2016-09-14 portfolio/text/ms.Rmd
Add to methods

@@
 - overview
-
 - describe predictors
-
 - We adjusted for inflation by setting all revenues in 2009 USD 
 based on data from TODO.
+- note the filtering and referred to the supplement
 
-We calculated effective species diversity for
+## Predictors
+
+Our models focused on the predictor species diversity, 
+while controlling for effort. We calculated effective species diversity for
 permit holder $i$ and year $t$ as the inverse of 
@@
 where $s$ indexes species from $1$ through $n$.
+We calculated effort or effective season length as
+the sum of days fished across each permit each year. 
+For example, if someone fished for 30 days
+on a salmon drift gillnet permit in Southeast Alaska
+and 40 days on a statewide longline halibut permit,
+that individual's season length 
+would be 70 days.
+
+## Hierarchical model
 

### 2016-09-14 portfolio/text/ms.Rmd
Play with wording in abstract

@@
 is usually (`r g2_dens`% probability) 
-associated with increased revenue variability.
+associated with increased uncertainty in revenue.
 At the same time, Alaskan commercial fisheries
@@
 Natural resources such as forests, land animals, and fishes
-provide a primary source of income for millions(?) of people worldwide [REFS].
+provide a primary source of income for millions of people worldwide [REFS].
 However, environmental,

### 2016-09-14 portfolio/text/ms.Rmd
Keep developing methods

@@
 - overview
-- describe predictors
-- We adjusted for inflation by setting all revenues in 2009 USD 
+
+## Data
+
+We obtained landings and revenue data for 
+all permit holders in Alaska 
+from `r range_yrs[1]` to `r range_yrs[2]`
+from the Alaskan Commercial Fisheries Entry Commission (CFEC).
+These data represent a total of 
+nearly `r round(ndat$initial_tot_rev/1e9)` billion 2009 USD
+and a combined `r round(ndat$initial/1e6,1)` million rows of data.
+We adjusted for inflation by setting all revenues in 2009 USD 
 based on data from TODO.
-- note the filtering and referred to the supplement
-
-## Predictors
+We implemented a number of filtering steps 
+to focus on individuals actively engaged in commercial fishing
+(Supporting Information). 
+In particular, we removed permit holders whose median revenue
+was less than $10,000 2009 USD.
 
 Our models focused on the predictor species diversity, 
-while controlling for effort. We calculated effective species diversity for
+while controlling for effort. 
+We calculated effective species diversity for
 permit holder $i$ and year $t$ as the inverse of 
 Simpson's diversity index $\lambda$ [@simpson1949]
-weighted by revenue $R$ (herein 'species diversity'):
+weighted by revenue $R$ ('species diversity'):
 
@@
 it has a number of advantages: 
-it eliminates having to deal with the nonstationary 
-timeseries and eliminates having to 
+it eliminates having to deal with the non-stationary 
+time series and eliminates having to 
 estimate thousands of people-specific intercepts.

### 2016-09-14 portfolio/text/ms.Rmd
Strip whitespace; add model checking note
Copy editing

@@
 ```
-
-\noindent
-Main message:
-Diversified portfolios of fishing permits and species reduce risk for fishermen,
-but changing species diversity can be risky, costly, and limited by regulations.
 
@@
 - in fact, specializing can be profitable within some fisheries
+
+\noindent
+Main message:
+Diversified portfolios of fishing permits and species reduce risk for fishermen,
+but changing species diversity can be risky, costly, and limited by regulations.
 
@@
 
-We obtained landings and revenue data for 
+We obtained fisheries landings and revenue data for
 all permit holders in Alaska
@@
 We calculated effective species diversity for
-permit holder $i$ and year $t$ as the inverse of 
+permit holder $i$ and time (year) $t$ as the inverse of
 Simpson's diversity index $\lambda$ [@simpson1949]
@@
 predicting $\log$ revenue, $\log(R_{i,t})$, for
-each fisher $i$ and time (year) $t$:
+each fisher $i$ and time $t$:
 
@@
   \beta_5 S_{i,t} D_{i,t} I_{i,t} +
-  \log(R_{i,t-1})
+  \log(R_{i,t-1}),
 \end{equation}
@@
 \beta_{1,j} &\sim \mathrm{N} \left(\beta_1, \sigma^2_{\beta_1}\right),\\
-\beta_{2,j} &\sim \mathrm{N} \left(\beta_1, \sigma^2_{\beta_2}\right).
+\beta_{2,j} &\sim \mathrm{N} \left(\beta_2, \sigma^2_{\beta_2}\right).
 \end{align}
@@
 greater than 200 for all parameters [@gelman2014].
+We verified that our model was coded correctly by
+comparing our estimates to an ad-hoc two-stage
+model fit with the lme4 R package [@bates2015],
+and by fitting our model to simulated data
+and recovering the true values.
 Code to reproduce our results (with a simulated data set)
@@
 
-Fishing can be big $ but is risky, for example, Alaska
-
-What's been done
-
 ## @kasperski2013

### 2016-09-14 portfolio/text/ms.Rmd
Add a section on priors

@@
 to perform Bayesian statistical inference.
+We assigned weakly informative priors on all parameters:
+$N(0, 2^2)$ priors on all 
+$\beta$, $\gamma$, and $\eta$ parameters,
+and half-$t(3, 0, 2)$ priors 
+(i.e. degrees of freedom of 3 and scale of 2)
+on all $\sigma$ parameters.
 We ran four chains and 2,000 iterations and discarded
@@
 We also ensured that $\hat{R}<1.05$ (the potential scale reduction factor)
-and that the effective sample size was
-greater than 200 for all parameters [@gelman2014].
-We verified that our model was coded correctly by
+and that the effective sample size,
+as calculated by the rstan package, 
+was greater than 200 for all parameters [@gelman2014].
+We verified that our model returned sensible estimates by
 comparing our estimates to an ad-hoc two-stage

### 2016-09-14 portfolio/text/ms.Rmd
Correct season length definition

@@
 We calculated effort or effective season length as
-the sum of days fished across each permit each year.
+the sum the season lengths for each permit each year.
+Season length for a given permit 
+was calculated as the the number of days between 
+the first instance of fishing a permit 
+and the last instance of fishing a permit each year.
 For example, if someone fished for 30 days
@@
 
-The basic model is a hierarchical linear regression with Gaussian errors
+The basic form of our model was a 
+hierarchical linear regression with Gaussian errors
 predicting $\log$ revenue, $\log(R_{i,t})$, for

### 2016-09-14 portfolio/text/ms.Rmd
Start drafting results

@@
 g2_dens <- readRDS("../data-generated/g2_prob_dens_above_zero.rds")
+n_grouped_permits <- readRDS("../data-generated/n_grouped_permits.rds")
 ```
@@
 
-Our models focused on the predictor species diversity,
+We combined similar permits into permit groupings 
+and identified sets of 
+permit groupings held by an individual 
+in a specific year as 'strategies' (Supporting Information). 
+For example, we combined permits that were the same except for vessel size,
+and we combined a number of permits that were targeting a single species 
+and differed only in region fished.
+Therefore, someone fishing herring roe in Kodiak
+and someone fishing herring roe in Alaska Peninsula
+were both considered to have a 'herring roe' permit group.
+For salmon permits, 
+we grouped permits using the same gear type across regions
+for groups that shared similar species composition. 
+After this process of combining permits,
+we were left with `r n_grouped_permits`
+unique fishing permit groups (Table S\ref{tab:strategies}),
+which we aggregated
+within each person-year combination 
+to form permit strategies.
+For example, if someone fished halibut
+and sablefish in the same year,
+their strategy for that year would be `halibut-sablefish'.
+Specific details on our strategy definitions are available in 
+the Supporting Information.
+
+Our models focused on species diversity
+as a possible predictor of revenue and revenue variability
 while controlling for effort.
@@
 Each year $t$ of revenue for fisher $i$
-is assigned a strategy indexed by $j$ and
-expected $\log(R_{i,t})$ is modelled as:
+is assigned a strategy indexed by $j$ 
+(Supporting Information, Table \ref{tab:strategies})
+and expected $\log(R_{i,t})$ is modelled as:
 
@@
 
+\noindent
+Fig. \ref{fig:ts} - limited entry + movement towards salmon and specialization in Alaska
+
+- move towards specialization, salmon fig 1D (overall declining trend disappears if you exclude salmon)
+- halibut most often combined with other permits (exclude permit holders that hold any halibut and about ninety percent of the remaining permit holders hold only one permit) fig 1c
+- Salmon the most often on its own (exclude salmon and the proportioned single permit decreases to about 70%) fig 1c
+- majority are single permit holders (70-90%) fig 1C, but the single permit holders make up only about half of the overall revenue (and surpassed half for the first time in 2013 and stayed that way) fig 1B
+
+\noindent
+Something about overall model fit
+
+- Parameters from hierarchical model (main: Fig. S\ref{fig:fe}, varying: S\ref{fig:re})
+- Residuals clearly show pattern that we see in the parameter estimates (Fig. S\ref{fig:resid})
+- Downside residuals look very similar and justify using the full data set
+(upside and downside) (Fig. S\ref{fig:resid-downside})
+- Mention the year-permit varying intercepts and what they mean 
+(Fig. S\ref{fig:str-yrs}) --- good and bad years where 
+everyone in that strategy tended to do better or worse ---
+this accounts for that and therefore compares everything to 
+the expected revenue for that year
+
+\noindent
+Fig. \ref{fig:spokes} - add permits
+
+- nearly universal reduction in variability (Fig. S\ref{fig:spoke-all})
+- e.g., `r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR] 
+lower expected standard deviation
+for halibut plus another permit vs. single halibut permits);
+- usually associated with higher median revenue 
+- but can be very expensive (provide some examples from website)
+
+\noindent
+Fig. \ref{fig:across} - across strategies
+
+- clear exponential decline in standardized variability with strategies that are inherently more species diverse 
+- not a clear effect of days fished at this level
+- fishing permits with higher species diversity reduces variability
+(`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction per unit
+increase in mean effective species diversity);
+
+\noindent
+Fig. \ref{fig:within} - within strategies
+
+- fish more days, make more money (Fig. S\ref{fig:fe})
+- after accounting for days fished, most commonly, 
+specializing is associated with higher revenue, 
+but sometimes the reverse (Fig. \ref{fig:within})
+- if we don't account for changes in effort, then higher species diversity 
+means more money fairly consistently (TODO make figure)
+- changing species diversity (number of species fished or evenness of species
+fished) tends to be associated with greater variability / uncertainty
+year-to-year (Figs. \ref{fig:within}, S\ref{fig:fe})
+- increasing species diversity within
+the constraints of a permit set has a fishery-dependent effect on revenue and
+is usually (`r g2_dens`% probability)
+associated with increased uncertainty in revenue.
+(Fig. \ref{fig:within})
+
 # Discussion

### 2016-09-14 portfolio/text/supp.Rnw
Keep developing methods
Start drafting results

@@
 
+\newlabel{eq:mean}{3}
+\newlabel{eq:cv}{7}
+
 \begin{document}
@@
 and differed only in region fished.
-For example, someone fishing herring roe in the Kodiak (G34K)
+For example, someone fishing herring roe in Kodiak (G34K)
 and someone fishing herring roe in Alaska Peninsula (G34M)
@@
 and sablefish in the same year
-their strategy for that year would be ``halibut-sablefish''.
+their strategy for that year would be `halibut-sablefish'.
 We removed people who switched strategies from one year to the next
@@
 
-\section{Hierarchical model}
+\section{Model checking with two-stage approach}
 
-We checked our models with a two-part modelling approach
-using the R package lme4.
+We checked our models with a two-stage modelling approach
+using the R package lme4 \citep{bates2015}.
 To do this, we fit the initial model of the mean component $\log(R_{i,t})$

### 2016-09-14 portfolio/text/ms.Rmd
Finish bringing in mindmap (mostly discussion)

@@
 
-\noindent
-Fig. \ref{fig:ts} - limited entry + movement towards salmon and specialization in Alaska
+### Fig. \ref{fig:ts} - limited entry + movement towards salmon and specialization in Alaska
 
@@
 
-\noindent
-Something about overall model fit
+### Something about overall model fit
 
@@
 
-\noindent
-Fig. \ref{fig:spokes} - add permits
+### Fig. \ref{fig:spokes} - add permits
 
@@
 
-\noindent
-Fig. \ref{fig:across} - across strategies
+### Fig. \ref{fig:across} - across strategies
 
@@
 
-\noindent
-Fig. \ref{fig:within} - within strategies
+### Fig. \ref{fig:within} - within strategies
 
@@
 # Discussion
+
+### Take homes / summary
+
+- What's novel
+    - Huge dataset
+    - Permit level data
+    - Account for covariates
+    - Accounting for strategies
+    - Keeping annual data (not condensing to mean and cv)
+    - Joint modelling
+    - look at different mechanisms of diversification
+- Suggests it's mostly an across strategy/permit effect (lowering variability),
+  which isn't always a possibility because of regulations and cost 
+- Adjusting within current permit constraints is risky... Not clearly risk
+  reducing to generalize from one year to the next
+- not as simple as increase diversity (within permit) and reduce risk
+- 2 dimensions to think about (revenue and variability); and the trade-off is fishery specific 
+    - effects within a permit strategy are permit specific  
+    - Importantly it's a tradeoff - 2 dimensions - sometimes make more money by specializing... Some highly variable single species permits make a ton of money... which preferable?
+
+### Reconciling the effects at different levels of species diversity
+
+- How fit in with K and H?
+- Year to year risky... But maybe evens out/better over time
+- diversity within permit - often bycatch, limited room for intentional diversification
+- Risk results not really risk, but unpredictability given our covariates ...
+don't change anything and it's most predictable how well you'll do
+- What's unpredictable in our model, might be more predictable on a year to
+year, species by species basis for individual permit holder
+- Equilibrium, average variability vs year to year effect for one person
+
+### Limitations + obvious extensions
+
+- limitations...
+- Vary by time
+- Vary by mean diversity
+- Spatial variation
+- Nonlinear effects
+- Response to specific events like EVOS and ITQs
+
+### Implications for limited entry, ITQs, and catch shares
+
+- if this is true, this is concerning because in many areas of the world and
+fisheries, regulations and market forces have encouraged specialization 
+- Permit diversity result: a privileged position - expensive to get to into
+(quote king crab license) 
+- Locking into permits limits ability to pick lower risk strategies and is
+getting people locked into low diversity permits that may be doing well now,
+but not necessarily always
+
+### Conclusions + recommendations
 

### 2016-09-19 portfolio/text/ms.Rmd
updated ms

@@
 
+Title should emphasize individuals, e.g. "Benefits of diversification and risks of changing strategies for individual fishermen/fishers"
 <!-- Fishing is inherently risky and we have long-sought regulations and fishing behaviors that reduce this risk. -->
@@
 Here, we simultaneously estimate the effects of three
-methods of diversification for Alaskan fishery permit holders
+methods of diversification for individual participants in Alaskan fisheries
 on revenue and revenue variability:
-(1) owning multiple permits,
-(2) owning permits that inherently fish higher species diversity,
+(1) diversifying by owning multiple permits,
+(2) diversifying by owning permits that inherently increase species diversity,
 and (3) increasing species diversity within the constraints of a permit set.
@@
 - in fact, specializing can be profitable within some fisheries
+- analog here in farming literature is that diversification for small scale farmers generally good, except when they specialize on high demand/high value crops (tobacco, cocoa).
+Seems like benefits of diversification are also smaller for large-scale industrial farmers.
 
@@
 for these individuals.
-For example, ...
+For example, drought and flooding are major sources of risk to agricultural food security (http://rstb.royalsocietypublishing.org/content/360/1463/2139.short),
+
 And EVOS example
@@
 
+- crop rotation has been proposed as metric for buffering against short term variation and long term trends (like climate change)
+
+BUT in some situations, external forces (prices, demand) may discourage the adoption of these diverse strategies and help promote specialization, 
+http://link.springer.com/article/10.1007/s10584-004-0710-z
+
 \bigskip
@@
 
+Most papers on fisheries have adopted the approach similar to Kasperski et al.
+and used data reported to the level of vessels or higher.
+This analysis is unique in that we're able to separate individual people from vessels (enabling us to track people that switch fishery sectors over time, or change strategies and utilize multiple vessels)
+
 \bigskip
@@
 Despite this existing work,
-key questions remain on the relationship between species diversity and
+key questions remain on the relationship between diversty of fish catches and
 fishing revenue variability.
@@
 annual revenue and revenue variability.
+
+I think we need a little here on the AK data and why that's both important and generalizable to fisheries around the world. 
+Fisheries in Alaska are the highest valued fisheries in N. America
+with landings valued at XX $ billion annually. There are a number of fishing
+strategies where both specialization and diversification are extremely profitable.
+There's a number of fishing permits that constrain the potential diversity of species landed (e.g.
+herring, halibut, sablefish, or crab permits that allow the user to only land that species) versus broader 'finfish' permits that allow users to land a wide variety of groundfish spp.
+- perhaps we also need to have a sentence or two on the implict constraints of geography and gear / vessel size on species diversification
+
 Our model makes use of the entire time series,
@@
 in a specific year as 'strategies' (Supporting Information). 
-For example, we combined permits that were the same except for vessel size,
+For example, we combined permits that were the same except for vessel size (e.g.
+a longliner fishing sablefish on a vessel under 60ft was considered to have the same strategy as a longliner fishing sablefish on a vessel over 60ft),
 and we combined a number of permits that were targeting a single species 
@@
 permit holder $i$ and time (year) $t$ as the inverse of
-Simpson's diversity index $\lambda$ [@simpson1949]
-weighted by revenue $R$ ('species diversity'):
+Simpson's diversity index (Jost 2006; also referred to in economics as the Herfindahl-Hirschman Index) $\lambda$ [@simpson1949]
+weighted by revenue $R$ ('species diversity') (also say we did this for consistency with previous analyses,e.g. Kasperski and Holland):
 
@@
 
-The basic form of our model was a 
-hierarchical linear regression with Gaussian errors
+The general framework of our model is known as a variance function regression model, which separately models change in the mean and variance as a function of covariates (Davidian and Carrol 1987 http://amstat.tandfonline.com/doi/abs/10.1080/01621459.1987.10478543; Hall and Carrol 1989, http://www.jstor.org/stable/2345837; ).
+Where traditional linear models or hierarchical linear models generally assume constant variance, the variance function regression model allows variance to be heteroscedastic.
+For application to economic data on individual fishing portfolios, this type of model allows certain strategies and levels of diversification to be more or less variable than others.
+
+We extended the basic variance function regression to a hierarchical 
+Bayesian variance function regression model with Gaussian errors
 predicting $\log$ revenue, $\log(R_{i,t})$, for
@@
 $\mathrm{div}_{i,t} < \mathrm{div}_{i,t-1}$
-(species diversity decreasing; specializing).
+(species diversity decreasing; specializing). Estimating a separate
+coefficient for the effect of increasing diversity and decreasing
+diversity allows the costs and benefits of these changes to be asymmetric.
 
-The term $\log(R_{i,t-1}$) is treated as an offset term:
-it is added on the right side of the equation
+The term $\log(R_{i,t-1}$) is treated as an offset,
 to essentially model percent change in revenue,
 i.e., $\log(R_{i,t}) - \log(R_{i,t-1}) = \log(R_{i,t}/R_{i,t-1})$.
-While this approach required filtering the data
+This formulation is necessary because of the non-stationary trends in 
+revenue (resulting from experience or other unmodeled factors). While this approach required filtering the data
 to remove person-year combinations without a subsequent year of revenue,
@@
 
-The $\beta$'s represent estimated coefficients, some of which are
-allowed to vary.
+The $\beta$'s represent estimated coefficients, some of which are allowed to vary as random effect terms.
 The intercepts $\beta_{0,j,t}$ are allowed to vary by

### 2016-09-19 portfolio/text/ms.Rmd
update

@@
     - Huge dataset
-    - Permit level data
+    - Permit level data (on individuals, not vessels)
     - Account for covariates
@@
     - Keeping annual data (not condensing to mean and cv)
-    - Joint modelling
+    - Joint modelling of mean and variance
     - look at different mechanisms of diversification
@@
 - How fit in with K and H?
+
+EW: part of the difference is individuals vs vessels.
+For example, K&H suggested that diversity declined following IFQs for halibut sablefish vessels.
+The reason for the decline was that the vessels were allowed to fish halibut for more days (because people were sharing boats).
+At the level of people, we wouldn't expect the diversity of someone who only had halibut/sablefish permit to be affected by IFQs.
+The person, not the vessel holds the quota share: https://alaskafisheries.noaa.gov/fisheries/ifq
+
 - Year to year risky... But maybe evens out/better over time
-- diversity within permit - often bycatch, limited room for intentional diversification
+- diversity within permit - often bycatch, limited room for intentional diversification (unless you have a M or salmon permit)
 - Risk results not really risk, but unpredictability given our covariates ...

### 2016-09-21 portfolio/text/ms.Rmd
Minor text edits

--- 2016-09-14 portfolio/text/ms.Rmd
@@
 for individual permit holders on a year-by-year basis is unknown ---
-existing approaches have worked at the vessel or community level
+existing approaches have worked at the vessel [@kasperski2013],
+fleet [@sethi2012b], or community level [@sethi2014],
 and condensed time series into single metrics of variability or risk.
@@
 species diversity and revenue variability ---
-a fisheries permit holder can purchase
+a fishery permit holder can purchase
 an additional permit to target other species,
@@
 simultaneously estimating the effects of
-these three approaches to species diversification
-for Alaskan fishery permit holders on
-annual revenue and revenue variability.
-Our model makes use of the entire time series,
+these three species diversification approaches
+on annual revenue and revenue variability
+for Alaskan fishery permit holders.
+Our model makes use of entire time series,
 controls for effort and latent time effects,
@@
 
-- overview
+Overview paragraph here.
 
@@
 \begin{align}
-\beta_{1,j} &\sim \mathrm{N} \left(\beta_1, \sigma^2_{\beta_1}\right),\\
-\beta_{2,j} &\sim \mathrm{N} \left(\beta_2, \sigma^2_{\beta_2}\right).
+\beta_{1,j} &\sim \mathrm{N} \left(\mu_{\beta_1}, \sigma^2_{\beta_1}\right),\\
+\beta_{2,j} &\sim \mathrm{N} \left(\mu_{\beta_2}, \sigma^2_{\beta_2}\right).
 \end{align}
@@
 \begin{align}
-\gamma_{0,j}  &\sim \mathrm{N} \left(\gamma_0 + \eta_1 M_j + \eta_2 E_j, \sigma^2_{\gamma_0}\right),\\
-\gamma_{1,j}  &\sim \mathrm{N} \left(\gamma_1, \sigma^2_{\gamma_1}\right),\\
-\gamma_{2,j}  &\sim \mathrm{N} \left(\gamma_2, \sigma^2_{\gamma_2}\right).
+\gamma_{0,j}  &\sim \mathrm{N} \left(\eta_0 + \eta_1 M_j + \eta_2 E_j, \sigma^2_{\gamma_0}\right),\\
+\gamma_{1,j}  &\sim \mathrm{N} \left(\mu_{\gamma_1}, \sigma^2_{\gamma_1}\right),\\
+\gamma_{2,j}  &\sim \mathrm{N} \left(\mu_{\gamma_2}, \sigma^2_{\gamma_2}\right).
 \end{align}
@@
 and by fitting our model to simulated data
-and recovering the true values.
+and recovering the true values (Supporting Information).
 Code to reproduce our results (with a simulated data set)

### 2016-09-21 portfolio/text/ms.Rmd

--- 2016-09-19 portfolio/text/ms.Rmd
@@
 
-Title should emphasize individuals, e.g. "Benefits of diversification and risks of changing strategies for individual fishermen/fishers"
+*Eric: Title should emphasize individuals, e.g. "Benefits of diversification and risks of changing strategies for individual fishermen/fishers"*
 <!-- Fishing is inherently risky and we have long-sought regulations and fishing behaviors that reduce this risk. -->
@@
 - in fact, specializing can be profitable within some fisheries
-- analog here in farming literature is that diversification for small scale farmers generally good, except when they specialize on high demand/high value crops (tobacco, cocoa).
-Seems like benefits of diversification are also smaller for large-scale industrial farmers.
+- analog here in farming literature is that diversification
+  for small scale farmers generally good, except when they
+  specialize on high demand/high value crops (tobacco, cocoa). 
+  Seems like benefits of diversification are also smaller for 
+  large-scale industrial farmers.
 
@@
 for these individuals.
-For example, drought and flooding are major sources of risk to agricultural food security (http://rstb.royalsocietypublishing.org/content/360/1463/2139.short),
+For example, drought and flooding are major sources of risk 
+to agricultural food security 
+(http://rstb.royalsocietypublishing.org/content/360/1463/2139.short),
 
@@
 
-- crop rotation has been proposed as metric for buffering against short term variation and long term trends (like climate change)
+- crop rotation has been proposed as metric for buffering against short term
+  variation and long term trends (like climate change)
 
@@
 
-Most papers on fisheries have adopted the approach similar to Kasperski et al.
-and used data reported to the level of vessels or higher.
-This analysis is unique in that we're able to separate individual people from vessels (enabling us to track people that switch fishery sectors over time, or change strategies and utilize multiple vessels)
+Most papers on fisheries have adopted the approach similar to Kasperski et
+al. and used data reported to the level of vessels or higher. This analysis
+is unique in that we're able to separate individual people from vessels
+(enabling us to track people that switch fishery sectors over time, or
+change strategies and utilize multiple vessels)
 
@@
 Despite this existing work,
-key questions remain on the relationship between diversty of fish catches and
+key questions remain on the relationship between 
+diversity of fish catches and
 fishing revenue variability.
@@
 annual revenue and revenue variability.
+Our model makes use of entire time series,
+controls for effort and latent time effects,
+and allows the effect of species diversity
+to vary by fishing strategy (sets of similar permits)
+in a hierarchical model.
 
-I think we need a little here on the AK data and why that's both important and generalizable to fisheries around the world. 
+*Eric: I think we need a little here on the AK data and why that's both important and generalizable to fisheries around the world. 
 Fisheries in Alaska are the highest valued fisheries in N. America
@@
 herring, halibut, sablefish, or crab permits that allow the user to only land that species) versus broader 'finfish' permits that allow users to land a wide variety of groundfish spp.
-- perhaps we also need to have a sentence or two on the implict constraints of geography and gear / vessel size on species diversification
-
-Our model makes use of the entire time series,
-controls for effort and latent time effects,
-and allows the effect of species diversity
-to vary by fishing strategy (sets of similar permits)
-in a hierarchical model.
-
-<!-- Multiple ways one can diversify: -->
-<!-- buy another permit, pick diverse permit, diversify within existing constraints. -->
+- perhaps we also need to have a sentence or two on the implicit constraints of geography and gear / vessel size on species diversification*
 
@@
 in a specific year as 'strategies' (Supporting Information). 
-For example, we combined permits that were the same except for vessel size (e.g.
-a longliner fishing sablefish on a vessel under 60ft was considered to have the same strategy as a longliner fishing sablefish on a vessel over 60ft),
+For example, we combined permits that were the same except for vessel size 
+(e.g. a longliner fishing sablefish on a vessel under 60ft was 
+considered to have the same strategy as a longliner fishing 
+sablefish on a vessel over 60ft),
 and we combined a number of permits that were targeting a single species 
@@
 permit holder $i$ and time (year) $t$ as the inverse of
-Simpson's diversity index (Jost 2006; also referred to in economics as the Herfindahl-Hirschman Index) $\lambda$ [@simpson1949]
-weighted by revenue $R$ ('species diversity') (also say we did this for consistency with previous analyses,e.g. Kasperski and Holland):
+Simpson's diversity index 
+(TODO - Jost 2006; also referred to in economics as the Herfindahl-Hirschman Index) 
+$\lambda$ [@simpson1949]
+weighted by revenue $R$ ('species diversity') 
+*Eric: (also say we did this for consistency with previous analyses,e.g. Kasperski and Holland):*
 
@@
 
-The general framework of our model is known as a variance function regression model, which separately models change in the mean and variance as a function of covariates (Davidian and Carrol 1987 http://amstat.tandfonline.com/doi/abs/10.1080/01621459.1987.10478543; Hall and Carrol 1989, http://www.jstor.org/stable/2345837; ).
-Where traditional linear models or hierarchical linear models generally assume constant variance, the variance function regression model allows variance to be heteroscedastic.
-For application to economic data on individual fishing portfolios, this type of model allows certain strategies and levels of diversification to be more or less variable than others.
+The general framework of our model is known as a variance function
+regression model, which separately models change in the mean and variance as
+a function of covariates (TODO - Davidian and Carrol 1987
+http://amstat.tandfonline.com/doi/abs/10.1080/01621459.1987.10478543; Hall
+and Carrol 1989, http://www.jstor.org/stable/2345837). Where traditional
+linear models or hierarchical linear models generally assume constant
+variance, the variance function regression model allows variance to be
+heteroskedastic. For application to economic data on individual fishing
+portfolios, this type of model allows certain strategies and levels of
+diversification to be more or less variable than others.
 
@@
 
-The $\beta$'s represent estimated coefficients, some of which are allowed to vary as random effect terms.
+The $\beta$'s represent estimated coefficients, some of which are allowed to
+vary as 'random effect' terms.
 The intercepts $\beta_{0,j,t}$ are allowed to vary by
@@
 
-EW: part of the difference is individuals vs vessels.
-For example, K&H suggested that diversity declined following IFQs for halibut sablefish vessels.
-The reason for the decline was that the vessels were allowed to fish halibut for more days (because people were sharing boats).
-At the level of people, we wouldn't expect the diversity of someone who only had halibut/sablefish permit to be affected by IFQs.
-The person, not the vessel holds the quota share: https://alaskafisheries.noaa.gov/fisheries/ifq
-
+- EW: part of the difference is individuals vs vessels. For example, K&H
+suggested that diversity declined following IFQs for halibut sablefish
+vessels. The reason for the decline was that the vessels were allowed to
+fish halibut for more days (because people were sharing boats). At the level
+of people, we wouldn't expect the diversity of someone who only had
+halibut/sablefish permit to be affected by IFQs. The person, not the vessel
+holds the quota share: https://alaskafisheries.noaa.gov/fisheries/ifq
 - Year to year risky... But maybe evens out/better over time

### 2016-09-21 portfolio/text/anderson-etal-rev-pfx.tex
Start drafting results
Tweak text after Eric edits

--- 2016-08-16 portfolio/text/anderson-etal-rev-pfx.tex
@@
 
-\title{Portfolios of fishing permits and species
-reduce risk but adjusting fishing strategy is risky}
+\title{Benefits of diversification and risks of 
+changing strategies for individual fishers\\
+Portfolios of fishing permits and species
+reduce risk but adjusting fishing strategy is risky\\
+Portfolio benefits to individual fishers
+depend on approach to diversification\\
+Portfolio benefits of income diversification
+are unavailable to many fishermen\\
+Effect of species diversity on income variability
+for individuals depends on fishing strategy\\
+Effect of species diversity on revenue risk
+for individual fishers
+depends on diversification approach\\
+}
 \author{
@@
 Jennifer C. Shriver,
-Jordan T. Watson
+Jordan T. Watson,
+\ldots
 }
+
+\newlabel{tab:strategies}{1}
+\newlabel{fig:fe}{1}
+\newlabel{fig:re}{2}
+\newlabel{fig:spoke-all}{3}
+\newlabel{fig:resid}{4}
+\newlabel{fig:resid-downside}{5}
+\newlabel{fig:str-yrs}{6}
 

### 2016-09-21 portfolio/text/ms.Rmd
Tweak text after Eric edits

@@
 
-*Eric: Title should emphasize individuals, e.g. "Benefits of diversification and risks of changing strategies for individual fishermen/fishers"*
 <!-- Fishing is inherently risky and we have long-sought regulations and fishing behaviors that reduce this risk. -->
@@
 - in fact, specializing can be profitable within some fisheries
-- analog here in farming literature is that diversification
+- Eric: analog here in farming literature is that diversification
   for small scale farmers generally good, except when they
@@
 (http://rstb.royalsocietypublishing.org/content/360/1463/2139.short),
-
 And EVOS example
@@
   variation and long term trends (like climate change)
-
-BUT in some situations, external forces (prices, demand) may discourage the adoption of these diverse strategies and help promote specialization, 
+- BUT in some situations, external forces (prices, demand) may discourage the
+  adoption of these diverse strategies and help promote specialization, 
 http://link.springer.com/article/10.1007/s10584-004-0710-z
@@
 At vessel level, ... [@kasperski2013] ...
-
-Most papers on fisheries have adopted the approach similar to Kasperski et
+- Eric: Most papers on fisheries have adopted the approach similar to Kasperski et
 al. and used data reported to the level of vessels or higher. This analysis
@@
 There's a number of fishing permits that constrain the potential diversity of species landed (e.g.
-herring, halibut, sablefish, or crab permits that allow the user to only land that species) versus broader 'finfish' permits that allow users to land a wide variety of groundfish spp.
-- perhaps we also need to have a sentence or two on the implicit constraints of geography and gear / vessel size on species diversification*
+herring, halibut, sablefish, or crab permits that allow the user to only land that species) versus broader 'finfish' permits that allow users to land a wide variety of groundfish spp.*
+*perhaps we also need to have a sentence or two on the implicit constraints of geography and gear / vessel size on species diversification*
 
@@
 In particular, we removed permit holders whose median revenue
-was less than $10,000 2009 USD.
+was less than \$10,000 2009 USD.
 
@@
 For example, we combined permits that were the same except for vessel size 
-(e.g. a longliner fishing sablefish on a vessel under 60ft was 
+(e.g. a longliner fishing sablefish on a vessel under 60 ft was 
 considered to have the same strategy as a longliner fishing 
-sablefish on a vessel over 60ft),
+sablefish on a vessel over 60 ft),
 and we combined a number of permits that were targeting a single species 

### 2016-09-21 portfolio/text/supp.Rnw
Tweak text after Eric edits

@@
 two times their standard deivations)
-to the magnitude of residual error.}
+to the magnitude of residual error. (TODO symbols need to be updated)}
 \label{fig:fe}

### 2016-09-21 portfolio/text/ms.Rmd
Add refs

@@
 For example, drought and flooding are major sources of risk 
-to agricultural food security 
-(http://rstb.royalsocietypublishing.org/content/360/1463/2139.short),
+to agricultural food security [@gregory2005],
 And EVOS example
@@
   variation and long term trends (like climate change)
-- BUT in some situations, external forces (prices, demand) may discourage the
-  adoption of these diverse strategies and help promote specialization, 
-http://link.springer.com/article/10.1007/s10584-004-0710-z
+- but in some situations, external forces (prices, demand) may discourage the
+  adoption of these diverse strategies and 
+  help promote specialization [@bradshaw2004]
 
@@
 At vessel level, ... [@kasperski2013] ...
-- Eric: Most papers on fisheries have adopted the approach similar to Kasperski et
-al. and used data reported to the level of vessels or higher. This analysis
+- Eric: Most papers on fisheries have adopted the approach similar to @kasperski2013 and used data reported to the level of vessels or higher. This analysis
 is unique in that we're able to separate individual people from vessels
@@
 while controlling for effort.
-We calculated effective species diversity for
+For consistency with previous analyses [e.g. @kasperski2013],
+we calculated effective species diversity for
 permit holder $i$ and time (year) $t$ as the inverse of
-Simpson's diversity index 
-(TODO - Jost 2006; also referred to in economics as the Herfindahl-Hirschman Index) 
-$\lambda$ [@simpson1949]
-weighted by revenue $R$ ('species diversity') 
-*Eric: (also say we did this for consistency with previous analyses,e.g. Kasperski and Holland):*
+Simpson's diversity index $\lambda$
+(\citeauthor{simpson1949} \citeyear{simpson1949},
+\citeauthor{jost2006} \citeyear{jost2006};
+also referred to in economics as the Herfindahl-Hirschman Index)
+weighted by revenue $R$:
 
@@
 regression model, which separately models change in the mean and variance as
-a function of covariates (TODO - Davidian and Carrol 1987
-http://amstat.tandfonline.com/doi/abs/10.1080/01621459.1987.10478543; Hall
-and Carrol 1989, http://www.jstor.org/stable/2345837). Where traditional
+a function of covariates [e.g. @davidian1987; @hall1989]. Where traditional
 linear models or hierarchical linear models generally assume constant

### 2016-09-22 portfolio/text/ms.Rmd
ole edits

@@
+---
+output: html_document
+---
 ```{r set-knitr-options, cache=FALSE, echo=FALSE}
@@
 
-# Abstract
+<!-- Ole likes the first or fourth title best, "
+Benefit of diversification are unavailable to many fishermen
+" -->
+
+# Ole Abstract
 
@@
 <!-- The increasing the diversity of species fished has been proposed to reduce risk for fishermen -->
+
+A fundamental challenge for individuals who rely on the extraction of natural resource for their livelihood is income variability.  
+Fishers, loggers, farmers, miners, and other  all face uncertainty in income driven by a mix of environmental, biological, and economic (e.g. price) varibility.
+A key component underlying the sustainability of these industries is identifying both regulatory actions and individual behavior affect income variability and by extension the economic stability and sustainable use of natural resources.
+
+For commercial fisheries, recent work has identified that targeting a diversity of species can reduce income variability.
+However, any diversity-variability relationship is embedded in a regulatory framework that restricts individual ability to unilaterally change fishing behavior.  
+Here we identify three primary axes by which fishers in Alaska, one of the world's most productive fishery regions, can potentailly diversify fishing activities and understand how individual actions along each diversity axis affect both revenue and revenue variability.
+We construct a novel statistical model to simultaneously assess the value to individuals fishers of choosing to 1) own permits that allow fishers that target a diverse fish species assemblage, 2) diversify by owning multiple permits, or 3) catch a greater species within the constraints of a given permit.
+
+Thus, we present a unique approach to understanding important aspects of diversity relevant to individaul fishers' decisions. 
+We show contrastings effects of diversity on revenue and revenue variability along the three axes.
+First, we show (results for choosing a permit).
+Next we show (results for owning multiple permits).
+Finally we show (results for individual permit diversification).
+
+Thus, while there are clear benefits to some flavors of diversification, others don't do much.
+Importantly, the most effective measure for increasing revenue and decreasing risk - purchasing additional permits to participate in additional fisheries - was also the most costly
+
+We find that
+fishing an additional permit is associated with higher revenue
+and lower variability (e.g.,
+`r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR] lower
+expected standard deviation
+for halibut plus another permit vs. single halibut permits);
+fishing permits with higher species diversity reduces variability
+(`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction per unit
+increase in mean effective species diversity);
+but that increasing species diversity within
+the constraints of a permit set has a fishery-dependent effect on revenue and
+is usually (`r g2_dens`% probability)
+associated with increased uncertainty in revenue.
+At the same time, Alaskan commercial fisheries
+have been increasingly focussed on
+single-permit (often salmon) fishing strategies
+with single-permit fishers earning
+more than multiple-permit fishers
+for the first time in 2013.
+Our results suggest that the benefit of species diversity
+for revenue variability
+is mostly an across-permit effect, not
+a benefit that can be easily harnessed by an individual permit holder.
+Effective methods of diversification ---
+purchasing additional permits or switching permits ---
+are costly and often limited by regulations such as limited entry
+programs and individual transferable quotas (ITQs).
+Take home ...
+
+# Original Abstract
 
@@
 
-Targeting a higher diversity of species
-has been proposed as a risk-buffering mechanism...
+Increasing diversity has been proposed as a risk-buffering mechanism...
 Give examples from other sectors
 
-- crop rotation has been proposed as metric for buffering against short term
+- Farming: crop rotation has been proposed as metric for buffering against short term
   variation and long term trends (like climate change)
+- Logging: Maintaining a diverse mix of tree species, age structure, and space to buffer against things like wildfires and price varition. 
 - but in some situations, external forces (prices, demand) may discourage the
@@
 At vessel level, ... [@kasperski2013] ...
+However, these papers look at aggregates across individuals and largely ignore how existing fisheries regulations limit the ability of individual fishers to diversify.
+Key uncertainties remain about the fundamental relationship between diversity, fishing revenue, and fishing revenue variability at the level of the individual.
+While the days of open, unregulated fisheries in most developed countries have passed - fishers must own a license to participate in virtually all fisheries - fishers retain several options for increasing diversity.
+For example, fishers can do thing X, thing Y, or thing Z.
+But uncertainty remains about the value of adopting different diversification strategies for improving revenue and reducing revenue variability.
+Furthermore, theory and empirical studies both suggest there can be substantial benefits to specialization (REFS?)
+
 - Eric: Most papers on fisheries have adopted the approach similar to @kasperski2013 and used data reported to the level of vessels or higher. This analysis
@@
 
-Despite this existing work,
-key questions remain on the relationship between 
-diversity of fish catches and
-fishing revenue variability.
+-Ole: I think this is the spot for intro to AK.
+I moved some of Eric's stuff up here and added to it.[*Eric: I think we need a little here on the AK data and why that's both important and generalizable to fisheries around the world.]
+
+Fisheries in Alaska are the highest valued fisheries in N. America
+with landings valued at XX $ billion annually representing harvests from YY species.
+Alaska fisheries support a diverse fishing fleets ranging from small, open, single operator vessels to multi-million dollar state of the art ships with large crews.
+Similar diversity in fishing strategies can be found within the Alaska fleet with different individuals adopting fishing strategies that range from highly specialized (e.g.
+fishing only one species with one gear type or in one area) to highly diverse (e.g.
+fishing multiple species with multiple gear types in multiple areas).
+This broad range of fishing types provide an unique opportunity for understanding the role diversity plays in determining both revenue and risk for individual fishers.
+
+We develop a novel statistical approach for understanding the effect of diveristy on both revenue and risk that builds on results from the vessel [@kasperski2013], fleet [@sethi2012b], and community levels [@sethi2014].
+First, we  summarize trends in diversification across fishing fleets in Alaska.
+Next, we identified three main axes by which fishers may diversify: 1), 2), 3).
+Finally, we constructed a model to simultaneously estimate the relationship between each of these three axes of diversity and both annual revenue and risk for individual fishers over nearly ZZ years.
+Our hierachical model makes use of entire time series,
+controls for effort and latent time effects,
+and allows the effect of species diversity
+to vary by fishing strategy (sets of similar permits).
+It is the first model that jointly explores the effect of diversity on revenue and risk at the individual level and the first considers the multiple axes of diversity simultaneously.
+
+strategies where both specialization and diversification are extremely profitable.
+There's a number of fishing permits that constrain the potential diversity of species landed (e.g.
+herring, halibut, sablefish, or crab permits that allow the user to only land that species) versus broader 'finfish' permits that allow users to land a wide variety of groundfish spp.*
+*perhaps we also need to have a sentence or two on the implicit constraints of geography and gear / vessel size on species diversification*
+
 First, the effect of changing species diversity
@@
 
-The general framework of our model is known as a variance function
-regression model, which separately models change in the mean and variance as
+-Ole: I'd start with a little reminder of what the purpose of the model is for. Like:
+To jointly model revenue and revenue variability as a function of individual-level covariates, we use a variance function regression model, which separately models change in the mean and variance as
 a function of covariates [e.g. @davidian1987; @hall1989]. Where traditional
 linear models or hierarchical linear models generally assume constant
-variance, the variance function regression model allows variance to be
+variance [across what? covariates?], the variance function regression model allows variance to be
 heteroskedastic. For application to economic data on individual fishing
@@
 and expected $\log(R_{i,t})$ is modelled as:
+OLE: I might swap out log(R)-hat for something like mu(or your favorit greek letter) just to avoid 
+potential confusion like using the term “expected log(R)”.  I also just don't like hats... but that's my problem.
 
@@
 
-The term $\log(R_{i,t-1}$) is treated as an offset,
-to essentially model percent change in revenue,
+We use $\log(R_{i,t-1}$) as an offset, and thus
+our model describes the percent change in revenue,
 i.e., $\log(R_{i,t}) - \log(R_{i,t-1}) = \log(R_{i,t}/R_{i,t-1})$.
-This formulation is necessary because of the non-stationary trends in 
-revenue (resulting from experience or other unmodeled factors). While this approach required filtering the data
-to remove person-year combinations without a subsequent year of revenue,
-it has a number of advantages:
+We model change in revenue because preliminary analysis revealed non-stationary trends in 
+revenue (resulting from experience or other unmodeled factors). This approach required filtering the data
+to remove person-year combinations without a subsequent year of revenue, but
+presents a number of advantages:
 it eliminates having to deal with the non-stationary
 time series and eliminates having to
-estimate thousands of people-specific intercepts.
+estimate thousands of intercepts associated individual fishers.
 
@@
 are centered on zero, and are constrained by a normal distribution:
+
+Ole: I'd use tau (or something) for the variances here so that people don't have to look at the subscripts to differentiate between variances at different hierarchical levels.
 

### 2016-09-26 portfolio/text/ms.Rmd
Incorporate Ole's edits

@@
----
-output: html_document
----
 ```{r set-knitr-options, cache=FALSE, echo=FALSE}
@@
 
-<!-- Ole likes the first or fourth title best, "
-Benefit of diversification are unavailable to many fishermen
-" -->
-
-# Ole Abstract
-
-<!-- Fishing is inherently risky and we have long-sought regulations and fishing behaviors that reduce this risk. -->
-<!-- The increasing the diversity of species fished has been proposed to reduce risk for fishermen -->
-
-A fundamental challenge for individuals who rely on the extraction of natural resource for their livelihood is income variability.  
-Fishers, loggers, farmers, miners, and other  all face uncertainty in income driven by a mix of environmental, biological, and economic (e.g. price) varibility.
-A key component underlying the sustainability of these industries is identifying both regulatory actions and individual behavior affect income variability and by extension the economic stability and sustainable use of natural resources.
-
-For commercial fisheries, recent work has identified that targeting a diversity of species can reduce income variability.
-However, any diversity-variability relationship is embedded in a regulatory framework that restricts individual ability to unilaterally change fishing behavior.  
-Here we identify three primary axes by which fishers in Alaska, one of the world's most productive fishery regions, can potentailly diversify fishing activities and understand how individual actions along each diversity axis affect both revenue and revenue variability.
-We construct a novel statistical model to simultaneously assess the value to individuals fishers of choosing to 1) own permits that allow fishers that target a diverse fish species assemblage, 2) diversify by owning multiple permits, or 3) catch a greater species within the constraints of a given permit.
-
-Thus, we present a unique approach to understanding important aspects of diversity relevant to individaul fishers' decisions. 
-We show contrastings effects of diversity on revenue and revenue variability along the three axes.
-First, we show (results for choosing a permit).
-Next we show (results for owning multiple permits).
-Finally we show (results for individual permit diversification).
-
-Thus, while there are clear benefits to some flavors of diversification, others don't do much.
-Importantly, the most effective measure for increasing revenue and decreasing risk - purchasing additional permits to participate in additional fisheries - was also the most costly
-
-We find that
-fishing an additional permit is associated with higher revenue
+A fundamental challenge for individuals who rely on the extraction of natural
+resource for their livelihood is income variability.
+Fishers, loggers, farmers, miners, and others
+all face uncertainty in income driven by a mix of
+environmental, biological, and economic (e.g. price) variability.
+A key component underlying the sustainability 
+of these industries is identifying both
+regulatory actions and individual behaviour affect income variability 
+and by extension the economic stability and 
+sustainable use of natural resources.   
+For commercial fisheries, recent work has identified that 
+targeting a diversity of species can reduce income variability.
+However, any diversity-variability relationship 
+is embedded in a regulatory framework that restricts individual ability 
+to unilaterally change fishing behaviour.  
+Here we identify three primary axes by which fishers in Alaska, 
+one of the world's most productive fishery regions, can 
+potentially diversify fishing activities and understand how 
+individual actions along each diversity axis affect 
+both revenue and revenue variability. 
+We construct a novel statistical model to simultaneously 
+assess the value to individuals fishers of choosing to 
+(1) own permits that allow fishers that target a diverse fish species assemblage, 
+(2) diversify by owning multiple permits, or 
+(3) catch a greater species within the constraints of a given permit.  
+Thus, we present a unique approach to understanding important aspects of
+diversity relevant to individual fishers' decisions. 
+We show contrasting effects of diversity on 
+revenue and revenue variability along the three axes. 
+First, we show that fishing permits with 
+higher species diversity reduces variability
+(`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction per unit
+increase in mean effective species diversity).
+Next, we show that fishing an additional permit 
+is associated with higher revenue
 and lower variability (e.g.,
@@
 for halibut plus another permit vs. single halibut permits);
-fishing permits with higher species diversity reduces variability
-(`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction per unit
-increase in mean effective species diversity);
-but that increasing species diversity within
+Finally, we show that increasing species diversity within
 the constraints of a permit set has a fishery-dependent effect on revenue and
 is usually (`r g2_dens`% probability)
-associated with increased uncertainty in revenue.
-At the same time, Alaskan commercial fisheries
-have been increasingly focussed on
-single-permit (often salmon) fishing strategies
-with single-permit fishers earning
-more than multiple-permit fishers
-for the first time in 2013.
-Our results suggest that the benefit of species diversity
-for revenue variability
-is mostly an across-permit effect, not
-a benefit that can be easily harnessed by an individual permit holder.
-Effective methods of diversification ---
-purchasing additional permits or switching permits ---
-are costly and often limited by regulations such as limited entry
-programs and individual transferable quotas (ITQs).
-Take home ...
+associated with increased uncertainty in revenue the following year.
 
-# Original Abstract
+Thus, while there are clear benefits 
+to some flavours of diversification, others don't do much.
+Importantly, the most effective measure for 
+increasing revenue and decreasing risk --- 
+purchasing additional permits to participate in additional fisheries --- 
+was also the most costly.
 
-Individuals relying on natural resources as their primary source of income
-often experience high levels of income variability
-due to the combined effects of environmental, biological, and price variation.
-Commercial fishers are a prime example and
-recent work has sought to identify regulations and fishing behaviours
-that can reduce this variability.
-One candidate has been that fishing higher levels of species diversity
-decreases income variability.
-However, it is unknown if this species-diversity effect
-on income variability holds
-across multiple approaches to diversifying catch,
-at the level of individual permit holders,
-and across different fishing strategies.
-Here, we simultaneously estimate the effects of three
-methods of diversification for individual participants in Alaskan fisheries
-on revenue and revenue variability:
-(1) diversifying by owning multiple permits,
-(2) diversifying by owning permits that inherently increase species diversity,
-and (3) increasing species diversity within the constraints of a permit set.
-We control for changes in effort and other
-latent effects through time.
-We find that
-fishing an additional permit is associated with higher revenue
-and lower variability (e.g.,
-`r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR] lower
-expected standard deviation
-for halibut plus another permit vs. single halibut permits);
-fishing permits with higher species diversity reduces variability
-(`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction per unit
-increase in mean effective species diversity);
-but that increasing species diversity within
-the constraints of a permit set has a fishery-dependent effect on revenue and
-is usually (`r g2_dens`% probability)
-associated with increased uncertainty in revenue.
 At the same time, Alaskan commercial fisheries
@@
   variation and long term trends (like climate change)
-- Logging: Maintaining a diverse mix of tree species, age structure, and space to buffer against things like wildfires and price varition. 
+- Logging: Maintaining a diverse mix of tree species, age structure, and space
+  to buffer against things like wildfires and price variation. 
 - but in some situations, external forces (prices, demand) may discourage the
@@
 At vessel level, ... [@kasperski2013] ...
-However, these papers look at aggregates across individuals and largely ignore how existing fisheries regulations limit the ability of individual fishers to diversify.
-Key uncertainties remain about the fundamental relationship between diversity, fishing revenue, and fishing revenue variability at the level of the individual.
-While the days of open, unregulated fisheries in most developed countries have passed - fishers must own a license to participate in virtually all fisheries - fishers retain several options for increasing diversity.
-For example, fishers can do thing X, thing Y, or thing Z.
-But uncertainty remains about the value of adopting different diversification strategies for improving revenue and reducing revenue variability.
-Furthermore, theory and empirical studies both suggest there can be substantial benefits to specialization (REFS?)
+However, these papers look at aggregates across individuals and largely ignore
+how existing fisheries regulations limit the ability of individual fishers to
+diversify. Key uncertainties remain about the fundamental relationship between
+diversity, fishing revenue, and fishing revenue variability at the level of the
+individual.  
+While the days of open, unregulated fisheries in most developed countries have
+passed - fishers must own a license to participate in virtually all fisheries
+- fishers retain several options for increasing diversity. For example, fishers
+can do thing X, thing Y, or thing Z. But uncertainty remains about the value
+of adopting different diversification strategies for improving revenue and
+reducing revenue variability. Furthermore, theory and empirical studies both
+suggest there can be substantial benefits to specialization (REFS?)
 
@@
 
--Ole: I think this is the spot for intro to AK.
-I moved some of Eric's stuff up here and added to it.[*Eric: I think we need a little here on the AK data and why that's both important and generalizable to fisheries around the world.]
+-Ole: I think this is the spot for intro to AK.  I moved some of Eric's stuff
+up here and added to it.[*Eric: I think we need a little here on the AK data
+and why that's both important and generalizable to fisheries around the world.]
 
 Fisheries in Alaska are the highest valued fisheries in N. America
-with landings valued at XX $ billion annually representing harvests from YY species.
-Alaska fisheries support a diverse fishing fleets ranging from small, open, single operator vessels to multi-million dollar state of the art ships with large crews.
-Similar diversity in fishing strategies can be found within the Alaska fleet with different individuals adopting fishing strategies that range from highly specialized (e.g.
-fishing only one species with one gear type or in one area) to highly diverse (e.g.
-fishing multiple species with multiple gear types in multiple areas).
-This broad range of fishing types provide an unique opportunity for understanding the role diversity plays in determining both revenue and risk for individual fishers.
+with landings valued at XX $ billion annually representing harvests from YY
+species.  Alaska fisheries support a diverse fishing fleets ranging from small,
+open, single operator vessels to multi-million dollar state of the art ships
+with large crews. Similar diversity in fishing strategies can be found within
+the Alaska fleet with different individuals adopting fishing strategies that
+range from highly specialized (e.g. fishing only one species with one gear type
+or in one area) to highly diverse (e.g. fishing multiple species with multiple
+gear types in multiple areas). This broad range of fishing types provide an
+unique opportunity for understanding the role diversity plays in determining
+both revenue and risk for individual fishers.
 
-We develop a novel statistical approach for understanding the effect of diveristy on both revenue and risk that builds on results from the vessel [@kasperski2013], fleet [@sethi2012b], and community levels [@sethi2014].
+We develop a novel statistical approach for understanding the effect of
+diversity on both revenue and risk that builds on results from the vessel
+[@kasperski2013], fleet [@sethi2012b], and community levels [@sethi2014].
 First, we  summarize trends in diversification across fishing fleets in Alaska.
 Next, we identified three main axes by which fishers may diversify: 1), 2), 3).
-Finally, we constructed a model to simultaneously estimate the relationship between each of these three axes of diversity and both annual revenue and risk for individual fishers over nearly ZZ years.
-Our hierachical model makes use of entire time series,
+Finally, we constructed a model to simultaneously estimate the relationship
+between each of these three axes of diversity and both annual revenue and risk
+for individual fishers over nearly ZZ years. 
+Our hierarchical model makes use of
+entire time series,
 controls for effort and latent time effects,
@@
 to vary by fishing strategy (sets of similar permits).  
-It is the first model that jointly explores the effect of diversity on revenue and risk at the individual level and the first considers the multiple axes of diversity simultaneously.
+It is the first model that jointly explores the effect of diversity on revenue
+and risk at the individual level and the first considers the multiple axes of
+diversity simultaneously.
 
-strategies where both specialization and diversification are extremely profitable.
-There's a number of fishing permits that constrain the potential diversity of species landed (e.g.
-herring, halibut, sablefish, or crab permits that allow the user to only land that species) versus broader 'finfish' permits that allow users to land a wide variety of groundfish spp.*
-*perhaps we also need to have a sentence or two on the implicit constraints of geography and gear / vessel size on species diversification*
+strategies where both specialization and diversification are extremely
+profitable. There's a number of fishing permits that constrain the potential
+diversity of species landed (e.g. herring, halibut, sablefish, or crab permits
+that allow the user to only land that species) versus broader 'finfish' permits
+that allow users to land a wide variety of groundfish spp.*
+
+*perhaps we also need to have a sentence or two on the implicit constraints of
+geography and gear / vessel size on species diversification*
 
@@
 -Ole: I'd start with a little reminder of what the purpose of the model is for. Like:
-To jointly model revenue and revenue variability as a function of individual-level covariates, we use a variance function regression model, which separately models change in the mean and variance as
+To jointly model revenue and revenue variability as a function of
+individual-level covariates, we use a variance function regression model, which
+separately models change in the mean and variance as
 a function of covariates [e.g. @davidian1987; @hall1989]. Where traditional
 linear models or hierarchical linear models generally assume constant
-variance [across what? covariates?], the variance function regression model allows variance to be
+variance [Ole: across what? covariates?], the variance function regression model allows variance to be
 heteroskedastic. For application to economic data on individual fishing
@@
 and expected $\log(R_{i,t})$ is modelled as:
-OLE: I might swap out log(R)-hat for something like mu(or your favorit greek letter) just to avoid 
+Ole: I might swap out log(R)-hat for something like mu(or your favorit greek letter) just to avoid 
 potential confusion like using the term “expected log(R)”.  I also just don't like hats... but that's my problem.

### 2016-09-26 portfolio/text/ms.Rmd
Strip trailing white space
Editing of paper

@@
 
+# Abstract
+
 A fundamental challenge for individuals who rely on the extraction of natural
-resource for their livelihood is income variability.
+resources for their livelihood is income variability.
 Fishers, loggers, farmers, miners, and others
 all face uncertainty in income driven by a mix of
-environmental, biological, and economic (e.g. price) variability.
+environmental, biological, and economic variability.
 A key component underlying the sustainability
 of these industries is identifying both
-regulatory actions and individual behaviour affect income variability 
+how regulatory actions and individual behaviour affect income variability
 and by extension the economic stability and
@@
 one of the world's most productive fishery regions, can
-potentially diversify fishing activities and understand how 
-individual actions along each diversity axis affect 
+potentially diversify fishing activities and
+investigate how individual actions along each diversity axis affect
 both revenue and revenue variability.
 We construct a novel statistical model to simultaneously
-assess the value to individuals fishers of choosing to 
-(1) own permits that allow fishers that target a diverse fish species assemblage, 
+assess the value to individual fishers of choosing to
+(1) own permits that allow targeting a diverse fish species assemblage,
 (2) diversify by owning multiple permits, or
-(3) catch a greater species within the constraints of a given permit.  
-Thus, we present a unique approach to understanding important aspects of
-diversity relevant to individual fishers' decisions. 
+(3) catch a greater species diversity within the constraints of a given permit.
 We show contrasting effects of diversity on
@@
 expected standard deviation
-for halibut plus another permit vs. single halibut permits);
+for halibut plus another permit vs. single halibut permits).
 Finally, we show that increasing species diversity within
@@
 associated with increased uncertainty in revenue the following year.
-
 Thus, while there are clear benefits
-to some flavours of diversification, others don't do much.
+to some axes of diversification, others are less effective.
 Importantly, the most effective measure for
@@
 have been increasingly focussed on
-single-permit (often salmon) fishing strategies
+single-permit (often salmon) fishing strategies,
 with single-permit fishers earning
@@
 
--Ole: I think this is the spot for intro to AK.  I moved some of Eric's stuff
-up here and added to it.[*Eric: I think we need a little here on the AK data
-and why that's both important and generalizable to fisheries around the world.]
+<!-- -Ole: I think this is the spot for intro to AK.  I moved some of Eric's stuff -->
+<!-- up here and added to it.[*Eric: I think we need a little here on the AK data -->
+<!-- and why that's both important and generalizable to fisheries around the world.] -->
 
@@
 in a hierarchical model.
-
-*Eric: I think we need a little here on the AK data and why that's both important and generalizable to fisheries around the world. 
-Fisheries in Alaska are the highest valued fisheries in N. America
-with landings valued at XX $ billion annually. There are a number of fishing
-strategies where both specialization and diversification are extremely profitable.
-There's a number of fishing permits that constrain the potential diversity of species landed (e.g.
-herring, halibut, sablefish, or crab permits that allow the user to only land that species) versus broader 'finfish' permits that allow users to land a wide variety of groundfish spp.*
-*perhaps we also need to have a sentence or two on the implicit constraints of geography and gear / vessel size on species diversification*
 
@@
 
--Ole: I'd start with a little reminder of what the purpose of the model is for. Like:
 To jointly model revenue and revenue variability as a function of
@@
 separately models change in the mean and variance as
-a function of covariates [e.g. @davidian1987; @hall1989]. Where traditional
-linear models or hierarchical linear models generally assume constant
-variance [Ole: across what? covariates?], the variance function regression model allows variance to be
-heteroskedastic. For application to economic data on individual fishing
+a function of covariates [e.g. @davidian1987; @hall1989]. 
+Whereas traditional linear models or 
+hierarchical linear models generally assume
+homoscedasticity --- that the residual variance 
+is the same for all values of the predictor variables ---
+the variance function regression model allows variance to be heteroskedastic. 
+For application to economic data on individual fishing
 portfolios, this type of model allows certain strategies and levels of
@@
 and expected $\log(R_{i,t})$ is modelled as:
-Ole: I might swap out log(R)-hat for something like mu(or your favorit greek letter) just to avoid 
-potential confusion like using the term “expected log(R)”.  I also just don't like hats... but that's my problem.
+<!-- Ole: I might swap out log(R)-hat for something like mu(or your favorit greek letter) just to avoid -->
+<!-- potential confusion like using the term "expected log(R)".   -->
+<!-- I also just don't like hats... but that's my problem. -->
 
@@
 
-Ole: I'd use tau (or something) for the variances here so that people don't have to look at the subscripts to differentiate between variances at different hierarchical levels.
+<!-- Ole: I'd use tau (or something) for the variances here so that people don't have to look at the subscripts to differentiate between variances at different hierarchical levels. -->
 

### 2016-09-27 portfolio/text/ms.Rmd
Reorder figures and results

@@
 
+### Fig. \ref{fig:across} - across strategies
+
+- clear exponential decline in standardized variability with strategies that are inherently more species diverse
+- not a clear effect of days fished at this level
+- fishing permits with higher species diversity reduces variability
+(`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction per unit
+increase in mean effective species diversity);
+
 ### Fig. \ref{fig:spokes} - add permits
@@
 - but can be very expensive (provide some examples from website)
-
-### Fig. \ref{fig:across} - across strategies
-
-- clear exponential decline in standardized variability with strategies that are inherently more species diverse
-- not a clear effect of days fished at this level
-- fishing permits with higher species diversity reduces variability
-(`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction per unit
-increase in mean effective species diversity);
 
@@
 
-- What's novel
+- Suggests it's mostly an across strategy/permit effect (lowering variability),
+  which isn't always a possibility because of regulations and cost
+- Adjusting within current permit constraints is risky... Not clearly risk
+  reducing to generalize from one year to the next
+- not as simple as increase diversity (within permit) and reduce risk
+- What's methodologically novel
     - Huge dataset
@@
     - look at different mechanisms of diversification
-- Suggests it's mostly an across strategy/permit effect (lowering variability),
-  which isn't always a possibility because of regulations and cost
-- Adjusting within current permit constraints is risky... Not clearly risk
-  reducing to generalize from one year to the next
-- not as simple as increase diversity (within permit) and reduce risk
 - 2 dimensions to think about (revenue and variability); and the trade-off is fishery specific
@@
 \begin{center}
-\includegraphics[width=0.9\textwidth]{../figs/stan-gg-spoke2.pdf}
-\caption{Diversifying permits reduces risk and usually increases revenue.
-Two examples of expected revenue variability
-and median revenue for
-owners of (A) individual halibut and (B) gillnet herring roe permits
-(dark blue)
-compared with owners of the same permit plus one more (medium blue)
-or two more permits (light blue).
-On the y-axis, dots and lines represent posterior median and 95\%
-credible intervals of expected standard deviation.}
-\label{fig:spokes}
-\end{center}
-\end{figure}
-
-\clearpage
-
-\begin{figure}[htbp]
-\begin{center}
 \includegraphics[width=0.7\textwidth]{../figs/stan-across-strategy-variability2.pdf}
@@
 \label{fig:across}
+\end{center}
+\end{figure}
+
+\clearpage
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=0.9\textwidth]{../figs/stan-gg-spoke2.pdf}
+\caption{Diversifying permits reduces risk and usually increases revenue.
+Two examples of expected revenue variability
+and median revenue for
+owners of (A) individual halibut and (B) gillnet herring roe permits
+(dark blue)
+compared with owners of the same permit plus one more (medium blue)
+or two more permits (light blue).
+On the y-axis, dots and lines represent posterior median and 95\%
+credible intervals of expected standard deviation.}
+\label{fig:spokes}
 \end{center}

### 2016-09-27 portfolio/text/ms.Rmd
Copy editing

@@
 
-The $\beta$'s represent estimated coefficients, some of which are allowed to
+The $\beta$ terms represent estimated coefficients, some of which are allowed to
 vary as 'random effect' terms.
@@
 \noindent
-The $\beta_{1,j}$'s represent how much an X\% decrease in species
+The $\beta_{1,j}$'s represent how much a X\% decrease in species
 diversity within a strategy translates to a Y\% change in revenue.

### 2016-09-27 portfolio/text/ms.Rmd
Major in progress reworking of introduction

@@
 resources for their livelihood is income variability.
-Fishers, loggers, farmers, miners, and others
-all face uncertainty in income driven by a mix of
+Fishers, loggers, farmers, and miners, among others,
+face income uncertainty driven by a mix of
 environmental, biological, and economic variability.
@@
 to unilaterally change fishing behaviour.
+
 Here we identify three primary axes by which fishers in Alaska,
@@
 (3) catch a greater species diversity within the constraints of a given permit.
+
 We show contrasting effects of diversity on
@@
 associated with increased uncertainty in revenue the following year.
+
 Thus, while there are clear benefits
@@
 was also the most costly.
-
 At the same time, Alaskan commercial fisheries
@@
 \noindent
+
+\clearpage
+
 Extra:
@@
 
-Natural resources such as forests, land animals, and fishes
-provide a primary source of income for millions of people worldwide [REFS].
-However, environmental,
-biological, and price variation
-combined with regulatory changes
-and human-caused environmental disasters
-generate high levels of income variability and risk
-for these individuals.
-For example, drought and flooding are major sources of risk
-to agricultural food security [@gregory2005],
-And EVOS example
+Managing natural resources and the 
+livelihoods of those that depend on their extraction is hard.
+Environmental, biological, and economic variation
+combine to generate high levels of income variability and risk.
+Drought and flooding is a major source of 
+risk to agricultural food security and farmers' incomes [@gregory2005].
+Bark beetle outbreaks can cause devastation to 
+the income of loggers.
+Price crashes in insert mineral here 
+are hard on miners' income variability ...
 
-\bigskip
+Individuals naturally develop strategies to reduce this income variation.
+For example, farmers rotate their crops to buffer short-term 
+climate variation and long-term environmental trends REF. 
+Loggers maintain a diverse mix of tree species, age structure, 
+geographical areas to buffer against 
+disease outbreak, wildfires, and price variation REF. 
+But, otherwise well-intentioned regulations
+can limit the basket of actions
+available to individuals to reduce income variability.
+For example, ...
+In other cases, external forces (prices, demand) may discourage the
+adoption of these diverse strategies and
+help promote specialization [@bradshaw2004].
 
-Increasing diversity has been proposed as a risk-buffering mechanism...
-Give examples from other sectors
+Fisheries provide a prime example...
+While the days of open, unregulated fisheries in most developed countries have
+passed --- fishers must own a license to participate in virtually all fisheries
+--- fishers retain several options for increasing diversity. 
+For example, fishers can
+fish multiple species as permitted by a license, 
+own multiple licenses,
+or fish in multiple regions.
 
-- Farming: crop rotation has been proposed as metric for buffering against short term
-  variation and long term trends (like climate change)
-- Logging: Maintaining a diverse mix of tree species, age structure, and space
-  to buffer against things like wildfires and price variation.
-- but in some situations, external forces (prices, demand) may discourage the
-  adoption of these diverse strategies and
-  help promote specialization [@bradshaw2004]
+But uncertainty remains about the value
+of adopting various diversification strategies for improving revenue and
+reducing revenue variability. 
+Furthermore, theory and empirical studies both
+suggest there can be substantial benefits to specialization (REFS).
 
-\bigskip
-
-This effect has been identified for fisheries.
-At community level, ... [@sethi2014]
-At vessel level, ... [@kasperski2013] ...
 However, these papers look at aggregates across individuals and largely ignore
 how existing fisheries regulations limit the ability of individual fishers to
-diversify. Key uncertainties remain about the fundamental relationship between
+diversify. 
+Key uncertainties remain about the fundamental relationship between
 diversity, fishing revenue, and fishing revenue variability at the level of the
 individual.
-While the days of open, unregulated fisheries in most developed countries have
-passed - fishers must own a license to participate in virtually all fisheries
-- fishers retain several options for increasing diversity. For example, fishers
-can do thing X, thing Y, or thing Z. But uncertainty remains about the value
-of adopting different diversification strategies for improving revenue and
-reducing revenue variability. Furthermore, theory and empirical studies both
-suggest there can be substantial benefits to specialization (REFS?)
+
+Diversity in the fishing strategies of individuals
+ranges from highly specialized (e.g. fishing only one species with one gear type
+or in one area) to highly diverse (e.g. fishing multiple species with multiple
+gear types in multiple areas).
+
+Fisheries in Alaska provide a unique opportunity 
+to understand the role of individuals' strategies 
+in determining income and income variability
+for those whose livelihood depends on the extraction of natural resources.
+Fisheries in Alaska are the highest-valued fisheries in North America
+with landings valued at XX $ billion annually representing harvests from YY
+species.
+The industry supports a range of fishing fleets ranging from small,
+open, single-operator vessels to multi-million-dollar state-of-the-art ships
+with large crews.
+
+However, the days of open, unregulated fisheries in most developed countries have
+passed --- fishers must own a license to participate in virtually all fisheries.
+And while such regulations have proven 
+helpful in reducing or eliminating overfishing,
+they do limit the available strategies to reduce income variability.
 
@@
 
-Fisheries in Alaska are the highest valued fisheries in N. America
-with landings valued at XX $ billion annually representing harvests from YY
-species.  Alaska fisheries support a diverse fishing fleets ranging from small,
-open, single operator vessels to multi-million dollar state of the art ships
-with large crews. Similar diversity in fishing strategies can be found within
-the Alaska fleet with different individuals adopting fishing strategies that
-range from highly specialized (e.g. fishing only one species with one gear type
-or in one area) to highly diverse (e.g. fishing multiple species with multiple
-gear types in multiple areas). This broad range of fishing types provide an
-unique opportunity for understanding the role diversity plays in determining
-both revenue and risk for individual fishers.
-
 We develop a novel statistical approach for understanding the effect of
@@
 First, we  summarize trends in diversification across fishing fleets in Alaska.
-Next, we identified three main axes by which fishers may diversify: 1), 2), 3).
-Finally, we constructed a model to simultaneously estimate the relationship
+Next, we identify three main axes by which individual fishers may diversify
+the species they target:
+(1) owning permits that allow targeting a diversity of fish species,
+(2) purchasing additional permits,
+or (3) catching a greater species diversity within the constraints of a given permit.
+
+Finally, we construct a model to simultaneously estimate the relationship
 between each of these three axes of diversity and both annual revenue and risk
@@
 in a hierarchical model.
+
+\clearpage
 

### 2016-09-27 portfolio/text/anderson-etal-rev-pfx.tex
Narrow down draft titles
Incorporate Jordan's suggestions

--- 2016-08-16 portfolio/text/anderson-etal-rev-pfx.tex
@@
 
-\title{Portfolios of fishing permits and species
-reduce risk but adjusting fishing strategy is risky}
+\title{Benefits of diversification and risks of changing strategies for individual fishers\\
+Benefits of diversification are unavailable to many fishermen\\
+}
 \author{
@@
 Jennifer C. Shriver,
-Jordan T. Watson
+Jordan T. Watson$^X$,
+\ldots
 }
+
+\newlabel{tab:strategies}{1}
+\newlabel{fig:fe}{1}
+\newlabel{fig:re}{2}
+\newlabel{fig:spoke-all}{3}
+\newlabel{fig:resid}{4}
+\newlabel{fig:resid-downside}{5}
+\newlabel{fig:str-yrs}{6}
 
@@
 \noindent
+$^X$Ted Stevens Marine Research Institute, Alaska Fisheries Science Center, 17109
+Pt Lena Loop Road, Juneau, AK 99801, USA
+
+\noindent
 Affiliations \ldots

### 2016-09-27 portfolio/text/ms.Rmd
Incorporate Jordan's suggestions

@@
 We calculated effort or effective season length as
-the sum the season lengths for each permit each year.
+the sum of the season lengths for each permit each year.
 Season length for a given permit
@@
 i.e., $\log(R_{i,t}) - \log(R_{i,t-1}) = \log(R_{i,t}/R_{i,t-1})$.
-We model change in revenue because preliminary analysis revealed non-stationary trends in
-revenue (resulting from experience or other unmodeled factors). This approach required filtering the data
-to remove person-year combinations without a subsequent year of revenue, but
-presents a number of advantages:
-it eliminates having to deal with the non-stationary
+We model change in revenue because preliminary analysis 
+revealed non-stationary trends in
+revenue (resulting from experience or other unmodeled factors). 
+This approach eliminates having to deal with the non-stationary
 time series and eliminates having to
-estimate thousands of intercepts associated individual fishers.
+estimate thousands of mean-revenue intercepts 
+associated with individual fishers.
+We removed person-year combinations without a subsequent year of revenue.
 
@@
 for a permit holder who does not change
-species diversity or days fished from year to year
-with the strategy-level effect of mean days fished removed.
+species diversity or days fished from year to year.
 }

### 2016-10-02 diversity/analysis/herring_case_study.Rmd
updated case study

@@
 
-herring = cfec[substr(cfec$p_fshy,1,1) %in% c("L","G"),]
+herring = cfec[substr(cfec$p_fshy,1,1) %in% c("L","G","H"),]
 
@@
   filter(maxpeople > 25)
+
+quitters = group_by(herring, p_holder, p_fshy) %>%
+  summarize(lastYear = max(year))
+
+group_by(quitters[quitters$lastYear<2014,],lastYear,p_fshy) %>% 
+  summarize(n = n()) %>% 
+  ggplot(aes(lastYear,n)) + geom_line() + facet_wrap(~p_fshy, scale="free_y") + xlab("Year") + ylab("People leaving the fishery (last year fished)")
+
+# look at what these people did after leaving
+cfec_quitters = cfec[which(cfec$p_holder%in%unique(quitters$p_holder)),]
+cfec_quitters = left_join(cfec_quitters, quitters[,c("p_holder","lastYear")])
+
+fates = as.data.frame(group_by(cfec_quitters, p_holder, p_fshy) %>% 
+  summarize(n = length(which(year > lastYear[1]))) %>% 
+  group_by(p_fshy) %>% 
+  summarize(n0 = length(which(n==0)), nOther = length(which(n>0))))
+
+trackQuitters = function(permit) {
+cfec2= cfec_quitters[which(cfec_quitters$p_holder %in% unique(quitters[quitters$p_fshy==permit,]$p_holder)),]
+cfec2 = cfec2[which(cfec2$year > cfec2$lastYear),]
+cfec2$SPEC = substr(cfec2$p_fshy,1,1)
+totalpermit = group_by(cfec2,SPEC) %>% 
+  summarize(g = sum(g_earn))
+totalpermit$g = totalpermit$g/sum(totalpermit$g)
+return(as.data.frame(arrange(totalpermit,-g)))
+
+}
+
+#trackQuitters("L 21C")
 ```

### 2016-10-05 salmon/archive/salmon_paper_combined.Rmd
salmon update

---
title: "Reduced diversity and increased specialzation of Alaska salmon fishermen: effects on variation in individual revenues"
author: ''
date: ''
output:
  pdf_document:
    fig_caption: yes
  html_document: default
  word_document: default
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=FALSE, tidy.opts=list(width.cutoff=60), warning = FALSE, message = FALSE)
library(dplyr)
library(knitr)
library(ggplot2)
library(date)
library(ggtern)
library(reshape2)
library(mgcv)
library(fpc)
library(gridExtra)
library(glmmTMB)
library(zoo)
```

## Abstract
\setlength\parindent{24pt}

On a large scale, individual fishers may experience positive outcomes from diversifying their catches; examples include reduced year-to-year variability in annual revenue, and increases in average revenue. Fishers can increase their species diversity by owning a permit that allows multiple species to be targeted, or by owning multiple permits. External forces such as management regulations (limited entry,  catch shares) or market forces (permit prices) may disincentive individuals to hold multiple permits. Thus, individuals with a single permit may increase species diversity within the confines of their permit (longer fishing seasons, broader spatial area). Using a long-term dataset of individual salmon fishers in the Alaska, we develop new models for understanding how the effects of diversification have changed through time. These changes allow us to compare patterns across regions and gear, as well as better understand how external forces have changed incentives for diversifying (hatchery production and price trends, EVOS?).

## Introduction

[broader paragraph on diversification with extensions to terrestrial agriculture etc, ]

Previous analyses have largely focused on patterns of diversification across ecosystems and the benefits to adopt more or less diverse fishing strategies (Kasperski and Holland 2013; Anderson et al. in review). Fisheries in Alaska have been studied extensively, because they are one of the most productive regions in North America (both by landed weight, and revenue cite) and they allow individuals to adopt a wide range of fishing strategies. Anderson et al. (in review) defined strategies as a permit of collection of permits held by single individuals. While adding new permits allows fishers to increase species diveristy, few individual permits allow individuals to harvest more than one species (examples include halibut, sablefish, herring, crab, shellfish, shrimp, sea cucumber). Exceptions to these single species permits incldue participants in federally managed groundfish fisheries (CFEC 'M' permits) and participants in state managed salmon fisheries (CFEC 'S' permits).

Understanding the effects of species diversification on salmon fishers is particularly important, as the number of salmon specialists (defined as fishers holding a single salmon permit) has increased over the last 30 years as a fraction of the total (Anderson et al. in review). For these limited entry salmon fisheries, general trends that have emerged are that more people are participating since the early 2000s (Fig. A1), median revenue is increasing (across individuals, Fig. A2), and landed species diversity declining over time (averaged across indivduals, Fig. A6). Across different salmon fisheries, other than a slight negative effect of diversification on variability (similar to Anderson et al. in review), there appears to be little correlation between diversity and revenue (Fig. A4). In terrestrial agriculture, there are numerous examples where diversification may affect the mean or CV but not both (Purdy, Barry M; Langemeier, Michael R; Featherstone, Allen M. Journal of Agricultural and Applied Economics29.1 (Jul 1, 1997): 149.), and the same result may be true for salmon fishers. Further, these impacts of diversification may be variable over time.

Five species of salmon are encountered in Alaska waters (Chinook, chum, pink, sockeye, coho), and the ability of individual salmon fishers to diversify across species is in part a function of geography and gears. For example, salmon may be caught with purse seine gear (n = 6 areas), drift gillnets (5 areas), and set gillnets (12 areas). Other gears include beach seine and troll gears (though these are generally restricted to Kodiak and Southeast Alaska, respectively). While some of these combinations of gears and areas allow fishers to diversify, there are many examples of permits where the catch is dominated by a single species. Perhaps the best examples include fisheries that target sockeye salmon (e.g. Bristol Bay set and drift gillnet fisheries). For these fisheries, specialization is associated with higher revenue and reduced CV in annual revenue (Anderson et al. in review).

A secondary motivation for focusing on the effects of diversification for salmon fishers is to better understand the long term consequences of hatchery salmon production on the revenue and variability in revenue of individuals. Following low returns and declining productivity in the mid 1980s, many hatchery programs in the state rapidly increased hatchery releases (Mahnken et al. 1998). Many of these hatchery programs increased production to stabilize the variability of fishers catch (and revenue), though these programs may also have costs to wild production (Ruggerone et al. 2015). [Other history here?]

In addition to the changing species composition of fish avaialble to be harvested, the benefits (or costs) of diversification are likely affected by a number of exogenous drivers, including marketing, prices, new processing technology, etc (cite Gunnar Knapp + others). [Break out as own paragraph+add more examples?] Following price crashes in the early 2000s, many of these factors have been positive for salmon fishermen (increasing trends in prices and catches, Fig. A3, A12). In contrast, the late 1980s - early 1990s experienced high abundance and low prices (in part due to collapses of foreign currencies). Combined with the increased trend in hatchery production and resulting change species compositon of adult salmon, these external factors may have also altered incentives for diversification. If there's an incentive to substitute high value species (Chinook, sockeye) with abundant low value species (pink, chum), the analog in the terrrestrial framring world would be crop substitution of low-volume high value crops (spices) with high-volume low value crops (cereals) (other examples in literature = cocoa, tobacco).

The objective of our analysis is to better understand the effects of diversification on single-permit salmon holders by extending the methods used by Anderson et al. (in review) to include time-varying effects. These time-varying effects will allow us to compare temporal trends across regions, identify periods when having a diverse catch portfolio was most beneficial, and allow us to evaluate support for effects of longer term drivers in the region.

## Methods

### Data

We obtained fisheries landings and revenue data for all permit holders in Alaska from 1985 - 2014 from the Alaskan Commercial Fisheries Entry Commission (CFEC). We adjusted for inflation by setting all revenues in 2009 USD. Because were were only interested in single permit fisheries, we adopted a simplified version of the filtering steps described in  Anderson et al. (in review). After aggregating individual fish tickets to annual data, we retained individuals who only fished one of the six salmon permits included in a year (Table 1, Fig. A5). Like Anderson et al. (in review) we also worked with a log-differenced dataset to remove non-stationary individuals; to avoid complications of individuals switching strategies, we only included pairs of years for a given person where they did not change strategies.

```{r, echo=FALSE}
m = matrix("", 6, 3)
m[1,] = c("S01A", "Purse Seine", "Southeast Alaska")
m[2,] = c("S01E", "Purse Seine", "Prince William Sound")
m[3,] = c("S01K", "Purse Seine", "Kodiak")
m[4,] = c("S01M", "Purse Seine", "Alaska Peninsula")
m[5,] = c("S03A", "Gillnet", "Southeast Alaska")
m[6,] = c("S03E", "Gillnet", "Prince William Sound")
colnames(m) = c("Permit", "Gear", "Region")
kable(m)
```

### As hatchery programs have increased production, have individual salmon fishermen experienced reduced variability?

Hatchery release trends for some regions included in our data have been summarized previously (e.g. Ward et al. in review summarized the trends for Prince William Sound). Release patterns have been dominated by pink and chum salmon (for PWS, these species represent 80% and 18% of total releases for 2012-2014 respectively, Ward et al. in review) and production has increased since the late 1970s (217 million juvenile pink salmon in PWS in 1985 versus 672 million in 2014). Though our analysis doesn't include , as a contrast we can do coarse comparisons across regions and species to examine potential effects of hatchery fish on variability. For example, we can compare the variation in pink salmon returns (earnings or weight) from Prince William Sound (largely hatchery) to those from Southeast Alaska (largely wild). Similarly, we can compare the variability in earnings from species dominated by hatchery returns (pink, chum) to those with less hatchery influence (Chinook, sockeye). To make these comparisons, we aggregated the earnings and landings by species and permit, then calculated 3-year rolling CVs. As a second comparison, we examined whether long-term participants in any of the fisheries had experienced a change in variability. Similar to Kasperski and Holland (2013) who calculated CVs of vessels before and after significant managment events, we identified individuals who fished the same permit for at least three years in the first decade of our data (1985-1995) and at least three years in the last decade of the data (2004-2014).

### Robustness of strategies to changing environments [stash in appendix?]

One of the challenges in modeling the effects of species diversification for individual fishers is that between two years, both the species composition of the catch and the species composition of the fish available to be caught may change (making it hard to disentangle the selectivity of fishermen versus portfolio of fish available). To isolate the effects of a changing environment from changing strategies, we used a subset of the data from fishery permits where discrete strategies were emergent.

Using purse seine permit holders in Prince William Sound as an example, we first calculated the species composition (weighted by revenue) for each active permit holder in each year. For each pair of years that an individual fished, we calculated the squared Euclidian distance between the species compositions of those years, $$\Delta { p }_{ ind,t }=\sum _{ sp=1 }^{ 5 }{ { \left| { p }_{ ind,sp,t }-{ p }_{ ind,sp,t-1 } \right|  }^{ 2 } } $$.

Larger values of $\Delta { p }_{ ind,t }$ indicate greater year-to-year discrepancy in species composition. To focus on individuals who had more constant strategies in adjacent years, we ommitted data corresponding to the upper third of this distribution (retaining $\Delta { p }_{ ind,t }$ < 0.334). Next, we used the species composition from each individual and pair of years to cluster individuals into strategies. Clustering was performed using the partitioning around medoids (pam) algorithm in R with an estimated number of clusters (using 'pamk' in fpc library), and selected 2 groups of individual strategies: pink salmon specialization (medoid proportion of pink salmon = 0.95), and a strategy focused on pink salmon and chum (medoid proportions of 0.64 and 0.31 respectively). For each group, we then calculated the mean and CV of revenue; we calculated the log difference of revenue for each individual, $\Delta { R }_{ i }=log\left( { R }_{ i,t+1 }/{ R }_{ i,t } \right)$ and then calculated the CV as the standard deviation over people, $sd(\Delta { R }_{ i=1:n })$.

### Analysis 3: [Core of paper] Does effect of diversification change over time?

Anderson et al. (in review) implemented a hierarchical Bayesian model that allowed group level predictors and covariates (diversity of catch, days fished) to simultaneously affect the mean and the variance of revenue for individual fishers. This model also included random intercepts by strategy and year; these terms allowed all permit holders in a given year to experience common fluctuations (external drivers including price changes, fluctuations in fish population abundance, or environmental disasters such as EVOS).

To ask whether how benefits of diversification may have changed over time, we extend the model of Anderson et al. (in review) to also include time-varying random effects in the slope coefficients of diversification. We define the variables $\Delta { S }_{ t }=log\left( { S }_{ t }/{ S }_{ t-1 } \right)$ to represent the percent change in species diversity of the catch between years and $\Delta { D }_{ t }=log\left( { D }_{ t }/{ D }_{ t-1 } \right)$ to represent the percent change in fishing effort (days fished) between years. The model for the the expected log revenue of individual *i* in year *t*, $log({ R }_{ i,t })$, can be written as $$\widehat { log\left( { R }_{ i,t } \right)  } ={ B }_{ 0,j,t }+{ B }_{ 1,j,t }\cdot \Delta S_{ t }+{ B }_{ 2,j,t }\cdot \left( \Delta S_{ t }\cdot { S }_{ t-1 } \right) +{ B }_{ 3,j,t }^{ 2 }\cdot \left( \Delta S_{ t }\cdot { S }_{ t-1 }^{ 2 } \right) +{ B }_{ 4,j }\cdot \Delta D_{ t }+log\left( { R }_{ i,t-1 } \right) $$, where *j* subscripts the salmon permit (Table 1). We model the time varying permit specific intercepts as random and exchangeable across years, ${ B }_{ 0,j,t }\sim Normal\left( { u }_{ { B }_{ 0 } },{ \sigma  }_{ { B }_{ 0 } } \right)$. The term ${ B }_{ 1,j,t }$ represents the permit level random effect of diversifying species. Our model extendeds that of Anderson et al. (in review) to include the interaction terms $\Delta S_{ t }\cdot { S }_{ t-1 }$ which allow the effect of diversification to be dependent on the previous level of diversification. The squared term allows this interaction to be quadratic, and the coefficients ${ B }_{ 2,j,t }$ and ${ B }_{ 3,j,t }$, allow this to vary by permit and through time. The flexibility in these terms allows the benefit of increasing species diversity to be positive in some years (generalizing benefits revenue), negative in other years (specializing benefits revenue) and neutral in other years (revenue is independent alterations to catch composition). We constrain the random effects associated with diversification to be modeled as autoregressive processes with shared variances across strategies, for example: ${ B }_{ 1,j,t }\sim Normal\left( { B }_{ 1,j,t-1 },{ \sigma  }_{ { B }_{ 1 } } \right)$. Like Anderson et al. (in review) we also allow changing effort (measured as percent change in days fished) ${ B }_{ 4,j }$ to vary by strategy, but we do not allow this effect to vary through time. Finally, the term $log\left( { R }_{ i,t-1 } \right)$ is included as an offset (Anderson et al. in review).

- Fit models from Sean's paper, but include time varying interactions on species diversity.  GLMM code below

Specific things that could be looked at: 
- EVOS (closure for Kodiak)  
- PWS pink salmon crash in 1992/1993: Pacific Salmon & their Ecosystems: Status and Future Options

Does diversification exist within permits and if so, do we see similar advantages of diversification on revenue + CV?

## Results

Our summaries of changing variability (aggregate for each fishery, across individuals, or within individuals) was not consistent with the hypothesis that variability in revenue or landings has decreased over time (Fig. A9-A10). For pink and chum salmon, fisheries in PWS generally experience higher year to year variability compared to SEAK. In looking at the changing CV of individuals, we see a slight increase in the CV of revenue for people fishing in the recent decade who were also fishing 1985-1995 (with the exception being S01A).

## Discussion

- Revenue is going up since 2000 -- but each permit holder is fishing more (more days, more total pounds) and the trend in weight is faster than the trend in price.

- Ability to specialize primarily driven by many forces beyond control of fishermen. But in some situations, diversification is possible by extending fishing season to exploit additional spp (fishing earlier for chum in PWS S01E). Or fishing different areas to exploit more spp in S03 (SEAK).

- Prices are generally synchronous for all spp (especially coho/chum/pink), and have increased at a faster rate compared to chum and pink salmons since 2000 (Fig A8-A9). We're also seeing a trend toward greater synchrony between species (Fig A11-A12, A14). Projecting future prices are complex -- Gunnar Knapp has a good presentation on this http://www.adfg.alaska.gov/static/fishing/PDFs/commercial/gk_trends_4-12-12.pdf. Synchrony in prices may largely be function of marketing, improved product, freezing, etc.

- We're seeing increased specialization within permits. People in PWS are trending toward increased specialization on pink salmon. People in SEAK trending toward increased specialization on chum. Both of these are switches from higher value species (sockeye, chum) to lower value species. Both pink and chum have higher variability in year to year returns than sockeye.

- One pontential exogenous driver related to management that can be ruled out as being responsible for increased specialization of salmon fishermen is changes in the timing or length of fishing seasons. Shifting seasons to later in the year may reduce catch of early run fish for example. However, we found no significant trend in either the season length or average season date for any fishery (Fig. A15).

\break

```{r, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
revenueThreshold = 10000

dat.annual = readRDS(file="../../portfolio/data-generated/cfec-annual-for-modeling.rds")

# single permit holders, salmon permits
dat.annual  = filter(dat.annual ,npermit==1 & substr(strategy_permit,1,1)=="S") %>%
  group_by(strategy_permit) %>% 
  mutate(nPeople = length(unique(p_holder))) %>% 
  filter(nPeople > 200) %>% 
  dplyr::select(-nPeople)
# drop troll
dat.annual = dat.annual[dat.annual$strategy_permit%in%c("S05B","S15B")==FALSE,]
# drop S04X, Y, Z because low participation (< 20) in many years
# S04W deleted because low median earnings
dat.annual = dat.annual[dat.annual$strategy_permit%in%c("S04W", "S04X","S04Y","S04Z")==FALSE,]

agg.annual = group_by(dat.annual, strategy_permit, year) %>% 
  summarize(people = length(unique(p_holder)),
    specDiv = mean(specDiv),
    meanrev = median(revenue),
    low = quantile(revenue,0.25),
    upp = quantile(revenue,0.75))

dat.diff = readRDS(file="../../portfolio/data-generated/cfec-diff-for-modeling.rds")
dat.diff = dat.diff[which(dat.diff$revenue >= revenueThreshold & dat.diff$revenue.prev >= revenueThreshold), ]
dat.diff = dat.diff[dat.diff$strategy_permit%in%c("S01A","S01E","S01K","S01L","S01M",
  "S03A","S03E","S03H","S03M","S03T","S04D","S04H","S04K","S04M","S04T"),]

agg.diff = group_by(dat.diff, strategy_permit) %>% 
  summarize(specDiv = mean(specDiv), meanrev = mean(revenue),
    cv = sd(log(revenue/revenue.prev)))

# Also load in raw CFEC data for plots
cfec <- feather::read_feather("../../portfolio/data-generated/cfec.feather")
cfec = cfec[cfec$p_fshy%in%c("S 01A", "S 01E", "S 01K", "S 03A", "S 03E", "S 01M"),]
cfec$year = as.numeric(substr(cfec$landdate, 1, 4))

# only use people who are using just that permit - diversification within permit
cfec.raw = cfec
dat.salmon = dat.diff[dat.diff$strategy_permit%in%c("S01A", "S01E", "S01K", "S03A", "S03E","S01M"),]

# Restrict analysis to people who keep same strategy - complicates days, etc
dat.salmon = dat.salmon[which(dat.salmon$strategy_permit==dat.salmon$strategy_permit.prev),]

dat.salmon$days.change = log((dat.salmon$days_permit+1) / (dat.salmon$days_permit.prev+1))
dat.salmon$year_permit = as.factor(paste(dat.salmon$year, dat.salmon$strategy_permit))

cfec = cfec[which(paste(cfec$p_holder,cfec$year) %in% paste(dat.salmon$p_holder,dat.salmon$year)),]

# Make come plots for the appendix of (1) median fishing dates over time, (2) length of fishing season over time (calculated as width of 95% intervals)
cfec.raw$month = as.numeric(substr(cfec.raw$day, 1,2))
cfec.raw$day = as.numeric(substr(cfec.raw$day, 4,5))
cfec.raw$jday = mdy.date(cfec.raw$month, cfec.raw$day, cfec.raw$year) - mdy.date(1, 1, cfec.raw$year)
# summarize the temporal distribution of the catches
seasons = group_by(cfec.raw, p_fshy, year) %>% 
  summarize(q = mean(sample(jday, size = 10000, replace=T, prob=g_earn)),
    lower = quantile(sample(jday, size = 10000, replace=T, prob=g_earn),0.025),
    upper = quantile(sample(jday, size = 10000, replace=T, prob=g_earn),0.975),
    length = upper - lower)
seasons = rename(seasons, permit = p_fshy)
appendixPlot0_p1 = ggplot(seasons, aes(year, q, color = permit, group=permit)) + geom_line() + xlab("Year") + ylab("Mean julian day")
appendixPlot0_p2 = ggplot(seasons, aes(year, length, color = permit, group=permit)) + geom_line() + xlab("Year") + ylab("Season length")
```

```{r, echo = FALSE, eval=FALSE}
dat = dat.salmon
# create other category, lumping coho/chinook
dat$salm = dat$coho+dat$chnk
dat$salm.prev = dat$coho.prev+dat$chnk.prev

dat$totsalm = dat$salm + dat$chum + dat$sock + dat$pink
dat$totsalm.prev = dat$salm.prev + dat$chum.prev + dat$sock.prev + dat$pink.prev

dat$salm = dat$salm/dat$totsalm
dat$sock = dat$sock/dat$totsalm
dat$pink = dat$pink/dat$totsalm
dat$chum = dat$chum/dat$totsalm
dat$chnk = dat$chnk/dat$totsalm
dat$coho = dat$coho/dat$totsalm
dat$salm.prev = dat$salm.prev/dat$totsalm.prev
dat$sock.prev = dat$sock.prev/dat$totsalm.prev
dat$pink.prev = dat$pink.prev/dat$totsalm.prev
dat$chum.prev = dat$chum.prev/dat$totsalm.prev
dat$chnk.prev = dat$chnk.prev/dat$totsalm.prev
dat$coho.prev = dat$coho.prev/dat$totsalm.prev

dat = dat[-which(is.na(dat$salm+dat$sock+dat$pink+dat$chum)),]
dat$logdiff = log(dat$revenue/dat$revenue.prev)

dat_old = dat
dat$sumP2 = abs(dat$salm - dat$salm.prev) + 
  abs(dat$pink-dat$pink.prev) + 
  abs(dat$chum-dat$chum.prev) + 
  abs(dat$sock-dat$sock.prev)
dat = dat[which(dat$sumP2 < 0.3),]
```

```{r, echo = FALSE, eval=FALSE, fig.cap="Results for PWS Purse seine fishery", fig.pos="placeHere", cache=TRUE}

df = dat[which(dat$strategy_permit=="S01E"),]
pamk_s01e <- pamk(df[,c("chum","pink","sock","salm")])
df$group = c("Pink-chum","Pink")[pamk_s01e$pamobject$clustering]

p3 = ggtern(data=df[,c("chum","pink","sock","group")], aes(x=chum,y=pink,sock,color=as.factor(group))) +
  geom_point(alpha=0.3) +
  labs(x="CHUM",y="PINK",z="SOCK") + theme(
        axis.title=element_text(size=8)) + theme(legend.position = "none")

p0 = group_by(df, year, group) %>% 
  summarize(n = length(unique(p_holder))) %>% 
ggplot(aes(x = year, y = n, group=group, col = group)) + geom_line() + ylab("People") + ggtitle("S01E")

p1 = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T)) %>% 
ggplot(aes(x = year, y = meanrev, group=group, col = group)) + geom_line() + ylab("Mean revenue") + ggtitle("S01E")

p2 = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T)) %>% 
ggplot(aes(x = year, y = cv, group=group, col = group)) + geom_line() + ylab("CV revenue") + ggtitle("S01E")

g = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T))

gridExtra::grid.arrange(p3,p0,p1,p2,ncol=2)
```

\break

```{r makePlots, echo = FALSE}
appendixPlot1 = ggplot(agg.annual, aes(year, people, colour = specDiv)) + geom_line() + facet_wrap(~ strategy_permit, scale="free") + ylab("Participation") + xlab("Year")

appendixPlot2_p1 = ggplot(agg.diff, aes(specDiv, log(meanrev), colour = cv, label=strategy_permit)) + xlab("Species diversity") +ylab("Ln (mean revenue)") + geom_text(size=2)
appendixPlot2_p2 = ggplot(agg.diff, aes(specDiv, cv, colour = log(meanrev), label=strategy_permit)) + xlab("Species diversity") +ylab("CV") + geom_text(size=2)
appendixPlot2_p3 = ggplot(agg.diff, aes(log(meanrev), cv, colour = specDiv, label=strategy_permit)) + ylab("CV") + xlab("Ln (mean revenue)") + geom_text(size=2)
```

\break

```{r, fig.cap="Species composition for 1988 and 2014 (each dot is a person) in the S03E fishery", fig.pos="placeHere", echo=FALSE, fig.height = 2, eval=FALSE}
### One way to visualize the change is with ternary plots over the 3 axes (chum / sockeye / pink)

daty = group_by(cfec[cfec$p_fshy=="S 03E" & cfec$spec %in% c("CHUM","PINK","SOCK","COHO","CHNK"),], spec, p_holder, year) %>%
  summarize(p = sum(g_earn)) %>%
  group_by(p_holder, year) %>% mutate(p = p/sum(p))
daty <- melt(daty, id.vars = c("spec", "p_holder", "year"))
daty <- dcast(daty, p_holder + year ~ spec + variable)[,-1]
daty[is.na(daty)]=0
ggtern(data=daty[which(daty$year%in%c(1988,2014)),c("year", "CHUM_p","PINK_p","SOCK_p")], aes(x=CHUM_p,y=PINK_p,SOCK_p,color=as.factor(year))) +
  geom_point(alpha=0.3,size=1) +
  labs(x="CHUM",y="PINK",z="SOCK",title="Proportion of annual revenue (S03E)")
```

\break

```{r, echo=FALSE, eval=FALSE}

mu.mod = glmmTMB(log(revenue) ~ specDiv.change + specDiv.change:specdiv.prev + specDiv.change:I(specdiv.prev^2) + days.change + (1+days.change|strategy_permit) + (-1+specDiv.change + specDiv.change:specdiv.prev + specDiv.change:I(specdiv.prev^2)|year_permit), offset = log(revenue.prev), dat.salmon)

cv.mod = glmmTMB(log(abs(residuals(mu.mod))) ~ specDiv.change + specDiv.change:specdiv.prev + specDiv.change:I(specdiv.prev^2) + days.change + (1+days.change|strategy_permit) + (-1+specDiv.change + specDiv.change:specdiv.prev + specDiv.change:I(specdiv.prev^2)|year_permit), dat.salmon)
# to ask whether benefit of diversificaiton is changing via strategy, we want to summarize (1) direction of 
# benefit (e.g. is the effect negative or positive), and (2) the magnitude (how steep / shallow it is). With the
# quadratic model, 
# 'squared term' indicates concave up/down
# derivative is magnitude, 2a * x + b

df = ranef(mu.mod)$cond$year_permit
colnames(df) = c("c","b","a")
df$year = as.numeric(substr(rownames(df), 1, 4))
df$strategy = substr(rownames(df), 6, 9)
df$deriv = 2*df$a + df$b
# also calculate change of going from 1 to 2 spp
df$pred1 = df$a*(1^2) + df$b*1 + df$c + fixef(mu.mod)$cond["specDiv.change"] + fixef(mu.mod)$cond["specDiv.change:specdiv.prev"]*1
# add fixed effects into predictions

# We can demonstrate the quadratic curves by year for a single strategy / set of years
df.s01e = df[df$strategy=="S01E",]
demo = expand.grid("speciesDiversity" = seq(1,4,length.out=100), "year" = 1986:2014)
demo = left_join(demo, df.s01e)
demo$effect = demo$a * (demo$speciesDiversity^2) + demo$b * (demo$speciesDiversity) + demo$c
```

```{r, echo = FALSE, warning=FALSE, message=FALSE, results='hide', eval=FALSE}
pdf("../../salmon/Salmon_residuals.pdf")
# Residuals for mean model
dat.salmon$mu.resid = residuals(mu.mod)
dat.salmon$mu.fitted = fitted(mu.mod)

ggplot(dat.salmon, aes(mu.fitted, mu.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Fitted") + ylab("Residuals") + ggtitle("Mean model")
ggplot(dat.salmon, aes(specDiv.change, mu.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Change in species diversification") + ylab("Residuals") + ggtitle("Mean model") + geom_hline(aes(yintercept=0),color="salmon1")
ggplot(dat.salmon, aes(specdiv.prev, mu.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Previous species diversity") + ylab("Residuals") + ggtitle("Mean model") + geom_hline(aes(yintercept=0),color="salmon1")
ggplot(dat.salmon, aes(log((revenue+revenue.prev)/2), mu.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Average revenue") + ylab("Residuals") + ggtitle("Mean model") + geom_hline(aes(yintercept=0),color="salmon1")
ggplot(dat.salmon, aes(specDiv.change, log(abs(mu.resid)), group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Change in species diversification") + ylab("Ln abs(residuals)") + ggtitle("Mean model")
ggplot(dat.salmon, aes(specdiv.prev, log(abs(mu.resid)), group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Previous species diversity") + ylab("Ln abs(residuals)") + ggtitle("Mean model")
ggplot(dat.salmon, aes(specdiv.prev, log(abs(mu.resid)), group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Previous species diversity") + ylab("Ln abs(residuals)") + ggtitle("Mean model")

# Residuals for CV model
dat.salmon$cv.resid = residuals(cv.mod)
dat.salmon$cv.fitted = fitted(cv.mod)

ggplot(dat.salmon, aes(cv.fitted, cv.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Fitted") + ylab("Residuals") + ggtitle("CV model")
ggplot(dat.salmon, aes(specDiv.change, cv.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Change in species diversification") + ylab("Residuals") + ggtitle("CV model") + geom_hline(aes(yintercept=0),color="salmon1")
ggplot(dat.salmon, aes(specdiv.prev, cv.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Previous species diversity") + ylab("Residuals") + ggtitle("CV model") + geom_hline(aes(yintercept=0),color="salmon1")
ggplot(dat.salmon, aes(log((revenue+revenue.prev)/2), cv.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Average revenue") + ylab("Residuals") + ggtitle("CV model") + geom_hline(aes(yintercept=0),color="salmon1")
dev.off()
```
```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Variation in effects of diversification by year for the S01E (PWS purse seine) permit holders. ", eval=FALSE}
###We can make quadratic plots of these effects by year and strategy. These plots represent the effect of increasing diversity (as relative change) on mean revenue. For example, in 1986 if you were fishing 3-4 species there would be little effect of adding more spp; but if you were fishing 1 species, there would be a slight negative effect of adding more. In 1991, there would be a huge benefit of increasing diversity, and that effect would taper off as more were added.

ggplot(demo, aes(speciesDiversity, effect, group = year)) + geom_line() + facet_wrap(~ year) + xlab("Species Diversity") + ylab("Effect of increasing diversity (relative change)")
```

\break

```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Time series of benefits of diversification by permit / strategy.", fig.height = 3, eval=FALSE}
###This plot shows the time series of becoming increasingly diversified from 1 species on mean revenue. For S01E for example, there's huge benefits of diversifying in 1991, and huge costs in 2010. 2010 was a huge pink salmon year for PWS (with high prices specialization was good) and 1991 was the start of pink salmon stock + price collapse. Across permits, perhaps slight declines in some of these in recent years?

ggplot(df, aes(year, pred1, group = strategy, colour = as.factor(strategy))) + geom_line() + xlab("Year") + ylab("Benefit of increasing species diversity from 1 species") + ggtitle("Effect on Mean")
```

```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Time series of benefits of diversification split out by permit / strategy.", fig.height = 3, eval=FALSE}
ggplot(df, aes(year, pred1, group = strategy, colour = as.factor(strategy))) + geom_line() + xlab("Year") + ylab("Benefit of increasing species diversity from 1 species") + facet_wrap(~ strategy, scale="free") + geom_smooth() + geom_hline(aes(yintercept=0))
```

```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Time series of effects of diversification on mean and CV by permit.", eval=FALSE}
df.cv = ranef(cv.mod)$cond$year_permit
colnames(df.cv) = c("c.cv","b.cv","a.cv")
df.cv$year = as.numeric(substr(rownames(df.cv), 1, 4))
df.cv$strategy = substr(rownames(df.cv), 6, 9)
# also calculate change of going from 1 to 2 spp
df.cv$pred1.cv = df.cv$a.cv*(1^2) + df.cv$b.cv*1 + df.cv$c.cv + fixef(cv.mod)$cond["specDiv.change"] + fixef(cv.mod)$cond["specDiv.change:specdiv.prev"]*1

ggplot(df.cv, aes(year, pred1.cv, group = strategy, colour = as.factor(strategy))) + geom_line(linetype="dashed") + geom_line(data=df, aes(year, pred1, group = strategy, colour = as.factor(strategy))) + xlab("Year") + ylab("Effect of increasing species diversity from 1 species") + facet_wrap(~ strategy, scale="free") + ggtitle("Effects on mean (solid) and CV (dashed)") + geom_hline(aes(yintercept=0), color="grey30")

```

\break

```{r, echo=FALSE, eval = FALSE}
# These are the run  timing plots per Milo's questions

# for 2008, create cumulative density plots of salmon landings by calday
pinkchum = cfec.raw[cfec.raw$p_fshy=="S 01E" & cfec.raw$spec%in%c("PINK","CHUM","SOCK"),]
#pinkchum = cfec.raw[cfec.raw$p_fshy=="S 03T" & cfec.raw$spec%in%c("PINK","SOCK","CHNK"),]

# aggregate by calendar day
pinkchum$month = as.numeric(substr(pinkchum$day, 1, 2))
pinkchum$day = as.numeric(substr(pinkchum$day, 4, 5))
pinkchum$jday = date::mdy.date(pinkchum$month, pinkchum$day, pinkchum$year) - date::mdy.date(1,1,pinkchum$year)

group_by(pinkchum, jday, year) %>% 
  summarize(pink = sum(g_earn[spec=="PINK"]), 
    chum = sum(g_earn[spec=="CHUM"]),
    sock = sum(g_earn[spec=="SOCK"])) %>% 
  group_by(year) %>% 
  mutate(tot = sum(pink) + sum(chum) + sum(sock), pink = pink/tot, chum = chum/tot, sock = sock/tot) %>% 
ggplot(aes(jday, pink)) + geom_line(color="pink") + geom_line(aes(jday, chum), color="darkgreen") + geom_line(aes(jday, sock), color="lightblue") + facet_wrap(~ year, scale="free")

# make plot of all earnings or weight by jday
salm = cfec.raw[cfec.raw$p_fshy=="S 01E",]
salm$month = as.numeric(substr(salm$day, 1, 2))
salm$day = as.numeric(substr(salm$day, 4, 5))
salm$jday = date::mdy.date(salm$month, salm$day, salm$year) - date::mdy.date(1,1,salm$year)

g = group_by(salm, jday, year) %>% 
  summarize(earn = sum(g_earn), 
    lbs = sum(g_pounds), 
    pink = sum(g_earn[spec=="PINK"])/earn,
    sock = sum(g_earn[spec=="SOCK"])/earn,
    chum = sum(g_earn[spec=="CHUM"])/earn,
    coho = sum(g_earn[spec=="COHO"])/earn,
    chnk = sum(g_earn[spec=="CHNK"])/earn) %>% 
ggplot(g, aes(jday, (earn))) + geom_line(colour = rgb(g$sock+g$coho+g$chnk, g$chum, g$pink)) + facet_wrap(~ year, scale="free_y")

```

```{r, echo = FALSE, fig.cap="SEAK purse seine fishery", cache=TRUE, fig.pos="placeHere", eval=FALSE}

df = dat[which(dat$strategy_permit=="S01A"),]
pamk_s01a <- pamk(df[,c("chum","pink","sock","salm")])
df$group = pamk_s01a$pamobject$clustering

p3 = ggtern(data=df[,c("chum","pink","sock","group")], aes(x=chum,y=pink,sock,color=as.factor(group))) +
  geom_point(alpha=0.3) +
  labs(x="CHUM",y="PINK",z="SOCK") + theme(
        axis.title=element_text(size=8)) + theme(legend.position = "none")

p0 = group_by(df, year, group) %>% 
  summarize(n = length(unique(p_holder))) %>% 
ggplot(aes(x = year, y = n, group=group, col = as.factor(group))) + geom_line() + ylab("People") + ggtitle("S01A")

p1 = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T)) %>% 
ggplot(aes(x = year, y = meanrev, group=group, col = as.factor(group))) + geom_line() + ylab("Mean revenue") + ggtitle("S01A")

p2 = group_by(df, year, group) %>% 
  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
    cv = sd(logdiff,na.rm=T)) %>% 
ggplot(aes(x = year, y = cv, group=group, col = as.factor(group))) + geom_line() + ylab("CV revenue") + ggtitle("S01A")

gridExtra::grid.arrange(p3, p0, p1, p2, ncol=2)
```

\pagebreak

# Appendix

## Overall trends in participation across permits

```{r, fig.cap="Participation (unique people)", fig.pos="placeHere", echo=FALSE}
appendixPlot1
```

\break

```{r, fig.cap="Trends in revenue and variability across people", fig.pos="placeHere", echo=FALSE, fig.height = 3.1}
group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], p_holder, year, p_fshy) %>%
  summarize(gearn = sum(g_earn)) %>%
  group_by(year, p_fshy) %>%
  summarize(med = median(gearn), low = quantile(gearn, 0.25),
    upp = quantile(gearn,0.75)) %>%
  ggplot(aes(year, log(med))) + geom_line() + facet_wrap(~p_fshy,scale="free") + 
  geom_ribbon(aes(ymin= log(low), ymax= log(upp)), alpha=0.4) + ylab("Ln quartiles of revenue (across people)") + ggtitle("Trends in revenue") + xlab("Year")
```

```{r, echo=FALSE, fig.height = 3.1, fig.cap="Trends in average landed weight per person", fig.pos="placeHere"}
group_by(cfec, p_fshy, year, p_holder) %>% 
  summarize(meanW = mean(g_pounds/2000, na.rm=T)) %>% 
  group_by(p_fshy, year) %>% 
  summarize(meanW = mean(meanW, na.rm=T)) %>% 
  ggplot(aes(year, meanW)) + geom_line() + facet_wrap(~ p_fshy, scale="free") + ylab("Mean total tons / person") + xlab("Year") + ggtitle("Trends in landed weight")
```

\break

```{r, fig.cap="Diversity, mean revenue and CV (aggregated across years)", fig.pos="placeHere", echo=FALSE}
gridExtra::grid.arrange(appendixPlot2_p1,appendixPlot2_p2,appendixPlot2_p3, ncol=2)
```

\break

## Species composition

```{r, fig.cap="Proportion of species landed, weighted by earnings", fig.pos="placeHere", echo=FALSE, fig.height = 3.3}
# plot proportion of species landed by salmon permit
group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p)) %>%
  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() + facet_wrap(~p_fshy) + 
  ylab("Species % (total) ") + xlab("")
```

```{r, fig.cap="Effective species diversity across people", fig.pos="placeHere", echo=FALSE, fig.height = 3.3}
group_by(dat.salmon, year, strategy_permit) %>%
  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
  ggplot(aes(year, meanDiv)) + geom_line() + facet_wrap(~strategy_permit, scale="free") + 
  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Effective diversity (across people)") + xlab("")
```

```{r, fig.cap="Proportion of total returns to Prince William Sound (hatchery + wild, catch + escapement). Years before 1980 and after 2011 aren't included because not all species have data.", fig.pos="placeHere", echo=FALSE, fig.height = 3.3}
returns = readxl::read_excel("salmon_returns_PWS.xlsx")
returns = returns[,c("Year","Sockeye_total","Pink_total","Chum_total","Coho","Chinook")]

returns = as.data.frame(returns[which(returns$Year%in%c(1980:2011)),])
for(i in 1:ncol(returns)) returns[,i] = as.numeric(returns[,i])
for(i in 1:nrow(returns)) {
    # normalize
    returns[i,-1] = returns[i,-1] / sum(returns[i,-1])
}

melted = reshape2::melt(returns,id.vars="Year")
names(melted)[which(names(melted)=="variable")]="species"
ggplot(melted,aes(Year,value, group = species, color=species)) + geom_line() + ylab("Proportion of total returns (by #)")
```

\break

## Trends in revenue and prices

Across regions, we see lots of synchrony in prices between coho, chum, and pink salmon, with them increasing more quickly than sockeye/Chinook. Sockeye/Chinook have been the highest priced historically, with coho > chum > pink.  For most permit types, median revenue / person has generally increased, especially since early 2000s. For some permits, variability appears to have changed through time (quartile range getting smaller).

```{r, fig.cap="Total revenue by area (across people and permits within each)", fig.pos="placeHere", echo=FALSE, fig.height=3.5}
# plot prices of species landed by salmon permit
cfec.raw$area = NA
cfec.raw$area[cfec.raw$p_fshy%in%c("S 01A", "S 03A")] = "Southeast"
cfec.raw$area[cfec.raw$p_fshy%in%c("S 01E", "S 03E")] = "PWS"
cfec.raw$area[cfec.raw$p_fshy%in%c("S 01K", "S 04K")] = "Kodiak"
cfec.raw$area[cfec.raw$p_fshy%in%c("S 01M")] = "Alaska Peninsula"

group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, area) %>%
  summarize(p = sum(g_earn)) %>% group_by(year, area) %>%
  ggplot(aes(year, (p)/1e6, group=spec, color = spec)) + geom_line() + facet_wrap(~area, scale="free") +
  ylab("Revenue (millions) ") + xlab("Year")
```

\break

## Trends in variability

```{r, fig.cap="Trends in annual variability by fishery", fig.pos="placeHere", echo=FALSE,  fig.height=3.3}
summary = group_by(cfec.raw[cfec.raw$spec%in%c("CHUM","COHO","CHNK","SOCK","PINK"),], year, p_fshy, spec) %>% 
  summarize(sumg = log(sum(g_earn)), sumw = log(sum(g_pounds))) %>% 
  arrange(p_fshy, spec, year) %>% group_by(p_fshy, spec) %>% 
  mutate(logdiffg = rollapply(c(NA, diff(sumg)), 3, sd, fill=c(NA, NULL, NA)), 
    logdiffw = rollapply(c(NA, diff(sumw)), 3, sd, fill=c(NA, NULL, NA)))

ggplot(summary, aes(year, logdiffg, color=spec)) + geom_line() + facet_wrap(~ p_fshy, scale="free") + ylab("3 year rolling CV (revenue)") + xlab("Year")
```

```{r, fig.cap="Trends in annual variability by fishery and species", fig.pos="placeHere",  echo=FALSE, fig.height=3.3}
g1 = ggplot(summary[which(summary$spec=="PINK" & summary$p_fshy%in%c("S 01A", "S 01E", "S 03A", "S 03E")),], aes(year, logdiffg, color=p_fshy)) + geom_line() + ylab("3 year rolling CV (revenue)") + xlab("Year") + ggtitle("Pink salmon")

g2 = ggplot(summary[which(summary$spec=="CHUM" & summary$p_fshy%in%c("S 01A", "S 01E", "S 03A", "S 03E")),], aes(year, logdiffg, color=p_fshy)) + geom_line() + ylab("3 year rolling CV (revenue)") + xlab("Year") + ggtitle("Chum salmon")

gridExtra::grid.arrange(g1,g2, ncol=1)
```

```{r, fig.cap="Trends in annual variability by fishery and species", fig.pos="placeHere", echo=FALSE, eval=FALSE}
g1 = ggplot(summary[which(summary$spec=="PINK" & summary$p_fshy%in%c("S 01A", "S 01E", "S 03A", "S 03E")),], aes(year, logdiffw, color=p_fshy)) + geom_line() + ylab("3 year rolling CV (weight)") + xlab("Year") + ggtitle("Pink salmon")

g2 = ggplot(summary[which(summary$spec=="CHUM" & summary$p_fshy%in%c("S 01A", "S 01E", "S 03A", "S 03E")),], aes(year, logdiffw, color=p_fshy)) + geom_line() + ylab("3 year rolling CV (weight)") + xlab("Year") + ggtitle("Chum salmon")

gridExtra::grid.arrange(g1,g2, ncol=1)
```

```{r, fig.cap="Trends in variability across people", fig.pos="placeHere", echo=FALSE, fig.height=3.5, eval=FALSE}
summary = group_by(cfec.raw[cfec.raw$spec%in%c("CHUM","COHO","CHNK","SOCK","PINK"),], year, p_fshy, p_holder) %>% 
  summarize(sumg = sum(g_earn), sumw = sum(g_pounds)) %>% 
  group_by(p_fshy, year) %>% 
  summarize(cvg = sd(sumg,na.rm=T)/mean(sumg,na.rm=T),
    cvw = sd(sumw,na.rm=T)/mean(sumw,na.rm=T))
ggplot(summary, aes(year, cvg )) + geom_line(color ="green" ) + facet_wrap(~ p_fshy, scale="free") + geom_line(aes(year, cvw), color = "blue") + ylab("3 year rolling CV") + xlab("Year") + ggtitle("Variability across people: total revenue (green) and weight (blue)")
```

```{r, fig.cap="Change in CV for individual people", fig.pos="placeHere", echo=FALSE, fig.height=3.5}
summary = group_by(cfec.raw[cfec.raw$spec%in%c("CHUM","COHO","CHNK","SOCK","PINK"),], year, p_fshy, p_holder) %>% 
  summarize(sumg = sum(g_earn)) %>% 
  group_by(p_fshy, p_holder) %>% 
  mutate(nPre = length(which(year%in%seq(1985,1995))),
    nPost = length(which(year%in%seq(2005,2014)))) %>% 
  filter(nPre > 3 & nPost > 3) %>% select(-nPre, -nPost) %>% 
  summarize(cv.pre = mean(sumg[which(year <= 1995)],na.rm=T)/sd(sumg[which(year <= 1995)],na.rm=T),
    cv.post = mean(sumg[which(year >= 2005)],na.rm=T)/sd(sumg[which(year >= 2005)],na.rm=T))

ggplot(summary, aes(log(cv.post/cv.pre))) + geom_histogram(bins=100) + facet_wrap(~ p_fshy) + geom_vline(aes(xintercept = 0)) + ylab("Count") + xlab("% Change in CV for individual people (2005-2014 / 1985-1995)") + ggtitle("Change in variability for individual people")
#summary = left_join(expand.grid(year = unique(summary$year), p_holder = unique(summary$p_holder)), summary)

# add rolling CV for each person
#summary = group_by(summary, p_holder) %>% 
#  mutate(logdiffg = rollapply(c(NA, diff(sumg)), 3, sd, fill=c(NA, NULL, NA))) %>%
#  do(fit = lm(logdiffg ~ year, data = .))
```

\break

```{r, fig.cap="Inflation adjusted prices / species by area, relative to 2002 (the year corresponding to the lowest average prices).", fig.pos="placeHere", echo=FALSE, fig.height = 3.1}

g = group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, area) %>%
  summarize(p = mean(g_price))
p1 = group_by(g, spec, area) %>% 
  mutate(p = p/p[which(year==2002)]) %>% 
  group_by(year, area) %>%
  ggplot(aes(year, log(p), group=spec, color = spec)) + geom_line() + facet_wrap(~area) +
  ylab("Ln avg price relative to 2002") + ggtitle("Synchrony in prices: relative change") + xlab("")
p1

```

```{r, fig.cap="Percent change (log-difference) in inflation adjusted prices / species by area", fig.pos="placeHere", echo=FALSE, fig.height = 3.1}

g = group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, area) %>%
  summarize(p = mean(g_price))
# do log-diff of prices -- get price year before
g$p_t1 = g$p[match(paste(g$year, g$spec, g$area), 
  paste(g$year+1, g$spec, g$area))]

p2 = group_by(g, spec, area) %>% 
  ggplot(aes(year, log(p/p_t1), group=spec, color = spec)) + geom_line() + facet_wrap(~area) +
  ylab("Ln price (t/t-1)") + ggtitle("Synchrony in prices: percent change") + xlab("")
p2

g$logdiff = log(g$p/g$p_t1)
p3 = group_by(g, year, area) %>% 
summarize(sd = sd(logdiff,na.rm=TRUE)) %>% ggplot(aes(year, sd, color=area)) + geom_line() + xlab("Year") + ylab("SD of price changes across species") + ggtitle("Synchrony in price fluctuations across species")

```

```{r, fig.pos="placeHere", fig.cap = "Standard deviation (across species, by area) for fluctuations in prices. Fluctuations are calculated as the log difference price data between sequential years. The standard deviation of log differenced data is interpreted as the coefficient of variation (CV)."}
p3
```

\break

### Season length

There appears to be no significant trend in season length (defined as the difference between 95% CIs) or mean julian day for any fishery. In all cases, the dates are weighted by the earnings associated with landed catch.

```{r, fig.cap="Trends in mean julian day and season length by fishery", fig.pos="placeHere", echo=FALSE, fig.height=3.5}
gridExtra::grid.arrange(appendixPlot0_p1, appendixPlot0_p2, ncol = 1)
```

### 2016-10-05 salmon/paper/salmon_paper.Rmd
salmon update

--- 2016-10-05 salmon/archive/salmon_paper_combined.Rmd
@@
 ---
-title: "Reduced diversity and increased specialzation of Alaska salmon fishermen: effects on variation in individual revenues"
+title: 'Reduced diversity and increased specialzation of Alaska salmon fishermen:
+  effects on variation in individual revenues'
 author: ''
@@
   html_document: default
-  word_document: default
+bibliography: refs.bib
 ---
@@
 
-Previous analyses have largely focused on patterns of diversification across ecosystems and the benefits to adopt more or less diverse fishing strategies (Kasperski and Holland 2013; Anderson et al.
-in review).
-Fisheries in Alaska have been studied extensively, because they are one of the most productive regions in North America (both by landed weight, and revenue cite) and they allow individuals to adopt a wide range of fishing strategies.
-Anderson et al.
-(in review) defined strategies as a permit of collection of permits held by single individuals.
-While adding new permits allows fishers to increase species diveristy, few individual permits allow individuals to harvest more than one species (examples include halibut, sablefish, herring, crab, shellfish, shrimp, sea cucumber).
-Exceptions to these single species permits incldue participants in federally managed groundfish fisheries (CFEC 'M' permits) and participants in state managed salmon fisheries (CFEC 'S' permits).
+Previous analyses have largely focused on patterns of diversification across ecosystems and the benefits to adopt more or less diverse fishing strategies [@kasperski_income_2013; @anderson_benefits].
+Fisheries in Alaska have been studied extensively, because they are one of the most productive regions in North America (both by landed weight, and revenue [@office_of_science_and_technology_fisheries_2016]) and they allow individuals to adopt a wide range of fishing strategies.
+[@anderson_benefits] defined strategies as a permit of collection of permits held by single individuals.
+While adding new permits allows fishers to increase species diveristy, few individual permits allow individuals to harvest more than one species (examples of single-species permits in the Gulf of Alaska include halibut, sablefish, herring, crab, shellfish, shrimp, sea cucumber).
+Some exceptions to these single species permits include participants in federally managed groundfish fisheries and participants in state managed salmon fisheries.
 
-Understanding the effects of species diversification on salmon fishers is particularly important, as the number of salmon specialists (defined as fishers holding a single salmon permit) has increased over the last 30 years as a fraction of the total (Anderson et al.
-in review).
+Understanding the effects of species diversification on salmon fishers is particularly important, as the number of salmon specialists (defined as fishers holding a single salmon permit) has increased over the last 30 years as a fraction of the total [@anderson_benefits].
 For these limited entry salmon fisheries, general trends that have emerged are that more people are participating since the early 2000s (Fig.
@@
 A6).
-Across different salmon fisheries, other than a slight negative effect of diversification on variability (similar to Anderson et al.
-in review), there appears to be little correlation between diversity and revenue (Fig.
+Across different salmon fisheries, other than a slight negative effect of diversification on variability (similar to [@anderson_benefits]), there appears to be little correlation between diversity and revenue (Fig.
 A4).
-In terrestrial agriculture, there are numerous examples where diversification may affect the mean or CV but not both (Purdy, Barry M; Langemeier, Michael R; Featherstone, Allen M.
-Journal of Agricultural and Applied Economics29.1 (Jul 1, 1997): 149.), and the same result may be true for salmon fishers.
+In terrestrial agriculture, there are numerous examples where diversification may affect the mean or CV but not both [@purdy_financial_1997], and the same result may be true for salmon fishers.
 Further, these impacts of diversification may be variable over time.
@@
 Bristol Bay set and drift gillnet fisheries).
-For these fisheries, specialization is associated with higher revenue and reduced CV in annual revenue (Anderson et al.
-in review).
+For these fisheries, specialization is associated with higher revenue and reduced CV in annual revenue [@anderson_benefits].
 
 A secondary motivation for focusing on the effects of diversification for salmon fishers is to better understand the long term consequences of hatchery salmon production on the revenue and variability in revenue of individuals.
-Following low returns and declining productivity in the mid 1980s, many hatchery programs in the state rapidly increased hatchery releases (Mahnken et al.
-1998).
-Many of these hatchery programs increased production to stabilize the variability of fishers catch (and revenue), though these programs may also have costs to wild production (Ruggerone et al.
-2015).
+Following low returns and declining productivity in the mid 1980s, many hatchery programs in the state rapidly increased hatchery releases [@mahnken_historical_1998; @leber_stock_2008].
+Many of these hatchery programs increased production to stabilize the variability of fishers catch (and revenue), though these programs may also have costs to wild production [@ruggerone_productivity_2015].
+The intra- and inter-specific effects of large scale hatchery programs has been debated [@hilborn_review_2001; @hilborn_hatcheries_1992; @wertheimer_comment:_2001; @heard_hatchery], but the effects of hatcheries have largely been focused on ecological interactions of aggregate yield to the fishery [@leber_stock_2008] or economic impacts [boyce_alaska_1993], rather than to variability experienced by individual fishermen.
 [Other history here?]
 
-In addition to the changing species composition of fish avaialble to be harvested, the benefits (or costs) of diversification are likely affected by a number of exogenous drivers, including marketing, prices, new processing technology, etc (cite Gunnar Knapp + others).
+In addition to the changing species composition of fish avaialble to be harvested, the benefits (or costs) of diversification are likely affected by a number of exogenous drivers, including marketing, prices, new processing technology [@knapp_trends].
 [Break out as own paragraph+add more examples?] Following price crashes in the early 2000s, many of these factors have been positive for salmon fishermen (increasing trends in prices and catches, Fig.
@@
 Combined with the increased trend in hatchery production and resulting change species compositon of adult salmon, these external factors may have also altered incentives for diversification.
-If there's an incentive to substitute high value species (Chinook, sockeye) with abundant low value species (pink, chum), the analog in the terrrestrial framring world would be crop substitution of low-volume high value crops (spices) with high-volume low value crops (cereals) (other examples in literature = cocoa, tobacco).
+If there's an incentive to substitute high value species (Chinook, sockeye) with abundant low value species (pink, chum), the analog in the terrestrial framring world would be crop substitution of low-volume high value crops (spices) with high-volume low value crops (cereals) (other examples in literature = cocoa, tobacco).
 
-The objective of our analysis is to better understand the effects of diversification on single-permit salmon holders by extending the methods used by Anderson et al.
-(in review) to include time-varying effects.
+The objective of our analysis is to better understand the effects of diversification on single-permit salmon holders by extending the methods used by [@anderson_benefits] to include time-varying effects.
 These time-varying effects will allow us to compare temporal trends across regions, identify periods when having a diverse catch portfolio was most beneficial, and allow us to evaluate support for effects of longer term drivers in the region.
@@
 We adjusted for inflation by setting all revenues in 2009 USD.
-Because were were only interested in single permit fisheries, we adopted a simplified version of the filtering steps described in  Anderson et al.
-(in review).
+Because were were only interested in single permit fisheries, we adopted a simplified version of the filtering steps described in [@anderson_benefits].
 After aggregating individual fish tickets to annual data, we retained individuals who only fished one of the six salmon permits included in a year (Table 1, Fig.
 A5).
-Like Anderson et al.
-(in review) we also worked with a log-differenced dataset to remove non-stationary individuals; to avoid complications of individuals switching strategies, we only included pairs of years for a given person where they did not change strategies.
+Like [@anderson_benefits] we also worked with a log-differenced dataset to remove non-stationary individuals; to avoid complications of individuals switching strategies, we only included pairs of years for a given person where they did not change strategies.
 
@@
 Hatchery release trends for some regions included in our data have been summarized previously (e.g.
-Ward et al.
-in review summarized the trends for Prince William Sound).
-Release patterns have been dominated by pink and chum salmon (for PWS, these species represent 80% and 18% of total releases for 2012-2014 respectively, Ward et al.
-in review) and production has increased since the late 1970s (217 million juvenile pink salmon in PWS in 1985 versus 672 million in 2014).
+[@ward_evaluating] summarized the trends for Prince William Sound).
+Release patterns have been dominated by pink and chum salmon (for PWS, these species represent 80% and 18% of total releases for 2012-2014 respectively, [@ward_evaluating]) and production has increased since the late 1970s (217 million juvenile pink salmon in PWS in 1985 versus 672 million in 2014).
 Though our analysis doesn't include , as a contrast we can do coarse comparisons across regions and species to examine potential effects of hatchery fish on variability.
@@
 As a second comparison, we examined whether long-term participants in any of the fisheries had experienced a change in variability.
-Similar to Kasperski and Holland (2013) who calculated CVs of vessels before and after significant managment events, we identified individuals who fished the same permit for at least three years in the first decade of our data (1985-1995) and at least three years in the last decade of the data (2004-2014).
+Similar to [@kasperski_income_2013] who calculated CVs of vessels before and after significant managment events, we identified individuals who fished the same permit for at least three years in the first decade of our data (1985-1995) and at least three years in the last decade of the data (2004-2014).
 
@@
 
-Anderson et al.
-(in review) implemented a hierarchical Bayesian model that allowed group level predictors and covariates (diversity of catch, days fished) to simultaneously affect the mean and the variance of revenue for individual fishers.
+[@anderson_benefits] implemented a hierarchical Bayesian model that allowed group level predictors and covariates (diversity of catch, days fished) to simultaneously affect the mean and the variance of revenue for individual fishers.
 This model also included random intercepts by strategy and year; these terms allowed all permit holders in a given year to experience common fluctuations (external drivers including price changes, fluctuations in fish population abundance, or environmental disasters such as EVOS).
 
-To ask whether how benefits of diversification may have changed over time, we extend the model of Anderson et al.
-(in review) to also include time-varying random effects in the slope coefficients of diversification.
+To ask whether how benefits of diversification may have changed over time, we extend the model of [@anderson_benefits] to also include time-varying random effects in the slope coefficients of diversification.
 We define the variables $\Delta { S }_{ t }=log\left( { S }_{ t }/{ S }_{ t-1 } \right)$ to represent the percent change in species diversity of the catch between years and $\Delta { D }_{ t }=log\left( { D }_{ t }/{ D }_{ t-1 } \right)$ to represent the percent change in fishing effort (days fished) between years.
-The model for the the expected log revenue of individual *i* in year *t*, $log({ R }_{ i,t })$, can be written as $$\widehat { log\left( { R }_{ i,t } \right)  } ={ B }_{ 0,j,t }+{ B }_{ 1,j,t }\cdot \Delta S_{ t }+{ B }_{ 2,j,t }\cdot \left( \Delta S_{ t }\cdot { S }_{ t-1 } \right) +{ B }_{ 3,j,t }^{ 2 }\cdot \left( \Delta S_{ t }\cdot { S }_{ t-1 }^{ 2 } \right) +{ B }_{ 4,j }\cdot \Delta D_{ t }+log\left( { R }_{ i,t-1 } \right) $$, where *j* subscripts the salmon permit (Table 1).
+The model for the the expected log revenue of individual *i* in year *t*, $$\widehat { log\left( { R }_{ i,t } \right)  } ={ B }_{ 0,j,t }+{ B }_{ 1,j,t }\cdot \Delta S_{ t }+{ B }_{ 2,j,t }\cdot \left( \Delta S_{ t }\cdot { S }_{ t-1 } \right) +{ B }_{ 3,j }\cdot \Delta D_{ t }+log\left( { R }_{ i,t-1 } \right) $$, where *j* subscripts the salmon permit (Table 1).
 We model the time varying permit specific intercepts as random and exchangeable across years, ${ B }_{ 0,j,t }\sim Normal\left( { u }_{ { B }_{ 0 } },{ \sigma  }_{ { B }_{ 0 } } \right)$.
 The term ${ B }_{ 1,j,t }$ represents the permit level random effect of diversifying species.
-Our model extendeds that of Anderson et al.
-(in review) to include the interaction terms $\Delta S_{ t }\cdot { S }_{ t-1 }$ which allow the effect of diversification to be dependent on the previous level of diversification.
-The squared term allows this interaction to be quadratic, and the coefficients ${ B }_{ 2,j,t }$ and ${ B }_{ 3,j,t }$, allow this to vary by permit and through time.
+Our model extendeds that of [@anderson_benefits] to include the interaction terms $\Delta S_{ t }\cdot { S }_{ t-1 }$ which allow the effect of diversification to be dependent on the previous level of diversification.
+We assume this relationship is linear, and the coefficients ${ B }_{ 1,j,t }$ and ${ B }_{ 2,j,t }$ allow this to vary by permit and through time.
 The flexibility in these terms allows the benefit of increasing species diversity to be positive in some years (generalizing benefits revenue), negative in other years (specializing benefits revenue) and neutral in other years (revenue is independent alterations to catch composition).
 We constrain the random effects associated with diversification to be modeled as autoregressive processes with shared variances across strategies, for example: ${ B }_{ 1,j,t }\sim Normal\left( { B }_{ 1,j,t-1 },{ \sigma  }_{ { B }_{ 1 } } \right)$.
-Like Anderson et al.
-(in review) we also allow changing effort (measured as percent change in days fished) ${ B }_{ 4,j }$ to vary by strategy, but we do not allow this effect to vary through time.
-Finally, the term $log\left( { R }_{ i,t-1 } \right)$ is included as an offset (Anderson et al.
-in review).
+Like [@anderson_benefits] we also allow changing effort (measured as percent change in days fished) ${ B }_{ 4,j }$ to vary by strategy, but we do not allow this effect to vary through time.
+Finally, the term $log\left( { R }_{ i,t-1 } \right)$ is included as an offset [@anderson_benefits].
 
-- Fit models from Sean's paper, but include time varying interactions on species diversity.  GLMM code below
+- Fit models from Sean's paper, but include time varying interactions on species diversity.
+All done in STAN [@stan_development_team_stan_2015] using the NUTS algorithm [@hoffman_no-u-turn_2014].
 
@@
 
+Figure 1 shows that the time varying intercepts on the mean, also in [@anderson_benefits] are basically flat. These represent common fluctuations in revenue experienced by all permit holders.
+
+Figure 2 is largley illustrative, but shows the benefit of including time varying coefficients.
+
+Figure 3 shows the temporal trends, by permit, in going from 1 to 2 species fished, and how those effects have changed through time.
+S01A, S01E, S03A appear to be decreasing.
+S01K and S01M increasing.
+S03E pretty flat?
+
+Figure 4 shows that temporal trends, by permit, in going from 1 to 2 species fished on the variance.
+From [@anderson_benefits] we'd expect diversification to redudce variance -- that seems to be the biggest signal for S01K and S01M in the years of low sockeye prices.
+
 ## Discussion
@@
 We're also seeing a trend toward greater synchrony between species (Fig A11-A12, A14).
-Projecting future prices are complex -- Gunnar Knapp has a good presentation on this http://www.adfg.alaska.gov/static/fishing/PDFs/commercial/gk_trends_4-12-12.pdf.
+Projecting future prices are complex [@knapp_trends].
 Synchrony in prices may largely be function of marketing, improved product, freezing, etc.
@@
 
-\break
+```{r, echo=FALSE, fig.pos="placeHere", fig.cap="Time varying intercepts on estimated mean revenue, by permit."}
+library(rstan)
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
-revenueThreshold = 10000
+# grab original data to look at permit names
+diffdat = readRDS(file="../../salmon/data-generated/salmon.rds")
+years = seq(min(diffdat$year), max(diffdat$year))
+permits = levels(as.factor(diffdat$permit))
 
-dat.annual = readRDS(file="../../portfolio/data-generated/cfec-annual-for-modeling.rds")
+yearpermit = data.frame("yrpermit"=levels(diffdat$year_permit))
+yearpermit$year = as.numeric(substr(yearpermit$yrpermit,1,4))
+yearpermit$permit = substr(yearpermit$yrpermit,6,9)
 
-# single permit holders, salmon permits
-dat.annual  = filter(dat.annual ,npermit==1 & substr(strategy_permit,1,1)=="S") %>%
-  group_by(strategy_permit) %>% 
-  mutate(nPeople = length(unique(p_holder))) %>% 
-  filter(nPeople > 200) %>% 
-  dplyr::select(-nPeople)
-# drop troll
-dat.annual = dat.annual[dat.annual$strategy_permit%in%c("S05B","S15B")==FALSE,]
-# drop S04X, Y, Z because low participation (< 20) in many years
-# S04W deleted because low median earnings
-dat.annual = dat.annual[dat.annual$strategy_permit%in%c("S04W", "S04X","S04Y","S04Z")==FALSE,]
+load("../../salmon/analysis/model-linear.Rdata")
 
-agg.annual = group_by(dat.annual, strategy_permit, year) %>% 
-  summarize(people = length(unique(p_holder)),
-    specDiv = mean(specDiv),
-    meanrev = median(revenue),
-    low = quantile(revenue,0.25),
-    upp = quantile(revenue,0.75))
+est = extract(mod, permute="TRUE")
 
-dat.diff = readRDS(file="../../portfolio/data-generated/cfec-diff-for-modeling.rds")
-dat.diff = dat.diff[which(dat.diff$revenue >= revenueThreshold & dat.diff$revenue.prev >= revenueThreshold), ]
-dat.diff = dat.diff[dat.diff$strategy_permit%in%c("S01A","S01E","S01K","S01L","S01M",
-  "S03A","S03E","S03H","S03M","S03T","S04D","S04H","S04K","S04M","S04T"),]
+yearpermit$b0 = apply(est$b0_str_yr,2,median)
+yearpermit$b0_low = apply(est$b0_str_yr,2,quantile,0.25)
+yearpermit$b0_hi = apply(est$b0_str_yr,2,quantile,0.75)
 
-agg.diff = group_by(dat.diff, strategy_permit) %>% 
-  summarize(specDiv = mean(specDiv), meanrev = mean(revenue),
-    cv = sd(log(revenue/revenue.prev)))
-
-# Also load in raw CFEC data for plots
-cfec <- feather::read_feather("../../portfolio/data-generated/cfec.feather")
-cfec = cfec[cfec$p_fshy%in%c("S 01A", "S 01E", "S 01K", "S 03A", "S 03E", "S 01M"),]
-cfec$year = as.numeric(substr(cfec$landdate, 1, 4))
-
-# only use people who are using just that permit - diversification within permit
-cfec.raw = cfec
-dat.salmon = dat.diff[dat.diff$strategy_permit%in%c("S01A", "S01E", "S01K", "S03A", "S03E","S01M"),]
-
-# Restrict analysis to people who keep same strategy - complicates days, etc
-dat.salmon = dat.salmon[which(dat.salmon$strategy_permit==dat.salmon$strategy_permit.prev),]
-
-dat.salmon$days.change = log((dat.salmon$days_permit+1) / (dat.salmon$days_permit.prev+1))
-dat.salmon$year_permit = as.factor(paste(dat.salmon$year, dat.salmon$strategy_permit))
-
-cfec = cfec[which(paste(cfec$p_holder,cfec$year) %in% paste(dat.salmon$p_holder,dat.salmon$year)),]
-
-# Make come plots for the appendix of (1) median fishing dates over time, (2) length of fishing season over time (calculated as width of 95% intervals)
-cfec.raw$month = as.numeric(substr(cfec.raw$day, 1,2))
-cfec.raw$day = as.numeric(substr(cfec.raw$day, 4,5))
-cfec.raw$jday = mdy.date(cfec.raw$month, cfec.raw$day, cfec.raw$year) - mdy.date(1, 1, cfec.raw$year)
-# summarize the temporal distribution of the catches
-seasons = group_by(cfec.raw, p_fshy, year) %>% 
-  summarize(q = mean(sample(jday, size = 10000, replace=T, prob=g_earn)),
-    lower = quantile(sample(jday, size = 10000, replace=T, prob=g_earn),0.025),
-    upper = quantile(sample(jday, size = 10000, replace=T, prob=g_earn),0.975),
-    length = upper - lower)
-seasons = rename(seasons, permit = p_fshy)
-appendixPlot0_p1 = ggplot(seasons, aes(year, q, color = permit, group=permit)) + geom_line() + xlab("Year") + ylab("Mean julian day")
-appendixPlot0_p2 = ggplot(seasons, aes(year, length, color = permit, group=permit)) + geom_line() + xlab("Year") + ylab("Season length")
-```
-
-```{r, echo = FALSE, eval=FALSE}
-dat = dat.salmon
-# create other category, lumping coho/chinook
-dat$salm = dat$coho+dat$chnk
-dat$salm.prev = dat$coho.prev+dat$chnk.prev
-
-dat$totsalm = dat$salm + dat$chum + dat$sock + dat$pink
-dat$totsalm.prev = dat$salm.prev + dat$chum.prev + dat$sock.prev + dat$pink.prev
-
-dat$salm = dat$salm/dat$totsalm
-dat$sock = dat$sock/dat$totsalm
-dat$pink = dat$pink/dat$totsalm
-dat$chum = dat$chum/dat$totsalm
-dat$chnk = dat$chnk/dat$totsalm
-dat$coho = dat$coho/dat$totsalm
-dat$salm.prev = dat$salm.prev/dat$totsalm.prev
-dat$sock.prev = dat$sock.prev/dat$totsalm.prev
-dat$pink.prev = dat$pink.prev/dat$totsalm.prev
-dat$chum.prev = dat$chum.prev/dat$totsalm.prev
-dat$chnk.prev = dat$chnk.prev/dat$totsalm.prev
-dat$coho.prev = dat$coho.prev/dat$totsalm.prev
-
-dat = dat[-which(is.na(dat$salm+dat$sock+dat$pink+dat$chum)),]
-dat$logdiff = log(dat$revenue/dat$revenue.prev)
-
-dat_old = dat
-dat$sumP2 = abs(dat$salm - dat$salm.prev) + 
-  abs(dat$pink-dat$pink.prev) + 
-  abs(dat$chum-dat$chum.prev) + 
-  abs(dat$sock-dat$sock.prev)
-dat = dat[which(dat$sumP2 < 0.3),]
-```
-
-```{r, echo = FALSE, eval=FALSE, fig.cap="Results for PWS Purse seine fishery", fig.pos="placeHere", cache=TRUE}
-
-df = dat[which(dat$strategy_permit=="S01E"),]
-pamk_s01e <- pamk(df[,c("chum","pink","sock","salm")])
-df$group = c("Pink-chum","Pink")[pamk_s01e$pamobject$clustering]
-
-p3 = ggtern(data=df[,c("chum","pink","sock","group")], aes(x=chum,y=pink,sock,color=as.factor(group))) +
-  geom_point(alpha=0.3) +
-  labs(x="CHUM",y="PINK",z="SOCK") + theme(
-        axis.title=element_text(size=8)) + theme(legend.position = "none")
-
-p0 = group_by(df, year, group) %>% 
-  summarize(n = length(unique(p_holder))) %>% 
-ggplot(aes(x = year, y = n, group=group, col = group)) + geom_line() + ylab("People") + ggtitle("S01E")
-
-p1 = group_by(df, year, group) %>% 
-  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
-    cv = sd(logdiff,na.rm=T)) %>% 
-ggplot(aes(x = year, y = meanrev, group=group, col = group)) + geom_line() + ylab("Mean revenue") + ggtitle("S01E")
-
-p2 = group_by(df, year, group) %>% 
-  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
-    cv = sd(logdiff,na.rm=T)) %>% 
-ggplot(aes(x = year, y = cv, group=group, col = group)) + geom_line() + ylab("CV revenue") + ggtitle("S01E")
-
-g = group_by(df, year, group) %>% 
-  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
-    cv = sd(logdiff,na.rm=T))
-
-gridExtra::grid.arrange(p3,p0,p1,p2,ncol=2)
-```
-
-\break
-
-```{r makePlots, echo = FALSE}
-appendixPlot1 = ggplot(agg.annual, aes(year, people, colour = specDiv)) + geom_line() + facet_wrap(~ strategy_permit, scale="free") + ylab("Participation") + xlab("Year")
-
-appendixPlot2_p1 = ggplot(agg.diff, aes(specDiv, log(meanrev), colour = cv, label=strategy_permit)) + xlab("Species diversity") +ylab("Ln (mean revenue)") + geom_text(size=2)
-appendixPlot2_p2 = ggplot(agg.diff, aes(specDiv, cv, colour = log(meanrev), label=strategy_permit)) + xlab("Species diversity") +ylab("CV") + geom_text(size=2)
-appendixPlot2_p3 = ggplot(agg.diff, aes(log(meanrev), cv, colour = specDiv, label=strategy_permit)) + ylab("CV") + xlab("Ln (mean revenue)") + geom_text(size=2)
-```
-
-\break
-
-```{r, fig.cap="Species composition for 1988 and 2014 (each dot is a person) in the S03E fishery", fig.pos="placeHere", echo=FALSE, fig.height = 2, eval=FALSE}
-### One way to visualize the change is with ternary plots over the 3 axes (chum / sockeye / pink)
-
-daty = group_by(cfec[cfec$p_fshy=="S 03E" & cfec$spec %in% c("CHUM","PINK","SOCK","COHO","CHNK"),], spec, p_holder, year) %>%
-  summarize(p = sum(g_earn)) %>%
-  group_by(p_holder, year) %>% mutate(p = p/sum(p))
-daty <- melt(daty, id.vars = c("spec", "p_holder", "year"))
-daty <- dcast(daty, p_holder + year ~ spec + variable)[,-1]
-daty[is.na(daty)]=0
-ggtern(data=daty[which(daty$year%in%c(1988,2014)),c("year", "CHUM_p","PINK_p","SOCK_p")], aes(x=CHUM_p,y=PINK_p,SOCK_p,color=as.factor(year))) +
-  geom_point(alpha=0.3,size=1) +
-  labs(x="CHUM",y="PINK",z="SOCK",title="Proportion of annual revenue (S03E)")
-```
-
-\break
-
-```{r, echo=FALSE, eval=FALSE}
-
-mu.mod = glmmTMB(log(revenue) ~ specDiv.change + specDiv.change:specdiv.prev + specDiv.change:I(specdiv.prev^2) + days.change + (1+days.change|strategy_permit) + (-1+specDiv.change + specDiv.change:specdiv.prev + specDiv.change:I(specdiv.prev^2)|year_permit), offset = log(revenue.prev), dat.salmon)
-
-cv.mod = glmmTMB(log(abs(residuals(mu.mod))) ~ specDiv.change + specDiv.change:specdiv.prev + specDiv.change:I(specdiv.prev^2) + days.change + (1+days.change|strategy_permit) + (-1+specDiv.change + specDiv.change:specdiv.prev + specDiv.change:I(specdiv.prev^2)|year_permit), dat.salmon)
-# to ask whether benefit of diversificaiton is changing via strategy, we want to summarize (1) direction of 
-# benefit (e.g. is the effect negative or positive), and (2) the magnitude (how steep / shallow it is). With the
-# quadratic model, 
-# 'squared term' indicates concave up/down
-# derivative is magnitude, 2a * x + b
-
-df = ranef(mu.mod)$cond$year_permit
-colnames(df) = c("c","b","a")
-df$year = as.numeric(substr(rownames(df), 1, 4))
-df$strategy = substr(rownames(df), 6, 9)
-df$deriv = 2*df$a + df$b
-# also calculate change of going from 1 to 2 spp
-df$pred1 = df$a*(1^2) + df$b*1 + df$c + fixef(mu.mod)$cond["specDiv.change"] + fixef(mu.mod)$cond["specDiv.change:specdiv.prev"]*1
-# add fixed effects into predictions
-
-# We can demonstrate the quadratic curves by year for a single strategy / set of years
-df.s01e = df[df$strategy=="S01E",]
-demo = expand.grid("speciesDiversity" = seq(1,4,length.out=100), "year" = 1986:2014)
-demo = left_join(demo, df.s01e)
-demo$effect = demo$a * (demo$speciesDiversity^2) + demo$b * (demo$speciesDiversity) + demo$c
-```
-
-```{r, echo = FALSE, warning=FALSE, message=FALSE, results='hide', eval=FALSE}
-pdf("../../salmon/Salmon_residuals.pdf")
-# Residuals for mean model
-dat.salmon$mu.resid = residuals(mu.mod)
-dat.salmon$mu.fitted = fitted(mu.mod)
-
-ggplot(dat.salmon, aes(mu.fitted, mu.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Fitted") + ylab("Residuals") + ggtitle("Mean model")
-ggplot(dat.salmon, aes(specDiv.change, mu.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Change in species diversification") + ylab("Residuals") + ggtitle("Mean model") + geom_hline(aes(yintercept=0),color="salmon1")
-ggplot(dat.salmon, aes(specdiv.prev, mu.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Previous species diversity") + ylab("Residuals") + ggtitle("Mean model") + geom_hline(aes(yintercept=0),color="salmon1")
-ggplot(dat.salmon, aes(log((revenue+revenue.prev)/2), mu.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Average revenue") + ylab("Residuals") + ggtitle("Mean model") + geom_hline(aes(yintercept=0),color="salmon1")
-ggplot(dat.salmon, aes(specDiv.change, log(abs(mu.resid)), group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Change in species diversification") + ylab("Ln abs(residuals)") + ggtitle("Mean model")
-ggplot(dat.salmon, aes(specdiv.prev, log(abs(mu.resid)), group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Previous species diversity") + ylab("Ln abs(residuals)") + ggtitle("Mean model")
-ggplot(dat.salmon, aes(specdiv.prev, log(abs(mu.resid)), group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Previous species diversity") + ylab("Ln abs(residuals)") + ggtitle("Mean model")
-
-# Residuals for CV model
-dat.salmon$cv.resid = residuals(cv.mod)
-dat.salmon$cv.fitted = fitted(cv.mod)
-
-ggplot(dat.salmon, aes(cv.fitted, cv.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Fitted") + ylab("Residuals") + ggtitle("CV model")
-ggplot(dat.salmon, aes(specDiv.change, cv.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Change in species diversification") + ylab("Residuals") + ggtitle("CV model") + geom_hline(aes(yintercept=0),color="salmon1")
-ggplot(dat.salmon, aes(specdiv.prev, cv.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Previous species diversity") + ylab("Residuals") + ggtitle("CV model") + geom_hline(aes(yintercept=0),color="salmon1")
-ggplot(dat.salmon, aes(log((revenue+revenue.prev)/2), cv.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Average revenue") + ylab("Residuals") + ggtitle("CV model") + geom_hline(aes(yintercept=0),color="salmon1")
-dev.off()
-```
-```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Variation in effects of diversification by year for the S01E (PWS purse seine) permit holders. ", eval=FALSE}
-###We can make quadratic plots of these effects by year and strategy.
-These plots represent the effect of increasing diversity (as relative change) on mean revenue.
-For example, in 1986 if you were fishing 3-4 species there would be little effect of adding more spp; but if you were fishing 1 species, there would be a slight negative effect of adding more.
-In 1991, there would be a huge benefit of increasing diversity, and that effect would taper off as more were added.
-
-ggplot(demo, aes(speciesDiversity, effect, group = year)) + geom_line() + facet_wrap(~ year) + xlab("Species Diversity") + ylab("Effect of increasing diversity (relative change)")
-```
-
-\break
-
-```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Time series of benefits of diversification by permit / strategy.", fig.height = 3, eval=FALSE}
-###This plot shows the time series of becoming increasingly diversified from 1 species on mean revenue.
-For S01E for example, there's huge benefits of diversifying in 1991, and huge costs in 2010.
-2010 was a huge pink salmon year for PWS (with high prices specialization was good) and 1991 was the start of pink salmon stock + price collapse.
-Across permits, perhaps slight declines in some of these in recent years?
-
-ggplot(df, aes(year, pred1, group = strategy, colour = as.factor(strategy))) + geom_line() + xlab("Year") + ylab("Benefit of increasing species diversity from 1 species") + ggtitle("Effect on Mean")
-```
-
-```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Time series of benefits of diversification split out by permit / strategy.", fig.height = 3, eval=FALSE}
-ggplot(df, aes(year, pred1, group = strategy, colour = as.factor(strategy))) + geom_line() + xlab("Year") + ylab("Benefit of increasing species diversity from 1 species") + facet_wrap(~ strategy, scale="free") + geom_smooth() + geom_hline(aes(yintercept=0))
-```
-
-```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Time series of effects of diversification on mean and CV by permit.", eval=FALSE}
-df.cv = ranef(cv.mod)$cond$year_permit
-colnames(df.cv) = c("c.cv","b.cv","a.cv")
-df.cv$year = as.numeric(substr(rownames(df.cv), 1, 4))
-df.cv$strategy = substr(rownames(df.cv), 6, 9)
-# also calculate change of going from 1 to 2 spp
-df.cv$pred1.cv = df.cv$a.cv*(1^2) + df.cv$b.cv*1 + df.cv$c.cv + fixef(cv.mod)$cond["specDiv.change"] + fixef(cv.mod)$cond["specDiv.change:specdiv.prev"]*1
-
-ggplot(df.cv, aes(year, pred1.cv, group = strategy, colour = as.factor(strategy))) + geom_line(linetype="dashed") + geom_line(data=df, aes(year, pred1, group = strategy, colour = as.factor(strategy))) + xlab("Year") + ylab("Effect of increasing species diversity from 1 species") + facet_wrap(~ strategy, scale="free") + ggtitle("Effects on mean (solid) and CV (dashed)") + geom_hline(aes(yintercept=0), color="grey30")
+ggplot(data=yearpermit) + 
+  geom_ribbon(aes(x=year,ymin=b0_low,ymax=b0_hi),alpha=0.3) +
+  geom_line(aes(x=year,y=b0)) + 
+  facet_wrap(~permit) + xlab("Year") + ylab("Intercepts on mean")
 
@@
 
-```{r, echo=FALSE, eval = FALSE}
-# These are the run  timing plots per Milo's questions
+```{r, fig.pos="placeHere", fig.cap="Time varying effects of adding 1 species on revenue (measured in ln dollars) for purse seine permit in Prince William Sound (S01E).
+This figure illustrates two things: (1) the benefit of adding more species diminishes as a function of past diversity, and (2) the economic benefit to adding more species has been reduced through time.", echo=FALSE}
+# grab original data to look at permit names
+diffdat = readRDS(file="../../salmon/data-generated/salmon.rds")
+years = seq(min(diffdat$year), max(diffdat$year))
+permits = levels(as.factor(diffdat$permit))
 
-# for 2008, create cumulative density plots of salmon landings by calday
-pinkchum = cfec.raw[cfec.raw$p_fshy=="S 01E" & cfec.raw$spec%in%c("PINK","CHUM","SOCK"),]
-#pinkchum = cfec.raw[cfec.raw$p_fshy=="S 03T" & cfec.raw$spec%in%c("PINK","SOCK","CHNK"),]
+load("../../salmon/analysis/model-linear.Rdata")
 
-# aggregate by calendar day
-pinkchum$month = as.numeric(substr(pinkchum$day, 1, 2))
-pinkchum$day = as.numeric(substr(pinkchum$day, 4, 5))
-pinkchum$jday = date::mdy.date(pinkchum$month, pinkchum$day, pinkchum$year) - date::mdy.date(1,1,pinkchum$year)
+est = extract(mod, permute="TRUE")
 
-group_by(pinkchum, jday, year) %>% 
-  summarize(pink = sum(g_earn[spec=="PINK"]), 
-    chum = sum(g_earn[spec=="CHUM"]),
-    sock = sum(g_earn[spec=="SOCK"])) %>% 
-  group_by(year) %>% 
-  mutate(tot = sum(pink) + sum(chum) + sum(sock), pink = pink/tot, chum = chum/tot, sock = sock/tot) %>% 
-ggplot(aes(jday, pink)) + geom_line(color="pink") + geom_line(aes(jday, chum), color="darkgreen") + geom_line(aes(jday, sock), color="lightblue") + facet_wrap(~ year, scale="free")
+# start with making time series of estimates on mean
+# challenge is that they're time varying quadratic functions
+# do facet_wrap of estimates by year
 
-# make plot of all earnings or weight by jday
-salm = cfec.raw[cfec.raw$p_fshy=="S 01E",]
-salm$month = as.numeric(substr(salm$day, 1, 2))
-salm$day = as.numeric(substr(salm$day, 4, 5))
-salm$jday = date::mdy.date(salm$month, salm$day, salm$year) - date::mdy.date(1,1,salm$year)
+# effect of adding more 1 species from 1. specdiv.prev = 1
 
-g = group_by(salm, jday, year) %>% 
-  summarize(earn = sum(g_earn), 
-    lbs = sum(g_pounds), 
-    pink = sum(g_earn[spec=="PINK"])/earn,
-    sock = sum(g_earn[spec=="SOCK"])/earn,
-    chum = sum(g_earn[spec=="CHUM"])/earn,
-    coho = sum(g_earn[spec=="COHO"])/earn,
-    chnk = sum(g_earn[spec=="CHNK"])/earn) %>% 
-ggplot(g, aes(jday, (earn))) + geom_line(colour = rgb(g$sock+g$coho+g$chnk, g$chum, g$pink)) + facet_wrap(~ year, scale="free_y")
+specdiv.prev = seq(1, 5, by=0.1)
+specdiv = specdiv.prev + 1 # always add 1 more in these comparisons
+x1 = log(specdiv/specdiv.prev)
+x2 = log(specdiv/specdiv.prev) * specdiv.prev
+x3 = log(specdiv/specdiv.prev) * specdiv.prev^2
 
-```
+df = expand.grid("year"=1:29, "diversity" = seq(1,length(x1)), "permit"=1:6)
+df$permits = permits[df$permit]
+df$years = years[df$year]
+df$x = specdiv.prev[df$diversity]
+df$pred = 0
+df$low = 0
+df$high = 0
+for(i in 1:nrow(df)) {
+  tmp = x1[df$diversity[i]] * (est$b_1[,df$year[i], df$permit[i]]+est$b1_str_yr_mu) +
+    x2[df$diversity[i]] * (est$b_2[,df$year[i], df$permit[i]]+est$b2_str_yr_mu)# +
+    #x3[df$diversity[i]] * (est$b_3[,df$year[i], df$permit[i]]+est$b3_str_yr_mu)
+  df$pred[i] = median(tmp)
+  df$low[i] = quantile(tmp,0.25)
+  df$high[i] = quantile(tmp,0.75)
+}
 
-```{r, echo = FALSE, fig.cap="SEAK purse seine fishery", cache=TRUE, fig.pos="placeHere", eval=FALSE}
+ggplot(df[df$permits=="S01E",], aes(x=x,y=pred,group=years)) +
+  #geom_ribbon(aes(ymin=low,ymax=high), alpha=0.4) +
+  geom_line() +
+  facet_wrap(~years) + xlab("Previous species diversity") + ylab("Predicted benefit of adding +1 spp (ln $)")
 
-df = dat[which(dat$strategy_permit=="S01A"),]
-pamk_s01a <- pamk(df[,c("chum","pink","sock","salm")])
-df$group = pamk_s01a$pamobject$clustering
-
-p3 = ggtern(data=df[,c("chum","pink","sock","group")], aes(x=chum,y=pink,sock,color=as.factor(group))) +
-  geom_point(alpha=0.3) +
-  labs(x="CHUM",y="PINK",z="SOCK") + theme(
-        axis.title=element_text(size=8)) + theme(legend.position = "none")
-
-p0 = group_by(df, year, group) %>% 
-  summarize(n = length(unique(p_holder))) %>% 
-ggplot(aes(x = year, y = n, group=group, col = as.factor(group))) + geom_line() + ylab("People") + ggtitle("S01A")
-
-p1 = group_by(df, year, group) %>% 
-  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
-    cv = sd(logdiff,na.rm=T)) %>% 
-ggplot(aes(x = year, y = meanrev, group=group, col = as.factor(group))) + geom_line() + ylab("Mean revenue") + ggtitle("S01A")
-
-p2 = group_by(df, year, group) %>% 
-  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
-    cv = sd(logdiff,na.rm=T)) %>% 
-ggplot(aes(x = year, y = cv, group=group, col = as.factor(group))) + geom_line() + ylab("CV revenue") + ggtitle("S01A")
-
-gridExtra::grid.arrange(p3, p0, p1, p2, ncol=2)
-```
-
-\pagebreak
-
-# Appendix
-
-## Overall trends in participation across permits
-
-```{r, fig.cap="Participation (unique people)", fig.pos="placeHere", echo=FALSE}
-appendixPlot1
 ```
@@
 
-```{r, fig.cap="Trends in revenue and variability across people", fig.pos="placeHere", echo=FALSE, fig.height = 3.1}
-group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], p_holder, year, p_fshy) %>%
-  summarize(gearn = sum(g_earn)) %>%
-  group_by(year, p_fshy) %>%
-  summarize(med = median(gearn), low = quantile(gearn, 0.25),
-    upp = quantile(gearn,0.75)) %>%
-  ggplot(aes(year, log(med))) + geom_line() + facet_wrap(~p_fshy,scale="free") + 
-  geom_ribbon(aes(ymin= log(low), ymax= log(upp)), alpha=0.4) + ylab("Ln quartiles of revenue (across people)") + ggtitle("Trends in revenue") + xlab("Year")
-```
+```{r, fig.pos="placeHere", fig.cap="Time varying effects of adding 1 species on revenue (measured in ln dollars) for hypothetical individuals with species diversity ~ 1.
+This shows declining benefits of diversification for S01A, S01E, S03A, increasing (but more uncertain) benefits for sockeye oriented fisheries S01M and S01K.", echo=FALSE}
+# grab original data to look at permit names
+diffdat = readRDS(file="../../salmon/data-generated/salmon.rds")
+years = seq(min(diffdat$year), max(diffdat$year))
+permits = levels(as.factor(diffdat$permit))
 
-```{r, echo=FALSE, fig.height = 3.1, fig.cap="Trends in average landed weight per person", fig.pos="placeHere"}
-group_by(cfec, p_fshy, year, p_holder) %>% 
-  summarize(meanW = mean(g_pounds/2000, na.rm=T)) %>% 
-  group_by(p_fshy, year) %>% 
-  summarize(meanW = mean(meanW, na.rm=T)) %>% 
-  ggplot(aes(year, meanW)) + geom_line() + facet_wrap(~ p_fshy, scale="free") + ylab("Mean total tons / person") + xlab("Year") + ggtitle("Trends in landed weight")
+load("../../salmon/analysis/model-linear.Rdata")
+
+est = extract(mod, permute="TRUE")
+
+# start with making time series of estimates on mean
+# challenge is that they're time varying quadratic functions
+# do facet_wrap of estimates by year
+
+# effect of adding more 1 species from 1. specdiv.prev = 1
+
+specdiv.prev = seq(1, 5, by=0.1)
+specdiv = specdiv.prev + 1 # always add 1 more in these comparisons
+x1 = log(specdiv/specdiv.prev)
+x2 = log(specdiv/specdiv.prev) * specdiv.prev
+x3 = log(specdiv/specdiv.prev) * specdiv.prev^2
+
+df = expand.grid("year"=1:29, "diversity" = seq(1,length(x1)), "permit"=1:6)
+df$permits = permits[df$permit]
+df$years = years[df$year]
+df$x = specdiv.prev[df$diversity]
+df$pred = 0
+df$low = 0
+df$high = 0
+for(i in 1:nrow(df)) {
+  tmp = x1[df$diversity[i]] * (est$b_1[,df$year[i], df$permit[i]]+est$b1_str_yr_mu) +
+    x2[df$diversity[i]] * (est$b_2[,df$year[i], df$permit[i]]+est$b2_str_yr_mu)# +
+    #x3[df$diversity[i]] * (est$b_3[,df$year[i], df$permit[i]]+est$b3_str_yr_mu)
+  df$pred[i] = median(tmp)
+  df$low[i] = quantile(tmp,0.25)
+  df$high[i] = quantile(tmp,0.75)
+}
+
+ggplot(df[df$diversity==1,], aes(x=years,y=pred,group=permits)) +
+  geom_line() +
+  geom_ribbon(aes(ymin=low,ymax=high), alpha=0.4) +
+  facet_wrap(~permits) + xlab("Year") + ylab("Predicted benefit of adding +1 spp (ln $)")
+
 ```
@@
 
-```{r, fig.cap="Diversity, mean revenue and CV (aggregated across years)", fig.pos="placeHere", echo=FALSE}
-gridExtra::grid.arrange(appendixPlot2_p1,appendixPlot2_p2,appendixPlot2_p3, ncol=2)
+```{r, fig.pos="placeHere", fig.cap="Time varying effects of adding 1 species on revenue (measured in ln dollars) for hypothetical individuals with species diversity ~ 1.
+This shows declining benefits of diversification for S01A, S01E, S03A, increasing (but more uncertain) benefits for sockeye oriented fisheries S01M and S01K.", echo=FALSE}
+# grab original data to look at permit names
+diffdat = readRDS(file="../../salmon/data-generated/salmon.rds")
+years = seq(min(diffdat$year), max(diffdat$year))
+permits = levels(as.factor(diffdat$permit))
+
+load("../../salmon/analysis/model-linear.Rdata")
+
+est = extract(mod, permute="TRUE")
+
+# start with making time series of estimates on mean
+# challenge is that they're time varying quadratic functions
+# do facet_wrap of estimates by year
+
+# effect of adding more 1 species from 1. specdiv.prev = 1
+
+specdiv.prev = seq(1, 5, by=0.1)
+specdiv = specdiv.prev + 1 # always add 1 more in these comparisons
+x1 = log(specdiv/specdiv.prev)
+x2 = log(specdiv/specdiv.prev) * specdiv.prev
+x3 = log(specdiv/specdiv.prev) * specdiv.prev^2
+
+df = expand.grid("year"=1:29, "diversity" = seq(1,length(x1)), "permit"=1:6)
+df$permits = permits[df$permit]
+df$years = years[df$year]
+df$x = specdiv.prev[df$diversity]
+df$pred = 0
+df$low = 0
+df$high = 0
+for(i in 1:nrow(df)) {
+  tmp = x1[df$diversity[i]] * (est$g_1[,df$year[i], df$permit[i]]+est$g1_str_yr_mu) +
+    x2[df$diversity[i]] * (est$g_2[,df$year[i], df$permit[i]]+est$g2_str_yr_mu)# +
+  #x3[df$diversity[i]] * (est$b_3[,df$year[i], df$permit[i]]+est$b3_str_yr_mu)
+  df$pred[i] = median(tmp)
+  df$low[i] = quantile(tmp,0.25)
+  df$high[i] = quantile(tmp,0.75)
+}
+
+ggplot(df[df$diversity==1,], aes(x=years,y=pred,group=permits)) +
+  geom_line() +
+  geom_ribbon(aes(ymin=low,ymax=high), alpha=0.4) +
+  facet_wrap(~permits) + xlab("Year") + ylab("Predicted benefit of adding +1 spp (ln $)")
+
 ```
@@
 
-## Species composition
-
-```{r, fig.cap="Proportion of species landed, weighted by earnings", fig.pos="placeHere", echo=FALSE, fig.height = 3.3}
-# plot proportion of species landed by salmon permit
-group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
-  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p)) %>%
-  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() + facet_wrap(~p_fshy) + 
-  ylab("Species % (total) ") + xlab("")
-```
-
-```{r, fig.cap="Effective species diversity across people", fig.pos="placeHere", echo=FALSE, fig.height = 3.3}
-group_by(dat.salmon, year, strategy_permit) %>%
-  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
-  ggplot(aes(year, meanDiv)) + geom_line() + facet_wrap(~strategy_permit, scale="free") + 
-  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Effective diversity (across people)") + xlab("")
-```
-
-```{r, fig.cap="Proportion of total returns to Prince William Sound (hatchery + wild, catch + escapement).
-Years before 1980 and after 2011 aren't included because not all species have data.", fig.pos="placeHere", echo=FALSE, fig.height = 3.3}
-returns = readxl::read_excel("salmon_returns_PWS.xlsx")
-returns = returns[,c("Year","Sockeye_total","Pink_total","Chum_total","Coho","Chinook")]
-
-returns = as.data.frame(returns[which(returns$Year%in%c(1980:2011)),])
-for(i in 1:ncol(returns)) returns[,i] = as.numeric(returns[,i])
-for(i in 1:nrow(returns)) {
-    # normalize
-    returns[i,-1] = returns[i,-1] / sum(returns[i,-1])
-}
-
-melted = reshape2::melt(returns,id.vars="Year")
-names(melted)[which(names(melted)=="variable")]="species"
-ggplot(melted,aes(Year,value, group = species, color=species)) + geom_line() + ylab("Proportion of total returns (by #)")
-```
-
-\break
-
-## Trends in revenue and prices
-
-Across regions, we see lots of synchrony in prices between coho, chum, and pink salmon, with them increasing more quickly than sockeye/Chinook.
-Sockeye/Chinook have been the highest priced historically, with coho > chum > pink.
-For most permit types, median revenue / person has generally increased, especially since early 2000s.
-For some permits, variability appears to have changed through time (quartile range getting smaller).
-
-```{r, fig.cap="Total revenue by area (across people and permits within each)", fig.pos="placeHere", echo=FALSE, fig.height=3.5}
-# plot prices of species landed by salmon permit
-cfec.raw$area = NA
-cfec.raw$area[cfec.raw$p_fshy%in%c("S 01A", "S 03A")] = "Southeast"
-cfec.raw$area[cfec.raw$p_fshy%in%c("S 01E", "S 03E")] = "PWS"
-cfec.raw$area[cfec.raw$p_fshy%in%c("S 01K", "S 04K")] = "Kodiak"
-cfec.raw$area[cfec.raw$p_fshy%in%c("S 01M")] = "Alaska Peninsula"
-
-group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, area) %>%
-  summarize(p = sum(g_earn)) %>% group_by(year, area) %>%
-  ggplot(aes(year, (p)/1e6, group=spec, color = spec)) + geom_line() + facet_wrap(~area, scale="free") +
-  ylab("Revenue (millions) ") + xlab("Year")
-```
-
-\break
-
-## Trends in variability
-
-```{r, fig.cap="Trends in annual variability by fishery", fig.pos="placeHere", echo=FALSE,  fig.height=3.3}
-summary = group_by(cfec.raw[cfec.raw$spec%in%c("CHUM","COHO","CHNK","SOCK","PINK"),], year, p_fshy, spec) %>% 
-  summarize(sumg = log(sum(g_earn)), sumw = log(sum(g_pounds))) %>% 
-  arrange(p_fshy, spec, year) %>% group_by(p_fshy, spec) %>% 
-  mutate(logdiffg = rollapply(c(NA, diff(sumg)), 3, sd, fill=c(NA, NULL, NA)), 
-    logdiffw = rollapply(c(NA, diff(sumw)), 3, sd, fill=c(NA, NULL, NA)))
-
-ggplot(summary, aes(year, logdiffg, color=spec)) + geom_line() + facet_wrap(~ p_fshy, scale="free") + ylab("3 year rolling CV (revenue)") + xlab("Year")
-```
-
-```{r, fig.cap="Trends in annual variability by fishery and species", fig.pos="placeHere",  echo=FALSE, fig.height=3.3}
-g1 = ggplot(summary[which(summary$spec=="PINK" & summary$p_fshy%in%c("S 01A", "S 01E", "S 03A", "S 03E")),], aes(year, logdiffg, color=p_fshy)) + geom_line() + ylab("3 year rolling CV (revenue)") + xlab("Year") + ggtitle("Pink salmon")
-
-g2 = ggplot(summary[which(summary$spec=="CHUM" & summary$p_fshy%in%c("S 01A", "S 01E", "S 03A", "S 03E")),], aes(year, logdiffg, color=p_fshy)) + geom_line() + ylab("3 year rolling CV (revenue)") + xlab("Year") + ggtitle("Chum salmon")
-
-gridExtra::grid.arrange(g1,g2, ncol=1)
-```
-
-```{r, fig.cap="Trends in annual variability by fishery and species", fig.pos="placeHere", echo=FALSE, eval=FALSE}
-g1 = ggplot(summary[which(summary$spec=="PINK" & summary$p_fshy%in%c("S 01A", "S 01E", "S 03A", "S 03E")),], aes(year, logdiffw, color=p_fshy)) + geom_line() + ylab("3 year rolling CV (weight)") + xlab("Year") + ggtitle("Pink salmon")
-
-g2 = ggplot(summary[which(summary$spec=="CHUM" & summary$p_fshy%in%c("S 01A", "S 01E", "S 03A", "S 03E")),], aes(year, logdiffw, color=p_fshy)) + geom_line() + ylab("3 year rolling CV (weight)") + xlab("Year") + ggtitle("Chum salmon")
-
-gridExtra::grid.arrange(g1,g2, ncol=1)
-```
-
-```{r, fig.cap="Trends in variability across people", fig.pos="placeHere", echo=FALSE, fig.height=3.5, eval=FALSE}
-summary = group_by(cfec.raw[cfec.raw$spec%in%c("CHUM","COHO","CHNK","SOCK","PINK"),], year, p_fshy, p_holder) %>% 
-  summarize(sumg = sum(g_earn), sumw = sum(g_pounds)) %>% 
-  group_by(p_fshy, year) %>% 
-  summarize(cvg = sd(sumg,na.rm=T)/mean(sumg,na.rm=T),
-    cvw = sd(sumw,na.rm=T)/mean(sumw,na.rm=T))
-ggplot(summary, aes(year, cvg )) + geom_line(color ="green" ) + facet_wrap(~ p_fshy, scale="free") + geom_line(aes(year, cvw), color = "blue") + ylab("3 year rolling CV") + xlab("Year") + ggtitle("Variability across people: total revenue (green) and weight (blue)")
-```
-
-```{r, fig.cap="Change in CV for individual people", fig.pos="placeHere", echo=FALSE, fig.height=3.5}
-summary = group_by(cfec.raw[cfec.raw$spec%in%c("CHUM","COHO","CHNK","SOCK","PINK"),], year, p_fshy, p_holder) %>% 
-  summarize(sumg = sum(g_earn)) %>% 
-  group_by(p_fshy, p_holder) %>% 
-  mutate(nPre = length(which(year%in%seq(1985,1995))),
-    nPost = length(which(year%in%seq(2005,2014)))) %>% 
-  filter(nPre > 3 & nPost > 3) %>% select(-nPre, -nPost) %>% 
-  summarize(cv.pre = mean(sumg[which(year <= 1995)],na.rm=T)/sd(sumg[which(year <= 1995)],na.rm=T),
-    cv.post = mean(sumg[which(year >= 2005)],na.rm=T)/sd(sumg[which(year >= 2005)],na.rm=T))
-
-ggplot(summary, aes(log(cv.post/cv.pre))) + geom_histogram(bins=100) + facet_wrap(~ p_fshy) + geom_vline(aes(xintercept = 0)) + ylab("Count") + xlab("% Change in CV for individual people (2005-2014 / 1985-1995)") + ggtitle("Change in variability for individual people")
-#summary = left_join(expand.grid(year = unique(summary$year), p_holder = unique(summary$p_holder)), summary)
-
-# add rolling CV for each person
-#summary = group_by(summary, p_holder) %>% 
-#  mutate(logdiffg = rollapply(c(NA, diff(sumg)), 3, sd, fill=c(NA, NULL, NA))) %>%
-#  do(fit = lm(logdiffg ~ year, data = .))
-```
-
-\break
-
-```{r, fig.cap="Inflation adjusted prices / species by area, relative to 2002 (the year corresponding to the lowest average prices).", fig.pos="placeHere", echo=FALSE, fig.height = 3.1}
-
-g = group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, area) %>%
-  summarize(p = mean(g_price))
-p1 = group_by(g, spec, area) %>% 
-  mutate(p = p/p[which(year==2002)]) %>% 
-  group_by(year, area) %>%
-  ggplot(aes(year, log(p), group=spec, color = spec)) + geom_line() + facet_wrap(~area) +
-  ylab("Ln avg price relative to 2002") + ggtitle("Synchrony in prices: relative change") + xlab("")
-p1
-
-```
-
-```{r, fig.cap="Percent change (log-difference) in inflation adjusted prices / species by area", fig.pos="placeHere", echo=FALSE, fig.height = 3.1}
-
-g = group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, area) %>%
-  summarize(p = mean(g_price))
-# do log-diff of prices -- get price year before
-g$p_t1 = g$p[match(paste(g$year, g$spec, g$area), 
-  paste(g$year+1, g$spec, g$area))]
-
-p2 = group_by(g, spec, area) %>% 
-  ggplot(aes(year, log(p/p_t1), group=spec, color = spec)) + geom_line() + facet_wrap(~area) +
-  ylab("Ln price (t/t-1)") + ggtitle("Synchrony in prices: percent change") + xlab("")
-p2
-
-g$logdiff = log(g$p/g$p_t1)
-p3 = group_by(g, year, area) %>% 
-summarize(sd = sd(logdiff,na.rm=TRUE)) %>% ggplot(aes(year, sd, color=area)) + geom_line() + xlab("Year") + ylab("SD of price changes across species") + ggtitle("Synchrony in price fluctuations across species")
-
-```
-
-```{r, fig.pos="placeHere", fig.cap = "Standard deviation (across species, by area) for fluctuations in prices.
-Fluctuations are calculated as the log difference price data between sequential years.
-The standard deviation of log differenced data is interpreted as the coefficient of variation (CV)."}
-p3
-```
-
-\break
-
-### Season length
-
-There appears to be no significant trend in season length (defined as the difference between 95% CIs) or mean julian day for any fishery.
-In all cases, the dates are weighted by the earnings associated with landed catch.
-
-```{r, fig.cap="Trends in mean julian day and season length by fishery", fig.pos="placeHere", echo=FALSE, fig.height=3.5}
-gridExtra::grid.arrange(appendixPlot0_p1, appendixPlot0_p2, ncol = 1)
-```
+# References

### 2016-10-05 salmon/paper/salmon_paper_appendix.Rmd
salmon update

--- 2016-10-05 salmon/archive/salmon_paper_combined.Rmd
@@
 ---
-title: "Reduced diversity and increased specialzation of Alaska salmon fishermen: effects on variation in individual revenues"
+title: ""
 author: ''
@@
 ```
-
-## Abstract
-\setlength\parindent{24pt}
-
-On a large scale, individual fishers may experience positive outcomes from diversifying their catches; examples include reduced year-to-year variability in annual revenue, and increases in average revenue.
-Fishers can increase their species diversity by owning a permit that allows multiple species to be targeted, or by owning multiple permits.
-External forces such as management regulations (limited entry,  catch shares) or market forces (permit prices) may disincentive individuals to hold multiple permits.
-Thus, individuals with a single permit may increase species diversity within the confines of their permit (longer fishing seasons, broader spatial area).
-Using a long-term dataset of individual salmon fishers in the Alaska, we develop new models for understanding how the effects of diversification have changed through time.
-These changes allow us to compare patterns across regions and gear, as well as better understand how external forces have changed incentives for diversifying (hatchery production and price trends, EVOS?).
-
-## Introduction
-
-[broader paragraph on diversification with extensions to terrestrial agriculture etc, ]
-
-Previous analyses have largely focused on patterns of diversification across ecosystems and the benefits to adopt more or less diverse fishing strategies (Kasperski and Holland 2013; Anderson et al.
-in review).
-Fisheries in Alaska have been studied extensively, because they are one of the most productive regions in North America (both by landed weight, and revenue cite) and they allow individuals to adopt a wide range of fishing strategies.
-Anderson et al.
-(in review) defined strategies as a permit of collection of permits held by single individuals.
-While adding new permits allows fishers to increase species diveristy, few individual permits allow individuals to harvest more than one species (examples include halibut, sablefish, herring, crab, shellfish, shrimp, sea cucumber).
-Exceptions to these single species permits incldue participants in federally managed groundfish fisheries (CFEC 'M' permits) and participants in state managed salmon fisheries (CFEC 'S' permits).
-
-Understanding the effects of species diversification on salmon fishers is particularly important, as the number of salmon specialists (defined as fishers holding a single salmon permit) has increased over the last 30 years as a fraction of the total (Anderson et al.
-in review).
-For these limited entry salmon fisheries, general trends that have emerged are that more people are participating since the early 2000s (Fig.
-A1), median revenue is increasing (across individuals, Fig.
-A2), and landed species diversity declining over time (averaged across indivduals, Fig.
-A6).
-Across different salmon fisheries, other than a slight negative effect of diversification on variability (similar to Anderson et al.
-in review), there appears to be little correlation between diversity and revenue (Fig.
-A4).
-In terrestrial agriculture, there are numerous examples where diversification may affect the mean or CV but not both (Purdy, Barry M; Langemeier, Michael R; Featherstone, Allen M.
-Journal of Agricultural and Applied Economics29.1 (Jul 1, 1997): 149.), and the same result may be true for salmon fishers.
-Further, these impacts of diversification may be variable over time.
-
-Five species of salmon are encountered in Alaska waters (Chinook, chum, pink, sockeye, coho), and the ability of individual salmon fishers to diversify across species is in part a function of geography and gears.
-For example, salmon may be caught with purse seine gear (n = 6 areas), drift gillnets (5 areas), and set gillnets (12 areas).
-Other gears include beach seine and troll gears (though these are generally restricted to Kodiak and Southeast Alaska, respectively).
-While some of these combinations of gears and areas allow fishers to diversify, there are many examples of permits where the catch is dominated by a single species.
-Perhaps the best examples include fisheries that target sockeye salmon (e.g.
-Bristol Bay set and drift gillnet fisheries).
-For these fisheries, specialization is associated with higher revenue and reduced CV in annual revenue (Anderson et al.
-in review).
-
-A secondary motivation for focusing on the effects of diversification for salmon fishers is to better understand the long term consequences of hatchery salmon production on the revenue and variability in revenue of individuals.
-Following low returns and declining productivity in the mid 1980s, many hatchery programs in the state rapidly increased hatchery releases (Mahnken et al.
-1998).
-Many of these hatchery programs increased production to stabilize the variability of fishers catch (and revenue), though these programs may also have costs to wild production (Ruggerone et al.
-2015).
-[Other history here?]
-
-In addition to the changing species composition of fish avaialble to be harvested, the benefits (or costs) of diversification are likely affected by a number of exogenous drivers, including marketing, prices, new processing technology, etc (cite Gunnar Knapp + others).
-[Break out as own paragraph+add more examples?] Following price crashes in the early 2000s, many of these factors have been positive for salmon fishermen (increasing trends in prices and catches, Fig.
-A3, A12).
-In contrast, the late 1980s - early 1990s experienced high abundance and low prices (in part due to collapses of foreign currencies).
-Combined with the increased trend in hatchery production and resulting change species compositon of adult salmon, these external factors may have also altered incentives for diversification.
-If there's an incentive to substitute high value species (Chinook, sockeye) with abundant low value species (pink, chum), the analog in the terrrestrial framring world would be crop substitution of low-volume high value crops (spices) with high-volume low value crops (cereals) (other examples in literature = cocoa, tobacco).
-
-The objective of our analysis is to better understand the effects of diversification on single-permit salmon holders by extending the methods used by Anderson et al.
-(in review) to include time-varying effects.
-These time-varying effects will allow us to compare temporal trends across regions, identify periods when having a diverse catch portfolio was most beneficial, and allow us to evaluate support for effects of longer term drivers in the region.
-
-## Methods
-
-### Data
-
-We obtained fisheries landings and revenue data for all permit holders in Alaska from 1985 - 2014 from the Alaskan Commercial Fisheries Entry Commission (CFEC).
-We adjusted for inflation by setting all revenues in 2009 USD.
-Because were were only interested in single permit fisheries, we adopted a simplified version of the filtering steps described in  Anderson et al.
-(in review).
-After aggregating individual fish tickets to annual data, we retained individuals who only fished one of the six salmon permits included in a year (Table 1, Fig.
-A5).
-Like Anderson et al.
-(in review) we also worked with a log-differenced dataset to remove non-stationary individuals; to avoid complications of individuals switching strategies, we only included pairs of years for a given person where they did not change strategies.
-
-```{r, echo=FALSE}
-m = matrix("", 6, 3)
-m[1,] = c("S01A", "Purse Seine", "Southeast Alaska")
-m[2,] = c("S01E", "Purse Seine", "Prince William Sound")
-m[3,] = c("S01K", "Purse Seine", "Kodiak")
-m[4,] = c("S01M", "Purse Seine", "Alaska Peninsula")
-m[5,] = c("S03A", "Gillnet", "Southeast Alaska")
-m[6,] = c("S03E", "Gillnet", "Prince William Sound")
-colnames(m) = c("Permit", "Gear", "Region")
-kable(m)
-```
-
-### As hatchery programs have increased production, have individual salmon fishermen experienced reduced variability?
-
-Hatchery release trends for some regions included in our data have been summarized previously (e.g.
-Ward et al.
-in review summarized the trends for Prince William Sound).
-Release patterns have been dominated by pink and chum salmon (for PWS, these species represent 80% and 18% of total releases for 2012-2014 respectively, Ward et al.
-in review) and production has increased since the late 1970s (217 million juvenile pink salmon in PWS in 1985 versus 672 million in 2014).
-Though our analysis doesn't include , as a contrast we can do coarse comparisons across regions and species to examine potential effects of hatchery fish on variability.
-For example, we can compare the variation in pink salmon returns (earnings or weight) from Prince William Sound (largely hatchery) to those from Southeast Alaska (largely wild).
-Similarly, we can compare the variability in earnings from species dominated by hatchery returns (pink, chum) to those with less hatchery influence (Chinook, sockeye).
-To make these comparisons, we aggregated the earnings and landings by species and permit, then calculated 3-year rolling CVs.
-As a second comparison, we examined whether long-term participants in any of the fisheries had experienced a change in variability.
-Similar to Kasperski and Holland (2013) who calculated CVs of vessels before and after significant managment events, we identified individuals who fished the same permit for at least three years in the first decade of our data (1985-1995) and at least three years in the last decade of the data (2004-2014).
-
-### Robustness of strategies to changing environments [stash in appendix?]
-
-One of the challenges in modeling the effects of species diversification for individual fishers is that between two years, both the species composition of the catch and the species composition of the fish available to be caught may change (making it hard to disentangle the selectivity of fishermen versus portfolio of fish available).
-To isolate the effects of a changing environment from changing strategies, we used a subset of the data from fishery permits where discrete strategies were emergent.
-
-Using purse seine permit holders in Prince William Sound as an example, we first calculated the species composition (weighted by revenue) for each active permit holder in each year.
-For each pair of years that an individual fished, we calculated the squared Euclidian distance between the species compositions of those years, $$\Delta { p }_{ ind,t }=\sum _{ sp=1 }^{ 5 }{ { \left| { p }_{ ind,sp,t }-{ p }_{ ind,sp,t-1 } \right|  }^{ 2 } } $$.
-
-Larger values of $\Delta { p }_{ ind,t }$ indicate greater year-to-year discrepancy in species composition.
-To focus on individuals who had more constant strategies in adjacent years, we ommitted data corresponding to the upper third of this distribution (retaining $\Delta { p }_{ ind,t }$ < 0.334).
-Next, we used the species composition from each individual and pair of years to cluster individuals into strategies.
-Clustering was performed using the partitioning around medoids (pam) algorithm in R with an estimated number of clusters (using 'pamk' in fpc library), and selected 2 groups of individual strategies: pink salmon specialization (medoid proportion of pink salmon = 0.95), and a strategy focused on pink salmon and chum (medoid proportions of 0.64 and 0.31 respectively).
-For each group, we then calculated the mean and CV of revenue; we calculated the log difference of revenue for each individual, $\Delta { R }_{ i }=log\left( { R }_{ i,t+1 }/{ R }_{ i,t } \right)$ and then calculated the CV as the standard deviation over people, $sd(\Delta { R }_{ i=1:n })$.
-
-### Analysis 3: [Core of paper] Does effect of diversification change over time?
-
-Anderson et al.
-(in review) implemented a hierarchical Bayesian model that allowed group level predictors and covariates (diversity of catch, days fished) to simultaneously affect the mean and the variance of revenue for individual fishers.
-This model also included random intercepts by strategy and year; these terms allowed all permit holders in a given year to experience common fluctuations (external drivers including price changes, fluctuations in fish population abundance, or environmental disasters such as EVOS).
-
-To ask whether how benefits of diversification may have changed over time, we extend the model of Anderson et al.
-(in review) to also include time-varying random effects in the slope coefficients of diversification.
-We define the variables $\Delta { S }_{ t }=log\left( { S }_{ t }/{ S }_{ t-1 } \right)$ to represent the percent change in species diversity of the catch between years and $\Delta { D }_{ t }=log\left( { D }_{ t }/{ D }_{ t-1 } \right)$ to represent the percent change in fishing effort (days fished) between years.
-The model for the the expected log revenue of individual *i* in year *t*, $log({ R }_{ i,t })$, can be written as $$\widehat { log\left( { R }_{ i,t } \right)  } ={ B }_{ 0,j,t }+{ B }_{ 1,j,t }\cdot \Delta S_{ t }+{ B }_{ 2,j,t }\cdot \left( \Delta S_{ t }\cdot { S }_{ t-1 } \right) +{ B }_{ 3,j,t }^{ 2 }\cdot \left( \Delta S_{ t }\cdot { S }_{ t-1 }^{ 2 } \right) +{ B }_{ 4,j }\cdot \Delta D_{ t }+log\left( { R }_{ i,t-1 } \right) $$, where *j* subscripts the salmon permit (Table 1).
-We model the time varying permit specific intercepts as random and exchangeable across years, ${ B }_{ 0,j,t }\sim Normal\left( { u }_{ { B }_{ 0 } },{ \sigma  }_{ { B }_{ 0 } } \right)$.
-The term ${ B }_{ 1,j,t }$ represents the permit level random effect of diversifying species.
-Our model extendeds that of Anderson et al.
-(in review) to include the interaction terms $\Delta S_{ t }\cdot { S }_{ t-1 }$ which allow the effect of diversification to be dependent on the previous level of diversification.
-The squared term allows this interaction to be quadratic, and the coefficients ${ B }_{ 2,j,t }$ and ${ B }_{ 3,j,t }$, allow this to vary by permit and through time.
-The flexibility in these terms allows the benefit of increasing species diversity to be positive in some years (generalizing benefits revenue), negative in other years (specializing benefits revenue) and neutral in other years (revenue is independent alterations to catch composition).
-We constrain the random effects associated with diversification to be modeled as autoregressive processes with shared variances across strategies, for example: ${ B }_{ 1,j,t }\sim Normal\left( { B }_{ 1,j,t-1 },{ \sigma  }_{ { B }_{ 1 } } \right)$.
-Like Anderson et al.
-(in review) we also allow changing effort (measured as percent change in days fished) ${ B }_{ 4,j }$ to vary by strategy, but we do not allow this effect to vary through time.
-Finally, the term $log\left( { R }_{ i,t-1 } \right)$ is included as an offset (Anderson et al.
-in review).
-
-- Fit models from Sean's paper, but include time varying interactions on species diversity.  GLMM code below
-
-Specific things that could be looked at: 
-- EVOS (closure for Kodiak)  
-- PWS pink salmon crash in 1992/1993: Pacific Salmon & their Ecosystems: Status and Future Options
-
-Does diversification exist within permits and if so, do we see similar advantages of diversification on revenue + CV?
-
-## Results
-
-Our summaries of changing variability (aggregate for each fishery, across individuals, or within individuals) was not consistent with the hypothesis that variability in revenue or landings has decreased over time (Fig.
-A9-A10).
-For pink and chum salmon, fisheries in PWS generally experience higher year to year variability compared to SEAK.
-In looking at the changing CV of individuals, we see a slight increase in the CV of revenue for people fishing in the recent decade who were also fishing 1985-1995 (with the exception being S01A).
-
-## Discussion
-
-- Revenue is going up since 2000 -- but each permit holder is fishing more (more days, more total pounds) and the trend in weight is faster than the trend in price.
-
-- Ability to specialize primarily driven by many forces beyond control of fishermen.
-But in some situations, diversification is possible by extending fishing season to exploit additional spp (fishing earlier for chum in PWS S01E).
-Or fishing different areas to exploit more spp in S03 (SEAK).
-
-- Prices are generally synchronous for all spp (especially coho/chum/pink), and have increased at a faster rate compared to chum and pink salmons since 2000 (Fig A8-A9).
-We're also seeing a trend toward greater synchrony between species (Fig A11-A12, A14).
-Projecting future prices are complex -- Gunnar Knapp has a good presentation on this http://www.adfg.alaska.gov/static/fishing/PDFs/commercial/gk_trends_4-12-12.pdf.
-Synchrony in prices may largely be function of marketing, improved product, freezing, etc.
-
-- We're seeing increased specialization within permits.
-People in PWS are trending toward increased specialization on pink salmon.
-People in SEAK trending toward increased specialization on chum.
-Both of these are switches from higher value species (sockeye, chum) to lower value species.
-Both pink and chum have higher variability in year to year returns than sockeye.
-
-- One pontential exogenous driver related to management that can be ruled out as being responsible for increased specialization of salmon fishermen is changes in the timing or length of fishing seasons.
-Shifting seasons to later in the year may reduce catch of early run fish for example.
-However, we found no significant trend in either the season length or average season date for any fishery (Fig.
-A15).
-
-\break
 

### 2016-10-05 salmon/analysis/1-run-model.R

@@ -1,0 +1,55 @@
+
+diffdat = readRDS(file="salmon/data-generated/salmon.rds")
+
+# prep data for stan
+N = nrow(diffdat) # number of rows of data
+#X_ij = diffdat[,c("days.change")] #"specDiv.change","specdiv.prev",
+#if(class(X_ij)=="numeric") X_ij = matrix(X_ij,ncol=1)
+J = ncol(X_ij) # number of covariates on mean
+X_sigma_ik = diffdat[,c("days.change")] #"specDiv.change","specdiv.prev",
+if(class(X_sigma_ik)=="numeric") X_sigma_ik = matrix(X_sigma_ik,ncol=1)
+K = ncol(X_sigma_ik) # number of covariates on variance
+
+offset = diffdat$revenue.prev
+y_i = diffdat$revenue
+n_strategy = length(unique(diffdat$permit))
+strategy_i = as.numeric(as.factor(diffdat$permit))
+n_str_yr = length(unique(diffdat$year_permit))
+str_yr_i = as.numeric(as.factor(diffdat$year_permit))
+n_yr = length(unique(diffdat$year))
+year_i = diffdat$year - min(diffdat$year) + 1
+
+b1_cov_i = diffdat$specDiv.change # log(percent change in diversity)
+b2_cov_i = diffdat$specDiv.change * diffdat$specdiv.prev # log(percent change in diversity)
+b3_cov_i = diffdat$specDiv.change * (diffdat$specdiv.prev^2)
+b4_cov_i = diffdat$days.change
+
+# create list for STAN
+stan_data = list("N"=N, "y_i"=y_i, "offset"=offset, "n_strategy"=n_strategy,
+  "n_str_yr"=n_str_yr,"str_yr_i"=str_yr_i, "n_yr"=n_yr, "year_i"=year_i,
+  "b1_cov_i"=b1_cov_i, "b2_cov_i"=b2_cov_i, "b3_cov_i"=b3_cov_i, "b4_cov_i"=b4_cov_i,
+  "g1_cov_i"=b1_cov_i, "g2_cov_i"=b2_cov_i, "g3_cov_i"=b3_cov_i, "g4_cov_i"=b4_cov_i)
+stan_pars = c("b0_str_yr", "b0_str_yr_tau", "g0_strategy", "g0_strategy_tau",
+  "b_1", "b_2", "b_3", "b_4", "b1_str_yr_tau", "b2_str_yr_tau",
[…]

### 2016-10-05 salmon/analysis/2-make-plots.R

@@ -1,0 +1,82 @@
+library(rstan)
+
+# grab original data to look at permit names
+diffdat = readRDS(file="salmon/data-generated/salmon.rds")
+years = seq(min(diffdat$year), max(diffdat$year))
+permits = levels(as.factor(diffdat$permit))
+
+yearpermit = data.frame("yrpermit"=levels(diffdat$year_permit))
+yearpermit$year = as.numeric(substr(yearpermit$yrpermit,1,4))
+yearpermit$permit = substr(yearpermit$yrpermit,6,9)
+
+load("salmon/analysis/model-linear.Rdata")
+
+est = extract(mod, permute="TRUE")
+
+yearpermit$b0 = apply(est$b0_str_yr,2,median)
+yearpermit$b0_low = apply(est$b0_str_yr,2,quantile,0.25)
+yearpermit$b0_hi = apply(est$b0_str_yr,2,quantile,0.75)
+
+ggplot(data=yearpermit) +
+  geom_ribbon(aes(x=year,ymin=b0_low,ymax=b0_hi),alpha=0.3) +
+  geom_line(aes(x=year,y=b0)) +
+  facet_wrap(~permit) + xlab("Year") + ylab("Intercepts on mean")
+
+# start with making time series of estimates on mean
+# challenge is that they're time varying quadratic functions
+# do facet_wrap of estimates by year
+
+# effect of adding more 1 species from 1. specdiv.prev = 1
+
+specdiv.prev = seq(1, 5, by=0.1)
+specdiv = specdiv.prev + 1 # always add 1 more in these comparisons
+x1 = log(specdiv/specdiv.prev)
+x2 = log(specdiv/specdiv.prev) * specdiv.prev
+x3 = log(specdiv/specdiv.prev) * specdiv.prev^2
+
+df = expand.grid("year"=1:29, "diversity" = seq(1,length(x1)), "permit"=1:6)
+df$permits = permits[df$permit]
+df$years = years[df$year]
+df$x = specdiv.prev[df$diversity]
+df$pred = 0
+df$low = 0
+df$high = 0
+for(i in 1:nrow(df)) {
+  tmp = x1[df$diversity[i]] * (est$b_1[,df$year[i], df$permit[i]]+est$b1_str_yr_mu) +
[…]

### 2016-10-05 salmon/analysis/0-process-data.R
salmon

@@ -1,0 +1,26 @@
+library(dplyr)
+revenueThreshold = 10000
+
+dat.diff = readRDS(file="portfolio/data-generated/cfec-diff-for-modeling.rds")
+
+length(unique(dat.diff$p_holder))
+
+dat.diff = group_by(dat.diff, p_holder) %>%
+  mutate(medRev = median(revenue, na.rm=T)) %>%
+  filter(medRev >= revenueThreshold) %>%
+  select(-medRev)
+
+length(unique(dat.diff$p_holder))
+
+dat.salmon = dat.diff[dat.diff$strategy_permit%in%c("S01A", "S01E", "S01K", "S03A", "S03E","S01M"),]
+length(unique(dat.salmon$p_holder))
+
+# Restrict analysis to people who keep same strategy - complicates days, etc
+dat.salmon = dat.salmon[which(dat.salmon$strategy_permit==dat.salmon$strategy_permit.prev),]
+
+dat.salmon$days.change = log((dat.salmon$days_permit+1) / (dat.salmon$days_permit.prev+1))
+dat.salmon$year_permit = as.factor(paste(dat.salmon$year, dat.salmon$strategy_permit))
+
+length(unique(dat.salmon$p_holder))
+
+saveRDS(dat.salmon, file="salmon/data-generated/salmon.rds")

### 2016-10-05 salmon/paper/salmon_paper.Rmd
paper change

@@
   effects on variation in individual revenues'
-author: ''
+author: "Eric, Sean, Ole, Rich, Milo, Jordan, others?"
 date: ''
 output:
-  word_document: default
   pdf_document:
@@
   html_document: default
+  word_document: default
 bibliography: refs.bib
@@
 Fisheries in Alaska have been studied extensively, because they are one of the most productive regions in North America (both by landed weight, and revenue [@office_of_science_and_technology_fisheries_2016]) and they allow individuals to adopt a wide range of fishing strategies.
-[@anderson_benefits] defined strategies as a permit of collection of permits held by single individuals.
-While adding new permits allows fishers to increase species diveristy, few individual permits allow individuals to harvest more than one species (examples of single-species permits in the Gulf of Alaska include halibut, sablefish, herring, crab, shellfish, shrimp, sea cucumber).
+[@anderson_benefits] defined strategies as a permit or collection of permits held by single individuals.
+While adding new permits allows fishers to increase species diversity, few individual permits allow individuals to harvest more than one species (examples of single-species permits in the Gulf of Alaska include halibut, sablefish, herring, crab, shellfish, shrimp, sea cucumber).
 Some exceptions to these single species permits include participants in federally managed groundfish fisheries and participants in state managed salmon fisheries.
@@
 
-Five species of salmon are encountered in Alaska waters (Chinook, chum, pink, sockeye, coho), and the ability of individual salmon fishers to diversify across species is in part a function of geography and gears.
+Five species of salmon are encountered in Alaska waters (Chinook [*Oncorhynchus tshawytscha*], chum [*O.
+keta*], pink [*O.
+gorbuscha*], sockeye [*O.
+nerka*], coho [*O.
+kisutch*]), and the ability of individual salmon fishers to diversify across species is in part a function of geography and gears.
 For example, salmon may be caught with purse seine gear (n = 6 areas), drift gillnets (5 areas), and set gillnets (12 areas).
@@
 
+[Jordan: Maybe a paragraph or notes about the importance of the role that salmon are serving in buffering or stabilizing fishing revenues in the state and how subsequently, understanding the responses of salmon to climate dynamics is especially critical (Mueter et al.
+2002, and maybe Schindler et al., 2010)]
+
 In addition to the changing species composition of fish avaialble to be harvested, the benefits (or costs) of diversification are likely affected by a number of exogenous drivers, including marketing, prices, new processing technology [@knapp_trends].
@@
 In contrast, the late 1980s - early 1990s experienced high abundance and low prices (in part due to collapses of foreign currencies).
-Combined with the increased trend in hatchery production and resulting change species compositon of adult salmon, these external factors may have also altered incentives for diversification.
-If there's an incentive to substitute high value species (Chinook, sockeye) with abundant low value species (pink, chum), the analog in the terrestrial framring world would be crop substitution of low-volume high value crops (spices) with high-volume low value crops (cereals) (other examples in literature = cocoa, tobacco).
+Combined with the increased trend in hatchery production and resulting changes in species compositon of adult salmon, these external factors may have also altered incentives for diversification.
+If there's an incentive to substitute high value species (Chinook, sockeye) with more abundant low value species (pink, chum), the analog in the terrestrial framring world would be crop substitution of low-volume high value crops (spices) with high-volume low value crops (cereals) (other examples in literature = cocoa, tobacco).
 
 The objective of our analysis is to better understand the effects of diversification on single-permit salmon holders by extending the methods used by [@anderson_benefits] to include time-varying effects.
-These time-varying effects will allow us to compare temporal trends across regions, identify periods when having a diverse catch portfolio was most beneficial, and allow us to evaluate support for effects of longer term drivers in the region.
+These time-varying effects will allow us to compare temporal trends across regions, identify periods when having diverse catch portfolio was most beneficial, and allow us to evaluate support for effects of longer term drivers in the region.
 
@@
 
-We obtained fisheries landings and revenue data for all permit holders in Alaska from 1985 - 2014 from the Alaskan Commercial Fisheries Entry Commission (CFEC).
+We obtained fisheries landings and revenue data for all permit holders in Alaska from 1985 - 2014 from the Commercial Fisheries Entry Commission (CFEC).
 We adjusted for inflation by setting all revenues in 2009 USD.
@@
 Release patterns have been dominated by pink and chum salmon (for PWS, these species represent 80% and 18% of total releases for 2012-2014 respectively, [@ward_evaluating]) and production has increased since the late 1970s (217 million juvenile pink salmon in PWS in 1985 versus 672 million in 2014).
-Though our analysis doesn't include , as a contrast we can do coarse comparisons across regions and species to examine potential effects of hatchery fish on variability.
+Though our analysis with CFEC data doesn't separate hatchery versus wild fish, as a contrast we can make coarse comparisons across regions and species to examine potential effects of hatchery fish on variability.
 For example, we can compare the variation in pink salmon returns (earnings or weight) from Prince William Sound (largely hatchery) to those from Southeast Alaska (largely wild).
@@
 To isolate the effects of a changing environment from changing strategies, we used a subset of the data from fishery permits where discrete strategies were emergent.
+For details see Appendix.
 
@@
 
-To ask whether how benefits of diversification may have changed over time, we extend the model of [@anderson_benefits] to also include time-varying random effects in the slope coefficients of diversification.
+To ask how benefits of diversification may have changed over time, we extend the model of [@anderson_benefits] to also include time-varying random effects in the slope coefficients of diversification.
 We define the variables $\Delta { S }_{ t }=log\left( { S }_{ t }/{ S }_{ t-1 } \right)$ to represent the percent change in species diversity of the catch between years and $\Delta { D }_{ t }=log\left( { D }_{ t }/{ D }_{ t-1 } \right)$ to represent the percent change in fishing effort (days fished) between years.
-The model for the the expected log revenue of individual *i* in year *t*, $$\widehat { log\left( { R }_{ i,t } \right)  } ={ B }_{ 0,j,t }+{ B }_{ 1,j,t }\cdot \Delta S_{ t }+{ B }_{ 2,j,t }\cdot \left( \Delta S_{ t }\cdot { S }_{ t-1 } \right) +{ B }_{ 3,j }\cdot \Delta D_{ t }+log\left( { R }_{ i,t-1 } \right) $$, where *j* subscripts the salmon permit (Table 1).
-We model the time varying permit specific intercepts as random and exchangeable across years, ${ B }_{ 0,j,t }\sim Normal\left( { u }_{ { B }_{ 0 } },{ \sigma  }_{ { B }_{ 0 } } \right)$.
+The model for the expected log revenue of individual *i* in year *t*, $$\widehat { log\left( { R }_{ i,t } \right)  } ={ B }_{ 0,j,t }+{ B }_{ 1,j,t }\cdot \Delta S_{ t }+{ B }_{ 2,j,t }\cdot \left( \Delta S_{ t }\cdot { S }_{ t-1 } \right) +{ B }_{ 3,j }\cdot \Delta D_{ t }+log\left( { R }_{ i,t-1 } \right) $$, where *j* subscripts the salmon permit (Table 1).
+We model the time-varying permit specific intercepts as random and exchangeable across years, ${ B }_{ 0,j,t }\sim Normal\left( { u }_{ { B }_{ 0 } },{ \sigma  }_{ { B }_{ 0 } } \right)$.
 The term ${ B }_{ 1,j,t }$ represents the permit level random effect of diversifying species.
-Our model extendeds that of [@anderson_benefits] to include the interaction terms $\Delta S_{ t }\cdot { S }_{ t-1 }$ which allow the effect of diversification to be dependent on the previous level of diversification.
+Our model extends that of [@anderson_benefits] to include the interaction terms $\Delta S_{ t }\cdot { S }_{ t-1 }$ which allow the effect of diversification to be dependent on the previous level of diversification.
 We assume this relationship is linear, and the coefficients ${ B }_{ 1,j,t }$ and ${ B }_{ 2,j,t }$ allow this to vary by permit and through time.
-The flexibility in these terms allows the benefit of increasing species diversity to be positive in some years (generalizing benefits revenue), negative in other years (specializing benefits revenue) and neutral in other years (revenue is independent alterations to catch composition).
+The flexibility in these terms allows the benefit of increasing species diversity to be positive in some years (generalizing benefits revenue), negative in other years (specializing benefits revenue) and neutral in other years (revenue is independent of alterations to catch composition).
 We constrain the random effects associated with diversification to be modeled as autoregressive processes with shared variances across strategies, for example: ${ B }_{ 1,j,t }\sim Normal\left( { B }_{ 1,j,t-1 },{ \sigma  }_{ { B }_{ 1 } } \right)$.
@@
 - PWS pink salmon crash in 1992/1993: Pacific Salmon & their Ecosystems: Status and Future Options
+- sockeye marketing 1983: [Jordan: So the Copper River salmon marketing ramped up in about 1983.
+I wonder if we could see a steady increase in value or movement towards sockeye in the early years of the data set]
 
@@
 df$high = 0
+b0_indx = match(paste(as.character(df$years), as.character(df$permits)), as.character(yearpermit$yrpermit))
+
 for(i in 1:nrow(df)) {
-  tmp = x1[df$diversity[i]] * (est$b_1[,df$year[i], df$permit[i]]+est$b1_str_yr_mu) +
+  tmp = est$b0_str_yr[,b0_indx[i]] + 
+    x1[df$diversity[i]] * (est$b_1[,df$year[i], df$permit[i]]+est$b1_str_yr_mu) +
     x2[df$diversity[i]] * (est$b_2[,df$year[i], df$permit[i]]+est$b2_str_yr_mu)# +
@@
 
-```{r, fig.pos="placeHere", fig.cap="Time varying effects of adding 1 species on revenue (measured in ln dollars) for hypothetical individuals with species diversity ~ 1.
-This shows declining benefits of diversification for S01A, S01E, S03A, increasing (but more uncertain) benefits for sockeye oriented fisheries S01M and S01K.", echo=FALSE}
+```{r, fig.pos="placeHere", fig.cap="Time varying effects of adding 1 species on log standard deviation of revenue for hypothetical individuals with species diversity ~ 1.", echo=FALSE}
 # grab original data to look at permit names
@@
 df$high = 0
+
 for(i in 1:nrow(df)) {
-  tmp = x1[df$diversity[i]] * (est$g_1[,df$year[i], df$permit[i]]+est$g1_str_yr_mu) +
+  tmp = est$g0_strategy[,df$permit[i]] + x1[df$diversity[i]] * (est$g_1[,df$year[i], df$permit[i]]+est$g1_str_yr_mu) +
     x2[df$diversity[i]] * (est$g_2[,df$year[i], df$permit[i]]+est$g2_str_yr_mu)# +

### 2016-10-05 salmon/paper/salmon_paper.Rmd
simple model

@@
 
-```{r, echo=FALSE, fig.pos="placeHere", fig.cap="Time varying intercepts on estimated mean revenue, by permit."}
+\break
+
+```{r, echo=FALSE, fig.pos="placeHere", fig.cap="Time varying species composition and diversity, by permit.
+Species are colored as follows: yellow (chum), pink (sockeye), red (Chinook), green (coho), blue (pink salmon).", fig.height=3.2}
+cfec <- feather::read_feather("../../portfolio/data-generated/salmon.feather")
+cfec$year = as.numeric(substr(cfec$landdate, 1, 4))
+
+library(dplyr)
+# plot proportion of species landed by salmon permit
+g1 = group_by(cfec[cfec$p_fshy=="S 01A" & cfec$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
+  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p)) %>%
+  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
+  ylab("Species %") + xlab("") + ggtitle("S01A") + theme(legend.position="none")
+
+g2 = group_by(cfec[cfec$p_fshy=="S 01E" & cfec$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
+  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p)) %>%
+  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
+  ylab("Species %") + xlab("") + ggtitle("S01E") + theme(legend.position="none")
+
+g3 = group_by(cfec[cfec$p_fshy=="S 01K" & cfec$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
+  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p)) %>%
+  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
+  ylab("Species %") + xlab("") + ggtitle("S01K") + theme(legend.position="none")
+
+g4 = group_by(cfec[cfec$p_fshy=="S 01M" & cfec$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
+  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p)) %>%
+  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
+  ylab("Species %") + xlab("") + ggtitle("S01M") + theme(legend.position="none")
+
+g5 = group_by(cfec[cfec$p_fshy=="S 03A" & cfec$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
+  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p)) %>%
+  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
+  ylab("Species %") + xlab("") + ggtitle("S03A") + theme(legend.position="none")
+
+g6 = group_by(cfec[cfec$p_fshy=="S 03E" & cfec$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
+  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p)) %>%
+  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
+  ylab("Species %") + xlab("") + ggtitle("S03E") + theme(legend.position="none")
+
+dat.diff = readRDS(file="../../portfolio/data-generated/cfec-diff-for-modeling.rds")
+dat.diff = dat.diff[which(dat.diff$revenue >= 10000 & dat.diff$revenue.prev >= 10000), ]
+dat.salmon = dat.diff[dat.diff$strategy_permit%in%c("S01A", "S01E", "S01K", "S03A", "S03E","S01M"),]
+
+h1 = group_by(dat.salmon[dat.salmon$strategy_permit=="S01A",], year, strategy_permit) %>%
+  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
+  ggplot(aes(year, meanDiv)) + geom_line() + 
+  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Diversity") + xlab("") + theme(legend.position="none")
+
+h2 = group_by(dat.salmon[dat.salmon$strategy_permit=="S01E",], year, strategy_permit) %>%
+  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
+  ggplot(aes(year, meanDiv)) + geom_line() + 
+  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Diversity") + xlab("") + theme(legend.position="none")
+
+h3 = group_by(dat.salmon[dat.salmon$strategy_permit=="S01K",], year, strategy_permit) %>%
+  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
+  ggplot(aes(year, meanDiv)) + geom_line() + 
+  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Diversity") + xlab("") + theme(legend.position="none")
+
+h4 = group_by(dat.salmon[dat.salmon$strategy_permit=="S01M",], year, strategy_permit) %>%
+  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
+  ggplot(aes(year, meanDiv)) + geom_line() + 
+  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Diversity") + xlab("") + theme(legend.position="none")
+
+h5 = group_by(dat.salmon[dat.salmon$strategy_permit=="S03A",], year, strategy_permit) %>%
+  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
+  ggplot(aes(year, meanDiv)) + geom_line() + 
+  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Diversity") + xlab("") + theme(legend.position="none")
+
+h6 = group_by(dat.salmon[dat.salmon$strategy_permit=="S03E",], year, strategy_permit) %>%
+  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
+  ggplot(aes(year, meanDiv)) + geom_line() + 
+  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Diversity") + xlab("") + theme(legend.position="none")
+
+gridExtra::grid.arrange(g1, g2, g3, h1, h2, h3, nrow=2)
+```
+
+```{r, echo=FALSE, fig.pos="placeHere", fig.cap="Time varying species composition and diversity, by permit.", fig.height=3.2} 
+gridExtra::grid.arrange(g4, g5, g6, h4, h5, h6, nrow=2)
+```
+
+\break
+
+```{r, echo=FALSE}
 library(rstan)
+```
 
+```{r, echo=FALSE, fig.pos="placeHere", fig.cap="Time varying intercepts on estimated mean revenue, by permit.", eval=FALSE}
 # grab original data to look at permit names
@@
   facet_wrap(~permit) + xlab("Year") + ylab("Intercepts on mean")
-
 ```
@@
 for(i in 1:nrow(df)) {
-  tmp = est$g0_strategy[,df$permit[i]] + x1[df$diversity[i]] * (est$g_1[,df$year[i], df$permit[i]]+est$g1_str_yr_mu) +
+  tmp = est$g0 + est$g0_strategy[,df$permit[i]] + x1[df$diversity[i]] * (est$g_1[,df$year[i], df$permit[i]]+est$g1_str_yr_mu) +
     x2[df$diversity[i]] * (est$g_2[,df$year[i], df$permit[i]]+est$g2_str_yr_mu)# +

### 2016-10-05 salmon/analysis/3-run-factor-model.R

@@ -1,0 +1,79 @@
+
+dat = readRDS(file="salmon/data-generated/salmon.rds")
+
+dat$salm = dat$coho+dat$chnk
+dat$salm.prev = dat$coho.prev+dat$chnk.prev
+
+dat$totsalm = dat$salm + dat$chum + dat$sock + dat$pink
+dat$totsalm.prev = dat$salm.prev + dat$chum.prev + dat$sock.prev + dat$pink.prev
+
+dat$salm = dat$salm/dat$totsalm
+dat$sock = dat$sock/dat$totsalm
+dat$pink = dat$pink/dat$totsalm
+dat$chum = dat$chum/dat$totsalm
+dat$chnk = dat$chnk/dat$totsalm
+dat$coho = dat$coho/dat$totsalm
+dat$salm.prev = dat$salm.prev/dat$totsalm.prev
+dat$sock.prev = dat$sock.prev/dat$totsalm.prev
+dat$pink.prev = dat$pink.prev/dat$totsalm.prev
+dat$chum.prev = dat$chum.prev/dat$totsalm.prev
+dat$chnk.prev = dat$chnk.prev/dat$totsalm.prev
+dat$coho.prev = dat$coho.prev/dat$totsalm.prev
+
+dat = dat[-which(is.na(dat$salm+dat$sock+dat$pink+dat$chum)),]
+dat$logdiff = log(dat$revenue/dat$revenue.prev)
+
+dat$sumP2 = abs(dat$salm - dat$salm.prev) +
+  abs(dat$pink-dat$pink.prev) +
+  abs(dat$chum-dat$chum.prev) +
+  abs(dat$sock-dat$sock.prev)
+dat = dat[which(dat$sumP2 < 0.3),]
+
+dat = dat[which(dat$strategy_permit=="S01E"),]
+
+pamk_s01e <- fpc::pamk(dat[,c("chum","pink","sock","salm")])
+dat$group = c("Pink-chum","Pink")[pamk_s01e$pamobject$clustering]
+
+# prep data for stan
+N = nrow(dat) # number of rows of data
+#X_ij = diffdat[,c("days.change")] #"specDiv.change","specdiv.prev",
+#if(class(X_ij)=="numeric") X_ij = matrix(X_ij,ncol=1)
+J = ncol(X_ij) # number of covariates on mean
+X_sigma_ik = dat[,c("days.change")] #"specDiv.change","specdiv.prev",
[…]

### 2016-10-06 salmon/paper/salmon_paper.Rmd
salmon paper

@@
 
-```{r, echo=FALSE, fig.pos="placeHere", fig.cap="Time varying intercepts on estimated mean revenue, by permit.", eval=FALSE}
+```{r, eval=FALSE,echo=FALSE}
+load("../../salmon/analysis/cv-model-simple.Rdata")
+# groups here are 80% pink vs 97% pink
+
+yrs = 1986:2014
+
+hist(extract(mod)[["g_offset"]], 40, col="grey", xlab="Effect of specializing on pink (sd)")
+
+xx = exp(extract(mod)[["sd_yr"]])
+df.gen = data.frame("year"=yrs, "mean" = apply(xx, 2, mean), "low" = apply(xx, 2, quantile, 0.025),
+  "hi" = apply(xx, 2, quantile, 0.975), "strategy"="pink:chum")
+
+offset = matrix(extract(mod)[["g_offset"]],ncol=1) %*% matrix(rep(1,length(yrs)), nrow=1)
+xx = exp(extract(mod)[["sd_yr"]] + offset)
+df.spec = data.frame("year"=yrs, "mean" = apply(xx, 2, mean), "low" = apply(xx, 2, quantile, 0.025),
+  "hi" = apply(xx, 2, quantile, 0.975), "strategy"="pink")
+
+ggplot(data=rbind(df.gen,df.spec), aes(year, mean, group=strategy, fill=strategy)) + geom_ribbon(aes(ymin=low,ymax=hi), alpha=0.5)
+
+```
+
+```{r, echo=FALSE, fig.pos="placeHere", fig.cap="Time varying intercepts on estimated mean revenue, by permit.
+These are interpreted as the mean change in revenue for each permit holder in each year (positive equates to increasing revenue, etc), after accounting for effects of catch diversification and days fished."}
 # grab original data to look at permit names

### 2016-10-06 salmon/analysis/1-run-model.R

@@ -3,15 +3,15 @@ diffdat = readRDS(file="salmon/data-generated/salmon.rds")
 
 # prep data for stan
 N = nrow(diffdat) # number of rows of data
-#X_ij = diffdat[,c("days.change")] #"specDiv.change","specdiv.prev",
+X_ij = diffdat[,c("days.change")] #"specDiv.change","specdiv.prev",
 #if(class(X_ij)=="numeric") X_ij = matrix(X_ij,ncol=1)
 J = ncol(X_ij) # number of covariates on mean
 X_sigma_ik = diffdat[,c("days.change")] #"specDiv.change","specdiv.prev",
-if(class(X_sigma_ik)=="numeric") X_sigma_ik = matrix(X_sigma_ik,ncol=1)
+#if(class(X_sigma_ik)=="numeric") X_sigma_ik = matrix(X_sigma_ik,ncol=1)
 K = ncol(X_sigma_ik) # number of covariates on variance
 
-offset = diffdat$revenue.prev
-y_i = diffdat$revenue
+offset = log(diffdat$revenue.prev)
+y_i = log(diffdat$revenue)
 n_strategy = length(unique(diffdat$permit))
 strategy_i = as.numeric(as.factor(diffdat$permit))
 n_str_yr = length(unique(diffdat$year_permit))

### 2016-10-06 salmon/analysis/3-run-simple-cv.R

@@ -44,8 +44,8 @@ X_sigma_ik = dat[,c("days.change")] #"specDiv.change","specdiv.prev",
 if(class(X_sigma_ik)=="numeric") X_sigma_ik = matrix(X_sigma_ik,ncol=1)
 K = ncol(X_sigma_ik) # number of covariates on variance
 
-offset = dat$revenue.prev
-y_i = dat$revenue
+offset = log(dat$revenue.prev)
+y_i = log(dat$revenue)
 n_strategy = length(unique(dat$group))
 strategy_i = as.numeric(as.factor(dat$group))
 n_yr = length(unique(dat$year))

### 2016-10-06 salmon/paper/salmon_paper.Rmd
paper

@@
 
-ggplot(data=rbind(df.gen,df.spec), aes(year, mean, group=strategy, fill=strategy)) + geom_ribbon(aes(ymin=low,ymax=hi), alpha=0.5)
+ggplot(data=rbind(df.gen,df.spec), aes(year, mean, group=strategy, fill=strategy)) + geom_ribbon(aes(ymin=low,ymax=hi), alpha=0.5) + geom_hline(yintercept=0, color="red")
 
@@
 yearpermit$b0 = apply(est$b0_str_yr,2,median)
-yearpermit$b0_low = apply(est$b0_str_yr,2,quantile,0.25)
-yearpermit$b0_hi = apply(est$b0_str_yr,2,quantile,0.75)
+yearpermit$b0_low = apply(est$b0_str_yr,2,quantile,0.025)
+yearpermit$b0_hi = apply(est$b0_str_yr,2,quantile,0.975)
 
@@
   geom_line(aes(x=year,y=b0)) + 
-  facet_wrap(~permit) + xlab("Year") + ylab("Intercepts on mean")
+  facet_wrap(~permit) + xlab("Year") + 
+  ylab("Intercepts on mean") + 
+  geom_hline(yintercept=0, color="red")
 ```
@@
 x2 = log(specdiv/specdiv.prev) * specdiv.prev
-x3 = log(specdiv/specdiv.prev) * specdiv.prev^2
 
@@
 for(i in 1:nrow(df)) {
-  tmp = est$b0_str_yr[,b0_indx[i]] + 
-    x1[df$diversity[i]] * (est$b_1[,df$year[i], df$permit[i]]+est$b1_str_yr_mu) +
-    x2[df$diversity[i]] * (est$b_2[,df$year[i], df$permit[i]]+est$b2_str_yr_mu)# +
-    #x3[df$diversity[i]] * (est$b_3[,df$year[i], df$permit[i]]+est$b3_str_yr_mu)
+  tmp = x1[df$diversity[i]] * (est$b_1[,df$year[i], df$permit[i]]+est$b1_str_yr_mu) + x2[df$diversity[i]] * (est$b_2[,df$year[i], df$permit[i]]+est$b2_str_yr_mu)
+
   df$pred[i] = median(tmp)
-  df$low[i] = quantile(tmp,0.25)
-  df$high[i] = quantile(tmp,0.75)
+  df$low[i] = quantile(tmp,0.025)
+  df$high[i] = quantile(tmp,0.975)
 }
@@
   geom_line() +
-  facet_wrap(~years) + xlab("Previous species diversity") + ylab("Predicted benefit of adding +1 spp (ln $)")
+  facet_wrap(~years) + xlab("Previous species diversity") + ylab("Predicted benefit of adding +1 spp (ln $)") + geom_hline(yintercept=0, color="red")
 
@@
   df$pred[i] = median(tmp)
-  df$low[i] = quantile(tmp,0.25)
-  df$high[i] = quantile(tmp,0.75)
+  df$low[i] = quantile(tmp,0.025)
+  df$high[i] = quantile(tmp,0.975)
 }
@@
   geom_ribbon(aes(ymin=low,ymax=high), alpha=0.4) +
-  facet_wrap(~permits) + xlab("Year") + ylab("Predicted benefit of adding +1 spp (ln $)")
+  facet_wrap(~permits) + xlab("Year") + ylab("Predicted benefit of adding +1 spp (ln $)") + geom_hline(yintercept=0, color="red")
 
@@
   geom_ribbon(aes(ymin=low,ymax=high), alpha=0.4) +
-  facet_wrap(~permits) + xlab("Year") + ylab("Predicted benefit of adding +1 spp (ln $)")
+  facet_wrap(~permits) + xlab("Year") + ylab("Predicted marginal benefit of adding +1 spp (ln $)")
 

### 2016-10-08 salmon/paper/salmon_paper.Rmd
salmon model

@@
 ---
-title: 'Reduced diversity and increased specialzation of Alaska salmon fishermen:
+title: 'Reduced diversity and increased specialization of Alaska salmon fishermen:
   effects on variation in individual revenues'
-author: "Eric, Sean, Ole, Rich, Milo, Jordan, others?"
+author: "Eric, Sean, Ole, Rich, Milo, Jordan, Jen, others"
 date: ''
@@
 library(zoo)
+modelname = "model-linear" # model-linear-nointeractions
 ```
@@
 We define the variables $\Delta { S }_{ t }=log\left( { S }_{ t }/{ S }_{ t-1 } \right)$ to represent the percent change in species diversity of the catch between years and $\Delta { D }_{ t }=log\left( { D }_{ t }/{ D }_{ t-1 } \right)$ to represent the percent change in fishing effort (days fished) between years.
-The model for the expected log revenue of individual *i* in year *t*, $$\widehat { log\left( { R }_{ i,t } \right)  } ={ B }_{ 0,j,t }+{ B }_{ 1,j,t }\cdot \Delta S_{ t }+{ B }_{ 2,j,t }\cdot \left( \Delta S_{ t }\cdot { S }_{ t-1 } \right) +{ B }_{ 3,j }\cdot \Delta D_{ t }+log\left( { R }_{ i,t-1 } \right) $$, where *j* subscripts the salmon permit (Table 1).
+The model for the expected log revenue of individual *i* in year *t*, $$\widehat { log\left( { R }_{ i,t } \right)  } ={ B }_{ 0,j,t }+{ B }_{ 1,j,t }\cdot \Delta S_{ i,t }+{ B }_{ 2,j,t }\cdot \left( \Delta S_{ i,t }\cdot { S }_{ i,t-1 } \right) +{ B }_{ 3,j }\cdot \Delta D_{ i,t }+log\left( { R }_{ i,t-1 } \right) $$, where *j* subscripts the salmon permit (Table 1).
 We model the time-varying permit specific intercepts as random and exchangeable across years, ${ B }_{ 0,j,t }\sim Normal\left( { u }_{ { B }_{ 0 } },{ \sigma  }_{ { B }_{ 0 } } \right)$.
@@
 Finally, the term $log\left( { R }_{ i,t-1 } \right)$ is included as an offset [@anderson_benefits].
+  Like [@anderson_benefits], we also included all of the above covariates in the variance model.
+We assumed the residuals from the mean model described above to be normally distributed, $$log\left( { R }_{ i,t } \right) \sim Normal\left( \widehat { log\left( { R }_{ i,t } \right)  } ,\widehat { { \sigma  }_{ i,t } }  \right) $$.
+While conventional hierarchcial linear regression assumes variances to be constant or shared by grouping variables, we also included fixed and random covariates, making the model akin to variance function regression or stochastic volatility models (CITE).
+Using the same notation as for the mean, we can express the log standard deviation as $$\widehat { log\left( { \sigma  }_{ i,t } \right)  } ={ G }_{ 0,j }+{ G }_{ 1,j,t }\cdot \Delta S_{ i,t }+{ G }_{ 2,j,t }\cdot \left( \Delta S_{ i,t }\cdot { S }_{ i,t-1 } \right) +G_{ 3,j }\cdot \Delta D_{ i,t }$$.
+Unlike the linear model for the mean, we did not include an offset term.
+Further, we constrained the variance intercepts ${ G }_{ 0,j }$ to be temporally constant by strategy, instead of the dynamic nature of the intercepts in the mean model.
+Like the mean model, the random effects were constrained to be autoregressive, ${ G }_{ 1,j,t }\sim Normal\left( { G }_{ 1,j,t-1 },{ \sigma  }_{ { G }_{ 1 } } \right)$, and covariates associated with days fished and species diversity were included as predictors.
 
-- Fit models from Sean's paper, but include time varying interactions on species diversity.
 All done in STAN [@stan_development_team_stan_2015] using the NUTS algorithm [@hoffman_no-u-turn_2014].
@@
 
-Figure 1 shows that the time varying intercepts on the mean, also in [@anderson_benefits] are basically flat. These represent common fluctuations in revenue experienced by all permit holders.
+Figure 1 and 2 shows the time series of species composition and effective spp diversity. Interesting patterns are (1) overall low diversity of S01E (pinks), (2) increasing chum in S03A
 
-Figure 2 is largley illustrative, but shows the benefit of including time varying coefficients.
+Figure 3: One component of variability is the year-to-year variation that individuals experience as a result of external forces (stock abundance, prices, etc).
+Or common fluctuations in revenue experienced by all permit holders [@anderson_benefits].
+This shows the time varying intercepts on the mean -- negative equals losing revenue, positive = increasing revenue.
+This shows big differences between permit types (gillnets tend to have lower year-to-year variation).
+Some permits have extreme variation - S01E (PWS seine) has fluctuations between exp(-1) and exp(+1), corresponding to year-to-year changes of 36-270%.
+Figure 4 shows the time varying nature of of the model for all permits.
+These are the time series of benefits of changing from effective species diversity of 1 to 2, and here we're basically isolating year-to-year differences in revenue (the intercepts) from diversification benefits.
+In some cases the effect is positive (fishing more spp = more revenue).
+But in many cases it's not, and the trend is negative (indicating that specialization is more profitable over time).
+In several cases, these time series mirror the aggregate diversity in the raw time series (Fig.
+1).
 
-Figure 3 shows the temporal trends, by permit, in going from 1 to 2 species fished, and how those effects have changed through time.
-S01A, S01E, S03A appear to be decreasing.
-S01K and S01M increasing.
-S03E pretty flat?
-
-Figure 4 shows that temporal trends, by permit, in going from 1 to 2 species fished on the variance.
-From [@anderson_benefits] we'd expect diversification to redudce variance -- that seems to be the biggest signal for S01K and S01M in the years of low sockeye prices.
+Figure 5 shows that temporal trends, by permit, in going from 1 to 2 species fished on the variance (again removing the overall intercepts).
+From [@anderson_benefits] we'd expect diversification to redudce variance.
+A few examples of how to read this: S01A is slightly positve for most of the time series, but rapidly increases in the last 10 years -- this means generalists with diverse portfolios have had increases in year-to-year variability.
+For S01E, the start and end of the time series are associated with these 'costs' to diversification.
+For S03E, we see a more gradual trend (cost of diversification increases).
 
@@
 
-load("../../salmon/analysis/model-linear.Rdata")
+load(paste0("../../salmon/analysis/",modelname,".Rdata"))
 
@@
 ```{r, fig.pos="placeHere", fig.cap="Time varying effects of adding 1 species on revenue (measured in ln dollars) for purse seine permit in Prince William Sound (S01E).
-This figure illustrates two things: (1) the benefit of adding more species diminishes as a function of past diversity, and (2) the economic benefit to adding more species has been reduced through time.", echo=FALSE}
+This figure illustrates two things: (1) the benefit of adding more species diminishes as a function of past diversity, and (2) the economic benefit to adding more species has been reduced through time.", echo=FALSE, eval=FALSE}
 # grab original data to look at permit names
@@
 for(i in 1:nrow(df)) {
-  tmp = x1[df$diversity[i]] * (est$b_1[,df$year[i], df$permit[i]]+est$b1_str_yr_mu) + x2[df$diversity[i]] * (est$b_2[,df$year[i], df$permit[i]]+est$b2_str_yr_mu)
-
+  tmp = x1[df$diversity[i]] * (est$b_1[,df$year[i], df$permit[i]])# + x2[df$diversity[i]] * (est$b_2[,df$year[i], df$permit[i]]+est$b2_str_yr_mu)
   df$pred[i] = median(tmp)
@@
 }
-
+df_s01e_b1 = df
 ggplot(df[df$permits=="S01E",], aes(x=x,y=pred,group=years)) +
@@
   geom_line() +
-  facet_wrap(~years) + xlab("Previous species diversity") + ylab("Predicted benefit of adding +1 spp (ln $)") + geom_hline(yintercept=0, color="red")
+  facet_wrap(~years) + xlab("Previous species diversity") + ylab("Predicted benefit of adding +1 spp (ln $)")
 
@@
 ```{r, fig.pos="placeHere", fig.cap="Time varying effects of adding 1 species on revenue (measured in ln dollars) for hypothetical individuals with species diversity ~ 1.
-This shows declining benefits of diversification for S01A, S01E, S03A, increasing (but more uncertain) benefits for sockeye oriented fisheries S01M and S01K.", echo=FALSE}
+This shows declining benefits of diversification for most fisheries, S01M.", echo=FALSE}
 # grab original data to look at permit names
@@
 for(i in 1:nrow(df)) {
-  tmp = est$b0_str_yr[,b0_indx[i]] + 
-    x1[df$diversity[i]] * (est$b_1[,df$year[i], df$permit[i]]+est$b1_str_yr_mu) +
-    x2[df$diversity[i]] * (est$b_2[,df$year[i], df$permit[i]]+est$b2_str_yr_mu)# +
+  #tmp = est$b0_str_yr[,b0_indx[i]] + 
+  tmp = 
+    x1[df$diversity[i]] * (est$b_1[,df$year[i], df$permit[i]])# +
+    #x2[df$diversity[i]] * (est$b_2[,df$year[i], df$permit[i]]+est$b2_str_yr_mu)# +
     #x3[df$diversity[i]] * (est$b_3[,df$year[i], df$permit[i]]+est$b3_str_yr_mu)
-  df$pred[i] = median(tmp)
+  df$pred[i] = mean(tmp)
   df$low[i] = quantile(tmp,0.025)
@@
 for(i in 1:nrow(df)) {
-  tmp = est$g0 + est$g0_strategy[,df$permit[i]] + x1[df$diversity[i]] * (est$g_1[,df$year[i], df$permit[i]]+est$g1_str_yr_mu) +
-    x2[df$diversity[i]] * (est$g_2[,df$year[i], df$permit[i]]+est$g2_str_yr_mu)# +
-  #x3[df$diversity[i]] * (est$b_3[,df$year[i], df$permit[i]]+est$b3_str_yr_mu)
+  #tmp = est$g0 + est$g0_strategy[,df$permit[i]] + 
+  tmp = 
+    x1[df$diversity[i]] * (est$g_1[,df$year[i], df$permit[i]])# +
+    #x2[df$diversity[i]] * (est$g_2[,df$year[i], df$permit[i]]+est$g2_str_yr_mu)
   df$pred[i] = median(tmp)
@@
   geom_ribbon(aes(ymin=low,ymax=high), alpha=0.4) +
-  facet_wrap(~permits) + xlab("Year") + ylab("Predicted marginal benefit of adding +1 spp (ln $)")
+  facet_wrap(~permits) + xlab("Year") + ylab("Predicted marginal benefit of adding +1 spp (ln $)") + geom_hline(yintercept=0, color="red")
+
+```
+
+```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "As model validation, we can compare the estimated intercepts for PWS S01E (purse seine) representing average year to year change in revenue versus the change year over year in ln(pink salmon returns).
+Correlation between them is ~ 0.8.
+This suggests (as expected) people on average make a lot more money in years when pink salmon returns are high."}
+
+S01E = yearpermit[yearpermit$permit=="S01E",]
+names(S01E)[which(names(S01E)=="year")]="Year"
+
+returns = readxl::read_excel("../../salmon/paper/salmon_returns_PWS.xlsx")
+pinkPWS = returns[,c("Year", "Pink_total", "Chum_total")]
+pinkPWS = pinkPWS[which(pinkPWS$Year >= 1985),]
+pinkPWS$diff = c(NA, diff(log(pinkPWS$Pink_total)))
+
+pinkPWS = dplyr::left_join(pinkPWS, S01E)
+
+# bring in slopes from above
+slopes = df_s01e_b1[df_s01e_b1$permits=="S01E" & df_s01e_b1$diversity==1,]
+slopes = slopes[,c("years","pred")]
+names(slopes)[which(names(slopes)=="years")] = "Year"
+names(slopes)[which(names(slopes)=="pred")] = "B1"
+pinkPWS = dplyr::left_join(pinkPWS, slopes)
+
+ggplot(data=pinkPWS, aes(b0, diff)) + geom_point() + stat_smooth(method=lm) + xlab("Estimated intercept from model") + ylab("Change in total pink salmon returns")
+
+pinkPWS$diffchum = c(NA, diff(log(pinkPWS$Chum_total)))
+
+prices = group_by(cfec[which(cfec$p_fshy%in%c("S 01E","S 03E") & cfec$spec%in%c("PINK","CHUM","SOCK")),], year, spec) %>% 
+  summarize(meanPrice = mean(g_price)) %>%
+  arrange(year)
+prices <- dcast(melt(prices, id.vars = c("spec", "year")), year ~ variable+spec)
+names(prices)[which(names(prices)=="year")]="Year"
+
+pinkPWS = dplyr::left_join(pinkPWS, prices)
+pinkPWS$pricediffchum = c(NA, diff(pinkPWS$meanPrice_CHUM))
+pinkPWS$pricediffpink = c(NA, diff(pinkPWS$meanPrice_PINK))
+pinkPWS$pricediffsock = c(NA, diff(pinkPWS$meanPrice_SOCK))
+
+#cor(pinkPWS[,-c(5,6)], use="pairwise.complete.obs")
+
+#ggplot(data=pinkPWS, aes(b0, B1)) + geom_point() + stat_smooth(method=lm) + xlab("Estimated intercept from model") + ylab("Change in total pink salmon returns")
+
+#diversity = dplyr::group_by(dat.salmon[dat.salmon$strategy_permit=="S01E",], year) %>%
+#  summarize(meanDiv = mean(specDiv))
+#names(diversity)[which(names(diversity)=="year")] = "Year"
+#pinkPWS = dplyr::left_join(pinkPWS, diversity)
 

### 2016-10-08 salmon/analysis/1-run-model.R

@@ -29,19 +29,76 @@ stan_data = list("N"=N, "y_i"=y_i, "offset"=offset, "n_strategy"=n_strategy,
   "n_str_yr"=n_str_yr,"str_yr_i"=str_yr_i, "n_yr"=n_yr, "year_i"=year_i,
   "b1_cov_i"=b1_cov_i, "b2_cov_i"=b2_cov_i, "b3_cov_i"=b3_cov_i, "b4_cov_i"=b4_cov_i,
   "g1_cov_i"=b1_cov_i, "g2_cov_i"=b2_cov_i, "g3_cov_i"=b3_cov_i, "g4_cov_i"=b4_cov_i)
-stan_pars = c("b0_str_yr", "b0_str_yr_tau", "g0_strategy", "g0_strategy_tau",
-  "b_1", "b_2", "b_3", "b_4", "b1_str_yr_tau", "b2_str_yr_tau",
-  "b3_str_yr_tau", "b4_str_yr_tau", "b1_str_yr_mu", "b2_str_yr_mu", "b3_str_yr_mu",
-  "b4_str_yr_mu", "g0", "g_1", "g_2", "g_3", "g_4", "g1_str_yr_tau", "g2_str_yr_tau",
-  "g3_str_yr_tau", "g4_str_yr_tau", "g1_str_yr_mu", "g2_str_yr_mu", "g3_str_yr_mu",
-  "g4_str_yr_mu")
 
+if(modelname=="quadratic") {
+
+  stan_pars = c("b0_str_yr", "b0_str_yr_tau", "g0_strategy", "g0_strategy_tau",
+    "b_1", "b_2", "b_3", "b_4", "b1_str_yr_tau", "b2_str_yr_tau", "b3_str_yr_tau",
+    "b4_str_yr_tau", "b1_str_yr_mu", "b2_str_yr_mu", "b3_str_yr_mu",
+    "b4_str_yr_mu", "g0", "g_1", "g_2", "g_3", "g_4", "g1_str_yr_tau", "g2_str_yr_tau",
+    "g3_str_yr_tau", "g4_str_yr_tau", "g1_str_yr_mu", "g2_str_yr_mu", "g3_str_yr_mu",
+    "g4_str_yr_mu")
+
+  library(rstan)
+  rstan_options(auto_write = TRUE)
+  options(mc.cores = parallel::detectCores())
+
[…]

### 2016-10-08 salmon/paper/salmon_paper.Rmd
paper

@@
 Or common fluctuations in revenue experienced by all permit holders [@anderson_benefits].
-This shows the time varying intercepts on the mean -- negative equals losing revenue, positive = increasing revenue.
+This shows the time varying intercepts on the mean -- negative equals losing revenue on average, positive = increasing revenue.
 This shows big differences between permit types (gillnets tend to have lower year-to-year variation).
 Some permits have extreme variation - S01E (PWS seine) has fluctuations between exp(-1) and exp(+1), corresponding to year-to-year changes of 36-270%.
+
 Figure 4 shows the time varying nature of of the model for all permits.
@@
 But in many cases it's not, and the trend is negative (indicating that specialization is more profitable over time).
-In several cases, these time series mirror the aggregate diversity in the raw time series (Fig.
+In some cases, these time series mirror the aggregate diversity in the raw time series (Fig.
 1).
+I think we also expect this result - fishermen increase diversity or not, because it's profitable.
 

### 2016-10-10 salmon/paper/salmon_paper.Rmd
log -> \log, Normal -> \mathrm{Normal}, STAN -> Stan

@@
 Clustering was performed using the partitioning around medoids (pam) algorithm in R with an estimated number of clusters (using 'pamk' in fpc library), and selected 2 groups of individual strategies: pink salmon specialization (medoid proportion of pink salmon = 0.95), and a strategy focused on pink salmon and chum (medoid proportions of 0.64 and 0.31 respectively).
-For each group, we then calculated the mean and CV of revenue; we calculated the log difference of revenue for each individual, $\Delta { R }_{ i }=log\left( { R }_{ i,t+1 }/{ R }_{ i,t } \right)$ and then calculated the CV as the standard deviation over people, $sd(\Delta { R }_{ i=1:n })$.
+For each group, we then calculated the mean and CV of revenue; we calculated the log difference of revenue for each individual, $\Delta { R }_{ i }=\log\left( { R }_{ i,t+1 }/{ R }_{ i,t } \right)$ and then calculated the CV as the standard deviation over people, $sd(\Delta { R }_{ i=1:n })$.
 
@@
 To ask how benefits of diversification may have changed over time, we extend the model of [@anderson_benefits] to also include time-varying random effects in the slope coefficients of diversification.
-We define the variables $\Delta { S }_{ t }=log\left( { S }_{ t }/{ S }_{ t-1 } \right)$ to represent the percent change in species diversity of the catch between years and $\Delta { D }_{ t }=log\left( { D }_{ t }/{ D }_{ t-1 } \right)$ to represent the percent change in fishing effort (days fished) between years.
-The model for the expected log revenue of individual *i* in year *t*, $$\widehat { log\left( { R }_{ i,t } \right)  } ={ B }_{ 0,j,t }+{ B }_{ 1,j,t }\cdot \Delta S_{ i,t }+{ B }_{ 2,j,t }\cdot \left( \Delta S_{ i,t }\cdot { S }_{ i,t-1 } \right) +{ B }_{ 3,j }\cdot \Delta D_{ i,t }+log\left( { R }_{ i,t-1 } \right) $$, where *j* subscripts the salmon permit (Table 1).
-We model the time-varying permit specific intercepts as random and exchangeable across years, ${ B }_{ 0,j,t }\sim Normal\left( { u }_{ { B }_{ 0 } },{ \sigma  }_{ { B }_{ 0 } } \right)$.
+We define the variables $\Delta { S }_{ t }=\log\left( { S }_{ t }/{ S }_{ t-1 } \right)$ to represent the percent change in species diversity of the catch between years and $\Delta { D }_{ t }=\log\left( { D }_{ t }/{ D }_{ t-1 } \right)$ to represent the percent change in fishing effort (days fished) between years.
+The model for the expected log revenue of individual *i* in year *t*, $$\widehat { \log\left( { R }_{ i,t } \right)  } ={ B }_{ 0,j,t }+{ B }_{ 1,j,t }\cdot \Delta S_{ i,t }+{ B }_{ 2,j,t }\cdot \left( \Delta S_{ i,t }\cdot { S }_{ i,t-1 } \right) +{ B }_{ 3,j }\cdot \Delta D_{ i,t }+\log\left( { R }_{ i,t-1 } \right) $$, where *j* subscripts the salmon permit (Table 1).
+We model the time-varying permit specific intercepts as random and exchangeable across years, ${ B }_{ 0,j,t }\sim \mathrm{Normal}\left( { u }_{ { B }_{ 0 } },{ \sigma  }_{ { B }_{ 0 } } \right)$.
 The term ${ B }_{ 1,j,t }$ represents the permit level random effect of diversifying species.
@@
 The flexibility in these terms allows the benefit of increasing species diversity to be positive in some years (generalizing benefits revenue), negative in other years (specializing benefits revenue) and neutral in other years (revenue is independent of alterations to catch composition).
-We constrain the random effects associated with diversification to be modeled as autoregressive processes with shared variances across strategies, for example: ${ B }_{ 1,j,t }\sim Normal\left( { B }_{ 1,j,t-1 },{ \sigma  }_{ { B }_{ 1 } } \right)$.
+We constrain the random effects associated with diversification to be modeled as autoregressive processes with shared variances across strategies, for example: ${ B }_{ 1,j,t }\sim \mathrm{Normal}\left( { B }_{ 1,j,t-1 },{ \sigma  }_{ { B }_{ 1 } } \right)$.
 Like [@anderson_benefits] we also allow changing effort (measured as percent change in days fished) ${ B }_{ 4,j }$ to vary by strategy, but we do not allow this effect to vary through time.
-Finally, the term $log\left( { R }_{ i,t-1 } \right)$ is included as an offset [@anderson_benefits].
+Finally, the term $\log\left( { R }_{ i,t-1 } \right)$ is included as an offset [@anderson_benefits].
   Like [@anderson_benefits], we also included all of the above covariates in the variance model.
-We assumed the residuals from the mean model described above to be normally distributed, $$log\left( { R }_{ i,t } \right) \sim Normal\left( \widehat { log\left( { R }_{ i,t } \right)  } ,\widehat { { \sigma  }_{ i,t } }  \right) $$.
+We assumed the residuals from the mean model described above to be normally distributed, $$\log\left( { R }_{ i,t } \right) \sim \mathrm{Normal}\left( \widehat { \log\left( { R }_{ i,t } \right)  } ,\widehat { { \sigma  }_{ i,t } }  \right) $$.
 While conventional hierarchcial linear regression assumes variances to be constant or shared by grouping variables, we also included fixed and random covariates, making the model akin to variance function regression or stochastic volatility models (CITE).
-Using the same notation as for the mean, we can express the log standard deviation as $$\widehat { log\left( { \sigma  }_{ i,t } \right)  } ={ G }_{ 0,j }+{ G }_{ 1,j,t }\cdot \Delta S_{ i,t }+{ G }_{ 2,j,t }\cdot \left( \Delta S_{ i,t }\cdot { S }_{ i,t-1 } \right) +G_{ 3,j }\cdot \Delta D_{ i,t }$$.
+Using the same notation as for the mean, we can express the log standard deviation as $$\widehat { \log\left( { \sigma  }_{ i,t } \right)  } ={ G }_{ 0,j }+{ G }_{ 1,j,t }\cdot \Delta S_{ i,t }+{ G }_{ 2,j,t }\cdot \left( \Delta S_{ i,t }\cdot { S }_{ i,t-1 } \right) +G_{ 3,j }\cdot \Delta D_{ i,t }$$.
 Unlike the linear model for the mean, we did not include an offset term.
 Further, we constrained the variance intercepts ${ G }_{ 0,j }$ to be temporally constant by strategy, instead of the dynamic nature of the intercepts in the mean model.
-Like the mean model, the random effects were constrained to be autoregressive, ${ G }_{ 1,j,t }\sim Normal\left( { G }_{ 1,j,t-1 },{ \sigma  }_{ { G }_{ 1 } } \right)$, and covariates associated with days fished and species diversity were included as predictors.
+Like the mean model, the random effects were constrained to be autoregressive, ${ G }_{ 1,j,t }\sim \mathrm{Normal}\left( { G }_{ 1,j,t-1 },{ \sigma  }_{ { G }_{ 1 } } \right)$, and covariates associated with days fished and species diversity were included as predictors.
 
-All done in STAN [@stan_development_team_stan_2015] using the NUTS algorithm [@hoffman_no-u-turn_2014].
+All done in Stan [@stan_development_team_stan_2015] using the NUTS algorithm [@hoffman_no-u-turn_2014].
 

### 2016-10-13 salmon/paper/salmon_paper_appendix.Rmd
updates to salmon paper and diversity paper

@@
 
-```{r, echo=FALSE, message=FALSE, warning=FALSE, cache=TRUE}
+```{r, echo=FALSE, message=FALSE, warning=FALSE}
 revenueThreshold = 10000
 
-dat.annual = readRDS(file="../../portfolio/data-generated/cfec-annual-for-modeling.rds")
+dat.annual = readRDS(file="../../portfolio/data-generated/salmon-annual-for-modeling.rds")
 
@@
 
-dat.diff = readRDS(file="../../portfolio/data-generated/cfec-diff-for-modeling.rds")
+dat.diff = readRDS(file="../../portfolio/data-generated/salmon-diff-for-modeling.rds")
 dat.diff = dat.diff[which(dat.diff$revenue >= revenueThreshold & dat.diff$revenue.prev >= revenueThreshold), ]
-dat.diff = dat.diff[dat.diff$strategy_permit%in%c("S01A","S01E","S01K","S01L","S01M",
-  "S03A","S03E","S03H","S03M","S03T","S04D","S04H","S04K","S04M","S04T"),]
+dat.diff = dat.diff[dat.diff$strategy_permit%in%c("S01A","S01E","S01K","S01M",
+  "S03A","S03E"),]
 
@@
 # Also load in raw CFEC data for plots
-cfec <- feather::read_feather("../../portfolio/data-generated/cfec.feather")
+cfec <- feather::read_feather("../../portfolio/data-generated/salmon.feather")
 cfec = cfec[cfec$p_fshy%in%c("S 01A", "S 01E", "S 01K", "S 03A", "S 03E", "S 01M"),]
@@
 # summarize the temporal distribution of the catches
+cfec.raw = cfec.raw[which(is.na(cfec.raw$jday)==FALSE),]
 seasons = group_by(cfec.raw, p_fshy, year) %>% 
@@
 
-```{r, echo = FALSE, eval=FALSE, fig.cap="Results for PWS Purse seine fishery", fig.pos="placeHere", cache=TRUE}
+```{r, echo = FALSE, eval=FALSE, fig.cap="Results for PWS Purse seine fishery", fig.pos="placeHere"}
 
@@
 
-```{r, echo = FALSE, fig.cap="SEAK purse seine fishery", cache=TRUE, fig.pos="placeHere", eval=FALSE}
+```{r, echo = FALSE, fig.cap="SEAK purse seine fishery", fig.pos="placeHere", eval=FALSE}
 

### 2016-10-13 diversity/plot_prices_revenue.R

@@ -1,0 +1,20 @@
+
+
+cfec <- feather::read_feather("portfolio/data-generated/cfec.feather")
+cfec$year <- as.numeric(cfec$year)
+
+species = c("CHNK","COHO","PINK","SOCK","CHUM","PHLB","SABL","PLCK","KCRB")
+
+library(dplyr)
+pdf("prices_revenue.pdf")
+prices = group_by(cfec[cfec$spec%in%species,], year, spec) %>%
+  summarize(priceDollars = sum(g_earn)/sum(g_pounds), revenueMillions = sum(g_earn)/1000000)
+
+%>%
+  ungroup() %>%
+  group_by(spec) %>%
+  mutate(scaleP = price/price[year==1985])
+
+  ggplot(scaled, aes(year, scaleP, color=spec)) + geom_line() +
+  xlab("Year") + ylab("Price ($)") + geom_point(aes(size=million))
+dev.off()

### 2016-10-13 salmon/analysis/0-process-data.R

@@ -1,7 +1,7 @@
 library(dplyr)
 revenueThreshold = 10000
 
-dat.diff = readRDS(file="portfolio/data-generated/cfec-diff-for-modeling.rds")
+dat.diff = readRDS(file="portfolio/data-generated/salmon-diff-for-modeling.rds")
 
 length(unique(dat.diff$p_holder))
 
@@ -11,6 +11,7 @@ dat.diff = group_by(dat.diff, p_holder) %>%
   select(-medRev)
 
 length(unique(dat.diff$p_holder))
+# 3483 people of 3592
 
 dat.salmon = dat.diff[dat.diff$strategy_permit%in%c("S01A", "S01E", "S01K", "S03A", "S03E","S01M"),]
 length(unique(dat.salmon$p_holder))

### 2016-10-13 salmon/analysis/1-run-model.R

@@ -1,6 +1,8 @@
 
 diffdat = readRDS(file="salmon/data-generated/salmon.rds")
 
+diffdat = diffdat[-which(is.na(diffdat$days.change)),]
+
 # prep data for stan
 N = nrow(diffdat) # number of rows of data
 X_ij = diffdat[,c("days.change")] #"specDiv.change","specdiv.prev",

### 2016-10-13 salmon/analysis/process_salmon_data.R

@@ -1,0 +1,78 @@
+
+setwd("/users/eric.ward/downloads/cfec_early")
+
+for(y in 1975:1984) {
+
+early = haven::read_sas(paste0("gross_earnings_",y,".sas7bdat"))
+
+early$F_PERMIT = paste0(early$ADFG_H_PERMIT_FISHERY,early$ADFG_H_PERMIT_SERIAL_NUMBER,early$ADFG_H_PERMIT_CHECK_DIGIT)
+
+early$startdt = paste0(substr(early$ADFG_H_DATE_FISHING_BEGAN,1,4), "-",
+  substr(early$ADFG_H_DATE_FISHING_BEGAN,5,6), "-",
+  substr(early$ADFG_H_DATE_FISHING_BEGAN,7,8))
+
+early$landdate = paste0(substr(early$ADFG_H_DATE_LANDED,1,4), "-",
+  substr(early$ADFG_H_DATE_LANDED,5,6), "-",
+  substr(early$ADFG_H_DATE_LANDED,7,8))
+
+early = rename(early, F_INPFC=ADFG_H_CDQ_CODE,
+  F_GEAR = ADFG_H_GEAR_CODE,
+  P_TYPE = CFEC_PMT_TYPE,
+  P_STATUS = CFEC_PMT_ID_STATUS,
+  A_RES = CFEC_ADR_RESIDENCY,
+  C_LANDST = CFEC_LANDING_STATUS,
+  F_DISPOS = ADFG_I_DISPOSITION_CODE,
+  F_MPGMID = ADFG_H_MGT_PROGRAM_ID,
+  F_DELIVR = ADFG_I_DELIVERY_CODE)
+
+early = early[,c("ADFG_B_BATCH_YEAR","ADFG_H_PORT_CODE","CFEC_FILE_NUMBER",
+  "CFEC_PMT_SERIAL","CFEC_VALUE","ADFG_I_POUNDS","ADFG_H_SEQ_TICKET_NUMBER",
+  "ADFG_H_ADFG_NUMBER","CFEC_CORRECTED_ADFG","ADFG_I_HARVEST_CODE","ADFG_H_PROCESSOR_CODE",
+  "CFEC_SPECIES_CODE","CFEC_STAT_AREA","ADFG_I_SPECIES_CODE","CFEC_PMT_FSHY",
[…]

### 2016-10-14 diversity/analysis/herring_case_study.Rmd
update

@@
 library(date)
+library(reshape2)
+library(cluster)
+```
+
+```{r}
+cfec <- feather::read_feather("../../portfolio/data-generated/cfec.feather")
+cfec$year <- as.numeric(cfec$year)
+
+this.permit = "S 03T"
+# S01E S03T - effective permit diversity
+
+simp.div = function(x) {
+  1/sum((x/sum(x))^2)
+}
+
+g1 = group_by(cfec, p_holder) %>% 
+  mutate(pws = ifelse(length(which(p_fshy==c(this.permit)))>0, 1, 0)) %>% 
+  filter(pws>0) %>%
+  select(-pws) %>% 
+  group_by(p_holder, year, p_fshy) %>% 
+  summarize(g = sum(g_earn)) %>%
+  group_by(p_holder, year) %>% 
+  summarize(div = simp.div(g)) %>% 
+  group_by(year) %>% 
+  summarize(mean = mean(div), lower=quantile(div,0.025), upper=quantile(div,0.975)) %>% 
+  ggplot(aes(year, mean)) + geom_line() + ggtitle(this.permit) + ylab("Mean diversity")
+
+# Identify dominant strategies associated with this permit
+g2 = group_by(cfec, p_holder) %>% 
+  mutate(pws = ifelse(length(which(p_fshy==c(this.permit)))>0, 1, 0)) %>% 
+  filter(pws>0) %>%
+  select(-pws) %>% 
+  group_by(p_holder, year, p_fshy) %>% 
+  summarize(g = sum(g_earn)) %>%
+  group_by(p_holder, year) %>% 
+  mutate(totRev = sum(g)) %>% 
+  group_by(p_holder, year, p_fshy) %>% 
+  mutate(g = g / totRev) %>% 
+  select(-totRev)
+
+# use reshape 2 to cast this to wide format
+aql <- melt(g2, id.vars = c("year", "p_fshy", "p_holder"))
+aqw <- dcast(aql, year + p_holder ~ p_fshy)
+# replace NAs w/0s
+for(i in 1:ncol(aqw)) {
+  aqw[which(is.na(aqw[,i])),i] = 0
+}
+
+indx = sample(seq(1,nrow(aqw)),size=500,replace=F)
+sub = aqw[indx,-c(1,2)]
+
+sub = aqw[,-c(1,2)]
+coarse_df = aqw[,-c(1,2)]
+coarse_df$S = apply(coarse_df[,which(substr(names(sub),1,1)=="S")],1,sum)
+coarse_df$B = apply(coarse_df[,which(substr(names(sub),1,1)=="B")],1,sum)
+coarse_df$GHL = apply(coarse_df[,which(substr(names(sub),1,1)%in%c("G","H","L"))],1,sum)
+coarse_df$KTD = apply(coarse_df[,which(substr(names(sub),1,1)%in%c("K","T","D"))],1,sum)
+coarse_df$M = apply(coarse_df[,which(substr(names(sub),1,1)%in%c("M"))],1,sum)
+coarse_df$P = apply(coarse_df[,which(substr(names(sub),1,1)%in%c("P"))],1,sum)
+coarse_df = coarse_df[,c("S","B","GHL","KTD","M","P")]
+
+clust = fpc::pamk(coarse_df[sample(seq(1,nrow(aqw)),size=500,replace=F),])# cluster based on subset to pick medoids
+medoids = clust$pamobject$medoids
+clust = pam(coarse_df, k = clust$nc) # cluster full dataset
+
+names_medoids = ""
+for(i in 1:nrow(medoids)) {
+  names_medoids[i] = paste0(colnames(medoids)[which(medoids[i,] > 0.05)], collapse=":")
+}
+
+aqw$clusters = clust$clustering
+totals = group_by(aqw, clusters, year) %>% 
+  summarize(n = n())
+totals$group = names_medoids[totals$clusters]
+
+ggplot(totals, aes(year, n, group=group, color=group)) + geom_line() + geom_point()
+
+ggplot(totals, aes(year, n)) + geom_area(aes(fill = group), position = "stack", colour = 1)
+
+```
+
+```{r}
+
+```
+
+```{r}
+cfec <- feather::read_feather("../../portfolio/data-generated/cfec.feather")
+cfec$year <- as.numeric(cfec$year)
+
+# grab people who ever held prince william sound herring permit
+# calculate unique permits by person-year
+pws.G34 = group_by(cfec, p_holder) %>% 
+  mutate(pws = ifelse(length(which(p_fshy==c("G 34E")))>0, 1, 0)) %>% 
+  filter(pws>0) %>%
+  select(-pws) %>% 
+  group_by(p_holder, year) %>% 
+  summarize(n = length(unique(p_fshy))) %>%
+  group_by(year) %>% 
+  summarize(p1 = length(which(n>1)) / length(which(n > 0)))
+
+pws.G01 = group_by(cfec, p_holder) %>% 
+  mutate(pws = ifelse(length(which(p_fshy==c("G 01E")))>0, 1, 0)) %>% 
+  filter(pws>0) %>%
+  select(-pws) %>% 
+  group_by(p_holder, year) %>% 
+  summarize(n = length(unique(p_fshy))) %>%
+  group_by(year) %>% 
+  summarize(p1 = length(which(n>1)) / length(which(n > 0)))
+
+pws.G34$permit="g34"
+pws.G01$permit="g01"
+pws = rbind(pws.G34, pws.G01)
+
+ggplot(pws, aes(year, p1, group=permit, color=permit)) + geom_line() + 
+  ylab("Proportion of people landing fish under > 1 permit") + xlab("Year")
+
+```
+
+```{r}
+# grab people who ever held prince william sound herring permit
+# calculate unique permits by person-year
+pws.G34 = group_by(cfec, p_holder) %>% 
+  mutate(pws = ifelse(length(which(p_fshy==c("G 34E")))>0, 1, 0)) %>% 
+  filter(pws>0) %>%
+  select(-pws) %>% 
+  group_by(p_holder, year) %>% 
+  summarize(n = length(unique(p_fshy))) %>%
+  group_by(year) %>% 
+  summarize(p1 = length(which(n>1)) / length(which(n > 0)))
+
+pws.G01 = group_by(cfec, p_holder) %>% 
+  mutate(pws = ifelse(length(which(p_fshy==c("G 01E")))>0, 1, 0)) %>% 
+  filter(pws>0) %>%
+  select(-pws) %>% 
+  group_by(p_holder, year) %>% 
+  summarize(n = length(unique(p_fshy))) %>%
+  group_by(year) %>% 
+  summarize(p1 = length(which(n>1)) / length(which(n > 0)))
+
+pws.G34$permit="g34"
+pws.G01$permit="g01"
+pws = rbind(pws.G34, pws.G01)
+
+ggplot(pws, aes(year, p1, group=permit, color=permit)) + geom_line() + 
+  ylab("Proportion of people landing fish under > 1 permit") + xlab("Year")
+
 ```
@@
 cfec$year <- as.numeric(cfec$year)
+
+# grab people who ever held prince william sound herring permit
+# calculate unique permits by person-year
+pws.G34 = group_by(cfec, p_holder) %>% 
+  mutate(pws = ifelse(length(which(p_fshy==c("G 34E")))>0, 1, 0)) %>% 
+  filter(pws>0) %>%
+  select(-pws) %>% 
+  group_by(p_holder, year) %>% 
+  summarize(n = length(unique(p_fshy))) %>%
+  group_by(year) %>% 
+  summarize(p1 = length(which(n>1)) / length(which(n > 0)))
+
+pws.G01 = group_by(cfec, p_holder) %>% 
+  mutate(pws = ifelse(length(which(p_fshy==c("G 01E")))>0, 1, 0)) %>% 
+  filter(pws>0) %>%
+  select(-pws) %>% 
+  group_by(p_holder, year) %>% 
+  summarize(n = length(unique(p_fshy))) %>%
+  group_by(year) %>% 
+  summarize(p1 = length(which(n>1)) / length(which(n > 0)))
+
+pws.G34$permit="g34"
+pws.G01$permit="g01"
+pws = rbind(pws.G34, pws.G01)
+
+ggplot(pws, aes(year, p1, group=permit, color=permit)) + geom_line() + 
+  ylab("Proportion of people landing fish under > 1 permit") + xlab("Year")
 

### 2016-10-14 diversity/general_plots_multipanel.R

@@ -1,0 +1,262 @@
+library(dplyr)
+library(ggplot2)
+library(reshape2)
+
+pdf(paste0("diversity/generalPlots_multipanel",".pdf"))
+
+cfec = feather::read_feather("portfolio/data-generated/cfec.feather")
+cfec$year <- as.numeric(cfec$year)
+simp.div = function(x) {
+  1/sum((x/sum(x))^2)
+}
+
+
+yourfishery <- c("G 01E","G 34E", "H 01E", "H 34E", "L 21E")
+mydat <- cfec[which(cfec$p_fshy%in%yourfishery),]
+# Identify dominant strategies associated with this permit
+g2 = group_by(cfec, p_holder) %>%
+  mutate(pws = ifelse(length(which(p_fshy%in%yourfishery))>0, 1, 0)) %>%
+  filter(pws>0) %>%
+  select(-pws) %>%
+  group_by(p_holder, year, p_fshy) %>%
+  summarize(g = sum(g_earn)) %>%
+  group_by(p_holder, year) %>%
+  mutate(totRev = sum(g)) %>%
+  group_by(p_holder, year, p_fshy) %>%
+  mutate(g = g / totRev) %>%
+  select(-totRev)
+
+# use reshape 2 to cast this to wide format
+aql <- melt(g2, id.vars = c("year", "p_fshy", "p_holder"))
+aqw <- dcast(aql, year + p_holder ~ p_fshy)
+# replace NAs w/0s
+for(i in 1:ncol(aqw)) {
+  aqw[which(is.na(aqw[,i])),i] = 0
+}
+
+indx = sample(seq(1,nrow(aqw)),size=1000,replace=F)
+sub = aqw[indx,-c(1,2)]
+
+sub = aqw[,-c(1,2)]
+coarse_df = aqw[,-c(1,2)]
+coarse_df$S = apply(coarse_df[,which(substr(names(sub),1,1)=="S")],1,sum)
+coarse_df$B = apply(coarse_df[,which(substr(names(sub),1,1)=="B")],1,sum)
+coarse_df$GHL = apply(coarse_df[,which(substr(names(sub),1,1)%in%c("G","H","L"))],1,sum)
+coarse_df$KTD = apply(coarse_df[,which(substr(names(sub),1,1)%in%c("K","T","D"))],1,sum)
+coarse_df$M = apply(coarse_df[,which(substr(names(sub),1,1)%in%c("M"))],1,sum)
[…]

### 2016-10-14 summaries_for_ole.R

@@ -3,10 +3,9 @@
 cfec <- feather::read_feather("portfolio/data-generated/cfec.feather")
 cfec$year <- as.numeric(cfec$year)
 
-areas = read.csv(file.choose())
-
-cfec = cfec[which(cfec$stat6 %in% areas$STAT6),]
 cfec = cfec[which(substr(cfec$p_fshy,1,1) %in% c("M","B","C")),]
+
+cfec = cfec[which(cfec$region%in%c("Kodiak","PWS","Alaska Peninsula","Cook Inlet")),]
 
 # 99.4% of the records are made up of B/C/M permits, 361619 / 363789
 library(dplyr)
@@ -14,10 +13,9 @@ g = group_by(cfec, year, stat6) %>%
   summarize(pounds = sum(g_pounds))
 
 # roll these stat6 areas into Ole's regions
-names(areas)[which(names(areas)=="STAT6")] = "stat6"
-g = left_join(g, areas)
+#names(areas)[which(names(areas)=="STAT6")] = "stat6"
+#g = left_join(g, areas)
+#totPounds = group_by(g, AREA, year) %>%
+#  summarize(totPounds = sum(pounds))
 
-totPounds = group_by(g, AREA, year) %>%
-  summarize(totPounds = sum(pounds))
-
-write.table(totPounds, "totPounds.csv", sep=",", row.names=FALSE)
+write.table(g, "pounds_by_stat6.csv", sep=",", row.names=FALSE)

### 2016-10-14 salmon/paper/salmon_paper.Rmd
updates

@@
 
+- In terms of participation, why did people drop out in 2000 in SEAK and Alaska Peninsula, but not in PWS?
+PWS largely seems like a boutique fishery on the Copper River -- which includes a qualtiy component (Copper River at forefront) and a timing issue relative to the rest of the Gulf of Alaska populations (buffers against market impacts)
+
 - Revenue is going up since 2000 -- but each permit holder is fishing more (more days, more total pounds) and the trend in weight is faster than the trend in price.
@@
 # grab original data to look at permit names
-diffdat = readRDS(file="../../salmon/data-generated/salmon.rds")
+diffdat = readRDS(file="../../portfolio/data-generated/salmon-diff-for-modeling.rds")
 years = seq(min(diffdat$year), max(diffdat$year))
@@
 
-yearpermit = data.frame("yrpermit"=levels(diffdat$year_permit))
+yearpermit = data.frame("yrpermit"=unique(diffdat$year_permit))
 yearpermit$year = as.numeric(substr(yearpermit$yrpermit,1,4))

### 2016-10-14 compare_sockeyePrices_bb_copper.R

@@ -1,0 +1,26 @@
+library(dplyr)
+library(ggplot2)
+library(reshape2)
+
+cfec = feather::read_feather("portfolio/data-generated/cfec.feather")
+cfec$year <- as.numeric(cfec$year)
+
+sock = filter(cfec, spec=="SOCK")
+
+copper = filter(sock, stat6 %in% c(20020, 21235, 20010, 21231, 21221, 21211, 21215, 21225))
+copper = group_by(copper, year) %>%
+  summarize(p = sum(g_earn)/sum(g_pounds))
+copper$area = "copper"
+
+bb = filter(sock, stat6 %in% c(32610, 32611, 32671, 32670,
+  32672, 32620, 32621, 32630, 32631, 32640, 32641, 32105,
+  32180, 32140, 32160, 32170, 32100, 32120, 32130, 32140,
+  32100, 32110))
+bb = group_by(bb, year) %>%
+  summarize(p = sum(g_earn)/sum(g_pounds))
+bb$area = "bristol"
+
+dat = bind_rows(copper, bb)
+
+ggplot(dat, aes(year, p, group = area, color = area)) +
+  geom_line() + xlab("Year") + ylab("Sockeye price ($/lb)")

### 2016-10-14 diversity/general_plots_multipanel.R

@@ -2,8 +2,6 @@ library(dplyr)
 library(ggplot2)
 library(reshape2)
 
-pdf(paste0("diversity/generalPlots_multipanel",".pdf"))
-
 cfec = feather::read_feather("portfolio/data-generated/cfec.feather")
 cfec$year <- as.numeric(cfec$year)
 simp.div = function(x) {
@@ -45,6 +43,7 @@ coarse_df$GHL = apply(coarse_df[,which(substr(names(sub),1,1)%in%c("G","H","L"))
 coarse_df$KTD = apply(coarse_df[,which(substr(names(sub),1,1)%in%c("K","T","D"))],1,sum)
 coarse_df$M = apply(coarse_df[,which(substr(names(sub),1,1)%in%c("M"))],1,sum)
 coarse_df$P = apply(coarse_df[,which(substr(names(sub),1,1)%in%c("P"))],1,sum)
+#coarse_df$Q = apply(coarse_df[,which(substr(names(sub),1,1)%in%c("Q"))],1,sum)
 coarse_df$OTHER = apply(coarse_df[,which(substr(names(sub),1,1)%in%c("S","B","G","H","L","K","T","D","M","P")==FALSE)],1,sum)
 coarse_df = coarse_df[,c("S","B","GHL","KTD","M","P","OTHER")]
 
@@ -67,8 +66,6 @@ p1.totals = group_by(totals, year, group) %>%
   summarize(n = sum(n))
 
 #p1 = ggplot(totals, aes(year, n)) + geom_area(aes(fill = group), position = "stack", colour = 1) + xlab("Year") + ylab("People")
-
-
 
 yourfishery <- c("S 01E")
 mydat <- cfec[which(cfec$p_fshy%in%yourfishery),]
@@ -104,8 +101,9 @@ coarse_df$GHL = apply(coarse_df[,which(substr(names(sub),1,1)%in%c("G","H","L"))
 coarse_df$KTD = apply(coarse_df[,which(substr(names(sub),1,1)%in%c("K","T","D"))],1,sum)
 coarse_df$M = apply(coarse_df[,which(substr(names(sub),1,1)%in%c("M"))],1,sum)
 coarse_df$P = apply(coarse_df[,which(substr(names(sub),1,1)%in%c("P"))],1,sum)
[…]

### 2016-10-14 salmon/paper/salmon_paper.Rmd
make plots

@@
 library(ggplot2)
+library(reshape2)
 library(date)
 library(ggtern)
-library(reshape2)
 library(mgcv)
@@
 library(zoo)
+library(rstan)
 modelname = "model-linear" # model-linear-nointeractions
@@
 
+```{r, echo=FALSE, fig.pos="placeHere", fig.cap="Harvest of hatchery produced salmon through time and by area.", fig.height=3.2}
+
+seak = read.csv("salmon/data-generated/stopha_table32.csv")
+pws = read.csv("salmon/data-generated/stopha_table33.csv")
+kodiak = read.csv("salmon/data-generated/stopha_table35.csv")
+
+seak = melt(seak, id.vars = c("Year"))
+pws = melt(pws, id.vars = c("Year"))
+kodiak = melt(kodiak, id.vars = c("Year"))
+
+seak$region = "Southeast"
+pws$region = "Prince William Sound"
+kodiak$region = "Kodiak"
+
+seak$value[which(is.na(seak$value))] = 0
+pws$value[which(is.na(pws$value))] = 0
+kodiak$value[which(is.na(kodiak$value))] = 0
+
+dat = bind_rows(seak, pws)
+dat = bind_rows(dat, kodiak)
+
+names(dat)[which(names(dat)=="variable")]="species"
+dat$value = dat$value / 1000000
+
+ggplot(dat, aes(Year, value)) +
+  geom_area(aes(fill = species), position = "stack", colour = 1) +
+  xlab("Year") + ylab("Harvest of hatchery produced salmon (millions)") +
+  facet_wrap(~region, scale="free_y")
+
+```
+
 ```{r, echo=FALSE, fig.pos="placeHere", fig.cap="Time varying species composition and diversity, by permit.
@@
 
-library(dplyr)
 # plot proportion of species landed by salmon permit
@@
 \break
-
-```{r, echo=FALSE}
-library(rstan)
-```
 

### 2016-10-14 portfolio/analysis/3.3-plot-within-strategy-effects.R
IFQs

@@ -37,7 +37,7 @@ ns <- dat %>% group_by(strategy) %>%
 res <- inner_join(res, ns)
 
 p1 <- res %>%
-  mutate(strategy_label = ifelse(nn > 350, str_label, NA)) %>%
+  mutate(strategy_label = ifelse(nn > 10, str_label, NA)) %>%
   mutate(inc = inc) %>%
   ggplot(aes(x = inc_rev, y = inc)) +
   geom_hline(yintercept = 0, lty = 2, col = "grey65") +
@@ -66,7 +66,7 @@ p2 <- res %>%
   mutate(dec = -dec, dec_rev = -dec_rev) %>%
   mutate(dec.l = -dec.l, dec_rev.l = -dec_rev.l) %>%
   mutate(dec.u = -dec.u, dec_rev.u = -dec_rev.u) %>%
-  mutate(strategy_label = ifelse(nn > 350, str_label, NA)) %>%
+  mutate(strategy_label = ifelse(nn > 10, str_label, NA)) %>%
   ggplot(aes(x = dec_rev, y = dec)) +
   geom_hline(yintercept = 0, lty = 2, col = "grey65") +
   geom_vline(xintercept = 0, lty = 2, col = "grey65") +
@@ -89,3 +89,15 @@ p2 <- res %>%
 pdf("portfolio/figs/stan-offset-break-anti-spaghetti.pdf", width = 10, height = 4)
 grid_arrange_shared_legend(p1, p2, ncol = 2, nrow = 1, position = "right")
 dev.off()
+
+# ifq
+
+ifq <- readr::read_csv("data/ifq_pfshy_year.csv")
+ifq <- ifq %>% filter(!is.na(year_catch_share))
+ifq
+
+ifq2 <- c("B61B", "B61B C61B", "B61B C61B S15B", "B61B S01", "B61B S03",
+  "B61B S03A", "B61B S03E", "B61B S15B", "C61B", "K91", "K91 T09",
+    "M7HB")
+
+res <- res %>% mutate(ifq = strategy %in% ifq2)

### 2016-10-18 data/process_raw_data/process_raw_cfec_data.Rmd
updates to salmon and diversity

--- 2016-04-07 process_raw_data/process_raw_cfec_data.Rmd
@@
 # drop records with no spp
-cfec = cfec[-which(is.na(cfec$spec)),]
+cfec = filter(cfec, !is.na(spec))
 rm(spec.tbl)
@@
 ```{r}
+cfec$p_holder_id = cfec$p_holder
+
 cfec$p_holder = as.numeric(as.factor(cfec$p_holder))

### 2016-10-18 salmon/paper/salmon_paper.Rmd
updates to salmon and diversity

@@
 
-seak = read.csv("salmon/data-generated/stopha_table32.csv")
-pws = read.csv("salmon/data-generated/stopha_table33.csv")
-kodiak = read.csv("salmon/data-generated/stopha_table35.csv")
+seak = read.csv("../../salmon/data-generated/stopha_table32.csv")
+pws = read.csv("../../salmon/data-generated/stopha_table33.csv")
+kodiak = read.csv("../../salmon/data-generated/stopha_table35.csv")
 
@@
 
-g6 = group_by(cfec[cfec$p_fshy=="S 03E" & cfec$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
+g6 = group_by(cfec[cfec$p_fshy=="S 03E" & cfec$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK") & cfec$year > 0,], year, spec, p_fshy) %>%
   summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p)) %>%
@@
 # grab original data to look at permit names
-diffdat = readRDS(file="../../portfolio/data-generated/salmon-diff-for-modeling.rds")
+diffdat = readRDS(file="../../salmon/data-generated/salmon.rds")
+
 years = seq(min(diffdat$year), max(diffdat$year))
@@
 
-yearpermit = data.frame("yrpermit"=unique(diffdat$year_permit))
+# note that we have to use factor() here instead of unique() because
+# that's the same as the ordering in the estimation code
+yearpermit = data.frame("yrpermit"=levels(as.factor(diffdat$year_permit)))
 yearpermit$year = as.numeric(substr(yearpermit$yrpermit,1,4))
@@
 
-load(paste0("../../salmon/analysis/",modelname,".Rdata"))
-
+yearpermitold=yearpermit
+load("../../salmon/analysis/model-linear.Rdata")
+yearpermit=yearpermitold
 est = extract(mod, permute="TRUE")
@@
 
-```{r, fig.pos="placeHere", fig.cap="Time varying effects of adding 1 species on revenue (measured in ln dollars) for purse seine permit in Prince William Sound (S01E).
-This figure illustrates two things: (1) the benefit of adding more species diminishes as a function of past diversity, and (2) the economic benefit to adding more species has been reduced through time.", echo=FALSE, eval=FALSE}
+```{r, echo=FALSE}
+# fig.pos="placeHere", fig.cap="Time varying effects of adding 1 species on revenue (measured in ln dollars) for purse seine permit in Prince William Sound (S01E).
+This figure illustrates two things: (1) the benefit of adding more species diminishes as a function of past diversity, and (2) the economic benefit to adding more species has been reduced through time."
 # grab original data to look at permit names
@@
 years = seq(min(diffdat$year), max(diffdat$year))
-permits = levels(as.factor(diffdat$permit))
+#permits = levels(as.factor(diffdat$permit))
 
-load(paste0("../../salmon/analysis/",modelname,".Rdata"))
+load("../../salmon/analysis/model-linear.Rdata")
+yearpermit=yearpermitold
 
@@
 df_s01e_b1 = df
-ggplot(df[df$permits=="S01E",], aes(x=x,y=pred,group=years)) +
-  #geom_ribbon(aes(ymin=low,ymax=high), alpha=0.4) +
-  geom_line() +
-  facet_wrap(~years) + xlab("Previous species diversity") + ylab("Predicted benefit of adding +1 spp (ln $)")
+#ggplot(df[df$permits=="S01E",], aes(x=x,y=pred,group=years)) +
+#  #geom_ribbon(aes(ymin=low,ymax=high), alpha=0.4) +
+#  geom_line() +
+#  facet_wrap(~years) + xlab("Previous species diversity") + ylab("Predicted #benefit of adding +1 spp (ln $)")
 
@@
 years = seq(min(diffdat$year), max(diffdat$year))
-permits = levels(as.factor(diffdat$permit))
+#permits = levels(as.factor(diffdat$permit))
 
-load(paste0("../../salmon/analysis/",modelname,".Rdata"))
+yearpermitold=yearpermit
+load("../../salmon/analysis/model-linear.Rdata")
+yearpermit=yearpermitold
 
@@
 
-df = expand.grid("year"=1:29, "diversity" = seq(1,length(x1)), "permit"=1:6)
+df = expand.grid("year"=1:length(years), "diversity" = seq(1,length(x1)), "permit"=1:6)
 df$permits = permits[df$permit]
@@
 
+yearpermit$b0 = apply(est$b0_str_yr,2,median)
+yearpermit$b0_low = apply(est$b0_str_yr,2,quantile,0.025)
+yearpermit$b0_hi = apply(est$b0_str_yr,2,quantile,0.975)
+
 S01E = yearpermit[yearpermit$permit=="S01E",]
@@
 pinkPWS = returns[,c("Year", "Pink_total", "Chum_total")]
-pinkPWS = pinkPWS[which(pinkPWS$Year >= 1985),]
+pinkPWS = pinkPWS[which(pinkPWS$Year >= 1975),]
 pinkPWS$diff = c(NA, diff(log(pinkPWS$Pink_total)))
@@
 
+ggplot(data=pinkPWS, aes(b0, diff,color=Year)) + geom_point() + stat_smooth(method=lm) + xlab("Estimated intercept from model") + ylab("Change in total pink salmon returns")
+
 # bring in slopes from above
-slopes = df_s01e_b1[df_s01e_b1$permits=="S01E" & df_s01e_b1$diversity==1,]
-slopes = slopes[,c("years","pred")]
-names(slopes)[which(names(slopes)=="years")] = "Year"
-names(slopes)[which(names(slopes)=="pred")] = "B1"
-pinkPWS = dplyr::left_join(pinkPWS, slopes)
+#slopes = df_s01e_b1[df_s01e_b1$permits=="S01E" & df_s01e_b1$diversity==1,]
+#slopes = slopes[,c("years","pred")]
+#names(slopes)[which(names(slopes)=="years")] = "Year"
+#names(slopes)[which(names(slopes)=="pred")] = "B1"
+#pinkPWS = dplyr::left_join(pinkPWS, slopes)
 
-ggplot(data=pinkPWS, aes(b0, diff)) + geom_point() + stat_smooth(method=lm) + xlab("Estimated intercept from model") + ylab("Change in total pink salmon returns")
+#pinkPWS$diffchum = c(NA, diff(log(pinkPWS$Chum_total)))
 
-pinkPWS$diffchum = c(NA, diff(log(pinkPWS$Chum_total)))
+#prices = group_by(cfec[which(cfec$p_fshy%in%c("S 01E","S 03E") & #cfec$spec%in%c("PINK","CHUM","SOCK")),], year, spec) %>% 
+#  summarize(meanPrice = mean(g_price)) %>%
+#  arrange(year)
+#prices <- dcast(melt(prices, id.vars = c("spec", "year")), year ~ #variable+spec)
+#names(prices)[which(names(prices)=="year")]="Year"
 
-prices = group_by(cfec[which(cfec$p_fshy%in%c("S 01E","S 03E") & cfec$spec%in%c("PINK","CHUM","SOCK")),], year, spec) %>% 
-  summarize(meanPrice = mean(g_price)) %>%
-  arrange(year)
-prices <- dcast(melt(prices, id.vars = c("spec", "year")), year ~ variable+spec)
-names(prices)[which(names(prices)=="year")]="Year"
-
-pinkPWS = dplyr::left_join(pinkPWS, prices)
-pinkPWS$pricediffchum = c(NA, diff(pinkPWS$meanPrice_CHUM))
-pinkPWS$pricediffpink = c(NA, diff(pinkPWS$meanPrice_PINK))
-pinkPWS$pricediffsock = c(NA, diff(pinkPWS$meanPrice_SOCK))
+#pinkPWS = dplyr::left_join(pinkPWS, prices)
+#pinkPWS$pricediffchum = c(NA, diff(pinkPWS$meanPrice_CHUM))
+#pinkPWS$pricediffpink = c(NA, diff(pinkPWS$meanPrice_PINK))
+#pinkPWS$pricediffsock = c(NA, diff(pinkPWS$meanPrice_SOCK))
 

### 2016-10-18 diversity/general_plots_multipanel_anne.R

@@ -8,12 +8,10 @@ simp.div = function(x) {
   1/sum((x/sum(x))^2)
 }
 
-
-yourfishery <- c("G 01E","G 34E", "H 01E", "H 34E", "L 21E")
 #mydat <- cfec[which(cfec$p_fshy%in%yourfishery),]
 # Identify dominant strategies associated with this permit
 g2 = group_by(cfec, p_holder) %>%
-  mutate(pws = ifelse(length(which(p_fshy%in%yourfishery))>0, 1, 0)) %>%
+  mutate(pws = ifelse(length(which(region%in%c("PWS")))>0, 1, 0)) %>%
   filter(pws>0) %>%
   select(-pws) %>%
   group_by(p_holder, year, p_fshy) %>%
@@ -67,11 +65,8 @@ p1.totals = group_by(totals, year, group) %>%
 
 #p1 = ggplot(totals, aes(year, n)) + geom_area(aes(fill = group), position = "stack", colour = 1) + xlab("Year") + ylab("People")
 
-yourfishery <- c("S 01E")
-#mydat <- cfec[which(cfec$p_fshy%in%yourfishery),]
-# Identify dominant strategies associated with this permit
 g2 = group_by(cfec, p_holder) %>%
-  mutate(pws = ifelse(length(which(p_fshy%in%yourfishery))>0, 1, 0)) %>%
+  mutate(pws = ifelse(length(which(region%in%c("Cook Inlet")))>0, 1, 0)) %>%
   filter(pws>0) %>%
   select(-pws) %>%
   group_by(p_holder, year, p_fshy) %>%
@@ -126,12 +121,8 @@ p2.totals = group_by(totals, year, group) %>%
 #p2 = ggplot(totals, aes(year, n)) + geom_area(aes(fill = group), position = "stack", colour = 1) + xlab("Year") + ylab("People")
 
 
-yourfishery <- c("S 03T")
-#
-#mydat <- cfec[which(cfec$p_fshy%in%yourfishery),]
-# Identify dominant strategies associated with this permit
 g2 = group_by(cfec, p_holder) %>%
-  mutate(pws = ifelse(length(which(p_fshy%in%yourfishery))>0, 1, 0)) %>%
+  mutate(pws = ifelse(length(which(region%in%c("Kodiak")))>0, 1, 0)) %>%
   filter(pws>0) %>%
   select(-pws) %>%
[…]

### 2016-10-18 diversity/small_large_halibut.R

@@ -8,6 +8,10 @@ simp.div = function(x) {
   1/sum((x/sum(x))^2)
 }
 
+# Seak Inside / Outside
+
+#
+
 # sum up halibut landings by region and permit
 pdf("small_vs_large_halibut.pdf")
 group_by(cfec[cfec$region%in%c("PWS","Kodiak","Cook Inlet","SEAK Outside"), ], year, region) %>%

### 2016-10-18 diversity/subset_people.R

@@ -2,7 +2,7 @@ library(dplyr)
 library(ggplot2)
 library(reshape2)
 
-cfec = feather::read_feather("portfolio/data-generated/cfec.feather")
+cfec = readRDS("portfolio/data-generated/cfec-diff-for-modeling.rds")
 cfec$year <- as.numeric(cfec$year)
 simp.div = function(x) {
   1/sum((x/sum(x))^2)
@@ -247,13 +247,12 @@ p4.totals = group_by(totals, year, group) %>%
 #p4 = ggplot(totals, aes(year, n)) + geom_area(aes(fill = group), position = "stack", colour = 1) + xlab("Year") + ylab("People")
 
 # Ole's halibut example
-#yourfishery <- c("SEAK Inside","SEAK Outside","PWS","Cook Inlet",
-#  "Kodiak","Yakutat","Alaska Peninsula")
-#yourfishery <- c("SEAK Inside","SEAK Outside")
+yourfishery <- c("SEAK Inside","SEAK Outside","PWS","Cook Inlet",
+  "Kodiak","Yakutat","Alaska Peninsula")
 #mydat <- cfec[which(cfec$region%in%yourfishery),]
 # Identify dominant strategies associated with this permit
 g2 = group_by(cfec, p_holder) %>%
-  mutate(pws = ifelse(length(which(p_fshy%in%c("B 06B","B 61B")))>0, 1, 0)) %>%
+  mutate(pws = ifelse(length(which(cfec$region%in%yourfishery & p_fshy%in%c("B 06B","B 61B")))>0, 1, 0)) %>%
   filter(pws>0) %>%
   select(-pws) %>%
   group_by(p_holder, year, p_fshy) %>%
@@ -272,7 +271,7 @@ for(i in 1:ncol(aqw)) {
   aqw[which(is.na(aqw[,i])),i] = 0
 }
 
-indx = sample(seq(1,nrow(aqw)),size=5000,replace=F)
+indx = sample(seq(1,nrow(aqw)),size=1000,replace=F)
 sub = aqw[indx,-c(1,2)]
 
 sub = aqw[,-c(1,2)]
@@ -288,7 +287,7 @@ coarse_df$Q = apply(coarse_df[,which(substr(names(sub),1,1)%in%c("Q"))],1,sum)
 coarse_df$OTHER = apply(coarse_df[,which(substr(names(sub),1,1)%in%c("S","B","G","H","L","K","T","D","M","P","Q")==FALSE)],1,sum)
[…]

### 2016-10-18 salmon/analysis/0-process-data.R

@@ -1,7 +1,14 @@
 library(dplyr)
 revenueThreshold = 10000
 
+#dat.diff = readRDS(file="portfolio/data-generated/cfec-diff-for-modeling.rds")
+
 dat.diff = readRDS(file="portfolio/data-generated/salmon-diff-for-modeling.rds")
+
+#group_by(cfec.diff[cfec.diff$permit=="S01E",], year) %>%
+#  summarize(g = sum(revenue))
+#group_by(dat.diff[dat.diff$permit=="S01E",], year) %>%
+#  summarize(g = sum(revenue))
 
 length(unique(dat.diff$p_holder))
 

### 2016-10-18 salmon/analysis/1-run-model.R

@@ -1,7 +1,8 @@
+library(dplyr)
 
 diffdat = readRDS(file="salmon/data-generated/salmon.rds")
 
-diffdat = diffdat[-which(is.na(diffdat$days.change)),]
+diffdat = filter(diffdat, !is.na(days.change))
 
 # prep data for stan
 N = nrow(diffdat) # number of rows of data

### 2016-10-18 salmon/analysis/process_salmon_data.R

@@ -51,7 +51,13 @@ early = rename(early, year=ADFG_B_BATCH_YEAR,
 
 early$g_price=ifelse(early$g_pounds>0,early$g_earn/early$g_pounds,0)
 
-early = early[early$p_fshy%in%c("S 01E", "S 03A", "S 01A", "S 03E", "S 01K", "S 01M"),]
+#early = early[early$p_fshy%in%c("S 01E", "S 03A", "S 01A", "S 03E", "S 01K", "S 01M"),]
+
+#early = group_by(early, p_holder) %>%
+#  mutate(npermit = length(unique(p_fshy))) %>%
+#  filter(npermit == 1) %>%
+#  select(-npermit) %>%
+#  filter(p_fshy %in% c("S 01E", "S 03A", "S 01A", "S 03E", "S 01K", "S 01M"))
 
 saveRDS(early, paste0("processed",y,".rds"))
 
@@ -67,12 +73,34 @@ for(y in 1976:1984) {
 saveRDS(dat, paste0("processed","_allYears",".rds"))
 
 # merge in the 1985+ data
-recent = readRDS("salmon_data.rds")
-recent = recent[,-1]
-recent$startdt = as.character(recent$startdt)
-recent$landdate = as.character(recent$landdate)
-dat = dat[,which(names(dat)%in%names(recent))]
-dat$f_ticket = as.character(dat$f_ticket)
+setwd("/users/eric.ward/downloads/cfec_early")
+older = readRDS("processed_allYears.rds")
 
-cfec = bind_rows(dat, recent)
+setwd("/users/eric.ward/documents/pfx-commercial/data")
+load("cfec_w_IDs")
+#cfecnew = group_by(cfecnew, p_holder, year) %>%
+#  mutate(npermit = length(unique(p_fshy))) %>%
+#  filter(npermit == 1) %>%
+#  select(-npermit) %>%
+#  ungroup() %>%
+#  filter(p_fshy %in% c("S 01E", "S 03A", "S 01A", "S 03E", "S 01K", "S 01M"))
+#cfecnew = cfecnew[which(cfecnew$p_fshy%in%c("S 01K", "S 01M", "S 01E", "S 01A","S 03E","S 03A")),]
+#saveRDS(cfecnew, "recent_salmon.rds")
+cfec = readRDS("recent_salmon.rds")
+
+# run process_raw_cfec_data header here
+
+cfec = cfec[,-1]
[…]

### 2016-10-24 salmon/paper/salmon_paper.Rmd
salmon paper edits from Alan

@@
   effects on variation in individual revenues'
-author: "Eric, Sean, Ole, Rich, Milo, Jordan, Jen, others"
+author: "Eric, Sean, Ole, Rich, Milo, Jordan, Jen, Alan, others"
 date: ''
@@
 
-On a large scale, individual fishers may experience positive outcomes from diversifying their catches; examples include reduced year-to-year variability in annual revenue, and increases in average revenue.
+On a large scale, individual fishers may experience positive outcomes from diversifying their catches; examples include reduced year-to-year variability in annual revenue and increases in average revenue.
 Fishers can increase their species diversity by owning a permit that allows multiple species to be targeted, or by owning multiple permits.
-External forces such as management regulations (limited entry,  catch shares) or market forces (permit prices) may disincentive individuals to hold multiple permits.
+External forces such as management regulations (limited entry,  catch shares) or market forces (permit or fish prices) may disincentive individuals to hold multiple permits.
 Thus, individuals with a single permit may increase species diversity within the confines of their permit (longer fishing seasons, broader spatial area).
-Using a long-term dataset of individual salmon fishers in the Alaska, we develop new models for understanding how the effects of diversification have changed through time.
+Using a long-term dataset of individual salmon fishers in Alaska, we develop new models for understanding how the effects of diversification have changed through time.
 These changes allow us to compare patterns across regions and gear, as well as better understand how external forces have changed incentives for diversifying (hatchery production and price trends, EVOS?).
@@
 Previous analyses have largely focused on patterns of diversification across ecosystems and the benefits to adopt more or less diverse fishing strategies [@kasperski_income_2013; @anderson_benefits].
-Fisheries in Alaska have been studied extensively, because they are one of the most productive regions in North America (both by landed weight, and revenue [@office_of_science_and_technology_fisheries_2016]) and they allow individuals to adopt a wide range of fishing strategies.
+Fisheries in Alaska have been studied extensively, because they are one of the most productive seafood regions in the world (both by landed weight, and revenue [@office_of_science_and_technology_fisheries_2016], FAO cite) and they allow individuals to adopt a wide range of fishing strategies.
 [@anderson_benefits] defined strategies as a permit or collection of permits held by single individuals.
@@
 In terrestrial agriculture, there are numerous examples where diversification may affect the mean or CV but not both [@purdy_financial_1997], and the same result may be true for salmon fishers.
-Further, these impacts of diversification may be variable over time.
+Further, these impacts of diversification are variable over time.
 
@@
 nerka*], coho [*O.
-kisutch*]), and the ability of individual salmon fishers to diversify across species is in part a function of geography and gears.
+kisutch*]), and the ability of individual salmon fishers to diversify across species is in part a function of geography and fishing gears.
 For example, salmon may be caught with purse seine gear (n = 6 areas), drift gillnets (5 areas), and set gillnets (12 areas).
@@
 Following low returns and declining productivity in the mid 1980s, many hatchery programs in the state rapidly increased hatchery releases [@mahnken_historical_1998; @leber_stock_2008].
-Many of these hatchery programs increased production to stabilize the variability of fishers catch (and revenue), though these programs may also have costs to wild production [@ruggerone_productivity_2015].
+Many of these hatchery programs increased production to stabilize the variability of fishers catch (and revenue), though these programs may also have impacts on wild production [@ruggerone_productivity_2015].
 The intra- and inter-specific effects of large scale hatchery programs has been debated [@hilborn_review_2001; @hilborn_hatcheries_1992; @wertheimer_comment:_2001; @heard_hatchery], but the effects of hatcheries have largely been focused on ecological interactions of aggregate yield to the fishery [@leber_stock_2008] or economic impacts [boyce_alaska_1993], rather than to variability experienced by individual fishermen.
@@
 
-In addition to the changing species composition of fish avaialble to be harvested, the benefits (or costs) of diversification are likely affected by a number of exogenous drivers, including marketing, prices, new processing technology [@knapp_trends].
+In addition to the changing species composition of fish avaialble to be harvested, the benefits (or costs) of diversification are affected by a number of exogenous drivers, including marketing, prices, new processing technology [@knapp_trends].
 [Break out as own paragraph+add more examples?] Following price crashes in the early 2000s, many of these factors have been positive for salmon fishermen (increasing trends in prices and catches, Fig.
@@
 We adjusted for inflation by setting all revenues in 2009 USD.
-Because were were only interested in single permit fisheries, we adopted a simplified version of the filtering steps described in [@anderson_benefits].
-After aggregating individual fish tickets to annual data, we retained individuals who only fished one of the six salmon permits included in a year (Table 1, Fig.
+Because we were only interested in single permit fisheries, we adopted a simplified version of the filtering steps described in [@anderson_benefits].
+After aggregating individual trip-level fish tickets to annual data, we retained individuals who only fished one of the six salmon permits included in a year (Table 1, Fig.
 A5).
@@
 [@ward_evaluating] summarized the trends for Prince William Sound).
-Release patterns have been dominated by pink and chum salmon (for PWS, these species represent 80% and 18% of total releases for 2012-2014 respectively, [@ward_evaluating]) and production has increased since the late 1970s (217 million juvenile pink salmon in PWS in 1985 versus 672 million in 2014).
+Release patterns have been dominated by pink and chum salmon (in recent years 2012-2014 in PWS, for example, pink and chum salmon represent 80% and 18% of total releases, [@ward_evaluating]) and production has increased since the late 1970s (217 million juvenile pink salmon in PWS in 1985 versus 672 million in 2014).
 Though our analysis with CFEC data doesn't separate hatchery versus wild fish, as a contrast we can make coarse comparisons across regions and species to examine potential effects of hatchery fish on variability.
-For example, we can compare the variation in pink salmon returns (earnings or weight) from Prince William Sound (largely hatchery) to those from Southeast Alaska (largely wild).
+For example, we can compare the variation in the largely hatchery pink salmon returns (earnings and weight) from Prince William Sound to the largely wild pink salmon returns from Southeast Alaska.
 Similarly, we can compare the variability in earnings from species dominated by hatchery returns (pink, chum) to those with less hatchery influence (Chinook, sockeye).
@@
 
-One of the challenges in modeling the effects of species diversification for individual fishers is that between two years, both the species composition of the catch and the species composition of the fish available to be caught may change (making it hard to disentangle the selectivity of fishermen versus portfolio of fish available).
+One of the challenges in modeling the effects of species diversification for individual fishers is that between two years, both the species composition of the catch and the species composition of the fish available to be caught change (making it hard to disentangle the selectivity of fishermen versus portfolio of fish available).
 To isolate the effects of a changing environment from changing strategies, we used a subset of the data from fishery permits where discrete strategies were emergent.
@@
 A9-A10).
-For pink and chum salmon, fisheries in PWS generally experience higher year to year variability compared to SEAK.
+For pink and chum salmon, fisheries in PWS generally experience higher year-to-year variability compared to SEAK.
 In looking at the changing CV of individuals, we see a slight increase in the CV of revenue for people fishing in the recent decade who were also fishing 1985-1995 (with the exception being S01A).
@@
 
-Figure 3: One component of variability is the year-to-year variation that individuals experience as a result of external forces (stock abundance, prices, etc).
-Or common fluctuations in revenue experienced by all permit holders [@anderson_benefits].
-This shows the time varying intercepts on the mean -- negative equals losing revenue on average, positive = increasing revenue.
-This shows big differences between permit types (gillnets tend to have lower year-to-year variation).
-Some permits have extreme variation - S01E (PWS seine) has fluctuations between exp(-1) and exp(+1), corresponding to year-to-year changes of 36-270%.
+Figure 3 shows that one component of variability is the year-to-year variation that individuals experience as a result of external forces (stock abundance, prices, etc) as wekk as common fluctuations in revenue experienced by all permit holders (Anderson et al., n.d.).
+This shows the time varying intercepts on the mean -- negative equals losing revenue on average, while positive implies increasing revenue.
+The panels of this figure show big differences between permit types (gillnets tend to have lower year-to-year variation).
+Some permits have extreme variation; for example, S01E (PWS seine) has fluctuations between exp(-1) and exp(+1), corresponding to year-to-year changes of 36-270%.
 
 Figure 4 shows the time varying nature of of the model for all permits.
-These are the time series of benefits of changing from effective species diversity of 1 to 2, and here we're basically isolating year-to-year differences in revenue (the intercepts) from diversification benefits.
-In some cases the effect is positive (fishing more spp = more revenue).
-But in many cases it's not, and the trend is negative (indicating that specialization is more profitable over time).
+These are the time series of benefits of changing from effective species diversity of 1 to 2, and here we are isolating year-to-year differences in revenue (the intercepts) from diversification benefits.
+In some cases the effect is positive (fishing more spp yields more revenue).
+But in many cases this is not the case, and the trend is negative (indicating that specialization is more profitable over time).
 In some cases, these time series mirror the aggregate diversity in the raw time series (Fig.
@@
 - Ability to specialize primarily driven by many forces beyond control of fishermen.
-But in some situations, diversification is possible by extending fishing season to exploit additional spp (fishing earlier for chum in PWS S01E).
+But in some situations, diversification is possible by extending the fishing season to exploit additional spp (fishing earlier for chum in PWS S01E).
 Or fishing different areas to exploit more spp in S03 (SEAK).
@@
 We're also seeing a trend toward greater synchrony between species (Fig A11-A12, A14).
-Projecting future prices are complex [@knapp_trends].
+Projecting future prices is notably complex [@knapp_trends].
 Synchrony in prices may largely be function of marketing, improved product, freezing, etc.
@@
 Correlation between them is ~ 0.8.
-This suggests (as expected) people on average make a lot more money in years when pink salmon returns are high."}
+This suggests (as expected) people on average make a lot more money in years when pink salmon returns are high.
+[side note: diversification may only be beneficial here when returns and prices are both low.]"}
 

### 2016-10-24 salmon/paper/salmon_paper.Rmd
typos

@@
 A1), median revenue is increasing (across individuals, Fig.
-A2), and landed species diversity declining over time (averaged across indivduals, Fig.
+A2), and landed species diversity declining over time (averaged across individuals, Fig.
 A6).
@@
 
-In addition to the changing species composition of fish avaialble to be harvested, the benefits (or costs) of diversification are affected by a number of exogenous drivers, including marketing, prices, new processing technology [@knapp_trends].
+In addition to the changing species composition of fish available to be harvested, the benefits (or costs) of diversification are affected by a number of exogenous drivers, including marketing, prices, new processing technology [@knapp_trends].
 [Break out as own paragraph+add more examples?] Following price crashes in the early 2000s, many of these factors have been positive for salmon fishermen (increasing trends in prices and catches, Fig.
@@
 In contrast, the late 1980s - early 1990s experienced high abundance and low prices (in part due to collapses of foreign currencies).
-Combined with the increased trend in hatchery production and resulting changes in species compositon of adult salmon, these external factors may have also altered incentives for diversification.
-If there's an incentive to substitute high value species (Chinook, sockeye) with more abundant low value species (pink, chum), the analog in the terrestrial framring world would be crop substitution of low-volume high value crops (spices) with high-volume low value crops (cereals) (other examples in literature = cocoa, tobacco).
+Combined with the increased trend in hatchery production and resulting changes in species composition of adult salmon, these external factors may have also altered incentives for diversification.
+If there's an incentive to substitute high value species (Chinook, sockeye) with more abundant low value species (pink, chum), the analog in the terrestrial farming world would be crop substitution of low-volume high value crops (spices) with high-volume low value crops (cereals) (other examples in literature = cocoa, tobacco).
 
@@
 As a second comparison, we examined whether long-term participants in any of the fisheries had experienced a change in variability.
-Similar to [@kasperski_income_2013] who calculated CVs of vessels before and after significant managment events, we identified individuals who fished the same permit for at least three years in the first decade of our data (1985-1995) and at least three years in the last decade of the data (2004-2014).
+Similar to [@kasperski_income_2013] who calculated CVs of vessels before and after significant management events, we identified individuals who fished the same permit for at least three years in the first decade of our data (1985-1995) and at least three years in the last decade of the data (2004-2014).
 
@@
 We assumed the residuals from the mean model described above to be normally distributed, $$\log\left( { R }_{ i,t } \right) \sim \mathrm{Normal}\left( \widehat { \log\left( { R }_{ i,t } \right)  } ,\widehat { { \sigma  }_{ i,t } }  \right) $$.
-While conventional hierarchcial linear regression assumes variances to be constant or shared by grouping variables, we also included fixed and random covariates, making the model akin to variance function regression or stochastic volatility models (CITE).
+While conventional hierarchical linear regression assumes variances to be constant or shared by grouping variables, we also included fixed and random covariates, making the model akin to variance function regression or stochastic volatility models (CITE).
 Using the same notation as for the mean, we can express the log standard deviation as $$\widehat { \log\left( { \sigma  }_{ i,t } \right)  } ={ G }_{ 0,j }+{ G }_{ 1,j,t }\cdot \Delta S_{ i,t }+{ G }_{ 2,j,t }\cdot \left( \Delta S_{ i,t }\cdot { S }_{ i,t-1 } \right) +G_{ 3,j }\cdot \Delta D_{ i,t }$$.
@@
 
-Figure 3 shows that one component of variability is the year-to-year variation that individuals experience as a result of external forces (stock abundance, prices, etc) as wekk as common fluctuations in revenue experienced by all permit holders (Anderson et al., n.d.).
+Figure 3 shows that one component of variability is the year-to-year variation that individuals experience as a result of external forces (stock abundance, prices, etc) as well as common fluctuations in revenue experienced by all permit holders (Anderson et al., n.d.).
 This shows the time varying intercepts on the mean -- negative equals losing revenue on average, while positive implies increasing revenue.
@@
 
-Figure 4 shows the time varying nature of of the model for all permits.
+Figure 4 shows the time varying nature of the model for all permits.
 These are the time series of benefits of changing from effective species diversity of 1 to 2, and here we are isolating year-to-year differences in revenue (the intercepts) from diversification benefits.
@@
 Figure 5 shows that temporal trends, by permit, in going from 1 to 2 species fished on the variance (again removing the overall intercepts).
-From [@anderson_benefits] we'd expect diversification to redudce variance.
-A few examples of how to read this: S01A is slightly positve for most of the time series, but rapidly increases in the last 10 years -- this means generalists with diverse portfolios have had increases in year-to-year variability.
+From [@anderson_benefits] we'd expect diversification to reduce variance.
+A few examples of how to read this: S01A is slightly positive for most of the time series, but rapidly increases in the last 10 years -- this means generalists with diverse portfolios have had increases in year-to-year variability.
 For S01E, the start and end of the time series are associated with these 'costs' to diversification.

### 2016-10-24 salmon/paper/salmon_paper.Rmd
updates from rich

@@
   word_document: default
-bibliography: refs.bib
+bibliography: salmon.bib
 ---
@@
 Fishers can increase their species diversity by owning a permit that allows multiple species to be targeted, or by owning multiple permits.
-External forces such as management regulations (limited entry,  catch shares) or market forces (permit or fish prices) may disincentive individuals to hold multiple permits.
-Thus, individuals with a single permit may increase species diversity within the confines of their permit (longer fishing seasons, broader spatial area).
+External forces such as management regulations (limited entry,  catch shares), the number of permits available, or market forces (permit or fish prices) may disincentive individuals to hold multiple permits.
+Thus, individuals holding a single permit may only be able to increase species diversity within the confines of their permit (longer fishing seasons, broader spatial area).
 Using a long-term dataset of individual salmon fishers in Alaska, we develop new models for understanding how the effects of diversification have changed through time.
-These changes allow us to compare patterns across regions and gear, as well as better understand how external forces have changed incentives for diversifying (hatchery production and price trends, EVOS?).
+These changes allow us to compare patterns across regions and gear, as well as better understand how external forces have changed incentives for diversifying (hatchery production, market forces, EVOS?).
 
@@
 
-Understanding the effects of species diversification on salmon fishers is particularly important, as the number of salmon specialists (defined as fishers holding a single salmon permit) has increased over the last 30 years as a fraction of the total [@anderson_benefits].
+Understanding the effects of species diversification on salmon fishers is particularly important, as the number of salmon specialists (defined as fishers holding a single salmon permit) has increased over the last 30 years as a fraction of the total [@anderson_benefits] (figure XX).
 For these limited entry salmon fisheries, general trends that have emerged are that more people are participating since the early 2000s (Fig.
@@
 A6).
-Across different salmon fisheries, other than a slight negative effect of diversification on variability (similar to [@anderson_benefits]), there appears to be little correlation between diversity and revenue (Fig.
+Across different salmon fisheries -- other than a slight negative effect of diversification on variability (similar to [@anderson_benefits]) -- there appears to be little correlation between diversity and revenue (Fig.
 A4).
-In terrestrial agriculture, there are numerous examples where diversification may affect the mean or CV but not both [@purdy_financial_1997], and the same result may be true for salmon fishers.
+In terrestrial agriculture, there are numerous examples where diversification may affect the mean or CV of revenue, but not both [@purdy_financial_1997], and the same result may be true for salmon fishers.
 Further, these impacts of diversification are variable over time.
 
-Five species of salmon are encountered in Alaska waters (Chinook [*Oncorhynchus tshawytscha*], chum [*O.
+Five species of salmon are harvested in Alaska waters (Chinook [*Oncorhynchus tshawytscha*], chum [*O.
 keta*], pink [*O.
@@
 A secondary motivation for focusing on the effects of diversification for salmon fishers is to better understand the long term consequences of hatchery salmon production on the revenue and variability in revenue of individuals.
-Following low returns and declining productivity in the mid 1980s, many hatchery programs in the state rapidly increased hatchery releases [@mahnken_historical_1998; @leber_stock_2008].
-Many of these hatchery programs increased production to stabilize the variability of fishers catch (and revenue), though these programs may also have impacts on wild production [@ruggerone_productivity_2015].
-The intra- and inter-specific effects of large scale hatchery programs has been debated [@hilborn_review_2001; @hilborn_hatcheries_1992; @wertheimer_comment:_2001; @heard_hatchery], but the effects of hatcheries have largely been focused on ecological interactions of aggregate yield to the fishery [@leber_stock_2008] or economic impacts [boyce_alaska_1993], rather than to variability experienced by individual fishermen.
-[Other history here?]
+The modern era of Alaska’s large-scale hatchery program began in the late 1970s (Stopha 2015, Rich will get a reference), with substantial increases in hatchery production during the 1980s following low returns [@mahnken_historical_1998; @leber_stock_2008-1].
+Many of these hatchery programs increased production to stabilize the variability of salmon returns [@hilborn_review_2001], though these programs may also have negative impacts on wild production [@wertheimer_relationship_2005; @ruggerone_productivity_2015].
+The intra- and inter-specific effects of large scale hatchery programs has been debated [@hilborn_review_2001; @hilborn_hatcheries_1992; @wertheimer_comment:_2001; @heard_hatchery], but the effects of hatcheries have largely been focused on ecological interactions of aggregate yield to the fishery [@leber_stock_2008-1] or economic impacts [@boyce_alaska_1993], rather than to variability experienced by individual fishermen.
 
@@
 
-In addition to the changing species composition of fish available to be harvested, the benefits (or costs) of diversification are affected by a number of exogenous drivers, including marketing, prices, new processing technology [@knapp_trends].
-[Break out as own paragraph+add more examples?] Following price crashes in the early 2000s, many of these factors have been positive for salmon fishermen (increasing trends in prices and catches, Fig.
+In addition to the changing species composition of fish available to be harvested, the benefits (or costs) of diversification are affected by a number of exogenous drivers, including the influx of farmed salmon, consolidation of salmon buyers, marketing, prices, new processing technology [@knapp_trends].
+[Break out as own paragraph+add more examples?] Following price crashes in the 1990s and early 2000s (cite CFEC), many of these factors have been positive for salmon fishermen (increasing trends in prices and catches, Fig.
 A3, A12).
@@
 
-We obtained fisheries landings and revenue data for all permit holders in Alaska from 1985 - 2014 from the Commercial Fisheries Entry Commission (CFEC).
+We obtained fisheries landings and revenue data for all permit holders in Alaska from 1985-2014 from the Commercial Fisheries Entry Commission (CFEC).
 We adjusted for inflation by setting all revenues in 2009 USD.
@@
 A5).
-Like [@anderson_benefits] we also worked with a log-differenced dataset to remove non-stationary individuals; to avoid complications of individuals switching strategies, we only included pairs of years for a given person where they did not change strategies.
+Like [@anderson_benefits] we also worked with a log-differenced dataset to remove non-stationary individuals; to avoid complications of individuals switching strategies, we only included pairs of years for a given person where they did not change strategies (i.e., didn’t change fisheries or add permits, etc.).
 
@@
 [@ward_evaluating] summarized the trends for Prince William Sound).
-Release patterns have been dominated by pink and chum salmon (in recent years 2012-2014 in PWS, for example, pink and chum salmon represent 80% and 18% of total releases, [@ward_evaluating]) and production has increased since the late 1970s (217 million juvenile pink salmon in PWS in 1985 versus 672 million in 2014).
+Release patterns have been dominated by pink and chum salmon ([@ward_evaluating], cite more specific #?) and production has increased since the late 1970s (217 million juvenile pink salmon in PWS in 1985 versus 672 million in 2014).
 Though our analysis with CFEC data doesn't separate hatchery versus wild fish, as a contrast we can make coarse comparisons across regions and species to examine potential effects of hatchery fish on variability.
@@
 
+[Rich's note: This and delta D might be tricky for pink salmon stocks, which alternate between strong even or odd years (depends on the area) and may dictate fishery participation.
+For these, could we, should we, use log(St/St-2)??]
+
 All done in Stan [@stan_development_team_stan_2015] using the NUTS algorithm [@hoffman_no-u-turn_2014].
@@
 Does diversification exist within permits and if so, do we see similar advantages of diversification on revenue + CV?
+
+PWS chum and sockeye (increasing hatchery % over time) vs. Bristol Bay (all wild).
+
+CV of variability for high value species (sockeye, Chinook) vs.
+low value species (pink and chum).
+Coho fall somewhere in the middle.
+In other words, how have choices of species to produce by hatcheries affected the CV of revenue for individual permits?
+% hatchery fish of the harvest could be used as a continuous variable in the model.
+
+Similar to above, do life-history characteristics influence the variability of revenue?
+For example, sockeye and chum have substantially more complicated life-history traits (1-5 ocean years) than pink salmon (1 ocean year only).
+For high-volume hatcheries such as pink salmon: do annual returns affect price and revenue the following year more than for high-value fisheries?
 
@@
 - In terms of participation, why did people drop out in 2000 in SEAK and Alaska Peninsula, but not in PWS?
-PWS largely seems like a boutique fishery on the Copper River -- which includes a qualtiy component (Copper River at forefront) and a timing issue relative to the rest of the Gulf of Alaska populations (buffers against market impacts)
+PWS largely seems like a boutique fishery on the Copper River -- which includes a qualtiy component (Copper River at forefront) and a timing issue relative to the rest of the Gulf of Alaska populations (buffers against market impacts).
+In addition, there are many indications that costs have increased over time as a result of requirements by fish buyers for fish of higher quality (refrigerated sea water, ice, increased handling time, more expensive vessels) more expensive permits, etc.
 
@@
 
-- One pontential exogenous driver related to management that can be ruled out as being responsible for increased specialization of salmon fishermen is changes in the timing or length of fishing seasons.
+- One potential exogenous driver related to management that can be ruled out as being responsible for increased specialization of salmon fishermen is changes in the timing or length of fishing seasons.
 Shifting seasons to later in the year may reduce catch of early run fish for example.
@@
 A15).
+Also, management actions over time have generally tried to eliminate interception fisheries.
+Examples include the closure of the South Kayak Island area where sockeye bound for Cook Inlet were being harvested off the Copper River.
+Rich can provide some citations.
+
+- The decisions to implement a hatchery program for a given species has economic consequences for overall revenue and the variability of that revenue to fishers and other participants.
+Salmon that have complex life histories in the ocean phase of their life cycle (sockeye and possibly chum salmon) may have an inherent portfolio that buffers against large annual swings in abundance compared to pink and coho salmon.
+Also, high-value species may experience less annual fluctuation than lower value species.
+
+- The potential to diversify may be limited for areas with a limited number of available species or by managements actions that prohibit fishing in certain areas (e.g., PWS where seiners are generally confined to areas/times with pink and chum salmon whereas gillnet fishers are able to fish in areas with sockeye, chum, Chinook, coho, and pink salmon).
+Rich can provide a citation for the PWS management plan.
 

### 2016-10-25 salmon/paper/salmon_paper.Rmd
paper and bib update

@@
 Combined with the increased trend in hatchery production and resulting changes in species composition of adult salmon, these external factors may have also altered incentives for diversification.
-If there's an incentive to substitute high value species (Chinook, sockeye) with more abundant low value species (pink, chum), the analog in the terrestrial farming world would be crop substitution of low-volume high value crops (spices) with high-volume low value crops (cereals) (other examples in literature = cocoa, tobacco).
+
+Tradeoffs between specialization and diversification may also be found in case studies of terrestrial agriculture.
+For small scale farmers, increasing crop diversity may increase revenue and stabilize year-to-year income variability [@di_falco_crop_2003].
+At larger scales or in the developed world, however, the incentives for diversification may be different.
+In the United States and Canada for example, crop diversity has declined since the late-1970s [@aguilar_crop_2015; @bradshaw_farm-level_2004]; some of the factors responsible include changing national policies, technology, and increased demand from other parts of the world (e.g.
+increased U.S.
+production of soybeans to meet demand from Asia).
+Shifts to from diverse of high value cash-crops to specializing on high-volume low-value crops such as cereals may have analogs for salmon fishermen in low value species with rapid life histories (pink and chum salmon).
 
@@
 
+# chinook chum coho pink sock
+cbPalette <- c("#009E73", "#4B2E83", "#999999", "#CC79A7", "#D55E00")
+
+dat$col = as.numeric(as.factor(dat$species))
 ggplot(dat, aes(Year, value)) +
-  geom_area(aes(fill = species), position = "stack", colour = 1) +
+  geom_area(aes(fill = cbPalette[col]), position = "stack") +
   xlab("Year") + ylab("Harvest of hatchery produced salmon (millions)") +
+  scale_fill_manual(values = cbPalette) + theme(legend.position="none") + 
   facet_wrap(~region, scale="free_y")
@@
 ```{r, echo=FALSE, fig.pos="placeHere", fig.cap="Time varying species composition and diversity, by permit.
-Species are colored as follows: yellow (chum), pink (sockeye), red (Chinook), green (coho), blue (pink salmon).", fig.height=3.2}
+Species are colored as follows: green (Chinook), purple (chum), silver  (coho), pink (pink), red (sockeye).", fig.height=3.2}
 cfec <- feather::read_feather("../../portfolio/data-generated/salmon.feather")
@@
 # plot proportion of species landed by salmon permit
+
 g1 = group_by(cfec[cfec$p_fshy=="S 01A" & cfec$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
-  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p)) %>%
+  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
   ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
-  ylab("Species %") + xlab("") + ggtitle("S01A") + theme(legend.position="none")
+  ylab("Species %") + xlab("") + ggtitle("S01A")+ theme(legend.position="none") + 
+scale_colour_manual(values=cbPalette)
 
 g2 = group_by(cfec[cfec$p_fshy=="S 01E" & cfec$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
-  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p)) %>%
+  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
   ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
-  ylab("Species %") + xlab("") + ggtitle("S01E") + theme(legend.position="none")
+  ylab("Species %") + xlab("") + ggtitle("S01E") + theme(legend.position="none") + 
+scale_colour_manual(values=cbPalette)
 
 g3 = group_by(cfec[cfec$p_fshy=="S 01K" & cfec$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
-  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p)) %>%
+  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
   ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
-  ylab("Species %") + xlab("") + ggtitle("S01K") + theme(legend.position="none")
+  ylab("Species %") + xlab("") + ggtitle("S01K") + theme(legend.position="none") + 
+scale_colour_manual(values=cbPalette)
 
 g4 = group_by(cfec[cfec$p_fshy=="S 01M" & cfec$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
-  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p)) %>%
+  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
   ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
-  ylab("Species %") + xlab("") + ggtitle("S01M") + theme(legend.position="none")
+  ylab("Species %") + xlab("") + ggtitle("S01M") + theme(legend.position="none") + 
+scale_colour_manual(values=cbPalette)
 
 g5 = group_by(cfec[cfec$p_fshy=="S 03A" & cfec$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
-  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p)) %>%
+  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
   ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
-  ylab("Species %") + xlab("") + ggtitle("S03A") + theme(legend.position="none")
+  ylab("Species %") + xlab("") + ggtitle("S03A") + theme(legend.position="none") + 
+scale_colour_manual(values=cbPalette)
 
 g6 = group_by(cfec[cfec$p_fshy=="S 03E" & cfec$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK") & cfec$year > 0,], year, spec, p_fshy) %>%
-  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p)) %>%
+  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
   ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
-  ylab("Species %") + xlab("") + ggtitle("S03E") + theme(legend.position="none")
+  ylab("Species %") + xlab("") + ggtitle("S03E") + theme(legend.position="none") + 
+scale_colour_manual(values=cbPalette)
 

### 2016-10-27 salmon/paper/salmon_paper.Rmd
updated Fig 1

@@
 
-seak$value[which(is.na(seak$value))] = 0
-pws$value[which(is.na(pws$value))] = 0
-kodiak$value[which(is.na(kodiak$value))] = 0
+seak$value[is.na(seak$value)] = 0
+pws$value[is.na(pws$value)] = 0
+kodiak$value[is.na(kodiak$value)] = 0
 
@@
 ggplot(dat, aes(Year, value)) +
-  geom_area(aes(fill = cbPalette[col]), position = "stack") +
+  geom_area(aes(fill = species), position = "stack") +
   xlab("Year") + ylab("Harvest of hatchery produced salmon (millions)") +
-  scale_fill_manual(values = cbPalette) + theme(legend.position="none") + 
+  scale_fill_manual(values = c('Chinook'=cbPalette[1], 'Chum'=cbPalette[2], 
+    'Coho'=cbPalette[3], 'Pink'=cbPalette[4], 'Sockeye'=cbPalette[5]), name="Species",
+    labels =c("Chinook","Sockeye","Coho","Pink","Chum")) + 
   facet_wrap(~region, scale="free_y")

### 2016-10-28 salmon/data-generated/process_wattum_data.R
data processing

@@ -1,0 +1,12 @@
+
+# this processes the kodiak data
+dat = read_excel("salmon/data-generated/Table39_Wattum2015.xlsx", col_names = FALSE)
+kodiak = matrix(0, nrow(dat), 7)
+for(i in 1:nrow(dat)) {
+  kodiak[i,] = unlist(strsplit(gsub(",","",dat[i,]), " "))
+}
+kodiak[20,1]=1989
+kodiak = as.data.frame(as.matrix(kodiak))
+names(kodiak) = c("Year","Chinook","Sockeye","Coho","Pink","Chum","Total")
+write.table(kodiak, "salmon/data-generated/kodiak_totalHarvest.csv",sep=",",row.names=F,col.names=T)
+

### 2016-10-28 salmon/paper/salmon_paper_appendix.Rmd
appendix update

@@
 # Also load in raw CFEC data for plots
-cfec <- feather::read_feather("../../portfolio/data-generated/salmon.feather")
+#cfec <- feather::read_feather("../../portfolio/data-generated/salmon.feather")
+cfec = readRDS("../../portfolio/data-generated/salmon_data_for_paper.rds")
 cfec = cfec[cfec$p_fshy%in%c("S 01A", "S 01E", "S 01K", "S 03A", "S 03E", "S 01M"),]
@@
 dat.salmon = dat.diff[dat.diff$strategy_permit%in%c("S01A", "S01E", "S01K", "S03A", "S03E","S01M"),]
+
+# summarize annual data - total revenue and CV (3-year rolling)
+annual_summary = group_by(cfec[which(cfec$year>0),], year, p_fshy) %>% 
+  summarize(tot = sum(g_earn)) %>% arrange(p_fshy) %>% 
+  group_by(p_fshy) %>% mutate(rollmean = c(NA,rollmean(tot,k=3),NA),
+    rollsd = c(NA,rollapply(tot,width=3,sd),NA), rollcv = rollsd/rollmean)
+#ggplot(annual_summary, aes(year, rollcv,color=p_fshy)) + geom_line()
+#ggplot(annual_summary, aes(log(rollmean), rollcv, color=p_fshy)) + geom_point()
+
+# calculate total earnings by person / permit / year
+indiv_summary = group_by(cfec[which(cfec$year>0),], year, p_holder) %>% 
+  mutate(n = length(unique(p_fshy))) %>% filter(n == 1) %>% 
+  group_by(year, p_holder) %>% 
+  summarize(g = sum(g_earn), p_fshy = p_fshy[1])
+# add in missing data
+people_year = expand.grid("year"=unique(cfec$year[cfec$year>0]), "p_holder"=unique(indiv_summary$p_holder))
+indiv_summary = left_join(people_year, indiv_summary)
+# apply rolling mean / sd to data
+indiv_summary = group_by(indiv_summary, p_holder) %>% 
+  mutate(rollmean = c(NA,rollapply(g,width=3,mean),NA),
+    rollsd = c(NA,rollapply(g,width=3,sd),NA),
+    rollcv = rollsd/rollmean)
+# calculate averages across individuals over time
+indiv_summary = group_by(indiv_summary, p_fshy, year) %>% 
+  summarize(ind_mean = mean(rollmean,na.rm=T), 
+    ind_cv = mean(rollcv,na.rm=T))
+
+alldat = left_join(annual_summary, indiv_summary)
+# this shows cvs that individuals experience are same as cvs of aggregate
+#ggplot(alldat, aes(rollcv, ind_cv,color=p_fshy)) + geom_point()
+#ggplot(alldat, aes(log(rollmean), ind_cv,color=p_fshy)) + geom_point()
+#ggplot(alldat, aes(log(rollmean),log(ind_mean), color=p_fshy)) + geom_point(aes(alpha=year))
+#ggplot(alldat, aes(year, log(rollmean))) + facet_wrap(~p_fshy) + geom_line()
+#ggplot(alldat, aes(year, rollcv)) + facet_wrap(~p_fshy) + geom_line() + geom_line(aes(year, ind_cv, color="red"))
+#ggplot(alldat, aes(year, log(ind_cv/rollcv), col=p_fshy)) + geom_line()
+#ggplot(alldat, aes(log(rollcv), log(ind_cv),color=p_fshy)) + geom_point()
 
@@
 
+## Species diversity
+
+```{r, fig.cap="Proportion of species landed, weighted by earnings", fig.pos="placeHere", echo=FALSE, fig.height = 3.3, eval=FALSE}
 ## Species composition
-
-```{r, fig.cap="Proportion of species landed, weighted by earnings", fig.pos="placeHere", echo=FALSE, fig.height = 3.3}
 # plot proportion of species landed by salmon permit

### 2016-10-28 salmon/paper/salmon_paper.Rmd
paper update

@@
   word_document: default
-bibliography: salmon.bib
+bibliography: CatchPortfolios_salmon.bib
 ---
@@
 knitr::opts_chunk$set(echo = TRUE, tidy=FALSE, tidy.opts=list(width.cutoff=60), warning = FALSE, message = FALSE)
-library(dplyr)
 library(knitr)
-library(ggplot2)
+library(tidyverse)
 library(reshape2)
@@
 modelname = "model-linear" # model-linear-nointeractions
+
 ```
@@
 
-### Robustness of strategies to changing environments [stash in appendix?]
-
-One of the challenges in modeling the effects of species diversification for individual fishers is that between two years, both the species composition of the catch and the species composition of the fish available to be caught change (making it hard to disentangle the selectivity of fishermen versus portfolio of fish available).
-To isolate the effects of a changing environment from changing strategies, we used a subset of the data from fishery permits where discrete strategies were emergent.
-For details see Appendix.
-
-Using purse seine permit holders in Prince William Sound as an example, we first calculated the species composition (weighted by revenue) for each active permit holder in each year.
-For each pair of years that an individual fished, we calculated the squared Euclidian distance between the species compositions of those years, $$\Delta { p }_{ ind,t }=\sum _{ sp=1 }^{ 5 }{ { \left| { p }_{ ind,sp,t }-{ p }_{ ind,sp,t-1 } \right|  }^{ 2 } } $$.
-
-Larger values of $\Delta { p }_{ ind,t }$ indicate greater year-to-year discrepancy in species composition.
-To focus on individuals who had more constant strategies in adjacent years, we ommitted data corresponding to the upper third of this distribution (retaining $\Delta { p }_{ ind,t }$ < 0.334).
-Next, we used the species composition from each individual and pair of years to cluster individuals into strategies.
-Clustering was performed using the partitioning around medoids (pam) algorithm in R with an estimated number of clusters (using 'pamk' in fpc library), and selected 2 groups of individual strategies: pink salmon specialization (medoid proportion of pink salmon = 0.95), and a strategy focused on pink salmon and chum (medoid proportions of 0.64 and 0.31 respectively).
-For each group, we then calculated the mean and CV of revenue; we calculated the log difference of revenue for each individual, $\Delta { R }_{ i }=\log\left( { R }_{ i,t+1 }/{ R }_{ i,t } \right)$ and then calculated the CV as the standard deviation over people, $sd(\Delta { R }_{ i=1:n })$.
-
 ### Analysis 3: [Core of paper] Does effect of diversification change over time?
@@
 For high-volume hatcheries such as pink salmon: do annual returns affect price and revenue the following year more than for high-value fisheries?
-
 ## Results
@@
 
+- One of the outputs that can be calculated from the raw data is variability of revenue over time (e.g.
+year to year, at the level of individual & permit).
+This gets at more of the social impacts of changing opportunities, but long term trends in rising income variability have been documented in the US (Hacker 2006).
+Understanding trends in revenue of individuals whose incomes are dependent on natural resources are interesting because they tend to be more variable than that of the larger population [@mishra_income_2002].
+Are there specific attributes that correlate with rising variability (for example the CBO has suggested variability in year to year income is negatively correlated with income [@cbo_recent_2008])
+
+### Robustness of strategies to changing environments [stash in appendix?]
+
+One of the challenges in modeling the effects of species diversification for individual fishers is that between two years, both the species composition of the catch and the species composition of the fish available to be caught change (making it hard to disentangle the selectivity of fishermen versus portfolio of fish available).
+To isolate the effects of a changing environment from changing strategies, we used a subset of the data from fishery permits where discrete strategies were emergent.
+For details see Appendix.
+
+Using purse seine permit holders in Prince William Sound as an example, we first calculated the species composition (weighted by revenue) for each active permit holder in each year.
+For each pair of years that an individual fished, we calculated the squared Euclidian distance between the species compositions of those years, $$\Delta { p }_{ ind,t }=\sum _{ sp=1 }^{ 5 }{ { \left| { p }_{ ind,sp,t }-{ p }_{ ind,sp,t-1 } \right|  }^{ 2 } } $$.
+
+Larger values of $\Delta { p }_{ ind,t }$ indicate greater year-to-year discrepancy in species composition.
+To focus on individuals who had more constant strategies in adjacent years, we ommitted data corresponding to the upper third of this distribution (retaining $\Delta { p }_{ ind,t }$ < 0.334).
+Next, we used the species composition from each individual and pair of years to cluster individuals into strategies.
+Clustering was performed using the partitioning around medoids (pam) algorithm in R with an estimated number of clusters (using 'pamk' in fpc library), and selected 2 groups of individual strategies: pink salmon specialization (medoid proportion of pink salmon = 0.95), and a strategy focused on pink salmon and chum (medoid proportions of 0.64 and 0.31 respectively).
+For each group, we then calculated the mean and CV of revenue; we calculated the log difference of revenue for each individual, $\Delta { R }_{ i }=\log\left( { R }_{ i,t+1 }/{ R }_{ i,t } \right)$ and then calculated the CV as the standard deviation over people, $sd(\Delta { R }_{ i=1:n })$.
+
 \break
@@
   facet_wrap(~region, scale="free_y")
-
 ```
@@
 Species are colored as follows: green (Chinook), purple (chum), silver  (coho), pink (pink), red (sockeye).", fig.height=3.2}
-cfec <- feather::read_feather("../../portfolio/data-generated/salmon.feather")
+#cfec <- feather::read_feather("../../portfolio/data-generated/salmon.feather")
+#saveRDS(cfec, "portfolio/data-generated/salmon_data_for_paper.rds")
+cfec = readRDS("../../portfolio/data-generated/salmon_data_for_paper.rds")
 cfec$year = as.numeric(substr(cfec$landdate, 1, 4))
@@
   ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
-  ylab("Species %") + xlab("") + ggtitle("S01A")+ theme(legend.position="none") + 
-scale_colour_manual(values=cbPalette)
+  ylab("Species %") + xlab("") + ggtitle("S01A") + 
+scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
+    'COHO'=cbPalette[3], 'PINK'=cbPalette[4], 'SOCK'=cbPalette[5]), name="Species",
+    labels =c("Chinook","Chum","Coho","Pink","Sockeye")) + theme(legend.position="none")
 
@@
   ylab("Species %") + xlab("") + ggtitle("S01E") + theme(legend.position="none") + 
-scale_colour_manual(values=cbPalette)
+scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
+    'COHO'=cbPalette[3], 'PINK'=cbPalette[4], 'SOCK'=cbPalette[5]))
 
@@
   ylab("Species %") + xlab("") + ggtitle("S01K") + theme(legend.position="none") + 
-scale_colour_manual(values=cbPalette)
+scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
+    'COHO'=cbPalette[3], 'PINK'=cbPalette[4], 'SOCK'=cbPalette[5]))
 
@@
   ylab("Species %") + xlab("") + ggtitle("S01M") + theme(legend.position="none") + 
-scale_colour_manual(values=cbPalette)
+scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
+    'COHO'=cbPalette[3], 'PINK'=cbPalette[4], 'SOCK'=cbPalette[5]), name = "Species", labels =c("Chinook","Chum","Coho","Pink","Sockeye")) + theme(legend.position="none")
 
@@
   ylab("Species %") + xlab("") + ggtitle("S03A") + theme(legend.position="none") + 
-scale_colour_manual(values=cbPalette)
+scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
+    'COHO'=cbPalette[3], 'PINK'=cbPalette[4], 'SOCK'=cbPalette[5]))
 
@@
   ylab("Species %") + xlab("") + ggtitle("S03E") + theme(legend.position="none") + 
-scale_colour_manual(values=cbPalette)
+scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
+    'COHO'=cbPalette[3], 'PINK'=cbPalette[4], 'SOCK'=cbPalette[5]))
 

### 2016-10-28 portfolio/text/ms.Rmd
Copy editing

@@
 and sablefish in the same year,
-their strategy for that year would be `halibut-sablefish'.
+their strategy for that year would be "halibut-sablefish".
 Specific details on our strategy definitions are available in
@@
 Season length for a given permit
-was calculated as the the number of days between
+was calculated as the number of days between
 the first instance of fishing a permit
@@
 is the same for all values of the predictor variables ---
-the variance function regression model allows variance to be heteroskedastic. 
+the variance-function regression model allows the variance to be heteroscedastic. 
 For application to economic data on individual fishing
@@
 
-We extended the basic variance function regression to a hierarchical
-Bayesian variance function regression model with Gaussian errors
+We extended the basic variance-function regression to a hierarchical
+Bayesian variance-function regression model with Gaussian errors
 predicting $\log$ revenue, $\log(R_{i,t})$, for
@@
 (i.e. degrees of freedom of 3 and scale of 2)
-on all $\sigma$ parameters.
+on all $\sigma$ parameters to allow for heavier tails.
 We ran four chains and 2,000 iterations and discarded
@@
 and that the effective sample size,
-as calculated by the rstan package,
+as calculated by the rstan R package [@standevelopmentteam2015],
 was greater than 200 for all parameters [@gelman2014].

### 2016-10-28 salmon/paper/salmon_paper.Rmd
update

@@
 
-seak = read.csv("../../salmon/data-generated/stopha_table32.csv")
-pws = read.csv("../../salmon/data-generated/stopha_table33.csv")
-kodiak = read.csv("../../salmon/data-generated/stopha_table35.csv")
+seak_hatch = read.csv("../../salmon/data-generated/seak_hatcheryHarvest.csv")
+pws_hatch = read.csv("../../salmon/data-generated/pws_hatcheryHarvest.csv")
+kodiak_hatch = read.csv("../../salmon/data-generated/kodiak_hatcheryHarvest.csv")
 
-seak = melt(seak, id.vars = c("Year"))
-pws = melt(pws, id.vars = c("Year"))
-kodiak = melt(kodiak, id.vars = c("Year"))
+seak_hatch = melt(seak_hatch, id.vars = c("Year"))
+pws_hatch = melt(pws_hatch, id.vars = c("Year"))
+kodiak_hatch = melt(kodiak_hatch, id.vars = c("Year"))
 
-seak$region = "Southeast"
-pws$region = "Prince William Sound"
-kodiak$region = "Kodiak"
+seak_hatch$region = "Southeast"
+pws_hatch$region = "Prince William Sound"
+kodiak_hatch$region = "Kodiak"
 
-seak$value[is.na(seak$value)] = 0
-pws$value[is.na(pws$value)] = 0
-kodiak$value[is.na(kodiak$value)] = 0
+seak_hatch$value[is.na(seak_hatch$value)] = 0
+pws_hatch$value[is.na(pws_hatch$value)] = 0
+kodiak_hatch$value[is.na(kodiak_hatch$value)] = 0
 
-dat = bind_rows(seak, pws)
-dat = bind_rows(dat, kodiak)
+dat = bind_rows(seak_hatch, pws_hatch)
+dat = bind_rows(dat, kodiak_hatch)
 
@@
   facet_wrap(~region, scale="free_y")
+
+# bring in total harvest data (hatchery + wild)
+#seak = read.csv("../../salmon/data-generated/stopha_table32.csv")
+pws = read.csv("../../salmon/data-generated/pws_totalHarvest.csv")
+kodiak = read.csv("../../salmon/data-generated/kodiak_totalHarvest.csv")
+
+pws = melt(pws[,c("Year","Chinook","Sockeye","Coho","Pink","Chum")], id.vars = c("Year"))
+kodiak = melt(kodiak[,c("Year","Chinook","Sockeye","Coho","Pink","Chum")], id.vars = c("Year"))
+pws$region = "Prince William Sound"
+kodiak$region = "Kodiak"
+datall = bind_rows(pws,kodiak)
+names(datall)[which(names(datall)=="variable")]="species"
+datall$value = datall$value / 1000000
+names(datall)[which(names(datall)=="value")]="totalharvest"
+
+# join in hatchery harvest with total harvest
+dat = left_join(dat, datall)
+
+dat$percent_hatch = dat$value/dat$totalharvest
+dat$percent_hatch[is.na(dat$percent_hatch)] = 0
+dat$percent_hatch[dat$percent_hatch>1] = 1
+# chinook chum coho pink sock
+#cbPalette <- c("#009E73", "#4B2E83", "#999999", "#CC79A7", "#D55E00")
+
+ggplot(dat[dat$region%in%c("Kodiak", "Prince William Sound") & dat$Year < 2015,], aes(Year, percent_hatch, color=species)) +
+  geom_line() +
+  xlab("Year") + 
+  ylab("Proportion of hatchery harvest") +
+  facet_wrap(~region, scale="free_y")
+#  
+#  scale_fill_manual(values = c('Chinook'=cbPalette[1], 'Chum'=cbPalette[2], 
+#    'Coho'=cbPalette[3], 'Pink'=cbPalette[4], 'Sockeye'=cbPalette[5]), name="Species",
+#    labels =c("Chinook","Sockeye","Coho","Pink","Chum")) +
+
 ```

### 2016-10-28 salmon/data-generated/process_wattum_data.R

@@ -1,12 +1,0 @@
-
-# this processes the kodiak data
-dat = read_excel("salmon/data-generated/Table39_Wattum2015.xlsx", col_names = FALSE)
-kodiak = matrix(0, nrow(dat), 7)
-for(i in 1:nrow(dat)) {
-  kodiak[i,] = unlist(strsplit(gsub(",","",dat[i,]), " "))
-}
-kodiak[20,1]=1989
-kodiak = as.data.frame(as.matrix(kodiak))
-names(kodiak) = c("Year","Chinook","Sockeye","Coho","Pink","Chum","Total")
-write.table(kodiak, "salmon/data-generated/kodiak_totalHarvest.csv",sep=",",row.names=F,col.names=T)
-

### 2016-10-28 salmon/paper/salmon_paper.Rmd
paper update

--- 2016-10-28 salmon/paper/salmon_paper.Rmd
@@
 dat$col = as.numeric(as.factor(dat$species))
-ggplot(dat, aes(Year, value)) +
+g1 = ggplot(dat, aes(Year, value)) +
   geom_area(aes(fill = species), position = "stack") +
@@
   facet_wrap(~region, scale="free_y")
+
+# bring in total harvest data (hatchery + wild)
+hatch = read.csv("salmon/data-generated/commonPropertyCommercialCatch.csv")
+hatch = hatch[,c("Year","Chinook","Sockeye","Coho","Pink","Chum","Region")]
+hatch$Region = as.character(hatch$Region)
+hatch$Region[which(hatch$Region=="PWS")] = "Prince William Sound"
+hatch = melt(hatch, id.vars = c("Year","Region"))
+names(hatch)[which(names(hatch)=="variable")]="Species"
+
+g2 = ggplot(hatch, aes(Year, value, color=Species)) + 
+  geom_line() + facet_wrap(~Region) + 
+  xlab("Year") + ylab("Hatchery contribution to catch (%)") + 
+  scale_color_manual(values = c('Chinook'=cbPalette[1], 'Chum'=cbPalette[2], 
+    'Coho'=cbPalette[3], 'Pink'=cbPalette[4], 'Sockeye'=cbPalette[5]), name="Species",
+    labels =c("Chinook","Sockeye","Coho","Pink","Chum"))
+
+gridExtra::grid.arrange(g1, g2)
+
 ```

### 2016-10-28 salmon/paper/salmon_paper.Rmd
update

@@
 # bring in total harvest data (hatchery + wild)
-hatch = read.csv("salmon/data-generated/commonPropertyCommercialCatch.csv")
+hatch = read.csv("../../salmon/data-generated/commonPropertyCommercialCatch.csv")
 hatch = hatch[,c("Year","Chinook","Sockeye","Coho","Pink","Chum","Region")]

### 2016-11-10 diversity/analysis/herring_case_study.Rmd
herring case study

@@
 library(cluster)
+library(feather)
 ```
 
-```{r}
+## The Boom and Bust of Pacific Herring
+
+Pacific herring (*Clupea pallasii*) have been one of the focal species associated with the Exxon Valdez Oil Spill (EVOS) (Carls et al. 2002, Thorne and Thomas 2008, )
+
+have been one of the most important commercial and subsistence resources in the history of Alaska, with commercial fisheries established in 1878 (Woodby et al.
+2005).
+Herring were initially harvested for oil in the 1880s at Killisnoo (U.S.
+Department of Interior 1893; Skud et al.
+1960).
+For much of the next 50 years, the herring reduction fishery continued to expand, producing both bait and oil.
+Statewide, the herring reduction harvest was highest in 1937 and declined to the mid-1960s (Reid 1971).
+Prince William Sound experienced similar dynamics, with commercial harvest beginning as a bait fishery in 1913 (Rounsefell and Dahlgren 1935), and the reduction fishery peaking in 1937 (Funk 1994).
+Unlike more recent herring fisheries, these bait and reduction fisheries generally targeted schools of herring in the ocean.
+Over time, their size-selectivity became smaller, and they started targeting smaller fish for reduction (Funk 1994).
+In the 1950s, less than 10 boats were participating in the reduction fishery, and the continued decline in herring abundance, combined with cheaper altenratives from foreign markets (Peruvian anchovetta) led to the closure of the fishery in the late 1960s (Funk 1994).
+
+## Shifting demand: the graying of the palette
+
+In the early 1970s, new herring fisheries targeting a new herring product once again emerged in Prince William Sound.
+Declines of herring in Russia and Japan (Carlson 2005; Woodby et al.
+2005) prompted increased demand for salted herring roe (and more recently, luxury products like spawned herring eggs on kelp).
+Unlike the reduction fisheries, herring roe fisheries targeted herring near shore immediately before they spawned (spring), using a variety of gears, including gillnets, beach seines, and purse seines.
+
+Trends in revenue and landings in Prince William Sound fisheries largely mirror those of the rest of the state, with peaks in 1988 (Carlson 2005).
+Coincident with EVOS in 1989, harvest and prices  and declines through the 1990s.
+
+In addition to being susceptible to changing demand, a related pressure on Alaskan sac roe fisheries has been external markets.
+Over the period 1994-2003 for example, roe from Alaska represented a small fraction of the total supplied to Japan, especially compared to roe from Canada and China (Carlson 2005).
+
+## Changing Participation
+
+Declines in herring abundnace, demand (and prices) and ex-vessel values in Prince William Sound have been well documented (Carlson 2005). Following peak
+
+Participation
+
+\break
+
+```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Fishery landings for herring in Prince William Sound.
+L12E and L21E permits are combined for spawn on kelp, H01E and H34E are combined for food and bait, and G01E and G34E are combined for sac roe fisheries."}
+
+cfec <- read_feather("../../portfolio/data-generated/cfec.feather")
+cfec$year <- as.numeric(cfec$year)
+
+simp.div = function(x) {
+  1/sum((x/sum(x))^2)
+}
+
+pws = dplyr::filter(cfec, p_fshy%in%c("L 12E", "L 21E", "G 01E", "G 34E", "H 01E", "H 34E"))
+pws$p_fshy[pws$p_fshy%in%c("L 12E","L 21E")] = "Spawn on kelp" 
+pws$p_fshy[pws$p_fshy%in%c("H 01E","H 34E")] = "Food and bait"
+pws$p_fshy[pws$p_fshy%in%c("G 01E","G 34E")] = "Roe"
+
+names(pws)[which(names(pws)=="p_fshy")]="Fishery"
+g1 = group_by(pws[pws$year<=1998,], Fishery, year) %>% 
+  summarize(w = sum(g_pounds)) %>% 
+  ggplot(aes(year, w, group=Fishery, color=Fishery)) + geom_line() + 
+  xlab("Year") + ylab("Pounds")
+
+g2 = group_by(pws[pws$year<=1998,], Fishery, year) %>% 
+  summarize(w = length(unique(p_holder))) %>% 
+  ggplot(aes(year, w, group=Fishery, color=Fishery)) + geom_line() + 
+  xlab("Year") + ylab("Active permits")
+
+gridExtra::grid.arrange(g1, g2)
+```
+
+## Changing Strategies
+
+```{r, eval=FALSE}
 cfec <- feather::read_feather("../../portfolio/data-generated/cfec.feather")
@@
 
-this.permit = "S 03T"
+this.permit = "G 01E"
 # S01E S03T - effective permit diversity
@@
 }
+```
 
+```{r, eval=FALSE}
 g1 = group_by(cfec, p_holder) %>% 
@@
 
-```{r}
-
-```
-
-```{r}
+```{r,eval=FALSE}
 cfec <- feather::read_feather("../../portfolio/data-generated/cfec.feather")
@@
 
-```{r}
+```{r eval=FALSE}
 # grab people who ever held prince william sound herring permit
@@
 
-```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide"}
+```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide", eval=FALSE}
 cfec <- feather::read_feather("../../portfolio/data-generated/cfec.feather")
@@
 
-```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Herring permit holders by year."}
+```{r, eval=FALSE, echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Herring permit holders by year."}
   ggplot(herring_summary, aes(year, people)) + geom_line() + facet_wrap(~ p_fshy, scale="free_y") + xlab("Year") + ylab("People")
@@
 
-```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Trends in landed catch by permit.", fig.height=3.5}
+```{r, eval=FALSE, echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Trends in landed catch by permit.", fig.height=3.5}
   ggplot(herring_summary, aes(year, log(lbs))) + geom_line() + 
@@
 
-```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Trends in prices by permit.", fig.height=3.5}
+```{r, eval=FALSE,echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Trends in prices by permit.", fig.height=3.5}
   ggplot(herring_summary, aes(year, p)) + geom_line() + 
@@
 
-```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Scatterplots of prices versus revenue by permit."}
+```{r, eval=FALSE,echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Scatterplots of prices versus revenue by permit."}
   ggplot(herring_summary, aes(p, log(lbs/2000), colour = year)) + geom_point() + facet_wrap(~ p_fshy, scale="free") + ylab("Ln (tons)") + xlab("Price (dollars / lb.)")
@@
 
-```{r, echo = FALSE}
+```{r,eval=FALSE, echo = FALSE}
 
@@
 
-```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Proportion of permit holders specializing on herring.", fig.height=3.5}
+```{r, eval=FALSE,echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Proportion of permit holders specializing on herring.", fig.height=3.5}
 ggplot(herring_summary, aes(year, (only_herring_single+only_herring_multiple)/(only_herring_single+only_herring_multiple+herring_combo))) + facet_wrap(~p_fshy) + geom_line() + xlab("Year") + ylab("Proportion of people only fishing herring")
@@
 
-```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Proportion of permit holders diversifying spatially.", fig.height=3.5}
+```{r, eval=FALSE,echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Proportion of permit holders diversifying spatially.", fig.height=3.5}
 p1 = ggplot(herring_summary, aes(year, (g34)/(only_herring_single+only_herring_multiple+herring_combo), group = p_fshy, colour = p_fshy)) + geom_line() + xlab("Year") + ylab("Proportion of people diversifying spatially")
@@
 
-```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Proportion of people diversifying w/diff herring permits.", fig.height=3.5}
+```{r, eval=FALSE,echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Proportion of people diversifying w/diff herring permits.", fig.height=3.5}
 ggplot(herring_summary, aes(year, (multiple_fisheries)/(only_herring_single+only_herring_multiple+herring_combo))) + facet_wrap(~p_fshy) + geom_line() + xlab("Year") + ylab("Proportion of people diversifying w/diff herring permits")
@@
 
-```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Ratio of effective diversity to total number of permit holders (as measure of evenness).
+```{r, eval=FALSE,echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Ratio of effective diversity to total number of permit holders (as measure of evenness).
 Values near 1 indicate everyone's revenue is similar.", fig.height=3.5}
@@
 
-```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Fraction of revenue made up of herring (median, of per-person fractions).",eval=FALSE}
+```{r, eval=FALSE,echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Fraction of revenue made up of herring (median, of per-person fractions).",eval=FALSE}
 
@@
 ```
+
+## References
+
+Carls, M.G., G.D.
+Marty, and J.E.
+Hose.
+2002.
+Synthesis of the toxicological impacts of the Exxon Valdez oil spill on Pacific herring (Clupea pallasi) in Prince William Sound, Alaska, U.S.A.
+Canadian Journal of Fisheries and Aquatic Sciences, 2002, 59(1): 153-172.
+
+Carlson, S. 2005. Changes in roe herring markets: a review of available evidence. CFEC report number 05-5N. Commercial Fisheries Entry Commission, Juneau, AK.
+
+Funk, F. 1994. Forecast of the Pacific herring biomass in Prince William Sound, Alaska, 1993. Regional Information Report, 5J94-04. ADFG, Commercial Fisheries and Managmenet Division.
+
+Reid, G.M. 1971. Age composition, weight, length, and sex of herring, Clupea pallasii, used for 
+reduction in Alaska, 1929-1966.
+
+Rounsefell, G.A. and E.H. Dahlgren. 1935. Fluctuations in the supply of herring, Clupea pallasii, in Prince William Sound, Alaska. Bulletin of the U.S. Bureau of Fisheries 47:263-291.
+
+Skud, B.E., H.M. Sakuda, and G.M Reid. 1960. Statistics of the Alaska herring fishery, 1878-1956. U.S. Fish and Wildlife Service, Statistical Digest 48, 21 p.
+
+Thorne, R. E., and Thomas, G. L. 2008. Herring and the "Exxon Valdez" oil spill: an investigation into historical data conflicts. – ICES Journal of Marine Science, 65: 44–50.
+
+U.S. Department of Interior 1893. Population and Resources of Alaska at the Eleventh Census: 1890. U.S. Department of Interior, Census Office.
+
+Woodby, D., D. Carlile, S. Siddeek, F. Funk, J.H. Clark, and L. Hulbert. 2005. Commercial Fisheries in Alaska. ADFG special publication 05-09.  http://www.sf.adfg.state.ak.us/FedAidPDFs/sp05-09.pdf

### 2016-11-10 diversity/analysis/herring_case_study_v2.Rmd
herring case study

--- 2016-10-02 diversity/analysis/herring_case_study.Rmd
@@
 ---
-title: "Herring: case study 1"
+title: 'Herring Case Study'
 author: ''
@@
 library(date)
+library(reshape2)
+library(cluster)
+library(feather)
 ```
 
-```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide"}
-cfec <- feather::read_feather("../../portfolio/data-generated/cfec.feather")
-cfec$year <- as.numeric(cfec$year)
+### The Boom and Bust of Pacific Herring
 
-herring = cfec[substr(cfec$p_fshy,1,1) %in% c("L","G","H"),]
+Pacific herring (*Clupea pallasii*) have been one of the focal species associated with the Exxon Valdez Oil Spill (EVOS) (Carls et al. 2002, Thorne and Thomas 2008, )
 
-herring_summary = group_by(herring, p_fshy, year) %>%
-  summarize(g = sum(g_earn), p = mean(g_price,na.rm=T), lbs = sum(g_pounds), people = length(unique(p_holder))) %>%
-  group_by(p_fshy) %>%
-  mutate(n = length(unique(year)), maxpeople = max(people,na.rm=T)) %>%
-  filter(n > 9) %>%
-  filter(maxpeople > 25)
+have been one of the most important commercial and subsistence resources in the history of Alaska, with commercial fisheries established in 1878 (Woodby et al.
+2005).
+Herring were initially harvested for oil in the 1880s at Killisnoo (U.S.
+Department of Interior 1893; Skud et al.
+1960).
+For much of the next 50 years, the herring reduction fishery continued to expand, producing both bait and oil.
+Statewide, the herring reduction harvest was highest in 1937 and declined to the mid-1960s (Reid 1971).
+Prince William Sound experienced similar dynamics, with commercial harvest beginning as a bait fishery in 1913 (Rounsefell and Dahlgren 1935), and the reduction fishery peaking in 1937 (Funk 1994).
+Unlike more recent herring fisheries, these bait and reduction fisheries generally targeted schools of herring in the ocean.
+Over time, their size-selectivity became smaller, and they started targeting smaller fish for reduction (Funk 1994).
+In the 1950s, less than 10 boats were participating in the reduction fishery, and the continued decline in herring abundance, combined with cheaper altenratives from foreign markets (Peruvian anchovetta) led to the closure of the fishery in the late 1960s (Funk 1994).
 
-quitters = group_by(herring, p_holder, p_fshy) %>%
-  summarize(lastYear = max(year))
+### Shifting demand: the graying of the palate
 
-group_by(quitters[quitters$lastYear<2014,],lastYear,p_fshy) %>% 
-  summarize(n = n()) %>% 
-  ggplot(aes(lastYear,n)) + geom_line() + facet_wrap(~p_fshy, scale="free_y") + xlab("Year") + ylab("People leaving the fishery (last year fished)")
+In the early 1970s, new herring fisheries targeting a new herring product once again emerged in Prince William Sound.
+Declines of herring in Russia and Japan (Carlson 2005; Woodby et al.
+2005) prompted increased demand for salted herring roe (and more recently, luxury products like spawned herring eggs on kelp).
+Unlike the reduction fisheries, herring roe fisheries targeted herring near shore immediately before they spawned (spring), using a variety of gears, including gillnets, beach seines, and purse seines.
 
-# look at what these people did after leaving
-cfec_quitters = cfec[which(cfec$p_holder%in%unique(quitters$p_holder)),]
-cfec_quitters = left_join(cfec_quitters, quitters[,c("p_holder","lastYear")])
+Trends in revenue, biomass, and landings in Prince William Sound fisheries largely mirror those of the rest of the state, with peaks in 1988 (Pearson et al.
+2012; Carlson 2005).
+Coincident with EVOS in 1989, prices and harvest declined through much of the 1990s, with limited fisheries openings, and fisheries have remained closed since 1999 because estimated biomass has remained lower than that required for reopening the fishery (20000 short tons).
+Though not affected by EVOS, harvest of herring roe in other Alaska fisheries have also declined since the late 1980s; these include fisheries in Norton Sound, Bristol Bay, and Togiak (Woodby et al.
+2005).
+Perhaps the exception to this declining trend is roe harvest is in Sitka Sound (Thynes et al.
+2016;  Woodby et al.
+2005).
+The timing of herring spawn and harvest typically follows a latitudinal gradient from south to north (Carlson 2005), so more southern fisheries and processors may have an advantage on fisheries in other areas.
 
-fates = as.data.frame(group_by(cfec_quitters, p_holder, p_fshy) %>% 
-  summarize(n = length(which(year > lastYear[1]))) %>% 
-  group_by(p_fshy) %>% 
-  summarize(n0 = length(which(n==0)), nOther = length(which(n>0))))
-
-trackQuitters = function(permit) {
-cfec2= cfec_quitters[which(cfec_quitters$p_holder %in% unique(quitters[quitters$p_fshy==permit,]$p_holder)),]
-cfec2 = cfec2[which(cfec2$year > cfec2$lastYear),]
-cfec2$SPEC = substr(cfec2$p_fshy,1,1)
-totalpermit = group_by(cfec2,SPEC) %>% 
-  summarize(g = sum(g_earn))
-totalpermit$g = totalpermit$g/sum(totalpermit$g)
-return(as.data.frame(arrange(totalpermit,-g)))
-
-}
-
-#trackQuitters("L 21C")
-```
-
-## Overview
-
-First, we can pull out the herring permits by all areas.
-We'll restrict the analysis to (1) fisheries with 10 or more years of data, and (2) fisheries that at their max had at least 25 people landing catch in a year.
-We see that people generally decline to 0, except fisheries in SE (G01A, G34A), and spawn on kelp fisheries (L21).
-Despite catch being flat or maybe increasing, we see prices go down or flat for nearly all permits (and spawn on kelp L21 permits have significantly higher price / lb compared to roe fisheries).
-
-In terms of the diversification of these permit holders, some questions to ask might be (1) if people stopped fishing herring, did they stop fishing entirely, or fish something else?
-(2) for the people that stayed, how many of them are diversified (other permits) and how much of their revenue comes from other permits?
-
-For a lot of permits, we see decreases in the proportion of people fishing only herring.
-In terms of spatial diversity, for several of the fisheries that haven't crashed, we see increases in diversity: G01M/T, G34N, L21.
-In terms of diversifying to other types of herring permits,
-
-```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Herring permit holders by year."}
-  ggplot(herring_summary, aes(year, people)) + geom_line() + facet_wrap(~ p_fshy, scale="free_y") + xlab("Year") + ylab("People")
-```
+In addition to being susceptible to changing demand, a related pressure on Alaskan sac roe fisheries has been external markets.
+Over the period 1994-2003 for example, roe from Alaska represented a small fraction of the total supplied to Japan, especially compared to roe from Canada and China (Carlson 2005).
+Though roe from Pacific herring are considered to be of the highest quality, there may also be competition from other roe fisheries offering substitutes (e.g.
+Atlantic herring).
+Finally, because the salted roe is preserved, supply and demand may be affected by surplus in previous years (Carlson 2005).
 
@@
 
-```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Trends in landed catch by permit.", fig.height=3.5}
-  ggplot(herring_summary, aes(year, log(lbs))) + geom_line() + 
-  facet_wrap(~ p_fshy) + ylab("Ln (tons)") + xlab("Year")
+```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Active permits for herring in Prince William Sound.
+L12E and L21E permits are combined for spawn on kelp, H01E and H34E are combined for food and bait, and G01E and G34E are combined for sac roe fisheries."}
+
+cfec <- read_feather("../../portfolio/data-generated/cfec.feather")
+cfec$year <- as.numeric(cfec$year)
+
+simp.div = function(x) {
+  1/sum((x/sum(x))^2)
+}
+
+pws = dplyr::filter(cfec, p_fshy%in%c("L 12E", "L 21E", "G 01E", "G 34E", "H 01E", "H 34E"))
+pws$p_fshy[pws$p_fshy%in%c("L 12E","L 21E")] = "Spawn on kelp" 
+pws$p_fshy[pws$p_fshy%in%c("H 01E","H 34E")] = "Food and bait"
+pws$p_fshy[pws$p_fshy%in%c("G 01E","G 34E")] = "Roe"
+
+names(pws)[which(names(pws)=="p_fshy")]="Fishery"
+g1 = group_by(pws[pws$year<=1998,], Fishery, year) %>% 
+  summarize(w = sum(g_pounds)) %>% 
+  ggplot(aes(year, w, group=Fishery, color=Fishery)) + geom_line() + 
+  xlab("Year") + ylab("Pounds")
+
+g2 = group_by(pws[pws$year<=1998,], Fishery, year) %>% 
+  summarize(w = length(unique(p_holder))) %>% 
+  ggplot(aes(year, w, group=Fishery, color=Fishery)) + geom_line() + 
+  xlab("Year") + ylab("Active permits")
+print(g2)
+#gridExtra::grid.arrange(g1, g2)
 ```
 
-```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Trends in prices by permit.", fig.height=3.5}
-  ggplot(herring_summary, aes(year, p)) + geom_line() + 
-  facet_wrap(~ p_fshy, scale="free_y") + ylab("Price (Dollars / lb.)") + xlab("Year")
+### Changing Participation
+
+Declines in herring abundance, demand (and prices) and ex-vessel values in Prince William Sound have been well documented (Carlson 2005).
+Following peak biomass in the late 1980s, the number of active permit holders declined significantly until the fisheries closed in the late 1990s.
+Permits for herring roe fisherines in Prince William Sound may be unique to other case studies in that although the fishery remains closed, the permits continue to hold value (and are bought and sold).
+Because of the closure, permit values have declined substantially to approximately 10% of the value 20 years ago (from 127000 dollars in February 1997 or 327000 in 2016 dollars to 31000 in February 2016) (CFEC Permit Value Report 2016).
+
+### Changing diversification: the salmon safety net
+
+With herring fisheries in PWS closed, permit holders had two options: stop fishing entirely, or fish other non-herring permits that remained open.
+The fate of herring permit holders in PWS was evaluated using the CFEC database, 1985-2014.
+We defined this 'population' of permit holders as being anyone who ever held one of the six herring permits in PWS.
+In each year, we summarized the number of active fishers (on any permit, including non-herring fisheries).
+As a proxy for individual permit holders leaving fishing entirely, we also calculated the last year that each permit holder was active (again, on any permit).
+These summaries show a long-term decline in the number of individuals in this population (an expected result, as many of the herring permit holders had herring contribute a large fraction of their revenue).
+We also see fishery closures as catalysts for people stopping participation in the fishery.
+
+```{r, echo=FALSE, fig.pos="placeHere", fig.cap="Histogram of last year individual permit holders holding Prince William Sound herring permits were active in any fishery throughout the state.
+The last years that herring fishing was permitted in Prince William Sound was 1997-1998."}
+pws = dplyr::filter(cfec, p_fshy%in%c("G 01E", "G 34E", "H 01E", "H 34E", "L 12E", "L 21E"))
+pws2 = cfec[which(cfec$p_holder%in%unique(pws$p_holder)),]
+
+g = group_by(pws2, p_holder) %>% 
+  summarize(maxY = max(year))
+
+g2 = group_by(pws2, year) %>% 
+  summarize(n = length(unique(p_holder)))
+par(mfrow=c(2,1), mgp=c(2,1,0),mai=c(0.7,0.7,0.1,0.2))
+plot(g2$year, g2$n, type="b", xlab="Year", ylab="Active permit holders", lwd=2, ylim=c(0,600))
+plot(table(g$maxY)[1:29], xlab="Last year fished", ylab="Permit holders")
 ```
 
-```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Scatterplots of prices versus revenue by permit."}
-  ggplot(herring_summary, aes(p, log(lbs/2000), colour = year)) + geom_point() + facet_wrap(~ p_fshy, scale="free") + ylab("Ln (tons)") + xlab("Price (dollars / lb.)")
+For the population of fishers who ever held a PWS herring permit, we summarized the revenue by year (revenue summed in dollars, using deflation corrected earnings).
+Individual permits were grouped to coarse taxonomic levels by the first character of their CFEC permit code (e.g.
+all 'G', 'H', and 'L' permits combined as herring, all 'S' permits combined for salmon, all 'B' permits combined for halibut; Figure XX).
+These aggregate groups indicate that as expected, herring make up a smaller fraction of the total revenue over time (particularly following fisheries closures).
+Herring continue to make up a small fraction of revenue for this population, because several fishers continue to be active in other areas throughout the state.
+Few of these individuals remain herring specialists however; a more common strategy is for fishers to actively fish both herring and salmon (Figure XX).
+For the majority of individuals in this population, the CFEC data indicate that the majority of active fishers abandonded herring entirely.
+The most common non-herring strategy for these fishers was to target salmon exclusively.
+A smaller number of fishers became specialized on halibut, or targeted both salmon and halibut in the same year.
+Unlike halibut, groundfish, or crab fisheries, where the ability to participate is more restrictive (permit holders must buy quota), limited entry salmon fisheries in the state may provide a safety net for fishers affected by the decline or closure of other fisheries in the state.
 
-```
+### References
 
-\break
+Carls, M.G., G.D.
+Marty, and J.E.
+Hose.
+2002.
+Synthesis of the toxicological impacts of the Exxon Valdez oil spill on Pacific herring (Clupea pallasi) in Prince William Sound, Alaska, U.S.A.
+Canadian Journal of Fisheries and Aquatic Sciences, 2002, 59(1): 153-172.
 
-## How much of your revenue is driven by non-herring?
+Carlson, S. 2005. Changes in roe herring markets: a review of available evidence. CFEC report number 05-5N. Commercial Fisheries Entry Commission, Juneau, AK.
 
-```{r, echo = FALSE}
+Funk, F. 1994. Forecast of the Pacific herring biomass in Prince William Sound, Alaska, 1993. Regional Information Report, 5J94-04. ADFG, Commercial Fisheries and Managmenet Division.
 
-herring_permits = unique(herring_summary$p_fshy)
+Pearson, W.H., R.B.
+Deriso, R.A.
+Elston, S.E.
+Hook, K.R.
+Parker, and J.W.
+Anderson.
+2012.
+Hypotheses concerning the decline and poor recovery of Pacific herring in Prince William Sound, Alaska.
+Reviews in Fish Biology and Fisheries, 22(1): 95–135.
 
-# the complicated thing here is that people may hold multiple herring permits, 
-# multiple permits, no permits. So for each permit type we need to calculate 
-# each of these groups independently. Some people may be counted multiple times.
+Reid, G.M. 1971. Age composition, weight, length, and sex of herring, Clupea pallasii, used for 
+reduction in Alaska, 1929-1966.
 
-# next calculate strategies for diversification
-calc_permit_summary = function(this_permit) {
-people_year = group_by(cfec[cfec$p_fshy == this_permit,], p_holder, year) %>% 
-  summarize(g = sum(g_earn)) %>% 
-  filter(g > 0) %>% 
-  select(-g)
-# grab all records pertaining to these people-year combinations
-herring = left_join(people_year, cfec)
+Rounsefell, G.A. and E.H. Dahlgren. 1935. Fluctuations in the supply of herring, Clupea pallasii, in Prince William Sound, Alaska. Bulletin of the U.S. Bureau of Fisheries 47:263-291.
 
-# aggregate fish tickets to permits
-herring = group_by(herring, p_holder, p_fshy, year) %>% 
-  summarize(g_earn = sum(g_earn))
+Skud, B.E., H.M. Sakuda, and G.M Reid. 1960. Statistics of the Alaska herring fishery, 1878-1956. U.S. Fish and Wildlife Service, Statistical Digest 48, 21 p.
 
-permit_summary = group_by(herring, p_holder, year) %>% 
-  summarize(npermit = length(unique(p_fshy)),
-    npermit_herring = length(which(p_fshy %in% herring_permits)),
-    npermit_g34 = length(which(p_fshy %in% herring_permits & substr(p_fshy,1,4)=="G 34")),
-    npermit_g01 = length(which(p_fshy %in% herring_permits & substr(p_fshy,1,4)=="G 01")),
-    npermit_l12 = length(which(p_fshy %in% herring_permits & substr(p_fshy,1,4)=="L 12")),
-    npermit_l21 = length(which(p_fshy %in% herring_permits & substr(p_fshy,1,4)=="L 21")),
-    nmult = ceiling(npermit_g34/100) + ceiling(npermit_g01/100) + ceiling(npermit_l21/100) + ceiling(npermit_l12/100)) %>% 
-    as.data.frame() %>%
-    group_by(year) %>% 
-  summarize(only_herring_single = length(which(npermit==1 & npermit_herring == 1)),
-    only_herring_multiple = length(which(npermit>1 & npermit_herring >1 & npermit == npermit_herring)), herring_combo = length(which(npermit > 1 & npermit_herring >= 1 & npermit > npermit_herring)), g34 = length(which(npermit_g34 > 1)), g01 = length(which(npermit_g01 > 1)),
-    l12 = length(which(npermit_l12 > 1)), l21 = length(which(npermit_l21 > 1)), 
-    multiple_fisheries = length(which(nmult > 1)))
+Thorne, R. E., and Thomas, G. L. 2008. Herring and the "Exxon Valdez" oil spill: an investigation into historical data conflicts. – ICES Journal of Marine Science, 65: 44–50.
 
-permit_summary$p_fshy = this_permit
-return(permit_summary)
-}
+Thynes, T., D. Gordon, D. Harris, and S. Walker. 2016. 2016 Southeast Alaska Sac Roe Herring Fishery Management Plan. Regional Information Report 1J16-02. ADFG, Division of Commercial Fisheries.
 
-herring_summary = calc_permit_summary(herring_permits[1])
-for(i in 2:length(herring_permits)) {
-  herring_summary = rbind(herring_summary, calc_permit_summary(herring_permits[i]))
-}
+U.S. Department of Interior 1893. Population and Resources of Alaska at the Eleventh Census: 1890. U.S. Department of Interior, Census Office.
 
-```
-
-```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Proportion of permit holders specializing on herring.", fig.height=3.5}
-ggplot(herring_summary, aes(year, (only_herring_single+only_herring_multiple)/(only_herring_single+only_herring_multiple+herring_combo))) + facet_wrap(~p_fshy) + geom_line() + xlab("Year") + ylab("Proportion of people only fishing herring")
-```
-
-```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Proportion of permit holders diversifying spatially.", fig.height=3.5}
-p1 = ggplot(herring_summary, aes(year, (g34)/(only_herring_single+only_herring_multiple+herring_combo), group = p_fshy, colour = p_fshy)) + geom_line() + xlab("Year") + ylab("Proportion of people diversifying spatially")
-
-p2 = ggplot(herring_summary, aes(year, (g01)/(only_herring_single+only_herring_multiple+herring_combo), group = p_fshy, colour = p_fshy)) + geom_line() + xlab("Year") + ylab("Proportion of people diversifying spatially")
-
-p3 = ggplot(herring_summary, aes(year, (l21)/(only_herring_single+only_herring_multiple+herring_combo), group = p_fshy, colour = p_fshy)) + geom_line() + xlab("Year") + ylab("Proportion of people diversifying spatially")
-
-p4 = ggplot(herring_summary, aes(year, (l12)/(only_herring_single+only_herring_multiple+herring_combo), group = p_fshy, colour = p_fshy)) + geom_line() + xlab("Year") + ylab("Proportion of people diversifying spatially")
-
-gridExtra::grid.arrange(p1, p2, p3, p4)
-```
-
-\break
-
-```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Proportion of people diversifying w/diff herring permits.", fig.height=3.5}
-ggplot(herring_summary, aes(year, (multiple_fisheries)/(only_herring_single+only_herring_multiple+herring_combo))) + facet_wrap(~p_fshy) + geom_line() + xlab("Year") + ylab("Proportion of people diversifying w/diff herring permits")
-```
-
-```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Ratio of effective diversity to total number of permit holders (as measure of evenness).
-Values near 1 indicate everyone's revenue is similar.", fig.height=3.5}
-# We can also calculate how herring catch is distributed across fisherment
-simp.div = function(x) {1/sum((x/sum(x))^2)}
-
-group_by(cfec[cfec$p_fshy%in%herring_permits,], year, p_holder, p_fshy) %>% 
-  summarize(g = sum(g_earn)) %>% 
-  group_by(year, p_fshy) %>% 
-  summarize(div = simp.div(g), n = length(unique(p_holder))) %>% 
-  ggplot(aes(year, div/n)) + facet_wrap(~p_fshy, scale="free_y") + geom_line() + xlab("Year") + ylab("Diversity across people")
-```
-
-\break
-
-```{r, echo = FALSE, warning = FALSE, message = FALSE, results = "hide", fig.pos="placeHere", fig.cap = "Fraction of revenue made up of herring (median, of per-person fractions).",eval=FALSE}
-
-herring_permits = unique(herring_summary$p_fshy)
-
-herring_people = group_by(cfec, p_fshy, year, p_holder) %>%
-  summarize(g = sum(g_earn),
-  herring_landing = ifelse(length(which(p_fshy %in% herring_permits)) == 0, 0, 1)) %>% 
-  data.frame() %>% 
-  group_by(p_holder, year) %>% 
-  mutate(herring_personyear = ifelse(length(which(p_fshy %in% herring_permits)) == 0, 0, 1),
-  g_herring = sum(g[herring_landing==1]), g_other = sum(g[herring_landing==0]),
-  p_fshy = p_fshy[which(p_fshy %in% herring_permits)][1]) %>% 
-  group_by(p_holder) %>% 
-  mutate(ever_herring = sum(herring_personyear)) %>%   
-  filter(ever_herring > 0) %>% 
-  select(-g, -herring_landing, -herring_personyear, -ever_herring)
-
-herring_people$p_herring = herring_people$g_herring / (herring_people$g_herring + herring_people$g_other)
-group_by(herring_people[herring_people$p_fshy %in% herring_permits, ], p_fshy, year) %>% 
-  summarize(g = median(p_herring)) %>% 
-  ggplot(aes(year, g)) + geom_line() + facet_wrap(~p_fshy) + ylab("Median % herring (as fraction of total revenue)") + xlab("Year")
-```
+Woodby, D., D. Carlile, S. Siddeek, F. Funk, J.H. Clark, and L. Hulbert. 2005. Commercial Fisheries in Alaska. ADFG special publication 05-09.  http://www.sf.adfg.state.ak.us/FedAidPDFs/sp05-09.pdf

### 2016-11-23 portfolio/text/ms.Rmd
edits to rebuild intro

@@
 
-Managing natural resources and the 
-livelihoods of those that depend on their extraction is hard.
-Environmental, biological, and economic variation
-combine to generate high levels of income variability and risk.
-Drought and flooding is a major source of 
-risk to agricultural food security and farmers' incomes [@gregory2005].
-Bark beetle outbreaks can cause devastation to 
-the income of loggers.
-Price crashes in insert mineral here 
-are hard on miners' income variability ...
+Managing natural resources and sustaining the livelihoods of those that depend on their extraction is difficult.
+These professions tend to have higher year-to-year variability compared to all others (@sethi2012b ; https://www.ers.usda.gov/webdocs/publications/aer812/31365_aer812f.pdf).
+In addition to economic risk, such as changes in prices, demand, or business cycles, individuals dependent on natural resources are also subject to biological and environmental variability.
+Examples include drought and flooding being a major source of risk to agricultural food security and farmers' incomes [@gregory2005], frequent catastrophic disease outbreaks or wildfires having negative impact on the logging industry (http://www.pnas.org/content/104/50/19697.full), or fluctuating prices [include mineral/gas example – Eric can’t find a good one].
 
-Individuals naturally develop strategies to reduce this income variation.
-For example, farmers rotate their crops to buffer short-term 
-climate variation and long-term environmental trends REF. 
-Loggers maintain a diverse mix of tree species, age structure, 
-geographical areas to buffer against 
-disease outbreak, wildfires, and price variation REF. 
-But, otherwise well-intentioned regulations
-can limit the basket of actions
-available to individuals to reduce income variability.
-For example, ...
-In other cases, external forces (prices, demand) may discourage the
-adoption of these diverse strategies and
-help promote specialization [@bradshaw2004].
+Individuals who rely on natural resources for income develop strategies to reduce this variability in year to year earnings.
+For example, farmers may rotate their crops to buffer short-term climate variation and long-term environmental trends (http://www.nrcresearchpress.com/doi/abs/10.4141/cjps88-001#.WDUQ5aIrLBI).
+Similarly, loggers maintain a diverse mix of tree species, age structure, and harvest in geographical areas to buffer against disease outbreak, wildfires, and price variation REF.
 
-Fisheries provide a prime example...
-While the days of open, unregulated fisheries in most developed countries have
-passed --- fishers must own a license to participate in virtually all fisheries
---- fishers retain several options for increasing diversity. 
-For example, fishers can
-fish multiple species as permitted by a license, 
-own multiple licenses,
-or fish in multiple regions.
+In many of these socio-ecological systems, otherwise well-intentioned regulations can limit the basket of actions available to individuals to reduce income variability, or incentivize against diversification.
+In other words, while diversification is generally seen as reducing variability, many individuals don’t.
+For example, in terrestrial agriculture, crop subsidies in the US may incentivize some farms to specialize (dairy, grains, soybeans,  http://ageconsearch.umn.edu/bitstream/33967/1/ae020812.pdf).
+In other instances, external market forces (prices, demand) may discourage the adoption of diverse strategies and help promote specialization [@bradshaw2004].
+In fisheries, a series of sequential changes around the world have generally limited participation and led to increased specialization (Hilborn et al.
+2003 State of the World’s Fisheries).
 
-But uncertainty remains about the value
-of adopting various diversification strategies for improving revenue and
-reducing revenue variability. 
-Furthermore, theory and empirical studies both
-suggest there can be substantial benefits to specialization (REFS).
+Commercial fishers present an interesting example of diversification, perhaps because of the extreme risk that some individuals experience, particularly relative to terrestrial agriculture (@sethi2012b).
+First, the majority of insurance programs available to terrestrial farmers are not available to commercial fishermen (e.g.
+crop insurance programs; @kasperski2013).
+Second, as both fish and fishers are spatially mobile, equating fishers to farmers would be possible if the farmers’ field became an amorphous blob that changed shape and moved unpredictably day to day.
+Fishermen still have the ability to lower risk, however.
+While the days of open, unregulated fisheries in most developed countries have passed (fishers must own a license to participate in virtually all fisheries), fishers can (1) fish multiple species as permitted by a license, (2) own multiple licenses, or (3) diversify spatially and fish in multiple regions.
 
-However, these papers look at aggregates across individuals and largely ignore
-how existing fisheries regulations limit the ability of individual fishers to
-diversify. 
-Key uncertainties remain about the fundamental relationship between
-diversity, fishing revenue, and fishing revenue variability at the level of the
-individual.
+Despite these ways fishers can diversify, the effect of adopting various diversification strategies for improving revenue and reducing revenue variability remains uncertain.
+Furthermore, theory and empirical studies both suggest there can be substantial benefits to specialization (REFS).
+The majority of papers that examine effects of diversification on revenue do so at aggregate levels (households, counties, communities), rather than focus on individuals.
+Because of this, key uncertainties remain about the fundamental relationship between diversity, fishing revenue, and fishing revenue variability at the level of the individual.
 
-Diversity in the fishing strategies of individuals
-ranges from highly specialized (e.g. fishing only one species with one gear type
-or in one area) to highly diverse (e.g. fishing multiple species with multiple
-gear types in multiple areas).
+One of the most well studied examples of diversification are the state and federal fisheries in Alaska.
+These provide a unique opportunity to understand the role of individuals' strategies in determining income and income variability for those whose livelihood depends on the extraction of natural resources.
+Fisheries in Alaska are among the highest-valued fisheries in the world with 6 billion landed pounds valued at nearly  $2 billion annually (NMFS 2016) [link]( http://www.st.nmfs.noaa.gov/Assets/commercial/fus/fus15/documents/FUS2015.pdf).
+The industry supports a range of fishing fleets ranging from small, open, single-operator vessels to multi-million-dollar state-of-the-art ships with large crews.
+Like much of the rest of the world (Hilborn et al.
+2003 state of the world’s fisheries), the days of open, unregulated fisheries in most developed countries have passed --- fishers must own a license to participate in virtually all fisheries.
+And while such regulations have proven helpful in reducing or eliminating overfishing, they do limit the available strategies to reduce income variability.
 
-Fisheries in Alaska provide a unique opportunity 
-to understand the role of individuals' strategies 
-in determining income and income variability
-for those whose livelihood depends on the extraction of natural resources.
-Fisheries in Alaska are the highest-valued fisheries in North America
-with landings valued at XX $ billion annually representing harvests from YY
-species.
-The industry supports a range of fishing fleets ranging from small,
-open, single-operator vessels to multi-million-dollar state-of-the-art ships
-with large crews.
+Using a unique dataset collected over 30 years (1985-2014), collected on individuals, rather than fleets or vessels, we are interesting in how revenue variability of individuals changes through time, and is affected by diversification.
+The benefit of using individuals rather than vessels is that we are able to track fishes that switch fishery sectors over time, or alter strategies and utilize multiple vessels.
+We developing and apply a novel statistical approach for understanding the effect of diversity on both revenue and risk that builds on results from the vessel [@kasperski2013], fleet [@sethi2012b], and community level variability [@sethi2014].
+To our knowledge, this is the first application of this type of model in any natural resource field that jointly explores the effect of diversity on revenue and risk at the individual level and the first considers the multiple axes of diversity simultaneously.
 
-However, the days of open, unregulated fisheries in most developed countries have
-passed --- fishers must own a license to participate in virtually all fisheries.
-And while such regulations have proven 
-helpful in reducing or eliminating overfishing,
-they do limit the available strategies to reduce income variability.
-
-- Eric: Most papers on fisheries have adopted the approach similar to @kasperski2013 and used data reported to the level of vessels or higher. This analysis
-is unique in that we're able to separate individual people from vessels
-(enabling us to track people that switch fishery sectors over time, or
-change strategies and utilize multiple vessels)
-
-\bigskip
-
-<!-- -Ole: I think this is the spot for intro to AK.  I moved some of Eric's stuff -->
-<!-- up here and added to it.[*Eric: I think we need a little here on the AK data -->
-<!-- and why that's both important and generalizable to fisheries around the world.] -->
-
-We develop a novel statistical approach for understanding the effect of
-diversity on both revenue and risk that builds on results from the vessel
-[@kasperski2013], fleet [@sethi2012b], and community levels [@sethi2014].
 First, we summarize trends in diversification across fishing fleets in Alaska.
-Next, we identify three main axes by which individual fishers may diversify
-the species they target:
-(1) owning permits that allow targeting a diversity of fish species,
-(2) purchasing additional permits,
+Next, we identify three main axes by which individuals may diversify the species they target [Eric: I think these kind of bleed together, perhaps make more clear or reduce to 2]:
+(1) owning permits that allow targeting a diversity of fish species (e.g. miscellaneous groundfish or salmon permits),
+(2) purchasing additional permits (the majority of permits in Alaska are single species – halibut, herring, etc),
 or (3) catching a greater species diversity within the constraints of a given permit.
-
-Finally, we construct a model to simultaneously estimate the relationship
-between each of these three axes of diversity and both annual revenue and risk
-for individual fishers over nearly ZZ years.
-Our hierarchical model makes use of
-entire time series,
-controls for effort and latent time effects,
-and allows the effect of species diversity
-to vary by fishing strategy (sets of similar permits).
-It is the first model that jointly explores the effect of diversity on revenue
-and risk at the individual level and the first considers the multiple axes of
-diversity simultaneously.
-
-strategies where both specialization and diversification are extremely
-profitable. There's a number of fishing permits that constrain the potential
-diversity of species landed (e.g. herring, halibut, sablefish, or crab permits
-that allow the user to only land that species) versus broader 'finfish' permits
-that allow users to land a wide variety of groundfish spp.*
-
 *perhaps we also need to have a sentence or two on the implicit constraints of

### 2016-11-23 portfolio/text/ms.Rmd
Add refs

@@
 Managing natural resources and sustaining the livelihoods of those that depend on their extraction is difficult.
-These professions tend to have higher year-to-year variability compared to all others (@sethi2012b ; https://www.ers.usda.gov/webdocs/publications/aer812/31365_aer812f.pdf).
+These professions tend to have higher year-to-year variability compared to all others [@mishra2002; @sethi2012b].
 In addition to economic risk, such as changes in prices, demand, or business cycles, individuals dependent on natural resources are also subject to biological and environmental variability.
-Examples include drought and flooding being a major source of risk to agricultural food security and farmers' incomes [@gregory2005], frequent catastrophic disease outbreaks or wildfires having negative impact on the logging industry (http://www.pnas.org/content/104/50/19697.full), or fluctuating prices [include mineral/gas example – Eric can’t find a good one].
+Examples include drought and flooding being a major source of risk to agricultural food security and farmers' incomes [@gregory2005], frequent catastrophic disease outbreaks or wildfires having negative impact on the logging industry [@kirilenko2007], or fluctuating prices [include mineral/gas example – Eric can’t find a good one].
 
 Individuals who rely on natural resources for income develop strategies to reduce this variability in year to year earnings.
-For example, farmers may rotate their crops to buffer short-term climate variation and long-term environmental trends (http://www.nrcresearchpress.com/doi/abs/10.4141/cjps88-001#.WDUQ5aIrLBI).
+For example, farmers may rotate their crops to buffer short-term climate variation and long-term environmental trends [@zentner1988].
 Similarly, loggers maintain a diverse mix of tree species, age structure, and harvest in geographical areas to buffer against disease outbreak, wildfires, and price variation REF.
@@
 In other words, while diversification is generally seen as reducing variability, many individuals don’t.
-For example, in terrestrial agriculture, crop subsidies in the US may incentivize some farms to specialize (dairy, grains, soybeans,  http://ageconsearch.umn.edu/bitstream/33967/1/ae020812.pdf).
+For example, in terrestrial agriculture, crop subsidies in the US may incentivize some farms to specialize [e.g.
+dairy, grains, soybeans;  @mishra2002].
 In other instances, external market forces (prices, demand) may discourage the adoption of diverse strategies and help promote specialization [@bradshaw2004].
-In fisheries, a series of sequential changes around the world have generally limited participation and led to increased specialization (Hilborn et al.
-2003 State of the World’s Fisheries).
+In fisheries, a series of sequential changes around the world have generally limited participation and led to increased specialization [@hilborn2003a].
 
@@
 
-One of the most well studied examples of diversification are the state and federal fisheries in Alaska.
+One of the most well-studied examples of diversification are the state and federal fisheries in Alaska.
 These provide a unique opportunity to understand the role of individuals' strategies in determining income and income variability for those whose livelihood depends on the extraction of natural resources.
-Fisheries in Alaska are among the highest-valued fisheries in the world with 6 billion landed pounds valued at nearly  $2 billion annually (NMFS 2016) [link]( http://www.st.nmfs.noaa.gov/Assets/commercial/fus/fus15/documents/FUS2015.pdf).
+Fisheries in Alaska are among the highest-valued fisheries in the world with 6 billion landed pounds valued at nearly 2 billion USD annually [@nmfs2016].
 The industry supports a range of fishing fleets ranging from small, open, single-operator vessels to multi-million-dollar state-of-the-art ships with large crews.
-Like much of the rest of the world (Hilborn et al.
-2003 state of the world’s fisheries), the days of open, unregulated fisheries in most developed countries have passed --- fishers must own a license to participate in virtually all fisheries.
+Like much of the rest of the world [@hilborn2003a], the days of open, unregulated fisheries in most developed countries have passed --- fishers must own a license to participate in virtually all fisheries.
 And while such regulations have proven helpful in reducing or eliminating overfishing, they do limit the available strategies to reduce income variability.
 
-Using a unique dataset collected over 30 years (1985-2014), collected on individuals, rather than fleets or vessels, we are interesting in how revenue variability of individuals changes through time, and is affected by diversification.
+Using a unique dataset collected over 30 years (1985--2014), collected on individuals, rather than fleets or vessels, we are interesting in how revenue variability of individuals changes through time, and is affected by diversification.
 The benefit of using individuals rather than vessels is that we are able to track fishes that switch fishery sectors over time, or alter strategies and utilize multiple vessels.

### 2016-11-23 portfolio/text/ms.Rmd
Some quick edits to intro

@@
 Managing natural resources and sustaining the livelihoods of those that depend on their extraction is difficult.
-These professions tend to have higher year-to-year variability compared to all others [@mishra2002; @sethi2012b].
-In addition to economic risk, such as changes in prices, demand, or business cycles, individuals dependent on natural resources are also subject to biological and environmental variability.
-Examples include drought and flooding being a major source of risk to agricultural food security and farmers' incomes [@gregory2005], frequent catastrophic disease outbreaks or wildfires having negative impact on the logging industry [@kirilenko2007], or fluctuating prices [include mineral/gas example – Eric can’t find a good one].
+These professions tend to have high year-to-year variability in income [@mishra2002; @sethi2012b].
+In addition to economic risk, such as changes in prices or demand, individuals dependent on natural resources are also subject to biological and environmental variability.
+For example, drought and flooding are a major source of risk to agricultural food security and farmers' incomes [@gregory2005], and catastrophic disease outbreaks or wildfires increase risk for the logging industry [@kirilenko2007].
 
-Individuals who rely on natural resources for income develop strategies to reduce this variability in year to year earnings.
+Individuals who rely on natural resources for income develop strategies to reduce this variability in income.
 For example, farmers may rotate their crops to buffer short-term climate variation and long-term environmental trends [@zentner1988].
@@
 
-In many of these socio-ecological systems, otherwise well-intentioned regulations can limit the basket of actions available to individuals to reduce income variability, or incentivize against diversification.
-In other words, while diversification is generally seen as reducing variability, many individuals don’t.
+In many of these socio-ecological systems, otherwise well-intentioned regulations can limit the actions available to individuals to reduce income variability, or incentivize against diversification.
 For example, in terrestrial agriculture, crop subsidies in the US may incentivize some farms to specialize [e.g.
@@
 
-Commercial fishers present an interesting example of diversification, perhaps because of the extreme risk that some individuals experience, particularly relative to terrestrial agriculture (@sethi2012b).
-First, the majority of insurance programs available to terrestrial farmers are not available to commercial fishermen (e.g.
-crop insurance programs; @kasperski2013).
+Commercial fishers present an interesting example of diversification, given the extreme risk that some individuals experience, particularly relative to terrestrial agriculture (@sethi2012b).
+First, the majority of insurance programs available to terrestrial farmers are not available to commercial fishermen [e.g.
+crop insurance programs; @kasperski2013].
 Second, as both fish and fishers are spatially mobile, equating fishers to farmers would be possible if the farmers’ field became an amorphous blob that changed shape and moved unpredictably day to day.
 Fishermen still have the ability to lower risk, however.
-While the days of open, unregulated fisheries in most developed countries have passed (fishers must own a license to participate in virtually all fisheries), fishers can (1) fish multiple species as permitted by a license, (2) own multiple licenses, or (3) diversify spatially and fish in multiple regions.
+While the days of open, unregulated fisheries in most developed countries have passed (fishers must own a license to participate in virtually all fisheries), fishers can fish multiple species as permitted by a license, own multiple licenses, or diversify spatially and fish in multiple regions.
 
@@
 Furthermore, theory and empirical studies both suggest there can be substantial benefits to specialization (REFS).
-The majority of papers that examine effects of diversification on revenue do so at aggregate levels (households, counties, communities), rather than focus on individuals.
+The majority of papers that examine effects of diversification on revenue do so at aggregate levels (e.g.
+households, counties, communities; REFS), rather than focus on individuals.
 Because of this, key uncertainties remain about the fundamental relationship between diversity, fishing revenue, and fishing revenue variability at the level of the individual.
@@
 These provide a unique opportunity to understand the role of individuals' strategies in determining income and income variability for those whose livelihood depends on the extraction of natural resources.
-Fisheries in Alaska are among the highest-valued fisheries in the world with 6 billion landed pounds valued at nearly 2 billion USD annually [@nmfs2016].
+Fisheries in Alaska are among the highest-valued fisheries in the world with 6 billion landed pounds [TODO kg] valued at nearly 2 billion USD annually [@nmfs2016].
 The industry supports a range of fishing fleets ranging from small, open, single-operator vessels to multi-million-dollar state-of-the-art ships with large crews.
@@
 
-Using a unique dataset collected over 30 years (1985--2014), collected on individuals, rather than fleets or vessels, we are interesting in how revenue variability of individuals changes through time, and is affected by diversification.
-The benefit of using individuals rather than vessels is that we are able to track fishes that switch fishery sectors over time, or alter strategies and utilize multiple vessels.
-We developing and apply a novel statistical approach for understanding the effect of diversity on both revenue and risk that builds on results from the vessel [@kasperski2013], fleet [@sethi2012b], and community level variability [@sethi2014].
+Using a unique dataset collected over 30 years (1985--2014) collected on individuals, rather than fleets or vessels, we examine how revenue variability of individuals changes through time, and is affected by diversification.
+The benefit of using individuals rather than vessels is that we are able to track fishers that switch fishery sectors over time, or alter strategies and use multiple vessels.
+We develop and apply a novel statistical approach for understanding the effect of diversity on both revenue and risk that builds on results from the vessel [@kasperski2013], fleet [@sethi2012b], and community level variability [@sethi2014].
 To our knowledge, this is the first application of this type of model in any natural resource field that jointly explores the effect of diversity on revenue and risk at the individual level and the first considers the multiple axes of diversity simultaneously.

### 2016-12-05 portfolio/text/ms.Rmd
Rework introduction and start cleaning up paper

@@
 However, any diversity-variability relationship
-is embedded in a regulatory framework that restricts individual ability
-to unilaterally change fishing behaviour.
-
-Here we identify three primary axes by which fishers in Alaska,
-one of the world's most productive fishery regions, can
-potentially diversify fishing activities and
-investigate how individual actions along each diversity axis affect
-both revenue and revenue variability.
-We construct a novel statistical model to simultaneously
-assess the value to individual fishers of choosing to
-(1) own permits that allow targeting a diverse fish species assemblage,
-(2) diversify by owning multiple permits, or
-(3) catch a greater species diversity within the constraints of a given permit.
-
+is embedded in a regulatory framework that restricts individual ability to
+change fishing behaviour. Here we identify three primary axes by which fishers
+in Alaska, one of the world's most productive fishery regions, can potentially
+diversify fishing activities and investigate how individual actions along each
+diversity axis affect both revenue and revenue variability. We construct a novel
+statistical model to simultaneously assess the value to individual fishers of
+choosing to (1) own permits that allow targeting a diverse fish species
+assemblage, (2) diversify by owning multiple permits, or (3) catch a greater
+species diversity within the constraints of a given permit.
 We show contrasting effects of diversity on
@@
 associated with increased uncertainty in revenue the following year.
-
 Thus, while there are clear benefits
@@
 
-Extra:
+<!-- Extra: -->
 
-- there is often limited room for intentional diversification within a permit
-- in fact, specializing can be profitable within some fisheries
-- Eric: analog here in farming literature is that diversification
-  for small scale farmers generally good, except when they
-  specialize on high demand/high value crops (tobacco, cocoa).
-  Seems like benefits of diversification are also smaller for
-  large-scale industrial farmers.
+<!-- - there is often limited room for intentional diversification within a permit -->
+<!-- - in fact, specializing can be profitable within some fisheries -->
+<!-- - Eric: analog here in farming literature is that diversification -->
+<!--   for small scale farmers generally good, except when they -->
+<!--   specialize on high demand/high value crops (tobacco, cocoa). -->
+<!--   Seems like benefits of diversification are also smaller for -->
+<!--   large-scale industrial farmers. -->
 
-\noindent
-Main message:
-Diversified portfolios of fishing permits and species reduce risk for fishermen,
-but changing species diversity can be risky, costly, and limited by regulations.
+<!-- \noindent -->
+<!-- Main message: -->
+<!-- Diversified portfolios of fishing permits and species reduce risk for fishermen, -->
+<!-- but changing species diversity can be risky, costly, and limited by regulations. -->
 
@@
 
-<!-- Introduction, Methods, Results, Discussion, Conclusion:
+<!-- Goal: Introduction, Methods, Results, Discussion, Conclusion:
 <!-- 600, 900, 500, 800, 200, -->
-
-<!-- write out each paragraph’s main message in 15 words or less -->
-<!-- circulate skeleton ... this is the time to fix problems with flow, -->
-<!-- ideas/content, main message -->
 
@@
 
-Managing natural resources and sustaining the livelihoods of those that depend on their extraction is difficult.
-These professions tend to have high year-to-year variability in income [@mishra2002; @sethi2012b].
-In addition to economic risk, such as changes in prices or demand, individuals dependent on natural resources are also subject to biological and environmental variability.
-For example, drought and flooding are a major source of risk to agricultural food security and farmers' incomes [@gregory2005], and catastrophic disease outbreaks or wildfires increase risk for the logging industry [@kirilenko2007].
+<!-- Papers to cite [@smith1986] -->
 
-Individuals who rely on natural resources for income develop strategies to reduce this variability in income.
-For example, farmers may rotate their crops to buffer short-term climate variation and long-term environmental trends [@zentner1988].
-Similarly, loggers maintain a diverse mix of tree species, age structure, and harvest in geographical areas to buffer against disease outbreak, wildfires, and price variation REF.
+Managing natural resources and sustaining the livelihoods of those that depend
+on their extraction is difficult. These professions tend to have high
+year-to-year variability in income [@mishra2002; @sethi2012b]. In addition to
+economic risk, such as changes in demand, individuals dependent on natural
+resources are also subject to biological and environmental variability. For
+example, drought and flooding are a major source of risk to agricultural food
+security and farmers' incomes [@gregory2005], and catastrophic disease outbreaks
+and wildfires increase risk for the logging industry [@kirilenko2007].
 
-In many of these socio-ecological systems, otherwise well-intentioned regulations can limit the actions available to individuals to reduce income variability, or incentivize against diversification.
-For example, in terrestrial agriculture, crop subsidies in the US may incentivize some farms to specialize [e.g.
-dairy, grains, soybeans;  @mishra2002].
-In other instances, external market forces (prices, demand) may discourage the adoption of diverse strategies and help promote specialization [@bradshaw2004].
-In fisheries, a series of sequential changes around the world have generally limited participation and led to increased specialization [@hilborn2003a].
+Individuals who rely on natural resources for income develop strategies to
+reduce income variability. For example, farmers may rotate their crops to buffer
+short-term climate variation and long-term environmental trends [@zentner1988].
+Similarly, loggers maintain a diverse mix of tree species, age structure, and
+harvest in geographical areas to buffer against disease outbreak, wildfires, and
+price variation REF. However, otherwise well-intentioned regulations can limit
+the actions available to individuals to reduce income variability, or
+incentivize against diversification. For example, in terrestrial agriculture,
+crop subsidies in the US may incentivize some farms to specialize [e.g. dairy,
+grains, soybeans; @mishra2002]. In other instances, external market forces
+(prices, demand) may discourage the adoption of diverse strategies and help
+promote specialization [@bradshaw2004].
 
-Commercial fishers present an interesting example of diversification, given the extreme risk that some individuals experience, particularly relative to terrestrial agriculture (@sethi2012b).
-First, the majority of insurance programs available to terrestrial farmers are not available to commercial fishermen [e.g.
-crop insurance programs; @kasperski2013].
-Second, as both fish and fishers are spatially mobile, equating fishers to farmers would be possible if the farmers’ field became an amorphous blob that changed shape and moved unpredictably day to day.
-Fishermen still have the ability to lower risk, however.
-While the days of open, unregulated fisheries in most developed countries have passed (fishers must own a license to participate in virtually all fisheries), fishers can fish multiple species as permitted by a license, own multiple licenses, or diversify spatially and fish in multiple regions.
+Commercial fishers present an interesting example of diversification, given the
+extreme risk that some individuals experience, particularly relative to
+terrestrial agriculture (@sethi2012b). The majority of insurance programs
+available to terrestrial farmers (e.g. crop insurance programs) are not
+available to commercial fishers [@kasperski2013] and both fish and fishers are
+spatially mobile. However, fishers still have the ability to lower risk. They
+can, for example, target multiple species as permitted by a license, own
+multiple licenses, or diversify spatially and fish in multiple regions. The
+effectiveness of these diversification strategies for improving revenue and
+reducing revenue variability remains uncertain. Furthermore, theory and
+empirical studies both suggest there can be substantial benefits to
+specialization (REFS). The majority of papers that examine effects of
+diversification on revenue do so at aggregate levels (e.g. households, counties,
+communities; REFS), rather than focus on individuals. Because of this, key
+uncertainties remain about the fundamental relationship between diversity,
+fishing revenue, and fishing revenue variability at the level of the individual.
 
-Despite these ways fishers can diversify, the effect of adopting various diversification strategies for improving revenue and reducing revenue variability remains uncertain.
-Furthermore, theory and empirical studies both suggest there can be substantial benefits to specialization (REFS).
-The majority of papers that examine effects of diversification on revenue do so at aggregate levels (e.g.
-households, counties, communities; REFS), rather than focus on individuals.
-Because of this, key uncertainties remain about the fundamental relationship between diversity, fishing revenue, and fishing revenue variability at the level of the individual.
+The state and federal fisheries of Alaska, provide a unique opportunity to
+understand the role of individuals' strategies in determining income and income
+variability for those whose livelihood depends on the extraction of natural
+resources. Fisheries in Alaska are among the highest-valued fisheries in the
+world with 2.7 billion kg landed biomass valued at nearly 2 billion USD annually
+[@nmfs2016]. The industry supports a range of fishing fleets ranging from small,
+open, single-operator vessels to multi-million-dollar state-of-the-art ships
+with large crews. Like much of the rest of the world [@hilborn2003a], the days
+of open, unregulated fisheries in most developed countries have passed ---
+fishers must own a license to participate in virtually all fisheries. And while
+such regulations have proven helpful in reducing or eliminating overfishing,
+they limit the available strategies to reduce income variability. For example,
+most permits limit fishing to a single species, diversifying by purchasing
+additional fishing permits can be expensive (e.g. ~\$XX for king crab), and some
+fisheries limit the entry of new participants through individual fishing quotas
+(IFQs) (REF).
 
-One of the most well-studied examples of diversification are the state and federal fisheries in Alaska.
-These provide a unique opportunity to understand the role of individuals' strategies in determining income and income variability for those whose livelihood depends on the extraction of natural resources.
-Fisheries in Alaska are among the highest-valued fisheries in the world with 6 billion landed pounds [TODO kg] valued at nearly 2 billion USD annually [@nmfs2016].
-The industry supports a range of fishing fleets ranging from small, open, single-operator vessels to multi-million-dollar state-of-the-art ships with large crews.
-Like much of the rest of the world [@hilborn2003a], the days of open, unregulated fisheries in most developed countries have passed --- fishers must own a license to participate in virtually all fisheries.
-And while such regulations have proven helpful in reducing or eliminating overfishing, they do limit the available strategies to reduce income variability.
-
-Using a unique dataset collected over 30 years (1985--2014) collected on individuals, rather than fleets or vessels, we examine how revenue variability of individuals changes through time, and is affected by diversification.
-The benefit of using individuals rather than vessels is that we are able to track fishers that switch fishery sectors over time, or alter strategies and use multiple vessels.
-We develop and apply a novel statistical approach for understanding the effect of diversity on both revenue and risk that builds on results from the vessel [@kasperski2013], fleet [@sethi2012b], and community level variability [@sethi2014].
-To our knowledge, this is the first application of this type of model in any natural resource field that jointly explores the effect of diversity on revenue and risk at the individual level and the first considers the multiple axes of diversity simultaneously.
-
-First, we summarize trends in diversification across fishing fleets in Alaska.
-Next, we identify three main axes by which individuals may diversify the species they target [Eric: I think these kind of bleed together, perhaps make more clear or reduce to 2]:
-(1) owning permits that allow targeting a diversity of fish species (e.g. miscellaneous groundfish or salmon permits),
-(2) purchasing additional permits (the majority of permits in Alaska are single species – halibut, herring, etc),
-or (3) catching a greater species diversity within the constraints of a given permit.
-*perhaps we also need to have a sentence or two on the implicit constraints of
-geography and gear / vessel size on species diversification*
-
-First, the effect of changing species diversity
-for individual permit holders on a year-by-year basis is unknown ---
-existing approaches have worked at the vessel [@kasperski2013],
-fleet [@sethi2012b], or community level [@sethi2014],
-and condensed time series into single metrics of variability or risk.
-Second, the effect of changing species diversity
-on revenue itself is unknown ---
-existing approaches have focused solely on variability or risk.
-Third, it is unknown if the approach to
-diversifying species fished affects the relationship between
-species diversity and revenue variability ---
-a fishery permit holder can purchase
-an additional permit to target other species,
-pick a permit that inherently targets many species,
-or increase species diversity
-within the constraints of his or her existing permits.
-Many permit holders may be limited to the last approach
-given that the first two may be costly and limited by regulations,
-but it is unclear if the benefits and costs are the same.
-Here, we address these questions by
-simultaneously estimating the effects of
-these three approaches to species diversification
-for Alaskan fishery permit holders on
-annual revenue and revenue variability.
-Our model makes use of entire time series,
-controls for effort and latent time effects,
-and allows the effect of species diversity
-to vary by fishing strategy (sets of similar permits)
-in a hierarchical model.
+Using a unique dataset collected over 30 years (1985--2014) on individuals,
+rather than fleets or vessels, we examine how revenue and revenue variability of
+individuals is affected by diversification. By tracking individuals rather than
+vessels we can track people that switch fishery sectors over time, or alter
+strategies and use multiple vessels. We develop and apply a novel statistical
+approach for understanding the effect of diversity on both revenue and revenue
+variability that builds on results from vessel [@kasperski2013], fleet
+[@sethi2012b], and community-level variability [@sethi2014]. Our model controls
+for fishing effort and latent time effects (e.g. changing environmental
+conditions), simultaneously estimates the effect of species diversity on
+individual revenue and revenue variability, and estimates effects both within
+and across fishing strategies (sets of similar licenses or permits). To our
+knowledge, this is the first application of this type of model in any natural
+resource field.
 
@@
 
-### Fig. \ref{fig:ts} - limited entry + movement towards salmon and specialization in Alaska
+**Fig. \ref{fig:ts} - limited entry + movement towards salmon and specialization
+in Alaska**
 
@@
 
-### Fig. \ref{fig:across} - across strategies
+**Fig. \ref{fig:across} - across strategies**
 
@@
 
-### Fig. \ref{fig:spokes} - add permits
+**Fig. \ref{fig:spokes} - add permits**
 
@@
 
-### Fig. \ref{fig:within} - within strategies
+**Fig. \ref{fig:within} - within strategies**
 
@@
 
-# Notes
+<!-- # Notes -->
 
-## @kasperski2013
-- vessel-level CFEC data
-- dome-shaped result
-- Herfindahl-Hirschman Index/HHI (Simpson)
-- diversity decreasing for those still fishing in 2010, but not otherwise
+<!-- ## @kasperski2013 -->
+<!-- - vessel-level CFEC data -->
+<!-- - dome-shaped result -->
+<!-- - Herfindahl-Hirschman Index/HHI (Simpson) -->
+<!-- - diversity decreasing for those still fishing in 2010, but not otherwise -->
 
-## @sethi2012b
-- risk measures to 90 Alaskan fishing fishery/fleets
-- revenue more variable than catch (+ price = extra variability)
-- large catches, big fleets = lower risk
+<!-- ## @sethi2012b -->
+<!-- - risk measures to 90 Alaskan fishing fishery/fleets -->
+<!-- - revenue more variable than catch (+ price = extra variability) -->
+<!-- - large catches, big fleets = lower risk -->
 
-## @sethi2010b
-- risk management for fisheries review
+<!-- ## @sethi2010b -->
+<!-- - risk management for fisheries review -->
 
-## @sethi2012
-- equal-share catch-pooling cooperatives in king and snow crab
-- "variability reduction depends on the degree of correlation amongst
-  participants’ catches"
+<!-- ## @sethi2012 -->
+<!-- - equal-share catch-pooling cooperatives in king and snow crab -->
+<!-- - "variability reduction depends on the degree of correlation amongst -->
+<!--   participants’ catches" -->
 
-## @sethi2014
-- community-level CFEC data
-- number of fisheries strongest predictor of (lower) CV/CVaR
-- also simpson index
-- suggest barriers to commercial entry increase community risk by restricting
-  access to 'variety of fisheries'
-- some communities have limited ability to diversify; may need to diversify
-  outside of fishing
+<!-- ## @sethi2014 -->
+<!-- - community-level CFEC data -->
+<!-- - number of fisheries strongest predictor of (lower) CV/CVaR -->
+<!-- - also simpson index -->
+<!-- - suggest barriers to commercial entry increase community risk by restricting -->
+<!--   access to 'variety of fisheries' -->
+<!-- - some communities have limited ability to diversify; may need to diversify -->
+<!--   outside of fishing -->
 
-[@perruso2005]
-[@hilborn2001]
-[@edwards2004]
-[@trenkel2013]
-[@mumford2009]
-[@oostenbrugge2002]
-[@baldursson1997]
-[@jardine2015]
-[@sanchirico2008]
+<!-- [@perruso2005] -->
+<!-- [@hilborn2001] -->
+<!-- [@edwards2004] -->
+<!-- [@trenkel2013] -->
+<!-- [@mumford2009] -->
+<!-- [@oostenbrugge2002] -->
+<!-- [@baldursson1997] -->
+<!-- [@jardine2015] -->
+<!-- [@sanchirico2008] -->

### 2016-12-05 portfolio/text/ms.Rmd
Start developing result section

@@
 
-The state and federal fisheries of Alaska, provide a unique opportunity to
+The state and federal fisheries of Alaska provide a unique opportunity to
 understand the role of individuals' strategies in determining income and income
@@
 
-**Fig. \ref{fig:ts} - limited entry + movement towards salmon and specialization
-in Alaska**
-
-- move towards specialization, salmon fig 1D (overall declining trend disappears if you exclude salmon)
-- halibut most often combined with other permits (exclude permit holders that hold any halibut and about ninety percent of the remaining permit holders hold only one permit) fig 1c
-- Salmon the most often on its own (exclude salmon and the proportioned single permit decreases to about 70%) fig 1c
-- majority are single permit holders (70-90%) fig 1C, but the single permit holders make up only about half of the overall revenue (and surpassed half for the first time in 2013 and stayed that way) fig 1B
+Trends in fishing revenue and diversification over time for individuals fishing
+commercially in Alaska (Fig. \ref{fig:ts}A) reveal some movement toward
+specialization, and in particular specialization on salmon. The majority
+(70--90%) of fishers hold single permits, with a slight increase in single
+permit holders since around 2000 (Fig. \ref{fig:ts}C). This overall trend is
+most sensitive to individuals with halibut or salmon licenses --- halibut permit
+owners most often have multiple permits whereas salmon permit holders tend to be
+single permit holders more than individuals with permits for other fisheries
+(Fig. \ref{fig:ts}C). Although the majority of individuals hold single permits,
+multiple permit holders make approximately half of the overall revenue (Fig.
+\ref{fig:ts}B). Switching focus from permits owned to species landed, mean
+fished species diversity declined somewhat from 1980--2000 but has since
+remained steady (Fig. \ref{fig:ts}D). However, this trend appears to be driven
+by an increase in salmon specialists in the 1980s and 1990s --- the trend
+disappears if salmon license holders are excluded (Fig. \ref{fig:ts}D).
 
@@
 
-**Fig. \ref{fig:across} - across strategies**
+We observed a clear exponential decrease in revenue variability for strategies
+(groups of permits) with higher mean fished species diversity (Fig.
+\ref{fig:across}). Strategies with one unit higher mean species diversity had
+on average
+`r h1[[2]]`% (`r h1[[1]]`--`r h1[[3]]`% 95% CI) 
+lower standardized variability. This pattern looks qualitatively similar with or
+without individuals fishing salmon. The gillnet fishery for herring roe and the
+king crab fishery had the highest estimated revenue variability. Most of the
+strategies with the lowest estimated revenue variability combined halibut and
+salmon permits and had high levels of fished species diversity (Fig.
+\ref{fig:across}).
 
-- clear exponential decline in standardized variability with strategies that are inherently more species diverse
-- not a clear effect of days fished at this level
-- fishing permits with higher species diversity reduces variability
-(`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction per unit
-increase in mean effective species diversity);
-
-**Fig. \ref{fig:spokes} - add permits**
-
-- nearly universal reduction in variability (Fig. S\ref{fig:spoke-all})
-- e.g., `r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR]
-lower expected standard deviation
-for halibut plus another permit vs. single halibut permits);
-- usually associated with higher median revenue
-- but can be very expensive (provide some examples from website)
+Comparing single permit holders to other individuals holding that same permit
+plus one or more other permits illustrates the expected reduction in revenue
+variability from perching additional permits (Fig. \ref{fig:spokes},
+Fig. S\ref{fig:spoke-all}). For example, individuals fishing on a halibut longline
+permit plus an additional permit always had lower expected revenue variability
+compared to individuals fishing only a halibut longline permit 
+(median reduction of `r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR], Fig.
+\ref{fig:spokes}A). Likewise, adding a permit to a herring roe gillnet permit
+substantially reduced expected revenue variability. Adding permits also usually
+resulted in higher expected revenue, although the permits themselves represent
+a significant expense (e.g. XX).
 

### 2016-12-06 portfolio/text/ms.Rmd
More writing

@@
 
+*TODO: for the following into the results section or and the method section with
+this:*
+
+Residuals from the mean component of our model clearly show the "V"-shaped
+pattern that the variance component of our model seeks to capture (Fig.
+S\ref{fig:resid}). Furthermore, the downside residuals (year-individual data
+points with lower than expected revenue) look qualitatively similar to the
+complete set of residuals (upside and downside) (Fig.
+S\ref{fig:resid-downside}). Therefore, it seemed appropriate to model the
+variability as a proxy for revenue risk.
+
+The estimated year-permit intercepts represent years when people
+with a particular permit set in a given year tended to do better or worse than
+the long-term average for that permit set (Fig. S\ref{fig:str-yrs}). Therefore,
+an individual's expected revenue and variability is after accounting for the
+general trend in revenue for that permit set.
+
+Our model estimates overall average effects (Fig. S\ref{fig:fe}) and effects
+that vary by permit set (Fig. S\ref{fig:re}).
+
 # Results
@@
 disappears if salmon license holders are excluded (Fig. \ref{fig:ts}D).
-
-### Something about overall model fit
-
-- Parameters from hierarchical model (main: Fig. S\ref{fig:fe}, varying: S\ref{fig:re})
-- Residuals clearly show pattern that we see in the parameter estimates (Fig. S\ref{fig:resid})
-- Downside residuals look very similar and justify using the full data set
-(upside and downside) (Fig. S\ref{fig:resid-downside})
-- Mention the year-permit varying intercepts and what they mean
-(Fig. S\ref{fig:str-yrs}) --- good and bad years where
-everyone in that strategy tended to do better or worse ---
-this accounts for that and therefore compares everything to
-the expected revenue for that year
 
@@
 
-**Fig. \ref{fig:within} - within strategies**
+In contrast to the clear effect of species diversity on revenue variability when
+comparing across fishing strategies, the effect of changing fished species
+diversity for an individual *within* the constraints of a given permit set is
+more variable (Fig. \ref{fig:within}). After accounting for fishing effort (Fig.
+S\ref{fig:fe}), specializing (decreasing species diversity) tends to be
+associated with increased revenue (`r g2_dens`% probability; Fig.
+\ref{fig:within}). This was particularly clear for permits where individuals
+tend to specialize on sockeye salmon: individuals with setnet permits for salmon
+in Prince William Sound, Cook Inlet, Kodiak, Bristol Bay, or the Alaska
+Peninsula experienced a ~1.5% increase in revenue for a 1% decrease in fished
+species diversity (Fig. \ref{fig:within}B). The converse was usually true ---
+generalizing tends to be associated with lower expected revenue after accounting
+for fishing effort (i.e. Fig. \ref{fig:within}A and \ref{fig:within}B are nearly
+mirror images left to right). This specializing benefit on revenue is only
+apparent after accounting for fishing effort. Given that fishing for more
+species tends to be associated with increased fishing effort, generalizing tends
+to be associated with higher revenue when not standardizing for fishing effort
+(Fig. S-TODO).
 
-- fish more days, make more money (Fig. S\ref{fig:fe})
-- after accounting for days fished, most commonly,
-specializing is associated with higher revenue,
-but sometimes the reverse (Fig. \ref{fig:within})
-- if we don't account for changes in effort, then higher species diversity
-means more money fairly consistently (TODO make figure)
-- changing species diversity (number of species fished or evenness of species
-fished) tends to be associated with greater variability / uncertainty
-year-to-year (Figs. \ref{fig:within}, S\ref{fig:fe})
-- increasing species diversity within
-the constraints of a permit set has a fishery-dependent effect on revenue and
-is usually (`r g2_dens`% probability)
-associated with increased uncertainty in revenue.
-(Fig. \ref{fig:within})
+After accounting for fishing effort, adjusting (increasing or decreasing) one's
+fished species diversity within the constraints of a permit set tends to be
+associated with greater expected variability or uncertainty in year-to-year
+revenue (Fig. \ref{fig:within} strategies above 0, Fig. S\ref{fig:fe}). This
+effect was strongest for individuals with a halibut longline permit or halibut
+and sablefish longline permits. Individuals with these strategies experienced
+~2% increase in standardized variability for a 1% change in fished species
+diversity. Fished species diversity for these permits is primarily in the form
+of bycatch.
+
+The Alaskan halibut and sablefish fisheries were rationalized in 199X --- they
+switched from XX to a individual fishery quota (IFQ) system. We refit our models
+allowing all effects to vary separately before and after 19XX for strategies
+involving these permits (i.e. treating them as separate strategies). Variability
+was reduced by ~25% (XX--XX 95% CI) after the introduction of these IFQs
+compared to before for individuals with only a halibut or sablefish permit (Fig.
+SX). The cost of changing fished species diversity for an individual in terms of
+the effect on revenue variability increased after the introduction of IFQs for
+all strategies, and especially for individuals fishing only halibut and
+sablefish (Fig. SX).
 

### 2016-12-06 portfolio/text/ms.Rmd
Create and add IFQ figure to supplement

@@
 The Alaskan halibut and sablefish fisheries were rationalized in 199X --- they
-switched from XX to a individual fishery quota (IFQ) system. We refit our models
-allowing all effects to vary separately before and after 19XX for strategies
-involving these permits (i.e. treating them as separate strategies). Variability
-was reduced by ~25% (XX--XX 95% CI) after the introduction of these IFQs
-compared to before for individuals with only a halibut or sablefish permit (Fig.
-SX). The cost of changing fished species diversity for an individual in terms of
-the effect on revenue variability increased after the introduction of IFQs for
-all strategies, and especially for individuals fishing only halibut and
-sablefish (Fig. SX).
+switched from XX to a individual fishery quota (IFQ) system. We refit our
+models allowing all effects to vary separately before and after 19XX for
+strategies involving these permits (i.e. treating them as separate strategies).
+Variability was reduced by ~25% (XX--XX 95% CI) after the introduction of these
+IFQs compared to before for individuals with only a halibut or sablefish permit
+(Fig. S\ref{fig:ifq}A). The cost of changing fished species diversity for an
+individual in terms of the effect on revenue variability increased after the
+introduction of IFQs for most strategies, and especially for individuals fishing
+only halibut and sablefish (Fig. S\ref{fig:ifq}B).
 

### 2016-12-06 portfolio/text/supp.Rnw
Create and add IFQ figure to supplement

@@
 
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=1\textwidth]{../figs/itq-combined.pdf}
+\caption{The effect of IFQs (individual fishing quotas) on model estimates. (A)
+  Posterior density of the ratio between strategy-level estimates of
+  standardized revenue variability ($\gamma_{0j}$'s) after the introduction of
+  IFQs vs.\ before the introduction of IFQs. (B) Estimates (medians) of the
+  effect of generalizing (increasing species diversity) on variability
+  ($\gamma_{1j}$'s) for the full model (``Grouped'') compared to a model where
+  separate effects are estimated before and after IFQs were introduced (``IFQ
+  split''). Pattern looks similar for the effect of specializing on variability
+  ($\gamma_{2j}$'s) and is not shown.}
+\label{fig:ifq}
+\end{center}
+\end{figure}
+
 % \begin{figure}[htbp]

### 2016-12-06 portfolio/analysis/3.3-compare-ifq.R

@@ -56,8 +56,8 @@ res_ifq <- mutate(res_ifq,
 # ----------
 # plot the effects between the 2 models
 
-plot_ifq_effect <- function(variable = "inc", title = "", flip = FALSE,
-  confidence_intervals = FALSE) {
+plot_ifq_effect <- function(variable = "inc", title = "", flip = FALSE, lab = "",
+  confidence_intervals = FALSE, cols = RColorBrewer::brewer.pal(8, "Greys")[c(4, 8)]) {
   d1 <- select_(res, "strategy", variable) %>%
     rename_(effect = variable)
   d2 <- select_(res_ifq, "strategy_ifq", "strategy", variable) %>%
@@ -65,7 +65,8 @@ plot_ifq_effect <- function(variable = "inc", title = "", flip = FALSE,
   d <- left_join(d1, d2) %>%
     group_by(strategy) %>%
     mutate(ifq = grepl("IFQ", strategy_ifq)) %>%
-    mutate(ifq_any = any(grepl("IFQ", strategy_ifq)))
+    mutate(ifq_any = any(grepl("IFQ", strategy_ifq))) %>%
+    left_join(select(md, strategy, str_label))
 
   if (flip) {
     d$effect <- -1 * d$effect
@@ -74,20 +75,31 @@ plot_ifq_effect <- function(variable = "inc", title = "", flip = FALSE,
 
   d <- d %>%
     filter(ifq_any)
-  g <- ggplot(d, aes(x = "Grouped", xend = "IFQ split", y = effect, yend = effect_ifq,
-    color = ifq)) + geom_segment() +
-    geom_text(data = unique(select(d, effect, strategy)),
-      aes(x = "Grouped", y = effect, label = strategy),
-      hjust = "right", color = "black", inherit.aes = FALSE) +
-    labs(title = title, color = "IFQ") +
-    theme_light()
+  g <- ggplot(d, aes(x = 1, xend = 2, y = effect, yend = effect_ifq,
+    color = ifq, linetype = ifq)) + geom_segment() +
+    geom_text(data = unique(select(d, effect, strategy, str_label)),
+      aes(x = 1, y = effect, label = str_label),
+      hjust = "right", color = "black", inherit.aes = FALSE, size = 3) +
+    labs(color = "IFQ", linetype = "IFQ") +
[…]

### 2016-12-06 portfolio/analysis/3.3-ifq-plots.R

@@ -1,0 +1,66 @@
+## The following code implements the posterior density ratio plots for all the parameters
+## I'm not currently using it
+
+## library(tidyverse)
+##
+## load("portfolio/data-generated/diff-dat-stan.rda")
+## load("portfolio/data-generated/m_ifq.rda")
+## load("portfolio/data-generated/m.rda")
+## devtools::load_all("pfxr")
+##
+## p <- stanhelpers::extract_df(m_ifq)
+##
+## get_ratio_posteriors <- function(effect_name, breaks = NULL, lab = "") {
+##   effects_dat <- p[[effect_name]] %>%
+##     tidyr::gather(strategy, posterior) %>%
+##     mutate(strategy_ifq_id =
+##              as.numeric(sub(paste0(effect_name, "_([0-9]+)"), "\\1", strategy))) %>%
+##     select(-strategy) %>%
+##     inner_join(md_ifq)
+##
+##   get_ifq_ratio <- function(strategy) {
+##     before <- filter(effects_dat, strategy_ifq == strategy)
+##     after <- filter(effects_dat, strategy_ifq == paste(strategy, "IFQ"))
+##     ratio <- exp(after$posterior) / exp(before$posterior)
+##     data.frame(strategy = strategy, ratio = ratio)
+##   }
+##
+##   s <- unique(effects_dat$strategy_ifq)
+##   strategies <- gsub(" IFQ", "", s[grepl("IFQ", s)])
+##
+##   out <- plyr::ldply(strategies, get_ifq_ratio)
+##   out <- out %>% filter(ratio < 200) %>%
+##     group_by(strategy) %>%
+##     mutate(median_ratio = median(ratio)) %>%
+##     ungroup() %>%
+##     left_join(select(md, strategy, str_label))
+##
+##   labs <- unique(select(out, median_ratio, str_label)) %>%
+##     mutate(str_label_ordered = factor(str_label,
+##                                       levels = str_label[order(-median_ratio)])) %>%
+##     select(-median_ratio)
+##
+##   out$str_label_ordered <- NULL
+##   out <- inner_join(out, labs)
+##
+##   g <- ggplot(out, aes(str_label_ordered, ratio)) +
+##     geom_hline(yintercept = 1, col = "grey65") +
+##     geom_violin() +
+##     scale_y_log10(breaks = breaks) +
[…]

### 2016-12-06 portfolio/text/ms.Rmd
Edit results

@@
 Trends in fishing revenue and diversification over time for individuals fishing
-commercially in Alaska (Fig. \ref{fig:ts}A) reveal some movement toward
-specialization, and in particular specialization on salmon. The majority
-(70--90%) of fishers hold single permits, with a slight increase in single
-permit holders since around 2000 (Fig. \ref{fig:ts}C). This overall trend is
-most sensitive to individuals with halibut or salmon licenses --- halibut permit
-owners most often have multiple permits whereas salmon permit holders tend to be
-single permit holders more than individuals with permits for other fisheries
-(Fig. \ref{fig:ts}C). Although the majority of individuals hold single permits,
-multiple permit holders make approximately half of the overall revenue (Fig.
-\ref{fig:ts}B). Switching focus from permits owned to species landed, mean
-fished species diversity declined somewhat from 1980--2000 but has since
-remained steady (Fig. \ref{fig:ts}D). However, this trend appears to be driven
-by an increase in salmon specialists in the 1980s and 1990s --- the trend
-disappears if salmon license holders are excluded (Fig. \ref{fig:ts}D).
+commercially in Alaska (Fig. \ref{fig:ts}A) reveal a transition toward
+specialization, and in particular specialization on fishing salmon. The majority
+(70--90%) of fishers hold single permits, with a slight increase in
+single-permit holders since ~2000 (Fig. \ref{fig:ts}C). This overall trend is
+most sensitive to halibut or salmon permit holders --- compared to other
+fisheries, halibut permit holders are more likely to hold multiple permits
+whereas salmon permit holders are more likely to be single permit holders (Fig.
+\ref{fig:ts}C). Despite making up only 10--30% of the individuals, multiple
+permit holders make approximately half the overall revenue (Fig. \ref{fig:ts}B).
+The mean fished species diversity of this revenue declined somewhat from
+1980--2000 but has since remained steady (Fig. \ref{fig:ts}D). However, this
+trend appears to be driven by an increase in salmon specialists in the 1980s and
+1990s --- the trend disappears if salmon license holders are excluded (Fig.
+\ref{fig:ts}D).
 
-We observed a clear exponential decrease in revenue variability for strategies
+We observe a clear exponential decrease in revenue variability for strategies
 (groups of permits) with higher mean fished species diversity (Fig.
-\ref{fig:across}). Strategies with one unit higher mean species diversity had
-on average
-`r h1[[2]]`% (`r h1[[1]]`--`r h1[[3]]`% 95% CI) 
-lower standardized variability. This pattern looks qualitatively similar with or
-without individuals fishing salmon. The gillnet fishery for herring roe and the
-king crab fishery had the highest estimated revenue variability. Most of the
-strategies with the lowest estimated revenue variability combined halibut and
-salmon permits and had high levels of fished species diversity (Fig.
+\ref{fig:across}). Strategies with one unit higher mean effective species
+diversity had on average `r h1[[2]]`% (`r h1[[1]]`--`r h1[[3]]`% 95% CI) lower
+standardized variability. This pattern looks qualitatively similar with or
+without individuals targeting salmon. The gillnet fishery for herring roe and
+the king crab fishery had the highest estimated revenue variability. Most of the
+strategies with the lowest estimated revenue variability combined halibut with
+a salmon permit and had high levels of fished species diversity (Fig.
 \ref{fig:across}).
@@
 plus one or more other permits illustrates the expected reduction in revenue
-variability from perching additional permits (Fig. \ref{fig:spokes},
-Fig. S\ref{fig:spoke-all}). For example, individuals fishing on a halibut longline
-permit plus an additional permit always had lower expected revenue variability
-compared to individuals fishing only a halibut longline permit 
-(median reduction of `r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR], Fig.
-\ref{fig:spokes}A). Likewise, adding a permit to a herring roe gillnet permit
-substantially reduced expected revenue variability. Adding permits also usually
-resulted in higher expected revenue, although the permits themselves represent
-a significant expense (e.g. XX).
+variability from purchasing additional permits (Fig. \ref{fig:spokes}, Fig.
+S\ref{fig:spoke-all}). For example, individuals fishing a halibut longline
+permit plus one additional permit always had lower expected revenue variability
+compared to individuals fishing only a halibut longline permit (median reduction
+of `r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR], Fig.
+\ref{fig:spokes}A). Likewise, adding a permit to a herring roe gillnet strategy
+substantially reduced expected revenue variability (Fig. \ref{fig:spokes}B).
+Adding permits also usually resulted in higher expected revenue, although the
+permits themselves represent a significant expense (e.g. XX).
 
-In contrast to the clear effect of species diversity on revenue variability when
-comparing across fishing strategies, the effect of changing fished species
-diversity for an individual *within* the constraints of a given permit set is
-more variable (Fig. \ref{fig:within}). After accounting for fishing effort (Fig.
-S\ref{fig:fe}), specializing (decreasing species diversity) tends to be
-associated with increased revenue (`r g2_dens`% probability; Fig.
-\ref{fig:within}). This was particularly clear for permits where individuals
-tend to specialize on sockeye salmon: individuals with setnet permits for salmon
-in Prince William Sound, Cook Inlet, Kodiak, Bristol Bay, or the Alaska
-Peninsula experienced a ~1.5% increase in revenue for a 1% decrease in fished
-species diversity (Fig. \ref{fig:within}B). The converse was usually true ---
-generalizing tends to be associated with lower expected revenue after accounting
-for fishing effort (i.e. Fig. \ref{fig:within}A and \ref{fig:within}B are nearly
-mirror images left to right). This specializing benefit on revenue is only
-apparent after accounting for fishing effort. Given that fishing for more
-species tends to be associated with increased fishing effort, generalizing tends
-to be associated with higher revenue when not standardizing for fishing effort
-(Fig. S-TODO).
+In contrast to the clear across-strategy effects of species diversity on revenue
+and revenue variability, the effect of changing fished species diversity for an
+individual *within* the constraints of a given permit set is more variable (Fig.
+\ref{fig:within}). After accounting for fishing effort (Fig. S\ref{fig:fe}),
+specializing (decreasing species diversity) tends to be associated with
+increased revenue (`r g2_dens`% probability; Fig. \ref{fig:within}). This is
+particularly clear for permits where individuals tend to specialize on sockeye
+salmon: individuals with setnet permits for salmon in Prince William Sound, Cook
+Inlet, Kodiak, Bristol Bay, or the Alaska Peninsula experience a ~1.5% increase
+in revenue for a 1% decrease in fished species diversity (Fig.
+\ref{fig:within}B). The converse is also usually true --- generalizing tends to
+be associated with lower expected revenue after accounting for fishing effort
+(i.e. Fig. \ref{fig:within}A and \ref{fig:within}B are near mirror images left
+to right). This specializing benefit for revenue is only apparent *after*
+accounting for fishing effort. Given that fishing for more species tends to be
+associated with increased fishing effort, generalizing tends to be associated
+with higher revenue when not standardizing for fishing effort (Fig. S-TODO).
 
-After accounting for fishing effort, adjusting (increasing or decreasing) one's
-fished species diversity within the constraints of a permit set tends to be
-associated with greater expected variability or uncertainty in year-to-year
-revenue (Fig. \ref{fig:within} strategies above 0, Fig. S\ref{fig:fe}). This
-effect was strongest for individuals with a halibut longline permit or halibut
-and sablefish longline permits. Individuals with these strategies experienced
-~2% increase in standardized variability for a 1% change in fished species
-diversity. Fished species diversity for these permits is primarily in the form
-of bycatch.
+After accounting for fishing effort, adjusting (increasing or decreasing) an
+individual's fished species diversity within the constraints of a permit set
+tends to be associated with greater expected variability or uncertainty in
+year-to-year revenue (Fig. \ref{fig:within} strategies above 0; Fig.
+S\ref{fig:fe}). This effect was strongest for individuals with a halibut
+longline permit or halibut and sablefish longline permits. Individuals with
+these strategies experienced ~2% increase in standardized variability for a 1%
+change in fished species diversity (Fig. \ref{fig:within}). Fished species
+diversity levels above one for these permits is primarily due to bycatch.
 
-The Alaskan halibut and sablefish fisheries were rationalized in 199X --- they
-switched from XX to a individual fishery quota (IFQ) system. We refit our
-models allowing all effects to vary separately before and after 19XX for
+The Alaskan halibut and sablefish fisheries were "rationalized" in 199X --- they
+switched from XX to an individual fishery quota (IFQ) system. We refit our
+models allowing all effects to vary separately before and after 1995 for
 strategies involving these permits (i.e. treating them as separate strategies).
-Variability was reduced by ~25% (XX--XX 95% CI) after the introduction of these
-IFQs compared to before for individuals with only a halibut or sablefish permit
-(Fig. S\ref{fig:ifq}A). The cost of changing fished species diversity for an
-individual in terms of the effect on revenue variability increased after the
-introduction of IFQs for most strategies, and especially for individuals fishing
-only halibut and sablefish (Fig. S\ref{fig:ifq}B).
+The introduction of IFQs reduced variability by ~25% (XX--XX 95% CI) for
+individuals with only a halibut or sablefish permit (Fig. S\ref{fig:ifq}A). For
+most strategies, the effect of increased revenue variability associated with
+changing fished species diversity became stronger after the introduction of IFQs
+(Fig. S\ref{fig:ifq}B). This was especially true for individuals fishing only
+halibut or sablefish (Fig. S\ref{fig:ifq}B).
 

### 2016-12-06 portfolio/text/ms.Rmd
Keep editing results

@@
 
-Trends in fishing revenue and diversification over time for individuals fishing
-commercially in Alaska (Fig. \ref{fig:ts}A) reveal a transition toward
-specialization, and in particular specialization on fishing salmon. The majority
-(70--90%) of fishers hold single permits, with a slight increase in
-single-permit holders since ~2000 (Fig. \ref{fig:ts}C). This overall trend is
+We evaluated trends in revenue and diversification over time for individuals
+fishing commercially in Alaska from 1985--2014 (Fig. \ref{fig:ts}A). The
+majority (70--90%) of fishers hold single permits, with a slight increase in
+single-permit holders since 2000 (Fig. \ref{fig:ts}C). This overall trend is
 most sensitive to halibut or salmon permit holders --- compared to other
-fisheries, halibut permit holders are more likely to hold multiple permits
-whereas salmon permit holders are more likely to be single permit holders (Fig.
-\ref{fig:ts}C). Despite making up only 10--30% of the individuals, multiple
-permit holders make approximately half the overall revenue (Fig. \ref{fig:ts}B).
-The mean fished species diversity of this revenue declined somewhat from
-1980--2000 but has since remained steady (Fig. \ref{fig:ts}D). However, this
-trend appears to be driven by an increase in salmon specialists in the 1980s and
-1990s --- the trend disappears if salmon license holders are excluded (Fig.
-\ref{fig:ts}D).
+fisheries, halibut permit holders are more likely to hold halibut plus other
+permits, whereas salmon permit holders are more likely to hold only one permit
+(Fig. \ref{fig:ts}C). Despite making up only 10--30% of the individuals,
+multiple permit holders make approximately half of the overall revenue (Fig.
+\ref{fig:ts}B). The mean effective fished species diversity of this revenue
+declined somewhat from 1980--2000 but has since remained steady (Fig.
+\ref{fig:ts}D). However, this trend appears to be driven by an increase in
+salmon specialists in the 1980s and 1990s --- the trend disappears if salmon
+license holders are excluded (Fig. \ref{fig:ts}D).
 
-We observe a clear exponential decrease in revenue variability for strategies
-(groups of permits) with higher mean fished species diversity (Fig.
-\ref{fig:across}). Strategies with one unit higher mean effective species
-diversity had on average `r h1[[2]]`% (`r h1[[1]]`--`r h1[[3]]`% 95% CI) lower
-standardized variability. This pattern looks qualitatively similar with or
-without individuals targeting salmon. The gillnet fishery for herring roe and
-the king crab fishery had the highest estimated revenue variability. Most of the
-strategies with the lowest estimated revenue variability combined halibut with
-a salmon permit and had high levels of fished species diversity (Fig.
-\ref{fig:across}).
+Revenue variability decreases exponentially with higher mean fished species
+diversity across permit sets (Fig. \ref{fig:across}). Controlling for effort,
+strategies with one unit higher mean effective species diversity had on average
+`r h1[[2]]`% (`r h1[[1]]`--`r h1[[3]]`% 95% CI) lower variability. This pattern
+looks qualitatively similar with or without individuals targeting salmon. As
+examples, variability was highest for the herring roe gillnet and king crab
+fisheries. Most of the strategies with the lowest estimated revenue variability
+combined halibut with a salmon permit and had high levels of fished species
+diversity (Fig. \ref{fig:across}).
 
-Comparing single permit holders to other individuals holding that same permit
-plus one or more other permits illustrates the expected reduction in revenue
-variability from purchasing additional permits (Fig. \ref{fig:spokes}, Fig.
+Comparing single permit holders to individuals holding that same permit plus
+additional permits illustrates the expected reduction in revenue variability
+from purchasing additional permits (Fig. \ref{fig:spokes}, Fig.
 S\ref{fig:spoke-all}). For example, individuals fishing a halibut longline
@@
 substantially reduced expected revenue variability (Fig. \ref{fig:spokes}B).
-Adding permits also usually resulted in higher expected revenue, although the
-permits themselves represent a significant expense (e.g. XX).
+Adding permits also usually increased expected revenue, although the permits
+themselves represent a significant expense (e.g. TODO).
 

### 2016-12-06 portfolio/text/ms.Rmd
Edit results of commercial portfolio paper

@@
 specializing (decreasing species diversity) tends to be associated with
-increased revenue (`r g2_dens`% probability; Fig. \ref{fig:within}). This is
+increased revenue (~0.5% increased revenue per 1% decrease in species diversity
+on average or TODO ... `r g2_dens`% probability; Fig. \ref{fig:within}). This is
 particularly clear for permits where individuals tend to specialize on sockeye
-salmon: individuals with setnet permits for salmon in Prince William Sound, Cook
-Inlet, Kodiak, Bristol Bay, or the Alaska Peninsula experience a ~1.5% increase
-in revenue for a 1% decrease in fished species diversity (Fig.
-\ref{fig:within}B). The converse is also usually true --- generalizing tends to
-be associated with lower expected revenue after accounting for fishing effort
-(i.e. Fig. \ref{fig:within}A and \ref{fig:within}B are near mirror images left
-to right). This specializing benefit for revenue is only apparent *after*
-accounting for fishing effort. Given that fishing for more species tends to be
-associated with increased fishing effort, generalizing tends to be associated
-with higher revenue when not standardizing for fishing effort (Fig. S-TODO).
+salmon. For example, individuals with setnet permits for salmon in Prince
+William Sound, Cook Inlet, Kodiak, Bristol Bay, or the Alaska Peninsula target
+primarily sockeye salmon and experience a ~1.5% increase in revenue for a 1%
+decrease in fished species diversity (Fig. \ref{fig:within}B). The converse is
+also usually true --- generalizing tends to be associated with lower expected
+revenue after accounting for fishing effort (i.e. Fig. \ref{fig:within}A and
+\ref{fig:within}B are near mirror images left to right). This benefit of
+specializing is only apparent after accounting for fishing effort. Given that
+fishing for more species tends to be associated with increased fishing effort
+(fishing more days), generalizing tends to be associated with higher revenue
+when not standardizing for days finished (Fig. S-TODO).
 
 After accounting for fishing effort, adjusting (increasing or decreasing) an
-individual's fished species diversity within the constraints of a permit set
-tends to be associated with greater expected variability or uncertainty in
-year-to-year revenue (Fig. \ref{fig:within} strategies above 0; Fig.
-S\ref{fig:fe}). This effect was strongest for individuals with a halibut
+individual's fished species diversity from year-to-year within the constraints
+of a permit set tends to be associated with greater expected variability or
+uncertainty in year-to-year revenue (Fig. \ref{fig:within} strategies above 0;
+Fig. S\ref{fig:fe}). This effect was strongest for individuals with a halibut
 longline permit or halibut and sablefish longline permits. Individuals with
 these strategies experienced ~2% increase in standardized variability for a 1%
-change in fished species diversity (Fig. \ref{fig:within}). Fished species
-diversity levels above one for these permits is primarily due to bycatch.
+change in fished species diversity (Fig. \ref{fig:within}). Levels of fished
+species diversity levels above one for these permits are primarily due to
+bycatch.
 
-The Alaskan halibut and sablefish fisheries were "rationalized" in 199X --- they
-switched from XX to an individual fishery quota (IFQ) system. We refit our
-models allowing all effects to vary separately before and after 1995 for
-strategies involving these permits (i.e. treating them as separate strategies).
+The Alaskan halibut and sablefish fisheries were converted to an individual
+fishery quota (IFQ) system in 1994. This catch share system... TODO. We refit
+our models allowing all permit-level effects to vary separately before and after
+the introduction of IFQs for strategies involving halibut and sablefish permits.
 The introduction of IFQs reduced variability by ~25% (XX--XX 95% CI) for
-individuals with only a halibut or sablefish permit (Fig. S\ref{fig:ifq}A). For
-most strategies, the effect of increased revenue variability associated with
-changing fished species diversity became stronger after the introduction of IFQs
-(Fig. S\ref{fig:ifq}B). This was especially true for individuals fishing only
-halibut or sablefish (Fig. S\ref{fig:ifq}B).
+individuals with only halibut or sablefish permits (Fig. S\ref{fig:ifq}A).
+Furthermore, for most strategies, the effect of increased revenue variability
+associated with changing species diversity became stronger after the
+introduction of IFQs (Fig. S\ref{fig:ifq}B). This was especially true for
+individuals fishing only halibut or sablefish and less so for individuals
+combining halibut or sable fish with a salmon permit (Fig. S\ref{fig:ifq}B).
 

### 2016-12-06 portfolio/text/ms.Rmd
Rewrap and edit paper

@@
 A fundamental challenge for individuals who rely on the extraction of natural
-resources for their livelihood is income variability.
-Fishers, loggers, farmers, and miners, among others,
-face income uncertainty driven by a mix of
-environmental, biological, and economic variability.
-A key component underlying the sustainability
-of these industries is identifying both
-how regulatory actions and individual behaviour affect income variability
-and by extension the economic stability and
-sustainable use of natural resources.
-For commercial fisheries, recent work has identified that
-targeting a diversity of species can reduce income variability.
-However, any diversity-variability relationship
-is embedded in a regulatory framework that restricts individual ability to
-change fishing behaviour. Here we identify three primary axes by which fishers
-in Alaska, one of the world's most productive fishery regions, can potentially
-diversify fishing activities and investigate how individual actions along each
-diversity axis affect both revenue and revenue variability. We construct a novel
-statistical model to simultaneously assess the value to individual fishers of
-choosing to (1) own permits that allow targeting a diverse fish species
-assemblage, (2) diversify by owning multiple permits, or (3) catch a greater
-species diversity within the constraints of a given permit.
-We show contrasting effects of diversity on
-revenue and revenue variability along the three axes.
-First, we show that fishing permits with
-higher species diversity reduces variability
-(`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction per unit
-increase in mean effective species diversity).
-Next, we show that fishing an additional permit
-is associated with higher revenue
-and lower variability (e.g.,
-`r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR] lower
-expected standard deviation
-for halibut plus another permit vs. single halibut permits).
-Finally, we show that increasing species diversity within
-the constraints of a permit set has a fishery-dependent effect on revenue and
-is usually (`r g2_dens`% probability)
-associated with increased uncertainty in revenue the following year.
-Thus, while there are clear benefits
-to some axes of diversification, others are less effective.
-Importantly, the most effective measure for
-increasing revenue and decreasing risk ---
-purchasing additional permits to participate in additional fisheries ---
-was also the most costly.
-At the same time, Alaskan commercial fisheries
-have been increasingly focussed on
-single-permit (often salmon) fishing strategies,
-with single-permit fishers earning
-more than multiple-permit fishers
-for the first time in 2013.
-Our results suggest that the benefit of species diversity
-for revenue variability
-is mostly an across-permit effect, not
-a benefit that can be easily harnessed by an individual permit holder.
-Effective methods of diversification ---
-purchasing additional permits or switching permits ---
-are costly and often limited by regulations such as limited entry
-programs and individual transferable quotas (ITQs).
-Take home ...
+resources for their livelihood is income variability. Fishers, loggers, farmers,
+and miners, among others, face high levels of income uncertainty driven by a mix
+of environmental, biological, and economic variability. A key component
+underlying the sustainability of these industries is identifying both how
+regulatory actions and individual behaviour affect income variability and by
+extension the economic stability and sustainable use of natural resources. For
+commercial fisheries, recent work has identified that targeting a diversity of
+species can reduce income variability. However, any diversity-variability
+relationship is embedded in a regulatory framework that restricts individual
+ability to change fishing behaviour. Here we identify three primary axes by
+which fishers in Alaska, one of the world's most productive fishery regions, can
+potentially diversify fishing activities and investigate how individual actions
+along each diversity axis affect both revenue and revenue variability. We
+construct a novel statistical model to simultaneously assess the value to
+individual fishers of choosing to (1) own permits that allow targeting a diverse
+fish species assemblage, (2) diversify by owning multiple permits, or (3) catch
+a greater species diversity within the constraints of a given permit. We show
+contrasting effects of diversity on revenue and revenue variability along the
+three axes. First, we show that fishing permits with higher species diversity
+reduces variability (`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction
+per unit increase in mean effective species diversity). Next, we show that
+fishing an additional permit is associated with higher revenue and lower
+variability (e.g., `r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`%
+IQR] lower expected standard deviation for halibut plus another permit vs.
+single halibut permits). Finally, we show that increasing species diversity
+within the constraints of a permit set has a fishery-dependent effect on revenue
+and is usually (`r g2_dens`% probability) associated with increased uncertainty
+in revenue the following year. Thus, while there are clear benefits to some axes
+of diversification, others are less effective. Importantly, the most effective
+measure for increasing revenue and decreasing risk --- purchasing additional
+permits to participate in additional fisheries --- was also the most costly. At
+the same time, Alaskan commercial fisheries have been increasingly focussed on
+single-permit (often salmon) fishing strategies. Our results suggest that the
+benefit of species diversity for revenue variability is mostly an across-permit
+effect, not a benefit that can be easily harnessed by an individual permit
+holder. Effective methods of diversification --- purchasing additional permits
+or switching permits --- are costly and often limited by regulations such as
+limited entry programs and individual transferable quotas (ITQs). Take home ...
 
@@
 incentivize against diversification. For example, in terrestrial agriculture,
-crop subsidies in the US may incentivize some farms to specialize [e.g. dairy,
-grains, soybeans; @mishra2002]. In other instances, external market forces
+crop subsidies in the US may incentivize some farms to specialize (e.g. dairy,
+grains, soybeans; @mishra2002). In other instances, external market forces
 (prices, demand) may discourage the adoption of diverse strategies and help
@@
 extreme risk that some individuals experience, particularly relative to
-terrestrial agriculture (@sethi2012b). The majority of insurance programs
+terrestrial agriculture [@sethi2012b]. The majority of insurance programs
 available to terrestrial farmers (e.g. crop insurance programs) are not
@@
 they limit the available strategies to reduce income variability. For example,
-most permits limit fishing to a single species, diversifying by purchasing
+many permits limit fishing to a single species, diversifying by purchasing
 additional fishing permits can be expensive (e.g. ~\$XX for king crab), and some
@@
 vessels we can track people that switch fishery sectors over time, or alter
-strategies and use multiple vessels. We develop and apply a novel statistical
-approach for understanding the effect of diversity on both revenue and revenue
-variability that builds on results from vessel [@kasperski2013], fleet
-[@sethi2012b], and community-level variability [@sethi2014]. Our model controls
-for fishing effort and latent time effects (e.g. changing environmental
+strategies and use multiple vessels. We develop and apply a hierarchical
+Bayesian model to understand the effect of diversity on both individual revenue
+and revenue variability that builds on results from vessel [@kasperski2013],
+fleet [@sethi2012b], and community-level variability [@sethi2014]. Our model
+controls for fishing effort and latent time effects (e.g. changing environmental
 conditions), simultaneously estimates the effect of species diversity on
@@
 
-Overview paragraph here.
+We developed a hierarchical variance-function regression model, which separately
+models changes in the mean and variance of individual revenue as a function of
+covariates [@davidian1987; @hall1989]. Whereas traditional hierarchical linear
+models assume homoscedasticity --- that the residual variance does not vary
+systematically --- the variance-function regression model allows the variance to
+be heteroscedastic and vary as a function of covariates. For application to
+economic data on individual fishing portfolios, this model allows certain
+strategies and levels of diversification to be more or less variable than
+others. Our model includes the number of days fished to represent fishing
+effort,
 
@@
 
-We obtained fisheries landings and revenue data for
-all permit holders in Alaska
-from `r range_yrs[1]` to `r range_yrs[2]`
-from the Alaskan Commercial Fisheries Entry Commission (CFEC).
-These data represent a total of
-nearly `r round(ndat$initial_tot_rev/1e9)` billion 2009 USD
-and a combined `r round(ndat$initial/1e6,1)` million rows of data.
-We adjusted for inflation by setting all revenues in 2009 USD
-based on data from TODO.
-We implemented a number of filtering steps
-to focus on individuals actively engaged in commercial fishing
-(Supporting Information).
-In particular, we removed permit holders whose median revenue
-was less than \$10,000 2009 USD.
+We obtained fisheries landings and revenue data for all permit holders in Alaska
+from `r range_yrs[1]` to `r range_yrs[2]` from the Alaskan Commercial Fisheries
+Entry Commission (CFEC). These data represent a total of nearly `r 
+round(ndat$initial_tot_rev/1e9)` billion 2009 USD and a combined `r 
+round(ndat$initial/1e6,1)` million rows of data. We adjusted for inflation by
+setting all revenues in 2009 USD [@bea2016]. We implemented a number of
+filtering steps to focus on individuals actively engaged in commercial fishing
+(Supporting Information). In particular, we removed permit holders whose median
+revenue was less than \$10,000 2009 USD.
 
-We combined similar permits into permit groupings
-and identified sets of
-permit groupings held by an individual
-in a specific year as 'strategies' (Supporting Information).
-For example, we combined permits that were the same except for vessel size
-(e.g. a longliner fishing sablefish on a vessel under 60 ft was
-considered to have the same strategy as a longliner fishing
-sablefish on a vessel over 60 ft),
-and we combined a number of permits that were targeting a single species
-and differed only in region fished.
-Therefore, someone fishing herring roe in Kodiak
-and someone fishing herring roe in Alaska Peninsula
-were both considered to have a 'herring roe' permit group.
-For salmon permits,
-we grouped permits using the same gear type across regions
-for groups that shared similar species composition.
-After this process of combining permits,
-we were left with `r n_grouped_permits`
-unique fishing permit groups (Table S\ref{tab:strategies}),
-which we aggregated
-within each person-year combination
-to form permit strategies.
-For example, if someone fished halibut
-and sablefish in the same year,
-their strategy for that year would be "halibut-sablefish".
-Specific details on our strategy definitions are available in
-the Supporting Information.
+We combined similar permits into permit groupings and identified sets of permit
+groupings held by an individual in a specific year as 'strategies' (Supporting
+Information). For example, we combined permits that were the same except for
+vessel size (e.g. a longliner fishing sablefish on a vessel under 60 ft was
+considered to have the same strategy as a longliner fishing sablefish on
+a vessel over 60 ft), and we combined a number of permits that were targeting
+a single species and differed only in region fished. Therefore, someone fishing
+herring roe in Kodiak and someone fishing herring roe in Alaska Peninsula were
+both considered to have a 'herring roe' permit group. For salmon permits, we
+grouped permits using the same gear type across regions for groups that shared
+similar species composition. After this process of combining permits, we were
+left with `r n_grouped_permits` unique fishing permit groups (Table
+S\ref{tab:strategies}), which we aggregated within each person-year combination
+to form permit strategies. For example, if someone fished halibut and sablefish
+in the same year, their strategy for that year would be "halibut-sablefish".
+Specific details on our strategy definitions are available in the Supporting
+Information.
 
-Our models focused on species diversity
-as a possible predictor of revenue and revenue variability
-while controlling for effort.
-For consistency with previous analyses [e.g. @kasperski2013],
-we calculated effective species diversity for
-permit holder $i$ and time (year) $t$ as the inverse of
-Simpson's diversity index $\lambda$
-(\citeauthor{simpson1949} \citeyear{simpson1949},
-\citeauthor{jost2006} \citeyear{jost2006};
-also referred to in economics as the Herfindahl-Hirschman Index)
-weighted by revenue $R$:
+Our models focused on species diversity as a possible predictor of revenue and
+revenue variability while controlling for effort. For consistency with previous
+analyses [@kasperski2013], we calculated effective species diversity for
+permit holder $i$ and time (year) $t$ as the inverse of Simpson's diversity
+index $\lambda$ [@simpson1949; @jost2006] (also referred to in economics as the
+Herfindahl-Hirschman Index) weighted by revenue $R$:
 
@@
 
-\noindent
-where $s$ indexes species from $1$ through $n$.
-We calculated effort or effective season length as
-the sum of the season lengths for each permit each year.
-Season length for a given permit
-was calculated as the number of days between
-the first instance of fishing a permit
-and the last instance of fishing a permit each year.
-For example, if someone fished for 30 days
-on a salmon drift gillnet permit in Southeast Alaska
-and 40 days on a statewide longline halibut permit,
-that individual's season length
-would be 70 days.
+\noindent where $s$ indexes species from $1$ through $n$. We calculated effort
+or effective season length as the sum of the season lengths for each permit each
+year. Season length for a given permit was calculated as the number of days
+between the first instance of fishing a permit and the last instance of fishing
+a permit each year. For example, if someone fished for 30 days on a salmon drift
+gillnet permit in Southeast Alaska and 40 days on a statewide longline halibut
+permit, that individual's season length would be 70 days.
 
@@
 To jointly model revenue and revenue variability as a function of
-individual-level covariates, we use a variance function regression model, which
-separately models change in the mean and variance as
-a function of covariates [e.g. @davidian1987; @hall1989]. 
-Whereas traditional linear models or 
-hierarchical linear models generally assume
-homoscedasticity --- that the residual variance 
-is the same for all values of the predictor variables ---
-the variance-function regression model allows the variance to be heteroscedastic. 
-For application to economic data on individual fishing
-portfolios, this type of model allows certain strategies and levels of
-diversification to be more or less variable than others.
-
-We extended the basic variance-function regression to a hierarchical
-Bayesian variance-function regression model with Gaussian errors
-predicting $\log$ revenue, $\log(R_{i,t})$, for
-each fisher $i$ and time $t$:
+individual-level covariates, we extended the basic variance-function regression
+to a hierarchical Bayesian variance-function regression model with Gaussian
+errors predicting $\log$ revenue, $\log(R_{i,t})$, for each fisher $i$ and time
+$t$:
 
@@
 
-\noindent
-Each year $t$ of revenue for fisher $i$
-is assigned a strategy indexed by $j$
-(Supporting Information, Table \ref{tab:strategies})
-and expected $\log(R_{i,t})$ is modelled as:
+\noindent Each year $t$ of revenue for fisher $i$ is assigned a strategy indexed
+by $j$ (Supporting Information, Table \ref{tab:strategies}) and expected
+$\log(R_{i,t})$ is modelled as:
+
 <!-- Ole: I might swap out log(R)-hat for something like mu(or your favorit greek letter) just to avoid -->
@@
 
-\noindent
-where $S_{i,t}$ is the percent change in species diversity,
-$\log \left(\mathrm{div}_{i,t}/\mathrm{div}_{i,t-1}\right)$,
-$D_{i,t}$ is the percent change in days fished,
-$\log \left(\mathrm{days}_{i,t}/\mathrm{days}_{i,t-1}\right)$, and
-$I_{i,t}$ is an indicator variable that takes a value $0$
-if $\mathrm{div}_{i,t} \ge \mathrm{div}_{i,t-1}$
-(species diversity increasing; generalizing)
-and takes a value $1$ if
-$\mathrm{div}_{i,t} < \mathrm{div}_{i,t-1}$
-(species diversity decreasing; specializing). Estimating a separate
-coefficient for the effect of increasing diversity and decreasing
-diversity allows the costs and benefits of these changes to be asymmetric.
+\noindent where $S_{i,t}$ is the percent change in species diversity, $\log
+\left(\mathrm{div}_{i,t}/\mathrm{div}_{i,t-1}\right)$, $D_{i,t}$ is the percent
+change in days fished, $\log
+\left(\mathrm{days}_{i,t}/\mathrm{days}_{i,t-1}\right)$, and $I_{i,t}$ is an
+indicator variable that takes a value $0$ if $\mathrm{div}_{i,t} \ge
+\mathrm{div}_{i,t-1}$ (species diversity increasing; generalizing) and takes
+a value $1$ if $\mathrm{div}_{i,t} < \mathrm{div}_{i,t-1}$ (species diversity
+decreasing; specializing). Estimating a separate coefficient for the effect of
+increasing diversity and decreasing diversity allows the costs and benefits of
+these changes to be asymmetric.
 
-We use $\log(R_{i,t-1}$) as an offset, and thus
-our model describes the percent change in revenue,
-i.e., $\log(R_{i,t}) - \log(R_{i,t-1}) = \log(R_{i,t}/R_{i,t-1})$.
-We model change in revenue because preliminary analysis 
-revealed non-stationary trends in
-revenue (resulting from experience or other unmodeled factors). 
-This approach eliminates having to deal with the non-stationary
-time series and eliminates having to
-estimate thousands of mean-revenue intercepts 
-associated with individual fishers.
-We removed person-year combinations without a subsequent year of revenue.
+We use $\log(R_{i,t-1}$) as an offset, and thus our model describes the percent
+change in revenue, i.e., $\log(R_{i,t}) - \log(R_{i,t-1})
+= \log(R_{i,t}/R_{i,t-1})$. We model change in revenue because preliminary
+analysis revealed non-stationary trends in revenue (resulting from experience or
+other unmodeled factors). This approach eliminates having to deal with the
+non-stationary time series and eliminates having to estimate thousands of
+mean-revenue intercepts associated with individual fishers. We removed
+person-year combinations without a subsequent year of revenue.
 
 The $\beta$ terms represent estimated coefficients, some of which are allowed to
-vary as 'random effect' terms.
-The intercepts $\beta_{0,j,t}$ are allowed to vary by
-strategy-year combinations,
-are centered on zero, and are constrained by a normal distribution:
+vary as 'random effect' terms. The intercepts $\beta_{0,j,t}$ are allowed to
+vary by strategy-year combinations, are centered on zero, and are constrained by
+a normal distribution:
 
@@
 
-\noindent
-This, combined with the lack of a global intercept, constrains
-the mean change in log revenue across all fishers and years to
-be zero. I.e., we assume that the data are stationary
-after accounting for all covariates and including
-an intercept that varies with strategy-year combinations.
-Initial versions of the model
-confirmed that the global intercept along with stragey-level
-intercepts would be estimated at almost exactly zero.
+\noindent This, combined with the lack of a global intercept, constrains the
+mean change in log revenue across all fishers and years to be zero. I.e., we
+assume that the data are stationary after accounting for all covariates and
+including an intercept that varies with strategy-year combinations. Initial
+versions of the model confirmed that the global intercept along with
+stragey-level intercepts would be estimated at almost exactly zero.
 
-The slopes $\beta_{1,j}$ and $\beta_{1,j}$
-are then allowed to vary by strategy with estimated means:
+The slopes $\beta_{1,j}$ and $\beta_{1,j}$ are then allowed to vary by strategy
+with estimated means:
 
@@
 
-\noindent
-The $\beta_{1,j}$'s represent how much a X\% decrease in species
-diversity within a strategy translates to a Y\% change in revenue.
-The $\beta_{2,j}$'s represent how much an X\% increase in species
-diversity within a strategy translates to a Y\% change in revenue.
-These are after controlling for percent change in days fished, $\beta_3$,
-and the interaction between days fished and percent change in species
-diversity, $\beta_4$ and $\beta_5$.
+\noindent The $\beta_{1,j}$'s represent how much a X\% decrease in species
+diversity within a strategy translates to a Y\% change in revenue. The
+$\beta_{2,j}$'s represent how much an X\% increase in species diversity within
+a strategy translates to a Y\% change in revenue. These are after controlling
+for percent change in days fished, $\beta_3$, and the interaction between days
+fished and percent change in species diversity, $\beta_4$ and $\beta_5$.
 
-Whereas a typical hierarchical linear regression model
-would estimate a single residual error scale term, $\sigma$,
-we model the scale of the residual error  with
-another hierarchical model with a similar form:
+Whereas a typical hierarchical linear regression model would estimate a single
+residual error scale term, $\sigma$, we model the scale of the residual error
+with another hierarchical model with a similar form:
 
@@
 
-\noindent
-Again, $S_{i,t}$ and $I_{i,t}$ represent, respectively,
-percent change in species
-diversity and an indicator variable for increasing or
-decreasing species diversity.
-We exponentiate the equation to ensure all scale parameters, $\sigma_{i,t}$,
-are positive. The $\gamma$'s are estimated coefficients, three of which
-are allowed to vary by strategy, constrained by normal distributions:
+\noindent Again, $S_{i,t}$ and $I_{i,t}$ represent, respectively, percent change
+in species diversity and an indicator variable for increasing or decreasing
+species diversity. We exponentiate the equation to ensure all scale parameters,
+$\sigma_{i,t}$, are positive. The $\gamma$'s are estimated coefficients, three
+of which are allowed to vary by strategy, constrained by normal distributions:
 
@@
 
-\noindent
-The mean of the $\gamma_{0,j}$ coefficients is
-modelled with
-two strategy-level predictors.
-The symbols $M_j$ and $E_j$ represent the raw mean species diversity and
-mean combined fishing season length (effort) for strategy $j$.
-The coefficients $\eta_1$ and $\eta_2$ represent predictors
-in this strategy-level regression
-estimating effects across strategies.
+\noindent The mean of the $\gamma_{0,j}$ coefficients is modelled with two
+strategy-level predictors. The symbols $M_j$ and $E_j$ represent the raw mean
+species diversity and mean combined fishing season length (effort) for strategy
+$j$. The coefficients $\eta_1$ and $\eta_2$ represent predictors in this
+strategy-level regression estimating effects across strategies.
 
-We fit our models with Stan 2.11.1
-[@hoffman2014; @standevelopmentteam2015; @standevelopmentteam2015a; @carpenter2015]
-and R 3.3.1 [@rcoreteam2016].
-Stan implements the
-No-U-Turn Hamiltonian Markov chain Monte Carlo algorithm [@hoffman2014]
-to perform Bayesian statistical inference.
-We assigned weakly informative priors on all parameters:
-$N(0, 2^2)$ priors on all
-$\beta$, $\gamma$, and $\eta$ parameters,
-and half-$t(3, 0, 2)$ priors
-(i.e. degrees of freedom of 3 and scale of 2)
-on all $\sigma$ parameters to allow for heavier tails.
-We ran four chains and 2,000 iterations and discarded
-the first 1,000 iterations of each chain
-as warm up (i.e. 4,000 final samples).
-We checked for chain convergence visually with trace plots.
-We also ensured that $\hat{R}<1.05$ (the potential scale reduction factor)
-and that the effective sample size,
-as calculated by the rstan R package [@standevelopmentteam2015],
-was greater than 200 for all parameters [@gelman2014].
-We verified that our model returned sensible estimates by
-comparing our estimates to an ad-hoc two-stage
-model fit with the lme4 R package [@bates2015],
-and by fitting our model to simulated data
-and recovering the true values (Supporting Information).
-Code to reproduce our results (with a simulated data set)
-is available at <https://github.com/NCEAS/pfx-commercial>.
+We fit our models with Stan 2.11.1 [@hoffman2014; @standevelopmentteam2015;
+@standevelopmentteam2015a; @carpenter2015] and R 3.3.1 [@rcoreteam2016]. Stan
+implements the No-U-Turn Hamiltonian Markov chain Monte Carlo algorithm
+[@hoffman2014] to perform Bayesian statistical inference. We assigned weakly
+informative priors on all parameters: $N(0, 2^2)$ priors on all $\beta$,
+$\gamma$, and $\eta$ parameters, and half-$t(3, 0, 2)$ priors (i.e. degrees of
+freedom of 3 and scale of 2) on all $\sigma$ parameters to allow for heavier
+tails. We ran four chains and 2,000 iterations and discarded the first 1,000
+iterations of each chain as warm up (i.e. 4,000 final samples). We checked for
+chain convergence visually with trace plots. We also ensured that $\hat{R}<1.05$
+(the potential scale reduction factor) and that the effective sample size, as
+calculated by the rstan R package [@standevelopmentteam2015], was greater than
+200 for all parameters [@gelman2014]. We verified that our model returned
+sensible estimates by comparing our estimates to an ad-hoc two-stage model fit
+with the lme4 R package [@bates2015], and by fitting our model to simulated data
+and recovering the true values (Supporting Information). Code to reproduce our
+results (with a simulated data set) is available at
+<https://github.com/NCEAS/pfx-commercial>.
 
@@
 
-The estimated year-permit intercepts represent years when people
-with a particular permit set in a given year tended to do better or worse than
-the long-term average for that permit set (Fig. S\ref{fig:str-yrs}). Therefore,
-an individual's expected revenue and variability is after accounting for the
+The estimated year-permit intercepts represent years when people with
+a particular permit set in a given year tended to do better or worse than the
+long-term average for that permit set (Fig. S\ref{fig:str-yrs}). Therefore, an
+individual's expected revenue and variability is after accounting for the
 general trend in revenue for that permit set.
@@
 
-\bibliographystyle{apalike}
+\bibliographystyle{pnas2011}
 \bibliography{refs}
@@
 \includegraphics[width=0.7\textwidth]{../figs/fig1-jk.pdf}
-\caption{
-Trends in revenue and diversification over time.
-(A) Map of fishing regions in Alaska, USA.
-(B) Total revenue in billions of 2009 U.S. dollars from `r range_yrs[1]` to
-`r range_yrs[2]` for multiple (grey) and single (blue) permit holders.
-(C) Proportion of single permit holders for all fisheries (thick black line)
-and with permit holders omitted each year who fished any of the
-`r n_jacknifed` top species groups (thin lines).
-(D) Mean effective diversity of species fished (Simpson's diversity index)
-for all fisheries (thick black line)
-and with permit holders omitted each year who fished any of the
-`r n_jacknifed` top species groups (thin lines).
-}
-\label{fig:ts}
-\end{center}
-\end{figure}
+
+\caption{ Trends in revenue and diversification over time. (A) Map of fishing
+regions in Alaska, USA. (B) Total revenue in billions of 2009 U.S. dollars from
+`r range_yrs[1]` to `r range_yrs[2]` for multiple (grey) and single (blue)
+permit holders. (C) Proportion of single permit holders for all fisheries (thick
+black line) and with permit holders omitted each year who fished any of the `r 
+n_jacknifed` top species groups (thin lines). (D) Mean effective diversity of
+species fished (Simpson's diversity index) for all fisheries (thick black line)
+and with permit holders omitted each year who fished any of the `r n_jacknifed`
+top species groups (thin lines). } \label{fig:ts} \end{center} \end{figure}
 
@@
 \includegraphics[width=0.7\textwidth]{../figs/stan-across-strategy-variability2.pdf}
-\caption{Individuals who fish
-permits and permit combinations (permit strategies) with higher species diversity
-have lower expected revenue variability.
-On the y-axis, dots, thick lines, and thin lines represent posterior medians,
-50\%, and 95\% credible intervals for
-the `r nstr` most common permit strategies.
-Thick black line, and dark and light grey shaded regions
-indicate median, 50\%, and 95\% credible intervals
-of strategy-level regression built into the hierarchical model.
-Estimated variability represents expected standard deviation
-for a permit holder who does not change
-species diversity or days fished from year to year.
-}
+
+\caption{Individuals who fish permits and permit combinations (permit
+strategies) with higher species diversity have lower expected revenue
+variability. On the y-axis, dots, thick lines, and thin lines represent
+posterior medians, 50\%, and 95\% credible intervals for the `r nstr` most
+common permit strategies. Thick black line, and dark and light grey shaded
+regions indicate median, 50\%, and 95\% credible intervals of strategy-level
+regression built into the hierarchical model. Estimated variability represents
+expected standard deviation for a permit holder who does not change species
+diversity or days fished from year to year. }
+
 \label{fig:across}
@@
 \includegraphics[width=0.9\textwidth]{../figs/stan-gg-spoke2.pdf}
-\caption{Diversifying permits reduces risk and usually increases revenue.
-Two examples of expected revenue variability
-and median revenue for
-owners of (A) individual halibut and (B) gillnet herring roe permits
-(dark blue)
-compared with owners of the same permit plus one more (medium blue)
-or two more permits (light blue).
-On the y-axis, dots and lines represent posterior median and 95\%
+
+\caption{Diversifying permits reduces risk and usually increases revenue. Two
+examples of expected revenue variability and median revenue for owners of (A)
+individual halibut and (B) gillnet herring roe permits (dark blue) compared with
+owners of the same permit plus one more (medium blue) or two more permits (light
+blue). On the y-axis, dots and lines represent posterior median and 95\%
 credible intervals of expected standard deviation.}
+
 \label{fig:spokes}
@@
 \includegraphics[width=1.1\textwidth]{../figs/stan-offset-break-anti-spaghetti.pdf}
+
 \caption{Effects of (A) generalizing and (B) specializing from year to year for
-individual permit holders across the `r nstr` most common strategies.
-Changing species diversity from year to year
-is associated with higher variability
-for individuals within the strategies above the zero line.
-Generalizing (A) or specializing (B) is associated with
-greater expected revenue to the right of the zero line
-and lower expected revenue to the left of the zero line.
-These effects are estimated holding
-fishing season length constant from year to year.
-Dots and line segments represent medians and
-50\% credible intervals of posteriors.
-}
+individual permit holders across the `r nstr` most common strategies. Changing
+species diversity from year to year is associated with higher variability for
+individuals within the strategies above the zero line. Generalizing (A) or
+specializing (B) is associated with greater expected revenue to the right of the
+zero line and lower expected revenue to the left of the zero line. These effects
+are estimated holding fishing season length constant from year to year. Dots and
+line segments represent medians and 50\% credible intervals of posteriors. }
+
 \label{fig:within}
@@
 <!-- [@sanchirico2008] -->
+
+<!-- vim: textwidth=80 
+-->

### 2016-12-06 portfolio/text/ms.Rmd
Copyediting

@@
 systematically --- the variance-function regression model allows the variance to
-be heteroscedastic and vary as a function of covariates. For application to
-economic data on individual fishing portfolios, this model allows certain
-strategies and levels of diversification to be more or less variable than
-others. Our model includes the number of days fished to represent fishing
-effort,
+be heteroscedastic and vary as a function of covariates. Our specific model
+allows certain fishing strategies and levels of species diversification to
+generate more or less revenue and be more or less variable than others.
 
@@
 
-We obtained fisheries landings and revenue data for all permit holders in Alaska
-from `r range_yrs[1]` to `r range_yrs[2]` from the Alaskan Commercial Fisheries
-Entry Commission (CFEC). These data represent a total of nearly `r 
-round(ndat$initial_tot_rev/1e9)` billion 2009 USD and a combined `r 
+We obtained commercial fisheries landings and revenue data for all permit
+holders in Alaska from `r range_yrs[1]` to `r range_yrs[2]` from the Alaskan
+Commercial Fisheries Entry Commission (CFEC). These data represent a total of
+nearly `r round(ndat$initial_tot_rev/1e9)` billion 2009 USD and a combined `r 
 round(ndat$initial/1e6,1)` million rows of data. We adjusted for inflation by
-setting all revenues in 2009 USD [@bea2016]. We implemented a number of
+setting all revenues to 2009 USD [@bea2016]. We implemented a number of
 filtering steps to focus on individuals actively engaged in commercial fishing
 (Supporting Information). In particular, we removed permit holders whose median
-revenue was less than \$10,000 2009 USD.
+revenue was less than \$10,000 2009 USD to focus on individuals for whom
+commercial fishing formed a substantial part of their income.
 
 We combined similar permits into permit groupings and identified sets of permit
-groupings held by an individual in a specific year as 'strategies' (Supporting
+groupings held by an individual in a specific year as "strategies" (Supporting
 Information). For example, we combined permits that were the same except for
@@
 a vessel over 60 ft), and we combined a number of permits that were targeting
-a single species and differed only in region fished. Therefore, someone fishing
-herring roe in Kodiak and someone fishing herring roe in Alaska Peninsula were
-both considered to have a 'herring roe' permit group. For salmon permits, we
-grouped permits using the same gear type across regions for groups that shared
-similar species composition. After this process of combining permits, we were
-left with `r n_grouped_permits` unique fishing permit groups (Table
-S\ref{tab:strategies}), which we aggregated within each person-year combination
-to form permit strategies. For example, if someone fished halibut and sablefish
-in the same year, their strategy for that year would be "halibut-sablefish".
-Specific details on our strategy definitions are available in the Supporting
-Information.
+a single species and differed only in region fished. Therefore, for example,
+someone fishing herring roe in Kodiak and someone fishing herring roe in Alaska
+Peninsula were both considered to have a 'herring roe' permit group. For salmon
+permits, we grouped permits using the same gear type across regions for groups
+that shared similar species composition. After this process of combining
+permits, we were left with `r n_grouped_permits` unique fishing permit groups
+(Table S\ref{tab:strategies}), which we aggregated within each person-year
+combination to form permit strategies. For example, if someone fished halibut
+and sablefish in the same year, their strategy for that year would be
+"halibut-sablefish". Specific details on our strategy definitions are available
+in the Supporting Information.
 
@@
 The $\beta$ terms represent estimated coefficients, some of which are allowed to
-vary as 'random effect' terms. The intercepts $\beta_{0,j,t}$ are allowed to
+vary as "random effect" terms. The intercepts $\beta_{0,j,t}$ are allowed to
 vary by strategy-year combinations, are centered on zero, and are constrained by
@@
 
-\noindent This, combined with the lack of a global intercept, constrains the
-mean change in log revenue across all fishers and years to be zero. I.e., we
-assume that the data are stationary after accounting for all covariates and
-including an intercept that varies with strategy-year combinations. Initial
-versions of the model confirmed that the global intercept along with
-stragey-level intercepts would be estimated at almost exactly zero.
+\noindent This approach, combined with the lack of a global intercept,
+constrains the mean change in log revenue across all fishers and years to be
+zero. I.e., we assume that the data are stationary after accounting for all
+covariates and including an intercept that varies with strategy-year
+combinations. Initial versions of the model confirmed that the global intercept
+along with strategy-level intercepts would be estimated at almost exactly zero.
 
@@
 
-\noindent The $\beta_{1,j}$'s represent how much a X\% decrease in species
+\noindent The $\beta_{1,j}$'s represent how much a 1\% decrease in species
 diversity within a strategy translates to a Y\% change in revenue. The
-$\beta_{2,j}$'s represent how much an X\% increase in species diversity within
+$\beta_{2,j}$'s represent how much a 1\% increase in species diversity within
 a strategy translates to a Y\% change in revenue. These are after controlling

### 2016-12-06 portfolio/text/ms.Rmd
More copyediting

@@
 expected standard deviation for a permit holder who does not change species
-diversity or days fished from year to year. }
+diversity or days fished from year-to-year. }
 
@@
 
-\caption{Effects of (A) generalizing and (B) specializing from year to year for
-individual permit holders across the `r nstr` most common strategies. Changing
-species diversity from year to year is associated with higher variability for
-individuals within the strategies above the zero line. Generalizing (A) or
-specializing (B) is associated with greater expected revenue to the right of the
-zero line and lower expected revenue to the left of the zero line. These effects
-are estimated holding fishing season length constant from year to year. Dots and
-line segments represent medians and 50\% credible intervals of posteriors. }
+\caption{Effects of (A) generalizing and (B) specializing from year-to-year for
+individual permit holders \emph{within} the `r nstr` most common strategies.
+Changing species diversity from year-to-year is associated with higher
+variability for individuals within the strategies above the zero line.
+Generalizing (A) or specializing (B) is associated with greater expected revenue
+to the right of the zero line and lower expected revenue to the left of the zero
+line. These effects are estimated holding fishing season length constant from
+year-to-year. Dots and line segments represent medians and 50\% credible
+intervals of posteriors. }
 

### 2016-12-06 portfolio/text/anderson-etal-rev-pfx.tex
Rework introduction and start cleaning up paper
Create and add IFQ figure to supplement
Switch to a numbered citation system
Tweak knitting, title, and copyediting

@@
 \usepackage{commath} % for \abs{}
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{4pt}\setlength{\parskip}{0pt}}
 
@@
 
-\usepackage[round,sectionbib]{natbib}
-\bibpunct{(}{)}{;}{a}{}{;}
+% \usepackage[round,sectionbib]{natbib}
+% \bibpunct{(}{)}{;}{a}{}{;}
 
@@
 
-\title{Benefits of diversification and risks of changing strategies for individual fishers\\
+\title{
+  Benefits and risks of diversification\\ for individual fishers\\
+  Benefits of diversification and risks of 
+  changing strategies for individual fishers\\
 Benefits of diversification are unavailable to many fishermen\\
@@
 \newlabel{fig:str-yrs}{6}
+\newlabel{fig:ifq}{7}
 

### 2016-12-06 portfolio/text/ms.Rmd
weird sync issue

@@
-```{r set-knitr-options, cache=FALSE, echo=FALSE}
-library("knitr")
-opts_chunk$set(cache=TRUE, echo=FALSE, message=FALSE, autodep=TRUE,
-  results='hide')
-opts_chunk$set(warning=FALSE, message=FALSE, cache.path="../cache/",
-  cache.lazy=TRUE)
-```
-```{r get-dat, cache=FALSE}
-load("../data-generated/output-values.rda")
-h1 <- readRDS("../data-generated/h1_effect.rds")
-spoke <- readRDS("../data-generated/spoke_fraction.rds")
-g2_dens <- readRDS("../data-generated/g2_prob_dens_above_zero.rds")
-n_grouped_permits <- readRDS("../data-generated/n_grouped_permits.rds")
-```
+<!DOCTYPE html>
+<html lang="en" class=" is-copy-enabled emoji-size-boost is-u2f-enabled">
+  <head prefix="og: http://ogp.me/ns# fb: http://ogp.me/ns/fb# object: http://ogp.me/ns/object# article: http://ogp.me/ns/article# profile: http://ogp.me/ns/profile#">
+    <meta charset='utf-8'>
 
-# Abstract
+    <link crossorigin="anonymous" href="https://assets-cdn.github.com/assets/frameworks-c64b206fb6104fd10a7b4f7460f970885149c62649a7414df084683d7a0f7bf3.css" integrity="sha256-xksgb7YQT9EKe090YPlwiFFJxiZJp0FN8IRoPXoPe/M=" media="all" rel="stylesheet" />
+    <link crossorigin="anonymous" href="https://assets-cdn.github.com/assets/github-10cfb952863ec274f56e591f95dd8cd5e2c2516671d7989a327f7bf753c08909.css" integrity="sha256-EM+5UoY+wnT1blkfld2M1eLCUWZx15iaMn9791PAiQk=" media="all" rel="stylesheet" />
 
-A fundamental challenge for individuals who rely on the extraction of natural
+    <meta http-equiv="X-UA-Compatible" content="IE=edge">
+    <meta http-equiv="Content-Language" content="en">
+    <meta name="viewport" content="width=device-width">
+
+    <title>pfx-commercial/ms.Rmd at master · NCEAS/pfx-commercial</title>
+    <link rel="search" type="application/opensearchdescription+xml" href="/opensearch.xml" title="GitHub">
+    <link rel="fluid-icon" href="https://github.com/fluidicon.png" title="GitHub">
+    <link rel="apple-touch-icon" href="/apple-touch-icon.png">
+    <link rel="apple-touch-icon" sizes="57x57" href="/apple-touch-icon-57x57.png">
+    <link rel="apple-touch-icon" sizes="60x60" href="/apple-touch-icon-60x60.png">
+    <link rel="apple-touch-icon" sizes="72x72" href="/apple-touch-icon-72x72.png">
+    <link rel="apple-touch-icon" sizes="76x76" href="/apple-touch-icon-76x76.png">
+    <link rel="apple-touch-icon" sizes="114x114" href="/apple-touch-icon-114x114.png">
+    <link rel="apple-touch-icon" sizes="120x120" href="/apple-touch-icon-120x120.png">
+    <link rel="apple-touch-icon" sizes="144x144" href="/apple-touch-icon-144x144.png">
+    <link rel="apple-touch-icon" sizes="152x152" href="/apple-touch-icon-152x152.png">
+    <link rel="apple-touch-icon" sizes="180x180" href="/apple-touch-icon-180x180.png">
+    <meta property="fb:app_id" content="1401488693436528">
+
+      <meta content="https://avatars2.githubusercontent.com/u/2159348?v=3&amp;s=400" name="twitter:image:src" /><meta content="@github" name="twitter:site" /><meta content="summary" name="twitter:card" /><meta content="NCEAS/pfx-commercial" name="twitter:title" /><meta content="Contribute to pfx-commercial development by creating an account on GitHub." name="twitter:description" />
+      <meta content="https://avatars2.githubusercontent.com/u/2159348?v=3&amp;s=400" property="og:image" /><meta content="GitHub" property="og:site_name" /><meta content="object" property="og:type" /><meta content="NCEAS/pfx-commercial" property="og:title" /><meta content="https://github.com/NCEAS/pfx-commercial" property="og:url" /><meta content="Contribute to pfx-commercial development by creating an account on GitHub." property="og:description" />
+      <meta name="browser-stats-url" content="https://api.github.com/_private/browser/stats">
+    <meta name="browser-errors-url" content="https://api.github.com/_private/browser/errors">
+    <link rel="assets" href="https://assets-cdn.github.com/">
+    <link rel="web-socket" href="wss://live.github.com/_sockets/NTA0Njg4NDo0OTE1ZGE4NjRjZjExYmU3YzgwMDFmMGEwN2M4ZTIxYzplM2U0Yjk4MWFjMjg5N2U4MTMyMDM2YzMxYWE3NzBmNzg0MmY0OTgwZmJhYTQ1Zjc2NzBhYzg4ZTdiNjM2MmI5--9f3c14941fd0bc15eb2fce42405037f6cb2a11f5">
+    <meta name="pjax-timeout" content="1000">
+    <link rel="sudo-modal" href="/sessions/sudo_modal">
+    <meta name="request-id" content="1812FC12:15610:276B4F4D:58473282" data-pjax-transient>
+
+    <meta name="msapplication-TileImage" content="/windows-tile.png">
+    <meta name="msapplication-TileColor" content="#ffffff">
+    <meta name="selected-link" value="repo_source" data-pjax-transient>
+
+    <meta name="google-site-verification" content="KT5gs8h0wvaagLKAVWq8bbeNwnZZK1r1XQysX3xurLU">
+<meta name="google-site-verification" content="ZzhVyEFwb7w3e0-uOTltm8Jsck2F5StVihD0exw2fsA">
+    <meta name="google-analytics" content="UA-3769691-2">
+
+<meta content="collector.githubapp.com" name="octolytics-host" /><meta content="github" name="octolytics-app-id" /><meta content="1812FC12:15610:276B4F4D:58473282" name="octolytics-dimension-request_id" /><meta content="5046884" name="octolytics-actor-id" /><meta content="eric-ward" name="octolytics-actor-login" /><meta content="d7c8c13b6921d5dbb2d08511e76e4d3922fb6f78877f69812f4e2cc5b901f5c9" name="octolytics-actor-hash" />
+<meta content="/&lt;user-name&gt;/&lt;repo-name&gt;/blob/show" data-pjax-transient="true" name="analytics-location" />
+
+  <meta class="js-ga-set" name="dimension1" content="Logged In">
+
+        <meta name="hostname" content="github.com">
+    <meta name="user-login" content="eric-ward">
+
+        <meta name="expected-hostname" content="github.com">
+      <meta name="js-proxy-site-detection-payload" content="OTZkMDAzYjQ5M2JjYTZlYzA5MmU0MTdkODA1ZDk5MWU1MWY3Y2RjNzhlMzdlN2FlNWZkZjUxZjc3ZDRhYmQ5Nnx7InJlbW90ZV9hZGRyZXNzIjoiMjQuMTguMjUyLjE4IiwicmVxdWVzdF9pZCI6IjE4MTJGQzEyOjE1NjEwOjI3NkI0RjREOjU4NDczMjgyIiwidGltZXN0YW1wIjoxNDgxMDYwOTk5LCJob3N0IjoiZ2l0aHViLmNvbSJ9">
+
+      <link rel="mask-icon" href="https://assets-cdn.github.com/pinned-octocat.svg" color="#000000">
+      <link rel="icon" type="image/x-icon" href="https://assets-cdn.github.com/favicon.ico">
+
+    <meta name="html-safe-nonce" content="306849f02c6158107b6119c3ea451ae14657b020">
+    <meta content="f2086a1d058512824537f14ace9f906f412f8160" name="form-nonce" />
+
+    <meta http-equiv="x-pjax-version" content="77135ce72af7054f190f2da8838d6ffb">
+
+  <meta name="description" content="Contribute to pfx-commercial development by creating an account on GitHub.">
+  <meta name="go-import" content="github.com/NCEAS/pfx-commercial git https://github.com/NCEAS/pfx-commercial.git">
+
+  <meta content="2159348" name="octolytics-dimension-user_id" /><meta content="NCEAS" name="octolytics-dimension-user_login" /><meta content="45410050" name="octolytics-dimension-repository_id" /><meta content="NCEAS/pfx-commercial" name="octolytics-dimension-repository_nwo" /><meta content="true" name="octolytics-dimension-repository_public" /><meta content="false" name="octolytics-dimension-repository_is_fork" /><meta content="45410050" name="octolytics-dimension-repository_network_root_id" /><meta content="NCEAS/pfx-commercial" name="octolytics-dimension-repository_network_root_nwo" />
+  <link href="https://github.com/NCEAS/pfx-commercial/commits/master.atom" rel="alternate" title="Recent Commits to pfx-commercial:master" type="application/atom+xml">
+
+      <link rel="canonical" href="https://github.com/NCEAS/pfx-commercial/blob/master/portfolio/text/ms.Rmd" data-pjax-transient>
+  </head>
+
+  <body class="logged-in  env-production macintosh vis-public page-blob">
+    <div id="js-pjax-loader-bar" class="pjax-loader-bar"><div class="progress"></div></div>
+    <a href="#start-of-content" tabindex="1" class="accessibility-aid js-skip-to-content">Skip to content</a>
+
+        <div class="header header-logged-in true" role="banner">
+  <div class="container clearfix">
+
+    <a class="header-logo-invertocat" href="https://github.com/" data-hotkey="g d" aria-label="Homepage" data-ga-click="Header, go to dashboard, icon:logo">
+  <svg aria-hidden="true" class="octicon octicon-mark-github" height="28" version="1.1" viewBox="0 0 16 16" width="28"><path fill-rule="evenodd" d="M8 0C3.58 0 0 3.58 0 8c0 3.54 2.29 6.53 5.47 7.59.4.07.55-.17.55-.38 0-.19-.01-.82-.01-1.49-2.01.37-2.53-.49-2.69-.94-.09-.23-.48-.94-.82-1.13-.28-.15-.68-.52-.01-.53.63-.01 1.08.58 1.23.82.72 1.21 1.87.87 2.33.66.07-.52.28-.87.51-1.07-1.78-.2-3.64-.89-3.64-3.95 0-.87.31-1.59.82-2.15-.08-.2-.36-1.02.08-2.12 0 0 .67-.21 2.2.82.64-.18 1.32-.27 2-.27.68 0 1.36.09 2 .27 1.53-1.04 2.2-.82 2.2-.82.44 1.1.16 1.92.08 2.12.51.56.82 1.27.82 2.15 0 3.07-1.87 3.75-3.65 3.95.29.25.54.73.54 1.48 0 1.07-.01 1.93-.01 2.2 0 .21.15.46.55.38A8.013 8.013 0 0 0 16 8c0-4.42-3.58-8-8-8z"/></svg>
+</a>
+
+        <div class="header-search scoped-search site-scoped-search js-site-search" role="search">
+  <!-- '"` --><!-- </textarea></xmp> --></option></form><form accept-charset="UTF-8" action="/NCEAS/pfx-commercial/search" class="js-site-search-form" data-scoped-search-url="/NCEAS/pfx-commercial/search" data-unscoped-search-url="/search" method="get"><div style="margin:0;padding:0;display:inline"><input name="utf8" type="hidden" value="&#x2713;" /></div>
+    <label class="form-control header-search-wrapper js-chromeless-input-container">
+      <div class="header-search-scope">This repository</div>
+      <input type="text"
+        class="form-control header-search-input js-site-search-focus js-site-search-field is-clearable"
+        data-hotkey="s"
+        name="q"
+        placeholder="Search"
+        aria-label="Search this repository"
+        data-unscoped-placeholder="Search GitHub"
+        data-scoped-placeholder="Search"
+        autocapitalize="off">
+    </label>
+</form></div>
+
+      <ul class="header-nav float-left" role="navigation">
+        <li class="header-nav-item">
+          <a href="/pulls" aria-label="Pull requests you created" class="js-selected-navigation-item header-nav-link" data-ga-click="Header, click, Nav menu - item:pulls context:user" data-hotkey="g p" data-selected-links="/pulls /pulls/assigned /pulls/mentioned /pulls">
+            Pull requests
+</a>        </li>
+        <li class="header-nav-item">
+          <a href="/issues" aria-label="Issues you created" class="js-selected-navigation-item header-nav-link" data-ga-click="Header, click, Nav menu - item:issues context:user" data-hotkey="g i" data-selected-links="/issues /issues/assigned /issues/mentioned /issues">
+            Issues
+</a>        </li>
+          <li class="header-nav-item">
+            <a class="header-nav-link" href="https://gist.github.com/" data-ga-click="Header, go to gist, text:gist">Gist</a>
+          </li>
+      </ul>
+
+<ul class="header-nav user-nav float-right" id="user-links">
+  <li class="header-nav-item">
+
+    <a href="/notifications" aria-label="You have unread notifications" class="header-nav-link notification-indicator tooltipped tooltipped-s js-socket-channel js-notification-indicator" data-channel="tenant:1:notification-changed:5046884" data-ga-click="Header, go to notifications, icon:unread" data-hotkey="g n">
+        <span class="mail-status unread"></span>
+        <svg aria-hidden="true" class="octicon octicon-bell" height="16" version="1.1" viewBox="0 0 14 16" width="14"><path fill-rule="evenodd" d="M14 12v1H0v-1l.73-.58c.77-.77.81-2.55 1.19-4.42C2.69 3.23 6 2 6 2c0-.55.45-1 1-1s1 .45 1 1c0 0 3.39 1.23 4.16 5 .38 1.88.42 3.66 1.19 4.42l.66.58H14zm-7 4c1.11 0 2-.89 2-2H5c0 1.11.89 2 2 2z"/></svg>
+</a>
+  </li>
+
+  <li class="header-nav-item dropdown js-menu-container">
+    <a class="header-nav-link tooltipped tooltipped-s js-menu-target" href="/new"
+       aria-label="Create new…"
+       data-ga-click="Header, create new, icon:add">
+      <svg aria-hidden="true" class="octicon octicon-plus float-left" height="16" version="1.1" viewBox="0 0 12 16" width="12"><path fill-rule="evenodd" d="M12 9H7v5H5V9H0V7h5V2h2v5h5z"/></svg>
+      <span class="dropdown-caret"></span>
+    </a>
+
+    <div class="dropdown-menu-content js-menu-content">
+      <ul class="dropdown-menu dropdown-menu-sw">
+
+<a class="dropdown-item" href="/new" data-ga-click="Header, create new repository">
+  New repository
+</a>
+
+  <a class="dropdown-item" href="/new/import" data-ga-click="Header, import a repository">
+    Import repository
+  </a>
+
+<a class="dropdown-item" href="https://gist.github.com/" data-ga-click="Header, create new gist">
+  New gist
+</a>
+
+  <a class="dropdown-item" href="/organizations/new" data-ga-click="Header, create new organization">
+    New organization
+  </a>
+
+  <div class="dropdown-divider"></div>
+  <div class="dropdown-header">
+    <span title="NCEAS/pfx-commercial">This repository</span>
+  </div>
+    <a class="dropdown-item" href="/NCEAS/pfx-commercial/issues/new" data-ga-click="Header, create new issue">
+      New issue
+    </a>
+
+      </ul>
+    </div>
+  </li>
+
+  <li class="header-nav-item dropdown js-menu-container">
+    <a class="header-nav-link name tooltipped tooltipped-sw js-menu-target" href="/eric-ward"
+       aria-label="View profile and more"
+       data-ga-click="Header, show menu, icon:avatar">
+      <img alt="@eric-ward" class="avatar" height="20" src="https://avatars2.githubusercontent.com/u/5046884?v=3&amp;s=40" width="20" />
+      <span class="dropdown-caret"></span>
+    </a>
+
+    <div class="dropdown-menu-content js-menu-content">
+      <div class="dropdown-menu dropdown-menu-sw">
+        <div class="dropdown-header header-nav-current-user css-truncate">
+          Signed in as <strong class="css-truncate-target">eric-ward</strong>
+        </div>
+
+        <div class="dropdown-divider"></div>
+
+        <a class="dropdown-item" href="/eric-ward" data-ga-click="Header, go to profile, text:your profile">
+          Your profile
+        </a>
+        <a class="dropdown-item" href="/eric-ward?tab=stars" data-ga-click="Header, go to starred repos, text:your stars">
+          Your stars
+        </a>
+        <a class="dropdown-item" href="/explore" data-ga-click="Header, go to explore, text:explore">
+          Explore
+        </a>
+          <a class="dropdown-item" href="/integrations" data-ga-click="Header, go to integrations, text:integrations">
+            Integrations
+          </a>
+        <a class="dropdown-item" href="https://help.github.com" data-ga-click="Header, go to help, text:help">
+          Help
+        </a>
+
+        <div class="dropdown-divider"></div>
+
+        <a class="dropdown-item" href="/settings/profile" data-ga-click="Header, go to settings, icon:settings">
+          Settings
+        </a>
+
+        <!-- '"` --><!-- </textarea></xmp> --></option></form><form accept-charset="UTF-8" action="/logout" class="logout-form" data-form-nonce="f2086a1d058512824537f14ace9f906f412f8160" method="post"><div style="margin:0;padding:0;display:inline"><input name="utf8" type="hidden" value="&#x2713;" /><input name="authenticity_token" type="hidden" value="PnbQhWTo21/c91t6711esJeC3XADkQpt6F5tc07PrUW8XnSQgyd36jUpfGEHM3UbQRUVuX3tNIMOckYBS+V6lg==" /></div>
+          <button type="submit" class="dropdown-item dropdown-signout" data-ga-click="Header, sign out, icon:logout">
+            Sign out
+          </button>
+</form>      </div>
+    </div>
+  </li>
+</ul>
+
+  </div>
+</div>
+
+    <div id="start-of-content" class="accessibility-aid"></div>
+
+      <div id="js-flash-container">
+</div>
+
+    <div role="main">
+        <div itemscope itemtype="http://schema.org/SoftwareSourceCode">
+    <div id="js-repo-pjax-container" data-pjax-container>
+
+<div class="pagehead repohead instapaper_ignore readability-menu experiment-repo-nav">
+  <div class="container repohead-details-container">
+
+<ul class="pagehead-actions">
+
+  <li>
+        <!-- '"` --><!-- </textarea></xmp> --></option></form><form accept-charset="UTF-8" action="/notifications/subscribe" class="js-social-container" data-autosubmit="true" data-form-nonce="f2086a1d058512824537f14ace9f906f412f8160" data-remote="true" method="post"><div style="margin:0;padding:0;display:inline"><input name="utf8" type="hidden" value="&#x2713;" /><input name="authenticity_token" type="hidden" value="hFCZaDumpg36s+UNPQAtiT2CABWrtFfxVyHcBMJBBc0GeD193GkKuBNtwhbVbgYi6xXI3NXIaR+xDfd2x2vSHg==" /></div>      <input class="form-control" id="repository_id" name="repository_id" type="hidden" value="45410050" />
+
+        <div class="select-menu js-menu-container js-select-menu">
+          <a href="/NCEAS/pfx-commercial/subscription"
+            class="btn btn-sm btn-with-count select-menu-button js-menu-target" role="button" tabindex="0" aria-haspopup="true"
+            data-ga-click="Repository, click Watch settings, action:blob#show">
+            <span class="js-select-button">
+              <svg aria-hidden="true" class="octicon octicon-eye" height="16" version="1.1" viewBox="0 0 16 16" width="16"><path fill-rule="evenodd" d="M8.06 2C3 2 0 8 0 8s3 6 8.06 6C13 14 16 8 16 8s-3-6-7.94-6zM8 12c-2.2 0-4-1.78-4-4 0-2.2 1.8-4 4-4 2.22 0 4 1.8 4 4 0 2.22-1.78 4-4 4zm2-4c0 1.11-.89 2-2 2-1.11 0-2-.89-2-2 0-1.11.89-2 2-2 1.11 0 2 .89 2 2z"/></svg>
+              Unwatch
+            </span>
+          </a>
+          <a class="social-count js-social-count"
+            href="/NCEAS/pfx-commercial/watchers"
+            aria-label="18 users are watching this repository">
+            18
+          </a>
+
+        <div class="select-menu-modal-holder">
+          <div class="select-menu-modal subscription-menu-modal js-menu-content" aria-hidden="true">
+            <div class="select-menu-header js-navigation-enable" tabindex="-1">
+              <svg aria-label="Close" class="octicon octicon-x js-menu-close" height="16" role="img" version="1.1" viewBox="0 0 12 16" width="12"><path fill-rule="evenodd" d="M7.48 8l3.75 3.75-1.48 1.48L6 9.48l-3.75 3.75-1.48-1.48L4.52 8 .77 4.25l1.48-1.48L6 6.52l3.75-3.75 1.48 1.48z"/></svg>
+              <span class="select-menu-title">Notifications</span>
+            </div>
+
+              <div class="select-menu-list js-navigation-container" role="menu">
+
+                <div class="select-menu-item js-navigation-item " role="menuitem" tabindex="0">
+                  <svg aria-hidden="true" class="octicon octicon-check select-menu-item-icon" height="16" version="1.1" viewBox="0 0 12 16" width="12"><path fill-rule="evenodd" d="M12 5l-8 8-4-4 1.5-1.5L4 10l6.5-6.5z"/></svg>
+                  <div class="select-menu-item-text">
+                    <input id="do_included" name="do" type="radio" value="included" />
+                    <span class="select-menu-item-heading">Not watching</span>
+                    <span class="description">Be notified when participating or @mentioned.</span>
+                    <span class="js-select-button-text hidden-select-button-text">
+                      <svg aria-hidden="true" class="octicon octicon-eye" height="16" version="1.1" viewBox="0 0 16 16" width="16"><path fill-rule="evenodd" d="M8.06 2C3 2 0 8 0 8s3 6 8.06 6C13 14 16 8 16 8s-3-6-7.94-6zM8 12c-2.2 0-4-1.78-4-4 0-2.2 1.8-4 4-4 2.22 0 4 1.8 4 4 0 2.22-1.78 4-4 4zm2-4c0 1.11-.89 2-2 2-1.11 0-2-.89-2-2 0-1.11.89-2 2-2 1.11 0 2 .89 2 2z"/></svg>
+                      Watch
+                    </span>
+                  </div>
+                </div>
+
+                <div class="select-menu-item js-navigation-item selected" role="menuitem" tabindex="0">
+                  <svg aria-hidden="true" class="octicon octicon-check select-menu-item-icon" height="16" version="1.1" viewBox="0 0 12 16" width="12"><path fill-rule="evenodd" d="M12 5l-8 8-4-4 1.5-1.5L4 10l6.5-6.5z"/></svg>
+                  <div class="select-menu-item-text">
+                    <input checked="checked" id="do_subscribed" name="do" type="radio" value="subscribed" />
+                    <span class="select-menu-item-heading">Watching</span>
+                    <span class="description">Be notified of all conversations.</span>
+                    <span class="js-select-button-text hidden-select-button-text">
+                      <svg aria-hidden="true" class="octicon octicon-eye" height="16" version="1.1" viewBox="0 0 16 16" width="16"><path fill-rule="evenodd" d="M8.06 2C3 2 0 8 0 8s3 6 8.06 6C13 14 16 8 16 8s-3-6-7.94-6zM8 12c-2.2 0-4-1.78-4-4 0-2.2 1.8-4 4-4 2.22 0 4 1.8 4 4 0 2.22-1.78 4-4 4zm2-4c0 1.11-.89 2-2 2-1.11 0-2-.89-2-2 0-1.11.89-2 2-2 1.11 0 2 .89 2 2z"/></svg>
+                      Unwatch
+                    </span>
+                  </div>
+                </div>
+
+                <div class="select-menu-item js-navigation-item " role="menuitem" tabindex="0">
+                  <svg aria-hidden="true" class="octicon octicon-check select-menu-item-icon" height="16" version="1.1" viewBox="0 0 12 16" width="12"><path fill-rule="evenodd" d="M12 5l-8 8-4-4 1.5-1.5L4 10l6.5-6.5z"/></svg>
+                  <div class="select-menu-item-text">
+                    <input id="do_ignore" name="do" type="radio" value="ignore" />
+                    <span class="select-menu-item-heading">Ignoring</span>
+                    <span class="description">Never be notified.</span>
+                    <span class="js-select-button-text hidden-select-button-text">
+                      <svg aria-hidden="true" class="octicon octicon-mute" height="16" version="1.1" viewBox="0 0 16 16" width="16"><path fill-rule="evenodd" d="M8 2.81v10.38c0 .67-.81 1-1.28.53L3 10H1c-.55 0-1-.45-1-1V7c0-.55.45-1 1-1h2l3.72-3.72C7.19 1.81 8 2.14 8 2.81zm7.53 3.22l-1.06-1.06-1.97 1.97-1.97-1.97-1.06 1.06L11.44 8 9.47 9.97l1.06 1.06 1.97-1.97 1.97 1.97 1.06-1.06L13.56 8l1.97-1.97z"/></svg>
+                      Stop ignoring
+                    </span>
+                  </div>
+                </div>
+
+              </div>
+
+            </div>
+          </div>
+        </div>
+</form>
+  </li>
+
+  <li>
+
+  <div class="js-toggler-container js-social-container starring-container ">
+
+    <!-- '"` --><!-- </textarea></xmp> --></option></form><form accept-charset="UTF-8" action="/NCEAS/pfx-commercial/unstar" class="starred" data-form-nonce="f2086a1d058512824537f14ace9f906f412f8160" data-remote="true" method="post"><div style="margin:0;padding:0;display:inline"><input name="utf8" type="hidden" value="&#x2713;" /><input name="authenticity_token" type="hidden" value="eZCbUBwMgs0G16UNQzNUF5vddtZCkLQ8m5cvsKhv47f7uD9F+8MueO8JgharXX+8TUq+HzzsitJ9uwTCrUU0ZA==" /></div>
+      <button
+        type="submit"
+        class="btn btn-sm btn-with-count js-toggler-target"
+        aria-label="Unstar this repository" title="Unstar NCEAS/pfx-commercial"
+        data-ga-click="Repository, click unstar button, action:blob#show; text:Unstar">
+        <svg aria-hidden="true" class="octicon octicon-star" height="16" version="1.1" viewBox="0 0 14 16" width="14"><path fill-rule="evenodd" d="M14 6l-4.9-.64L7 1 4.9 5.36 0 6l3.6 3.26L2.67 14 7 11.67 11.33 14l-.93-4.74z"/></svg>
+        Unstar
+      </button>
+        <a class="social-count js-social-count" href="/NCEAS/pfx-commercial/stargazers"
+           aria-label="1 user starred this repository">
+          1
+        </a>
+</form>
+    <!-- '"` --><!-- </textarea></xmp> --></option></form><form accept-charset="UTF-8" action="/NCEAS/pfx-commercial/star" class="unstarred" data-form-nonce="f2086a1d058512824537f14ace9f906f412f8160" data-remote="true" method="post"><div style="margin:0;padding:0;display:inline"><input name="utf8" type="hidden" value="&#x2713;" /><input name="authenticity_token" type="hidden" value="JzXPzK6G/zH59e4dmw67mMnB7McXzaR4xJPQ3Uj4OlalHWvZSUlThBAryQZzYJAzH1YkDmmxmpYiv/uvTdLthQ==" /></div>
+      <button
+        type="submit"
+        class="btn btn-sm btn-with-count js-toggler-target"
+        aria-label="Star this repository" title="Star NCEAS/pfx-commercial"
+        data-ga-click="Repository, click star button, action:blob#show; text:Star">
+        <svg aria-hidden="true" class="octicon octicon-star" height="16" version="1.1" viewBox="0 0 14 16" width="14"><path fill-rule="evenodd" d="M14 6l-4.9-.64L7 1 4.9 5.36 0 6l3.6 3.26L2.67 14 7 11.67 11.33 14l-.93-4.74z"/></svg>
+        Star
+      </button>
+        <a class="social-count js-social-count" href="/NCEAS/pfx-commercial/stargazers"
+           aria-label="1 user starred this repository">
+          1
+        </a>
+</form>  </div>
+
+  </li>
+
+  <li>
+          <a href="#fork-destination-box" class="btn btn-sm btn-with-count"
+              title="Fork your own copy of NCEAS/pfx-commercial to your account"
+              aria-label="Fork your own copy of NCEAS/pfx-commercial to your account"
+              rel="facebox"
+              data-ga-click="Repository, show fork modal, action:blob#show; text:Fork">
+              <svg aria-hidden="true" class="octicon octicon-repo-forked" height="16" version="1.1" viewBox="0 0 10 16" width="10"><path fill-rule="evenodd" d="M8 1a1.993 1.993 0 0 0-1 3.72V6L5 8 3 6V4.72A1.993 1.993 0 0 0 2 1a1.993 1.993 0 0 0-1 3.72V6.5l3 3v1.78A1.993 1.993 0 0 0 5 15a1.993 1.993 0 0 0 1-3.72V9.5l3-3V4.72A1.993 1.993 0 0 0 8 1zM2 4.2C1.34 4.2.8 3.65.8 3c0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2 0 .65-.55 1.2-1.2 1.2zm3 10c-.66 0-1.2-.55-1.2-1.2 0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2 0 .65-.55 1.2-1.2 1.2zm3-10c-.66 0-1.2-.55-1.2-1.2 0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2 0 .65-.55 1.2-1.2 1.2z"/></svg>
+            Fork
+          </a>
+
+          <div id="fork-destination-box" style="display: none;">
+            <h2 class="facebox-header" data-facebox-id="facebox-header">Where should we fork this repository?</h2>
+            <include-fragment src=""
+                class="js-fork-select-fragment fork-select-fragment"
+                data-url="/NCEAS/pfx-commercial/fork?fragment=1">
+              <img alt="Loading" height="64" src="https://assets-cdn.github.com/images/spinners/octocat-spinner-128.gif" width="64" />
+            </include-fragment>
+          </div>
+
+    <a href="/NCEAS/pfx-commercial/network" class="social-count"
+       aria-label="1 user forked this repository">
+      1
+    </a>
+  </li>
+</ul>
+
+    <h1 class="public ">
+  <svg aria-hidden="true" class="octicon octicon-repo" height="16" version="1.1" viewBox="0 0 12 16" width="12"><path fill-rule="evenodd" d="M4 9H3V8h1v1zm0-3H3v1h1V6zm0-2H3v1h1V4zm0-2H3v1h1V2zm8-1v12c0 .55-.45 1-1 1H6v2l-1.5-1.5L3 16v-2H1c-.55 0-1-.45-1-1V1c0-.55.45-1 1-1h10c.55 0 1 .45 1 1zm-1 10H1v2h2v-1h3v1h5v-2zm0-10H2v9h9V1z"/></svg>
+  <span class="author" itemprop="author"><a href="/NCEAS" class="url fn" rel="author">NCEAS</a></span><!--
+--><span class="path-divider">/</span><!--
+--><strong itemprop="name"><a href="/NCEAS/pfx-commercial" data-pjax="#js-repo-pjax-container">pfx-commercial</a></strong>
+
+</h1>
+
+  </div>
+  <div class="container">
+
+<nav class="reponav js-repo-nav js-sidenav-container-pjax"
+     itemscope
+     itemtype="http://schema.org/BreadcrumbList"
+     role="navigation"
+     data-pjax="#js-repo-pjax-container">
+
+  <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
+    <a href="/NCEAS/pfx-commercial" class="js-selected-navigation-item selected reponav-item" data-hotkey="g c" data-selected-links="repo_source repo_downloads repo_commits repo_releases repo_tags repo_branches /NCEAS/pfx-commercial" itemprop="url">
+      <svg aria-hidden="true" class="octicon octicon-code" height="16" version="1.1" viewBox="0 0 14 16" width="14"><path fill-rule="evenodd" d="M9.5 3L8 4.5 11.5 8 8 11.5 9.5 13 14 8 9.5 3zm-5 0L0 8l4.5 5L6 11.5 2.5 8 6 4.5 4.5 3z"/></svg>
+      <span itemprop="name">Code</span>
+      <meta itemprop="position" content="1">
+</a>  </span>
+
+    <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
+      <a href="/NCEAS/pfx-commercial/issues" class="js-selected-navigation-item reponav-item" data-hotkey="g i" data-selected-links="repo_issues repo_labels repo_milestones /NCEAS/pfx-commercial/issues" itemprop="url">
+        <svg aria-hidden="true" class="octicon octicon-issue-opened" height="16" version="1.1" viewBox="0 0 14 16" width="14"><path fill-rule="evenodd" d="M7 2.3c3.14 0 5.7 2.56 5.7 5.7s-2.56 5.7-5.7 5.7A5.71 5.71 0 0 1 1.3 8c0-3.14 2.56-5.7 5.7-5.7zM7 1C3.14 1 0 4.14 0 8s3.14 7 7 7 7-3.14 7-7-3.14-7-7-7zm1 3H6v5h2V4zm0 6H6v2h2v-2z"/></svg>
+        <span itemprop="name">Issues</span>
+        <span class="counter">0</span>
+        <meta itemprop="position" content="2">
+</a>    </span>
+
+  <span itemscope itemtype="http://schema.org/ListItem" itemprop="itemListElement">
+    <a href="/NCEAS/pfx-commercial/pulls" class="js-selected-navigation-item reponav-item" data-hotkey="g p" data-selected-links="repo_pulls /NCEAS/pfx-commercial/pulls" itemprop="url">
+      <svg aria-hidden="true" class="octicon octicon-git-pull-request" height="16" version="1.1" viewBox="0 0 12 16" width="12"><path fill-rule="evenodd" d="M11 11.28V5c-.03-.78-.34-1.47-.94-2.06C9.46 2.35 8.78 2.03 8 2H7V0L4 3l3 3V4h1c.27.02.48.11.69.31.21.2.3.42.31.69v6.28A1.993 1.993 0 0 0 10 15a1.993 1.993 0 0 0 1-3.72zm-1 2.92c-.66 0-1.2-.55-1.2-1.2 0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2 0 .65-.55 1.2-1.2 1.2zM4 3c0-1.11-.89-2-2-2a1.993 1.993 0 0 0-1 3.72v6.56A1.993 1.993 0 0 0 2 15a1.993 1.993 0 0 0 1-3.72V4.72c.59-.34 1-.98 1-1.72zm-.8 10c0 .66-.55 1.2-1.2 1.2-.65 0-1.2-.55-1.2-1.2 0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2zM2 4.2C1.34 4.2.8 3.65.8 3c0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2 0 .65-.55 1.2-1.2 1.2z"/></svg>
+      <span itemprop="name">Pull requests</span>
+      <span class="counter">0</span>
+      <meta itemprop="position" content="3">
+</a>  </span>
+
+  <a href="/NCEAS/pfx-commercial/projects" class="js-selected-navigation-item reponav-item" data-selected-links="repo_projects new_repo_project repo_project /NCEAS/pfx-commercial/projects">
+    <svg aria-hidden="true" class="octicon octicon-project" height="16" version="1.1" viewBox="0 0 15 16" width="15"><path fill-rule="evenodd" d="M10 12h3V2h-3v10zm-4-2h3V2H6v8zm-4 4h3V2H2v12zm-1 1h13V1H1v14zM14 0H1a1 1 0 0 0-1 1v14a1 1 0 0 0 1 1h13a1 1 0 0 0 1-1V1a1 1 0 0 0-1-1z"/></svg>
+    Projects
+    <span class="counter">0</span>
+</a>
+    <a href="/NCEAS/pfx-commercial/wiki" class="js-selected-navigation-item reponav-item" data-hotkey="g w" data-selected-links="repo_wiki /NCEAS/pfx-commercial/wiki">
+      <svg aria-hidden="true" class="octicon octicon-book" height="16" version="1.1" viewBox="0 0 16 16" width="16"><path fill-rule="evenodd" d="M3 5h4v1H3V5zm0 3h4V7H3v1zm0 2h4V9H3v1zm11-5h-4v1h4V5zm0 2h-4v1h4V7zm0 2h-4v1h4V9zm2-6v9c0 .55-.45 1-1 1H9.5l-1 1-1-1H2c-.55 0-1-.45-1-1V3c0-.55.45-1 1-1h5.5l1 1 1-1H15c.55 0 1 .45 1 1zm-8 .5L7.5 3H2v9h6V3.5zm7-.5H9.5l-.5.5V12h6V3z"/></svg>
+      Wiki
+</a>
+
+  <a href="/NCEAS/pfx-commercial/pulse" class="js-selected-navigation-item reponav-item" data-selected-links="pulse /NCEAS/pfx-commercial/pulse">
+    <svg aria-hidden="true" class="octicon octicon-pulse" height="16" version="1.1" viewBox="0 0 14 16" width="14"><path fill-rule="evenodd" d="M11.5 8L8.8 5.4 6.6 8.5 5.5 1.6 2.38 8H0v2h3.6l.9-1.8.9 5.4L9 8.5l1.6 1.5H14V8z"/></svg>
+    Pulse
+</a>
+  <a href="/NCEAS/pfx-commercial/graphs" class="js-selected-navigation-item reponav-item" data-selected-links="repo_graphs repo_contributors /NCEAS/pfx-commercial/graphs">
+    <svg aria-hidden="true" class="octicon octicon-graph" height="16" version="1.1" viewBox="0 0 16 16" width="16"><path fill-rule="evenodd" d="M16 14v1H0V0h1v14h15zM5 13H3V8h2v5zm4 0H7V3h2v10zm4 0h-2V6h2v7z"/></svg>
+    Graphs
+</a>
+
+</nav>
+
+  </div>
+</div>
+
+<div class="container new-discussion-timeline experiment-repo-nav">
+  <div class="repository-content">
+
+<a href="/NCEAS/pfx-commercial/blob/292702045bcfdd7fae64dd9db5f50d64fd75b852/portfolio/text/ms.Rmd" class="d-none js-permalink-shortcut" data-hotkey="y">Permalink</a>
+
+<!-- blob contrib key: blob_contributors:v21:10db3054131f537f888c1dbda8c49f7a -->
+
+<div class="file-navigation js-zeroclipboard-container">
+
+<div class="select-menu branch-select-menu js-menu-container js-select-menu float-left">
+  <button class="btn btn-sm select-menu-button js-menu-target css-truncate" data-hotkey="w"
+
+    type="button" aria-label="Switch branches or tags" tabindex="0" aria-haspopup="true">
+    <i>Branch:</i>
+    <span class="js-select-button css-truncate-target">master</span>
+  </button>
+
+  <div class="select-menu-modal-holder js-menu-content js-navigation-container" data-pjax aria-hidden="true">
+
+    <div class="select-menu-modal">
+      <div class="select-menu-header">
+        <svg aria-label="Close" class="octicon octicon-x js-menu-close" height="16" role="img" version="1.1" viewBox="0 0 12 16" width="12"><path fill-rule="evenodd" d="M7.48 8l3.75 3.75-1.48 1.48L6 9.48l-3.75 3.75-1.48-1.48L4.52 8 .77 4.25l1.48-1.48L6 6.52l3.75-3.75 1.48 1.48z"/></svg>
+        <span class="select-menu-title">Switch branches/tags</span>
+      </div>
+
+      <div class="select-menu-filters">
+        <div class="select-menu-text-filter">
+          <input type="text" aria-label="Find or create a branch…" id="context-commitish-filter-field" class="form-control js-filterable-field js-navigation-enable" placeholder="Find or create a branch…">
+        </div>
+        <div class="select-menu-tabs">
+          <ul>
+            <li class="select-menu-tab">
+              <a href="#" data-tab-filter="branches" data-filter-placeholder="Find or create a branch…" class="js-select-menu-tab" role="tab">Branches</a>
+            </li>
+            <li class="select-menu-tab">
+              <a href="#" data-tab-filter="tags" data-filter-placeholder="Find a tag…" class="js-select-menu-tab" role="tab">Tags</a>
+            </li>
+          </ul>
+        </div>
+      </div>
+
+      <div class="select-menu-list select-menu-tab-bucket js-select-menu-tab-bucket" data-tab-filter="branches" role="menu">
+
+        <div data-filterable-for="context-commitish-filter-field" data-filterable-type="substring">
+
+            <a class="select-menu-item js-navigation-item js-navigation-open "
+               href="/NCEAS/pfx-commercial/blob/eric-edits/portfolio/text/ms.Rmd"
+               data-name="eric-edits"
+               data-skip-pjax="true"
+               rel="nofollow">
+              <svg aria-hidden="true" class="octicon octicon-check select-menu-item-icon" height="16" version="1.1" viewBox="0 0 12 16" width="12"><path fill-rule="evenodd" d="M12 5l-8 8-4-4 1.5-1.5L4 10l6.5-6.5z"/></svg>
+              <span class="select-menu-item-text css-truncate-target js-select-menu-filter-text">
+                eric-edits
+              </span>
+            </a>
+            <a class="select-menu-item js-navigation-item js-navigation-open selected"
+               href="/NCEAS/pfx-commercial/blob/master/portfolio/text/ms.Rmd"
+               data-name="master"
+               data-skip-pjax="true"
+               rel="nofollow">
+              <svg aria-hidden="true" class="octicon octicon-check select-menu-item-icon" height="16" version="1.1" viewBox="0 0 12 16" width="12"><path fill-rule="evenodd" d="M12 5l-8 8-4-4 1.5-1.5L4 10l6.5-6.5z"/></svg>
+              <span class="select-menu-item-text css-truncate-target js-select-menu-filter-text">
+                master
+              </span>
+            </a>
+            <a class="select-menu-item js-navigation-item js-navigation-open "
+               href="/NCEAS/pfx-commercial/blob/matt-trick/portfolio/text/ms.Rmd"
+               data-name="matt-trick"
+               data-skip-pjax="true"
+               rel="nofollow">
+              <svg aria-hidden="true" class="octicon octicon-check select-menu-item-icon" height="16" version="1.1" viewBox="0 0 12 16" width="12"><path fill-rule="evenodd" d="M12 5l-8 8-4-4 1.5-1.5L4 10l6.5-6.5z"/></svg>
+              <span class="select-menu-item-text css-truncate-target js-select-menu-filter-text">
+                matt-trick
+              </span>
+            </a>
+            <a class="select-menu-item js-navigation-item js-navigation-open "
+               href="/NCEAS/pfx-commercial/blob/ole-branch/portfolio/text/ms.Rmd"
+               data-name="ole-branch"
+               data-skip-pjax="true"
+               rel="nofollow">
+              <svg aria-hidden="true" class="octicon octicon-check select-menu-item-icon" height="16" version="1.1" viewBox="0 0 12 16" width="12"><path fill-rule="evenodd" d="M12 5l-8 8-4-4 1.5-1.5L4 10l6.5-6.5z"/></svg>
+              <span class="select-menu-item-text css-truncate-target js-select-menu-filter-text">
+                ole-branch
+              </span>
+            </a>
+        </div>
+
+          <!-- '"` --><!-- </textarea></xmp> --></option></form><form accept-charset="UTF-8" action="/NCEAS/pfx-commercial/branches" class="js-create-branch select-menu-item select-menu-new-item-form js-navigation-item js-new-item-form" data-form-nonce="f2086a1d058512824537f14ace9f906f412f8160" method="post"><div style="margin:0;padding:0;display:inline"><input name="utf8" type="hidden" value="&#x2713;" /><input name="authenticity_token" type="hidden" value="JoVkgzNEbLfCImtzP0yDwV1X66bdufGrmHvasWsvlTikrcCW1IvAAiv8TGjXIqhqi8Ajb6PFz0V+V/HDbgVC6w==" /></div>
+          <svg aria-hidden="true" class="octicon octicon-git-branch select-menu-item-icon" height="16" version="1.1" viewBox="0 0 10 16" width="10"><path fill-rule="evenodd" d="M10 5c0-1.11-.89-2-2-2a1.993 1.993 0 0 0-1 3.72v.3c-.02.52-.23.98-.63 1.38-.4.4-.86.61-1.38.63-.83.02-1.48.16-2 .45V4.72a1.993 1.993 0 0 0-1-3.72C.88 1 0 1.89 0 3a2 2 0 0 0 1 1.72v6.56c-.59.35-1 .99-1 1.72 0 1.11.89 2 2 2 1.11 0 2-.89 2-2 0-.53-.2-1-.53-1.36.09-.06.48-.41.59-.47.25-.11.56-.17.94-.17 1.05-.05 1.95-.45 2.75-1.25S8.95 7.77 9 6.73h-.02C9.59 6.37 10 5.73 10 5zM2 1.8c.66 0 1.2.55 1.2 1.2 0 .65-.55 1.2-1.2 1.2C1.35 4.2.8 3.65.8 3c0-.65.55-1.2 1.2-1.2zm0 12.41c-.66 0-1.2-.55-1.2-1.2 0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2 0 .65-.55 1.2-1.2 1.2zm6-8c-.66 0-1.2-.55-1.2-1.2 0-.65.55-1.2 1.2-1.2.65 0 1.2.55 1.2 1.2 0 .65-.55 1.2-1.2 1.2z"/></svg>
+            <div class="select-menu-item-text">
+              <span class="select-menu-item-heading">Create branch: <span class="js-new-item-name"></span></span>
+              <span class="description">from ‘master’</span>
+            </div>
+            <input type="hidden" name="name" id="name" class="js-new-item-value">
+            <input type="hidden" name="branch" id="branch" value="master">
+            <input type="hidden" name="path" id="path" value="portfolio/text/ms.Rmd">
+</form>
+      </div>
+
+      <div class="select-menu-list select-menu-tab-bucket js-select-menu-tab-bucket" data-tab-filter="tags">
+        <div data-filterable-for="context-commitish-filter-field" data-filterable-type="substring">
+
+        </div>
+
+        <div class="select-menu-no-results">Nothing to show</div>
+      </div>
+
+    </div>
+  </div>
+</div>
+
+  <div class="BtnGroup float-right">
+    <a href="/NCEAS/pfx-commercial/find/master"
+          class="js-pjax-capture-input btn btn-sm BtnGroup-item"
+          data-pjax
+          data-hotkey="t">
+      Find file
+    </a>
+    <button aria-label="Copy file path to clipboard" class="js-zeroclipboard btn btn-sm BtnGroup-item tooltipped tooltipped-s" data-copied-hint="Copied!" type="button">Copy path</button>
+  </div>
+  <div class="breadcrumb js-zeroclipboard-target">
+    <span class="repo-root js-repo-root"><span class="js-path-segment"><a href="/NCEAS/pfx-commercial"><span>pfx-commercial</span></a></span></span><span class="separator">/</span><span class="js-path-segment"><a href="/NCEAS/pfx-commercial/tree/master/portfolio"><span>portfolio</span></a></span><span class="separator">/</span><span class="js-path-segment"><a href="/NCEAS/pfx-commercial/tree/master/portfolio/text"><span>text</span></a></span><span class="separator">/</span><strong class="final-path">ms.Rmd</strong>
+  </div>
+</div>
+
+  <div class="commit-tease">
+      <span class="float-right">
+        <a class="commit-tease-sha" href="/NCEAS/pfx-commercial/commit/292702045bcfdd7fae64dd9db5f50d64fd75b852" data-pjax>
+          2927020
+        </a>
+        <relative-time datetime="2016-12-06T18:41:14Z">Dec 6, 2016</relative-time>
+      </span>
+      <div>
+        <img alt="@seananderson" class="avatar" height="20" src="https://avatars0.githubusercontent.com/u/19349?v=3&amp;s=40" width="20" />
+        <a href="/seananderson" class="user-mention" rel="contributor">seananderson</a>
+          <a href="/NCEAS/pfx-commercial/commit/292702045bcfdd7fae64dd9db5f50d64fd75b852" class="message" data-pjax="true" title="Tweak knitting, title, and copyediting">Tweak knitting, title, and copyediting</a>
+      </div>
+
+    <div class="commit-tease-contributors">
+      <button type="button" class="btn-link muted-link contributors-toggle" data-facebox="#blob_contributors_box">
+        <strong>3</strong>
+         contributors
+      </button>
+          <a class="avatar-link tooltipped tooltipped-s" aria-label="seananderson" href="/NCEAS/pfx-commercial/commits/master/portfolio/text/ms.Rmd?author=seananderson"><img alt="@seananderson" class="avatar" height="20" src="https://avatars0.githubusercontent.com/u/19349?v=3&amp;s=40" width="20" /> </a>
+    <a class="avatar-link tooltipped tooltipped-s" aria-label="eric-ward" href="/NCEAS/pfx-commercial/commits/master/portfolio/text/ms.Rmd?author=eric-ward"><img alt="@eric-ward" class="avatar" height="20" src="https://avatars2.githubusercontent.com/u/5046884?v=3&amp;s=40" width="20" /> </a>
+    <a class="avatar-link tooltipped tooltipped-s" aria-label="OleShelton" href="/NCEAS/pfx-commercial/commits/master/portfolio/text/ms.Rmd?author=OleShelton"><img alt="@OleShelton" class="avatar" height="20" src="https://avatars0.githubusercontent.com/u/9326475?v=3&amp;s=40" width="20" /> </a>
+
+    </div>
+
+    <div id="blob_contributors_box" style="display:none">
+      <h2 class="facebox-header" data-facebox-id="facebox-header">Users who have contributed to this file</h2>
+      <ul class="facebox-user-list" data-facebox-id="facebox-description">
+          <li class="facebox-user-list-item">
+            <img alt="@seananderson" height="24" src="https://avatars2.githubusercontent.com/u/19349?v=3&amp;s=48" width="24" />
+            <a href="/seananderson">seananderson</a>
+          </li>
+          <li class="facebox-user-list-item">
+            <img alt="@eric-ward" height="24" src="https://avatars0.githubusercontent.com/u/5046884?v=3&amp;s=48" width="24" />
+            <a href="/eric-ward">eric-ward</a>
+          </li>
+          <li class="facebox-user-list-item">
+            <img alt="@OleShelton" height="24" src="https://avatars2.githubusercontent.com/u/9326475?v=3&amp;s=48" width="24" />
+            <a href="/OleShelton">OleShelton</a>
+          </li>
+      </ul>
+    </div>
+  </div>
+
+<div class="file">
+  <div class="file-header">
+  <div class="file-actions">
+
+    <div class="BtnGroup">
+      <a href="/NCEAS/pfx-commercial/raw/master/portfolio/text/ms.Rmd" class="btn btn-sm BtnGroup-item" id="raw-url">Raw</a>
+        <a href="/NCEAS/pfx-commercial/blame/master/portfolio/text/ms.Rmd" class="btn btn-sm js-update-url-with-hash BtnGroup-item">Blame</a>
+      <a href="/NCEAS/pfx-commercial/commits/master/portfolio/text/ms.Rmd" class="btn btn-sm BtnGroup-item" rel="nofollow">History</a>
+    </div>
+
+        <a class="btn-octicon tooltipped tooltipped-nw"
+           href="github-mac://openRepo/https://github.com/NCEAS/pfx-commercial?branch=master&amp;filepath=portfolio%2Ftext%2Fms.Rmd"
+           aria-label="Open this file in GitHub Desktop"
+           data-ga-click="Repository, open with desktop, type:mac">
+            <svg aria-hidden="true" class="octicon octicon-device-desktop" height="16" version="1.1" viewBox="0 0 16 16" width="16"><path fill-rule="evenodd" d="M15 2H1c-.55 0-1 .45-1 1v9c0 .55.45 1 1 1h5.34c-.25.61-.86 1.39-2.34 2h8c-1.48-.61-2.09-1.39-2.34-2H15c.55 0 1-.45 1-1V3c0-.55-.45-1-1-1zm0 9H1V3h14v8z"/></svg>
+        </a>
+
+        <!-- '"` --><!-- </textarea></xmp> --></option></form><form accept-charset="UTF-8" action="/NCEAS/pfx-commercial/edit/master/portfolio/text/ms.Rmd" class="inline-form js-update-url-with-hash" data-form-nonce="f2086a1d058512824537f14ace9f906f412f8160" method="post"><div style="margin:0;padding:0;display:inline"><input name="utf8" type="hidden" value="&#x2713;" /><input name="authenticity_token" type="hidden" value="E7wosrszmT41XhSnK9KI5k3J5ywwliRq9n+oQGn6mBqRlIynXPw1i9yAM7zDvKNNm14v5U7qGoQQU4MybNBPyQ==" /></div>
+          <button class="btn-octicon tooltipped tooltipped-nw" type="submit"
+            aria-label="Edit this file" data-hotkey="e" data-disable-with>
+            <svg aria-hidden="true" class="octicon octicon-pencil" height="16" version="1.1" viewBox="0 0 14 16" width="14"><path fill-rule="evenodd" d="M0 12v3h3l8-8-3-3-8 8zm3 2H1v-2h1v1h1v1zm10.3-9.3L12 6 9 3l1.3-1.3a.996.996 0 0 1 1.41 0l1.59 1.59c.39.39.39 1.02 0 1.41z"/></svg>
+          </button>
+</form>        <!-- '"` --><!-- </textarea></xmp> --></option></form><form accept-charset="UTF-8" action="/NCEAS/pfx-commercial/delete/master/portfolio/text/ms.Rmd" class="inline-form" data-form-nonce="f2086a1d058512824537f14ace9f906f412f8160" method="post"><div style="margin:0;padding:0;display:inline"><input name="utf8" type="hidden" value="&#x2713;" /><input name="authenticity_token" type="hidden" value="8ajy8vMOPnFn9X7xnvc7AE7RJesuQLEtAsVBYl7oU6FzgFbnFMGSxI4rWep2mRCrmEbtIlA8j8Pk6WoQW8KEcg==" /></div>
+          <button class="btn-octicon btn-octicon-danger tooltipped tooltipped-nw" type="submit"
+            aria-label="Delete this file" data-disable-with>
+            <svg aria-hidden="true" class="octicon octicon-trashcan" height="16" version="1.1" viewBox="0 0 12 16" width="12"><path fill-rule="evenodd" d="M11 2H9c0-.55-.45-1-1-1H5c-.55 0-1 .45-1 1H2c-.55 0-1 .45-1 1v1c0 .55.45 1 1 1v9c0 .55.45 1 1 1h7c.55 0 1-.45 1-1V5c.55 0 1-.45 1-1V3c0-.55-.45-1-1-1zm-1 12H3V5h1v8h1V5h1v8h1V5h1v8h1V5h1v9zm1-10H2V3h9v1z"/></svg>
+          </button>
+</form>  </div>
+
+  <div class="file-info">
+      658 lines (558 sloc)
+      <span class="file-info-divider"></span>
+    34.1 KB
+  </div>
+</div>
+
+  <div id="readme" class="readme blob instapaper_body">
+    <article class="markdown-body entry-content" itemprop="text"><div class="highlight highlight-source-r"><pre>library(<span class="pl-s"><span class="pl-pds">"</span>knitr<span class="pl-pds">"</span></span>)
+<span class="pl-smi">opts_chunk</span><span class="pl-k">$</span>set(<span class="pl-v">echo</span><span class="pl-k">=</span><span class="pl-c1">FALSE</span>, <span class="pl-v">message</span><span class="pl-k">=</span><span class="pl-c1">FALSE</span>, <span class="pl-v">warning</span><span class="pl-k">=</span><span class="pl-c1">FALSE</span>, <span class="pl-v">results</span><span class="pl-k">=</span><span class="pl-s"><span class="pl-pds">'</span>hide<span class="pl-pds">'</span></span>)</pre></div>
+
+<div class="highlight highlight-source-r"><pre>load(<span class="pl-s"><span class="pl-pds">"</span>../data-generated/output-values.rda<span class="pl-pds">"</span></span>)
+<span class="pl-smi">h1</span> <span class="pl-k">&lt;-</span> readRDS(<span class="pl-s"><span class="pl-pds">"</span>../data-generated/h1_effect.rds<span class="pl-pds">"</span></span>)
+<span class="pl-smi">spoke</span> <span class="pl-k">&lt;-</span> readRDS(<span class="pl-s"><span class="pl-pds">"</span>../data-generated/spoke_fraction.rds<span class="pl-pds">"</span></span>)
+<span class="pl-smi">g2_dens</span> <span class="pl-k">&lt;-</span> readRDS(<span class="pl-s"><span class="pl-pds">"</span>../data-generated/g2_prob_dens_above_zero.rds<span class="pl-pds">"</span></span>)
+<span class="pl-smi">n_grouped_permits</span> <span class="pl-k">&lt;-</span> readRDS(<span class="pl-s"><span class="pl-pds">"</span>../data-generated/n_grouped_permits.rds<span class="pl-pds">"</span></span>)</pre></div>
+
+<h1><a id="user-content-abstract" class="anchor" href="#abstract" aria-hidden="true"><svg aria-hidden="true" class="octicon octicon-link" height="16" version="1.1" viewBox="0 0 16 16" width="16"><path fill-rule="evenodd" d="M4 9h1v1H4c-1.5 0-3-1.69-3-3.5S2.55 3 4 3h4c1.45 0 3 1.69 3 3.5 0 1.41-.91 2.72-2 3.25V8.59c.58-.45 1-1.27 1-2.09C10 5.22 8.98 4 8 4H4c-.98 0-2 1.22-2 2.5S3 9 4 9zm9-3h-1v1h1c1 0 2 1.22 2 2.5S13.98 12 13 12H9c-.98 0-2-1.22-2-2.5 0-.83.42-1.64 1-2.09V6.25c-1.09.53-2 1.84-2 3.25C6 11.31 7.55 13 9 13h4c1.45 0 3-1.69 3-3.5S14.5 6 13 6z"></path></svg></a>Abstract</h1>
+
+<p>A fundamental challenge for individuals who rely on the extraction of natural
 resources for their livelihood is income variability. Fishers, loggers, farmers,
@@
 three axes. First, we show that fishing permits with higher species diversity
-reduces variability (`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction
+reduces variability (<code>r h1[[2]]</code>% [<code>r h1[[1]]</code>--<code>r h1[[3]]</code>% 95% CI] reduction
 per unit increase in mean effective species diversity). Next, we show that
 fishing an additional permit is associated with higher revenue and lower
-variability (e.g., `r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`%
+variability (e.g., <code>r 100-spoke[[2]]</code>% [<code>r 100-spoke[[3]]</code>--<code>r 100-spoke[[1]]</code>%
 IQR] lower expected standard deviation for halibut plus another permit vs.
@@
 within the constraints of a permit set has a fishery-dependent effect on revenue
-and is usually (`r g2_dens`% probability) associated with increased uncertainty
+and is usually (<code>r g2_dens</code>% probability) associated with increased uncertainty
 in revenue the following year. Thus, while there are clear benefits to some axes
@@
 or switching permits --- are costly and often limited by regulations such as
-limited entry programs and individual transferable quotas (ITQs). Take home ...
+limited entry programs and individual transferable quotas (ITQs). Take home ...</p>
 
-\bigskip
-\noindent
+<p>\bigskip
+\noindent</p>
 
-\clearpage
+<p>\clearpage</p>
 
-<!-- Extra: -->
+<h1><a id="user-content-introduction" class="anchor" href="#introduction" aria-hidden="true"><svg aria-hidden="true" class="octicon octicon-link" height="16" version="1.1" viewBox="0 0 16 16" width="16"><path fill-rule="evenodd" d="M4 9h1v1H4c-1.5 0-3-1.69-3-3.5S2.55 3 4 3h4c1.45 0 3 1.69 3 3.5 0 1.41-.91 2.72-2 3.25V8.59c.58-.45 1-1.27 1-2.09C10 5.22 8.98 4 8 4H4c-.98 0-2 1.22-2 2.5S3 9 4 9zm9-3h-1v1h1c1 0 2 1.22 2 2.5S13.98 12 13 12H9c-.98 0-2-1.22-2-2.5 0-.83.42-1.64 1-2.09V6.25c-1.09.53-2 1.84-2 3.25C6 11.31 7.55 13 9 13h4c1.45 0 3-1.69 3-3.5S14.5 6 13 6z"></path></svg></a>Introduction</h1>
 
-<!-- - there is often limited room for intentional diversification within a permit -->
-<!-- - in fact, specializing can be profitable within some fisheries -->
-<!-- - Eric: analog here in farming literature is that diversification -->
-<!--   for small scale farmers generally good, except when they -->
-<!--   specialize on high demand/high value crops (tobacco, cocoa). -->
-<!--   Seems like benefits of diversification are also smaller for -->
-<!--   large-scale industrial farmers. -->
-
-<!-- \noindent -->
-<!-- Main message: -->
-<!-- Diversified portfolios of fishing permits and species reduce risk for fishermen, -->
-<!-- but changing species diversity can be risky, costly, and limited by regulations. -->
-
-<!-- , which is concerning  -->
-<!-- given that some fishing regulations and market forces -->
-<!-- may be increasingly promoting specialization. -->
-<!-- Main message in 25 words or less: -->
-
-<!-- Take 1: -->
-<!-- - There is a trend towards specializing on single salmon permits in Alaska -->
-<!-- - Diversified portfolios of fishing permits and species reduce risk, -->
-<!--   but changing species diversity for an individual permit holder can be risky -->
-<!-- - The types of diversity needed to reduce risk are often restricted by -->
-<!--   cost and regulations -->
-
-<!-- Take 2: -->
-<!-- - The effect of species diversification on revenue and risk for fishermen -->
-<!-- depends on the type of diversification and the fishery. -->
-
-<!-- Goal: Introduction, Methods, Results, Discussion, Conclusion:
-<!-- 600, 900, 500, 800, 200, -->
-
-# Introduction
-
-<!-- Papers to cite [@smith1986] -->
-
-Managing natural resources and sustaining the livelihoods of those that depend
+<p>Managing natural resources and sustaining the livelihoods of those that depend
 on their extraction is difficult. These professions tend to have high
@@
 security and farmers' incomes [@gregory2005], and catastrophic disease outbreaks
-and wildfires increase risk for the logging industry [@kirilenko2007].
+and wildfires increase risk for the logging industry [@kirilenko2007].</p>
 
-Individuals who rely on natural resources for income develop strategies to
+<p>Individuals who rely on natural resources for income develop strategies to
 reduce income variability. For example, farmers may rotate their crops to buffer
@@
 (prices, demand) may discourage the adoption of diverse strategies and help
-promote specialization [@bradshaw2004].
+promote specialization [@bradshaw2004]. </p>
 
-Commercial fishers present an interesting example of diversification, given the
+<p>Commercial fishers present an interesting example of diversification, given the
 extreme risk that some individuals experience, particularly relative to
@@
 uncertainties remain about the fundamental relationship between diversity,
-fishing revenue, and fishing revenue variability at the level of the individual.
+fishing revenue, and fishing revenue variability at the level of the individual.</p>
 
-The state and federal fisheries of Alaska provide a unique opportunity to
+<p>The state and federal fisheries of Alaska provide a unique opportunity to
 understand the role of individuals' strategies in determining income and income
@@
 fisheries limit the entry of new participants through individual fishing quotas
-(IFQs) (REF).
+(IFQs) (REF). </p>
 
-Using a unique dataset collected over 30 years (1985--2014) on individuals,
+<p>Using a unique dataset collected over 30 years (1985--2014) on individuals,
 rather than fleets or vessels, we examine how revenue and revenue variability of
@@
 and revenue variability that builds on results from vessel [@kasperski2013],
-fleet [@sethi2012b], and community-level variability [@sethi2014]. Our model
+fleet [@sethi2012b], and community-level [@sethi2014] variability. Our model
 controls for fishing effort and latent time effects (e.g. changing environmental
@@
 knowledge, this is the first application of this type of model in any natural
-resource field.
+resource field.</p>
 
-\clearpage
+<p>\clearpage</p>
 
-# Methods
+<h1><a id="user-content-methods" class="anchor" href="#methods" aria-hidden="true"><svg aria-hidden="true" class="octicon octicon-link" height="16" version="1.1" viewBox="0 0 16 16" width="16"><path fill-rule="evenodd" d="M4 9h1v1H4c-1.5 0-3-1.69-3-3.5S2.55 3 4 3h4c1.45 0 3 1.69 3 3.5 0 1.41-.91 2.72-2 3.25V8.59c.58-.45 1-1.27 1-2.09C10 5.22 8.98 4 8 4H4c-.98 0-2 1.22-2 2.5S3 9 4 9zm9-3h-1v1h1c1 0 2 1.22 2 2.5S13.98 12 13 12H9c-.98 0-2-1.22-2-2.5 0-.83.42-1.64 1-2.09V6.25c-1.09.53-2 1.84-2 3.25C6 11.31 7.55 13 9 13h4c1.45 0 3-1.69 3-3.5S14.5 6 13 6z"></path></svg></a>Methods</h1>
 
-We developed a hierarchical variance-function regression model, which separately
-models changes in the mean and variance of individual revenue as a function of
-covariates [@davidian1987; @hall1989]. Whereas traditional hierarchical linear
-models assume homoscedasticity --- that the residual variance does not vary
-systematically --- the variance-function regression model allows the variance to
-be heteroscedastic and vary as a function of covariates. Our specific model
-allows certain fishing strategies and levels of species diversification to
-generate more or less revenue and be more or less variable than others.
+<p>We developed a hierarchical variance-function regression model [@davidian1987;
+@hall1989], which separately models changes in the mean and variance of
+individual revenue as a function of covariates. Whereas traditional
+hierarchical linear models assume homoscedasticity --- that the residual
+variance does not vary systematically --- the variance-function regression model
+allows the variance to be heteroscedastic and vary as a function of predictors.
+Our specific model allows certain fishing strategies and levels of species
+diversification to generate more or less revenue and be more or less variable
+than others.</p>
 
-## Data
+<h2><a id="user-content-data" class="anchor" href="#data" aria-hidden="true"><svg aria-hidden="true" class="octicon octicon-link" height="16" version="1.1" viewBox="0 0 16 16" width="16"><path fill-rule="evenodd" d="M4 9h1v1H4c-1.5 0-3-1.69-3-3.5S2.55 3 4 3h4c1.45 0 3 1.69 3 3.5 0 1.41-.91 2.72-2 3.25V8.59c.58-.45 1-1.27 1-2.09C10 5.22 8.98 4 8 4H4c-.98 0-2 1.22-2 2.5S3 9 4 9zm9-3h-1v1h1c1 0 2 1.22 2 2.5S13.98 12 13 12H9c-.98 0-2-1.22-2-2.5 0-.83.42-1.64 1-2.09V6.25c-1.09.53-2 1.84-2 3.25C6 11.31 7.55 13 9 13h4c1.45 0 3-1.69 3-3.5S14.5 6 13 6z"></path></svg></a>Data</h2>
 
-We obtained commercial fisheries landings and revenue data for all permit
-holders in Alaska from `r range_yrs[1]` to `r range_yrs[2]` from the Alaskan
+<p>We obtained commercial fisheries landings and revenue data for all permit
+holders in Alaska from <code>r range_yrs[1]</code> to <code>r range_yrs[2]</code> from the Alaskan
 Commercial Fisheries Entry Commission (CFEC). These data represent a total of
-nearly `r round(ndat$initial_tot_rev/1e9)` billion 2009 USD and a combined `r 
-round(ndat$initial/1e6,1)` million rows of data. We adjusted for inflation by
+nearly <code>r round(ndat$initial_tot_rev/1e9)</code> billion 2009 USD and a combined <code>r 
+round(ndat$initial/1e6,1)</code> million rows of data. We adjusted for inflation by
 setting all revenues to 2009 USD [@bea2016]. We implemented a number of
@@
 revenue was less than \$10,000 2009 USD to focus on individuals for whom
-commercial fishing formed a substantial part of their income.
+commercial fishing formed a substantial part of their income.</p>
 
-We combined similar permits into permit groupings and identified sets of permit
+<p>We combined similar permits into permit groupings and identified sets of permit
 groupings held by an individual in a specific year as "strategies" (Supporting
@@
 that shared similar species composition. After this process of combining
-permits, we were left with `r n_grouped_permits` unique fishing permit groups
+permits, we were left with <code>r n_grouped_permits</code> unique fishing permit groups
 (Table S\ref{tab:strategies}), which we aggregated within each person-year
@@
 "halibut-sablefish". Specific details on our strategy definitions are available
-in the Supporting Information.
+in the Supporting Information.</p>
 
-Our models focused on species diversity as a possible predictor of revenue and
+<p>Our models focused on species diversity as a possible predictor of revenue and
 revenue variability while controlling for effort. For consistency with previous
@@
 permit holder $i$ and time (year) $t$ as the inverse of Simpson's diversity
-index $\lambda$ [@simpson1949; @jost2006] (also referred to in economics as the
-Herfindahl-Hirschman Index) weighted by revenue $R$:
+index $\lambda$ <a href="/NCEAS/pfx-commercial/blob/master/portfolio/text/also%20referred%20to%20in%20economics%20as%20the%0AHerfindahl-Hirschman%20Index">@simpson1949; @jost2006</a> weighted by revenue $R$:</p>
 
-\begin{equation}
+<p>\begin{equation}
 1/\lambda_{i,t} = 1 / \left(\sum\limits_{s=1}^{n_{i,t}}
   \left(R_{i,s,t}/ \sum\limits_{s=1}^{n_{i,t}} R_{i,s,t}\right)\right)^2,
-\end{equation}
+\end{equation}</p>
 
-\noindent where $s$ indexes species from $1$ through $n$. We calculated effort
+<p>\noindent where $s$ indexes species from $1$ through $n$. We calculated effort
 or effective season length as the sum of the season lengths for each permit each
@@
 gillnet permit in Southeast Alaska and 40 days on a statewide longline halibut
-permit, that individual's season length would be 70 days.
+permit, that individual's season length would be 70 days.</p>
 
-## Hierarchical model
+<h2><a id="user-content-hierarchical-model" class="anchor" href="#hierarchical-model" aria-hidden="true"><svg aria-hidden="true" class="octicon octicon-link" height="16" version="1.1" viewBox="0 0 16 16" width="16"><path fill-rule="evenodd" d="M4 9h1v1H4c-1.5 0-3-1.69-3-3.5S2.55 3 4 3h4c1.45 0 3 1.69 3 3.5 0 1.41-.91 2.72-2 3.25V8.59c.58-.45 1-1.27 1-2.09C10 5.22 8.98 4 8 4H4c-.98 0-2 1.22-2 2.5S3 9 4 9zm9-3h-1v1h1c1 0 2 1.22 2 2.5S13.98 12 13 12H9c-.98 0-2-1.22-2-2.5 0-.83.42-1.64 1-2.09V6.25c-1.09.53-2 1.84-2 3.25C6 11.31 7.55 13 9 13h4c1.45 0 3-1.69 3-3.5S14.5 6 13 6z"></path></svg></a>Hierarchical model</h2>
 
-To jointly model revenue and revenue variability as a function of
+<p>To jointly model revenue and revenue variability as a function of
 individual-level covariates, we extended the basic variance-function regression
@@
 errors predicting $\log$ revenue, $\log(R_{i,t})$, for each fisher $i$ and time
-$t$:
+$t$:</p>
 
-\begin{equation}
+<p>\begin{equation}
 \log(R_{i,t}) \sim \mathrm{N} \left(\widehat{\log(R_{i,t})}, \widehat{\sigma_{i,t}^2} \right).
-\end{equation}
+\end{equation}</p>
 
-\noindent Each year $t$ of revenue for fisher $i$ is assigned a strategy indexed
+<p>\noindent Each year $t$ of revenue for fisher $i$ is assigned a strategy indexed
 by $j$ (Supporting Information, Table \ref{tab:strategies}) and expected
-$\log(R_{i,t})$ is modelled as:
+$\log(R_{i,t})$ is modelled as:</p>
 
-<!-- Ole: I might swap out log(R)-hat for something like mu(or your favorit greek letter) just to avoid -->
-<!-- potential confusion like using the term "expected log(R)".   -->
-<!-- I also just don't like hats... but that's my problem. -->
-
-\begin{equation}
+<p>\begin{equation}
 \label{eq:mean}
@@
   \log(R_{i,t-1}),
-\end{equation}
+\end{equation}</p>
 
-\noindent where $S_{i,t}$ is the percent change in species diversity, $\log
-\left(\mathrm{div}_{i,t}/\mathrm{div}_{i,t-1}\right)$, $D_{i,t}$ is the percent
+<p>\noindent where $S_{i,t}$ is the percent change in species diversity, $\log
+\left(\mathrm{div}<em>{i,t}/\mathrm{div}</em>{i,t-1}\right)$, $D_{i,t}$ is the percent
 change in days fished, $\log
-\left(\mathrm{days}_{i,t}/\mathrm{days}_{i,t-1}\right)$, and $I_{i,t}$ is an
-indicator variable that takes a value $0$ if $\mathrm{div}_{i,t} \ge
-\mathrm{div}_{i,t-1}$ (species diversity increasing; generalizing) and takes
-a value $1$ if $\mathrm{div}_{i,t} < \mathrm{div}_{i,t-1}$ (species diversity
+\left(\mathrm{days}<em>{i,t}/\mathrm{days}</em>{i,t-1}\right)$, and $I_{i,t}$ is an
+indicator variable that takes a value $0$ if $\mathrm{div}<em>{i,t} \ge
+\mathrm{div}</em>{i,t-1}$ (species diversity increasing; generalizing) and takes
+a value $1$ if $\mathrm{div}<em>{i,t} &lt; \mathrm{div}</em>{i,t-1}$ (species diversity
 decreasing; specializing). Estimating a separate coefficient for the effect of
 increasing diversity and decreasing diversity allows the costs and benefits of
-these changes to be asymmetric.
+these changes to be asymmetric.</p>
 
-We use $\log(R_{i,t-1}$) as an offset, and thus our model describes the percent
+<p>We use $\log(R_{i,t-1}$) as an offset, and thus our model describes the percent
 change in revenue, i.e., $\log(R_{i,t}) - \log(R_{i,t-1})
@@
 mean-revenue intercepts associated with individual fishers. We removed
-person-year combinations without a subsequent year of revenue.
+person-year combinations without a subsequent year of revenue.</p>
 
-The $\beta$ terms represent estimated coefficients, some of which are allowed to
+<p>The $\beta$ terms represent estimated coefficients, some of which are allowed to
 vary as "random effect" terms. The intercepts $\beta_{0,j,t}$ are allowed to
 vary by strategy-year combinations, are centered on zero, and are constrained by
-a normal distribution:
+a normal distribution:</p>
 
-<!-- Ole: I'd use tau (or something) for the variances here so that people don't have to look at the subscripts to differentiate between variances at different hierarchical levels. -->
+<p>\begin{equation}
+\beta_{0,j,t} \sim \mathrm{N} \left(0, \sigma^2_{\beta_{j,t}} \right).
+\end{equation}</p>
 
-\begin{equation}
-\beta_{0,j,t} \sim \mathrm{N} \left(0, \sigma^2_{\beta_{j,t}} \right).
-\end{equation}
-
-\noindent This approach, combined with the lack of a global intercept,
+<p>\noindent This approach, combined with the lack of a global intercept,
 constrains the mean change in log revenue across all fishers and years to be
@@
 combinations. Initial versions of the model confirmed that the global intercept
-along with strategy-level intercepts would be estimated at almost exactly zero.
+along with strategy-level intercepts would be estimated at almost exactly zero.</p>
 
-The slopes $\beta_{1,j}$ and $\beta_{1,j}$ are then allowed to vary by strategy
-with estimated means:
+<p>The slopes $\beta_{1,j}$ and $\beta_{1,j}$ are then allowed to vary by strategy
+with estimated means:</p>
 
-\begin{align}
-\beta_{1,j} &\sim \mathrm{N} \left(\mu_{\beta_1}, \sigma^2_{\beta_1}\right),\\
-\beta_{2,j} &\sim \mathrm{N} \left(\mu_{\beta_2}, \sigma^2_{\beta_2}\right).
-\end{align}
+<p>\begin{align}
+\beta_{1,j} &amp;\sim \mathrm{N} \left(\mu_{\beta_1}, \sigma^2_{\beta_1}\right),\
+\beta_{2,j} &amp;\sim \mathrm{N} \left(\mu_{\beta_2}, \sigma^2_{\beta_2}\right).
+\end{align}</p>
 
-\noindent The $\beta_{1,j}$'s represent how much a 1\% decrease in species
+<p>\noindent The $\beta_{1,j}$'s represent how much a 1\% decrease in species
 diversity within a strategy translates to a Y\% change in revenue. The
@@
 for percent change in days fished, $\beta_3$, and the interaction between days
-fished and percent change in species diversity, $\beta_4$ and $\beta_5$.
+fished and percent change in species diversity, $\beta_4$ and $\beta_5$.</p>
 
-Whereas a typical hierarchical linear regression model would estimate a single
+<p>Whereas a typical hierarchical linear regression model would estimate a single
 residual error scale term, $\sigma$, we model the scale of the residual error
-with another hierarchical model with a similar form:
+with another hierarchical model with a similar form:</p>
 
-\begin{equation}
+<p>\begin{equation}
 \label{eq:cv}
@@
 \gamma_5 S_{i,t} D_{i,t} I_{i,t}\right).
-\end{equation}
+\end{equation}</p>
 
-\noindent Again, $S_{i,t}$ and $I_{i,t}$ represent, respectively, percent change
+<p>\noindent Again, $S_{i,t}$ and $I_{i,t}$ represent, respectively, percent change
 in species diversity and an indicator variable for increasing or decreasing
@@
 $\sigma_{i,t}$, are positive. The $\gamma$'s are estimated coefficients, three
-of which are allowed to vary by strategy, constrained by normal distributions:
+of which are allowed to vary by strategy, constrained by normal distributions:</p>
 
-\begin{align}
-\gamma_{0,j}  &\sim \mathrm{N} \left(\eta_0 + \eta_1 M_j + \eta_2 E_j, \sigma^2_{\gamma_0}\right),\\
-\gamma_{1,j}  &\sim \mathrm{N} \left(\mu_{\gamma_1}, \sigma^2_{\gamma_1}\right),\\
-\gamma_{2,j}  &\sim \mathrm{N} \left(\mu_{\gamma_2}, \sigma^2_{\gamma_2}\right).
-\end{align}
+<p>\begin{align}
+\gamma_{0,j}  &amp;\sim \mathrm{N} \left(\eta_0 + \eta_1 M_j + \eta_2 E_j, \sigma^2_{\gamma_0}\right),\
+\gamma_{1,j}  &amp;\sim \mathrm{N} \left(\mu_{\gamma_1}, \sigma^2_{\gamma_1}\right),\
+\gamma_{2,j}  &amp;\sim \mathrm{N} \left(\mu_{\gamma_2}, \sigma^2_{\gamma_2}\right).
+\end{align}</p>
 
-\noindent The mean of the $\gamma_{0,j}$ coefficients is modelled with two
+<p>\noindent The mean of the $\gamma_{0,j}$ coefficients is modelled with two
 strategy-level predictors. The symbols $M_j$ and $E_j$ represent the raw mean
@@
 $j$. The coefficients $\eta_1$ and $\eta_2$ represent predictors in this
-strategy-level regression estimating effects across strategies.
+strategy-level regression estimating effects across strategies.</p>
 
-We fit our models with Stan 2.11.1 [@hoffman2014; @standevelopmentteam2015;
+<p>We fit our models with Stan 2.11.1 [@hoffman2014; @standevelopmentteam2015;
 @standevelopmentteam2015a; @carpenter2015] and R 3.3.1 [@rcoreteam2016]. Stan
@@
 iterations of each chain as warm up (i.e. 4,000 final samples). We checked for
-chain convergence visually with trace plots. We also ensured that $\hat{R}<1.05$
+chain convergence visually with trace plots. We also ensured that $\hat{R}&lt;1.05$
 (the potential scale reduction factor) and that the effective sample size, as
@@
 results (with a simulated data set) is available at
-<https://github.com/NCEAS/pfx-commercial>.
+.</p>
 
-*TODO: for the following into the results section or and the method section with
-this:*
+<p><em>TODO: for the following into the results section or and the method section with
+this:</em></p>
 
-Residuals from the mean component of our model clearly show the "V"-shaped
+<p>Residuals from the mean component of our model clearly show the "V"-shaped
 pattern that the variance component of our model seeks to capture (Fig.
@@
 S\ref{fig:resid-downside}). Therefore, it seemed appropriate to model the
-variability as a proxy for revenue risk.
+variability as a proxy for revenue risk. </p>
 
-The estimated year-permit intercepts represent years when people with
+<p>The estimated year-permit intercepts represent years when people with
 a particular permit set in a given year tended to do better or worse than the
@@
 individual's expected revenue and variability is after accounting for the
-general trend in revenue for that permit set.
+general trend in revenue for that permit set.</p>
 
-Our model estimates overall average effects (Fig. S\ref{fig:fe}) and effects
-that vary by permit set (Fig. S\ref{fig:re}).
+<p>Our model estimates overall average effects (Fig. S\ref{fig:fe}) and effects
+that vary by permit set (Fig. S\ref{fig:re}).</p>
 
-# Results
+<h1><a id="user-content-results" class="anchor" href="#results" aria-hidden="true"><svg aria-hidden="true" class="octicon octicon-link" height="16" version="1.1" viewBox="0 0 16 16" width="16"><path fill-rule="evenodd" d="M4 9h1v1H4c-1.5 0-3-1.69-3-3.5S2.55 3 4 3h4c1.45 0 3 1.69 3 3.5 0 1.41-.91 2.72-2 3.25V8.59c.58-.45 1-1.27 1-2.09C10 5.22 8.98 4 8 4H4c-.98 0-2 1.22-2 2.5S3 9 4 9zm9-3h-1v1h1c1 0 2 1.22 2 2.5S13.98 12 13 12H9c-.98 0-2-1.22-2-2.5 0-.83.42-1.64 1-2.09V6.25c-1.09.53-2 1.84-2 3.25C6 11.31 7.55 13 9 13h4c1.45 0 3-1.69 3-3.5S14.5 6 13 6z"></path></svg></a>Results</h1>
 
-We evaluated trends in revenue and diversification over time for individuals
+<p>We evaluated trends in revenue and diversification over time for individuals
 fishing commercially in Alaska from 1985--2014 (Fig. \ref{fig:ts}A). The
@@
 salmon specialists in the 1980s and 1990s --- the trend disappears if salmon
-license holders are excluded (Fig. \ref{fig:ts}D).
+license holders are excluded (Fig. \ref{fig:ts}D).</p>
 
-Revenue variability decreases exponentially with higher mean fished species
+<p>Revenue variability decreases exponentially with higher mean fished species
 diversity across permit sets (Fig. \ref{fig:across}). Controlling for effort,
 strategies with one unit higher mean effective species diversity had on average
-`r h1[[2]]`% (`r h1[[1]]`--`r h1[[3]]`% 95% CI) lower variability. This pattern
+<code>r h1[[2]]</code>% (<code>r h1[[1]]</code>--<code>r h1[[3]]</code>% 95% CI) lower variability. This pattern
 looks qualitatively similar with or without individuals targeting salmon. As
@@
 combined halibut with a salmon permit and had high levels of fished species
-diversity (Fig. \ref{fig:across}).
+diversity (Fig. \ref{fig:across}).</p>
 
-Comparing single permit holders to individuals holding that same permit plus
+<p>Comparing single permit holders to individuals holding that same permit plus
 additional permits illustrates the expected reduction in revenue variability
@@
 compared to individuals fishing only a halibut longline permit (median reduction
-of `r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR], Fig.
+of <code>r 100-spoke[[2]]</code>% [<code>r 100-spoke[[3]]</code>--<code>r 100-spoke[[1]]</code>% IQR], Fig.
 \ref{fig:spokes}A). Likewise, adding a permit to a herring roe gillnet strategy
@@
 Adding permits also usually increased expected revenue, although the permits
-themselves represent a significant expense (e.g. TODO).
+themselves represent a significant expense (e.g. TODO).</p>
 
-In contrast to the clear across-strategy effects of species diversity on revenue
+<p>In contrast to the clear across-strategy effects of species diversity on revenue
 and revenue variability, the effect of changing fished species diversity for an
-individual *within* the constraints of a given permit set is more variable (Fig.
+individual <em>within</em> the constraints of a given permit set is more variable (Fig.
 \ref{fig:within}). After accounting for fishing effort (Fig. S\ref{fig:fe}),
@@
 increased revenue (~0.5% increased revenue per 1% decrease in species diversity
-on average or TODO ... `r g2_dens`% probability; Fig. \ref{fig:within}). This is
+on average or TODO ... <code>r g2_dens</code>% probability; Fig. \ref{fig:within}). This is
 particularly clear for permits where individuals tend to specialize on sockeye
@@
 (fishing more days), generalizing tends to be associated with higher revenue
-when not standardizing for days finished (Fig. S-TODO).
+when not standardizing for days finished (Fig. S-TODO). </p>
 
-After accounting for fishing effort, adjusting (increasing or decreasing) an
+<p>After accounting for fishing effort, adjusting (increasing or decreasing) an
 individual's fished species diversity from year-to-year within the constraints
@@
 species diversity levels above one for these permits are primarily due to
-bycatch.
+bycatch. </p>
 
-The Alaskan halibut and sablefish fisheries were converted to an individual
+<p>The Alaskan halibut and sablefish fisheries were converted to an individual
 fishery quota (IFQ) system in 1994. This catch share system... TODO. We refit
@@
 individuals fishing only halibut or sablefish and less so for individuals
-combining halibut or sable fish with a salmon permit (Fig. S\ref{fig:ifq}B).
+combining halibut or sable fish with a salmon permit (Fig. S\ref{fig:ifq}B).</p>
 
-# Discussion
+<h1><a id="user-content-discussion" class="anchor" href="#discussion" aria-hidden="true"><svg aria-hidden="true" class="octicon octicon-link" height="16" version="1.1" viewBox="0 0 16 16" width="16"><path fill-rule="evenodd" d="M4 9h1v1H4c-1.5 0-3-1.69-3-3.5S2.55 3 4 3h4c1.45 0 3 1.69 3 3.5 0 1.41-.91 2.72-2 3.25V8.59c.58-.45 1-1.27 1-2.09C10 5.22 8.98 4 8 4H4c-.98 0-2 1.22-2 2.5S3 9 4 9zm9-3h-1v1h1c1 0 2 1.22 2 2.5S13.98 12 13 12H9c-.98 0-2-1.22-2-2.5 0-.83.42-1.64 1-2.09V6.25c-1.09.53-2 1.84-2 3.25C6 11.31 7.55 13 9 13h4c1.45 0 3-1.69 3-3.5S14.5 6 13 6z"></path></svg></a>Discussion</h1>
 
-### Take homes / summary
+<h3><a id="user-content-take-homes--summary" class="anchor" href="#take-homes--summary" aria-hidden="true"><svg aria-hidden="true" class="octicon octicon-link" height="16" version="1.1" viewBox="0 0 16 16" width="16"><path fill-rule="evenodd" d="M4 9h1v1H4c-1.5 0-3-1.69-3-3.5S2.55 3 4 3h4c1.45 0 3 1.69 3 3.5 0 1.41-.91 2.72-2 3.25V8.59c.58-.45 1-1.27 1-2.09C10 5.22 8.98 4 8 4H4c-.98 0-2 1.22-2 2.5S3 9 4 9zm9-3h-1v1h1c1 0 2 1.22 2 2.5S13.98 12 13 12H9c-.98 0-2-1.22-2-2.5 0-.83.42-1.64 1-2.09V6.25c-1.09.53-2 1.84-2 3.25C6 11.31 7.55 13 9 13h4c1.45 0 3-1.69 3-3.5S14.5 6 13 6z"></path></svg></a>Take homes / summary</h3>
 
-- Suggests it's mostly an across strategy/permit effect (lowering variability),
-  which isn't always a possibility because of regulations and cost
-- Adjusting within current permit constraints is risky... Not clearly risk
-  reducing to generalize from one year to the next
-- not as simple as increase diversity (within permit) and reduce risk
-- What's methodologically novel
-    - Huge dataset
-    - Permit level data (on individuals, not vessels)
-    - Account for covariates
-    - Accounting for strategies
-    - Keeping annual data (not condensing to mean and cv)
-    - Joint modelling of mean and variance
-    - look at different mechanisms of diversification
-- 2 dimensions to think about (revenue and variability); and the trade-off is fishery specific
-    - effects within a permit strategy are permit specific
-    - Importantly it's a tradeoff - 2 dimensions - sometimes make more money by specializing... Some highly variable single species permits make a ton of money... which preferable?
+<ul>
+<li>Suggests it's mostly an across strategy/permit effect (lowering variability),
+which isn't always a possibility because of regulations and cost</li>
+<li>Adjusting within current permit constraints is risky... Not clearly risk
+reducing to generalize from one year to the next</li>
+<li>not as simple as increase diversity (within permit) and reduce risk</li>
+<li>What's methodologically novel
 
-### Reconciling the effects at different levels of species diversity
+<ul>
+<li>Huge dataset</li>
+<li>Permit level data (on individuals, not vessels)</li>
+<li>Account for covariates</li>
+<li>Accounting for strategies</li>
+<li>Keeping annual data (not condensing to mean and cv)</li>
+<li>Joint modelling of mean and variance</li>
+<li>look at different mechanisms of diversification</li>
+</ul></li>
+<li>2 dimensions to think about (revenue and variability); and the trade-off is fishery specific
 
-- How fit in with K and H?
+<ul>
+<li>effects within a permit strategy are permit specific</li>
+<li>Importantly it's a tradeoff - 2 dimensions - sometimes make more money by specializing... Some highly variable single species permits make a ton of money... which preferable?</li>
+</ul></li>
+</ul>
 
-- EW: part of the difference is individuals vs vessels. For example, K&H
+<h3><a id="user-content-reconciling-the-effects-at-different-levels-of-species-diversity" class="anchor" href="#reconciling-the-effects-at-different-levels-of-species-diversity" aria-hidden="true"><svg aria-hidden="true" class="octicon octicon-link" height="16" version="1.1" viewBox="0 0 16 16" width="16"><path fill-rule="evenodd" d="M4 9h1v1H4c-1.5 0-3-1.69-3-3.5S2.55 3 4 3h4c1.45 0 3 1.69 3 3.5 0 1.41-.91 2.72-2 3.25V8.59c.58-.45 1-1.27 1-2.09C10 5.22 8.98 4 8 4H4c-.98 0-2 1.22-2 2.5S3 9 4 9zm9-3h-1v1h1c1 0 2 1.22 2 2.5S13.98 12 13 12H9c-.98 0-2-1.22-2-2.5 0-.83.42-1.64 1-2.09V6.25c-1.09.53-2 1.84-2 3.25C6 11.31 7.55 13 9 13h4c1.45 0 3-1.69 3-3.5S14.5 6 13 6z"></path></svg></a>Reconciling the effects at different levels of species diversity</h3>
+
+<ul>
+<li><p>How fit in with K and H?</p></li>
+<li><p>EW: part of the difference is individuals vs vessels. For example, K&amp;H
 suggested that diversity declined following IFQs for halibut sablefish
@@
 halibut/sablefish permit to be affected by IFQs. The person, not the vessel
-holds the quota share: https://alaskafisheries.noaa.gov/fisheries/ifq
-- Year to year risky... But maybe evens out/better over time
-- diversity within permit - often bycatch, limited room for intentional diversification (unless you have a M or salmon permit)
-- Risk results not really risk, but unpredictability given our covariates ...
-don't change anything and it's most predictable how well you'll do
-- What's unpredictable in our model, might be more predictable on a year to
-year, species by species basis for individual permit holder
-- Equilibrium, average variability vs year to year effect for one person
+holds the quota share: <a href="https://alaskafisheries.noaa.gov/fisheries/ifq">https://alaskafisheries.noaa.gov/fisheries/ifq</a></p></li>
+<li>Year to year risky... But maybe evens out/better over time</li>
+<li>diversity within permit - often bycatch, limited room for intentional diversification (unless you have a M or salmon permit)</li>
+<li>Risk results not really risk, but unpredictability given our covariates ...
+don't change anything and it's most predictable how well you'll do</li>
+<li>What's unpredictable in our model, might be more predictable on a year to
+year, species by species basis for individual permit holder</li>
+<li>Equilibrium, average variability vs year to year effect for one person</li>
+</ul>
 
-### Limitations + obvious extensions
+<h3><a id="user-content-limitations--obvious-extensions" class="anchor" href="#limitations--obvious-extensions" aria-hidden="true"><svg aria-hidden="true" class="octicon octicon-link" height="16" version="1.1" viewBox="0 0 16 16" width="16"><path fill-rule="evenodd" d="M4 9h1v1H4c-1.5 0-3-1.69-3-3.5S2.55 3 4 3h4c1.45 0 3 1.69 3 3.5 0 1.41-.91 2.72-2 3.25V8.59c.58-.45 1-1.27 1-2.09C10 5.22 8.98 4 8 4H4c-.98 0-2 1.22-2 2.5S3 9 4 9zm9-3h-1v1h1c1 0 2 1.22 2 2.5S13.98 12 13 12H9c-.98 0-2-1.22-2-2.5 0-.83.42-1.64 1-2.09V6.25c-1.09.53-2 1.84-2 3.25C6 11.31 7.55 13 9 13h4c1.45 0 3-1.69 3-3.5S14.5 6 13 6z"></path></svg></a>Limitations + obvious extensions</h3>
 
-- limitations...
-- Vary by time
-- Vary by mean diversity
-- Spatial variation
-- Nonlinear effects
-- Response to specific events like EVOS and ITQs
+<ul>
+<li>limitations...</li>
+<li>Vary by time</li>
+<li>Vary by mean diversity</li>
+<li>Spatial variation</li>
+<li>Nonlinear effects</li>
+<li>Response to specific events like EVOS and ITQs</li>
+</ul>
 
-### Implications for limited entry, ITQs, and catch shares
+<h3><a id="user-content-implications-for-limited-entry-itqs-and-catch-shares" class="anchor" href="#implications-for-limited-entry-itqs-and-catch-shares" aria-hidden="true"><svg aria-hidden="true" class="octicon octicon-link" height="16" version="1.1" viewBox="0 0 16 16" width="16"><path fill-rule="evenodd" d="M4 9h1v1H4c-1.5 0-3-1.69-3-3.5S2.55 3 4 3h4c1.45 0 3 1.69 3 3.5 0 1.41-.91 2.72-2 3.25V8.59c.58-.45 1-1.27 1-2.09C10 5.22 8.98 4 8 4H4c-.98 0-2 1.22-2 2.5S3 9 4 9zm9-3h-1v1h1c1 0 2 1.22 2 2.5S13.98 12 13 12H9c-.98 0-2-1.22-2-2.5 0-.83.42-1.64 1-2.09V6.25c-1.09.53-2 1.84-2 3.25C6 11.31 7.55 13 9 13h4c1.45 0 3-1.69 3-3.5S14.5 6 13 6z"></path></svg></a>Implications for limited entry, ITQs, and catch shares</h3>
 
-- if this is true, this is concerning because in many areas of the world and
-fisheries, regulations and market forces have encouraged specialization
-- Permit diversity result: a privileged position - expensive to get to into
-(quote king crab license)
-- Locking into permits limits ability to pick lower risk strategies and is
+<ul>
+<li>if this is true, this is concerning because in many areas of the world and
+fisheries, regulations and market forces have encouraged specialization</li>
+<li>Permit diversity result: a privileged position - expensive to get to into
+(quote king crab license)</li>
+<li>Locking into permits limits ability to pick lower risk strategies and is
 getting people locked into low diversity permits that may be doing well now,
-but not necessarily always
+but not necessarily always</li>
+</ul>
 
-### Conclusions + recommendations
+<h3><a id="user-content-conclusions--recommendations" class="anchor" href="#conclusions--recommendations" aria-hidden="true"><svg aria-hidden="true" class="octicon octicon-link" height="16" version="1.1" viewBox="0 0 16 16" width="16"><path fill-rule="evenodd" d="M4 9h1v1H4c-1.5 0-3-1.69-3-3.5S2.55 3 4 3h4c1.45 0 3 1.69 3 3.5 0 1.41-.91 2.72-2 3.25V8.59c.58-.45 1-1.27 1-2.09C10 5.22 8.98 4 8 4H4c-.98 0-2 1.22-2 2.5S3 9 4 9zm9-3h-1v1h1c1 0 2 1.22 2 2.5S13.98 12 13 12H9c-.98 0-2-1.22-2-2.5 0-.83.42-1.64 1-2.09V6.25c-1.09.53-2 1.84-2 3.25C6 11.31 7.55 13 9 13h4c1.45 0 3-1.69 3-3.5S14.5 6 13 6z"></path></svg></a>Conclusions + recommendations</h3>
 
-\bibliographystyle{pnas2011}
-\bibliography{refs}
+<p>\bibliographystyle{pnas2011}
+\bibliography{refs}</p>
 
-\clearpage
+<p>\clearpage</p>
 
-# Figures
+<h1><a id="user-content-figures" class="anchor" href="#figures" aria-hidden="true"><svg aria-hidden="true" class="octicon octicon-link" height="16" version="1.1" viewBox="0 0 16 16" width="16"><path fill-rule="evenodd" d="M4 9h1v1H4c-1.5 0-3-1.69-3-3.5S2.55 3 4 3h4c1.45 0 3 1.69 3 3.5 0 1.41-.91 2.72-2 3.25V8.59c.58-.45 1-1.27 1-2.09C10 5.22 8.98 4 8 4H4c-.98 0-2 1.22-2 2.5S3 9 4 9zm9-3h-1v1h1c1 0 2 1.22 2 2.5S13.98 12 13 12H9c-.98 0-2-1.22-2-2.5 0-.83.42-1.64 1-2.09V6.25c-1.09.53-2 1.84-2 3.25C6 11.31 7.55 13 9 13h4c1.45 0 3-1.69 3-3.5S14.5 6 13 6z"></path></svg></a>Figures</h1>
 
-\begin{figure}[htbp]
+<p>\begin{figure}[htbp]
 \begin{center}
-\includegraphics[width=0.7\textwidth]{../figs/fig1-jk.pdf}
+\includegraphics[width=0.7\textwidth]{../figs/fig1-jk.pdf}</p>
 
-\caption{ Trends in revenue and diversification over time. (A) Map of fishing
+<p>\caption{ Trends in revenue and diversification over time. (A) Map of fishing
 regions in Alaska, USA. (B) Total revenue in billions of 2009 U.S. dollars from
-`r range_yrs[1]` to `r range_yrs[2]` for multiple (grey) and single (blue)
+<code>r range_yrs[1]</code> to <code>r range_yrs[2]</code> for multiple (grey) and single (blue)
 permit holders. (C) Proportion of single permit holders for all fisheries (thick
-black line) and with permit holders omitted each year who fished any of the `r 
-n_jacknifed` top species groups (thin lines). (D) Mean effective diversity of
+black line) and with permit holders omitted each year who fished any of the <code>r 
+n_jacknifed</code> top species groups (thin lines). (D) Mean effective diversity of
 species fished (Simpson's diversity index) for all fisheries (thick black line)
-and with permit holders omitted each year who fished any of the `r n_jacknifed`
-top species groups (thin lines). } \label{fig:ts} \end{center} \end{figure}
+and with permit holders omitted each year who fished any of the <code>r n_jacknifed</code>
+top species groups (thin lines). } \label{fig:ts} \end{center} \end{figure}</p>
 
-\clearpage
+<p>\clearpage</p>
 
-\begin{figure}[htbp]
+<p>\begin{figure}[htbp]
 \begin{center}
-\includegraphics[width=0.7\textwidth]{../figs/stan-across-strategy-variability2.pdf}
+\includegraphics[width=0.7\textwidth]{../figs/stan-across-strategy-variability2.pdf}</p>
 
-\caption{Individuals who fish permits and permit combinations (permit
+<p>\caption{Individuals who fish permits and permit combinations (permit
 strategies) with higher species diversity have lower expected revenue
 variability. On the y-axis, dots, thick lines, and thin lines represent
-posterior medians, 50\%, and 95\% credible intervals for the `r nstr` most
+posterior medians, 50\%, and 95\% credible intervals for the <code>r nstr</code> most
 common permit strategies. Thick black line, and dark and light grey shaded
@@
 expected standard deviation for a permit holder who does not change species
-diversity or days fished from year-to-year. }
+diversity or days fished from year-to-year. }</p>
 
-\label{fig:across}
+<p>\label{fig:across}
 \end{center}
-\end{figure}
+\end{figure}</p>
 
-\clearpage
+<p>\clearpage</p>
 
-\begin{figure}[htbp]
+<p>\begin{figure}[htbp]
 \begin{center}
-\includegraphics[width=0.9\textwidth]{../figs/stan-gg-spoke2.pdf}
+\includegraphics[width=0.9\textwidth]{../figs/stan-gg-spoke2.pdf}</p>
 
-\caption{Diversifying permits reduces risk and usually increases revenue. Two
+<p>\caption{Diversifying permits reduces risk and usually increases revenue. Two
 examples of expected revenue variability and median revenue for owners of (A)
@@
 blue). On the y-axis, dots and lines represent posterior median and 95\%
-credible intervals of expected standard deviation.}
+credible intervals of expected standard deviation.}</p>
 
-\label{fig:spokes}
+<p>\label{fig:spokes}
 \end{center}
-\end{figure}
+\end{figure}</p>
 
-\clearpage
+<p>\clearpage</p>
 
-\begin{figure}[htbp]
+<p>\begin{figure}[htbp]
 \begin{center}
-\includegraphics[width=1.1\textwidth]{../figs/stan-offset-break-anti-spaghetti.pdf}
+\includegraphics[width=1.1\textwidth]{../figs/stan-offset-break-anti-spaghetti.pdf}</p>
 
-\caption{Effects of (A) generalizing and (B) specializing from year-to-year for
-individual permit holders \emph{within} the `r nstr` most common strategies.
+<p>\caption{Effects of (A) generalizing and (B) specializing from year-to-year for
+individual permit holders \emph{within} the <code>r nstr</code> most common strategies.
 Changing species diversity from year-to-year is associated with higher
@@
 year-to-year. Dots and line segments represent medians and 50\% credible
-intervals of posteriors. }
+intervals of posteriors. }</p>
 
-\label{fig:within}
+<p>\label{fig:within}
 \end{center}
-\end{figure}
+\end{figure}</p>
 
-\clearpage
+<p>\clearpage</p>
 
-<!-- # Notes -->
+</article>
+  </div>
 
-<!-- ## @kasperski2013 -->
-<!-- - vessel-level CFEC data -->
-<!-- - dome-shaped result -->
-<!-- - Herfindahl-Hirschman Index/HHI (Simpson) -->
-<!-- - diversity decreasing for those still fishing in 2010, but not otherwise -->
+</div>
 
-<!-- ## @sethi2012b -->
-<!-- - risk measures to 90 Alaskan fishing fishery/fleets -->
-<!-- - revenue more variable than catch (+ price = extra variability) -->
-<!-- - large catches, big fleets = lower risk -->
+<button type="button" data-facebox="#jump-to-line" data-facebox-class="linejump" data-hotkey="l" class="d-none">Jump to Line</button>
+<div id="jump-to-line" style="display:none">
+  <!-- '"` --><!-- </textarea></xmp> --></option></form><form accept-charset="UTF-8" action="" class="js-jump-to-line-form" method="get"><div style="margin:0;padding:0;display:inline"><input name="utf8" type="hidden" value="&#x2713;" /></div>
+    <input class="form-control linejump-input js-jump-to-line-field" type="text" placeholder="Jump to line&hellip;" aria-label="Jump to line" autofocus>
+    <button type="submit" class="btn">Go</button>
+</form></div>
 
-<!-- ## @sethi2010b -->
-<!-- - risk management for fisheries review -->
+  </div>
+  <div class="modal-backdrop js-touch-events"></div>
+</div>
 
-<!-- ## @sethi2012 -->
-<!-- - equal-share catch-pooling cooperatives in king and snow crab -->
-<!-- - "variability reduction depends on the degree of correlation amongst -->
-<!--   participants’ catches" -->
+    </div>
+  </div>
 
-<!-- ## @sethi2014 -->
-<!-- - community-level CFEC data -->
-<!-- - number of fisheries strongest predictor of (lower) CV/CVaR -->
-<!-- - also simpson index -->
-<!-- - suggest barriers to commercial entry increase community risk by restricting -->
-<!--   access to 'variety of fisheries' -->
-<!-- - some communities have limited ability to diversify; may need to diversify -->
-<!--   outside of fishing -->
+    </div>
 
-<!-- [@perruso2005] -->
-<!-- [@hilborn2001] -->
-<!-- [@edwards2004] -->
-<!-- [@trenkel2013] -->
-<!-- [@mumford2009] -->
-<!-- [@oostenbrugge2002] -->
-<!-- [@baldursson1997] -->
-<!-- [@jardine2015] -->
-<!-- [@sanchirico2008] -->
+        <div class="container site-footer-container">
+  <div class="site-footer" role="contentinfo">
+    <ul class="site-footer-links float-right">
+        <li><a href="https://github.com/contact" data-ga-click="Footer, go to contact, text:contact">Contact GitHub</a></li>
+      <li><a href="https://developer.github.com" data-ga-click="Footer, go to api, text:api">API</a></li>
+      <li><a href="https://training.github.com" data-ga-click="Footer, go to training, text:training">Training</a></li>
+      <li><a href="https://shop.github.com" data-ga-click="Footer, go to shop, text:shop">Shop</a></li>
+        <li><a href="https://github.com/blog" data-ga-click="Footer, go to blog, text:blog">Blog</a></li>
+        <li><a href="https://github.com/about" data-ga-click="Footer, go to about, text:about">About</a></li>
 
-<!-- vim: textwidth=80 
--->
+    </ul>
+
+    <a href="https://github.com" aria-label="Homepage" class="site-footer-mark" title="GitHub">
+      <svg aria-hidden="true" class="octicon octicon-mark-github" height="24" version="1.1" viewBox="0 0 16 16" width="24"><path fill-rule="evenodd" d="M8 0C3.58 0 0 3.58 0 8c0 3.54 2.29 6.53 5.47 7.59.4.07.55-.17.55-.38 0-.19-.01-.82-.01-1.49-2.01.37-2.53-.49-2.69-.94-.09-.23-.48-.94-.82-1.13-.28-.15-.68-.52-.01-.53.63-.01 1.08.58 1.23.82.72 1.21 1.87.87 2.33.66.07-.52.28-.87.51-1.07-1.78-.2-3.64-.89-3.64-3.95 0-.87.31-1.59.82-2.15-.08-.2-.36-1.02.08-2.12 0 0 .67-.21 2.2.82.64-.18 1.32-.27 2-.27.68 0 1.36.09 2 .27 1.53-1.04 2.2-.82 2.2-.82.44 1.1.16 1.92.08 2.12.51.56.82 1.27.82 2.15 0 3.07-1.87 3.75-3.65 3.95.29.25.54.73.54 1.48 0 1.07-.01 1.93-.01 2.2 0 .21.15.46.55.38A8.013 8.013 0 0 0 16 8c0-4.42-3.58-8-8-8z"/></svg>
+</a>
+    <ul class="site-footer-links">
+      <li>&copy; 2016 <span title="0.12418s from github-fe116-cp1-prd.iad.github.net">GitHub</span>, Inc.</li>
+        <li><a href="https://github.com/site/terms" data-ga-click="Footer, go to terms, text:terms">Terms</a></li>
+        <li><a href="https://github.com/site/privacy" data-ga-click="Footer, go to privacy, text:privacy">Privacy</a></li>
+        <li><a href="https://github.com/security" data-ga-click="Footer, go to security, text:security">Security</a></li>
+        <li><a href="https://status.github.com/" data-ga-click="Footer, go to status, text:status">Status</a></li>
+        <li><a href="https://help.github.com" data-ga-click="Footer, go to help, text:help">Help</a></li>
+    </ul>
+  </div>
+</div>
+
+    <div id="ajax-error-message" class="ajax-error-message flash flash-error">
+      <svg aria-hidden="true" class="octicon octicon-alert" height="16" version="1.1" viewBox="0 0 16 16" width="16"><path fill-rule="evenodd" d="M8.865 1.52c-.18-.31-.51-.5-.87-.5s-.69.19-.87.5L.275 13.5c-.18.31-.18.69 0 1 .19.31.52.5.87.5h13.7c.36 0 .69-.19.86-.5.17-.31.18-.69.01-1L8.865 1.52zM8.995 13h-2v-2h2v2zm0-3h-2V6h2v4z"/></svg>
+      <button type="button" class="flash-close js-flash-close js-ajax-error-dismiss" aria-label="Dismiss error">
+        <svg aria-hidden="true" class="octicon octicon-x" height="16" version="1.1" viewBox="0 0 12 16" width="12"><path fill-rule="evenodd" d="M7.48 8l3.75 3.75-1.48 1.48L6 9.48l-3.75 3.75-1.48-1.48L4.52 8 .77 4.25l1.48-1.48L6 6.52l3.75-3.75 1.48 1.48z"/></svg>
+      </button>
+      You can't perform that action at this time.
+    </div>
+
+      <script crossorigin="anonymous" integrity="sha256-eGrxxkowQBvwoW6v7VFBW5vLA/cv8xg6H6YAfXyw+Xk=" src="https://assets-cdn.github.com/assets/frameworks-786af1c64a30401bf0a16eafed51415b9bcb03f72ff3183a1fa6007d7cb0f979.js"></script>
+      <script async="async" crossorigin="anonymous" integrity="sha256-HOK6rPd2DwVwRK5mpbjjntq1ZYO5/9ZD6rvtdTOK9iA=" src="https://assets-cdn.github.com/assets/github-1ce2baacf7760f057044ae66a5b8e39edab56583b9ffd643eabbed75338af620.js"></script>
+
+    <div class="js-stale-session-flash stale-session-flash flash flash-warn flash-banner d-none">
+      <svg aria-hidden="true" class="octicon octicon-alert" height="16" version="1.1" viewBox="0 0 16 16" width="16"><path fill-rule="evenodd" d="M8.865 1.52c-.18-.31-.51-.5-.87-.5s-.69.19-.87.5L.275 13.5c-.18.31-.18.69 0 1 .19.31.52.5.87.5h13.7c.36 0 .69-.19.86-.5.17-.31.18-.69.01-1L8.865 1.52zM8.995 13h-2v-2h2v2zm0-3h-2V6h2v4z"/></svg>
+      <span class="signed-in-tab-flash">You signed in with another tab or window. <a href="">Reload</a> to refresh your session.</span>
+      <span class="signed-out-tab-flash">You signed out in another tab or window. <a href="">Reload</a> to refresh your session.</span>
+    </div>
+    <div class="facebox" id="facebox" style="display:none;">
+  <div class="facebox-popup">
+    <div class="facebox-content" role="dialog" aria-labelledby="facebox-header" aria-describedby="facebox-description">
+    </div>
+    <button type="button" class="facebox-close js-facebox-close" aria-label="Close modal">
+      <svg aria-hidden="true" class="octicon octicon-x" height="16" version="1.1" viewBox="0 0 12 16" width="12"><path fill-rule="evenodd" d="M7.48 8l3.75 3.75-1.48 1.48L6 9.48l-3.75 3.75-1.48-1.48L4.52 8 .77 4.25l1.48-1.48L6 6.52l3.75-3.75 1.48 1.48z"/></svg>
+    </button>
+  </div>
+</div>
+
+  </body>
+</html>

### 2016-12-06 portfolio/text/ms.Rmd
Revert "weird sync issue"
This reverts commit be32164a98a0818fc1cbb2b83070a1ef78a97abe.

--- 2016-12-06 portfolio/text/ms.Rmd
@@
 library("knitr")
-opts_chunk$set(cache=TRUE, echo=FALSE, message=FALSE, autodep=TRUE,
-  results='hide')
-opts_chunk$set(warning=FALSE, message=FALSE, cache.path="../cache/",
-  cache.lazy=TRUE)
+opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, results='hide')
 ```
-```{r get-dat, cache=FALSE}
+```{r get-dat}
 load("../data-generated/output-values.rda")
@@
 
-We developed a hierarchical variance-function regression model, which separately
-models changes in the mean and variance of individual revenue as a function of
-covariates [@davidian1987; @hall1989]. Whereas traditional hierarchical linear
-models assume homoscedasticity --- that the residual variance does not vary
-systematically --- the variance-function regression model allows the variance to
-be heteroscedastic and vary as a function of covariates. Our specific model
-allows certain fishing strategies and levels of species diversification to
-generate more or less revenue and be more or less variable than others.
+We developed a hierarchical variance-function regression model [@davidian1987;
+@hall1989], which separately models changes in the mean and variance of
+individual revenue as a function of covariates. Whereas traditional
+hierarchical linear models assume homoscedasticity --- that the residual
+variance does not vary systematically --- the variance-function regression model
+allows the variance to be heteroscedastic and vary as a function of predictors.
+Our specific model allows certain fishing strategies and levels of species
+diversification to generate more or less revenue and be more or less variable
+than others.
 

### 2016-12-06 portfolio/text/ms.Rmd
Add Eric's edits

@@
 harvest in geographical areas to buffer against disease outbreak, wildfires, and
-price variation REF. However, otherwise well-intentioned regulations can limit
-the actions available to individuals to reduce income variability, or
-incentivize against diversification. For example, in terrestrial agriculture,
-crop subsidies in the US may incentivize some farms to specialize (e.g. dairy,
-grains, soybeans; @mishra2002). In other instances, external market forces
-(prices, demand) may discourage the adoption of diverse strategies and help
-promote specialization [@bradshaw2004].
+price variation REF. However, external forces may restrict the options available
+to individuals to reduce variability. For example, otherwise well-intentioned
+regulations may limit how individuals diversify, or incentivize against
+diversification. For example, in terrestrial agriculture, crop subsidies in the
+US may incentivize some farms to specialize (e.g. dairy, grains, soybeans;
+@mishra2002). In other instances, external market forces (prices, demand) may
+discourage the adoption of diverse strategies and help promote specialization
+[@bradshaw2004].
 
@@
 extreme risk that some individuals experience, particularly relative to
-terrestrial agriculture [@sethi2012b]. The majority of insurance programs
-available to terrestrial farmers (e.g. crop insurance programs) are not
-available to commercial fishers [@kasperski2013] and both fish and fishers are
-spatially mobile. However, fishers still have the ability to lower risk. They
-can, for example, target multiple species as permitted by a license, own
-multiple licenses, or diversify spatially and fish in multiple regions. The
-effectiveness of these diversification strategies for improving revenue and
+terrestrial agriculture [@sethi2012b]. The majority of risk insurance programs
+available to terrestrial farmers (e.g. crop insurance programs) have not been
+obtainable by commercial fishers [@kasperski2013], possibly because both fish
+and fishers are spatially mobile. However, fishers still have the ability to
+lower risk. They can, for example, target as many species as permitted by
+a license, they may own multiple licenses, or diversify spatially and fish in
+multiple regions (increasing their total catch, or numbers of species caught).
+The effectiveness of these diversification strategies for improving revenue and
 reducing revenue variability remains uncertain. Furthermore, theory and
 empirical studies both suggest there can be substantial benefits to
-specialization (REFS). The majority of papers that examine effects of
-diversification on revenue do so at aggregate levels (e.g. households, counties,
-communities; REFS), rather than focus on individuals. Because of this, key
-uncertainties remain about the fundamental relationship between diversity,
-fishing revenue, and fishing revenue variability at the level of the individual.
+specialization for both farmers and fishers [@abdulai2000; @arnason2005]. The
+majority of papers that examine effects of diversification on revenue do so at
+aggregate levels that combine individuals (e.g. households, counties,
+communities; @sethi2014; @kasperski2013). Because of this, key uncertainties
+remain about the fundamental relationship between diversity, fishing revenue,
+and fishing revenue variability at the level of the individual.
 
@@
 understand the role of individuals' strategies in determining income and income
-variability for those whose livelihood depends on the extraction of natural
-resources. Fisheries in Alaska are among the highest-valued fisheries in the
+variability. Fisheries in Alaska are among the highest-valued fisheries in the
 world with 2.7 billion kg landed biomass valued at nearly 2 billion USD annually
@@
 many permits limit fishing to a single species, diversifying by purchasing
-additional fishing permits can be expensive (e.g. ~\$XX for king crab), and some
-fisheries limit the entry of new participants through individual fishing quotas
-(IFQs) (REF).
+additional fishing permits can be expensive (e.g. ~\$192K USD for a king and
+tanner crab permit in Southeast Alaska in 2016;
+<https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>), and some fisheries limit the
+entry of new participants through individual fishing quotas (IFQs) [@grimm2012].
 
@@
 vessels we can track people that switch fishery sectors over time, or alter
-strategies and use multiple vessels. We develop and apply a hierarchical
-Bayesian model to understand the effect of diversity on both individual revenue
-and revenue variability that builds on results from vessel [@kasperski2013],
-fleet [@sethi2012b], and community-level [@sethi2014] variability. Our model
-controls for fishing effort and latent time effects (e.g. changing environmental
+strategies and use multiple vessels. Similarly, we are able to disentagnle
+fishers that share the same vessel. We develop and apply a hierarchical Bayesian
+model to understand the effect of diversity on both individual revenue and
+revenue variability that builds on results from vessel [@kasperski2013], fleet
+[@sethi2012b], and community-level [@sethi2014] variability. Our model controls
+for fishing effort and latent time effects (e.g. changing environmental
 conditions), simultaneously estimates the effect of species diversity on
@@
 We developed a hierarchical variance-function regression model [@davidian1987;
-@hall1989], which separately models changes in the mean and variance of
+@hall1989], which separately but simultaneously models changes in the mean and variance of
 individual revenue as a function of covariates. Whereas traditional
@@
 
-We combined similar permits into permit groupings and identified sets of permit
-groupings held by an individual in a specific year as "strategies" (Supporting
-Information). For example, we combined permits that were the same except for
-vessel size (e.g. a longliner fishing sablefish on a vessel under 60 ft was
-considered to have the same strategy as a longliner fishing sablefish on
-a vessel over 60 ft), and we combined a number of permits that were targeting
-a single species and differed only in region fished. Therefore, for example,
-someone fishing herring roe in Kodiak and someone fishing herring roe in Alaska
-Peninsula were both considered to have a 'herring roe' permit group. For salmon
-permits, we grouped permits using the same gear type across regions for groups
-that shared similar species composition. After this process of combining
+We used combinations of state and federal fishing permits held by an individual
+in in a specific year to define fishing "strategies", representing the species
+targeted (Supporting Information). For example, we combined permits that were
+otherwise the same except for vessel size (e.g. a longliner fishing sablefish
+on a vessel under 60 ft was considered to have the same strategy as a longliner
+fishing sablefish on a vessel over 60 ft), and we combined a number of permits
+that were targeting a single species and differed only in region fished.
+Therefore, for example, someone fishing herring roe in Kodiak and someone
+fishing herring roe in Alaska Peninsula were both considered to have a 'herring
+roe' permit group. For salmon permits, we formed salmon strategies based on
+similarities in gear type and species compositon (Supporting Information).
+
+After this process of combining
 permits, we were left with `r n_grouped_permits` unique fishing permit groups
@@
 \noindent where $s$ indexes species from $1$ through $n$. We calculated effort
-or effective season length as the sum of the season lengths for each permit each
-year. Season length for a given permit was calculated as the number of days
-between the first instance of fishing a permit and the last instance of fishing
-a permit each year. For example, if someone fished for 30 days on a salmon drift
-gillnet permit in Southeast Alaska and 40 days on a statewide longline halibut
-permit, that individual's season length would be 70 days.
+or effective season length as the sum of the season lengths for each permit
+each year. Season length for a given permit was calculated as the number of
+days between the first instance of fishing a permit and the last instance of
+fishing a permit each year. For example, if someone fished for 30 days on a
+salmon drift gillnet permit in Southeast Alaska and 40 days on a statewide
+longline halibut permit, that individual's season length was calculated as 70
+days.
 
@@
 iterations of each chain as warm up (i.e. 4,000 final samples). We checked for
-chain convergence visually with trace plots. We also ensured that $\hat{R}<1.05$
-(the potential scale reduction factor) and that the effective sample size, as
-calculated by the rstan R package [@standevelopmentteam2015], was greater than
-200 for all parameters [@gelman2014]. We verified that our model returned
-sensible estimates by comparing our estimates to an ad-hoc two-stage model fit
-with the lme4 R package [@bates2015], and by fitting our model to simulated data
-and recovering the true values (Supporting Information). Code to reproduce our
-results (with a simulated data set) is available at
+chain convergence visually with trace plots. We also ensured that
+$\hat{R}<1.05$ (the potential scale reduction factor) and that the effective
+sample size, as calculated by the rstan R package [@standevelopmentteam2015],
+was greater than 200 for all parameters [@gelman2014]. We verified that our
+model returned sensible estimates by comparing our estimates to an ad-hoc
+two-stage model fit with the lme4 R package [@bates2015], where we first fit a
+mixed effect regression model to the mean, and secondly modeled the absolute
+residuals from that first model in a second mixed effects model representing
+the variance component. We also validated the model with simulated data, and
+were able to recover the true values (Supporting Information). Code to
+reproduce our results (with a simulated data set) is available at
 <https://github.com/NCEAS/pfx-commercial>.

### 2016-12-07 portfolio/text/ms.Rmd
Add remaining edits from Eric

@@
 on their extraction is difficult. These professions tend to have high
-year-to-year variability in income [@mishra2002; @sethi2012b]. In addition to
-economic risk, such as changes in demand, individuals dependent on natural
-resources are also subject to biological and environmental variability. For
-example, drought and flooding are a major source of risk to agricultural food
-security and farmers' incomes [@gregory2005], and catastrophic disease outbreaks
-and wildfires increase risk for the logging industry [@kirilenko2007].
+year-to-year variability in income relative to individuals who don't rely on
+natural resources as a source of income  [@mishra2002; @sethi2012b]. In
+addition to economic risk, such as changes in demand or prices, individuals
+dependent on natural resources are also subject to biological and environmental
+variability. For example, drought and flooding are a major source of risk to
+agricultural food security and farmers' incomes [@gregory2005], and
+catastrophic disease outbreaks and wildfires increase risk for the logging
+industry [@kirilenko2007].
 

### 2016-12-08 portfolio/text/supp.Rnw
Move dual fig 4 to supp

@@
 load("../data-generated/output-values.rda")
+h1 <- readRDS("../data-generated/h1_effect.rds")
+spoke <- readRDS("../data-generated/spoke_fraction.rds")
+g2_dens <- readRDS("../data-generated/g2_prob_dens_above_zero.rds")
 n_grouped_permits <- readRDS("../data-generated/n_grouped_permits.rds")
@@
 
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=1.1\textwidth]{../figs/stan-offset-break-anti-spaghetti.pdf}
+\caption{Effects of (A) generalizing and (B) specializing from year-to-year for
+individual permit holders \emph{within} the `r nstr` most common strategies.
+Changing species diversity from year-to-year is associated with higher
+variability for individuals within the strategies above the zero line.
+Generalizing (A) or specializing (B) is associated with greater expected revenue
+to the right of the zero line and lower expected revenue to the left of the zero
+line. These effects are estimated holding fishing season length constant from
+year-to-year. Dots and line segments represent medians and 50\% credible
+intervals of posteriors. }
+\label{fig:mirror}
+\end{center}
+\end{figure}
+
 % \begin{figure}[htbp]

### 2016-12-08 portfolio/analysis/3.3-plot-within-strategy-effects.R
Make new fig4

@@ -37,7 +37,7 @@ ns <- dat %>% group_by(strategy) %>%
 res <- inner_join(res, ns)
 
 p1 <- res %>%
-  mutate(strategy_label = ifelse(nn > 10, str_label, NA)) %>%
+  mutate(strategy_label = ifelse(nn > 250, str_label, NA)) %>%
   mutate(inc = inc) %>%
   ggplot(aes(x = inc_rev, y = inc)) +
   geom_hline(yintercept = 0, lty = 2, col = "grey65") +
@@ -51,7 +51,7 @@ p1 <- res %>%
     point.padding = unit(0.3, "lines"), max.iter = 6e3, segment.size = 0.3) +
   geom_point(aes(color = strategy_mean_div, size = nn)) +
   xlab("Effect of generalizing on revenue") +
-  ylab("Effect of generalizing on variability") +
+  ylab("Effect of generalizing on revenue variability") +
   scale_color_viridis() +
   theme_gg() +
   guides(
@@ -62,11 +62,12 @@ p1 <- res %>%
   labs(colour = "Mean sp.\ndiversity", size = "Number of\npermit holders") +
   theme(legend.title = element_text(size = rel(0.85)))
 
+ann_col <- "grey40"
 p2 <- res %>%
   mutate(dec = -dec, dec_rev = -dec_rev) %>%
   mutate(dec.l = -dec.l, dec_rev.l = -dec_rev.l) %>%
   mutate(dec.u = -dec.u, dec_rev.u = -dec_rev.u) %>%
-  mutate(strategy_label = ifelse(nn > 10, str_label, NA)) %>%
+  mutate(strategy_label = ifelse(nn > 250, str_label, NA)) %>%
   ggplot(aes(x = dec_rev, y = dec)) +
   geom_hline(yintercept = 0, lty = 2, col = "grey65") +
   geom_vline(xintercept = 0, lty = 2, col = "grey65") +
@@ -75,29 +76,40 @@ p2 <- res %>%
   geom_segment(aes(x = dec_rev, xend = dec_rev, y = dec.l, yend = dec.u),
     alpha = 0.1, lwd = 0.4) +
   geom_text_repel(aes(label = strategy_label, x = dec_rev, y = dec),
-    size = 2.8, colour = "grey60", #segment.color = "grey80",
+    size = 2.9, colour = "grey60", #segment.color = "grey80",
     point.padding = unit(0.3, "lines"), max.iter = 6e3, segment.size = 0.3) +
[…]

### 2016-12-08 portfolio/analysis/3.3-plot-within-strategy-effects-ifqs-temp.R
Cleanup

@@ -1,104 +1,0 @@
-library(dplyr)
-library(ggplot2)
-library(viridis)
-library(ggrepel)
-load("portfolio/data-generated/diff-dat-stan.rda")
-load("portfolio/data-generated/m_ifq.rda")
-devtools::load_all("pfxr")
-
-b <- broom::tidyMCMC(m_ifq, conf.int = T, estimate.method = "median",
-  conf.level = 0.5, conf.method = "quantile")
-
-g1 <- b[grepl("coef_g1", b$term), ]
-b1 <- b[grepl("coef_b1", b$term), ]
-g2 <- b[grepl("coef_g2", b$term), ]
-b2 <- b[grepl("coef_b1", b$term), ]
-
-res <- data.frame(strategy_ifq_id = 1:nrow(g1)) %>%
-  inner_join(md_ifq) %>%
-  mutate(strategy = as.character(strategy_ifq))
-res$inc <- g2$estimate
-res$dec <- g1$estimate
-res$inc_rev <- b2$estimate
-res$dec_rev <- b1$estimate
-
-res$inc.l <- g2$conf.low
-res$dec.l <- g1$conf.low
-res$inc_rev.l <- b2$conf.low
-res$dec_rev.l <- b1$conf.low
-
-res$inc.u <- g2$conf.high
-res$dec.u <- g1$conf.high
-res$inc_rev.u <- b2$conf.high
-res$dec_rev.u <- b1$conf.high
-
-ns <- dat %>% group_by(strategy_ifq) %>%
-  summarise(nn = length(unique(p_holder))) %>%
-  rename(strategy = strategy_ifq)
-res <- inner_join(res, ns)
-
-p1 <- res %>%
-  mutate(strategy_label = ifelse(nn > 30, strategy_ifq, NA)) %>%
-  mutate(inc = inc) %>%
-  ggplot(aes(x = inc_rev, y = inc)) +
-  geom_hline(yintercept = 0, lty = 2, col = "grey65") +
-  geom_vline(xintercept = 0, lty = 2, col = "grey65") +
-  geom_segment(aes(x = inc_rev.l, xend = inc_rev.u, y = inc, yend = inc),
-    alpha = 0.1, lwd = 0.4) +
-  geom_segment(aes(x = inc_rev, xend = inc_rev, y = inc.l, yend = inc.u),
-    alpha = 0.1, lwd = 0.4) +
[…]

### 2016-12-08 portfolio/text/ms.Rmd
Edit

@@
 Managing natural resources and sustaining the livelihoods of those that depend
-on their extraction is difficult. These professions tend to have high
-year-to-year variability in income relative to individuals who don't rely on
-natural resources as a source of income  [@mishra2002; @sethi2012b]. In
-addition to economic risk, such as changes in demand or prices, individuals
+on their extraction is hard. These professions tend to have high year-to-year
+variability in income relative to other professions [@mishra2002; @sethi2012b].
+In addition to economic risk, such as changes in demand or prices, individuals
 dependent on natural resources are also subject to biological and environmental
-variability. For example, drought and flooding are a major source of risk to
-agricultural food security and farmers' incomes [@gregory2005], and
-catastrophic disease outbreaks and wildfires increase risk for the logging
-industry [@kirilenko2007].
+variability. For example, drought and flooding are a major source of risk for
+agricultural food security and farmers' incomes [@gregory2005], and catastrophic
+disease outbreaks and wildfires increase risk for the logging industry
+[@kirilenko2007].
 
@@
 harvest in geographical areas to buffer against disease outbreak, wildfires, and
-price variation REF. However, external forces may restrict the options available
-to individuals to reduce variability. For example, otherwise well-intentioned
-regulations may limit how individuals diversify, or incentivize against
-diversification. For example, in terrestrial agriculture, crop subsidies in the
-US may incentivize some farms to specialize (e.g. dairy, grains, soybeans;
-@mishra2002). In other instances, external market forces (prices, demand) may
-discourage the adoption of diverse strategies and help promote specialization
-[@bradshaw2004].
+price variation (REF). However, external forces may restrict the available
+options to dampen income variability. Otherwise well-intentioned regulations may
+limit how individuals diversify, or incentivize against diversification. For
+example, in terrestrial agriculture, crop subsidies in the US may incentivize
+some farms to specialize (e.g. dairy, grains, soybeans; @mishra2002). In other
+instances, external market forces (e.g. prices or demand) may discourage the
+adoption of diverse strategies and help promote specialization [@bradshaw2004].
 
@@
 terrestrial agriculture [@sethi2012b]. The majority of risk insurance programs
-available to terrestrial farmers (e.g. crop insurance programs) have not been
+available to terrestrial farmers (e.g. crop insurance programs) are not
 obtainable by commercial fishers [@kasperski2013], possibly because both fish
@@
 lower risk. They can, for example, target as many species as permitted by
-a license, they may own multiple licenses, or diversify spatially and fish in
-multiple regions (increasing their total catch, or numbers of species caught).
-The effectiveness of these diversification strategies for improving revenue and
-reducing revenue variability remains uncertain. Furthermore, theory and
-empirical studies both suggest there can be substantial benefits to
-specialization for both farmers and fishers [@abdulai2000; @arnason2005]. The
-majority of papers that examine effects of diversification on revenue do so at
-aggregate levels that combine individuals (e.g. households, counties,
-communities; @sethi2014; @kasperski2013). Because of this, key uncertainties
+a license, own multiple licenses, or diversify spatially and fish in multiple
+regions.
+
+The effectiveness of these diversification strategies for increasing revenue and
+reducing revenue variability remains uncertain. For example, theory and
+empirical studies suggest there can be substantial benefits to specialization
+for both farmers and fishers [@abdulai2000; @arnason2005]. Furthermore, the
+majority of research examining the effects of diversification on revenue do so
+at aggregate levels that combine individuals (e.g. households, counties, or
+communities; [@kasperski2013; @sethi2014]). Because of this, key uncertainties
 remain about the fundamental relationship between diversity, fishing revenue,
@@
 variability. Fisheries in Alaska are among the highest-valued fisheries in the
-world with 2.7 billion kg landed biomass valued at nearly 2 billion USD annually
-[@nmfs2016]. The industry supports a range of fishing fleets ranging from small,
-open, single-operator vessels to multi-million-dollar state-of-the-art ships
-with large crews. Like much of the rest of the world [@hilborn2003a], the days
-of open, unregulated fisheries in most developed countries have passed ---
-fishers must own a license to participate in virtually all fisheries. And while
-such regulations have proven helpful in reducing or eliminating overfishing,
-they limit the available strategies to reduce income variability. For example,
-many permits limit fishing to a single species, diversifying by purchasing
-additional fishing permits can be expensive (e.g. ~\$192K USD for a king and
-tanner crab permit in Southeast Alaska in 2016;
-<https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>), and some fisheries limit the
-entry of new participants through individual fishing quotas (IFQs) [@grimm2012].
+world with ~2.7 billion kg landed biomass valued at nearly 2 billion USD
+annually [@nmfs2016]. Like much of the rest of the world, the days of open,
+unregulated fisheries in most developed countries have passed --- fishers must
+own a license to participate in virtually all fisheries [@hilborn2003a]. And
+while such regulations have helped reduce overfishing, they limit the available
+strategies to reduce income variability. For example, many permits limit fishing
+to a single species, diversifying by purchasing additional fishing permits can
+be expensive (e.g. ~\$192K USD for a king and tanner crab permit in Southeast
+Alaska in 2016; <https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>), and some
+fisheries limit the entry of new participants through individual fishing quotas
+(IFQs) [@grimm2012].
 
@@
 individuals is affected by diversification. By tracking individuals rather than
-vessels we can track people that switch fishery sectors over time, or alter
-strategies and use multiple vessels. Similarly, we are able to disentagnle
-fishers that share the same vessel. We develop and apply a hierarchical Bayesian
-model to understand the effect of diversity on both individual revenue and
-revenue variability that builds on results from vessel [@kasperski2013], fleet
-[@sethi2012b], and community-level [@sethi2014] variability. Our model controls
-for fishing effort and latent time effects (e.g. changing environmental
-conditions), simultaneously estimates the effect of species diversity on
-individual revenue and revenue variability, and estimates effects both within
-and across fishing strategies (sets of similar licenses or permits). To our
-knowledge, this is the first application of this type of model in any natural
-resource field.
+vessels we can track people that switch fishery permits over time or use
+multiple vessels. Similarly, we are able to disentangle fishers that share the
+same vessel. We develop and apply a hierarchical Bayesian model to understand
+the effect of diversity on both individual revenue and revenue variability that
+builds on results from vessel [@kasperski2013], fleet [@sethi2012b], and
+community-level [@sethi2014] variability. Our model controls for fishing effort
+and latent time effects (e.g. changing environmental conditions), simultaneously
+estimates the effect of species diversity on individual revenue and revenue
+variability, and estimates effects both within and across fishing strategies
+(sets of similar licenses or permits). To our knowledge, this is the first
+application of this type of model in any natural resource field.
 
-\clearpage
-
-# Methods
-
-We developed a hierarchical variance-function regression model [@davidian1987;
-@hall1989], which separately but simultaneously models changes in the mean and variance of
-individual revenue as a function of covariates. Whereas traditional
-hierarchical linear models assume homoscedasticity --- that the residual
-variance does not vary systematically --- the variance-function regression model
-allows the variance to be heteroscedastic and vary as a function of predictors.
-Our specific model allows certain fishing strategies and levels of species
-diversification to generate more or less revenue and be more or less variable
-than others.
-
-## Data
-
-We obtained commercial fisheries landings and revenue data for all permit
-holders in Alaska from `r range_yrs[1]` to `r range_yrs[2]` from the Alaskan
-Commercial Fisheries Entry Commission (CFEC). These data represent a total of
-nearly `r round(ndat$initial_tot_rev/1e9)` billion 2009 USD and a combined `r 
-round(ndat$initial/1e6,1)` million rows of data. We adjusted for inflation by
-setting all revenues to 2009 USD [@bea2016]. We implemented a number of
-filtering steps to focus on individuals actively engaged in commercial fishing
-(Supporting Information). In particular, we removed permit holders whose median
-revenue was less than \$10,000 2009 USD to focus on individuals for whom
-commercial fishing formed a substantial part of their income.
-
-We used combinations of state and federal fishing permits held by an individual
-in in a specific year to define fishing "strategies", representing the species
-targeted (Supporting Information). For example, we combined permits that were
-otherwise the same except for vessel size (e.g. a longliner fishing sablefish
-on a vessel under 60 ft was considered to have the same strategy as a longliner
-fishing sablefish on a vessel over 60 ft), and we combined a number of permits
-that were targeting a single species and differed only in region fished.
-Therefore, for example, someone fishing herring roe in Kodiak and someone
-fishing herring roe in Alaska Peninsula were both considered to have a 'herring
-roe' permit group. For salmon permits, we formed salmon strategies based on
-similarities in gear type and species compositon (Supporting Information).
-
-After this process of combining
-permits, we were left with `r n_grouped_permits` unique fishing permit groups
-(Table S\ref{tab:strategies}), which we aggregated within each person-year
-combination to form permit strategies. For example, if someone fished halibut
-and sablefish in the same year, their strategy for that year would be
-"halibut-sablefish". Specific details on our strategy definitions are available
-in the Supporting Information.
-
-Our models focused on species diversity as a possible predictor of revenue and
-revenue variability while controlling for effort. For consistency with previous
-analyses [@kasperski2013], we calculated effective species diversity for
-permit holder $i$ and time (year) $t$ as the inverse of Simpson's diversity
-index $\lambda$ [@simpson1949; @jost2006] (also referred to in economics as the
-Herfindahl-Hirschman Index) weighted by revenue $R$:
-
-\begin{equation}
-1/\lambda_{i,t} = 1 / \left(\sum\limits_{s=1}^{n_{i,t}}
-  \left(R_{i,s,t}/ \sum\limits_{s=1}^{n_{i,t}} R_{i,s,t}\right)\right)^2,
-\end{equation}
-
-\noindent where $s$ indexes species from $1$ through $n$. We calculated effort
-or effective season length as the sum of the season lengths for each permit
-each year. Season length for a given permit was calculated as the number of
-days between the first instance of fishing a permit and the last instance of
-fishing a permit each year. For example, if someone fished for 30 days on a
-salmon drift gillnet permit in Southeast Alaska and 40 days on a statewide
-longline halibut permit, that individual's season length was calculated as 70
-days.
-
-## Hierarchical model
-
-To jointly model revenue and revenue variability as a function of
-individual-level covariates, we extended the basic variance-function regression
-to a hierarchical Bayesian variance-function regression model with Gaussian
-errors predicting $\log$ revenue, $\log(R_{i,t})$, for each fisher $i$ and time
-$t$:
-
-\begin{equation}
-\log(R_{i,t}) \sim \mathrm{N} \left(\widehat{\log(R_{i,t})}, \widehat{\sigma_{i,t}^2} \right).
-\end{equation}
-
-\noindent Each year $t$ of revenue for fisher $i$ is assigned a strategy indexed
-by $j$ (Supporting Information, Table \ref{tab:strategies}) and expected
-$\log(R_{i,t})$ is modelled as:
-
-<!-- Ole: I might swap out log(R)-hat for something like mu(or your favorit greek letter) just to avoid -->
-<!-- potential confusion like using the term "expected log(R)".   -->
-<!-- I also just don't like hats... but that's my problem. -->
-
-\begin{equation}
-\label{eq:mean}
-\widehat{\log(R_{i,t})} = \beta_{0,j,t} -
-  \beta_{1,j} S_{i,t} I_{i,t} +
-  \beta_{2,j} S_{i,t} I_{i,t} +
-  \beta_3 D_{i,t} -
-  \beta_4 S_{i,t} D_{i,t} I_{i,t} +
-  \beta_5 S_{i,t} D_{i,t} I_{i,t} +
-  \log(R_{i,t-1}),
-\end{equation}
-
-\noindent where $S_{i,t}$ is the percent change in species diversity, $\log
-\left(\mathrm{div}_{i,t}/\mathrm{div}_{i,t-1}\right)$, $D_{i,t}$ is the percent
-change in days fished, $\log
-\left(\mathrm{days}_{i,t}/\mathrm{days}_{i,t-1}\right)$, and $I_{i,t}$ is an
-indicator variable that takes a value $0$ if $\mathrm{div}_{i,t} \ge
-\mathrm{div}_{i,t-1}$ (species diversity increasing; generalizing) and takes
-a value $1$ if $\mathrm{div}_{i,t} < \mathrm{div}_{i,t-1}$ (species diversity
-decreasing; specializing). Estimating a separate coefficient for the effect of
-increasing diversity and decreasing diversity allows the costs and benefits of
-these changes to be asymmetric.
-
-We use $\log(R_{i,t-1}$) as an offset, and thus our model describes the percent
-change in revenue, i.e., $\log(R_{i,t}) - \log(R_{i,t-1})
-= \log(R_{i,t}/R_{i,t-1})$. We model change in revenue because preliminary
-analysis revealed non-stationary trends in revenue (resulting from experience or
-other unmodeled factors). This approach eliminates having to deal with the
-non-stationary time series and eliminates having to estimate thousands of
-mean-revenue intercepts associated with individual fishers. We removed
-person-year combinations without a subsequent year of revenue.
-
-The $\beta$ terms represent estimated coefficients, some of which are allowed to
-vary as "random effect" terms. The intercepts $\beta_{0,j,t}$ are allowed to
-vary by strategy-year combinations, are centered on zero, and are constrained by
-a normal distribution:
-
-<!-- Ole: I'd use tau (or something) for the variances here so that people don't have to look at the subscripts to differentiate between variances at different hierarchical levels. -->
-
-\begin{equation}
-\beta_{0,j,t} \sim \mathrm{N} \left(0, \sigma^2_{\beta_{j,t}} \right).
-\end{equation}
-
-\noindent This approach, combined with the lack of a global intercept,
-constrains the mean change in log revenue across all fishers and years to be
-zero. I.e., we assume that the data are stationary after accounting for all
-covariates and including an intercept that varies with strategy-year
-combinations. Initial versions of the model confirmed that the global intercept
-along with strategy-level intercepts would be estimated at almost exactly zero.
-
-The slopes $\beta_{1,j}$ and $\beta_{1,j}$ are then allowed to vary by strategy
-with estimated means:
-
-\begin{align}
-\beta_{1,j} &\sim \mathrm{N} \left(\mu_{\beta_1}, \sigma^2_{\beta_1}\right),\\
-\beta_{2,j} &\sim \mathrm{N} \left(\mu_{\beta_2}, \sigma^2_{\beta_2}\right).
-\end{align}
-
-\noindent The $\beta_{1,j}$'s represent how much a 1\% decrease in species
-diversity within a strategy translates to a Y\% change in revenue. The
-$\beta_{2,j}$'s represent how much a 1\% increase in species diversity within
-a strategy translates to a Y\% change in revenue. These are after controlling
-for percent change in days fished, $\beta_3$, and the interaction between days
-fished and percent change in species diversity, $\beta_4$ and $\beta_5$.
-
-Whereas a typical hierarchical linear regression model would estimate a single
-residual error scale term, $\sigma$, we model the scale of the residual error
-with another hierarchical model with a similar form:
-
-\begin{equation}
-\label{eq:cv}
-\widehat{\sigma_{i,t}} =
-\exp \left( \gamma_{0,j} -
-\gamma_{1,j} S_{i,t} I_{i,t} +
-\gamma_{2,j} S_{i,t} I_{i,t} +
-\gamma_3 D_{i,t} -
-\gamma_4 S_{i,t} D_{i,t} I_{i,t} +
-\gamma_5 S_{i,t} D_{i,t} I_{i,t}\right).
-\end{equation}
-
-\noindent Again, $S_{i,t}$ and $I_{i,t}$ represent, respectively, percent change
-in species diversity and an indicator variable for increasing or decreasing
-species diversity. We exponentiate the equation to ensure all scale parameters,
-$\sigma_{i,t}$, are positive. The $\gamma$'s are estimated coefficients, three
-of which are allowed to vary by strategy, constrained by normal distributions:
-
-\begin{align}
-\gamma_{0,j}  &\sim \mathrm{N} \left(\eta_0 + \eta_1 M_j + \eta_2 E_j, \sigma^2_{\gamma_0}\right),\\
-\gamma_{1,j}  &\sim \mathrm{N} \left(\mu_{\gamma_1}, \sigma^2_{\gamma_1}\right),\\
-\gamma_{2,j}  &\sim \mathrm{N} \left(\mu_{\gamma_2}, \sigma^2_{\gamma_2}\right).
-\end{align}
-
-\noindent The mean of the $\gamma_{0,j}$ coefficients is modelled with two
-strategy-level predictors. The symbols $M_j$ and $E_j$ represent the raw mean
-species diversity and mean combined fishing season length (effort) for strategy
-$j$. The coefficients $\eta_1$ and $\eta_2$ represent predictors in this
-strategy-level regression estimating effects across strategies.
-
-We fit our models with Stan 2.11.1 [@hoffman2014; @standevelopmentteam2015;
-@standevelopmentteam2015a; @carpenter2015] and R 3.3.1 [@rcoreteam2016]. Stan
-implements the No-U-Turn Hamiltonian Markov chain Monte Carlo algorithm
-[@hoffman2014] to perform Bayesian statistical inference. We assigned weakly
-informative priors on all parameters: $N(0, 2^2)$ priors on all $\beta$,
-$\gamma$, and $\eta$ parameters, and half-$t(3, 0, 2)$ priors (i.e. degrees of
-freedom of 3 and scale of 2) on all $\sigma$ parameters to allow for heavier
-tails. We ran four chains and 2,000 iterations and discarded the first 1,000
-iterations of each chain as warm up (i.e. 4,000 final samples). We checked for
-chain convergence visually with trace plots. We also ensured that
-$\hat{R}<1.05$ (the potential scale reduction factor) and that the effective
-sample size, as calculated by the rstan R package [@standevelopmentteam2015],
-was greater than 200 for all parameters [@gelman2014]. We verified that our
-model returned sensible estimates by comparing our estimates to an ad-hoc
-two-stage model fit with the lme4 R package [@bates2015], where we first fit a
-mixed effect regression model to the mean, and secondly modeled the absolute
-residuals from that first model in a second mixed effects model representing
-the variance component. We also validated the model with simulated data, and
-were able to recover the true values (Supporting Information). Code to
-reproduce our results (with a simulated data set) is available at
-<https://github.com/NCEAS/pfx-commercial>.
-
-*TODO: for the following into the results section or and the method section with
-this:*
-
-Residuals from the mean component of our model clearly show the "V"-shaped
-pattern that the variance component of our model seeks to capture (Fig.
-S\ref{fig:resid}). Furthermore, the downside residuals (year-individual data
-points with lower than expected revenue) look qualitatively similar to the
-complete set of residuals (upside and downside) (Fig.
-S\ref{fig:resid-downside}). Therefore, it seemed appropriate to model the
-variability as a proxy for revenue risk.
-
-The estimated year-permit intercepts represent years when people with
-a particular permit set in a given year tended to do better or worse than the
-long-term average for that permit set (Fig. S\ref{fig:str-yrs}). Therefore, an
-individual's expected revenue and variability is after accounting for the
-general trend in revenue for that permit set.
-
-Our model estimates overall average effects (Fig. S\ref{fig:fe}) and effects
-that vary by permit set (Fig. S\ref{fig:re}).
+Our hierarchical variance-function regression model [@davidian1987; @hall1989]
+simultaneously models changes in the mean and variance of individual revenue as
+a function of covariates. Whereas traditional hierarchical linear models assume
+homoscedasticity --- that the residual variance does not vary systematically ---
+the variance-function regression model allows the variance to be heteroscedastic
+and vary as a function of predictors. Our specific model allows the revenue rate
+of return (log of the ratio of year-to-year revenue) and the revenue variability
+(the standard deviation of this revenue rate of return) to vary by fishing
+strategy (fishing permit sets), fishing effort (season length), and levels of
+species diversification.
 
@@
 
-We evaluated trends in revenue and diversification over time for individuals
-fishing commercially in Alaska from 1985--2014 (Fig. \ref{fig:ts}A). The
-majority (70--90%) of fishers hold single permits, with a slight increase in
+We first evaluated trends in revenue and diversification over time for
+individuals fishing commercially in Alaska from 1985--2014 (Fig. \ref{fig:ts}A).
+The majority (70--90%) of fishers hold single permits, with a slight increase in
 single-permit holders since 2000 (Fig. \ref{fig:ts}C). This overall trend is
@@
 Adding permits also usually increased expected revenue, although the permits
-themselves represent a significant expense (e.g. TODO).
+themselves represent a significant expense (e.g. ~\$192K USD for a king and
+tanner crab permit in Southeast Alaska in 2016;
+<https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>).
 
@@
 individual *within* the constraints of a given permit set is more variable (Fig.
-\ref{fig:within}). After accounting for fishing effort (Fig. S\ref{fig:fe}),
-specializing (decreasing species diversity) tends to be associated with
-increased revenue (~0.5% increased revenue per 1% decrease in species diversity
-on average or TODO ... `r g2_dens`% probability; Fig. \ref{fig:within}). This is
-particularly clear for permits where individuals tend to specialize on sockeye
-salmon. For example, individuals with setnet permits for salmon in Prince
-William Sound, Cook Inlet, Kodiak, Bristol Bay, or the Alaska Peninsula target
-primarily sockeye salmon and experience a ~1.5% increase in revenue for a 1%
-decrease in fished species diversity (Fig. \ref{fig:within}B). The converse is
-also usually true --- generalizing tends to be associated with lower expected
-revenue after accounting for fishing effort (i.e. Fig. \ref{fig:within}A and
-\ref{fig:within}B are near mirror images left to right). This benefit of
-specializing is only apparent after accounting for fishing effort. Given that
-fishing for more species tends to be associated with increased fishing effort
-(fishing more days), generalizing tends to be associated with higher revenue
-when not standardizing for days finished (Fig. S-TODO).
+\ref{fig:within}). After accounting for fishing effort, specializing tends to be
+associated with increased revenue (Fig. \ref{fig:within}; Fig. S\ref{fig:fe}).
+On average, there is a ~0.5% increase in the year-to-year revenue ratio per 1%
+decrease in species diversity, and across strategies, `r g2_dens`% of the
+probability density supports a positive relationship between specialization and
+revenue. The converse is also usually true --- generalizing tends to be
+associated with lower expected revenue after accounting for fishing effort (i.e.
+Fig. S\ref{fig:mirror}A and \ref{fig:mirror}B are near mirror images left to
+right). This benefit of specializing is only apparent after accounting for
+fishing effort. Given that fishing for more species tends to be associated with
+increased fishing effort (fishing more days), generalizing tends to be
+associated with higher revenue when not standardizing for days finished (Fig.
+S-TODO).
 
 After accounting for fishing effort, adjusting (increasing or decreasing) an
-individual's fished species diversity from year-to-year within the constraints
-of a permit set tends to be associated with greater expected variability or
-uncertainty in year-to-year revenue (Fig. \ref{fig:within} strategies above 0;
-Fig. S\ref{fig:fe}). This effect was strongest for individuals with a halibut
-longline permit or halibut and sablefish longline permits. Individuals with
-these strategies experienced ~2% increase in standardized variability for a 1%
-change in fished species diversity (Fig. \ref{fig:within}). Levels of fished
-species diversity levels above one for these permits are primarily due to
-bycatch.
+individual's fished species diversity within the constraints of a permit set
+tends to be associated with greater expected variability or uncertainty in
+year-to-year revenue (Fig. \ref{fig:within} strategies above 0; Fig.
+S\ref{fig:fe}). This effect was strongest for individuals with a halibut permit
+or halibut and sablefish permits. Individuals with these strategies experienced
+~2% increase in variability for a 1% change in the year-to-year ratio of species
+diversity (Fig. \ref{fig:within}). Levels of fished species diversity above one
+for these particular permits are primarily due to bycatch.
 
-The Alaskan halibut and sablefish fisheries were converted to an individual
-fishery quota (IFQ) system in 1994. This catch share system... TODO. We refit
-our models allowing all permit-level effects to vary separately before and after
-the introduction of IFQs for strategies involving halibut and sablefish permits.
-The introduction of IFQs reduced variability by ~25% (XX--XX 95% CI) for
-individuals with only halibut or sablefish permits (Fig. S\ref{fig:ifq}A).
+The management of the Alaskan halibut and sablefish fisheries was converted to
+individual fishery quotas (IFQ) in 1994. This catch share system... TODO. We
+refit our models allowing all permit-level effects to vary separately before and
+after the introduction of IFQs for strategies involving halibut and sablefish
+permits. The introduction of IFQs reduced variability by ~25% (XX--XX 95% CI)
+for individuals with only halibut or sablefish permits (Fig. S\ref{fig:ifq}A).
 Furthermore, for most strategies, the effect of increased revenue variability
@@
 
-\caption{ Trends in revenue and diversification over time. (A) Map of fishing
+\caption{Trends in revenue and diversification over time. (A) Map of fishing
 regions in Alaska, USA. (B) Total revenue in billions of 2009 U.S. dollars from
@@
 individual halibut and (B) gillnet herring roe permits (dark blue) compared with
-owners of the same permit plus one more (medium blue) or two more permits (light
-blue). On the y-axis, dots and lines represent posterior median and 95\%
-credible intervals of expected standard deviation.}
+owners of the same permit plus one more (medium blue) or two more (light
+blue) permits. On the y-axis, dots and lines represent posterior median and 95\%
+credible intervals of expected standard deviations of year-to-year revenue rates
+of return.}
 
@@
 \begin{center}
-\includegraphics[width=1.1\textwidth]{../figs/stan-offset-break-anti-spaghetti.pdf}
+\includegraphics[width=0.6\textwidth]{../figs/stan-offset-break-anti-spaghetti-specializing.pdf}
 
-\caption{Effects of (A) generalizing and (B) specializing from year-to-year for
-individual permit holders \emph{within} the `r nstr` most common strategies.
-Changing species diversity from year-to-year is associated with higher
-variability for individuals within the strategies above the zero line.
-Generalizing (A) or specializing (B) is associated with greater expected revenue
-to the right of the zero line and lower expected revenue to the left of the zero
-line. These effects are estimated holding fishing season length constant from
-year-to-year. Dots and line segments represent medians and 50\% credible
-intervals of posteriors. }
+\caption{Effect of specializing on year-to-year changes in revenue and revenue
+variability for individual permit holders \emph{within} the `r nstr` most common
+strategies. Changing species diversity from year-to-year is associated with
+higher variability for individuals within the strategies above the zero line.
+Specializing is associated with greater expected revenue to the right of the
+zero line and lower expected revenue to the left of the zero line. For example,
+individuals with setnet permits for salmon in Prince William Sound, Cook Inlet,
+Kodiak, Bristol Bay, or the Alaska Peninsula target primarily sockeye salmon and
+experience a $\sim$1.5\% increase in their year-to-year revenue ratio for a 1\%
+decrease in their year-to-year species diversity. These effects are estimated
+holding fishing season length constant from year-to-year. Dots and line segments
+represent medians and 50\% credible intervals of posteriors. The effect of
+generalizing (increasing species diversity) is similar (Fig. S\ref{fig:mirror})}.
 
@@
 \clearpage
+
+# Materials and methods
+
+We obtained commercial fisheries landings and revenue data for all permit
+holders in Alaska from `r range_yrs[1]` to `r range_yrs[2]` from the Alaskan
+Commercial Fisheries Entry Commission (CFEC). These data represent a total of
+nearly `r round(ndat$initial_tot_rev/1e9)` billion 2009 USD and a combined `r
+round(ndat$initial/1e6,1)` million rows of data. We adjusted for inflation by
+setting all revenues to 2009 USD [@bea2016]. We implemented a number of
+filtering steps to focus on individuals actively engaged in commercial fishing
+(Supporting Information). In particular, we removed permit holders whose median
+revenue was less than \$10,000 2009 USD to focus on individuals for whom
+commercial fishing formed a substantial part of their income.
+
+We used combinations of state and federal fishing permits held by an individual
+in in a specific year to define fishing "strategies", representing the species
+targeted (Supporting Information). For example, we combined permits that were
+otherwise the same except for vessel size (e.g. a longliner fishing sablefish
+on a vessel under 60 ft was considered to have the same strategy as a longliner
+fishing sablefish on a vessel over 60 ft), and we combined a number of permits
+that were targeting a single species and differed only in region fished.
+Therefore, for example, someone fishing herring roe in Kodiak and someone
+fishing herring roe in Alaska Peninsula were both considered to have a "herring
+roe" permit group. For salmon permits, we formed salmon strategies based on
+similarities in gear type and species compositon (Supporting Information).
+
+After this process of combining permits, we were left with `r n_grouped_permits`
+unique fishing permit groups (Table S\ref{tab:strategies}), which we aggregated
+within each person-year combination to form permit strategies. For example, if
+someone fished halibut and sablefish in the same year, their strategy for that
+year would be "halibut-sablefish". Specific details on our strategy definitions
+are available in the Supporting Information.
+
+Our models focused on species diversity as a possible predictor of revenue and
+revenue variability while controlling for effort. For consistency with previous
+analyses [@kasperski2013], we calculated effective species diversity for
+permit holder $i$ and time (year) $t$ as the inverse of Simpson's diversity
+index $\lambda$ [@simpson1949; @jost2006] (also referred to in economics as the
+Herfindahl-Hirschman Index) weighted by revenue $R$:
+
+\begin{equation}
+1/\lambda_{i,t} = 1 / \left(\sum\limits_{s=1}^{n_{i,t}}
+  \left(R_{i,s,t}/ \sum\limits_{s=1}^{n_{i,t}} R_{i,s,t}\right)\right)^2,
+\end{equation}
+
+\noindent where $s$ indexes species from $1$ through $n$. We calculated effort
+or effective season length as the sum of the season lengths for each permit
+each year. Season length for a given permit was calculated as the number of
+days between the first instance of fishing a permit and the last instance of
+fishing a permit each year. For example, if someone fished for 30 days on a
+salmon drift gillnet permit in Southeast Alaska and 40 days on a statewide
+longline halibut permit, that individual's season length was calculated as 70
+days.
+
+## Hierarchical model
+
+To jointly model revenue and revenue variability as a function of
+individual-level covariates, we extended the basic variance-function regression
+to a hierarchical Bayesian variance-function regression model with Gaussian
+errors predicting $\log$ revenue, $\log(R_{i,t})$, for each fisher $i$ and time
+$t$:
+
+\begin{equation}
+\log(R_{i,t}) \sim \mathrm{N} \left(\mu_{i,t}, \sigma_{i,t}^2 \right).
+\end{equation}
+
+\noindent Each year $t$ of revenue for fisher $i$ is assigned a strategy (permit
+set) indexed by $j$ (Supporting Information, Table \ref{tab:strategies}) and
+$\mu_{i,t}$ is modelled as:
+
+\begin{equation}
+\label{eq:mean}
+\mu_{i,t} = \beta_{0,j,t} -
+  \beta_{1,j} S_{i,t} I_{i,t} +
+  \beta_{2,j} S_{i,t} I_{i,t} +
+  \beta_3 D_{i,t} -
+  \beta_4 S_{i,t} D_{i,t} I_{i,t} +
+  \beta_5 S_{i,t} D_{i,t} I_{i,t} +
+  \log(R_{i,t-1}),
+\end{equation}
+
+\noindent where $S_{i,t}$ is the log ratio of species diversity from
+year-to-year, $\log \left(\mathrm{div}_{i,t}/\mathrm{div}_{i,t-1}\right)$,
+$D_{i,t}$ is the log ratio of days fished from year-to-year, $\log
+\left(\mathrm{days}_{i,t}/\mathrm{days}_{i,t-1}\right)$, and $I_{i,t}$ is an
+indicator variable that takes a value $0$ if $\mathrm{div}_{i,t} \ge
+\mathrm{div}_{i,t-1}$ (species diversity increasing; generalizing) and takes
+a value $1$ if $\mathrm{div}_{i,t} < \mathrm{div}_{i,t-1}$ (species diversity
+decreasing; specializing). Estimating a separate coefficient for the effect of
+increasing diversity and decreasing diversity allows the costs and benefits of
+these changes to be asymmetric.
+
+We use $\log(R_{i,t-1}$) as an offset, and thus our model describes the log
+ratio of revenue from year-to-year, i.e., $\log(R_{i,t}) - \log(R_{i,t-1})
+= \log(R_{i,t}/R_{i,t-1})$. We model change in revenue because preliminary
+analysis revealed non-stationary trends in revenue (resulting from experience or
+other unmodeled factors). This approach eliminates having to deal with the
+non-stationary time series and eliminates having to estimate thousands of
+mean-revenue intercepts associated with individual fishers. We removed
+person-year combinations without a subsequent year of revenue.
+
+The $\beta$ terms represent estimated coefficients, some of which are allowed to
+vary as "random effect" terms. The intercepts $\beta_{0,j,t}$ are allowed to
+vary by strategy-year combinations, are centred on zero, and are constrained by
+a normal distribution:
+
+<!-- Ole: I'd use tau (or something) for the variances here so that people don't have to look at the subscripts to differentiate between variances at different hierarchical levels. -->
+
+\begin{equation}
+\beta_{0,j,t} \sim \mathrm{N} \left(0, \sigma^2_{\beta_{j,t}} \right).
+\end{equation}
+
+\noindent This approach, combined with the lack of a global intercept,
+constrains the mean change in log revenue across all fishers and years to be
+zero. I.e., we assume that the data are stationary after accounting for all
+covariates and including an intercept that varies with strategy-year
+combinations. Initial versions of the model confirmed that the global intercept
+along with strategy-level intercepts would be estimated at almost exactly zero.
+
+The slopes $\beta_{1,j}$ and $\beta_{1,j}$ are then allowed to vary by strategy
+with estimated means:
+
+\begin{align}
+\beta_{1,j} &\sim \mathrm{N} \left(\mu_{\beta_1}, \tau^2_{\beta_1}\right),\\
+\beta_{2,j} &\sim \mathrm{N} \left(\mu_{\beta_2}, \tau^2_{\beta_2}\right).
+\end{align}
+
+\noindent The $\beta_{1,j}$'s represent how much a 1\% decrease in species
+diversity within a strategy translates to a Y\% change in revenue. The
+$\beta_{2,j}$'s represent how much a 1\% increase in species diversity within
+a strategy translates to a Y\% change in revenue. These are after controlling
+for changes in effort, $\beta_3$, and the interaction between changes in effort
+and changes in species diversity, $\beta_4$ and $\beta_5$.
+
+Whereas a typical hierarchical linear regression model would estimate a single
+residual error scale term, $\sigma$, we model the scale of the residual error
+with another hierarchical model with a similar form:
+
+\begin{equation}
+\label{eq:cv}
+\sigma_{i,t} =
+\exp \left( \gamma_{0,j} -
+\gamma_{1,j} S_{i,t} I_{i,t} +
+\gamma_{2,j} S_{i,t} I_{i,t} +
+\gamma_3 D_{i,t} -
+\gamma_4 S_{i,t} D_{i,t} I_{i,t} +
+\gamma_5 S_{i,t} D_{i,t} I_{i,t}\right).
+\end{equation}
+
+\noindent Again, $S_{i,t}$ and $I_{i,t}$ represent, respectively, percent change
+in species diversity and an indicator variable for increasing or decreasing
+species diversity. We exponentiate the equation to ensure all scale parameters,
+$\sigma_{i,t}$, are positive. The $\gamma$'s are estimated coefficients, three
+of which are allowed to vary by strategy, constrained by normal distributions:
+
+\begin{align}
+\gamma_{0,j}  &\sim \mathrm{N} \left(\eta_0 + \eta_1 M_j + \eta_2 E_j, \tau^2_{\gamma_0}\right),\\
+\gamma_{1,j}  &\sim \mathrm{N} \left(\mu_{\gamma_1}, \tau^2_{\gamma_1}\right),\\
+\gamma_{2,j}  &\sim \mathrm{N} \left(\mu_{\gamma_2}, \tau^2_{\gamma_2}\right).
+\end{align}
+
+\noindent The mean of the $\gamma_{0,j}$ coefficients is modelled with two
+strategy-level predictors. The symbols $M_j$ and $E_j$ represent the mean
+species diversity and mean combined fishing season length (effort) for strategy
+$j$. The coefficients $\eta_1$ and $\eta_2$ represent predictors in this
+strategy-level regression estimating effects across strategies.
+
+We fit our models with Stan 2.11.1 [@hoffman2014; @standevelopmentteam2015;
+@standevelopmentteam2015a; @carpenter2015] and R 3.3.1 [@rcoreteam2016]. Stan
+implements the No-U-Turn Hamiltonian Markov chain Monte Carlo algorithm
+[@hoffman2014] to perform Bayesian statistical inference. We assigned weakly
+informative priors on all parameters: $N(0, 2^2)$ priors on all $\beta$,
+$\gamma$, and $\eta$ parameters, and half-$t(3, 0, 2)$ priors (i.e. degrees of
+freedom of 3 and scale of 2) on all $\sigma$ and $\tau$. We ran four chains and
+2,000 iterations and discarded the first 1,000 iterations of each chain as warm
+up. We checked for chain convergence visually with trace plots, ensured that
+$\hat{R}<1.05$ (the potential scale reduction factor) and that the effective
+sample size, as calculated by the rstan R package [@standevelopmentteam2015],
+was greater than 200 for all parameters [@gelman2014]. We verified that our
+model returned sensible estimates by comparing our estimates to an ad-hoc
+two-stage model fit with the lme4 R package [@bates2015], where we first fit
+a mixed-effect regression model to the mean, and then modelled the absolute
+residuals from that first model in a second mixed effects model representing the
+variance component. We also validated the model with simulated data, and were
+able to recover the true values (Supporting Information). Code used in our
+analysis is available at <https://github.com/NCEAS/pfx-commercial/portfolios/>.
+
+*TODO: move the following into the results section or into the methods section:*
+
+Residuals from the mean component of our model clearly show the "V"-shaped
+pattern that the variance component of our model seeks to capture (Fig.
+S\ref{fig:resid}). Furthermore, the downside residuals (year-individual data
+points with lower than expected revenue) look qualitatively similar to the
+complete set of residuals (upside and downside) (Fig.
+S\ref{fig:resid-downside}). Therefore, it seemed appropriate to model the
+variability as a proxy for revenue risk.
+
+The estimated year-permit intercepts represent years when people with
+a particular permit set in a given year tended to do better or worse than the
+long-term average for that permit set (Fig. S\ref{fig:str-yrs}). Therefore, an
+individual's expected revenue and variability is after accounting for the
+general trend in revenue for that permit set.
+
+Our model estimates overall average effects (Fig. S\ref{fig:fe}) and effects
+that vary by permit set (Fig. S\ref{fig:re}).
 

### 2016-12-11 portfolio/text/ms.Rmd
Edit intro/results

@@
 rather than fleets or vessels, we examine how revenue and revenue variability of
-individuals is affected by diversification. By tracking individuals rather than
-vessels we can track people that switch fishery permits over time or use
-multiple vessels. Similarly, we are able to disentangle fishers that share the
-same vessel. We develop and apply a hierarchical Bayesian model to understand
-the effect of diversity on both individual revenue and revenue variability that
-builds on results from vessel [@kasperski2013], fleet [@sethi2012b], and
-community-level [@sethi2014] variability. Our model controls for fishing effort
-and latent time effects (e.g. changing environmental conditions), simultaneously
-estimates the effect of species diversity on individual revenue and revenue
-variability, and estimates effects both within and across fishing strategies
-(sets of similar licenses or permits). To our knowledge, this is the first
-application of this type of model in any natural resource field.
+individuals is affected by diversification. By tracking individuals we can track
+people that switch fishery permits over time or use multiple vessels. Similarly,
+we can disentangle fishers that share the same vessel. We develop and apply
+a hierarchical Bayesian model to understand the effect of fished species
+diversity on individual revenue and revenue variability that builds on results
+from vessel [@kasperski2013], fleet [@sethi2012b], and community-level
+[@sethi2014] variability.
 
-Our hierarchical variance-function regression model [@davidian1987; @hall1989]
-simultaneously models changes in the mean and variance of individual revenue as
-a function of covariates. Whereas traditional hierarchical linear models assume
-homoscedasticity --- that the residual variance does not vary systematically ---
-the variance-function regression model allows the variance to be heteroscedastic
-and vary as a function of predictors. Our specific model allows the revenue rate
-of return (log of the ratio of year-to-year revenue) and the revenue variability
-(the standard deviation of this revenue rate of return) to vary by fishing
-strategy (fishing permit sets), fishing effort (season length), and levels of
-species diversification.
+Our model is a variance-function regression model [@davidian1987; @hall1989],
+which simultaneously models changes in the mean and variance of individual
+revenue as a function of covariates. Whereas traditional hierarchical linear
+models assume homoscedasticity --- that the residual variance does not vary
+systematically --- the variance-function regression model allows the variance to
+be heteroscedastic and vary as a function of predictors. Our specific model
+simultaneously estimates the effect of species diversity on individual revenue
+and revenue variability, and estimates effects both within and across fishing
+strategies (sets of similar licenses or permits). At the same time, it controls
+for fishing effort and latent time effects (e.g. changing environmental
+conditions). To our knowledge, this is the first application of this type of
+model in any natural resource field.
 
@@
 
-We first evaluated trends in revenue and diversification over time for
-individuals fishing commercially in Alaska from 1985--2014 (Fig. \ref{fig:ts}A).
-The majority (70--90%) of fishers hold single permits, with a slight increase in
+We evaluated trends in revenue and diversification over time for individuals
+fishing commercially in Alaska from 1985--2014 (Fig. \ref{fig:ts}A). The
+majority (70--90%) of fishers hold single permits, with a slight increase in
 single-permit holders since 2000 (Fig. \ref{fig:ts}C). This overall trend is
@@
 (Fig. \ref{fig:ts}C). Despite making up only 10--30% of the individuals,
-multiple permit holders make approximately half of the overall revenue (Fig.
+multiple permit holders earn approximately half the overall revenue (Fig.
 \ref{fig:ts}B). The mean effective fished species diversity of this revenue
@@
 
-Revenue variability decreases exponentially with higher mean fished species
-diversity across permit sets (Fig. \ref{fig:across}). Controlling for effort,
-strategies with one unit higher mean effective species diversity had on average
-`r h1[[2]]`% (`r h1[[1]]`--`r h1[[3]]`% 95% CI) lower variability. This pattern
-looks qualitatively similar with or without individuals targeting salmon. As
-examples, variability was highest for the herring roe gillnet and king crab
-fisheries. Most of the strategies with the lowest estimated revenue variability
-combined halibut with a salmon permit and had high levels of fished species
-diversity (Fig. \ref{fig:across}).
+Individual revenue variability decreases exponentially with higher mean fished
+species diversity across permit sets (Fig. \ref{fig:across}). Controlling for
+effort, strategies with one unit higher mean effective species diversity had on
+average `r h1[[2]]`% (`r h1[[1]]`--`r h1[[3]]`% 95% CI) lower variability. This
+pattern looks qualitatively similar with or without individuals targeting
+salmon. As examples, variability was highest for the herring roe gillnet and
+king crab fisheries. Most of the strategies with the lowest estimated revenue
+variability combined halibut with a salmon permit and had high levels of fished
+species diversity (Fig. \ref{fig:across}).
 
@@
 
-The management of the Alaskan halibut and sablefish fisheries was converted to
-individual fishery quotas (IFQ) in 1994. This catch share system... TODO. We
-refit our models allowing all permit-level effects to vary separately before and
-after the introduction of IFQs for strategies involving halibut and sablefish
-permits. The introduction of IFQs reduced variability by ~25% (XX--XX 95% CI)
-for individuals with only halibut or sablefish permits (Fig. S\ref{fig:ifq}A).
-Furthermore, for most strategies, the effect of increased revenue variability
-associated with changing species diversity became stronger after the
-introduction of IFQs (Fig. S\ref{fig:ifq}B). This was especially true for
-individuals fishing only halibut or sablefish and less so for individuals
-combining halibut or sable fish with a salmon permit (Fig. S\ref{fig:ifq}B).
+Alaskan halibut and sablefish fisheries were converted to individual fishery
+quota (IFQ) management in 1994. This catch share system... TODO. We refit our
+models allowing all permit-level effects to vary separately before and after the
+introduction of IFQs for strategies involving halibut and sablefish permits. The
+introduction of IFQs reduced variability by ~25% (XX--XX 95% CI) for individuals
+with only halibut or sablefish permits (Fig. S\ref{fig:ifq}A). Furthermore, for
+most strategies, the association changing species diversity with increased
+revenue variability became stronger after the introduction of IFQs (Fig.
+S\ref{fig:ifq}B). This was especially true for individuals fishing only halibut
+or sablefish and less so for individuals combining halibut or sablefish with
+a salmon permit (Fig. S\ref{fig:ifq}B).
 

### 2016-12-11 portfolio/text/ms.Rmd
Editing portfolio paper

@@
-```{r set-knitr-options, cache=FALSE, echo=FALSE}
+```{r set-knitr-options, echo=FALSE}
 library("knitr")
-opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, results='hide')
+opts_chunk$set(echo=FALSE, message=FALSE, warning=FALSE, results="hide")
 ```
@@
 ```
+
+<!-- Diversified portfolios of fishing permits and species reduce risk for fishermen, -->
+<!-- but changing species diversity can be risky, costly, and limited by regulations. -->
+<!-- Goal: Introduction, Methods, Results, Discussion, Conclusion:
+<!-- 600, 900, 500, 800, 200, -->
 
@@
 
-\bigskip
-\noindent
-
 \clearpage
-
-<!-- Extra: -->
-
-<!-- - there is often limited room for intentional diversification within a permit -->
-<!-- - in fact, specializing can be profitable within some fisheries -->
-<!-- - Eric: analog here in farming literature is that diversification -->
-<!--   for small scale farmers generally good, except when they -->
-<!--   specialize on high demand/high value crops (tobacco, cocoa). -->
-<!--   Seems like benefits of diversification are also smaller for -->
-<!--   large-scale industrial farmers. -->
-
-<!-- \noindent -->
-<!-- Main message: -->
-<!-- Diversified portfolios of fishing permits and species reduce risk for fishermen, -->
-<!-- but changing species diversity can be risky, costly, and limited by regulations. -->
-
-<!-- , which is concerning  -->
-<!-- given that some fishing regulations and market forces -->
-<!-- may be increasingly promoting specialization. -->
-<!-- Main message in 25 words or less: -->
-
-<!-- Take 1: -->
-<!-- - There is a trend towards specializing on single salmon permits in Alaska -->
-<!-- - Diversified portfolios of fishing permits and species reduce risk, -->
-<!--   but changing species diversity for an individual permit holder can be risky -->
-<!-- - The types of diversity needed to reduce risk are often restricted by -->
-<!--   cost and regulations -->
-
-<!-- Take 2: -->
-<!-- - The effect of species diversification on revenue and risk for fishermen -->
-<!-- depends on the type of diversification and the fishery. -->
-
-<!-- Goal: Introduction, Methods, Results, Discussion, Conclusion:
-<!-- 600, 900, 500, 800, 200, -->
 
@@
 
-### Take homes / summary
+### Summary
 
-- Suggests it's mostly an across strategy/permit effect (lowering variability),
+- suggests it's mostly an across strategy/permit effect (lowering variability),
   which isn't always a possibility because of regulations and cost
-- Adjusting within current permit constraints is risky... Not clearly risk
-  reducing to generalize from one year to the next
+- adjusting within current permit constraints is risky... Not clearly risk
+  reducing to generalize from one year to the next, and accounting for effort,
+  specializing can increase revenue
 - not as simple as increase diversity (within permit) and reduce risk
-- What's methodologically novel
-    - Huge dataset
-    - Permit level data (on individuals, not vessels)
-    - Account for covariates
-    - Accounting for strategies
-    - Keeping annual data (not condensing to mean and cv)
-    - Joint modelling of mean and variance
-    - look at different mechanisms of diversification
-- 2 dimensions to think about (revenue and variability); and the trade-off is fishery specific
-    - effects within a permit strategy are permit specific
-    - Importantly it's a tradeoff - 2 dimensions - sometimes make more money by specializing... Some highly variable single species permits make a ton of money... which preferable?
+- methodologically novel: individuals, huge dataset, account for covariates,
+  accounting for strategies, keeping annual data (not condensing to mean and
+  cv), joint modelling of mean and variance, evaluating within and across
+  strategies (different mechanisms of diversification)
+- 2 dimensions to think about (revenue and variability); and the trade-off is
+  fishery specific: effects within a permit strategy are permit specific;
+  sometimes make more money by specializing... some highly variable single
+  species permits make a ton of money
 
@@
 
-- How fit in with K and H?
-
-- EW: part of the difference is individuals vs vessels. For example, K&H
+- why variability/uncertainty might increase with changes in diversity...
+- and how fit in with @kasperski2013, @sethi2012b, @sethi2014
+- part of the difference is individuals vs vessels. E.g. @kasperski2013
   suggested that diversity declined following IFQs for halibut sablefish
-vessels. The reason for the decline was that the vessels were allowed to
-fish halibut for more days (because people were sharing boats). At the level
-of people, we wouldn't expect the diversity of someone who only had
+  vessels. The reason for the decline was that the vessels were allowed to fish
+  halibut for more days (because people were sharing boats). At the level of
+  people, we wouldn't expect the diversity of someone who only had
   halibut/sablefish permit to be affected by IFQs. The person, not the vessel
   holds the quota share: https://alaskafisheries.noaa.gov/fisheries/ifq
-- Year to year risky... But maybe evens out/better over time
-- diversity within permit - often bycatch, limited room for intentional diversification (unless you have a M or salmon permit)
-- Risk results not really risk, but unpredictability given our covariates ...
-don't change anything and it's most predictable how well you'll do
-- What's unpredictable in our model, might be more predictable on a year to
-year, species by species basis for individual permit holder
-- Equilibrium, average variability vs year to year effect for one person
+- year to year risky... But maybe evens out/better over time
+- diversity within permit - often bycatch, limited room for intentional
+  diversification (unless you have a M or salmon permit)
+- risk results not really risk, but unpredictability given our covariates ...
+  don't change anything and it's most predictable how well you'll do. What's
+  unpredictable in our model, might be more predictable on a year to year,
+  species by species basis for individual permit holder
+- CV vs year-to-year effect for one person
+
+Other:
+
+- specializing can be profitable within some fisheries. Eric: analog here in
+  farming literature is that diversification for small scale farmers generally
+  good, except when they specialize on high demand/high value crops (tobacco,
+  cocoa). Seems like benefits of diversification are also smaller for
+  large-scale industrial farmers.
 
@@
 
-- limitations...
-- Vary by time
-- Vary by mean diversity
-- Spatial variation
-- Nonlinear effects
-- Response to specific events like EVOS and ITQs
+- could vary by time and vary by mean diversity
+- spatial/geographic variation
+- nonlinear effects
+- response to specific events like EVOS
 
@@
 
-- if this is true, this is concerning because in many areas of the world and
-fisheries, regulations and market forces have encouraged specialization
-- Permit diversity result: a privileged position - expensive to get to into
-(quote king crab license)
-- Locking into permits limits ability to pick lower risk strategies and is
+- concerning because in many areas of the world and fisheries, regulations and
+  market forces have encouraged specialization
+- permit diversity result: a privileged position --- expensive to get into
+- locking into permits limits ability to pick lower risk strategies and is
   getting people locked into low diversity permits that may be doing well now,
@@
 
-<!-- Ole: I'd use tau (or something) for the variances here so that people don't have to look at the subscripts to differentiate between variances at different hierarchical levels. -->
-
 \begin{equation}

### 2016-12-12 portfolio/text/ms.Rmd
Edit results + methods

@@
 - concerning because in many areas of the world and fisheries, regulations and
-  market forces have encouraged specialization
-- permit diversity result: a privileged position --- expensive to get into
+  market forces have encouraged specialization / limited (effective)
+  diversification (e.g. multiple permits or diverse permits)
 - locking into permits limits ability to pick lower risk strategies and is
@@
 # Materials and methods
+
+<!-- TODO: get fig:re in results -->
+<!-- Our model estimates overall average effects (Fig. S\ref{fig:fe}) and effects -->
+<!-- that vary by permit set (Fig. S\ref{fig:re}). -->
 
@@
 along with strategy-level intercepts would be estimated at almost exactly zero.
+The estimated $\beta_{j,t}$'s represent years when people with a particular
+permit set tended to do better or worse than the long-term average for that
+permit set (Fig. S\ref{fig:str-yrs}). Therefore, an individual's expected
+revenue (and variability) is after accounting for the general trend in revenue
+for that permit set.
 
@@
 species diversity. We exponentiate the equation to ensure all scale parameters,
-$\sigma_{i,t}$, are positive. The $\gamma$'s are estimated coefficients, three
-of which are allowed to vary by strategy, constrained by normal distributions:
+$\sigma_{i,t}$, are positive. Residuals from the mean component of our model
+clearly show a "V"-shaped pattern that the variance component of our model can
+capture (Fig. S\ref{fig:resid}). Furthermore, the downside residuals
+(year-individual data points with lower than expected revenue) look
+qualitatively similar to the complete set of residuals (upside and downside)
+(Fig. S\ref{fig:resid-downside}). Therefore, it seemed appropriate to model the
+variability as a proxy for revenue risk.
+
+The $\gamma$'s are estimated coefficients, three of which are allowed to vary by
+strategy, constrained by normal distributions:
 
@@
 two-stage model fit with the lme4 R package [@bates2015], where we first fit
-a mixed-effect regression model to the mean, and then modelled the absolute
-residuals from that first model in a second mixed effects model representing the
+a mixed-effects regression model to the mean, and then modelled the absolute
+residuals from that first model in a second mixed-effects model representing the
 variance component. We also validated the model with simulated data, and were
@@
 analysis is available at <https://github.com/NCEAS/pfx-commercial/portfolios/>.
-
-*TODO: move the following into the results section or into the methods section:*
-
-Residuals from the mean component of our model clearly show the "V"-shaped
-pattern that the variance component of our model seeks to capture (Fig.
-S\ref{fig:resid}). Furthermore, the downside residuals (year-individual data
-points with lower than expected revenue) look qualitatively similar to the
-complete set of residuals (upside and downside) (Fig.
-S\ref{fig:resid-downside}). Therefore, it seemed appropriate to model the
-variability as a proxy for revenue risk.
-
-The estimated year-permit intercepts represent years when people with
-a particular permit set in a given year tended to do better or worse than the
-long-term average for that permit set (Fig. S\ref{fig:str-yrs}). Therefore, an
-individual's expected revenue and variability is after accounting for the
-general trend in revenue for that permit set.
-
-Our model estimates overall average effects (Fig. S\ref{fig:fe}) and effects
-that vary by permit set (Fig. S\ref{fig:re}).
 

### 2016-12-12 portfolio/text/ms.Rmd
Correct methods

@@
 The estimated $\beta_{j,t}$'s represent years when people with a particular
-permit set tended to do better or worse than the long-term average for that
-permit set (Fig. S\ref{fig:str-yrs}). Therefore, an individual's expected
-revenue (and variability) is after accounting for the general trend in revenue
-for that permit set.
+permit set tended to collectively do better or worse than the long-term average
+(Fig. S\ref{fig:str-yrs}). Therefore, an individual's expected change in revenue
+(and variability) is after accounting for the general trend in revenue for that
+permit set.
 
@@
 
-\noindent The $\beta_{1,j}$'s represent how much a 1\% decrease in species
-diversity within a strategy translates to a Y\% change in revenue. The
-$\beta_{2,j}$'s represent how much a 1\% increase in species diversity within
-a strategy translates to a Y\% change in revenue. These are after controlling
-for changes in effort, $\beta_3$, and the interaction between changes in effort
-and changes in species diversity, $\beta_4$ and $\beta_5$.
+\noindent The $\beta_{1,j}$'s and $\beta_{2,j}$'s represent how much a 1\%
+decrease and 1\% increase, respectively, in the ratio of species diversity from
+year-to-year within a strategy translates to a given percent change in the ratio
+of revenue from year-to-year. These are after controlling for changes in effort,
+$\beta_3$, and the interaction between changes in effort and changes in species
+diversity, $\beta_4$ and $\beta_5$.
 

### 2016-12-12 portfolio/text/ms.Rmd
Fix caption

@@
 higher variability for individuals within the strategies above the zero line.
-Specializing is associated with greater expected revenue to the right of the
-zero line and lower expected revenue to the left of the zero line. For example,
-individuals with setnet permits for salmon in Prince William Sound, Cook Inlet,
-Kodiak, Bristol Bay, or the Alaska Peninsula target primarily sockeye salmon and
-experience a $\sim$1.5\% increase in their year-to-year revenue ratio for a 1\%
-decrease in their year-to-year species diversity. These effects are estimated
-holding fishing season length constant from year-to-year. Dots and line segments
-represent medians and 50\% credible intervals of posteriors. The effect of
-generalizing (increasing species diversity) is similar (Fig. S\ref{fig:mirror})}.
+Specializing is associated with greater expected revenue ratios to the right of
+the zero line and lower expected revenue ratios to the left of the zero line.
+For example, individuals with setnet permits for salmon in Prince William Sound,
+Cook Inlet, Kodiak, Bristol Bay, or the Alaska Peninsula (second purple dot from
+right) target primarily sockeye salmon and experience a $\sim$1.5\% increase in
+their year-to-year revenue ratio (horizontal axis) and a $\sim$0.5\% increase in
+their revenue variability (vertical axis) for a 1\% decrease in their ratio of
+year-to-year species diversity. These effects are estimated holding fishing
+season length constant from year-to-year. Dots and line segments represent
+medians and 50\% credible intervals of posteriors. The effect of generalizing
+(increasing species diversity) is similar (Fig. S\ref{fig:mirror}).}
 

### 2016-12-12 portfolio/text/ms.Rmd
Results and supp edits

@@
 \ref{fig:within}). After accounting for fishing effort, specializing tends to be
-associated with increased revenue (Fig. \ref{fig:within}; Fig. S\ref{fig:fe}).
-On average, there is a ~0.5% increase in the year-to-year revenue ratio per 1%
-decrease in species diversity, and across strategies, `r g2_dens`% of the
-probability density supports a positive relationship between specialization and
-revenue. The converse is also usually true --- generalizing tends to be
-associated with lower expected revenue after accounting for fishing effort (i.e.
-Fig. S\ref{fig:mirror}A and \ref{fig:mirror}B are near mirror images left to
-right). This benefit of specializing is only apparent after accounting for
-fishing effort. Given that fishing for more species tends to be associated with
-increased fishing effort (fishing more days), generalizing tends to be
-associated with higher revenue when not standardizing for days finished (Fig.
-S-TODO).
+associated with increased revenue (Fig. \ref{fig:within}; Fig. S\ref{fig:fe},
+Fig. S\ref{fig:re}). On average, there is a ~0.5% increase in the year-to-year
+revenue ratio per 1% decrease in species diversity, and across strategies, `r
+g2_dens`% of the probability density supports a positive relationship between
+specialization and revenue. The converse is also usually true --- generalizing
+tends to be associated with lower expected revenue after accounting for fishing
+effort (i.e. Fig. S\ref{fig:mirror}A and \ref{fig:mirror}B are near mirror
+images left to right). This benefit of specializing is only apparent after
+accounting for fishing effort. Given that fishing for more species tends to be
+associated with increased fishing effort (fishing more days), generalizing tends
+to be associated with higher revenue when not standardizing for days finished
+(Fig. S-TODO).
 
@@
 tends to be associated with greater expected variability or uncertainty in
-year-to-year revenue (Fig. \ref{fig:within} strategies above 0; Fig.
-S\ref{fig:fe}). This effect was strongest for individuals with a halibut permit
-or halibut and sablefish permits. Individuals with these strategies experienced
-~2% increase in variability for a 1% change in the year-to-year ratio of species
-diversity (Fig. \ref{fig:within}). Levels of fished species diversity above one
-for these particular permits are primarily due to bycatch.
+year-to-year revenue (Fig. \ref{fig:within} strategies above 0; Figs.
+S\ref{fig:fe}, \ref{fig:re}, \ref{fig:mirror}). This effect was strongest for
+individuals with a halibut permit or halibut and sablefish permits. Individuals
+with these strategies experienced ~2% increase in variability for a 1% change in
+the year-to-year ratio of species diversity (Fig. \ref{fig:within}). Levels of
+fished species diversity above one for these particular permits are primarily
+due to bycatch.
 
@@
 quota (IFQ) management in 1994. This catch share system... TODO. We refit our
-models allowing all permit-level effects to vary separately before and after the
+models allowing all strategy-level effects to vary separately before and after the
 introduction of IFQs for strategies involving halibut and sablefish permits. The

### 2016-12-12 portfolio/text/supp.Rnw
Results and supp edits

@@
 \begin{center}
-\includegraphics[width=1.1\textwidth]{../figs/stan-offset-break-anti-spaghetti.pdf}
+\includegraphics[width=1.05\textwidth]{../figs/stan-offset-break-anti-spaghetti.pdf}
 \caption{Effects of (A) generalizing and (B) specializing from year-to-year for
-individual permit holders \emph{within} the `r nstr` most common strategies.
-Changing species diversity from year-to-year is associated with higher
-variability for individuals within the strategies above the zero line.
-Generalizing (A) or specializing (B) is associated with greater expected revenue
-to the right of the zero line and lower expected revenue to the left of the zero
-line. These effects are estimated holding fishing season length constant from
-year-to-year. Dots and line segments represent medians and 50\% credible
-intervals of posteriors. }
+  individual permit holders \emph{within} the \Sexpr{nstr} most common
+  strategies. Changing species diversity from year-to-year is associated with
+  higher variability for individuals within the strategies above the zero line.
+  Generalizing (A) or specializing (B) is associated with greater expected
+  revenue ratios to the right of the zero line and lower expected revenue ratios
+  to the left of the zero line. These effects are estimated holding fishing
+  season length constant from year-to-year. Dots and line segments represent
+  medians and 50\% credible intervals of posteriors. Panel B and Fig.~4 are the
+  same.}
 \label{fig:mirror}

### 2016-12-12 portfolio/text/ms.Rmd
Edit intro

@@
 a license, own multiple licenses, or diversify spatially and fish in multiple
-regions.
-
-The effectiveness of these diversification strategies for increasing revenue and
-reducing revenue variability remains uncertain. For example, theory and
-empirical studies suggest there can be substantial benefits to specialization
-for both farmers and fishers [@abdulai2000; @arnason2005]. Furthermore, the
-majority of research examining the effects of diversification on revenue do so
-at aggregate levels that combine individuals (e.g. households, counties, or
-communities; [@kasperski2013; @sethi2014]). Because of this, key uncertainties
-remain about the fundamental relationship between diversity, fishing revenue,
-and fishing revenue variability at the level of the individual.
+regions. The effectiveness of these diversification strategies for increasing
+revenue and reducing revenue variability remains uncertain. For example, theory
+and empirical studies suggest there can be substantial benefits to
+specialization for both farmers and fishers [@abdulai2000; @arnason2005].
+Furthermore, the majority of research examining the effects of diversification
+on revenue do so at aggregate levels that combine individuals (e.g. households,
+counties, or communities; [@kasperski2013; @sethi2014]). Because of this, key
+uncertainties remain about the fundamental relationship between diversity,
+fishing revenue, and fishing revenue variability at the level of the individual.
 
@@
 models assume homoscedasticity --- that the residual variance does not vary
-systematically --- the variance-function regression model allows the variance to
-be heteroscedastic and vary as a function of predictors. Our specific model
-simultaneously estimates the effect of species diversity on individual revenue
-and revenue variability, and estimates effects both within and across fishing
-strategies (sets of similar licenses or permits). At the same time, it controls
-for fishing effort and latent time effects (e.g. changing environmental
+systematically --- the variance-function regression model allows the residual
+variance to be heteroscedastic and vary as a function of predictors. Our
+specific model simultaneously estimates the effect of species diversity (the
+inverse of Simpson's diversity index [@simpson1949; @jost2006]) on individual
+revenue and revenue variability, and estimates effects both within and across
+fishing strategies (sets of similar licenses or permits). At the same time, it
+controls for fishing effort and latent time effects (e.g. changing environmental
 conditions). To our knowledge, this is the first application of this type of

### 2016-12-13 portfolio/text/ms.Rmd
Copyediting

@@
 Managing natural resources and sustaining the livelihoods of those that depend
-on their extraction is hard. These professions tend to have high year-to-year
+on their extraction is hard. These livelihoods tend to have high year-to-year
 variability in income relative to other professions [@mishra2002; @sethi2012b].
@@
 tends to be associated with lower expected revenue after accounting for fishing
-effort (i.e. Fig. S\ref{fig:mirror}A and \ref{fig:mirror}B are near mirror
+effort (i.e. Fig. S\ref{fig:mirror}A and S\ref{fig:mirror}B are near mirror
 images left to right). This benefit of specializing is only apparent after

### 2016-12-13 portfolio/analysis/3.1-offset-stan.R
Add model without effort

@@ -39,7 +39,7 @@ m <- stan("portfolio/analysis/portfolio-offset.stan",
   pars = c("mu", "sigma"), include = FALSE)
 save(m, standat, file = "portfolio/data-generated/m.rda")
 
-b <- broom::tidy(m1, conf.int = T, estimate.method = "median", rhat = T, ess = T)
+b <- broom::tidy(m1, conf.int = TRUE, estimate.method = "median", rhat = TRUE, ess = TRUE)
 filter(b, rhat > 1.05)
 filter(b, ess < 100)
 filter(b, grepl("^h1", term))
@@ -48,6 +48,22 @@ filter(b, grepl("^g0", term))
 filter(b, grepl("^g_k", term))
 filter(b, grepl("^b_j", term))
 filter(b, grepl("*_tau$", term))
+
+# Fit the same model without the effort predictor
+
+standat_noeffort <- standat
+effort_columns <- grep("days", names(as.data.frame(standat_noeffort$X_ij)))
+standat_noeffort$X_ij <- standat_noeffort$X_ij[, -effort_columns]
+standat_noeffort$X_sigma_ik <- standat_noeffort$X_sigma_ik[, -effort_columns]
+standat_noeffort$J <- ncol(standat_noeffort$X_ij)
+standat_noeffort$K <- ncol(standat_noeffort$X_sigma_ik)
+
+system.time({
+m_noeffort <- stan("portfolio/analysis/portfolio-offset.stan",
+  data = standat, iter = 400, chains = 2,
+  pars = c("mu", "sigma"), include = FALSE)
+save(m_noeffort, standat, file = "portfolio/data-generated/m_noeffort.rda")
+})
 
 ## m_ns <- stan("portfolio/analysis/portfolio-offset-nosigma.stan",
 ##   data = standat, iter = 120, chains = 2,

### 2016-12-14 portfolio/text/ms.Rmd
Add discussion paragraph

@@
 
-### Summary
-
-- suggests it's mostly an across strategy/permit effect (lowering variability),
-  which isn't always a possibility because of regulations and cost
-- adjusting within current permit constraints is risky... Not clearly risk
-  reducing to generalize from one year to the next, and accounting for effort,
-  specializing can increase revenue
-- not as simple as increase diversity (within permit) and reduce risk
-- methodologically novel: individuals, huge dataset, account for covariates,
-  accounting for strategies, keeping annual data (not condensing to mean and
-  cv), joint modelling of mean and variance, evaluating within and across
-  strategies (different mechanisms of diversification)
-- 2 dimensions to think about (revenue and variability); and the trade-off is
-  fishery specific: effects within a permit strategy are permit specific;
-  sometimes make more money by specializing... some highly variable single
-  species permits make a ton of money
+We fit a hierarchical variance-function regression model to 30 years of
+commercial fisheries revenue data in Alaska to simultaneously estimate the
+effect of changes in species diversity on individual's revenue and revenue
+variability. We find that individuals fishing permit sets that allow high levels
+of species diversity tend to have lower revenue variability. Adding an
+additional fisheries permit usually increases revenue and decreases revenue
+variability. However, within the constraints of a permit and accounting for
+effort, specializing (reducing species diversity) tends to increase revenue and
+specializing or generalizing tends to be associated with greater revenue
+variability, with considerable differences across fisheries. Thus, the most
+effective mechanism by which species diversity can lower revenue variability for
+individual fishers --- switching or adding fishing permits --- is often costly,
+limited by regulations, and unavailable to many individuals.
 
@@
 - why variability/uncertainty might increase with changes in diversity...
+
 - and how fit in with @kasperski2013, @sethi2012b, @sethi2014
+
 - part of the difference is individuals vs vessels. E.g. @kasperski2013
@@
   holds the quota share: https://alaskafisheries.noaa.gov/fisheries/ifq
+
 - year to year risky... But maybe evens out/better over time
+
 - diversity within permit - often bycatch, limited room for intentional
   diversification (unless you have a M or salmon permit)
+
 - risk results not really risk, but unpredictability given our covariates ...
@@
   species by species basis for individual permit holder
+
 - CV vs year-to-year effect for one person
@@
 
+There are a number of limitations to our model and extensions that could further
+our understanding of the role of diversification...
+
 - could vary by time and vary by mean diversity
+
 - spatial/geographic variation
+
 - nonlinear effects
+
 - response to specific events like EVOS
@@
   diversification (e.g. multiple permits or diverse permits)
+
 - locking into permits limits ability to pick lower risk strategies and is
@@
 ### Conclusions + recommendations
+
+- bring it out to wider than fisheries (farming, timber industry, ...)
+
+- 2 dimensions to think about (revenue and variability); and the trade-off is
+  fishery specific: effects within a permit strategy are permit specific;
+  sometimes make more money by specializing... some highly variable single
+  species permits make a ton of money
+
+- broadly, diversification can provide a powerful form of protection from
+  multiple sources of risk to income for those that rely on extracting natural
+  resources, but not all forms of diversification are equal, and on an
+  individual level there can be economic benefits to specializing and risks to
+  adjusting ones strategy
+
+- with increasing environmental variability and extremes due to climate change
+  those that rely on natural resources for their livelihoods need approaches to
+  buffer risk more than ever
+
+- regulations that limit the ability to diversify can have other considerable
+  benefits but it's important to be aware of trade-offs inherent in these
+  regulations when making management decisions
 

### 2016-12-14 portfolio/analysis/3.3-plot-within-strategy-effects.R
Fix label

@@ -79,8 +79,8 @@ p2 <- res %>%
     size = 2.9, colour = "grey60", #segment.color = "grey80",
     point.padding = unit(0.3, "lines"), max.iter = 6e3, segment.size = 0.3) +
   geom_point(aes(color = strategy_mean_div, size = nn)) +
-  xlab("Effect of specializing on revenue variability") +
-  ylab("Effect of specializing on variability") +
+  xlab("Effect of specializing on revenue") +
+  ylab("Effect of specializing on revenue variability") +
   scale_color_viridis() +
   labs(colour = "Mean sp.\ndiversity", size = "Number of\npermit holders") +
   theme(legend.title = element_text(size = rel(0.85))) +

### 2016-12-14 portfolio/text/ms.Rmd
Add effort fig

@@
 effort (i.e. Fig. S\ref{fig:mirror}A and S\ref{fig:mirror}B are near mirror
-images left to right). This benefit of specializing is only apparent after
+images left to right). This benefit of specializing is most apparent after
 accounting for fishing effort. Given that fishing for more species tends to be
-associated with increased fishing effort (fishing more days), generalizing tends
-to be associated with higher revenue when not standardizing for days finished
-(Fig. S-TODO).
+associated with increased fishing effort (fishing more days), specializing tends
+to be ...TODO (Fig. S\ref{fig:noeffort}).
 

### 2016-12-16 portfolio/text/ms.Rmd

@@
 # Abstract
+Ole would drop the confidence bounds for the abstract.  I added a bit but this seems like a long abstract for PNAS...
 
@@
 variability (e.g., `r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`%
-IQR] lower expected standard deviation for halibut plus another permit vs.
-single halibut permits). Finally, we show that increasing species diversity
+IQR] decline in standard deviation for fishers with halibut and another permit vs.
+only a halibut permit). Finally, we show that increasing species diversity
 within the constraints of a permit set has a fishery-dependent effect on revenue
@@
 of diversification, others are less effective. Importantly, the most effective
-measure for increasing revenue and decreasing risk --- purchasing additional
-permits to participate in additional fisheries --- was also the most costly. At
-the same time, Alaskan commercial fisheries have been increasingly focussed on
-single-permit (often salmon) fishing strategies. Our results suggest that the
-benefit of species diversity for revenue variability is mostly an across-permit
+mechanism for increasing revenue and decreasing risk --- purchasing additional
+permits to participate in additional fisheries --- was also the most costly.  Furthermore, 
+the least costly option of increasing catch diversity within an existing permit
+was associated with increased revenue variability. Thus, over the past 30 years as
+the Alaskan commercial fishing fleet has become increasingly specialized and most individuals 
+only participate under a single permit, the practical benefits accrued
+from increased species diversification are limited.
+
+Our results suggest that thebenefit of species diversity for revenue variability is mostly an across-permit
 effect, not a benefit that can be easily harnessed by an individual permit
@@
 obtainable by commercial fishers [@kasperski2013], possibly because both fish
-and fishers are spatially mobile. However, fishers still have the ability to
-lower risk. They can, for example, target as many species as permitted by
+and fishers are spatially mobile 
+[Ole is pretty sure it is because they don't have sufficient political clout, suggest deleting the clause after the comma in previous sentence]. 
+However, fishers maintain the ability to
+lower risk by diversifying their fishing portfolio. They can, for example, target as many species as permitted by
 a license, own multiple licenses, or diversify spatially and fish in multiple
@@
 on revenue do so at aggregate levels that combine individuals (e.g. households,
-counties, or communities; [@kasperski2013; @sethi2014]). Because of this, key
+counties, or communities; [@kasperski2013; @sethi2014]). Thus, key
 uncertainties remain about the fundamental relationship between diversity,
@@
 a hierarchical Bayesian model to understand the effect of fished species
-diversity on individual revenue and revenue variability that builds on results
+diversity [OLE: Call this diversity of catch rather than fished?] on individual 
+revenue and revenue variability that builds on results
 from vessel [@kasperski2013], fleet [@sethi2012b], and community-level
@@
 
-Our model is a variance-function regression model [@davidian1987; @hall1989],
-which simultaneously models changes in the mean and variance of individual
-revenue as a function of covariates. Whereas traditional hierarchical linear
-models assume homoscedasticity --- that the residual variance does not vary
+We develop a variance-function regression model [@davidian1987; @hall1989],
+to simultaneously models changes in the mean and variance of individual
+revenue as a function of covariates.
+[OLE: I think we leave this sentence for the methods somewhere- Whereas traditional hierarchical linear models assume homoscedasticity --- that the residual variance does not vary
 systematically --- the variance-function regression model allows the residual
-variance to be heteroscedastic and vary as a function of predictors. Our
-specific model simultaneously estimates the effect of species diversity (the
+variance to be heteroscedastic and vary as a function of predictors.] 
+We simultaneously estimate the effect of species diversity (the
 inverse of Simpson's diversity index [@simpson1949; @jost2006]) on individual
-revenue and revenue variability, and estimates effects both within and across
-fishing strategies (sets of similar licenses or permits). At the same time, it
-controls for fishing effort and latent time effects (e.g. changing environmental
-conditions). To our knowledge, this is the first application of this type of
+revenue and revenue variability, and estimate effects both within and across
+fishing strategies while controling for fishing effort and latent time effects 
+(e.g. changing environmental conditions). To our knowledge, this is the first application of this type of
 model in any natural resource field.
@@
 associated with increased revenue (Fig. \ref{fig:within}; Fig. S\ref{fig:fe},
-Fig. S\ref{fig:re}). On average, there is a ~0.5% increase in the year-to-year
+Fig. S\ref{fig:re}) but also increased variability. On average, there is a ~0.5% increase in the year-to-year
 revenue ratio per 1% decrease in species diversity, and across strategies, `r
 g2_dens`% of the probability density supports a positive relationship between
-specialization and revenue. The converse is also usually true --- generalizing
+specialization and revenue.
+
+However, adjusting (increasing or decreasing) an
+individual's fished species diversity within the constraints of a permit set
+tends to be associated with greater expected variability or uncertainty in
+year-to-year revenue (Fig. \ref{fig:within} strategies above 0; Figs.
+S\ref{fig:fe}, \ref{fig:re}, \ref{fig:mirror}). This effect was strongest for
+individuals with a halibut permit or halibut and sablefish permits. Individuals
+with these strategies experienced ~2% increase in variability for a 1% change in
+the year-to-year ratio of species diversity (Fig.
+\ref{fig:within}).
+Interestingly, becoming more generalized in species diversity also leads to increased revenue variability (FIG SUPP), suggesting that the act of changing strategies is a contributor to increased variability of revenue.
+
+Levels of fished species diversity above one for these particular permits are primarily
+due to bycatch.
+
+[OLE: I would just say - see supplement for discussion of consequences of generalizing and not include this section in the main text]
+The converse is also usually true --- generalizing
 tends to be associated with lower expected revenue after accounting for fishing
@@
 
-After accounting for fishing effort, adjusting (increasing or decreasing) an
-individual's fished species diversity within the constraints of a permit set
-tends to be associated with greater expected variability or uncertainty in
-year-to-year revenue (Fig. \ref{fig:within} strategies above 0; Figs.
-S\ref{fig:fe}, \ref{fig:re}, \ref{fig:mirror}). This effect was strongest for
-individuals with a halibut permit or halibut and sablefish permits. Individuals
-with these strategies experienced ~2% increase in variability for a 1% change in
-the year-to-year ratio of species diversity (Fig. \ref{fig:within}). Levels of
-fished species diversity above one for these particular permits are primarily
-due to bycatch.
-
 Alaskan halibut and sablefish fisheries were converted to individual fishery
-quota (IFQ) management in 1994. This catch share system... TODO. We refit our
-models allowing all strategy-level effects to vary separately before and after the
-introduction of IFQs for strategies involving halibut and sablefish permits. The
-introduction of IFQs reduced variability by ~25% (XX--XX 95% CI) for individuals
-with only halibut or sablefish permits (Fig. S\ref{fig:ifq}A). Furthermore, for
-most strategies, the association changing species diversity with increased
+quota (IFQ) management in 1994, which potentially confounds our estimates of the effect of
+diversity and a major regulatory change. We refit our models allowing all strategy-level effects
+to vary separately before and after the introduction of IFQs for strategies involving halibut
+and sablefish permits. The introduction of IFQs reduced variability by ~25% (XX--XX 95% CI)
+for individuals with only halibut or sablefish permits (Fig. S\ref{fig:ifq}A). 
+Furthermore, for most strategies, the association changing species diversity with increased
 revenue variability became stronger after the introduction of IFQs (Fig.
@@
 # Discussion
+[Ole thinks we should start bigger here to reflect the big main themes in the intro]
 
-We fit a hierarchical variance-function regression model to 30 years of
-commercial fisheries revenue data in Alaska to simultaneously estimate the
-effect of changes in species diversity on individual's revenue and revenue
-variability. We find that individuals fishing permit sets that allow high levels
+For individuals deriving livelihoods from natural resources, theory suggests that adopting strategies that utilize
+a diversity of assets [species?] provide substantial benefits over strategies with lower diversity [REFS].
+However, the ability to effectively diversify in practice is limited by regulation and market forces.
+Here we document the effect of changing species diversity on revenue and revenue variability across 
+three scale. Each scale reflects individual level choices that produce varied levels of species diversity.  
+We find that individuals fishing permit sets that allow high levels
 of species diversity tend to have lower revenue variability. Adding an
@@
 individual fishers --- switching or adding fishing permits --- is often costly,
-limited by regulations, and unavailable to many individuals.
+limited by regulations, and consequently unavailable to many individuals in practice.
+
+Our results shows additional support for the idea that individual fishers can reduce 
+their income variability by diversifying their catch portfolios. However, most individuals
+don’t appear to change their catch diversity significantly from year to year, and 
+most strategies adopted by Alaskan fishermen lie in a high-risk, high-reward quadrant
+in which reducing diversification increases both revenue variability and expected revenue (Fig. 4).
+Previous work has suggested that changes to fisheries management such as limited entry or
+catch share programs may inadvertently act as barrier to diversification and promote 
+specializing (Hilborn et al. 2003; Kasperski and Holland 2013). While some of these 
+specialized fishing strategies may be associated with increased revenue variability,
+the benefit may be that they are also associated with increased revenue.
+External market forces may interact in concert with management regulations to promote 
+specialization for fishers or other individuals dependent on natural resources. 
+In terrestrial agriculture for instance, there are also many instances of specializing 
+being profitable; drivers include the ability to specialize on high-profit cash crops
+(tobacco, cocoa), the availability of government agricultural subsidies (soybeans, corn in the USA),
+and farming on a large industrial scale. With increasing environmental variability and 
+extreme events in a changing climate, it is particularly important that individuals who 
+rely on natural resources as a primary source of income have strategies to reduce risk.
+Though variability in individual revenues may not be actively incorporated into many management
+decisions, analyses like ours help to understand the range of diversification strategies 
+available and how their benefits may change over time.
+
+Our work compliments previous results that have examined the effects of species diversity at the 
+community (Ref) and vessel(Ref) level that show THIS (Sethi, Kasperski, others)
 
@@
   regulations when making management decisions
+
+We fit a hierarchical variance-function regression model to 30 years of
+commercial fisheries revenue data in Alaska to simultaneously estimate the
+effect of changes in species diversity on individual's revenue and revenue
+variability.
+
+##3 ERIC
 

### 2016-12-16 portfolio/text/ms.Rmd
Edits incorporating Ole's edits

@@
 # Abstract
-Ole would drop the confidence bounds for the abstract.  I added a bit but this seems like a long abstract for PNAS...
+
+<!-- TODO slash length of abstract -->
 
@@
 variability (e.g., `r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`%
-IQR] decline in standard deviation for fishers with halibut and another permit vs.
-only a halibut permit). Finally, we show that increasing species diversity
+IQR] decline in standard deviation for fishers with halibut and another permit
+vs. only a halibut permit). Finally, we show that increasing species diversity
 within the constraints of a permit set has a fishery-dependent effect on revenue
@@
 mechanism for increasing revenue and decreasing risk --- purchasing additional
-permits to participate in additional fisheries --- was also the most costly.  Furthermore, 
-the least costly option of increasing catch diversity within an existing permit
-was associated with increased revenue variability. Thus, over the past 30 years as
-the Alaskan commercial fishing fleet has become increasingly specialized and most individuals 
-only participate under a single permit, the practical benefits accrued
-from increased species diversification are limited.
+permits to participate in additional fisheries --- was also the most costly.
+Furthermore, the least costly option of increasing catch diversity within an
+existing permit was associated with increased revenue variability. Thus, over
+the past 30 years as the Alaskan commercial fishing fleet has become
+increasingly specialized and most individuals only participate under a single
+permit, the practical benefits accrued from increased species diversification
+are limited.
 
-Our results suggest that thebenefit of species diversity for revenue variability is mostly an across-permit
-effect, not a benefit that can be easily harnessed by an individual permit
-holder. Effective methods of diversification --- purchasing additional permits
-or switching permits --- are costly and often limited by regulations such as
-limited entry programs and individual transferable quotas (ITQs). Take home ...
+Our results suggest that the benefit of species diversity for revenue
+variability is mostly an across-permit effect, not a benefit that can be easily
+harnessed by an individual permit holder. Effective methods of diversification
+--- purchasing additional permits or switching permits --- are costly and often
+limited by regulations such as limited entry programs and individual
+transferable quotas (ITQs). Take home ...
 
@@
 available to terrestrial farmers (e.g. crop insurance programs) are not
-obtainable by commercial fishers [@kasperski2013], possibly because both fish
-and fishers are spatially mobile 
-[Ole is pretty sure it is because they don't have sufficient political clout, suggest deleting the clause after the comma in previous sentence]. 
-However, fishers maintain the ability to
-lower risk by diversifying their fishing portfolio. They can, for example, target as many species as permitted by
-a license, own multiple licenses, or diversify spatially and fish in multiple
-regions. The effectiveness of these diversification strategies for increasing
-revenue and reducing revenue variability remains uncertain. For example, theory
-and empirical studies suggest there can be substantial benefits to
-specialization for both farmers and fishers [@abdulai2000; @arnason2005].
-Furthermore, the majority of research examining the effects of diversification
-on revenue do so at aggregate levels that combine individuals (e.g. households,
-counties, or communities; [@kasperski2013; @sethi2014]). Thus, key
-uncertainties remain about the fundamental relationship between diversity,
-fishing revenue, and fishing revenue variability at the level of the individual.
+obtainable by commercial fishers [@kasperski2013]. However, fishers maintain the
+ability to lower risk by diversifying their fishing portfolio. They can, for
+example, target as many species as permitted by a license, own multiple
+licenses, or diversify spatially and fish in multiple regions. The effectiveness
+of these diversification strategies for increasing revenue and reducing revenue
+variability remains uncertain. For example, theory and empirical studies suggest
+there can be substantial benefits to specialization for both farmers and fishers
+[@abdulai2000; @arnason2005]. Furthermore, the majority of research examining
+the effects of diversification on revenue do so at aggregate levels that combine
+individuals (e.g. households, counties, or communities; [@kasperski2013;
+@sethi2014]). Thus, key uncertainties remain about the fundamental relationship
+between diversity, fishing revenue, and fishing revenue variability at the level
+of the individual.
 
@@
 
-We develop a variance-function regression model [@davidian1987; @hall1989],
-to simultaneously models changes in the mean and variance of individual
-revenue as a function of covariates.
-[OLE: I think we leave this sentence for the methods somewhere- Whereas traditional hierarchical linear models assume homoscedasticity --- that the residual variance does not vary
-systematically --- the variance-function regression model allows the residual
-variance to be heteroscedastic and vary as a function of predictors.] 
-We simultaneously estimate the effect of species diversity (the
-inverse of Simpson's diversity index [@simpson1949; @jost2006]) on individual
-revenue and revenue variability, and estimate effects both within and across
-fishing strategies while controling for fishing effort and latent time effects 
-(e.g. changing environmental conditions). To our knowledge, this is the first application of this type of
-model in any natural resource field.
+We develop a variance-function regression model [@davidian1987; @hall1989], to
+simultaneously models changes in the mean and variance of individual revenue as
+a function of covariates. We simultaneously estimate the effect of species
+diversity (the inverse of Simpson's diversity index [@simpson1949; @jost2006])
+on individual revenue and revenue variability, and estimate effects both within
+and across fishing strategies while controlling for fishing effort and latent
+time effects (e.g. changing environmental conditions). To our knowledge, this is
+the first application of this type of model in any natural resource field.
 
@@
 associated with increased revenue (Fig. \ref{fig:within}; Fig. S\ref{fig:fe},
-Fig. S\ref{fig:re}) but also increased variability. On average, there is a ~0.5% increase in the year-to-year
-revenue ratio per 1% decrease in species diversity, and across strategies, `r
-g2_dens`% of the probability density supports a positive relationship between
-specialization and revenue.
+Fig. S\ref{fig:re}) but also increased variability. On average, there is a ~0.5%
+increase in the year-to-year revenue ratio per 1% decrease in species diversity,
+and across strategies, `r g2_dens`% of the probability density supports
+a positive relationship between specialization and revenue.
 
-However, adjusting (increasing or decreasing) an
-individual's fished species diversity within the constraints of a permit set
-tends to be associated with greater expected variability or uncertainty in
-year-to-year revenue (Fig. \ref{fig:within} strategies above 0; Figs.
-S\ref{fig:fe}, \ref{fig:re}, \ref{fig:mirror}). This effect was strongest for
-individuals with a halibut permit or halibut and sablefish permits. Individuals
-with these strategies experienced ~2% increase in variability for a 1% change in
-the year-to-year ratio of species diversity (Fig.
-\ref{fig:within}).
-Interestingly, becoming more generalized in species diversity also leads to increased revenue variability (FIG SUPP), suggesting that the act of changing strategies is a contributor to increased variability of revenue.
-
-Levels of fished species diversity above one for these particular permits are primarily
-due to bycatch.
-
-[OLE: I would just say - see supplement for discussion of consequences of generalizing and not include this section in the main text]
-The converse is also usually true --- generalizing
-tends to be associated with lower expected revenue after accounting for fishing
-effort (i.e. Fig. S\ref{fig:mirror}A and S\ref{fig:mirror}B are near mirror
-images left to right). This benefit of specializing is most apparent after
-accounting for fishing effort. Given that fishing for more species tends to be
-associated with increased fishing effort (fishing more days), specializing tends
-to be ...TODO (Fig. S\ref{fig:noeffort}).
+However, adjusting (increasing or decreasing) an individual's fished species
+diversity within the constraints of a permit set tends to be associated with
+greater expected variability or uncertainty in year-to-year revenue (Fig.
+\ref{fig:within} strategies above 0; Figs. S\ref{fig:fe}, \ref{fig:re},
+\ref{fig:mirror}). This effect was strongest for individuals with a halibut
+permit or halibut and sablefish permits. Individuals with these strategies
+experienced a ~2% increase in variability for a 1% change in the year-to-year
+ratio of species diversity (Fig. \ref{fig:within}). Becoming more generalized in
+species diversity also leads to increased revenue variability (Fig.
+S\ref{fig:mirror}A), suggesting that the act of changing strategies is
+a contributor to increased variability of revenue. This benefit of specializing
+is most apparent after accounting for fishing effort (Fig. S\ref{fig:noeffort}).
 
 Alaskan halibut and sablefish fisheries were converted to individual fishery
-quota (IFQ) management in 1994, which potentially confounds our estimates of the effect of
-diversity and a major regulatory change. We refit our models allowing all strategy-level effects
-to vary separately before and after the introduction of IFQs for strategies involving halibut
-and sablefish permits. The introduction of IFQs reduced variability by ~25% (XX--XX 95% CI)
-for individuals with only halibut or sablefish permits (Fig. S\ref{fig:ifq}A). 
-Furthermore, for most strategies, the association changing species diversity with increased
+quota (IFQ) management in 1994, which potentially confounds our estimates of the
+effect of diversity and a major regulatory change. We refit our models allowing
+all strategy-level effects to vary separately before and after the introduction
+of IFQs for strategies involving halibut and sablefish permits. The introduction
+of IFQs reduced variability by ~25% (XX--XX 95% CI) for individuals with only
+halibut or sablefish permits (Fig. S\ref{fig:ifq}A). Furthermore, for most
+strategies, the association between changing species diversity and increased
 revenue variability became stronger after the introduction of IFQs (Fig.
@@
 # Discussion
-[Ole thinks we should start bigger here to reflect the big main themes in the intro]
 
-For individuals deriving livelihoods from natural resources, theory suggests that adopting strategies that utilize
-a diversity of assets [species?] provide substantial benefits over strategies with lower diversity [REFS].
-However, the ability to effectively diversify in practice is limited by regulation and market forces.
-Here we document the effect of changing species diversity on revenue and revenue variability across 
-three scale. Each scale reflects individual level choices that produce varied levels of species diversity.  
-We find that individuals fishing permit sets that allow high levels
-of species diversity tend to have lower revenue variability. Adding an
-additional fisheries permit usually increases revenue and decreases revenue
-variability. However, within the constraints of a permit and accounting for
-effort, specializing (reducing species diversity) tends to increase revenue and
-specializing or generalizing tends to be associated with greater revenue
-variability, with considerable differences across fisheries. Thus, the most
-effective mechanism by which species diversity can lower revenue variability for
-individual fishers --- switching or adding fishing permits --- is often costly,
-limited by regulations, and consequently unavailable to many individuals in practice.
+For individuals deriving livelihoods from natural resources, theory suggests
+that adopting strategies with a diversity of assets provide substantial benefits
+over strategies with lower diversity [REFS]. However, the ability to effectively
+diversify in practice is limited by regulation and market forces. Here we
+document the effect of changing species diversity on revenue and revenue
+variability across three scale. Each scale reflects individual-level choices
+that produce varied levels of species diversity. We find that individuals
+fishing permit sets that allow high levels of species diversity tend to have
+lower revenue variability. Adding an additional fisheries permit usually
+increases revenue and decreases revenue variability. However, within the
+constraints of a permit and accounting for effort, specializing (reducing
+species diversity) tends to increase revenue and specializing or generalizing
+tends to be associated with greater revenue variability, with considerable
+differences across fisheries. Thus, the most effective mechanism by which
+species diversity can lower revenue variability for individual fishers ---
+switching or adding fishing permits --- is often costly, limited by regulations,
+and consequently unavailable to many individuals in practice.
 
-Our results shows additional support for the idea that individual fishers can reduce 
-their income variability by diversifying their catch portfolios. However, most individuals
-don’t appear to change their catch diversity significantly from year to year, and 
-most strategies adopted by Alaskan fishermen lie in a high-risk, high-reward quadrant
-in which reducing diversification increases both revenue variability and expected revenue (Fig. 4).
-Previous work has suggested that changes to fisheries management such as limited entry or
-catch share programs may inadvertently act as barrier to diversification and promote 
-specializing (Hilborn et al. 2003; Kasperski and Holland 2013). While some of these 
-specialized fishing strategies may be associated with increased revenue variability,
-the benefit may be that they are also associated with increased revenue.
-External market forces may interact in concert with management regulations to promote 
-specialization for fishers or other individuals dependent on natural resources. 
-In terrestrial agriculture for instance, there are also many instances of specializing 
-being profitable; drivers include the ability to specialize on high-profit cash crops
-(tobacco, cocoa), the availability of government agricultural subsidies (soybeans, corn in the USA),
-and farming on a large industrial scale. With increasing environmental variability and 
-extreme events in a changing climate, it is particularly important that individuals who 
-rely on natural resources as a primary source of income have strategies to reduce risk.
-Though variability in individual revenues may not be actively incorporated into many management
-decisions, analyses like ours help to understand the range of diversification strategies 
-available and how their benefits may change over time.
+<!-- TODO mention the quadrants in figure 4 -->
+
+Our results shows additional support for the idea that individual fishers can
+reduce their income variability by diversifying their catch portfolios. However,
+most individuals do not appear to change their catch diversity substantially
+from year to year, and most strategies adopted by Alaskan fishermen lie in
+a high-risk, high-reward quadrant in which reducing diversification increases
+both revenue variability and expected revenue. Previous work has suggested that
+changes to fisheries management such as limited entry or catch share programs
+may inadvertently act as barrier to diversification and promote specializing
+(Hilborn et al. 2003; Kasperski and Holland 2013). While some of these
+specialized fishing strategies may be associated with increased revenue
+variability, the benefit may be that they are also associated with increased
+revenue. External market forces may interact in concert with management
+regulations to promote specialization for fishers or other individuals dependent
+on natural resources. In terrestrial agriculture for instance, there are also
+many instances of specializing being profitable; drivers include the ability to
+specialize on high-profit cash crops (tobacco, cocoa), the availability of
+government agricultural subsidies (soybeans, corn in the USA), and farming on
+a large industrial scale. With increasing environmental variability and extreme
+events in a changing climate (REF), it is particularly important that
+individuals who rely on natural resources as a primary source of income have
+strategies to reduce risk. Though variability in individual revenues may not be
+actively incorporated into many management decisions, analyses like ours help to
+understand the range of diversification strategies available and how their
+benefits may change over time.
 
@@
 variability.
-
-##3 ERIC
 
@@
 
-Whereas a typical hierarchical linear regression model would estimate a single
-residual error scale term, $\sigma$, we model the scale of the residual error
-with another hierarchical model with a similar form:
+Whereas traditional hierarchical linear models assume homoscedasticity --- that
+the residual variance does not vary systematically --- our variance-function
+regression model allows the residual variance to be heteroscedastic and vary as
+a function of predictors. Instead of estimating a single residual error scale
+term, $\sigma$, we model the scale of the residual error with another
+hierarchical model with a similar form:
 

### 2016-12-16 portfolio/text/ms.Rmd
Edits

@@
 fishing commercially in Alaska from 1985--2014 (Fig. \ref{fig:ts}A). The
-majority (70--90%) of fishers hold single permits, with a slight increase in
-single-permit holders since 2000 (Fig. \ref{fig:ts}C). This overall trend is
-most sensitive to halibut or salmon permit holders --- compared to other
-fisheries, halibut permit holders are more likely to hold halibut plus other
-permits, whereas salmon permit holders are more likely to hold only one permit
-(Fig. \ref{fig:ts}C). Despite making up only 10--30% of the individuals,
-multiple permit holders earn approximately half the overall revenue (Fig.
-\ref{fig:ts}B). The mean effective fished species diversity of this revenue
-declined somewhat from 1980--2000 but has since remained steady (Fig.
-\ref{fig:ts}D). However, this trend appears to be driven by an increase in
-salmon specialists in the 1980s and 1990s --- the trend disappears if salmon
-license holders are excluded (Fig. \ref{fig:ts}D).
+majority (70--90%) hold single permits, with a slight increase in single-permit
+holders since 2000 (Fig. \ref{fig:ts}C). This overall trend is most sensitive to
+halibut or salmon permit holders --- compared to other fisheries, halibut permit
+holders are more likely to hold halibut plus other permits, whereas salmon
+permit holders are more likely to hold a single permit (Fig. \ref{fig:ts}C).
+Despite making up only 10--30% of the individuals, multiple permit holders earn
+approximately half the overall revenue (Fig. \ref{fig:ts}B). The mean effective
+species diversity of this revenue declined somewhat from 1980--2000 but has
+since remained steady (Fig. \ref{fig:ts}D). However, this trend appears largely
+driven by an increase in salmon specialists in the 1980s and 1990s --- the trend
+disappears if salmon license holders are excluded (Fig. \ref{fig:ts}D).
 
@@
 salmon. As examples, variability was highest for the herring roe gillnet and
-king crab fisheries. Most of the strategies with the lowest estimated revenue
+king crab fisheries. Most strategies with the lowest estimated revenue
 variability combined halibut with a salmon permit and had high levels of fished
@@
 from purchasing additional permits (Fig. \ref{fig:spokes}, Fig.
-S\ref{fig:spoke-all}). For example, individuals fishing a halibut longline
-permit plus one additional permit always had lower expected revenue variability
-compared to individuals fishing only a halibut longline permit (median reduction
-of `r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR], Fig.
+S\ref{fig:spoke-all}). For example, individuals fishing a halibut permit plus
+one additional permit always had lower expected revenue variability compared to
+individuals fishing only a halibut permit (median reduction of `r
+100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`% IQR], Fig.
 \ref{fig:spokes}A). Likewise, adding a permit to a herring roe gillnet strategy
@@
 themselves represent a significant expense (e.g. ~\$192K USD for a king and
-tanner crab permit in Southeast Alaska in 2016;
+tanner crab permit in Southeast Alaska in 2016 REF;
 <https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>).
@@
 \ref{fig:within}). After accounting for fishing effort, specializing tends to be
-associated with increased revenue (Fig. \ref{fig:within}; Fig. S\ref{fig:fe},
-Fig. S\ref{fig:re}) but also increased variability. On average, there is a ~0.5%
-increase in the year-to-year revenue ratio per 1% decrease in species diversity,
-and across strategies, `r g2_dens`% of the probability density supports
-a positive relationship between specialization and revenue.
+associated with increased revenue but also increased variability (Fig.
+\ref{fig:within}; Figs S\ref{fig:fe}, S\ref{fig:re}, S\ref{fig:noeffort}). On
+average, there is a ~0.5% increase in the year-to-year revenue ratio per 1%
+decrease in species diversity, and across strategies, `r g2_dens`% of the
+probability density supports a positive relationship between specialization and
+revenue.
 
-However, adjusting (increasing or decreasing) an individual's fished species
+<!--This benefit of specializing is most apparent after accounting for fishing
+effort ().-->
+
+However, changing (increasing or decreasing) an individual's fished species
 diversity within the constraints of a permit set tends to be associated with
 greater expected variability or uncertainty in year-to-year revenue (Fig.
-\ref{fig:within} strategies above 0; Figs. S\ref{fig:fe}, \ref{fig:re},
-\ref{fig:mirror}). This effect was strongest for individuals with a halibut
+\ref{fig:within} strategies above zero; Figs. S\ref{fig:fe}, S\ref{fig:re},
+S\ref{fig:mirror}). This effect was strongest for individuals with a halibut
 permit or halibut and sablefish permits. Individuals with these strategies
@@
 S\ref{fig:mirror}A), suggesting that the act of changing strategies is
-a contributor to increased variability of revenue. This benefit of specializing
-is most apparent after accounting for fishing effort (Fig. S\ref{fig:noeffort}).
+a contributor to increased variability of revenue.
+
+<!-- TODO mention the quadrants in figure 4 -->
 
 Alaskan halibut and sablefish fisheries were converted to individual fishery
-quota (IFQ) management in 1994, which potentially confounds our estimates of the
-effect of diversity and a major regulatory change. We refit our models allowing
-all strategy-level effects to vary separately before and after the introduction
-of IFQs for strategies involving halibut and sablefish permits. The introduction
-of IFQs reduced variability by ~25% (XX--XX 95% CI) for individuals with only
+quota (IFQ) management in 1994, which potentially confounds our estimates with
+a major regulatory change. Therefore, we refit our models allowing all
+strategy-level effects to vary separately before and after the introduction of
+IFQs for strategies involving halibut and sablefish permits. The introduction of
+IFQs reduced variability by ~25% (XX--XX 95% CI) for individuals with only
 halibut or sablefish permits (Fig. S\ref{fig:ifq}A). Furthermore, for most
@@
 For individuals deriving livelihoods from natural resources, theory suggests
-that adopting strategies with a diversity of assets provide substantial benefits
-over strategies with lower diversity [REFS]. However, the ability to effectively
-diversify in practice is limited by regulation and market forces. Here we
-document the effect of changing species diversity on revenue and revenue
+that adopting strategies with a diversity of assets provides substantial
+benefits over strategies with lower diversity [REFS]. However, the ability to
+effectively diversify in practice is limited by regulation and market forces.
+Here we document the effect of changing species diversity on revenue and revenue
 variability across three scale. Each scale reflects individual-level choices
@@
 and consequently unavailable to many individuals in practice.
-
-<!-- TODO mention the quadrants in figure 4 -->
 

### 2016-12-21 portfolio/text/ms.Rmd
Develop discussion

@@
 that adopting strategies with a diversity of assets provides substantial
-benefits over strategies with lower diversity [REFS]. However, the ability to
-effectively diversify in practice is limited by regulation and market forces.
-Here we document the effect of changing species diversity on revenue and revenue
-variability across three scale. Each scale reflects individual-level choices
-that produce varied levels of species diversity. We find that individuals
-fishing permit sets that allow high levels of species diversity tend to have
-lower revenue variability. Adding an additional fisheries permit usually
-increases revenue and decreases revenue variability. However, within the
-constraints of a permit and accounting for effort, specializing (reducing
-species diversity) tends to increase revenue and specializing or generalizing
-tends to be associated with greater revenue variability, with considerable
-differences across fisheries. Thus, the most effective mechanism by which
-species diversity can lower revenue variability for individual fishers ---
-switching or adding fishing permits --- is often costly, limited by regulations,
-and consequently unavailable to many individuals in practice.
+benefits over strategies with lower diversity [REFS]. However, the ability of
+individuals to effectively diversify is limited by regulation and market forces
+in practice. Here we document the effect of changing species diversity on
+revenue and revenue variability across three scale. Each scale reflects
+individual-level choices that can affect species diversity. We find that
+individuals with fishing permit sets that allow high levels of species diversity
+tend to have lower revenue variability. Adding a fisheries permit usually
+increases revenue and decreases revenue variability. Within the constraints of
+a permit and accounting for effort, specializing (reducing species diversity)
+tends to increase revenue while both specializing or generalizing tends to be
+associated with greater revenue variability - though there are considerable
+differences among fisheries. Thus, the most effective mechanism by which species
+diversity can lower revenue variability for individual fishers --- switching or
+adding fishing permits --- is costly, limited by regulations, and consequently
+unavailable to many individuals in practice.
 
-Our results shows additional support for the idea that individual fishers can
+TODO: cut apart and recombine the following couple chunks:
+
+Our results provide additional support for the idea that individual fishers can
 reduce their income variability by diversifying their catch portfolios. However,
 most individuals do not appear to change their catch diversity substantially
-from year to year, and most strategies adopted by Alaskan fishermen lie in
+from year to year [OLE: is there a supplemental figure or statistic that speaks
+to this point?], and most strategies adopted by Alaskan fishermen lie in
 a high-risk, high-reward quadrant in which reducing diversification increases
@@
 variability, the benefit may be that they are also associated with increased
-revenue. External market forces may interact in concert with management
-regulations to promote specialization for fishers or other individuals dependent
-on natural resources. In terrestrial agriculture for instance, there are also
-many instances of specializing being profitable; drivers include the ability to
-specialize on high-profit cash crops (tobacco, cocoa), the availability of
-government agricultural subsidies (soybeans, corn in the USA), and farming on
-a large industrial scale. With increasing environmental variability and extreme
-events in a changing climate (REF), it is particularly important that
-individuals who rely on natural resources as a primary source of income have
-strategies to reduce risk. Though variability in individual revenues may not be
-actively incorporated into many management decisions, analyses like ours help to
-understand the range of diversification strategies available and how their
-benefits may change over time.
+revenue.
 
-Our work compliments previous results that have examined the effects of species diversity at the 
-community (Ref) and vessel(Ref) level that show THIS (Sethi, Kasperski, others)
+Worldwide, fisheries confront the challenge of sustainable management (). While
+many fisheries are well managed and productive, many are also rebuilding from
+periods of over-exploitation (Worm et al. Science 2009). A common tactic to aid
+in fisheries rebuilding is to modify regulations to reduce the unintended
+harvest rebuilding stocks while allowing the continued harvest of healthy stocks
+(need REF). Such bycatch reduction is an important goal for both ecological
+(Crowder and Murawski 1998, Lewison et al. 2004) and economic (Curtis and Hicks
+2000; Chakravorty and Nemoto 2001) reasons. Cumulatively, though, regulatory
+changes for bycatch reduction as well as other reasons gradually lead to
+fisheries that target one or a few species and individual fishers wishing to
+participate in multiple fisheries must own (or purchase) multiple permits. Thus
+a general consequence of fisheries regulations that are shown to improve
+sustainability of fish stocks and avoid overfishing - including improved
+assessments (Hilborn and Ovando 2014) and individual quota systems (Costello et
+al. 2008 ) - may ultimately restrict individual diversification or at least make
+developing a diverse fishing portfolio more costly. Add Alaska example of
+IFQs...?
 
-### Reconciling the effects at different levels of species diversity
+<!-- Chakravorty, U. and Nemoto, K. (2001) Modeling the effects of area closure
+and tax policies: a spatial-temporal model of the hawaii longline fishery.
+Marine Resource Economics 15, 179–204.
 
-- why variability/uncertainty might increase with changes in diversity...
+Costello C, Gaines SD, Lynham J (2008) Can catch shares prevent fisheries
+collapse? Science 321(5896):1678–1681
 
-- and how fit in with @kasperski2013, @sethi2012b, @sethi2014
+Crowder, L.B. and Murawski, S.A. (1998) Fisheries bycatch: implications for
+management. Fisheries Management 23, 8–17.
 
-- part of the difference is individuals vs vessels. E.g. @kasperski2013
-  suggested that diversity declined following IFQs for halibut sablefish
-  vessels. The reason for the decline was that the vessels were allowed to fish
-  halibut for more days (because people were sharing boats). At the level of
-  people, we wouldn't expect the diversity of someone who only had
-  halibut/sablefish permit to be affected by IFQs. The person, not the vessel
-  holds the quota share: https://alaskafisheries.noaa.gov/fisheries/ifq
+Curtis, R. and Hicks, R.L. (2000) The Cost of sea turtle preservation: the case
+of Hawaii’s Pelagic longliners. American Journal of Agricultural Economics 82,
+1191– 1197.
 
-- year to year risky... But maybe evens out/better over time
+Hilborn R, Ovando D. 2014 Reflections on the success of traditional fisheries
+management. ICES J. Mar. Sci. 71, 1040–1046.
 
-- diversity within permit - often bycatch, limited room for intentional
-  diversification (unless you have a M or salmon permit)
+Lewison, R.L., Crowder, L.B., Rimprove fishing gear ead, A.J. and Freeman, S.A.
+(2004) Understanding impacts of fisheries bycatch on marine megafauna. Trends in
+Ecology & Evolution 19, 598–604.
 
-- risk results not really risk, but unpredictability given our covariates ...
-  don't change anything and it's most predictable how well you'll do. What's
-  unpredictable in our model, might be more predictable on a year to year,
-  species by species basis for individual permit holder
+Read, A.J., Drinker, P. and Northridge, S. (2006) Bycatch of marine mammals in
+US and global fisheries. Conservation Biology 20, 163–169. -->
 
-- CV vs year-to-year effect for one person
+<!-- Reconciling the different levels of effects -->
 
-Other:
+Our findings across fishing strategies are in line with previous findings
+[@kasperski2013, @sethi2012b, @sethi2014], but we also found that, for
+individuals, specializing tended to be associated with increased revenue and
+altering species diversity tended to be associated with increased revenue
+variability. Specializing has been shown to be profitable in some circumstances
+for other natural resource sectors. For example, while diversification is
+generally helpful for small-scale farmers, farmers that specialize on
+high-demand and high-value crops such as tobacco or cocoa can benefit, at least
+in the short term, from specializing on these crops (REF). In terms of variance,
+there are number of reasons by annual revenue variability might be expected to
+increase with changes to individual fished species diversity. First, changes to
+fishing patterns will sometimes be associated with a learning or adaptation
+phase as individuals become accustomed to new locations, new gear, or new
+species. Secondly, changes in diversity might reflect years in which targeted
+species were less available and fishers were forced to target other species or
+landed more bycatch. Most importantly, this increased variability technically
+reflects an inability for our large-scale model to predict an individual's
+revenue in a subsequent year. There are many reasons, often unique to an
+individual permit holder and therefore not included in a large-scale model, that
+might make that specific individual's revenue less predictable given a change in
+species diversity compared to not changing species diversity from one year to
+the next.
 
-- specializing can be profitable within some fisheries. Eric: analog here in
-  farming literature is that diversification for small scale farmers generally
-  good, except when they specialize on high demand/high value crops (tobacco,
-  cocoa). Seems like benefits of diversification are also smaller for
-  large-scale industrial farmers.
+<!--In terrestrial agriculture for instance, there are many instances of
+specializing being profitable; drivers include the ability to specialize on
+high-profit cash crops (tobacco, cocoa), the availability of government
+agricultural subsidies (soybeans, corn in the USA), and farming on a large
+industrial scale (REF).-->
 
-### Limitations + obvious extensions
+<!-- Limitations + obvious extensions paragraph -->
 
-There are a number of limitations to our model and extensions that could further
-our understanding of the role of diversification...
-
-- could vary by time and vary by mean diversity
-
-- spatial/geographic variation
-
-- nonlinear effects
-
-- response to specific events like EVOS
+Our analysis provides a synthetic view of the effects of species diversity on
+individuals over 30 years in one of the most productive fishing regions in the
+world. However, our analysis simultaneously uses data from across the broad
+geography of Alaska and over 30 years, potentially obscuring regional or
+time-variation in the consequences of diversification on revenue and revenue
+variability. We expect future research examining diversification over finer
+spatial and temporal scales will uncover additional interesting relationships.
+Furthermore, in addressing the overall pattern, we have averaged out several
+major discrete events (e.g., the Exxon Valdez oil spill in 1989) that had
+significant consequences for the ecosystem and social system in Alaska (general
+EVOS consequences REF). We do not wish to minimize the important of such events,
+but note that it is important to understand the benefits of diversification in
+the context of rare and extreme disturbances which occur in marine systems
+worldwide.
 
@@
 
-### Conclusions + recommendations
+<!-- Conclusions + recommendations
 
 - bring it out to wider than fisheries (farming, timber industry, ...)
-
 - 2 dimensions to think about (revenue and variability); and the trade-off is
@@
   species permits make a ton of money
-
 - broadly, diversification can provide a powerful form of protection from
@@
   adjusting ones strategy 
-
 - with increasing environmental variability and extremes due to climate change
@@
   buffer risk more than ever 
-
 - regulations that limit the ability to diversify can have other considerable
@@
   regulations when making management decisions 
+-->
 
-We fit a hierarchical variance-function regression model to 30 years of
-commercial fisheries revenue data in Alaska to simultaneously estimate the
-effect of changes in species diversity on individual's revenue and revenue
-variability.
+External market forces may interact in concert with management regulations to
+promote specialization for fishers or other individuals dependent on natural
+resources. With increasing environmental variability and extreme events in
+a changing climate (REF), it is particularly important that individuals who rely
+on natural resources as a primary source of income have strategies to reduce
+risk. Though variability in individual revenues may not be actively incorporated
+into many management decisions, analyses like ours help to understand the range
+of diversification strategies available and how their benefits may change over
+time.
 

### 2016-12-22 portfolio/text/ms.Rmd
Editing

@@
 
+Our findings across fishing strategies are in line with previous findings
+[@kasperski2013, @sethi2012b, @sethi2014], but we also found that within
+strategies, for individuals, specializing tended to be associated with increased
+revenue and altering species diversity tended to be associated with increased
+revenue variability. Specializing has been shown to be profitable in some
+circumstances for other natural resource sectors. For example, while
+diversification is generally helpful for small-scale farmers, farmers that
+specialize on high-demand and high-value crops such as tobacco or cocoa can
+benefit, at least in the short term, from specializing on these crops (REF). In
+terms of variance, there are number of reasons by annual revenue variability
+might be expected to increase with changes to individual fished species
+diversity. First, changes to fishing patterns will sometimes be associated with
+a learning or adaptation phase as individuals become accustomed to new
+locations, new gear, or new species. Secondly, changes in diversity might
+reflect years in which targeted species were less available and fishers were
+forced to target other species or landed more bycatch. Most importantly, this
+increased variability technically reflects an inability for our large-scale
+model to predict an individual's revenue in a subsequent year. There are many
+reasons, often unique to an individual permit holder and therefore not included
+in a large-scale model, that might make that specific individual's revenue less
+predictable given a change in species diversity compared to not changing species
+diversity from one year to the next.
+
+Our analysis provides a synthetic view of the effects of species diversity on
+individuals over 30 years in one of the most productive fishing regions in the
+world. However, this large temporal and geographical scope potentially obscures
+time or regional-variation in the consequences of diversification on revenue and
+revenue variability. We expect future research examining diversification over
+finer scales will uncover additional interesting relationships. Furthermore, in
+addressing the overall pattern, we have averaged out several major discrete
+events (e.g. the Exxon Valdez oil spill in 1989) that had significant
+consequences for the ecosystem and social system in Alaska (REF). In addition to
+variability, ... it is important to understand the benefits of diversification
+in the context of rare and extreme disturbances which occur in marine systems
+worldwide.
+
 TODO: cut apart and recombine the following couple chunks:
@@
 reduce their income variability by diversifying their catch portfolios. However,
-most individuals do not appear to change their catch diversity substantially
-from year to year [OLE: is there a supplemental figure or statistic that speaks
-to this point?], and most strategies adopted by Alaskan fishermen lie in
-a high-risk, high-reward quadrant in which reducing diversification increases
-both revenue variability and expected revenue. Previous work has suggested that
-changes to fisheries management such as limited entry or catch share programs
-may inadvertently act as barrier to diversification and promote specializing
+most individuals do not change their catch diversity substantially from year to
+year [OLE: is there a supplemental figure or statistic that speaks to this
+point?], and most strategies adopted by Alaskan fishermen lie in a high-risk,
+high-reward quadrant in which reducing diversification increases both revenue
+variability and expected revenue. Previous work has suggested that changes to
+fisheries management such as limited entry or catch share programs may
+inadvertently act as barrier to diversification and promote specializing
 (Hilborn et al. 2003; Kasperski and Holland 2013). While some of these
@@
 
-Our findings across fishing strategies are in line with previous findings
-[@kasperski2013, @sethi2012b, @sethi2014], but we also found that, for
-individuals, specializing tended to be associated with increased revenue and
-altering species diversity tended to be associated with increased revenue
-variability. Specializing has been shown to be profitable in some circumstances
-for other natural resource sectors. For example, while diversification is
-generally helpful for small-scale farmers, farmers that specialize on
-high-demand and high-value crops such as tobacco or cocoa can benefit, at least
-in the short term, from specializing on these crops (REF). In terms of variance,
-there are number of reasons by annual revenue variability might be expected to
-increase with changes to individual fished species diversity. First, changes to
-fishing patterns will sometimes be associated with a learning or adaptation
-phase as individuals become accustomed to new locations, new gear, or new
-species. Secondly, changes in diversity might reflect years in which targeted
-species were less available and fishers were forced to target other species or
-landed more bycatch. Most importantly, this increased variability technically
-reflects an inability for our large-scale model to predict an individual's
-revenue in a subsequent year. There are many reasons, often unique to an
-individual permit holder and therefore not included in a large-scale model, that
-might make that specific individual's revenue less predictable given a change in
-species diversity compared to not changing species diversity from one year to
-the next.
-
 <!--In terrestrial agriculture for instance, there are many instances of
@@
 <!-- Limitations + obvious extensions paragraph -->
-
-Our analysis provides a synthetic view of the effects of species diversity on
-individuals over 30 years in one of the most productive fishing regions in the
-world. However, our analysis simultaneously uses data from across the broad
-geography of Alaska and over 30 years, potentially obscuring regional or
-time-variation in the consequences of diversification on revenue and revenue
-variability. We expect future research examining diversification over finer
-spatial and temporal scales will uncover additional interesting relationships.
-Furthermore, in addressing the overall pattern, we have averaged out several
-major discrete events (e.g., the Exxon Valdez oil spill in 1989) that had
-significant consequences for the ecosystem and social system in Alaska (general
-EVOS consequences REF). We do not wish to minimize the important of such events,
-but note that it is important to understand the benefits of diversification in
-the context of rare and extreme disturbances which occur in marine systems
-worldwide.
 

### 2017-01-02 portfolio/text/ms.Rmd
Strip whitespace and edit discussion

@@
 that adopting strategies with a diversity of assets provides substantial
-benefits over strategies with lower diversity [REFS]. However, the ability of
-individuals to effectively diversify is limited by regulation and market forces
-in practice. Here we document the effect of changing species diversity on
-revenue and revenue variability across three scale. Each scale reflects
-individual-level choices that can affect species diversity. We find that
-individuals with fishing permit sets that allow high levels of species diversity
-tend to have lower revenue variability. Adding a fisheries permit usually
-increases revenue and decreases revenue variability. Within the constraints of
-a permit and accounting for effort, specializing (reducing species diversity)
-tends to increase revenue while both specializing or generalizing tends to be
-associated with greater revenue variability - though there are considerable
-differences among fisheries. Thus, the most effective mechanism by which species
-diversity can lower revenue variability for individual fishers --- switching or
-adding fishing permits --- is costly, limited by regulations, and consequently
-unavailable to many individuals in practice.
+benefits over strategies with lower diversity [REFS]. However, in practice, the
+ability of individuals to effectively diversify is limited by regulation and
+market forces. Here we document the effect of individual-level changes in
+species diversity on revenue and revenue variability across three scales. We
+find that (1) individuals with fishing permit sets that allow high levels of
+species diversity tend to have lower revenue variability. (2) Adding a fisheries
+permit usually increases revenue and decreases revenue variability. However, (3)
+within the constraints of a permit and accounting for effort, specializing tends
+to increase revenue while both specializing and generalizing tend to correlate
+with greater revenue variability with considerable differences amongst
+fisheries. Thus, the most effective mechanism by which species diversity can
+lower revenue variability for individual fishers --- switching or adding fishing
+permits --- is costly, limited by regulations, and consequently unavailable to
+many individuals.
 
@@
 specialize on high-demand and high-value crops such as tobacco or cocoa can
-benefit, at least in the short term, from specializing on these crops (REF). In
-terms of variance, there are number of reasons by annual revenue variability
-might be expected to increase with changes to individual fished species
-diversity. First, changes to fishing patterns will sometimes be associated with
-a learning or adaptation phase as individuals become accustomed to new
-locations, new gear, or new species. Secondly, changes in diversity might
-reflect years in which targeted species were less available and fishers were
-forced to target other species or landed more bycatch. Most importantly, this
-increased variability technically reflects an inability for our large-scale
-model to predict an individual's revenue in a subsequent year. There are many
-reasons, often unique to an individual permit holder and therefore not included
-in a large-scale model, that might make that specific individual's revenue less
-predictable given a change in species diversity compared to not changing species
-diversity from one year to the next.
+benefit, at least in the short term, from specializing (REF). In terms of
+variance, there are number of reasons by annual revenue variability might
+increase with changes to individual fished species diversity. First, changes to
+fishing patterns will sometimes be associated with a learning or adaptation
+phase as individuals become accustomed to new locations, new gear, or new
+species. Secondly, changes in diversity might reflect years in which targeted
+species were less available and fishers were forced to target other species or
+landed more bycatch. Most importantly, this increased variability technically
+reflects an inability for our large-scale model to predict an individual's
+revenue in a subsequent year. There are many reasons, often unique to an
+individual permit holder and therefore not included in a large-scale model, that
+might make that specific individual's revenue less predictable if species
+diversity changes.
 
@@
 on natural resources as a primary source of income have strategies to reduce
-risk. Though variability in individual revenues may not be actively incorporated
+risk. Though variability in individual revenue may not be actively incorporated
 into many management decisions, analyses like ours help to understand the range
@@
 <!-- vim: textwidth=80
+
 -->

### 2017-01-03 portfolio/text/ms.Rmd
Slash abstract

@@
 
-<!-- Diversified portfolios of fishing permits and species reduce risk for fishermen, -->
-<!-- but changing species diversity can be risky, costly, and limited by regulations. -->
-<!-- Goal: Introduction, Methods, Results, Discussion, Conclusion:
+<!-- Introduction, Methods, Results, Discussion, Conclusion -->
 <!-- 600, 900, 500, 800, 200, -->
@@
 
-<!-- TODO slash length of abstract -->
+<!-- TODO < 250 words -->
 
-A fundamental challenge for individuals who rely on the extraction of natural
-resources for their livelihood is income variability. Fishers, loggers, farmers,
-and miners, among others, face high levels of income uncertainty driven by a mix
-of environmental, biological, and economic variability. A key component
-underlying the sustainability of these industries is identifying both how
-regulatory actions and individual behaviour affect income variability and by
-extension the economic stability and sustainable use of natural resources. For
-commercial fisheries, recent work has identified that targeting a diversity of
-species can reduce income variability. However, any diversity-variability
-relationship is embedded in a regulatory framework that restricts individual
-ability to change fishing behaviour. Here we identify three primary axes by
-which fishers in Alaska, one of the world's most productive fishery regions, can
-potentially diversify fishing activities and investigate how individual actions
-along each diversity axis affect both revenue and revenue variability. We
-construct a novel statistical model to simultaneously assess the value to
-individual fishers of choosing to (1) own permits that allow targeting a diverse
-fish species assemblage, (2) diversify by owning multiple permits, or (3) catch
-a greater species diversity within the constraints of a given permit. We show
-contrasting effects of diversity on revenue and revenue variability along the
-three axes. First, we show that fishing permits with higher species diversity
-reduces variability (`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction
-per unit increase in mean effective species diversity). Next, we show that
-fishing an additional permit is associated with higher revenue and lower
-variability (e.g., `r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`%
-IQR] decline in standard deviation for fishers with halibut and another permit
-vs. only a halibut permit). Finally, we show that increasing species diversity
-within the constraints of a permit set has a fishery-dependent effect on revenue
-and is usually (`r g2_dens`% probability) associated with increased uncertainty
-in revenue the following year. Thus, while there are clear benefits to some axes
-of diversification, others are less effective. Importantly, the most effective
-mechanism for increasing revenue and decreasing risk --- purchasing additional
-permits to participate in additional fisheries --- was also the most costly.
-Furthermore, the least costly option of increasing catch diversity within an
-existing permit was associated with increased revenue variability. Thus, over
-the past 30 years as the Alaskan commercial fishing fleet has become
-increasingly specialized and most individuals only participate under a single
-permit, the practical benefits accrued from increased species diversification
-are limited.
+Individuals who rely on the extraction of natural resources for their
+livelihood --- fishers, farmers, and loggers, among others --- face high levels
+of income variability driven by a mix of environmental, biological, and
+economic variability.
+A key component to managing these industries is identifying how regulatory
+actions and individual behaviour affect income variability and by extension the
+economic stability and sustainable use of natural resources. For commercial
+fisheries, recent work has identified that communities and vessels can lower
+revenue variability by increasing the diversity of species fished. 
+<!-- fix: One sentence clearly stating the general problem being addressed by
+this particular study --> 
+However, it is unclear if ... extend to individual... year to year changes
+<!-- cut down, 1 sentence: -->
+Here we identify three axes by which fishers in Alaska, one of the world's most
+productive fishery regions, can potentially diversify fishing activities and
+investigate how individual actions along each diversity axis affect revenue and
+revenue variability. 
+<!-- We simultaneously assess the value to individual fishers of choosing to (1) own -->
+<!-- permits that allow targeting a diverse fish species assemblage, (2) diversify -->
+<!-- by adding additional permits, or (3) catch a greater species diversity within -->
+<!-- the constraints of a given permit. --> 
+We show that fishing permits with higher species diversity reduces revenue
+variability and fishing an additional permit is associated with higher revenue
+and lower variability. However, increasing species diversity within the
+constraints of a permit set has a fishery-dependent effect on revenue and is
+usually (`r g2_dens`% probability) associated with increased uncertainty in
+revenue the following year. 
+Our results suggest that the benefit of species diversity for reducing revenue
+variability is mostly an across-permit effect, not a benefit that can be easily
+harnessed by individual permit holders. The most effective mechanism for
+increasing revenue and decreasing variability --- purchasing permits to
+participate in additional fisheries --- is also the most costly and often
+limited by regulations such as limited entry and catch share programs.
 
-Our results suggest that the benefit of species diversity for revenue
-variability is mostly an across-permit effect, not a benefit that can be easily
-harnessed by an individual permit holder. Effective methods of diversification
---- purchasing additional permits or switching permits --- are costly and often
-limited by regulations such as limited entry programs and individual
-transferable quotas (ITQs). Take home ...
+<!-- Thus, over the past 30 years as the Alaskan commercial fishing fleet has become -->
+<!-- increasingly specialized, -->
 

### 2017-01-03 portfolio/text/ms.Rmd
More abstract editing

@@
 Individuals who rely on the extraction of natural resources for their
-livelihood --- fishers, farmers, and loggers, among others --- face high levels
-of income variability driven by a mix of environmental, biological, and
-economic variability.
+livelihood --- fishers, farmers, miners, and loggers, among others --- face
+high levels of income variability driven by a mix of environmental, biological,
+and economic variability.
 A key component to managing these industries is identifying how regulatory
@@
 Here we identify three axes by which fishers in Alaska, one of the world's most
-productive fishery regions, can potentially diversify fishing activities and
-investigate how individual actions along each diversity axis affect revenue and
-revenue variability. 
+productive fishing regions, can potentially diversify fishing activities and
+investigate how individual actions along each axis affect revenue and revenue
+variability. 
 <!-- We simultaneously assess the value to individual fishers of choosing to (1) own -->
@@
 constraints of a permit set has a fishery-dependent effect on revenue and is
-usually (`r g2_dens`% probability) associated with increased uncertainty in
-revenue the following year. 
-Our results suggest that the benefit of species diversity for reducing revenue
-variability is mostly an across-permit effect, not a benefit that can be easily
-harnessed by individual permit holders. The most effective mechanism for
-increasing revenue and decreasing variability --- purchasing permits to
-participate in additional fisheries --- is also the most costly and often
-limited by regulations such as limited entry and catch share programs.
+usually (`r g2_dens`% probability) associated with increased revenue uncertainty
+the following year. 
+Our results demonstrate that the most effective option for individuals to
+increase revenue and decrease variability is to purchase permits to participate
+in additional or more diverse fisheries. However, this option is costly
+and often limited by regulations such as limited entry and catch share
+programs.
+With increasing environmental variability in a changing climate, it will be
+particularly important that individuals who rely on natural resources for
+their livelihood have effective strategies to reduce income variability.
 

### 2017-01-03 portfolio/text/ms.Rmd
Fix bib; add outline

@@
 diversity changes.
+
+increased revenue with specialization? 
+why might someone specialize? but this has long-term risks 
+	efficiency 
+		and individual behavior important in fleet dynamics and efficiency of catch @hilborn1985
+		"Specialization, as measured by the share of total cultivated land devoted to rice production, also tends to lower inefficiency," @abdulai2000
+	cash crops 
+		sockeye e.g.
+		cocoa: most profitable crop export in Nigeria @mustapha1999
+			when cocoa prices were high, most cocoa farmers were highly specialized; now there is much more diversification to other sources of income...
+for example trading and electronic repairs @mustapha1999
+		tobacco: high profitability and low variability of income from tobacco mean that the optimal farming strategy (regardless of risk profile) is to specialize as much as possible on tobacco (given alternatives such as soybean, alfalfa, chickpeas, ...) -- great reference to use -- quadratic optimization @chavez2014
+	conversely, diversification can happen as a consequence of non-ideal situations 
+		poor fishing years mean the fishermen must seek out other options - this paper 
+		downwards economic shocks tend to be associated with increasing diversity for households relying on forests for income @debela2012
+		for example, many small farms use diversification to maximize revenue given their limited land @barbieri2008
+	but has long-term risks 
+		"...most graphic, recent example of such risks was the reliance of U.S.
+crop producers on export markets for wheat, corn, and soybeans during the 1970s.
+Many farmers borrowed large sums of money to buy additional land and buy specialized equipment ....
+
+		Specialized farmers producing export commodities were hardest hit by the financial crisis of American agriculture in the early 1980s...." @ikerd1990
+		"Historically, cocoa farmers rarely engaged in non-farm commercial or artisanal activities.
+When cocoa farmers took to non-farm occupations in the 1970s and 19803, it was as a result of the severe crisis of accumulation confronting them (Berry 1985).
+Indeed, only 5 per cent of the 1964 Alade sample had any non-farm occupation (Oloko 1964: 57).
+My 1992 study established that by the early 1980s, that is just before SAP, about 35 per cent of the Lower stratum in that survey, 30 per cent of the Middle, and 33 percent of the Upper had non‐farm occupations." @mustapha1999
+
+barriers to diversification 
+	cost of entry
+		"poorer households have fewer opportunities in non-cropping activities ...
+and hence less diversified incomes.
+...
+reflect[s] their relative lack of capital, which makes it difficult for them to diversify away from subsistence agriculture...
+" @abdulai2001
+		entry barriers mean that poor farmers tend to be less diversified and are forced to experience lower returns and higher variability in earnings @barrett2001
+			e.g.
+"Block and Webb and More income diversification associated with higher welfare measured in both income and nutritional terms in Ethiopia, and that the least diversified ft hard to increase their diversification over time.
+...
+Barrett et al.
+report a strong association between greater income diversification ˆ and higher wealth and income in Cote d’Ivoire and Kenya." @barrett2001
+		access to financial capital important to income diversification for fishing communities in Kenya @olale2012
+		overall, diversification was positively related with total income for those living in "forest margins" @illukpitiya2008
+	ifqs
+		catch shares may help prevent fisheries collapse @costello2008
+		entry expensive
+		diversification declined at a time following catch share programs on the West Coast @holland2016
+		vessels that entered West Coast fisheries less diversified than those that entered before IFQs (and of course cost is increased to get into these fisheries for those that were not already participants) @holland2016
+		catch share programs may reduce variability and offset the effects of lower diversity -- found that year-to-year vessel revenue decreased after catch shares (and average revenue increased for most) @holland2016
 
@@
 
-We fit our models with Stan 2.11.1 [@hoffman2014; @standevelopmentteam2015;
-@standevelopmentteam2015a; @carpenter2015] and R 3.3.1 [@rcoreteam2016]. Stan
+We fit our models with Stan 2.14.1 [@standevelopmentteam2016;
+@standevelopmentteam2016a] and R 3.3.1 [@rcoreteam2016]. Stan
 implements the No-U-Turn Hamiltonian Markov chain Monte Carlo algorithm
@@
 $\hat{R}<1.05$ (the potential scale reduction factor) and that the effective
-sample size, as calculated by the rstan R package [@standevelopmentteam2015],
+sample size, as calculated by the rstan R package [@standevelopmentteam2016],
 was greater than 200 for all parameters [@gelman2014]. We verified that our

### 2017-01-03 portfolio/text/ms.Rmd
Edits

@@
 revenue variability by increasing the diversity of species fished. 
-<!-- fix: One sentence clearly stating the general problem being addressed by
-this particular study --> 
 However, it is unclear if ... extend to individual... year to year changes
-<!-- cut down, 1 sentence: -->
 Here we identify three axes by which fishers in Alaska, one of the world's most
-productive fishing regions, can potentially diversify fishing activities and
+productive fishing regions, can diversify fishing activities and
 investigate how individual actions along each axis affect revenue and revenue
 variability. 
-<!-- We simultaneously assess the value to individual fishers of choosing to (1) own -->
-<!-- permits that allow targeting a diverse fish species assemblage, (2) diversify -->
-<!-- by adding additional permits, or (3) catch a greater species diversity within -->
-<!-- the constraints of a given permit. --> 
-We show that fishing permits with higher species diversity reduces revenue
-variability and fishing an additional permit is associated with higher revenue
-and lower variability. However, increasing species diversity within the
+We show that fishing permits with higher species diversity reduces individual
+revenue variability and fishing an additional permit is associated with higher
+revenue and lower variability. However, increasing species diversity within the
 constraints of a permit set has a fishery-dependent effect on revenue and is
-usually (`r g2_dens`% probability) associated with increased revenue uncertainty
-the following year. 
+usually (`r g2_dens`% probability) associated with increased revenue
+uncertainty the following year. 
 Our results demonstrate that the most effective option for individuals to
@@
 dependent on natural resources are also subject to biological and environmental
-variability. For example, drought and flooding are a major source of risk for
-agricultural food security and farmers' incomes [@gregory2005], and catastrophic
-disease outbreaks and wildfires increase risk for the logging industry
-[@kirilenko2007].
+variability [@smith1986]. For example, drought and flooding are a major source
+of risk for agricultural food security and farmers' incomes [@gregory2005], and
+catastrophic disease outbreaks and wildfires increase risk for the logging
+industry [@kirilenko2007].
 
@@
 there can be substantial benefits to specialization for both farmers and fishers
-[@abdulai2000; @arnason2005]. Furthermore, the majority of research examining
-the effects of diversification on revenue do so at aggregate levels that combine
-individuals (e.g. households, counties, or communities; [@kasperski2013;
-@sethi2014]). Thus, key uncertainties remain about the fundamental relationship
-between diversity, fishing revenue, and fishing revenue variability at the level
-of the individual.
+[@smith1986; @abdulai2000; @arnason2005]. Furthermore, the majority of research
+examining the effects of diversification on revenue do so at aggregate levels
+that combine individuals (e.g. counties or communities;
+[@kasperski2013; @sethi2014]). Thus, key uncertainties remain about the
+fundamental relationship between diversity, fishing revenue, and fishing
+revenue variability at the level of the individual.
 
@@
 world with ~2.7 billion kg landed biomass valued at nearly 2 billion USD
-annually [@nmfs2016]. Like much of the rest of the world, the days of open,
+annually [@nmfs2016]. The days of open,
 unregulated fisheries in most developed countries have passed --- fishers must
-own a license to participate in virtually all fisheries [@hilborn2003a]. And
-while such regulations have helped reduce overfishing, they limit the available
-strategies to reduce income variability. For example, many permits limit fishing
-to a single species, diversifying by purchasing additional fishing permits can
-be expensive (e.g. ~\$192K USD for a king and tanner crab permit in Southeast
-Alaska in 2016; <https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>), and some
-fisheries limit the entry of new participants through individual fishing quotas
-(IFQs) [@grimm2012].
+own a license to participate in virtually all fisheries [@hilborn2003a]. While
+such regulations have helped reduce overfishing, they limit the available
+strategies to reduce income variability. For example, many permits limit
+fishing to a single species, diversifying by purchasing additional fishing
+permits can be expensive (e.g. ~\$192K USD for a king and tanner crab permit in
+Southeast Alaska in 2016; <https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>),
+and some fisheries limit the entry of new participants through individual
+fishing quotas (IFQs) [@grimm2012].
 
@@
 we can disentangle fishers that share the same vessel. We develop and apply
-a hierarchical Bayesian model to understand the effect of fished species
-diversity [OLE: Call this diversity of catch rather than fished?] on individual
-revenue and revenue variability that builds on results
-from vessel [@kasperski2013], fleet [@sethi2012b], and community-level
+a hierarchical Bayesian model to understand the effect of fisheries catch
+species diversity on individual revenue and revenue variability that builds on
+results from vessel [@kasperski2013], fleet [@sethi2012b], and community-level
 [@sethi2014] variability.
@@
 Despite making up only 10--30% of the individuals, multiple permit holders earn
-approximately half the overall revenue (Fig. \ref{fig:ts}B). The mean effective
-species diversity of this revenue declined somewhat from 1980--2000 but has
-since remained steady (Fig. \ref{fig:ts}D). However, this trend appears largely
-driven by an increase in salmon specialists in the 1980s and 1990s --- the trend
-disappears if salmon license holders are excluded (Fig. \ref{fig:ts}D).
+approximately half the overall revenue (Fig. \ref{fig:ts}B). The mean species
+diversity of this revenue declined somewhat from 1980--2000 but has since
+remained steady (Fig. \ref{fig:ts}D). However, this trend appears largely
+driven by an increase in salmon specialists in the 1980s and 1990s --- the
+trend disappears if salmon license holders are excluded (Fig. \ref{fig:ts}D).
 
@@
 themselves represent a significant expense (e.g. ~\$192K USD for a king and
-tanner crab permit in Southeast Alaska in 2016 REF;
+tanner crab permit in Southeast Alaska in 2016
 <https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>).
@@
 strategy-level effects to vary separately before and after the introduction of
-IFQs for strategies involving halibut and sablefish permits. The introduction of
-IFQs reduced variability by ~25% (XX--XX 95% CI) for individuals with only
-halibut or sablefish permits (Fig. S\ref{fig:ifq}A). Furthermore, for most
+IFQs for strategies involving halibut and sablefish permits. The introduction
+of IFQs reduced variability by ~25% (TODO XX--XX 95% CI) for individuals with
+only halibut or sablefish permits (Fig. S\ref{fig:ifq}A). Furthermore, for most
 strategies, the association between changing species diversity and increased
@@
 
-### Implications for limited entry, ITQs, and catch shares
+<!-- ### Implications for limited entry, ITQs, and catch shares -->
 
-- concerning because in many areas of the world and fisheries, regulations and
-  market forces have encouraged specialization / limited (effective)
-  diversification (e.g. multiple permits or diverse permits)
+<!-- - concerning because in many areas of the world and fisheries, regulations and -->
+<!--   market forces have encouraged specialization / limited (effective) -->
+<!--   diversification (e.g. multiple permits or diverse permits) -->
 
-- locking into permits limits ability to pick lower risk strategies and is
-  getting people locked into low diversity permits that may be doing well now,
-  but not necessarily always
+<!-- - locking into permits limits ability to pick lower risk strategies and is -->
+<!--   getting people locked into low diversity permits that may be doing well now, -->
+<!--   but not necessarily always -->
 
@@
 \caption{Trends in revenue and diversification over time. (A) Map of fishing
-regions in Alaska, USA. (B) Total revenue in billions of 2009 U.S. dollars from
-`r range_yrs[1]` to `r range_yrs[2]` for multiple (grey) and single (blue)
-permit holders. (C) Proportion of single permit holders for all fisheries (thick
-black line) and with permit holders omitted each year who fished any of the `r
+regions in Alaska, USA. (B) Total revenue in billions of 2009 USD from `r
+range_yrs[1]` to `r range_yrs[2]` for multiple (grey) and single (blue) permit
+holders. (C) Proportion of single permit holders for all fisheries (thick black
+line) and with permit holders omitted each year who fished any of the `r
 n_jacknifed` top species groups (thin lines). (D) Mean effective diversity of
@@
 Commercial Fisheries Entry Commission (CFEC). These data represent a total of
-nearly `r round(ndat$initial_tot_rev/1e9)` billion 2009 USD and a combined `r
-round(ndat$initial/1e6,1)` million rows of data. We adjusted for inflation by
-setting all revenues to 2009 USD [@bea2016]. We implemented a number of
-filtering steps to focus on individuals actively engaged in commercial fishing
-(Supporting Information). In particular, we removed permit holders whose median
-revenue was less than \$10,000 2009 USD to focus on individuals for whom
-commercial fishing formed a substantial part of their income.
+nearly `r round(ndat$initial_tot_rev/1e9)` billion 2009 USD. We adjusted for
+inflation by setting all revenues to 2009 USD [@bea2016]. We implemented
+a number of filtering steps to focus on individuals actively engaged in
+commercial fishing (Supporting Information). In particular, we removed permit
+holders whose median revenue was less than \$10,000 2009 USD to focus on
+individuals for whom commercial fishing formed a substantial part of their
+income.
 
 We used combinations of state and federal fishing permits held by an individual
-in in a specific year to define fishing "strategies", representing the species
+in a specific year to define fishing "strategies", representing the species
 targeted (Supporting Information). For example, we combined permits that were
@@
 roe" permit group. For salmon permits, we formed salmon strategies based on
-similarities in gear type and species compositon (Supporting Information).
+similarities in gear type and species composition (Supporting Information).
 
@@
 revenue variability while controlling for effort. For consistency with previous
-analyses [@kasperski2013], we calculated effective species diversity for
-permit holder $i$ and time (year) $t$ as the inverse of Simpson's diversity
-index $\lambda$ [@simpson1949; @jost2006] (also referred to in economics as the
+analyses [@vedeld2007; @illukpitiya2008; @debela2012; @kasperski2013;
+sethi2014], we calculated effective species diversity for permit holder $i$ and
+time (year) $t$ as the inverse of Simpson's diversity index $\lambda$
+[@simpson1949; @jost2006] (also referred to in economics as the
 Herfindahl-Hirschman Index) weighted by revenue $R$:
@@
 
-\noindent Again, $S_{i,t}$ and $I_{i,t}$ represent, respectively, percent change
-in species diversity and an indicator variable for increasing or decreasing
-species diversity. We exponentiate the equation to ensure all scale parameters,
-$\sigma_{i,t}$, are positive. Residuals from the mean component of our model
-clearly show a "V"-shaped pattern that the variance component of our model can
-capture (Fig. S\ref{fig:resid}). Furthermore, the downside residuals
-(year-individual data points with lower than expected revenue) look
-qualitatively similar to the complete set of residuals (upside and downside)
-(Fig. S\ref{fig:resid-downside}). Therefore, it seemed appropriate to model the
-variability as a proxy for revenue risk.
+\noindent Again, $S_{i,t}$ and $I_{i,t}$ represent, respectively, the log ratio
+of species diversity from year-to-year and an indicator variable for increasing
+or decreasing species diversity. We exponentiate the equation to ensure all
+scale parameters, $\sigma_{i,t}$, are positive. Residuals from the mean
+component of our model clearly show a "V"-shaped pattern, which the variance
+component of our model can capture (Fig. S\ref{fig:resid}). Furthermore, the
+downside residuals (year-individual data points with lower than expected
+revenue) look qualitatively similar to the complete set of residuals (upside
+and downside) (Fig. S\ref{fig:resid-downside}). Therefore, it seemed
+appropriate to model the variability as a proxy for revenue risk.
 
@@
 
-\noindent The mean of the $\gamma_{0,j}$ coefficients is modelled with two
-strategy-level predictors. The symbols $M_j$ and $E_j$ represent the mean
-species diversity and mean combined fishing season length (effort) for strategy
-$j$. The coefficients $\eta_1$ and $\eta_2$ represent predictors in this
-strategy-level regression estimating effects across strategies.
+\noindent The $\gamma_{0,j}$ coefficients are modelled with two strategy-level
+predictors. The symbols $M_j$ and $E_j$ represent the mean species diversity
+and mean combined fishing season length (effort) for strategy $j$. The
+coefficients $\eta_1$ and $\eta_2$ represent predictors in this strategy-level
+regression estimating effects across strategies.
+
+<!-- TODO: rerun with 2.14.1 -->
 
 We fit our models with Stan 2.14.1 [@standevelopmentteam2016;
-@standevelopmentteam2016a] and R 3.3.1 [@rcoreteam2016]. Stan
-implements the No-U-Turn Hamiltonian Markov chain Monte Carlo algorithm
-[@hoffman2014] to perform Bayesian statistical inference. We assigned weakly
-informative priors on all parameters: $N(0, 2^2)$ priors on all $\beta$,
-$\gamma$, and $\eta$ parameters, and half-$t(3, 0, 2)$ priors (i.e. degrees of
-freedom of 3 and scale of 2) on all $\sigma$ and $\tau$. We ran four chains and
-2,000 iterations and discarded the first 1,000 iterations of each chain as warm
-up. We checked for chain convergence visually with trace plots, ensured that
-$\hat{R}<1.05$ (the potential scale reduction factor) and that the effective
-sample size, as calculated by the rstan R package [@standevelopmentteam2016],
-was greater than 200 for all parameters [@gelman2014]. We verified that our
-model returned sensible estimates by comparing our estimates to an ad-hoc
-two-stage model fit with the lme4 R package [@bates2015], where we first fit
-a mixed-effects regression model to the mean, and then modelled the absolute
-residuals from that first model in a second mixed-effects model representing the
-variance component. We also validated the model with simulated data, and were
-able to recover the true values (Supporting Information). Code used in our
-analysis is available at <https://github.com/NCEAS/pfx-commercial/portfolios/>.
+@standevelopmentteam2016a] and R 3.3.2 [@rcoreteam2016]. Stan implements the
+No-U-Turn Hamiltonian Markov chain Monte Carlo algorithm [@hoffman2014] to
+perform Bayesian statistical inference. We assigned weakly informative priors
+on all parameters: $N(0, 2^2)$ priors on all $\beta$, $\gamma$, and $\eta$
+parameters, and half-$t(3, 0, 2)$ priors (i.e. degrees of freedom of 3 and
+scale of 2) on all $\sigma$ and $\tau$. We ran four chains and 2,000 iterations
+and discarded the first 1,000 iterations of each chain as warm up. We checked
+for chain convergence visually with trace plots, ensured that $\hat{R}<1.05$
+(the potential scale reduction factor) and that the effective sample size, as
+calculated by the rstan R package [@standevelopmentteam2016], was greater than
+200 for all parameters [@gelman2014]. We verified that our model returned
+sensible estimates by comparing our estimates to an ad-hoc two-stage model fit
+with the lme4 R package [@bates2015], where we first fit a mixed-effects
+regression model to the mean, and then modelled the absolute residuals from
+that first model in a second mixed-effects model representing the variance
+component. We also validated the model with simulated data, and were able to
+recover the true values (Supporting Information). Code used in our analysis is
+available at <https://github.com/NCEAS/pfx-commercial/portfolios/>.
 

### 2017-01-03 portfolio/text/ms.Rmd
Develop discussion

@@
 high levels of income variability driven by a mix of environmental, biological,
-and economic variability.
-A key component to managing these industries is identifying how regulatory
-actions and individual behaviour affect income variability and by extension the
-economic stability and sustainable use of natural resources. For commercial
-fisheries, recent work has identified that communities and vessels can lower
-revenue variability by increasing the diversity of species fished. 
-However, it is unclear if ... extend to individual... year to year changes
+and economic variability. A key component to managing these industries is
+identifying how regulatory actions and individual behaviour affect income
+variability and by extension the economic stability and sustainable use of
+natural resources. For commercial fisheries, recent work has identified that
+communities and vessels that fish a greater diversity of species have less
+revenue variability than those that fish a lower species diversity. However, it
+is unclear if these benefits extend to the level of the individual and how
+year-to-year changes in diversification affect revenue and revenue variability.
 Here we identify three axes by which fishers in Alaska, one of the world's most
-productive fishing regions, can diversify fishing activities and
-investigate how individual actions along each axis affect revenue and revenue
-variability. 
+productive fishing regions, can diversify fishing activities and investigate
+how individual actions along each axis affect revenue and revenue variability.
 We show that fishing permits with higher species diversity reduces individual
@@
 usually (`r g2_dens`% probability) associated with increased revenue
-uncertainty the following year. 
-Our results demonstrate that the most effective option for individuals to
-increase revenue and decrease variability is to purchase permits to participate
-in additional or more diverse fisheries. However, this option is costly
-and often limited by regulations such as limited entry and catch share
-programs.
-With increasing environmental variability in a changing climate, it will be
-particularly important that individuals who rely on natural resources for
-their livelihood have effective strategies to reduce income variability.
+uncertainty the following year. Our results demonstrate that the most effective
+option for individuals to decrease revenue variability is to purchase permits
+to participate in additional or more diverse fisheries. However, this option is
+costly and often limited by regulations such as limited entry and catch share
+programs. With increasing environmental variability in a changing climate, it
+will be particularly important that individuals who rely on natural resources
+for their livelihood have effective strategies to reduce income variability.
 
-<!-- Thus, over the past 30 years as the Alaskan commercial fishing fleet has become -->
-<!-- increasingly specialized, -->
+<!-- TODO: fit something about specialization over time into the abstract.
+Thus, over the past 30 years as the Alaskan commercial fishing fleet has become
+increasingly specialized, -->
 
@@
 world with ~2.7 billion kg landed biomass valued at nearly 2 billion USD
-annually [@nmfs2016]. The days of open,
-unregulated fisheries in most developed countries have passed --- fishers must
-own a license to participate in virtually all fisheries [@hilborn2003a]. While
-such regulations have helped reduce overfishing, they limit the available
-strategies to reduce income variability. For example, many permits limit
-fishing to a single species, diversifying by purchasing additional fishing
-permits can be expensive (e.g. ~\$192K USD for a king and tanner crab permit in
-Southeast Alaska in 2016; <https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>),
-and some fisheries limit the entry of new participants through individual
-fishing quotas (IFQs) [@grimm2012].
+annually [@nmfs2016]. The days of open, unregulated fisheries in most developed
+countries have passed --- fishers must own a license to participate in
+virtually all fisheries [@hilborn2003a]. While such regulations have helped
+reduce overfishing, they limit the available strategies to reduce income
+variability. For example, many permits limit fishing to a single species,
+diversifying by purchasing additional fishing permits can be expensive (e.g.
+~\$192K USD for a king and tanner crab permit in Southeast Alaska in 2016;
+<https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>), and some fisheries limit
+the entry of new participants through individual fishing quotas (IFQs)
+[@grimm2012].
 
@@
 
-<!--This benefit of specializing is most apparent after accounting for fishing
-effort ().-->
-
 However, changing (increasing or decreasing) an individual's fished species
@@
 
-Our findings across fishing strategies are in line with previous findings
-[@kasperski2013, @sethi2012b, @sethi2014], but we also found that within
-strategies, for individuals, specializing tended to be associated with increased
-revenue and altering species diversity tended to be associated with increased
-revenue variability. Specializing has been shown to be profitable in some
-circumstances for other natural resource sectors. For example, while
-diversification is generally helpful for small-scale farmers, farmers that
-specialize on high-demand and high-value crops such as tobacco or cocoa can
-benefit, at least in the short term, from specializing (REF). In terms of
-variance, there are number of reasons by annual revenue variability might
-increase with changes to individual fished species diversity. First, changes to
-fishing patterns will sometimes be associated with a learning or adaptation
-phase as individuals become accustomed to new locations, new gear, or new
-species. Secondly, changes in diversity might reflect years in which targeted
-species were less available and fishers were forced to target other species or
-landed more bycatch. Most importantly, this increased variability technically
-reflects an inability for our large-scale model to predict an individual's
-revenue in a subsequent year. There are many reasons, often unique to an
-individual permit holder and therefore not included in a large-scale model, that
+Our results across fishing strategies are in line with previous findings
+[@kasperski2013, @sethi2012b, @sethi2014] at the community and vessel levels,
+but our results with respect to individual year-to-year behaviour provide new
+insights. Within strategies, specializing tended to be associated with
+increased revenue. This could be for a number of reasons. First, specialization
+can increase efficiency [@smith1986] and individual behaviour is an important
+component to the efficiency of catch [@hilborn1985]. Indeed, specialization can
+also increase efficiency in other natural resource industries, such as rice
+production [@abdulai2000]. Second, given the availability of "cash crops",
+specializing can be highly profitable. For example, when cocoa prices have been
+high, cocoa farmers in Nigeria have been highly specialized on the crop
+[@mustapha1999] and specializing on valuable tobacco farming can be the optimal
+strategy regardless of alternative crops [@chavez2014]. In our analyses,
+fishers with available sockeye salmon, a highly valuable species, benefited the
+most from specializing. However, specializing can have long-term risks if
+marker, biological, or environmental conditions change. For instance, U.S.
+farmers specialized on wheat, corn, or soybeans were hardest hit in the early
+1980s when a financial crisis hit [@ikerd1990]. Third, a lack of specialization
+can happen as a consequence of non-ideal circumstances such as poor
+availability of a more valuable fish stock in a given year. For example,
+downward economic shocks have been associated with increased diversity for
+households relying on forests for income [@debela2012].
+
+We also found that altering species diversity tended to be associated with
+increased revenue variability. There are number of reasons by annual revenue
+variability might increase with changes to individual fished species diversity.
+First, changes to fishing patterns will sometimes be associated with a learning
+or adaptation phase as individuals become accustomed to new locations, new
+gear, or new species [REF]. Secondly, changes in diversity might reflect years
+in which targeted species were less available and fishers were forced to target
+other species or landed more bycatch adding stochasticity to their revenue.
+Most importantly, this increased variability technically reflects an inability
+for our large-scale model to predict an individual's revenue in a subsequent
+year. There are many reasons, often unique to an individual permit holder's
+behaviour [@hilborn1985] and therefore not included in a large-scale model, that
 might make that specific individual's revenue less predictable if species
@@
 
-increased revenue with specialization? 
-why might someone specialize? but this has long-term risks 
-	efficiency 
-		and individual behavior important in fleet dynamics and efficiency of catch @hilborn1985
-		"Specialization, as measured by the share of total cultivated land devoted to rice production, also tends to lower inefficiency," @abdulai2000
-	cash crops 
-		sockeye e.g.
-		cocoa: most profitable crop export in Nigeria @mustapha1999
-			when cocoa prices were high, most cocoa farmers were highly specialized; now there is much more diversification to other sources of income...
-for example trading and electronic repairs @mustapha1999
-		tobacco: high profitability and low variability of income from tobacco mean that the optimal farming strategy (regardless of risk profile) is to specialize as much as possible on tobacco (given alternatives such as soybean, alfalfa, chickpeas, ...) -- great reference to use -- quadratic optimization @chavez2014
-	conversely, diversification can happen as a consequence of non-ideal situations 
-		poor fishing years mean the fishermen must seek out other options - this paper 
-		downwards economic shocks tend to be associated with increasing diversity for households relying on forests for income @debela2012
-		for example, many small farms use diversification to maximize revenue given their limited land @barbieri2008
-	but has long-term risks 
-		"...most graphic, recent example of such risks was the reliance of U.S.
-crop producers on export markets for wheat, corn, and soybeans during the 1970s.
-Many farmers borrowed large sums of money to buy additional land and buy specialized equipment ....
-
-		Specialized farmers producing export commodities were hardest hit by the financial crisis of American agriculture in the early 1980s...." @ikerd1990
-		"Historically, cocoa farmers rarely engaged in non-farm commercial or artisanal activities.
-When cocoa farmers took to non-farm occupations in the 1970s and 19803, it was as a result of the severe crisis of accumulation confronting them (Berry 1985).
-Indeed, only 5 per cent of the 1964 Alade sample had any non-farm occupation (Oloko 1964: 57).
-My 1992 study established that by the early 1980s, that is just before SAP, about 35 per cent of the Lower stratum in that survey, 30 per cent of the Middle, and 33 percent of the Upper had non‐farm occupations." @mustapha1999
-
-barriers to diversification 
-	cost of entry
-		"poorer households have fewer opportunities in non-cropping activities ...
-and hence less diversified incomes.
-...
-reflect[s] their relative lack of capital, which makes it difficult for them to diversify away from subsistence agriculture...
-" @abdulai2001
-		entry barriers mean that poor farmers tend to be less diversified and are forced to experience lower returns and higher variability in earnings @barrett2001
-			e.g.
-"Block and Webb and More income diversification associated with higher welfare measured in both income and nutritional terms in Ethiopia, and that the least diversified ft hard to increase their diversification over time.
-...
-Barrett et al.
-report a strong association between greater income diversification ˆ and higher wealth and income in Cote d’Ivoire and Kenya." @barrett2001
-		access to financial capital important to income diversification for fishing communities in Kenya @olale2012
-		overall, diversification was positively related with total income for those living in "forest margins" @illukpitiya2008
-	ifqs
-		catch shares may help prevent fisheries collapse @costello2008
-		entry expensive
-		diversification declined at a time following catch share programs on the West Coast @holland2016
-		vessels that entered West Coast fisheries less diversified than those that entered before IFQs (and of course cost is increased to get into these fisheries for those that were not already participants) @holland2016
-		catch share programs may reduce variability and offset the effects of lower diversity -- found that year-to-year vessel revenue decreased after catch shares (and average revenue increased for most) @holland2016
+There are many possible barriers to diversification for individuals relying on
+natural resources for their livelihood. A primary barrier is the cost of
+diversifying. For commercial fishers, purchasing additional licenses and gear
+typically represents a substantial cost. Related barriers are apparent in other
+natural resource sectors. For example, diversification has been positively
+linked with total income or access to financial capital for farmers
+[@abdulai2001; @barrett2001], small-scale fishing communities [@olale2012], and
+those relying on forest resources [@illukpitiya2008]. Another major barrier to
+diversification can be management restrictions. In commercial fisheries, while
+catch shares [@costello2008] and limited season lengths [REF] can be effective
+tools to avoid overfishing and distribute social capital, they can also limit
+opportunities for diversification [@smith1986; @kasperski2013]. However, catch
+shares, in particular, may also reduce income variability (Fig.
+S\ref{fig:ifq}A) perhaps offsetting the effects of reduced diversification
+[@holland2016].
 
@@
 world. However, this large temporal and geographical scope potentially obscures
-time or regional-variation in the consequences of diversification on revenue and
-revenue variability. We expect future research examining diversification over
-finer scales will uncover additional interesting relationships. Furthermore, in
+time or regional-variation in the consequences of diversification on revenue
+and revenue variability. We expect future research examining diversification
+over finer scales will uncover additional relationships. Furthermore, in
 addressing the overall pattern, we have averaged out several major discrete
-events (e.g. the Exxon Valdez oil spill in 1989) that had significant
-consequences for the ecosystem and social system in Alaska (REF). In addition to
-variability, ... it is important to understand the benefits of diversification
-in the context of rare and extreme disturbances which occur in marine systems
-worldwide.
+events, such as the Exxon Valdez oil spill in 1989, that had significant
+consequences for the ecological and social systems in Alaska [@paine1996]. In
+addition to revenue or income variability, it is important to understand the
+benefits of diversification in the context of rare and extreme disturbances
+which occur in natural systems worldwide [REF].
 
-TODO: cut apart and recombine the following couple chunks:
+<!-- TODO: cut apart and recombine the following couple chunks:-->
 
-Our results provide additional support for the idea that individual fishers can
+<!--Our results provide additional support for the idea that individual fishers can
 reduce their income variability by diversifying their catch portfolios. However,
@@
 variability, the benefit may be that they are also associated with increased
-revenue.
+revenue.-->
 
+<!--
 Worldwide, fisheries confront the challenge of sustainable management (). While
@@
 IFQs...?
+-->
 

### 2017-01-03 portfolio/text/ms.Rmd
Edits

@@
 Managing natural resources and sustaining the livelihoods of those that depend
-on their extraction is hard. These livelihoods tend to have high year-to-year
+on their extraction is hard. These livelihoods tend to have high annual
 variability in income relative to other professions [@mishra2002; @sethi2012b].
@@
 limit how individuals diversify, or incentivize against diversification. For
-example, in terrestrial agriculture, crop subsidies in the US may incentivize
+example, in terrestrial agriculture, crop subsidies in the U.S. may incentivize
 some farms to specialize (e.g. dairy, grains, soybeans; @mishra2002). In other
@@
 
-Our results across fishing strategies are in line with previous findings
-[@kasperski2013, @sethi2012b, @sethi2014] at the community and vessel levels,
-but our results with respect to individual year-to-year behaviour provide new
-insights. Within strategies, specializing tended to be associated with
-increased revenue. This could be for a number of reasons. First, specialization
-can increase efficiency [@smith1986] and individual behaviour is an important
-component to the efficiency of catch [@hilborn1985]. Indeed, specialization can
-also increase efficiency in other natural resource industries, such as rice
-production [@abdulai2000]. Second, given the availability of "cash crops",
-specializing can be highly profitable. For example, when cocoa prices have been
-high, cocoa farmers in Nigeria have been highly specialized on the crop
+Our results across fishing strategies are in line with previous findings at the
+community and vessel levels [@kasperski2013; @sethi2012b; @sethi2014], but our
+results with respect to individual year-to-year behaviour provide new insights.
+Within strategies, specializing tended to be associated with increased revenue.
+This could be for a number of reasons. First, specialization can increase
+fishing efficiency [@smith1986]. Efficiency gains from specialization can be
+seen in other natural resource sectors. For example, rice farmers in Ghana can
+increase their efficiency by devoting more land and labor to rice production
+compared to other crops [@abdulai2000]. Second, given the availability of "cash
+crops", specializing can be highly profitable. For instance, when cocoa prices
+have been high, farmers in Nigeria have been highly specialized on the crop
 [@mustapha1999] and specializing on valuable tobacco farming can be the optimal
 strategy regardless of alternative crops [@chavez2014]. In our analyses,
-fishers with available sockeye salmon, a highly valuable species, benefited the
-most from specializing. However, specializing can have long-term risks if
-marker, biological, or environmental conditions change. For instance, U.S.
-farmers specialized on wheat, corn, or soybeans were hardest hit in the early
-1980s when a financial crisis hit [@ikerd1990]. Third, a lack of specialization
-can happen as a consequence of non-ideal circumstances such as poor
-availability of a more valuable fish stock in a given year. For example,
-downward economic shocks have been associated with increased diversity for
-households relying on forests for income [@debela2012].
+fishers with the opportunity to fish highly valuable sockeye salmon benefited
+the most from specializing. However, specializing can have longterm risks if
+market, biological, or environmental conditions change. For example, U.S.
+farmers specialized on wheat, corn, or soybeans were hit harder than
+diversified farmers in the early 1980s when a financial crisis hit
+[@ikerd1990]. Third, a lack of specialization can happen as a consequence of
+non-ideal circumstances such as poor availability of a more valuable fish stock
+in a given year. For example, downward economic shocks have been associated
+with increased diversity for households relying on forests for income
+[@debela2012].
 
 We also found that altering species diversity tended to be associated with
-increased revenue variability. There are number of reasons by annual revenue
-variability might increase with changes to individual fished species diversity.
-First, changes to fishing patterns will sometimes be associated with a learning
-or adaptation phase as individuals become accustomed to new locations, new
-gear, or new species [REF]. Secondly, changes in diversity might reflect years
-in which targeted species were less available and fishers were forced to target
-other species or landed more bycatch adding stochasticity to their revenue.
-Most importantly, this increased variability technically reflects an inability
-for our large-scale model to predict an individual's revenue in a subsequent
-year. There are many reasons, often unique to an individual permit holder's
-behaviour [@hilborn1985] and therefore not included in a large-scale model, that
-might make that specific individual's revenue less predictable if species
-diversity changes.
+increased revenue variability. There are number of possible reasons for this
+relationship. For instance, changes to fishing patterns will sometimes be
+associated with a learning or adaptation phase as individuals become accustomed
+to new locations, new gear, or new species [REF]. Additionally, changes in
+diversity might reflect years in which targeted species were less available and
+fishers were forced to target other species or landed more bycatch adding
+stochasticity to their revenue. Most importantly, this increased variability
+reflects an inability for our large-scale model to predict an individual's
+revenue in a subsequent year. There are many reasons, often unique to an
+individual permit holder's behaviour [@hilborn1985] and therefore not included
+in our large-scale model, that might make that specific individual's revenue
+less predictable if species diversity changes.
 

### 2017-01-03 portfolio/text/ms.Rmd
Edits

@@
 communities and vessels that fish a greater diversity of species have less
-revenue variability than those that fish a lower species diversity. However, it
+revenue variability than those that fish a lower diversity. However, it
 is unclear if these benefits extend to the level of the individual and how
@@
 Here we identify three axes by which fishers in Alaska, one of the world's most
-productive fishing regions, can diversify fishing activities and investigate
-how individual actions along each axis affect revenue and revenue variability.
+productive fishing regions, can diversify fishing activities.
 We show that fishing permits with higher species diversity reduces individual
@@
 
-Using a unique dataset collected over 30 years (1985--2014) on individuals,
+Using a unique dataset collected over 30 years on individuals,
 rather than fleets or vessels, we examine how revenue and revenue variability of
@@
 [@sethi2014] variability.
-
 We develop a variance-function regression model [@davidian1987; @hall1989], to
@@
 species diversity across permit sets (Fig. \ref{fig:across}). Controlling for
-effort, strategies with one unit higher mean effective species diversity had on
+effort, strategies with one unit higher mean species diversity had on
 average `r h1[[2]]`% (`r h1[[1]]`--`r h1[[3]]`% 95% CI) lower variability. This
@@
 average, there is a ~0.5% increase in the year-to-year revenue ratio per 1%
-decrease in species diversity, and across strategies, `r g2_dens`% of the
+decrease in species diversity, and across strategies weighted equally, `r g2_dens`% of the
 probability density supports a positive relationship between specialization and
@@
 
-However, changing (increasing or decreasing) an individual's fished species
+Changing (increasing or decreasing) an individual's fished species
 diversity within the constraints of a permit set tends to be associated with
@@
 a major regulatory change. Therefore, we refit our models allowing all
-strategy-level effects to vary separately before and after the introduction of
-IFQs for strategies involving halibut and sablefish permits. The introduction
+strategy-level effects involving halibut and sablefish permits
+to vary separately before and after the introduction of IFQs. The introduction
 of IFQs reduced variability by ~25% (TODO XX--XX 95% CI) for individuals with
@@
 increase their efficiency by devoting more land and labor to rice production
-compared to other crops [@abdulai2000]. Second, given the availability of "cash
-crops", specializing can be highly profitable. For instance, when cocoa prices
-have been high, farmers in Nigeria have been highly specialized on the crop
+compared to other crops [@abdulai2000]. Second, given the availability of
+valuable resources, specializing can be highly profitable. For instance, when cocoa prices
+have been high, Nigerian farmers have tended to specialize on the crop
 [@mustapha1999] and specializing on valuable tobacco farming can be the optimal
@@
 diversified farmers in the early 1980s when a financial crisis hit
-[@ikerd1990]. Third, a lack of specialization can happen as a consequence of
+[@ikerd1990]. Third, generalization --- a lack of specialization ---
+can happen as a consequence of
 non-ideal circumstances such as poor availability of a more valuable fish stock
@@
 and revenue variability. We expect future research examining diversification
-over finer scales will uncover additional relationships. Furthermore, in
+over finer scales will uncover additional interesting relationships. Furthermore, in
 addressing the overall pattern, we have averaged out several major discrete
@@
 
+# Acknowledgments
+
+TODO. National Center for Ecological Analysis and Synthesis (NCEAS) working
+group *Applying portfolio effects to the Gulf of Alaska ecosystem: Did
+multi-scale diversity buffer against the Exxon Valdez oil spill?*. Other
+participants... Smith...
+
 <!-- # Notes -->

### 2017-01-03 portfolio/text/ms.Rmd
Clean up

@@
 # Introduction
-
-<!-- Papers to cite [@smith1986] -->
 
@@
 compared to other crops [@abdulai2000]. Second, given the availability of
-valuable resources, specializing can be highly profitable. For instance, when cocoa prices
-have been high, Nigerian farmers have tended to specialize on the crop
-[@mustapha1999] and specializing on valuable tobacco farming can be the optimal
-strategy regardless of alternative crops [@chavez2014]. In our analyses,
+valuable resources, specializing can be highly profitable. For instance, when
+cocoa prices have been high, Nigerian farmers have tended to specialize on the
+crop [@mustapha1999] and specializing on valuable tobacco farming can be the
+optimal strategy regardless of alternative crops [@chavez2014]. In our analyses,
 fishers with the opportunity to fish highly valuable sockeye salmon benefited
@@
 market, biological, or environmental conditions change. For example, U.S.
-farmers specialized on wheat, corn, or soybeans were hit harder than
-diversified farmers in the early 1980s when a financial crisis hit
-[@ikerd1990]. Third, generalization --- a lack of specialization ---
-can happen as a consequence of
+farmers specialized on wheat, corn, or soybeans were hit harder than diversified
+farmers in the early 1980s when a financial crisis hit [@ikerd1990]. Third,
+generalization --- a lack of specialization --- can happen as a consequence of
 non-ideal circumstances such as poor availability of a more valuable fish stock
-in a given year. For example, downward economic shocks have been associated
-with increased diversity for households relying on forests for income
-[@debela2012].
+in a given year. For example, downward economic shocks have been associated with
+increased diversity for households relying on forests for income [@debela2012].
 
@@
 
-<!-- TODO: cut apart and recombine the following couple chunks:-->
-
-<!--Our results provide additional support for the idea that individual fishers can
-reduce their income variability by diversifying their catch portfolios. However,
-most individuals do not change their catch diversity substantially from year to
-year [OLE: is there a supplemental figure or statistic that speaks to this
-point?], and most strategies adopted by Alaskan fishermen lie in a high-risk,
-high-reward quadrant in which reducing diversification increases both revenue
-variability and expected revenue. Previous work has suggested that changes to
-fisheries management such as limited entry or catch share programs may
-inadvertently act as barrier to diversification and promote specializing
-(Hilborn et al. 2003; Kasperski and Holland 2013). While some of these
-specialized fishing strategies may be associated with increased revenue
-variability, the benefit may be that they are also associated with increased
-revenue.-->
-
 <!--
@@
 -->
-
-<!-- Chakravorty, U. and Nemoto, K. (2001) Modeling the effects of area closure
-and tax policies: a spatial-temporal model of the hawaii longline fishery.
-Marine Resource Economics 15, 179–204.
-
-Costello C, Gaines SD, Lynham J (2008) Can catch shares prevent fisheries
-collapse? Science 321(5896):1678–1681
-
-Crowder, L.B. and Murawski, S.A. (1998) Fisheries bycatch: implications for
-management. Fisheries Management 23, 8–17.
-
-Curtis, R. and Hicks, R.L. (2000) The Cost of sea turtle preservation: the case
-of Hawaii’s Pelagic longliners. American Journal of Agricultural Economics 82,
-1191– 1197.
-
-Hilborn R, Ovando D. 2014 Reflections on the success of traditional fisheries
-management. ICES J. Mar. Sci. 71, 1040–1046.
-
-Lewison, R.L., Crowder, L.B., Rimprove fishing gear ead, A.J. and Freeman, S.A.
-(2004) Understanding impacts of fisheries bycatch on marine megafauna. Trends in
-Ecology & Evolution 19, 598–604.
-
-Read, A.J., Drinker, P. and Northridge, S. (2006) Bycatch of marine mammals in
-US and global fisheries. Conservation Biology 20, 163–169. -->
-
-<!-- Reconciling the different levels of effects -->
-
-<!--In terrestrial agriculture for instance, there are many instances of
-specializing being profitable; drivers include the ability to specialize on
-high-profit cash crops (tobacco, cocoa), the availability of government
-agricultural subsidies (soybeans, corn in the USA), and farming on a large
-industrial scale (REF).-->
-
-<!-- Limitations + obvious extensions paragraph -->
-
-<!-- ### Implications for limited entry, ITQs, and catch shares -->
-
-<!-- - concerning because in many areas of the world and fisheries, regulations and -->
-<!--   market forces have encouraged specialization / limited (effective) -->
-<!--   diversification (e.g. multiple permits or diverse permits) -->
-
-<!-- - locking into permits limits ability to pick lower risk strategies and is -->
-<!--   getting people locked into low diversity permits that may be doing well now, -->
-<!--   but not necessarily always -->
 
@@
 
+<!--Our results provide additional support for the idea that individual fishers can
+reduce their income variability by diversifying their catch portfolios. However,
+most individuals do not change their catch diversity substantially from year to
+year [OLE: is there a supplemental figure or statistic that speaks to this
+point?], and most strategies adopted by Alaskan fishermen lie in a high-risk,
+high-reward quadrant in which reducing diversification increases both revenue
+variability and expected revenue. Previous work has suggested that changes to
+fisheries management such as limited entry or catch share programs may
+inadvertently act as barrier to diversification and promote specializing
+(Hilborn et al. 2003; Kasperski and Holland 2013). While some of these
+specialized fishing strategies may be associated with increased revenue
+variability, the benefit may be that they are also associated with increased
+revenue.-->
+
+TODO: finish this:
 External market forces may interact in concert with management regulations to
@@
 
-<!-- # Notes -->
-
-<!-- ## @kasperski2013 -->
-<!-- - vessel-level CFEC data -->
-<!-- - dome-shaped result -->
-<!-- - Herfindahl-Hirschman Index/HHI (Simpson) -->
-<!-- - diversity decreasing for those still fishing in 2010, but not otherwise -->
-
-<!-- ## @sethi2012b -->
-<!-- - risk measures to 90 Alaskan fishing fishery/fleets -->
-<!-- - revenue more variable than catch (+ price = extra variability) -->
-<!-- - large catches, big fleets = lower risk -->
-
-<!-- ## @sethi2010b -->
-<!-- - risk management for fisheries review -->
-
-<!-- ## @sethi2012 -->
-<!-- - equal-share catch-pooling cooperatives in king and snow crab -->
-<!-- - "variability reduction depends on the degree of correlation amongst -->
-<!--   participants’ catches" -->
-
-<!-- ## @sethi2014 -->
-<!-- - community-level CFEC data -->
-<!-- - number of fisheries strongest predictor of (lower) CV/CVaR -->
-<!-- - also simpson index -->
-<!-- - suggest barriers to commercial entry increase community risk by restricting -->
-<!--   access to 'variety of fisheries' -->
-<!-- - some communities have limited ability to diversify; may need to diversify -->
-<!--   outside of fishing -->
-
-<!-- [@perruso2005] -->
-<!-- [@hilborn2001] -->
-<!-- [@edwards2004] -->
-<!-- [@trenkel2013] -->
-<!-- [@mumford2009] -->
-<!-- [@oostenbrugge2002] -->
-<!-- [@baldursson1997] -->
-<!-- [@jardine2015] -->
-<!-- [@sanchirico2008] -->
-
-<!-- vim: textwidth=80
-
--->
+<!-- vim: textwidth=80 -->

### 2017-01-04 portfolio/text/ms.Rmd
Edits

@@
 
-Individuals who rely on the extraction of natural resources for their
-livelihood --- fishers, farmers, miners, and loggers, among others --- face
-high levels of income variability driven by a mix of environmental, biological,
-and economic variability. A key component to managing these industries is
-identifying how regulatory actions and individual behaviour affect income
-variability and by extension the economic stability and sustainable use of
-natural resources. For commercial fisheries, recent work has identified that
-communities and vessels that fish a greater diversity of species have less
-revenue variability than those that fish a lower diversity. However, it
-is unclear if these benefits extend to the level of the individual and how
-year-to-year changes in diversification affect revenue and revenue variability.
-Here we identify three axes by which fishers in Alaska, one of the world's most
-productive fishing regions, can diversify fishing activities.
-We show that fishing permits with higher species diversity reduces individual
-revenue variability and fishing an additional permit is associated with higher
-revenue and lower variability. However, increasing species diversity within the
-constraints of a permit set has a fishery-dependent effect on revenue and is
-usually (`r g2_dens`% probability) associated with increased revenue
-uncertainty the following year. Our results demonstrate that the most effective
-option for individuals to decrease revenue variability is to purchase permits
-to participate in additional or more diverse fisheries. However, this option is
-costly and often limited by regulations such as limited entry and catch share
-programs. With increasing environmental variability in a changing climate, it
-will be particularly important that individuals who rely on natural resources
-for their livelihood have effective strategies to reduce income variability.
+Individuals who rely on the extraction of natural resources for their livelihood
+--- fishers, farmers, miners, and loggers, among others --- face high levels of
+income variability driven by a mix of environmental, biological, and economic
+variability. A key component to managing these industries is identifying how
+regulatory actions and individual behaviour affect income variability and by
+extension the economic stability and sustainable use of natural resources. For
+commercial fisheries, recent work has identified that communities and vessels
+that fish a greater diversity of species have less revenue variability than
+those that fish a lower diversity. However, it is unclear if these benefits
+extend to the level of the individual and how year-to-year changes in
+diversification affect revenue and revenue variability. Here we identify three
+axes by which fishers in Alaska, one of the world's most productive fishing
+regions, can diversify fishing activities. We show that fishing permits with
+higher species diversity reduces individual revenue variability and fishing an
+additional permit is associated with higher revenue and lower variability.
+However, increasing species diversity within the constraints of existing permits
+has a fishery-dependent effect on revenue and is usually (`r g2_dens`%
+probability) associated with increased revenue uncertainty the following year.
+Our results demonstrate that the most effective option for individuals to
+decrease revenue variability is to purchase permits to participate in additional
+or more diverse fisheries. However, this option is costly and often limited by
+regulations such as limited entry and catch share programs. With increasing
+environmental variability in a changing climate, it will be particularly
+important that individuals who rely on natural resources for their livelihood
+have effective strategies to reduce income variability.
 
@@
 examining the effects of diversification on revenue do so at aggregate levels
-that combine individuals (e.g. counties or communities;
-[@kasperski2013; @sethi2014]). Thus, key uncertainties remain about the
-fundamental relationship between diversity, fishing revenue, and fishing
-revenue variability at the level of the individual.
+that combine individuals (e.g. counties or communities; [@kasperski2013;
+@sethi2014]). Thus, key uncertainties remain about the fundamental relationship
+between diversity, fishing revenue, and fishing revenue variability at the level
+of the individual.
 
@@
 annually [@nmfs2016]. The days of open, unregulated fisheries in most developed
-countries have passed --- fishers must own a license to participate in
-virtually all fisheries [@hilborn2003a]. While such regulations have helped
-reduce overfishing, they limit the available strategies to reduce income
-variability. For example, many permits limit fishing to a single species,
-diversifying by purchasing additional fishing permits can be expensive (e.g.
-~\$192K USD for a king and tanner crab permit in Southeast Alaska in 2016;
-<https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>), and some fisheries limit
-the entry of new participants through individual fishing quotas (IFQs)
+countries have passed --- fishers must own a license to participate in virtually
+all fisheries [@hilborn2003a]. While such regulations have helped reduce
+overfishing, they limit the available strategies to reduce income variability.
+For example, many permits limit fishing to a single species, diversifying by
+purchasing additional fishing permits can be expensive[^1], and some fisheries
+limit the entry of new participants through individual fishing quotas (IFQs)
 [@grimm2012].
 
-Using a unique dataset collected over 30 years on individuals,
-rather than fleets or vessels, we examine how revenue and revenue variability of
+[^1]: For example, it cost ~\$192K USD for a king and tanner crab permit in
+Southeast Alaska in 2016: <https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>.
+
+Using a unique dataset collected over 30 years on individual fishers, rather
+than fleets or vessels, we examine how revenue and revenue variability of
 individuals is affected by diversification. By tracking individuals we can track
@@
 results from vessel [@kasperski2013], fleet [@sethi2012b], and community-level
-[@sethi2014] variability.
-We develop a variance-function regression model [@davidian1987; @hall1989], to
-simultaneously models changes in the mean and variance of individual revenue as
-a function of covariates. We simultaneously estimate the effect of species
-diversity (the inverse of Simpson's diversity index [@simpson1949; @jost2006])
-on individual revenue and revenue variability, and estimate effects both within
-and across fishing strategies while controlling for fishing effort and latent
-time effects (e.g. changing environmental conditions). To our knowledge, this is
-the first application of this type of model in any natural resource field.
+[@sethi2014] variability. We develop a variance-function regression model
+[@davidian1987; @hall1989], to simultaneously model changes in the mean and
+variance of individual revenue as a function of covariates. We simultaneously
+estimate the effect of species diversity (the inverse of Simpson's diversity
+index [@simpson1949; @jost2006]) on individual revenue and revenue variability,
+and estimate effects both within and across fishing strategies while controlling
+for fishing effort and latent time effects (e.g. changing environmental
+conditions). To our knowledge, this is the first application of this type of
+model in any natural resource field.
 
@@
 fishing commercially in Alaska from 1985--2014 (Fig. \ref{fig:ts}A). The
-majority (70--90%) hold single permits, with a slight increase in single-permit
+majority (70--90%) hold a single permit, with a slight increase in single-permit
 holders since 2000 (Fig. \ref{fig:ts}C). This overall trend is most sensitive to
@@
 permit holders are more likely to hold a single permit (Fig. \ref{fig:ts}C).
-Despite making up only 10--30% of the individuals, multiple permit holders earn
+Despite making up only 10--30% of the individuals, multiple-permit holders earn
 approximately half the overall revenue (Fig. \ref{fig:ts}B). The mean species
 diversity of this revenue declined somewhat from 1980--2000 but has since
-remained steady (Fig. \ref{fig:ts}D). However, this trend appears largely
-driven by an increase in salmon specialists in the 1980s and 1990s --- the
-trend disappears if salmon license holders are excluded (Fig. \ref{fig:ts}D).
+remained steady (Fig. \ref{fig:ts}D). However, this trend appears largely driven
+by an increase in salmon specialists in the 1980s and 1990s --- the trend
+disappears if salmon license holders are excluded (Fig. \ref{fig:ts}D).
 
@@
 species diversity across permit sets (Fig. \ref{fig:across}). Controlling for
-effort, strategies with one unit higher mean species diversity had on
-average `r h1[[2]]`% (`r h1[[1]]`--`r h1[[3]]`% 95% CI) lower variability. This
-pattern looks qualitatively similar with or without individuals targeting
-salmon. As examples, variability was highest for the herring roe gillnet and
-king crab fisheries. Most strategies with the lowest estimated revenue
-variability combined halibut with a salmon permit and had high levels of fished
-species diversity (Fig. \ref{fig:across}).
+effort, strategies with one unit higher mean species diversity had on average `r
+h1[[2]]`% (`r h1[[1]]`--`r h1[[3]]`% 95% CI) lower variability. This pattern
+looks qualitatively similar with or without individuals targeting salmon. As
+examples, variability was highest for the herring roe gillnet and king crab
+fisheries. Most strategies with the lowest estimated revenue variability
+combined halibut with a salmon permit and had high levels of fished species
+diversity (Fig. \ref{fig:across}).
 
@@
 Adding permits also usually increased expected revenue, although the permits
-themselves represent a significant expense (e.g. ~\$192K USD for a king and
-tanner crab permit in Southeast Alaska in 2016
-<https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>).
+themselves represent a significant expense.
 
@@
 average, there is a ~0.5% increase in the year-to-year revenue ratio per 1%
-decrease in species diversity, and across strategies weighted equally, `r g2_dens`% of the
-probability density supports a positive relationship between specialization and
-revenue.
+decrease in species diversity, and across strategies weighted equally, `r
+g2_dens`% of the probability density supports a positive relationship between
+specialization and revenue.
 
-Changing (increasing or decreasing) an individual's fished species
-diversity within the constraints of a permit set tends to be associated with
-greater expected variability or uncertainty in year-to-year revenue (Fig.
+Changing (increasing or decreasing) an individual's fished species diversity
+within the constraints of a permit set tends to be associated with greater
+expected variability or uncertainty in year-to-year revenue (Fig.
 \ref{fig:within} strategies above zero; Figs. S\ref{fig:fe}, S\ref{fig:re},
@@
 a major regulatory change. Therefore, we refit our models allowing all
-strategy-level effects involving halibut and sablefish permits
-to vary separately before and after the introduction of IFQs. The introduction
-of IFQs reduced variability by ~25% (TODO XX--XX 95% CI) for individuals with
-only halibut or sablefish permits (Fig. S\ref{fig:ifq}A). Furthermore, for most
+strategy-level effects involving halibut and sablefish permits to vary
+separately before and after the introduction of IFQs. The introduction of IFQs
+reduced variability by ~25% (TODO XX--XX 95% CI) for individuals with only
+halibut or sablefish permits (Fig. S\ref{fig:ifq}A). Furthermore, for most
 strategies, the association between changing species diversity and increased
@@
 For individuals deriving livelihoods from natural resources, theory suggests
-that adopting strategies with a diversity of assets provides substantial
+that adopting strategies that exploit a diversity of assets provides substantial
 benefits over strategies with lower diversity [REFS]. However, in practice, the
@@
 diversity might reflect years in which targeted species were less available and
-fishers were forced to target other species or landed more bycatch adding
+fishers were forced to target other species or landed more bycatch, adding
 stochasticity to their revenue. Most importantly, this increased variability
@@
 opportunities for diversification [@smith1986; @kasperski2013]. However, catch
-shares, in particular, may also reduce income variability (Fig.
-S\ref{fig:ifq}A) perhaps offsetting the effects of reduced diversification
-[@holland2016].
+shares, in particular, may also reduce income variability (Fig. S\ref{fig:ifq}A)
+perhaps offsetting the effects of reduced diversification [@holland2016].
 
@@
 world. However, this large temporal and geographical scope potentially obscures
-time or regional-variation in the consequences of diversification on revenue
-and revenue variability. We expect future research examining diversification
-over finer scales will uncover additional interesting relationships. Furthermore, in
+time or regional-variation in the consequences of diversification on revenue and
+revenue variability. We expect future research examining diversification over
+finer scales will uncover additional interesting relationships. Furthermore, in
 addressing the overall pattern, we have averaged out several major discrete
@@
 
-<!-- vim: textwidth=80 -->
+<!-- vim: textwidth=80 
+-->

### 2017-01-04 portfolio/text/ms.Rmd
Edit to clarify

@@
 Individual revenue variability decreases exponentially with higher mean fished
-species diversity across permit sets (Fig. \ref{fig:across}). Controlling for
-effort, strategies with one unit higher mean species diversity had on average `r
-h1[[2]]`% (`r h1[[1]]`--`r h1[[3]]`% 95% CI) lower variability. This pattern
-looks qualitatively similar with or without individuals targeting salmon. As
-examples, variability was highest for the herring roe gillnet and king crab
-fisheries. Most strategies with the lowest estimated revenue variability
-combined halibut with a salmon permit and had high levels of fished species
-diversity (Fig. \ref{fig:across}).
+species diversity across combinations of state and federal fishing permits
+("strategies"; Fig. \ref{fig:across}). Controlling for effort, strategies with
+one unit higher mean species diversity had on average `r h1[[2]]`% (`r
+h1[[1]]`--`r h1[[3]]`% 95% CI) lower variability. This pattern looks
+qualitatively similar with or without individuals targeting salmon. As examples,
+variability was highest for the herring roe gillnet and king crab fisheries.
+Most strategies with the lowest estimated revenue variability combined halibut
+with a salmon permit and had high levels of fished species diversity (Fig.
+\ref{fig:across}).
 
@@
 and revenue variability, the effect of changing fished species diversity for an
-individual *within* the constraints of a given permit set is more variable (Fig.
-\ref{fig:within}). After accounting for fishing effort, specializing tends to be
-associated with increased revenue but also increased variability (Fig.
+individual *within* the constraints of a given set of permits is more variable
+(Fig. \ref{fig:within}). After accounting for fishing effort, specializing tends
+to be associated with increased revenue but also increased variability (Fig.
 \ref{fig:within}; Figs S\ref{fig:fe}, S\ref{fig:re}, S\ref{fig:noeffort}). On
@@
 Changing (increasing or decreasing) an individual's fished species diversity
-within the constraints of a permit set tends to be associated with greater
+within the constraints of a set of permits tends to be associated with greater
 expected variability or uncertainty in year-to-year revenue (Fig.
@@
 species diversity on revenue and revenue variability across three scales. We
-find that (1) individuals with fishing permit sets that allow high levels of
-species diversity tend to have lower revenue variability. (2) Adding a fisheries
-permit usually increases revenue and decreases revenue variability. However, (3)
-within the constraints of a permit and accounting for effort, specializing tends
-to increase revenue while both specializing and generalizing tend to correlate
-with greater revenue variability with considerable differences amongst
-fisheries. Thus, the most effective mechanism by which species diversity can
-lower revenue variability for individual fishers --- switching or adding fishing
-permits --- is costly, limited by regulations, and consequently unavailable to
-many individuals.
+find that (1) individuals with sets of permits that allow high levels of species
+diversity tend to have lower revenue variability. (2) Adding a fisheries permit
+usually increases revenue and decreases revenue variability. However, (3) within
+the constraints of a permit and accounting for effort, specializing tends to
+increase revenue while both specializing and generalizing tend to correlate with
+greater revenue variability with considerable differences amongst fisheries.
+Thus, the most effective mechanism by which species diversity can lower revenue
+variability for individual fishers --- switching or adding fishing permits ---
+is costly, limited by regulations, and consequently unavailable to many
+individuals.
 
-Our results across fishing strategies are in line with previous findings at the
+Our findings across fishing strategies are in line with previous research at the
 community and vessel levels [@kasperski2013; @sethi2012b; @sethi2014], but our
-results with respect to individual year-to-year behaviour provide new insights.
-Within strategies, specializing tended to be associated with increased revenue.
-This could be for a number of reasons. First, specialization can increase
-fishing efficiency [@smith1986]. Efficiency gains from specialization can be
-seen in other natural resource sectors. For example, rice farmers in Ghana can
-increase their efficiency by devoting more land and labor to rice production
-compared to other crops [@abdulai2000]. Second, given the availability of
-valuable resources, specializing can be highly profitable. For instance, when
-cocoa prices have been high, Nigerian farmers have tended to specialize on the
-crop [@mustapha1999] and specializing on valuable tobacco farming can be the
-optimal strategy regardless of alternative crops [@chavez2014]. In our analyses,
-fishers with the opportunity to fish highly valuable sockeye salmon benefited
-the most from specializing. However, specializing can have longterm risks if
-market, biological, or environmental conditions change. For example, U.S.
-farmers specialized on wheat, corn, or soybeans were hit harder than diversified
-farmers in the early 1980s when a financial crisis hit [@ikerd1990]. Third,
-generalization --- a lack of specialization --- can happen as a consequence of
-non-ideal circumstances such as poor availability of a more valuable fish stock
-in a given year. For example, downward economic shocks have been associated with
-increased diversity for households relying on forests for income [@debela2012].
+findings within strategies, with respect to individual year-to-year behaviour,
+provide new insights. Here, specializing tended to be associated with increased
+revenue. This could be for a number of reasons. First, specialization can
+increase fishing efficiency [@smith1986]. Efficiency gains from specialization
+can be seen in other natural resource sectors. For example, rice farmers in
+Ghana can increase their efficiency by devoting more land and labor to rice
+production compared to other crops [@abdulai2000]. Second, given the
+availability of valuable resources, specializing can be highly profitable. For
+instance, when cocoa prices have been high, Nigerian farmers have tended to
+specialize on the crop [@mustapha1999] and specializing on valuable tobacco
+farming can be the optimal strategy regardless of alternative crops
+[@chavez2014]. In our analyses, fishers with the opportunity to fish highly
+valuable sockeye salmon benefited the most from specializing. However,
+specializing can have longterm risks if market, biological, or environmental
+conditions change. For example, U.S. farmers specialized on wheat, corn, or
+soybeans were hit harder than diversified farmers in the early 1980s when
+a financial crisis hit [@ikerd1990]. Third, diversifying --- a lack of
+specialization --- can happen as a consequence of non-ideal circumstances such
+as poor availability of a more valuable resource in a given year. For example,
+downward economic shocks have been associated with increased diversity for
+households relying on forests for income [@debela2012].
 
-We also found that altering species diversity tended to be associated with
-increased revenue variability. There are number of possible reasons for this
-relationship. For instance, changes to fishing patterns will sometimes be
-associated with a learning or adaptation phase as individuals become accustomed
-to new locations, new gear, or new species [REF]. Additionally, changes in
-diversity might reflect years in which targeted species were less available and
-fishers were forced to target other species or landed more bycatch, adding
-stochasticity to their revenue. Most importantly, this increased variability
-reflects an inability for our large-scale model to predict an individual's
-revenue in a subsequent year. There are many reasons, often unique to an
-individual permit holder's behaviour [@hilborn1985] and therefore not included
-in our large-scale model, that might make that specific individual's revenue
-less predictable if species diversity changes.
+We also found that, within strategies, altering species diversity tended to be
+associated with increased revenue variability. There are number of possible
+reasons for this relationship. For instance, changes to fishing patterns will
+sometimes be associated with a learning or adaptation phase as individuals
+become accustomed to new locations, new gear, or new species [REF].
+Additionally, changes in diversity might reflect years in which targeted species
+were less available and fishers were forced to target other species or landed
+more bycatch, adding stochasticity to their revenue. Most importantly, this
+increased variability reflects an inability for our large-scale model to predict
+an individual's revenue in a subsequent year. There are many reasons, often
+unique to an individual permit holder's behaviour [@hilborn1985] and therefore
+not included in our large-scale model, that might make a specific individual's
+revenue less predictable if species diversity changes.
 
@@
 <!-- TODO: get fig:re in results -->
-<!-- Our model estimates overall average effects (Fig. S\ref{fig:fe}) and effects -->
-<!-- that vary by permit set (Fig. S\ref{fig:re}). -->
+<!-- Our model estimates overall average effects (Fig. S\ref{fig:fe}) and
+effects --> <!-- that vary by strategy (Fig. S\ref{fig:re}). -->
 
@@
 along with strategy-level intercepts would be estimated at almost exactly zero.
-The estimated $\beta_{j,t}$'s represent years when people with a particular
-permit set tended to collectively do better or worse than the long-term average
+The estimated $\beta_{j,t}$'s represent years when people with a particular set
+of permits tended to collectively do better or worse than the long-term average
 (Fig. S\ref{fig:str-yrs}). Therefore, an individual's expected change in revenue
 (and variability) is after accounting for the general trend in revenue for that
-permit set.
+set of permits.
 

### 2017-01-04 portfolio/text/ms.Rmd
More edits and references

@@
 
-Individuals who rely on the extraction of natural resources for their livelihood
+Individuals relying on the extraction of natural resources for their livelihood
 --- fishers, farmers, miners, and loggers, among others --- face high levels of
-income variability driven by a mix of environmental, biological, and economic
-variability. A key component to managing these industries is identifying how
-regulatory actions and individual behaviour affect income variability and by
-extension the economic stability and sustainable use of natural resources. For
-commercial fisheries, recent work has identified that communities and vessels
-that fish a greater diversity of species have less revenue variability than
-those that fish a lower diversity. However, it is unclear if these benefits
-extend to the level of the individual and how year-to-year changes in
-diversification affect revenue and revenue variability. Here we identify three
+income variability driven by a mix of environmental, biological, management, and
+economic variability. A key component to managing these industries is
+identifying how regulatory actions and individual behaviour affect income
+variability and by extension the economic stability and sustainable use of
+natural resources. For commercial fisheries, recent work has identified that
+communities and vessels fishing a greater diversity of species have less revenue
+variability than those fishing a lower diversity. However, it is unclear if
+these benefits extend to individual fishers and how year-to-year changes in
+diversification affect revenue and revenue variability. Here, we identify three
 axes by which fishers in Alaska, one of the world's most productive fishing
-regions, can diversify fishing activities. We show that fishing permits with
-higher species diversity reduces individual revenue variability and fishing an
-additional permit is associated with higher revenue and lower variability.
-However, increasing species diversity within the constraints of existing permits
-has a fishery-dependent effect on revenue and is usually (`r g2_dens`%
-probability) associated with increased revenue uncertainty the following year.
-Our results demonstrate that the most effective option for individuals to
-decrease revenue variability is to purchase permits to participate in additional
-or more diverse fisheries. However, this option is costly and often limited by
-regulations such as limited entry and catch share programs. With increasing
-environmental variability in a changing climate, it will be particularly
-important that individuals who rely on natural resources for their livelihood
-have effective strategies to reduce income variability.
+regions, can diversify fishing activities. We show that, despite increasing
+specialization over the last 30 years, fishing permits with higher species
+diversity reduces individual revenue variability and fishing an additional
+permit is associated with higher revenue and lower variability. However,
+increasing species diversity within the constraints of existing permits has
+a fishery-dependent effect on revenue and is usually (`r g2_dens`% probability)
+associated with increased revenue uncertainty the following year. Our results
+demonstrate that the most effective option for individuals to decrease revenue
+variability is to purchase permits to participate in additional or more diverse
+fisheries. However, this option is expensive, often limited by regulations such
+as catch share programs, and consequently unavailable to many individuals. With
+increasing environmental variability in a changing climate, it will be
+particularly important that individuals relying on natural resources for their
+livelihood have effective strategies to reduce income variability.
 
-<!-- TODO: fit something about specialization over time into the abstract.
-Thus, over the past 30 years as the Alaskan commercial fishing fleet has become
-increasingly specialized, -->
+<!-- TODO: take home is the management context is key and controls show that
+it's more complicated. TODO mentioned that specializing tends to be profitable
+-->
 
@@
 individual *within* the constraints of a given set of permits is more variable
-(Fig. \ref{fig:within}). After accounting for fishing effort, specializing tends
-to be associated with increased revenue but also increased variability (Fig.
-\ref{fig:within}; Figs S\ref{fig:fe}, S\ref{fig:re}, S\ref{fig:noeffort}). On
-average, there is a ~0.5% increase in the year-to-year revenue ratio per 1%
+(Fig. \ref{fig:within}). After accounting for fishing effort and changes in
+annual average revenue for fishers with the same set of permits, specializing
+tends to be associated with increased revenue but also increased variability
+(Fig. \ref{fig:within}; Figs S\ref{fig:fe}, S\ref{fig:re}, S\ref{fig:noeffort}).
+On average, there is a ~0.5% increase in the year-to-year revenue ratio per 1%
 decrease in species diversity, and across strategies weighted equally, `r
@@
 that adopting strategies that exploit a diversity of assets provides substantial
-benefits over strategies with lower diversity [REFS]. However, in practice, the
-ability of individuals to effectively diversify is limited by regulation and
-market forces. Here we document the effect of individual-level changes in
-species diversity on revenue and revenue variability across three scales. We
-find that (1) individuals with sets of permits that allow high levels of species
-diversity tend to have lower revenue variability. (2) Adding a fisheries permit
-usually increases revenue and decreases revenue variability. However, (3) within
-the constraints of a permit and accounting for effort, specializing tends to
-increase revenue while both specializing and generalizing tend to correlate with
-greater revenue variability with considerable differences amongst fisheries.
-Thus, the most effective mechanism by which species diversity can lower revenue
-variability for individual fishers --- switching or adding fishing permits ---
-is costly, limited by regulations, and consequently unavailable to many
-individuals.
+benefits over strategies with lower diversity [@baldursson1997; @hilborn2001;
+@edwards2004]. However, in practice, the ability of individuals to effectively
+diversify is limited by regulation and market forces. Here we document the
+effect of individual-level changes in species diversity on revenue and revenue
+variability across three scales. We find that (1) individuals with sets of
+permits that allow high levels of species diversity tend to have lower revenue
+variability. (2) Adding a fisheries permit usually increases revenue and
+decreases revenue variability. However, (3) within the constraints of a permit
+and accounting for effort, specializing tends to increase revenue while both
+specializing and generalizing tend to correlate with greater revenue variability
+with considerable differences amongst fisheries. Thus, the most effective
+mechanism by which species diversity can lower revenue variability for
+individual fishers --- switching or adding fishing permits --- is expensive,
+limited by regulations, and consequently unavailable to many individuals.
 
@@
 benefits of diversification in the context of rare and extreme disturbances
-which occur in natural systems worldwide [REF].
+which occur in natural systems worldwide [@anderson2016].
 
@@
 al. 2008 ) - may ultimately restrict individual diversification or at least make
-developing a diverse fishing portfolio more costly. Add Alaska example of
+developing a diverse fishing portfolio more expensive. Add Alaska example of
 IFQs...?
@@
 resources. With increasing environmental variability and extreme events in
-a changing climate (REF), it is particularly important that individuals who rely
-on natural resources as a primary source of income have strategies to reduce
-risk. Though variability in individual revenue may not be actively incorporated
-into many management decisions, analyses like ours help to understand the range
-of diversification strategies available and how their benefits may change over
-time.
+a changing climate [@easterling2000], it is particularly important that
+individuals who rely on natural resources as a primary source of income have
+strategies to reduce risk. Though variability in individual revenue may not be
+actively incorporated into many management decisions, analyses like ours help to
+understand the range of diversification strategies available and how their
+benefits may change over time.
 
@@
 [@simpson1949; @jost2006] (also referred to in economics as the
-Herfindahl-Hirschman Index) weighted by revenue $R$:
+Herfindahl-Hirschman Index [@hirschman1964]) weighted by revenue $R$:
 
@@
 increasing diversity and decreasing diversity allows the costs and benefits of
-these changes to be asymmetric.
+these changes to be asymmetric (Fig. S\ref{fig:resid}).
 

### 2017-01-04 portfolio/text/ms.Rmd
More edits and references

@@
 variability than those fishing a lower diversity. However, it is unclear if
-these benefits extend to individual fishers and how year-to-year changes in
-diversification affect revenue and revenue variability. Here, we identify three
-axes by which fishers in Alaska, one of the world's most productive fishing
-regions, can diversify fishing activities. We show that, despite increasing
-specialization over the last 30 years, fishing permits with higher species
-diversity reduces individual revenue variability and fishing an additional
-permit is associated with higher revenue and lower variability. However,
-increasing species diversity within the constraints of existing permits has
-a fishery-dependent effect on revenue and is usually (`r g2_dens`% probability)
-associated with increased revenue uncertainty the following year. Our results
-demonstrate that the most effective option for individuals to decrease revenue
-variability is to purchase permits to participate in additional or more diverse
-fisheries. However, this option is expensive, often limited by regulations such
-as catch share programs, and consequently unavailable to many individuals. With
-increasing environmental variability in a changing climate, it will be
-particularly important that individuals relying on natural resources for their
-livelihood have effective strategies to reduce income variability.
+these benefits extend to the actions of individual fishers and how year-to-year
+changes in diversification affect revenue and revenue variability. Here, we
+identify three axes by which fishers in Alaska, one of the world's most
+productive fishing regions, can diversify fishing activities. We show that,
+despite increasing specialization over the last 30 years, fishing permits with
+higher species diversity reduces individual revenue variability and fishing an
+additional permit is associated with higher revenue and lower variability.
+However, increasing species diversity within the constraints of existing permits
+has a fishery-dependent effect on revenue and is usually (`r g2_dens`%
+probability) associated with increased revenue uncertainty the following year.
+Our results demonstrate that the most effective option for individuals to
+decrease revenue variability is to purchase permits to participate in additional
+or more diverse fisheries. However, this option is expensive, often limited by
+regulations such as catch share programs, and consequently unavailable to many
+individuals. With increasing environmental variability in a changing climate, it
+will be particularly important that individuals relying on natural resources for
+their livelihood have effective strategies to reduce income variability.
 
@@
 valuable sockeye salmon benefited the most from specializing. However,
-specializing can have longterm risks if market, biological, or environmental
-conditions change. For example, U.S. farmers specialized on wheat, corn, or
-soybeans were hit harder than diversified farmers in the early 1980s when
-a financial crisis hit [@ikerd1990]. Third, diversifying --- a lack of
+specializing can have longterm risks if market, management, biological, or
+environmental conditions change. For example, U.S. farmers specialized on wheat,
+corn, or soybeans were hit harder than diversified farmers in the early 1980s
+when a financial crisis hit [@ikerd1990]. Third, diversifying --- a lack of
 specialization --- can happen as a consequence of non-ideal circumstances such
-as poor availability of a more valuable resource in a given year. For example,
-downward economic shocks have been associated with increased diversity for
-households relying on forests for income [@debela2012].
+as poor availability of a more valuable resource. For example, Ugandan
+households relying on forestry resources for income increase their income
+portfolio diversity following downward economic shocks [@debela2012].
 
@@
 associated with increased revenue variability. There are number of possible
-reasons for this relationship. For instance, changes to fishing patterns will
-sometimes be associated with a learning or adaptation phase as individuals
-become accustomed to new locations, new gear, or new species [REF].
+reasons for this relationship. For instance, local knowledge and experience are
+a major determinant of fishing efficiency [@hilborn1985a]. Therefore, changes to
+fishing patterns may sometimes be associated with a learning or adaptation phase
+as individuals become accustomed to new locations, gear, or species.
 Additionally, changes in diversity might reflect years in which targeted species
@@
 an individual's revenue in a subsequent year. There are many reasons, often
-unique to an individual permit holder's behaviour [@hilborn1985] and therefore
-not included in our large-scale model, that might make a specific individual's
-revenue less predictable if species diversity changes.
+unique to an individual permit holder's behaviour [@hilborn1985; @hilborn1985a;
+@branch2006] and therefore not included in our large-scale model, that might
+make a specific individual's revenue less predictable if species diversity
+changes.
 
@@
 diversifying. For commercial fishers, purchasing additional licenses and gear
-typically represents a substantial cost. Related barriers are apparent in other
-natural resource sectors. For example, diversification has been positively
+typically represents a substantial expense. Related barriers are apparent in
+other natural resource sectors. For example, diversification has been positively
 linked with total income or access to financial capital for farmers
@@
 those relying on forest resources [@illukpitiya2008]. Another major barrier to
-diversification can be management restrictions. In commercial fisheries, while
-catch shares [@costello2008] and limited season lengths [REF] can be effective
-tools to avoid overfishing and distribute social capital, they can also limit
-opportunities for diversification [@smith1986; @kasperski2013]. However, catch
-shares, in particular, may also reduce income variability (Fig. S\ref{fig:ifq}A)
-perhaps offsetting the effects of reduced diversification [@holland2016].
+diversification can be management restrictions. In commercial fisheries,
+regulations that you have been shown to reduce overfishing --- including
+improved assessments [@melnychuk2017] and individual quota systems
+[@costello2008] --- may ultimately restrict individual diversification
+[@smith1986; @kasperski2013] or at least make developing a diverse fishing
+portfolio more expensive. However, individual quota systems, in particular, may
+also reduce income variability (Fig. S\ref{fig:ifq}A) perhaps offsetting the
+effects of reduced diversification [@holland2016].
 
@@
 individuals who rely on natural resources as a primary source of income have
-strategies to reduce risk. Though variability in individual revenue may not be
+strategies to reduce risk. Although variability in individual revenue may not be
 actively incorporated into many management decisions, analyses like ours help to
@@
 benefits may change over time.
+
+<!--take home is the management context is key and controls show that
+it's more complicated-->
 

### 2017-01-04 portfolio/text/ms.Rmd
Finish discussion

@@
 
-<!--
-Worldwide, fisheries confront the challenge of sustainable management (). While
-many fisheries are well managed and productive, many are also rebuilding from
-periods of over-exploitation (Worm et al. Science 2009). A common tactic to aid
-in fisheries rebuilding is to modify regulations to reduce the unintended
-harvest rebuilding stocks while allowing the continued harvest of healthy stocks
-(need REF). Such bycatch reduction is an important goal for both ecological
-(Crowder and Murawski 1998, Lewison et al. 2004) and economic (Curtis and Hicks
-2000; Chakravorty and Nemoto 2001) reasons. Cumulatively, though, regulatory
-changes for bycatch reduction as well as other reasons gradually lead to
-fisheries that target one or a few species and individual fishers wishing to
-participate in multiple fisheries must own (or purchase) multiple permits. Thus
-a general consequence of fisheries regulations that are shown to improve
-sustainability of fish stocks and avoid overfishing - including improved
-assessments (Hilborn and Ovando 2014) and individual quota systems (Costello et
-al. 2008 ) - may ultimately restrict individual diversification or at least make
-developing a diverse fishing portfolio more expensive. Add Alaska example of
-IFQs...?
--->
-
-<!-- Conclusions + recommendations
-
-- bring it out to wider than fisheries (farming, timber industry, ...)
-- 2 dimensions to think about (revenue and variability); and the trade-off is
-  fishery specific: effects within a permit strategy are permit specific;
-  sometimes make more money by specializing... some highly variable single
-  species permits make a ton of money
-- broadly, diversification can provide a powerful form of protection from
-  multiple sources of risk to income for those that rely on extracting natural
-  resources, but not all forms of diversification are equal, and on an
-  individual level there can be economic benefits to specializing and risks to
-  adjusting ones strategy
-- with increasing environmental variability and extremes due to climate change
-  those that rely on natural resources for their livelihoods need approaches to
-  buffer risk more than ever
-- regulations that limit the ability to diversify can have other considerable
-  benefits but it's important to be aware of trade-offs inherent in these
-  regulations when making management decisions
--->
-
-<!--Our results provide additional support for the idea that individual fishers can
-reduce their income variability by diversifying their catch portfolios. However,
-most individuals do not change their catch diversity substantially from year to
-year [OLE: is there a supplemental figure or statistic that speaks to this
-point?], and most strategies adopted by Alaskan fishermen lie in a high-risk,
-high-reward quadrant in which reducing diversification increases both revenue
-variability and expected revenue. Previous work has suggested that changes to
-fisheries management such as limited entry or catch share programs may
-inadvertently act as barrier to diversification and promote specializing
-(Hilborn et al. 2003; Kasperski and Holland 2013). While some of these
-specialized fishing strategies may be associated with increased revenue
-variability, the benefit may be that they are also associated with increased
-revenue.-->
-
-TODO: finish this:
-External market forces may interact in concert with management regulations to
-promote specialization for fishers or other individuals dependent on natural
-resources. With increasing environmental variability and extreme events in
-a changing climate [@easterling2000], it is particularly important that
-individuals who rely on natural resources as a primary source of income have
-strategies to reduce risk. Although variability in individual revenue may not be
-actively incorporated into many management decisions, analyses like ours help to
-understand the range of diversification strategies available and how their
-benefits may change over time.
-
-<!--take home is the management context is key and controls show that
-it's more complicated-->
+With increasing environmental variability and extreme events in a changing
+climate [@easterling2000], it is particularly important that individuals who
+rely on natural resources as a primary source of income, such as fishers,
+farmers, and forestry workers, have effective strategies to reduce risk. While
+diversification can provide a powerful form of risk buffering, we have shown
+that the relationship between individual changes to diversification and income
+variability is complex and context specific. Furthermore, external market forces
+can interact with management regulations, which have other considerable
+benefits, to restrict individual diversification opportunities and promote
+specialization. Although variability in individual revenue may not be actively
+incorporated into management decisions, analyses such as ours help to understand
+the range of diversification strategies available, trade-offs inherent in such
+management decisions, and how these trade-offs may change over time.
 
@@
 # Materials and methods
-
-<!-- TODO: get fig:re in results -->
-<!-- Our model estimates overall average effects (Fig. S\ref{fig:fe}) and
-effects --> <!-- that vary by strategy (Fig. S\ref{fig:re}). -->
 

### 2017-01-06 portfolio/text/anderson-etal-rev-pfx.tex
Move dual fig 4 to supp
Fix bib; add outline
Edits
Jordan and Milo edits

@@
 \title{
-  Benefits and risks of diversification\\ for individual fishers\\
-  Benefits of diversification and risks of 
-  changing strategies for individual fishers\\
-Benefits of diversification are unavailable to many fishermen\\
+  Benefits and risks of diversification\\ for individual fishers
+  % Benefits of diversification and risks of 
+  % changing strategies for individual fishers\\
+% Benefits of diversification are unavailable to many fishermen\\
 }
@@
 Eric J. Ward$^2$,
+Andrew O. Shelton$^2$,\\
 Milo D. Adkison$^3$,
@@
 Alan C. Haynie$^5$,
-Andrew O. Shelton$^2$,\\
-Jennifer C. Shriver,
-Jordan T. Watson$^X$,
+Jennifer C. Shriver$^4$,
+Jordan T. Watson$^6$,
 \ldots
@@
 \newlabel{fig:ifq}{7}
+\newlabel{fig:mirror}{8}
+\newlabel{fig:noeffort}{9}
 
@@
 
+% TODO: standardize and complete addresses
+
 \noindent
@@
 \noindent
+$^3$College of Fisheries and Ocean Sciences, University of Alaska Fairbanks,
+17101 Point Lena Loop Rd, Juneau, AK, 99801, USA
+
+\noindent
+$^4$Alaska Department of Fish and Game, Commercial Fisheries Division, Cordova, USA
+
+\noindent
 $^5$Alaska Fisheries Science Center,
@@
 \noindent
-$^X$Ted Stevens Marine Research Institute, Alaska Fisheries Science Center, 17109
-Pt Lena Loop Road, Juneau, AK 99801, USA
-
-\noindent
-Affiliations \ldots
+$^6$NOAA / National Marine Fisheries Service, Alaska Fisheries Science Center, Auke
+Bay Laboratories, 17109 Pt. Lena Loop Rd., Juneau, AK 99801, USA
 
@@
 \setlength\parskip{0in}
-
-% \vspace{-1cm}
-
-%\input{../analysis/values} % R output
 

### 2017-01-06 portfolio/text/ms.Rmd
Jordan and Milo edits

@@
 income variability driven by a mix of environmental, biological, management, and
-economic variability. A key component to managing these industries is
-identifying how regulatory actions and individual behaviour affect income
-variability and by extension the economic stability and sustainable use of
-natural resources. For commercial fisheries, recent work has identified that
-communities and vessels fishing a greater diversity of species have less revenue
-variability than those fishing a lower diversity. However, it is unclear if
-these benefits extend to the actions of individual fishers and how year-to-year
-changes in diversification affect revenue and revenue variability. Here, we
-identify three axes by which fishers in Alaska, one of the world's most
-productive fishing regions, can diversify fishing activities. We show that,
-despite increasing specialization over the last 30 years, fishing permits with
-higher species diversity reduces individual revenue variability and fishing an
-additional permit is associated with higher revenue and lower variability.
-However, increasing species diversity within the constraints of existing permits
-has a fishery-dependent effect on revenue and is usually (`r g2_dens`%
-probability) associated with increased revenue uncertainty the following year.
-Our results demonstrate that the most effective option for individuals to
-decrease revenue variability is to purchase permits to participate in additional
-or more diverse fisheries. However, this option is expensive, often limited by
-regulations such as catch share programs, and consequently unavailable to many
-individuals. With increasing environmental variability in a changing climate, it
-will be particularly important that individuals relying on natural resources for
-their livelihood have effective strategies to reduce income variability.
+economic dynamics. A key component to managing these industries is identifying
+how regulatory actions and individual behaviour affect income variability and by
+extension economic stability and the sustainable use of natural resources. For
+commercial fisheries, communities and vessels fishing a greater diversity of
+species have less revenue variability than those fishing a lower diversity.
+However, it is unclear if these benefits extend to the actions of individual
+fishers and how year-to-year changes in diversification affect revenue and
+revenue variability. Here, we identify three axes by which fishers in Alaska,
+one of the world's most productive fishing regions, can diversify fishing
+activities. We show that, despite increasing specialization over the last 30
+years, fishing permits with higher species diversity reduces individual revenue
+variability and fishing an additional permit is associated with higher revenue
+and lower variability. However, increasing species diversity within the
+constraints of existing permits has a fishery-dependent effect on revenue and is
+usually (`r g2_dens`% probability) associated with increased revenue uncertainty
+the following year. Our results demonstrate that the most effective option for
+individuals to decrease revenue variability is to purchase permits to
+participate in additional or more diverse fisheries. However, this option is
+expensive, often limited by regulations such as catch share programs, and
+consequently unavailable to many individuals. With increasing environmental
+variability in a changing climate, it will be particularly important that
+individuals relying on natural resources for their livelihood have effective
+strategies to reduce income variability.
 
@@
 
+<!-- rotate -> plant multiple? -->
+
 Individuals who rely on natural resources for income develop strategies to
@@
 Similarly, loggers maintain a diverse mix of tree species, age structure, and
-harvest in geographical areas to buffer against disease outbreak, wildfires, and
-price variation (REF). However, external forces may restrict the available
-options to dampen income variability. Otherwise well-intentioned regulations may
-limit how individuals diversify, or incentivize against diversification. For
-example, in terrestrial agriculture, crop subsidies in the U.S. may incentivize
-some farms to specialize (e.g. dairy, grains, soybeans; @mishra2002). In other
+harvest in multiple areas to buffer against disease outbreak, wildfires, and
+price variation (REF). However, otherwise well-intentioned regulations may limit
+how individuals diversify, or incentivize against diversification. For example,
+in terrestrial agriculture, crop subsidies in the U.S. may incentivize some
+farms to specialize (e.g. dairy, grains, soybeans; @mishra2002). In other
 instances, external market forces (e.g. prices or demand) may discourage the
@@
 
+<!-- important to acknowledge that IFQ regs have been implemented partly for economic -->
+<!-- reasons and have been successful by some economic measures. Might even want to -->
+<!-- insert the above sentence. -->
+
 The state and federal fisheries of Alaska provide a unique opportunity to
@@
 all fisheries [@hilborn2003a]. While such regulations have helped reduce
-overfishing, they limit the available strategies to reduce income variability.
-For example, many permits limit fishing to a single species, diversifying by
-purchasing additional fishing permits can be expensive[^1], and some fisheries
-limit the entry of new participants through individual fishing quotas (IFQs)
-[@grimm2012].
+overfishing and in many fisheries have increased profits, they limit the
+available strategies to reduce income variability. For example, many permits
+limit fishing to a single species, diversifying by purchasing additional fishing
+permits can be expensive[^1], and some fisheries limit the entry of new
+participants through individual fishing quotas (IFQs) [@grimm2012].
 
@@
 we can disentangle fishers that share the same vessel. We develop and apply
-a hierarchical Bayesian model to understand the effect of fisheries catch
-species diversity on individual revenue and revenue variability that builds on
-results from vessel [@kasperski2013], fleet [@sethi2012b], and community-level
-[@sethi2014] variability. We develop a variance-function regression model
-[@davidian1987; @hall1989], to simultaneously model changes in the mean and
-variance of individual revenue as a function of covariates. We simultaneously
-estimate the effect of species diversity (the inverse of Simpson's diversity
-index [@simpson1949; @jost2006]) on individual revenue and revenue variability,
-and estimate effects both within and across fishing strategies while controlling
-for fishing effort and latent time effects (e.g. changing environmental
-conditions). To our knowledge, this is the first application of this type of
-model in any natural resource field.
+a hierarchical Bayesian model to understand the effect of species diversity
+within fisheries catches on individual revenue and revenue variability. Our
+model builds on results from vessel [@kasperski2013], fleet [@sethi2012b], and
+community-level [@sethi2014] variability. We develop a variance-function
+regression model [@davidian1987; @hall1989], to simultaneously model changes in
+the mean and variance of individual revenue as a function of covariates. We
+simultaneously estimate the effect of species diversity (the inverse of
+Simpson's diversity index [@simpson1949; @jost2006]) on individual revenue and
+revenue variability, and estimate effects both within and across fishing
+strategies while controlling for fishing effort and latent time effects (e.g.
+changing environmental conditions). To our knowledge, this is the first
+application of this type of model in any natural resource field.
 
@@
 approximately half the overall revenue (Fig. \ref{fig:ts}B). The mean species
-diversity of this revenue declined somewhat from 1980--2000 but has since
+diversity of this revenue declined somewhat from 1985--2000 but has since
 remained steady (Fig. \ref{fig:ts}D). However, this trend appears largely driven
@@
 separately before and after the introduction of IFQs. The introduction of IFQs
-reduced variability by ~25% (TODO XX--XX 95% CI) for individuals with only
-halibut or sablefish permits (Fig. S\ref{fig:ifq}A). Furthermore, for most
+reduced variability in revenue by ~25% (TODO XX--XX 95% CI) for individuals with
+only halibut or sablefish permits (Fig. S\ref{fig:ifq}A). Furthermore, for most
 strategies, the association between changing species diversity and increased
@@
 diversify is limited by regulation and market forces. Here we document the
-effect of individual-level changes in species diversity on revenue and revenue
-variability across three scales. We find that (1) individuals with sets of
-permits that allow high levels of species diversity tend to have lower revenue
-variability. (2) Adding a fisheries permit usually increases revenue and
+effect of individual-level changes in fished species diversity on revenue and
+revenue variability across three scales. We find that (1) individuals with sets
+of permits that allow high levels of species diversity tend to have lower
+revenue variability. (2) Adding a fisheries permit usually increases revenue and
 decreases revenue variability. However, (3) within the constraints of a permit
@@
 diversification can be management restrictions. In commercial fisheries,
-regulations that you have been shown to reduce overfishing --- including
-improved assessments [@melnychuk2017] and individual quota systems
-[@costello2008] --- may ultimately restrict individual diversification
-[@smith1986; @kasperski2013] or at least make developing a diverse fishing
-portfolio more expensive. However, individual quota systems, in particular, may
-also reduce income variability (Fig. S\ref{fig:ifq}A) perhaps offsetting the
-effects of reduced diversification [@holland2016].
+regulations that have been shown to reduce overfishing --- including improved
+assessments [@melnychuk2017] and individual quota systems [@costello2008] ---
+may ultimately restrict individual diversification [@smith1986; @kasperski2013]
+or at least make developing a diverse fishing portfolio more expensive. However,
+individual quota systems, in particular, may also reduce income variability
+(Fig. S\ref{fig:ifq}A) perhaps offsetting the effects of reduced diversification
+[@holland2016].
 
@@
 \caption{Trends in revenue and diversification over time. (A) Map of fishing
-regions in Alaska, USA. (B) Total revenue in billions of 2009 USD from `r
+regions in Alaska, USA. (B) Total revenue in billions of USD from `r
 range_yrs[1]` to `r range_yrs[2]` for multiple (grey) and single (blue) permit
@@
 We obtained commercial fisheries landings and revenue data for all permit
-holders in Alaska from `r range_yrs[1]` to `r range_yrs[2]` from the Alaskan
+holders in Alaska from `r range_yrs[1] - 1` to `r range_yrs[2]` from the Alaskan
 Commercial Fisheries Entry Commission (CFEC). These data represent a total of
@@
 commercial fishing (Supporting Information). In particular, we removed permit
-holders whose median revenue was less than \$10,000 2009 USD to focus on
-individuals for whom commercial fishing formed a substantial part of their
-income.
+holders whose median revenue was less than \$10,000 USD to focus on individuals
+for whom commercial fishing formed a substantial part of their income.
+
+<!-- TODO Jordan: held by active fishers were not included in permit counts -->
+<!-- / portfolios? Disregard this comment as you please but I thought I’d mention. -->
 
@@
 fishing sablefish on a vessel over 60 ft), and we combined a number of permits
-that were targeting a single species and differed only in region fished.
-Therefore, for example, someone fishing herring roe in Kodiak and someone
-fishing herring roe in Alaska Peninsula were both considered to have a "herring
-roe" permit group. For salmon permits, we formed salmon strategies based on
-similarities in gear type and species composition (Supporting Information).
+that were targeting a single species and differed only in region fished. For
+example, someone fishing herring roe in Kodiak and someone fishing herring roe
+in Alaska Peninsula were both considered to have a "herring roe" permit group.
+For salmon permits, we formed salmon strategies based on similarities in gear
+type and species composition (Supporting Information).
 
@@
 analysis revealed non-stationary trends in revenue (resulting from experience or
-other unmodeled factors). This approach eliminates having to deal with the
+other unmodelled factors). This approach eliminates having to deal with the
 non-stationary time series and eliminates having to estimate thousands of
@@
 (Fig. S\ref{fig:str-yrs}). Therefore, an individual's expected change in revenue
-(and variability) is after accounting for the general trend in revenue for that
-set of permits.
+(and variability) represents the expectation after accounting for the general
+trend in revenue for that set of permits.
 

### 2017-01-06 portfolio/text/supp.Rnw
Add effort fig
Trivial tweaks
Jordan and Milo edits

@@
 We combined permits for
-otter troll,
+otter trawl,
 longline finfish,
@@
 sea cucumber,
-sea urchin,
-tanner crabs
+sea urchin, and
+tanner crabs.
 
@@
 
+% TODO: Are you saying that PWS sockeye permits were combined with BB sockeye permits?I can't remember but that's what it sounds like so I just wanted to be clear.
+
 After this process of combining permits,
@@
 
-TODO length(unique(pholders))
+% TODO length(unique(pholders))
 % Rows of data: \Sexpr{ndat$final}
@@
 \includegraphics[width=1\textwidth]{../figs/stan-gg-spoke-all.pdf}
-\caption{The effect of adding a permit to ones strategy
+\caption{The effect of adding a permit to one's strategy
 on estimated revenue variability and median revenue.
@@
 \caption{The log-absolute value residuals from the mean-component model (Eq.~\ref{eq:mean}).
-It is these residuals that are modeled in the variance-component model
+These residuals are modelled in the variance-component model
 (Eq.~\ref{eq:cv}).
@@
 
+\clearpage
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=0.6\textwidth]{../figs/stan-offset-break-anti-spaghetti-specializing-no-effort.pdf}
+\caption{Same as Fig.~4 but without the effort (days fished) predictor and
+  without its interaction with species diversity.}
+\label{fig:noeffort}
+\end{center}
+\end{figure}
+
 % \begin{figure}[htbp]

### 2017-01-10 portfolio/text/ms.Rmd

--- 2017-01-04 portfolio/text/ms.Rmd
@@
 
-<!-- Introduction, Methods, Results, Discussion, Conclusion -->
+<!-- Diversified portfolios of fishing permits and species reduce risk for fishermen, -->
+<!-- but changing species diversity can be risky, costly, and limited by regulations. -->
+<!-- Goal: Introduction, Methods, Results, Discussion, Conclusion:
 <!-- 600, 900, 500, 800, 200, -->
@@
 
-<!-- TODO < 250 words -->
+<!-- TODO slash length of abstract -->
 
-Individuals relying on the extraction of natural resources for their livelihood
---- fishers, farmers, miners, and loggers, among others --- face high levels of
-income variability driven by a mix of environmental, biological, management, and
-economic variability. A key component to managing these industries is
-identifying how regulatory actions and individual behaviour affect income
-variability and by extension the economic stability and sustainable use of
-natural resources. For commercial fisheries, recent work has identified that
-communities and vessels fishing a greater diversity of species have less revenue
-variability than those fishing a lower diversity. However, it is unclear if
-these benefits extend to the actions of individual fishers and how year-to-year
-changes in diversification affect revenue and revenue variability. Here, we
-identify three axes by which fishers in Alaska, one of the world's most
-productive fishing regions, can diversify fishing activities. We show that,
-despite increasing specialization over the last 30 years, fishing permits with
-higher species diversity reduces individual revenue variability and fishing an
-additional permit is associated with higher revenue and lower variability.
-However, increasing species diversity within the constraints of existing permits
-has a fishery-dependent effect on revenue and is usually (`r g2_dens`%
-probability) associated with increased revenue uncertainty the following year.
-Our results demonstrate that the most effective option for individuals to
-decrease revenue variability is to purchase permits to participate in additional
-or more diverse fisheries. However, this option is expensive, often limited by
-regulations such as catch share programs, and consequently unavailable to many
-individuals. With increasing environmental variability in a changing climate, it
-will be particularly important that individuals relying on natural resources for
-their livelihood have effective strategies to reduce income variability.
+A fundamental challenge for individuals who rely on the extraction of natural
+resources for their livelihood is income variability. Fishers, loggers, farmers,
+and miners, among others, face high levels of income uncertainty driven by a mix
+of environmental, biological, and economic variability. A key component
+underlying the sustainability of these industries is identifying both how
+regulatory actions and individual behaviour affect income variability and by
+extension the economic stability and sustainable use of natural resources. For
+commercial fisheries, recent work has identified that targeting a diversity of
+species can reduce income variability. However, any diversity-variability
+relationship is embedded in a regulatory framework that restricts individual
+ability to change fishing behaviour. Here we identify three primary axes by
+which fishers in Alaska, one of the world's most productive fishery regions, can
+potentially diversify fishing activities and investigate how individual actions
+along each diversity axis affect both revenue and revenue variability. We
+construct a novel statistical model to simultaneously assess the value to
+individual fishers of choosing to (1) own permits that allow targeting a diverse
+fish species assemblage, (2) diversify by owning multiple permits, or (3) catch
+a greater species diversity within the constraints of a given permit. We show
+contrasting effects of diversity on revenue and revenue variability along the
+three axes. First, we show that fishing permits with higher species diversity
+reduces variability (`r h1[[2]]`% [`r h1[[1]]`--`r h1[[3]]`% 95% CI] reduction
+per unit increase in mean effective species diversity). Next, we show that
+fishing an additional permit is associated with higher revenue and lower
+variability (e.g., `r 100-spoke[[2]]`% [`r 100-spoke[[3]]`--`r 100-spoke[[1]]`%
+IQR] decline in standard deviation for fishers with halibut and another permit
+vs. only a halibut permit). Finally, we show that increasing species diversity
+within the constraints of a permit set has a fishery-dependent effect on revenue
+and is usually (`r g2_dens`% probability) associated with increased uncertainty
+in revenue the following year. Thus, while there are clear benefits to some axes
+of diversification, others are less effective. Importantly, the most effective
+mechanism for increasing revenue and decreasing risk --- purchasing additional
+permits to participate in additional fisheries --- was also the most costly.
+Furthermore, the least costly option of increasing catch diversity within an
+existing permit was associated with increased revenue variability. Thus, over
+the past 30 years as the Alaskan commercial fishing fleet has become
+increasingly specialized and most individuals only participate under a single
+permit, the practical benefits accrued from increased species diversification
+are limited.
 
-<!-- TODO: take home is the management context is key and controls show that
-it's more complicated. TODO mentioned that specializing tends to be profitable
--->
+Our results suggest that the benefit of species diversity for revenue
+variability is mostly an across-permit effect, not a benefit that can be easily
+harnessed by an individual permit holder. Effective methods of diversification
+--- purchasing additional permits or switching permits --- are costly and often
+limited by regulations such as limited entry programs and individual
+transferable quotas (ITQs). Take home ...
 
@@
 
+<!-- Papers to cite [@smith1986] -->
+
 Managing natural resources and sustaining the livelihoods of those that depend
-on their extraction is hard. These livelihoods tend to have high annual
+on their extraction is hard. These livelihoods tend to have high year-to-year
 variability in income relative to other professions [@mishra2002; @sethi2012b].
@@
 dependent on natural resources are also subject to biological and environmental
-variability [@smith1986]. For example, drought and flooding are a major source
-of risk for agricultural food security and farmers' incomes [@gregory2005], and
-catastrophic disease outbreaks and wildfires increase risk for the logging
-industry [@kirilenko2007].
+variability. For example, drought and flooding are a major source of risk for
+agricultural food security and farmers' incomes [@gregory2005], and catastrophic
+disease outbreaks and wildfires increase risk for the logging industry
+[@kirilenko2007].
 
@@
 limit how individuals diversify, or incentivize against diversification. For
-example, in terrestrial agriculture, crop subsidies in the U.S. may incentivize
+example, in terrestrial agriculture, crop subsidies in the US may incentivize
 some farms to specialize (e.g. dairy, grains, soybeans; @mishra2002). In other
@@
 there can be substantial benefits to specialization for both farmers and fishers
-[@smith1986; @abdulai2000; @arnason2005]. Furthermore, the majority of research
-examining the effects of diversification on revenue do so at aggregate levels
-that combine individuals (e.g. counties or communities; [@kasperski2013;
+[@abdulai2000; @arnason2005]. Furthermore, the majority of research examining
+the effects of diversification on revenue do so at aggregate levels that combine
+individuals (e.g. households, counties, or communities; [@kasperski2013;
 @sethi2014]). Thus, key uncertainties remain about the fundamental relationship
@@
 world with ~2.7 billion kg landed biomass valued at nearly 2 billion USD
-annually [@nmfs2016]. The days of open, unregulated fisheries in most developed
-countries have passed --- fishers must own a license to participate in virtually
-all fisheries [@hilborn2003a]. While such regulations have helped reduce
-overfishing, they limit the available strategies to reduce income variability.
-For example, many permits limit fishing to a single species, diversifying by
-purchasing additional fishing permits can be expensive[^1], and some fisheries
-limit the entry of new participants through individual fishing quotas (IFQs)
-[@grimm2012].
+annually [@nmfs2016]. Like much of the rest of the world, the days of open,
+unregulated fisheries in most developed countries have passed --- fishers must
+own a license to participate in virtually all fisheries [@hilborn2003a]. And
+while such regulations have helped reduce overfishing, they limit the available
+strategies to reduce income variability. For example, many permits limit fishing
+to a single species, diversifying by purchasing additional fishing permits can
+be expensive (e.g. ~\$192K USD for a king and tanner crab permit in Southeast
+Alaska in 2016; <https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>), and some
+fisheries limit the entry of new participants through individual fishing quotas
+(IFQs) [@grimm2012].
 
-[^1]: For example, it cost ~\$192K USD for a king and tanner crab permit in
-Southeast Alaska in 2016: <https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>.
-
-Using a unique dataset collected over 30 years on individual fishers, rather
-than fleets or vessels, we examine how revenue and revenue variability of
+Using a unique dataset collected over 30 years (1985--2014) on individuals,
+rather than fleets or vessels, we examine how revenue and revenue variability of
 individuals is affected by diversification. By tracking individuals we can track
@@
 we can disentangle fishers that share the same vessel. We develop and apply
-a hierarchical Bayesian model to understand the effect of fisheries catch
-species diversity on individual revenue and revenue variability that builds on
-results from vessel [@kasperski2013], fleet [@sethi2012b], and community-level
-[@sethi2014] variability. We develop a variance-function regression model
-[@davidian1987; @hall1989], to simultaneously model changes in the mean and
-variance of individual revenue as a function of covariates. We simultaneously
-estimate the effect of species diversity (the inverse of Simpson's diversity
-index [@simpson1949; @jost2006]) on individual revenue and revenue variability,
-and estimate effects both within and across fishing strategies while controlling
-for fishing effort and latent time effects (e.g. changing environmental
-conditions). To our knowledge, this is the first application of this type of
-model in any natural resource field.
+a hierarchical Bayesian model to understand the effect of fished species
+diversity [OLE: Call this diversity of catch rather than fished?] on individual 
+revenue and revenue variability that builds on results
+from vessel [@kasperski2013], fleet [@sethi2012b], and community-level
+[@sethi2014] variability.
+
+We develop a variance-function regression model [@davidian1987; @hall1989], to
+simultaneously models changes in the mean and variance of individual revenue as
+a function of covariates. We simultaneously estimate the effect of species
+diversity (the inverse of Simpson's diversity index [@simpson1949; @jost2006])
+on individual revenue and revenue variability, and estimate effects both within
+and across fishing strategies while controlling for fishing effort and latent
+time effects (e.g. changing environmental conditions). To our knowledge, this is
+the first application of this type of model in any natural resource field.
 
@@
 fishing commercially in Alaska from 1985--2014 (Fig. \ref{fig:ts}A). The
-majority (70--90%) hold a single permit, with a slight increase in single-permit
+majority (70--90%) hold single permits, with a slight increase in single-permit
 holders since 2000 (Fig. \ref{fig:ts}C). This overall trend is most sensitive to
@@
 permit holders are more likely to hold a single permit (Fig. \ref{fig:ts}C).
-Despite making up only 10--30% of the individuals, multiple-permit holders earn
-approximately half the overall revenue (Fig. \ref{fig:ts}B). The mean species
-diversity of this revenue declined somewhat from 1980--2000 but has since
-remained steady (Fig. \ref{fig:ts}D). However, this trend appears largely driven
-by an increase in salmon specialists in the 1980s and 1990s --- the trend
+Despite making up only 10--30% of the individuals, multiple permit holders earn
+approximately half the overall revenue (Fig. \ref{fig:ts}B). The mean effective
+species diversity of this revenue declined somewhat from 1980--2000 but has
+since remained steady (Fig. \ref{fig:ts}D). However, this trend appears largely
+driven by an increase in salmon specialists in the 1980s and 1990s --- the trend
 disappears if salmon license holders are excluded (Fig. \ref{fig:ts}D).
@@
 Individual revenue variability decreases exponentially with higher mean fished
-species diversity across combinations of state and federal fishing permits
-("strategies"; Fig. \ref{fig:across}). Controlling for effort, strategies with
-one unit higher mean species diversity had on average `r h1[[2]]`% (`r
-h1[[1]]`--`r h1[[3]]`% 95% CI) lower variability. This pattern looks
-qualitatively similar with or without individuals targeting salmon. As examples,
-variability was highest for the herring roe gillnet and king crab fisheries.
-Most strategies with the lowest estimated revenue variability combined halibut
-with a salmon permit and had high levels of fished species diversity (Fig.
-\ref{fig:across}).
+species diversity across permit sets (Fig. \ref{fig:across}). Controlling for
+effort, strategies with one unit higher mean effective species diversity had on
+average `r h1[[2]]`% (`r h1[[1]]`--`r h1[[3]]`% 95% CI) lower variability. This
+pattern looks qualitatively similar with or without individuals targeting
+salmon. As examples, variability was highest for the herring roe gillnet and
+king crab fisheries. Most strategies with the lowest estimated revenue
+variability combined halibut with a salmon permit and had high levels of fished
+species diversity (Fig. \ref{fig:across}).
 
@@
 Adding permits also usually increased expected revenue, although the permits
-themselves represent a significant expense.
+themselves represent a significant expense (e.g. ~\$192K USD for a king and
+tanner crab permit in Southeast Alaska in 2016 REF;
+<https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>).
 
@@
 and revenue variability, the effect of changing fished species diversity for an
-individual *within* the constraints of a given set of permits is more variable
-(Fig. \ref{fig:within}). After accounting for fishing effort and changes in
-annual average revenue for fishers with the same set of permits, specializing
-tends to be associated with increased revenue but also increased variability
-(Fig. \ref{fig:within}; Figs S\ref{fig:fe}, S\ref{fig:re}, S\ref{fig:noeffort}).
-On average, there is a ~0.5% increase in the year-to-year revenue ratio per 1%
-decrease in species diversity, and across strategies weighted equally, `r
-g2_dens`% of the probability density supports a positive relationship between
-specialization and revenue.
+individual *within* the constraints of a given permit set is more variable (Fig.
+\ref{fig:within}). After accounting for fishing effort, specializing tends to be
+associated with increased revenue but also increased variability (Fig.
+\ref{fig:within}; Figs S\ref{fig:fe}, S\ref{fig:re}, S\ref{fig:noeffort}). On
+average, there is a ~0.5% increase in the year-to-year revenue ratio per 1%
+decrease in species diversity, and across strategies, `r g2_dens`% of the
+probability density supports a positive relationship between specialization and
+revenue.
 
-Changing (increasing or decreasing) an individual's fished species diversity
-within the constraints of a set of permits tends to be associated with greater
-expected variability or uncertainty in year-to-year revenue (Fig.
+<!--This benefit of specializing is most apparent after accounting for fishing
+effort ().-->
+
+However, changing (increasing or decreasing) an individual's fished species
+diversity within the constraints of a permit set tends to be associated with
+greater expected variability or uncertainty in year-to-year revenue (Fig.
 \ref{fig:within} strategies above zero; Figs. S\ref{fig:fe}, S\ref{fig:re},
@@
 a major regulatory change. Therefore, we refit our models allowing all
-strategy-level effects involving halibut and sablefish permits to vary
-separately before and after the introduction of IFQs. The introduction of IFQs
-reduced variability by ~25% (TODO XX--XX 95% CI) for individuals with only
+strategy-level effects to vary separately before and after the introduction of
+IFQs for strategies involving halibut and sablefish permits. The introduction of
+IFQs reduced variability by ~25% (XX--XX 95% CI) for individuals with only
 halibut or sablefish permits (Fig. S\ref{fig:ifq}A). Furthermore, for most
@@
 For individuals deriving livelihoods from natural resources, theory suggests
-that adopting strategies that exploit a diversity of assets provides substantial
-benefits over strategies with lower diversity [@baldursson1997; @hilborn2001;
-@edwards2004]. However, in practice, the ability of individuals to effectively
-diversify is limited by regulation and market forces. Here we document the
-effect of individual-level changes in species diversity on revenue and revenue
-variability across three scales. We find that (1) individuals with sets of
-permits that allow high levels of species diversity tend to have lower revenue
-variability. (2) Adding a fisheries permit usually increases revenue and
-decreases revenue variability. However, (3) within the constraints of a permit
-and accounting for effort, specializing tends to increase revenue while both
-specializing and generalizing tend to correlate with greater revenue variability
-with considerable differences amongst fisheries. Thus, the most effective
-mechanism by which species diversity can lower revenue variability for
-individual fishers --- switching or adding fishing permits --- is expensive,
-limited by regulations, and consequently unavailable to many individuals.
+that adopting strategies with a diversity of assets provides substantial
+benefits over strategies with lower diversity [REFS]. However, the ability of individuals to
+effectively diversify is limited by regulation and market forces in practice.
+Here we document the effect of changing species diversity on revenue and revenue
+variability across three scale. Each scale reflects individual-level choices
+that can affect species diversity. We find that individuals with
+fishing permit sets that allow high levels of species diversity tend to have
+lower revenue variability. Adding a fisheries permit usually
+increases revenue and decreases revenue variability. Within the
+constraints of a permit and accounting for effort, specializing (reducing
+species diversity) tends to increase revenue while both specializing or generalizing
+tends to be associated with greater revenue variability - though there are considerable
+differences among fisheries. Thus, the most effective mechanism by which
+species diversity can lower revenue variability for individual fishers ---
+switching or adding fishing permits --- is costly, limited by regulations,
+and consequently unavailable to many individuals in practice.
 
-Our findings across fishing strategies are in line with previous research at the
-community and vessel levels [@kasperski2013; @sethi2012b; @sethi2014], but our
-findings within strategies, with respect to individual year-to-year behaviour,
-provide new insights. Here, specializing tended to be associated with increased
-revenue. This could be for a number of reasons. First, specialization can
-increase fishing efficiency [@smith1986]. Efficiency gains from specialization
-can be seen in other natural resource sectors. For example, rice farmers in
-Ghana can increase their efficiency by devoting more land and labor to rice
-production compared to other crops [@abdulai2000]. Second, given the
-availability of valuable resources, specializing can be highly profitable. For
-instance, when cocoa prices have been high, Nigerian farmers have tended to
-specialize on the crop [@mustapha1999] and specializing on valuable tobacco
-farming can be the optimal strategy regardless of alternative crops
-[@chavez2014]. In our analyses, fishers with the opportunity to fish highly
-valuable sockeye salmon benefited the most from specializing. However,
-specializing can have longterm risks if market, management, biological, or
-environmental conditions change. For example, U.S. farmers specialized on wheat,
-corn, or soybeans were hit harder than diversified farmers in the early 1980s
-when a financial crisis hit [@ikerd1990]. Third, diversifying --- a lack of
-specialization --- can happen as a consequence of non-ideal circumstances such
-as poor availability of a more valuable resource. For example, Ugandan
-households relying on forestry resources for income increase their income
-portfolio diversity following downward economic shocks [@debela2012].
+Our results shows additional support for the idea that individual fishers can
+reduce their income variability by diversifying their catch portfolios. However,
+most individuals do not appear to change their catch diversity substantially
+from year to year [OLE: is there a supplemental figure or statistic that speaks to this point?], and most strategies adopted by Alaskan fishermen lie in
+a high-risk, high-reward quadrant in which reducing diversification increases
+both revenue variability and expected revenue. Previous work has suggested that
+changes to fisheries management such as limited entry or catch share programs
+may inadvertently act as barrier to diversification and promote specializing
+(Hilborn et al. 2003; Kasperski and Holland 2013). While some of these
+specialized fishing strategies may be associated with increased revenue
+variability, the benefit may be that they are also associated with increased
+revenue.
 
-We also found that, within strategies, altering species diversity tended to be
-associated with increased revenue variability. There are number of possible
-reasons for this relationship. For instance, local knowledge and experience are
-a major determinant of fishing efficiency [@hilborn1985a]. Therefore, changes to
-fishing patterns may sometimes be associated with a learning or adaptation phase
-as individuals become accustomed to new locations, gear, or species.
-Additionally, changes in diversity might reflect years in which targeted species
-were less available and fishers were forced to target other species or landed
-more bycatch, adding stochasticity to their revenue. Most importantly, this
-increased variability reflects an inability for our large-scale model to predict
-an individual's revenue in a subsequent year. There are many reasons, often
-unique to an individual permit holder's behaviour [@hilborn1985; @hilborn1985a;
-@branch2006] and therefore not included in our large-scale model, that might
-make a specific individual's revenue less predictable if species diversity
-changes.
+External market forces may interact in concert with management
+regulations to promote specialization for fishers or other individuals dependent
+on natural resources. In terrestrial agriculture for instance, there are 
+many instances of specializing being profitable; drivers include the ability to
+specialize on high-profit cash crops (tobacco, cocoa), the availability of
+government agricultural subsidies (soybeans, corn in the USA), and farming on
+a large industrial scale (REF). With increasing environmental variability and extreme
+events in a changing climate (REF), it is particularly important that
+individuals who rely on natural resources as a primary source of income have
+strategies to reduce risk. Though variability in individual revenues may not be
+actively incorporated into many management decisions, analyses like ours help to
+understand the range of diversification strategies available and how their
+benefits may change over time.
 
-There are many possible barriers to diversification for individuals relying on
-natural resources for their livelihood. A primary barrier is the cost of
-diversifying. For commercial fishers, purchasing additional licenses and gear
-typically represents a substantial expense. Related barriers are apparent in
-other natural resource sectors. For example, diversification has been positively
-linked with total income or access to financial capital for farmers
-[@abdulai2001; @barrett2001], small-scale fishing communities [@olale2012], and
-those relying on forest resources [@illukpitiya2008]. Another major barrier to
-diversification can be management restrictions. In commercial fisheries,
-regulations that you have been shown to reduce overfishing --- including
-improved assessments [@melnychuk2017] and individual quota systems
-[@costello2008] --- may ultimately restrict individual diversification
-[@smith1986; @kasperski2013] or at least make developing a diverse fishing
-portfolio more expensive. However, individual quota systems, in particular, may
-also reduce income variability (Fig. S\ref{fig:ifq}A) perhaps offsetting the
-effects of reduced diversification [@holland2016].
+Worldwide, fisheries confront the challenge of sustainable management ().
+While many fisheries are well managed and productive, many are also rebuilding from periods of over-exploitation (Worm et al.
+Science 2009).
+A common tactic to aid in fisheries rebuilding is to improve fishing gear modify regulations to reduce the unintended harvest rebuilding stocks while allowing the continued harvest of healthy stocks (need REF).
+Such bycatch reduction is an important goal for both ecological (Crowder and Murawski 1998, Lewison et al.
+2004) and economic (Curtis and Hicks 2000; Chakravorty and Nemoto 2001) reasons.
+Cumulatively, though, regulatory changes for bycatch reduction as well as other reasons gradually lead to fisheries that target one or a few species and individual fishers wishing to participate in multiple fisheries must own (or purchase) multiple permits.
+Thus a general consequence of fisheries regulations that are shown to improve sustainability of fish stocks and avoid overfishing - including improved assessments (Hilborn and Ovando 2014) and individual quota systems (Costello et al.
+2008 ) - may ultimately restrict individual diversification or at least make developing a diverse fishing portfolio more costly.
+Add Alaska example of IFQs...?
 
-Our analysis provides a synthetic view of the effects of species diversity on
-individuals over 30 years in one of the most productive fishing regions in the
-world. However, this large temporal and geographical scope potentially obscures
-time or regional-variation in the consequences of diversification on revenue and
-revenue variability. We expect future research examining diversification over
-finer scales will uncover additional interesting relationships. Furthermore, in
-addressing the overall pattern, we have averaged out several major discrete
-events, such as the Exxon Valdez oil spill in 1989, that had significant
-consequences for the ecological and social systems in Alaska [@paine1996]. In
-addition to revenue or income variability, it is important to understand the
-benefits of diversification in the context of rare and extreme disturbances
-which occur in natural systems worldwide [@anderson2016].
+Chakravorty, U. and Nemoto, K. (2001) Modeling the effects of area closure and tax policies: a spatial-temporal model of the hawaii longline fishery. Marine Resource Economics 15, 179–204.
 
-With increasing environmental variability and extreme events in a changing
-climate [@easterling2000], it is particularly important that individuals who
-rely on natural resources as a primary source of income, such as fishers,
-farmers, and forestry workers, have effective strategies to reduce risk. While
-diversification can provide a powerful form of risk buffering, we have shown
-that the relationship between individual changes to diversification and income
-variability is complex and context specific. Furthermore, external market forces
-can interact with management regulations, which have other considerable
-benefits, to restrict individual diversification opportunities and promote
-specialization. Although variability in individual revenue may not be actively
-incorporated into management decisions, analyses such as ours help to understand
-the range of diversification strategies available, trade-offs inherent in such
-management decisions, and how these trade-offs may change over time.
+Costello C, Gaines SD, Lynham J (2008) Can catch shares prevent fisheries collapse? Science 321(5896):1678–1681
+
+Crowder, L.B. and Murawski, S.A. (1998) Fisheries bycatch: implications for management. Fisheries Management 23, 8–17.
+
+Curtis, R. and Hicks, R.L. (2000) The Cost of sea turtle preservation: the case of Hawaii’s Pelagic longliners. American Journal of Agricultural Economics 82, 1191– 1197.
+
+Hilborn R, Ovando D. 2014 Reflections on the success of traditional fisheries management. ICES J. Mar. Sci. 71, 1040–1046.
+
+Lewison, R.L., Crowder, L.B., Read, A.J. and Freeman, S.A. (2004) Understanding impacts of fisheries bycatch on marine megafauna. Trends in Ecology & Evolution 19, 598–604.
+
+Read, A.J., Drinker, P. and Northridge, S. (2006) Bycatch of marine mammals in US and global fisheries. Conservation Biology 20, 163–169.
+
+### Reconciling the effects at different levels of species diversity
+
+- why variability/uncertainty might increase with changes in diversity...
+
+- and how fit in with @kasperski2013, @sethi2012b, @sethi2014
+
+- part of the difference is individuals vs vessels. E.g. @kasperski2013
+  suggested that diversity declined following IFQs for halibut sablefish
+  vessels. The reason for the decline was that the vessels were allowed to fish
+  halibut for more days (because people were sharing boats). At the level of
+  people, we wouldn't expect the diversity of someone who only had
+  halibut/sablefish permit to be affected by IFQs. The person, not the vessel
+  holds the quota share: https://alaskafisheries.noaa.gov/fisheries/ifq
+
+- year to year risky... But maybe evens out/better over time
+
+- diversity within permit - often bycatch, limited room for intentional
+  diversification (unless you have a M or salmon permit)
+
+- risk results not really risk, but unpredictability given our covariates ...
+  don't change anything and it's most predictable how well you'll do. What's
+  unpredictable in our model, might be more predictable on a year to year,
+  species by species basis for individual permit holder
+
+- CV vs year-to-year effect for one person
+
+Other:
+
+- specializing can be profitable within some fisheries. Eric: analog here in
+  farming literature is that diversification for small scale farmers generally
+  good, except when they specialize on high demand/high value crops (tobacco,
+  cocoa). Seems like benefits of diversification are also smaller for
+  large-scale industrial farmers.
+
+### Limitations + obvious extensions paragraph by Ole:
+
+Our analysis provides a synthetic view of the effects of species diversity on individuals over 30 years in one of the most productive fisheries regions in the world.
+However, our analysis simultaneously uses data from across the broad geography of Alaska and over 30 years, potentially obscuring regional or time-variation in the consequences of diversification on revenue and revenue variability.
+We expect future research examining diversification over finer spatial and temporal scales will uncover additional interesting relationships.
+Furthermore, in addressing the overall pattern, we have averaged out several important of major discrete events (e.g., the Exxon Valdez oil spill in 1989) that had significant consequences for the ecosystem and social system in Alaska (general EVOS consequences REF).
+We do not wish to minimize the important of such events, but note that it is important to understand the benefits of diversification in the context of rare and extreme disturbances which occur in marine systems worldwide.
+
+There are a number of limitations to our model and extensions that could further
+our understanding of the role of diversification...
+
+- could vary by time and vary by mean diversity
+
+- spatial/geographic variation
+
+- nonlinear effects
+
+- response to specific events like EVOS
+
+### Implications for limited entry, ITQs, and catch shares
+
+- concerning because in many areas of the world and fisheries, regulations and
+  market forces have encouraged specialization / limited (effective)
+  diversification (e.g. multiple permits or diverse permits)
+
+- locking into permits limits ability to pick lower risk strategies and is
+  getting people locked into low diversity permits that may be doing well now,
+  but not necessarily always
+
+### Conclusions + recommendations
+
+- bring it out to wider than fisheries (farming, timber industry, ...)
+
+- 2 dimensions to think about (revenue and variability); and the trade-off is
+  fishery specific: effects within a permit strategy are permit specific;
+  sometimes make more money by specializing... some highly variable single
+  species permits make a ton of money
+
+- broadly, diversification can provide a powerful form of protection from
+  multiple sources of risk to income for those that rely on extracting natural
+  resources, but not all forms of diversification are equal, and on an
+  individual level there can be economic benefits to specializing and risks to
+  adjusting ones strategy
+
+- with increasing environmental variability and extremes due to climate change
+  those that rely on natural resources for their livelihoods need approaches to
+  buffer risk more than ever
+
+- regulations that limit the ability to diversify can have other considerable
+  benefits but it's important to be aware of trade-offs inherent in these
+  regulations when making management decisions
+
+We fit a hierarchical variance-function regression model to 30 years of
+commercial fisheries revenue data in Alaska to simultaneously estimate the
+effect of changes in species diversity on individual's revenue and revenue
+variability.
 
@@
 \caption{Trends in revenue and diversification over time. (A) Map of fishing
-regions in Alaska, USA. (B) Total revenue in billions of 2009 USD from `r
-range_yrs[1]` to `r range_yrs[2]` for multiple (grey) and single (blue) permit
-holders. (C) Proportion of single permit holders for all fisheries (thick black
-line) and with permit holders omitted each year who fished any of the `r
+regions in Alaska, USA. (B) Total revenue in billions of 2009 U.S. dollars from
+`r range_yrs[1]` to `r range_yrs[2]` for multiple (grey) and single (blue)
+permit holders. (C) Proportion of single permit holders for all fisheries (thick
+black line) and with permit holders omitted each year who fished any of the `r 
 n_jacknifed` top species groups (thin lines). (D) Mean effective diversity of
@@
 
+<!-- TODO: get fig:re in results -->
+<!-- Our model estimates overall average effects (Fig. S\ref{fig:fe}) and effects -->
+<!-- that vary by permit set (Fig. S\ref{fig:re}). -->
+
 We obtained commercial fisheries landings and revenue data for all permit
@@
 Commercial Fisheries Entry Commission (CFEC). These data represent a total of
-nearly `r round(ndat$initial_tot_rev/1e9)` billion 2009 USD. We adjusted for
-inflation by setting all revenues to 2009 USD [@bea2016]. We implemented
-a number of filtering steps to focus on individuals actively engaged in
-commercial fishing (Supporting Information). In particular, we removed permit
-holders whose median revenue was less than \$10,000 2009 USD to focus on
-individuals for whom commercial fishing formed a substantial part of their
-income.
+nearly `r round(ndat$initial_tot_rev/1e9)` billion 2009 USD and a combined `r
+round(ndat$initial/1e6,1)` million rows of data. We adjusted for inflation by
+setting all revenues to 2009 USD [@bea2016]. We implemented a number of
+filtering steps to focus on individuals actively engaged in commercial fishing
+(Supporting Information). In particular, we removed permit holders whose median
+revenue was less than \$10,000 2009 USD to focus on individuals for whom
+commercial fishing formed a substantial part of their income.
 
 We used combinations of state and federal fishing permits held by an individual
-in a specific year to define fishing "strategies", representing the species
+in in a specific year to define fishing "strategies", representing the species
 targeted (Supporting Information). For example, we combined permits that were
@@
 roe" permit group. For salmon permits, we formed salmon strategies based on
-similarities in gear type and species composition (Supporting Information).
+similarities in gear type and species compositon (Supporting Information).
 
@@
 revenue variability while controlling for effort. For consistency with previous
-analyses [@vedeld2007; @illukpitiya2008; @debela2012; @kasperski2013;
-sethi2014], we calculated effective species diversity for permit holder $i$ and
-time (year) $t$ as the inverse of Simpson's diversity index $\lambda$
-[@simpson1949; @jost2006] (also referred to in economics as the
-Herfindahl-Hirschman Index [@hirschman1964]) weighted by revenue $R$:
+analyses [@kasperski2013], we calculated effective species diversity for
+permit holder $i$ and time (year) $t$ as the inverse of Simpson's diversity
+index $\lambda$ [@simpson1949; @jost2006] (also referred to in economics as the
+Herfindahl-Hirschman Index) weighted by revenue $R$:
 
@@
 increasing diversity and decreasing diversity allows the costs and benefits of
-these changes to be asymmetric (Fig. S\ref{fig:resid}).
+these changes to be asymmetric.
 
@@
 along with strategy-level intercepts would be estimated at almost exactly zero.
-The estimated $\beta_{j,t}$'s represent years when people with a particular set
-of permits tended to collectively do better or worse than the long-term average
+The estimated $\beta_{j,t}$'s represent years when people with a particular
+permit set tended to collectively do better or worse than the long-term average
 (Fig. S\ref{fig:str-yrs}). Therefore, an individual's expected change in revenue
 (and variability) is after accounting for the general trend in revenue for that
-set of permits.
+permit set.
 
@@
 
-\noindent Again, $S_{i,t}$ and $I_{i,t}$ represent, respectively, the log ratio
-of species diversity from year-to-year and an indicator variable for increasing
-or decreasing species diversity. We exponentiate the equation to ensure all
-scale parameters, $\sigma_{i,t}$, are positive. Residuals from the mean
-component of our model clearly show a "V"-shaped pattern, which the variance
-component of our model can capture (Fig. S\ref{fig:resid}). Furthermore, the
-downside residuals (year-individual data points with lower than expected
-revenue) look qualitatively similar to the complete set of residuals (upside
-and downside) (Fig. S\ref{fig:resid-downside}). Therefore, it seemed
-appropriate to model the variability as a proxy for revenue risk.
+\noindent Again, $S_{i,t}$ and $I_{i,t}$ represent, respectively, percent change
+in species diversity and an indicator variable for increasing or decreasing
+species diversity. We exponentiate the equation to ensure all scale parameters,
+$\sigma_{i,t}$, are positive. Residuals from the mean component of our model
+clearly show a "V"-shaped pattern that the variance component of our model can
+capture (Fig. S\ref{fig:resid}). Furthermore, the downside residuals
+(year-individual data points with lower than expected revenue) look
+qualitatively similar to the complete set of residuals (upside and downside)
+(Fig. S\ref{fig:resid-downside}). Therefore, it seemed appropriate to model the
+variability as a proxy for revenue risk.
 
@@
 
-\noindent The $\gamma_{0,j}$ coefficients are modelled with two strategy-level
-predictors. The symbols $M_j$ and $E_j$ represent the mean species diversity
-and mean combined fishing season length (effort) for strategy $j$. The
-coefficients $\eta_1$ and $\eta_2$ represent predictors in this strategy-level
-regression estimating effects across strategies.
+\noindent The mean of the $\gamma_{0,j}$ coefficients is modelled with two
+strategy-level predictors. The symbols $M_j$ and $E_j$ represent the mean
+species diversity and mean combined fishing season length (effort) for strategy
+$j$. The coefficients $\eta_1$ and $\eta_2$ represent predictors in this
+strategy-level regression estimating effects across strategies.
 
-<!-- TODO: rerun with 2.14.1 -->
+We fit our models with Stan 2.11.1 [@hoffman2014; @standevelopmentteam2015;
+@standevelopmentteam2015a; @carpenter2015] and R 3.3.1 [@rcoreteam2016]. Stan
+implements the No-U-Turn Hamiltonian Markov chain Monte Carlo algorithm
+[@hoffman2014] to perform Bayesian statistical inference. We assigned weakly
+informative priors on all parameters: $N(0, 2^2)$ priors on all $\beta$,
+$\gamma$, and $\eta$ parameters, and half-$t(3, 0, 2)$ priors (i.e. degrees of
+freedom of 3 and scale of 2) on all $\sigma$ and $\tau$. We ran four chains and
+2,000 iterations and discarded the first 1,000 iterations of each chain as warm
+up. We checked for chain convergence visually with trace plots, ensured that
+$\hat{R}<1.05$ (the potential scale reduction factor) and that the effective
+sample size, as calculated by the rstan R package [@standevelopmentteam2015],
+was greater than 200 for all parameters [@gelman2014]. We verified that our
+model returned sensible estimates by comparing our estimates to an ad-hoc
+two-stage model fit with the lme4 R package [@bates2015], where we first fit
+a mixed-effects regression model to the mean, and then modelled the absolute
+residuals from that first model in a second mixed-effects model representing the
+variance component. We also validated the model with simulated data, and were
+able to recover the true values (Supporting Information). Code used in our
+analysis is available at <https://github.com/NCEAS/pfx-commercial/portfolios/>.
 
-We fit our models with Stan 2.14.1 [@standevelopmentteam2016;
-@standevelopmentteam2016a] and R 3.3.2 [@rcoreteam2016]. Stan implements the
-No-U-Turn Hamiltonian Markov chain Monte Carlo algorithm [@hoffman2014] to
-perform Bayesian statistical inference. We assigned weakly informative priors
-on all parameters: $N(0, 2^2)$ priors on all $\beta$, $\gamma$, and $\eta$
-parameters, and half-$t(3, 0, 2)$ priors (i.e. degrees of freedom of 3 and
-scale of 2) on all $\sigma$ and $\tau$. We ran four chains and 2,000 iterations
-and discarded the first 1,000 iterations of each chain as warm up. We checked
-for chain convergence visually with trace plots, ensured that $\hat{R}<1.05$
-(the potential scale reduction factor) and that the effective sample size, as
-calculated by the rstan R package [@standevelopmentteam2016], was greater than
-200 for all parameters [@gelman2014]. We verified that our model returned
-sensible estimates by comparing our estimates to an ad-hoc two-stage model fit
-with the lme4 R package [@bates2015], where we first fit a mixed-effects
-regression model to the mean, and then modelled the absolute residuals from
-that first model in a second mixed-effects model representing the variance
-component. We also validated the model with simulated data, and were able to
-recover the true values (Supporting Information). Code used in our analysis is
-available at <https://github.com/NCEAS/pfx-commercial/portfolios/>.
+<!-- # Notes -->
 
-# Acknowledgments
+<!-- ## @kasperski2013 -->
+<!-- - vessel-level CFEC data -->
+<!-- - dome-shaped result -->
+<!-- - Herfindahl-Hirschman Index/HHI (Simpson) -->
+<!-- - diversity decreasing for those still fishing in 2010, but not otherwise -->
 
-TODO. National Center for Ecological Analysis and Synthesis (NCEAS) working
-group *Applying portfolio effects to the Gulf of Alaska ecosystem: Did
-multi-scale diversity buffer against the Exxon Valdez oil spill?*. Other
-participants... Smith...
+<!-- ## @sethi2012b -->
+<!-- - risk measures to 90 Alaskan fishing fishery/fleets -->
+<!-- - revenue more variable than catch (+ price = extra variability) -->
+<!-- - large catches, big fleets = lower risk -->
+
+<!-- ## @sethi2010b -->
+<!-- - risk management for fisheries review -->
+
+<!-- ## @sethi2012 -->
+<!-- - equal-share catch-pooling cooperatives in king and snow crab -->
+<!-- - "variability reduction depends on the degree of correlation amongst -->
+<!--   participants’ catches" -->
+
+<!-- ## @sethi2014 -->
+<!-- - community-level CFEC data -->
+<!-- - number of fisheries strongest predictor of (lower) CV/CVaR -->
+<!-- - also simpson index -->
+<!-- - suggest barriers to commercial entry increase community risk by restricting -->
+<!--   access to 'variety of fisheries' -->
+<!-- - some communities have limited ability to diversify; may need to diversify -->
+<!--   outside of fishing -->
+
+<!-- [@perruso2005] -->
+<!-- [@hilborn2001] -->
+<!-- [@edwards2004] -->
+<!-- [@trenkel2013] -->
+<!-- [@mumford2009] -->
+<!-- [@oostenbrugge2002] -->
+<!-- [@baldursson1997] -->
+<!-- [@jardine2015] -->
+<!-- [@sanchirico2008] -->
 

### 2017-01-13 portfolio/text/anderson-etal-rev-pfx.tex
Add address; tweak makefile
Format paper

--- 2016-09-27 portfolio/text/anderson-etal-rev-pfx.tex
@@
-\documentclass[12pt,english]{article}
-\usepackage{geometry}
-\geometry{verbose, letterpaper, tmargin = 2.54cm, bmargin = 2.54cm,
-  lmargin = 2.54cm, rmargin = 2.54cm}
-\geometry{letterpaper}
-\usepackage{graphicx}
-\usepackage{amsmath}
-\usepackage{setspace}
-\usepackage{booktabs}
-\usepackage{ragged2e}
-\usepackage{longtable}
-\usepackage{multirow}
-\usepackage{array}
-\usepackage{tabu} % for spacing between rows in longtable
-\usepackage{titlesec}
-\usepackage{url}
-\usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
-\usepackage{commath} % for \abs{}
+\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
+% Use the lineno option to display guide line numbers if required.
+% Note that the use of elements such as single-column equations
+% may affect the guide line number alignment.
 
-\usepackage{lineno}
-\usepackage{xcolor}
-\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
-\modulolinenumbers[2]
+\templatetype{pnasresearcharticle} % Choose template
+% {pnasresearcharticle} = Template for a two-column research article
+% {pnasmathematics} = Template for a one-column mathematics article
+% {pnasinvited} = Template for a PNAS invited submission
 
-% Linux Libertine:
-\usepackage{textcomp}
-\usepackage[sb]{libertine}
-\usepackage[varqu,varl]{inconsolata}% sans serif typewriter
-\usepackage[libertine,bigdelims,vvarbb]{newtxmath} % bb from STIX
-\usepackage[cal=boondoxo]{mathalfa} % mathcal
-\useosf % osf for text, not math
-\usepackage[supstfm=libertinesups,%
-  supscaled=1.2,%
-  raised=-.13em]{superiors}
+\title{Benefits and risks of diversification for individual fishers}
 
-\textheight 22.0cm
+% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
+\author[a]{Sean C. Anderson}
+\author[b]{Eric J. Ward}
+\author[b]{Andrew O. Shelton}
+\author[c]{Milo D. Adkison}
+\author[c]{Anne H. Beaudreau}
+\author[d]{Richard E. Brenner}
+\author[e]{Alan C. Haynie}
+\author[d]{Jennifer C. Shriver}
+\author[f]{Jordan T. Watson}
+\author[d]{Benjamin C. Williams}
 
-\usepackage[round,sectionbib]{natbib}
-\bibpunct{(}{)}{;}{a}{}{;}
+\affil[a]{School of Aquatic and Fishery Sciences, Box 355020,
+University of Washington, Seattle, WA 98195, USA}
 
-% Fix line numbering and align environment
-% http://phaseportrait.blogspot.ca/2007/08/lineno-and-amsmath-compatibility.html
-\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
-  \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
-  \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
-  \renewenvironment{#1}%
-     {\linenomath\csname old#1\endcsname}%
-     {\csname oldend#1\endcsname\endlinenomath}}%
-\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
-  \patchAmsMathEnvironmentForLineno{#1}%
-  \patchAmsMathEnvironmentForLineno{#1*}}%
-\AtBeginDocument{%
-\patchBothAmsMathEnvironmentsForLineno{equation}%
-\patchBothAmsMathEnvironmentsForLineno{align}%
-\patchBothAmsMathEnvironmentsForLineno{flalign}%
-\patchBothAmsMathEnvironmentsForLineno{alignat}%
-\patchBothAmsMathEnvironmentsForLineno{gather}%
-\patchBothAmsMathEnvironmentsForLineno{multline}%
+\affil[b]{Conservation Biology Division, Northwest Fisheries Science Center,
+National Marine Fisheries Service, National Oceanographic and
+Atmospheric Administration, 2725 Montlake Blvd E, Seattle, WA 98112, USA}
+
+\affil[c]{College of Fisheries and Ocean Sciences, University of Alaska Fairbanks,
+17101 Point Lena Loop Rd, Juneau, AK, 99801, USA}
+
+\affil[d]{Alaska Department of Fish and Game,
+Division of Commercial Fisheries,
+PO BOX 115526,
+Juneau, AK, USA 99811-5526, USA}
+
+\affil[e]{Alaska Fisheries Science Center,
+National Marine Fisheries Service, National Oceanographic and
+Atmospheric Administration, Bldg 4,
+7600 Sand Point Way NE, Seattle, WA, 98115, USA}
+
+\affil[f]{NOAA / National Marine Fisheries Service, Alaska Fisheries Science Center, Auke
+Bay Laboratories, 17109 Pt. Lena Loop Rd., Juneau, AK 99801, USA}
+
+% Please give the surname of the lead author for the running footer
+\leadauthor{Anderson}
+
+% Please add here a significance statement to explain the relevance of your work
+\significancestatement{TODO}
+
+% Please include corresponding author, author contribution and author declaration information
+\authorcontributions{
+S.C.A., TODO designed the study; S.C.A., TODO analyzed the
+data and wrote the paper with input from all authors.}
+\authordeclaration{The authors declare no conflict of interest.}
+\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: sandrsn@uw.edu}
+
+% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
+\keywords{
+TODO $|$
+TODO
 }
 
-\title{Benefits of diversification and risks of changing strategies for individual fishers\\
-Benefits of diversification are unavailable to many fishermen\\
-}
-\author{
-Sean C. Anderson$^1$,
-Eric J. Ward$^2$,
-Milo D. Adkison$^3$,
-Anne H. Beaudreau$^3$,\\
-Richard E. Brenner$^4$,
-Alan C. Haynie$^5$,
-Andrew O. Shelton$^2$,\\
-Jennifer C. Shriver,
-Jordan T. Watson$^X$,
-\ldots
-}
+\newcommand{\tabstrategies}{1}
+\newcommand{\figfe}{1}
+\newcommand{\figre}{2}
+\newcommand{\figspokeall}{3}
+\newcommand{\figresid}{4}
+\newcommand{\figresiddownside}{5}
+\newcommand{\figstryrs}{6}
+\newcommand{\figifq}{7}
+\newcommand{\figmirror}{8}
+\newcommand{\fignoeffort}{9}
 
-\newlabel{tab:strategies}{1}
-\newlabel{fig:fe}{1}
-\newlabel{fig:re}{2}
-\newlabel{fig:spoke-all}{3}
-\newlabel{fig:resid}{4}
-\newlabel{fig:resid-downside}{5}
-\newlabel{fig:str-yrs}{6}
+% 250:
+\begin{abstract}
+Individuals relying on the extraction of natural resources for their
+livelihood---fishers, farmers, miners, and loggers, among others---face high
+levels of income variability driven by a mix of environmental, biological,
+management, and
+economic dynamics. A key component to managing these industries is identifying
+how regulatory actions and individual behaviour affect income variability and by
+extension economic stability and the sustainable use of natural resources. For
+commercial fisheries, communities and vessels fishing a greater diversity of
+species have less revenue variability than those fishing a lower diversity.
+However, it is unclear if these benefits extend to the actions of individual
+fishers and how year-to-year changes in diversification affect revenue and
+revenue variability. Here, we identify three axes by which fishers in Alaska,
+one of the world's most productive fishing regions, can diversify fishing
+activities. We show that, despite increasing specialization over the last 30
+years, fishing permits with higher species diversity reduces individual revenue
+variability and fishing an additional permit is associated with higher revenue
+and lower variability. However, increasing species diversity within the
+constraints of existing permits has a fishery-dependent effect on revenue and is
+usually (93\% probability) associated with increased revenue uncertainty
+the following year. Our results demonstrate that the most effective option for
+individuals to decrease revenue variability is to purchase permits to
+participate in additional or more diverse fisheries. However, this option is
+expensive, often limited by regulations such as catch share programs, and
+consequently unavailable to many individuals. With increasing environmental
+variability in a changing climate, it will be particularly important that
+individuals relying on natural resources for their livelihood have effective
+strategies to reduce income variability.
+ % (`r g2_dens`% probability)
+\end{abstract}
 
-\date{}
-
-% remove numbers in front of sections:
-\makeatletter
-\renewcommand\@seccntformat[1]{}
-\makeatother
+\dates{This manuscript was compiled on \today}
+\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
 
 \begin{document}
-% \begin{flushleft}
+
+% Optional adjustment to line up main text (after abstract) of first page with line numbers, when using both lineno and twocolumn options.
+% You should only change this length when you've finalised the article contents.
+\verticaladjustment{-2pt}
 
 \maketitle
+\thispagestyle{firststyle}
+\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
 
-\noindent
-$^1$School of Aquatic and Fishery Sciences, Box 355020,
-University of Washington, Seattle, WA 98195, USA
+% If your first paragraph (i.e.
+with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation.
+If this is the case, add \parshape=0 to the end of the list environment.
 
-\noindent
-$^2$Conservation Biology Division, Northwest Fisheries Science Center,
-National Marine Fisheries Service, National Oceanographic and
-Atmospheric Administration, 2725 Montlake Blvd E, Seattle, WA 98112, USA
-
-\noindent
-$^5$Alaska Fisheries Science Center,
-National Marine Fisheries Service, National Oceanographic and
-Atmospheric Administration, Bldg 4,
-7600 Sand Point Way NE, Seattle, WA, 98115, USA
-
-\noindent
-$^X$Ted Stevens Marine Research Institute, Alaska Fisheries Science Center, 17109
-Pt Lena Loop Road, Juneau, AK 99801, USA
-
-\noindent
-Affiliations \ldots
-
-\setlength{\parindent}{0.9cm}
-\setlength\parskip{0in}
-
-% \vspace{-1cm}
-
-%\input{../analysis/values} % R output
-
-\thispagestyle{empty}
-
-\clearpage
-
-\linenumbers
-
-\begin{spacing}{1.4}
 \input{ms}
 
-\end{spacing}
-% \end{flushleft}
+\matmethods{
+We obtained commercial fisheries landings and revenue data for all permit holders in Alaska from 1985 to 2014 from the Alaskan Commercial Fisheries Entry Commission (CFEC).
+These data represent a total of nearly 43 billion 2009 USD.
+We adjusted for inflation by setting all revenues to 2009 USD \cite{bea2016}.
+We implemented a number of filtering steps to focus on individuals actively engaged in commercial fishing (Supporting Information).
+In particular, we removed permit holders whose median revenue was less than \$10,000 USD to focus on individuals for whom commercial fishing formed a substantial part of their income.
+
+We used combinations of state and federal fishing permits held by an individual in a specific year to define fishing ``strategies'', representing the species targeted (Supporting Information).
+For example, we combined permits that were otherwise the same except for vessel size (e.g.~a longliner fishing sablefish on a vessel under 60 ft was considered to have the same strategy as a longliner fishing sablefish on a vessel over 60 ft), and we combined a number of permits that were targeting a single species and differed only in region fished.
+For example, someone fishing herring roe in Kodiak and someone fishing herring roe in Alaska Peninsula were both considered to have a ``herring roe'' permit group.
+For salmon permits, we formed salmon strategies based on similarities in gear type and species composition (Supporting Information).
+
+After this process of combining permits, we were left with 23 unique fishing permit groups (Table~S\tabstrategies), which we aggregated within each person-year combination to form permit strategies.
+For example, if someone fished halibut and sablefish in the same year, their strategy for that year would be ``halibut-sablefish''.
+Specific details on our strategy definitions are available in the Supporting Information.
+
+Our models focused on species diversity as a possible predictor of revenue and revenue variability while controlling for effort.
+For consistency with previous analyses {[}\cite{vedeld2007}; \cite{illukpitiya2008}; \cite{debela2012}; \cite{kasperski2013}; sethi2014{]}, we calculated effective species diversity for permit holder \(i\) and time (year) \(t\) as the inverse of Simpson's diversity index \(\lambda\) \cite{simpson1949, jost2006} (also referred to in economics as the Herfindahl-Hirschman Index \cite{hirschman1964}) weighted by revenue \(R\):
+
+\begin{equation}
+1/\lambda_{i,t} = 1 / \left(\sum\limits_{s=1}^{n_{i,t}}
+  \left(R_{i,s,t}/ \sum\limits_{s=1}^{n_{i,t}} R_{i,s,t}\right)\right)^2,
+\end{equation}
+
+\noindent where \(s\) indexes species from \(1\) through \(n\).
+We calculated effort or effective season length as the sum of the season lengths for each permit each year.
+Season length for a given permit was calculated as the number of days between the first instance of fishing a permit and the last instance of fishing a permit each year.
+For example, if someone fished for 30 days on a salmon drift gillnet permit in Southeast Alaska and 40 days on a statewide longline halibut permit, that individual's season length was calculated as 70 days.
+
+\subsection{Hierarchical model}\label{hierarchical-model}
+
+To jointly model revenue and revenue variability as a function of individual-level covariates, we extended the basic variance-function regression to a hierarchical Bayesian variance-function regression model with Gaussian errors predicting \(\log\) revenue, \(\log(R_{i,t})\), for each fisher \(i\) and time \(t\):
+
+\begin{equation}
+\log(R_{i,t}) \sim \mathrm{N} \left(\mu_{i,t}, \sigma_{i,t}^2 \right).
+\end{equation}
+
+\noindent Each year \(t\) of revenue for fisher \(i\) is assigned a strategy (permit set) indexed by \(j\) (Supporting Information, Table~\tabstrategies) and \(\mu_{i,t}\) is modelled as:
+
+\begin{equation}
+\label{eq:mean}
+\mu_{i,t} = \beta_{0,j,t} -
+  \beta_{1,j} S_{i,t} I_{i,t} +
+  \beta_{2,j} S_{i,t} I_{i,t} +
+  \beta_3 D_{i,t} -
+  \beta_4 S_{i,t} D_{i,t} I_{i,t} +
+  \beta_5 S_{i,t} D_{i,t} I_{i,t} +
+  \log(R_{i,t-1}),
+\end{equation}
+
+\noindent where \(S_{i,t}\) is the log ratio of species diversity from year-to-year, \(\log \left(\mathrm{div}_{i,t}/\mathrm{div}_{i,t-1}\right)\), \(D_{i,t}\) is the log ratio of days fished from year-to-year, \(\log \left(\mathrm{days}_{i,t}/\mathrm{days}_{i,t-1}\right)\), and \(I_{i,t}\) is an indicator variable that takes a value \(0\) if \(\mathrm{div}_{i,t} \ge \mathrm{div}_{i,t-1}\) (species diversity increasing; generalizing) and takes a value \(1\) if \(\mathrm{div}_{i,t} < \mathrm{div}_{i,t-1}\) (species diversity decreasing; specializing).
+Estimating a separate coefficient for the effect of increasing diversity and decreasing diversity allows the costs and benefits of these changes to be asymmetric (Fig.~S\figresid).
+
+We use \(\log(R_{i,t-1}\)) as an offset, and thus our model describes the log ratio of revenue from year-to-year, i.e., \(\log(R_{i,t}) - \log(R_{i,t-1}) = \log(R_{i,t}/R_{i,t-1})\).
+We model change in revenue because preliminary analysis revealed non-stationary trends in revenue (resulting from experience or other unmodelled factors).
+This approach eliminates having to deal with the non-stationary time series and eliminates having to estimate thousands of mean-revenue intercepts associated with individual fishers.
+We removed person-year combinations without a subsequent year of revenue.
+
+The \(\beta\) terms represent estimated coefficients, some of which are allowed to vary as ``random effect'' terms.
+The intercepts \(\beta_{0,j,t}\) are allowed to vary by strategy-year combinations, are centred on zero, and are constrained by a normal distribution:
+
+\begin{equation}
+\beta_{0,j,t} \sim \mathrm{N} \left(0, \sigma^2_{\beta_{j,t}} \right).
+\end{equation}
+
+\noindent This approach, combined with the lack of a global intercept, constrains the mean change in log revenue across all fishers and years to be zero.
+I.e., we assume that the data are stationary after accounting for all covariates and including an intercept that varies with strategy-year combinations.
+Initial versions of the model confirmed that the global intercept along with strategy-level intercepts would be estimated at almost exactly zero.
+The estimated \(\beta_{j,t}\)'s represent years when people with a particular set of permits tended to collectively do better or worse than the long-term average (Fig.~S\figstryrs).
+Therefore, an individual's expected change in revenue (and variability) represents the expectation after accounting for the general trend in revenue for that set of permits.
+
+The slopes \(\beta_{1,j}\) and \(\beta_{1,j}\) are then allowed to vary by strategy with estimated means:
+
+\begin{align}
+\beta_{1,j} &\sim \mathrm{N} \left(\mu_{\beta_1}, \tau^2_{\beta_1}\right),\\
+\beta_{2,j} &\sim \mathrm{N} \left(\mu_{\beta_2}, \tau^2_{\beta_2}\right).
+\end{align}
+
+\noindent The \(\beta_{1,j}\)'s and \(\beta_{2,j}\)'s represent how much a 1\% decrease and 1\% increase, respectively, in the ratio of species diversity from year-to-year within a strategy translates to a given percent change in the ratio of revenue from year-to-year.
+These are after controlling for changes in effort, \(\beta_3\), and the interaction between changes in effort and changes in species diversity, \(\beta_4\) and \(\beta_5\).
+
+Whereas traditional hierarchical linear models assume homoscedasticity --- that the residual variance does not vary systematically --- our variance-function regression model allows the residual variance to be heteroscedastic and vary as a function of predictors.
+Instead of estimating a single residual error scale term, \(\sigma\), we model the scale of the residual error with another hierarchical model with a similar form:
+
+\begin{equation}
+\label{eq:cv}
+\sigma_{i,t} =
+\exp \left( \gamma_{0,j} -
+\gamma_{1,j} S_{i,t} I_{i,t} +
+\gamma_{2,j} S_{i,t} I_{i,t} +
+\gamma_3 D_{i,t} -
+\gamma_4 S_{i,t} D_{i,t} I_{i,t} +
+\gamma_5 S_{i,t} D_{i,t} I_{i,t}\right).
+\end{equation}
+
+\noindent Again, \(S_{i,t}\) and \(I_{i,t}\) represent, respectively, the log ratio of species diversity from year-to-year and an indicator variable for increasing or decreasing species diversity.
+We exponentiate the equation to ensure all scale parameters, \(\sigma_{i,t}\), are positive.
+Residuals from the mean component of our model clearly show a ``V''-shaped pattern, which the variance component of our model can capture (Fig.~S\figresid).
+Furthermore, the downside residuals (year-individual data points with lower than expected revenue) look qualitatively similar to the complete set of residuals (upside and downside) (Fig.~S\figresiddownside).
+Therefore, it seemed appropriate to model the variability as a proxy for revenue risk.
+
+The \(\gamma\)'s are estimated coefficients, three of which are allowed to vary by strategy, constrained by normal distributions:
+
+\begin{align}
+\gamma_{0,j}  &\sim \mathrm{N} \left(\eta_0 + \eta_1 M_j + \eta_2 E_j, \tau^2_{\gamma_0}\right),\\
+\gamma_{1,j}  &\sim \mathrm{N} \left(\mu_{\gamma_1}, \tau^2_{\gamma_1}\right),\\
+\gamma_{2,j}  &\sim \mathrm{N} \left(\mu_{\gamma_2}, \tau^2_{\gamma_2}\right).
+\end{align}
+
+\noindent The \(\gamma_{0,j}\) coefficients are modelled with two strategy-level predictors.
+The symbols \(M_j\) and \(E_j\) represent the mean species diversity and mean combined fishing season length (effort) for strategy \(j\).
+The coefficients \(\eta_1\) and \(\eta_2\) represent predictors in this strategy-level regression estimating effects across strategies.
+
+We fit our models with Stan 2.14.1 \cite{standevelopmentteam2016, standevelopmentteam2016a} and R 3.3.2 \cite{rcoreteam2016}.
+Stan implements the No-U-Turn Hamiltonian Markov chain Monte Carlo algorithm \cite{hoffman2014} to perform Bayesian statistical inference.
+We assigned weakly informative priors on all parameters: \(N(0, 2^2)\) priors on all \(\beta\), \(\gamma\), and \(\eta\) parameters, and half-\(t(3, 0, 2)\) priors (i.e.~degrees of freedom of 3 and scale of 2) on all \(\sigma\) and \(\tau\).
+We ran four chains and 2,000 iterations and discarded the first 1,000 iterations of each chain as warm up.
+We checked for chain convergence visually with trace plots, ensured that \(\hat{R}<1.05\) (the potential scale reduction factor) and that the effective sample size, as calculated by the rstan R package \cite{standevelopmentteam2016}, was greater than 200 for all parameters \cite{gelman2014}.
+We verified that our model returned sensible estimates by comparing our estimates to an ad-hoc two-stage model fit with the lme4 R package \cite{bates2015}, where we first fit a mixed-effects regression model to the mean, and then modelled the absolute residuals from that first model in a second mixed-effects model representing the variance component.
+We also validated the model with simulated data, and were able to recover the true values (Supporting Information).
+Code used in our analysis is available at \url{https://github.com/NCEAS/pfx-commercial/portfolios/}.
+}
+\showmatmethods
+
+\showacknow % Display the acknowledgments section
+
+% \pnasbreak splits and balances the columns before the references.
+% If you see unexpected formatting errors, try commenting out this line
+% as it can run into problems with floats and footnotes on the final page.
+\pnasbreak
+
+% Bibliography
+\bibliography{refs}
+
 \end{document}

### 2017-01-13 portfolio/text/ms.Rmd
Format paper

--- 2017-01-06 portfolio/text/ms.Rmd
@@
 
-<!-- Introduction, Methods, Results, Discussion, Conclusion -->
-<!-- 600, 900, 500, 800, 200, -->
-
-# Abstract
-
-<!-- TODO < 250 words -->
-
-Individuals relying on the extraction of natural resources for their livelihood
---- fishers, farmers, miners, and loggers, among others --- face high levels of
-income variability driven by a mix of environmental, biological, management, and
-economic dynamics. A key component to managing these industries is identifying
-how regulatory actions and individual behaviour affect income variability and by
-extension economic stability and the sustainable use of natural resources. For
-commercial fisheries, communities and vessels fishing a greater diversity of
-species have less revenue variability than those fishing a lower diversity.
-However, it is unclear if these benefits extend to the actions of individual
-fishers and how year-to-year changes in diversification affect revenue and
-revenue variability. Here, we identify three axes by which fishers in Alaska,
-one of the world's most productive fishing regions, can diversify fishing
-activities. We show that, despite increasing specialization over the last 30
-years, fishing permits with higher species diversity reduces individual revenue
-variability and fishing an additional permit is associated with higher revenue
-and lower variability. However, increasing species diversity within the
-constraints of existing permits has a fishery-dependent effect on revenue and is
-usually (`r g2_dens`% probability) associated with increased revenue uncertainty
-the following year. Our results demonstrate that the most effective option for
-individuals to decrease revenue variability is to purchase permits to
-participate in additional or more diverse fisheries. However, this option is
-expensive, often limited by regulations such as catch share programs, and
-consequently unavailable to many individuals. With increasing environmental
-variability in a changing climate, it will be particularly important that
-individuals relying on natural resources for their livelihood have effective
-strategies to reduce income variability.
-
-<!-- TODO: take home is the management context is key and controls show that
-it's more complicated. TODO mentioned that specializing tends to be profitable
--->
-
-\clearpage
-
 # Introduction
 
-Managing natural resources and sustaining the livelihoods of those that depend
+\dropcap{M}anaging natural resources and sustaining the livelihoods of those that depend
 on their extraction is hard. These livelihoods tend to have high annual
@@
 
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=0.5\textwidth]{../figs/fig1-jk.pdf}
+
+\caption{Trends in revenue and diversification over time. (A) Map of fishing
+regions in Alaska, USA. (B) Total revenue in billions of USD from `r
+range_yrs[1]` to `r range_yrs[2]` for multiple (grey) and single (blue) permit
+holders. (C) Proportion of single permit holders for all fisheries (thick black
+line) and with permit holders omitted each year who fished any of the `r
+n_jacknifed` top species groups (thin lines). (D) Mean effective diversity of
+species fished (Simpson's diversity index) for all fisheries (thick black line)
+and with permit holders omitted each year who fished any of the `r n_jacknifed`
+top species groups (thin lines). } \label{fig:ts} \end{center}
+\end{figure}
+
+\begin{figure}[htb]
+\begin{center}
+\includegraphics[width=0.5\textwidth]{../figs/stan-across-strategy-variability2.pdf}
+
+\caption{Individuals who fish permits and permit combinations (permit
+strategies) with higher species diversity have lower expected revenue
+variability. On the y-axis, dots, thick lines, and thin lines represent
+posterior medians, 50\%, and 95\% credible intervals for the `r nstr` most
+common permit strategies. Thick black line, and dark and light grey shaded
+regions indicate median, 50\%, and 95\% credible intervals of strategy-level
+regression built into the hierarchical model. Estimated variability represents
+expected standard deviation for a permit holder who does not change species
+diversity or days fished from year-to-year. }
+
+\label{fig:across}
+\end{center}
+\end{figure}
+
 Individual revenue variability decreases exponentially with higher mean fished
@@
 from purchasing additional permits (Fig. \ref{fig:spokes}, Fig.
-S\ref{fig:spoke-all}). For example, individuals fishing a halibut permit plus
+S\figspokeall). For example, individuals fishing a halibut permit plus
 one additional permit always had lower expected revenue variability compared to
@@
 
+\begin{figure*}[htbp]
+\begin{center}
+\includegraphics[width=0.8\textwidth]{../figs/stan-gg-spoke2.pdf}
+
+\caption{Diversifying permits reduces risk and usually increases revenue. Two
+examples of expected revenue variability and median revenue for owners of (A)
+individual halibut and (B) gillnet herring roe permits (dark blue) compared with
+owners of the same permit plus one more (medium blue) or two more (light
+blue) permits. On the y-axis, dots and lines represent posterior median and 95\%
+credible intervals of expected standard deviations of year-to-year revenue rates
+of return.}
+
+\label{fig:spokes}
+\end{center}
+\end{figure*}
+
 In contrast to the clear across-strategy effects of species diversity on revenue
@@
 tends to be associated with increased revenue but also increased variability
-(Fig. \ref{fig:within}; Figs S\ref{fig:fe}, S\ref{fig:re}, S\ref{fig:noeffort}).
+(Fig. \ref{fig:within}; Figs S\figfe, S\figre, S\ref{fig:noeffort}).
 On average, there is a ~0.5% increase in the year-to-year revenue ratio per 1%
@@
 
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=0.5\textwidth]{../figs/stan-offset-break-anti-spaghetti-specializing.pdf}
+
+\caption{Effect of specializing on year-to-year changes in revenue and revenue
+variability for individual permit holders \emph{within} the `r nstr` most common
+strategies. Changing species diversity from year-to-year is associated with
+higher variability for individuals within the strategies above the zero line.
+Specializing is associated with greater expected revenue ratios to the right of
+the zero line and lower expected revenue ratios to the left of the zero line.
+For example, individuals with setnet permits for salmon in Prince William Sound,
+Cook Inlet, Kodiak, Bristol Bay, or the Alaska Peninsula (second purple dot from
+right) target primarily sockeye salmon and experience a $\sim$1.5\% increase in
+their year-to-year revenue ratio (horizontal axis) and a $\sim$0.5\% increase in
+their revenue variability (vertical axis) for a 1\% decrease in their ratio of
+year-to-year species diversity. These effects are estimated holding fishing
+season length constant from year-to-year. Dots and line segments represent
+medians and 50\% credible intervals of posteriors. The effect of generalizing
+(increasing species diversity) is similar (Fig. S\figmirror).}
+
+\label{fig:within}
+\end{center}
+\end{figure}
+
 Changing (increasing or decreasing) an individual's fished species diversity
@@
 expected variability or uncertainty in year-to-year revenue (Fig.
-\ref{fig:within} strategies above zero; Figs. S\ref{fig:fe}, S\ref{fig:re},
-S\ref{fig:mirror}). This effect was strongest for individuals with a halibut
+\ref{fig:within} strategies above zero; Figs. S\figfe, S\figre,
+S\figmirror). This effect was strongest for individuals with a halibut
 permit or halibut and sablefish permits. Individuals with these strategies
@@
 reduced variability in revenue by ~25% (TODO XX--XX 95% CI) for individuals with
-only halibut or sablefish permits (Fig. S\ref{fig:ifq}A). Furthermore, for most
+only halibut or sablefish permits (Fig. S\figifq A). Furthermore, for most
 strategies, the association between changing species diversity and increased
 revenue variability became stronger after the introduction of IFQs (Fig.
-S\ref{fig:ifq}B). This was especially true for individuals fishing only halibut
+S\figifq B). This was especially true for individuals fishing only halibut
 or sablefish and less so for individuals combining halibut or sablefish with
-a salmon permit (Fig. S\ref{fig:ifq}B).
+a salmon permit (Fig. S\figifq B).
 
@@
 individual quota systems, in particular, may also reduce income variability
-(Fig. S\ref{fig:ifq}A) perhaps offsetting the effects of reduced diversification
+(Fig. S\figifq A) perhaps offsetting the effects of reduced diversification
 [@holland2016].
@@
 
-\bibliographystyle{pnas2011}
-\bibliography{refs}
-
-\clearpage
-
-# Figures
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[width=0.7\textwidth]{../figs/fig1-jk.pdf}
-
-\caption{Trends in revenue and diversification over time. (A) Map of fishing
-regions in Alaska, USA. (B) Total revenue in billions of USD from `r
-range_yrs[1]` to `r range_yrs[2]` for multiple (grey) and single (blue) permit
-holders. (C) Proportion of single permit holders for all fisheries (thick black
-line) and with permit holders omitted each year who fished any of the `r
-n_jacknifed` top species groups (thin lines). (D) Mean effective diversity of
-species fished (Simpson's diversity index) for all fisheries (thick black line)
-and with permit holders omitted each year who fished any of the `r n_jacknifed`
-top species groups (thin lines). } \label{fig:ts} \end{center} \end{figure}
-
-\clearpage
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[width=0.7\textwidth]{../figs/stan-across-strategy-variability2.pdf}
-
-\caption{Individuals who fish permits and permit combinations (permit
-strategies) with higher species diversity have lower expected revenue
-variability. On the y-axis, dots, thick lines, and thin lines represent
-posterior medians, 50\%, and 95\% credible intervals for the `r nstr` most
-common permit strategies. Thick black line, and dark and light grey shaded
-regions indicate median, 50\%, and 95\% credible intervals of strategy-level
-regression built into the hierarchical model. Estimated variability represents
-expected standard deviation for a permit holder who does not change species
-diversity or days fished from year-to-year. }
-
-\label{fig:across}
-\end{center}
-\end{figure}
-
-\clearpage
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[width=0.9\textwidth]{../figs/stan-gg-spoke2.pdf}
-
-\caption{Diversifying permits reduces risk and usually increases revenue. Two
-examples of expected revenue variability and median revenue for owners of (A)
-individual halibut and (B) gillnet herring roe permits (dark blue) compared with
-owners of the same permit plus one more (medium blue) or two more (light
-blue) permits. On the y-axis, dots and lines represent posterior median and 95\%
-credible intervals of expected standard deviations of year-to-year revenue rates
-of return.}
-
-\label{fig:spokes}
-\end{center}
-\end{figure}
-
-\clearpage
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[width=0.6\textwidth]{../figs/stan-offset-break-anti-spaghetti-specializing.pdf}
-
-\caption{Effect of specializing on year-to-year changes in revenue and revenue
-variability for individual permit holders \emph{within} the `r nstr` most common
-strategies. Changing species diversity from year-to-year is associated with
-higher variability for individuals within the strategies above the zero line.
-Specializing is associated with greater expected revenue ratios to the right of
-the zero line and lower expected revenue ratios to the left of the zero line.
-For example, individuals with setnet permits for salmon in Prince William Sound,
-Cook Inlet, Kodiak, Bristol Bay, or the Alaska Peninsula (second purple dot from
-right) target primarily sockeye salmon and experience a $\sim$1.5\% increase in
-their year-to-year revenue ratio (horizontal axis) and a $\sim$0.5\% increase in
-their revenue variability (vertical axis) for a 1\% decrease in their ratio of
-year-to-year species diversity. These effects are estimated holding fishing
-season length constant from year-to-year. Dots and line segments represent
-medians and 50\% credible intervals of posteriors. The effect of generalizing
-(increasing species diversity) is similar (Fig. S\ref{fig:mirror}).}
-
-\label{fig:within}
-\end{center}
-\end{figure}
-
-\clearpage
-
-# Materials and methods
-
-We obtained commercial fisheries landings and revenue data for all permit
-holders in Alaska from `r range_yrs[1] - 1` to `r range_yrs[2]` from the Alaskan
-Commercial Fisheries Entry Commission (CFEC). These data represent a total of
-nearly `r round(ndat$initial_tot_rev/1e9)` billion 2009 USD. We adjusted for
-inflation by setting all revenues to 2009 USD [@bea2016]. We implemented
-a number of filtering steps to focus on individuals actively engaged in
-commercial fishing (Supporting Information). In particular, we removed permit
-holders whose median revenue was less than \$10,000 USD to focus on individuals
-for whom commercial fishing formed a substantial part of their income.
-
-<!-- TODO Jordan: held by active fishers were not included in permit counts -->
-<!-- / portfolios? Disregard this comment as you please but I thought I’d mention. -->
-
-We used combinations of state and federal fishing permits held by an individual
-in a specific year to define fishing "strategies", representing the species
-targeted (Supporting Information). For example, we combined permits that were
-otherwise the same except for vessel size (e.g. a longliner fishing sablefish
-on a vessel under 60 ft was considered to have the same strategy as a longliner
-fishing sablefish on a vessel over 60 ft), and we combined a number of permits
-that were targeting a single species and differed only in region fished. For
-example, someone fishing herring roe in Kodiak and someone fishing herring roe
-in Alaska Peninsula were both considered to have a "herring roe" permit group.
-For salmon permits, we formed salmon strategies based on similarities in gear
-type and species composition (Supporting Information).
-
-After this process of combining permits, we were left with `r n_grouped_permits`
-unique fishing permit groups (Table S\ref{tab:strategies}), which we aggregated
-within each person-year combination to form permit strategies. For example, if
-someone fished halibut and sablefish in the same year, their strategy for that
-year would be "halibut-sablefish". Specific details on our strategy definitions
-are available in the Supporting Information.
-
-Our models focused on species diversity as a possible predictor of revenue and
-revenue variability while controlling for effort. For consistency with previous
-analyses [@vedeld2007; @illukpitiya2008; @debela2012; @kasperski2013;
-sethi2014], we calculated effective species diversity for permit holder $i$ and
-time (year) $t$ as the inverse of Simpson's diversity index $\lambda$
-[@simpson1949; @jost2006] (also referred to in economics as the
-Herfindahl-Hirschman Index [@hirschman1964]) weighted by revenue $R$:
-
-\begin{equation}
-1/\lambda_{i,t} = 1 / \left(\sum\limits_{s=1}^{n_{i,t}}
-  \left(R_{i,s,t}/ \sum\limits_{s=1}^{n_{i,t}} R_{i,s,t}\right)\right)^2,
-\end{equation}
-
-\noindent where $s$ indexes species from $1$ through $n$. We calculated effort
-or effective season length as the sum of the season lengths for each permit
-each year. Season length for a given permit was calculated as the number of
-days between the first instance of fishing a permit and the last instance of
-fishing a permit each year. For example, if someone fished for 30 days on a
-salmon drift gillnet permit in Southeast Alaska and 40 days on a statewide
-longline halibut permit, that individual's season length was calculated as 70
-days.
-
-## Hierarchical model
-
-To jointly model revenue and revenue variability as a function of
-individual-level covariates, we extended the basic variance-function regression
-to a hierarchical Bayesian variance-function regression model with Gaussian
-errors predicting $\log$ revenue, $\log(R_{i,t})$, for each fisher $i$ and time
-$t$:
-
-\begin{equation}
-\log(R_{i,t}) \sim \mathrm{N} \left(\mu_{i,t}, \sigma_{i,t}^2 \right).
-\end{equation}
-
-\noindent Each year $t$ of revenue for fisher $i$ is assigned a strategy (permit
-set) indexed by $j$ (Supporting Information, Table \ref{tab:strategies}) and
-$\mu_{i,t}$ is modelled as:
-
-\begin{equation}
-\label{eq:mean}
-\mu_{i,t} = \beta_{0,j,t} -
-  \beta_{1,j} S_{i,t} I_{i,t} +
-  \beta_{2,j} S_{i,t} I_{i,t} +
-  \beta_3 D_{i,t} -
-  \beta_4 S_{i,t} D_{i,t} I_{i,t} +
-  \beta_5 S_{i,t} D_{i,t} I_{i,t} +
-  \log(R_{i,t-1}),
-\end{equation}
-
-\noindent where $S_{i,t}$ is the log ratio of species diversity from
-year-to-year, $\log \left(\mathrm{div}_{i,t}/\mathrm{div}_{i,t-1}\right)$,
-$D_{i,t}$ is the log ratio of days fished from year-to-year, $\log
-\left(\mathrm{days}_{i,t}/\mathrm{days}_{i,t-1}\right)$, and $I_{i,t}$ is an
-indicator variable that takes a value $0$ if $\mathrm{div}_{i,t} \ge
-\mathrm{div}_{i,t-1}$ (species diversity increasing; generalizing) and takes
-a value $1$ if $\mathrm{div}_{i,t} < \mathrm{div}_{i,t-1}$ (species diversity
-decreasing; specializing). Estimating a separate coefficient for the effect of
-increasing diversity and decreasing diversity allows the costs and benefits of
-these changes to be asymmetric (Fig. S\ref{fig:resid}).
-
-We use $\log(R_{i,t-1}$) as an offset, and thus our model describes the log
-ratio of revenue from year-to-year, i.e., $\log(R_{i,t}) - \log(R_{i,t-1})
-= \log(R_{i,t}/R_{i,t-1})$. We model change in revenue because preliminary
-analysis revealed non-stationary trends in revenue (resulting from experience or
-other unmodelled factors). This approach eliminates having to deal with the
-non-stationary time series and eliminates having to estimate thousands of
-mean-revenue intercepts associated with individual fishers. We removed
-person-year combinations without a subsequent year of revenue.
-
-The $\beta$ terms represent estimated coefficients, some of which are allowed to
-vary as "random effect" terms. The intercepts $\beta_{0,j,t}$ are allowed to
-vary by strategy-year combinations, are centred on zero, and are constrained by
-a normal distribution:
-
-\begin{equation}
-\beta_{0,j,t} \sim \mathrm{N} \left(0, \sigma^2_{\beta_{j,t}} \right).
-\end{equation}
-
-\noindent This approach, combined with the lack of a global intercept,
-constrains the mean change in log revenue across all fishers and years to be
-zero. I.e., we assume that the data are stationary after accounting for all
-covariates and including an intercept that varies with strategy-year
-combinations. Initial versions of the model confirmed that the global intercept
-along with strategy-level intercepts would be estimated at almost exactly zero.
-The estimated $\beta_{j,t}$'s represent years when people with a particular set
-of permits tended to collectively do better or worse than the long-term average
-(Fig. S\ref{fig:str-yrs}). Therefore, an individual's expected change in revenue
-(and variability) represents the expectation after accounting for the general
-trend in revenue for that set of permits.
-
-The slopes $\beta_{1,j}$ and $\beta_{1,j}$ are then allowed to vary by strategy
-with estimated means:
-
-\begin{align}
-\beta_{1,j} &\sim \mathrm{N} \left(\mu_{\beta_1}, \tau^2_{\beta_1}\right),\\
-\beta_{2,j} &\sim \mathrm{N} \left(\mu_{\beta_2}, \tau^2_{\beta_2}\right).
-\end{align}
-
-\noindent The $\beta_{1,j}$'s and $\beta_{2,j}$'s represent how much a 1\%
-decrease and 1\% increase, respectively, in the ratio of species diversity from
-year-to-year within a strategy translates to a given percent change in the ratio
-of revenue from year-to-year. These are after controlling for changes in effort,
-$\beta_3$, and the interaction between changes in effort and changes in species
-diversity, $\beta_4$ and $\beta_5$.
-
-Whereas traditional hierarchical linear models assume homoscedasticity --- that
-the residual variance does not vary systematically --- our variance-function
-regression model allows the residual variance to be heteroscedastic and vary as
-a function of predictors. Instead of estimating a single residual error scale
-term, $\sigma$, we model the scale of the residual error with another
-hierarchical model with a similar form:
-
-\begin{equation}
-\label{eq:cv}
-\sigma_{i,t} =
-\exp \left( \gamma_{0,j} -
-\gamma_{1,j} S_{i,t} I_{i,t} +
-\gamma_{2,j} S_{i,t} I_{i,t} +
-\gamma_3 D_{i,t} -
-\gamma_4 S_{i,t} D_{i,t} I_{i,t} +
-\gamma_5 S_{i,t} D_{i,t} I_{i,t}\right).
-\end{equation}
-
-\noindent Again, $S_{i,t}$ and $I_{i,t}$ represent, respectively, the log ratio
-of species diversity from year-to-year and an indicator variable for increasing
-or decreasing species diversity. We exponentiate the equation to ensure all
-scale parameters, $\sigma_{i,t}$, are positive. Residuals from the mean
-component of our model clearly show a "V"-shaped pattern, which the variance
-component of our model can capture (Fig. S\ref{fig:resid}). Furthermore, the
-downside residuals (year-individual data points with lower than expected
-revenue) look qualitatively similar to the complete set of residuals (upside
-and downside) (Fig. S\ref{fig:resid-downside}). Therefore, it seemed
-appropriate to model the variability as a proxy for revenue risk.
-
-The $\gamma$'s are estimated coefficients, three of which are allowed to vary by
-strategy, constrained by normal distributions:
-
-\begin{align}
-\gamma_{0,j}  &\sim \mathrm{N} \left(\eta_0 + \eta_1 M_j + \eta_2 E_j, \tau^2_{\gamma_0}\right),\\
-\gamma_{1,j}  &\sim \mathrm{N} \left(\mu_{\gamma_1}, \tau^2_{\gamma_1}\right),\\
-\gamma_{2,j}  &\sim \mathrm{N} \left(\mu_{\gamma_2}, \tau^2_{\gamma_2}\right).
-\end{align}
-
-\noindent The $\gamma_{0,j}$ coefficients are modelled with two strategy-level
-predictors. The symbols $M_j$ and $E_j$ represent the mean species diversity
-and mean combined fishing season length (effort) for strategy $j$. The
-coefficients $\eta_1$ and $\eta_2$ represent predictors in this strategy-level
-regression estimating effects across strategies.
-
-<!-- TODO: rerun with 2.14.1 -->
-
-We fit our models with Stan 2.14.1 [@standevelopmentteam2016;
-@standevelopmentteam2016a] and R 3.3.2 [@rcoreteam2016]. Stan implements the
-No-U-Turn Hamiltonian Markov chain Monte Carlo algorithm [@hoffman2014] to
-perform Bayesian statistical inference. We assigned weakly informative priors
-on all parameters: $N(0, 2^2)$ priors on all $\beta$, $\gamma$, and $\eta$
-parameters, and half-$t(3, 0, 2)$ priors (i.e. degrees of freedom of 3 and
-scale of 2) on all $\sigma$ and $\tau$. We ran four chains and 2,000 iterations
-and discarded the first 1,000 iterations of each chain as warm up. We checked
-for chain convergence visually with trace plots, ensured that $\hat{R}<1.05$
-(the potential scale reduction factor) and that the effective sample size, as
-calculated by the rstan R package [@standevelopmentteam2016], was greater than
-200 for all parameters [@gelman2014]. We verified that our model returned
-sensible estimates by comparing our estimates to an ad-hoc two-stage model fit
-with the lme4 R package [@bates2015], where we first fit a mixed-effects
-regression model to the mean, and then modelled the absolute residuals from
-that first model in a second mixed-effects model representing the variance
-component. We also validated the model with simulated data, and were able to
-recover the true values (Supporting Information). Code used in our analysis is
-available at <https://github.com/NCEAS/pfx-commercial/portfolios/>.
-
-# Acknowledgments
-
-TODO. National Center for Ecological Analysis and Synthesis (NCEAS) working
+\acknow{TODO. National Center for Ecological Analysis and Synthesis (NCEAS) working
 group *Applying portfolio effects to the Gulf of Alaska ecosystem: Did
@@
 participants... Smith...
+}
 

### 2017-01-14 portfolio/text/anderson-etal-rev-pfx.tex
Edit methods

@@
 \matmethods{
-We obtained commercial fisheries landings and revenue data for all permit holders in Alaska from 1985 to 2014 from the Alaskan Commercial Fisheries Entry Commission (CFEC).
-These data represent a total of nearly 43 billion 2009 USD.
+
+We obtained commercial fisheries landings and revenue data for all permit
+holders in Alaska from 1985 to 2014 from the Alaskan Commercial Fisheries Entry
+Commission (CFEC). These data represent a total of nearly 43 billion 2009 USD.
 We adjusted for inflation by setting all revenues to 2009 USD \cite{bea2016}.
-We implemented a number of filtering steps to focus on individuals actively engaged in commercial fishing (Supporting Information).
-In particular, we removed permit holders whose median revenue was less than \$10,000 USD to focus on individuals for whom commercial fishing formed a substantial part of their income.
+We implemented a number of filtering steps to focus on individuals actively
+engaged in commercial fishing (Supporting Information). In particular, we
+removed permit holders whose median revenue was less than \$10,000 USD to focus
+on individuals for whom commercial fishing formed a substantial part of their
+income.
 
-We used combinations of state and federal fishing permits held by an individual in a specific year to define fishing ``strategies'', representing the species targeted (Supporting Information).
-For example, we combined permits that were otherwise the same except for vessel size (e.g.~a longliner fishing sablefish on a vessel under 60 ft was considered to have the same strategy as a longliner fishing sablefish on a vessel over 60 ft), and we combined a number of permits that were targeting a single species and differed only in region fished.
-For example, someone fishing herring roe in Kodiak and someone fishing herring roe in Alaska Peninsula were both considered to have a ``herring roe'' permit group.
-For salmon permits, we formed salmon strategies based on similarities in gear type and species composition (Supporting Information).
+We used combinations of state and federal fishing permits held by an individual
+in a specific year to define fishing ``strategies'', representing the species
+targeted (Supporting Information). For example, we combined permits that were
+otherwise the same except for vessel size (e.g.~a longliner fishing sablefish
+on a vessel under 60 ft was considered to have the same strategy as a longliner
+fishing sablefish on a vessel over 60 ft), and we combined a number of permits
+that were targeting a single species and differed only in region fished. For
+example, someone fishing herring roe in Kodiak and someone fishing herring roe
+in Alaska Peninsula were both considered to have a ``herring roe'' permit
+group. For salmon permits, we formed salmon strategies based on similarities in
+gear type and species composition (Supporting Information).
 
-After this process of combining permits, we were left with 23 unique fishing permit groups (Table~S\tabstrategies), which we aggregated within each person-year combination to form permit strategies.
-For example, if someone fished halibut and sablefish in the same year, their strategy for that year would be ``halibut-sablefish''.
-Specific details on our strategy definitions are available in the Supporting Information.
+After this process of combining permits, we were left with 23 unique fishing
+permit groups (Table~S\tabstrategies), which we aggregated within each
+person-year combination to form permit strategies. For example, if someone
+fished halibut and sablefish in the same year, their strategy for that year
+would be ``halibut-sablefish''. Specific details on our strategy definitions
+are available in the Supporting Information.
 
-Our models focused on species diversity as a possible predictor of revenue and revenue variability while controlling for effort.
-For consistency with previous analyses {[}\cite{vedeld2007}; \cite{illukpitiya2008}; \cite{debela2012}; \cite{kasperski2013}; sethi2014{]}, we calculated effective species diversity for permit holder \(i\) and time (year) \(t\) as the inverse of Simpson's diversity index \(\lambda\) \cite{simpson1949, jost2006} (also referred to in economics as the Herfindahl-Hirschman Index \cite{hirschman1964}) weighted by revenue \(R\):
+Our models focused on species diversity as a possible predictor of revenue and
+revenue variability while controlling for effort. For consistency with previous
+analyses {[}\cite{vedeld2007}; \cite{illukpitiya2008}; \cite{debela2012};
+\cite{kasperski2013}; sethi2014{]}, we calculated effective species diversity
+for permit holder \(i\) and time (year) \(t\) as the inverse of Simpson's
+diversity index \(\lambda\) \cite{simpson1949, jost2006} (also referred to in
+economics as the Herfindahl-Hirschman Index \cite{hirschman1964}) weighted by
+revenue \(R\):
 
@@
 
-\noindent where \(s\) indexes species from \(1\) through \(n\).
-We calculated effort or effective season length as the sum of the season lengths for each permit each year.
-Season length for a given permit was calculated as the number of days between the first instance of fishing a permit and the last instance of fishing a permit each year.
-For example, if someone fished for 30 days on a salmon drift gillnet permit in Southeast Alaska and 40 days on a statewide longline halibut permit, that individual's season length was calculated as 70 days.
+\noindent where \(s\) indexes species from \(1\) through \(n\). We calculated
+effort or effective season length as the sum of the season lengths for each
+permit each year. Season length for a given permit was calculated as the number
+of days between the first instance of fishing a permit and the last instance of
+fishing a permit each year. For example, if someone fished for 30 days on a
+salmon drift gillnet permit in Southeast Alaska and 40 days on a statewide
+longline halibut permit, that individual's season length was calculated as 70
+days.
 
@@
 
-To jointly model revenue and revenue variability as a function of individual-level covariates, we extended the basic variance-function regression to a hierarchical Bayesian variance-function regression model with Gaussian errors predicting \(\log\) revenue, \(\log(R_{i,t})\), for each fisher \(i\) and time \(t\):
+To jointly model revenue and revenue variability as a function of
+individual-level covariates, we extended the basic variance-function regression
+to a hierarchical Bayesian variance-function regression model with Gaussian
+errors predicting \(\log\) revenue, \(\log(R_{i,t})\), for each fisher \(i\)
+and time \(t\):
 
@@
 
-\begin{equation}
+\begin{equation*}
+  \begin{split}
 \label{eq:mean}
@@
   \beta_{2,j} S_{i,t} I_{i,t} +
-  \beta_3 D_{i,t} -
+  \beta_3 D_{i,t} -\\
   \beta_4 S_{i,t} D_{i,t} I_{i,t} +
@@
   \log(R_{i,t-1}),
-\end{equation}
+  \end{split}
+\end{equation*}
 
-\noindent where \(S_{i,t}\) is the log ratio of species diversity from year-to-year, \(\log \left(\mathrm{div}_{i,t}/\mathrm{div}_{i,t-1}\right)\), \(D_{i,t}\) is the log ratio of days fished from year-to-year, \(\log \left(\mathrm{days}_{i,t}/\mathrm{days}_{i,t-1}\right)\), and \(I_{i,t}\) is an indicator variable that takes a value \(0\) if \(\mathrm{div}_{i,t} \ge \mathrm{div}_{i,t-1}\) (species diversity increasing; generalizing) and takes a value \(1\) if \(\mathrm{div}_{i,t} < \mathrm{div}_{i,t-1}\) (species diversity decreasing; specializing).
-Estimating a separate coefficient for the effect of increasing diversity and decreasing diversity allows the costs and benefits of these changes to be asymmetric (Fig.~S\figresid).
+\noindent where \(S_{i,t}\) is the log ratio of species diversity from
+year-to-year, \(\log \left(\mathrm{div}_{i,t}/\mathrm{div}_{i,t-1}\right)\),
+\(D_{i,t}\) is the log ratio of days fished from year-to-year, \(\log
+  \left(\mathrm{days}_{i,t}/\mathrm{days}_{i,t-1}\right)\), and \(I_{i,t}\) is
+an indicator variable that takes a value \(0\) if \(\mathrm{div}_{i,t} \ge
+  \mathrm{div}_{i,t-1}\) (species diversity increasing; generalizing) and takes
+a value \(1\) if \(\mathrm{div}_{i,t} < \mathrm{div}_{i,t-1}\) (species
+diversity decreasing; specializing). Estimating a separate coefficient for the
+effect of increasing diversity and decreasing diversity allows the costs and
+benefits of these changes to be asymmetric (Fig.~S\figresid).
 
-We use \(\log(R_{i,t-1}\)) as an offset, and thus our model describes the log ratio of revenue from year-to-year, i.e., \(\log(R_{i,t}) - \log(R_{i,t-1}) = \log(R_{i,t}/R_{i,t-1})\).
-We model change in revenue because preliminary analysis revealed non-stationary trends in revenue (resulting from experience or other unmodelled factors).
-This approach eliminates having to deal with the non-stationary time series and eliminates having to estimate thousands of mean-revenue intercepts associated with individual fishers.
-We removed person-year combinations without a subsequent year of revenue.
+We use \(\log(R_{i,t-1}\)) as an offset, and thus our model describes the log
+ratio of revenue from year-to-year, i.e., \(\log(R_{i,t}) - \log(R_{i,t-1}) =
+  \log(R_{i,t}/R_{i,t-1})\). We model change in revenue because preliminary
+analysis revealed non-stationary trends in revenue (resulting from experience
+or other unmodelled factors). This approach eliminates having to deal with the
+non-stationary time series and eliminates having to estimate thousands of
+mean-revenue intercepts associated with individual fishers. We removed
+person-year combinations without a subsequent year of revenue.
 
-The \(\beta\) terms represent estimated coefficients, some of which are allowed to vary as ``random effect'' terms.
-The intercepts \(\beta_{0,j,t}\) are allowed to vary by strategy-year combinations, are centred on zero, and are constrained by a normal distribution:
+The \(\beta\) terms represent estimated coefficients, some of which are allowed
+to vary as ``random effect'' terms. The intercepts \(\beta_{0,j,t}\) are
+allowed to vary by strategy-year combinations, are centred on zero, and are
+constrained by a normal distribution:
 
@@
 
-\noindent This approach, combined with the lack of a global intercept, constrains the mean change in log revenue across all fishers and years to be zero.
-I.e., we assume that the data are stationary after accounting for all covariates and including an intercept that varies with strategy-year combinations.
-Initial versions of the model confirmed that the global intercept along with strategy-level intercepts would be estimated at almost exactly zero.
-The estimated \(\beta_{j,t}\)'s represent years when people with a particular set of permits tended to collectively do better or worse than the long-term average (Fig.~S\figstryrs).
-Therefore, an individual's expected change in revenue (and variability) represents the expectation after accounting for the general trend in revenue for that set of permits.
+\noindent This approach, combined with the lack of a global intercept,
+constrains the mean change in log revenue across all fishers and years to be
+zero. I.e., we assume that the data are stationary after accounting for all
+covariates and including an intercept that varies with strategy-year
+combinations. Initial versions of the model confirmed that the global intercept
+along with strategy-level intercepts would be estimated at almost exactly zero.
+The estimated \(\beta_{j,t}\)'s represent years when people with a particular
+set of permits tended to collectively do better or worse than the long-term
+average (Fig.~S\figstryrs). Therefore, an individual's expected change in
+revenue (and variability) represents the expectation after accounting for the
+general trend in revenue for that set of permits.
 
-The slopes \(\beta_{1,j}\) and \(\beta_{1,j}\) are then allowed to vary by strategy with estimated means:
+The slopes \(\beta_{1,j}\) and \(\beta_{1,j}\) are then allowed to vary by
+strategy with estimated means:
 
@@
 
-\noindent The \(\beta_{1,j}\)'s and \(\beta_{2,j}\)'s represent how much a 1\% decrease and 1\% increase, respectively, in the ratio of species diversity from year-to-year within a strategy translates to a given percent change in the ratio of revenue from year-to-year.
-These are after controlling for changes in effort, \(\beta_3\), and the interaction between changes in effort and changes in species diversity, \(\beta_4\) and \(\beta_5\).
+\noindent The \(\beta_{1,j}\)'s and \(\beta_{2,j}\)'s represent how much a 1\%
+decrease and 1\% increase, respectively, in the ratio of species diversity from
+year-to-year within a strategy translates to a given percent change in the
+ratio of revenue from year-to-year. These are after controlling for changes in
+effort, \(\beta_3\), and the interaction between changes in effort and changes
+in species diversity, \(\beta_4\) and \(\beta_5\).
 
-Whereas traditional hierarchical linear models assume homoscedasticity --- that the residual variance does not vary systematically --- our variance-function regression model allows the residual variance to be heteroscedastic and vary as a function of predictors.
-Instead of estimating a single residual error scale term, \(\sigma\), we model the scale of the residual error with another hierarchical model with a similar form:
+Whereas traditional hierarchical linear models assume homoscedasticity --- that
+the residual variance does not vary systematically --- our variance-function
+regression model allows the residual variance to be heteroscedastic and vary as
+a function of predictors. Instead of estimating a single residual error scale
+term, \(\sigma\), we model the scale of the residual error with another
+hierarchical model with a similar form:
 
-\begin{equation}
+\begin{equation*}
+\begin{split}
 \label{eq:cv}
 \sigma_{i,t} =
-\exp \left( \gamma_{0,j} -
+\exp (\gamma_{0,j} -
 \gamma_{1,j} S_{i,t} I_{i,t} +
 \gamma_{2,j} S_{i,t} I_{i,t} +
-\gamma_3 D_{i,t} -
+\gamma_3 D_{i,t} -\\
 \gamma_4 S_{i,t} D_{i,t} I_{i,t} +
-\gamma_5 S_{i,t} D_{i,t} I_{i,t}\right).
-\end{equation}
+\gamma_5 S_{i,t} D_{i,t} I_{i,t}).
+\end{split}
+\end{equation*}
 
-\noindent Again, \(S_{i,t}\) and \(I_{i,t}\) represent, respectively, the log ratio of species diversity from year-to-year and an indicator variable for increasing or decreasing species diversity.
-We exponentiate the equation to ensure all scale parameters, \(\sigma_{i,t}\), are positive.
-Residuals from the mean component of our model clearly show a ``V''-shaped pattern, which the variance component of our model can capture (Fig.~S\figresid).
-Furthermore, the downside residuals (year-individual data points with lower than expected revenue) look qualitatively similar to the complete set of residuals (upside and downside) (Fig.~S\figresiddownside).
-Therefore, it seemed appropriate to model the variability as a proxy for revenue risk.
+\noindent Again, \(S_{i,t}\) and \(I_{i,t}\) represent, respectively, the log
+ratio of species diversity from year-to-year and an indicator variable for
+increasing or decreasing species diversity. We exponentiate the equation to
+ensure all scale parameters, \(\sigma_{i,t}\), are positive. Residuals from the
+mean component of our model clearly show a ``V''-shaped pattern, which the
+variance component of our model can capture (Fig.~S\figresid). Furthermore, the
+downside residuals (year-individual data points with lower than expected
+revenue) look qualitatively similar to the complete set of residuals (upside
+and downside) (Fig.~S\figresiddownside). Therefore, it seemed appropriate to
+model the variability as a proxy for revenue risk.
 
-The \(\gamma\)'s are estimated coefficients, three of which are allowed to vary by strategy, constrained by normal distributions:
+The \(\gamma\)'s are estimated coefficients, three of which are allowed to vary
+by strategy, constrained by normal distributions:
 
@@
 
-\noindent The \(\gamma_{0,j}\) coefficients are modelled with two strategy-level predictors.
-The symbols \(M_j\) and \(E_j\) represent the mean species diversity and mean combined fishing season length (effort) for strategy \(j\).
-The coefficients \(\eta_1\) and \(\eta_2\) represent predictors in this strategy-level regression estimating effects across strategies.
+\noindent The \(\gamma_{0,j}\) coefficients are modelled with two
+strategy-level predictors. The symbols \(M_j\) and \(E_j\) represent the mean
+species diversity and mean combined fishing season length (effort) for strategy
+\(j\). The coefficients \(\eta_1\) and \(\eta_2\) represent predictors in this
+strategy-level regression estimating effects across strategies.
 
-We fit our models with Stan 2.14.1 \cite{standevelopmentteam2016, standevelopmentteam2016a} and R 3.3.2 \cite{rcoreteam2016}.
-Stan implements the No-U-Turn Hamiltonian Markov chain Monte Carlo algorithm \cite{hoffman2014} to perform Bayesian statistical inference.
-We assigned weakly informative priors on all parameters: \(N(0, 2^2)\) priors on all \(\beta\), \(\gamma\), and \(\eta\) parameters, and half-\(t(3, 0, 2)\) priors (i.e.~degrees of freedom of 3 and scale of 2) on all \(\sigma\) and \(\tau\).
-We ran four chains and 2,000 iterations and discarded the first 1,000 iterations of each chain as warm up.
-We checked for chain convergence visually with trace plots, ensured that \(\hat{R}<1.05\) (the potential scale reduction factor) and that the effective sample size, as calculated by the rstan R package \cite{standevelopmentteam2016}, was greater than 200 for all parameters \cite{gelman2014}.
-We verified that our model returned sensible estimates by comparing our estimates to an ad-hoc two-stage model fit with the lme4 R package \cite{bates2015}, where we first fit a mixed-effects regression model to the mean, and then modelled the absolute residuals from that first model in a second mixed-effects model representing the variance component.
-We also validated the model with simulated data, and were able to recover the true values (Supporting Information).
-Code used in our analysis is available at \url{https://github.com/NCEAS/pfx-commercial/portfolios/}.
-}
-\showmatmethods
+We fit our models with Stan 2.14.1 \cite{standevelopmentteam2016,
+  standevelopmentteam2016a} and R 3.3.2 \cite{rcoreteam2016}. Stan implements
+the No-U-Turn Hamiltonian Markov chain Monte Carlo algorithm \cite{hoffman2014}
+to perform Bayesian statistical inference. We assigned weakly informative
+priors on all parameters: \(N(0, 2^2)\) priors on all \(\beta\), \(\gamma\),
+and \(\eta\) parameters, and half-\(t(3, 0, 2)\) priors (i.e.~degrees of
+freedom of 3 and scale of 2) on all \(\sigma\) and \(\tau\). We ran four chains
+and 2,000 iterations and discarded the first 1,000 iterations of each chain as
+warm up. We checked for chain convergence visually with trace plots, ensured
+that \(\hat{R}<1.05\) (the potential scale reduction factor) and that the
+effective sample size, as calculated by the rstan R package
+\cite{standevelopmentteam2016}, was greater than 200 for all parameters
+\cite{gelman2014}. We verified that our model returned sensible estimates by
+comparing our estimates to an ad-hoc two-stage model fit with the lme4 R
+package \cite{bates2015}, where we first fit a mixed-effects regression model
+to the mean, and then modelled the absolute residuals from that first model in
+a second mixed-effects model representing the variance component. We also
+validated the model with simulated data, and were able to recover the true
+values (Supporting Information). Code used in our analysis is available at
+\url{https://github.com/NCEAS/pfx-commercial/portfolios/}. } \showmatmethods
 

### 2017-01-24 portfolio/text/anderson-etal-rev-pfx.tex
Incorporate edits

@@
 \authorcontributions{
-S.C.A., TODO designed the study; S.C.A., TODO analyzed the
-data and wrote the paper with input from all authors.}
+S.C.A, E.J.W, and A.O.S designed the study with contributions from all authors;
+J.C.S. also provided data;
+S.C.A. and E.J.W. analyzed the data and wrote the paper with input from all
+authors.}
 \authordeclaration{The authors declare no conflict of interest.}
@@
 management, and
-economic dynamics. A key component to managing these industries is identifying
-how regulatory actions and individual behaviour affect income variability and by
-extension economic stability and the sustainable use of natural resources. For
+economic factors. Key to managing these industries is identifying
+how regulatory actions and individual behaviour affect income variability and, by
+extension, economic stability and the sustainable use of natural resources. In
 commercial fisheries, communities and vessels fishing a greater diversity of
-species have less revenue variability than those fishing a lower diversity.
+species have less revenue variability than those fishing fewer species.
 However, it is unclear if these benefits extend to the actions of individual
 fishers and how year-to-year changes in diversification affect revenue and
-revenue variability. Here, we identify three axes by which fishers in Alaska,
-one of the world's most productive fishing regions, can diversify fishing
+revenue variability. Here, we evaluate two axes by which fishers in Alaska,
+can diversify fishing
 activities. We show that, despite increasing specialization over the last 30
-years, fishing permits with higher species diversity reduces individual revenue
+years, fishing a set of permits with higher species diversity reduces individual revenue
 variability and fishing an additional permit is associated with higher revenue
@@
 participate in additional or more diverse fisheries. However, this option is
-expensive, often limited by regulations such as catch share programs, and
+expensive, often limited by regulations such as license limitation or catch share programs, and
 consequently unavailable to many individuals. With increasing environmental

### 2017-01-24 portfolio/text/ms.Rmd
Incorporate edits

@@
 \dropcap{M}anaging natural resources and sustaining the livelihoods of those that depend
-on their extraction is hard. These livelihoods tend to have high annual
+on their extraction is challenged by multiple sources of variability in the coupled human-natural system. These livelihoods tend to have high annual
 variability in income relative to other professions [@mishra2002; @sethi2012b].
@@
 farms to specialize (e.g. dairy, grains, soybeans; @mishra2002). In other
-instances, external market forces (e.g. prices or demand) may discourage the
+instances, market forces (e.g. prices or demand) may discourage the
 adoption of diverse strategies and help promote specialization [@bradshaw2004].
 
-Commercial fishers present an interesting example of diversification, given the
-extreme risk that some individuals experience, particularly relative to
+Diversification may benefit commercial fishers, given the
+extreme economic risk that some individuals experience, particularly relative to
 terrestrial agriculture [@sethi2012b]. The majority of risk insurance programs
@@
 ability to lower risk by diversifying their fishing portfolio. They can, for
-example, target as many species as permitted by a license, own multiple
-licenses, or diversify spatially and fish in multiple regions. The effectiveness
+example, target as many species as allowed by a permit, own multiple
+permits, or diversify spatially and fish in multiple regions. The effectiveness
 of these diversification strategies for increasing revenue and reducing revenue
-variability remains uncertain. For example, theory and empirical studies suggest
+variability remains uncertain. For example, theory and empirical studies demonstrate that
 there can be substantial benefits to specialization for both farmers and fishers
@@
 examining the effects of diversification on revenue do so at aggregate levels
-that combine individuals (e.g. counties or communities; [@kasperski2013;
-@sethi2014]). Thus, key uncertainties remain about the fundamental relationship
-between diversity, fishing revenue, and fishing revenue variability at the level
+that combine individuals (e.g., counties or communities; [@kasperski2013;
+@sethi2014; @cline2017]). Thus, key uncertainties remain about the relationship
+among diversity, fishing revenue, and fishing revenue variability at the level
 of the individual.
@@
 annually [@nmfs2016]. The days of open, unregulated fisheries in most developed
-countries have passed --- fishers must own a license to participate in virtually
+countries have passed --- fishers must own a permit to participate in virtually
 all fisheries [@hilborn2003a]. While such regulations have helped reduce
@@
 
+<!-- TODO AH: I think our fisher apporach is good, but I wonder if it’s worth a footnote to address the potential critique that at the indivdual level we don’t observe their wages in non-fishing activities.
+I’m not sure if this is worth bringing up, but this is always the case anyway with vessels where the skipper may do other things.
+-->
+
 Using a unique dataset collected over 30 years on individual fishers, rather
 than fleets or vessels, we examine how revenue and revenue variability of
-individuals is affected by diversification. By tracking individuals we can track
-people that switch fishery permits over time or use multiple vessels. Similarly,
-we can disentangle fishers that share the same vessel. We develop and apply
+individuals is affected by diversification. By tracking individuals, it is possible to track
+people that switch fishery permits over time or use multiple vessels and differentiate fishers that share the same vessel. We developped and applied
 a hierarchical Bayesian model to understand the effect of species diversity
@@
 model builds on results from vessel [@kasperski2013], fleet [@sethi2012b], and
-community-level [@sethi2014] variability. We develop a variance-function
+community-level [@sethi2014; @cline2017] variability. We developped a variance-function
 regression model [@davidian1987; @hall1989], to simultaneously model changes in
@@
 strategies while controlling for fishing effort and latent time effects (e.g.
-changing environmental conditions). To our knowledge, this is the first
+changing environmental conditions).
+The model was used to evaluate two axes by which fishers in Alaska can diversify fishing activities: (1) choice of permit strategy (i.e., permit or set of permits fished), and (2) choice of what species to fish *within* a permit strategy.
+To our knowledge, this is the first
 application of this type of model in any natural resource field.
@@
 # Results
+
+<!-- TODO AH: do we need to make a comment on the early 80s data since  we mention 1980 here but say that we starting in 1985?
+If I were reading,  I would wonder if you have the data, why don’t you use it from the start?
+on page 15 you say 1986 -->
 
@@
 approximately half the overall revenue (Fig. \ref{fig:ts}B). The mean species
-diversity of this revenue declined somewhat from 1985--2000 but has since
-remained steady (Fig. \ref{fig:ts}D). However, this trend appears largely driven
-by an increase in salmon specialists in the 1980s and 1990s --- the trend
-disappears if salmon license holders are excluded (Fig. \ref{fig:ts}D).
+diversity of the catch that generated this revenue declined somewhat from 1985--2000 but has since
+remained steady (Fig. \ref{fig:ts}D). This trend appears largely driven
+by an increase in salmon specialists in the 1980s and 1990s, disappearing
+if salmon license holders are excluded (Fig. \ref{fig:ts}D).
 
@@
 
+<!-- TODO AB: To me it is more intuitive that the darker, hotter colors would represent higher values.
+Consdier reversing the color scale to go from red (highest # of permit holders) to yellow (lowest #).
+-->
+
 Individual revenue variability decreases exponentially with higher mean fished
@@
 one unit higher mean species diversity had on average `r h1[[2]]`% (`r
-h1[[1]]`--`r h1[[3]]`% 95% CI) lower variability. This pattern looks
+h1[[1]]`--`r h1[[3]]`% 95% CI) lower variability. This pattern is
 qualitatively similar with or without individuals targeting salmon. As examples,
-variability was highest for the herring roe gillnet and king crab fisheries.
+revenue variability was highest for the herring roe gillnet and king crab fisheries.
 Most strategies with the lowest estimated revenue variability combined halibut
@@
 and revenue variability, the effect of changing fished species diversity for an
-individual *within* the constraints of a given set of permits is more variable
+individual *within* the constraints of a given strategy (set of permits) is more variable
 (Fig. \ref{fig:within}). After accounting for fishing effort and changes in
@@
 species diversity also leads to increased revenue variability (Fig.
-S\ref{fig:mirror}A), suggesting that the act of changing strategies is
+S\ref{fig:mirror}A), suggesting that the act of changing the number of species fished for a given set of permits is
 a contributor to increased variability of revenue.
@@
 Alaskan halibut and sablefish fisheries were converted to individual fishery
-quota (IFQ) management in 1994, which potentially confounds our estimates with
+quota (IFQ) management in 1995, which potentially confounds our estimates with
 a major regulatory change. Therefore, we refit our models allowing all
@@
 separately before and after the introduction of IFQs. The introduction of IFQs
-reduced variability in revenue by ~25% (TODO XX--XX 95% CI) for individuals with
-only halibut or sablefish permits (Fig. S\figifq A). Furthermore, for most
+was associated with a reduction in revenue variability of ~25% (TODO XX--XX 95% CI) for individuals with
+only halibut or sablefish permits (Fig. S\figifq A).
+<!-- TODO AH: We need to note that the populatoin changed -->
+Furthermore, for individuals within most
 strategies, the association between changing species diversity and increased
@@
 effect of individual-level changes in fished species diversity on revenue and
-revenue variability across three scales. We find that (1) individuals with sets
+revenue variability across two scales. We find that individuals with sets
 of permits that allow high levels of species diversity tend to have lower
-revenue variability. (2) Adding a fisheries permit usually increases revenue and
-decreases revenue variability. However, (3) within the constraints of a permit
-and accounting for effort, specializing tends to increase revenue while both
-specializing and generalizing tend to correlate with greater revenue variability
-with considerable differences amongst fisheries. Thus, the most effective
+revenue variability and adding a fisheries permit usually increases revenue and
+decreases revenue variability. However, within the constraints of a permit strategy
+and accounting for effort, specializing tends to increase revenue while changes to fished species diversity tend to correlate with greater revenue variability
+although there are considerable differences amongst fisheries. Thus, the most effective
 mechanism by which species diversity can lower revenue variability for
@@
 Our findings across fishing strategies are in line with previous research at the
-community and vessel levels [@kasperski2013; @sethi2012b; @sethi2014], but our
-findings within strategies, with respect to individual year-to-year behaviour,
+community and vessel levels [@kasperski2013; @sethi2012b; @sethi2014; @cline2017], but our
+findings within strategies, with respect to individual year-to-year behavior,
 provide new insights. Here, specializing tended to be associated with increased
@@
 corn, or soybeans were hit harder than diversified farmers in the early 1980s
-when a financial crisis hit [@ikerd1990]. Third, diversifying --- a lack of
+when a financial crisis hit [@ikerd1990].
+<!-- TODO AB: We might want to include more examples / references from fisheries. -->
+Third, diversifying --- a lack of
 specialization --- can happen as a consequence of non-ideal circumstances such
@@
 portfolio diversity following downward economic shocks [@debela2012].
+Because specialization inherently holds risk from ``putting all one's eggs in one
+basket,'' it is not suprising that a decline in the most valuable resource
+would lead to a shift towards diversification.
 
@@
 as individuals become accustomed to new locations, gear, or species.
+<!-- TODO AB: I think in this paragraph it would be good to pick an example or two and speculate on potential reasons specific to those permit strategies.
+Otherwise, statements like this one aren’t really anchored to anything concrete.
+For example, what does increased specialization look like in the Hal ll permit, which showed one of the greatest effects of specialization on revenue diversity?
+How many permits could a Hal ll permit holder potentially fish?
+This goes back to my comment re: Table 1 (see p.
+15) -->
 Additionally, changes in diversity might reflect years in which targeted species
-were less available and fishers were forced to target other species or landed
-more bycatch, adding stochasticity to their revenue. Most importantly, this
-increased variability reflects an inability for our large-scale model to predict
+were less available and fishers instead chose to target other species or landed
+more bycatch,
+<!-- TODO AB: This is the first time bycatch is explicitly mentioned in the study and it raises some questions about how target vs.
+bycatch species are defined and considered in the analysis.
+Some clarification in the methods would be helpful.
+-->
+adding stochasticity to their revenue. Most importantly, this
+increased variability reflects an inability for our model to predict
 an individual's revenue in a subsequent year. There are many reasons, often
-unique to an individual permit holder's behaviour [@hilborn1985; @hilborn1985a;
-@branch2006] and therefore not included in our large-scale model, that might
+unique to an individual permit holder's behavior [@hilborn1985; @hilborn1985a;
+@branch2006] and therefore not in our model, that might
 make a specific individual's revenue less predictable if species diversity
@@
 time or regional-variation in the consequences of diversification on revenue and
-revenue variability. We expect future research examining diversification over
-finer scales will uncover additional interesting relationships. Furthermore, in
-addressing the overall pattern, we have averaged out several major discrete
+revenue variability. Furthermore, in
+addressing the overall pattern, we have averaged out the potential signal of several major discrete
 events, such as the Exxon Valdez oil spill in 1989, that had significant
@@
 benefits of diversification in the context of rare and extreme disturbances
-which occur in natural systems worldwide [@anderson2016].
+which occur in natural systems worldwide [@anderson2017].
+<!-- TODO AH: Isn’t the variabiity you mention in this sentence part of this? -->
 
@@
 specialization. Although variability in individual revenue may not be actively
-incorporated into management decisions, analyses such as ours help to understand
+incorporated into management decisions, analyses such as ours help to explain
 the range of diversification strategies available, trade-offs inherent in such

### 2017-01-24 salmon/paper/salmon_paper_AOS.Rmd
Ole's additions to Paper.  Sorry but I forgot to make a new branch.

--- 2016-10-28 salmon/paper/salmon_paper.Rmd
@@
 output:
+  html_document: default
   pdf_document:
     fig_caption: yes
-  html_document: default
   word_document: default
@@
 library(gridExtra)
-library(glmmTMB)
+#library(glmmTMB)
 library(zoo)
@@
 Previous analyses have largely focused on patterns of diversification across ecosystems and the benefits to adopt more or less diverse fishing strategies [@kasperski_income_2013; @anderson_benefits].
-Fisheries in Alaska have been studied extensively, because they are one of the most productive seafood regions in the world (both by landed weight, and revenue [@office_of_science_and_technology_fisheries_2016], FAO cite) and they allow individuals to adopt a wide range of fishing strategies.
+Fisheries in Alaska are one of the most productive seafood regions in the world (both by landed weight, and revenue [@office_of_science_and_technology_fisheries_2016], FAO cite) and they allow individuals to adopt a wide range of fishing strategies.
 [@anderson_benefits] defined strategies as a permit or collection of permits held by single individuals.
@@
 
-Understanding the effects of species diversification on salmon fishers is particularly important, as the number of salmon specialists (defined as fishers holding a single salmon permit) has increased over the last 30 years as a fraction of the total [@anderson_benefits] (figure XX).
-For these limited entry salmon fisheries, general trends that have emerged are that more people are participating since the early 2000s (Fig.
-A1), median revenue is increasing (across individuals, Fig.
-A2), and landed species diversity declining over time (averaged across individuals, Fig.
-A6).
-Across different salmon fisheries -- other than a slight negative effect of diversification on variability (similar to [@anderson_benefits]) -- there appears to be little correlation between diversity and revenue (Fig.
+Salmon fisheries are particularly important in Alaska both in terms of participation and value.
+The number of fishers participating only in salmon fisheries has increased over the last 30 years as a fraction of the total [@anderson_benefits] (figure XX) and overall participation, total value?, and median revenue has increased since the early 2000s (Fig.
+A1 and A2).
+Yet the landed species diversity has declined over time (averaged across individuals, Fig.
+A6) and across different salmon fisheries there appears to be little correlation between diversity and revenue (Fig.
 A4).
-In terrestrial agriculture, there are numerous examples where diversification may affect the mean or CV of revenue, but not both [@purdy_financial_1997], and the same result may be true for salmon fishers.
-Further, these impacts of diversification are variable over time.
+Thus the relationship between harvest diversity and revenue is neither simple or straight-forward.
+As seen in terrestrial agriculture[@purdy_financial_1997], diversification may affect the mean or CV of revenue, but not both, or the effect of diversification for revenue may vary over time [bad last sentence].
+
+Salmon permits that allow the catch of multiple species provide an unusual opportunity for examining how the availablity of resources and fisher choice interact to affect both the total revenue and variability in revenue achieved by individual fishers.
+All individuals owning a single permit have the same opportunites to fish and to choose different fishing strategies.
+Individual fishers can therefore elect to utilize a diverse set of available species or be highly specialized on a single species.
+Thus the variability in fisher behavior among indivdiual fishers exist a range of strategies that allow us to understand the relationship between diversity of harvest and economic outcomes at the scale of the individual making decisions.
+Most previous analyses have focused on either vessel or community-level effects of diversification which are aggregate properties one step away from actual decision makers.
 
@@
 Other gears include beach seine and troll gears (though these are generally restricted to Kodiak and Southeast Alaska, respectively).
-While some of these combinations of gears and areas allow fishers to diversify, there are many examples of permits where the catch is dominated by a single species.
-Perhaps the best examples include fisheries that target sockeye salmon (e.g.
-Bristol Bay set and drift gillnet fisheries).
+Most of these gear types allow fishers the opportunity to harvest a diversity of salmon species.
+In practice, though, the catch is dominated by a single species for some permits.
+Perhaps the best examples are the Bristol Bay set and drift gillnet fisheries where all five salmon species are present but fishers target sockeye salmon almost exclusively (REF).
 For these fisheries, specialization is associated with higher revenue and reduced CV in annual revenue [@anderson_benefits].
 
-A secondary motivation for focusing on the effects of diversification for salmon fishers is to better understand the long term consequences of hatchery salmon production on the revenue and variability in revenue of individuals.
-The modern era of Alaska’s large-scale hatchery program began in the late 1970s (Stopha 2015, Rich will get a reference), with substantial increases in hatchery production during the 1980s following low returns [@mahnken_historical_1998; @leber_stock_2008-1].
+An additional, underappreciated consideration for understanding the relationship between diversification and fisheries value is how the availablity of different species in the environment affects the harvest diversity realized by individual fishers.
+For salmon one of the most imporant factors determining salmon abundance has been the expansion of artifical hatchery propagation of salmon species.
+Alaska’s large-scale hatchery program began in the late 1970s (Stopha 2015, Rich will get a reference), with substantial increases in hatchery production during the 1980s following low returns [@mahnken_historical_1998; @leber_stock_2008-1].
 Many of these hatchery programs increased production to stabilize the variability of salmon returns [@hilborn_review_2001], though these programs may also have negative impacts on wild production [@wertheimer_relationship_2005; @ruggerone_productivity_2015].
@@
 
+To understand the effects of diversification on revenue and revenue variability, we analyze single-permit salmon holders landings in Alaska between 1985 and 2015.
+We focus on XXX mixed species salmon fisheries in Alaska (ZZZ seine and YYY gillnet fisheries) and explore three aspects of the diversity-revenue relationship.
+First, we describe salmon catches - including total biomass, species composition, and diversity of catches -, fisheries participation, and revenue in Alaska over the past 30 years for each region and fishery.
+Second, we construct statistical time-series models to understand patterns of reveune and variability for each fishery.
+Specifically, we extend the methods used by [@anderson_benefits] to estimate time-varying effects of diversity on revenue and revenue variability.
+We then use estimated time-varying parameters to ask if the effects of diversity are constistent across time and similar among gear types and regions.
+Finally, we link our results to available information on economic and environmental data to uncover potential connections between the diversity-revenue relationship and longer term drivers in the region.
+
+*Ole's proposed flow of figures.(based on Eric's figure numbers circa 1//24/2017)
+  + Generally, I would add Bristol Bay sockeye at least in the time-series figures because it is such a well known and well studied fishery.
+It also would be a nice contrast to the others just to show that it has effectively no diversity at all.
+It may not warrant inclusion in the modeling or later analysis, but I think not mentioning the most valuable salmon fishery in Alaska is a bit odd.
+Also, it should be mentioned that Bristol Bay is weird and not really representative of most salmon fisheries in Alaska.
+  + A Suggest putting Appendix Figs 1,2,3 as a 3 panel main text figure- pick out the focal fisheries and plot trends in participation in one panel, trends in per-capita landed weight in a second, and trends in revenue in the third (may be best to put a smoother through all of the raw data to make it palatable to the eye).
+Point is that despite being in different regions and different gears, there are broad similarities in all fisheries: falling participation, rising catches, relatively flat revenue.
+  + B I like figure 1, but would add Bristol Bay (see above.).
+May need to switch to areas as rows, and the two panels as columns.
+Also I would add a total catch line to the top panel (maybe add a third panel?
+not sure).
+  + C Fig 2 & 3.
+Suggest making the percent composition plots one array of figures and having all 6 or 7 of the diversity plots in one figure with different colors or line-types.
+Maybe make the diversity figure big and then the composition for each area smaller and to the side?
+Main point is areas differ in composition and differ in overall diversity & trend in diversity.
+The diversity panel may be better served in the bullet point A, showing raw trends in diversity with other raw attributes.
+  + D Fig.
+4.
+Important points here seem to be about the variation across years (some gear-regions are more variable than others) and I wonder if this would be better presented as an esimated among-year variance or SD for each area and then possibly a pairwise correlations tile-plot among areas.
+Suggest putting Fig.
+4 time-series in the appendix.
+  + E Fig 5 and 6.
+Suggest collapsing multi-panel in Fig.5 to one panel and Fig.6 to one panel.
+Point of Fig.
+5 is choosing to add species is negative on revenue and increasingly negative in recent years.
+Story of Fig 6.
+is a little more confused, and no clear signal of diversity on revenue variability
+  + Use final figure or two to relate external information to changes in diversity-revenue or diversity-revenue variability relationship.
+Possibilities: total hatchery production, price of permits.
+Also could break each area into revenue-variability bi-plots and plot each fishery in that bivariate space by time block (e.g.
+5 year time blocks)
+
+Extra stuff:
+
 Tradeoffs between specialization and diversification may also be found in case studies of terrestrial agriculture.
@@
 Shifts to from diverse of high value cash-crops to specializing on high-volume low-value crops such as cereals may have analogs for salmon fishermen in low value species with rapid life histories (pink and chum salmon).
-
-The objective of our analysis is to better understand the effects of diversification on single-permit salmon holders by extending the methods used by [@anderson_benefits] to include time-varying effects.
-These time-varying effects will allow us to compare temporal trends across regions, identify periods when having diverse catch portfolio was most beneficial, and allow us to evaluate support for effects of longer term drivers in the region.
 
@@
 Also, high-value species may experience less annual fluctuation than lower value species.
+
+- For small scale farmers, economists have often wondered why farmers don't diversify more.
+Diversification can be profitable, but in the short term risky or expensive.
+Bigger than cost seems to be the risk [@karlan_agricultural_2012]
 

### 2017-01-24 portfolio/text/anderson-etal-rev-pfx.tex

@@
 removed permit holders whose median revenue was less than \$10,000 USD to focus
-on individuals for whom commercial fishing formed a substantial part of their
-income.
+on individuals with substantial income from commercial fishing.
 
@@
 in a specific year to define fishing ``strategies'', representing the species
-targeted (Supporting Information). For example, we combined permits that were
+caught (Supporting Information). For example, we combined permits that were
 otherwise the same except for vessel size (e.g.~a longliner fishing sablefish
@@
 fishing sablefish on a vessel over 60 ft), and we combined a number of permits
-that were targeting a single species and differed only in region fished. For
+that were caught a single species and differed only in region fished. For
 example, someone fishing herring roe in Kodiak and someone fishing herring roe
 in Alaska Peninsula were both considered to have a ``herring roe'' permit
-group. For salmon permits, we formed salmon strategies based on similarities in
+strategy. For salmon permits, we formed salmon strategies based on similarities in
 gear type and species composition (Supporting Information).
@@
 are available in the Supporting Information.
+% TODO AB: Since Table 1 defines all of the fishng strategies that are then used as figure labels (and not defined elsewhere in the text), I think it should go into the main body of the text rather than Suppl Info.
+This would also help clarify any confusion a reader might have about the difference between a permit strategy and an individual permit.
+I suggest 2 additional columns in Table 1: 1) potential species that could be harvested under each permit strategy, 2) number of unique permits contained within each strategy.
+This would really help to show how much flexibility someone within each permit strategy might have.
 
@@
 revenue variability while controlling for effort. For consistency with previous
-analyses {[}\cite{vedeld2007}; \cite{illukpitiya2008}; \cite{debela2012};
-\cite{kasperski2013}; sethi2014{]}, we calculated effective species diversity
+analyses \cite{vedeld2007, illukpitiya2008, debela2012, kasperski2013, sethi2014, cline2017}, we calculated effective species diversity
 for permit holder \(i\) and time (year) \(t\) as the inverse of Simpson's
 diversity index \(\lambda\) \cite{simpson1949, jost2006} (also referred to in
-economics as the Herfindahl-Hirschman Index \cite{hirschman1964}) weighted by
+economics as the Herfindahl-Hirschman Index (HHI) \cite{hirschman1964}) weighted by
 revenue \(R\):
@@
 
-\noindent where \(s\) indexes species from \(1\) through \(n\). We calculated
+\noindent where \(s\) indexes species from \(1\) through \(n\).
+
+We calculated
 effort or effective season length as the sum of the season lengths for each
@@
 of days between the first instance of fishing a permit and the last instance of
-fishing a permit each year. For example, if someone fished for 30 days on a
+fishing a permit each year. For example, if an individual's season lated for 30 days on a
 salmon drift gillnet permit in Southeast Alaska and 40 days on a statewide
@@
 days.
+% AH: I imagine at our scale this may not matter much, but did you use actualy fishing days too?
+If people are fishing another permit at the start and end of their main permit, this would show up as a huge increase in effort, even though it’s really small.
 
@@
 an indicator variable that takes a value \(0\) if \(\mathrm{div}_{i,t} \ge
-  \mathrm{div}_{i,t-1}\) (species diversity increasing; generalizing) and takes
+  \mathrm{div}_{i,t-1}\) (species diversity increasing year-to-year; generalizing) and takes
 a value \(1\) if \(\mathrm{div}_{i,t} < \mathrm{div}_{i,t-1}\) (species
-diversity decreasing; specializing). Estimating a separate coefficient for the
+diversity decreasing year-to-year; specializing). Estimating a separate coefficient for the
 effect of increasing diversity and decreasing diversity allows the costs and
@@
 non-stationary time series and eliminates having to estimate thousands of
-mean-revenue intercepts associated with individual fishers. We removed
+mean-revenue intercepts associated with individual fishers.
+% TODO AH: Should you note that you tested for it?  You say that you assume stationarity after you differnece, but why don’t you test and report it?
+We removed
 person-year combinations without a subsequent year of revenue.
+% TODO AH: what proportion of the data is this?
 
@@
 along with strategy-level intercepts would be estimated at almost exactly zero.
-The estimated \(\beta_{j,t}\)'s represent years when people with a particular
+The estimated \(\beta_{j,t}\) coefficients represent years when people with a particular
 set of permits tended to collectively do better or worse than the long-term
@@
 
-The slopes \(\beta_{1,j}\) and \(\beta_{1,j}\) are then allowed to vary by
+The slopes \(\beta_{1,j}\) and \(\beta_{2,j}\) are then allowed to vary by
 strategy with estimated means:
@@
 
-\noindent The \(\beta_{1,j}\)'s and \(\beta_{2,j}\)'s represent how much a 1\%
+\noindent The \(\beta_{1,j}\) and \(\beta_{2,j}\) coefficients represent how much a 1\%
 decrease and 1\% increase, respectively, in the ratio of species diversity from
@@
 hierarchical model with a similar form:
+
+% TODO AH: explain here somewhere why gamma 1 &4 are negative
 
@@
 downside residuals (year-individual data points with lower than expected
-revenue) look qualitatively similar to the complete set of residuals (upside
+revenue) appear qualitatively similar to the complete set of residuals (upside
 and downside) (Fig.~S\figresiddownside). Therefore, it seemed appropriate to

### 2017-01-25 portfolio/figs/strategy-table.tex
Incorporate comments

@@
-% latex table generated in R 3.3.1 by xtable 1.8-2 package
-Label & Description \\ 
+% latex table generated in R 3.3.2 by xtable 1.8-2 package
+Label & Description & Mean spp. diversity \\ 
   \midrule
-Dun C & Dungeness crab \\ 
-  Fin - ll & Finfish, longline, statewide \\ 
-  Fin - otter & Finfish, otter trawl, statewide \\ 
-  Fin - pot & Finfish, pot gear, statewide \\ 
-  Hal - ll & Halibut, longline \\ 
-  Herr - pound & Herring, pound \\ 
-  Herr roe - gill & Herring roe, gillnet \\ 
-  King C & King crab \\ 
-  S - drift - CI, AP, BB (sock) & Salmon (mostly sockeye), drift gillnet, CI, AP, BB \\ 
-  S - drift - PWS & Salmon, drift gillnet, PWS \\ 
-  S - drift - SE & Salmon, drift gillnet, Southeast Alaska \\ 
-  S - seine - AP & Salmon, purse seine, Alaska Penninsula \\ 
-  S - seine - CH & Salmon, purse seine, Chignik \\ 
-  S - seine - K & Salmon, purse seine, Cook Inlet and Kodiak \\ 
-  S - seine - PWS & Salmon, purse seine, PWS \\ 
-  S - seine - SE & Salmon, purse seine, Southeast Alaska \\ 
-  S - set - NS/K & Salmon, set gillnet, Norton Sound and Kuskokwim \\ 
-  S - set - PWS, CI, K, AP, BB (sock) & Salmon (mostly sockeye), set gillnet, PWS, CI, K,  AP, BB \\ 
-  S - set - YAK & Salmon, set gillnet, Yakutat \\ 
-  S - troll & Salmon, power troll, statewide \\ 
-  Sab & Sablefish, longline, statewide \\ 
-  Sea cuc & Sea cucumber \\ 
-  Tan C & Tanner crab, pot gear \\ 
+Dun C & Dungeness crab & 1.0 \\ 
+  King C & King crab & 1.0 \\ 
+  Herr - pound & Herring, pound & 1.0 \\ 
+  Sea cuc & Sea cucumber & 1.0 \\ 
+  Tan C & Tanner crab, pot gear & 1.0 \\ 
+  Hal - ll & Halibut, longline & 1.0 \\ 
+  Fin - pot & Finfish, pot gear, statewide & 1.0 \\ 
+  Herr roe - gill & Herring roe, gillnet & 1.0 \\ 
+  Sab & Sablefish, longline, statewide & 1.0 \\ 
+  S - drift - CI, AP, BB (sock) & Salmon (mostly sockeye), drift gillnet, CI, AP, BB & 1.1 \\ 
+  Fin - ll & Finfish, longline, statewide & 1.1 \\ 
+  S - set - PWS, CI, K, AP, BB (sock) & Salmon (mostly sockeye), set gillnet, PWS, CI, K,  AP, BB & 1.2 \\ 
+  S - seine - CH & Salmon, purse seine, Chignik & 1.3 \\ 
+  S - seine - PWS & Salmon, purse seine, PWS & 1.4 \\ 
+  Fin - otter & Finfish, otter trawl, statewide & 1.5 \\ 
+  S - set - YAK & Salmon, set gillnet, Yakutat & 1.7 \\ 
+  S - troll & Salmon, power troll, statewide & 1.8 \\ 
+  S - seine - K & Salmon, purse seine, Cook Inlet and Kodiak & 1.9 \\ 
+  S - drift - PWS & Salmon, drift gillnet, PWS & 2.0 \\ 
+  S - seine - SE & Salmon, purse seine, Southeast Alaska & 2.1 \\ 
+  S - seine - AP & Salmon, purse seine, Alaska Penninsula & 2.2 \\ 
+  S - set - NS/K & Salmon, set gillnet, Norton Sound and Kuskokwim & 2.5 \\ 
+  S - drift - SE & Salmon, drift gillnet, Southeast Alaska & 2.5 \\ 
    \bottomrule

### 2017-01-25 portfolio/text/anderson-etal-rev-pfx.tex
Incorporate comments

@@
 
-\affil[d]{Alaska Department of Fish and Game,
-Division of Commercial Fisheries,
-PO BOX 115526,
-Juneau, AK, USA 99811-5526, USA}
+\affil[d]{Division of Commercial Fisheries,
+Alaska Department of Fish and Game, PO BOX 115526, Juneau, AK 99811, USA}
 
@@
 expensive, often limited by regulations such as license limitation or catch share programs, and
-consequently unavailable to many individuals. With increasing environmental
-variability in a changing climate, it will be particularly important that
+consequently unavailable to many individuals. With increasing climatic variability,
+it will be particularly important that
 individuals relying on natural resources for their livelihood have effective
@@
 mean-revenue intercepts associated with individual fishers.
-% TODO AH: Should you note that you tested for it?  You say that you assume stationarity after you differnece, but why don’t you test and report it?
+% TODO AH: Should you note that you tested for it?  You say that you assume stationarity after you differnece, but why don’t you test and report it? reduces...?
 We removed

### 2017-01-25 portfolio/text/ms.Rmd
Remaining edits
Incorporate comments

@@
 
-\dropcap{M}anaging natural resources and sustaining the livelihoods of those that depend
-on their extraction is challenged by multiple sources of variability in the coupled human-natural system. These livelihoods tend to have high annual
+\dropcap{It} can be difficult for individuals to sustain a 
+livelihood from natural resource extraction.
+These livelihoods tend to have high annual
 variability in income relative to other professions [@mishra2002; @sethi2012b].
@@
 industry [@kirilenko2007].
+<!-- TODO Could add anthropogenic perturbations (e.g., Exxon Valdez) to this list or in the next paragraph. -->
 
@@
 Individuals who rely on natural resources for income develop strategies to
-reduce income variability. For example, farmers may rotate their crops to buffer
+reduce income variability. For example, farmers may TODO 
+to buffer
 short-term climate variation and long-term environmental trends [@zentner1988].
@@
 in terrestrial agriculture, crop subsidies in the U.S. may incentivize some
-farms to specialize (e.g. dairy, grains, soybeans; @mishra2002). In other
+farms to specialize (e.g. dairy, grains, soybeans; @mishra2002). <!-- TODO LP: ? -->
+In other
 instances, market forces (e.g. prices or demand) may discourage the
@@
 obtainable by commercial fishers [@kasperski2013]. However, fishers maintain the
-ability to lower risk by diversifying their fishing portfolio. They can, for
+ability to lower risk by diversifying their fishing portfolio.
+ <!-- TODO LP: this is the question you are trying to answer -->
+They can, for
 example, target as many species as allowed by a permit, own multiple
@@
 I’m not sure if this is worth bringing up, but this is always the case anyway with vessels where the skipper may do other things.
--->
+triue, and for others-->
 
@@
 
-\caption{Trends in revenue and diversification over time. (A) Map of fishing
-regions in Alaska, USA. (B) Total revenue in billions of USD from `r
+\caption{Trends in revenue and diversification over time. (A) Map of major fishing
+areas in Alaska, USA. (B) Total revenue in billions of USD from `r
 range_yrs[1]` to `r range_yrs[2]` for multiple (grey) and single (blue) permit
@@
 Comparing single permit holders to individuals holding that same permit plus
-additional permits illustrates the expected reduction in revenue variability
+additional permits illustrates the expected reduction in year-to-year revenue variability
 from purchasing additional permits (Fig. \ref{fig:spokes}, Fig.
@@
 (Fig. \ref{fig:within}; Figs S\figfe, S\figre, S\ref{fig:noeffort}).
+ <!-- TODO LP: effort is endogenous? -->
 On average, there is a ~0.5% increase in the year-to-year revenue ratio per 1%
@@
 \ref{fig:within} strategies above zero; Figs. S\figfe, S\figre,
-S\figmirror). This effect was strongest for individuals with a halibut
-permit or halibut and sablefish permits. Individuals with these strategies
+S\figmirror). This effect was strongest for individuals with a halibut and sablefish permit. 
+Individuals with this strategies
 experienced a ~2% increase in variability for a 1% change in the year-to-year
-ratio of species diversity (Fig. \ref{fig:within}). Becoming more generalized in
+ratio of species diversity (Fig. \ref{fig:within}).
+Becoming more generalized in
 species diversity also leads to increased revenue variability (Fig.
@@
 Ghana can increase their efficiency by devoting more land and labor to rice
-production compared to other crops [@abdulai2000]. Second, given the
+production compared to other crops [@abdulai2000]. <!-- TODO: weak -->
+Second, given the
 availability of valuable resources, specializing can be highly profitable. For
@@
 revenue variability. Furthermore, in
-addressing the overall pattern, we have averaged out the potential signal of several major discrete
-events, such as the Exxon Valdez oil spill in 1989, that had significant
-consequences for the ecological and social systems in Alaska [@paine1996]. In
+addressing the overall pattern, we have averaged out the potential 
+signal of several major discrete
+events, such as the *Exxon Valdez* oil spill in 1989, that was a significant
+disruption to the management, ecological, 
+and social structure of fisheries in Alaska [@paine1996]. In
 addition to revenue or income variability, it is important to understand the
@@
 multi-scale diversity buffer against the Exxon Valdez oil spill?*. Other
-participants... Smith...
+participants... Smith... Support of Jordan Watson came from NOAA Fisheries' Spatial Economics Toolbox for Fisheries (FishSET) Project.
 }

### 2017-01-25 portfolio/text/supp.Rnw
Incorporate comments

@@
 longline finfish,
-pot gear.
+pot gear,
 king crab,
@@
 We combined permits across each of
-dungeness crab,
+Dungeness crab,
 herring roe purse seining,
@@
 there was a wide variety in species revenue by
-area ($\sim$ 85\% pink in Prince William Sound to
-85\% sockeye on the Alaskan Peninsula).
+area ($\sim$ 85\% pink salmon in Prince William Sound to
+85\% sockeye salmon on the Alaskan Peninsula).
 Of the five permits, only two had similar species compositions,
@@
 their strategy for that year would be `halibut-sablefish'.
-We removed people who switched strategies from one year to the next
-since we were working with changes in revenue from year to year
+We removed data for pairs of years in which people switched
+strategies from one year to the next
+since we were working with changes in revenue from year-to-year
 and so needed consistent strategies across any pair of years
@@
 across all years to focus on the major strategies
-leaving us with a total of \Sexpr{nstr} strategies.
+leaving us with a total of \Sexpr{nstr} (unique or combined) strategies.
 Finally, we removed all person-year combinations without
@@
 \clearpage
-
+% TODO: AB: I suggest 2 additional columns: 1) potential species that could be harvested under each permit strategy, 2) number of unique permits contained within each strategy.
+This would really help to show how much flexibility someone within each permit strategy might have.
 \begin{table}[h!]
+\footnotesize
   \centering
-  \caption{Labels and descriptions of fishing strategies. In many cases,
+  \caption{Labels and descriptions of unique fishing strategies. In many cases,
 a particular strategy represents a group of similar permits across areas
@@
   \label{tab:strategies}
-  \begin{tabular}{ll}
+  \begin{tabular}{lll}
   \toprule
@@
 \caption{Estimates of within-strategy effects from the hierarchical model.
-Panels refer to parameters in Eqns \ref{eq:mean} and \ref{eq:cv}.
+Panels refer to parameters in equations \ref{eq:mean} and \ref{eq:cv}.
 Shown are medians (dots) and 50\% and 95\% credible intervals
@@
 
+% TODO: tweak labels to fit
 \begin{figure}[htbp]
@@
 \end{figure}
+
+% TODO: all figs referenced?
 

### 2017-01-25 portfolio/analysis/3.3-make-strategy-table.R

@@ -4,16 +4,27 @@ d <- readr::read_csv("data/strategies-labels.csv")
 load("portfolio/data-generated/diff-dat-stan.rda")
 d <- d %>% mutate(modeled = ifelse(strategy %in% md$strategy, "Yes", "No"))
 
-d <- select(d, str_label, description, modeled) %>%
+# spp <- group_by(dat, strategy) %>%
+  # summarise(
+    # spp_l = round(quantile(specDiv, prob = 0.25), 1),
+    # spp_h = round(quantile(specDiv, prob = 0.75), 1))
+
+# d <- inner_join(d, spp)
+d <- inner_join(d, select(md, strategy, strategy_mean_div))
+
+d <- select(d, str_label, description, strategy_mean_div, modeled) %>%
   rename(Label = str_label, Description = description,
-    `Over 100 permit holders` = modeled) %>%
-  arrange(Label)
+    `Over 100 permit holders` = modeled,
+    `Mean spp. diversity` = strategy_mean_div) %>%
+  arrange(`Mean spp. diversity`) %>%
+  mutate(`Mean spp. diversity` = round(`Mean spp. diversity`, 1))
+
 d <- d %>% filter(`Over 100 permit holders` == "Yes") %>%
   select(-`Over 100 permit holders`)
 
 xtable::print.xtable(
   xtable::xtable(d,
-    caption = ""),
+    caption = "", digits = 1),
   include.rownames = FALSE,
   file = "portfolio/figs/strategy-table.tex",
   booktabs = TRUE,

### 2017-01-26 portfolio/text/anderson-etal-rev-pfx.tex
Incorporating more edits

@@
 \keywords{
-TODO $|$
-TODO
+diversity-stability $|$
+Bayesian variance-function regression $|$
+income variability $|$
+natural resource management $|$
+ecological portfolio effects
 }
@@
 the following year. Our results demonstrate that the most effective option for
-individuals to decrease revenue variability is to purchase permits to
+individuals to decrease revenue variability is to
 participate in additional or more diverse fisheries. However, this option is
@@
 are available in the Supporting Information.
-% TODO AB: Since Table 1 defines all of the fishng strategies that are then used as figure labels (and not defined elsewhere in the text), I think it should go into the main body of the text rather than Suppl Info.
-This would also help clarify any confusion a reader might have about the difference between a permit strategy and an individual permit.
-I suggest 2 additional columns in Table 1: 1) potential species that could be harvested under each permit strategy, 2) number of unique permits contained within each strategy.
-This would really help to show how much flexibility someone within each permit strategy might have.
 
@@
 
-\noindent Each year \(t\) of revenue for fisher \(i\) is assigned a strategy (permit set) indexed by \(j\) (Supporting Information, Table~\tabstrategies) and \(\mu_{i,t}\) is modelled as:
+\noindent Each year \(t\) of revenue for fisher \(i\) is assigned a strategy
+(permit set) indexed by \(j\) (Supporting Information, Table~\tabstrategies)
+and \(\mu_{i,t}\) is modelled as:
 
@@
 an indicator variable that takes a value \(0\) if \(\mathrm{div}_{i,t} \ge
-  \mathrm{div}_{i,t-1}\) (species diversity increasing year-to-year; generalizing) and takes
-a value \(1\) if \(\mathrm{div}_{i,t} < \mathrm{div}_{i,t-1}\) (species
-diversity decreasing year-to-year; specializing). Estimating a separate coefficient for the
-effect of increasing diversity and decreasing diversity allows the costs and
-benefits of these changes to be asymmetric (Fig.~S\figresid).
+  \mathrm{div}_{i,t-1}\) (species diversity increasing year-to-year;
+generalizing) and takes a value \(1\) if \(\mathrm{div}_{i,t} <
+  \mathrm{div}_{i,t-1}\) (species diversity decreasing year-to-year;
+specializing). Estimating a separate coefficient for the effect of increasing
+diversity and decreasing diversity allows the costs and benefits of these
+changes to be asymmetric (Fig.~S\figresid). The negative signs in front of
+$\beta_{1,j}$ and $\beta_4$ makes these coefficients interpretable as effects of
+increasing specialization (as opposed to decreasing specialization).
 
@@
   \log(R_{i,t}/R_{i,t-1})\). We model change in revenue because preliminary
-analysis revealed non-stationary trends in revenue (resulting from experience
-or other unmodelled factors). This approach eliminates having to deal with the
-non-stationary time series and eliminates having to estimate thousands of
+analyses revealed non-stationary trends in revenue (resulting from experience
+or other unmodelled factors). This approach reduces the
+non-stationarity of the time series and allows us to avoid estimating thousands of
 mean-revenue intercepts associated with individual fishers.
-% TODO AH: Should you note that you tested for it?  You say that you assume stationarity after you differnece, but why don’t you test and report it? reduces...?
-We removed
-person-year combinations without a subsequent year of revenue.
-% TODO AH: what proportion of the data is this?
+Because we were modelling year-to-year changes,
+we removed person-year combinations without a subsequent year of revenue.
 
@@
 along with strategy-level intercepts would be estimated at almost exactly zero.
-The estimated \(\beta_{j,t}\) coefficients represent years when people with a particular
-set of permits tended to collectively do better or worse than the long-term
-average (Fig.~S\figstryrs). Therefore, an individual's expected change in
-revenue (and variability) represents the expectation after accounting for the
-general trend in revenue for that set of permits.
+The estimated \(\beta_{j,t}\) coefficients represent years when people with a
+particular set of permits tended to collectively do better or worse than the
+long-term average (Fig.~S\figstryrs). Therefore, an individual's expected
+change in revenue (and variability) represents the expectation after accounting
+for the general trend in revenue for that set of permits.
 
@@
 
-Whereas traditional hierarchical linear models assume homoscedasticity --- that
-the residual variance does not vary systematically --- our variance-function
+Whereas traditional hierarchical linear models assume homoscedasticity---that
+the residual variance does not vary systematically---our variance-function
 regression model allows the residual variance to be heteroscedastic and vary as
@@
 hierarchical model with a similar form:
-
-% TODO AH: explain here somewhere why gamma 1 &4 are negative
 
@@
 
+\acknow{This work was conducted as part of the National Center for Ecological
+  Analysis and Synthesis working group "Applying portfolio effects to the Gulf
+  of Alaska ecosystem: Did multi-scale diversity buffer against the Exxon
+  Valdez oil spill?", funded by TODO. We thank other participants of the
+  working group for helpful discussions. Additional support was provided by a
+  David H. Smith Conservation Research Fellowship (S.C.A.), the NOAA Fisheries'
+  Spatial Economics Toolbox for Fisheries (FishSET) Project (J.T.W), \ldots }
+
 \showacknow % Display the acknowledgments section

### 2017-01-26 portfolio/text/ms.Rmd
Incorporating more edits

@@
 
-# Introduction
+<!-- # Introduction -->
 
-\dropcap{It} can be difficult for individuals to sustain a 
-livelihood from natural resource extraction.
-These livelihoods tend to have high annual
+\dropcap{I}t can be difficult for individuals to sustain a livelihood from
+natural resource extraction. These livelihoods tend to have high annual 
 variability in income relative to other professions [@mishra2002; @sethi2012b]. 
@@
 industry [@kirilenko2007].
-<!-- TODO Could add anthropogenic perturbations (e.g., Exxon Valdez) to this list or in the next paragraph. -->
-
-<!-- rotate -> plant multiple? -->
+<!-- TODO Could add anthropogenic perturbations (e.g., Exxon Valdez) to this
+list or in the next paragraph. -->
 
 Individuals who rely on natural resources for income develop strategies to
-reduce income variability. For example, farmers may TODO 
-to buffer
-short-term climate variation and long-term environmental trends [@zentner1988].
-Similarly, loggers maintain a diverse mix of tree species, age structure, and
-harvest in multiple areas to buffer against disease outbreak, wildfires, and
-price variation (REF). However, otherwise well-intentioned regulations may limit
-how individuals diversify, or incentivize against diversification. For example,
-in terrestrial agriculture, crop subsidies in the U.S. may incentivize some
-farms to specialize (e.g. dairy, grains, soybeans; @mishra2002). <!-- TODO LP: ? -->
-In other
-instances, market forces (e.g. prices or demand) may discourage the
-adoption of diverse strategies and help promote specialization [@bradshaw2004].
+reduce income variability. For example, farmers may diversify their crops or
+include off-farm income sources to buffer against environmental and market
+shocks, as well as longterm climatic trends and seasonality [@ellis2000;
+@barrett2001; @ellis2004]. However, otherwise well-intentioned regulations may
+limit how individuals diversify, or incentivize against diversification. For
+instance, crop subsidies in the U.S. may incentivize some farms to specialize
+(e.g. dairy, grains, soybeans; @mishra2002). In other instances, market forces
+(e.g. prices or demand) may discourage the adoption of diverse strategies and
+help promote specialization [@bradshaw2004].
 
-Diversification may benefit commercial fishers, given the
-extreme economic risk that some individuals experience, particularly relative to
-terrestrial agriculture [@sethi2012b]. The majority of risk insurance programs
-available to terrestrial farmers (e.g. crop insurance programs) are not
-obtainable by commercial fishers [@kasperski2013]. However, fishers maintain the
-ability to lower risk by diversifying their fishing portfolio.
- <!-- TODO LP: this is the question you are trying to answer -->
-They can, for
-example, target as many species as allowed by a permit, own multiple
-permits, or diversify spatially and fish in multiple regions. The effectiveness
-of these diversification strategies for increasing revenue and reducing revenue
-variability remains uncertain. For example, theory and empirical studies demonstrate that
-there can be substantial benefits to specialization for both farmers and fishers
-[@smith1986; @abdulai2000; @arnason2005]. Furthermore, the majority of research
-examining the effects of diversification on revenue do so at aggregate levels
-that combine individuals (e.g., counties or communities; [@kasperski2013;
+Diversification may benefit commercial fishers, given the extreme economic risk
+that some individuals experience, particularly relative to terrestrial
+agriculture [@sethi2012b]. The majority of risk insurance programs available to
+terrestrial farmers (e.g. crop insurance programs) are not obtainable by
+commercial fishers [@kasperski2013]. However, fishers may be able to lower risk
+by diversifying their fishing portfolio. They can, for example, target as many
+species as allowed by a permit, own multiple permits, or diversify spatially and
+fish in multiple regions. The effectiveness of these diversification strategies
+for increasing revenue and reducing revenue variability remains uncertain. For
+example, theory and empirical studies demonstrate that there can be substantial
+benefits to specialization for both farmers and fishers [@smith1986;
+@abdulai2000; @arnason2005]. Furthermore, the majority of research examining the
+effects of diversification on revenue do so at aggregate levels that combine
+individuals (e.g., counties or communities; [@kasperski2013;
 @sethi2014; @cline2017]). Thus, key uncertainties remain about the relationship 
@@
 annually [@nmfs2016]. The days of open, unregulated fisheries in most developed
-countries have passed --- fishers must own a permit to participate in virtually
+countries have passed---fishers must own a permit to participate in virtually
 all fisheries [@hilborn2003a]. While such regulations have helped reduce
@@
 
-<!-- TODO AH: I think our fisher apporach is good, but I wonder if it’s worth a footnote to address the potential critique that at the indivdual level we don’t observe their wages in non-fishing activities.
-I’m not sure if this is worth bringing up, but this is always the case anyway with vessels where the skipper may do other things.
-triue, and for others-->
+<!-- TODO AH: I think our fisher apporach is good, but I wonder if it’s worth
+a footnote to address the potential critique that at the indivdual level we
+don’t observe their wages in non-fishing activities.  I’m not sure if this is
+worth bringing up, but this is always the case anyway with vessels where the
+skipper may do other things. triue, and for others-->
 
@@
 than fleets or vessels, we examine how revenue and revenue variability of 
-individuals is affected by diversification. By tracking individuals, it is possible to track
-people that switch fishery permits over time or use multiple vessels and differentiate fishers that share the same vessel. We developped and applied
-a hierarchical Bayesian model to understand the effect of species diversity
-within fisheries catches on individual revenue and revenue variability. Our
-model builds on results from vessel [@kasperski2013], fleet [@sethi2012b], and
-community-level [@sethi2014; @cline2017] variability. We developped a variance-function
-regression model [@davidian1987; @hall1989], to simultaneously model changes in
-the mean and variance of individual revenue as a function of covariates. We
-simultaneously estimate the effect of species diversity (the inverse of
-Simpson's diversity index [@simpson1949; @jost2006]) on individual revenue and
-revenue variability, and estimate effects both within and across fishing
-strategies while controlling for fishing effort and latent time effects (e.g.
-changing environmental conditions).
-The model was used to evaluate two axes by which fishers in Alaska can diversify fishing activities: (1) choice of permit strategy (i.e., permit or set of permits fished), and (2) choice of what species to fish *within* a permit strategy.
-To our knowledge, this is the first
-application of this type of model in any natural resource field.
+individuals is affected by diversification. By tracking individuals, it is
+possible to track people that switch fishery permits over time or use multiple
+vessels and differentiate fishers that share the same vessel. We developped and
+applied a hierarchical Bayesian model to understand the effect of species
+diversity within fisheries catches on individual revenue and revenue
+variability. Our model builds on results from vessel [@kasperski2013], fleet
+[@sethi2012b], and community-level [@sethi2014; @cline2017] variability. We
+developped a variance-function regression model [@davidian1987; @hall1989], to
+simultaneously model changes in the mean and variance of individual revenue as a
+function of covariates. We simultaneously estimate the effect of species
+diversity (the inverse of Simpson's diversity index [@simpson1949; @jost2006])
+on individual revenue and revenue variability, and estimate effects both within
+and across fishing strategies while controlling for fishing effort and latent
+time effects (e.g. changing environmental conditions). The model was used to
+evaluate two axes by which fishers in Alaska can diversify fishing activities:
+(1) choice of permit strategy (i.e., permit or set of permits fished), and (2)
+choice of what species to fish *within* a permit strategy. To our knowledge,
+this is the first application of this type of model in any natural resource
+field.
 
-# Results
-
-<!-- TODO AH: do we need to make a comment on the early 80s data since  we mention 1980 here but say that we starting in 1985?
-If I were reading,  I would wonder if you have the data, why don’t you use it from the start?
-on page 15 you say 1986 -->
+\section*{Results}
 
@@
 holders since 2000 (Fig. \ref{fig:ts}C). This overall trend is most sensitive to
-halibut or salmon permit holders --- compared to other fisheries, halibut permit
+halibut or salmon permit holders---compared to other fisheries, halibut permit
 holders are more likely to hold halibut plus other permits, whereas salmon
@@
 approximately half the overall revenue (Fig. \ref{fig:ts}B). The mean species
-diversity of the catch that generated this revenue declined somewhat from 1985--2000 but has since
-remained steady (Fig. \ref{fig:ts}D). This trend appears largely driven
-by an increase in salmon specialists in the 1980s and 1990s, disappearing
-if salmon license holders are excluded (Fig. \ref{fig:ts}D).
+diversity of the catch that generated this revenue declined somewhat from
+1985--2000 but has since remained steady (Fig. \ref{fig:ts}D). This trend
+appears largely driven by an increase in salmon specialists in the late 1980s and the
+1990s, disappearing if salmon license holders are excluded (Fig. \ref{fig:ts}D).
 
@@
 
-<!-- TODO AB: To me it is more intuitive that the darker, hotter colors would represent higher values.
-Consdier reversing the color scale to go from red (highest # of permit holders) to yellow (lowest #).
--->
+<!-- TODO AB: To me it is more intuitive that the darker, hotter colors would
+represent higher values. Consdier reversing the color scale to go from red
+(highest # of permit holders) to yellow (lowest #).  -->
 
@@
 expected variability or uncertainty in year-to-year revenue (Fig. 
-\ref{fig:within} strategies above zero; Figs. S\figfe, S\figre,
-S\figmirror). This effect was strongest for individuals with a halibut and sablefish permit. 
-Individuals with this strategies
-experienced a ~2% increase in variability for a 1% change in the year-to-year
-ratio of species diversity (Fig. \ref{fig:within}).
-Becoming more generalized in
-species diversity also leads to increased revenue variability (Fig.
-S\ref{fig:mirror}A), suggesting that the act of changing the number of species fished for a given set of permits is
-a contributor to increased variability of revenue.
+\ref{fig:within} strategies above zero; Figs. S\figfe, S\figre, S\figmirror).
+This effect was strongest for individuals with a halibut and sablefish permit. 
+Individuals with this strategies experienced a ~2% increase in variability for a
+1% change in the year-to-year ratio of species diversity (Fig.
+\ref{fig:within}). Becoming more generalized in species diversity also leads to
+increased revenue variability (Fig. S\ref{fig:mirror}A), suggesting that the act
+of changing the number of species fished for a given set of permits is a
+contributor to increased variability of revenue.
 
@@
 Alaskan halibut and sablefish fisheries were converted to individual fishery 
-quota (IFQ) management in 1995, which potentially confounds our estimates with
-a major regulatory change. Therefore, we refit our models allowing all
+quota (IFQ) management in 1995, which potentially confounds our estimates with a
+major regulatory change. Therefore, we refit our models allowing all 
 strategy-level effects involving halibut and sablefish permits to vary 
 separately before and after the introduction of IFQs. The introduction of IFQs 
-was associated with a reduction in revenue variability of ~25% (TODO XX--XX 95% CI) for individuals with
-only halibut or sablefish permits (Fig. S\figifq A).
+was associated with a reduction in revenue variability of ~25% (TODO XX--XX 95%
+CI) for individuals with only halibut or sablefish permits (Fig. S\figifq A).
 <!-- TODO AH: We need to note that the populatoin changed -->
@@
 
-# Discussion
+\section*{Discussion}
 
@@
 mechanism by which species diversity can lower revenue variability for
-individual fishers --- switching or adding fishing permits --- is expensive,
+individual fishers---switching or adding fishing permits---is expensive,
 limited by regulations, and consequently unavailable to many individuals.
@@
 <!-- TODO AB: We might want to include more examples / references from fisheries. -->
-Third, diversifying --- a lack of
-specialization --- can happen as a consequence of non-ideal circumstances such
+Third, diversifying---a lack of
+specialization---can happen as a consequence of non-ideal circumstances such
 as poor availability of a more valuable resource. For example, Ugandan
@@
 diversification can be management restrictions. In commercial fisheries,
-regulations that have been shown to reduce overfishing --- including improved
-assessments [@melnychuk2017] and individual quota systems [@costello2008] ---
-may ultimately restrict individual diversification [@smith1986; @kasperski2013]
+regulations that have been shown to reduce overfishing---including improved
+assessments [@melnychuk2017] and individual quota systems 
+[@costello2008]---may ultimately restrict individual diversification [@smith1986; @kasperski2013]
 or at least make developing a diverse fishing portfolio more expensive. However,
@@
 time or regional-variation in the consequences of diversification on revenue and
-revenue variability. Furthermore, in
-addressing the overall pattern, we have averaged out the potential 
-signal of several major discrete
-events, such as the *Exxon Valdez* oil spill in 1989, that was a significant
-disruption to the management, ecological, 
-and social structure of fisheries in Alaska [@paine1996]. In
-addition to revenue or income variability, it is important to understand the
-benefits of diversification in the context of rare and extreme disturbances
-which occur in natural systems worldwide [@anderson2017].
-<!-- TODO AH: Isn’t the variabiity you mention in this sentence part of this? -->
+revenue variability. Furthermore, in addressing the overall pattern, we have
+averaged out the potential signal of several major discrete events, such as the
+*Exxon Valdez* oil spill in 1989, that was a significant disruption to the
+management, ecological, and social structure of fisheries in Alaska
+[@paine1996]. In addition to revenue or income variability, it is important to
+understand the benefits of diversification in the context of rare and extreme
+disturbances which occur in natural systems worldwide [@anderson2017].
 
@@
 
-\acknow{TODO. National Center for Ecological Analysis and Synthesis (NCEAS) working
-group *Applying portfolio effects to the Gulf of Alaska ecosystem: Did
-multi-scale diversity buffer against the Exxon Valdez oil spill?*. Other
-participants... Smith... Support of Jordan Watson came from NOAA Fisheries' Spatial Economics Toolbox for Fisheries (FishSET) Project.
-}
-
 <!-- vim: textwidth=80

### 2017-01-26 portfolio/analysis/3.3-plot-across-strategy-effects.R

@@ -119,13 +119,13 @@ pl <-  g0 %>% mutate(str_label = ifelse(nn > 150, str_label, NA)) %>%
   alpha = 1, segment.color = "grey80",
   point.padding = unit(0.0001, "lines"), max.iter = 3e4, segment.size = 0.2) +
   geom_point(aes(bg = nn), pch = 21, size = 1.9) +
-  scale_fill_distiller(palette = "YlOrRd", trans = "log10", direction = -1,
+  scale_fill_distiller(palette = "YlOrRd", trans = "log10", direction = 1,
     breaks = c(200, 1000, 5000)) +
   coord_cartesian(ylim = c(0.2, 1.2)) +
   theme_gg() +
   labs(x = "Mean species diversity", y = "Estimated revenue variability",
     fill = "Number of\npermit holders") +
-  theme(legend.justification = c(1, 1), legend.position = c(1, 1),
+  theme(legend.justification = c(1, 1), legend.position = c(0.95, 0.95),
     legend.title = element_text(size = rel(0.75)))
 
 ggsave("portfolio/figs/stan-across-strategy-variability2.pdf", width = 5.4, height = 4.3)

### 2017-01-26 portfolio/text/ms.Rmd
More ms edits

@@
 
-<!-- TODO mention the quadrants in figure 4 -->
+<!-- TODO mention the quadrants in figure 4? -->
 
@@
 separately before and after the introduction of IFQs. The introduction of IFQs 
-was associated with a reduction in revenue variability of ~25% (TODO XX--XX 95%
-CI) for individuals with only halibut or sablefish permits (Fig. S\figifq A).
+was associated with a reduction in revenue variability of ~20--30% for
+individuals with only halibut or sablefish permits (Fig. S\figifq A).
 <!-- TODO AH: We need to note that the populatoin changed -->

### 2017-01-27 portfolio/text/anderson-etal-rev-pfx.tex
Order figs; edits

@@
 \newcommand{\tabstrategies}{1}
-\newcommand{\figfe}{1}
-\newcommand{\figre}{2}
-\newcommand{\figspokeall}{3}
-\newcommand{\figresid}{4}
-\newcommand{\figresiddownside}{5}
-\newcommand{\figstryrs}{6}
-\newcommand{\figifq}{7}
-\newcommand{\figmirror}{8}
-\newcommand{\fignoeffort}{9}
+\newcommand{\figspokeall}{1}
+\newcommand{\figfe}{2}
+\newcommand{\figre}{3}
+\newcommand{\figmirror}{4}
+\newcommand{\fignoeffort}{5}
+\newcommand{\figifq}{6}
+\newcommand{\figstryrs}{7}
+\newcommand{\figresid}{8}
+\newcommand{\figresiddownside}{9}
 
@@
 
-\begin{equation*}
+\begin{equation}
   \begin{split}
@@
   \end{split}
-\end{equation*}
+\end{equation}
 
@@
 
-\begin{equation*}
+\begin{equation}
 \begin{split}
@@
 \end{split}
-\end{equation*}
+\end{equation}
 
@@
 warm up. We checked for chain convergence visually with trace plots, ensured
-that \(\hat{R}<1.05\) (the potential scale reduction factor) and that the
+that \(\hat{R}<1.05\) (the potential scale reduction factor), and that the
 effective sample size, as calculated by the rstan R package

### 2017-01-27 portfolio/text/ms.Rmd
Order figs; edits

@@
 \end{figure}
-
 <!-- TODO AB: To me it is more intuitive that the darker, hotter colors would
@@
 Comparing single permit holders to individuals holding that same permit plus
-additional permits illustrates the expected reduction in year-to-year revenue variability
-from purchasing additional permits (Fig. \ref{fig:spokes}, Fig.
-S\figspokeall). For example, individuals fishing a halibut permit plus
-one additional permit always had lower expected revenue variability compared to
+additional permits illustrates the expected reduction in year-to-year revenue
+variability from purchasing additional permits (Fig. \ref{fig:spokes}, Fig.
+S\figspokeall). For example, individuals fishing a halibut permit plus one
+additional permit always had lower expected revenue variability compared to
 individuals fishing only a halibut permit (median reduction of `r
@@
 and revenue variability, the effect of changing fished species diversity for an
-individual *within* the constraints of a given strategy (set of permits) is more variable
-(Fig. \ref{fig:within}). After accounting for fishing effort and changes in
-annual average revenue for fishers with the same set of permits, specializing
-tends to be associated with increased revenue but also increased variability
-(Fig. \ref{fig:within}; Figs S\figfe, S\figre, S\ref{fig:noeffort}).
- <!-- TODO LP: effort is endogenous? -->
-On average, there is a ~0.5% increase in the year-to-year revenue ratio per 1%
-decrease in species diversity, and across strategies weighted equally, `r
-g2_dens`% of the probability density supports a positive relationship between
-specialization and revenue.
+individual *within* the constraints of a given strategy (set of permits) is more
+variable (Fig. \ref{fig:within}). After accounting for fishing effort and
+changes in annual average revenue for fishers with the same set of permits,
+specializing tends to be associated with increased revenue but also increased
+variability (Fig. \ref{fig:within}; Figs S\figfe, S\figre, S\figmirror). <!--
+TODO LP: effort is endogenous? --> On average, there is a ~0.5% increase in the
+year-to-year revenue ratio per 1% decrease in species diversity, and across
+strategies weighted equally, `r g2_dens`% of the probability density supports
+a positive relationship between specialization and revenue. However, when
+not controlling for effort, the average effect of specializing on revenue
+becomes more neutral (Fig. S\fignoeffort).
 
@@
 Changing (increasing or decreasing) an individual's fished species diversity
-within the constraints of a set of permits tends to be associated with greater 
+within the constraints of a strategy tends to be associated with greater
 expected variability or uncertainty in year-to-year revenue (Fig.
@@
 This effect was strongest for individuals with a halibut and sablefish permit.
-Individuals with this strategies experienced a ~2% increase in variability for a
-1% change in the year-to-year ratio of species diversity (Fig.
-\ref{fig:within}). Becoming more generalized in species diversity also leads to
-increased revenue variability (Fig. S\ref{fig:mirror}A), suggesting that the act
-of changing the number of species fished for a given set of permits is a
-contributor to increased variability of revenue.
+Individuals with this strategy experienced a ~2% increase in variability for
+a 1% change in the year-to-year ratio of species diversity (Figs
+\ref{fig:within}, S\figmirror). 
+<!-- Becoming more generalized in species diversity -->
+<!-- also leads to increased revenue variability (Fig. S\figmirror A), suggesting -->
+<!-- that the act of changing the number of species fished for a given set of permits -->
+<!-- is a contributor to increased variability of revenue. -->
 
@@
 individuals with only halibut or sablefish permits (Fig. S\figifq A).
-<!-- TODO AH: We need to note that the populatoin changed -->
+<!-- TODO AH: We need to note that the population changed -->
 Furthermore, for individuals within most

### 2017-01-27 portfolio/text/supp.Rnw
Order figs; edits

@@
 \clearpage
-% TODO: AB: I suggest 2 additional columns: 1) potential species that could be harvested under each permit strategy, 2) number of unique permits contained within each strategy.
-This would really help to show how much flexibility someone within each permit strategy might have.
+
 \begin{table}[h!]
@@
 \begin{center}
+\includegraphics[width=1\textwidth]{../figs/stan-gg-spoke-all.pdf}
+\caption{The effect of adding a permit to one's strategy
+on estimated revenue variability and median revenue.
+This is an extended version of Fig.~2 that includes all major
+strategy combinations.}
+\label{fig:spoke-all}
+\end{center}
+\end{figure}
+
+\begin{figure}[htbp]
+\begin{center}
 \includegraphics[width=5in]{../figs/stan-main-effects.pdf}
@@
 predicting revenue ($\beta$) and variability ($\gamma$) in revenue.
-The $\eta$ coefficients represents strategy-level predictors
+The $\eta$ coefficients represent strategy-level predictors
 relating mean species diversity and mean days fished
@@
 two times their standard deivations)
-to the magnitude of residual error. (TODO symbols need to be updated)}
+to the magnitude of residual error.}
 \label{fig:fe}
@@
 
-% TODO: tweak labels to fit
+\clearpage
+
 \begin{figure}[htbp]
 \begin{center}
-\includegraphics[width=1\textwidth]{../figs/stan-gg-spoke-all.pdf}
-\caption{The effect of adding a permit to one's strategy
-on estimated revenue variability and median revenue.
-This is an extended version of Fig.~2 that includes all major
-strategy combinations.}
-\label{fig:spoke-all}
+\includegraphics[width=1.05\textwidth]{../figs/stan-offset-break-anti-spaghetti.pdf}
+\caption{Effects of (A) generalizing and (B) specializing from year-to-year for
+  individual permit holders \emph{within} the \Sexpr{nstr} most common
+  strategies. Changing species diversity from year-to-year is associated with
+  higher variability for individuals within the strategies above the zero line.
+  Generalizing (A) or specializing (B) is associated with greater expected
+  revenue ratios to the right of the zero line and lower expected revenue ratios
+  to the left of the zero line. These effects are estimated holding fishing
+  season length constant from year-to-year. Dots and line segments represent
+  medians and 50\% credible intervals of posteriors. Panel B and Fig.~4 are the
+  same.}
+\label{fig:mirror}
 \end{center}
@@
 
-% TODO: all figs referenced?
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=0.6\textwidth]{../figs/stan-offset-break-anti-spaghetti-specializing-no-effort.pdf}
+\caption{Same as Fig.~4 but without the effort (days fished) predictor and
+  without its interaction with species diversity.}
+\label{fig:noeffort}
+\end{center}
+\end{figure}
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=1\textwidth]{../figs/itq-combined.pdf}
+\caption{The effect of IFQs (individual fishing quotas) on model estimates. (A)
+  Posterior density of the ratio between strategy-level estimates of
+  standardized revenue variability ($\gamma_{0j}$'s) after the introduction of
+  IFQs vs.\ before the introduction of IFQs. (B) Estimates (medians) of the
+  effect of generalizing (increasing species diversity) on variability
+  ($\gamma_{1j}$'s) for the full model (``Grouped'') compared to a model where
+  separate effects are estimated before and after IFQs were introduced (``IFQ
+  split''). Pattern looks similar for the effect of specializing on variability
+  ($\gamma_{2j}$'s) and is not shown.}
+\label{fig:ifq}
+\end{center}
+\end{figure}
+
+\begin{figure}[htbp]
+\begin{center}
+\includegraphics[width=1\textwidth]{../figs/strategy-year-effects.pdf}
+\caption{Strategy-year intercept estimates ($\beta_{0,j,t}$ in
+Eq.~\ref{eq:mean}) from the hierarchical model.
+Shown are 40 samples from the posterior distribution.}
+\label{fig:str-yrs}
+\end{center}
+\end{figure}
 
@@
 
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[width=1\textwidth]{../figs/strategy-year-effects.pdf}
-\caption{Strategy-year intercept estimates ($\beta_{0,j,t}$ in
-Eq.~\ref{eq:mean}) from the hierarchical model.
-Shown are 40 samples from the posterior distribution.}
-\label{fig:str-yrs}
-\end{center}
-\end{figure}
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[width=1\textwidth]{../figs/itq-combined.pdf}
-\caption{The effect of IFQs (individual fishing quotas) on model estimates. (A)
-  Posterior density of the ratio between strategy-level estimates of
-  standardized revenue variability ($\gamma_{0j}$'s) after the introduction of
-  IFQs vs.\ before the introduction of IFQs. (B) Estimates (medians) of the
-  effect of generalizing (increasing species diversity) on variability
-  ($\gamma_{1j}$'s) for the full model (``Grouped'') compared to a model where
-  separate effects are estimated before and after IFQs were introduced (``IFQ
-  split''). Pattern looks similar for the effect of specializing on variability
-  ($\gamma_{2j}$'s) and is not shown.}
-\label{fig:ifq}
-\end{center}
-\end{figure}
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[width=1.05\textwidth]{../figs/stan-offset-break-anti-spaghetti.pdf}
-\caption{Effects of (A) generalizing and (B) specializing from year-to-year for
-  individual permit holders \emph{within} the \Sexpr{nstr} most common
-  strategies. Changing species diversity from year-to-year is associated with
-  higher variability for individuals within the strategies above the zero line.
-  Generalizing (A) or specializing (B) is associated with greater expected
-  revenue ratios to the right of the zero line and lower expected revenue ratios
-  to the left of the zero line. These effects are estimated holding fishing
-  season length constant from year-to-year. Dots and line segments represent
-  medians and 50\% credible intervals of posteriors. Panel B and Fig.~4 are the
-  same.}
-\label{fig:mirror}
-\end{center}
-\end{figure}
-
-\clearpage
-
-\begin{figure}[htbp]
-\begin{center}
-\includegraphics[width=0.6\textwidth]{../figs/stan-offset-break-anti-spaghetti-specializing-no-effort.pdf}
-\caption{Same as Fig.~4 but without the effort (days fished) predictor and
-  without its interaction with species diversity.}
-\label{fig:noeffort}
-\end{center}
-\end{figure}
-
 % \begin{figure}[htbp]

### 2017-01-27 portfolio/analysis/3.3-plot-main-effects.R

@@ -42,12 +42,14 @@ term_lu <- data.frame(
     "η0 (global intercept, variability)",
     "η1 (strategy-level diversity effect, variability)",
     "η2 (strategy-level days-fished effect, variability)"
-  )
+  ), stringsAsFactors = FALSE
 )
+term_lu$order <- seq_len(nrow(term_lu))
 p$term_clean <- NULL
+p$order <- NULL
 p <- inner_join(p, term_lu)
 p$term_clean <- factor(p$term_clean,
-  levels = rev(sort(as.character(p$term_clean))))
+  levels = p$term_clean[rev(p$order)])
 
 # invert specializing coefs to match fig 4:
 p[grepl("specializing", p$term_clean),c("l", "l.5", "m", "u.5", "u")] <-
@@ -60,7 +62,7 @@ p1 <- ggplot(p, aes(x = term_clean, y = m)) +
   ylab("Parameter estimate") +
   geom_hline(aes(yintercept = 0), lty = 2, col = "grey60") +
   geom_vline(aes(xintercept = 8.5), lty = 2, col = "grey60") +
-  geom_vline(aes(xintercept = 2.5), lty = 2, col = "grey60") +
+  geom_vline(aes(xintercept = 3.5), lty = 2, col = "grey60") +
   coord_flip() + theme_gg() + xlab("")
 # ggsave("portfolio/figs/stan-main-effects.pdf", width = 5.6, height = 4)
 cairo_pdf("portfolio/figs/stan-main-effects.pdf", width = 5.6, height = 4)

### 2017-01-31 portfolio/text/anderson-etal-rev-pfx.tex
Text edits

@@
 fishers and how year-to-year changes in diversification affect revenue and
-revenue variability. Here, we evaluate two axes by which fishers in Alaska,
+revenue variability. Here, we evaluate two axes by which fishers in Alaska
 can diversify fishing
@@
 constraints of existing permits has a fishery-dependent effect on revenue and is
-usually (93\% probability) associated with increased revenue uncertainty
+usually (XX\% probability) associated with increased revenue uncertainty
 the following year. Our results demonstrate that the most effective option for
@@
 \acknow{This work was conducted as part of the National Center for Ecological
-  Analysis and Synthesis working group "Applying portfolio effects to the Gulf
+  Analysis and Synthesis working group ``Applying portfolio effects to the Gulf
   of Alaska ecosystem: Did multi-scale diversity buffer against the Exxon
-  Valdez oil spill?", funded by TODO. We thank other participants of the
+  Valdez oil spill?'', funded by TODO. We thank other participants of the
   working group for helpful discussions. Additional support was provided by a

### 2017-01-31 portfolio/text/ms.Rmd
Text edits

@@
 industry [@kirilenko2007].
-<!-- TODO Could add anthropogenic perturbations (e.g., Exxon Valdez) to this
+<!-- Could add anthropogenic perturbations (e.g., Exxon Valdez) to this
 list or in the next paragraph. -->
@@
 worth bringing up, but this is always the case anyway with vessels where the
-skipper may do other things. triue, and for others-->
+skipper may do other things. triue, and for others - KH and Cl et al. ...-->
 
@@
 specializing tends to be associated with increased revenue but also increased
-variability (Fig. \ref{fig:within}; Figs S\figfe, S\figre, S\figmirror). <!--
-TODO LP: effort is endogenous? --> On average, there is a ~0.5% increase in the
+variability (Fig. \ref{fig:within}; Figs S\figfe, S\figre, S\figmirror).
+On average, there is a ~0.5% increase in the
 year-to-year revenue ratio per 1% decrease in species diversity, and across
@@
 individuals with only halibut or sablefish permits (Fig. S\figifq A).
-<!-- TODO AH: We need to note that the population changed -->
+<!-- TODO AH: although ... We need to note that the population changed -->
 Furthermore, for individuals within most
@@
 when a financial crisis hit [@ikerd1990].
-<!-- TODO AB: We might want to include more examples / references from fisheries. -->
+<!-- AB: We might want to include more examples / references from fisheries.
+herring - affected by closures? value declined...-->
 Third, diversifying---a lack of
@@
 This goes back to my comment re: Table 1 (see p.
-15) -->
+15)
+sea cuc - H W book - learning...
+herring - alaska 1970s - learn to fish for roe ... [ref]
+salmon - hatchery production - change target from diverse to most abundant... PWS seine - pink salmon or SE
+Hal ll + S troll...
+-->
 Additionally, changes in diversity might reflect years in which targeted species
 were less available and fishers instead chose to target other species or landed
-more bycatch,
+more non-target species,
 <!-- TODO AB: This is the first time bycatch is explicitly mentioned in the study and it raises some questions about how target vs.
@@
 Some clarification in the methods would be helpful.
--->
+all catch including bycatch non-target species-->
 adding stochasticity to their revenue. Most importantly, this

### 2017-02-09 portfolio/text/ms.Rmd
Edit text

@@
 was associated with a reduction in revenue variability of ~20--30% for
-individuals with only halibut or sablefish permits (Fig. S\figifq A).
-<!-- TODO AH: although ... We need to note that the population changed -->
+individuals with only halibut or sablefish permits (Fig. S\figifq A;
+although other changes in fish populations occurred concurrently).
 Furthermore, for individuals within most
@@
 revenue. This could be for a number of reasons. First, specialization can
-increase fishing efficiency [@smith1986]. Efficiency gains from specialization
-can be seen in other natural resource sectors. For example, rice farmers in
-Ghana can increase their efficiency by devoting more land and labor to rice
-production compared to other crops [@abdulai2000]. <!-- TODO: weak -->
+increase fishing efficiency [@smith1986]. 
+Similar efficiency gains from specialization are well known in 
+agricultural fields [@ikerd1990; @abdulai2000].
+<!-- For example, rice farmers in -->
+<!-- Ghana can increase their efficiency by devoting more land and labor to rice -->
+<!-- production compared to other crops [@abdulai2000]. -->
 Second, given the
@@
 as individuals become accustomed to new locations, gear, or species.
-<!-- TODO AB: I think in this paragraph it would be good to pick an example or two and speculate on potential reasons specific to those permit strategies.
+For example, many individuals fishing herring in Alaska in the 1970s 
+<!-- AB: I think in this paragraph it would be good to pick an example or two and speculate on potential reasons specific to those permit strategies.
 Otherwise, statements like this one aren’t really anchored to anything concrete.
@@
 more non-target species,
-<!-- TODO AB: This is the first time bycatch is explicitly mentioned in the study and it raises some questions about how target vs.
-bycatch species are defined and considered in the analysis.
-Some clarification in the methods would be helpful.
-all catch including bycatch non-target species-->
 adding stochasticity to their revenue. Most importantly, this

### 2017-02-09 portfolio/text/supp.Rnw
Text edits
Integrate main species into supplemental table

@@
 
-% TODO: Are you saying that PWS sockeye permits were combined with BB sockeye permits?I can't remember but that's what it sounds like so I just wanted to be clear.
-
 After this process of combining permits,
@@
 \begin{table}[h!]
-\footnotesize
+\scriptsize
   \centering
@@
 a particular strategy represents a group of similar permits across areas
-or vessel sizes.}
+or vessel sizes. The column ``Main species'' indicates species that make up at
+least 10\% of the total revenue across all years and individuals for that
+fishing strategy. ``Sock'' refers to sockeye salmon.}
   \label{tab:strategies}
-  \begin{tabular}{lll}
+  \begin{tabular}{llll}
   \toprule

### 2017-02-09 portfolio/analysis/3.3-make-strategy-table.R

@@ -12,6 +12,87 @@ d <- d %>% mutate(modeled = ifelse(strategy %in% md$strategy, "Yes", "No"))
 # d <- inner_join(d, spp)
 d <- inner_join(d, select(md, strategy, strategy_mean_div))
 
+#######
+dat <- readRDS(file="portfolio/data-generated/cfec-annual-for-modeling.rds")
+dat$permit_group = NA
+dat$permit_group[dat$permit%in%c("S05B","S15B")] = "S - troll"
+dat$permit_group[which(dat$permit%in%c("S03H","S03M","S03T"))] = "S - drift - CI, AP, BB (sock)"
+dat$permit_group[which(dat$permit%in%c("S01H","S01K"))] = "S - seine - CI, K"
+dat$permit_group[which(dat$permit%in%c("S04E","S04H","S04K","S04M","S04T"))] = "S - set - PWS, CI, K, AP, BB (sock)"
+dat$permit_group[which(dat$permit%in%c("S04W","S04Z"))] = "S - set - NS/KU"
+dat$permit_group[which(dat$permit%in%c("S01L"))] = "S - seine - CH"
+dat$permit_group[which(dat$permit%in%c("S01A"))] = "S - seine - SE"
+dat$permit_group[which(dat$permit%in%c("S01E"))] = "S - seine - PWS"
+dat$permit_group[which(dat$permit%in%c("S03E"))] = "S - drift - PWS"
+dat$permit_group[which(dat$permit%in%c("S01M"))] = "S - seine - AP"
+dat$permit_group[which(dat$permit%in%c("S03A"))] = "S - drift - SE"
+dat$permit_group[which(dat$permit%in%c("S04D"))] = "S - set - YAK"
+
+salmon_permit_table = group_by(dat[which(!is.na(dat$permit_group)),], permit_group) %>%
+  summarize(tot = sum(chnk) + sum(coho) + sum(pink) + sum(sock) + sum(chum),
+    chnk = round(sum(chnk)/tot,2), coho = round(sum(coho)/tot,2), pink = round(sum(pink)/tot,2),
+    sock = round(sum(sock)/tot,2), chum = round(sum(chum)/tot,2)) %>%
+  select(-tot)
+
+concatenate_species = function(x) {
+  tmp = x[c("chnk","coho","pink","sock","chum")] > 0.1
+  paste(names(tmp)[tmp], collapse=", ")
+}
+
[…]

### 2017-02-09 portfolio/analysis/3.5-output-values.R
Calculate probability of increased variance with diversity change

@@ -45,6 +45,12 @@ n_jacknifed <- names(nums[n_jacknifed])
 
 mean_permit_days <- round(mean(md$strategy_mean_days)/10) * 10
 
+e1 <- -1 * rstan::extract(m)$coef_g1_strategy
+e2 <- rstan::extract(m)$coef_g2_strategy
+
+e <- rbind(e1, e2)
+prob_change_inc_variance <- round(sum(e > 0) / (nrow(e)*ncol(e)), 2) * 100
+
 ndat <- readRDS("portfolio/data-generated/ndat.rds")
 ndat_cull <- readRDS("portfolio/data-generated/ndat-cull.rds")
 ndat <- c(ndat, ndat_cull)

### 2017-02-09 portfolio/text/anderson-etal-rev-pfx.tex
Edit text

@@
 S.C.A, E.J.W, and A.O.S designed the study with contributions from all authors;
-J.C.S. also provided data;
-S.C.A. and E.J.W. analyzed the data and wrote the paper with input from all
-authors.}
+J.C.S. and J.T.W. also provided and cleaned the data;
+S.C.A. and E.J.W. analyzed the data and wrote the paper with input from all authors.}
 \authordeclaration{The authors declare no conflict of interest.}
@@
 constraints of existing permits has a fishery-dependent effect on revenue and is
-usually (XX\% probability) associated with increased revenue uncertainty
+usually (87\% probability) associated with increased revenue uncertainty
 the following year. Our results demonstrate that the most effective option for
@@
 
-We obtained commercial fisheries landings and revenue data for all permit
+We obtained commercial fisheries landings and revenue data for all target and
+non-target species for permit
 holders in Alaska from 1985 to 2014 from the Alaskan Commercial Fisheries Entry
@@
   of Alaska ecosystem: Did multi-scale diversity buffer against the Exxon
-  Valdez oil spill?'', funded by TODO. We thank other participants of the
+  Valdez oil spill?''. We thank other participants of the
   working group for helpful discussions. Additional support was provided by a
-  David H. Smith Conservation Research Fellowship (S.C.A.), the NOAA Fisheries'
-  Spatial Economics Toolbox for Fisheries (FishSET) Project (J.T.W), \ldots }
+  David H. Smith Conservation Research Fellowship (S.C.A.) and the NOAA Fisheries'
+  Spatial Economics Toolbox for Fisheries (FishSET) Project (J.T.W).}
 

### 2017-02-09 salmon/paper/salmon_paper.Rmd
herring case study
added model for rw in variance (doesn't converge)

@@
   effects on variation in individual revenues'
-author: "Eric, Sean, Ole, Rich, Milo, Jordan, Jen, Alan, others"
-date: ''
+author: "Eric, Sean, Ole, Rich, Milo, Jordan, Jen, Alan, Ben, etc."
+date: '2017-02-08'
 output:
@@
 library(date)
-library(ggtern)
+#library(ggtern)
 library(mgcv)
-library(fpc)
+#library(fpc)
 library(gridExtra)
-library(glmmTMB)
+#library(glmmTMB)
 library(zoo)
 library(rstan)
+library(ggrepel)
 modelname = "model-linear" # model-linear-nointeractions
-
 ```
@@
 
-We obtained fisheries landings and revenue data for all permit holders in Alaska from 1985-2014 from the Commercial Fisheries Entry Commission (CFEC).
-We adjusted for inflation by setting all revenues in 2009 USD.
-Because we were only interested in single permit fisheries, we adopted a simplified version of the filtering steps described in [@anderson_benefits].
-After aggregating individual trip-level fish tickets to annual data, we retained individuals who only fished one of the six salmon permits included in a year (Table 1, Fig.
+We obtained fisheries landings and revenue data for all iundividual permit holders in Alaska from 1985-2014 from the Commercial Fisheries Entry Commission (CFEC).
+Each permit holder has a unique identifier that allows us to track quantities including the species caught and revenue received.
+We adjusted for inflation by converting all revenues to 2009 USD.
+Because of our focus on salmon fishers, we were only interested in single permit fisheries, and we adopted a simplified version of the filtering steps described in [@anderson_benefits].
+After aggregating individual trip-level catch and revenue to annual data, we retained permit holders who fished only one of the six salmon permits included in a year (Table 1, Fig.
 A5).
-Like [@anderson_benefits] we also worked with a log-differenced dataset to remove non-stationary individuals; to avoid complications of individuals switching strategies, we only included pairs of years for a given person where they did not change strategies (i.e., didn’t change fisheries or add permits, etc.).
+Like [@anderson_benefits] we also worked with a log-differenced dataset to remove non-stationary individuals; to avoid complications of individuals switching strategies (e.g.
+changing from a seine fishery in Kodiak to a gillnet fishery in Bristol Bay), we only included pairs of years for a given person where they did not change strategies (i.e., didn’t change fisheries or add permits, etc.).
 
@@
 [@ward_evaluating] summarized the trends for Prince William Sound).
-Release patterns have been dominated by pink and chum salmon ([@ward_evaluating], cite more specific #?) and production has increased since the late 1970s (217 million juvenile pink salmon in PWS in 1985 versus 672 million in 2014).
-Though our analysis with CFEC data doesn't separate hatchery versus wild fish, as a contrast we can make coarse comparisons across regions and species to examine potential effects of hatchery fish on variability.
+Release patterns have been dominated by pink and chum salmon ([@stopha2016]) and production has increased since the late 1970s (217 million juvenile pink salmon in PWS in 1985 versus 672 million in 2014).
+Though our analysis with CFEC data doesn't separate hatchery versus wild fish, as a contrast we can make comparisons across regions and species to examine potential effects of hatchery fish on variability.
 For example, we can compare the variation in the largely hatchery pink salmon returns (earnings and weight) from Prince William Sound to the largely wild pink salmon returns from Southeast Alaska.
@@
 
+- For small scale farmers, economists have often wondered why farmers don't diversify more.
+Diversification can be profitable, but in the short term risky or expensive.
+Bigger than cost seems to be the risk [@karlan_agricultural_2012]
+
 - The potential to diversify may be limited for areas with a limited number of available species or by managements actions that prohibit fishing in certain areas (e.g., PWS where seiners are generally confined to areas/times with pink and chum salmon whereas gillnet fishers are able to fish in areas with sockeye, chum, Chinook, coho, and pink salmon).
@@
 \break
+
+```{r, echo=FALSE, fig.pos="placeHere", fig.cap="CV and mean annual revenue by salmon permit, 1975-2014.
+Average participation (unique fishers) given by circle size, and effective species diversity is shown with color shading.", fig.height=3.2, eval=FALSE}
+dat.annual = readRDS(file="../../portfolio/data-generated/salmon-annual-for-modeling.rds")
+
+cfec_all = readRDS("../../portfolio/data-generated/salmon_data_for_paper.rds")
+people = group_by(cfec_all, p_fshy, year) %>% 
+  summarize(n = length(unique(p_holder))) %>% 
+  group_by(p_fshy) %>% summarize(n = mean(n,na.rm=T))
+names(people)[which(names(people)=="p_fshy")] = "strategy_permit"
+people$strategy_permit = paste0(substr(people$strategy_permit,1,1), substr(people$strategy_permit,3,5))
+
+dat.annual = filter(dat.annual, npermit==1 & substr(strategy_permit,1,1)=="S") %>%
+  group_by(strategy_permit) %>% 
+  mutate(nPeople = length(unique(p_holder))) %>% 
+  filter(nPeople > 200) %>% 
+  dplyr::select(-nPeople) %>% 
+  group_by(strategy_permit, year) %>% 
+  summarize(meanrev = mean(revenue), meandiv = mean(specDiv)) %>% 
+  arrange(strategy_permit, year) %>% 
+  group_by(strategy_permit) %>% 
+  summarize(mean_rev = mean(meanrev,na.rm=T), cv_rev = sd(diff(log(meanrev)),na.rm=T),
+    diversity = mean(meandiv,na.rm=T)) %>% as.data.frame
+
+dat.annual = left_join(dat.annual, people)
+dat.annual$gear = substr(dat.annual$strategy_permit,2,3)
+dat.annual$gear[which(dat.annual$gear=="01")]="Purse seine"
+dat.annual$gear[which(dat.annual$gear=="03")]="Drift gillnet"
+dat.annual$gear[which(dat.annual$gear=="04")]="Set gillnet"
+dat.annual$gear[which(dat.annual$gear=="05")]="Hand troll"
+dat.annual$gear[which(dat.annual$gear=="15")]="Power troll"
+
+dat.annual$price = NA
+dat.annual$price[dat.annual$strategy_permit=="S01A"]=250000
+dat.annual$price[dat.annual$strategy_permit=="S01E"]=186700
+dat.annual$price[dat.annual$strategy_permit=="S01H"]=84800
+dat.annual$price[dat.annual$strategy_permit=="S01K"]=40000
+dat.annual$price[dat.annual$strategy_permit=="S01L"]=227500
+dat.annual$price[dat.annual$strategy_permit=="S01M"]=56900
+
+dat.annual$price[dat.annual$strategy_permit=="S03A"]=88900
+dat.annual$price[dat.annual$strategy_permit=="S03E"]=224200
+dat.annual$price[dat.annual$strategy_permit=="S03H"]=63500
+dat.annual$price[dat.annual$strategy_permit=="S03M"]=119500
+dat.annual$price[dat.annual$strategy_permit=="S03T"]=148200
+dat.annual$price[dat.annual$strategy_permit=="S03W"]=NA
+
+dat.annual$price[dat.annual$strategy_permit=="S04D"]=88900
+dat.annual$price[dat.annual$strategy_permit=="S04H"]=15300
+dat.annual$price[dat.annual$strategy_permit=="S04K"]=77000
+dat.annual$price[dat.annual$strategy_permit=="S04M"]=55900
+dat.annual$price[dat.annual$strategy_permit=="S04T"]=38500
+dat.annual$price[dat.annual$strategy_permit=="S04W"]=7300
+dat.annual$price[dat.annual$strategy_permit=="S04X"]=5400
+dat.annual$price[dat.annual$strategy_permit=="S04Y"]=9900
+dat.annual$price[dat.annual$strategy_permit=="S04Z"]=11100
+
+dat.annual$price[dat.annual$strategy_permit=="S05B"]=11000
+dat.annual$price[dat.annual$strategy_permit=="S15B"]=38300
+
+g1 = ggplot(dat.annual, aes(x=log(mean_rev), y=log(cv_rev), label = dat.annual$strategy_permit)) + geom_point(aes(col=gear,size=diversity,alpha=0.7)) + scale_size(range = c(0, 20)) + geom_text_repel() + ylab("CV annual revenue") + xlab("Mean revenue (dollars)") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
+panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
+print(g1)
+pdf("Figure_01.pdf")
+print(g1)
+dev.off()
+```
 

### 2017-02-09 salmon/analysis/1-run-model.R

@@ -73,6 +73,25 @@ if(modelname=="linear") {
   save.image("salmon/analysis/model-linear.Rdata")
 }
 
+if(modelname=="linear_varrw") {
+  stan_pars = c("b0_str_yr", "b0_str_yr_tau", "g_0", "g0_strategy_tau",
+    "b_1", "b_2", "b_4", "b1_str_yr_tau", "b2_str_yr_tau",
+    "b4_str_yr_tau", "b1_str_yr_mu", "b2_str_yr_mu",
+    "b4_str_yr_mu", "g0", "g_1", "g_2", "g_4", "g1_str_yr_tau", "g2_str_yr_tau",
+    "g4_str_yr_tau", "g1_str_yr_mu", "g2_str_yr_mu",
+    "g4_str_yr_mu")
+
+  library(rstan)
+  rstan_options(auto_write = TRUE)
+  options(mc.cores = parallel::detectCores())
+
+  # estimate model. This model is modified from the simulation model by (1) including indices to allow NAs in the inputted data, and (2) including estimated year effects (intercepts)
+  mod = stan(file = 'salmon/analysis/portfolio-offset-linear-varrw.stan',data = stan_data,
+    verbose = TRUE, chains = 3, thin = 1, warmup = 1000, iter = 2000, pars = stan_pars)
+
+  save.image("salmon/analysis/model-linear-varrw.Rdata")
+}
+
 if(modelname=="interceptonly") {
 stan_pars = c("b0_str_yr", "b0_str_yr_tau", "g0_strategy", "g0_strategy_tau",
   "b_1", "b_4", "b1_str_yr_tau", "b4_str_yr_tau",

### 2017-02-09 script_for_salmon_tables.R

@@ -1,0 +1,31 @@
+library(dplyr)
+
+dat <- readRDS(file="portfolio/data-generated/cfec-annual-for-modeling.rds")
+dat$permit_group = NA
+dat$permit_group[dat$permit%in%c("S05B","S15B")] = "salmon_troll"
+dat$permit_group[which(dat$permit%in%c("S03H","S03M","S03T"))] = "drift_gillnet" 
+dat$permit_group[which(dat$permit%in%c("S01H","S01K"))] = "purse_seine"
+dat$permit_group[which(dat$permit%in%c("S04E","S04H","S04K","S04M","S04T"))] = "set_gillnet"
+dat$permit_group[which(dat$permit%in%c("S04W","S04Z"))] = "set_gillnet2"
+
+salmon_permit_table = group_by(dat[which(!is.na(dat$permit_group)),], permit_group) %>% 
+  summarize(tot = sum(chnk) + sum(coho) + sum(pink) + sum(sock) + sum(chum),
+    chnk = round(sum(chnk)/tot,2), coho = round(sum(coho)/tot,2), pink = round(sum(pink)/tot,2), 
+    sock = round(sum(sock)/tot,2), chum = round(sum(chum)/tot,2)) %>% 
+  select(-tot)
+
+# alternatively do it by permit
+salmon_permit_table2 = group_by(dat[which(dat$permit %in% c("S05B","S15B","S03H",
+  "S03M","S03T","S01H","S01K","S04E","S04H","S04K","S04M","S04T","S04W","S04Z")),], permit, year) %>% 
+  summarize(tot = sum(chnk,na.rm=T) + sum(coho,na.rm=T) + sum(pink,na.rm=T) + sum(sock,na.rm=T) + sum(chum,na.rm=T),
+    chnk = sum(chnk)/tot, coho = sum(coho)/tot, pink = sum(pink)/tot, 
+    sock = sum(sock)/tot, chum = sum(chum)/tot,
+    permit_group = permit_group[1]) %>% 
+  select(-tot) %>% 
+  group_by(permit) %>% 
+  summarize(chnk = round(mean(chnk,na.rm=T),2), pink = round(mean(pink,na.rm=T),2), coho = round(mean(coho,na.rm=T),2),
+    chum = round(mean(chum,na.rm=T),2), sock = round(mean(sock,na.rm=T),2), permit_group = permit_group[1])
+
+
+
+

### 2017-02-09 portfolio/analysis/3.6-sim.R
Adjust simulation test

@@ -97,38 +97,9 @@ standat <- list(
   mean_day_str = xh1
 )
 
-beta_init <- function() rnorm(standat$J)
-sigma_init <- function() rnorm(standat$K)
-dev_str_init <- function() rnorm(standat$n_strategy, 0, 0.05)
-dev_yr_init <- function() rnorm(standat$n_str_yr, 0, 0.05)
-tau_init <- function() runif(1, 0.05, 0.1)
-init_fun <- function() {
-  list(
-    b0 = beta_init()[1],
-    b0_strategy = dev_str_init(),
-    b0_str_yr = dev_yr_init(),
-    b0_strategy_tau = tau_init(),
-    b0_str_yr_tau = tau_init(),
-    b_j = beta_init(),
-    h1 = beta_init()[1],
-    h2 = beta_init()[1],
-    b1_strategy = dev_str_init(),
-    b1_strategy_tau = tau_init(),
-    b2_strategy = dev_str_init(),
-    b2_strategy_tau = tau_init(),
-    g0 = sigma_init()[1],
-    g0_strategy = dev_str_init(),
-    g0_strategy_tau = tau_init(),
-    g_k = sigma_init(),
-    g1_strategy = dev_str_init(),
-    g1_strategy_tau = tau_init(),
-    g2_strategy = dev_str_init(),
-    g2_strategy_tau = tau_init())
-}
-
 msim <- stan("portfolio/analysis/portfolio-offset-sim.stan",
-  data = standat, iter = 400, chains = 4,
-  pars = c("mu", "sigma"), include = FALSE, init = init_fun)
+  data = standat, iter = 500, chains = 2,
+  pars = c("mu", "sigma"), include = FALSE)
 save(msim, file = "portfolio/data-generated/nsim.rda")
 load("portfolio/data-generated/nsim.rda")
 

### 2017-02-09 portfolio/text/anderson-etal-rev-pfx.tex
Draft a significance statement

@@
 % Please add here a significance statement to explain the relevance of your work
-\significancestatement{TODO}
+% <120 words
+\significancestatement{
+Individuals relying on natural resources to sustain their livelihoods,
+such as fishers, farmers, and forestry workers,
+face high levels of income variability.
+For fishers, targeting a diversity of species has been shown to
+reduce community- and vessel-level revenue variability
+but the individual-level consequences are unknown.
+Our work demonstrates that individuals in fisheries targeting
+a diversity of species
+and individuals that participate in multiple fisheries
+buffer income variability compared to less diverse individuals.
+However, adjusting (increasing or decreasing) species diversity from year-to-year
+usually increases individual-level revenue variability.
+\textbf{OR}
+However, specializing from year-to-year
+(reducing species diversity)
+tends to be associated with increased revenue
+as well as increased revenue variability
+when controlling for effort.
+The most effective option for individuals to
+reduce revenue variability via diversification
+--- purchasing additional permits ---
+is also expensive, often limited by regulations,
+and therefore unavailable to many.
+}
+% We ask how multiple forms of
+% species diversification affect individual-level
+% revenue and revenue variability for Alaskan fishers.
 
@@
 Individuals relying on the extraction of natural resources for their
-livelihood---fishers, farmers, miners, and loggers, among others---face high
+livelihood face high
 levels of income variability driven by a mix of environmental, biological,
@@
   Analysis and Synthesis working group ``Applying portfolio effects to the Gulf
-  of Alaska ecosystem: Did multi-scale diversity buffer against the Exxon
-  Valdez oil spill?''. We thank other participants of the
+  of Alaska ecosystem: Did multi-scale diversity buffer against the \textit{Exxon
+  Valdez} oil spill?''. We thank other participants of the
   working group for helpful discussions. Additional support was provided by a

### 2017-02-10 portfolio/text/anderson-etal-rev-pfx.tex
eric's suggested changes

@@
 face high levels of income variability.
-For fishers, targeting a diversity of species has been shown to
-reduce community- and vessel-level revenue variability
+For fishers in particular, targeting multiple species has been shown to
+reduce revenue variability at large scales (vessels, communities)
 but the individual-level consequences are unknown.
@@
 buffer income variability compared to less diverse individuals.
-However, adjusting (increasing or decreasing) species diversity from year-to-year
-usually increases individual-level revenue variability.
+However, large adjustments in diversification strategies from year-to-year appear risky and
+usually increases revenue variability.
 \textbf{OR}

### 2017-02-10 portfolio/text/anderson-etal-rev-pfx.tex
Edits

--- 2017-02-09 portfolio/text/anderson-etal-rev-pfx.tex
@@
 face high levels of income variability.
-For fishers, targeting a diversity of species has been shown to
-reduce community- and vessel-level revenue variability
+For fishers in particular, targeting multiple species has been shown to
+reduce revenue variability at large scales (vessels and communities)
 but the individual-level consequences are unknown.
@@
 buffer income variability compared to less diverse individuals.
-However, adjusting (increasing or decreasing) species diversity from year-to-year
-usually increases individual-level revenue variability.
-\textbf{OR}
-However, specializing from year-to-year
-(reducing species diversity)
-tends to be associated with increased revenue
-as well as increased revenue variability
-when controlling for effort.
+However, large adjustments in diversification strategies from year-to-year appear risky and
+usually increase revenue variability.
+% \textbf{OR}
+% However, specializing from year-to-year
+% (reducing species diversity)
+% tends to be associated with increased revenue
+% as well as increased revenue variability
+% when controlling for effort.
 The most effective option for individuals to

### 2017-02-11 portfolio/text/anderson-etal-rev-pfx.tex
Ole tiny edits to Sig Statement.

@@
 \significancestatement{
-Individuals relying on natural resources to sustain their livelihoods,
+% AOS tiny edits:
+Individuals that rely on natural resources for their livelihoods,
 such as fishers, farmers, and forestry workers,
 face high levels of income variability.
-For fishers in particular, targeting multiple species has been shown to
+For fishers, harvesting multiple species has been shown to
 reduce revenue variability at large scales (vessels and communities)
-but the individual-level consequences are unknown.
+but the individual-level consequences of maintaining catch diversity are unknown.
 Our work demonstrates that individuals in fisheries targeting
@@
 buffer income variability compared to less diverse individuals.
-However, large adjustments in diversification strategies from year-to-year appear risky and
+However, large adjustments in diversification strategies from year-to-year are risky and
+usually increase revenue variability. The most effective option to
+reduce revenue variability via diversification
+--- purchasing additional permits ---
+is also expensive, often limited by regulations,
+and therefore unavailable to many.
+
+Individuals relying on natural resources to sustain their livelihoods,
+such as fishers, farmers, and forestry workers,
+face high levels of income variability.
+For fishers in particular, harvesting multiple species has been shown to
+reduce revenue variability at large scales (vessels and communities)
+but the individual-level consequences of maintaining catch diversity are unknown.
+Our work demonstrates that individuals in fisheries targeting
+a diversity of species
+and individuals that participate in multiple fisheries
+buffer income variability compared to less diverse individuals.
+However, large adjustments in diversification strategies from year-to-year are risky and
 usually increase revenue variability.

### 2017-02-11 portfolio/text/anderson-etal-rev-pfx.tex
Update significance statement

@@
 \significancestatement{
-% AOS tiny edits:
 Individuals that rely on natural resources for their livelihoods,
@@
 face high levels of income variability.
-For fishers, harvesting multiple species has been shown to
+For fishers, catching multiple species has been shown to
 reduce revenue variability at large scales (vessels and communities)
@@
 usually increase revenue variability. The most effective option to
-reduce revenue variability via diversification
---- purchasing additional permits ---
-is also expensive, often limited by regulations,
-and therefore unavailable to many.
-
-Individuals relying on natural resources to sustain their livelihoods,
-such as fishers, farmers, and forestry workers,
-face high levels of income variability.
-For fishers in particular, harvesting multiple species has been shown to
-reduce revenue variability at large scales (vessels and communities)
-but the individual-level consequences of maintaining catch diversity are unknown.
-Our work demonstrates that individuals in fisheries targeting
-a diversity of species
-and individuals that participate in multiple fisheries
-buffer income variability compared to less diverse individuals.
-However, large adjustments in diversification strategies from year-to-year are risky and
-usually increase revenue variability.
-% \textbf{OR}
-% However, specializing from year-to-year
-% (reducing species diversity)
-% tends to be associated with increased revenue
-% as well as increased revenue variability
-% when controlling for effort.
-The most effective option for individuals to
 reduce revenue variability via diversification

### 2017-02-11 portfolio/text/anderson-etal-rev-pfx.tex
Edits

@@
 participate in additional or more diverse fisheries. However, this option is
-expensive, often limited by regulations such as license limitation or catch share programs, and
-consequently unavailable to many individuals. With increasing climatic variability,
+expensive, often limited by regulations such as catch share programs, and
+consequently unavailable to many individuals. With increasing climatic variability
 it will be particularly important that

### 2017-02-18 portfolio/text/anderson-etal-rev-pfx.tex
Editing tweaks and add data statement

@@
 
-\affil[f]{NOAA / National Marine Fisheries Service, Alaska Fisheries Science Center, Auke
+\affil[f]{Alaska Fisheries Science Center, National Marine Fisheries Service, National Oceanographic and
+Atmospheric Administration, Auke
 Bay Laboratories, 17109 Pt. Lena Loop Rd., Juneau, AK 99801, USA}
@@
 
-% Please add here a significance statement to explain the relevance of your work
-% <120 words
-\significancestatement{
-Individuals that rely on natural resources for their livelihoods,
-such as fishers, farmers, and forestry workers,
-face high levels of income variability.
-For fishers, catching multiple species has been shown to
-reduce revenue variability at large scales (vessels and communities)
-but the individual-level consequences of maintaining catch diversity are unknown.
-Our work demonstrates that individuals in fisheries targeting
-a diversity of species
-and individuals that participate in multiple fisheries
-buffer income variability compared to less diverse individuals.
-However, large adjustments in diversification strategies from year-to-year are risky and
-usually increase revenue variability. The most effective option to
-reduce revenue variability via diversification
---- purchasing additional permits ---
-is also expensive, often limited by regulations,
-and therefore unavailable to many.
-}
-% We ask how multiple forms of
-% species diversification affect individual-level
-% revenue and revenue variability for Alaskan fishers.
+% Please add here a significance statement to explain the relevance of your work % <120 words \significancestatement{
+Individuals that rely on natural resources for their livelihoods, such as fishers, farmers, and forestry workers, face high levels of income variability.
+For fishers, catching multiple species has been shown to reduce revenue variability at large scales (vessels and communities) but the individual-level consequences of maintaining catch diversity are unknown.
+Our work demonstrates that individuals in fisheries targeting a diversity of species and individuals that participate in multiple fisheries buffer income variability compared to less diverse individuals.
+However, large adjustments in diversification strategies from year-to-year are risky and usually increase revenue variability.
+The most effective option to reduce revenue variability via diversification --- purchasing additional permits --- is also expensive, often limited by regulations, and therefore unavailable to many.
+} % We ask how multiple forms of % species diversification affect individual-level % revenue and revenue variability for Alaskan fishers.
 
@@
 \begin{abstract}
-Individuals relying on the extraction of natural resources for their
-livelihood face high
-levels of income variability driven by a mix of environmental, biological,
-management, and
-economic factors. Key to managing these industries is identifying
-how regulatory actions and individual behaviour affect income variability and, by
-extension, economic stability and the sustainable use of natural resources. In
-commercial fisheries, communities and vessels fishing a greater diversity of
-species have less revenue variability than those fishing fewer species.
-However, it is unclear if these benefits extend to the actions of individual
-fishers and how year-to-year changes in diversification affect revenue and
-revenue variability. Here, we evaluate two axes by which fishers in Alaska
-can diversify fishing
-activities. We show that, despite increasing specialization over the last 30
-years, fishing a set of permits with higher species diversity reduces individual revenue
-variability and fishing an additional permit is associated with higher revenue
-and lower variability. However, increasing species diversity within the
-constraints of existing permits has a fishery-dependent effect on revenue and is
-usually (87\% probability) associated with increased revenue uncertainty
-the following year. Our results demonstrate that the most effective option for
-individuals to decrease revenue variability is to
-participate in additional or more diverse fisheries. However, this option is
-expensive, often limited by regulations such as catch share programs, and
-consequently unavailable to many individuals. With increasing climatic variability
-it will be particularly important that
-individuals relying on natural resources for their livelihood have effective
-strategies to reduce income variability.
+Individuals relying on the extraction of natural resources for their livelihood face high levels of income variability driven by a mix of environmental, biological, management, and economic factors.
+Key to managing these industries is identifying how regulatory actions and individual behaviour affect income variability and, by extension, economic stability and the sustainable use of natural resources.
+In commercial fisheries, communities and vessels fishing a greater diversity of species have less revenue variability than those fishing fewer species.
+However, it is unclear if these benefits extend to the actions of individual fishers and how year-to-year changes in diversification affect revenue and revenue variability.
+Here, we evaluate two axes by which fishers in Alaska can diversify fishing activities.
+We show that, despite increasing specialization over the last 30 years, fishing a set of permits with higher species diversity reduces individual revenue variability and fishing an additional permit is associated with higher revenue and lower variability.
+However, increasing species diversity within the constraints of existing permits has a fishery-dependent effect on revenue and is usually (87\% probability) associated with increased revenue uncertainty the following year.
+Our results demonstrate that the most effective option for individuals to decrease revenue variability is to participate in additional or more diverse fisheries.
+However, this option is expensive, often limited by regulations such as catch share programs, and consequently unavailable to many individuals.
+With increasing climatic variability it will be particularly important that individuals relying on natural resources for their livelihood have effective strategies to reduce income variability.
 % (`r g2_dens`% probability)
@@
 validated the model with simulated data, and were able to recover the true
-values (Supporting Information). Code used in our analysis is available at
+values (Supporting Information).
+
+The fisheries data that support the findings of this study, though
+confidential, are available from the Commercial Fisheries Entry
+Commission (<https://www.cfec.state.ak.us/>) and summaries used in
+this paper, along with all code, are archived in a publicly available repository at
 \url{https://github.com/NCEAS/pfx-commercial/portfolios/}. } \showmatmethods

### 2017-02-22 salmon/paper/archive/salmon_paper.Rmd
move files

--- 2017-01-24 salmon/paper/salmon_paper_AOS.Rmd
@@
 library(date)
-library(ggtern)
+#library(ggtern)
 library(mgcv)
-library(fpc)
+#library(fpc)
 library(gridExtra)
@@
 library(rstan)
+library(ggrepel)
 modelname = "model-linear" # model-linear-nointeractions
-
 ```

### 2017-02-22 salmon/paper/salmon_paper_v3.Rmd
Revert "Revert "local change/""
This reverts commit 240797038b8aada993df17a517b92bf49f992ec8.

--- 2017-02-09 salmon/paper/salmon_paper.Rmd
@@
 author: "Eric, Sean, Ole, Rich, Milo, Jordan, Jen, Alan, Ben, etc."
-date: '2017-02-08'
+date: ''
 output:
@@
 Using a long-term dataset of individual salmon fishers in Alaska, we develop new models for understanding how the effects of diversification have changed through time.
-These changes allow us to compare patterns across regions and gear, as well as better understand how external forces have changed incentives for diversifying (hatchery production, market forces, EVOS?).
+These changes allow us to compare patterns across regions and gear, as well as better understand how external forces have changed incentives for diversifying (hatchery production, market forces, EVOS).
 
@@
 
-Previous analyses have largely focused on patterns of diversification across ecosystems and the benefits to adopt more or less diverse fishing strategies [@kasperski_income_2013; @anderson_benefits].
-Fisheries in Alaska have been studied extensively, because they are one of the most productive seafood regions in the world (both by landed weight, and revenue [@office_of_science_and_technology_fisheries_2016], FAO cite) and they allow individuals to adopt a wide range of fishing strategies.
+Previous analyses have largely focused on patterns of diversification across ecosystems and the benefits to adopt more or less diverse fishing strategies [@kasperski2013; @anderson_benefits].
+Fisheries in Alaska have been studied extensively, because they are one of the most productive seafood regions in the world (both by landed weight, and revenue [@officeofscienceandtechnology2016], FAO cite) and they allow individuals to adopt a wide range of fishing strategies.
 [@anderson_benefits] defined strategies as a permit or collection of permits held by single individuals.
@@
 
-Understanding the effects of species diversification on salmon fishers is particularly important, as the number of salmon specialists (defined as fishers holding a single salmon permit) has increased over the last 30 years as a fraction of the total [@anderson_benefits] (figure XX).
+Understanding the effects of species diversification on salmon fishers is particularly important, as the number of salmon specialists (defined as fishers holding a single salmon permit) has increased over the last 30 years as a fraction of the total [@anderson_benefits].
 For these limited entry salmon fisheries, general trends that have emerged are that more people are participating since the early 2000s (Fig.
@@
 Across different salmon fisheries -- other than a slight negative effect of diversification on variability (similar to [@anderson_benefits]) -- there appears to be little correlation between diversity and revenue (Fig.
-A4).
-In terrestrial agriculture, there are numerous examples where diversification may affect the mean or CV of revenue, but not both [@purdy_financial_1997], and the same result may be true for salmon fishers.
+2).
+In terrestrial agriculture, there are numerous examples where diversification may affect the mean or CV of revenue, but not both [@purdy1997], and the same result may be true for salmon fishers.
 Further, these impacts of diversification are variable over time.
@@
 Bristol Bay set and drift gillnet fisheries).
-For these fisheries, specialization is associated with higher revenue and reduced CV in annual revenue [@anderson_benefits].
+For these fisheries, specialization is associated with higher revenue and reduced CV in year-to-year revenue [@anderson_benefits].
 
 A secondary motivation for focusing on the effects of diversification for salmon fishers is to better understand the long term consequences of hatchery salmon production on the revenue and variability in revenue of individuals.
-The modern era of Alaska’s large-scale hatchery program began in the late 1970s (Stopha 2015, Rich will get a reference), with substantial increases in hatchery production during the 1980s following low returns [@mahnken_historical_1998; @leber_stock_2008-1].
-Many of these hatchery programs increased production to stabilize the variability of salmon returns [@hilborn_review_2001], though these programs may also have negative impacts on wild production [@wertheimer_relationship_2005; @ruggerone_productivity_2015].
-The intra- and inter-specific effects of large scale hatchery programs has been debated [@hilborn_review_2001; @hilborn_hatcheries_1992; @wertheimer_comment:_2001; @heard_hatchery], but the effects of hatcheries have largely been focused on ecological interactions of aggregate yield to the fishery [@leber_stock_2008-1] or economic impacts [@boyce_alaska_1993], rather than to variability experienced by individual fishermen.
+The modern era of Alaska’s large-scale hatchery program began in the late 1970s [@stopha2016], with substantial increases in hatchery production during the 1980s following low returns [@mahnken1998; @leber2008].
+Many of these hatchery programs increased production to stabilize the variability of salmon returns [@hilborn2001a], though these programs may also have negative impacts on wild production [@wertheimer2005; @ruggerone2015].
+The intra- and inter-specific effects of large scale hatchery programs has been debated [@hilborn2001a; @hilborn1992a; @wertheimer2001; @heard_hatchery], but the effects of hatcheries have largely been focused on ecological interactions of aggregate yield to the fishery [@leber2008] or economic impacts [@boyce1993], rather than to variability experienced by individual fishermen.
+Understanding the role salmon have in buffering or stabilizing fishing revenues is critical given the responses of salmon to variable environments [@mueter2002; @schindler2010].
 
-[Jordan: Maybe a paragraph or notes about the importance of the role that salmon are serving in buffering or stabilizing fishing revenues in the state and how subsequently, understanding the responses of salmon to climate dynamics is especially critical (Mueter et al.
-2002, and maybe Schindler et al., 2010)]
-
-In addition to the changing species composition of fish available to be harvested, the benefits (or costs) of diversification are affected by a number of exogenous drivers, including the influx of farmed salmon, consolidation of salmon buyers, marketing, prices, new processing technology [@knapp_trends].
-[Break out as own paragraph+add more examples?] Following price crashes in the 1990s and early 2000s (cite CFEC), many of these factors have been positive for salmon fishermen (increasing trends in prices and catches, Fig.
+In addition to the changing species composition of fish available to be harvested, the benefits or costs of diversification are affected by a number of exogenous drivers, including competition from markets producing farmed salmon, consolidation of salmon buyers, marketing, prices, new processing technology [@knapp_trends].
+Following price declines in the 1990s and early 2000s [@adfg2007], many of these factors have been positive for salmon fishermen (increasing trends in prices and catches, Fig.
 A3, A12).
 In contrast, the late 1980s - early 1990s experienced high abundance and low prices (in part due to collapses of foreign currencies).
-Combined with the increased trend in hatchery production and resulting changes in species composition of adult salmon, these external factors may have also altered incentives for diversification.
-
+Combined with the increased trend in hatchery production and resulting changes in species composition of adult salmon, these external factors may have also altered incentives or opportunities for diversification.
 Tradeoffs between specialization and diversification may also be found in case studies of terrestrial agriculture.
-For small scale farmers, increasing crop diversity may increase revenue and stabilize year-to-year income variability [@di_falco_crop_2003].
+For small scale farmers, increasing crop diversity may increase revenue and stabilize year-to-year income variability [@difalco2003].
 At larger scales or in the developed world, however, the incentives for diversification may be different.
-In the United States and Canada for example, crop diversity has declined since the late-1970s [@aguilar_crop_2015; @bradshaw_farm-level_2004]; some of the factors responsible include changing national policies, technology, and increased demand from other parts of the world (e.g.
+In the United States and Canada for example, crop diversity has declined since the late-1970s [@aguilar2015; @bradshaw2004]; some of the factors responsible include changing national policies, technology, and increased demand from other parts of the world (e.g.
 increased U.S.
 production of soybeans to meet demand from Asia).
-Shifts to from diverse of high value cash-crops to specializing on high-volume low-value crops such as cereals may have analogs for salmon fishermen in low value species with rapid life histories (pink and chum salmon).
+Shifts from more diverse high value crops to specializing on high-volume low-value crops such as cereals may have analogs for salmon fishermen in low value species with rapid life histories (pink and chum salmon).
 
-The objective of our analysis is to better understand the effects of diversification on single-permit salmon holders by extending the methods used by [@anderson_benefits] to include time-varying effects.
-These time-varying effects will allow us to compare temporal trends across regions, identify periods when having diverse catch portfolio was most beneficial, and allow us to evaluate support for effects of longer term drivers in the region.
+The objectives of our analysis are to (1) evaluate long term trends in stability for fleets and individuals to evaluate whether hatchery production has increased stability, and (2) to better understand the effects of diversification on single-permit salmon holders by extending the time series models of individual revenues used by [@anderson_benefits] to include time-varying effects.
+These time-varying effects of diversification will allow us to compare temporal trends across regions and identify periods when having diverse catch portfolio was most beneficial.
 
@@
 
-We obtained fisheries landings and revenue data for all iundividual permit holders in Alaska from 1985-2014 from the Commercial Fisheries Entry Commission (CFEC).
-Each permit holder has a unique identifier that allows us to track quantities including the species caught and revenue received.
-We adjusted for inflation by converting all revenues to 2009 USD.
-Because of our focus on salmon fishers, we were only interested in single permit fisheries, and we adopted a simplified version of the filtering steps described in [@anderson_benefits].
-After aggregating individual trip-level catch and revenue to annual data, we retained permit holders who fished only one of the six salmon permits included in a year (Table 1, Fig.
-A5).
-Like [@anderson_benefits] we also worked with a log-differenced dataset to remove non-stationary individuals; to avoid complications of individuals switching strategies (e.g.
-changing from a seine fishery in Kodiak to a gillnet fishery in Bristol Bay), we only included pairs of years for a given person where they did not change strategies (i.e., didn’t change fisheries or add permits, etc.).
+We obtained fisheries landings and revenue data for all permit holders in Alaska from 1985-2014 from the Commercial Fisheries Entry Commission (CFEC).
+Reported gross earnings (revenue) associated with each fish ticket were adjusted for inflation by converting all revenues too 2009 USD.
+Because our modeling focus involves estimating time varying effects of diversification, we limited our analysis to salmon fisheries with species diversity > 1 (omitting fisheries such as the Bristol Bay gillnet fisheries that specialize on sockeye salmon).
+To better understand the influence of changing hatchery production, we also restricted fisheries to those that have been influenced by hatchery salmon production.
 
@@
 
-### As hatchery programs have increased production, have individual salmon fishermen experienced reduced variability?
+### Has increased hatchery production stabilized Alaskan salmon fisheries?
 
@@
 [@ward_evaluating] summarized the trends for Prince William Sound).
-Release patterns have been dominated by pink and chum salmon ([@stopha2016]) and production has increased since the late 1970s (217 million juvenile pink salmon in PWS in 1985 versus 672 million in 2014).
-Though our analysis with CFEC data doesn't separate hatchery versus wild fish, as a contrast we can make comparisons across regions and species to examine potential effects of hatchery fish on variability.
-For example, we can compare the variation in the largely hatchery pink salmon returns (earnings and weight) from Prince William Sound to the largely wild pink salmon returns from Southeast Alaska.
-Similarly, we can compare the variability in earnings from species dominated by hatchery returns (pink, chum) to those with less hatchery influence (Chinook, sockeye).
-To make these comparisons, we aggregated the earnings and landings by species and permit, then calculated 3-year rolling CVs.
-As a second comparison, we examined whether long-term participants in any of the fisheries had experienced a change in variability.
-Similar to [@kasperski_income_2013] who calculated CVs of vessels before and after significant management events, we identified individuals who fished the same permit for at least three years in the first decade of our data (1985-1995) and at least three years in the last decade of the data (2004-2014).
+Release patterns have been dominated by pink and chum salmon ([@stopha2016], Fig.
+3) and production has increased since the late 1970s (217 million juvenile pink salmon in PWS in 1985 versus 672 million in 2014).
+In addition to examining temporal trends, we can use the spatial contrast between areas to quantify potential changes resulting from increased hatchery production (Prince William Sound, Kodiak, Southeast Alaska).
 
-### Analysis 3: [Core of paper] Does effect of diversification change over time?
+To examine long term stability in salmon fisheries, we summarized the trends in landings (total pounds) and revenue (2009 USD) at two levels.
+First, we calculated annual summaries of catch and revenue at the fishery level (annual values summed across all permit holders) and calculated 3-year running coefficients of variation (CV) by fishery.
+Examining fishery-level summaries alone may be problematic, because increases in individual variability may have a stabilizing effect at larger scales (communities, fisheries).
+As a second metric of variability, we calculated the 3-year running CV for every unique permit holder in each fishery, and averaged those individual CVs across permit holders.
+To contrast patterns across species, we repeated these calculations, computing the annual CV of revenue and prices separately for each species and fishery combination.
 
-[@anderson_benefits] implemented a hierarchical Bayesian model that allowed group level predictors and covariates (diversity of catch, days fished) to simultaneously affect the mean and the variance of revenue for individual fishers.
-This model also included random intercepts by strategy and year; these terms allowed all permit holders in a given year to experience common fluctuations (external drivers including price changes, fluctuations in fish population abundance, or environmental disasters such as EVOS).
+### Have effects of diversification changed over time for salmon fishers?
 
-To ask how benefits of diversification may have changed over time, we extend the model of [@anderson_benefits] to also include time-varying random effects in the slope coefficients of diversification.
+Because we were only interested in single permit fisheries, we applied a simplified version of the CFEC data filtering steps described in [@anderson_benefits].
+Individual trip-level catch and revenue statistics were aggregated to annual data, and we retained individual fishers who only participated in one fishery in a year (Table 1).
+Annual revenues were log-transformed to remove non-stationary individuals; to avoid complications in modeling individual switching permits, we only included pairs of years for a given person where they did not change permits (e.g.
+didn’t change fisheries between two years).
+[@anderson_benefits] implemented a hierarchical Bayesian time series model that incorporated group level predictors and individual covariates (diversity of catch, days fished) to simultaneously affect the mean and the variance of revenue for individual fishers.
+This model also included random intercepts by fishery and year, allowing all permit holders in a given year to experience common fluctuations (external drivers including price changes, fluctuations in fish population abundance, or environmental disasters such as EVOS).
+
+To ask how benefits of diversification may have changed over time for salmon fisheries in Alaska, we extend the model of [@anderson_benefits] to include time-varying random effects in the slope coefficients of diversification.
 We define the variables $\Delta { S }_{ t }=\log\left( { S }_{ t }/{ S }_{ t-1 } \right)$ to represent the percent change in species diversity of the catch between years and $\Delta { D }_{ t }=\log\left( { D }_{ t }/{ D }_{ t-1 } \right)$ to represent the percent change in fishing effort (days fished) between years.
@@
 Finally, the term $\log\left( { R }_{ i,t-1 } \right)$ is included as an offset [@anderson_benefits].
+
 Like [@anderson_benefits], we also included all of the above covariates in the variance model.
@@
 
-[Rich's note: This and delta D might be tricky for pink salmon stocks, which alternate between strong even or odd years (depends on the area) and may dictate fishery participation.
-For these, could we, should we, use log(St/St-2)??]
-
-All done in Stan [@stan_development_team_stan_2015] using the NUTS algorithm [@hoffman_no-u-turn_2014].
+All estimation was done in a Bayesian framework using R [@rcoreteam2016] and the package 'rstan' implementing using the NUTS algorithm [@hoffman2014].
 
 Specific things that could be looked at: 
-- EVOS (closure for Kodiak)  
-- PWS pink salmon crash in 1992/1993: Pacific Salmon & their Ecosystems: Status and Future Options
-- sockeye marketing 1983: [Jordan: So the Copper River salmon marketing ramped up in about 1983.
-I wonder if we could see a steady increase in value or movement towards sockeye in the early years of the data set]
+[Rich: maybe look at PWS pink salmon crash in 1992/1993: Pacific Salmon & their Ecosystems: Status and Future Options.
+EW: pink harvest generally falls from 89/90 to early 90s.
+It's a little of a chicken and egg problem too - because by the early 90s hatchery production was 75% of the total pink catch and also declined a little bit.
+From the CV perspective, the pink crash may be synchronizing -- stability was increased because there were several back to back bad years]
+[Jordan: So the Copper River salmon marketing ramped up in about 1983.
+I wonder if we could see a steady increase in value or movement towards sockeye in the early years of the data set.
+EW: Fig 3 largely shows the pattern in those early years -- it ramps up in PWS, but is less steep than for say chum]
 
-Does diversification exist within permits and if so, do we see similar advantages of diversification on revenue + CV?
-
-PWS chum and sockeye (increasing hatchery % over time) vs. Bristol Bay (all wild).
-
-CV of variability for high value species (sockeye, Chinook) vs.
-low value species (pink and chum).
-Coho fall somewhere in the middle.
-In other words, how have choices of species to produce by hatcheries affected the CV of revenue for individual permits?
-% hatchery fish of the harvest could be used as a continuous variable in the model.
-
-Similar to above, do life-history characteristics influence the variability of revenue?
-For example, sockeye and chum have substantially more complicated life-history traits (1-5 ocean years) than pink salmon (1 ocean year only).
-For high-volume hatcheries such as pink salmon: do annual returns affect price and revenue the following year more than for high-value fisheries?
 ## Results
 
-Our summaries of changing variability (aggregate for each fishery, across individuals, or within individuals) was not consistent with the hypothesis that variability in revenue or landings has decreased over time (Fig.
+Our summaries of changing variability (aggregate for each fishery, or averaged across individuals) are not consistent with the hypothesis that variability in revenue or landings has decreased over time (Fig.
 A9-A10).
@@
 Diversification can be profitable, but in the short term risky or expensive.
-Bigger than cost seems to be the risk [@karlan_agricultural_2012]
+Bigger than cost seems to be the risk [@karlan2012]
 
@@
 This gets at more of the social impacts of changing opportunities, but long term trends in rising income variability have been documented in the US (Hacker 2006).
-Understanding trends in revenue of individuals whose incomes are dependent on natural resources are interesting because they tend to be more variable than that of the larger population [@mishra_income_2002].
-Are there specific attributes that correlate with rising variability (for example the CBO has suggested variability in year to year income is negatively correlated with income [@cbo_recent_2008])
-
-### Robustness of strategies to changing environments [stash in appendix?]
-
-One of the challenges in modeling the effects of species diversification for individual fishers is that between two years, both the species composition of the catch and the species composition of the fish available to be caught change (making it hard to disentangle the selectivity of fishermen versus portfolio of fish available).
-To isolate the effects of a changing environment from changing strategies, we used a subset of the data from fishery permits where discrete strategies were emergent.
-For details see Appendix.
-
-Using purse seine permit holders in Prince William Sound as an example, we first calculated the species composition (weighted by revenue) for each active permit holder in each year.
-For each pair of years that an individual fished, we calculated the squared Euclidian distance between the species compositions of those years, $$\Delta { p }_{ ind,t }=\sum _{ sp=1 }^{ 5 }{ { \left| { p }_{ ind,sp,t }-{ p }_{ ind,sp,t-1 } \right|  }^{ 2 } } $$.
-
-Larger values of $\Delta { p }_{ ind,t }$ indicate greater year-to-year discrepancy in species composition.
-To focus on individuals who had more constant strategies in adjacent years, we ommitted data corresponding to the upper third of this distribution (retaining $\Delta { p }_{ ind,t }$ < 0.334).
-Next, we used the species composition from each individual and pair of years to cluster individuals into strategies.
-Clustering was performed using the partitioning around medoids (pam) algorithm in R with an estimated number of clusters (using 'pamk' in fpc library), and selected 2 groups of individual strategies: pink salmon specialization (medoid proportion of pink salmon = 0.95), and a strategy focused on pink salmon and chum (medoid proportions of 0.64 and 0.31 respectively).
-For each group, we then calculated the mean and CV of revenue; we calculated the log difference of revenue for each individual, $\Delta { R }_{ i }=\log\left( { R }_{ i,t+1 }/{ R }_{ i,t } \right)$ and then calculated the CV as the standard deviation over people, $sd(\Delta { R }_{ i=1:n })$.
+Understanding trends in revenue of individuals whose incomes are dependent on natural resources are interesting because they tend to be more variable than that of the larger population [@mishra2002].
+Are there specific attributes that correlate with rising variability (for example the CBO has suggested variability in year to year income is negatively correlated with income [@cbo2008])
 
@@
 
-```{r, echo=FALSE, fig.pos="placeHere", fig.cap="CV and mean annual revenue by salmon permit, 1975-2014.
-Average participation (unique fishers) given by circle size, and effective species diversity is shown with color shading.", fig.height=3.2, eval=FALSE}
+```{r fig1, echo=FALSE, fig.pos="placeHere", fig.cap="Trends in participation, landed catch, mean annual revenue, and effective species diversity for eight selected salmon fisheries.
+[EW notes: weight per person increasing more than revenue.
+Species diversity for most fisheries declining]"}
+# trends in participation, 
+cfec_all = readRDS("../../portfolio/data-generated/salmon_data_for_paper.rds")
+people = group_by(cfec_all, p_fshy, year) %>% 
+  summarize(n = length(unique(p_holder))) %>% 
+  filter(substr(p_fshy,1,1)=="S")
+people$p_fshy = paste0(substr(people$p_fshy,1,1), substr(people$p_fshy,3,5))
+names(people)[which(names(people)=="p_fshy")]="Permit"
+
+g1 = ggplot(people[people$Permit%in%c("S01K","S01M","S01E","S01A","S03A","S03E","S03T","S05B"),], aes(year, log10(n), color = Permit)) + geom_line() + ylab("Log10 unique people") + xlab("Year") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
+
+# trends in landed weight - metric tons
+weight = group_by(cfec_all, p_fshy, year) %>% 
+  summarize(w = sum(g_pounds,na.rm=T) * 0.000453592) %>% 
+  filter(substr(p_fshy,1,1)=="S")
+weight$p_fshy = paste0(substr(weight$p_fshy,1,1), substr(weight$p_fshy,3,5))
+names(weight)[which(names(weight)=="p_fshy")]="Permit"
+
+weight = left_join(weight, people)
+g2 = ggplot(weight[weight$Permit%in%c("S01K","S01M","S01E","S01A","S03A","S03E","S03T","S05B"),], aes(year, log10(w/n), color = Permit)) + geom_line() + ylab("Log10 tons per person") + xlab("Year") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
+
+# trends in revenue
+dat.annual = readRDS(file="../../portfolio/data-generated/salmon-annual-for-modeling.rds")
+dat.annual = group_by(dat.annual[which(dat.annual$strategy_permit %in% c("S01A", "S01E", "S01K", "S01M", "S03A", "S03E", "S03T", "S05B")),], year, strategy_permit) %>% 
+  summarize(meanrev = mean(revenue), meandiv = mean(specDiv))
+names(dat.annual)[which(names(dat.annual)=="strategy_permit")]="Permit"
+dat.annual = left_join(dat.annual, weight)
+
+g3 = ggplot(dat.annual[dat.annual$Permit%in%c("S01K","S01M","S01E","S01A","S03A","S03E","S03T","S05B"),], aes(year, log10(meanrev), color = Permit)) + geom_line() + ylab("Log10 revenue per person") + xlab("Year") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
+
+g4 = ggplot(dat.annual[dat.annual$Permit%in%c("S01K","S01M","S01E","S01A","S03A","S03E","S03T","S05B"),], aes(year, meandiv, color = Permit)) + geom_line() + ylab("Species diversity") + xlab("Year") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
+
+gridExtra::grid.arrange(g1,g2,g3, g4, ncol=2)
+
+```
+
+```{r fig2, echo=FALSE, fig.pos="placeHere", fig.cap="CV and mean annual revenue by salmon permit, 1975-2014.
+Average participation (unique fishers) given by circle size, and effective species diversity is shown with color shading.
+[EW notes: pretty cool clustering by permits here - highlighted by Ole.
+The modeling in the rest of the paper is largely comparing the blue seine fisheries to the orange gillnets in PWS / SEAK]", warning=FALSE, message=FALSE}
 dat.annual = readRDS(file="../../portfolio/data-generated/salmon-annual-for-modeling.rds")
 
-cfec_all = readRDS("../../portfolio/data-generated/salmon_data_for_paper.rds")
+#cfec_all = readRDS("../../portfolio/data-generated/salmon_data_for_paper.rds")
 people = group_by(cfec_all, p_fshy, year) %>% 
@@
 
-dat.annual = filter(dat.annual, npermit==1 & substr(strategy_permit,1,1)=="S") %>%
+records_s03h = which((dat.annual$strategy_permit == "S03H" & dat.annual$year %in% c(1989,1990) | dat.annual$strategy_permit == "S04X"))
+
+dat.annual = filter(dat.annual[-records_s03h,], npermit==1 & substr(strategy_permit,1,1)=="S") %>%
   group_by(strategy_permit) %>% 
@@
 
-g1 = ggplot(dat.annual, aes(x=log(mean_rev), y=log(cv_rev), label = dat.annual$strategy_permit)) + geom_point(aes(col=gear,size=diversity,alpha=0.7)) + scale_size(range = c(0, 20)) + geom_text_repel() + ylab("CV annual revenue") + xlab("Mean revenue (dollars)") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
+g1 = ggplot(dat.annual, aes(x=mean_rev, y=cv_rev, label = dat.annual$strategy_permit)) + geom_point(alpha=0.7, aes(col=gear,size=diversity)) + scale_size(range = c(0, 20)) + geom_text_repel() + ylab("CV annual revenue") + xlab("Mean revenue (dollars)") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
 panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
 print(g1)
-pdf("Figure_01.pdf")
-print(g1)
-dev.off()
+#pdf("Figure_01.pdf")
+#print(g1)
+#dev.off()
 ```
 
-```{r, echo=FALSE, fig.pos="placeHere", fig.cap="Harvest of hatchery produced salmon through time and by area.", fig.height=3.2}
+```{r fig3, echo=FALSE, fig.pos="placeHere", fig.cap="Harvest of hatchery produced salmon through time and by area.
+Shown are the numbers of hatchery fish caught by fisheries, as well as the contribution of hatchery fish to total catch, by species.
+[EW notes: purpose of this plot is to show (1) increasing production in top row and (2) increasing contribution of hatchery fish to commercial catches in bottom row]"}
 
@@
   geom_area(aes(fill = species), position = "stack") +
-  xlab("Year") + ylab("Harvest of hatchery produced salmon (millions)") +
+  xlab("Year") + ylab("Hatchery harvest (millions)") +
   scale_fill_manual(values = c('Chinook'=cbPalette[1], 'Chum'=cbPalette[2], 
@@
   geom_line() + facet_wrap(~Region) + 
-  xlab("Year") + ylab("Hatchery contribution to catch (%)") + 
+  xlab("Year") + ylab("% of total catch") + 
   scale_color_manual(values = c('Chinook'=cbPalette[1], 'Chum'=cbPalette[2], 
@@
 gridExtra::grid.arrange(g1, g2)
+```
+
+```{r fig4, fig.pos="placeHere", fig.cap="CV of annual revenue and prices, by permit and fishery.
+Variability in year to year revenue is explained more by permits than species, while variability in prices is affected more by species.
+", echo=FALSE}
+#cfec_all = readRDS("../../portfolio/data-generated/salmon_data_for_paper.rds")
+cfec_all$year = as.numeric(substr(cfec_all$landdate, 1, 4))
+
+cfec_salmon = cfec_all[which(cfec_all$p_fshy%in% c("S 01A", "S 01E", "S 01K","S 01M","S 03A", "S 03E")),]
+cfec_salmon = filter(cfec_salmon, year != 0)
+
+s0 = group_by(cfec_salmon[cfec_salmon$spec %in% c("CHNK","COHO","PINK","SOCK","CHUM"),], p_fshy, spec, year) %>% 
+  summarize(totrev = sum(g_earn), totlbs = sum(g_pounds), totprice = mean(g_price)) %>% 
+  arrange(p_fshy, spec, year) %>% 
+  group_by(p_fshy, spec) %>% 
+  summarize(srev = sd(diff(log(totrev))), slbs = sd(diff(log(totlbs))), sprice = sd(diff(log(totprice))))
+names(s0)[which(names(s0)=="p_fshy")] = "Permit"
+names(s0)[which(names(s0)=="spec")] = "Species"
+ggplot(s0, aes(srev,sprice,col=Species,shape = Permit)) + geom_point(size=3,alpha=0.8) + ylab("CV of year-to-year price per pound") + xlab("CV of year-to-year revenue") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
+
+s0 = group_by(cfec_salmon[cfec_salmon$spec %in% c("CHNK","COHO","PINK","SOCK","CHUM"),], p_fshy, spec, year) %>% 
+  summarize(totrev = sum(g_earn), totlbs = sum(g_pounds), totprice = mean(g_price))
+write.table(s0, "../../salmon/data-generated/species_level_data.csv", row.names=FALSE, col.names=TRUE)
+```
+
+```{r fig5, fig.pos="placeHere", fig.cap="Three-year running CVs of revenue and landings, calculated at the fleetwide level (summed over all people) and level of individual permit holders (CV averaged across permit holders).", echo=FALSE}
+# we want to plot 4 summaries, by fishery:
+# 1) variability in year-to-year total revenue
+# 2) variability in year-to-year total landings
+# 3) variability in year-to-year earnings of individuals
+# 4) variability in year-to-year landings of individuals
+
+s1 = group_by(cfec_salmon, p_fshy, year) %>%
+  summarize(landings = sum(g_pounds), revenue = sum(g_earn)) %>% 
+  group_by(p_fshy) %>% 
+  mutate(cvlandings = c(NA,rollapply(landings,width=3,sd),NA)/c(NA,rollmean(landings,k=3),NA), cvrevenue = c(NA,rollapply(revenue,width=3,sd),NA)/c(NA,rollmean(revenue,k=3),NA))
+names(s1)[which(names(s1)=="p_fshy")] = "Permit"
+s1$Permit = paste0(substr(s1$Permit,1,1), substr(s1$Permit,3,5))
+write.table(s1, "../../salmon/data-generated/fishery_level_cvs.csv", row.names=FALSE, col.names=TRUE)
+
+g1 = ggplot(s1, aes(year, cvlandings, col=Permit)) + geom_line() + xlab("Year") + ylab("CV (fleetwide landings)") + xlim(1985,2014) + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
+
+g2 = ggplot(s1, aes(year, cvrevenue, col=Permit)) + geom_line() + xlab("Year") + ylab("CV (fleetwide revenue)") + xlim(1985,2014) + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
+
+# CV of individual landings and revenue
+s2 = group_by(cfec_salmon, p_fshy, year, p_holder) %>%
+  summarize(landings = sum(g_pounds), revenue = sum(g_earn)) %>% 
+  arrange(p_fshy, p_holder, year)
+
+# fill in missing years with NAs
+people_fisheries = unique(paste0(s2$p_holder, ":", s2$p_fshy))
+df_nas = expand.grid("year"=sort(unique(s2$year)), "p_holder" = sort(unique(s2$p_holder)), "p_fshy" = unique(s2$p_fshy)) 
+s2 = left_join(df_nas,s2)
+
+# remove people with less than 3 years of fishing
+s2 = group_by(s2, p_holder) %>% 
+  mutate(nYear = length(unique(year))) %>% 
+  filter(nYear >= 3) %>% 
+  select(-nYear)
+
+s2 = group_by(s2, p_holder, p_fshy) %>% 
+  mutate(cvlandings = c(NA,rollapply(landings,width=3,sd),NA)/c(NA,rollmean(landings,k=3),NA), 
+    cvrevenue = c(NA,rollapply(revenue,width=3,sd),NA)/c(NA,rollmean(revenue,k=3),NA)) %>% 
+  group_by(p_fshy, year) %>% 
+      summarize(cvlandings=mean(cvlandings,na.rm=T), 
+        cvrevenue = mean(cvrevenue,na.rm=T))
+
+names(s2)[which(names(s2)=="p_fshy")] = "Permit"
+s2$Permit = paste0(substr(s2$Permit,1,1), substr(s2$Permit,3,5))
+write.table(s2, "../../salmon/data-generated/avg_individual_level_cvs.csv", row.names=FALSE, col.names=TRUE)
+
+g3 = ggplot(s2, aes(year, cvlandings, col=Permit)) + geom_line() + xlab("Year") + ylab("CV (individual landings)") + xlim(1985,2014) + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
+
+g4 = ggplot(s2, aes(year, cvrevenue, col=Permit)) + geom_line() + xlab("Year") + ylab("CV (individual revenue)") + theme_bw() + xlim(1985,2014) + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
+
+gridExtra::grid.arrange(g1,g2,g3,g4,ncol=2)
 
@@
 
-```{r, echo=FALSE, fig.pos="placeHere", fig.cap="Time varying species composition and diversity, by permit.
-Species are colored as follows: green (Chinook), purple (chum), silver  (coho), pink (pink), red (sockeye).", fig.height=3.2}
+```{r fig6, fig.pos="placeHere", fig.cap="Time varying effects of adding 1 species on revenue (measured in ln dollars) for hypothetical individuals with species diversity ~ 1 (in other words, the effect of switching from effective catch diversity = 1 to 2).
+This shows declining benefits of diversification for most fisheries, S01M.", echo=FALSE}
+# make yearpermit df
+diffdat = readRDS(file="../../salmon/data-generated/salmon.rds")
+years = seq(min(diffdat$year), max(diffdat$year))
+permits = levels(as.factor(diffdat$permit))
+# note that we have to use factor() here instead of unique() because
+# that's the same as the ordering in the estimation code
+yearpermit = data.frame("yrpermit"=levels(as.factor(diffdat$year_permit)))
+yearpermit$year = as.numeric(substr(yearpermit$yrpermit,1,4))
+yearpermit$permit = substr(yearpermit$yrpermit,6,9)
+
+# grab original data to look at permit names
+diffdat = readRDS(file="../../portfolio/data-generated/salmon-diff-for-modeling.rds")
+years = seq(min(diffdat$year), max(diffdat$year)) 
+#permits = levels(as.factor(diffdat$permit))
+
+yearpermitold=yearpermit
+load("../../salmon/analysis/model-linear.Rdata")
+yearpermit=yearpermitold
+
+est = extract(mod, permute="TRUE")
+
+# start with making time series of estimates on mean
+# challenge is that they're time varying quadratic functions
+# do facet_wrap of estimates by year
+
+# effect of adding more 1 species from 1. specdiv.prev = 1
+
+specdiv.prev = seq(1, 5, by=0.1)
+specdiv = specdiv.prev + 1 # always add 1 more in these comparisons
+x1 = log(specdiv/specdiv.prev)
+x2 = log(specdiv/specdiv.prev) * specdiv.prev
+x3 = log(specdiv/specdiv.prev) * specdiv.prev^2
+
+df = expand.grid("year"=1:length(years), "diversity" = seq(1,length(x1)), "permit"=1:6)
+df$permits = permits[df$permit]
+df$years = years[df$year]
+df$x = specdiv.prev[df$diversity]
+df$pred = 0
+df$low = 0
+df$high = 0
+b0_indx = match(paste(as.character(df$years), as.character(df$permits)), as.character(yearpermit$yrpermit)) 
+for(i in 1:nrow(df)) {
+  #tmp = est$b0_str_yr[,b0_indx[i]] + 
+  tmp = 
+    x1[df$diversity[i]] * (est$b_1[,df$year[i], df$permit[i]])# +
+    #x2[df$diversity[i]] * (est$b_2[,df$year[i], df$permit[i]]+est$b2_str_yr_mu)# +
+    #x3[df$diversity[i]] * (est$b_3[,df$year[i], df$permit[i]]+est$b3_str_yr_mu)
+  df$pred[i] = mean(tmp)
+  df$low[i] = quantile(tmp,0.025)
+  df$high[i] = quantile(tmp,0.975)
+}
+
+ggplot(df[df$diversity==1,], aes(x=years,y=pred,group=permits)) +
+  geom_line() +
+  geom_ribbon(aes(ymin=low,ymax=high), alpha=0.4) +
+  facet_wrap(~permits) + xlab("Year") + ylab("Predicted benefit of adding +1 spp (ln $)") + geom_hline(yintercept=0, color="red")
+
+```
+
+# References
+
+```{r figa1, echo=FALSE, fig.pos="placeHere", fig.cap="Time varying species composition and diversity, by permit.
+Species are colored as follows: green (Chinook), purple (chum), silver  (coho), pink (pink), red (sockeye)."}
 #cfec <- feather::read_feather("../../portfolio/data-generated/salmon.feather")
 #saveRDS(cfec, "portfolio/data-generated/salmon_data_for_paper.rds")
-cfec = readRDS("../../portfolio/data-generated/salmon_data_for_paper.rds")
-cfec$year = as.numeric(substr(cfec$landdate, 1, 4))
+#cfec_all = readRDS("../../portfolio/data-generated/salmon_data_for_paper.rds")
+#cfec_all$year = as.numeric(substr(cfec_all$landdate, 1, 4))
 
@@
 
-g1 = group_by(cfec[cfec$p_fshy=="S 01A" & cfec$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
+g1 = group_by(cfec_all[cfec_all$p_fshy=="S 01A" & cfec_all$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
   summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
@@
 
-g2 = group_by(cfec[cfec$p_fshy=="S 01E" & cfec$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
+g2 = group_by(cfec_all[cfec_all$p_fshy=="S 01E" & cfec_all$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
   summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
@@
 
-g3 = group_by(cfec[cfec$p_fshy=="S 01K" & cfec$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
+g3 = group_by(cfec_all[cfec_all$p_fshy=="S 01K" & cfec_all$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
   summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
@@
 
-g4 = group_by(cfec[cfec$p_fshy=="S 01M" & cfec$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
+g4 = group_by(cfec_all[cfec_all$p_fshy=="S 01M" & cfec_all$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
   summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
@@
 
-g5 = group_by(cfec[cfec$p_fshy=="S 03A" & cfec$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
+g5 = group_by(cfec_all[cfec_all$p_fshy=="S 03A" & cfec_all$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
   summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
@@
 
-g6 = group_by(cfec[cfec$p_fshy=="S 03E" & cfec$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK") & cfec$year > 0,], year, spec, p_fshy) %>%
+g6 = group_by(cfec_all[cfec_all$p_fshy=="S 03E" & cfec_all$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK") & cfec_all$year > 0,], year, spec, p_fshy) %>%
   summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
@@
 
-```{r, echo=FALSE, fig.pos="placeHere", fig.cap="Time varying species composition and diversity, by permit.", fig.height=3.2} 
+```{r figa2, echo=FALSE, fig.pos="placeHere", fig.cap="Time varying species composition and diversity, by permit.", fig.height=3.2} 
 gridExtra::grid.arrange(g4, g5, g6, h4, h5, h6, nrow=2)
@@
 
-\break
-
-```{r, eval=FALSE,echo=FALSE}
-load("../../salmon/analysis/cv-model-simple.Rdata")
-# groups here are 80% pink vs 97% pink
-
-yrs = 1986:2014
-
-hist(extract(mod)[["g_offset"]], 40, col="grey", xlab="Effect of specializing on pink (sd)")
-
-xx = exp(extract(mod)[["sd_yr"]])
-df.gen = data.frame("year"=yrs, "mean" = apply(xx, 2, mean), "low" = apply(xx, 2, quantile, 0.025),
-  "hi" = apply(xx, 2, quantile, 0.975), "strategy"="pink:chum")
-
-offset = matrix(extract(mod)[["g_offset"]],ncol=1) %*% matrix(rep(1,length(yrs)), nrow=1)
-xx = exp(extract(mod)[["sd_yr"]] + offset)
-df.spec = data.frame("year"=yrs, "mean" = apply(xx, 2, mean), "low" = apply(xx, 2, quantile, 0.025),
-  "hi" = apply(xx, 2, quantile, 0.975), "strategy"="pink")
-
-ggplot(data=rbind(df.gen,df.spec), aes(year, mean, group=strategy, fill=strategy)) + geom_ribbon(aes(ymin=low,ymax=hi), alpha=0.5) + geom_hline(yintercept=0, color="red")
-
-```
-
-```{r, echo=FALSE, fig.pos="placeHere", fig.cap="Time varying intercepts on estimated mean revenue, by permit.
-These are interpreted as the mean change in revenue for each permit holder in each year (positive equates to increasing revenue, etc), after accounting for effects of catch diversification and days fished."}
-# grab original data to look at permit names
-diffdat = readRDS(file="../../salmon/data-generated/salmon.rds")
-
-years = seq(min(diffdat$year), max(diffdat$year))
-permits = levels(as.factor(diffdat$permit))
-
-# note that we have to use factor() here instead of unique() because
-# that's the same as the ordering in the estimation code
-yearpermit = data.frame("yrpermit"=levels(as.factor(diffdat$year_permit)))
-yearpermit$year = as.numeric(substr(yearpermit$yrpermit,1,4))
-yearpermit$permit = substr(yearpermit$yrpermit,6,9)
-
-yearpermitold=yearpermit
-load("../../salmon/analysis/model-linear.Rdata")
-yearpermit=yearpermitold
-est = extract(mod, permute="TRUE")
-
-yearpermit$b0 = apply(est$b0_str_yr,2,median)
-yearpermit$b0_low = apply(est$b0_str_yr,2,quantile,0.025)
-yearpermit$b0_hi = apply(est$b0_str_yr,2,quantile,0.975)
-
-ggplot(data=yearpermit) + 
-  geom_ribbon(aes(x=year,ymin=b0_low,ymax=b0_hi),alpha=0.3) +
-  geom_line(aes(x=year,y=b0)) + 
-  facet_wrap(~permit) + xlab("Year") + 
-  ylab("Intercepts on mean") + 
-  geom_hline(yintercept=0, color="red")
-```
-
-\break
-
-```{r, echo=FALSE}
-# fig.pos="placeHere", fig.cap="Time varying effects of adding 1 species on revenue (measured in ln dollars) for purse seine permit in Prince William Sound (S01E).
-This figure illustrates two things: (1) the benefit of adding more species diminishes as a function of past diversity, and (2) the economic benefit to adding more species has been reduced through time."
-# grab original data to look at permit names
-diffdat = readRDS(file="../../salmon/data-generated/salmon.rds")
-years = seq(min(diffdat$year), max(diffdat$year))
-#permits = levels(as.factor(diffdat$permit))
-
-load("../../salmon/analysis/model-linear.Rdata")
-yearpermit=yearpermitold
-
-est = extract(mod, permute="TRUE")
-
-# start with making time series of estimates on mean
-# challenge is that they're time varying quadratic functions
-# do facet_wrap of estimates by year
-
-# effect of adding more 1 species from 1. specdiv.prev = 1
-
-specdiv.prev = seq(1, 5, by=0.1)
-specdiv = specdiv.prev + 1 # always add 1 more in these comparisons
-x1 = log(specdiv/specdiv.prev)
-x2 = log(specdiv/specdiv.prev) * specdiv.prev
-
-df = expand.grid("year"=1:29, "diversity" = seq(1,length(x1)), "permit"=1:6)
-df$permits = permits[df$permit]
-df$years = years[df$year]
-df$x = specdiv.prev[df$diversity]
-df$pred = 0
-df$low = 0
-df$high = 0
-b0_indx = match(paste(as.character(df$years), as.character(df$permits)), as.character(yearpermit$yrpermit))
-
-for(i in 1:nrow(df)) {
-  tmp = x1[df$diversity[i]] * (est$b_1[,df$year[i], df$permit[i]])# + x2[df$diversity[i]] * (est$b_2[,df$year[i], df$permit[i]]+est$b2_str_yr_mu)
-  df$pred[i] = median(tmp)
-  df$low[i] = quantile(tmp,0.025)
-  df$high[i] = quantile(tmp,0.975)
-}
-df_s01e_b1 = df
-#ggplot(df[df$permits=="S01E",], aes(x=x,y=pred,group=years)) +
-#  #geom_ribbon(aes(ymin=low,ymax=high), alpha=0.4) +
-#  geom_line() +
-#  facet_wrap(~years) + xlab("Previous species diversity") + ylab("Predicted #benefit of adding +1 spp (ln $)")
-
-```
-
-\break
-
-```{r, fig.pos="placeHere", fig.cap="Time varying effects of adding 1 species on revenue (measured in ln dollars) for hypothetical individuals with species diversity ~ 1.
-This shows declining benefits of diversification for most fisheries, S01M.", echo=FALSE}
-# grab original data to look at permit names
-diffdat = readRDS(file="../../portfolio/data-generated/salmon-diff-for-modeling.rds")
-years = seq(min(diffdat$year), max(diffdat$year))
-#permits = levels(as.factor(diffdat$permit))
-
-yearpermitold=yearpermit
-load("../../salmon/analysis/model-linear.Rdata")
-yearpermit=yearpermitold
-
-est = extract(mod, permute="TRUE")
-
-# start with making time series of estimates on mean
-# challenge is that they're time varying quadratic functions
-# do facet_wrap of estimates by year
-
-# effect of adding more 1 species from 1. specdiv.prev = 1
-
-specdiv.prev = seq(1, 5, by=0.1)
-specdiv = specdiv.prev + 1 # always add 1 more in these comparisons
-x1 = log(specdiv/specdiv.prev)
-x2 = log(specdiv/specdiv.prev) * specdiv.prev
-x3 = log(specdiv/specdiv.prev) * specdiv.prev^2
-
-df = expand.grid("year"=1:length(years), "diversity" = seq(1,length(x1)), "permit"=1:6)
-df$permits = permits[df$permit]
-df$years = years[df$year]
-df$x = specdiv.prev[df$diversity]
-df$pred = 0
-df$low = 0
-df$high = 0
-b0_indx = match(paste(as.character(df$years), as.character(df$permits)), as.character(yearpermit$yrpermit))
-
-for(i in 1:nrow(df)) {
-  #tmp = est$b0_str_yr[,b0_indx[i]] + 
-  tmp = 
-    x1[df$diversity[i]] * (est$b_1[,df$year[i], df$permit[i]])# +
-    #x2[df$diversity[i]] * (est$b_2[,df$year[i], df$permit[i]]+est$b2_str_yr_mu)# +
-    #x3[df$diversity[i]] * (est$b_3[,df$year[i], df$permit[i]]+est$b3_str_yr_mu)
-  df$pred[i] = mean(tmp)
-  df$low[i] = quantile(tmp,0.025)
-  df$high[i] = quantile(tmp,0.975)
-}
-
-ggplot(df[df$diversity==1,], aes(x=years,y=pred,group=permits)) +
-  geom_line() +
-  geom_ribbon(aes(ymin=low,ymax=high), alpha=0.4) +
-  facet_wrap(~permits) + xlab("Year") + ylab("Predicted benefit of adding +1 spp (ln $)") + geom_hline(yintercept=0, color="red")
-
-```
-
-\break
-
-```{r, fig.pos="placeHere", fig.cap="Time varying effects of adding 1 species on log standard deviation of revenue for hypothetical individuals with species diversity ~ 1.", echo=FALSE}
+```{r figa3, fig.pos="placeHere", fig.cap="Time varying effects of adding 1 species on log standard deviation of revenue for hypothetical individuals with species diversity ~ 1.", echo=FALSE}
 # grab original data to look at permit names
@@
 
-```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "As model validation, we can compare the estimated intercepts for PWS S01E (purse seine) representing average year to year change in revenue versus the change year over year in ln(pink salmon returns).
+```{r figa4, echo=FALSE, fig.pos="placeHere", fig.cap="Time varying intercepts on estimated mean revenue, by permit.
+These are interpreted as the mean change in revenue for each permit holder in each year (positive equates to increasing revenue, etc), after accounting for effects of catch diversification and days fished."}
+# grab original data to look at permit names
+diffdat = readRDS(file="../../salmon/data-generated/salmon.rds")
+
+years = seq(min(diffdat$year), max(diffdat$year))
+permits = levels(as.factor(diffdat$permit))
+
+# note that we have to use factor() here instead of unique() because
+# that's the same as the ordering in the estimation code
+yearpermit = data.frame("yrpermit"=levels(as.factor(diffdat$year_permit)))
+yearpermit$year = as.numeric(substr(yearpermit$yrpermit,1,4))
+yearpermit$permit = substr(yearpermit$yrpermit,6,9)
+
+yearpermitold=yearpermit
+load("../../salmon/analysis/model-linear.Rdata")
+yearpermit=yearpermitold
+est = extract(mod, permute="TRUE")
+
+yearpermit$b0 = apply(est$b0_str_yr,2,median)
+yearpermit$b0_low = apply(est$b0_str_yr,2,quantile,0.025)
+yearpermit$b0_hi = apply(est$b0_str_yr,2,quantile,0.975)
+
+ggplot(data=yearpermit) + 
+  geom_ribbon(aes(x=year,ymin=b0_low,ymax=b0_hi),alpha=0.3) +
+  geom_line(aes(x=year,y=b0)) + 
+  facet_wrap(~permit) + xlab("Year") + 
+  ylab("Intercepts on mean") + 
+  geom_hline(yintercept=0, color="red")
+```
+
+```{r figa5, eval=FALSE,echo=FALSE}
+load("../../salmon/analysis/cv-model-simple.Rdata")
+# groups here are 80% pink vs 97% pink
+
+yrs = 1986:2014
+
+hist(extract(mod)[["g_offset"]], 40, col="grey", xlab="Effect of specializing on pink (sd)")
+
+xx = exp(extract(mod)[["sd_yr"]])
+df.gen = data.frame("year"=yrs, "mean" = apply(xx, 2, mean), "low" = apply(xx, 2, quantile, 0.025),
+  "hi" = apply(xx, 2, quantile, 0.975), "strategy"="pink:chum")
+
+offset = matrix(extract(mod)[["g_offset"]],ncol=1) %*% matrix(rep(1,length(yrs)), nrow=1)
+xx = exp(extract(mod)[["sd_yr"]] + offset)
+df.spec = data.frame("year"=yrs, "mean" = apply(xx, 2, mean), "low" = apply(xx, 2, quantile, 0.025),
+  "hi" = apply(xx, 2, quantile, 0.975), "strategy"="pink")
+
+ggplot(data=rbind(df.gen,df.spec), aes(year, mean, group=strategy, fill=strategy)) + geom_ribbon(aes(ymin=low,ymax=hi), alpha=0.5) + geom_hline(yintercept=0, color="red")
+
+```
+
+```{r figa6, echo=FALSE, eval=FALSE}
+# fig.pos="placeHere", fig.cap="Time varying effects of adding 1 species on revenue (measured in ln dollars) for purse seine permit in Prince William Sound (S01E).
+This figure illustrates two things: (1) the benefit of adding more species diminishes as a function of past diversity, and (2) the economic benefit to adding more species has been reduced through time."
+# grab original data to look at permit names
+diffdat = readRDS(file="../../salmon/data-generated/salmon.rds")
+years = seq(min(diffdat$year), max(diffdat$year))
+#permits = levels(as.factor(diffdat$permit))
+
+load("../../salmon/analysis/model-linear.Rdata")
+yearpermit=yearpermitold
+
+est = extract(mod, permute="TRUE")
+
+# start with making time series of estimates on mean
+# challenge is that they're time varying quadratic functions
+# do facet_wrap of estimates by year
+
+# effect of adding more 1 species from 1. specdiv.prev = 1
+
+specdiv.prev = seq(1, 5, by=0.1)
+specdiv = specdiv.prev + 1 # always add 1 more in these comparisons
+x1 = log(specdiv/specdiv.prev)
+x2 = log(specdiv/specdiv.prev) * specdiv.prev
+
+df = expand.grid("year"=1:29, "diversity" = seq(1,length(x1)), "permit"=1:6)
+df$permits = permits[df$permit]
+df$years = years[df$year]
+df$x = specdiv.prev[df$diversity]
+df$pred = 0
+df$low = 0
+df$high = 0
+b0_indx = match(paste(as.character(df$years), as.character(df$permits)), as.character(yearpermit$yrpermit))
+
+for(i in 1:nrow(df)) {
+  tmp = x1[df$diversity[i]] * (est$b_1[,df$year[i], df$permit[i]])# + x2[df$diversity[i]] * (est$b_2[,df$year[i], df$permit[i]]+est$b2_str_yr_mu)
+  df$pred[i] = median(tmp)
+  df$low[i] = quantile(tmp,0.025)
+  df$high[i] = quantile(tmp,0.975)
+}
+df_s01e_b1 = df
+#ggplot(df[df$permits=="S01E",], aes(x=x,y=pred,group=years)) +
+#  #geom_ribbon(aes(ymin=low,ymax=high), alpha=0.4) +
+#  geom_line() +
+#  facet_wrap(~years) + xlab("Previous species diversity") + ylab("Predicted #benefit of adding +1 spp (ln $)")
+
+```
+
+```{r figa7, echo=FALSE, fig.pos="placeHere", fig.cap = "As model validation, we can compare the estimated intercepts for PWS S01E (purse seine) representing average year to year change in revenue versus the change year over year in ln(pink salmon returns).
 Correlation between them is ~ 0.8.
 This suggests (as expected) people on average make a lot more money in years when pink salmon returns are high.
-[side note: diversification may only be beneficial here when returns and prices are both low.]"}
+[side note: diversification may only be beneficial here when returns and prices are both low.]", eval=FALSE}
 
@@
 ggplot(data=pinkPWS, aes(b0, diff,color=Year)) + geom_point() + stat_smooth(method=lm) + xlab("Estimated intercept from model") + ylab("Change in total pink salmon returns")
-
-# bring in slopes from above
-#slopes = df_s01e_b1[df_s01e_b1$permits=="S01E" & df_s01e_b1$diversity==1,]
-#slopes = slopes[,c("years","pred")]
-#names(slopes)[which(names(slopes)=="years")] = "Year"
-#names(slopes)[which(names(slopes)=="pred")] = "B1"
-#pinkPWS = dplyr::left_join(pinkPWS, slopes)
-
-#pinkPWS$diffchum = c(NA, diff(log(pinkPWS$Chum_total)))
-
-#prices = group_by(cfec[which(cfec$p_fshy%in%c("S 01E","S 03E") & #cfec$spec%in%c("PINK","CHUM","SOCK")),], year, spec) %>% 
-#  summarize(meanPrice = mean(g_price)) %>%
-#  arrange(year)
-#prices <- dcast(melt(prices, id.vars = c("spec", "year")), year ~ #variable+spec)
-#names(prices)[which(names(prices)=="year")]="Year"
-
-#pinkPWS = dplyr::left_join(pinkPWS, prices)
-#pinkPWS$pricediffchum = c(NA, diff(pinkPWS$meanPrice_CHUM))
-#pinkPWS$pricediffpink = c(NA, diff(pinkPWS$meanPrice_PINK))
-#pinkPWS$pricediffsock = c(NA, diff(pinkPWS$meanPrice_SOCK))
-
-#cor(pinkPWS[,-c(5,6)], use="pairwise.complete.obs")
-
-#ggplot(data=pinkPWS, aes(b0, B1)) + geom_point() + stat_smooth(method=lm) + xlab("Estimated intercept from model") + ylab("Change in total pink salmon returns")
-
-#diversity = dplyr::group_by(dat.salmon[dat.salmon$strategy_permit=="S01E",], year) %>%
-#  summarize(meanDiv = mean(specDiv))
-#names(diversity)[which(names(diversity)=="year")] = "Year"
-#pinkPWS = dplyr::left_join(pinkPWS, diversity)
-
 ```
-
-\break
-
-# References

### 2017-02-22 salmon/paper/salmon_paper_appendix_v3.Rmd
Revert "Revert "test""
This reverts commit 5085fdd090082f92c5e3c1796c48806f08a6e80d.

--- 2016-10-28 salmon/paper/salmon_paper_appendix.Rmd
@@
 ---
-title: ""
+title: "Supplement for Ward et al., 'Reduced diversity and increased specialization of Alaska salmon fishermen...' "
 author: ''
@@
 
-```{r, echo = FALSE, eval=FALSE}
-dat = dat.salmon
-# create other category, lumping coho/chinook
-dat$salm = dat$coho+dat$chnk
-dat$salm.prev = dat$coho.prev+dat$chnk.prev
-
-dat$totsalm = dat$salm + dat$chum + dat$sock + dat$pink
-dat$totsalm.prev = dat$salm.prev + dat$chum.prev + dat$sock.prev + dat$pink.prev
-
-dat$salm = dat$salm/dat$totsalm
-dat$sock = dat$sock/dat$totsalm
-dat$pink = dat$pink/dat$totsalm
-dat$chum = dat$chum/dat$totsalm
-dat$chnk = dat$chnk/dat$totsalm
-dat$coho = dat$coho/dat$totsalm
-dat$salm.prev = dat$salm.prev/dat$totsalm.prev
-dat$sock.prev = dat$sock.prev/dat$totsalm.prev
-dat$pink.prev = dat$pink.prev/dat$totsalm.prev
-dat$chum.prev = dat$chum.prev/dat$totsalm.prev
-dat$chnk.prev = dat$chnk.prev/dat$totsalm.prev
-dat$coho.prev = dat$coho.prev/dat$totsalm.prev
-
-dat = dat[-which(is.na(dat$salm+dat$sock+dat$pink+dat$chum)),]
-dat$logdiff = log(dat$revenue/dat$revenue.prev)
-
-dat_old = dat
-dat$sumP2 = abs(dat$salm - dat$salm.prev) + 
-  abs(dat$pink-dat$pink.prev) + 
-  abs(dat$chum-dat$chum.prev) + 
-  abs(dat$sock-dat$sock.prev)
-dat = dat[which(dat$sumP2 < 0.3),]
-```
-
-```{r, echo = FALSE, eval=FALSE, fig.cap="Results for PWS Purse seine fishery", fig.pos="placeHere"}
-
-df = dat[which(dat$strategy_permit=="S01E"),]
-pamk_s01e <- pamk(df[,c("chum","pink","sock","salm")])
-df$group = c("Pink-chum","Pink")[pamk_s01e$pamobject$clustering]
-
-p3 = ggtern(data=df[,c("chum","pink","sock","group")], aes(x=chum,y=pink,sock,color=as.factor(group))) +
-  geom_point(alpha=0.3) +
-  labs(x="CHUM",y="PINK",z="SOCK") + theme(
-        axis.title=element_text(size=8)) + theme(legend.position = "none")
-
-p0 = group_by(df, year, group) %>% 
-  summarize(n = length(unique(p_holder))) %>% 
-ggplot(aes(x = year, y = n, group=group, col = group)) + geom_line() + ylab("People") + ggtitle("S01E")
-
-p1 = group_by(df, year, group) %>% 
-  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
-    cv = sd(logdiff,na.rm=T)) %>% 
-ggplot(aes(x = year, y = meanrev, group=group, col = group)) + geom_line() + ylab("Mean revenue") + ggtitle("S01E")
-
-p2 = group_by(df, year, group) %>% 
-  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
-    cv = sd(logdiff,na.rm=T)) %>% 
-ggplot(aes(x = year, y = cv, group=group, col = group)) + geom_line() + ylab("CV revenue") + ggtitle("S01E")
-
-g = group_by(df, year, group) %>% 
-  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
-    cv = sd(logdiff,na.rm=T))
-
-gridExtra::grid.arrange(p3,p0,p1,p2,ncol=2)
-```
-
 \break
@@
 
-```{r, fig.cap="Species composition for 1988 and 2014 (each dot is a person) in the S03E fishery", fig.pos="placeHere", echo=FALSE, fig.height = 2, eval=FALSE}
-### One way to visualize the change is with ternary plots over the 3 axes (chum / sockeye / pink)
+# Appendix
 
-daty = group_by(cfec[cfec$p_fshy=="S 03E" & cfec$spec %in% c("CHUM","PINK","SOCK","COHO","CHNK"),], spec, p_holder, year) %>%
-  summarize(p = sum(g_earn)) %>%
-  group_by(p_holder, year) %>% mutate(p = p/sum(p))
-daty <- melt(daty, id.vars = c("spec", "p_holder", "year"))
-daty <- dcast(daty, p_holder + year ~ spec + variable)[,-1]
-daty[is.na(daty)]=0
-ggtern(data=daty[which(daty$year%in%c(1988,2014)),c("year", "CHUM_p","PINK_p","SOCK_p")], aes(x=CHUM_p,y=PINK_p,SOCK_p,color=as.factor(year))) +
-  geom_point(alpha=0.3,size=1) +
-  labs(x="CHUM",y="PINK",z="SOCK",title="Proportion of annual revenue (S03E)")
+## Overall trends in participation across permits
+
+```{r figa1, fig.cap="Participation (unique people)", fig.pos="placeHere", echo=FALSE}
+print(appendixPlot1)
 ```
@@
 
-```{r, echo=FALSE, eval=FALSE}
-
-mu.mod = glmmTMB(log(revenue) ~ specDiv.change + specDiv.change:specdiv.prev + specDiv.change:I(specdiv.prev^2) + days.change + (1+days.change|strategy_permit) + (-1+specDiv.change + specDiv.change:specdiv.prev + specDiv.change:I(specdiv.prev^2)|year_permit), offset = log(revenue.prev), dat.salmon)
-
-cv.mod = glmmTMB(log(abs(residuals(mu.mod))) ~ specDiv.change + specDiv.change:specdiv.prev + specDiv.change:I(specdiv.prev^2) + days.change + (1+days.change|strategy_permit) + (-1+specDiv.change + specDiv.change:specdiv.prev + specDiv.change:I(specdiv.prev^2)|year_permit), dat.salmon)
-# to ask whether benefit of diversificaiton is changing via strategy, we want to summarize (1) direction of 
-# benefit (e.g. is the effect negative or positive), and (2) the magnitude (how steep / shallow it is). With the
-# quadratic model, 
-# 'squared term' indicates concave up/down
-# derivative is magnitude, 2a * x + b
-
-df = ranef(mu.mod)$cond$year_permit
-colnames(df) = c("c","b","a")
-df$year = as.numeric(substr(rownames(df), 1, 4))
-df$strategy = substr(rownames(df), 6, 9)
-df$deriv = 2*df$a + df$b
-# also calculate change of going from 1 to 2 spp
-df$pred1 = df$a*(1^2) + df$b*1 + df$c + fixef(mu.mod)$cond["specDiv.change"] + fixef(mu.mod)$cond["specDiv.change:specdiv.prev"]*1
-# add fixed effects into predictions
-
-# We can demonstrate the quadratic curves by year for a single strategy / set of years
-df.s01e = df[df$strategy=="S01E",]
-demo = expand.grid("speciesDiversity" = seq(1,4,length.out=100), "year" = 1986:2014)
-demo = left_join(demo, df.s01e)
-demo$effect = demo$a * (demo$speciesDiversity^2) + demo$b * (demo$speciesDiversity) + demo$c
+```{r figa2, fig.cap="Diversity, mean revenue and CV (aggregated across years)", fig.pos="placeHere", echo=FALSE}
+gridExtra::grid.arrange(appendixPlot2_p1,appendixPlot2_p2,appendixPlot2_p3, ncol=2)
 ```
 
-```{r, echo = FALSE, warning=FALSE, message=FALSE, results='hide', eval=FALSE}
-pdf("../../salmon/Salmon_residuals.pdf")
-# Residuals for mean model
-dat.salmon$mu.resid = residuals(mu.mod)
-dat.salmon$mu.fitted = fitted(mu.mod)
+```{r, echo=FALSE}
+#cfec <- feather::read_feather("../../portfolio/data-generated/salmon.feather")
+#saveRDS(cfec, "portfolio/data-generated/salmon_data_for_paper.rds")
+#cfec_all = readRDS("../../portfolio/data-generated/salmon_data_for_paper.rds")
+#cfec_all$year = as.numeric(substr(cfec_all$landdate, 1, 4))
 
-ggplot(dat.salmon, aes(mu.fitted, mu.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Fitted") + ylab("Residuals") + ggtitle("Mean model")
-ggplot(dat.salmon, aes(specDiv.change, mu.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Change in species diversification") + ylab("Residuals") + ggtitle("Mean model") + geom_hline(aes(yintercept=0),color="salmon1")
-ggplot(dat.salmon, aes(specdiv.prev, mu.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Previous species diversity") + ylab("Residuals") + ggtitle("Mean model") + geom_hline(aes(yintercept=0),color="salmon1")
-ggplot(dat.salmon, aes(log((revenue+revenue.prev)/2), mu.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Average revenue") + ylab("Residuals") + ggtitle("Mean model") + geom_hline(aes(yintercept=0),color="salmon1")
-ggplot(dat.salmon, aes(specDiv.change, log(abs(mu.resid)), group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Change in species diversification") + ylab("Ln abs(residuals)") + ggtitle("Mean model")
-ggplot(dat.salmon, aes(specdiv.prev, log(abs(mu.resid)), group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Previous species diversity") + ylab("Ln abs(residuals)") + ggtitle("Mean model")
-ggplot(dat.salmon, aes(specdiv.prev, log(abs(mu.resid)), group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Previous species diversity") + ylab("Ln abs(residuals)") + ggtitle("Mean model")
+# plot proportion of species landed by salmon permit
+cbPalette <- c("#009E73", "#4B2E83", "#999999", "#CC79A7", "#D55E00")
 
-# Residuals for CV model
-dat.salmon$cv.resid = residuals(cv.mod)
-dat.salmon$cv.fitted = fitted(cv.mod)
+#cfec = readRDS("../../portfolio/data-generated/salmon_data_for_paper.rds")
+#cfec.raw = cfec[cfec$p_fshy%in%c("S 01A", "S 01E", "S 01K", "S 03A", "S 03E", "S 01M"),]
 
-ggplot(dat.salmon, aes(cv.fitted, cv.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Fitted") + ylab("Residuals") + ggtitle("CV model")
-ggplot(dat.salmon, aes(specDiv.change, cv.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Change in species diversification") + ylab("Residuals") + ggtitle("CV model") + geom_hline(aes(yintercept=0),color="salmon1")
-ggplot(dat.salmon, aes(specdiv.prev, cv.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Previous species diversity") + ylab("Residuals") + ggtitle("CV model") + geom_hline(aes(yintercept=0),color="salmon1")
-ggplot(dat.salmon, aes(log((revenue+revenue.prev)/2), cv.resid, group = strategy_permit)) + geom_point(alpha=0.05,size=1) + facet_wrap(~strategy_permit) + xlab("Average revenue") + ylab("Residuals") + ggtitle("CV model") + geom_hline(aes(yintercept=0),color="salmon1")
-dev.off()
+g1 = group_by(cfec.raw[cfec.raw$p_fshy=="S 01A" & cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
+  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
+  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
+  ylab("Species %") + xlab("") + ggtitle("S01A") + 
+scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
+    'COHO'=cbPalette[3], 'PINK'=cbPalette[4], 'SOCK'=cbPalette[5]), name="Species",
+    labels =c("Chinook","Chum","Coho","Pink","Sockeye")) + theme(legend.position="none")
+
+g2 = group_by(cfec.raw[cfec.raw$p_fshy=="S 01E" & cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
+  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
+  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
+  ylab("Species %") + xlab("") + ggtitle("S01E") + theme(legend.position="none") + 
+scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
+    'COHO'=cbPalette[3], 'PINK'=cbPalette[4], 'SOCK'=cbPalette[5]))
+
+g3 = group_by(cfec.raw[cfec.raw$p_fshy=="S 01K" & cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
+  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
+  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
+  ylab("Species %") + xlab("") + ggtitle("S01K") + theme(legend.position="none") + 
+scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
+    'COHO'=cbPalette[3], 'PINK'=cbPalette[4], 'SOCK'=cbPalette[5]))
+
+g4 = group_by(cfec.raw[cfec.raw$p_fshy=="S 01M" & cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
+  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
+  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
+  ylab("Species %") + xlab("") + ggtitle("S01M") + theme(legend.position="none") + 
+scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
+    'COHO'=cbPalette[3], 'PINK'=cbPalette[4], 'SOCK'=cbPalette[5]), name = "Species", labels =c("Chinook","Chum","Coho","Pink","Sockeye")) + theme(legend.position="none")
+
+g5 = group_by(cfec.raw[cfec.raw$p_fshy=="S 03A" & cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
+  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
+  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
+  ylab("Species %") + xlab("") + ggtitle("S03A") + theme(legend.position="none") + 
+scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
+    'COHO'=cbPalette[3], 'PINK'=cbPalette[4], 'SOCK'=cbPalette[5]))
+
+g6 = group_by(cfec.raw[cfec.raw$p_fshy=="S 03E" & cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK") & cfec.raw$year > 0,], year, spec, p_fshy) %>%
+  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
+  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
+  ylab("Species %") + xlab("") + ggtitle("S03E") + theme(legend.position="none") + 
+scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
+    'COHO'=cbPalette[3], 'PINK'=cbPalette[4], 'SOCK'=cbPalette[5]))
+
+dat.diff = readRDS(file="../../portfolio/data-generated/salmon-diff-for-modeling.rds")
+dat.diff = dat.diff[which(dat.diff$revenue >= 10000 & dat.diff$revenue.prev >= 10000), ]
+dat.salmon = dat.diff[dat.diff$strategy_permit%in%c("S01A", "S01E", "S01K", "S03A", "S03E","S01M"),]
+
+h1 = group_by(dat.salmon[dat.salmon$strategy_permit=="S01A",], year, strategy_permit) %>%
+  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
+  ggplot(aes(year, meanDiv)) + geom_line() + 
+  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Diversity") + xlab("") + theme(legend.position="none")
+
+h2 = group_by(dat.salmon[dat.salmon$strategy_permit=="S01E",], year, strategy_permit) %>%
+  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
+  ggplot(aes(year, meanDiv)) + geom_line() + 
+  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Diversity") + xlab("") + theme(legend.position="none")
+
+h3 = group_by(dat.salmon[dat.salmon$strategy_permit=="S01K",], year, strategy_permit) %>%
+  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
+  ggplot(aes(year, meanDiv)) + geom_line() + 
+  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Diversity") + xlab("") + theme(legend.position="none")
+
+h4 = group_by(dat.salmon[dat.salmon$strategy_permit=="S01M",], year, strategy_permit) %>%
+  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
+  ggplot(aes(year, meanDiv)) + geom_line() + 
+  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Diversity") + xlab("") + theme(legend.position="none")
+
+h5 = group_by(dat.salmon[dat.salmon$strategy_permit=="S03A",], year, strategy_permit) %>%
+  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
+  ggplot(aes(year, meanDiv)) + geom_line() + 
+  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Diversity") + xlab("") + theme(legend.position="none")
+
+h6 = group_by(dat.salmon[dat.salmon$strategy_permit=="S03E",], year, strategy_permit) %>%
+  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
+  ggplot(aes(year, meanDiv)) + geom_line() + 
+  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Diversity") + xlab("") + theme(legend.position="none")
+
 ```
-```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Variation in effects of diversification by year for the S01E (PWS purse seine) permit holders. ", eval=FALSE}
-###We can make quadratic plots of these effects by year and strategy.
-These plots represent the effect of increasing diversity (as relative change) on mean revenue.
-For example, in 1986 if you were fishing 3-4 species there would be little effect of adding more spp; but if you were fishing 1 species, there would be a slight negative effect of adding more.
-In 1991, there would be a huge benefit of increasing diversity, and that effect would taper off as more were added.
 
-ggplot(demo, aes(speciesDiversity, effect, group = year)) + geom_line() + facet_wrap(~ year) + xlab("Species Diversity") + ylab("Effect of increasing diversity (relative change)")
+```{r figa3, echo=FALSE, fig.pos="placeHere", fig.cap="Time varying species composition and effective species diversity (1 / Simpson's Diversity), by permit for seine fisheries in Southeast Alaska (S01A), Prince William Sound (S01E) and Kodiak (S01K).
+Species are colored as follows: green (Chinook), purple (chum), silver  (coho), pink (pink), red (sockeye)", fig.height=3.2}
+gridExtra::grid.arrange(g1, g2, g3, h1, h2, h3, nrow=2)
+```
+
+```{r figa4, echo=FALSE, fig.pos="placeHere", fig.cap="Time varying species composition and effective species diversity (1 / Simpson's Diversity), by permit, for seine fisheries on the Alaska Peninsula (S01M), and gillnet fisheries in Southeast Alaska (S03A) and Prince William Sound (S03E).", fig.height=3.2}
+gridExtra::grid.arrange(g4, g5, g6, h4, h5, h6, nrow=2)
 ```
@@
 
-```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Time series of benefits of diversification by permit / strategy.", fig.height = 3, eval=FALSE}
-###This plot shows the time series of becoming increasingly diversified from 1 species on mean revenue.
-For S01E for example, there's huge benefits of diversifying in 1991, and huge costs in 2010.
-2010 was a huge pink salmon year for PWS (with high prices specialization was good) and 1991 was the start of pink salmon stock + price collapse.
-Across permits, perhaps slight declines in some of these in recent years?
+## Model Results
 
-ggplot(df, aes(year, pred1, group = strategy, colour = as.factor(strategy))) + geom_line() + xlab("Year") + ylab("Benefit of increasing species diversity from 1 species") + ggtitle("Effect on Mean")
-```
+```{r figa5, fig.pos="placeHere", fig.cap="Time varying effects of adding 1 species on log standard deviation of revenue for hypothetical individuals with species diversity ~ 1.", echo=FALSE}
 
-```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Time series of benefits of diversification split out by permit / strategy.", fig.height = 3, eval=FALSE}
-ggplot(df, aes(year, pred1, group = strategy, colour = as.factor(strategy))) + geom_line() + xlab("Year") + ylab("Benefit of increasing species diversity from 1 species") + facet_wrap(~ strategy, scale="free") + geom_smooth() + geom_hline(aes(yintercept=0))
-```
+diffdat = readRDS(file="../../salmon/data-generated/salmon.rds")
+years = seq(min(diffdat$year), max(diffdat$year))
+permits = levels(as.factor(diffdat$permit))
+# note that we have to use factor() here instead of unique() because
+# that's the same as the ordering in the estimation code
+yearpermit = data.frame("yrpermit"=levels(as.factor(diffdat$year_permit)))
+yearpermit$year = as.numeric(substr(yearpermit$yrpermit,1,4))
+yearpermit$permit = substr(yearpermit$yrpermit,6,9)
+yearpermitold=yearpermit
 
-```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Time series of effects of diversification on mean and CV by permit.", eval=FALSE}
-df.cv = ranef(cv.mod)$cond$year_permit
-colnames(df.cv) = c("c.cv","b.cv","a.cv")
-df.cv$year = as.numeric(substr(rownames(df.cv), 1, 4))
-df.cv$strategy = substr(rownames(df.cv), 6, 9)
-# also calculate change of going from 1 to 2 spp
-df.cv$pred1.cv = df.cv$a.cv*(1^2) + df.cv$b.cv*1 + df.cv$c.cv + fixef(cv.mod)$cond["specDiv.change"] + fixef(cv.mod)$cond["specDiv.change:specdiv.prev"]*1
+# grab original data to look at permit names
+diffdat = readRDS(file="../../portfolio/data-generated/salmon-diff-for-modeling.rds")
+years = seq(min(diffdat$year), max(diffdat$year))
+#permits = levels(as.factor(diffdat$permit))
 
-ggplot(df.cv, aes(year, pred1.cv, group = strategy, colour = as.factor(strategy))) + geom_line(linetype="dashed") + geom_line(data=df, aes(year, pred1, group = strategy, colour = as.factor(strategy))) + xlab("Year") + ylab("Effect of increasing species diversity from 1 species") + facet_wrap(~ strategy, scale="free") + ggtitle("Effects on mean (solid) and CV (dashed)") + geom_hline(aes(yintercept=0), color="grey30")
+load("../../salmon/analysis/model-linear.Rdata")
+yearpermit=yearpermitold
+
+est = rstan::extract(mod, permute="TRUE")
+
+# start with making time series of estimates on mean
+# challenge is that they're time varying quadratic functions
+# do facet_wrap of estimates by year
+
+# effect of adding more 1 species from 1. specdiv.prev = 1
+
+specdiv.prev = seq(1, 5, by=0.1)
+specdiv = specdiv.prev + 1 # always add 1 more in these comparisons
+x1 = log(specdiv/specdiv.prev)
+x2 = log(specdiv/specdiv.prev) * specdiv.prev
+x3 = log(specdiv/specdiv.prev) * specdiv.prev^2
+
+df = expand.grid("year"=1:length(years), "diversity" = seq(1,length(x1)), "permit"=1:6)
+df$permits = permits[df$permit]
+df$years = years[df$year]
+df$x = specdiv.prev[df$diversity]
+df$pred = 0
+df$low = 0
+df$high = 0
+
+for(i in 1:nrow(df)) {
+  #tmp = est$g0 + est$g0_strategy[,df$permit[i]] + 
+  tmp = 
+    x1[df$diversity[i]] * (est$g_1[,df$year[i], df$permit[i]])# +
+    #x2[df$diversity[i]] * (est$g_2[,df$year[i], df$permit[i]]+est$g2_str_yr_mu)
+  df$pred[i] = median(tmp)
+  df$low[i] = quantile(tmp,0.025)
+  df$high[i] = quantile(tmp,0.975)
+}
+
+ggplot(df[df$diversity==1,], aes(x=years,y=pred,group=permits)) +
+  geom_line() +
+  geom_ribbon(aes(ymin=low,ymax=high), alpha=0.4) +
+  facet_wrap(~permits) + xlab("Year") + ylab("Predicted marginal benefit of adding +1 spp (ln $)") + geom_hline(yintercept=0, color="red")
 
@@
 
-\break
+```{r figa6, echo=FALSE, fig.pos="placeHere", fig.cap="Time varying intercepts on estimated mean revenue, by permit.
+These are interpreted as the mean change in revenue for each permit holder in each year (positive equates to increasing revenue, etc), after accounting for effects of catch diversification and days fished."}
+# grab original data to look at permit names
+diffdat = readRDS(file="../../salmon/data-generated/salmon.rds")
 
-```{r, echo=FALSE, eval = FALSE}
-# These are the run  timing plots per Milo's questions
+years = seq(min(diffdat$year), max(diffdat$year))
+permits = levels(as.factor(diffdat$permit))
 
-# for 2008, create cumulative density plots of salmon landings by calday
-pinkchum = cfec.raw[cfec.raw$p_fshy=="S 01E" & cfec.raw$spec%in%c("PINK","CHUM","SOCK"),]
-#pinkchum = cfec.raw[cfec.raw$p_fshy=="S 03T" & cfec.raw$spec%in%c("PINK","SOCK","CHNK"),]
+# note that we have to use factor() here instead of unique() because
+# that's the same as the ordering in the estimation code
+yearpermit = data.frame("yrpermit"=levels(as.factor(diffdat$year_permit)))
+yearpermit$year = as.numeric(substr(yearpermit$yrpermit,1,4))
+yearpermit$permit = substr(yearpermit$yrpermit,6,9)
 
-# aggregate by calendar day
-pinkchum$month = as.numeric(substr(pinkchum$day, 1, 2))
-pinkchum$day = as.numeric(substr(pinkchum$day, 4, 5))
-pinkchum$jday = date::mdy.date(pinkchum$month, pinkchum$day, pinkchum$year) - date::mdy.date(1,1,pinkchum$year)
+yearpermitold=yearpermit
+load("../../salmon/analysis/model-linear.Rdata")
+yearpermit=yearpermitold
+est = rstan::extract(mod, permute="TRUE")
 
-group_by(pinkchum, jday, year) %>% 
-  summarize(pink = sum(g_earn[spec=="PINK"]), 
-    chum = sum(g_earn[spec=="CHUM"]),
-    sock = sum(g_earn[spec=="SOCK"])) %>% 
-  group_by(year) %>% 
-  mutate(tot = sum(pink) + sum(chum) + sum(sock), pink = pink/tot, chum = chum/tot, sock = sock/tot) %>% 
-ggplot(aes(jday, pink)) + geom_line(color="pink") + geom_line(aes(jday, chum), color="darkgreen") + geom_line(aes(jday, sock), color="lightblue") + facet_wrap(~ year, scale="free")
+yearpermit$b0 = apply(est$b0_str_yr,2,median)
+yearpermit$b0_low = apply(est$b0_str_yr,2,quantile,0.025)
+yearpermit$b0_hi = apply(est$b0_str_yr,2,quantile,0.975)
 
-# make plot of all earnings or weight by jday
-salm = cfec.raw[cfec.raw$p_fshy=="S 01E",]
-salm$month = as.numeric(substr(salm$day, 1, 2))
-salm$day = as.numeric(substr(salm$day, 4, 5))
-salm$jday = date::mdy.date(salm$month, salm$day, salm$year) - date::mdy.date(1,1,salm$year)
+ggplot(data=yearpermit) + 
+  geom_ribbon(aes(x=year,ymin=b0_low,ymax=b0_hi),alpha=0.3) +
+  geom_line(aes(x=year,y=b0)) + 
+  facet_wrap(~permit) + xlab("Year") + 
+  ylab("Intercepts on mean") + 
+  geom_hline(yintercept=0, color="red")
+```
 
-g = group_by(salm, jday, year) %>% 
-  summarize(earn = sum(g_earn), 
-    lbs = sum(g_pounds), 
-    pink = sum(g_earn[spec=="PINK"])/earn,
-    sock = sum(g_earn[spec=="SOCK"])/earn,
-    chum = sum(g_earn[spec=="CHUM"])/earn,
-    coho = sum(g_earn[spec=="COHO"])/earn,
-    chnk = sum(g_earn[spec=="CHNK"])/earn) %>% 
-ggplot(g, aes(jday, (earn))) + geom_line(colour = rgb(g$sock+g$coho+g$chnk, g$chum, g$pink)) + facet_wrap(~ year, scale="free_y")
+```{r figa7, eval=FALSE,echo=FALSE}
+load("../../salmon/analysis/cv-model-simple.Rdata")
+# groups here are 80% pink vs 97% pink
+
+yrs = 1986:2014
+
+hist(rstan::extract(mod)[["g_offset"]], 40, col="grey", xlab="Effect of specializing on pink (sd)")
+
+xx = exp(rstan::extract(mod)[["sd_yr"]])
+df.gen = data.frame("year"=yrs, "mean" = apply(xx, 2, mean), "low" = apply(xx, 2, quantile, 0.025),
+  "hi" = apply(xx, 2, quantile, 0.975), "strategy"="pink:chum")
+
+offset = matrix(rstan::extract(mod)[["g_offset"]],ncol=1) %*% matrix(rep(1,length(yrs)), nrow=1)
+xx = exp(extract(mod)[["sd_yr"]] + offset)
+df.spec = data.frame("year"=yrs, "mean" = apply(xx, 2, mean), "low" = apply(xx, 2, quantile, 0.025),
+  "hi" = apply(xx, 2, quantile, 0.975), "strategy"="pink")
+
+ggplot(data=rbind(df.gen,df.spec), aes(year, mean, group=strategy, fill=strategy)) + geom_ribbon(aes(ymin=low,ymax=hi), alpha=0.5) + geom_hline(yintercept=0, color="red")
 
@@
 
-```{r, echo = FALSE, fig.cap="SEAK purse seine fishery", fig.pos="placeHere", eval=FALSE}
+```{r figa8, echo=FALSE, eval=FALSE}
+# fig.pos="placeHere", fig.cap="Time varying effects of adding 1 species on revenue (measured in ln dollars) for purse seine permit in Prince William Sound (S01E).
+This figure illustrates two things: (1) the benefit of adding more species diminishes as a function of past diversity, and (2) the economic benefit to adding more species has been reduced through time."
+# grab original data to look at permit names
+diffdat = readRDS(file="../../salmon/data-generated/salmon.rds")
+years = seq(min(diffdat$year), max(diffdat$year))
+#permits = levels(as.factor(diffdat$permit))
 
-df = dat[which(dat$strategy_permit=="S01A"),]
-pamk_s01a <- pamk(df[,c("chum","pink","sock","salm")])
-df$group = pamk_s01a$pamobject$clustering
+load("../../salmon/analysis/model-linear.Rdata")
+yearpermit=yearpermitold
 
-p3 = ggtern(data=df[,c("chum","pink","sock","group")], aes(x=chum,y=pink,sock,color=as.factor(group))) +
-  geom_point(alpha=0.3) +
-  labs(x="CHUM",y="PINK",z="SOCK") + theme(
-        axis.title=element_text(size=8)) + theme(legend.position = "none")
+est = extract(mod, permute="TRUE")
 
-p0 = group_by(df, year, group) %>% 
-  summarize(n = length(unique(p_holder))) %>% 
-ggplot(aes(x = year, y = n, group=group, col = as.factor(group))) + geom_line() + ylab("People") + ggtitle("S01A")
+# start with making time series of estimates on mean
+# challenge is that they're time varying quadratic functions
+# do facet_wrap of estimates by year
 
-p1 = group_by(df, year, group) %>% 
-  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
-    cv = sd(logdiff,na.rm=T)) %>% 
-ggplot(aes(x = year, y = meanrev, group=group, col = as.factor(group))) + geom_line() + ylab("Mean revenue") + ggtitle("S01A")
+# effect of adding more 1 species from 1. specdiv.prev = 1
 
-p2 = group_by(df, year, group) %>% 
-  summarize(meanrev = log(mean(revenue + revenue.prev)/2), 
-    cv = sd(logdiff,na.rm=T)) %>% 
-ggplot(aes(x = year, y = cv, group=group, col = as.factor(group))) + geom_line() + ylab("CV revenue") + ggtitle("S01A")
+specdiv.prev = seq(1, 5, by=0.1)
+specdiv = specdiv.prev + 1 # always add 1 more in these comparisons
+x1 = log(specdiv/specdiv.prev)
+x2 = log(specdiv/specdiv.prev) * specdiv.prev
 
-gridExtra::grid.arrange(p3, p0, p1, p2, ncol=2)
+df = expand.grid("year"=1:29, "diversity" = seq(1,length(x1)), "permit"=1:6)
+df$permits = permits[df$permit]
+df$years = years[df$year]
+df$x = specdiv.prev[df$diversity]
+df$pred = 0
+df$low = 0
+df$high = 0
+b0_indx = match(paste(as.character(df$years), as.character(df$permits)), as.character(yearpermit$yrpermit))
+
+for(i in 1:nrow(df)) {
+  tmp = x1[df$diversity[i]] * (est$b_1[,df$year[i], df$permit[i]])# + x2[df$diversity[i]] * (est$b_2[,df$year[i], df$permit[i]]+est$b2_str_yr_mu)
+  df$pred[i] = median(tmp)
+  df$low[i] = quantile(tmp,0.025)
+  df$high[i] = quantile(tmp,0.975)
+}
+df_s01e_b1 = df
+#ggplot(df[df$permits=="S01E",], aes(x=x,y=pred,group=years)) +
+#  #geom_ribbon(aes(ymin=low,ymax=high), alpha=0.4) +
+#  geom_line() +
+#  facet_wrap(~years) + xlab("Previous species diversity") + ylab("Predicted #benefit of adding +1 spp (ln $)")
+
 ```
 
-\pagebreak
+```{r figa9, echo=FALSE, fig.pos="placeHere", fig.cap = "As model validation, we can compare the estimated intercepts for PWS S01E (purse seine) representing average year to year change in revenue versus the change year over year in ln(pink salmon returns).
+Correlation between them is ~ 0.8.
+This suggests (as expected) people on average make a lot more money in years when pink salmon returns are high.
+[side note: diversification may only be beneficial here when returns and prices are both low.]", eval=FALSE}
 
-# Appendix
+yearpermit$b0 = apply(est$b0_str_yr,2,median)
+yearpermit$b0_low = apply(est$b0_str_yr,2,quantile,0.025)
+yearpermit$b0_hi = apply(est$b0_str_yr,2,quantile,0.975)
 
-## Overall trends in participation across permits
+S01E = yearpermit[yearpermit$permit=="S01E",]
+names(S01E)[which(names(S01E)=="year")]="Year"
 
-```{r, fig.cap="Participation (unique people)", fig.pos="placeHere", echo=FALSE}
-appendixPlot1
-```
+returns = readxl::read_excel("../../salmon/paper/salmon_returns_PWS.xlsx")
+pinkPWS = returns[,c("Year", "Pink_total", "Chum_total")]
+pinkPWS = pinkPWS[which(pinkPWS$Year >= 1975),]
+pinkPWS$diff = c(NA, diff(log(pinkPWS$Pink_total)))
 
-\break
+pinkPWS = dplyr::left_join(pinkPWS, S01E)
 
-```{r, fig.cap="Trends in revenue and variability across people", fig.pos="placeHere", echo=FALSE, fig.height = 3.1}
-group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], p_holder, year, p_fshy) %>%
-  summarize(gearn = sum(g_earn)) %>%
-  group_by(year, p_fshy) %>%
-  summarize(med = median(gearn), low = quantile(gearn, 0.25),
-    upp = quantile(gearn,0.75)) %>%
-  ggplot(aes(year, log(med))) + geom_line() + facet_wrap(~p_fshy,scale="free") + 
-  geom_ribbon(aes(ymin= log(low), ymax= log(upp)), alpha=0.4) + ylab("Ln quartiles of revenue (across people)") + ggtitle("Trends in revenue") + xlab("Year")
-```
-
-```{r, echo=FALSE, fig.height = 3.1, fig.cap="Trends in average landed weight per person", fig.pos="placeHere"}
-group_by(cfec, p_fshy, year, p_holder) %>% 
-  summarize(meanW = mean(g_pounds/2000, na.rm=T)) %>% 
-  group_by(p_fshy, year) %>% 
-  summarize(meanW = mean(meanW, na.rm=T)) %>% 
-  ggplot(aes(year, meanW)) + geom_line() + facet_wrap(~ p_fshy, scale="free") + ylab("Mean total tons / person") + xlab("Year") + ggtitle("Trends in landed weight")
-```
-
-\break
-
-```{r, fig.cap="Diversity, mean revenue and CV (aggregated across years)", fig.pos="placeHere", echo=FALSE}
-gridExtra::grid.arrange(appendixPlot2_p1,appendixPlot2_p2,appendixPlot2_p3, ncol=2)
-```
-
-\break
-
-## Species diversity
-
-```{r, fig.cap="Proportion of species landed, weighted by earnings", fig.pos="placeHere", echo=FALSE, fig.height = 3.3, eval=FALSE}
-## Species composition
-# plot proportion of species landed by salmon permit
-group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
-  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p)) %>%
-  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() + facet_wrap(~p_fshy) + 
-  ylab("Species % (total) ") + xlab("")
-```
-
-```{r, fig.cap="Effective species diversity across people", fig.pos="placeHere", echo=FALSE, fig.height = 3.3}
-group_by(dat.salmon, year, strategy_permit) %>%
-  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
-  ggplot(aes(year, meanDiv)) + geom_line() + facet_wrap(~strategy_permit, scale="free") + 
-  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Effective diversity (across people)") + xlab("")
-```
-
-```{r, fig.cap="Proportion of total returns to Prince William Sound (hatchery + wild, catch + escapement).
-Years before 1980 and after 2011 aren't included because not all species have data.", fig.pos="placeHere", echo=FALSE, fig.height = 3.3}
-returns = readxl::read_excel("salmon_returns_PWS.xlsx")
-returns = returns[,c("Year","Sockeye_total","Pink_total","Chum_total","Coho","Chinook")]
-
-returns = as.data.frame(returns[which(returns$Year%in%c(1980:2011)),])
-for(i in 1:ncol(returns)) returns[,i] = as.numeric(returns[,i])
-for(i in 1:nrow(returns)) {
-    # normalize
-    returns[i,-1] = returns[i,-1] / sum(returns[i,-1])
-}
-
-melted = reshape2::melt(returns,id.vars="Year")
-names(melted)[which(names(melted)=="variable")]="species"
-ggplot(melted,aes(Year,value, group = species, color=species)) + geom_line() + ylab("Proportion of total returns (by #)")
+ggplot(data=pinkPWS, aes(b0, diff,color=Year)) + geom_point() + stat_smooth(method=lm) + xlab("Estimated intercept from model") + ylab("Change in total pink salmon returns")
 ```
@@
   ylab("Revenue (millions) ") + xlab("Year")
-```
-
-\break
-
-## Trends in variability
-
-```{r, fig.cap="Trends in annual variability by fishery", fig.pos="placeHere", echo=FALSE,  fig.height=3.3}
-summary = group_by(cfec.raw[cfec.raw$spec%in%c("CHUM","COHO","CHNK","SOCK","PINK"),], year, p_fshy, spec) %>% 
-  summarize(sumg = log(sum(g_earn)), sumw = log(sum(g_pounds))) %>% 
-  arrange(p_fshy, spec, year) %>% group_by(p_fshy, spec) %>% 
-  mutate(logdiffg = rollapply(c(NA, diff(sumg)), 3, sd, fill=c(NA, NULL, NA)), 
-    logdiffw = rollapply(c(NA, diff(sumw)), 3, sd, fill=c(NA, NULL, NA)))
-
-ggplot(summary, aes(year, logdiffg, color=spec)) + geom_line() + facet_wrap(~ p_fshy, scale="free") + ylab("3 year rolling CV (revenue)") + xlab("Year")
-```
-
-```{r, fig.cap="Trends in annual variability by fishery and species", fig.pos="placeHere",  echo=FALSE, fig.height=3.3}
-g1 = ggplot(summary[which(summary$spec=="PINK" & summary$p_fshy%in%c("S 01A", "S 01E", "S 03A", "S 03E")),], aes(year, logdiffg, color=p_fshy)) + geom_line() + ylab("3 year rolling CV (revenue)") + xlab("Year") + ggtitle("Pink salmon")
-
-g2 = ggplot(summary[which(summary$spec=="CHUM" & summary$p_fshy%in%c("S 01A", "S 01E", "S 03A", "S 03E")),], aes(year, logdiffg, color=p_fshy)) + geom_line() + ylab("3 year rolling CV (revenue)") + xlab("Year") + ggtitle("Chum salmon")
-
-gridExtra::grid.arrange(g1,g2, ncol=1)
-```
-
-```{r, fig.cap="Trends in annual variability by fishery and species", fig.pos="placeHere", echo=FALSE, eval=FALSE}
-g1 = ggplot(summary[which(summary$spec=="PINK" & summary$p_fshy%in%c("S 01A", "S 01E", "S 03A", "S 03E")),], aes(year, logdiffw, color=p_fshy)) + geom_line() + ylab("3 year rolling CV (weight)") + xlab("Year") + ggtitle("Pink salmon")
-
-g2 = ggplot(summary[which(summary$spec=="CHUM" & summary$p_fshy%in%c("S 01A", "S 01E", "S 03A", "S 03E")),], aes(year, logdiffw, color=p_fshy)) + geom_line() + ylab("3 year rolling CV (weight)") + xlab("Year") + ggtitle("Chum salmon")
-
-gridExtra::grid.arrange(g1,g2, ncol=1)
-```
-
-```{r, fig.cap="Trends in variability across people", fig.pos="placeHere", echo=FALSE, fig.height=3.5, eval=FALSE}
-summary = group_by(cfec.raw[cfec.raw$spec%in%c("CHUM","COHO","CHNK","SOCK","PINK"),], year, p_fshy, p_holder) %>% 
-  summarize(sumg = sum(g_earn), sumw = sum(g_pounds)) %>% 
-  group_by(p_fshy, year) %>% 
-  summarize(cvg = sd(sumg,na.rm=T)/mean(sumg,na.rm=T),
-    cvw = sd(sumw,na.rm=T)/mean(sumw,na.rm=T))
-ggplot(summary, aes(year, cvg )) + geom_line(color ="green" ) + facet_wrap(~ p_fshy, scale="free") + geom_line(aes(year, cvw), color = "blue") + ylab("3 year rolling CV") + xlab("Year") + ggtitle("Variability across people: total revenue (green) and weight (blue)")
-```
-
-```{r, fig.cap="Change in CV for individual people", fig.pos="placeHere", echo=FALSE, fig.height=3.5}
-summary = group_by(cfec.raw[cfec.raw$spec%in%c("CHUM","COHO","CHNK","SOCK","PINK"),], year, p_fshy, p_holder) %>% 
-  summarize(sumg = sum(g_earn)) %>% 
-  group_by(p_fshy, p_holder) %>% 
-  mutate(nPre = length(which(year%in%seq(1985,1995))),
-    nPost = length(which(year%in%seq(2005,2014)))) %>% 
-  filter(nPre > 3 & nPost > 3) %>% select(-nPre, -nPost) %>% 
-  summarize(cv.pre = mean(sumg[which(year <= 1995)],na.rm=T)/sd(sumg[which(year <= 1995)],na.rm=T),
-    cv.post = mean(sumg[which(year >= 2005)],na.rm=T)/sd(sumg[which(year >= 2005)],na.rm=T))
-
-ggplot(summary, aes(log(cv.post/cv.pre))) + geom_histogram(bins=100) + facet_wrap(~ p_fshy) + geom_vline(aes(xintercept = 0)) + ylab("Count") + xlab("% Change in CV for individual people (2005-2014 / 1985-1995)") + ggtitle("Change in variability for individual people")
-#summary = left_join(expand.grid(year = unique(summary$year), p_holder = unique(summary$p_holder)), summary)
-
-# add rolling CV for each person
-#summary = group_by(summary, p_holder) %>% 
-#  mutate(logdiffg = rollapply(c(NA, diff(sumg)), 3, sd, fill=c(NA, NULL, NA))) %>%
-#  do(fit = lm(logdiffg ~ year, data = .))
 ```

### 2017-02-22 salmon/paper/salmon_paper_v3.Rmd
Revert "Revert "Revert "local change/"""
This reverts commit 67ac571a582d1a80cb839a0492cdfe5fbb9e320f.

@@
   effects on variation in individual revenues'
-author: "Eric, Sean, Ole, Rich, Milo, Jordan, Jen, Alan, Ben, etc."
+author: "Eric, Sean, Ole, Rich, Milo, Jordan, Jen, Alan, Ben..."
 date: ''

### 2017-02-22 salmon/paper/salmon_paper_appendix_v3.Rmd
commit lost files

@@
 ---
-title: "Supplement for Ward et al., 'Reduced diversity and increased specialization of Alaska salmon fishermen...' "
+title: "Supplement for Ward et al., 'Reduced diversity and increased specialization of Alaska salmon fishermen' "
 author: ''

### 2017-02-22 salmon/paper/salmon_paper_v3.Rmd
commit lost files

@@
   effects on variation in individual revenues'
-author: "Eric, Sean, Ole, Rich, Milo, Jordan, Jen, Alan, Ben..."
+author: "Eric, Sean, Ole, Rich, Jordan, Jen, Alan, Milo, Ben"
 date: ''
@@
 library(date)
-#library(ggtern)
 library(mgcv)
-#library(fpc)
 library(gridExtra)
-#library(glmmTMB)
 library(zoo)
@@
 Thus, individuals holding a single permit may only be able to increase species diversity within the confines of their permit (longer fishing seasons, broader spatial area).
-Using a long-term dataset of individual salmon fishers in Alaska, we develop new models for understanding how the effects of diversification have changed through time.
+Using a 40 year dataset of individual salmon fishers in Alaska, we develop new models for understanding how the effects of diversification have changed through time.
 These changes allow us to compare patterns across regions and gear, as well as better understand how external forces have changed incentives for diversifying (hatchery production, market forces, EVOS).
@@
 
-[broader paragraph on diversification with extensions to terrestrial agriculture etc, ]
+[broader paragraph on diversification with extensions -- somewhat depends on journal, ]
 
@@
 Understanding the effects of species diversification on salmon fishers is particularly important, as the number of salmon specialists (defined as fishers holding a single salmon permit) has increased over the last 30 years as a fraction of the total [@anderson_benefits].
-For these limited entry salmon fisheries, general trends that have emerged are that more people are participating since the early 2000s (Fig.
-A1), median revenue is increasing (across individuals, Fig.
-A2), and landed species diversity declining over time (averaged across individuals, Fig.
-A6).
-Across different salmon fisheries -- other than a slight negative effect of diversification on variability (similar to [@anderson_benefits]) -- there appears to be little correlation between diversity and revenue (Fig.
+For these limited entry salmon fisheries, general trends since the early 2000s are that that more people are participating, landings are increasing, revenue is level or increasing slightly, and effective species diversity is declining over time (Fig.
+1; Figs.
+A1-A4).
+Across different salmon fisheries, there appears to be little correlation between diversity and revenue, though fisheries cluster similarly by gear on the risk - return axes (Fig.
 2).
 In terrestrial agriculture, there are numerous examples where diversification may affect the mean or CV of revenue, but not both [@purdy1997], and the same result may be true for salmon fishers.
-Further, these impacts of diversification are variable over time.
+Further, the costs or benefits of diversification may be variable over time.
 
@@
 kisutch*]), and the ability of individual salmon fishers to diversify across species is in part a function of geography and fishing gears.
-For example, salmon may be caught with purse seine gear (n = 6 areas), drift gillnets (5 areas), and set gillnets (12 areas).
+For example, salmon may be caught with purse seine gear (n = 6 areas), drift gillnets (5 areas), or set gillnets (12 areas).
 Other gears include beach seine and troll gears (though these are generally restricted to Kodiak and Southeast Alaska, respectively).
@@
 Following price declines in the 1990s and early 2000s [@adfg2007], many of these factors have been positive for salmon fishermen (increasing trends in prices and catches, Fig.
-A3, A12).
+A8-A9).
 In contrast, the late 1980s - early 1990s experienced high abundance and low prices (in part due to collapses of foreign currencies).
 Combined with the increased trend in hatchery production and resulting changes in species composition of adult salmon, these external factors may have also altered incentives or opportunities for diversification.
+
 Tradeoffs between specialization and diversification may also be found in case studies of terrestrial agriculture.
@@
 
-The objectives of our analysis are to (1) evaluate long term trends in stability for fleets and individuals to evaluate whether hatchery production has increased stability, and (2) to better understand the effects of diversification on single-permit salmon holders by extending the time series models of individual revenues used by [@anderson_benefits] to include time-varying effects.
+The objectives of our analysis are to (1) evaluate long term trends in stability and to evaluate whether hatchery production has increased stability, and (2) to better understand the effects of diversification on single-permit salmon holders by extending the time series models of individual revenues used by [@anderson_benefits] to include time-varying effects.
 These time-varying effects of diversification will allow us to compare temporal trends across regions and identify periods when having diverse catch portfolio was most beneficial.
@@
 Reported gross earnings (revenue) associated with each fish ticket were adjusted for inflation by converting all revenues too 2009 USD.
+Each fish ticket was associated with both a vessel and individual fisher, and because of our focus on individuals, we aggregated data by individuals to generate annual summaries of catch diversity and revenue.
 Because our modeling focus involves estimating time varying effects of diversification, we limited our analysis to salmon fisheries with species diversity > 1 (omitting fisheries such as the Bristol Bay gillnet fisheries that specialize on sockeye salmon).
@@
 m[6,] = c("S03E", "Gillnet", "Prince William Sound")
-colnames(m) = c("Permit", "Gear", "Region")
+colnames(m) = c("Fishery (permit)", "Gear", "Region")
 kable(m)
@@
 
+To understanding long term change in variability of Alaska salmon fisheries, it is important to understand how effects change across scales, from individuals to fisheries to communities.
+For consistency with previous work, we examined the stability in the catch or landings in weight of salmon species [@hilborn2001a; @knapp2007], rather than revenue.
+We quantified the effects of changing hatchery production at three levels of data aggregation: (1) regions (combining fisheries), (2) between fisheries (defined by permits), and (3) within fisheries, for average fishers.
+We expect different results in examining the effects of fisheries versus individuals, because for a given fishery with no difference in landings over time, but changing participation, individuals would experience changing variability.
+We relied on previously published data as the response variable in comparisons across fisheries (Fig.
+2), but the CFEC data for the remaining calculations.
 Hatchery release trends for some regions included in our data have been summarized previously (e.g.
-[@ward_evaluating] summarized the trends for Prince William Sound).
-Release patterns have been dominated by pink and chum salmon ([@stopha2016], Fig.
-3) and production has increased since the late 1970s (217 million juvenile pink salmon in PWS in 1985 versus 672 million in 2014).
-In addition to examining temporal trends, we can use the spatial contrast between areas to quantify potential changes resulting from increased hatchery production (Prince William Sound, Kodiak, Southeast Alaska).
+[@ward_evaluating] summarized the trends for Prince William Sound; [@stopha2016]; Fig.
+3).
+Because hatchery contribution estimates to individual fisheries are calculated by region, rather than by fishery, we assumed that fisheries in the same region shared a common proportion of hatchery fish (e.g.
+seine and gillnet gears in Southeast Alaska).
 
-To examine long term stability in salmon fisheries, we summarized the trends in landings (total pounds) and revenue (2009 USD) at two levels.
-First, we calculated annual summaries of catch and revenue at the fishery level (annual values summed across all permit holders) and calculated 3-year running coefficients of variation (CV) by fishery.
-Examining fishery-level summaries alone may be problematic, because increases in individual variability may have a stabilizing effect at larger scales (communities, fisheries).
-As a second metric of variability, we calculated the 3-year running CV for every unique permit holder in each fishery, and averaged those individual CVs across permit holders.
-To contrast patterns across species, we repeated these calculations, computing the annual CV of revenue and prices separately for each species and fishery combination.
+To examine the effects of hatchery production on stability, we used a hierarchical variance function regression model.
+Variance function models (or heteroscedastic regression models, [@smyth1989]) are typically regression models that also allow covariates to be included (similar to performing sequential regressions on the mean and residuals [@western2009]).
+For a given level of aggregation (e.g.
+across fisheries) we modeled the mean catch of each species in each fishery (or region) as a AR(1) process, $$E[{ Y }_{ s,f,t }]={ Y }_{ s,f,t-1 }+{ b }_{ s,f }{ X }_{ s,f,t }+{ \phi  }_{ s,f }{ \varepsilon  }_{ s,f,t-1 }$$ where ${ Y }_{ s,f,t-1 }$ is the observed catch of species *s* in fishery *f* in the previous time step, ${ b }_{ s,f }$ is an estimated coefficient relating hatchery influence to the mean catch, and ${ X }_{ s,f,t }$ is the proportion of hatchery contribution to the total catch for species *s* in time *t* (0 to 1).
+The hatchery effect coefficients were modeled hierarchically, and assumed that ${ b }_{ s,f }\sim N\left( { u }_{ b },{ \sigma  }_{ b } \right)$, where ${ u }_{ b }$ and ${ \sigma  }_{ b }$ represent the location and scale hyperparameters.
+Finally, to reduce negative correlation in the residuals at lag-1, we included a moving average component with estimated parameter ${ \phi  }_{ s,f }$, where ${ \varepsilon  }_{ s,f,t-1 }$ represents the residual from the previous time step, ${ { \varepsilon  }_{ s,f,t-1 }=Y }_{ s,f,t-1 }-E\left[ { Y }_{ s,f,t-1 } \right]$.
+
+To incorporate potential hatchery effects in the variability of the system, the second component of our model included a hierarchical component on the process variance terms, $$log\left( { \sigma  }_{ s,f,t } \right) ={ g  }_{ 0,s,f }+{ g  }_{ 1,s,f }{ X }_{ s,f,t }$$ where ${ g  }_{ 0,s,f }$ represents an estimated intercept unique to each fishery and species, and ${ g  }_{ 1,s,f }$ represents a fishery and species specific slope.
+We included the intercepts { g  }_{ 0,s,f } as fixed effects (representing differences by species) and again modeled the hatchery effect as hierarchical, ${ g }_{ 1,s,f }\sim N\left( { u }_{ g },{ \sigma  }_{ g } \right)$.
+The observed data were log-transformed and linked to the predicted mean and standard deviation with a normal likelihood, $ln\left( { Y }_{ s,f,t } \right) \sim N\left( E\left[ ln\left( { Y }_{ s,f,t } \right)  \right] ,{ \sigma  }_{ s,f,t } \right)$.
+
+We focued inference on the slopes of the fraction of hatchery contribution to the total catch, ${ g  }_{ 1,s }$.
+To allow for comparison across fisheries and scales of data, we converted these estimates to effect sizes, by calculating ${ \kappa  }_{ s }=ln\left( \frac { { g }_{ 0,s }+{ \left( 1/3 \right) \cdot g }_{ 1,s } }{ { g }_{ 0,s } }  \right)$.
+The log ratio of these effect sizes represent the approximate natural log of the percent change in the standard deviation resulting from a change in the hatchery contribution of zero to 1/3 of the catch.
+95% credible intervals were generated by repeating this calculation for each MCMC draw, and taking the quantiles of the resulting distribution.
 
@@
 Because we were only interested in single permit fisheries, we applied a simplified version of the CFEC data filtering steps described in [@anderson_benefits].
-Individual trip-level catch and revenue statistics were aggregated to annual data, and we retained individual fishers who only participated in one fishery in a year (Table 1).
-Annual revenues were log-transformed to remove non-stationary individuals; to avoid complications in modeling individual switching permits, we only included pairs of years for a given person where they did not change permits (e.g.
+Individual trip-level catch and revenue statistics were aggregated to annual data, and to clarify our comparison across fisheries we retained individual fishers who only participated in one fishery in a year (Table 1).
+Annual revenues were log-transformed to make the data approximately normally distributed; to avoid complications in modeling individual switching permits, we only included pairs of years for a given person where they did not change permits (e.g.
 didn’t change fisheries between two years).
@@
 We assumed the residuals from the mean model described above to be normally distributed, $$\log\left( { R }_{ i,t } \right) \sim \mathrm{Normal}\left( \widehat { \log\left( { R }_{ i,t } \right)  } ,\widehat { { \sigma  }_{ i,t } }  \right) $$.
-While conventional hierarchical linear regression assumes variances to be constant or shared by grouping variables, we also included fixed and random covariates, making the model akin to variance function regression or stochastic volatility models (CITE).
+While conventional hierarchical linear regression assumes variances to be constant or shared by grouping variables, we also included fixed and random covariates, making the model akin to variance function regression or stochastic volatility models [@smyth1989; @western2009].
 Using the same notation as for the mean, we can express the log standard deviation as $$\widehat { \log\left( { \sigma  }_{ i,t } \right)  } ={ G }_{ 0,j }+{ G }_{ 1,j,t }\cdot \Delta S_{ i,t }+{ G }_{ 2,j,t }\cdot \left( \Delta S_{ i,t }\cdot { S }_{ i,t-1 } \right) +G_{ 3,j }\cdot \Delta D_{ i,t }$$.
@@
 
-All estimation was done in a Bayesian framework using R [@rcoreteam2016] and the package 'rstan' implementing using the NUTS algorithm [@hoffman2014].
+### Estimation
 
-Specific things that could be looked at: 
+Estimation for both sets of models was done in a Bayesian framework using R [@rcoreteam2016] and the package 'rstan', implementing Markov chain Monte Carlo (MCMC) using the No-U Turn Sampling (NUTS) algorithm in Stan [@hoffman2014].
+Code to replicate our analysis, including additional diagnostics plots, and additional models is available at [https://github.com/NCEAS/pfx-commercial/tree/master/salmon].
+
+## Results
+
+Across species and scales of data (regions, fisheries, average fishers) we found no consistent support for the hypothesis that salmon catches have been stabilized by hatchery production.
+When total catches are aggregated across fisheries in the four regions in our analysis, we estimate generally positive effects of increased hatchery production on year-to-year variability in catches (Fig.
+4).
+Other than Chinook salmon, estimates (+/- standard errors) are positive for all species in PWS and Southeast Alaska (Fig.
+4).
+When we examine the same effects at finer scales (individual fisheries, or average fishers), the standard errors generally increase, and effects become weaker in magnitude.
+Some of the estimated effects switch signs entirely -- for example, hatchery production of chum salmon is estimated to have a positive effect on variability in catches in Prince William Sound, but at the level of fisheries or average fishers, increased contributions of hatchery chum salmon appear to reduce variability (Fig.
+4).
+Other effect sizes appear consistent across scales of data; for example increased sockeye salmon in PWS is expected to increase year-to-year variability across fisheries, within the seine and gillnet fisheries (S01E, S03E) and for average fishers (Fig.
+4).
+
+Our model of changing effects of diversification over time suggests that as diversity in these six fisheries have declined over time (Figs.
+A3-A4), higher revenue is associated with increased specialization (Fig.
+5).
+The negative estimated effects on mean revenue of increasing diversity implies that for the majority of these fisheries in most years, revenue increases when individuals specialize, rather than diversify.
+Though gillnet fisheries in Southeast Alaska and Prince William Sound have also had reduced benefits of diversifying, each fishery had several periods where it was more profitable to increase catch diversity (mid 1980s-early 1990s for Southeast Alaska, mid-1990s for PWS).
+With the exception of the Southeast Alaska seine fishery (S01A) in recent years, the negative trends in these estimated effects for most fisheries suggest that the benefits of specializing on revenue have also increased over time.
+
+[Maybe tie in to Anderson et al.] We found little support for increasing catch diversity reducing variability in revenue for most fisheries (Fig.
+A5).
+For seine fisheries in Southeast Alaska and Prince William Sound, there have been periods over the last several decades where diversification has increased variability, but this effect has been strongest over the most recent decade (Fig.
+A5).
+Perhaps the only example of a fishery where individuals have seen reduced variability from increasing catch diversity is the Kodiak seine fishery (this effect is strongest in the 1990s and last 5 years).
+While these effects of diversifying are near zero for most fisheries and don't exhibit a long term trend, an exception is the PWS seine fishery, which appears to have increasing risk associated with diversifying over the time series (Fig.
+A5).
+
+## Discussion
+
+Understanding the mechanisms responsible for trends or variability in the revenue of individuals whose incomes are dependent on natural resources is important for coupled social-ecological systems, because their income variability is typically greater than the larger population [@mishra2002].
+There are a number of ways that fishers in particular may reduce their exposure to risk; examples include holding permits that allow multiple species to be targeted, or buying multiple single-species permits to increase the number of species in their catch portfolio [@anderson_benefits; @kasperski2013].
+Just as [@anderson_benefits] found that catch diversity in Alaska was decreasing over time as a greater fraction of the fisher population held single salmon permits, we found within-fishery declines in catch diversity for the six fisheries included in our analysis.
+
+As catch diversity has declined, average revenues have remained relatively flat, but average landed weight has increased over the last 40 years.
+In other words, fishers have had to catch more to yield the same revenue, and their catches tend to be driven by 1 or 2 species.
+Like in other fisheries, the ability of fishers to diversify may be largely beyond their control.
+Options for diversification include fishing in different periods of the season (e.g.
+fishing earlier in the year in the PWS seine fishery to target chum salmon), or fishing in different areas to increase catch diversity.
+For example in the PWS seine fishery, individuals are generally confied to areas where pink and chum salmon are present, while gillnet fishers may fish in areas with all five species of salmon present (Rich can provide a citation for the PWS management plan).
+
 [Rich: maybe look at PWS pink salmon crash in 1992/1993: Pacific Salmon & their Ecosystems: Status and Future Options.
@@
 
-## Results
-
-Our summaries of changing variability (aggregate for each fishery, or averaged across individuals) are not consistent with the hypothesis that variability in revenue or landings has decreased over time (Fig.
-A9-A10).
-For pink and chum salmon, fisheries in PWS generally experience higher year-to-year variability compared to SEAK.
-In looking at the changing CV of individuals, we see a slight increase in the CV of revenue for people fishing in the recent decade who were also fishing 1985-1995 (with the exception being S01A).
-
-Figure 1 and 2 shows the time series of species composition and effective spp diversity. Interesting patterns are (1) overall low diversity of S01E (pinks), (2) increasing chum in S03A
-
-Figure 3 shows that one component of variability is the year-to-year variation that individuals experience as a result of external forces (stock abundance, prices, etc) as well as common fluctuations in revenue experienced by all permit holders (Anderson et al., n.d.).
-This shows the time varying intercepts on the mean -- negative equals losing revenue on average, while positive implies increasing revenue.
-The panels of this figure show big differences between permit types (gillnets tend to have lower year-to-year variation).
-Some permits have extreme variation; for example, S01E (PWS seine) has fluctuations between exp(-1) and exp(+1), corresponding to year-to-year changes of 36-270%.
-
-Figure 4 shows the time varying nature of the model for all permits.
-These are the time series of benefits of changing from effective species diversity of 1 to 2, and here we are isolating year-to-year differences in revenue (the intercepts) from diversification benefits.
-In some cases the effect is positive (fishing more spp yields more revenue).
-But in many cases this is not the case, and the trend is negative (indicating that specialization is more profitable over time).
-In some cases, these time series mirror the aggregate diversity in the raw time series (Fig.
-1).
-I think we also expect this result - fishermen increase diversity or not, because it's profitable.
-
-Figure 5 shows that temporal trends, by permit, in going from 1 to 2 species fished on the variance (again removing the overall intercepts).
-From [@anderson_benefits] we'd expect diversification to reduce variance.
-A few examples of how to read this: S01A is slightly positive for most of the time series, but rapidly increases in the last 10 years -- this means generalists with diverse portfolios have had increases in year-to-year variability.
-For S01E, the start and end of the time series are associated with these 'costs' to diversification.
-For S03E, we see a more gradual trend (cost of diversification increases).
-
-## Discussion
-
 - In terms of participation, why did people drop out in 2000 in SEAK and Alaska Peninsula, but not in PWS?
@@
 In addition, there are many indications that costs have increased over time as a result of requirements by fish buyers for fish of higher quality (refrigerated sea water, ice, increased handling time, more expensive vessels) more expensive permits, etc.
-
-- Revenue is going up since 2000 -- but each permit holder is fishing more (more days, more total pounds) and the trend in weight is faster than the trend in price.
-
-- Ability to specialize primarily driven by many forces beyond control of fishermen.
-But in some situations, diversification is possible by extending the fishing season to exploit additional spp (fishing earlier for chum in PWS S01E).
-Or fishing different areas to exploit more spp in S03 (SEAK).
 
@@
 
-- One potential exogenous driver related to management that can be ruled out as being responsible for increased specialization of salmon fishermen is changes in the timing or length of fishing seasons.
-Shifting seasons to later in the year may reduce catch of early run fish for example.
-However, we found no significant trend in either the season length or average season date for any fishery (Fig.
-A15).
-Also, management actions over time have generally tried to eliminate interception fisheries.
-Examples include the closure of the South Kayak Island area where sockeye bound for Cook Inlet were being harvested off the Copper River.
-Rich can provide some citations.
-
 - The decisions to implement a hatchery program for a given species has economic consequences for overall revenue and the variability of that revenue to fishers and other participants.
@@
 Bigger than cost seems to be the risk [@karlan2012]
-
-- The potential to diversify may be limited for areas with a limited number of available species or by managements actions that prohibit fishing in certain areas (e.g., PWS where seiners are generally confined to areas/times with pink and chum salmon whereas gillnet fishers are able to fish in areas with sockeye, chum, Chinook, coho, and pink salmon).
-Rich can provide a citation for the PWS management plan.
-
-- One of the outputs that can be calculated from the raw data is variability of revenue over time (e.g.
-year to year, at the level of individual & permit).
-This gets at more of the social impacts of changing opportunities, but long term trends in rising income variability have been documented in the US (Hacker 2006).
-Understanding trends in revenue of individuals whose incomes are dependent on natural resources are interesting because they tend to be more variable than that of the larger population [@mishra2002].
-Are there specific attributes that correlate with rising variability (for example the CBO has suggested variability in year to year income is negatively correlated with income [@cbo2008])
 
@@
 g1 = ggplot(people[people$Permit%in%c("S01K","S01M","S01E","S01A","S03A","S03E","S03T","S05B"),], aes(year, log10(n), color = Permit)) + geom_line() + ylab("Log10 unique people") + xlab("Year") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
+
+# trends in landed weight - metric tons
+#weight_by_species = group_by(cfec_all[cfec_all$p_fshy %in% c("S 01A", "S 01E", "S 01K", "S 01M", "S 03A", "S 03E") & cfec_all$spec %in% c("CHNK","COHO","CHUM","PINK","SOCK"), ], p_fshy, year, spec) %>%
+#summarize(w = sum(g_pounds,na.rm=T) * 0.000453592) %>% 
+#filter(substr(p_fshy,1,1)=="S")
+#saveRDS(weight_by_species, "../../portfolio/data-generated/weight_data_by_species_fishery.rds")
+#weight_by_species_people = left_join(weight_by_species, people)
+#saveRDS(weight_by_species_people, "../../portfolio/data-generated/weight_data_by_species_people.rds")
+
+# also calculate for each person, the number of unique permits held per year
+#people1 = group_by(cfec_all, year, p_holder) %>% 
+#  summarize(n = length(unique(p_fshy))) %>% 
+#  filter(n==1)
+#people1 = left_join(people1, cfec_all) %>% 
+#  filter(p_fshy %in% c("S 01A", "S 01E", "S 01K", "S 01M", "S 03A", "S 03E")) %>% 
+#  filter(spec %in% c("CHNK","COHO","PINK","SOCK","CHUM"))
+#people1 = group_by(people1, p_holder, year, p_fshy, spec) %>% 
+#  summarize(w = sum(g_earn)) %>% 
+#  group_by(p_fshy, year, spec) %>% 
+#  summarize(w = mean(w, na.rm=T))
+#saveRDS(people1, "../../salmon/data-generated/weight_data_by_species_people2.rds")
 
@@
 dat.annual$gear[which(dat.annual$gear=="04")]="Set gillnet"
-dat.annual$gear[which(dat.annual$gear=="05")]="Hand troll"
+dat.annual$gear[which(dat.annual$gear%in%c("05"))]="Hand troll"
 dat.annual$gear[which(dat.annual$gear=="15")]="Power troll"
@@
 
-```{r fig3, echo=FALSE, fig.pos="placeHere", fig.cap="Harvest of hatchery produced salmon through time and by area.
-Shown are the numbers of hatchery fish caught by fisheries, as well as the contribution of hatchery fish to total catch, by species.
-[EW notes: purpose of this plot is to show (1) increasing production in top row and (2) increasing contribution of hatchery fish to commercial catches in bottom row]"}
+\break
+
+```{r fig3, echo=FALSE, fig.pos="placeHere", fig.cap="Harvest of hatchery produced salmon through time, species, and area.
+Shown are the numbers of hatchery fish caught by fisheries (top row), and the hatchery contribution each species to total catch (bottom row).
+The Alaska Peninsula is not included, because hatchery production has been approximately zero."}
 
@@
     labels =c("Chinook","Sockeye","Coho","Pink","Chum")) + 
-  facet_wrap(~region, scale="free_y")
+  facet_wrap(~region, scale="free_y") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
 
@@
     'Coho'=cbPalette[3], 'Pink'=cbPalette[4], 'Sockeye'=cbPalette[5]), name="Species",
-    labels =c("Chinook","Sockeye","Coho","Pink","Chum"))
+    labels =c("Chinook","Sockeye","Coho","Pink","Chum")) + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
 
 gridExtra::grid.arrange(g1, g2)
+
+# join datasets together for modeling
+names(dat) = tolower(names(dat))
+names(hatch) = tolower(names(hatch))
+names(dat)[which(names(dat)=="value")] = "hatchery_harvest_millions"
+names(hatch)[which(names(hatch)=="value")] = "percent_hatchery"
+dat = left_join(dat, hatch)
+dat$total_harvest = dat$hatchery_harvest_millions / (dat$percent_hatchery/100)
+saveRDS(dat, file="../../salmon/data-generated/hatchery_harvest_millions.rds")
 ```
 
-```{r fig4, fig.pos="placeHere", fig.cap="CV of annual revenue and prices, by permit and fishery.
-Variability in year to year revenue is explained more by permits than species, while variability in prices is affected more by species.
-", echo=FALSE}
-#cfec_all = readRDS("../../portfolio/data-generated/salmon_data_for_paper.rds")
-cfec_all$year = as.numeric(substr(cfec_all$landdate, 1, 4))
+\break
 
-cfec_salmon = cfec_all[which(cfec_all$p_fshy%in% c("S 01A", "S 01E", "S 01K","S 01M","S 03A", "S 03E")),]
-cfec_salmon = filter(cfec_salmon, year != 0)
+```{r fig4, echo=FALSE, fig.pos="placeHere", fig.cap="Effects of the fraction of hatchery fish on the year-to-year variability in landed weight at three scales: (1) Combining fisheries across regions, (2) Across fisheries, and (3) Average variability of individual fishers within fisheries.
+Effects of variability are presented in terms of effect sizes, or the natural log of the standard deviation of year-to-year catches in a scenario with hatchery fish constituting 1/3 of the catch to the standard deviation in a scenario with no hatchery fish in the catch.
+Dots represent mean estimates, and error bars represent +/- 1 standard error"}
 
-s0 = group_by(cfec_salmon[cfec_salmon$spec %in% c("CHNK","COHO","PINK","SOCK","CHUM"),], p_fshy, spec, year) %>% 
-  summarize(totrev = sum(g_earn), totlbs = sum(g_pounds), totprice = mean(g_price)) %>% 
-  arrange(p_fshy, spec, year) %>% 
-  group_by(p_fshy, spec) %>% 
-  summarize(srev = sd(diff(log(totrev))), slbs = sd(diff(log(totlbs))), sprice = sd(diff(log(totprice))))
-names(s0)[which(names(s0)=="p_fshy")] = "Permit"
-names(s0)[which(names(s0)=="spec")] = "Species"
-ggplot(s0, aes(srev,sprice,col=Species,shape = Permit)) + geom_point(size=3,alpha=0.8) + ylab("CV of year-to-year price per pound") + xlab("CV of year-to-year revenue") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
+df = readRDS(file=paste0("../../salmon/data-generated/","region","_df.rds"))
+cbPalette <- c("#009E73", "#4B2E83", "#999999", "#CC79A7", "#D55E00")
 
-s0 = group_by(cfec_salmon[cfec_salmon$spec %in% c("CHNK","COHO","PINK","SOCK","CHUM"),], p_fshy, spec, year) %>% 
-  summarize(totrev = sum(g_earn), totlbs = sum(g_pounds), totprice = mean(g_price))
-write.table(s0, "../../salmon/data-generated/species_level_data.csv", row.names=FALSE, col.names=TRUE)
-```
+df$fishery[which(df$fishery=="Prince William Sound")] = "            PWS"
+df$fishery[which(df$fishery=="Southeast")] = "            SEAK"
+df$fishery[which(df$fishery=="Kodiak")] = "            Kodiak"
 
-```{r fig5, fig.pos="placeHere", fig.cap="Three-year running CVs of revenue and landings, calculated at the fleetwide level (summed over all people) and level of individual permit holders (CV averaged across permit holders).", echo=FALSE}
-# we want to plot 4 summaries, by fishery:
-# 1) variability in year-to-year total revenue
-# 2) variability in year-to-year total landings
-# 3) variability in year-to-year earnings of individuals
-# 4) variability in year-to-year landings of individuals
+p1 = ggplot(df, aes(x=fishery, y=estimate, group=species, color=species)) +
+  geom_point(position=position_dodge(0.5), alpha = 0.5, size=2) +
+  ylab("Ln ratio (effect size)") +
+  xlab("") +
+  ggtitle("Regions") +
+  geom_errorbar(aes(ymin=estimate-sd, ymax=estimate+sd),
+    position=position_dodge(0.5), alpha=0.5) +
+  scale_color_manual(values = c('Chinook'=cbPalette[1], 'Chum'=cbPalette[2],
+    'Coho'=cbPalette[3], 'Pink'=cbPalette[4], 'Sockeye'=cbPalette[5]), name="Species",
+    labels =c("Chinook","Sockeye","Coho","Pink","Chum"))+
+  geom_hline(aes(yintercept=0), linetype = 2) +
+  theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
 
-s1 = group_by(cfec_salmon, p_fshy, year) %>%
-  summarize(landings = sum(g_pounds), revenue = sum(g_earn)) %>% 
-  group_by(p_fshy) %>% 
-  mutate(cvlandings = c(NA,rollapply(landings,width=3,sd),NA)/c(NA,rollmean(landings,k=3),NA), cvrevenue = c(NA,rollapply(revenue,width=3,sd),NA)/c(NA,rollmean(revenue,k=3),NA))
-names(s1)[which(names(s1)=="p_fshy")] = "Permit"
-s1$Permit = paste0(substr(s1$Permit,1,1), substr(s1$Permit,3,5))
-write.table(s1, "../../salmon/data-generated/fishery_level_cvs.csv", row.names=FALSE, col.names=TRUE)
+df = readRDS(file=paste0("../../salmon/data-generated/","fishery","_df.rds"))
+df$fishery[which(df$fishery=="S 01A")]="SEAK-seine" 
+df$fishery[which(df$fishery=="S 01E")]="PWS-seine" 
+df$fishery[which(df$fishery=="S 01K")]="Kodiak-seine" 
+df$fishery[which(df$fishery=="S 01M")]="AK Pen.-seine" 
+df$fishery[which(df$fishery=="S 03A")]="SEAK-gillnet" 
+df$fishery[which(df$fishery=="S 03E")]="PWS-gillnet" 
+cbPalette <- c("#009E73", "#4B2E83", "#999999", "#CC79A7", "#D55E00")
 
-g1 = ggplot(s1, aes(year, cvlandings, col=Permit)) + geom_line() + xlab("Year") + ylab("CV (fleetwide landings)") + xlim(1985,2014) + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
+p2 = ggplot(df[df$fishery!="AK Pen.-seine",], aes(x=fishery, y=estimate, group=species, color=species)) +
+  geom_point(position=position_dodge(0.5), alpha = 0.5, size=2) +
+  ylab("Ln ratio (effect size)") +
+  xlab("") +
+  ggtitle("Fisheries") +
+  geom_errorbar(aes(ymin=estimate-sd, ymax=estimate+sd),
+    position=position_dodge(0.5), alpha=0.5) +
+  scale_color_manual(values = c('Chinook'=cbPalette[1], 'Chum'=cbPalette[2],
+    'Coho'=cbPalette[3], 'Pink'=cbPalette[4], 'Sockeye'=cbPalette[5]), name="Species",
+    labels =c("Chinook","Sockeye","Coho","Pink","Chum"))+
+  geom_hline(aes(yintercept=0), linetype = 2) +
+  theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
 
-g2 = ggplot(s1, aes(year, cvrevenue, col=Permit)) + geom_line() + xlab("Year") + ylab("CV (fleetwide revenue)") + xlim(1985,2014) + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
+df = readRDS(file=paste0("../../salmon/data-generated/","individual","_df.rds"))
+df$fishery[which(df$fishery=="S 01A")]="SEAK-seine" 
+df$fishery[which(df$fishery=="S 01E")]="PWS-seine" 
+df$fishery[which(df$fishery=="S 01K")]="Kodiak-seine" 
+df$fishery[which(df$fishery=="S 01M")]="AK Pen.-seine" 
+df$fishery[which(df$fishery=="S 03A")]="SEAK-gillnet" 
+df$fishery[which(df$fishery=="S 03E")]="PWS-gillnet"
 
-# CV of individual landings and revenue
-s2 = group_by(cfec_salmon, p_fshy, year, p_holder) %>%
-  summarize(landings = sum(g_pounds), revenue = sum(g_earn)) %>% 
-  arrange(p_fshy, p_holder, year)
+p3 = ggplot(df[df$fishery!="AK Pen.-seine",], aes(x=fishery, y=estimate, group=species, color=species)) +
+  geom_point(position=position_dodge(0.5), alpha = 0.5, size=2) +
+  ylab("Ln ratio (effect size)") +
+  xlab("") +
+  ggtitle("Within fisheries") +
+  geom_errorbar(aes(ymin=estimate-sd, ymax=estimate+sd),
+    position=position_dodge(0.5), alpha=0.5) +
+  scale_color_manual(values = c('Chinook'=cbPalette[1], 'Chum'=cbPalette[2],
+    'Coho'=cbPalette[3], 'Pink'=cbPalette[4], 'Sockeye'=cbPalette[5]), name="Species",
+    labels =c("Chinook","Sockeye","Coho","Pink","Chum"))+
+  geom_hline(aes(yintercept=0), linetype = 2) +
+  theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
 
-# fill in missing years with NAs
-people_fisheries = unique(paste0(s2$p_holder, ":", s2$p_fshy))
-df_nas = expand.grid("year"=sort(unique(s2$year)), "p_holder" = sort(unique(s2$p_holder)), "p_fshy" = unique(s2$p_fshy)) 
-s2 = left_join(df_nas,s2)
-
-# remove people with less than 3 years of fishing
-s2 = group_by(s2, p_holder) %>% 
-  mutate(nYear = length(unique(year))) %>% 
-  filter(nYear >= 3) %>% 
-  select(-nYear)
-
-s2 = group_by(s2, p_holder, p_fshy) %>% 
-  mutate(cvlandings = c(NA,rollapply(landings,width=3,sd),NA)/c(NA,rollmean(landings,k=3),NA), 
-    cvrevenue = c(NA,rollapply(revenue,width=3,sd),NA)/c(NA,rollmean(revenue,k=3),NA)) %>% 
-  group_by(p_fshy, year) %>% 
-      summarize(cvlandings=mean(cvlandings,na.rm=T), 
-        cvrevenue = mean(cvrevenue,na.rm=T))
-
-names(s2)[which(names(s2)=="p_fshy")] = "Permit"
-s2$Permit = paste0(substr(s2$Permit,1,1), substr(s2$Permit,3,5))
-write.table(s2, "../../salmon/data-generated/avg_individual_level_cvs.csv", row.names=FALSE, col.names=TRUE)
-
-g3 = ggplot(s2, aes(year, cvlandings, col=Permit)) + geom_line() + xlab("Year") + ylab("CV (individual landings)") + xlim(1985,2014) + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
-
-g4 = ggplot(s2, aes(year, cvrevenue, col=Permit)) + geom_line() + xlab("Year") + ylab("CV (individual revenue)") + theme_bw() + xlim(1985,2014) + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
-
-gridExtra::grid.arrange(g1,g2,g3,g4,ncol=2)
+gridExtra::grid.arrange(p1, p2, p3, ncol=2)
 
@@
 
-```{r fig6, fig.pos="placeHere", fig.cap="Time varying effects of adding 1 species on revenue (measured in ln dollars) for hypothetical individuals with species diversity ~ 1 (in other words, the effect of switching from effective catch diversity = 1 to 2).
+\break
+
+```{r fig5, fig.pos="placeHere", fig.cap="Time varying effects of adding 1 species on revenue (measured in ln dollars) for hypothetical individuals with species diversity ~ 1 (in other words, the effect of switching from effective catch diversity = 1 to 2).
 This shows declining benefits of diversification for most fisheries, S01M.", echo=FALSE}
@@
   geom_ribbon(aes(ymin=low,ymax=high), alpha=0.4) +
-  facet_wrap(~permits) + xlab("Year") + ylab("Predicted benefit of adding +1 spp (ln $)") + geom_hline(yintercept=0, color="red")
+  facet_wrap(~permits) + xlab("Year") + ylab("Predicted benefit of adding +1 spp (ln $)") + geom_hline(yintercept=0, color="red") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
 
@@
 
+\break 
+\break
+
 # References
-
-```{r figa1, echo=FALSE, fig.pos="placeHere", fig.cap="Time varying species composition and diversity, by permit.
-Species are colored as follows: green (Chinook), purple (chum), silver  (coho), pink (pink), red (sockeye)."}
-#cfec <- feather::read_feather("../../portfolio/data-generated/salmon.feather")
-#saveRDS(cfec, "portfolio/data-generated/salmon_data_for_paper.rds")
-#cfec_all = readRDS("../../portfolio/data-generated/salmon_data_for_paper.rds")
-#cfec_all$year = as.numeric(substr(cfec_all$landdate, 1, 4))
-
-# plot proportion of species landed by salmon permit
-
-g1 = group_by(cfec_all[cfec_all$p_fshy=="S 01A" & cfec_all$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
-  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
-  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
-  ylab("Species %") + xlab("") + ggtitle("S01A") + 
-scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
-    'COHO'=cbPalette[3], 'PINK'=cbPalette[4], 'SOCK'=cbPalette[5]), name="Species",
-    labels =c("Chinook","Chum","Coho","Pink","Sockeye")) + theme(legend.position="none")
-
-g2 = group_by(cfec_all[cfec_all$p_fshy=="S 01E" & cfec_all$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
-  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
-  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
-  ylab("Species %") + xlab("") + ggtitle("S01E") + theme(legend.position="none") + 
-scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
-    'COHO'=cbPalette[3], 'PINK'=cbPalette[4], 'SOCK'=cbPalette[5]))
-
-g3 = group_by(cfec_all[cfec_all$p_fshy=="S 01K" & cfec_all$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
-  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
-  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
-  ylab("Species %") + xlab("") + ggtitle("S01K") + theme(legend.position="none") + 
-scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
-    'COHO'=cbPalette[3], 'PINK'=cbPalette[4], 'SOCK'=cbPalette[5]))
-
-g4 = group_by(cfec_all[cfec_all$p_fshy=="S 01M" & cfec_all$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
-  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
-  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
-  ylab("Species %") + xlab("") + ggtitle("S01M") + theme(legend.position="none") + 
-scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
-    'COHO'=cbPalette[3], 'PINK'=cbPalette[4], 'SOCK'=cbPalette[5]), name = "Species", labels =c("Chinook","Chum","Coho","Pink","Sockeye")) + theme(legend.position="none")
-
-g5 = group_by(cfec_all[cfec_all$p_fshy=="S 03A" & cfec_all$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, p_fshy) %>%
-  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
-  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
-  ylab("Species %") + xlab("") + ggtitle("S03A") + theme(legend.position="none") + 
-scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
-    'COHO'=cbPalette[3], 'PINK'=cbPalette[4], 'SOCK'=cbPalette[5]))
-
-g6 = group_by(cfec_all[cfec_all$p_fshy=="S 03E" & cfec_all$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK") & cfec_all$year > 0,], year, spec, p_fshy) %>%
-  summarize(p = sum(g_earn)) %>% group_by(year, p_fshy) %>% mutate(p = p/sum(p), col = as.numeric(as.factor(spec))) %>%
-  ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
-  ylab("Species %") + xlab("") + ggtitle("S03E") + theme(legend.position="none") + 
-scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
-    'COHO'=cbPalette[3], 'PINK'=cbPalette[4], 'SOCK'=cbPalette[5]))
-
-dat.diff = readRDS(file="../../portfolio/data-generated/cfec-diff-for-modeling.rds")
-dat.diff = dat.diff[which(dat.diff$revenue >= 10000 & dat.diff$revenue.prev >= 10000), ]
-dat.salmon = dat.diff[dat.diff$strategy_permit%in%c("S01A", "S01E", "S01K", "S03A", "S03E","S01M"),]
-
-h1 = group_by(dat.salmon[dat.salmon$strategy_permit=="S01A",], year, strategy_permit) %>%
-  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
-  ggplot(aes(year, meanDiv)) + geom_line() + 
-  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Diversity") + xlab("") + theme(legend.position="none")
-
-h2 = group_by(dat.salmon[dat.salmon$strategy_permit=="S01E",], year, strategy_permit) %>%
-  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
-  ggplot(aes(year, meanDiv)) + geom_line() + 
-  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Diversity") + xlab("") + theme(legend.position="none")
-
-h3 = group_by(dat.salmon[dat.salmon$strategy_permit=="S01K",], year, strategy_permit) %>%
-  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
-  ggplot(aes(year, meanDiv)) + geom_line() + 
-  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Diversity") + xlab("") + theme(legend.position="none")
-
-h4 = group_by(dat.salmon[dat.salmon$strategy_permit=="S01M",], year, strategy_permit) %>%
-  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
-  ggplot(aes(year, meanDiv)) + geom_line() + 
-  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Diversity") + xlab("") + theme(legend.position="none")
-
-h5 = group_by(dat.salmon[dat.salmon$strategy_permit=="S03A",], year, strategy_permit) %>%
-  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
-  ggplot(aes(year, meanDiv)) + geom_line() + 
-  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Diversity") + xlab("") + theme(legend.position="none")
-
-h6 = group_by(dat.salmon[dat.salmon$strategy_permit=="S03E",], year, strategy_permit) %>%
-  summarize(meanDiv = mean(specDiv), low = quantile(specDiv,0.25), upp = quantile(specDiv,0.75)) %>%
-  ggplot(aes(year, meanDiv)) + geom_line() + 
-  geom_ribbon(aes(ymin= (low), ymax= (upp)), alpha=0.4) + ylab("Diversity") + xlab("") + theme(legend.position="none")
-
-gridExtra::grid.arrange(g1, g2, g3, h1, h2, h3, nrow=2)
-```
-
-```{r figa2, echo=FALSE, fig.pos="placeHere", fig.cap="Time varying species composition and diversity, by permit.", fig.height=3.2} 
-gridExtra::grid.arrange(g4, g5, g6, h4, h5, h6, nrow=2)
-```
-
-```{r figa3, fig.pos="placeHere", fig.cap="Time varying effects of adding 1 species on log standard deviation of revenue for hypothetical individuals with species diversity ~ 1.", echo=FALSE}
-# grab original data to look at permit names
-diffdat = readRDS(file="../../portfolio/data-generated/salmon-diff-for-modeling.rds")
-years = seq(min(diffdat$year), max(diffdat$year))
-#permits = levels(as.factor(diffdat$permit))
-
-load("../../salmon/analysis/model-linear.Rdata")
-yearpermit=yearpermitold
-
-est = extract(mod, permute="TRUE")
-
-# start with making time series of estimates on mean
-# challenge is that they're time varying quadratic functions
-# do facet_wrap of estimates by year
-
-# effect of adding more 1 species from 1. specdiv.prev = 1
-
-specdiv.prev = seq(1, 5, by=0.1)
-specdiv = specdiv.prev + 1 # always add 1 more in these comparisons
-x1 = log(specdiv/specdiv.prev)
-x2 = log(specdiv/specdiv.prev) * specdiv.prev
-x3 = log(specdiv/specdiv.prev) * specdiv.prev^2
-
-df = expand.grid("year"=1:length(years), "diversity" = seq(1,length(x1)), "permit"=1:6)
-df$permits = permits[df$permit]
-df$years = years[df$year]
-df$x = specdiv.prev[df$diversity]
-df$pred = 0
-df$low = 0
-df$high = 0
-
-for(i in 1:nrow(df)) {
-  #tmp = est$g0 + est$g0_strategy[,df$permit[i]] + 
-  tmp = 
-    x1[df$diversity[i]] * (est$g_1[,df$year[i], df$permit[i]])# +
-    #x2[df$diversity[i]] * (est$g_2[,df$year[i], df$permit[i]]+est$g2_str_yr_mu)
-  df$pred[i] = median(tmp)
-  df$low[i] = quantile(tmp,0.025)
-  df$high[i] = quantile(tmp,0.975)
-}
-
-ggplot(df[df$diversity==1,], aes(x=years,y=pred,group=permits)) +
-  geom_line() +
-  geom_ribbon(aes(ymin=low,ymax=high), alpha=0.4) +
-  facet_wrap(~permits) + xlab("Year") + ylab("Predicted marginal benefit of adding +1 spp (ln $)") + geom_hline(yintercept=0, color="red")
-
-```
-
-```{r figa4, echo=FALSE, fig.pos="placeHere", fig.cap="Time varying intercepts on estimated mean revenue, by permit.
-These are interpreted as the mean change in revenue for each permit holder in each year (positive equates to increasing revenue, etc), after accounting for effects of catch diversification and days fished."}
-# grab original data to look at permit names
-diffdat = readRDS(file="../../salmon/data-generated/salmon.rds")
-
-years = seq(min(diffdat$year), max(diffdat$year))
-permits = levels(as.factor(diffdat$permit))
-
-# note that we have to use factor() here instead of unique() because
-# that's the same as the ordering in the estimation code
-yearpermit = data.frame("yrpermit"=levels(as.factor(diffdat$year_permit)))
-yearpermit$year = as.numeric(substr(yearpermit$yrpermit,1,4))
-yearpermit$permit = substr(yearpermit$yrpermit,6,9)
-
-yearpermitold=yearpermit
-load("../../salmon/analysis/model-linear.Rdata")
-yearpermit=yearpermitold
-est = extract(mod, permute="TRUE")
-
-yearpermit$b0 = apply(est$b0_str_yr,2,median)
-yearpermit$b0_low = apply(est$b0_str_yr,2,quantile,0.025)
-yearpermit$b0_hi = apply(est$b0_str_yr,2,quantile,0.975)
-
-ggplot(data=yearpermit) + 
-  geom_ribbon(aes(x=year,ymin=b0_low,ymax=b0_hi),alpha=0.3) +
-  geom_line(aes(x=year,y=b0)) + 
-  facet_wrap(~permit) + xlab("Year") + 
-  ylab("Intercepts on mean") + 
-  geom_hline(yintercept=0, color="red")
-```
-
-```{r figa5, eval=FALSE,echo=FALSE}
-load("../../salmon/analysis/cv-model-simple.Rdata")
-# groups here are 80% pink vs 97% pink
-
-yrs = 1986:2014
-
-hist(extract(mod)[["g_offset"]], 40, col="grey", xlab="Effect of specializing on pink (sd)")
-
-xx = exp(extract(mod)[["sd_yr"]])
-df.gen = data.frame("year"=yrs, "mean" = apply(xx, 2, mean), "low" = apply(xx, 2, quantile, 0.025),
-  "hi" = apply(xx, 2, quantile, 0.975), "strategy"="pink:chum")
-
-offset = matrix(extract(mod)[["g_offset"]],ncol=1) %*% matrix(rep(1,length(yrs)), nrow=1)
-xx = exp(extract(mod)[["sd_yr"]] + offset)
-df.spec = data.frame("year"=yrs, "mean" = apply(xx, 2, mean), "low" = apply(xx, 2, quantile, 0.025),
-  "hi" = apply(xx, 2, quantile, 0.975), "strategy"="pink")
-
-ggplot(data=rbind(df.gen,df.spec), aes(year, mean, group=strategy, fill=strategy)) + geom_ribbon(aes(ymin=low,ymax=hi), alpha=0.5) + geom_hline(yintercept=0, color="red")
-
-```
-
-```{r figa6, echo=FALSE, eval=FALSE}
-# fig.pos="placeHere", fig.cap="Time varying effects of adding 1 species on revenue (measured in ln dollars) for purse seine permit in Prince William Sound (S01E).
-This figure illustrates two things: (1) the benefit of adding more species diminishes as a function of past diversity, and (2) the economic benefit to adding more species has been reduced through time."
-# grab original data to look at permit names
-diffdat = readRDS(file="../../salmon/data-generated/salmon.rds")
-years = seq(min(diffdat$year), max(diffdat$year))
-#permits = levels(as.factor(diffdat$permit))
-
-load("../../salmon/analysis/model-linear.Rdata")
-yearpermit=yearpermitold
-
-est = extract(mod, permute="TRUE")
-
-# start with making time series of estimates on mean
-# challenge is that they're time varying quadratic functions
-# do facet_wrap of estimates by year
-
-# effect of adding more 1 species from 1. specdiv.prev = 1
-
-specdiv.prev = seq(1, 5, by=0.1)
-specdiv = specdiv.prev + 1 # always add 1 more in these comparisons
-x1 = log(specdiv/specdiv.prev)
-x2 = log(specdiv/specdiv.prev) * specdiv.prev
-
-df = expand.grid("year"=1:29, "diversity" = seq(1,length(x1)), "permit"=1:6)
-df$permits = permits[df$permit]
-df$years = years[df$year]
-df$x = specdiv.prev[df$diversity]
-df$pred = 0
-df$low = 0
-df$high = 0
-b0_indx = match(paste(as.character(df$years), as.character(df$permits)), as.character(yearpermit$yrpermit))
-
-for(i in 1:nrow(df)) {
-  tmp = x1[df$diversity[i]] * (est$b_1[,df$year[i], df$permit[i]])# + x2[df$diversity[i]] * (est$b_2[,df$year[i], df$permit[i]]+est$b2_str_yr_mu)
-  df$pred[i] = median(tmp)
-  df$low[i] = quantile(tmp,0.025)
-  df$high[i] = quantile(tmp,0.975)
-}
-df_s01e_b1 = df
-#ggplot(df[df$permits=="S01E",], aes(x=x,y=pred,group=years)) +
-#  #geom_ribbon(aes(ymin=low,ymax=high), alpha=0.4) +
-#  geom_line() +
-#  facet_wrap(~years) + xlab("Previous species diversity") + ylab("Predicted #benefit of adding +1 spp (ln $)")
-
-```
-
-```{r figa7, echo=FALSE, fig.pos="placeHere", fig.cap = "As model validation, we can compare the estimated intercepts for PWS S01E (purse seine) representing average year to year change in revenue versus the change year over year in ln(pink salmon returns).
-Correlation between them is ~ 0.8.
-This suggests (as expected) people on average make a lot more money in years when pink salmon returns are high.
-[side note: diversification may only be beneficial here when returns and prices are both low.]", eval=FALSE}
-
-yearpermit$b0 = apply(est$b0_str_yr,2,median)
-yearpermit$b0_low = apply(est$b0_str_yr,2,quantile,0.025)
-yearpermit$b0_hi = apply(est$b0_str_yr,2,quantile,0.975)
-
-S01E = yearpermit[yearpermit$permit=="S01E",]
-names(S01E)[which(names(S01E)=="year")]="Year"
-
-returns = readxl::read_excel("../../salmon/paper/salmon_returns_PWS.xlsx")
-pinkPWS = returns[,c("Year", "Pink_total", "Chum_total")]
-pinkPWS = pinkPWS[which(pinkPWS$Year >= 1975),]
-pinkPWS$diff = c(NA, diff(log(pinkPWS$Pink_total)))
-
-pinkPWS = dplyr::left_join(pinkPWS, S01E)
-
-ggplot(data=pinkPWS, aes(b0, diff,color=Year)) + geom_point() + stat_smooth(method=lm) + xlab("Estimated intercept from model") + ylab("Change in total pink salmon returns")
-```

### 2017-02-22 salmon/analysis/varfunc_hierarchical.R

@@ -1,0 +1,130 @@
+library(reshape2)
+library(dplyr)
+library(rstan)
+library(shinystan)
+rstan_options(auto_write = TRUE)
+options(mc.cores = parallel::detectCores())
+
+for(model_name in c("individual","fishery","region")) {
+#model_name = "individual" # fishery, individual
+#weight_by_species = readRDS("portfolio/data-generated/weight_data_by_species_fishery.rds")
+#weight_by_species = readRDS("salmon/data-generated/weight_data_by_species_people2.rds")
+weight_by_species = readRDS("portfolio/data-generated/weight_data_by_species_people.rds")
+
+names(weight_by_species)[which(names(weight_by_species)=="spec")]="species"
+names(weight_by_species)[which(names(weight_by_species)=="p_fshy")]="fishery"
+weight_by_species$species[which(weight_by_species$species=="COHO")] = "Coho"
+weight_by_species$species[which(weight_by_species$species=="SOCK")] = "Sockeye"
+weight_by_species$species[which(weight_by_species$species=="CHNK")] = "Chinook"
+weight_by_species$species[which(weight_by_species$species=="PINK")] = "Pink"
+weight_by_species$species[which(weight_by_species$species=="CHUM")] = "Chum"
+weight_by_species$region = NA
+weight_by_species$region[which(weight_by_species$fishery == "S 01A")] = "Southeast"
+weight_by_species$region[which(weight_by_species$fishery == "S 03A")] = "Southeast"
+weight_by_species$region[which(weight_by_species$fishery == "S 03E")] = "Prince William Sound"
+weight_by_species$region[which(weight_by_species$fishery == "S 01E")] = "Prince William Sound"
+weight_by_species$region[which(weight_by_species$fishery == "S 01K")] = "Kodiak"
+weight_by_species$region[which(weight_by_species$fishery == "S 01M")] = "Alaska Peninsula"
+
+#weight_by_species$w = weight_by_species$w / weight_by_species$n
+
+if(model_name == "fishery") {
+  weight_by_species$w = log(weight_by_species$w)
+}
+if(model_name == "individual") {
[…]

### 2017-03-03 portfolio/text/anderson-etal-rev-pfx.tex
Unwrap and fix links

--- 2017-02-11 portfolio/text/anderson-etal-rev-pfx.tex
@@
 
-% Please add here a significance statement to explain the relevance of your work
-% <120 words
+% Please add here a significance statement to explain the relevance of your work % <120 words
 \significancestatement{
-Individuals that rely on natural resources for their livelihoods,
-such as fishers, farmers, and forestry workers,
-face high levels of income variability.
-For fishers, catching multiple species has been shown to
-reduce revenue variability at large scales (vessels and communities)
-but the individual-level consequences of maintaining catch diversity are unknown.
-Our work demonstrates that individuals in fisheries targeting
-a diversity of species
-and individuals that participate in multiple fisheries
-buffer income variability compared to less diverse individuals.
-However, large adjustments in diversification strategies from year-to-year are risky and
-usually increase revenue variability. The most effective option to
-reduce revenue variability via diversification
---- purchasing additional permits ---
-is also expensive, often limited by regulations,
-and therefore unavailable to many.
+Individuals that rely on natural resources for their livelihoods, such as fishers, farmers, and forestry workers, face high levels of income variability.
+For fishers, catching multiple species has been shown to reduce revenue variability at large scales (vessels and communities) but the individual-level consequences of maintaining catch diversity are unknown.
+Our work demonstrates that individuals in fisheries targeting a diversity of species and individuals that participate in multiple fisheries buffer income variability compared to less diverse individuals.
+However, large adjustments in diversification strategies from year-to-year are risky and usually increase revenue variability.
+The most effective option to reduce revenue variability via diversification --- purchasing additional permits --- is also expensive, often limited by regulations, and therefore unavailable to many.
 }
-% We ask how multiple forms of
-% species diversification affect individual-level
-% revenue and revenue variability for Alaskan fishers.
+% We ask how multiple forms of % species diversification affect individual-level % revenue and revenue variability for Alaskan fishers.
 
@@
 validated the model with simulated data, and were able to recover the true
-values (Supporting Information). Code used in our analysis is available at
-\url{https://github.com/NCEAS/pfx-commercial/portfolios/}. } \showmatmethods
+values (Supporting Information).
+
+The fisheries data that support the findings of this study, though
+confidential, are available from the Commercial Fisheries Entry
+Commission (\url{https://www.cfec.state.ak.us/}) and summaries used in
+this paper, along with all code, are archived in a publicly available repository at
+\url{https://github.com/NCEAS/pfx-commercial/tree/master/portfolio}. } \showmatmethods
 

### 2017-03-03 portfolio/text/anderson-etal-rev-pfx.tex
Comment out part of methods for now
Not really documented well

@@
 to the mean, and then modelled the absolute residuals from that first model in
-a second mixed-effects model representing the variance component. We also
-validated the model with simulated data, and were able to recover the true
-values (Supporting Information).
+a second mixed-effects model representing the variance component (Supporting Information).
+%We also
+%validated the model with simulated data, and were able to recover the true
+%values (Supporting Information).
 

### 2017-03-05 salmon/analysis/0b-create-modelling-data.Rmd
paper and appendix changes

--- 2016-08-16 portfolio/analysis/1-create-modelling-data.Rmd
@@
 library(date)
+library(reshape2)
+cfec <- feather::read_feather("portfolio/data-generated/salmon.feather")
 
-cfec <- feather::read_feather("portfolio/data-generated/cfec.feather")
+maxYear = 2015
+cfec = filter(cfec, year %in% seq(1975,maxYear))
+
+saveRDS(cfec, "portfolio/data-generated/salmon_data_for_paper.rds")
 
@@
 ndat$initial <- nrow(cfecnew)
+ndat$initial_tot_rev <- sum(cfecnew$g_earn)
 ```
@@
 codes$pfshy_species <- sub(" +$", "", codes$pfshy_species)
-readr::write_csv(codes, path = "portfolio/data-generated/cfec-codes.csv")
+#readr::write_csv(codes, path = "portfolio/data-generated/cfec-codes.csv")
 # as.data.frame(codes)
@@
 cfecnew = left_join(cfecnew, topstrategies) # adds strategy column
-write.csv(rev(sort(table(topstrategies$strategy))), file = "portfolio/data-generated/strategies.csv")
+#write.csv(rev(sort(table(topstrategies$strategy))), file = "portfolio/data-generated/strategies.csv")
 
@@
 # convert to wide format
-library(reshape2)
+
 melted_sbs <- melt(species_by_strategy, id.vars = c("strategy.permit", "spec"))
 species_by_strategy <- dcast(melted_sbs, strategy.permit ~ spec + variable)
-write.csv(species_by_strategy, file = "portfolio/data-generated/species_by_strategy.csv")
+#write.csv(species_by_strategy, file = "portfolio/data-generated/species_by_strategy.csv")
 
@@
       arrange(-n)
-gears %>% readr::write_csv(path = "portfolio/data-generated/gear-strategies.csv")
+#gears %>% readr::write_csv(path = "portfolio/data-generated/gear-strategies.csv")
 
@@
     weight = sum(g_pounds),
-    length = mean(vlength, na.rm=TRUE),
     strategy = strategy[1],
@@
     nregion = length(unique(region)))
+
+# for the synthesis paper :
+#saveRDS(cfecnew, file = "portfolio/data-generated/cfec-cleaned.rds")
 
@@
 
-saveRDS(cfecAnnual, file="portfolio/data-generated/cfec-annual-for-modeling.rds")
+saveRDS(cfecAnnual, file="portfolio/data-generated/salmon-annual-for-modeling.rds")
 
@@
 cfecAnnual.diff = cfecAnnual[is.na(cfecAnnual$specDiv.change)==FALSE, ]
-saveRDS(cfecAnnual.diff, file="portfolio/data-generated/cfec-diff-for-modeling.rds")
-ndat$annual_diff <- nrow(cfecAnnual.diff)
-saveRDS(ndat, file = "portfolio/data-generated/ndat.rds")
+saveRDS(cfecAnnual.diff, file="portfolio/data-generated/salmon-diff-for-modeling.rds")
+#ndat$annual_diff <- nrow(cfecAnnual.diff)
+#saveRDS(ndat, file = "portfolio/data-generated/ndat.rds")
 ```

### 2017-03-05 salmon/paper/salmon_paper_appendix_v3.Rmd
paper and appendix changes

@@
     length = upper - lower)
+
 seasons = rename(seasons, permit = p_fshy)
+
+seasons$permit[which(seasons$permit=="S 01A")] = "SEAK-seine"
+seasons$permit[which(seasons$permit=="S 01E")] = "PWS-seine"
+seasons$permit[which(seasons$permit=="S 01K")] = "Kodiak-seine"
+seasons$permit[which(seasons$permit=="S 01M")] = "AKP-seine"
+seasons$permit[which(seasons$permit=="S 03A")] = "SEAK-gillnet"
+seasons$permit[which(seasons$permit=="S 03E")] = "PWS-gillnet"
+
 appendixPlot0_p1 = ggplot(seasons, aes(year, q, color = permit, group=permit)) + geom_line() + xlab("Year") + ylab("Mean julian day")
@@
 
-```{r figa1, fig.cap="Participation (unique people)", fig.pos="placeHere", echo=FALSE}
+```{r figa1, fig.cap="Participation (unique people).
+Permits are coded by gear ('01'=seine, '03' = drift gillnet, '04' = set gillnet) and region ('A'=SEAK, 'D'=Yakutat, 'E'=PWS, 'H'=Cook Inlet, 'K'=Kodiak', 'L'=Chignik, 'M' = Alaskan Peninsula', 'T'=Bristol Bay', 'W'=Kotzebue).
+Effective species diversity is calculated as 1/Simpson's Diversity.", fig.pos="placeHere", echo=FALSE}
 print(appendixPlot1)
@@
 
-```{r figa2, fig.cap="Diversity, mean revenue and CV (aggregated across years)", fig.pos="placeHere", echo=FALSE}
+```{r figa2, fig.cap="Diversity, mean revenue and CV (aggregated across years).
+Permits are coded by gear ('01'=seine, '03' = drift gillnet, '04' = set gillnet) and region ('A'=SEAK, 'D'=Yakutat, 'E'=PWS, 'H'=Cook Inlet, 'K'=Kodiak', 'L'=Chignik, 'M' = Alaskan Peninsula', 'T'=Bristol Bay', 'W'=Kotzebue).
+Effective species diversity is calculated as 1/Simpson's Diversity.", fig.pos="placeHere", echo=FALSE}
 gridExtra::grid.arrange(appendixPlot2_p1,appendixPlot2_p2,appendixPlot2_p3, ncol=2)
@@
   ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
-  ylab("Species %") + xlab("") + ggtitle("S01A") + 
+  ylab("Species %") + xlab("") + ggtitle("SEAK-seine") + 
 scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
@@
   ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
-  ylab("Species %") + xlab("") + ggtitle("S01E") + theme(legend.position="none") + 
+  ylab("Species %") + xlab("") + ggtitle("PWS-seine") + theme(legend.position="none") + 
 scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
@@
   ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
-  ylab("Species %") + xlab("") + ggtitle("S01K") + theme(legend.position="none") + 
+  ylab("Species %") + xlab("") + ggtitle("Kodiak-seine") + theme(legend.position="none") + 
 scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
@@
   ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
-  ylab("Species %") + xlab("") + ggtitle("S01M") + theme(legend.position="none") + 
+  ylab("Species %") + xlab("") + ggtitle("AKP-seine") + theme(legend.position="none") + 
 scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
@@
   ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
-  ylab("Species %") + xlab("") + ggtitle("S03A") + theme(legend.position="none") + 
+  ylab("Species %") + xlab("") + ggtitle("SEAK-gillnet") + theme(legend.position="none") + 
 scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
@@
   ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
-  ylab("Species %") + xlab("") + ggtitle("S03E") + theme(legend.position="none") + 
+  ylab("Species %") + xlab("") + ggtitle("PWS-gillnet") + theme(legend.position="none") + 
 scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
@@
 
-```{r figa5, fig.pos="placeHere", fig.cap="Time varying effects of adding 1 species on log standard deviation of revenue for hypothetical individuals with species diversity ~ 1.", echo=FALSE}
+```{r figa5, fig.pos="placeHere", fig.cap="Time varying effects of increasing the effective species diversity of catch from 1 to 2 species, on log standard deviation of revenue.", echo=FALSE}
 
@@
 
+df$permits[which(df$permits=="S01A")] = "SEAK-seine"
+df$permits[which(df$permits=="S01E")] = "PWS-seine"
+df$permits[which(df$permits=="S01K")] = "Kodiak-seine"
+df$permits[which(df$permits=="S01M")] = "AKP-seine"
+df$permits[which(df$permits=="S03A")] = "SEAK-gillnet"
+df$permits[which(df$permits=="S03E")] = "PWS-gillnet"
+
 for(i in 1:nrow(df)) {
@@
   geom_ribbon(aes(ymin=low,ymax=high), alpha=0.4) +
-  facet_wrap(~permits) + xlab("Year") + ylab("Predicted marginal benefit of adding +1 spp (ln $)") + geom_hline(yintercept=0, color="red")
+  facet_wrap(~permits) + xlab("Year") + ylab("Predicted marginal benefit (ln $)") + geom_hline(yintercept=0, color="red")
 
@@
 
-ggplot(data=yearpermit) + 
+yearpermit2 = yearpermit
+yearpermit2$permit[which(yearpermit2$permit=="S01A")] = "SEAK-seine"
+yearpermit2$permit[which(yearpermit2$permit=="S01E")] = "PWS-seine"
+yearpermit2$permit[which(yearpermit2$permit=="S01K")] = "Kodiak-seine"
+yearpermit2$permit[which(yearpermit2$permit=="S01M")] = "AKP-seine"
+yearpermit2$permit[which(yearpermit2$permit=="S03A")] = "SEAK-gillnet"
+yearpermit2$permit[which(yearpermit2$permit=="S03E")] = "PWS-gillnet"
+
+ggplot(data=yearpermit2) + 
   geom_ribbon(aes(x=year,ymin=b0_low,ymax=b0_hi),alpha=0.3) +
@@
 This suggests (as expected) people on average make a lot more money in years when pink salmon returns are high.
-[side note: diversification may only be beneficial here when returns and prices are both low.]", eval=FALSE}
+[EW side note: diversification may only be beneficial here when returns and prices are both low.]"}
 

### 2017-03-05 salmon/paper/salmon_paper_v3.Rmd
paper and appendix changes

@@
 
-A secondary motivation for focusing on the effects of diversification for salmon fishers is to better understand the long term consequences of hatchery salmon production on the revenue and variability in revenue of individuals.
-The modern era of Alaska’s large-scale hatchery program began in the late 1970s [@stopha2016], with substantial increases in hatchery production during the 1980s following low returns [@mahnken1998; @leber2008].
-Many of these hatchery programs increased production to stabilize the variability of salmon returns [@hilborn2001a], though these programs may also have negative impacts on wild production [@wertheimer2005; @ruggerone2015].
-The intra- and inter-specific effects of large scale hatchery programs has been debated [@hilborn2001a; @hilborn1992a; @wertheimer2001; @heard_hatchery], but the effects of hatcheries have largely been focused on ecological interactions of aggregate yield to the fishery [@leber2008] or economic impacts [@boyce1993], rather than to variability experienced by individual fishermen.
-Understanding the role salmon have in buffering or stabilizing fishing revenues is critical given the responses of salmon to variable environments [@mueter2002; @schindler2010].
-
-In addition to the changing species composition of fish available to be harvested, the benefits or costs of diversification are affected by a number of exogenous drivers, including competition from markets producing farmed salmon, consolidation of salmon buyers, marketing, prices, new processing technology [@knapp_trends].
+The benefits or costs of catch diversification for fishermen are affected by a number of exogenous drivers, including competition from markets producing farmed salmon, consolidation of salmon buyers, marketing, prices, new processing technology [@knapp_trends].
 Following price declines in the 1990s and early 2000s [@adfg2007], many of these factors have been positive for salmon fishermen (increasing trends in prices and catches, Fig.
@@
 In contrast, the late 1980s - early 1990s experienced high abundance and low prices (in part due to collapses of foreign currencies).
-Combined with the increased trend in hatchery production and resulting changes in species composition of adult salmon, these external factors may have also altered incentives or opportunities for diversification.
-
-Tradeoffs between specialization and diversification may also be found in case studies of terrestrial agriculture.
-For small scale farmers, increasing crop diversity may increase revenue and stabilize year-to-year income variability [@difalco2003].
-At larger scales or in the developed world, however, the incentives for diversification may be different.
-In the United States and Canada for example, crop diversity has declined since the late-1970s [@aguilar2015; @bradshaw2004]; some of the factors responsible include changing national policies, technology, and increased demand from other parts of the world (e.g.
-increased U.S.
-production of soybeans to meet demand from Asia).
-Shifts from more diverse high value crops to specializing on high-volume low-value crops such as cereals may have analogs for salmon fishermen in low value species with rapid life histories (pink and chum salmon).
+Perhaps one of the largest changes affecting the availability or incentives for fishermen to diversify has been the increase in the production of hatchery fish.
+The modern era of Alaska’s large-scale hatchery program began in the late 1970s [@stopha2016; @ward_evaluating], with substantial increases in hatchery production during the 1980s following low returns [@mahnken1998; @leber2008].
+Many of these hatchery programs increased production to stabilize the variability of salmon catches [@hilborn2001a], though these programs may also have negative impacts on wild production [@wertheimer2005; @ruggerone2015].
+The intra- and inter-specific effects of large scale hatchery programs has been debated [@hilborn2001a; @hilborn1992a; @wertheimer2001; @heard_hatchery], but the effects of hatcheries have largely been focused on ecological interactions of aggregate yield to the fishery [@leber2008] or economic impacts [@boyce1993], rather than to variability experienced by fisheries or individual fishermen.
+Understanding the role salmon have in buffering or stabilizing fishing revenues is critical given the responses of salmon to variable environments [@mueter2002; @schindler2010].
 
@@
 
-We obtained fisheries landings and revenue data for all permit holders in Alaska from 1985-2014 from the Commercial Fisheries Entry Commission (CFEC).
+We obtained fisheries landings and revenue data for all permit holders in Alaska from 1985-2016 from the Commercial Fisheries Entry Commission (CFEC).
 Reported gross earnings (revenue) associated with each fish ticket were adjusted for inflation by converting all revenues too 2009 USD.
@@
 
-To understanding long term change in variability of Alaska salmon fisheries, it is important to understand how effects change across scales, from individuals to fisheries to communities.
-For consistency with previous work, we examined the stability in the catch or landings in weight of salmon species [@hilborn2001a; @knapp2007], rather than revenue.
-We quantified the effects of changing hatchery production at three levels of data aggregation: (1) regions (combining fisheries), (2) between fisheries (defined by permits), and (3) within fisheries, for average fishers.
-We expect different results in examining the effects of fisheries versus individuals, because for a given fishery with no difference in landings over time, but changing participation, individuals would experience changing variability.
-We relied on previously published data as the response variable in comparisons across fisheries (Fig.
-2), but the CFEC data for the remaining calculations.
-Hatchery release trends for some regions included in our data have been summarized previously (e.g.
-[@ward_evaluating] summarized the trends for Prince William Sound; [@stopha2016]; Fig.
-3).
-Because hatchery contribution estimates to individual fisheries are calculated by region, rather than by fishery, we assumed that fisheries in the same region shared a common proportion of hatchery fish (e.g.
-seine and gillnet gears in Southeast Alaska).
+To understanding long term change in variability of Alaska salmon fisheries, it is important to understand how hatchery production has affected the mean revenue and variability in revenue.
+For consistency with previous work, we examined the stability in the catch (weight of salmon species [@hilborn2001a; @knapp2007], rather than revenue.
+Within each region in our analysis (PWS, Southeast, Kodiak, Alaska Peninsula), we used the hatchery contribution to the Common Property Commercial Harvest as the predictor variable, and total harvest as the response.
+These quantities are reported by region ([@stopha2016]; Fig.
+3), but not at the level of fisheries or individuals.
+Because reporting was not done separately for the Kodiak and Alaska Peninsula regions until after 1995, we grouped these areas into a single region.
 
@@
 Variance function models (or heteroscedastic regression models, [@smyth1989]) are typically regression models that also allow covariates to be included (similar to performing sequential regressions on the mean and residuals [@western2009]).
-For a given level of aggregation (e.g.
-across fisheries) we modeled the mean catch of each species in each fishery (or region) as a AR(1) process, $$E[{ Y }_{ s,f,t }]={ Y }_{ s,f,t-1 }+{ b }_{ s,f }{ X }_{ s,f,t }+{ \phi  }_{ s,f }{ \varepsilon  }_{ s,f,t-1 }$$ where ${ Y }_{ s,f,t-1 }$ is the observed catch of species *s* in fishery *f* in the previous time step, ${ b }_{ s,f }$ is an estimated coefficient relating hatchery influence to the mean catch, and ${ X }_{ s,f,t }$ is the proportion of hatchery contribution to the total catch for species *s* in time *t* (0 to 1).
+We modeled the mean catch of each species in each region as an AR(1) process, $$E[{ Y }_{ s,f,t }]={ Y }_{ s,f,t-1 }+{ b }_{ s,f }{ X }_{ s,f,t }$$ where ${ Y }_{ s,f,t-1 }$ is the observed catch of species *s* in region *f* in the previous time step, ${ b }_{ s,f }$ is an estimated coefficient relating hatchery influence to the mean catch, and ${ X }_{ s,f,t }$ is the change in proportion of hatchery contribution to the total catch for species *s* in between times *t* and *t-1*.
+We used the difference in hatchery contribution, rather than raw values because the response variable was also first differenced.
 The hatchery effect coefficients were modeled hierarchically, and assumed that ${ b }_{ s,f }\sim N\left( { u }_{ b },{ \sigma  }_{ b } \right)$, where ${ u }_{ b }$ and ${ \sigma  }_{ b }$ represent the location and scale hyperparameters.
-Finally, to reduce negative correlation in the residuals at lag-1, we included a moving average component with estimated parameter ${ \phi  }_{ s,f }$, where ${ \varepsilon  }_{ s,f,t-1 }$ represents the residual from the previous time step, ${ { \varepsilon  }_{ s,f,t-1 }=Y }_{ s,f,t-1 }-E\left[ { Y }_{ s,f,t-1 } \right]$.
 
-To incorporate potential hatchery effects in the variability of the system, the second component of our model included a hierarchical component on the process variance terms, $$log\left( { \sigma  }_{ s,f,t } \right) ={ g  }_{ 0,s,f }+{ g  }_{ 1,s,f }{ X }_{ s,f,t }$$ where ${ g  }_{ 0,s,f }$ represents an estimated intercept unique to each fishery and species, and ${ g  }_{ 1,s,f }$ represents a fishery and species specific slope.
+To incorporate potential hatchery effects in the variability of the system, the second component of our model included a hierarchical component on the process variance terms, $$log\left( { \sigma  }_{ s,f,t } \right) ={ g  }_{ 0,s,f }+{ g  }_{ 1,s,f }{ X }_{ s,f,t }$$ where ${ g  }_{ 0,s,f }$ represents an estimated intercept unique to each region and species, and ${ g  }_{ 1,s,f }$ represents a region and species specific slope.
+The covariate ${ X }_{ s,f,t }$ is identical to that used in the mean model.
 We included the intercepts { g  }_{ 0,s,f } as fixed effects (representing differences by species) and again modeled the hatchery effect as hierarchical, ${ g }_{ 1,s,f }\sim N\left( { u }_{ g },{ \sigma  }_{ g } \right)$.
@@
 We focued inference on the slopes of the fraction of hatchery contribution to the total catch, ${ g  }_{ 1,s }$.
-To allow for comparison across fisheries and scales of data, we converted these estimates to effect sizes, by calculating ${ \kappa  }_{ s }=ln\left( \frac { { g }_{ 0,s }+{ \left( 1/3 \right) \cdot g }_{ 1,s } }{ { g }_{ 0,s } }  \right)$.
-The log ratio of these effect sizes represent the approximate natural log of the percent change in the standard deviation resulting from a change in the hatchery contribution of zero to 1/3 of the catch.
+To allow for comparison across regions and scales of data, we converted these estimates to effect sizes, by calculating ${ \kappa  }_{ s }=ln\left( \frac { { g }_{ 0,s }+{ \left( 1/3 \right) \cdot g }_{ 1,s } }{ { g }_{ 0,s } }  \right)$.
+The log ratio of these effect sizes represent the approximate natural log of the percent change in the standard deviation resulting from a change in the hatchery contribution of zero to 10% of the catch.
 95% credible intervals were generated by repeating this calculation for each MCMC draw, and taking the quantiles of the resulting distribution.
 
-### Have effects of diversification changed over time for salmon fishers?
+### Have effects of diversification changed over time for individual salmon fishers?
 
 Because we were only interested in single permit fisheries, we applied a simplified version of the CFEC data filtering steps described in [@anderson_benefits].
-Individual trip-level catch and revenue statistics were aggregated to annual data, and to clarify our comparison across fisheries we retained individual fishers who only participated in one fishery in a year (Table 1).
+Individual trip-level catch and revenue data 1975-2016 were aggregated to annual data for each fisher, and to clarify our comparison across fisheries we retained individual fishers who only participated in one fishery in a year (Table 1).
 Annual revenues were log-transformed to make the data approximately normally distributed; to avoid complications in modeling individual switching permits, we only included pairs of years for a given person where they did not change permits (e.g.
@@
 
+Tradeoffs between specialization and diversification may also be found in case studies of terrestrial agriculture.
+For small scale farmers, increasing crop diversity may increase revenue and stabilize year-to-year income variability [@difalco2003].
+At larger scales or in the developed world, however, the incentives for diversification may be different.
+In the United States and Canada for example, crop diversity has declined since the late-1970s [@aguilar2015; @bradshaw2004]; some of the factors responsible include changing national policies, technology, and increased demand from other parts of the world (e.g.
+increased U.S.
+production of soybeans to meet demand from Asia).
+Shifts from more diverse high value crops to specializing on high-volume low-value crops such as cereals may have analogs for salmon fishermen in low value species with rapid life histories (pink and chum salmon).
+
 [Rich: maybe look at PWS pink salmon crash in 1992/1993: Pacific Salmon & their Ecosystems: Status and Future Options.
@@
 
-g1 = ggplot(people[people$Permit%in%c("S01K","S01M","S01E","S01A","S03A","S03E","S03T","S05B"),], aes(year, log10(n), color = Permit)) + geom_line() + ylab("Log10 unique people") + xlab("Year") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
+people2 = people[people$Permit%in%c("S01K","S01M","S01E","S01A","S03A","S03E","S03T","S05B"),]
+people2$Permit[which(people2$Permit=="S01K")] = "Kodiak-seine"
+people2$Permit[which(people2$Permit=="S01M")] = "AKP-seine"
+people2$Permit[which(people2$Permit=="S01E")] = "PWS-seine"
+people2$Permit[which(people2$Permit=="S01A")] = "SEAK-seine"
+people2$Permit[which(people2$Permit=="S03A")] = "SEAK-gillnet"
+people2$Permit[which(people2$Permit=="S03E")] = "PWS-gillnet"
+people2$Permit[which(people2$Permit=="S03T")] = "BB-gillnet"
+people2$Permit[which(people2$Permit=="S05B")] = "State-troll"
+
+g1 = ggplot(people2, aes(year, log10(n), color = Permit)) + geom_line() + ylab("Log10 unique people") + xlab("Year") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
 
@@
 weight = left_join(weight, people)
-g2 = ggplot(weight[weight$Permit%in%c("S01K","S01M","S01E","S01A","S03A","S03E","S03T","S05B"),], aes(year, log10(w/n), color = Permit)) + geom_line() + ylab("Log10 tons per person") + xlab("Year") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
+
+weight2 = weight[weight$Permit%in%c("S01K","S01M","S01E","S01A","S03A","S03E","S03T","S05B"),]
+weight2$Permit[which(weight2$Permit=="S01K")] = "Kodiak-seine"
+weight2$Permit[which(weight2$Permit=="S01M")] = "AKP-seine"
+weight2$Permit[which(weight2$Permit=="S01E")] = "PWS-seine"
+weight2$Permit[which(weight2$Permit=="S01A")] = "SEAK-seine"
+weight2$Permit[which(weight2$Permit=="S03A")] = "SEAK-gillnet"
+weight2$Permit[which(weight2$Permit=="S03E")] = "PWS-gillnet"
+weight2$Permit[which(weight2$Permit=="S03T")] = "BB-gillnet"
+weight2$Permit[which(weight2$Permit=="S05B")] = "State-troll"
+
+g2 = ggplot(weight2, aes(year, log10(w/n), color = Permit)) + geom_line() + ylab("Log10 tons per person") + xlab("Year") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
 
@@
 
-g3 = ggplot(dat.annual[dat.annual$Permit%in%c("S01K","S01M","S01E","S01A","S03A","S03E","S03T","S05B"),], aes(year, log10(meanrev), color = Permit)) + geom_line() + ylab("Log10 revenue per person") + xlab("Year") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
+dat.annual2 = dat.annual[dat.annual$Permit%in%c("S01K","S01M","S01E","S01A","S03A","S03E","S03T","S05B"),]
+dat.annual2$Permit[which(dat.annual2$Permit=="S01K")] = "Kodiak-seine"
+dat.annual2$Permit[which(dat.annual2$Permit=="S01M")] = "AKP-seine"
+dat.annual2$Permit[which(dat.annual2$Permit=="S01E")] = "PWS-seine"
+dat.annual2$Permit[which(dat.annual2$Permit=="S01A")] = "SEAK-seine"
+dat.annual2$Permit[which(dat.annual2$Permit=="S03A")] = "SEAK-gillnet"
+dat.annual2$Permit[which(dat.annual2$Permit=="S03E")] = "PWS-gillnet"
+dat.annual2$Permit[which(dat.annual2$Permit=="S03T")] = "BB-gillnet"
+dat.annual2$Permit[which(dat.annual2$Permit=="S05B")] = "State-troll"
 
-g4 = ggplot(dat.annual[dat.annual$Permit%in%c("S01K","S01M","S01E","S01A","S03A","S03E","S03T","S05B"),], aes(year, meandiv, color = Permit)) + geom_line() + ylab("Species diversity") + xlab("Year") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
+g3 = ggplot(dat.annual2, aes(year, log10(meanrev), color = Permit)) + geom_line() + ylab("Log10 revenue per person") + xlab("Year") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
+
+g4 = ggplot(dat.annual2, aes(year, meandiv, color = Permit)) + geom_line() + ylab("Species diversity") + xlab("Year") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
 
@@
 
-```{r fig2, echo=FALSE, fig.pos="placeHere", fig.cap="CV and mean annual revenue by salmon permit, 1975-2014.
+```{r fig2, echo=FALSE, fig.pos="placeHere", fig.cap="CV and mean annual revenue by salmon permit, 1975-2016.
 Average participation (unique fishers) given by circle size, and effective species diversity is shown with color shading.
@@
 pws_hatch$region = "Prince William Sound"
-kodiak_hatch$region = "Kodiak"
+kodiak_hatch$region = "Kodiak - Alaska Peninsula"
 
@@
 # bring in total harvest data (hatchery + wild)
-hatch = read.csv("../../salmon/data-generated/commonPropertyCommercialCatch.csv")
-hatch = hatch[,c("Year","Chinook","Sockeye","Coho","Pink","Chum","Region")]
+hatch = read.csv("../../salmon/data-generated/commonPropertyCommercialCatch.csv",
+  stringsAsFactors = FALSE)
+
+kodap1 = hatch[which(hatch$Region%in%c("Kodiak") & hatch$Year <= 1995),]
+kodap2 = hatch[which(hatch$Region%in%c("AP", "Kodiak") & hatch$Year > 1995),]
+kodap2 = dplyr::group_by(kodap2, Year) %>% 
+  summarize(Chinook_harvest = sum(Chinook_harvest),
+    Sockeye_harvest = sum(Sockeye_harvest),
+    Coho_harvest = sum(Coho_harvest),
+    Pink_harvest = sum(Pink_harvest),
+    Chum_harvest = sum(Chum_harvest),
+    Region = Region[1],
+    Source = Source[1],
+    ReportNumber = ReportNumber[1],
+    Chinook_enhanced = sum(Chinook_enhanced),
+    Sockeye_enhanced = sum(Sockeye_enhanced),
+    Coho_enhanced = sum(Coho_enhanced),
+    Pink_enhanced = sum(Pink_enhanced),
+    Chum_enhanced = sum(Chum_enhanced))
+kodap = rbind(as.data.frame(kodap1), as.data.frame(kodap2))
+kodap$Region = as.character(kodap$Region)
+kodap$Source = as.character(kodap$Source)
+kodap$ReportNumber = as.character(kodap$ReportNumber)
+
+hatch = rbind(hatch[which(hatch$Region %in% c("PWS","Southeast")),], kodap)
+
+hatch$pChinook = hatch$Chinook_enhanced/hatch$Chinook_harvest
+hatch$pSockeye = hatch$Sockeye_enhanced/hatch$Sockeye_harvest
+hatch$pCoho = hatch$Coho_enhanced/hatch$Coho_harvest
+hatch$pPink = hatch$Pink_enhanced/hatch$Pink_harvest
+hatch$pChum = hatch$Chum_enhanced/hatch$Chum_harvest
+
+hatch = hatch[,c("Year","pChinook","pSockeye","pCoho","pPink","pChum","Region")]
+hatch = dplyr::rename(hatch, Chinook=pChinook,Sockeye=pSockeye, Coho=pCoho, Pink=pPink, Chum=pChum)
 hatch$Region = as.character(hatch$Region)
 hatch$Region[which(hatch$Region=="PWS")] = "Prince William Sound"
+hatch$Region[which(hatch$Region=="Kodiak")] = "Kodiak - Alaska Peninsula"
 hatch = melt(hatch, id.vars = c("Year","Region"))
@@
   geom_line() + facet_wrap(~Region) + 
-  xlab("Year") + ylab("% of total catch") + 
+  xlab("Year") + ylab("Hatchery fraction of total catch") + 
   scale_color_manual(values = c('Chinook'=cbPalette[1], 'Chum'=cbPalette[2], 
@@
 ```{r fig4, echo=FALSE, fig.pos="placeHere", fig.cap="Effects of the fraction of hatchery fish on the year-to-year variability in landed weight at three scales: (1) Combining fisheries across regions, (2) Across fisheries, and (3) Average variability of individual fishers within fisheries.
-Effects of variability are presented in terms of effect sizes, or the natural log of the standard deviation of year-to-year catches in a scenario with hatchery fish constituting 1/3 of the catch to the standard deviation in a scenario with no hatchery fish in the catch.
+Effects of variability are presented in terms of effect sizes, or the natural log of the standard deviation of year-to-year catches in a scenario with hatchery fish constituting 10% of the catch to the standard deviation in a scenario with no hatchery fish in the catch.
 Dots represent mean estimates, and error bars represent +/- 1 standard error"}
@@
 
-df$fishery[which(df$fishery=="Prince William Sound")] = "            PWS"
-df$fishery[which(df$fishery=="Southeast")] = "            SEAK"
-df$fishery[which(df$fishery=="Kodiak")] = "            Kodiak"
+df$fishery[which(df$fishery=="Prince William Sound")] = "PWS"
+df$fishery[which(df$fishery=="Southeast")] = "SEAK"
+df$fishery[which(df$fishery=="Kodiak")] = "Kodiak"
 
-p1 = ggplot(df, aes(x=fishery, y=estimate, group=species, color=species)) +
-  geom_point(position=position_dodge(0.5), alpha = 0.5, size=2) +
-  ylab("Ln ratio (effect size)") +
+df$fishery[which(df$fishery=="Prince William Sound")] = "PWS"
+df$fishery[which(df$fishery=="Southeast")] = "SEAK"
+df$fishery[which(df$fishery=="Kodiak")] = "Kodiak"
+
+p1 = ggplot(df, aes(x=fishery, y=bestimate, group=species, color=species)) +
+  geom_point(position=position_dodge(0.5), alpha = 0.5, size=3) +
+  ylab("Effect size (mean)") +
   xlab("") +
-  ggtitle("Regions") +
+  geom_errorbar(aes(ymin=bestimate-bsd, ymax=bestimate+bsd),
+    position=position_dodge(0.5), alpha=0.5) +
+  scale_color_manual(values = c('Chinook'=cbPalette[1], 'Chum'=cbPalette[2],
+    'Coho'=cbPalette[3], 'Pink'=cbPalette[4], 'Sockeye'=cbPalette[5]), name="Species",
+    labels =c("Chinook","Sockeye","Coho","Pink","Chum"))+
+  geom_hline(aes(yintercept=0), linetype = 2) +
+  theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
+  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
+    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
+
+p2 = ggplot(df, aes(x=fishery, y=estimate, group=species, color=species)) +
+  geom_point(position=position_dodge(0.5), alpha = 0.5, size=3) +
+  ylab("Effect size (sd)") +
+  xlab("") +
   geom_errorbar(aes(ymin=estimate-sd, ymax=estimate+sd),
@@
 
-df = readRDS(file=paste0("../../salmon/data-generated/","fishery","_df.rds"))
-df$fishery[which(df$fishery=="S 01A")]="SEAK-seine" 
-df$fishery[which(df$fishery=="S 01E")]="PWS-seine" 
-df$fishery[which(df$fishery=="S 01K")]="Kodiak-seine" 
-df$fishery[which(df$fishery=="S 01M")]="AK Pen.-seine" 
-df$fishery[which(df$fishery=="S 03A")]="SEAK-gillnet" 
-df$fishery[which(df$fishery=="S 03E")]="PWS-gillnet" 
-cbPalette <- c("#009E73", "#4B2E83", "#999999", "#CC79A7", "#D55E00")
-
-p2 = ggplot(df[df$fishery!="AK Pen.-seine",], aes(x=fishery, y=estimate, group=species, color=species)) +
-  geom_point(position=position_dodge(0.5), alpha = 0.5, size=2) +
-  ylab("Ln ratio (effect size)") +
-  xlab("") +
-  ggtitle("Fisheries") +
-  geom_errorbar(aes(ymin=estimate-sd, ymax=estimate+sd),
-    position=position_dodge(0.5), alpha=0.5) +
-  scale_color_manual(values = c('Chinook'=cbPalette[1], 'Chum'=cbPalette[2],
-    'Coho'=cbPalette[3], 'Pink'=cbPalette[4], 'Sockeye'=cbPalette[5]), name="Species",
-    labels =c("Chinook","Sockeye","Coho","Pink","Chum"))+
-  geom_hline(aes(yintercept=0), linetype = 2) +
-  theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
-  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
-
-df = readRDS(file=paste0("../../salmon/data-generated/","individual","_df.rds"))
-df$fishery[which(df$fishery=="S 01A")]="SEAK-seine" 
-df$fishery[which(df$fishery=="S 01E")]="PWS-seine" 
-df$fishery[which(df$fishery=="S 01K")]="Kodiak-seine" 
-df$fishery[which(df$fishery=="S 01M")]="AK Pen.-seine" 
-df$fishery[which(df$fishery=="S 03A")]="SEAK-gillnet" 
-df$fishery[which(df$fishery=="S 03E")]="PWS-gillnet"
-
-p3 = ggplot(df[df$fishery!="AK Pen.-seine",], aes(x=fishery, y=estimate, group=species, color=species)) +
-  geom_point(position=position_dodge(0.5), alpha = 0.5, size=2) +
-  ylab("Ln ratio (effect size)") +
-  xlab("") +
-  ggtitle("Within fisheries") +
-  geom_errorbar(aes(ymin=estimate-sd, ymax=estimate+sd),
-    position=position_dodge(0.5), alpha=0.5) +
-  scale_color_manual(values = c('Chinook'=cbPalette[1], 'Chum'=cbPalette[2],
-    'Coho'=cbPalette[3], 'Pink'=cbPalette[4], 'Sockeye'=cbPalette[5]), name="Species",
-    labels =c("Chinook","Sockeye","Coho","Pink","Chum"))+
-  geom_hline(aes(yintercept=0), linetype = 2) +
-  theme_bw() + theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
-  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
-    panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
-
-gridExtra::grid.arrange(p1, p2, p3, ncol=2)
-
+gridExtra::grid.arrange(p1, p2, ncol=1)
 ```
@@
 
-```{r fig5, fig.pos="placeHere", fig.cap="Time varying effects of adding 1 species on revenue (measured in ln dollars) for hypothetical individuals with species diversity ~ 1 (in other words, the effect of switching from effective catch diversity = 1 to 2).
-This shows declining benefits of diversification for most fisheries, S01M.", echo=FALSE}
+```{r fig5, fig.pos="placeHere", fig.cap="Time varying effects of increasing the effective diversity of catch from 1 to 2 species (measured in ln dollars).
+Effective species diversity is calculated as 1/Simpson's Diversity.
+This shows declining benefits of diversification for most fisheries.", echo=FALSE}
 # make yearpermit df
@@
 
+df$permits[which(df$permits=="S01A")] = "SEAK-seine"
+df$permits[which(df$permits=="S01E")] = "PWS-seine"
+df$permits[which(df$permits=="S01K")] = "Kodiak-seine"
+df$permits[which(df$permits=="S01M")] = "AKP-seine"
+df$permits[which(df$permits=="S03A")] = "SEAK-gillnet"
+df$permits[which(df$permits=="S03E")] = "PWS-gillnet"
+
 ggplot(df[df$diversity==1,], aes(x=years,y=pred,group=permits)) +
-  geom_line() +
-  geom_ribbon(aes(ymin=low,ymax=high), alpha=0.4) +
-  facet_wrap(~permits) + xlab("Year") + ylab("Predicted benefit of adding +1 spp (ln $)") + geom_hline(yintercept=0, color="red") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
+  geom_ribbon(aes(ymin=low,ymax=high), fill="blue",alpha=0.4) +
+  geom_line(col="blue") + 
+  facet_wrap(~permits) + xlab("Year") + ylab("Benefit of catch diversity increasing 1 to 2") + geom_hline(yintercept=0, color="red", linetype = 2) + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
 

### 2017-03-05 salmon/analysis/0a-inflation-adjustment.R

@@ -1,0 +1,39 @@
+# Read in, cleanup, and write out a version of the data
+
+library(dplyr)
+load("data/cfecCleaned2_salmon.Rdata")
+names(cfec) <- tolower(names(cfec))
+cfec <- as_data_frame(cfec)
+names(cfec) <- gsub("\\.", "_", names(cfec))
+cfec$year <- as.numeric(cfec$year)
+cfec$specn_orig <- as.numeric(cfec$specn_orig)
+
+deflationTable <- read.csv("/users/eric.ward/documents/CFEC/data/deflation.csv")
+
+# Adjust price and g_earn for inflation
+cfec$year <- as.numeric(cfec$year)
+cfec$day <- substr(cfec$landdate, 6, 10)
+cfec <- inner_join(cfec, deflationTable)
+cfec <- mutate(cfec, g_price = g_price / defl, g_earn = g_earn / defl)
+cfec$defl <- NULL
+
+groupings <- read.csv("data/species-groupings.csv", stringsAsFactors = FALSE)
+names(groupings) <- tolower(names(groupings))
+names(groupings) <- gsub(" \\(spec\\)", "", names(groupings))
+groupings <- select(groupings, specn, group1)
+groupings <- rename(groupings, taxa_broad = group1)
+cfec <- left_join(cfec, groupings)
+cfec <- mutate(cfec, salmon = ifelse(spec %in%
+  c("CHNK", "SOCK", "COHO", "PINK", "CHUM"), TRUE, FALSE))
+cfec <- mutate(cfec, pollock = ifelse(spec == "PLCK", TRUE, FALSE))
+
+region <- read.csv("data/regions.csv", stringsAsFactors = FALSE)
+
+region <- region[,c("stat6","final")] #don’t include “assigned by” column
+names(region)[2] = "region"
+cfec$region <- NULL
+cfec$stat6 = as.numeric(cfec$stat6)
+cfec <- left_join(cfec, region)
+
+# devtools::install_github("wesm/feather/R")
+feather::write_feather(cfec, "portfolio/data-generated/salmon.feather")

### 2017-03-05 salmon/analysis/4b - varfunc_hierarchical.R

@@ -1,0 +1,119 @@
+library(reshape2)
+library(dplyr)
+library(rstan)
+
+rstan_options(auto_write = TRUE)
+options(mc.cores = parallel::detectCores())
+
+dat = read.csv("salmon/data-generated/commonPropertyCommercialCatch.csv")
+dat = dplyr::rename(dat, year=Year, region=Region)
+
+harvest <- melt(dat[,c("year","region","Chinook_harvest","Sockeye_harvest",
+  "Coho_harvest","Pink_harvest","Chum_harvest")], id.vars = c("year","region"))
+harvest$variable = unlist(lapply(strsplit(as.character(harvest$variable),"_"), getElement, 1))
+harvest = dplyr::rename(harvest, species=variable, w = value)
+
+enhance <- melt(dat[,c("year","region","Chinook_enhanced","Sockeye_enhanced",
+  "Coho_enhanced","Pink_enhanced","Chum_enhanced")], id.vars = c("year","region"))
+enhance$variable = unlist(lapply(strsplit(as.character(enhance$variable),"_"), getElement, 1))
+enhance = dplyr::rename(enhance, species=variable, enhanced = value)
+
+dat = dplyr::left_join(harvest, enhance)
+# combine ap and kodiak - they were combined in reports up to 1995
+kodap_1 = dat[which(dat$region=="Kodiak" & dat$year <= 1995),]
+kodap_1$region = as.character(kodap_1$region)
+kodap_1$region = "KodAP"
+kodap_2 = group_by(dat[which(dat$year > 1995 & dat$region %in% c("Kodiak","AP")),], species, year) %>%
+  summarize("region"="KodAP", w = sum(w), enhanced = sum(enhanced)) %>%
+  as.data.frame
+kodap = rbind(kodap_1[,c("year","region","species","w","enhanced")],
+  kodap_2[,c("year","region","species","w","enhanced")])
+
+dat = dat[-which(dat$region%in% c("Kodiak","AP")),]
+dat = rbind(dat, kodap)
+dat$region = as.character(dat$region)
+dat$species = as.character(dat$species)
+
+dat$percent_hatchery = dat$enhanced/dat$w
+dat = select(dat, -enhanced)
+
+dat$w = log(dat$w)
+
[…]

### 2017-03-05 salmon/analysis/process_salmon_data.R

@@ -72,35 +72,317 @@ for(y in 1976:1984) {
 }
 saveRDS(dat, paste0("processed","_allYears",".rds"))
 
-# merge in the 1985+ data
-setwd("/users/eric.ward/downloads/cfec_early")
-older = readRDS("processed_allYears.rds")
+# Process 2015 data
+#dat_2015 = haven::read_sas(paste0("/users/eric.ward/documents/CFEC/data/ge_short_column_names_2015.sas7bdat"))
+dat_2015 = haven::read_sas(paste0("/users/eric.ward/documents/CFEC/data/ge_long_column_names_2015.sas7bdat"))
 
-setwd("/users/eric.ward/documents/pfx-commercial/data")
-load("cfec_w_IDs")
-#cfecnew = group_by(cfecnew, p_holder, year) %>%
-#  mutate(npermit = length(unique(p_fshy))) %>%
-#  filter(npermit == 1) %>%
-#  select(-npermit) %>%
-#  ungroup() %>%
-#  filter(p_fshy %in% c("S 01E", "S 03A", "S 01A", "S 03E", "S 01K", "S 01M"))
-#cfecnew = cfecnew[which(cfecnew$p_fshy%in%c("S 01K", "S 01M", "S 01E", "S 01A","S 03E","S 03A")),]
-#saveRDS(cfecnew, "recent_salmon.rds")
-cfec = readRDS("recent_salmon.rds")
+dat_2015$F_PERMIT = paste0(dat_2015$ADFG_H_PERMIT_FISHERY,dat_2015$ADFG_H_PERMIT_SERIAL_NUMBER,dat_2015$ADFG_H_PERMIT_CHECK_DIGIT)
 
-# run process_raw_cfec_data header here
+dat_2015$startdt = paste0(substr(dat_2015$ADFG_H_DATE_FISHING_BEGAN,1,4), "-",
+  substr(dat_2015$ADFG_H_DATE_FISHING_BEGAN,5,6), "-",
+  substr(dat_2015$ADFG_H_DATE_FISHING_BEGAN,7,8))
 
-cfec = cfec[,-1]
-cfec$startdt = as.character(cfec$startdt)
-cfec$landdate = as.character(cfec$landdate)
+dat_2015$landdate = paste0(substr(dat_2015$ADFG_H_DATE_LANDED,1,4), "-",
+  substr(dat_2015$ADFG_H_DATE_LANDED,5,6), "-",
[…]

### 2017-03-09 salmon/analysis/06 granger_rmd.Rmd
add granger stuff for price - abundance

---
title: "granger_test"
author: ""
date: "March 6, 2017"
output: pdf_document
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message=FALSE)
library(dplyr)
library(lmtest)
```

### Load data

```{r loaddat}
# inflation adjusted price data
cfec_all = readRDS("portfolio/data-generated/salmon_data_for_paper.rds")
```

### Filter and difference data
```{r datafilter}
# which is leading/lagging between prices / abundance harvest
g = filter(cfec_all, substr(p_fshy,1,1) == "S") %>%
  filter(spec %in% c("CHUM","COHO","CHNK","SOCK","PINK")) %>%
  group_by(year, spec, p_fshy) %>%
  summarize(earn = sum(g_earn, na.rm=T),
    pounds = sum(g_pounds, na.rm=T),
    price = earn/pounds)

g = filter(g, p_fshy %in% c("S 01A", "S 01E", "S 01H", "S 01L", "S 01M", "S 03A",
  "S 03E", "S 03M", "S 03T", "S 04D", "S 04H", "S 04K", "S 04M", "S 04T", "S 05B",
  "S 15B"))

# first difference all the data so it's stationary
gdiff = arrange(g, p_fshy, spec, year) %>%
  group_by(p_fshy, spec) %>%
  mutate(dearn = c(NA,diff(log(earn))),
    dpounds = c(NA,diff(log(pounds))),
    dprice = c(NA,diff(log(price)))) %>%
  filter(!is.na(dearn))
```

### Run granger causality tests

```{r}
#out = matrix("0", length(unique(gdiff$p_fshy)), 
#  length(unique(gdiff$spec)))

df = data.frame("Fishery" = rep(NA,40), "Species" = rep(NA,40), "Lags" = rep(NA,40))
indx = 0
for(i in 1:nrow(out)) {
  for(j in 1:ncol(out)) {
    Z = filter(gdiff, p_fshy == unique(gdiff$p_fshy)[i] & spec == unique(gdiff$spec)[j]) %>% 
      data.frame()

    #em <- dynlm(egg ~ L(egg, 1) + L(egg, 2) + L(egg, 3), data = ChickEgg)
    #em2 <- update(em, . ~ . + L(chicken, 1) + L(chicken, 2) + L(chicken, 3))
    #waldtest(em, em2)

    mod1_p = grangertest(y=Z$dprice, x=Z$dpounds, order=1)[2,4]
    mod2_p = grangertest(y=Z$dprice, x=Z$dpounds, order=2)[2,4]

    mod3_p = grangertest(y=Z$dpounds, x=Z$dprice, order=1)[2,4]
    mod4_p = grangertest(y=Z$dpounds, x=Z$dprice, order=2)[2,4]    
    #mod3_p = grangertest(y=Z$dprice, x=Z$dpounds, order=3)[2,4]
    str = paste0(ifelse(mod1_p <= 0.05 & mod3_p > 0.05, "1", ""), 
      ifelse(mod2_p <= 0.05 & mod4_p > 0.05, "2", ""))
    if(str != "") {
      indx = indx + 1
      df$Lags[indx] = str
      df$Fishery[indx] = unique(gdiff$p_fshy)[i]
      df$Species[indx] = unique(gdiff$spec)[j]
    }
  }
}
df = df[!is.na(df$Fishery),]
write.table(df, "salmon/data-generated/granger_table.csv", row.names=F, col.names=T)

# make fitted values for each fishery
Z = filter(gdiff, p_fshy == "S 01A" & spec == "CHNK") %>% data.frame()
Z$lag_price = Z$dprice - lag(Z$dprice)
Z$lag_pounds = lag(Z$dpounds)
write.table(Z, "salmon/data-generated/granger_table_s01a.csv", row.names=F, col.names=T)
#em <- dynlm(dprice ~ L(dprice, 1), data = Z)
#em2 = update(em, . ~ . + L(dpounds, 1))

```

### 2017-03-09 salmon/paper/salmon_paper_appendix_v3.Rmd
rearrange appendix

@@
 dat.diff = dat.diff[which(dat.diff$revenue >= revenueThreshold & dat.diff$revenue.prev >= revenueThreshold), ]
-dat.diff = dat.diff[dat.diff$strategy_permit%in%c("S01A","S01E","S01K","S01M",
-  "S03A","S03E"),]
 
@@
     cv = sd(log(revenue/revenue.prev)))
+
+dat.diff = dat.diff[dat.diff$strategy_permit%in%c("S01A","S01E","S01K","S01M",
+  "S03A","S03E"),]
 
@@
 
-seasons$permit[which(seasons$permit=="S 01A")] = "SEAK-seine"
+seasons$permit[which(seasons$permit=="S 01A")] = "SE-seine"
 seasons$permit[which(seasons$permit=="S 01E")] = "PWS-seine"
-seasons$permit[which(seasons$permit=="S 01K")] = "Kodiak-seine"
-seasons$permit[which(seasons$permit=="S 01M")] = "AKP-seine"
-seasons$permit[which(seasons$permit=="S 03A")] = "SEAK-gillnet"
+seasons$permit[which(seasons$permit=="S 01K")] = "KO-seine"
+seasons$permit[which(seasons$permit=="S 01M")] = "AP-seine"
+seasons$permit[which(seasons$permit=="S 03A")] = "SE-gillnet"
 seasons$permit[which(seasons$permit=="S 03E")] = "PWS-gillnet"
 
-appendixPlot0_p1 = ggplot(seasons, aes(year, q, color = permit, group=permit)) + geom_line() + xlab("Year") + ylab("Mean julian day")
-appendixPlot0_p2 = ggplot(seasons, aes(year, length, color = permit, group=permit)) + geom_line() + xlab("Year") + ylab("Season length")
+appendixPlot0_p1 = ggplot(seasons, aes(year, q, color = permit, group=permit)) + geom_line() + xlab("Year") + ylab("Mean julian day") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
+panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
+
+appendixPlot0_p2 = ggplot(seasons, aes(year, length, color = permit, group=permit)) + geom_line() + xlab("Year") + ylab("Season length") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
+panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
 ```
@@
 ```{r makePlots, echo = FALSE}
-appendixPlot1 = ggplot(agg.annual, aes(year, people, colour = specDiv)) + geom_line() + facet_wrap(~ strategy_permit, scale="free") + ylab("Participation") + xlab("Year")
+# rename
+agg.annual$label=NA
+agg.annual$label[which(agg.annual$strategy_permit == "S01A")] = "SE-seine"
+agg.annual$label[which(agg.annual$strategy_permit == "S01E")] = "PWS-seine"
+agg.annual$label[which(agg.annual$strategy_permit == "S01H")] = "CI-seine"
+agg.annual$label[which(agg.annual$strategy_permit == "S01K")] = "KO-seine"
+agg.annual$label[which(agg.annual$strategy_permit == "S01L")] = "CH-seine"
+agg.annual$label[which(agg.annual$strategy_permit == "S01M")] = "AP-seine"
+agg.annual$label[which(agg.annual$strategy_permit == "S03A")] = "SE-drift"
+agg.annual$label[which(agg.annual$strategy_permit == "S03E")] = "PWS-drift"
+agg.annual$label[which(agg.annual$strategy_permit == "S03H")] = "CI-drift"
+agg.annual$label[which(agg.annual$strategy_permit == "S03M")] = "AP-drift"
+agg.annual$label[which(agg.annual$strategy_permit == "S03T")] = "BB-drift"
+agg.annual$label[which(agg.annual$strategy_permit == "S03W")] = "KU-drift"
+agg.annual$label[which(agg.annual$strategy_permit == "S04D")] = "YA-drift"
+agg.annual$label[which(agg.annual$strategy_permit == "S04H")] = "CI-set"
+agg.annual$label[which(agg.annual$strategy_permit == "S04K")] = "KO-set"
+agg.annual$label[which(agg.annual$strategy_permit == "S04M")] = "AP-set"
+agg.annual$label[which(agg.annual$strategy_permit == "S04T")] = "BB-set"
 
-appendixPlot2_p1 = ggplot(agg.diff, aes(specDiv, log(meanrev), colour = cv, label=strategy_permit)) + xlab("Species diversity") +ylab("Ln (mean revenue)") + geom_text(size=2)
-appendixPlot2_p2 = ggplot(agg.diff, aes(specDiv, cv, colour = log(meanrev), label=strategy_permit)) + xlab("Species diversity") +ylab("CV") + geom_text(size=2)
-appendixPlot2_p3 = ggplot(agg.diff, aes(log(meanrev), cv, colour = specDiv, label=strategy_permit)) + ylab("CV") + xlab("Ln (mean revenue)") + geom_text(size=2)
+agg.annual = agg.annual[-which(agg.annual$label=="KU-drift"),]
+appendixPlot1 = ggplot(agg.annual, aes(year, people, colour = specDiv)) + geom_line() + facet_wrap(~ label, scale="free") + ylab("Participation") + xlab("Year")
+
+aggdiff = agg.diff[which(agg.diff$strategy_permit %in% unique(agg.annual$strategy_permit)),]
+aggdiff$label = agg.annual$label[match(aggdiff$strategy_permit, agg.annual$strategy_permit)]
+
+appendixPlot2_p1 = ggplot(aggdiff, aes(specDiv, log(meanrev), colour = cv, label=label)) + xlab("Species diversity") +ylab("Ln (mean revenue)") + geom_text(size=2) + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
+panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
+
+appendixPlot2_p2 = ggplot(aggdiff, aes(specDiv, cv, colour = log(meanrev), label=label)) + xlab("Species diversity") +ylab("CV") + geom_text(size=2) + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
+panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
+
+appendixPlot2_p3 = ggplot(aggdiff, aes(log(meanrev), cv, colour = specDiv, label=label)) + ylab("CV") + xlab("Ln (mean revenue)") + geom_text(size=2) + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
+panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
 ```
@@
 ```{r figa1, fig.cap="Participation (unique people).
-Permits are coded by gear ('01'=seine, '03' = drift gillnet, '04' = set gillnet) and region ('A'=SEAK, 'D'=Yakutat, 'E'=PWS, 'H'=Cook Inlet, 'K'=Kodiak', 'L'=Chignik, 'M' = Alaskan Peninsula', 'T'=Bristol Bay', 'W'=Kotzebue).
+Permits are coded by gear (purse seine, set and drift gillnets) and region ('AP' = Alaska Peninsula, 'BB' = Bristol Bay, 'CH' = Chignik, 'CI' = Cook Inlet, 'KO' = Kodiak, 'KU' = Kuskokwim, 'PWS' = Prince William Sound, 'SE' = Southeast, 'YA' = Yakutat).
 Effective species diversity is calculated as 1/Simpson's Diversity.", fig.pos="placeHere", echo=FALSE}
@@
 ```{r figa2, fig.cap="Diversity, mean revenue and CV (aggregated across years).
-Permits are coded by gear ('01'=seine, '03' = drift gillnet, '04' = set gillnet) and region ('A'=SEAK, 'D'=Yakutat, 'E'=PWS, 'H'=Cook Inlet, 'K'=Kodiak', 'L'=Chignik, 'M' = Alaskan Peninsula', 'T'=Bristol Bay', 'W'=Kotzebue).
+Permits are coded by gear (purse seine, set and drift gillnets) and region ('AP' = Alaska Peninsula, 'BB' = Bristol Bay, 'CH' = Chignik, 'CI' = Cook Inlet, 'KO' = Kodiak, 'KU' = Kuskokwim, 'PWS' = Prince William Sound, 'SE' = Southeast, 'YA' = Yakutat).
 Effective species diversity is calculated as 1/Simpson's Diversity.", fig.pos="placeHere", echo=FALSE}
@@
   ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
-  ylab("Species %") + xlab("") + ggtitle("SEAK-seine") + 
+  ylab("Species %") + xlab("") + ggtitle("SE-seine") + 
 scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
@@
   ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
-  ylab("Species %") + xlab("") + ggtitle("Kodiak-seine") + theme(legend.position="none") + 
+  ylab("Species %") + xlab("") + ggtitle("KO-seine") + theme(legend.position="none") + 
 scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
@@
   ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
-  ylab("Species %") + xlab("") + ggtitle("AKP-seine") + theme(legend.position="none") + 
+  ylab("Species %") + xlab("") + ggtitle("AP-seine") + theme(legend.position="none") + 
 scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
@@
   ggplot(aes(year, p, group=spec, color = spec)) + geom_line() +  
-  ylab("Species %") + xlab("") + ggtitle("SEAK-gillnet") + theme(legend.position="none") + 
+  ylab("Species %") + xlab("") + ggtitle("SE-gillnet") + theme(legend.position="none") + 
 scale_colour_manual(values = c('CHNK'=cbPalette[1], 'CHUM'=cbPalette[2], 
@@
 
-```{r figa3, echo=FALSE, fig.pos="placeHere", fig.cap="Time varying species composition and effective species diversity (1 / Simpson's Diversity), by permit for seine fisheries in Southeast Alaska (S01A), Prince William Sound (S01E) and Kodiak (S01K).
+```{r figa3, echo=FALSE, fig.pos="placeHere", fig.cap="Time varying species composition and effective species diversity (1 / Simpson's Diversity), by permit for seine fisheries in Southeast Alaska ('SE', S01A), Prince William Sound ('PWS', S01E) and Kodiak ('KO', S01K).
 Species are colored as follows: green (Chinook), purple (chum), silver  (coho), pink (pink), red (sockeye)", fig.height=3.2}
@@
 
-```{r figa4, echo=FALSE, fig.pos="placeHere", fig.cap="Time varying species composition and effective species diversity (1 / Simpson's Diversity), by permit, for seine fisheries on the Alaska Peninsula (S01M), and gillnet fisheries in Southeast Alaska (S03A) and Prince William Sound (S03E).", fig.height=3.2}
+```{r figa4, echo=FALSE, fig.pos="placeHere", fig.cap="Time varying species composition and effective species diversity (1 / Simpson's Diversity), by permit, for seine fisheries on the Alaska Peninsula ('AP', S01M), and gillnet fisheries in Southeast Alaska ('SE', S03A) and Prince William Sound ('PWS', S03E).", fig.height=3.2}
 gridExtra::grid.arrange(g4, g5, g6, h4, h5, h6, nrow=2)
+```
+
+\break
+
+## Trends in revenue and prices
+
+Across regions, we see lots of synchrony in prices between coho, chum, and pink salmon, with them increasing more quickly than sockeye/Chinook.
+Sockeye/Chinook have been the highest priced historically, with coho > chum > pink.
+For most permit types, median revenue / person has generally increased, especially since early 2000s.
+For some permits, variability appears to have changed through time (quartile range getting smaller).
+
+```{r, fig.cap="Total revenue by area (across people and permits within each)", fig.pos="placeHere", echo=FALSE, fig.height=3.5}
+# plot prices of species landed by salmon permit
+cfec.raw$area = NA
+cfec.raw$area[cfec.raw$p_fshy%in%c("S 01A", "S 03A")] = "Southeast"
+cfec.raw$area[cfec.raw$p_fshy%in%c("S 01E", "S 03E")] = "PWS"
+cfec.raw$area[cfec.raw$p_fshy%in%c("S 01K", "S 04K")] = "Kodiak"
+cfec.raw$area[cfec.raw$p_fshy%in%c("S 01M")] = "Alaska Peninsula"
+
+g1 = group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, area) %>%
+  summarize(p = sum(g_earn)) %>% group_by(year, area) %>%
+  ggplot(aes(year, (p)/1e6, group=spec, color = spec)) + geom_line() + facet_wrap(~area, scale="free") +
+  ylab("Revenue (millions) ") + xlab("Year") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
+panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
+g1
+```
+
+\break
+
+```{r, fig.cap="Inflation adjusted prices / species by area, relative to 2002 (the year corresponding to the lowest average prices).", fig.pos="placeHere", echo=FALSE, fig.height = 3.1}
+
+g = group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, area) %>%
+  summarize(p = mean(g_price))
+p1 = group_by(g, spec, area) %>% 
+  mutate(p = p/p[which(year==2002)]) %>% 
+  group_by(year, area) %>%
+  ggplot(aes(year, log(p), group=spec, color = spec)) + geom_line() + facet_wrap(~area) +
+  ylab("Ln avg price relative to 2002") + ggtitle("Synchrony in prices: relative change") + xlab("") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
+panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
+p1
+
+```
+
+```{r, fig.cap="Percent change (log-difference) in inflation adjusted prices / species by area", fig.pos="placeHere", echo=FALSE, fig.height = 3.1}
+
+g = group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, area) %>%
+  summarize(p = mean(g_price))
+# do log-diff of prices -- get price year before
+g$p_t1 = g$p[match(paste(g$year, g$spec, g$area), 
+  paste(g$year+1, g$spec, g$area))]
+
+p2 = group_by(g, spec, area) %>% 
+  ggplot(aes(year, log(p/p_t1), group=spec, color = spec)) + geom_line() + facet_wrap(~area) +
+  ylab("Ln price (t/t-1)") + ggtitle("Synchrony in prices: percent change") + xlab("") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
+panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
+p2
+
+g$logdiff = log(g$p/g$p_t1)
+p3 = group_by(g, year, area) %>% 
+summarize(sd = sd(logdiff,na.rm=TRUE)) %>% ggplot(aes(year, sd, color=area)) + geom_line() + xlab("Year") + ylab("SD of price changes across species") + ggtitle("Synchrony in price fluctuations across species") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
+panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
+
+```
+
+```{r, fig.pos="placeHere", fig.cap = "Standard deviation (across species, by area) for fluctuations in prices.
+Fluctuations are calculated as the log difference price data between sequential years.
+The standard deviation of log differenced data is interpreted as the coefficient of variation (CV)."}
+p3
+```
+
+\break
+
+## Causality between lagged abundance and prices
+
+Previous work on long term trends in Alaska salmon markets have highlighted effects of changing production practices, including freezing (Knapp et al.
+2012).
+Caches of frozen salmon from large harvest may have competitive effects on prices 1-2 years later.
+Using a Granger test for causality between first-differenced time series of harvest and prices, this effect is particularly true for coho and sockeye salmon, where abundance at time *t-1* or *t-2* improves predictions of prices at time *t*.
+
+```{r, echo=FALSE}
+library(knitr)
+d = read.table("../../salmon/data-generated/granger_table.csv", sep=",", header=TRUE)
+kable(d)
+```
+
+```{r echo=FALSE, eval=FALSE}
+#Z = read.table("../../salmon/data-generated/granger_table_s01a.csv", header=TRUE)
 ```
@@
 
-df$permits[which(df$permits=="S01A")] = "SEAK-seine"
+df$permits[which(df$permits=="S01A")] = "SE-seine"
 df$permits[which(df$permits=="S01E")] = "PWS-seine"
-df$permits[which(df$permits=="S01K")] = "Kodiak-seine"
-df$permits[which(df$permits=="S01M")] = "AKP-seine"
-df$permits[which(df$permits=="S03A")] = "SEAK-gillnet"
+df$permits[which(df$permits=="S01K")] = "KO-seine"
+df$permits[which(df$permits=="S01M")] = "AP-seine"
+df$permits[which(df$permits=="S03A")] = "SE-gillnet"
 df$permits[which(df$permits=="S03E")] = "PWS-gillnet"
@@
 
-ggplot(df[df$diversity==1,], aes(x=years,y=pred,group=permits)) +
+g1 = ggplot(df[df$diversity==1,], aes(x=years,y=pred,group=permits)) +
   geom_line() +
   geom_ribbon(aes(ymin=low,ymax=high), alpha=0.4) +
-  facet_wrap(~permits) + xlab("Year") + ylab("Predicted marginal benefit (ln $)") + geom_hline(yintercept=0, color="red")
-
+  facet_wrap(~permits) + xlab("Year") + ylab("Predicted marginal benefit") + geom_hline(yintercept=0, color="red") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
+panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
+g1
 ```
@@
 yearpermit2 = yearpermit
-yearpermit2$permit[which(yearpermit2$permit=="S01A")] = "SEAK-seine"
+yearpermit2$permit[which(yearpermit2$permit=="S01A")] = "SE-seine"
 yearpermit2$permit[which(yearpermit2$permit=="S01E")] = "PWS-seine"
-yearpermit2$permit[which(yearpermit2$permit=="S01K")] = "Kodiak-seine"
-yearpermit2$permit[which(yearpermit2$permit=="S01M")] = "AKP-seine"
-yearpermit2$permit[which(yearpermit2$permit=="S03A")] = "SEAK-gillnet"
+yearpermit2$permit[which(yearpermit2$permit=="S01K")] = "KO-seine"
+yearpermit2$permit[which(yearpermit2$permit=="S01M")] = "AP-seine"
+yearpermit2$permit[which(yearpermit2$permit=="S03A")] = "SE-gillnet"
 yearpermit2$permit[which(yearpermit2$permit=="S03E")] = "PWS-gillnet"
 
-ggplot(data=yearpermit2) + 
+g1 = ggplot(data=yearpermit2) + 
   geom_ribbon(aes(x=year,ymin=b0_low,ymax=b0_hi),alpha=0.3) +
@@
   ylab("Intercepts on mean") + 
-  geom_hline(yintercept=0, color="red")
+  geom_hline(yintercept=0, color="red") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
+panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
+g1
 ```
@@
 
-ggplot(data=rbind(df.gen,df.spec), aes(year, mean, group=strategy, fill=strategy)) + geom_ribbon(aes(ymin=low,ymax=hi), alpha=0.5) + geom_hline(yintercept=0, color="red")
+g1 = ggplot(data=rbind(df.gen,df.spec), aes(year, mean, group=strategy, fill=strategy)) + geom_ribbon(aes(ymin=low,ymax=hi), alpha=0.5) + geom_hline(yintercept=0, color="red") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
+panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
+g1
+```
 
+```{r, echo=FALSE, fig.pos="placeHere", fig.cap = "Change in natural log of landed weight versus change in effective species diversity, by permit.
+Decreases in landed weight is often accompanied by reduced catch diversity."}
+diffdat = readRDS(file="../../salmon/data-generated/salmon.rds")
+diffdat = filter(diffdat, !is.na(days.change))
+
+g1 = ggplot(diffdat, aes(log(weight)-log(weight.prev), specDiv.change)) + geom_point(alpha=0.15, col="blue", size=0.4) + facet_wrap(~ strategy_permit) + xlab("Change in ln annual landed weight") + ylab("Change in effective species diversity") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
+panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
+g1
 ```
@@
 
-ggplot(data=pinkPWS, aes(b0, diff,color=Year)) + geom_point() + stat_smooth(method=lm) + xlab("Estimated intercept from model") + ylab("Change in total pink salmon returns")
-```
-
-\break
-
-## Trends in revenue and prices
-
-Across regions, we see lots of synchrony in prices between coho, chum, and pink salmon, with them increasing more quickly than sockeye/Chinook.
-Sockeye/Chinook have been the highest priced historically, with coho > chum > pink.
-For most permit types, median revenue / person has generally increased, especially since early 2000s.
-For some permits, variability appears to have changed through time (quartile range getting smaller).
-
-```{r, fig.cap="Total revenue by area (across people and permits within each)", fig.pos="placeHere", echo=FALSE, fig.height=3.5}
-# plot prices of species landed by salmon permit
-cfec.raw$area = NA
-cfec.raw$area[cfec.raw$p_fshy%in%c("S 01A", "S 03A")] = "Southeast"
-cfec.raw$area[cfec.raw$p_fshy%in%c("S 01E", "S 03E")] = "PWS"
-cfec.raw$area[cfec.raw$p_fshy%in%c("S 01K", "S 04K")] = "Kodiak"
-cfec.raw$area[cfec.raw$p_fshy%in%c("S 01M")] = "Alaska Peninsula"
-
-group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, area) %>%
-  summarize(p = sum(g_earn)) %>% group_by(year, area) %>%
-  ggplot(aes(year, (p)/1e6, group=spec, color = spec)) + geom_line() + facet_wrap(~area, scale="free") +
-  ylab("Revenue (millions) ") + xlab("Year")
-```
-
-\break
-
-```{r, fig.cap="Inflation adjusted prices / species by area, relative to 2002 (the year corresponding to the lowest average prices).", fig.pos="placeHere", echo=FALSE, fig.height = 3.1}
-
-g = group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, area) %>%
-  summarize(p = mean(g_price))
-p1 = group_by(g, spec, area) %>% 
-  mutate(p = p/p[which(year==2002)]) %>% 
-  group_by(year, area) %>%
-  ggplot(aes(year, log(p), group=spec, color = spec)) + geom_line() + facet_wrap(~area) +
-  ylab("Ln avg price relative to 2002") + ggtitle("Synchrony in prices: relative change") + xlab("")
-p1
-
-```
-
-```{r, fig.cap="Percent change (log-difference) in inflation adjusted prices / species by area", fig.pos="placeHere", echo=FALSE, fig.height = 3.1}
-
-g = group_by(cfec.raw[cfec.raw$spec %in% c("CHUM","COHO","SOCK","PINK","CHNK"),], year, spec, area) %>%
-  summarize(p = mean(g_price))
-# do log-diff of prices -- get price year before
-g$p_t1 = g$p[match(paste(g$year, g$spec, g$area), 
-  paste(g$year+1, g$spec, g$area))]
-
-p2 = group_by(g, spec, area) %>% 
-  ggplot(aes(year, log(p/p_t1), group=spec, color = spec)) + geom_line() + facet_wrap(~area) +
-  ylab("Ln price (t/t-1)") + ggtitle("Synchrony in prices: percent change") + xlab("")
-p2
-
-g$logdiff = log(g$p/g$p_t1)
-p3 = group_by(g, year, area) %>% 
-summarize(sd = sd(logdiff,na.rm=TRUE)) %>% ggplot(aes(year, sd, color=area)) + geom_line() + xlab("Year") + ylab("SD of price changes across species") + ggtitle("Synchrony in price fluctuations across species")
-
-```
-
-```{r, fig.pos="placeHere", fig.cap = "Standard deviation (across species, by area) for fluctuations in prices.
-Fluctuations are calculated as the log difference price data between sequential years.
-The standard deviation of log differenced data is interpreted as the coefficient of variation (CV)."}
-p3
+g1 = ggplot(data=pinkPWS, aes(b0, diff,color=Year)) + geom_point() + stat_smooth(method=lm) + xlab("Estimated intercept from model") + ylab("Change in total pink salmon returns") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
+panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
+g1
 ```

### 2017-03-11 salmon/paper/salmon_paper_v3.Rmd
updated paper

@@
 library(tidyverse)
+library(ggsidekick)
+library(ggrepel)
 library(reshape2)
@@
 
-On a large scale, individual fishers may experience positive outcomes from diversifying their catches; examples include reduced year-to-year variability in annual revenue and increases in average revenue.
-Fishers can increase their species diversity by owning a permit that allows multiple species to be targeted, or by owning multiple permits.
-External forces such as management regulations (limited entry,  catch shares), the number of permits available, or market forces (permit or fish prices) may disincentive individuals to hold multiple permits.
+On a large scale, individual fishers may experience positive outcomes from diversifying their catches; benefits may include reduced year-to-year variability in annual revenue or increases in average revenue.
+Fishers can increase the species diversity of their catch by owning fishing permits that allow multiple species to be targeted, or by owning multiple permits.
+External factors such as management regulations (limited entry,  catch share programs), the number of permits available, or market forces (permit or fish prices) may disincentive individuals to hold multiple permits.
 Thus, individuals holding a single permit may only be able to increase species diversity within the confines of their permit (longer fishing seasons, broader spatial area).
-Using a 40 year dataset of individual salmon fishers in Alaska, we develop new models for understanding how the effects of diversification have changed through time.
-These changes allow us to compare patterns across regions and gear, as well as better understand how external forces have changed incentives for diversifying (hatchery production, market forces, EVOS).
+Using a 41 year dataset of individual salmon fishers in Alaska, we develop new models for understanding how the effects of diversification have changed through time.
+Applying these models to six diverse seine and gillnet fisheries, we find that the majority of these fisheries have experienced reduced diversity in catches through time, and increasing benefits on specialization on mean individual revenues.
+One of the mechanisms that has been hypothesized to reduce catch variability in Alaska has been increased hatchery production, however our regional models suggest that fisheries experience little change in variability as a result of increased hatchery influence.
 
@@
 Previous analyses have largely focused on patterns of diversification across ecosystems and the benefits to adopt more or less diverse fishing strategies [@kasperski2013; @anderson_benefits].
-Fisheries in Alaska have been studied extensively, because they are one of the most productive seafood regions in the world (both by landed weight, and revenue [@officeofscienceandtechnology2016], FAO cite) and they allow individuals to adopt a wide range of fishing strategies.
-[@anderson_benefits] defined strategies as a permit or collection of permits held by single individuals.
-While adding new permits allows fishers to increase species diversity, few individual permits allow individuals to harvest more than one species (examples of single-species permits in the Gulf of Alaska include halibut, sablefish, herring, crab, shellfish, shrimp, sea cucumber).
-Some exceptions to these single species permits include participants in federally managed groundfish fisheries and participants in state managed salmon fisheries.
+Fisheries in Alaska have been studied extensively, because they are one of the most productive seafood regions in the world (both by landed weight, and revenue [@officeofscienceandtechnology2016]) and they allow individuals to adopt a wide range of fishing strategies.
+[@anderson_benefits] defined fishing strategies as a permit or collection of permits held by single individuals.
+Buying new permits allows individuals to fish in more regions or target additional species.
+Some of these permits in Alaskan waters allow only one species to be targeted (e.g.
+sablefish, herring, crab, shellfish, shrimp, and sea cucumber), while others allow for multiple species (federally managed groundfish fisheries, state managed salmon fisheries).
 
-Understanding the effects of species diversification on salmon fishers is particularly important, as the number of salmon specialists (defined as fishers holding a single salmon permit) has increased over the last 30 years as a fraction of the total [@anderson_benefits].
-For these limited entry salmon fisheries, general trends since the early 2000s are that that more people are participating, landings are increasing, revenue is level or increasing slightly, and effective species diversity is declining over time (Fig.
+Understanding the effects of species diversification for individuals participating in salmon fisheries is particularly important, because the proportion of salmon specialists (defined as individuals whose fishing revenue comes from a single salmon permit) have increased over the last 30 years, and annual revenue from these salmon specialists has recently exceeded all other fisheries combined for the first time ever [@anderson_benefits].
+General trends in limited entry salmon fisheries since the early 2000s are that that more individuals have been participating, landings are increasing, revenue is increasing slightly, and effective species diversity is declining over time (Fig.
 1; Figs.
 A1-A4).
-Across different salmon fisheries, there appears to be little correlation between diversity and revenue, though fisheries cluster similarly by gear on the risk - return axes (Fig.
+Across salmon fisheries, there appears to be little correlation between catch diversity and revenue (Fig.
+A2), though fisheries cluster similarly by gear on axes of risk (CV of annual revenue) and return  (mean annual revenue, Fig.
 2).
-In terrestrial agriculture, there are numerous examples where diversification may affect the mean or CV of revenue, but not both [@purdy1997], and the same result may be true for salmon fishers.
-Further, the costs or benefits of diversification may be variable over time.
+Like in terrestrial agriculture [@purdy1997], it's possible that fishing diversification may affect the mean or CV of revenue, but not both.
+Further, it's possible that any benefits or costs of fishers diversifying may be variable through time.
 
@@
 kisutch*]), and the ability of individual salmon fishers to diversify across species is in part a function of geography and fishing gears.
-For example, salmon may be caught with purse seine gear (n = 6 areas), drift gillnets (5 areas), or set gillnets (12 areas).
-Other gears include beach seine and troll gears (though these are generally restricted to Kodiak and Southeast Alaska, respectively).
-While some of these combinations of gears and areas allow fishers to diversify, there are many examples of permits where the catch is dominated by a single species.
-Perhaps the best examples include fisheries that target sockeye salmon (e.g.
+For example, salmon may be caught with purse seine gear (6 areas), drift gillnets (5 areas), or set gillnets (12 areas).
+Other gears include beach seine and troll gears (though these are generally restricted to Kodiak and Southeast Alaska).
+While multiple salmon species are targeted in some of these fisheries, there are also examples of fisheries where catch is dominated by a single species.
+Examples of specialization include fisheries that target sockeye salmon (e.g.
 Bristol Bay set and drift gillnet fisheries).
@@
 
-The benefits or costs of catch diversification for fishermen are affected by a number of exogenous drivers, including competition from markets producing farmed salmon, consolidation of salmon buyers, marketing, prices, new processing technology [@knapp_trends].
-Following price declines in the 1990s and early 2000s [@adfg2007], many of these factors have been positive for salmon fishermen (increasing trends in prices and catches, Fig.
-A8-A9).
-In contrast, the late 1980s - early 1990s experienced high abundance and low prices (in part due to collapses of foreign currencies).
-Perhaps one of the largest changes affecting the availability or incentives for fishermen to diversify has been the increase in the production of hatchery fish.
-The modern era of Alaska’s large-scale hatchery program began in the late 1970s [@stopha2016; @ward_evaluating], with substantial increases in hatchery production during the 1980s following low returns [@mahnken1998; @leber2008].
-Many of these hatchery programs increased production to stabilize the variability of salmon catches [@hilborn2001a], though these programs may also have negative impacts on wild production [@wertheimer2005; @ruggerone2015].
-The intra- and inter-specific effects of large scale hatchery programs has been debated [@hilborn2001a; @hilborn1992a; @wertheimer2001; @heard_hatchery], but the effects of hatcheries have largely been focused on ecological interactions of aggregate yield to the fishery [@leber2008] or economic impacts [@boyce1993], rather than to variability experienced by fisheries or individual fishermen.
-Understanding the role salmon have in buffering or stabilizing fishing revenues is critical given the responses of salmon to variable environments [@mueter2002; @schindler2010].
+Mechanisms responsible for incentivizing specialization or diversification are complex, but include a number of exogenous drivers, such as competition from markets producing farmed salmon, consolidation of salmon buyers, marketing, prices, and new processing technology [@knapp_trends].
+In the late 1980s and early 1990s for example, salmon abundance was high and prices were low (in part due to collapses of foreign currencies).
+Following price declines into the early 2000s [@adfg2007], some of these trends have more recently reversed (increasing prices and catches, Figs.
+A5-A8).
+New technology such as freezing may increase competition within fisheries [@knapp_trends]; for example, large harvests of some species may have a negative impact on future prices (Supplementary Information).
+
+Perhaps one of the largest changes affecting the availability or incentives of fishers to diversify has been a general increase in the production of hatchery fish.
+The modern era of Alaska’s large-scale hatchery program began in the late 1970s [@stopha2016; @ward_evaluating], with substantial increases in hatchery releases during the 1980s following low returns [@mahnken1998; @leber2008].
+Many of these hatchery programs increased production in an effort to stabilize the variability of salmon catches [@hilborn2001a], though quantitative support for a reduction in variability (at the level of fisheries or individuals) remains unclear.
+Previous work has largely concentrated ecological effects of hatcheries [@hilborn2001a; @hilborn1992a; @wertheimer2001; @heard_hatchery; @ruggerone2015], or economic yield [@leber2008; @boyce1993].
+Understanding the role salmon have in buffering or stabilizing fishing revenues is critical given the responses of salmon to variable ocean environments [@mueter2002; @schindler2010].
 
@@
 Because our modeling focus involves estimating time varying effects of diversification, we limited our analysis to salmon fisheries with species diversity > 1 (omitting fisheries such as the Bristol Bay gillnet fisheries that specialize on sockeye salmon).
-To better understand the influence of changing hatchery production, we also restricted fisheries to those that have been influenced by hatchery salmon production.
+To better understand the influence of changing hatchery production, we also restricted fisheries to those that have been influenced by hatcheries.
 
@@
 For consistency with previous work, we examined the stability in the catch (weight of salmon species [@hilborn2001a; @knapp2007], rather than revenue.
-Within each region in our analysis (PWS, Southeast, Kodiak, Alaska Peninsula), we used the hatchery contribution to the Common Property Commercial Harvest as the predictor variable, and total harvest as the response.
-These quantities are reported by region ([@stopha2016]; Fig.
+Within each region in our analysis (Prince William Sound, Southeast, Kodiak, Alaska Peninsula), we used the hatchery contribution to the Common Property Commercial Harvest as the predictor variable, and total harvest as the response ([@stopha2016]).
+These quantities are estimated at the coarse scale by region (Fig.
 3), but not at the level of fisheries or individuals.
@@
 
-To incorporate potential hatchery effects in the variability of the system, the second component of our model included a hierarchical component on the process variance terms, $$log\left( { \sigma  }_{ s,f,t } \right) ={ g  }_{ 0,s,f }+{ g  }_{ 1,s,f }{ X }_{ s,f,t }$$ where ${ g  }_{ 0,s,f }$ represents an estimated intercept unique to each region and species, and ${ g  }_{ 1,s,f }$ represents a region and species specific slope.
+To incorporate potential hatchery effects in the variability of salmon catches, the second component of our model included a hierarchical component on the process variance terms, $$log\left( { \sigma  }_{ s,f,t } \right) ={ g  }_{ 0,s,f }+{ g  }_{ 1,s,f }{ X }_{ s,f,t }$$ where ${ g  }_{ 0,s,f }$ represents an estimated intercept unique to each region and species, and ${ g  }_{ 1,s,f }$ represents a region and species specific slope.
 The covariate ${ X }_{ s,f,t }$ is identical to that used in the mean model.
@@
 
-We focued inference on the slopes of the fraction of hatchery contribution to the total catch, ${ g  }_{ 1,s }$.
-To allow for comparison across regions and scales of data, we converted these estimates to effect sizes, by calculating ${ \kappa  }_{ s }=ln\left( \frac { { g }_{ 0,s }+{ \left( 1/3 \right) \cdot g }_{ 1,s } }{ { g }_{ 0,s } }  \right)$.
-The log ratio of these effect sizes represent the approximate natural log of the percent change in the standard deviation resulting from a change in the hatchery contribution of zero to 10% of the catch.
+We focued inference on the slopes of the fraction of hatchery contribution to the mean and standard deviation of estimated annual catches, ${ b }_{ s,f }$ and ${ g }_{ 1,s,f }$.
+To allow for comparison across regions and scales of data, we converted these estimates to effect sizes, by calculating ${ \kappa  }_{ s }=ln\left( \frac { { g }_{ 0,s }+{ 0.2 \cdot g }_{ 1,s } }{ { g }_{ 0,s } }  \right)$.
+The term ${ \kappa  }_{ s }$ then represents the approximate percent change in the standard deviation resulting from a 20% increase in the hatchery contribution.
 95% credible intervals were generated by repeating this calculation for each MCMC draw, and taking the quantiles of the resulting distribution.
@@
 
-Because we were only interested in single permit fisheries, we applied a simplified version of the CFEC data filtering steps described in [@anderson_benefits].
-Individual trip-level catch and revenue data 1975-2016 were aggregated to annual data for each fisher, and to clarify our comparison across fisheries we retained individual fishers who only participated in one fishery in a year (Table 1).
-Annual revenues were log-transformed to make the data approximately normally distributed; to avoid complications in modeling individual switching permits, we only included pairs of years for a given person where they did not change permits (e.g.
+Because we are focused on single permit fisheries, we applied a simplified version of the CFEC data filtering steps described in [@anderson_benefits].
+Individual trip-level catch and revenue data 1975-2016 were aggregated to annual data for each fisher, and to clarify our comparison across fisheries we only included individual fishers who only participated in one fishery in a year (Table 1).
+To avoid complications in modeling individual switching permits, we only included pairs of years for a given person where they did not change permits (e.g.
 didn’t change fisheries between two years).
@@
 The model for the expected log revenue of individual *i* in year *t*, $$\widehat { \log\left( { R }_{ i,t } \right)  } ={ B }_{ 0,j,t }+{ B }_{ 1,j,t }\cdot \Delta S_{ i,t }+{ B }_{ 2,j,t }\cdot \left( \Delta S_{ i,t }\cdot { S }_{ i,t-1 } \right) +{ B }_{ 3,j }\cdot \Delta D_{ i,t }+\log\left( { R }_{ i,t-1 } \right) $$, where *j* subscripts the salmon permit (Table 1).
-We model the time-varying permit specific intercepts as random and exchangeable across years, ${ B }_{ 0,j,t }\sim \mathrm{Normal}\left( { u }_{ { B }_{ 0 } },{ \sigma  }_{ { B }_{ 0 } } \right)$.
+We model the time-varying intercepts, specific to each permit, as random and exchangeable across years, ${ B }_{ 0,j,t }\sim \mathrm{Normal}\left( { u }_{ { B }_{ 0 } },{ \sigma  }_{ { B }_{ 0 } } \right)$.
 The term ${ B }_{ 1,j,t }$ represents the permit level random effect of diversifying species.
-Our model extends that of [@anderson_benefits] to include the interaction terms $\Delta S_{ t }\cdot { S }_{ t-1 }$ which allow the effect of diversification to be dependent on the previous level of diversification.
+We also include the interaction terms $\Delta S_{ t }\cdot { S }_{ t-1 }$ which allow the effect of diversification to be dependent on the previous level of diversification.
 We assume this relationship is linear, and the coefficients ${ B }_{ 1,j,t }$ and ${ B }_{ 2,j,t }$ allow this to vary by permit and through time.
@@
 Like [@anderson_benefits] we also allow changing effort (measured as percent change in days fished) ${ B }_{ 4,j }$ to vary by strategy, but we do not allow this effect to vary through time.
-Finally, the term $\log\left( { R }_{ i,t-1 } \right)$ is included as an offset [@anderson_benefits].
+Finally, the term $\log\left( { R }_{ i,t-1 } \right)$ is included as an offset.
 
 Like [@anderson_benefits], we also included all of the above covariates in the variance model.
-We assumed the residuals from the mean model described above to be normally distributed, $$\log\left( { R }_{ i,t } \right) \sim \mathrm{Normal}\left( \widehat { \log\left( { R }_{ i,t } \right)  } ,\widehat { { \sigma  }_{ i,t } }  \right) $$.
+We assumed the residuals from the mean model described above to be normally distributed, $$\log\left( { R }_{ i,t } \right) \sim \mathrm{Normal}\left( \widehat { \log\left( { R }_{ i,t } \right)  } ,\widehat { { \sigma  }_{ i,t } }  \right)$$.
 While conventional hierarchical linear regression assumes variances to be constant or shared by grouping variables, we also included fixed and random covariates, making the model akin to variance function regression or stochastic volatility models [@smyth1989; @western2009].
-Using the same notation as for the mean, we can express the log standard deviation as $$\widehat { \log\left( { \sigma  }_{ i,t } \right)  } ={ G }_{ 0,j }+{ G }_{ 1,j,t }\cdot \Delta S_{ i,t }+{ G }_{ 2,j,t }\cdot \left( \Delta S_{ i,t }\cdot { S }_{ i,t-1 } \right) +G_{ 3,j }\cdot \Delta D_{ i,t }$$.
-Unlike the linear model for the mean, we did not include an offset term.
-Further, we constrained the variance intercepts ${ G }_{ 0,j }$ to be temporally constant by strategy, instead of the dynamic nature of the intercepts in the mean model.
+We can express the log standard deviation as $$\widehat { \log\left( { \sigma  }_{ i,t } \right)  } ={ G }_{ 0,j }+{ G }_{ 1,j,t }\cdot \Delta S_{ i,t }+{ G }_{ 2,j,t }\cdot \left( \Delta S_{ i,t }\cdot { S }_{ i,t-1 } \right) +G_{ 3,j }\cdot \Delta D_{ i,t }$$.
+Unlike the linear model for the mean, an offset term is not included for the variance.
+Further, we constrained the variance intercepts ${G}_{0,j}$ to be temporally constant by strategy, instead of the dynamic nature of the intercepts in the mean model.
 Like the mean model, the random effects were constrained to be autoregressive, ${ G }_{ 1,j,t }\sim \mathrm{Normal}\left( { G }_{ 1,j,t-1 },{ \sigma  }_{ { G }_{ 1 } } \right)$, and covariates associated with days fished and species diversity were included as predictors.
@@
 Estimation for both sets of models was done in a Bayesian framework using R [@rcoreteam2016] and the package 'rstan', implementing Markov chain Monte Carlo (MCMC) using the No-U Turn Sampling (NUTS) algorithm in Stan [@hoffman2014].
-Code to replicate our analysis, including additional diagnostics plots, and additional models is available at [https://github.com/NCEAS/pfx-commercial/tree/master/salmon].
+Code to replicate our analysis, including additional diagnostics plots, and additional models is available at [https://github.com/NCEAS/pfx-commercial/salmon].
+[Eric will add more here]
 
@@
 
-Across species and scales of data (regions, fisheries, average fishers) we found no consistent support for the hypothesis that salmon catches have been stabilized by hatchery production.
-When total catches are aggregated across fisheries in the four regions in our analysis, we estimate generally positive effects of increased hatchery production on year-to-year variability in catches (Fig.
+Across species and regions we found no consistent support for the hypothesis that variability in salmon catches has been stabilized by increased hatchery production.
+Across fisheries in the four regions in our analysis, most estimates of hatchery production on changes in the mean total catch is near zero (Fig.
 4).
-Other than Chinook salmon, estimates (+/- standard errors) are positive for all species in PWS and Southeast Alaska (Fig.
+One exception is a negative estimated effect of hatcheries on Chinook catches in Southeast Alaska (reduced catches as the hatchery contribution increases).
+Similarly, most of our estimates of hatchery effects on year-to-year variability are slightly positive (Fig.
 4).
-When we examine the same effects at finer scales (individual fisheries, or average fishers), the standard errors generally increase, and effects become weaker in magnitude.
-Some of the estimated effects switch signs entirely -- for example, hatchery production of chum salmon is estimated to have a positive effect on variability in catches in Prince William Sound, but at the level of fisheries or average fishers, increased contributions of hatchery chum salmon appear to reduce variability (Fig.
-4).
-Other effect sizes appear consistent across scales of data; for example increased sockeye salmon in PWS is expected to increase year-to-year variability across fisheries, within the seine and gillnet fisheries (S01E, S03E) and for average fishers (Fig.
-4).
+The strongest positive effects appear to be the contributions of hatchery pink and coho salmon to variability in catches, particularly in Prince William Sound.
 
@@
 The negative estimated effects on mean revenue of increasing diversity implies that for the majority of these fisheries in most years, revenue increases when individuals specialize, rather than diversify.
-Though gillnet fisheries in Southeast Alaska and Prince William Sound have also had reduced benefits of diversifying, each fishery had several periods where it was more profitable to increase catch diversity (mid 1980s-early 1990s for Southeast Alaska, mid-1990s for PWS).
+Though gillnet fisheries in Southeast Alaska and Prince William Sound have also had reduced benefits of diversifying, each fishery had several periods where it was more profitable to increase catch diversity (mid 1980s-early 1990s for Southeast Alaska, mid-1990s for Prince William Sound).
 With the exception of the Southeast Alaska seine fishery (S01A) in recent years, the negative trends in these estimated effects for most fisheries suggest that the benefits of specializing on revenue have also increased over time.
 
-[Maybe tie in to Anderson et al.] We found little support for increasing catch diversity reducing variability in revenue for most fisheries (Fig.
-A5).
+Our results also showed contrasts in variability between seine and gillnet fisheries (Fig.
+A10), with individuals in seine fisheries experiencing higher annual variability in revenue.
+These results are expected, because when individuals are aggregated, seine fisheres are generally higher on axes of risk and return (CV of annual revenue, mean annual revenue; Fig.
+2).
+Though the catch composition of each of these fisheries has fluctuated through time, in areas where seine and gillnet fisheries exist (Prince William Sound, Southeast Alaska), effective catch diversity is slightly higher in gillnet fisheries (Figs A3-A4).
+
+We found little support for increasing catch diversity reducing variability in revenue for most fisheries (Fig.
+A9).
 For seine fisheries in Southeast Alaska and Prince William Sound, there have been periods over the last several decades where diversification has increased variability, but this effect has been strongest over the most recent decade (Fig.
-A5).
+A9).
 Perhaps the only example of a fishery where individuals have seen reduced variability from increasing catch diversity is the Kodiak seine fishery (this effect is strongest in the 1990s and last 5 years).
-While these effects of diversifying are near zero for most fisheries and don't exhibit a long term trend, an exception is the PWS seine fishery, which appears to have increasing risk associated with diversifying over the time series (Fig.
-A5).
+While these effects of diversifying are near zero for most fisheries and don't exhibit a long term trend, an exception is the Prince William Sound seine fishery, which appears to have increasing risk associated with diversifying over the time series (Fig.
+A9).
 
@@
 
-Understanding the mechanisms responsible for trends or variability in the revenue of individuals whose incomes are dependent on natural resources is important for coupled social-ecological systems, because their income variability is typically greater than the larger population [@mishra2002].
-There are a number of ways that fishers in particular may reduce their exposure to risk; examples include holding permits that allow multiple species to be targeted, or buying multiple single-species permits to increase the number of species in their catch portfolio [@anderson_benefits; @kasperski2013].
-Just as [@anderson_benefits] found that catch diversity in Alaska was decreasing over time as a greater fraction of the fisher population held single salmon permits, we found within-fishery declines in catch diversity for the six fisheries included in our analysis.
+Understanding the mechanisms responsible for trends or variability in the revenue of individuals whose incomes are dependent on natural resources is important for coupled social-ecological systems, because their income variability is typically greater than the larger population [@mishra2002; @sethi2012].
+Unlike terrestrial farmers, commercial fishers are often constrained by regulations that can limit the number of individuals who can participate, where fishing occurs, what type of gear is used, or when fishing occurs.
+Individuals participating in many of the commercial salmon fisheries in Alaska may have the option of increasing the diversity of their catches.
+Examining long terms trends in catch diversity for seine and gillnet fisheries in our analysis illustrates that most fisheries have become less diverse over time.
+In recent years, catches from several of these fisheries have been dominated by single species (sockeye salmon in the Prince William Sound gillnet fishery, pink salmon in the Prince William Sound seine fishery, chum salmon in the Southeast Alaska gillnet fishery; Figs.
+A3-A4).
 
-As catch diversity has declined, average revenues have remained relatively flat, but average landed weight has increased over the last 40 years.
-In other words, fishers have had to catch more to yield the same revenue, and their catches tend to be driven by 1 or 2 species.
-Like in other fisheries, the ability of fishers to diversify may be largely beyond their control.
-Options for diversification include fishing in different periods of the season (e.g.
-fishing earlier in the year in the PWS seine fishery to target chum salmon), or fishing in different areas to increase catch diversity.
-For example in the PWS seine fishery, individuals are generally confied to areas where pink and chum salmon are present, while gillnet fishers may fish in areas with all five species of salmon present (Rich can provide a citation for the PWS management plan).
+Like in other commercial fisheries, the ability of salmon fishers to diversify within the constraints of their permits may be largely driven by factors beyond their control.
+Options for catching other species may include fishing in different months or different regions.
+For example, individuals participating in the Prince William Sound seine fishery may fish earlier in the year to target chum salmon.
+Incentives to diversify are complex, and include a suite of external economic factors: competition from foreign markets, prices from processors, fuel costs, and marketing (e.g.
+increased marketing of Copper River sockeye salmon beginning in the early 1980s) [@knapp_trends].
+A potential driver unique to Alaska salmon fisheries are the potential impacts of hatchery salmon.
+Increased hatchery production may alter the species composition that individual fishers encounter, as well as the prices set by processors.
 
-Tradeoffs between specialization and diversification may also be found in case studies of terrestrial agriculture.
-For small scale farmers, increasing crop diversity may increase revenue and stabilize year-to-year income variability [@difalco2003].
-At larger scales or in the developed world, however, the incentives for diversification may be different.
-In the United States and Canada for example, crop diversity has declined since the late-1970s [@aguilar2015; @bradshaw2004]; some of the factors responsible include changing national policies, technology, and increased demand from other parts of the world (e.g.
-increased U.S.
-production of soybeans to meet demand from Asia).
-Shifts from more diverse high value crops to specializing on high-volume low-value crops such as cereals may have analogs for salmon fishermen in low value species with rapid life histories (pink and chum salmon).
+Results from our variance function model of individual revenues over the period 1975-2016 suggests that there have been long term increases in the benefits of individual fishers specializing (expected revenue increases when catch diversity decreases; Fig.
+5).
+For most of the time series, the impacts of changes to catch diversity on variability in year-to-year revenue are less clear.
+A counter-intuitive result is that since the mid-2000s, fisheries in Prince William Sound and seine fisheries in Southeast Alaska have had an estimated increase in revenue variability associated with diversification.
+Some of this relationship may be spurious, in that individuals who have large reductions in catch diversity may harvest less (Fig.
+A11).
+But another driver may be increased consolidation from processors.
+[EW: perhaps include Jen's figure in the appendix of processors / fishery declining through time?].
+Prices of all salmon have generally increased since the early 2000s, with some of the largest changes being in pink and chum salmon prices.
+In areas with high pink or chum salmon runs, the benefits of diversification may have been greatest when these species had low returns.
 
-[Rich: maybe look at PWS pink salmon crash in 1992/1993: Pacific Salmon & their Ecosystems: Status and Future Options.
-EW: pink harvest generally falls from 89/90 to early 90s.
-It's a little of a chicken and egg problem too - because by the early 90s hatchery production was 75% of the total pink catch and also declined a little bit.
-From the CV perspective, the pink crash may be synchronizing -- stability was increased because there were several back to back bad years]
-[Jordan: So the Copper River salmon marketing ramped up in about 1983.
-I wonder if we could see a steady increase in value or movement towards sockeye in the early years of the data set.
-EW: Fig 3 largely shows the pattern in those early years -- it ramps up in PWS, but is less steep than for say chum]
+An additional potential impact of increased hatchery production may be changes in the variability of catches or revenues, to fisheries or individuals.
+Across multiple regions in Alaska, we found little support for catches becoming more stable as hatchery programs have increased.
+One example of a positive correlation between the hatchery contribution to harvest and variability is pink salmon in Prince William Sound (Fig.
+4).
+Our model describing the revenues of individual fishers also suggests that inividuals in the Prince William Sound seine fishery have higher year to year variability relative to other fisheries (Fig.
+A10), with year to year changes being correlated with fluctuations in pink salmon returns (Fig.
+A12).
 
-- In terms of participation, why did people drop out in 2000 in SEAK and Alaska Peninsula, but not in PWS?
-PWS largely seems like a boutique fishery on the Copper River -- which includes a qualtiy component (Copper River at forefront) and a timing issue relative to the rest of the Gulf of Alaska populations (buffers against market impacts).
-In addition, there are many indications that costs have increased over time as a result of requirements by fish buyers for fish of higher quality (refrigerated sea water, ice, increased handling time, more expensive vessels) more expensive permits, etc.
+The increased specialization of Alaska fisheries has been highlighted in several recent analyses [@kasperski2013; @anderson_benefits], and our analysis extends this work to show similar trends within individual Alaskan salmon fisheries.
+In terrestrial agriculture, there are multiple examples of year-to-year income variability being stabilized by diversification [@difalco2003], but these incentives become more complex at larger scales.
+Crop diversification in developed countries is shaped by a number of drivers including changing national policies, technology, and increased demand from other parts of the world (e.g.
+U.S.
+production of soybeans for Asian countries) [@aguilar2015; @bradshaw2004].
+Shifts in terrestrial agriculture from a diverse mix of low- and high-value crops to focusing on high-volume low-value crops (such as soybeans or cereals) may have parallels for Alaskan salmon fisheries -- particularly those fisheries that have more recently targeted low value species with rapid life histories (pink, chum salmon).
 
-- Prices are generally synchronous for all spp (especially coho/chum/pink), and have increased at a faster rate compared to chum and pink salmons since 2000 (Fig A8-A9).
-We're also seeing a trend toward greater synchrony between species (Fig A11-A12, A14).
-Projecting future prices is notably complex [@knapp_trends].
-Synchrony in prices may largely be function of marketing, improved product, freezing, etc.
-
-- We're seeing increased specialization within permits.
-People in PWS are trending toward increased specialization on pink salmon.
-People in SEAK trending toward increased specialization on chum.
-Both of these are switches from higher value species (sockeye, chum) to lower value species.
-Both pink and chum have higher variability in year to year returns than sockeye.
-
-- The decisions to implement a hatchery program for a given species has economic consequences for overall revenue and the variability of that revenue to fishers and other participants.
-Salmon that have complex life histories in the ocean phase of their life cycle (sockeye and possibly chum salmon) may have an inherent portfolio that buffers against large annual swings in abundance compared to pink and coho salmon.
-Also, high-value species may experience less annual fluctuation than lower value species.
-
-- For small scale farmers, economists have often wondered why farmers don't diversify more.
-Diversification can be profitable, but in the short term risky or expensive.
-Bigger than cost seems to be the risk [@karlan2012]
+Increasing trends toward specialization in Alaska salmon fisheries may be acceptable for fishers, if individuals are willing to accept the additional risk that may accompany large revenues.
+Variability in the revenues that individuals or communities experience is by product of many complex drivers, including management, market forces, and environmental variation.
+While reducing risks to individual fisher revenues is not something that can currently be managed for, understanding patterns of variability that individuals experience is essential for quantifying changing revenue patterns at higher levels of fisheries and coastal communities, and protecting individuals from future economic shocks.
 
@@
 
-```{r fig1, echo=FALSE, fig.pos="placeHere", fig.cap="Trends in participation, landed catch, mean annual revenue, and effective species diversity for eight selected salmon fisheries.
-[EW notes: weight per person increasing more than revenue.
-Species diversity for most fisheries declining]"}
+```{r fig1, echo=FALSE, fig.pos="placeHere", fig.cap="Trends in participation, landed catch, mean annual revenue, and effective species diversity for eight selected Alaska salmon fisheries. "}
 # trends in participation, 
@@
 people2 = people[people$Permit%in%c("S01K","S01M","S01E","S01A","S03A","S03E","S03T","S05B"),]
-people2$Permit[which(people2$Permit=="S01K")] = "Kodiak-seine"
-people2$Permit[which(people2$Permit=="S01M")] = "AKP-seine"
+people2$Permit[which(people2$Permit=="S01K")] = "KO-seine"
+people2$Permit[which(people2$Permit=="S01M")] = "AP-seine"
 people2$Permit[which(people2$Permit=="S01E")] = "PWS-seine"
-people2$Permit[which(people2$Permit=="S01A")] = "SEAK-seine"
-people2$Permit[which(people2$Permit=="S03A")] = "SEAK-gillnet"
-people2$Permit[which(people2$Permit=="S03E")] = "PWS-gillnet"
-people2$Permit[which(people2$Permit=="S03T")] = "BB-gillnet"
+people2$Permit[which(people2$Permit=="S01A")] = "SE-seine"
+people2$Permit[which(people2$Permit=="S03A")] = "SE-drift"
+people2$Permit[which(people2$Permit=="S03E")] = "PWS-drift"
+people2$Permit[which(people2$Permit=="S03T")] = "BB-drift"
 people2$Permit[which(people2$Permit=="S05B")] = "State-troll"
@@
 weight2 = weight[weight$Permit%in%c("S01K","S01M","S01E","S01A","S03A","S03E","S03T","S05B"),]
-weight2$Permit[which(weight2$Permit=="S01K")] = "Kodiak-seine"
-weight2$Permit[which(weight2$Permit=="S01M")] = "AKP-seine"
+weight2$Permit[which(weight2$Permit=="S01K")] = "KO-seine"
+weight2$Permit[which(weight2$Permit=="S01M")] = "AP-seine"
 weight2$Permit[which(weight2$Permit=="S01E")] = "PWS-seine"
-weight2$Permit[which(weight2$Permit=="S01A")] = "SEAK-seine"
-weight2$Permit[which(weight2$Permit=="S03A")] = "SEAK-gillnet"
-weight2$Permit[which(weight2$Permit=="S03E")] = "PWS-gillnet"
-weight2$Permit[which(weight2$Permit=="S03T")] = "BB-gillnet"
+weight2$Permit[which(weight2$Permit=="S01A")] = "SE-seine"
+weight2$Permit[which(weight2$Permit=="S03A")] = "SE-drift"
+weight2$Permit[which(weight2$Permit=="S03E")] = "PWS-drift"
+weight2$Permit[which(weight2$Permit=="S03T")] = "BB-drift"
 weight2$Permit[which(weight2$Permit=="S05B")] = "State-troll"
@@
 dat.annual2 = dat.annual[dat.annual$Permit%in%c("S01K","S01M","S01E","S01A","S03A","S03E","S03T","S05B"),]
-dat.annual2$Permit[which(dat.annual2$Permit=="S01K")] = "Kodiak-seine"
-dat.annual2$Permit[which(dat.annual2$Permit=="S01M")] = "AKP-seine"
+dat.annual2$Permit[which(dat.annual2$Permit=="S01K")] = "KO-seine"
+dat.annual2$Permit[which(dat.annual2$Permit=="S01M")] = "AP-seine"
 dat.annual2$Permit[which(dat.annual2$Permit=="S01E")] = "PWS-seine"
-dat.annual2$Permit[which(dat.annual2$Permit=="S01A")] = "SEAK-seine"
-dat.annual2$Permit[which(dat.annual2$Permit=="S03A")] = "SEAK-gillnet"
-dat.annual2$Permit[which(dat.annual2$Permit=="S03E")] = "PWS-gillnet"
-dat.annual2$Permit[which(dat.annual2$Permit=="S03T")] = "BB-gillnet"
+dat.annual2$Permit[which(dat.annual2$Permit=="S01A")] = "SE-seine"
+dat.annual2$Permit[which(dat.annual2$Permit=="S03A")] = "SE-drift"
+dat.annual2$Permit[which(dat.annual2$Permit=="S03E")] = "PWS-drift"
+dat.annual2$Permit[which(dat.annual2$Permit=="S03T")] = "BB-drift"
 dat.annual2$Permit[which(dat.annual2$Permit=="S05B")] = "State-troll"
@@
 ```{r fig2, echo=FALSE, fig.pos="placeHere", fig.cap="CV and mean annual revenue by salmon permit, 1975-2016.
-Average participation (unique fishers) given by circle size, and effective species diversity is shown with color shading.
-[EW notes: pretty cool clustering by permits here - highlighted by Ole.
-The modeling in the rest of the paper is largely comparing the blue seine fisheries to the orange gillnets in PWS / SEAK]", warning=FALSE, message=FALSE}
-dat.annual = readRDS(file="../../portfolio/data-generated/salmon-annual-for-modeling.rds")
-
-#cfec_all = readRDS("../../portfolio/data-generated/salmon_data_for_paper.rds")
-people = group_by(cfec_all, p_fshy, year) %>% 
-  summarize(n = length(unique(p_holder))) %>% 
-  group_by(p_fshy) %>% summarize(n = mean(n,na.rm=T))
-names(people)[which(names(people)=="p_fshy")] = "strategy_permit"
-people$strategy_permit = paste0(substr(people$strategy_permit,1,1), substr(people$strategy_permit,3,5))
-
-records_s03h = which((dat.annual$strategy_permit == "S03H" & dat.annual$year %in% c(1989,1990) | dat.annual$strategy_permit == "S04X"))
-
-dat.annual = filter(dat.annual[-records_s03h,], npermit==1 & substr(strategy_permit,1,1)=="S") %>%
-  group_by(strategy_permit) %>% 
-  mutate(nPeople = length(unique(p_holder))) %>% 
-  filter(nPeople > 200) %>% 
-  dplyr::select(-nPeople) %>% 
-  group_by(strategy_permit, year) %>% 
-  summarize(meanrev = mean(revenue), meandiv = mean(specDiv)) %>% 
-  arrange(strategy_permit, year) %>% 
-  group_by(strategy_permit) %>% 
-  summarize(mean_rev = mean(meanrev,na.rm=T), cv_rev = sd(diff(log(meanrev)),na.rm=T),
-    diversity = mean(meandiv,na.rm=T)) %>% as.data.frame
-
-dat.annual = left_join(dat.annual, people)
-dat.annual$gear = substr(dat.annual$strategy_permit,2,3)
-dat.annual$gear[which(dat.annual$gear=="01")]="Purse seine"
-dat.annual$gear[which(dat.annual$gear=="03")]="Drift gillnet"
-dat.annual$gear[which(dat.annual$gear=="04")]="Set gillnet"
-dat.annual$gear[which(dat.annual$gear%in%c("05"))]="Hand troll"
-dat.annual$gear[which(dat.annual$gear=="15")]="Power troll"
-
-dat.annual$price = NA
-dat.annual$price[dat.annual$strategy_permit=="S01A"]=250000
-dat.annual$price[dat.annual$strategy_permit=="S01E"]=186700
-dat.annual$price[dat.annual$strategy_permit=="S01H"]=84800
-dat.annual$price[dat.annual$strategy_permit=="S01K"]=40000
-dat.annual$price[dat.annual$strategy_permit=="S01L"]=227500
-dat.annual$price[dat.annual$strategy_permit=="S01M"]=56900
-
-dat.annual$price[dat.annual$strategy_permit=="S03A"]=88900
-dat.annual$price[dat.annual$strategy_permit=="S03E"]=224200
-dat.annual$price[dat.annual$strategy_permit=="S03H"]=63500
-dat.annual$price[dat.annual$strategy_permit=="S03M"]=119500
-dat.annual$price[dat.annual$strategy_permit=="S03T"]=148200
-dat.annual$price[dat.annual$strategy_permit=="S03W"]=NA
-
-dat.annual$price[dat.annual$strategy_permit=="S04D"]=88900
-dat.annual$price[dat.annual$strategy_permit=="S04H"]=15300
-dat.annual$price[dat.annual$strategy_permit=="S04K"]=77000
-dat.annual$price[dat.annual$strategy_permit=="S04M"]=55900
-dat.annual$price[dat.annual$strategy_permit=="S04T"]=38500
-dat.annual$price[dat.annual$strategy_permit=="S04W"]=7300
-dat.annual$price[dat.annual$strategy_permit=="S04X"]=5400
-dat.annual$price[dat.annual$strategy_permit=="S04Y"]=9900
-dat.annual$price[dat.annual$strategy_permit=="S04Z"]=11100
-
-dat.annual$price[dat.annual$strategy_permit=="S05B"]=11000
-dat.annual$price[dat.annual$strategy_permit=="S15B"]=38300
-
-g1 = ggplot(dat.annual, aes(x=mean_rev, y=cv_rev, label = dat.annual$strategy_permit)) + geom_point(alpha=0.7, aes(col=gear,size=diversity)) + scale_size(range = c(0, 20)) + geom_text_repel() + ylab("CV annual revenue") + xlab("Mean revenue (dollars)") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
-panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
-print(g1)
+Effective species diversity of catch are shown by circle size (effective species diversity = 1 / Simpson's diversity), and fishing gears are shown with color shading.", warning=FALSE, message=FALSE}
+# dat.annual = readRDS(file="../../portfolio/data-generated/salmon-annual-for-modeling.rds")
+# 
+# people = group_by(cfec_all, p_fshy, year) %>% 
+#   summarize(n = length(unique(p_holder))) %>% 
+#   group_by(p_fshy) %>% summarize(n = mean(n,na.rm=T))
+# names(people)[which(names(people)=="p_fshy")] = "strategy_permit"
+# people$strategy_permit = paste0(substr(people$strategy_permit,1,1), substr(people$strategy_permit,3,5))
+# 
+# records_s03h = which((dat.annual$strategy_permit == "S03H" & dat.annual$year %in% c(1989,1990) | dat.annual$strategy_permit == "S04X"))
+# 
+# dat.annual = filter(dat.annual[-records_s03h,], npermit==1 & substr(strategy_permit,1,1)=="S") %>%
+#   group_by(strategy_permit) %>%
+#   mutate(nPeople = length(unique(p_holder))) %>%
+#   filter(nPeople > 200) %>%
+#   dplyr::select(-nPeople) %>%
+#   group_by(strategy_permit, year) %>%
+#   summarize(meanrev = mean(revenue), meandiv = mean(specDiv)) %>%
+#   arrange(strategy_permit, year) %>%
+#   group_by(strategy_permit) %>%
+#   summarize(mean_rev = mean(meanrev,na.rm=T), cv_rev = sd(diff(log(meanrev)),na.rm=T),
+#     diversity = mean(meandiv,na.rm=T)) %>% as.data.frame
+# 
+# dat.annual = left_join(dat.annual, people)
+# dat.annual$gear = substr(dat.annual$strategy_permit,2,3)
+# dat.annual$gear[which(dat.annual$gear=="01")]="Purse seine"
+# dat.annual$gear[which(dat.annual$gear=="03")]="Drift gillnet"
+# dat.annual$gear[which(dat.annual$gear=="04")]="Set gillnet"
+# dat.annual$gear[which(dat.annual$gear%in%c("05"))]="Hand troll"
+# dat.annual$gear[which(dat.annual$gear=="15")]="Power troll"
+# 
+# dat.annual$price = NA
+# dat.annual$price[dat.annual$strategy_permit=="S01A"]=250000
+# dat.annual$price[dat.annual$strategy_permit=="S01E"]=186700
+# dat.annual$price[dat.annual$strategy_permit=="S01H"]=84800
+# dat.annual$price[dat.annual$strategy_permit=="S01K"]=40000
+# dat.annual$price[dat.annual$strategy_permit=="S01L"]=227500
+# dat.annual$price[dat.annual$strategy_permit=="S01M"]=56900
+# 
+# dat.annual$price[dat.annual$strategy_permit=="S03A"]=88900
+# dat.annual$price[dat.annual$strategy_permit=="S03E"]=224200
+# dat.annual$price[dat.annual$strategy_permit=="S03H"]=63500
+# dat.annual$price[dat.annual$strategy_permit=="S03M"]=119500
+# dat.annual$price[dat.annual$strategy_permit=="S03T"]=148200
+# dat.annual$price[dat.annual$strategy_permit=="S03W"]=NA
+# 
+# dat.annual$price[dat.annual$strategy_permit=="S04D"]=88900
+# dat.annual$price[dat.annual$strategy_permit=="S04H"]=15300
+# dat.annual$price[dat.annual$strategy_permit=="S04K"]=77000
+# dat.annual$price[dat.annual$strategy_permit=="S04M"]=55900
+# dat.annual$price[dat.annual$strategy_permit=="S04T"]=38500
+# dat.annual$price[dat.annual$strategy_permit=="S04W"]=7300
+# dat.annual$price[dat.annual$strategy_permit=="S04X"]=5400
+# dat.annual$price[dat.annual$strategy_permit=="S04Y"]=9900
+# dat.annual$price[dat.annual$strategy_permit=="S04Z"]=11100
+# 
+# dat.annual$price[dat.annual$strategy_permit=="S05B"]=11000
+# dat.annual$price[dat.annual$strategy_permit=="S15B"]=38300
+# 
+# g1 = ggplot(dat.annual, aes(x=mean_rev, y=cv_rev, label = dat.annual$strategy_permit)) + geom_point(alpha=0.7, aes(col=gear,size=diversity)) + scale_size(range = c(0, 20)) + geom_text_repel() + ylab("CV annual revenue") + xlab("Mean revenue (dollars)") + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(),
+# panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
+# print(g1)
 #pdf("Figure_01.pdf")
@@
 #dev.off()
+
+dat.annual <- read.table("../../salmon/data-generated/bubble_plot_data_fig2.csv", stringsAsFactors = FALSE, header = TRUE)
+
+dat.annual <- mutate(dat.annual,
+  gear = factor(gear, levels = c("Drift gillnet", "Set gillnet", "Hand troll",
+    "Power troll", "Purse seine")))
+
+l <- tibble::tribble(~let, ~label,
+  "H", "CI",
+  "E", "PWS",
+  "L", "CH",
+  "K", "KO",
+  "M", "AP",
+  "T", "BB",
+  "A", "SE",
+  "B", "State",
+  "Z", "NS",
+  "W", "KU",
+  "D", "YAK",
+  "Y", "YUK")
+
+dat.annual <- mutate(dat.annual,
+  let = sub("S[0-9]+([A-Z])", "\\1", dat.annual$strategy_permit)) %>%
+  left_join(l) %>%
+  filter(strategy_permit != "S04Y")
+
+g <- ggplot(dat.annual,
+  aes(x = mean_rev/1e3, y = cv_rev,
+    label = label)) +
+  geom_point(alpha = 0.5, aes(fill = gear, size = diversity), pch = 21,
+    colour = "grey20") +
+  scale_size(range = c(1.5, 15), breaks = seq(1.2, 2.4, 0.4)) +
+  guides(fill = guide_legend(override.aes = list(size = 4, alpha = 0.7), order = 1)) +
+  geom_text_repel(size = 2.6, colour = "grey20",
+    alpha = 1, segment.color = "grey70",
+    point.padding = unit(0, "lines"), max.iter = 1e4, segment.size = 0.3) +
+  scale_fill_brewer(palette = "Set3") +
+  ylab("CV annual revenue") + xlab("Mean revenue ($1000)") +
+  labs(fill = "Gear", size = "Diversity") +
+  scale_y_continuous(breaks = seq(0.2, 1, 0.2)) +
+  theme_sleek() #+
+  # scale_x_log10()
+print(g)
 ```
@@
 Shown are the numbers of hatchery fish caught by fisheries (top row), and the hatchery contribution each species to total catch (bottom row).
-The Alaska Peninsula is not included, because hatchery production has been approximately zero."}
+Areas represents combined Kodiak and Alaska Peninsula ('KO - AP'), Southeast ('SE') and Prince William Sound ('PWS').
+"}
 
@@
 
-seak_hatch$region = "Southeast"
-pws_hatch$region = "Prince William Sound"
-kodiak_hatch$region = "Kodiak - Alaska Peninsula"
+seak_hatch$region = "SE"
+pws_hatch$region = "PWS"
+kodiak_hatch$region = "KO - AP"
 
@@
 hatch$Region = as.character(hatch$Region)
-hatch$Region[which(hatch$Region=="PWS")] = "Prince William Sound"
-hatch$Region[which(hatch$Region=="Kodiak")] = "Kodiak - Alaska Peninsula"
+hatch$Region[which(hatch$Region=="PWS")] = "PWS"
+hatch$Region[which(hatch$Region=="Kodiak")] = "KO - AP"
+hatch$Region[which(hatch$Region=="Southeast")] = "SE"
 hatch = melt(hatch, id.vars = c("Year","Region"))
@@
 
-```{r fig4, echo=FALSE, fig.pos="placeHere", fig.cap="Effects of the fraction of hatchery fish on the year-to-year variability in landed weight at three scales: (1) Combining fisheries across regions, (2) Across fisheries, and (3) Average variability of individual fishers within fisheries.
-Effects of variability are presented in terms of effect sizes, or the natural log of the standard deviation of year-to-year catches in a scenario with hatchery fish constituting 10% of the catch to the standard deviation in a scenario with no hatchery fish in the catch.
+```{r fig4, echo=FALSE, fig.pos="placeHere", fig.cap="Effects of the fraction of hatchery fish on (1) mean annual revenue, and (2) sd of annual revenue, for three areas ('KO' = Kodiak, 'PWS' = Prince William Sound, 'SE' = Southeast).
+Effect sizes are presented as the effects of a change of 0% hatchery fish to 20% hatchery fish.
 Dots represent mean estimates, and error bars represent +/- 1 standard error"}
@@
 df$fishery[which(df$fishery=="Prince William Sound")] = "PWS"
-df$fishery[which(df$fishery=="Southeast")] = "SEAK"
-df$fishery[which(df$fishery=="Kodiak")] = "Kodiak"
-
-df$fishery[which(df$fishery=="Prince William Sound")] = "PWS"
-df$fishery[which(df$fishery=="Southeast")] = "SEAK"
-df$fishery[which(df$fishery=="Kodiak")] = "Kodiak"
+df$fishery[which(df$fishery=="SEAK")] = "SE"
+df$fishery[which(df$fishery=="Kodiak")] = "KO"
 
@@
 
-```{r fig5, fig.pos="placeHere", fig.cap="Time varying effects of increasing the effective diversity of catch from 1 to 2 species (measured in ln dollars).
-Effective species diversity is calculated as 1/Simpson's Diversity.
-This shows declining benefits of diversification for most fisheries.", echo=FALSE}
+```{r fig5, fig.pos="placeHere", fig.cap="Time varying effects of increasing the effective diversity (1/Simpson's diversity) of catch from 1 to 2 species (presented as log ratio).
+This shows declining benefits of diversification on mean revenue for individuals in most fisheries.", echo=FALSE}
 # make yearpermit df
@@
 
-df$permits[which(df$permits=="S01A")] = "SEAK-seine"
+df$permits[which(df$permits=="S01A")] = "SE-seine"
 df$permits[which(df$permits=="S01E")] = "PWS-seine"
-df$permits[which(df$permits=="S01K")] = "Kodiak-seine"
-df$permits[which(df$permits=="S01M")] = "AKP-seine"
-df$permits[which(df$permits=="S03A")] = "SEAK-gillnet"
-df$permits[which(df$permits=="S03E")] = "PWS-gillnet"
+df$permits[which(df$permits=="S01K")] = "KO-seine"
+df$permits[which(df$permits=="S01M")] = "AP-seine"
+df$permits[which(df$permits=="S03A")] = "SE-drift"
+df$permits[which(df$permits=="S03E")] = "PWS-drift"
 
@@
   geom_line(col="blue") + 
-  facet_wrap(~permits) + xlab("Year") + ylab("Benefit of catch diversity increasing 1 to 2") + geom_hline(yintercept=0, color="red", linetype = 2) + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
+  facet_wrap(~permits) + xlab("Year") + ylab("Benefit of catch diversity increasing from 1 to 2") + geom_hline(yintercept=0, color="red", linetype = 2, alpha = 0.3) + theme_bw() + theme(panel.border = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))
 

### 2017-03-11 salmon/paper/salmon_paper_v3.Rmd
updated bib and paper

@@
 
-Increasing trends toward specialization in Alaska salmon fisheries may be acceptable for fishers, if individuals are willing to accept the additional risk that may accompany large revenues.
+Increasing trends toward specialization in Alaska salmon fisheries may be acceptable for fishers, if individuals are willing to accept the additional risk that may accompany large revenues, particularly when prices and abundances are high.
 Variability in the revenues that individuals or communities experience is by product of many complex drivers, including management, market forces, and environmental variation.

### 2017-03-11 salmon/paper/salmon_paper_v3.Rmd
paper

@@
   effects on variation in individual revenues'
-author: "Eric, Sean, Ole, Rich, Jordan, Jen, Alan, Milo, Ben"
+author: "Eric, Sean, Ole, Rich, Jordan, Jen, Anne, Alan, Milo, Ben"
 date: ''

### 2017-05-05 portfolio/text/anderson-etal-rev-pfx.tex
Start revising paper

@@
 for permit holder \(i\) and time (year) \(t\) as the inverse of Simpson's
-diversity index \(\lambda\) \cite{simpson1949, jost2006} (also referred to in
+concentration index \(\lambda\) \cite{simpson1949, jost2006} (this index is also referred to in
 economics as the Herfindahl-Hirschman Index (HHI) \cite{hirschman1964}) weighted by

### 2017-05-05 portfolio/text/ms.Rmd
Start revising paper

@@
 species diversity across combinations of state and federal fishing permits
-("strategies"; Fig. \ref{fig:across}). Controlling for effort, strategies with
+("strategies"; Fig. \ref{fig:across}). This pattern is
+qualitatively similar with or without individuals targeting salmon. Controlling for effort, strategies with
 one unit higher mean species diversity had on average `r h1[[2]]`% (`r
-h1[[1]]`--`r h1[[3]]`% 95% CI) lower variability. This pattern is
-qualitatively similar with or without individuals targeting salmon. As examples,
+h1[[1]]`--`r h1[[3]]`% 95% CI) lower variability. As examples,
 revenue variability was highest for the herring roe gillnet and king crab fisheries.
@@
 credible intervals of expected standard deviations of year-to-year revenue rates
-of return.}
+of return. Angled lines connects permit sets that differ by one permit.}
 
@@
 as individuals become accustomed to new locations, gear, or species.
-For example, many individuals fishing herring in Alaska in the 1970s 
+<!-- For example, many individuals fishing herring in Alaska in the 1970s --> 
 <!-- AB: I think in this paragraph it would be good to pick an example or two and speculate on potential reasons specific to those permit strategies.

### 2017-05-15 portfolio/text/anderson-etal-rev-pfx.tex
Editing for reviewer comments

@@
 gear type and species composition (Supporting Information).
+Gear types and permits are closely linked since in most cases a fisher would require a different permit to use different gear.
+The ability to diversify without changing gear or permits is therefore driven by biological availability and small-scale choices by fishers, such as fishing early versus late in a year or fishing in particular locations.
+% Response to reviewer comment A5
 

### 2017-05-15 portfolio/text/ms.Rmd
Editing for reviewer comments

@@
 season length constant from year-to-year. Dots and line segments represent
-medians and 50\% credible intervals of posteriors. The effect of generalizing
+medians and 50\% credible intervals of posteriors. 
+In other words, the axes represent coefficient values from the strategy-level slopes.
+Moving along an axis (say from left to right on the horizontal axis representing the effect of specializing on revenue) represents an increase in how steep the slope is between specialization and revenue.
+The effect of generalizing
 (increasing species diversity) is similar (Fig. S\figmirror).}
@@
 agricultural fields [@ikerd1990; @abdulai2000].
-<!-- For example, rice farmers in -->
-<!-- Ghana can increase their efficiency by devoting more land and labor to rice -->
-<!-- production compared to other crops [@abdulai2000]. -->
 Second, given the
@@
 when a financial crisis hit [@ikerd1990].
-<!-- AB: We might want to include more examples / references from fisheries.
-herring - affected by closures? value declined...-->
 Third, diversifying---a lack of
@@
 would lead to a shift towards diversification.
+<!-- Response to A1 -->
+Conversely, a specialization may appear as a response to increased revenue.
+A fisher who achieves high revenue from a hotspot of a particular fish species may choose to stop or pause fishing thereby indirectly limiting catch diversity.
+For example, pink and chum salmon are the dominant species caught in purse seine fisheries in Prince william Sound and their run timing rarely overlaps.
+If a fisher achieves high revenue with one species that fisher may cut the season short.
 
@@
 
+<!-- there are highly variable costs among fisheries with gear and permits in particular 
+but it is not clear that costs vary systematically with the diversity of the catch 
+there are opportunity costs to ... -->
+<!-- gear costs incur considerable capital investment and some maintenance costs 
+Permit types and gear types are closely linked and therefore fishers largely rely on the same gear regardless of whether you fish a specialized or diverse catch 
+people have limited flexibility to modify gear within permits which are largely determined by regulations 
+some permit combinations would allow a fisher to fish multiple permits with a single boat while others would require multiple boats
+just expand our list of cost factors and call it a day 
+costs such as leasing vessels from other fishers
+maintaining equipment  -->
+
 There are many possible barriers to diversification for individuals relying on
@@
 [@abdulai2001; @barrett2001], small-scale fishing communities [@olale2012], and
-those relying on forest resources [@illukpitiya2008]. Another major barrier to
+those relying on forest resources [@illukpitiya2008].
+<!-- Response to reviewer comment A2 -->
+While we were unable to incorporate costs, it is possible that including license and gear purchase costs along with others such as costs of gear maintenance, catch transportation, and selling to a processor would unveil a different relationship between catch diversification and profits than the relationships between catch diversification and revenue shown here.
+Another major barrier to
 diversification can be management restrictions. In commercial fisheries,
@@
 individuals over 30 years in one of the most productive fishing regions in the 
-world. However, this large temporal and geographical scope potentially obscures 
+world. 
+<!-- Response to reviewer comment B2 -->
+Although we focus on Alaska, and expect our results are largely generalizable to other productive and intensely managed fishing regions, it is possible that the benefits of diversification would be stronger in regions where fish stocks are highly variable or in less intensely managed regions where permit costs are lower and more freely available.
+Overall, the large temporal and geographical scope of our analysis potentially obscures 
 time or regional-variation in the consequences of diversification on revenue and
@@
 benefits, to restrict individual diversification opportunities and promote
-specialization. Although variability in individual revenue may not be actively
+specialization. 
+From a management perspective, metrics such as interannual variability in individual fishing revenue or fraction of single permit holders within communities could serve as bioeconomic indicators.
+Although variability in individual revenue is not actively
 incorporated into management decisions, analyses such as ours help to explain

### 2017-05-16 portfolio/text/anderson-etal-rev-pfx.tex
More edits incorporating reviewer comments

--- 2017-05-05 portfolio/text/anderson-etal-rev-pfx.tex
@@
 gear type and species composition (Supporting Information).
+Gear types and permits are closely linked since in most cases a fisher would require a different permit to use different gear.
+The ability to diversify without changing or adding gear or permits is therefore driven by biological availability and choices by fishers, such as fishing early versus late in a year or fishing in particular locations.
+% Response to reviewer comment A5
 
@@
 increasing or decreasing species diversity. We exponentiate the equation to
-ensure all scale parameters, \(\sigma_{i,t}\), are positive. Residuals from the
-mean component of our model clearly show a ``V''-shaped pattern, which the
-variance component of our model can capture (Fig.~S\figresid). Furthermore, the
-downside residuals (year-individual data points with lower than expected
+ensure all scale parameters, \(\sigma_{i,t}\), are positive. Residuals from
+the mean component of our model clearly show a ``V''-shaped pattern, which the
+variance component of our model can capture (Fig.~S\figresid). Furthermore,
+the downside residuals (year-individual data points with lower than expected
 revenue) appear qualitatively similar to the complete set of residuals (upside
 and downside) (Fig.~S\figresiddownside). Therefore, it seemed appropriate to
-model the variability as a proxy for revenue risk.
+model variability as a proxy for revenue risk. Even though individual upward
+spikes in revenue may be desirable and would be reflected by increased revenue
+variability, increases in average revenue variability are unlikely to be
+desirable.
 

### 2017-05-16 portfolio/text/ms.Rmd
More edits incorporating reviewer comments

@@
 fish in multiple regions. The effectiveness of these diversification strategies
-for increasing revenue and reducing revenue variability remains uncertain. For
-example, theory and empirical studies demonstrate that there can be substantial
+for increasing revenue and reducing revenue variability remains uncertain. Theory and empirical studies demonstrate that there can be substantial
 benefits to specialization for both farmers and fishers [@smith1986;
@@
 world with ~2.7 billion kg landed biomass valued at nearly 2 billion USD
-annually [@nmfs2016]. The days of open, unregulated fisheries in most developed
-countries have passed---fishers must own a permit to participate in virtually
+annually [@nmfs2016]. Fishers must own a permit to participate in virtually
 all fisheries [@hilborn2003a]. While such regulations have helped reduce
@@
 and across fishing strategies while controlling for fishing effort and latent
-time effects (e.g. changing environmental conditions). The model was used to
-evaluate two axes by which fishers in Alaska can diversify fishing activities:
+time effects (e.g. changing environmental conditions). We use the model to evaluate two axes by which fishers in Alaska can diversify fishing activities:
 (1) choice of permit strategy (i.e., permit or set of permits fished), and (2)
@@
 \begin{center}
-\includegraphics[width=0.5\textwidth]{../figs/fig1-jk.pdf}
+\includegraphics[width=0.45\textwidth]{../figs/fig1-jk.pdf}
 
@@
 \begin{center}
-\includegraphics[width=0.5\textwidth]{../figs/stan-across-strategy-variability2.pdf}
+\includegraphics[width=0.45\textwidth]{../figs/stan-across-strategy-variability2.pdf}
 
@@
 \begin{center}
-\includegraphics[width=0.8\textwidth]{../figs/stan-gg-spoke2.pdf}
+\includegraphics[width=0.65\textwidth]{../figs/stan-gg-spoke2.pdf}
 
@@
 \begin{center}
-\includegraphics[width=0.5\textwidth]{../figs/stan-offset-break-anti-spaghetti-specializing.pdf}
+\includegraphics[width=0.46\textwidth]{../figs/stan-offset-break-anti-spaghetti-specializing.pdf}
 
@@
 In other words, the axes represent coefficient values from the strategy-level slopes.
-Moving along an axis (say from left to right on the horizontal axis representing the effect of specializing on revenue) represents an increase in how steep the slope is between specialization and revenue.
+Moving along an axis (say from left to right on the horizontal axis) represents an increase in how steep the slope is between specialization and the year-to-year revenue ratio.
 The effect of generalizing
@@
 decreases revenue variability. However, within the constraints of a permit strategy
-and accounting for effort, specializing tends to increase revenue while changes to fished species diversity tend to correlate with greater revenue variability
+and accounting for effort, specializing tends to increase revenue while changes to fished species diversity tend to correlate with greater revenue variability,
 although there are considerable differences amongst fisheries. Thus, the most effective
@@
 Conversely, a specialization may appear as a response to increased revenue.
-A fisher who achieves high revenue from a hotspot of a particular fish species may choose to stop or pause fishing thereby indirectly limiting catch diversity.
+A fisher who achieves high revenue from a hotspot of a particular fish species one year may choose to stop or pause fishing thereby indirectly limiting catch diversity.
 For example, pink and chum salmon are the dominant species caught in purse seine fisheries in Prince william Sound and their run timing rarely overlaps.
@@
 <!-- Response to reviewer comment A2 -->
-While we were unable to incorporate costs, it is possible that including license and gear purchase costs along with others such as costs of gear maintenance, catch transportation, and selling to a processor would unveil a different relationship between catch diversification and profits than the relationships between catch diversification and revenue shown here.
+While we were unable to incorporate costs, it is possible that including them 
+ <!-- license and gear purchase costs along with others such as costs of gear maintenance, catch transportation, and selling to a processor -->
+would unveil a different relationship between catch diversification and profits than the relationships between catch diversification and revenue shown here.
 Another major barrier to
@@
 <!-- Response to reviewer comment B2 -->
-Although we focus on Alaska, and expect our results are largely generalizable to other productive and intensely managed fishing regions, it is possible that the benefits of diversification would be stronger in regions where fish stocks are highly variable or in less intensely managed regions where permit costs are lower and more freely available.
-Overall, the large temporal and geographical scope of our analysis potentially obscures 
+Although we focus on Alaska, and expect our results are largely generalizable to other productive and intensely managed fishing regions, it is possible that the benefits of diversification would be stronger in regions where fish stocks are highly variable or in regions where permit costs are lower and more freely available.
+<!--Overall, the large temporal and geographical scope of our analysis potentially obscures 
 time or regional-variation in the consequences of diversification on revenue and
-revenue variability. Furthermore, in addressing the overall pattern, we have
+revenue variability. -->
+Furthermore, in addressing the overall pattern, we have
 averaged out the potential signal of several major discrete events, such as the
@@
 From a management perspective, metrics such as interannual variability in individual fishing revenue or fraction of single permit holders within communities could serve as bioeconomic indicators.
-Although variability in individual revenue is not actively
+Although variability in individual revenue is not yet actively
 incorporated into management decisions, analyses such as ours help to explain

### 2017-05-16 portfolio/text/ms.Rmd
PWS seine sentence

@@
 A fisher who achieves high revenue from a hotspot of a particular fish species one year may choose to stop or pause fishing thereby indirectly limiting catch diversity.
-For example, pink and chum salmon are the dominant species caught in purse seine fisheries in Prince william Sound and their run timing rarely overlaps.
-If a fisher achieves high revenue with one species that fisher may cut the season short.
+For example, pink and chum salmon are the dominant species caught in purse seine fisheries in Prince William Sound, with chum salmon generally arriving
+before pink salmon. If a fisher targeting chum achieves high revenue early in 
+the season, they may choose to fish less later in the year.
 

### 2017-05-19 portfolio/text/ms.Rmd
Fixed typos

--- 2017-05-16 portfolio/text/ms.Rmd
@@
 possible to track people that switch fishery permits over time or use multiple
-vessels and differentiate fishers that share the same vessel. We developped and
+vessels and differentiate fishers that share the same vessel. We developed and
 applied a hierarchical Bayesian model to understand the effect of species
@@
 Conversely, a specialization may appear as a response to increased revenue.
-A fisher who achieves high revenue from a hotspot of a particular fish species one year may choose to stop or pause fishing thereby indirectly limiting catch diversity.
-For example, pink and chum salmon are the dominant species caught in purse seine fisheries in Prince william Sound and their run timing rarely overlaps.
-If a fisher achieves high revenue with one species that fisher may cut the season short.
+A fisher who achieves high revenue from a hotspot of a particular fish species
+one year may choose to stop or pause fishing thereby indirectly limiting catch
+diversity. For example, pink and chum salmon are the dominant species caught in
+purse seine fisheries in Prince William Sound, with chum salmon generally
+harvested earlier in the season than pink salmon. If fishers targeting chum salmon
+achieve high revenue early in the season, they may choose to fish less later in
+the season.
 

### 2017-05-24 portfolio/text/ms.Rmd
Final edits before resubmission

@@
 \begin{center}
-\includegraphics[width=0.45\textwidth]{../figs/fig1-jk.pdf}
+\includegraphics[width=0.45\textwidth]{fig1-jk.pdf}
 
@@
 \begin{center}
-\includegraphics[width=0.45\textwidth]{../figs/stan-across-strategy-variability2.pdf}
+\includegraphics[width=0.45\textwidth]{stan-across-strategy-variability2.pdf}
 
@@
 \begin{center}
-\includegraphics[width=0.65\textwidth]{../figs/stan-gg-spoke2.pdf}
+\includegraphics[width=0.65\textwidth]{stan-gg-spoke2.pdf}
 
@@
 \begin{center}
-\includegraphics[width=0.46\textwidth]{../figs/stan-offset-break-anti-spaghetti-specializing.pdf}
+\includegraphics[width=0.46\textwidth]{stan-offset-break-anti-spaghetti-specializing.pdf}
 
@@
 <!-- Response to reviewer comment A2 -->
-While we were unable to incorporate costs, it is possible that including them 
- <!-- license and gear purchase costs along with others such as costs of gear maintenance, catch transportation, and selling to a processor -->
-would unveil a different relationship between catch diversification and profits than the relationships between catch diversification and revenue shown here.
+While we were unable to incorporate costs, it is possible that 
+including license and gear purchase costs along with others such as costs of
+gear maintenance, catch transportation, and selling to a processor
+would alter the relationship between catch diversification and profits.
+<!-- unveil a different relationship between catch diversification and profits than the relationships between catch diversification and revenue shown here. -->
 Another major barrier to

### 2017-05-29 portfolio/text/strategy-table.tex
Update strategy table compilation

% latex table generated in R 3.3.2 by xtable 1.8-2 package
Label & Description & Mean spp. diversity & Main species \\ 
  \midrule
Dun C & Dungeness crab & 1.0 & crab \\ 
  King C & King crab & 1.0 & crab \\ 
  Herr - pound & Herring, pound & 1.0 & herring \\ 
  Sea cuc & Sea cucumber & 1.0 & sea cucumber \\ 
  Tan C & Tanner crab, pot gear & 1.0 & crab \\ 
  Hal - ll & Halibut, longline & 1.0 & halibut \\ 
  Fin - pot & Finfish, pot gear, statewide & 1.0 & cod \\ 
  Herr roe - gill & Herring roe, gillnet & 1.0 & herring \\ 
  Sab & Sablefish, longline, statewide & 1.0 & sablefish \\ 
  S - drift - CI, AP, BB (sock) & Salmon (mostly sockeye), drift gillnet, CI, AP, BB & 1.1 & sock \\ 
  Fin - ll & Finfish, longline, statewide & 1.1 & pollock \\ 
  S - set - PWS, CI, K, AP, BB (sock) & Salmon (mostly sockeye), set gillnet, PWS, CI, K,  AP, BB & 1.2 & sock \\ 
  S - seine - CH & Salmon, purse seine, Chignik & 1.3 & sock \\ 
  S - seine - PWS & Salmon, purse seine, PWS & 1.4 & pink, chum \\ 
  Fin - otter & Finfish, otter trawl, statewide & 1.5 & cod \\ 
  S - set - YAK & Salmon, set gillnet, Yakutat & 1.7 & coho, sock \\ 
  S - troll & Salmon, power troll, statewide & 1.8 & chnk, coho \\ 
  S - seine - CI, K & Salmon, purse seine, Cook Inlet and Kodiak & 1.9 & pink, sock \\ 
  S - drift - PWS & Salmon, drift gillnet, PWS & 2.0 & coho, pink, sock, chum \\ 
  S - seine - SE & Salmon, purse seine, Southeast Alaska & 2.1 & pink, sock, chum \\ 
  S - seine - AP & Salmon, purse seine, Alaska Penninsula & 2.2 & pink, sock, chum \\ 
  S - set - NS/KU & Salmon, set gillnet, Norton Sound and Kuskokwim & 2.5 & chnk, coho, sock, chum \\ 
  S - drift - SE & Salmon, drift gillnet, Southeast Alaska & 2.5 & coho, pink, sock, chum \\ 
   \bottomrule

### 2017-05-29 portfolio/text/supp.Rnw
Tweak format for editorial office

@@
 \usepackage{booktabs}
-\usepackage{verbatim}
+% \usepackage{verbatim}
 \usepackage{amsmath}
@@
 \usepackage{mathtools}
-\usepackage{lineno}
-\usepackage{xcolor}
-\renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
-\modulolinenumbers[2]
+% \usepackage{lineno}
+% \usepackage{xcolor}
+% \renewcommand\linenumberfont{\normalfont\tiny\sffamily\color{gray}}
+% \modulolinenumbers[2]
 
-% Linux Libertine:
-\usepackage{textcomp}
-\usepackage[sb]{libertine}
-\usepackage[varqu,varl]{inconsolata}% sans serif typewriter
-\usepackage[libertine,bigdelims,vvarbb]{newtxmath} % bb from STIX
-\usepackage[cal=boondoxo]{mathalfa} % mathcal
-%\useosf % osf for text, not math
-\usepackage[supstfm=libertinesups,%
-  supscaled=1.2,%
-  raised=-.13em]{superiors}
+% % Linux Libertine:
+% \usepackage{textcomp}
+% \usepackage[sb]{libertine}
+% \usepackage[varqu,varl]{inconsolata}% sans serif typewriter
+% \usepackage[libertine,bigdelims,vvarbb]{newtxmath} % bb from STIX
+% \usepackage[cal=boondoxo]{mathalfa} % mathcal
+% %\useosf % osf for text, not math
+% \usepackage[supstfm=libertinesups,%
+%   supscaled=1.2,%
+%   raised=-.13em]{superiors}
 
@@
 % http://phaseportrait.blogspot.ca/2007/08/lineno-and-amsmath-compatibility.html
-\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
-  \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
-  \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
-  \renewenvironment{#1}%
-     {\linenomath\csname old#1\endcsname}%
-     {\csname oldend#1\endcsname\endlinenomath}}%
-\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
-  \patchAmsMathEnvironmentForLineno{#1}%
-  \patchAmsMathEnvironmentForLineno{#1*}}%
-\AtBeginDocument{%
-\patchBothAmsMathEnvironmentsForLineno{equation}%
-\patchBothAmsMathEnvironmentsForLineno{align}%
-\patchBothAmsMathEnvironmentsForLineno{flalign}%
-\patchBothAmsMathEnvironmentsForLineno{alignat}%
-\patchBothAmsMathEnvironmentsForLineno{gather}%
-\patchBothAmsMathEnvironmentsForLineno{multline}%
-}
+% \newcommand*\patchAmsMathEnvironmentForLineno[1]{%
+%   \expandafter\let\csname old#1\expandafter\endcsname\csname #1\endcsname
+%   \expandafter\let\csname oldend#1\expandafter\endcsname\csname end#1\endcsname
+%   \renewenvironment{#1}%
+%      {\linenomath\csname old#1\endcsname}%
+%      {\csname oldend#1\endcsname\endlinenomath}}%
+% \newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
+%   \patchAmsMathEnvironmentForLineno{#1}%
+%   \patchAmsMathEnvironmentForLineno{#1*}}%
+% \AtBeginDocument{%
+% \patchBothAmsMathEnvironmentsForLineno{equation}%
+% \patchBothAmsMathEnvironmentsForLineno{align}%
+% \patchBothAmsMathEnvironmentsForLineno{flalign}%
+% \patchBothAmsMathEnvironmentsForLineno{alignat}%
+% \patchBothAmsMathEnvironmentsForLineno{gather}%
+% \patchBothAmsMathEnvironmentsForLineno{multline}%
+% }
 
@@
 We checked our models with a two-stage modelling approach
-using the R package lme4 \citep{bates2015}.
+using the R package lme4 [49].
 To do this, we fit the initial model of the mean component $\log(R_{i,t})$
@@
 
-\bibliographystyle{apalike}
-\bibliography{refs}
+% \bibliographystyle{apalike}
+% \bibliography{refs}
 
@@
   \toprule
-\input{../figs/strategy-table.tex}
+\input{strategy-table.tex}
   \end{tabular}
@@
 \begin{center}
-\includegraphics[width=1\textwidth]{../figs/stan-gg-spoke-all.pdf}
+\includegraphics[width=1\textwidth]{stan-gg-spoke-all.pdf}
 \caption{The effect of adding a permit to one's strategy
@@
 \begin{center}
-\includegraphics[width=5in]{../figs/stan-main-effects.pdf}
+\includegraphics[width=5in]{stan-main-effects.pdf}
 \caption{Main effects from the hierarchical model
@@
 \begin{center}
-\includegraphics[width=1.06\textwidth]{../figs/stan-str-posteriors-dot.pdf}
+\includegraphics[width=1.06\textwidth]{stan-str-posteriors-dot.pdf}
 \caption{Estimates of within-strategy effects from the hierarchical model.
@@
 \begin{center}
-\includegraphics[width=1.05\textwidth]{../figs/stan-offset-break-anti-spaghetti.pdf}
+\includegraphics[width=1.05\textwidth]{stan-offset-break-anti-spaghetti.pdf}
 \caption{Effects of (A) generalizing and (B) specializing from year-to-year for
@@
 \begin{center}
-\includegraphics[width=0.6\textwidth]{../figs/stan-offset-break-anti-spaghetti-specializing-no-effort.pdf}
+\includegraphics[width=0.6\textwidth]{stan-offset-break-anti-spaghetti-specializing-no-effort.pdf}
 \caption{Same as Fig.~4 but without the effort (days fished) predictor and
@@
 \begin{center}
-\includegraphics[width=1\textwidth]{../figs/itq-combined.pdf}
+\includegraphics[width=1\textwidth]{itq-combined.pdf}
 \caption{The effect of IFQs (individual fishing quotas) on model estimates. (A)
@@
 \begin{center}
-\includegraphics[width=1\textwidth]{../figs/strategy-year-effects.pdf}
+\includegraphics[width=1\textwidth]{strategy-year-effects.pdf}
 \caption{Strategy-year intercept estimates ($\beta_{0,j,t}$ in
@@
 \begin{center}
-\includegraphics[width=1\textwidth]{../figs/stan-offset-panel-mean-resid.png}
+\includegraphics[width=1\textwidth]{stan-offset-panel-mean-resid.png}
 \caption{The log-absolute value residuals from the mean-component model (Eq.~\ref{eq:mean}).
@@
 \begin{center}
-\includegraphics[width=1\textwidth]{../figs/stan-offset-panel-mean-resid-down.png}
+\includegraphics[width=1\textwidth]{stan-offset-panel-mean-resid-down.png}
 \caption{The log-absolute value of downside (negative) residuals from the main
@@
 % \begin{center}
-% \includegraphics[width=0.9\textwidth]{../figs/lmer-vs-stan2.pdf}
+% \includegraphics[width=0.9\textwidth]{lmer-vs-stan2.pdf}
 % \caption{Comparison of }

### 2017-07-07 portfolio/text/anderson-etal-rev-pfx.tex
Tweak format for editorial office
Wrap abstract and add alternative title

@@
 \title{Benefits and risks of diversification for individual fishers}
+% \title{Diversification trade-offs for individual fishers}
 
@@
 \begin{abstract}
-Individuals relying on the extraction of natural resources for their livelihood face high levels of income variability driven by a mix of environmental, biological, management, and economic factors.
-Key to managing these industries is identifying how regulatory actions and individual behaviour affect income variability and, by extension, economic stability and the sustainable use of natural resources.
-In commercial fisheries, communities and vessels fishing a greater diversity of species have less revenue variability than those fishing fewer species.
-However, it is unclear if these benefits extend to the actions of individual fishers and how year-to-year changes in diversification affect revenue and revenue variability.
-Here, we evaluate two axes by which fishers in Alaska can diversify fishing activities.
-We show that, despite increasing specialization over the last 30 years, fishing a set of permits with higher species diversity reduces individual revenue variability and fishing an additional permit is associated with higher revenue and lower variability.
-However, increasing species diversity within the constraints of existing permits has a fishery-dependent effect on revenue and is usually (87\% probability) associated with increased revenue uncertainty the following year.
-Our results demonstrate that the most effective option for individuals to decrease revenue variability is to participate in additional or more diverse fisheries.
-However, this option is expensive, often limited by regulations such as catch share programs, and consequently unavailable to many individuals.
-With increasing climatic variability it will be particularly important that individuals relying on natural resources for their livelihood have effective strategies to reduce income variability.
+Individuals relying on the extraction of natural resources for their livelihood
+face high levels of income variability driven by a mix of environmental,
+biological, management, and economic factors. Key to managing these industries
+is identifying how regulatory actions and individual behaviour affect income
+variability and risk and, by extension, economic stability and the sustainable use of
+natural resources. In commercial fisheries, communities and vessels fishing a
+greater diversity of species have less revenue variability than those fishing
+fewer species. However, it is unclear if these benefits extend to the actions of
+individual fishers and how year-to-year changes in diversification affect
+revenue and revenue variability. Here, we evaluate two axes by which fishers in
+Alaska can diversify fishing activities. We show that, despite increasing
+specialization over the last 30 years, fishing a set of permits with higher
+species diversity reduces individual revenue variability and fishing an
+additional permit is associated with higher revenue and lower variability.
+However, increasing species diversity within the constraints of existing permits
+has a fishery-dependent effect on revenue and is usually (87\% probability)
+associated with increased revenue uncertainty the following year. Our results
+demonstrate that the most effective option for individuals to decrease revenue
+variability is to participate in additional or more diverse fisheries. However,
+this option is expensive, often limited by regulations such as catch share
+programs, and consequently unavailable to many individuals. With increasing
+climatic variability it will be particularly important that individuals relying
+on natural resources for their livelihood have effective strategies to reduce
+income variability.
 % (`r g2_dens`% probability)
@@
 this paper, along with all code, are archived in a publicly available repository at
-\url{https://github.com/NCEAS/pfx-commercial/tree/master/portfolio}. } \showmatmethods
+\url{https://github.com/NCEAS/pfx-commercial/tree/master/portfolio}.
+
+\nocite{bates2015}
+} \showmatmethods
 

### 2017-07-07 portfolio/text/ms.Rmd
Clean up a couple variability risk words in the introductory paragraph

@@
 variability in income relative to other professions [@mishra2002; @sethi2012b].
-In addition to economic risk, such as changes in demand or prices, individuals 
+In addition to income variability from economic sources,
+such as changes in demand or prices, individuals
 dependent on natural resources are also subject to biological and environmental

### 2017-07-07 portfolio/text/ms.Rmd
obtainable by -> available to
simpler and more parallel
Edit last paragraph of introduction; dealing with reviewer comment

@@
 agriculture [@sethi2012b]. The majority of risk insurance programs available to
-terrestrial farmers (e.g. crop insurance programs) are not obtainable by
+terrestrial farmers (e.g. crop insurance programs) are not available to
 commercial fishers [@kasperski2013]. However, fishers may be able to lower risk
@@
 possible to track people that switch fishery permits over time or use multiple
-vessels and differentiate fishers that share the same vessel. We developed and
-applied a hierarchical Bayesian model to understand the effect of species
-diversity within fisheries catches on individual revenue and revenue
-variability. Our model builds on results from vessel [@kasperski2013], fleet
-[@sethi2012b], and community-level [@sethi2014; @cline2017] variability. We
-developped a variance-function regression model [@davidian1987; @hall1989], to
-simultaneously model changes in the mean and variance of individual revenue as a
-function of covariates. We simultaneously estimate the effect of species
+vessels and differentiate fishers that share the same vessel. We develop
+a hierarchical variance-function regression model [@davidian1987; @hall1989] to
+simultaneously understand the effect of species
+diversity within fisheries catches on individual revenue and revenue variability.
+Our model builds on results from vessel [@kasperski2013], fleet [@sethi2012b],
+and community-level [@sethi2014; @cline2017] variability,
+and although we model revenue *variability*, we herein equate revenue
+variability with risk for consistency with Kasperski and Holland [@kasperski2013].
+We simultaneously estimate the effect of species
 diversity (the inverse of Simpson's diversity index [@simpson1949; @jost2006])
@@
 and across fishing strategies while controlling for fishing effort and latent
-time effects (e.g. changing environmental conditions). We use the model to evaluate two axes by which fishers in Alaska can diversify fishing activities:
+time effects (e.g. changing environmental conditions).
+We use the model to evaluate two axes by which fishers in Alaska can
+diversify fishing activities:
 (1) choice of permit strategy (i.e., permit or set of permits fished), and (2)

### 2017-07-07 portfolio/text/ms.Rmd
Wrapping text, deleting trailing white space, deleting comments

@@
 species as allowed by a permit, own multiple permits, or diversify spatially and
-fish in multiple regions. The effectiveness of these diversification strategies
-for increasing revenue and reducing revenue variability remains uncertain. Theory and empirical studies demonstrate that there can be substantial
+fish in multiple regions.
+
+The effectiveness of these diversification strategies
+for increasing revenue and reducing revenue variability remains uncertain.
+Theory and empirical studies demonstrate that there can be substantial
 benefits to specialization for both farmers and fishers [@smith1986;
@@
 effects of diversification on revenue do so at aggregate levels that combine
-individuals (e.g., counties or communities; [@kasperski2013;
+individuals (e.g. counties or communities; [@kasperski2013;
 @sethi2014; @cline2017]). Thus, key uncertainties remain about the relationship
@@
 of the individual.
-
-<!-- important to acknowledge that IFQ regs have been implemented partly for economic -->
-<!-- reasons and have been successful by some economic measures. Might even want to -->
-<!-- insert the above sentence. -->
 
@@
 Southeast Alaska in 2016: <https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>.
-
-<!-- TODO AH: I think our fisher apporach is good, but I wonder if it’s worth
-a footnote to address the potential critique that at the indivdual level we
-don’t observe their wages in non-fishing activities.  I’m not sure if this is
-worth bringing up, but this is always the case anyway with vessels where the
-skipper may do other things. triue, and for others - KH and Cl et al. ...-->
 
@@
 decreases revenue variability. However, within the constraints of a permit strategy
-and accounting for effort, specializing tends to increase revenue while changes to fished species diversity tend to correlate with greater revenue variability,
+and accounting for effort, specializing tends to increase revenue while changes
+to fished species diversity tend to correlate with greater revenue variability,
 although there are considerable differences amongst fisheries. Thus, the most effective
@@
 portfolio diversity following downward economic shocks [@debela2012].
-Because specialization inherently holds risk from ``putting all one's eggs in one
-basket,'' it is not suprising that a decline in the most valuable resource
+Because specialization inherently holds risk from "putting all one's eggs in one
+basket," it is not suprising that a decline in the most valuable resource
 would lead to a shift towards diversification.
-<!-- Response to A1 -->
 Conversely, a specialization may appear as a response to increased revenue.
@@
 as individuals become accustomed to new locations, gear, or species.
-<!-- For example, many individuals fishing herring in Alaska in the 1970s --> 
-<!-- AB: I think in this paragraph it would be good to pick an example or two and speculate on potential reasons specific to those permit strategies.
-Otherwise, statements like this one aren’t really anchored to anything concrete.
-For example, what does increased specialization look like in the Hal ll permit, which showed one of the greatest effects of specialization on revenue diversity?
-How many permits could a Hal ll permit holder potentially fish?
-This goes back to my comment re: Table 1 (see p.
-15)
-sea cuc - H W book - learning...
-herring - alaska 1970s - learn to fish for roe ... [ref]
-salmon - hatchery production - change target from diverse to most abundant... PWS seine - pink salmon or SE
-Hal ll + S troll...
--->
 Additionally, changes in diversity might reflect years in which targeted species
@@
 
-<!-- there are highly variable costs among fisheries with gear and permits in particular 
-but it is not clear that costs vary systematically with the diversity of the catch 
-there are opportunity costs to ... -->
-<!-- gear costs incur considerable capital investment and some maintenance costs 
-Permit types and gear types are closely linked and therefore fishers largely rely on the same gear regardless of whether you fish a specialized or diverse catch 
-people have limited flexibility to modify gear within permits which are largely determined by regulations 
-some permit combinations would allow a fisher to fish multiple permits with a single boat while others would require multiple boats
-just expand our list of cost factors and call it a day 
-costs such as leasing vessels from other fishers
-maintaining equipment  -->
-
 There are many possible barriers to diversification for individuals relying on
@@
 those relying on forest resources [@illukpitiya2008].
-<!-- Response to reviewer comment A2 -->
 While we were unable to incorporate costs, it is possible that
@@
 would alter the relationship between catch diversification and profits.
-<!-- unveil a different relationship between catch diversification and profits than the relationships between catch diversification and revenue shown here. -->
 Another major barrier to
@@
 assessments [@melnychuk2017] and individual quota systems
-[@costello2008]---may ultimately restrict individual diversification [@smith1986; @kasperski2013]
+[@costello2008]---may ultimately restrict individual diversification
+[@smith1986; @kasperski2013]
 or at least make developing a diverse fishing portfolio more expensive. However,
@@
 world.
-<!-- Response to reviewer comment B2 -->
-Although we focus on Alaska, and expect our results are largely generalizable to other productive and intensely managed fishing regions, it is possible that the benefits of diversification would be stronger in regions where fish stocks are highly variable or in regions where permit costs are lower and more freely available.
-<!--Overall, the large temporal and geographical scope of our analysis potentially obscures 
-time or regional-variation in the consequences of diversification on revenue and
-revenue variability. -->
+Although we focus on Alaska, and expect our results are largely generalizable to
+other productive and intensely managed fishing regions, it is possible that the
+benefits of diversification would be stronger in regions where fish stocks are
+highly variable or in regions where permit costs are lower and more freely
+available.
 Furthermore, in addressing the overall pattern, we have
@@
 specialization.
-From a management perspective, metrics such as interannual variability in individual fishing revenue or fraction of single permit holders within communities could serve as bioeconomic indicators.
+From a management perspective, metrics such as interannual variability in
+individual fishing revenue or fraction of single permit holders within
+communities could serve as bioeconomic indicators.
 Although variability in individual revenue is not yet actively

### 2017-07-07 portfolio/text/anderson-etal-rev-pfx.tex
Add financial risk to abstract

@@
 \begin{abstract}
+
 Individuals relying on the extraction of natural resources for their livelihood
@@
 is identifying how regulatory actions and individual behaviour affect income
-variability and risk and, by extension, economic stability and the sustainable use of
-natural resources. In commercial fisheries, communities and vessels fishing a
-greater diversity of species have less revenue variability than those fishing
-fewer species. However, it is unclear if these benefits extend to the actions of
-individual fishers and how year-to-year changes in diversification affect
-revenue and revenue variability. Here, we evaluate two axes by which fishers in
-Alaska can diversify fishing activities. We show that, despite increasing
-specialization over the last 30 years, fishing a set of permits with higher
-species diversity reduces individual revenue variability and fishing an
-additional permit is associated with higher revenue and lower variability.
-However, increasing species diversity within the constraints of existing permits
-has a fishery-dependent effect on revenue and is usually (87\% probability)
-associated with increased revenue uncertainty the following year. Our results
-demonstrate that the most effective option for individuals to decrease revenue
-variability is to participate in additional or more diverse fisheries. However,
-this option is expensive, often limited by regulations such as catch share
-programs, and consequently unavailable to many individuals. With increasing
-climatic variability it will be particularly important that individuals relying
-on natural resources for their livelihood have effective strategies to reduce
-income variability.
+variability, financial risk, and, by extension, the economic stability and the
+sustainable use of natural resources. In commercial fisheries, communities and
+vessels fishing a greater diversity of species have less revenue variability
+than those fishing fewer species. However, it is unclear if these benefits
+extend to the actions of individual fishers and how year-to-year changes in
+diversification affect revenue and revenue variability. Here, we evaluate two
+axes by which fishers in Alaska can diversify fishing activities. We show that,
+despite increasing specialization over the last 30 years, fishing a set of
+permits with higher species diversity reduces individual revenue variability and
+fishing an additional permit is associated with higher revenue and lower
+variability. However, increasing species diversity within the constraints of
+existing permits has a fishery-dependent effect on revenue and is usually (87\%
+probability) associated with increased revenue uncertainty the following year.
+Our results demonstrate that the most effective option for individuals to
+decrease revenue variability is to participate in additional or more diverse
+fisheries. However, this option is expensive, often limited by regulations such
+as catch share programs, and consequently unavailable to many individuals. With
+increasing climatic variability it will be particularly important that
+individuals relying on natural resources for their livelihood have effective
+strategies to reduce financial risk.
+
 % (`r g2_dens`% probability)
@@
 revenue variability while controlling for effort. For consistency with previous
-analyses \cite{vedeld2007, illukpitiya2008, debela2012, kasperski2013, sethi2014, cline2017}, we calculated effective species diversity
+analyses \cite{vedeld2007, illukpitiya2008, debela2012, kasperski2013,
+  sethi2014, cline2017}, we calculated effective species diversity
 for permit holder \(i\) and time (year) \(t\) as the inverse of Simpson's
@@
 and downside) (Fig.~S\figresiddownside). Therefore, it seemed appropriate to
-model variability as a proxy for revenue risk. Even though individual upward
-spikes in revenue may be desirable and would be reflected by increased revenue
-variability, increases in average revenue variability are unlikely to be
-desirable.
+model variability as a proxy for revenue risk. Ideally, we would model
+a more direct measure of risk such as the probability of bankruptcy from low
+revenue; however, data on costs and other sources of income were not available.
 

### 2017-07-07 portfolio/text/ms.Rmd
2016 -> 2017 for permit prices

@@
 [^1]: For example, it cost ~\$192K USD for a king and tanner crab permit in
-Southeast Alaska in 2016: <https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>.
+Southeast Alaska in 2017: <https://www.cfec.state.ak.us/pmtvalue/X_K69A.HTM>.
 

### 2017-07-07 portfolio/text/ms.Rmd
More edits for reviewer response
Clearly defined risk
Add footnote
Add note about modeling risk in the methods

@@
 fish in multiple regions.
-
 The effectiveness of these diversification strategies
@@
 possible to track people that switch fishery permits over time or use multiple
-vessels and differentiate fishers that share the same vessel. We develop
-a hierarchical variance-function regression model [@davidian1987; @hall1989] to
-simultaneously understand the effect of species
-diversity within fisheries catches on individual revenue and revenue variability.
+vessels and differentiate fishers that share the same vessel.
+To simultaneously understand the effect of species
+diversity within fisheries catches on individual revenue and revenue variability,
+we develop a hierarchical variance-function regression model [@davidian1987; @hall1989].
 Our model builds on results from vessel [@kasperski2013], fleet [@sethi2012b],
 and community-level [@sethi2014; @cline2017] variability,
-and although we model revenue *variability*, we herein equate revenue
-variability with risk for consistency with Kasperski and Holland [@kasperski2013].
+and, although we model revenue variability, we herein equate revenue
+variability with financial risk for consistency 
+with Kasperski and Holland [@kasperski2013][^2].
 We simultaneously estimate the effect of species
@@
 field.
+
+[^2]: More specifically, variability is one component of risk. Risk combines
+elements of variability, uncertainty, and loss to define the "possibility of
+a bad thing happening" [@sethi2010b].
 
@@
 \end{figure}
-<!-- TODO AB: To me it is more intuitive that the darker, hotter colors would
-represent higher values. Consdier reversing the color scale to go from red
-(highest # of permit holders) to yellow (lowest #).  -->
 
@@
 
-\caption{Diversifying permits reduces risk and usually increases revenue. Two
+\caption{Diversifying permits reduces revenue variability and usually increases revenue. Two
 examples of expected revenue variability and median revenue for owners of (A)

### 2017-07-07 portfolio/text/supp.Rnw
Stop using minus signs in equations
And add appropriate notes in the captions

@@
 % \usepackage{verbatim}
-\usepackage{amsmath}
+% \usepackage{amsmath}
 \usepackage{url}
-\usepackage{mathtools}
+% \usepackage{mathtools}
 % \usepackage{lineno}
@@
 
+\renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\thetable}{S\arabic{table}}
+
 \begin{document}
 
-\linenumbers
+% \linenumbers
 
@@
 \maketitle
-
-\renewcommand{\thefigure}{S\arabic{figure}}
-\renewcommand{\thetable}{S\arabic{table}}
 
@@
 two times their standard deivations)
-to the magnitude of residual error.}
+to the magnitude of residual error.
+Note that $\mu_{\beta_1}$, $\beta_4$, $\mu_{\gamma_1}$, and $\gamma_4$
+have been multiplied by $-1$
+to make these coefficients interpretable as effects of
+increasing specialization (as opposed to decreasing specialization).
+}
 \label{fig:fe}
@@
 Strategies are ordered from high to low mean
-species diversity from top to bottom.}
+species diversity from top to bottom.
+Note that $\gamma_{1,j}$ and $\beta_{1,j}$ have been multiplied by $-1$
+to make these coefficients interpretable as effects of
+increasing specialization (as opposed to decreasing specialization).}
 \label{fig:re}
@@
   medians and 50\% credible intervals of posteriors. Panel B and Fig.~4 are the
-  same.}
+  same.
+This is a bivariate display of the $\gamma_{1,j}$, $\gamma_{2,j}$, $\beta_{1,j}$,
+  and $\beta_{2,j}$ coefficient estimates displayed in Fig.~\ref{fig:re}.
+  Note that $\gamma_{1,j}$ (panel B vertical axis) and $\beta_{1,j}$ (panel B
+  horizontal axis) have been multiplied by $-1$ to make these coefficients
+  interpretable as effects of increasing specialization (as opposed to
+  decreasing specialization).
+  }
 \label{fig:mirror}

### 2017-07-07 portfolio/text/supp.Rnw
Tweak formatting of supplement

@@
 \geometry{verbose, letterpaper, tmargin = 2.54cm, bmargin = 2.54cm,
-  lmargin = 2.54cm, rmargin = 2.54cm}
+  lmargin = 2.64cm, rmargin = 2.64cm}
 \geometry{letterpaper}
@@
 % \modulolinenumbers[2]
+\usepackage{lscape}
 
@@
 
-\title{Supporting Information}
+\title{Supporting Information Appendix:\\
+  Benefits and risks of diversification for individual fishers}
 
 \author{
+Sean C. Anderson \and
+Eric J. Ward \and
+Andrew O. Shelton \and
+Milo D. Adkison \and
+Anne H. Beaudreau \and
+Richard E. Brenner \and
+Alan C. Haynie \and
+Jennifer C. Shriver \and
+Jordan T. Watson \and
+Benjamin C. Williams
 }
@@
 
+\begin{landscape}
 \begin{table}[h!]
@@
 \end{table}
+\end{landscape}
 
@@
 
+\clearpage
+
 \begin{figure}[htbp]
